### 2010-05-25 Spatial_adaptation/standing-variation.tex
WorkingWiki replaced project description

\documentclass{article}
\begin{document}

Let the local rate of growth at location $x$ and time $t$ be given by $f(x,t)$,
and for each $x$ and $t$ let $(X^{(x,t)}_{s})_{s \ge t}$ have the distribution of a branching process begun at $x$ and time $t$,
and let $\mu^{(x,t)}_s(dz) = \P\{ |X^{(x,t)}_s| \in dz \}$  be the marginal distribution of the size of $X^{(x,t)}$.
For a given mutation rate $\mu$ and population density $\rho(dx)$ we have a Poisson process of inflorescences of rate $\mu$,
each of which having distribution $\mu$.  
Then the set of inflorescences still alive at time $t$ is Poisson, 
and the mean measure of such inflorescences that began at location $x$ is
\[ \nu_x(A) = \int_{-\infty}^t \P\{ X^{(x,s)}_t \in A \} ds .\]
The first thing we'd want to do is to ignore the spatial extent of $X$ and only look at the total number of individuals,
so we need to know
\[ \int_{-\infty}^t \mu^{(x,s)}_t(dz) ds . \]

Now assume everything is homogeneous, and let $Z_t$ be a continuous-time branching process begun at zero 
and having generator
\[ Lg(z) .\]
Suppose that $Z$ has total lifetime $\tau$ and that $\tau$ is finite almost surely.
If we treat each inflorescence as a point mass, then the set of (location, mass) coordinates $(x,z)$
are Poisson with intensity $dx \otimes q(z) dz$ where
\[  
  q(z) = \E \left[ \int_0^\tau \bone(Z_t = z) dt \right] .
\]
We will be interested in the probability that the offspring of any of the initial (assumed to be small) number of individuals in an inflorescence survive,
which is $1 - \widetilde q(1-p_s)$, where $p_s$ is the probability of survival (so $p_s \approx 2s/\sigma^2$) and
\[
  \widetilde q(u) = \sum_z q(z) u^z .
\]

\section{Poisson--Vorono\u\i\ facts}

Fill in some facts about PV tesselations.

\end{document}

### 2010-07-31 Spatial_adaptation/standing-variation.tex
/* Standing variation */  added notes from planning session
/* Standing variation */  some updates

@@
 \documentclass{article}
+\usepackage{amsmath,amssymb}
+
+\newcommand{\var}{\mathop{\mbox{Var}}}
+\renewcommand{\P}{\mathbb{P}}
+\newcommand{\E}{\mathbb{E}}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\one}{\mathbf{1}}
+
 \begin{document}
 
+\section{Size of inflorescences}
+
 Let the local rate of growth at location $x$ and time $t$ be given by $f(x,t)$,
-and for each $x$ and $t$ let $(X^{(x,t)}_{s})_{s \ge t}$ have the distribution of a branching process begun at $x$ and time $t$,
-and let $\mu^{(x,t)}_s(dz) = \P\{ |X^{(x,t)}_s| \in dz \}$  be the marginal distribution of the size of $X^{(x,t)}$.
-For a given mutation rate $\mu$ and population density $\rho(dx)$ we have a Poisson process of inflorescences of rate $\mu$,
-each of which having distribution $\mu$.  
-Then the set of inflorescences still alive at time $t$ is Poisson, 
-and the mean measure of such inflorescences that began at location $x$ is
-\[ \nu_x(A) = \int_{-\infty}^t \P\{ X^{(x,s)}_t \in A \} ds .\]
+and for each $x$ and $t$ let $(X^{(x,t)}_{s})_{s \ge t}$ have the distribution of a branching process begun with one individual at $x$ and time $t$,
+and let $\kappa^{(x,t)}_s(dz) = \P\{ |X^{(x,t)}_s| \in dz \}$  be the marginal distribution of the size of $X^{(x,t)}$.
+For a given mutation rate $\mu$ and population density $\rho(dx)$ we have a Poisson process of inflorescences of rate $\mu \rho$,
+each of which having distribution $\kappa$.  
+Then the set of inflorescences alive at time $t$ is Poisson, 
+and the mean measure of such inflorescences is
+\[ \nu(dx,A) = \mu \rho(dx) \int_{-\infty}^t \P\{ X^{(x,s)}_t \in A \} ds .\]
 The first thing we'd want to do is to ignore the spatial extent of $X$ and only look at the total number of individuals,
 so we need to know
-\[ \int_{-\infty}^t \mu^{(x,s)}_t(dz) ds . \]
+\[ \int_{-\infty}^t \kappa^{(x,s)}_t(dz) ds . \]
 
 Now assume everything is homogeneous, and let $Z_t$ be a continuous-time branching process begun at zero 
-and having generator
-\[ Lg(z) .\]
+and having branching rate $1$ and offspring distribution $\{p_k : k\ge 0\}$.
 Suppose that $Z$ has total lifetime $\tau$ and that $\tau$ is finite almost surely.
 If we treat each inflorescence as a point mass, then the set of (location, mass) coordinates $(x,z)$
-are Poisson with intensity $dx \otimes q(z) dz$ where
+are Poisson with intensity $dx \otimes q(1,z) dz$ where $q(1,z)=q(z)$ is the Green function of $Z$, namely
 \[  
-  q(z) = \E \left[ \int_0^\tau \bone(Z_t = z) dt \right] .
+  q(y,z) = \E^y \left[ \int_0^\tau \one(Z_t = z) dt \right] = \int_0^\infty P^t_{yz} dt .
 \]
-We will be interested in the probability that the offspring of any of the initial (assumed to be small) number of individuals in an inflorescence survive,
-which is $1 - \widetilde q(1-p_s)$, where $p_s$ is the probability of survival (so $p_s \approx 2s/\sigma^2$) and
+and $q$ solves $L q(y,z) = y \sum_k p_k (q(y+k-1,z)-q(y,z)) = \delta_{y=z}$.
+
+If we are interested in the density of ultimately successful influorescences,
+we need to futhermore weight by the probability that the offspring of any of the initial (assumed to be small) number of individuals in an inflorescence survive.
+If $p_s$ is the probability of survival in the new selective regime,
+this leads to the mean measure for successful influorescences
 \[
-  \widetilde q(u) = \sum_z q(z) u^z .
+   dx \int_0^\infty F(p_s,t) dt ,
+\]
+where $F$ is the generating function for $Z_t$,
+\[
+  F(a,t) = \E[a^{Z_t}] ,
+\]
+and let $\phi$ be the generating function for the offspring distribution,
+\[
+  \phi(a) = \sum_k p_k a^k .
+\]
+Now, $p_s$ is the minimal solution to $\phi(a) = a$,
+and more generally, $F$ solves
+\[
+   \partial_t F(a,t) = \phi(F(a,t)) - F(a,t).
 \]

### 2010-09-04 Spatial_adaptation/standing-variation.tex
/* Standing variation */ input of standing versus new case including hypergeometric evaluation of integral

@@
 
+\section{Properties of the mixed model}
+
+Suppose that there are (point masses of) mutations present at time 0 as a Poisson process of rate $\lambda_0$ (calculated above),
+and that after time 0 mutations arise at rate $\lambda$.
+Suppose that the mutations spread distance $f(t)$ in time $t$, 
+and that the volume of the cone of height $t$ is given by $h(t) = \int_0^t \omega_d f(s)^d ds$.
+
+Let $\tau$ denote the time until the origin is adapted.
+Then
+\begin{align}
+   \P \{ \tau > t \} &= \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\}, \quad \mbox{and} \\
+   \E [ \tau ] &= \int_0^\infty \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} .
+\end{align}
+As before, the mean density $\nu$ is equal to $1/\E[\tau]$.
+
+Let $z_0$ be the mean proportion of space covered by mutations present at time 0.
+Then, by the general result on competing inhomogeneous exponentials,
+\begin{align}
+    z_0 &= \P\{ \mbox{origin reached by mutations from $\lambda_0$ before $\lambda$} \} \\
+        &= \int_0^\infty \lambda_0 d \omega_d f'(t) f(t)^{d-1} \exp \left\{ - \lamba_0 \omega_d f(t)^d - \lambda h(t) \right\} dt .
+\end{align}
+The mean density of new mutations that would have taken off but were prevented by standing variation is
+\begin{align}
+    \nu_* = \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt
+\end{align}
+and so the mean proportion of new mutations that would be successful in the absence of standing variation, but are prevented by standing variation,
+is 
+\begin{align}
+    \z_* = \frac{ \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt }
+                { \int_0^\infty \lambda \exp\{ - \lambda h(t) \} dt } .
+\end{align}
+
+If we rescale time by $\alpha$ and space by $\beta$, 
+then this transforms $(\lambda_0, \lambda, f(\cdot))$ into $(\lambda_0/\beta^d, \lambda/(\alpha \beta^d), f(\cdot/\alpha)/\beta)$.  
+Setting $\beta = \lambda_0^{1/d}$ and $\alpha = \lambda / \lambda_0$, this is mapped to $(1,1,f(\cdot \lambda_0/\lambda)/\lambda_0^{1/d})$.
+
+\subsection{Constant speed}
+
+Now let $f(t) = vt$, so $h(t) = \frac{ \omega_d v^d }{ d+1 } t^{d+1}$.
+The integrals that appear above are of the following form, for $\alpha$, $\beta$, $a$, $b$, and $c$ all nonnegative:
+\begin{align}
+  \int_0^\infty t^c \exp \left( - \alpha t^a - \beta t^b \right) dt 
+            &= \left( b^{-1} \beta^{ (1-c)/b } \int_0^\infty u^{(c+1-b)/b} \exp\left( - \alpha \beta^{-a/b} u^{a/b} - u \right) du ,
+\end{align}
+so it suffices to learn something about the function
+\begin{equation}
+    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t ) dt ,
+\end{equation}
+for $a$, $b$, and $x$ nonnegative.  We will only be interested in this as a function of $x$; 
+the coefficients $a$ and $b$ will merely depend on the dimension and what quantity we are calculating.
+We can construct a power series for $G$ in $x$:
+\begin{align}
+    \partial_x^n G(a,b,x) &= \int_0^\infty (-1)^n t^{a+nb} \exp\left( -x t^b - t ) dt \\
+            &= (-1)^n G(a+nb,b,x) , \quad \mbox{so} \\
+    \partial_x^n G(a,b,x) \vert_{x=0} &= \Gamma(a+nb+1) .
+\end{align}
+So, if everything works out, we will have that
+\begin{align}
+    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
+\end{align}
+
+A few facts, first.
+Using repeatedly the identity $\Gamma(x+1) = x\Gamma(x)$
+and the Pochammer symbol $(x)_n = x(x+1)\cdots(x+n-1)$,
+note that for $m \in \N$,
+\[
+    \Gamma(m+y) = (y)_m \Gamma(y).
+\]
+Also note that for $c$ and $m$ integers,
+\[
+    (y)_{cm} = c^{cm} \prod_{k=0}^{c-1} (\frac{y+k}{c})_{m} ,
+\]
+and in particular, that
+\[
+    (cm+\ell)! = \ell! (\ell+1)_{cm} = \ell! c^{cm} \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m .
+\]
+
+Now suppose that $b$ is rational (as will be the case), letting $b = b'/c$ with $b'$ and $c$ positive integers.
+Therefore, setting $n = mc+\ell$,
+\begin{align}
+    G(a,b,x) &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} \Gamma(a+\ell b'/c+mb'+1) \\
+            &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} (a+\ell b')_{mb'} \Gamma(a + \ell b') \\
+            &= \sum_{\ell=0}^{c-1} \frac{ (-x)^{\ell} }{ \ell! } \sum_{m \ge 0} ( (-cx)^c / (b')^{b'} )^m 
+                    \frac{ \prod_{j=0}^{b'-1} (\frac{ a+b'\ell+j }{ c b' } )_m }{ \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m } ,
+\end{align}
+which is the sum of $c$ (generalized) hypergeometric functions.
+
 \section{Poisson--Vorono\u\i\ facts}

### 2010-09-04 Spatial_adaptation/standing-variation.tex
/* Standing variation */ fixed errors.

@@
     z_0 &= \P\{ \mbox{origin reached by mutations from $\lambda_0$ before $\lambda$} \} \\
-        &= \int_0^\infty \lambda_0 d \omega_d f'(t) f(t)^{d-1} \exp \left\{ - \lamba_0 \omega_d f(t)^d - \lambda h(t) \right\} dt .
+        &= \int_0^\infty \lambda_0 d \omega_d f'(t) f(t)^{d-1} \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} dt .
 \end{align}
@@
 \begin{align}
-    \z_* = \frac{ \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt }
+    z_* = \frac{ \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt }
                 { \int_0^\infty \lambda \exp\{ - \lambda h(t) \} dt } .
@@
   \int_0^\infty t^c \exp \left( - \alpha t^a - \beta t^b \right) dt 
-            &= \left( b^{-1} \beta^{ (1-c)/b } \int_0^\infty u^{(c+1-b)/b} \exp\left( - \alpha \beta^{-a/b} u^{a/b} - u \right) du ,
+            &= \left( b^{-1} \beta^{ (1-c)/b } \right) \int_0^\infty u^{(c+1-b)/b} \exp\left( - \alpha \beta^{-a/b} u^{a/b} - u \right) du ,
 \end{align}
@@
 \begin{equation}
-    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t ) dt ,
+    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
 \end{equation}
@@
 \begin{align}
-    \partial_x^n G(a,b,x) &= \int_0^\infty (-1)^n t^{a+nb} \exp\left( -x t^b - t ) dt \\
+    \partial_x^n G(a,b,x) &= \int_0^\infty (-1)^n t^{a+nb} \exp\left( -x t^b - t \right) dt \\
             &= (-1)^n G(a+nb,b,x) , \quad \mbox{so} \\

### 2010-09-04 Spatial_adaptation/standing-variation.tex
/* Standing variation */  adjustments and constant-speed beginning

@@
 Now let $f(t) = vt$, so $h(t) = \frac{ \omega_d v^d }{ d+1 } t^{d+1}$.
-The integrals that appear above are of the following form, for $\alpha$, $\beta$, $a$, $b$, and $c$ all nonnegative:
+The integrals that appear above are of the following form:
 \begin{align}
@@
 \begin{equation}
-    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
+    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt .
 \end{equation}
-for $a$, $b$, and $x$ nonnegative.  We will only be interested in this as a function of $x$; 
+We will only be interested in this as a function of $x$; 
 the coefficients $a$ and $b$ will merely depend on the dimension and what quantity we are calculating.
@@
             &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} (a+\ell b')_{mb'} \Gamma(a + \ell b') \\
-            &= \sum_{\ell=0}^{c-1} \frac{ (-x)^{\ell} }{ \ell! } \sum_{m \ge 0} ( (-cx)^c / (b')^{b'} )^m 
-                    \frac{ \prod_{j=0}^{b'-1} (\frac{ a+b'\ell+j }{ c b' } )_m }{ \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m } ,
+            &= \sum_{\ell=0}^{c-1} \frac{ (-x)^{\ell} }{ \ell! } \Gamma(a + \ell b') \sum_{m \ge 0} ( (-cx)^c / (b')^{b'} )^m 
+                    \frac{ \prod_{j=0}^{b'-1} (\frac{ a+b'\ell+j }{ b' } )_m }{ \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m } ,
 \end{align}
-which is the sum of $c$ (generalized) hypergeometric functions.
+which is the sum of $c$ (generalized) hypergeometric series.
+For these to converge, we need $b' \le c$, e.g.\ $b\le 1$.
+
+In particular, if we let $A=\lambda_0 \omega_d v^d $ and $B=\lambda h(1)$, then 
+\begin{align}
+    \E[\tau] &= \int_0^\infty e^{-A t^d - B t^{d+1} dt \\
+            &= \frac{B^{1/(d+1)}}{d+1} \int_0^\infty u^{-d/(d+1))} \exp\left( -AB^{-d/(d+1))} u^{d/(d+1)} -u \right) du \\
+            &= \frac{B^{1/(d+1)}}{d+1} G \left(-d/(d+1), d/(d+1), A B^{-d/(d+1)} \right) .
+\end{align}
 

### 2011-02-25 Spatial_adaptation/standing-variation.tex
/* Standing variation */ fixed some latex.
/* Standing variation */
/* Standing variation */  wikipedia ref added =)
minor fixes

@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\N}{\mathbb{N}}
 \newcommand{\one}{\mathbf{1}}
@@
 If $p_s$ is the probability of survival in the new selective regime,
-this leads to the mean measure for successful influorescences
+since the probability that an influorescence of size $z$ survives in the new selective regime is $1 - (1-p_s)^z$,
+the mean measure for successful influorescences is
 \[
-   dx \int_0^\infty F(p_s,t) dt ,
+   \mu dx \int_0^\infty (1-F(1-p_s,t)) dt ,
 \]
-where $F$ is the generating function for $Z_t$,
+where $F$ is the generating function for $Z_t$ (in the old selective regime),
 \[
-  F(a,t) = \E[a^{Z_t}] ,
+  F(a,t) = \E[a^{Z_t} \; | \; Z_0 = 1] .
 \]
-and let $\phi$ be the generating function for the offspring distribution,
-\[
-  \phi(a) = \sum_k p_k a^k .
-\]
-Now, $p_s$ is the minimal solution to $\phi(a) = a$,
-and more generally, $F$ solves
+Note that 
+if $\phi$ be the generating function for the offspring distribution in the old selective regime,
+then $F$ solves
 \[
@@
 \]
+If $\phi_+$ be the generating function for the offspring distribution in the new selective regime,
+\[
+  \phi_+(a) = \sum_k p_k a^k .
+\]
+Now, $p_s$ is the minimal solution to $\phi_+(a) = a$.
 
@@
 \end{align}
+Something simliar is noted at the Wikipedia page on the Gaussian integral, and claims such integrals show up in quantum field theory.
 
@@
 \begin{align}
-    \E[\tau] &= \int_0^\infty e^{-A t^d - B t^{d+1} dt \\
+    \E[\tau] &= \int_0^\infty e^{-A t^d - B t^{d+1}} dt \\
             &= \frac{B^{1/(d+1)}}{d+1} \int_0^\infty u^{-d/(d+1))} \exp\left( -AB^{-d/(d+1))} u^{d/(d+1)} -u \right) du \\
             &= \frac{B^{1/(d+1)}}{d+1} G \left(-d/(d+1), d/(d+1), A B^{-d/(d+1)} \right) .
+\end{align}
+So, if $d=1$, then we have {\bf (verify this by usual methods!)}, with $x = AB^{-1/2}$,
+\begin{align}
+    \E[\tau] &= \frac{\lambda^{1/(2)}}{2} G \left( -1/2, 1/2, x \right) \\
+            &= \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) \sum_{m\ge0} (4x^2)^m \frac{ (-1/2)_m }{ (1/2)_m (1)_m } 
+                        - x \Gamma(1/2) \sum_{m\ge0} (4x^2)^m \frac{ (1/2)_m }{ (1)_m (3/2)_m }  \right\} \\
+            &=  \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) {}_1F_1(-1/2;1/2;4x^2) - x \Gamma(1/2) {}_1F_1(1/2;3/2;4x^2) \right\} \\
+            &= \frac{\lambda^{1/(2)}}{2} \left\{ - \Gamma(-1/2) (1/x) \gamma(-1/2,4x^2) - \Gamma(1/2) (1/x) (\sqrt{\pi}/8) \erf(4x^2) \right\} .
+\end{align}
+
+And in $d=2$, since
+\begin{align}
+    G(-2/3,2/3,x) &= \Gamma(-2/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (-1/3)_m (1/6)_m }{ (1/3)_m (2/3)_m (1)_m }
+                    - x \Gamma(4/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (2/3)_m (7/6)_m }{ (2/3)_m (1)_m (4/3)_m }
+                        + \frac{1}{2} x^2 \Gamma(10/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (5/3)_m (13/6)_m }{ (2)_m (4/3)_m (5/3)_m } \\
+                &= \Gamma(-2/3) {}_2F_2(-1/3,1/6;1/3,2/3;-\frac{27x^2}{4})
+                    -x \Gamma(4/3) {}_1F_1(7/6;4/3;-\frac{27x^2}{4})
+                       -\frac{1}{2}x^2 \Gamma(10/3) {}_1F_1(13/6;4/3;-\frac{27x^2}{4}) .
 \end{align}

### 2011-02-25 standing-patches-paper.tex
WorkingWiki replaced project description

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}

\begin{document}

\section{Introduction}

Recall questions of origin of adaptations

Summarize previous paper

Question assumptions of no standing variation and homogenous selection

Stick to constant-speed waves.

\section{Standing variation}

\subsection{Model}

selection pressure changes at $t=0$ from $1-s_d$ to $1+s_b$ -- homogeneous

rescale to two parameters 
-- a characteristic length of regions and a proportion of haplotypes (or of space?) that are from standing variation

Branching-process-sized point mass approx to local groups present at $t=0$ 
-- depends on local population density large enough relative to migration (to be "branching")
-- and migration not too large (since we approximate as a point mass).
Find mean density of clusters of successful standing mutants.
Is greater-than-one size of clusters likely to make a difference?

recall PPP computations; draw cones of opportunity for new mutations

\subsection{Results}

proportion of haplotypes (or of space?) that are from standing variation

mean (and variance?) of distance between haplotypes

influence of migration on things -- plot these versus migration

\section{Patchy selection}

\subsection{Model}

patches where selection is $1+s_b$ separated by $1-s_d$ -- separation large relative to migration.  
(or relative to $s_d/m$?)

Describe simple discrete model.

Approximation in continuous model by branching Brownian motion with killing to find migration rate between patches --
good approximation neglecting interaction (if local population size is large).
Look at Barton.

\subsection{Results}

Pull out results from other paper on discrete model.

when are patches effectively isolated?

\section{Applications}

Look at some human parameters?

\section{Discussion}

if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography.
recall others' work.

patchy selection helps this happen;
effective migration rate looks like what;
strength of (positive?) selection suprisingly does not affect results (?)

standing variation will be important under these circumstances;
migration rate (and so geographic structure) will not affect numbers of types if this is true;
however migration always affcts spatial patterns.

discussion of applications.

\end{document}

### 2011-03-07 standing-patches-paper.tex
/* Hitchhiking */

@@
 Recall questions of origin of adaptations
+
+Reference on parallel adapation, standing variation, etc.:
+ coop,
+ pennings and hermisson,
 
@@
 selection pressure changes at $t=0$ from $1-s_d$ to $1+s_b$ -- homogeneous
+
+For now we ignore geographic hetergeneity, and assume that the species range
+is a homogenous, two-dimensional region.  (results for a one-dimensional region also appear)
+There are two selective types -- the neutral type, and the mutated type.
+We assume that different mutations are distinguishable --
+either as selectively equivalent mutations, or by linked variation.
+At first, the mutated type is at a seletive disadvantage, for long enough to be at selection-mutation equilibrium,
+but at a certain time the selective regime changes, so that the mutated type has a selective advantage and quickly spreads to fixation.
+After fixation, alleles are either descended
+from families of mutants present as standing variation when the selective regime changed,
+or from new mutants arising since that time.
+
+We will suppose that before time $t=0$,
+the mutant type has fitness $1-s_d$ relative to the neutral type,
+and that after time $t=0$,
+the mutant type has fitness $1+s_b$,
+for positive numbers $s_d$ and $s_b$.
+We assume that diploid fitness is additive, or at lesat that the important dynamics are determined by the haploid fitness.
+If the mutation is recessive, XXX more complicated things happen.
+More concretely, suppose that the species occupies an area with mean density
 

### 2011-04-01 Spatial_adaptation/standing-variation.tex
/* Standing variation */

@@
 Now, $p_s$ is the minimal solution to $\phi_+(a) = a$.
+
+\section{Notes on branching process approximation}
+
+The branching process approximation will be good as long as the probability of "interactions" are small,
+i.e.\ we can couple the branching process with the actual density--regulated model.
+Let $Z_t(x)$ be the numbers of selected individuals in deme $x$.
+If we assume that $Z_t$ is smooth enough, 
+the numbers of nearby individuals to interact with is approximately $Z_t(x) \sigma^d$,
+and the total number of individuals is about $\rho \sigma^d$,
+so the rate of interactions is about $(Z_t(x) \sigma^d)^2/(\rho \sigma^d)$.
+Now if $Z$ is a spatial branching process beginning from a single individual at $x=0$ at $t=0$, 
+and growing at rate $r$, then 
+\[ 
+    Z_t(x) \approx \exp(rt - \|x\|^2/(2\sigma^2 t))/ (2 \pi \sigma^2)^{d/2} 
+        \le e^{rt} / ( \sigma^2 t)^{d/2}
+\]
+and so we need $t$ such that
+\[
+    e^{2rt} t^{-d} << \rho \sigma^{3d} .
+\]
+or equivalently,
+\[
+    e^{2rt/d} t^{-1} << \rho^{1/d} \sigma^3 .
+\]
+This will be true if $t$ is of smaller order than the logarithm of the effective population density, $\log \rho \sigma^d$.
 

### 2011-04-01 standing-patches-paper.tex
/* Standing variation */

@@
 either as selectively equivalent mutations, or by linked variation.
-At first, the mutated type is at a seletive disadvantage, for long enough to be at selection-mutation equilibrium,
+At first, the mutated type is at a selective disadvantage, for long enough to be at selection-mutation equilibrium,
 but at a certain time the selective regime changes, so that the mutated type has a selective advantage and quickly spreads to fixation.
@@
 for positive numbers $s_d$ and $s_b$.
-We assume that diploid fitness is additive, or at lesat that the important dynamics are determined by the haploid fitness.
-If the mutation is recessive, XXX more complicated things happen.
-More concretely, suppose that the species occupies an area with mean density
+We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
+(If the mutation is recessive, XXX more complicated things happen.)
+
+For the other parameters,
+suppose that each offspring of a neutral parent is of the mutant type with probability $\mu$,
+and that the XXX dispersal variance is $\sigma^2$.
+The species occupies an area with mean density
+$\rho$ chromosomes per unit area.
+(XXX or do diploid? below is haploid.)
+
+We make use of the commonly used approximation
+that neglects spatial structure and competition between relatives,
+treating the offspring of a new mutant that appears in an area not already occupied by the mutated type
+as a (nonspatial) branching process.
+After $t=0$, the offspring of a single mutant form (approximately) a branching process with growth rate $s_b$,
+so each new mutant establishes locally with probability $p_s \approx 2s_b$.
+Therefore, the total rate at which new mutants appear and establish locally after $t=0$ in areas not already occupied by the mutant type
+is approximately $\lambda = 2 \mu \rho s_b$.
+Before $t=0$, the mutation is deleterious, so the genetic descendents of each new mutation are doomed to extinction,
+but may persist for some time.
+Consider a mutation that arose at time $-T<0$, and let $Z_s$ be the number of its descendants at time $s-T$.
+At time $t=0$, there are $Z_T$ individuals present with the mutation,
+and each has probability $p_s \approx 2s_b$ of establishing, approximately independently.
+Therefore, the probability that some descendants of this mutation establish and fix locally is
+$1-(1-p_s)^Z_T$, XXX continue here
+
+If $Z_t$ is a branching process with $Z_0 = 1$ and $\E[Z_t] = \exp(-s_d t)$, then
+
+For $t \ge 0$, new mutations appear at rate approximately $\lambda = 4 \mu \rho s_b$.
 

### 2011-04-04 Spatial_adaptation/standing-variation.tex
/* Standing variation */ random delay notes added

@@
 This will be true if $t$ is of smaller order than the logarithm of the effective population density, $\log \rho \sigma^d$.
+
+\subsection{On the random delay}
+
+Let $Y_t$ be the total number of offspring coming from a single individual
+(in the spatial, density-regulated model).
+Then we know that $(Y_t/t)^{1/d} \to v$, 
+but we saw from simulations that it looked like $Y_t^{1/d} \approx v(t-\tau)$,
+where $\tau$ is the (random) establishment time.
+(It would be useful to look at the coefficient of variation of $\tau$.)
+Well, we know that at first, $Y_t \approx W e^{rt}$, 
+where $W= \lim e^{-rt} Z_t$, and $Z_t$ is the associated branching process,
+and $r$ is the growth rate (er, $r=s$?).
+And eventually, $\partial_t Y_t^{1/d} \approx v$.
+If this switchover happened at a fixed size $C$,
+the size at which wavelike behavior took over,
+i.e.\ when $W e^{rt} = C$,
+then for $\tau = \frac{1}{r}\log(C/W)$ we'd have 
+\[
+    Y_t = \begin{cases}
+        W e^{rt} \quad & t \le \tau \\
+        (v (t-\tau))^d + C \quad & t > \tau ,
+\]
+which if $C$ is small relative to the scale we look on will look as described above.
+So the random delay is $(1/r)\log(C/W)$; and so the mean delay is something like $(1/r)\log C$
+and the randomness is determined by $(1/r)\log W$.
 

### 2011-04-04 Spatial_adaptation/standing-variation.tex
/* Standing variation */ fixed latex bug
/* Standing variation */ characteristic length etc notes

@@
         (v (t-\tau))^d + C \quad & t > \tau ,
+\end{cases}
 \]
@@
 So the random delay is $(1/r)\log(C/W)$; and so the mean delay is something like $(1/r)\log C$
-and the randomness is determined by $(1/r)\log W$.
+and the randomness is determined by $(1/r)\log (1/W)$.
 
@@
 
+\section{Characteristic lengths}
+
+How to pick appropriate rule-of-thumb parameters?
+Recall that before we had $\chi$ being the smallest solution to
+\[
+    f\left( \frac{1}{\lambda \chi^d \omega_d} \right) = \chi.
+\]
+Now, suppose we draw a cylinder in space-time with radius $\chi$ and height $t=f^{-1}(\chi)$;
+as before we choose $\chi$ so that the mean number of points in this cylinder is unity.
+Then
+\[
+    1 = \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d ,
+\]
+and so we choose $\chi$ to be the smallest solution to
+\[
+    f\left( \frac{1 - \lambda_0 \omega_d}{\lambda \chi^d \omega_d} \right) = \chi.
+\]
+
+As for the proportion of mutations coming from standing variation,
+it seems a natural measure would be the mean number of such mutations in such a cylinder,
+which is using the definition of $\chi$,
+\begin{align*}
+    \pi_0 &= \frac{\lambda_0 \omega_d \chi^d }{ \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d} \\
+        & = \frac{ 1 }{ 1 + \lambda f^{-1}(\chi) / \lambda_0 } \\
+        &= \lambda_0 \omega_d \chi^d .
+\end{align*}
+
+Now, if $f(t) = vt$, then $f^{-1}(r) = r/v$,
+and so we have
+\[
+    \chi = \frac{ v }{ \lambda \chi^d \omega_d } = \left( \frac{ v(1-\lambda_0 \omega_d) }{ \lambda \omega_d } \right)^{1/(d+1)} ,
+\]
+and 
+\[
+    \pi_0 = \frac{\lambda_0}{\lambda} v^{d/(d+1)} (1-\lambda_0 \omega_d)^{d/(d+1)} (\lambda \omega_d)^{1/(d+1)} .
+\]
+
 \section{Poisson--Vorono\u\i\ facts}

### 2011-04-05 Spatial_adaptation/standing-variation.tex
/* Standing variation */ math oops
/* Standing variation */ added about remaining variance in branching process
/* Standing variation */ entered relations between various quantities (ie how to find proportion of new mutations, etc)

@@
 and the randomness is determined by $(1/r)\log (1/W)$.
+
+What about how much variance is left in the branching process?
+Well, since $Z_{t+s} = \sum_{k=1}^{Z_t} Z^k_{s}$, where $Z^k_s$ are iid copies of $Z_s$,
+we have that $\var[Z_{t+s}|Z_t] = Z_t \var[Z_s]$,
+and so the coefficient of variation of $Z_{t+s}$ given $Z_t$ is
+\[
+   \mathop{CV}[Z_{t+s}|Z_t] &= \frac{\sqrt{Z_t \var[Z_s]}}{ Z_t \E[Z_s] } \\
+        &= e^{-rs} \sqrt{ \var[Z_s] } Z^{-1/2} \\
+        &\simeq e^{-rt/2} \sqrt{ \var[W] } .
+\]
 
@@
 \[
-    \chi = \frac{ v }{ \lambda \chi^d \omega_d } = \left( \frac{ v(1-\lambda_0 \omega_d) }{ \lambda \omega_d } \right)^{1/(d+1)} ,
+    \chi = \frac{ v (1-\lambda_0 \omega_d) }{ \lambda \chi^d \omega_d } = \left( \frac{ v(1-\lambda_0 \omega_d) }{ \lambda \omega_d } \right)^{1/(d+1)} ,
 \]
@@
 
+\section{General relationship between various quantities}
+
+We know that there are $\lamba_0$ successful standing variants per unit area;
+suppose that there are $\nu_+$ successful new mutants per unit area.
+Therefore, $\lambda_0+\nu_+$ is the mean number of successful mutants per unit area (duh),
+and $1/(\lambda_0+\nu_+)$ is the mean area associated with a randomly chosen patch.
+Let $p_0$ be the proportion of space occupied by standing variants, 
+let $a_0$ be the mean area associated with a randomly chosen standing variant,
+and let $a_+$ be the mean area associated with a randomly chosen new variant,
+so that
+\[
+    p_0 = \frac{ \lambda_0 a_0 }{ \lambda_0 a_0 + \nu_+ a_+ } = \lambda_0 a_0 ,
+\]
+since the denominator is equal to 1.
+
+We know how to calculate $p_0$, above, from thinking about the competing exponentials,
+which allows us to find $a_0$.
+
+Now, as in the first paper, the mean density of new patches is easy to calculate -- 
+$\lambda \nu_+ = \E[\tau]$, where $\tau$ is the mean adaptation time of a point;
+this is because 
+\begin{align*}
+    \P\{ \tau > t \} &= \P\{ \mbox{no points in $t$-cone} \} \\
+            &= (1/\lambda) \P\{ \mbox{new successful adaptation arises in non-occupied $(t+dt,x+dx)$} \} ;
+\end{align*}
+integrating over $t$ produces the result.
+
+Finally, the proportion of patches that come from new mutation is $\nu_+ / (\nu_+ + \lambda_0)$,
+and we know that, conditioning on the type of a randomly chosen patch,
+\[
+    \frac{1}{\lambda_0+\nu_+} = \frac{\lambda_0}{\lambda_0+\nu_+} a_0 + \frac{\nu_+}{\lambda_0+\nu_+} a_+ .
+\]
+So, given $\lambda_0$, $\nu_+$, and $a_0$, we can find $a_+$.
+
 \section{Poisson--Vorono\u\i\ facts}

### 2011-04-06 Spatial_adaptation/standing-variation.tex
/* Standing variation */ fixed char length error
/* Standing variation */ removed erroneous statement

@@
 \end{align}
-As before, the mean density $\nu$ is equal to $1/\E[\tau]$.
 
@@
 \]
-and so we choose $\chi$ to be the smallest solution to
+and so we choose $\chi$ to be the unique positive solution to 
 \[
-    f\left( \frac{1 - \lambda_0 \omega_d}{\lambda \chi^d \omega_d} \right) = \chi.
+    f\left( \frac{1 - \lambda_0 \chi^d \omega_d}{\lambda \chi^d \omega_d} \right) = \chi.
 \]
@@
 Now, if $f(t) = vt$, then $f^{-1}(r) = r/v$,
-and so we have
+and so we have $\chi$ the positive root of
 \[
-    \chi = \frac{ v (1-\lambda_0 \omega_d) }{ \lambda \chi^d \omega_d } = \left( \frac{ v(1-\lambda_0 \omega_d) }{ \lambda \omega_d } \right)^{1/(d+1)} ,
+  \omega_d \chi^d (\lambda_0 + \lambda v \chi) = 1 .
 \]
-and 
+If $d=1$ then this is
 \[
-    \pi_0 = \frac{\lambda_0}{\lambda} v^{d/(d+1)} (1-\lambda_0 \omega_d)^{d/(d+1)} (\lambda \omega_d)^{1/(d+1)} .
+    \chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v } .
 \]
+In $d=2$ ??
 

### 2011-04-06 Spatial_adaptation/standing-variation.tex
/* Standing variation */ added info about moments of X
/* Standing variation */ latex errors

@@
 
+Let $X$ be the distance to the eventually enclosing mutation and $R=f(\tau)$ (see first paper).
+Then if $g(t) = \lambda_0 \omega_d f(t)^d + \lambda h(t)$ then $(X,R)$ has density
+\[
+    d \omega_d x^{d-1} \exp(-g(f^{-1}(r)) dx dr,
+\]
+so integrating over $r$,
+\[
+    \P\{ X \in dx \} = dx d \omega_d x^{d-1} \int_x^\infty \exp( - g(f^{-1}(r)) ) dr ,
+\]
+and
+\[
+    \E[ X^n ] = \frac{d\omega_d}{n+d} \int_0^\infty r^{n+d} \exp( - g(f^{-1}(r)) ) dr .
+\]
+
 If we rescale time by $\alpha$ and space by $\beta$, 
@@
 
-We know that there are $\lamba_0$ successful standing variants per unit area;
+We know that there are $\lambda_0$ successful standing variants per unit area;
 suppose that there are $\nu_+$ successful new mutants per unit area.

### 2011-04-07 standing-patches-paper.tex
added much of standing variation methods

--- 2011-02-25 standing-patches-paper.tex
@@
 \usepackage{amsmath,amssymb}
+\usepackage{natbib}
+\usepackage{marginnote}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\usepackage{latexml} % for \iflatexml
+
+\newcommand{\var}{\mathop{\mbox{Var}}}
+\renewcommand{\P}{\mathbb{P}}
+\newcommand{\E}{\mathbb{E}}
+\newcommand{\R}{\mathbb{R}}
 
@@
 
+Reference on parallel adapation, standing variation, etc.:
+ coop,
+ pennings and hermisson,
+
 Summarize previous paper
@@
 
+\section{General model}
+
+outline here assumptions common to both models.
+
 \section{Standing variation}
 
-\subsection{Model}
+\subsection{The set-up}
 
-selection pressure changes at $t=0$ from $1-s_d$ to $1+s_b$ -- homogeneous
+% selection pressure changes at $t=0$ from $1-s_d$ to $1+s_b$ -- homogeneous
 
-rescale to two parameters 
--- a characteristic length of regions and a proportion of haplotypes (or of space?) that are from standing variation
+For now we ignore geographic hetergeneity, and assume that the species range
+is a homogenous, one- or two-dimensional region. 
+There are two selective classes -- the {\em neutral} type, and the {\em mutated} type.
+We assume that different mutations are distinguishable --
+either as selectively equivalent mutations, or by linked variation.
+For a long time in the past, the mutated type has been at a selective disadvantage (long enough to be at selection-mutation equilibrium),
+but at a certain time the selective regime changes, so that the mutated type has a selective advantage and quickly spreads to fixation.
+After fixation, alleles are either descended
+from families of mutants present as standing variation when the selective regime changed,
+or from new mutants arising since that time.
 
-Branching-process-sized point mass approx to local groups present at $t=0$ 
--- depends on local population density large enough relative to migration (to be "branching")
--- and migration not too large (since we approximate as a point mass).
-Find mean density of clusters of successful standing mutants.
-Is greater-than-one size of clusters likely to make a difference?
+For concreteness, suppose that before time $t=0$,
+the mutant type has fitness $1-s_d$ relative to the neutral type,
+and that after time $t=0$,
+the mutant type has fitness $1+s_b$,
+where $s_d$ and $s_b$ are positive numbers.
+We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
+(If the mutation is recessive, XXX more complicated things happen.)
+As for the other parameters,
+suppose that each offspring of a neutral parent is of the mutant type with probability $\mu$,
+and that the mean squared distance between parent and child is $\sigma^2$ (the {\em dispersal variance}).
+The species occupies an area with mean density
+$\rho$ haploid individuals (or chromosomes) per unit area.
+\marginnote{XXX or do diploid? below is haploid.}
+
+% Branching-process-sized point mass approx to local groups present at $t=0$
+% -- depends on local population density large enough relative to migration (to be "branching")
+% -- and migration not too large (since we approximate as a point mass).
+% Find mean density of clusters of successful standing mutants. 
+% Is greater-than-one size of clusters likely to make a difference?
+
+We make use of the commonly used approximation that neglects competition between relatives,
+treating the offspring of a new mutant that appears in an area not already occupied by the mutated type
+as a branching process.
+After $t=0$, the offspring of a single mutant form (approximately) a branching process with growth rate $s_b$,
+so each new mutant establishes locally with probability $p_s \approx 2s_b$.
+As in \cite{ralphcoop2010}, each new offspring has a very small probability of being a mutant and establishing locally,
+so the collections of times and locations at which mutants appear and establish locally 
+is well--approximated by a Poisson process in space and time.
+This the rate of this Poisson process, 
+the rate per unit area at which new mutants appear and establish locally after $t=0$ in areas not already occupied by the mutant type
+is approximately $\lambda = 2 \mu \rho s_b$.
+
+Before $t=0$, on the other hand, the mutation is deleterious, so the genetic descendents of each new mutation are (with high probability) doomed to extinction,
+but may persist for some time.
+Since the times and locations of mutants before $t=0$ also well-approximated by a Poisson process with rate $\mu \rho$,
+the locations of all mutant families extant at $t=0$ whose descendents are destined to fix locally is also,
+by the Mapping theorem, a Poisson process with rate we define to be $\lambda_0$.
+If we assume that the descendants of at most only a few members of any extant mutant family at $t=0$ will survive,
+and that these progenitors are near to each other in space, we can then treat each such family as a single mutant,
+equivalent to the mutants arising later.
+(The approximation will be good if the log of the size of an extant families is small relative to the establishment time,
+and the spatial distribution of the family is small relative to the spread between them.)
+To find $\lambda_0$, consider a mutation that arose at time $-T<0$, and let $Z_s$ be the number of its descendants at time $s-T$.
+At time $t=0$, there are $Z_T$ individuals present with the mutation,
+and each has probability $p_s \approx 2s_b$ of establishing, approximately independently.
+Therefore, the probability that some descendants of this mutation establish and fix locally is $1-(1-p_s)^{Z_T}$,
+so defining $\zeta(u,t) = \E[u^{Z_t} | Z_0=1 ]$ to be the generating function of $Z_T$,
+the rate is
+\begin{align*}
+    \mu \rho \int_0^\infty \left( 1- \zeta(1-p_s,T) \right) dT .
+\end{align*}
+Now for $s_b$ small, and using $p_s \approx 2s_b$,
+we know that $\zeta(1-2s_b,T) \approx 1-2s_b \E[Z_T] = 1-2s_b e^{-Ts_d}$,
+resulting in the approximation
+\begin{align}
+    \lambda_0 &= \mu \rho \int_0^\infty 2s_b e^{-s_d T} dT \\
+        &= \frac{ 2 \mu \rho s_b }{ s_d } .
+\end{align}
+
+XXX More discussion on whether this is a good approximation?  
+Compute typical size and radius of a standing cluster?
 
@@
 
+\subsection{Characteristic parameters}
+
+% a characteristic length of regions and a proportion of haplotypes (or of space?) that are from standing variation
+
+In \citet{ralphcoop2010}, studying the model without standing variation,
+we defined a {\em characteristic length} which gave the spatial scale across with mutants with distinct origins would establish.
+This was proportional to the mean distance between neighboring established mutants,
+but had the advantage of being easier to calculate.
+Furthermore, the time scale over which adaptation occurred could be found by dividing the characteristic length 
+by the speed at which the mutants spread.
+We will define the characteristic length for this model,
+as well as a similar compound parameter describing the relative importance of standing variation to the process of adaptation.
+
+Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
+As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s}$ is the speed of the wave of advance.
+The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
+$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
+(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
+% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
+Therefore, if we define $\chi$ to be the unique positive solution to
+$\lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1$,
+then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
+before it is expected that one other successful mutation would have arisen in the area covered so far.
+If we define $u = 2 \pi^3 \lambda_0^3 - 27 \pi^2 v^2 \lambda^2$,
+The solution to this is {\color{red} XXX what?}
+(In one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
+
+By the above calculation, we know that the relevant mutations occur about distance $\chi$ apart, 
+and occur within the first $\chi/v$ generations.
+Said another way, if we look in a circular region of space of radius $\chi$ over $\chi/d$ generations,
+we expect to find one mutational origin;
+we denote the probability that this mutation is from standing variation by $\pi_0$.
+Therefore, $\pi_0$ is roughly the proportion of establishing mutations that come from standing variation,
+and it is an elementary consequence of the Poisson process that
+\begin{align} \label{eqn:pizero}
+    \pi_0 = \lambda_0 \pi \chi^2 .
+\end{align}
+\marginnote{put in value of $\chi$ computed above.}
+(In one dimension, $\pi_0 = 2 \lambda_0 \chi$.)
+
 \subsection{Results}
 
-proportion of haplotypes (or of space?) that are from standing variation
+The quantity that is perhaps easiest to compute is the mean time until adaptation.
+Fix some geographic location, and let $\tau\ge0$ be the time at which the point is reached by the advantageous mutations.
+Then, as shown in Figure XX,
+$\tau > t$ if and only if the cone with point at $(x,t)$ and slope $v$ extending back to $t=0$ is empty of successful mutations.
+Since we assume these are a Poisson process, 
+\[
+    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 ) ,
+\]
+and so we have
+\begin{align}
+    \E[\tau] % &= \int_0^\infty \P\{ \tau > t \} dt \\
+        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 ) dt .
+\end{align}
 
-mean (and variance?) of distance between haplotypes
+% proportion of haplotypes (and of space) that are from standing variation
+
+We have defined $\lambda_0$ to be the mean density of standing variants that reach local fixation.
+Define $\nu_+$ to be the mean density of new mutants whose offspring fix locally.
+Since the probability that a mutant arising at $(x,t)$ is lucky enough to be born in a location not already occupied by mutants
+is $\P\{ \tau > t \}$,
+we can see  $\nu_+ = \int_0^\infty \lambda \P\{\tau>t\} dt$, and hence
+\begin{equation}
+    \nu_+ = (\mbox{mean density of new patches}) = \lambda \E[\tau] .
+\end{equation}
+Since the mean proportion of patches that come from standing variation is $\lambda_0 / (\lambda_0 + \nu_+)$,
+this gives us the exact quantity corresponding to $\pi_0$ from \eqref{eqn:pizero}.
+There are $\lambda_0 + \nu_+$ patches per unit area, so
+the typical patch (i.e.\ with distribution given by the Palm measure) occupies area $1/(\lambda_0 + \nu_+)$.
+
+We can also find the mean proportion of space covered by standing variants.
+If by time $t$ a geographic location has not yet been reached by the mutation,
+the probability it will be reached by time $t+dt$ 
+by a standing variant is approximately $2 \lambda_0 \pi vt dt$, 
+and the probability it is reached by a new variant is $\lambda \pi v^2 t^2 dt$.
+Therefore, as is standard for competing exponentials,
+the probability a given location is reached first by a standing variant,
+and therefore the mean proportion of space covered  by standing variants,
+is
+\begin{equation}
+    z_0 = \int_0^\infty {2 \lambda_0 \pi v t} \exp \left( - 2 \lambda_0 \pi v t - \lambda \pi v^2 t^2 \right) dt .
+\end{equation}
+
+Furthermore, if we define $a_0$ to be the mean area occupied by a typical standing variant, 
+then $a_0$ is given by the proportion of the range occupied by standing variants over the mean density of unique standing variants,
+i.e.\ $a_0 = z_0 / \lambda_0$.
+We can compare this to the corresponding quantity $a_+$ for a new variant, since the formula
+$a_0 / a_+ = z_0 / (1-z_0)$.
+
+\marginnote{do mean (and variance?) of distance between haplotypes?}
 

### 2011-04-07 standing-patches-paper.tex
added convergence notes

@@
 If we define $u = 2 \pi^3 \lambda_0^3 - 27 \pi^2 v^2 \lambda^2$,
-The solution to this is {\color{red} XXX what?}
+The solution to this is {\bf XXX what?}
 (In one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
@@
 \[
-    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 ) ,
+    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 \right) ,
 \]
@@
     \E[\tau] % &= \int_0^\infty \P\{ \tau > t \} dt \\
-        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 ) dt .
+        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 \right) dt .
 \end{align}

### 2011-04-07 Spatial_adaptation/standing-variation.tex
added integral appendix

@@
 \end{align}
+To check for convergence, note that by Stirling's formula $\Gamma(z) \simeq \sqrt{2\pi/z} (z/e)^z$,
+\begin{align}
+    \left( \frac{x^n \Gamma(a+nb+1) }{ n! } \right)^{1/n} &\simeq x \frac{ (a+nb+1)^{b+(a+1)/n} }{ (n+1)^{1+1/n} } e^{(1-b)-a/n} \\
+        &\simeq C n^{b-1} b^b \\
+        &\to 0  \quad \mbox{as } n \to \infty ,
+\end{align}
+so that the sum converges for all $x$ if $b<1$ (which is the case for us as $b=d/(d+1)$).
 Something simliar is noted at the Wikipedia page on the Gaussian integral, and claims such integrals show up in quantum field theory.

### 2011-04-07 standing-patches-paper.tex
added integral appendix

@@
     \E[\tau] % &= \int_0^\infty \P\{ \tau > t \} dt \\
-        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 \right) dt .
+        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 \right) dt.
 \end{align}
+For means of evaluting this integral, see Appendix \ref{apx:integrals}.
 
@@
 \end{equation}
+To evaluate this integral, again see Appendix \ref{apx:integrals}.
 
@@
 i.e.\ $a_0 = z_0 / \lambda_0$.
-We can compare this to the corresponding quantity $a_+$ for a new variant, since the formula
+We can compare this to the corresponding quantity $a_+$ for a new variant through the formula
 $a_0 / a_+ = z_0 / (1-z_0)$.
@@
 
+\appendix
+
+\section{Integrals appearing in the text}
+    \label{apx:integrals}
+
+In the text at several points appear integrals of the form
+\begin{align}
+  \int_0^\infty t^c \exp \left( - \alpha t^d - \beta t^{d+1} \right) dt 
+\end{align}
+where $c$ is a positive integer, $d$ is the dimension, and $\alpha$ and $\beta$ are positive real numbers.
+This could be evaluated through standard numerical methods; below we describe a power series expansion.
+Changing variables to $u = \beta t^{d+1}$, this becomes
+\begin{align}
+    \left( {d+1}^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
+\end{align}
+so it suffices to evaluate the function
+\begin{equation}
+    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
+\end{equation}
+in the case that $a=(c-d+1)/(d+1)$, $b=d/(d+1)$, and $x$ is a function of the demographic paramters.
+Since $a$ and $b$ only depend on the dimension and the quantity being computed,
+we are interested in $G$ as a function of $x$.
+At least in the case $b=d/(d+1)$ it is possible to express $G$ as a finite sum of gamma functions,
+but we proceed with a simpler method.
+Note that $\partial_x G(a,b,x) = -G(a+b,b,x)$,
+and that at $x=0$, the function $G$ is the gamma function $G(a,b,0) = \Gamma(a+1)$.
+Therefore, a Taylor series for $G$ would be
+\[
+    G(a,b,x) = \sum_{n \ge 0} = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
+\]
+It is easy to check using Stirling's formula that $\limsup_{n \to \infty} ( x^n \Gamma(a+nb+1)/n! )^{1/n} = 0$
+if $b<1$, so the sum converges.
+
 \end{document}

### 2011-04-07 Spatial_adaptation/standing-variation-code.R
WorkingWiki replaced project description

@@ -1,0 +1,62 @@
+
+# Implement theory from standing variation parts
+
+# mu is mutation rate
+# rho is population density
+# sigma is mean squared dispersal distance
+# sb >0 is advantage of mutation for t>0
+# sd >0 is disadvantage of mutation for t<0
+# 
+# ... and
+# lambda <- 2 * mu * rho * sb    # is rate of new successful mutations
+# lambda0 <- lambda / sd         # is approximate rate of standing successful mutations
+# v <- sigma * sqrt(2*sb)        # speed of wave of advance
+
+meanTime <- function (mu, rho, sb, sd, sigma) {
+    # mean time til adaptation of a point;
+    # E[\tau] is int_0^\infty exp(-lambda0 pi v^2 t^2 - lambda pi v^2 t^3) dt
+    f <- function (t) {
+        exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/sd + t) )
+    }
+    return( integrate(f, 0, Inf) )
+}
+
+meanNewDensity <- function (mu, rho, sb, sd, sigma) {
+    # mean density of patches arising from new mutations
+    #  is lambda * E[\tau]
+    x <- meanTime(mu, rho, sb, sd, sigma)
+    x$value <- x$value * 2 * mu * rho * sb 
+    return( x )
+}
+
+standingProportionArea <- function (mu, rho, sb, sd, sigma) {
+    # proportion of space arising from standing variation
+    # is 2 * lambda0 * pi * v * t * exp( - 2 * lambda0 * pi * v * t - lambda * pi * v^2 * t^2 )
+    f <- function (t) {
+        (1/sd) * 4 * sqrt(2) * mu * rho * sb^(3/2) * pi * t * exp( - 4 * mu * rho * sb^(3/2) * sigma * pi * t ( sqrt(2) / sd + sigma * t ) )
+    }
+    return( integrate(f, 0, Inf) )
+}
+
+standingProportionNumbers <- function (mu, rho, sb, sd, sigma) {
+    # proportion of patches arising from standing variation
+    #    is 1/(1+sd * E[\tau])
+    x <- meanTime(mu, rho, sb, sd, sigma)
+    x$value <- 1 / (1 + sd * x$value)
+    return( x )
+}
+
+mu <- 10^{-8}
+rho <- 10^6
+sb <- .01
+sd <- .01
+sigmavals <- c(0.1,1,10,20,50)
+
[…]

### 2011-04-07 Spatial_adaptation/standing-variation-code.R
split out functions; producing a plot now

@@ -1,79 +1,0 @@
-#
-# Implement theory from standing variation parts
-
-# mu is mutation rate
-# rho is population density
-# sigma is mean squared dispersal distance
-# sb >0 is advantage of mutation for t>0
-# sd >0 is disadvantage of mutation for t<0
-# 
-# ... and
-# lambda <- 2 * mu * rho * sb    # is rate of new successful mutations
-# lambda0 <- lambda / sd         # is approximate rate of standing successful mutations
-# v <- sigma * sqrt(2*sb)        # speed of wave of advance
-
-meanTime <- function (mu, rho, sb, sd, sigma) {
-    # mean time til adaptation of a point;
-    # E[\tau] is int_0^\infty exp(-lambda0 pi v^2 t^2 - lambda pi v^2 t^3) dt
-    f <- function (t) {
-        exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/sd + t) )
-    }
-    return( integrate(f, 0, Inf) )
-}
-
-meanNewDensity <- function (mu, rho, sb, sd, sigma) {
-    # mean density of patches arising from new mutations
-    #  is lambda * E[\tau]
-    x <- meanTime(mu, rho, sb, sd, sigma)
-    x$value <- x$value * 2 * mu * rho * sb 
-    return( x )
-}
-
-standingProportionArea <- function (mu, rho, sb, sd, sigma) {
-    # proportion of space arising from standing variation
-    # is 2 * lambda0 * pi * v * t * exp( - 2 * lambda0 * pi * v * t - lambda * pi * v^2 * t^2 )
-    f <- function (t) {
-        (1/sd) * 4 * sqrt(2) * mu * rho * sb^(3/2) * pi * t * exp( - 4 * mu * rho * sb^(3/2) * sigma * pi * t ( sqrt(2) / sd + sigma * t ) )
-    }
-    return( integrate(f, 0, Inf) )
-}
-
-standingProportionNumbers <- function (mu, rho, sb, sd, sigma) {
-    # proportion of patches arising from standing variation
-    #    is 1/(1+sd * E[\tau])
-    x <- meanTime(mu, rho, sb, sd, sigma)
-    x$value <- 1 / (1 + sd * x$value)
-    return( x )
-}
-
-# playing around
-mu <- 10^(-8)
-rho <- 10
-sb <- .01
-sd <- .01
-sigmavals <- c(0.1,1,10,20,50)
-
[…]

### 2011-04-07 Spatial_adaptation/standing-variation-fns.R
WorkingWiki replaced project description

@@ -1,0 +1,63 @@
+#
+# Implement theory from standing variation parts
+
+# mu is mutation rate
+# rho is population density
+# sigma is mean squared dispersal distance
+# sb >0 is advantage of mutation for t>0
+# sd >0 is disadvantage of mutation for t<0
+# 
+# ... and
+# lambda <- 2 * mu * rho * sb    # is rate of new successful mutations
+# lambda0 <- lambda / sd         # is approximate rate of standing successful mutations
+# v <- sigma * sqrt(2*sb)        # speed of wave of advance
+
+meanTime <- function (mu, rho, sb, sd, sigma) {
+    # mean time til adaptation of a point;
+    # E[\tau] is int_0^\infty exp(-lambda0 pi v^2 t^2 - lambda pi v^2 t^3) dt
+    f <- function (t) {
+        exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/sd + t) )
+    }
+    return( integrate(f, 0, Inf) )
+}
+
+meanNewDensity <- function (mu, rho, sb, sd, sigma) {
+    # mean density of patches arising from new mutations
+    #  is lambda * E[\tau]
+    x <- meanTime(mu, rho, sb, sd, sigma)
+    x$value <- x$value * 2 * mu * rho * sb 
+    return( x )
+}
+
+standingProportionArea <- function (mu, rho, sb, sd, sigma) {
+    # proportion of space arising from standing variation
+    # is 2 * lambda0 * pi * v * t * exp( - 2 * lambda0 * pi * v * t - lambda * pi * v^2 * t^2 )
+    f <- function (t) {
+        (1/sd) * 4 * sqrt(2) * mu * rho * sb^(3/2) * pi * t * exp( - 4 * mu * rho * sb^(3/2) * sigma * pi * t ( sqrt(2) / sd + sigma * t ) )
+    }
+    return( integrate(f, 0, Inf) )
+}
+
+standingProportionNumbers <- function (mu, rho, sb, sd, sigma) {
+    # proportion of patches arising from standing variation
+    #    is 1/(1+sd * E[\tau])
+    x <- meanTime(mu, rho, sb, sd, sigma)
+    x$value <- 1 / (1 + sd * x$value)
+    return( x )
+}
+
+paramString <- function(ps=c("mu","rho","sb","sd","sigma"), pos=-1) {
+    # return a string for use in plots of the parameters.
[…]

### 2011-04-07 Spatial_adaptation/standing-variation-plots.R
WorkingWiki replaced project description

@@ -1,0 +1,34 @@
+
+# playing around
+mu <- 10^(-8)
+rho <- 10
+sb <- .01
+sd <- .01
+sigmavals <- c(0.1,1,10,20,50)
+
+# plot Standing Proportion Numbers as a function of sigma
+sPvals <- sapply(sigmavals, function (sigma) { standingProportionNumbers(mu,rho,sb,sd,sigma)$value } )
+plot(sigmavals, sPvals, type="b", xlab=expression(paste(sigma, " = dispersal SD")), ylab=expression(paste(z[0], " = proportion of patches from standing variation")), ylim=c(0,1))
+
+
+# sickle cell case from ralphcoop2010
+mu <- 10^(-8)
+rhovals <- c(2.5,25)
+rhocols <- c("black","red")
+rholabs <- c(expression(rho==2.5), expression(rho==25))
+sb <- .05
+sd <- .01
+sigmavals <- c(0.1,1,10,20,50,100)
+otherlabs <- expression(paste( mu == 10^-8, "  ", s[b] == .05, "  ", s[d] == .01 ))
+
+# plot Standing Proportion Numbers as a function of sigma
+setEPS(horizontal = FALSE, onefile = FALSE, paper = "special")
+postscript( file="proportion-by-sigma.eps", width=6.5, height=3, title=paramString() )
+plot(0, 0, type="n", xlab=expression(paste(sigma, " = dispersal SD")), ylab=expression(paste(z[0], " = proportion of patches from standing variation")), xlim=range(sigmavals), ylim=c(0,1), main=otherlabs )
+legend("topleft", lty=1, col=rhocols, legend=rholabs)
+for (k in 1:length(rhovals)) {
+    rho <- rhovals[k]
+    sPvals <- sapply(sigmavals, function (sigma) { standingProportionNumbers(mu,rho,sb,sd,sigma)$value } )
+    lines(sigmavals, sPvals, col=rhocols[k])
+}
+

### 2011-04-07 standing-patches-paper.tex
/* Standing variation */
added figure

@@
 Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
-As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s}$ is the speed of the wave of advance.
+As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
 The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
@@
 we can see  $\nu_+ = \int_0^\infty \lambda \P\{\tau>t\} dt$, and hence
+$\nu_+ = \lambda \E[\tau] $.
+Using that $\lambda_0 = \lambda / s_d$, this gives us that
 \begin{equation}
-    \nu_+ = (\mbox{mean density of new patches}) = \lambda \E[\tau] .
+    \nu_+ = (\mbox{mean density of new patches}) = \frac{1}{1+s_d \E[\tau]} .
 \end{equation}
@@
 \begin{equation}
-    z_0 = \int_0^\infty {2 \lambda_0 \pi v t} \exp \left( - 2 \lambda_0 \pi v t - \lambda \pi v^2 t^2 \right) dt .
+    z_0 = \int_0^\infty {2 \lambda_0 \pi v t} \; \exp \left( - 2 \lambda_0 \pi v t - \lambda \pi v^2 t^2 \right) dt .
 \end{equation}
+\marginnote{put in this value in terms of Gaussian CDF?}
 To evaluate this integral, again see Appendix \ref{apx:integrals}.
@@
 $a_0 / a_+ = z_0 / (1-z_0)$.
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=1.0\textwidth]{proportion-by-sigma.eps}
+\caption{ %
+{\bf Mean proportion of patches arising from standing variation.} The paramters given are for the sickle-cell example.  XXX say more.
+}
+\label{fig:cartoon}
+\end{center}
+
+\end{figure}
 
@@
 \begin{align}
-    \left( {d+1}^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
+    \left( (d+1)^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
 \end{align}

### 2011-04-07 Spatial_adaptation/standing-variation-plots.R
/* Plots */

@@ -1,4 +1,6 @@
 
+source("Spatial_adaptation/standing-variation-fns.R")
+
 # playing around
 mu <- 10^(-8)
 rho <- 10

### 2011-04-07 Spatial_adaptation/standing-variation-plots.R
/* Plots */

@@ -9,8 +9,8 @@ sd <- .01
 sigmavals <- c(0.1,1,10,20,50)
 
 # plot Standing Proportion Numbers as a function of sigma
-sPvals <- sapply(sigmavals, function (sigma) { standingProportionNumbers(mu,rho,sb,sd,sigma)$value } )
-plot(sigmavals, sPvals, type="b", xlab=expression(paste(sigma, " = dispersal SD")), ylab=expression(paste(z[0], " = proportion of patches from standing variation")), ylim=c(0,1))
+# sPvals <- sapply(sigmavals, function (sigma) { standingProportionNumbers(mu,rho,sb,sd,sigma)$value } )
+# plot(sigmavals, sPvals, type="b", xlab=expression(paste(sigma, " = dispersal SD")), ylab=expression(paste(z[0], " = proportion of patches from standing variation")), ylim=c(0,1))
 
 
 # sickle cell case from ralphcoop2010
@@ -25,7 +25,7 @@ otherlabs <- expression(paste( mu == 10^-8, "  ", s[b] == .05, "  ", s[d] == .01
 
 # plot Standing Proportion Numbers as a function of sigma
 setEPS(horizontal = FALSE, onefile = FALSE, paper = "special")
-postscript( file="proportion-by-sigma.eps", width=6.5, height=3, title=paramString() )
+postscript( file="proportion-by-sigma.eps", width=6.5, height=5, title=paramString() )
 plot(0, 0, type="n", xlab=expression(paste(sigma, " = dispersal SD")), ylab=expression(paste(z[0], " = proportion of patches from standing variation")), xlim=range(sigmavals), ylim=c(0,1), main=otherlabs )
 legend("topleft", lty=1, col=rhocols, legend=rholabs)
 for (k in 1:length(rhovals)) {

### 2011-04-11 Spatial_adaptation/standing-variation.tex
/* Standing variation */ mathematica solves a cubic
/* Standing variation */ mathematica does integrals

@@
 
+\subsection{Mathematica says}
+
+...that 
+\begin{align}
+    \int_0^\infty \exp\left( - \lambda \pi v^2 (t^2/s_d + t^3) \right) dt 
+       = \frac{ 1 }{ 27 s_d } \left(
+            \exp\left( - \frac{2 \pi v^2 \lambda}{27 s_d^3} \right) 2 \sqrt{3} \pi 
+                \left( I_{-1/3}\left( \frac{2 \pi v^2 \lambda}{27 s_d^3} \right) 
+                    + I_{1/3}\left( \frac{2 \pi v^2 \lambda}{27 s_d^3} \right) 
+            \right)
+            - {}_2F_2\left(\frac{1}{2},1;\frac{2}{3},\frac{4}{3};-\frac{4 \pi v^2 \lambda}{27 s_d^3}\right)
+        \right) ,
+\end{align}
+where $I_n(x)$ is the modified Bessel function of the first kind (solves $z^2 y''(z) + z y'(z) - (z^2 + n^2) y = 0$),
+and ${}_2F_2(\cdot)$ is the generalized hypergeometric function.
+
+...and
+\begin{align}
+    \int_0^\infty \frac{2\lambda \pi v}{s_d} t \exp\left( - \lambda \pi v^2 (t^2/s_d + t^3) \right) dt \right) dt 
+        = \frac{1}{v s_d} - \frac{1}{2 s_d^2} \exp{ \frac{\pi v^2 \lambda}{4 s_d^2} \pi \sqrt{\lambda} \left( 1 - \Phi\left( \frac{ v \sqrt{\pi v} }{ 2 s_d } \right) \right) .
+\end{align}
+
 \section{Characteristic lengths}
@@
 In $d=2$ ??
+
+\subsection{Mathematica Says}
+
+... (and Wikipedia too) that the solutions to $ax^2 + bx^3 = 1$ are:
+let 
+\[
+   u = \left( -2a^3 + 27b^2 + 3 \sqrt{3} \sqrt{ -4a^3b^2+27b^4} \right)
+\]
+and $(1, r, \bar r)$ are the cube roots of $1$, with $r=(1+i\sqrt{3})/2$ then
+\[
+    x = \begin{cases} 
+        -\frac{a}{3b} + \frac{ 2^{1/3} a^2 }{ 3 b u^{1/3} } + \frac{ u^{1/3} }{ 3 b 2^{1/3} } \\
+        -\frac{a}{3b} + r \frac{ 2^{1/3} a^2 }{ 3 b u^{1/3} } + \bar r \frac{ u^{1/3} }{ 3 b 2^{1/3} } \\
+        -\frac{a}{3b} + \bar r \frac{ 2^{1/3} a^2 }{ 3 b u^{1/3} } + r \frac{ u^{1/3} }{ 3 b 2^{1/3} }  .
+    \end{cases}
+\]
+Plugging in $a=\pi \lambda_0$ and $b=\pi \lambda v$ we get
+$u = - 2 \pi^3 \lambda_0^3 + 3 \pi^2 \left( 9 v^2 \lambda^2 + \sqrt{ 81 b^4 \lambda^4 - 12 \pi v^2 \lambda^2 \lambda_0^3 } \right)$
+... and the whole expression is ugly.
 

### 2011-04-13 Spatial_adaptation/local-adaptation.tex
WorkingWiki replaced project description

\documentclass{article}
\usepackage{amsmath,amssymb}

\newcommand{\var}{\mathop{\mbox{Var}}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\one}{\mathbf{1}}

\begin{document}

Here is one idea for computing the migration rate.
Suppose that there is a patch at the origin with $N$ individuals,
and we want to find the ``migration rate'' to another patch at distance $r$.
Each generation, offspring are divided into two classes:
those living in the good patch (i.e.\ replacing some of those $N$ individuals)
and new migrants, that might eventually make it to the other patch.
Suppose that in $dy$ there are approximately $m(y) dy$ of these latter.
Let $p(y,r)$ be the probability that the branching process begun at $y$ survives
{\em in the selective environment with all patches except the one at $r$ erased}.
If we can assume that the two are independent,
we can approximate the probability that the new patch is colonized by ``new migrants'' in a certain generation by
$\int m(y) p(y,r) dy$.

Here is another.  
\[
    \partial_t \P\{ Z_t(A) = 0 \} \approx - \sum_x (1+s(x)) \kappa(x,A) \E[Z_t(x) \one_{Z_t(A)=0}] ,
\]
so if $\E[Z_t(x) | {Z_t(A)=0}] \approx \E[Z_t(x)]$ 
then a good approximation for the rate may be
\[
    \int (1+s(x)) \kappa(x,A) \E[Z_t(x)] dx .
\]
{\bf Or,} let $\hat Z_t$ be the process with $s(A)=-\infty$.
Then $\E[ \hat Z_t(A) ] \approx \E[Z_t(A) \one_{Z_t(A)=0}]$,
so
\[
    \partial_t \P\{ Z_t(A) = 0 \} \approx - \sum_x (1+s(x)) \kappa(x,A) \E[\hat Z_t(x)] \P\{ Z_t(A) = 0\},
\]
and the approximation is
\[
    \int (1+s(x)) \kappa(x,A) \E[\hat Z_t(x)] dx ,
\]
which could by found as in Slatkin.

What about the stable distribution?
Suppose that mutants have selective (dis-)advantage $s(x)$ at location $x$,
that the symmetric dispersal kernel is $\kappa(\cdot)$, and that the mean-squared dispersal distance is $\sum_y y^2 \kappa(y) = \sigma^2$.
Then in the large population limit, if $\phi(t,x)$ is the proportion mutants at location $x$ at time $t$,
if $s$ and $\sigma^2$ are of the same (small) order,
(see Peter's slides from Berkeley probability seminar)
\begin{align}
    \partial_t \phi(t,x) &\approx \sum_y s(x) \kappa(y-x) (1-\one_{x=y}) \phi(t,y)(1-\phi(t,x)) 
            + \sum_y \kappa(y-x) (1-\frac{1}{N}\one_{x=y}) (\phi(t,x)-\phi(t,y))  \\
        &\approx s(x) \phi(t,x) (1-\phi(t,x)) + \frac{1}{2} \sigma^2 \partial_x^2 \phi(t,x) .
\end{align}
...

{\bf Note:} In some applications, 
it may be better to begin with the migration rates rather than using the theory above to estimate them.

\end{document}

### 2011-04-14 Spatial_adaptation/local-adaptation.tex
Undo revision 3036 by [[Special:Contributions/Peter|Peter]] ([[User talk:Peter|talk]])

@@
 
-Here is one idea for computing the migration rate.
-Suppose that there is a patch at the origin with $N$ individuals,
-and we want to find the ``migration rate'' to another patch at distance $r$.
-Each generation, offspring are divided into two classes:
-those living in the good patch (i.e.\ replacing some of those $N$ individuals)
-and new migrants, that might eventually make it to the other patch.
-Suppose that in $dy$ there are approximately $m(y) dy$ of these latter.
-Let $p(y,r)$ be the probability that the branching process begun at $y$ survives
-{\em in the selective environment with all patches except the one at $r$ erased}.
-If we can assume that the two are independent,
-we can approximate the probability that the new patch is colonized by ``new migrants'' in a certain generation by
-$\int m(y) p(y,r) dy$.
-
-Here is another.  
-\[
-    \partial_t \P\{ Z_t(A) = 0 \} \approx - \sum_x (1+s(x)) \kappa(x,A) \E[Z_t(x) \one_{Z_t(A)=0}] ,
-\]
-so if $\E[Z_t(x) | {Z_t(A)=0}] \approx \E[Z_t(x)]$ 
-then a good approximation for the rate may be
-\[
-    \int (1+s(x)) \kappa(x,A) \E[Z_t(x)] dx .
-\]
-{\bf Or,} let $\hat Z_t$ be the process with $s(A)=-\infty$.
-Then $\E[ \hat Z_t(A) ] \approx \E[Z_t(A) \one_{Z_t(A)=0}]$,
-so
-\[
-    \partial_t \P\{ Z_t(A) = 0 \} \approx - \sum_x (1+s(x)) \kappa(x,A) \E[\hat Z_t(x)] \P\{ Z_t(A) = 0\},
-\]
-and the approximation is
-\[
-    \int (1+s(x)) \kappa(x,A) \E[\hat Z_t(x)] dx ,
-\]
-which could by found as in Slatkin.
-
-What about the stable distribution?
 Suppose that mutants have selective (dis-)advantage $s(x)$ at location $x$,
@@
 \end{align}
-...
-
-{\bf Note:} In some applications, 
-it may be better to begin with the migration rates rather than using the theory above to estimate them.
 

### 2011-04-14 standing-patches-paper.tex
added slatkin characteristic length ref
better interpretation of characteristic length eqn

@@
 \newcommand{\R}{\mathbb{R}}
+
+\bibliographystyle{plain}
 
@@
 as well as a similar compound parameter describing the relative importance of standing variation to the process of adaptation.
+(Note that a similar approach, summarizing with a {\em characteristic length} was taken by \citep{slatkin1973geneflow} for a different problem.)
 
@@
 Therefore, if we define $\chi$ to be the unique positive solution to
+\[
     $\lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1$,
+\]
 then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
 before it is expected that one other successful mutation would have arisen in the area covered so far.
-If we define $u = 2 \pi^3 \lambda_0^3 - 27 \pi^2 v^2 \lambda^2$,
-The solution to this is {\bf XXX what?}
-(In one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
+The explicit formula for $\chi$ is cumbersome but straightforward; here we omit the lengthy expression.
+(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
+This can be rewritten 
+\begin{equation} \label{eqn:defines_chi}
+    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^2 = (\pi \rho \mu)^{-1} .
+\end{equation}
+From this we see that $\chi$ increases linearly with $\sqrt{s_b}$,
+and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
+If $\sigma s_d \gg 1$, then standing variation is relatively unimportant, 
+and this reduces to the characteristic length of \citet{ralphcoop2010}, $\chi_+ = (v/\lambda \pi)^{1/3}$.
+On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
+and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-2}$. 
+% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.
 
@@
 
+\bibliography{standing_patches_refs}
+
 \appendix

### 2011-04-14 Spatial_adaptation/local-adaptation.tex
/* Local Adaptation */ more notes

@@
 
+Note that if $\tau$ is the first time that the new patch $A$ is occupied,
+and $Z_t$ is the state of the proces, then
+\[
+    \partial_t \P\{ \tau > t \} = \int \int_A s(x) \kappa(y-x) \E[ Z_t(dx) | \tau > t ] dy \P\{ \tau > t \}
+\]
+... so we want to find the first integral (and show that for $t>\epsilon$ it is approximately constant).
+Since we are also taking the small-dispersal limit, this seems tricky.
+
+Another option: we are trying to find the first time that a mutant appears in patch $A$ {\em and fixes locally};
+if the benefit is small then there may be many unsuccessful migrants first.
+Suppose we modify the process so that offspring appearing in $A$ have distribution {\em conditioned on local extinction};
+then we suppose that each such individual has probability ${}\approx 2s_b$ of fixation,
+so that the mean number of such offspring per generation times $2s_b$ gives us the rate we are looking for.
+A supercritical branching process conditioned on extinction is a subcritical branching process,
+with mean offspring number $\partial_z \E[z^N]\vert_{z=p}$, where $N$ is the offspring number and $p$ is the probability of extinction.
+In the case of a binomial (check this?) branching, this changes $s$ to $-s$.
+
 \end{document}

### 2011-04-19 Spatial_adaptation/local-adaptation.tex
/* Local Adaptation */

@@
 
+\section{On solving FKPP with varying selection}
+
+Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
+so that (setting $\sigma^2=2$)
+\[
+    \partial_t p(t,x) = \partial_x^2 p(t,x) + s(x) p(t,x) (1-p(t,x)) .
+\]
+The stable distribution $\phi(x) = \lim_{t\to\infty} p(t,x)$ then solves
+\begin{align} \label{eqn:definingphi}
+    \partial_x^2 \phi(x) = - s(x) \phi(x) (1-\phi(x)) ,
+\end{align}
+with appropriate boundary conditions.
+If we now assume that $s(x)$ is piecewise constant,
+$s(x) = s_i$ for $x \in [x_i,x_{i+1})$, with $x_0=-\infty$ and $x_{n+2}=\infty$,
+then the equation is integrable: if we multiply through by $\partial_x \phi(x)$ and integrate, then we get that
+\begin{align} \label{eqn:conservation}
+    ( \partial_x \phi(x) )^2  &= - \int^{x} s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
+        &= - \frac{1}{2} s_i \phi(x)^2 \left( 1 - \frac{1}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) \\
+        &:= - V_i(\phi(x)) \quad \mbox{for } x \in [x_i,x_{i+1}) ,
+\end{align}
+where $K_i$ is a constant of integration and we define
+\[
+        V_i(\phi) = \frac{1}{2} s_i \phi^2 \left( 1 - \frac{1}{3} \phi \right) + K_i .
+\]
+Note that $V_i'(0)=V_i'(1)=0$, that $V_i(0)=K_i$ and $V_i(1) = K_i-s_i/6$.  
+We will always have that $V(\phi) \le 0$.
+(We have then that $\phi'' = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
+Also note that this implies ``conservation of energy'', i.e.\ $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
+Rearranging, we get that $dx = d\phi / \sqrt{-V(\phi)}$, so where $\phi(x)$ is monotone, the inverse is
+\[
+    x(\phi) = x(\phi_0) \pm \int_{\phi_0}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } .
+\]
+For each $i$ then define the elliptic function
+\[
+    F_i(\phi) = \int_{\phi_i^*}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } ,
+\]
+where take the positive branch of the square root, and $\phi_i^*$ will be chosen later.
+Then we have that
+$x(\phi) - x(\phi_0) = \pm( F(\phi) - F(\phi_0))$,
+or for an appropriate $x_0$,
+\[
+    \phi(x) = F^{-1}\left( F(\phi(x_0)) \pm (x - x_0) \right).
+\]
+
+We clearly want $\lim_{x \to \infty} \partial_x \phi(\pm x) = 0$, 
+and $\lim_{x \to \infty} \phi(\pm x)$ to be zero or one depending on the sign of $s_0$ and $s_{n+1}$.
+Since $(\partial_x \phi(x))^2 = -V(x)$, this implies that if $s_0<0$, then $K_0 = 0$,
+while if $s_0>0$ then $K_0 = s_0/6$; and likewise for $K_{n+1}$.
+
+We also require that $\phi(x)$ and $\phi'(x)$ are continuous.
+Continuity of $\phi'(x)$ is equivalent to $V_i(x_{i+1}) = V_{i+1}(x_{i+1})$,
+which we can rearrange to find an equation for $K_{i+1}$ in terms of $K_i$ and $\phi(x_{i+1})$:
+\[
+    K_{i+1} - K_i = (s_i - s_{i+1}) \frac{1}{2} \phi(x_{i+1})^2 (1-\phi(x_{i+1})/3) .
+\]
+What about $\phi(x_i)$?  Well, if $\phi(x)$ is monotone on $[x_i,x_{i+1})$ then we can wolog take $\phi_i^*=0$ or $1$ depending on the sign of $s_i$.
+Otherwise, let $\phi_i^*$ be the (unique) root of $V_i$ in $[x_i,x_{i+1})$, so $V_i(\phi_i^*)=0$.
+In this case, $\phi_i^*$ is the maximum or minimum of $\phi$ in the interval:
+if $s>0$, then $\phi_i = \max\{ \phi(x) : x \in [x_i,x_{i+1})\}$. 
+Recall we defined $F_i$ using $\phi_i^*$; now using the fact that $\phi$ is monotone with the opposite sign on either side of $\phi_i^*$,
+$x_{i+1} - x_i = F(\phi(x_{i+1})) - F(\phi_i^*) + F(\phi(x_{i})) - F(\phi_i^*)$,
+and that $F(\phi_i^*) = 0$,
+we know that the length of the $i$th stretch is
+\[
+    x_{i+1} - x_i = \pm \left( F(\phi(x_i)) + F(\phi(x_{i+1})) \right).
+\]
+
+If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be all set.
+Unfortunately, we are given $\phi(-\infty)$ and $\phi(\infty)$,
+
 \end{document}

### 2011-04-20 Spatial_adaptation/local-adaptation.tex
/* Local Adaptation */ elliptic integral info

@@
 \[
-    x_{i+1} - x_i = \pm \left( F(\phi(x_i)) + F(\phi(x_{i+1})) \right).
+    x_{i+1} - x_i = \pm \left( F_i(\phi(x_i)) + F_i(\phi(x_{i+1})) \right).
 \]
 
+Note that all the $\pm{}$'s are easily relatable to the signs of $s_i$.
 If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be all set.
-Unfortunately, we are given $\phi(-\infty)$ and $\phi(\infty)$,
+Unfortunately, we are given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
+Monty deals with this by using symmetry, so he knows $\phi(0)$ and $\phi'(0)$.
+
+We are interested in the following case:
+\[
+    s(x) = \begin{cases}
+        +s_b \qquad \mbox{for } |x|<\epsilon \\
+        -s_d \qquad \mbox{for } x<-\epsilon \; \mbox{or}\; \epsilon \le x \le R \\
+        -s_* \qquad \mbox{for } R < x < R+2\epsilon \\
+        -s_d \qquad \mbox{for } x>R+2\epsilon .
+    \end{cases}
+\]
+We can solve the case with $s_* = s_d$, since $\phi'(0) = 0$ (as can Monty, although he doesn't say what it is).
+We should be able to argue that if $R$ is large then the value of $s_*$ won't make much difference.
+
+\subsection{Doing the integrals}
+
+This from ``NEQwiki, the nonlinear equations encyclopedia'', \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
+We want to integrate
+\[
+    \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
+        \sqrt{2}{s} \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } ,
+\]
+with $C=2K/s$.
+Let $\phi^2(1-\phi/3)+C = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
+and change variables first to $y^2=(\alpha-\phi$, 
+and then to $x = y/\sqrt{\alpha-\beta}$, so that
+\begin{align*}
+    \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } 
+        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-p^2) } } \\
+        = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
+\end{align*}
+with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
+Now Jacobi's ``incomplete elliptic integral of the first kind'' is
+\[
+    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) } .
+\]
 

### 2011-04-22 standing-patches-paper.tex
latex bib oops
latex oops
bib works now?
minor notes

@@
 \[
-    $\lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1$,
+    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
 \]
@@
 
-Describe simple discrete model.
+Describe simple discrete model, for general migration rates.
 
-Approximation in continuous model by branching Brownian motion with killing to find migration rate between patches -- 
-good approximation neglecting interaction (if local population size is large). 
-Look at Barton.
+Get migration and mutation rates from a homogeneous continuous model:
+The migration rate from patch $i$ to a patch at distance $r$ is $C_i \exp(-\gamma r \sqrt{s} /\sigma)$,
+where $C_i$ depends on the size of the patch and (to a lesser degree?) the local structural details of the selective difference.
+The mutation rate is more or less (size of patch) times $\mu$; 
+more accurately we can integrate using Barton assuming step selection.
 
@@
 
-Pull out results from other paper on discrete model.
+Pull out results from other paper on discrete model:
+if migration rates are equal then we get Chinese Restaurant/Ewens;
+in any case is independent of selection.
 
 when are patches effectively isolated?   
+This occurs when the effective migration rate is much smaller than the mutation rate, namely...
 

### 2011-04-26 Spatial_adaptation/local-adaptation.tex
/* Local Adaptation */
/* Local Adaptation */ latex oops
/* Local Adaptation */ small addition
oops equation

@@
 \newcommand{\one}{\mathbf{1}}
+
+\newcommand{\sn}{\mathop{\mbox{sn}}}
 
@@
 \[
-    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) } .
+    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) }} ,
 \]
+and the Jacobian elliptic function $\sn(x;k)$ is the inverse: $F(\sn(x;k);k) = x$.
+
+As $k \to 0$, $\sn(x;k) \to \sin(x)$, while as $k \to 1$, $\sn(x;k) \to \sinh(x)$.
+
+\section{Mutation rates}
+
+Barton (Gen Res 87), has expressions for the probability of establishment as a function of spatial location in fixed landscape of advantageous pockets:
+The probability of establishment beginning at state p is <latex>F(p) = 1 - \exp(-4Ns\int \phi(x) p(dx))</latex>,
+where 
+\[ \phi(x)^2 = L\phi(x) ; \qquad \mbox{and| \qquad \frac{\partial \phi}{\partial x} = 0 \qquad \mbox{as}\; x \to \infty .\]
+Here $L\phi = \Delta\phi + f\phi$, where $$sf(x)$$ is the selective advantage is location $x$.
+This form follows by assuming that the offspring of each individual go extinct independently (as in a branching process),
+and so the probability of extinction is a product over the extant individuals;
+there are $2N p(x) dx$ individuals at $x$
+and the probability of exinction of a single individual at $x$ is $\exp (2s \phi(x))$;
+Barton says this is derived by [http://www.bibsonomy.org/bibtex/23c3cf150aea743a3bfe52155d00cbf7b/peter.ralph Pollack] (J Appl Prob '66) using branching processes (and assuming Poisson offspring); 
+Pollack actually does it in a discrete world; Barton is (silently) taking a continuum limit.
+Barton also motivates by linearizing about establishment, and solves some interesting cases.
+
+So, let $f(x) = s(x)/s_b$ be the rescaled selective benefit, and we want to solve
+\[
+    \phi''(x) = - f(x) \phi(x) + \phi(x)^2 .
+\]
+This fits into the framework above, if $f(x)$ is piecewise constant,
+since
+\[
+    d \phi = \pm dx \sqrt{ K + f \phi + \phi^2 } .
+\]
+Barton gives the example we want...
 

### 2011-04-26 standing-patches-paper.tex
oops equation

@@
 \begin{equation} \label{eqn:defines_chi}
-    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^2 = (\pi \rho \mu)^{-1} .
+    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
 \end{equation}
-From this we see that $\chi$ increases linearly with $\sqrt{s_b}$,
+From this we see that $\chi$ increases linearly with $1/\sqrt{s_b}$,
 and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.

### 2011-04-26 standing-patches-paper.tex
minor notes

@@
 On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
-and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-2}$. 
+and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-1/2}$. 
+\marginnote{XXX check this from below.}
+% Note also substituting $z = \chi (\sqrt{2s_b} / (s_d \sigma)$ gets $z^2 + z^3 = (s_d^3 \sigma^2)/(\pi \rho \mu)$.
 % More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.
@@
 \[
-    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 \right) ,
+    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) ,
 \]
@@
     \E[\tau] % &= \int_0^\infty \P\{ \tau > t \} dt \\
-        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 \right) dt.
+        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) dt.
 \end{align}

### 2011-04-27 standing-patches-paper.tex
fixes to characteristic section

@@
 For concreteness, suppose that before time $t=0$,
-the mutant type has fitness $1-s_d$ relative to the neutral type,
+the mutant type has fitness $1-s_d$ relative to the neutral type
+(i.e.\ it produces on average $1-s_d$ times the number of offspring per generation),
 and that after time $t=0$,
 the mutant type has fitness $1+s_b$,
-where $s_d$ and $s_b$ are positive numbers.
+where $0<s_d<1$, and $s_b>0$ will usually be assumed to be small.
 We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
+The number of offspring has finite variance, which to keep the formulas simple we assume to be equal to one.
 (If the mutation is recessive, XXX more complicated things happen.)
@@
 before it is expected that one other successful mutation would have arisen in the area covered so far.
-The explicit formula for $\chi$ is cumbersome but straightforward; here we omit the lengthy expression.
+The explicit formula for $\chi$ is cumbersome; here we omit it.
 (This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
@@
 \begin{equation} \label{eqn:defines_chi}
-    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
+%    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
+\left( \chi s_d \sigma \sqrt{2s_b} \right)^2 + \left( \chi s_d \sigma \sqrt{2s_b} \right)^3 
+        = (s_d^3 \sigma^2)/(\pi \rho \mu).
 \end{equation}
-From this we see that $\chi$ increases linearly with $1/\sqrt{s_b}$,
-and that the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
-If $\sigma s_d \gg 1$, then standing variation is relatively unimportant, 
-and this reduces to the characteristic length of \citet{ralphcoop2010}, $\chi_+ = (v/\lambda \pi)^{1/3}$.
-On the other hand, if $\sigma s_d \ll 1$, then most adaptation come from standing variation, 
-and the solution to \eqref{eqn:defines_chi} is close to $\chi_0 = (\lambda_0 \pi)^{-1/2}$. 
-\marginnote{XXX check this from below.}
-% Note also substituting $z = \chi (\sqrt{2s_b} / (s_d \sigma)$ gets $z^2 + z^3 = (s_d^3 \sigma^2)/(\pi \rho \mu)$.
+From this we see that $\chi$ increases {\em linearly} with $1/\sqrt{s_b}$,
+and decreases with $\rho$ and $\mu$.
+Furthermore, the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
+If $\sigma^3 \s_d^3/(\pi \rho \mu)$ is small, then $\chi \approx \sqrt{ s_d / (2 s_b \pi \rho \mu)} = 1/\sqrt{\pi \lambda_0}$,
+% + O( \gamma ) with $\gamma = \sigma^3 \s_d^3/(\pi \rho \mu)$
+and nearly all mutants come from standing variation (note that $\sigma$ does not enter into it).
+On the other hand, if $\sigma^3 \s_d^3/(\pi \rho \mu)$ is large, then $\chi \approx (2s_b)^{-1/2} \left( \sigma \pi \rho \mu)\right)^{-1/3}$,
+% + O( 1 ) .
+the characteristic length from \citet{ralphcoop2010},
+in which case standing variation is relatively unimportant.
 % More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.
@@
 and occur within the first $\chi/v$ generations.
-Said another way, if we look in a circular region of space of radius $\chi$ over $\chi/d$ generations,
+Said another way, if we look in a circular region of space of radius $\chi$ over $\chi/v$ generations,
 we expect to find one mutational origin;
@@
 Since the mean proportion of patches that come from standing variation is $\lambda_0 / (\lambda_0 + \nu_+)$,
-this gives us the exact quantity corresponding to $\pi_0$ from \eqref{eqn:pizero}.
+this gives us the exact quantity we approximated by $\pi_0$ in \eqref{eqn:pizero}.
 There are $\lambda_0 + \nu_+$ patches per unit area, so

### 2011-04-27 standing-patches-paper.tex
added examples section

@@
 \left( \chi s_d \sigma \sqrt{2s_b} \right)^2 + \left( \chi s_d \sigma \sqrt{2s_b} \right)^3 
-        = (s_d^3 \sigma^2)/(\pi \rho \mu).
+        = \frac{ s_d^3 \sigma^2 }{ \pi \rho \mu } .
 \end{equation}
-From this we see that $\chi$ increases {\em linearly} with $1/\sqrt{s_b}$,
+From this we see that $\chi$ increases linearly with $1/\sqrt{s_b}$,
 and decreases with $\rho$ and $\mu$.
 Furthermore, the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
-If $\sigma^3 \s_d^3/(\pi \rho \mu)$ is small, then $\chi \approx \sqrt{ s_d / (2 s_b \pi \rho \mu)} = 1/\sqrt{\pi \lambda_0}$,
-% + O( \gamma ) with $\gamma = \sigma^3 \s_d^3/(\pi \rho \mu)$
-and nearly all mutants come from standing variation (note that $\sigma$ does not enter into it).
-On the other hand, if $\sigma^3 \s_d^3/(\pi \rho \mu)$ is large, then $\chi \approx (2s_b)^{-1/2} \left( \sigma \pi \rho \mu)\right)^{-1/3}$,
+For instance, if $\sigma^3 s_d^3/(\pi \rho \mu)$ is small, then $\chi \approx \sqrt{ s_d / (2 s_b \pi \rho \mu)} = 1/\sqrt{\pi \lambda_0}$,
+% + O( \gamma ) with $\gamma = \sigma^3 s_d^3/(\pi \rho \mu)$
+and nearly all mutants come from standing variation.
+On the other hand, if $\sigma^3 s_d^3/(\pi \rho \mu)$ is large, then $\chi \approx (2s_b)^{-1/2} \left( \sigma \pi \rho \mu)\right)^{-1/3}$,
 % + O( 1 ) .

### 2011-05-03 Spatial_adaptation/standing-variation-inflorescences.tex
WorkingWiki replaced project description

--- 2011-04-06 Spatial_adaptation/standing-variation.tex
@@
 \documentclass{article}
-\usepackage{amsmath,amssymb}
-
-\newcommand{\var}{\mathop{\mbox{Var}}}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\one}{\mathbf{1}}
+\input{Spatial_adaptation/preamble}
 
@@
 
-\section{Notes on branching process approximation}
-
-The branching process approximation will be good as long as the probability of "interactions" are small,
-i.e.\ we can couple the branching process with the actual density--regulated model.
-Let $Z_t(x)$ be the numbers of selected individuals in deme $x$.
-If we assume that $Z_t$ is smooth enough, 
-the numbers of nearby individuals to interact with is approximately $Z_t(x) \sigma^d$,
-and the total number of individuals is about $\rho \sigma^d$,
-so the rate of interactions is about $(Z_t(x) \sigma^d)^2/(\rho \sigma^d)$.
-Now if $Z$ is a spatial branching process beginning from a single individual at $x=0$ at $t=0$, 
-and growing at rate $r$, then 
-\[ 
-    Z_t(x) \approx \exp(rt - \|x\|^2/(2\sigma^2 t))/ (2 \pi \sigma^2)^{d/2} 
-        \le e^{rt} / ( \sigma^2 t)^{d/2}
-\]
-and so we need $t$ such that
-\[
-    e^{2rt} t^{-d} << \rho \sigma^{3d} .
-\]
-or equivalently,
-\[
-    e^{2rt/d} t^{-1} << \rho^{1/d} \sigma^3 .
-\]
-This will be true if $t$ is of smaller order than the logarithm of the effective population density, $\log \rho \sigma^d$.
-
-\subsection{On the random delay}
-
-Let $Y_t$ be the total number of offspring coming from a single individual
-(in the spatial, density-regulated model).
-Then we know that $(Y_t/t)^{1/d} \to v$, 
-but we saw from simulations that it looked like $Y_t^{1/d} \approx v(t-\tau)$,
-where $\tau$ is the (random) establishment time.
-(It would be useful to look at the coefficient of variation of $\tau$.)
-Well, we know that at first, $Y_t \approx W e^{rt}$, 
-where $W= \lim e^{-rt} Z_t$, and $Z_t$ is the associated branching process,
-and $r$ is the growth rate (er, $r=s$?).
-And eventually, $\partial_t Y_t^{1/d} \approx v$.
-If this switchover happened at a fixed size $C$,
-the size at which wavelike behavior took over,
-i.e.\ when $W e^{rt} = C$,
-then for $\tau = \frac{1}{r}\log(C/W)$ we'd have 
-\[
-    Y_t = \begin{cases}
-        W e^{rt} \quad & t \le \tau \\
-        (v (t-\tau))^d + C \quad & t > \tau ,
-\end{cases}
-\]
-which if $C$ is small relative to the scale we look on will look as described above.
-So the random delay is $(1/r)\log(C/W)$; and so the mean delay is something like $(1/r)\log C$
-and the randomness is determined by $(1/r)\log (1/W)$.
-
-What about how much variance is left in the branching process?
-Well, since $Z_{t+s} = \sum_{k=1}^{Z_t} Z^k_{s}$, where $Z^k_s$ are iid copies of $Z_s$,
-we have that $\var[Z_{t+s}|Z_t] = Z_t \var[Z_s]$,
-and so the coefficient of variation of $Z_{t+s}$ given $Z_t$ is
-\[
-   \mathop{CV}[Z_{t+s}|Z_t] &= \frac{\sqrt{Z_t \var[Z_s]}}{ Z_t \E[Z_s] } \\
-        &= e^{-rs} \sqrt{ \var[Z_s] } Z^{-1/2} \\
-        &\simeq e^{-rt/2} \sqrt{ \var[W] } .
-\]
-
-\section{Properties of the mixed model}
-
-Suppose that there are (point masses of) mutations present at time 0 as a Poisson process of rate $\lambda_0$ (calculated above),
-and that after time 0 mutations arise at rate $\lambda$.
-Suppose that the mutations spread distance $f(t)$ in time $t$, 
-and that the volume of the cone of height $t$ is given by $h(t) = \int_0^t \omega_d f(s)^d ds$.
-
-Let $\tau$ denote the time until the origin is adapted.
-Then
-\begin{align}
-   \P \{ \tau > t \} &= \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\}, \quad \mbox{and} \\
-   \E [ \tau ] &= \int_0^\infty \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} .
-\end{align}
-
-Let $z_0$ be the mean proportion of space covered by mutations present at time 0.
-Then, by the general result on competing inhomogeneous exponentials,
-\begin{align}
-    z_0 &= \P\{ \mbox{origin reached by mutations from $\lambda_0$ before $\lambda$} \} \\
-        &= \int_0^\infty \lambda_0 d \omega_d f'(t) f(t)^{d-1} \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} dt .
-\end{align}
-The mean density of new mutations that would have taken off but were prevented by standing variation is
-\begin{align}
-    \nu_* = \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt
-\end{align}
-and so the mean proportion of new mutations that would be successful in the absence of standing variation, but are prevented by standing variation,
-is 
-\begin{align}
-    z_* = \frac{ \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt }
-                { \int_0^\infty \lambda \exp\{ - \lambda h(t) \} dt } .
-\end{align}
-
-Let $X$ be the distance to the eventually enclosing mutation and $R=f(\tau)$ (see first paper).
-Then if $g(t) = \lambda_0 \omega_d f(t)^d + \lambda h(t)$ then $(X,R)$ has density
-\[
-    d \omega_d x^{d-1} \exp(-g(f^{-1}(r)) dx dr,
-\]
-so integrating over $r$,
-\[
-    \P\{ X \in dx \} = dx d \omega_d x^{d-1} \int_x^\infty \exp( - g(f^{-1}(r)) ) dr ,
-\]
-and
-\[
-    \E[ X^n ] = \frac{d\omega_d}{n+d} \int_0^\infty r^{n+d} \exp( - g(f^{-1}(r)) ) dr .
-\]
-
-If we rescale time by $\alpha$ and space by $\beta$, 
-then this transforms $(\lambda_0, \lambda, f(\cdot))$ into $(\lambda_0/\beta^d, \lambda/(\alpha \beta^d), f(\cdot/\alpha)/\beta)$.  
-Setting $\beta = \lambda_0^{1/d}$ and $\alpha = \lambda / \lambda_0$, this is mapped to $(1,1,f(\cdot \lambda_0/\lambda)/\lambda_0^{1/d})$.
-
-\subsection{Constant speed}
-
-Now let $f(t) = vt$, so $h(t) = \frac{ \omega_d v^d }{ d+1 } t^{d+1}$.
-The integrals that appear above are of the following form:
-\begin{align}
-  \int_0^\infty t^c \exp \left( - \alpha t^a - \beta t^b \right) dt 
-            &= \left( b^{-1} \beta^{ (1-c)/b } \right) \int_0^\infty u^{(c+1-b)/b} \exp\left( - \alpha \beta^{-a/b} u^{a/b} - u \right) du ,
-\end{align}
-so it suffices to learn something about the function
-\begin{equation}
-    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt .
-\end{equation}
-We will only be interested in this as a function of $x$; 
-the coefficients $a$ and $b$ will merely depend on the dimension and what quantity we are calculating.
-We can construct a power series for $G$ in $x$:
-\begin{align}
-    \partial_x^n G(a,b,x) &= \int_0^\infty (-1)^n t^{a+nb} \exp\left( -x t^b - t \right) dt \\
-            &= (-1)^n G(a+nb,b,x) , \quad \mbox{so} \\
-    \partial_x^n G(a,b,x) \vert_{x=0} &= \Gamma(a+nb+1) .
-\end{align}
-So, if everything works out, we will have that
-\begin{align}
-    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
-\end{align}
-Something simliar is noted at the Wikipedia page on the Gaussian integral, and claims such integrals show up in quantum field theory.
-
-A few facts, first.
-Using repeatedly the identity $\Gamma(x+1) = x\Gamma(x)$
-and the Pochammer symbol $(x)_n = x(x+1)\cdots(x+n-1)$,
-note that for $m \in \N$,
-\[
-    \Gamma(m+y) = (y)_m \Gamma(y).
-\]
-Also note that for $c$ and $m$ integers,
-\[
-    (y)_{cm} = c^{cm} \prod_{k=0}^{c-1} (\frac{y+k}{c})_{m} ,
-\]
-and in particular, that
-\[
-    (cm+\ell)! = \ell! (\ell+1)_{cm} = \ell! c^{cm} \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m .
-\]
-
-Now suppose that $b$ is rational (as will be the case), letting $b = b'/c$ with $b'$ and $c$ positive integers.
-Therefore, setting $n = mc+\ell$,
-\begin{align}
-    G(a,b,x) &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} \Gamma(a+\ell b'/c+mb'+1) \\
-            &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} (a+\ell b')_{mb'} \Gamma(a + \ell b') \\
-            &= \sum_{\ell=0}^{c-1} \frac{ (-x)^{\ell} }{ \ell! } \Gamma(a + \ell b') \sum_{m \ge 0} ( (-cx)^c / (b')^{b'} )^m 
-                    \frac{ \prod_{j=0}^{b'-1} (\frac{ a+b'\ell+j }{ b' } )_m }{ \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m } ,
-\end{align}
-which is the sum of $c$ (generalized) hypergeometric series.
-For these to converge, we need $b' \le c$, e.g.\ $b\le 1$.
-
-In particular, if we let $A=\lambda_0 \omega_d v^d $ and $B=\lambda h(1)$, then 
-\begin{align}
-    \E[\tau] &= \int_0^\infty e^{-A t^d - B t^{d+1}} dt \\
-            &= \frac{B^{1/(d+1)}}{d+1} \int_0^\infty u^{-d/(d+1))} \exp\left( -AB^{-d/(d+1))} u^{d/(d+1)} -u \right) du \\
-            &= \frac{B^{1/(d+1)}}{d+1} G \left(-d/(d+1), d/(d+1), A B^{-d/(d+1)} \right) .
-\end{align}
-So, if $d=1$, then we have {\bf (verify this by usual methods!)}, with $x = AB^{-1/2}$,
-\begin{align}
-    \E[\tau] &= \frac{\lambda^{1/(2)}}{2} G \left( -1/2, 1/2, x \right) \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) \sum_{m\ge0} (4x^2)^m \frac{ (-1/2)_m }{ (1/2)_m (1)_m } 
-                        - x \Gamma(1/2) \sum_{m\ge0} (4x^2)^m \frac{ (1/2)_m }{ (1)_m (3/2)_m }  \right\} \\
-            &=  \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) {}_1F_1(-1/2;1/2;4x^2) - x \Gamma(1/2) {}_1F_1(1/2;3/2;4x^2) \right\} \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ - \Gamma(-1/2) (1/x) \gamma(-1/2,4x^2) - \Gamma(1/2) (1/x) (\sqrt{\pi}/8) \erf(4x^2) \right\} .
-\end{align}
-
-And in $d=2$, since
-\begin{align}
-    G(-2/3,2/3,x) &= \Gamma(-2/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (-1/3)_m (1/6)_m }{ (1/3)_m (2/3)_m (1)_m }
-                    - x \Gamma(4/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (2/3)_m (7/6)_m }{ (2/3)_m (1)_m (4/3)_m }
-                        + \frac{1}{2} x^2 \Gamma(10/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (5/3)_m (13/6)_m }{ (2)_m (4/3)_m (5/3)_m } \\
-                &= \Gamma(-2/3) {}_2F_2(-1/3,1/6;1/3,2/3;-\frac{27x^2}{4})
-                    -x \Gamma(4/3) {}_1F_1(7/6;4/3;-\frac{27x^2}{4})
-                       -\frac{1}{2}x^2 \Gamma(10/3) {}_1F_1(13/6;4/3;-\frac{27x^2}{4}) .
-\end{align}
-
-\section{Characteristic lengths}
-
-How to pick appropriate rule-of-thumb parameters?
-Recall that before we had $\chi$ being the smallest solution to
-\[
-    f\left( \frac{1}{\lambda \chi^d \omega_d} \right) = \chi.
-\]
-Now, suppose we draw a cylinder in space-time with radius $\chi$ and height $t=f^{-1}(\chi)$;
-as before we choose $\chi$ so that the mean number of points in this cylinder is unity.
-Then
-\[
-    1 = \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d ,
-\]
-and so we choose $\chi$ to be the unique positive solution to 
-\[
-    f\left( \frac{1 - \lambda_0 \chi^d \omega_d}{\lambda \chi^d \omega_d} \right) = \chi.
-\]
-
-As for the proportion of mutations coming from standing variation,
-it seems a natural measure would be the mean number of such mutations in such a cylinder,
-which is using the definition of $\chi$,
-\begin{align*}
-    \pi_0 &= \frac{\lambda_0 \omega_d \chi^d }{ \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d} \\
-        & = \frac{ 1 }{ 1 + \lambda f^{-1}(\chi) / \lambda_0 } \\
-        &= \lambda_0 \omega_d \chi^d .
-\end{align*}
-
-Now, if $f(t) = vt$, then $f^{-1}(r) = r/v$,
-and so we have $\chi$ the positive root of
-\[
-  \omega_d \chi^d (\lambda_0 + \lambda v \chi) = 1 .
-\]
-If $d=1$ then this is
-\[
-    \chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v } .
-\]
-In $d=2$ ??
-
-\section{General relationship between various quantities}
-
-We know that there are $\lambda_0$ successful standing variants per unit area;
-suppose that there are $\nu_+$ successful new mutants per unit area.
-Therefore, $\lambda_0+\nu_+$ is the mean number of successful mutants per unit area (duh),
-and $1/(\lambda_0+\nu_+)$ is the mean area associated with a randomly chosen patch.
-Let $p_0$ be the proportion of space occupied by standing variants, 
-let $a_0$ be the mean area associated with a randomly chosen standing variant,
-and let $a_+$ be the mean area associated with a randomly chosen new variant,
-so that
-\[
-    p_0 = \frac{ \lambda_0 a_0 }{ \lambda_0 a_0 + \nu_+ a_+ } = \lambda_0 a_0 ,
-\]
-since the denominator is equal to 1.
-
-We know how to calculate $p_0$, above, from thinking about the competing exponentials,
-which allows us to find $a_0$.
-
-Now, as in the first paper, the mean density of new patches is easy to calculate -- 
-$\lambda \nu_+ = \E[\tau]$, where $\tau$ is the mean adaptation time of a point;
-this is because 
-\begin{align*}
-    \P\{ \tau > t \} &= \P\{ \mbox{no points in $t$-cone} \} \\
-            &= (1/\lambda) \P\{ \mbox{new successful adaptation arises in non-occupied $(t+dt,x+dx)$} \} ;
-\end{align*}
-integrating over $t$ produces the result.
-
-Finally, the proportion of patches that come from new mutation is $\nu_+ / (\nu_+ + \lambda_0)$,
-and we know that, conditioning on the type of a randomly chosen patch,
-\[
-    \frac{1}{\lambda_0+\nu_+} = \frac{\lambda_0}{\lambda_0+\nu_+} a_0 + \frac{\nu_+}{\lambda_0+\nu_+} a_+ .
-\]
-So, given $\lambda_0$, $\nu_+$, and $a_0$, we can find $a_+$.
-
-\section{Poisson--Vorono\u\i\ facts}
-
-Fill in some facts about PV tesselations.
-
 \end{document}

### 2011-05-03 Spatial_adaptation/standing-variation-bp-approx.tex
WorkingWiki replaced project description

--- 2011-04-06 Spatial_adaptation/standing-variation.tex
@@
 \documentclass{article}
-\usepackage{amsmath,amssymb}
-
-\newcommand{\var}{\mathop{\mbox{Var}}}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\one}{\mathbf{1}}
-
-\begin{document}
-
-\section{Size of inflorescences}
-
-Let the local rate of growth at location $x$ and time $t$ be given by $f(x,t)$,
-and for each $x$ and $t$ let $(X^{(x,t)}_{s})_{s \ge t}$ have the distribution of a branching process begun with one individual at $x$ and time $t$,
-and let $\kappa^{(x,t)}_s(dz) = \P\{ |X^{(x,t)}_s| \in dz \}$  be the marginal distribution of the size of $X^{(x,t)}$.
-For a given mutation rate $\mu$ and population density $\rho(dx)$ we have a Poisson process of inflorescences of rate $\mu \rho$,
-each of which having distribution $\kappa$.  
-Then the set of inflorescences alive at time $t$ is Poisson, 
-and the mean measure of such inflorescences is
-\[ \nu(dx,A) = \mu \rho(dx) \int_{-\infty}^t \P\{ X^{(x,s)}_t \in A \} ds .\]
-The first thing we'd want to do is to ignore the spatial extent of $X$ and only look at the total number of individuals,
-so we need to know
-\[ \int_{-\infty}^t \kappa^{(x,s)}_t(dz) ds . \]
-
-Now assume everything is homogeneous, and let $Z_t$ be a continuous-time branching process begun at zero 
-and having branching rate $1$ and offspring distribution $\{p_k : k\ge 0\}$.
-Suppose that $Z$ has total lifetime $\tau$ and that $\tau$ is finite almost surely.
-If we treat each inflorescence as a point mass, then the set of (location, mass) coordinates $(x,z)$
-are Poisson with intensity $dx \otimes q(1,z) dz$ where $q(1,z)=q(z)$ is the Green function of $Z$, namely
-\[  
-  q(y,z) = \E^y \left[ \int_0^\tau \one(Z_t = z) dt \right] = \int_0^\infty P^t_{yz} dt .
-\]
-and $q$ solves $L q(y,z) = y \sum_k p_k (q(y+k-1,z)-q(y,z)) = \delta_{y=z}$.
-
-If we are interested in the density of ultimately successful influorescences,
-we need to futhermore weight by the probability that the offspring of any of the initial (assumed to be small) number of individuals in an inflorescence survive.
-If $p_s$ is the probability of survival in the new selective regime,
-since the probability that an influorescence of size $z$ survives in the new selective regime is $1 - (1-p_s)^z$,
-the mean measure for successful influorescences is
-\[
-   \mu dx \int_0^\infty (1-F(1-p_s,t)) dt ,
-\]
-where $F$ is the generating function for $Z_t$ (in the old selective regime),
-\[
-  F(a,t) = \E[a^{Z_t} \; | \; Z_0 = 1] .
-\]
-Note that 
-if $\phi$ be the generating function for the offspring distribution in the old selective regime,
-then $F$ solves
-\[
-   \partial_t F(a,t) = \phi(F(a,t)) - F(a,t).
-\]
-If $\phi_+$ be the generating function for the offspring distribution in the new selective regime,
-\[
-  \phi_+(a) = \sum_k p_k a^k .
-\]
-Now, $p_s$ is the minimal solution to $\phi_+(a) = a$.
-
+\input{Spatial_adaptation/preamble}
 \section{Notes on branching process approximation}
@@
 and so the coefficient of variation of $Z_{t+s}$ given $Z_t$ is
-\[
+\begin{align}
    \mathop{CV}[Z_{t+s}|Z_t] &= \frac{\sqrt{Z_t \var[Z_s]}}{ Z_t \E[Z_s] } \\
@@
         &\simeq e^{-rt/2} \sqrt{ \var[W] } .
-\]
-
-\section{Properties of the mixed model}
-
-Suppose that there are (point masses of) mutations present at time 0 as a Poisson process of rate $\lambda_0$ (calculated above),
-and that after time 0 mutations arise at rate $\lambda$.
-Suppose that the mutations spread distance $f(t)$ in time $t$, 
-and that the volume of the cone of height $t$ is given by $h(t) = \int_0^t \omega_d f(s)^d ds$.
-
-Let $\tau$ denote the time until the origin is adapted.
-Then
-\begin{align}
-   \P \{ \tau > t \} &= \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\}, \quad \mbox{and} \\
-   \E [ \tau ] &= \int_0^\infty \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} .
 \end{align}
 
-Let $z_0$ be the mean proportion of space covered by mutations present at time 0.
-Then, by the general result on competing inhomogeneous exponentials,
-\begin{align}
-    z_0 &= \P\{ \mbox{origin reached by mutations from $\lambda_0$ before $\lambda$} \} \\
-        &= \int_0^\infty \lambda_0 d \omega_d f'(t) f(t)^{d-1} \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} dt .
-\end{align}
-The mean density of new mutations that would have taken off but were prevented by standing variation is
-\begin{align}
-    \nu_* = \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt
-\end{align}
-and so the mean proportion of new mutations that would be successful in the absence of standing variation, but are prevented by standing variation,
-is 
-\begin{align}
-    z_* = \frac{ \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt }
-                { \int_0^\infty \lambda \exp\{ - \lambda h(t) \} dt } .
-\end{align}
-
-Let $X$ be the distance to the eventually enclosing mutation and $R=f(\tau)$ (see first paper).
-Then if $g(t) = \lambda_0 \omega_d f(t)^d + \lambda h(t)$ then $(X,R)$ has density
-\[
-    d \omega_d x^{d-1} \exp(-g(f^{-1}(r)) dx dr,
-\]
-so integrating over $r$,
-\[
-    \P\{ X \in dx \} = dx d \omega_d x^{d-1} \int_x^\infty \exp( - g(f^{-1}(r)) ) dr ,
-\]
-and
-\[
-    \E[ X^n ] = \frac{d\omega_d}{n+d} \int_0^\infty r^{n+d} \exp( - g(f^{-1}(r)) ) dr .
-\]
-
-If we rescale time by $\alpha$ and space by $\beta$, 
-then this transforms $(\lambda_0, \lambda, f(\cdot))$ into $(\lambda_0/\beta^d, \lambda/(\alpha \beta^d), f(\cdot/\alpha)/\beta)$.  
-Setting $\beta = \lambda_0^{1/d}$ and $\alpha = \lambda / \lambda_0$, this is mapped to $(1,1,f(\cdot \lambda_0/\lambda)/\lambda_0^{1/d})$.
-
-\subsection{Constant speed}
-
-Now let $f(t) = vt$, so $h(t) = \frac{ \omega_d v^d }{ d+1 } t^{d+1}$.
-The integrals that appear above are of the following form:
-\begin{align}
-  \int_0^\infty t^c \exp \left( - \alpha t^a - \beta t^b \right) dt 
-            &= \left( b^{-1} \beta^{ (1-c)/b } \right) \int_0^\infty u^{(c+1-b)/b} \exp\left( - \alpha \beta^{-a/b} u^{a/b} - u \right) du ,
-\end{align}
-so it suffices to learn something about the function
-\begin{equation}
-    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt .
-\end{equation}
-We will only be interested in this as a function of $x$; 
-the coefficients $a$ and $b$ will merely depend on the dimension and what quantity we are calculating.
-We can construct a power series for $G$ in $x$:
-\begin{align}
-    \partial_x^n G(a,b,x) &= \int_0^\infty (-1)^n t^{a+nb} \exp\left( -x t^b - t \right) dt \\
-            &= (-1)^n G(a+nb,b,x) , \quad \mbox{so} \\
-    \partial_x^n G(a,b,x) \vert_{x=0} &= \Gamma(a+nb+1) .
-\end{align}
-So, if everything works out, we will have that
-\begin{align}
-    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
-\end{align}
-Something simliar is noted at the Wikipedia page on the Gaussian integral, and claims such integrals show up in quantum field theory.
-
-A few facts, first.
-Using repeatedly the identity $\Gamma(x+1) = x\Gamma(x)$
-and the Pochammer symbol $(x)_n = x(x+1)\cdots(x+n-1)$,
-note that for $m \in \N$,
-\[
-    \Gamma(m+y) = (y)_m \Gamma(y).
-\]
-Also note that for $c$ and $m$ integers,
-\[
-    (y)_{cm} = c^{cm} \prod_{k=0}^{c-1} (\frac{y+k}{c})_{m} ,
-\]
-and in particular, that
-\[
-    (cm+\ell)! = \ell! (\ell+1)_{cm} = \ell! c^{cm} \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m .
-\]
-
-Now suppose that $b$ is rational (as will be the case), letting $b = b'/c$ with $b'$ and $c$ positive integers.
-Therefore, setting $n = mc+\ell$,
-\begin{align}
-    G(a,b,x) &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} \Gamma(a+\ell b'/c+mb'+1) \\
-            &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} (a+\ell b')_{mb'} \Gamma(a + \ell b') \\
-            &= \sum_{\ell=0}^{c-1} \frac{ (-x)^{\ell} }{ \ell! } \Gamma(a + \ell b') \sum_{m \ge 0} ( (-cx)^c / (b')^{b'} )^m 
-                    \frac{ \prod_{j=0}^{b'-1} (\frac{ a+b'\ell+j }{ b' } )_m }{ \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m } ,
-\end{align}
-which is the sum of $c$ (generalized) hypergeometric series.
-For these to converge, we need $b' \le c$, e.g.\ $b\le 1$.
-
-In particular, if we let $A=\lambda_0 \omega_d v^d $ and $B=\lambda h(1)$, then 
-\begin{align}
-    \E[\tau] &= \int_0^\infty e^{-A t^d - B t^{d+1}} dt \\
-            &= \frac{B^{1/(d+1)}}{d+1} \int_0^\infty u^{-d/(d+1))} \exp\left( -AB^{-d/(d+1))} u^{d/(d+1)} -u \right) du \\
-            &= \frac{B^{1/(d+1)}}{d+1} G \left(-d/(d+1), d/(d+1), A B^{-d/(d+1)} \right) .
-\end{align}
-So, if $d=1$, then we have {\bf (verify this by usual methods!)}, with $x = AB^{-1/2}$,
-\begin{align}
-    \E[\tau] &= \frac{\lambda^{1/(2)}}{2} G \left( -1/2, 1/2, x \right) \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) \sum_{m\ge0} (4x^2)^m \frac{ (-1/2)_m }{ (1/2)_m (1)_m } 
-                        - x \Gamma(1/2) \sum_{m\ge0} (4x^2)^m \frac{ (1/2)_m }{ (1)_m (3/2)_m }  \right\} \\
-            &=  \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) {}_1F_1(-1/2;1/2;4x^2) - x \Gamma(1/2) {}_1F_1(1/2;3/2;4x^2) \right\} \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ - \Gamma(-1/2) (1/x) \gamma(-1/2,4x^2) - \Gamma(1/2) (1/x) (\sqrt{\pi}/8) \erf(4x^2) \right\} .
-\end{align}
-
-And in $d=2$, since
-\begin{align}
-    G(-2/3,2/3,x) &= \Gamma(-2/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (-1/3)_m (1/6)_m }{ (1/3)_m (2/3)_m (1)_m }
-                    - x \Gamma(4/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (2/3)_m (7/6)_m }{ (2/3)_m (1)_m (4/3)_m }
-                        + \frac{1}{2} x^2 \Gamma(10/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (5/3)_m (13/6)_m }{ (2)_m (4/3)_m (5/3)_m } \\
-                &= \Gamma(-2/3) {}_2F_2(-1/3,1/6;1/3,2/3;-\frac{27x^2}{4})
-                    -x \Gamma(4/3) {}_1F_1(7/6;4/3;-\frac{27x^2}{4})
-                       -\frac{1}{2}x^2 \Gamma(10/3) {}_1F_1(13/6;4/3;-\frac{27x^2}{4}) .
-\end{align}
-
-\section{Characteristic lengths}
-
-How to pick appropriate rule-of-thumb parameters?
-Recall that before we had $\chi$ being the smallest solution to
-\[
-    f\left( \frac{1}{\lambda \chi^d \omega_d} \right) = \chi.
-\]
-Now, suppose we draw a cylinder in space-time with radius $\chi$ and height $t=f^{-1}(\chi)$;
-as before we choose $\chi$ so that the mean number of points in this cylinder is unity.
-Then
-\[
-    1 = \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d ,
-\]
-and so we choose $\chi$ to be the unique positive solution to 
-\[
-    f\left( \frac{1 - \lambda_0 \chi^d \omega_d}{\lambda \chi^d \omega_d} \right) = \chi.
-\]
-
-As for the proportion of mutations coming from standing variation,
-it seems a natural measure would be the mean number of such mutations in such a cylinder,
-which is using the definition of $\chi$,
-\begin{align*}
-    \pi_0 &= \frac{\lambda_0 \omega_d \chi^d }{ \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d} \\
-        & = \frac{ 1 }{ 1 + \lambda f^{-1}(\chi) / \lambda_0 } \\
-        &= \lambda_0 \omega_d \chi^d .
-\end{align*}
-
-Now, if $f(t) = vt$, then $f^{-1}(r) = r/v$,
-and so we have $\chi$ the positive root of
-\[
-  \omega_d \chi^d (\lambda_0 + \lambda v \chi) = 1 .
-\]
-If $d=1$ then this is
-\[
-    \chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v } .
-\]
-In $d=2$ ??
-
-\section{General relationship between various quantities}
-
-We know that there are $\lambda_0$ successful standing variants per unit area;
-suppose that there are $\nu_+$ successful new mutants per unit area.
-Therefore, $\lambda_0+\nu_+$ is the mean number of successful mutants per unit area (duh),
-and $1/(\lambda_0+\nu_+)$ is the mean area associated with a randomly chosen patch.
-Let $p_0$ be the proportion of space occupied by standing variants, 
-let $a_0$ be the mean area associated with a randomly chosen standing variant,
-and let $a_+$ be the mean area associated with a randomly chosen new variant,
-so that
-\[
-    p_0 = \frac{ \lambda_0 a_0 }{ \lambda_0 a_0 + \nu_+ a_+ } = \lambda_0 a_0 ,
-\]
-since the denominator is equal to 1.
-
-We know how to calculate $p_0$, above, from thinking about the competing exponentials,
-which allows us to find $a_0$.
-
-Now, as in the first paper, the mean density of new patches is easy to calculate -- 
-$\lambda \nu_+ = \E[\tau]$, where $\tau$ is the mean adaptation time of a point;
-this is because 
-\begin{align*}
-    \P\{ \tau > t \} &= \P\{ \mbox{no points in $t$-cone} \} \\
-            &= (1/\lambda) \P\{ \mbox{new successful adaptation arises in non-occupied $(t+dt,x+dx)$} \} ;
-\end{align*}
-integrating over $t$ produces the result.
-
-Finally, the proportion of patches that come from new mutation is $\nu_+ / (\nu_+ + \lambda_0)$,
-and we know that, conditioning on the type of a randomly chosen patch,
-\[
-    \frac{1}{\lambda_0+\nu_+} = \frac{\lambda_0}{\lambda_0+\nu_+} a_0 + \frac{\nu_+}{\lambda_0+\nu_+} a_+ .
-\]
-So, given $\lambda_0$, $\nu_+$, and $a_0$, we can find $a_+$.
-
-\section{Poisson--Vorono\u\i\ facts}
-
-Fill in some facts about PV tesselations.
-
 \end{document}

### 2011-05-03 Spatial_adaptation/standing-variation-mixed-model.tex
WorkingWiki replaced project description

--- 2011-04-07 Spatial_adaptation/standing-variation.tex
@@
 \documentclass{article}
-\usepackage{amsmath,amssymb}
-
-\newcommand{\var}{\mathop{\mbox{Var}}}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\one}{\mathbf{1}}
-
-\begin{document}
-
-\section{Size of inflorescences}
-
-Let the local rate of growth at location $x$ and time $t$ be given by $f(x,t)$,
-and for each $x$ and $t$ let $(X^{(x,t)}_{s})_{s \ge t}$ have the distribution of a branching process begun with one individual at $x$ and time $t$,
-and let $\kappa^{(x,t)}_s(dz) = \P\{ |X^{(x,t)}_s| \in dz \}$  be the marginal distribution of the size of $X^{(x,t)}$.
-For a given mutation rate $\mu$ and population density $\rho(dx)$ we have a Poisson process of inflorescences of rate $\mu \rho$,
-each of which having distribution $\kappa$.  
-Then the set of inflorescences alive at time $t$ is Poisson, 
-and the mean measure of such inflorescences is
-\[ \nu(dx,A) = \mu \rho(dx) \int_{-\infty}^t \P\{ X^{(x,s)}_t \in A \} ds .\]
-The first thing we'd want to do is to ignore the spatial extent of $X$ and only look at the total number of individuals,
-so we need to know
-\[ \int_{-\infty}^t \kappa^{(x,s)}_t(dz) ds . \]
-
-Now assume everything is homogeneous, and let $Z_t$ be a continuous-time branching process begun at zero 
-and having branching rate $1$ and offspring distribution $\{p_k : k\ge 0\}$.
-Suppose that $Z$ has total lifetime $\tau$ and that $\tau$ is finite almost surely.
-If we treat each inflorescence as a point mass, then the set of (location, mass) coordinates $(x,z)$
-are Poisson with intensity $dx \otimes q(1,z) dz$ where $q(1,z)=q(z)$ is the Green function of $Z$, namely
-\[  
-  q(y,z) = \E^y \left[ \int_0^\tau \one(Z_t = z) dt \right] = \int_0^\infty P^t_{yz} dt .
-\]
-and $q$ solves $L q(y,z) = y \sum_k p_k (q(y+k-1,z)-q(y,z)) = \delta_{y=z}$.
-
-If we are interested in the density of ultimately successful influorescences,
-we need to futhermore weight by the probability that the offspring of any of the initial (assumed to be small) number of individuals in an inflorescence survive.
-If $p_s$ is the probability of survival in the new selective regime,
-since the probability that an influorescence of size $z$ survives in the new selective regime is $1 - (1-p_s)^z$,
-the mean measure for successful influorescences is
-\[
-   \mu dx \int_0^\infty (1-F(1-p_s,t)) dt ,
-\]
-where $F$ is the generating function for $Z_t$ (in the old selective regime),
-\[
-  F(a,t) = \E[a^{Z_t} \; | \; Z_0 = 1] .
-\]
-Note that 
-if $\phi$ be the generating function for the offspring distribution in the old selective regime,
-then $F$ solves
-\[
-   \partial_t F(a,t) = \phi(F(a,t)) - F(a,t).
-\]
-If $\phi_+$ be the generating function for the offspring distribution in the new selective regime,
-\[
-  \phi_+(a) = \sum_k p_k a^k .
-\]
-Now, $p_s$ is the minimal solution to $\phi_+(a) = a$.
-
-\section{Notes on branching process approximation}
-
-The branching process approximation will be good as long as the probability of "interactions" are small,
-i.e.\ we can couple the branching process with the actual density--regulated model.
-Let $Z_t(x)$ be the numbers of selected individuals in deme $x$.
-If we assume that $Z_t$ is smooth enough, 
-the numbers of nearby individuals to interact with is approximately $Z_t(x) \sigma^d$,
-and the total number of individuals is about $\rho \sigma^d$,
-so the rate of interactions is about $(Z_t(x) \sigma^d)^2/(\rho \sigma^d)$.
-Now if $Z$ is a spatial branching process beginning from a single individual at $x=0$ at $t=0$, 
-and growing at rate $r$, then 
-\[ 
-    Z_t(x) \approx \exp(rt - \|x\|^2/(2\sigma^2 t))/ (2 \pi \sigma^2)^{d/2} 
-        \le e^{rt} / ( \sigma^2 t)^{d/2}
-\]
-and so we need $t$ such that
-\[
-    e^{2rt} t^{-d} << \rho \sigma^{3d} .
-\]
-or equivalently,
-\[
-    e^{2rt/d} t^{-1} << \rho^{1/d} \sigma^3 .
-\]
-This will be true if $t$ is of smaller order than the logarithm of the effective population density, $\log \rho \sigma^d$.
-
-\subsection{On the random delay}
-
-Let $Y_t$ be the total number of offspring coming from a single individual
-(in the spatial, density-regulated model).
-Then we know that $(Y_t/t)^{1/d} \to v$, 
-but we saw from simulations that it looked like $Y_t^{1/d} \approx v(t-\tau)$,
-where $\tau$ is the (random) establishment time.
-(It would be useful to look at the coefficient of variation of $\tau$.)
-Well, we know that at first, $Y_t \approx W e^{rt}$, 
-where $W= \lim e^{-rt} Z_t$, and $Z_t$ is the associated branching process,
-and $r$ is the growth rate (er, $r=s$?).
-And eventually, $\partial_t Y_t^{1/d} \approx v$.
-If this switchover happened at a fixed size $C$,
-the size at which wavelike behavior took over,
-i.e.\ when $W e^{rt} = C$,
-then for $\tau = \frac{1}{r}\log(C/W)$ we'd have 
-\[
-    Y_t = \begin{cases}
-        W e^{rt} \quad & t \le \tau \\
-        (v (t-\tau))^d + C \quad & t > \tau ,
-\end{cases}
-\]
-which if $C$ is small relative to the scale we look on will look as described above.
-So the random delay is $(1/r)\log(C/W)$; and so the mean delay is something like $(1/r)\log C$
-and the randomness is determined by $(1/r)\log (1/W)$.
-
-What about how much variance is left in the branching process?
-Well, since $Z_{t+s} = \sum_{k=1}^{Z_t} Z^k_{s}$, where $Z^k_s$ are iid copies of $Z_s$,
-we have that $\var[Z_{t+s}|Z_t] = Z_t \var[Z_s]$,
-and so the coefficient of variation of $Z_{t+s}$ given $Z_t$ is
-\[
-   \mathop{CV}[Z_{t+s}|Z_t] &= \frac{\sqrt{Z_t \var[Z_s]}}{ Z_t \E[Z_s] } \\
-        &= e^{-rs} \sqrt{ \var[Z_s] } Z^{-1/2} \\
-        &\simeq e^{-rt/2} \sqrt{ \var[W] } .
-\]
-
+\input{Spatial_adaptation/preamble}
 \section{Properties of the mixed model}
@@
             &=  \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) {}_1F_1(-1/2;1/2;4x^2) - x \Gamma(1/2) {}_1F_1(1/2;3/2;4x^2) \right\} \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ - \Gamma(-1/2) (1/x) \gamma(-1/2,4x^2) - \Gamma(1/2) (1/x) (\sqrt{\pi}/8) \erf(4x^2) \right\} .
+            &= \frac{\lambda^{1/(2)}}{2} \left\{ - \Gamma(-1/2) (1/x) \gamma(-1/2,4x^2) - \Gamma(1/2) (1/x) (\sqrt{\pi}/8) (1-\Phi(4x^2)) \right\} .
 \end{align}
+where $\Phi$ is the error function.
 
@@
 
-\section{Characteristic lengths}
+\subsection{Mathematica says}
 
-How to pick appropriate rule-of-thumb parameters?
-Recall that before we had $\chi$ being the smallest solution to
-\[
-    f\left( \frac{1}{\lambda \chi^d \omega_d} \right) = \chi.
-\]
-Now, suppose we draw a cylinder in space-time with radius $\chi$ and height $t=f^{-1}(\chi)$;
-as before we choose $\chi$ so that the mean number of points in this cylinder is unity.
-Then
-\[
-    1 = \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d ,
-\]
-and so we choose $\chi$ to be the unique positive solution to 
-\[
-    f\left( \frac{1 - \lambda_0 \chi^d \omega_d}{\lambda \chi^d \omega_d} \right) = \chi.
-\]
+...that 
+\begin{align}
+    \int_0^\infty \exp\left( - \lambda \pi v^2 (t^2/s_d + t^3) \right) dt 
+       = \frac{ 1 }{ 27 s_d } \left(
+            \exp\left( - \frac{2 \pi v^2 \lambda}{27 s_d^3} \right) 2 \sqrt{3} \pi 
+                \left( I_{-1/3}\left( \frac{2 \pi v^2 \lambda}{27 s_d^3} \right) 
+                    + I_{1/3}\left( \frac{2 \pi v^2 \lambda}{27 s_d^3} \right) 
+            \right)
+            - 9 {}_2F_2\left(\frac{1}{2},1;\frac{2}{3},\frac{4}{3};-\frac{4 \pi v^2 \lambda}{27 s_d^3}\right)
+        \right) ,
+\end{align}
+where $I_n(x)$ is the modified Bessel function of the first kind (solves $z^2 y''(z) + z y'(z) - (z^2 + n^2) y = 0$),
+and ${}_2F_2(\cdot)$ is the generalized hypergeometric function.
 
-As for the proportion of mutations coming from standing variation,
-it seems a natural measure would be the mean number of such mutations in such a cylinder,
-which is using the definition of $\chi$,
-\begin{align*}
-    \pi_0 &= \frac{\lambda_0 \omega_d \chi^d }{ \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d} \\
-        & = \frac{ 1 }{ 1 + \lambda f^{-1}(\chi) / \lambda_0 } \\
-        &= \lambda_0 \omega_d \chi^d .
-\end{align*}
-
-Now, if $f(t) = vt$, then $f^{-1}(r) = r/v$,
-and so we have $\chi$ the positive root of
-\[
-  \omega_d \chi^d (\lambda_0 + \lambda v \chi) = 1 .
-\]
-If $d=1$ then this is
-\[
-    \chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v } .
-\]
-In $d=2$ ??
-
-\section{General relationship between various quantities}
-
-We know that there are $\lambda_0$ successful standing variants per unit area;
-suppose that there are $\nu_+$ successful new mutants per unit area.
-Therefore, $\lambda_0+\nu_+$ is the mean number of successful mutants per unit area (duh),
-and $1/(\lambda_0+\nu_+)$ is the mean area associated with a randomly chosen patch.
-Let $p_0$ be the proportion of space occupied by standing variants, 
-let $a_0$ be the mean area associated with a randomly chosen standing variant,
-and let $a_+$ be the mean area associated with a randomly chosen new variant,
-so that
-\[
-    p_0 = \frac{ \lambda_0 a_0 }{ \lambda_0 a_0 + \nu_+ a_+ } = \lambda_0 a_0 ,
-\]
-since the denominator is equal to 1.
-
-We know how to calculate $p_0$, above, from thinking about the competing exponentials,
-which allows us to find $a_0$.
-
-Now, as in the first paper, the mean density of new patches is easy to calculate -- 
-$\lambda \nu_+ = \E[\tau]$, where $\tau$ is the mean adaptation time of a point;
-this is because 
-\begin{align*}
-    \P\{ \tau > t \} &= \P\{ \mbox{no points in $t$-cone} \} \\
-            &= (1/\lambda) \P\{ \mbox{new successful adaptation arises in non-occupied $(t+dt,x+dx)$} \} ;
-\end{align*}
-integrating over $t$ produces the result.
-
-Finally, the proportion of patches that come from new mutation is $\nu_+ / (\nu_+ + \lambda_0)$,
-and we know that, conditioning on the type of a randomly chosen patch,
-\[
-    \frac{1}{\lambda_0+\nu_+} = \frac{\lambda_0}{\lambda_0+\nu_+} a_0 + \frac{\nu_+}{\lambda_0+\nu_+} a_+ .
-\]
-So, given $\lambda_0$, $\nu_+$, and $a_0$, we can find $a_+$.
-
-\section{Poisson--Vorono\u\i\ facts}
-
-Fill in some facts about PV tesselations.
+...and
+\begin{align}
+    \int_0^\infty \frac{2\lambda \pi v}{s_d} t \exp\left( - \lambda \pi v^2 (t^2/s_d + t^3) \right) dt 
+        = \frac{1}{v s_d} - \frac{1}{2 s_d^2} \exp\left( \frac{\pi v^2 \lambda}{4 s_d^2} \right) \pi \sqrt{\lambda} \left( 1 - \Phi\left( \frac{ v \sqrt{\pi \lambda} }{ 2 s_d } \right) \right) .
+\end{align}
+where $\Phi$ is the Gaussian CDF.
 

### 2011-05-03 Spatial_adaptation/standing-variation-char-lens.tex
WorkingWiki replaced project description

--- 2011-04-06 Spatial_adaptation/standing-variation.tex
@@
 \documentclass{article}
-\usepackage{amsmath,amssymb}
-
-\newcommand{\var}{\mathop{\mbox{Var}}}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\one}{\mathbf{1}}
-
-\begin{document}
-
-\section{Size of inflorescences}
-
-Let the local rate of growth at location $x$ and time $t$ be given by $f(x,t)$,
-and for each $x$ and $t$ let $(X^{(x,t)}_{s})_{s \ge t}$ have the distribution of a branching process begun with one individual at $x$ and time $t$,
-and let $\kappa^{(x,t)}_s(dz) = \P\{ |X^{(x,t)}_s| \in dz \}$  be the marginal distribution of the size of $X^{(x,t)}$.
-For a given mutation rate $\mu$ and population density $\rho(dx)$ we have a Poisson process of inflorescences of rate $\mu \rho$,
-each of which having distribution $\kappa$.  
-Then the set of inflorescences alive at time $t$ is Poisson, 
-and the mean measure of such inflorescences is
-\[ \nu(dx,A) = \mu \rho(dx) \int_{-\infty}^t \P\{ X^{(x,s)}_t \in A \} ds .\]
-The first thing we'd want to do is to ignore the spatial extent of $X$ and only look at the total number of individuals,
-so we need to know
-\[ \int_{-\infty}^t \kappa^{(x,s)}_t(dz) ds . \]
-
-Now assume everything is homogeneous, and let $Z_t$ be a continuous-time branching process begun at zero 
-and having branching rate $1$ and offspring distribution $\{p_k : k\ge 0\}$.
-Suppose that $Z$ has total lifetime $\tau$ and that $\tau$ is finite almost surely.
-If we treat each inflorescence as a point mass, then the set of (location, mass) coordinates $(x,z)$
-are Poisson with intensity $dx \otimes q(1,z) dz$ where $q(1,z)=q(z)$ is the Green function of $Z$, namely
-\[  
-  q(y,z) = \E^y \left[ \int_0^\tau \one(Z_t = z) dt \right] = \int_0^\infty P^t_{yz} dt .
-\]
-and $q$ solves $L q(y,z) = y \sum_k p_k (q(y+k-1,z)-q(y,z)) = \delta_{y=z}$.
-
-If we are interested in the density of ultimately successful influorescences,
-we need to futhermore weight by the probability that the offspring of any of the initial (assumed to be small) number of individuals in an inflorescence survive.
-If $p_s$ is the probability of survival in the new selective regime,
-since the probability that an influorescence of size $z$ survives in the new selective regime is $1 - (1-p_s)^z$,
-the mean measure for successful influorescences is
-\[
-   \mu dx \int_0^\infty (1-F(1-p_s,t)) dt ,
-\]
-where $F$ is the generating function for $Z_t$ (in the old selective regime),
-\[
-  F(a,t) = \E[a^{Z_t} \; | \; Z_0 = 1] .
-\]
-Note that 
-if $\phi$ be the generating function for the offspring distribution in the old selective regime,
-then $F$ solves
-\[
-   \partial_t F(a,t) = \phi(F(a,t)) - F(a,t).
-\]
-If $\phi_+$ be the generating function for the offspring distribution in the new selective regime,
-\[
-  \phi_+(a) = \sum_k p_k a^k .
-\]
-Now, $p_s$ is the minimal solution to $\phi_+(a) = a$.
-
-\section{Notes on branching process approximation}
-
-The branching process approximation will be good as long as the probability of "interactions" are small,
-i.e.\ we can couple the branching process with the actual density--regulated model.
-Let $Z_t(x)$ be the numbers of selected individuals in deme $x$.
-If we assume that $Z_t$ is smooth enough, 
-the numbers of nearby individuals to interact with is approximately $Z_t(x) \sigma^d$,
-and the total number of individuals is about $\rho \sigma^d$,
-so the rate of interactions is about $(Z_t(x) \sigma^d)^2/(\rho \sigma^d)$.
-Now if $Z$ is a spatial branching process beginning from a single individual at $x=0$ at $t=0$, 
-and growing at rate $r$, then 
-\[ 
-    Z_t(x) \approx \exp(rt - \|x\|^2/(2\sigma^2 t))/ (2 \pi \sigma^2)^{d/2} 
-        \le e^{rt} / ( \sigma^2 t)^{d/2}
-\]
-and so we need $t$ such that
-\[
-    e^{2rt} t^{-d} << \rho \sigma^{3d} .
-\]
-or equivalently,
-\[
-    e^{2rt/d} t^{-1} << \rho^{1/d} \sigma^3 .
-\]
-This will be true if $t$ is of smaller order than the logarithm of the effective population density, $\log \rho \sigma^d$.
-
-\subsection{On the random delay}
-
-Let $Y_t$ be the total number of offspring coming from a single individual
-(in the spatial, density-regulated model).
-Then we know that $(Y_t/t)^{1/d} \to v$, 
-but we saw from simulations that it looked like $Y_t^{1/d} \approx v(t-\tau)$,
-where $\tau$ is the (random) establishment time.
-(It would be useful to look at the coefficient of variation of $\tau$.)
-Well, we know that at first, $Y_t \approx W e^{rt}$, 
-where $W= \lim e^{-rt} Z_t$, and $Z_t$ is the associated branching process,
-and $r$ is the growth rate (er, $r=s$?).
-And eventually, $\partial_t Y_t^{1/d} \approx v$.
-If this switchover happened at a fixed size $C$,
-the size at which wavelike behavior took over,
-i.e.\ when $W e^{rt} = C$,
-then for $\tau = \frac{1}{r}\log(C/W)$ we'd have 
-\[
-    Y_t = \begin{cases}
-        W e^{rt} \quad & t \le \tau \\
-        (v (t-\tau))^d + C \quad & t > \tau ,
-\end{cases}
-\]
-which if $C$ is small relative to the scale we look on will look as described above.
-So the random delay is $(1/r)\log(C/W)$; and so the mean delay is something like $(1/r)\log C$
-and the randomness is determined by $(1/r)\log (1/W)$.
-
-What about how much variance is left in the branching process?
-Well, since $Z_{t+s} = \sum_{k=1}^{Z_t} Z^k_{s}$, where $Z^k_s$ are iid copies of $Z_s$,
-we have that $\var[Z_{t+s}|Z_t] = Z_t \var[Z_s]$,
-and so the coefficient of variation of $Z_{t+s}$ given $Z_t$ is
-\[
-   \mathop{CV}[Z_{t+s}|Z_t] &= \frac{\sqrt{Z_t \var[Z_s]}}{ Z_t \E[Z_s] } \\
-        &= e^{-rs} \sqrt{ \var[Z_s] } Z^{-1/2} \\
-        &\simeq e^{-rt/2} \sqrt{ \var[W] } .
-\]
-
-\section{Properties of the mixed model}
-
-Suppose that there are (point masses of) mutations present at time 0 as a Poisson process of rate $\lambda_0$ (calculated above),
-and that after time 0 mutations arise at rate $\lambda$.
-Suppose that the mutations spread distance $f(t)$ in time $t$, 
-and that the volume of the cone of height $t$ is given by $h(t) = \int_0^t \omega_d f(s)^d ds$.
-
-Let $\tau$ denote the time until the origin is adapted.
-Then
-\begin{align}
-   \P \{ \tau > t \} &= \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\}, \quad \mbox{and} \\
-   \E [ \tau ] &= \int_0^\infty \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} .
-\end{align}
-
-Let $z_0$ be the mean proportion of space covered by mutations present at time 0.
-Then, by the general result on competing inhomogeneous exponentials,
-\begin{align}
-    z_0 &= \P\{ \mbox{origin reached by mutations from $\lambda_0$ before $\lambda$} \} \\
-        &= \int_0^\infty \lambda_0 d \omega_d f'(t) f(t)^{d-1} \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} dt .
-\end{align}
-The mean density of new mutations that would have taken off but were prevented by standing variation is
-\begin{align}
-    \nu_* = \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt
-\end{align}
-and so the mean proportion of new mutations that would be successful in the absence of standing variation, but are prevented by standing variation,
-is 
-\begin{align}
-    z_* = \frac{ \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt }
-                { \int_0^\infty \lambda \exp\{ - \lambda h(t) \} dt } .
-\end{align}
-
-Let $X$ be the distance to the eventually enclosing mutation and $R=f(\tau)$ (see first paper).
-Then if $g(t) = \lambda_0 \omega_d f(t)^d + \lambda h(t)$ then $(X,R)$ has density
-\[
-    d \omega_d x^{d-1} \exp(-g(f^{-1}(r)) dx dr,
-\]
-so integrating over $r$,
-\[
-    \P\{ X \in dx \} = dx d \omega_d x^{d-1} \int_x^\infty \exp( - g(f^{-1}(r)) ) dr ,
-\]
-and
-\[
-    \E[ X^n ] = \frac{d\omega_d}{n+d} \int_0^\infty r^{n+d} \exp( - g(f^{-1}(r)) ) dr .
-\]
-
-If we rescale time by $\alpha$ and space by $\beta$, 
-then this transforms $(\lambda_0, \lambda, f(\cdot))$ into $(\lambda_0/\beta^d, \lambda/(\alpha \beta^d), f(\cdot/\alpha)/\beta)$.  
-Setting $\beta = \lambda_0^{1/d}$ and $\alpha = \lambda / \lambda_0$, this is mapped to $(1,1,f(\cdot \lambda_0/\lambda)/\lambda_0^{1/d})$.
-
-\subsection{Constant speed}
-
-Now let $f(t) = vt$, so $h(t) = \frac{ \omega_d v^d }{ d+1 } t^{d+1}$.
-The integrals that appear above are of the following form:
-\begin{align}
-  \int_0^\infty t^c \exp \left( - \alpha t^a - \beta t^b \right) dt 
-            &= \left( b^{-1} \beta^{ (1-c)/b } \right) \int_0^\infty u^{(c+1-b)/b} \exp\left( - \alpha \beta^{-a/b} u^{a/b} - u \right) du ,
-\end{align}
-so it suffices to learn something about the function
-\begin{equation}
-    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt .
-\end{equation}
-We will only be interested in this as a function of $x$; 
-the coefficients $a$ and $b$ will merely depend on the dimension and what quantity we are calculating.
-We can construct a power series for $G$ in $x$:
-\begin{align}
-    \partial_x^n G(a,b,x) &= \int_0^\infty (-1)^n t^{a+nb} \exp\left( -x t^b - t \right) dt \\
-            &= (-1)^n G(a+nb,b,x) , \quad \mbox{so} \\
-    \partial_x^n G(a,b,x) \vert_{x=0} &= \Gamma(a+nb+1) .
-\end{align}
-So, if everything works out, we will have that
-\begin{align}
-    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
-\end{align}
-Something simliar is noted at the Wikipedia page on the Gaussian integral, and claims such integrals show up in quantum field theory.
-
-A few facts, first.
-Using repeatedly the identity $\Gamma(x+1) = x\Gamma(x)$
-and the Pochammer symbol $(x)_n = x(x+1)\cdots(x+n-1)$,
-note that for $m \in \N$,
-\[
-    \Gamma(m+y) = (y)_m \Gamma(y).
-\]
-Also note that for $c$ and $m$ integers,
-\[
-    (y)_{cm} = c^{cm} \prod_{k=0}^{c-1} (\frac{y+k}{c})_{m} ,
-\]
-and in particular, that
-\[
-    (cm+\ell)! = \ell! (\ell+1)_{cm} = \ell! c^{cm} \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m .
-\]
-
-Now suppose that $b$ is rational (as will be the case), letting $b = b'/c$ with $b'$ and $c$ positive integers.
-Therefore, setting $n = mc+\ell$,
-\begin{align}
-    G(a,b,x) &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} \Gamma(a+\ell b'/c+mb'+1) \\
-            &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} (a+\ell b')_{mb'} \Gamma(a + \ell b') \\
-            &= \sum_{\ell=0}^{c-1} \frac{ (-x)^{\ell} }{ \ell! } \Gamma(a + \ell b') \sum_{m \ge 0} ( (-cx)^c / (b')^{b'} )^m 
-                    \frac{ \prod_{j=0}^{b'-1} (\frac{ a+b'\ell+j }{ b' } )_m }{ \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m } ,
-\end{align}
-which is the sum of $c$ (generalized) hypergeometric series.
-For these to converge, we need $b' \le c$, e.g.\ $b\le 1$.
-
-In particular, if we let $A=\lambda_0 \omega_d v^d $ and $B=\lambda h(1)$, then 
-\begin{align}
-    \E[\tau] &= \int_0^\infty e^{-A t^d - B t^{d+1}} dt \\
-            &= \frac{B^{1/(d+1)}}{d+1} \int_0^\infty u^{-d/(d+1))} \exp\left( -AB^{-d/(d+1))} u^{d/(d+1)} -u \right) du \\
-            &= \frac{B^{1/(d+1)}}{d+1} G \left(-d/(d+1), d/(d+1), A B^{-d/(d+1)} \right) .
-\end{align}
-So, if $d=1$, then we have {\bf (verify this by usual methods!)}, with $x = AB^{-1/2}$,
-\begin{align}
-    \E[\tau] &= \frac{\lambda^{1/(2)}}{2} G \left( -1/2, 1/2, x \right) \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) \sum_{m\ge0} (4x^2)^m \frac{ (-1/2)_m }{ (1/2)_m (1)_m } 
-                        - x \Gamma(1/2) \sum_{m\ge0} (4x^2)^m \frac{ (1/2)_m }{ (1)_m (3/2)_m }  \right\} \\
-            &=  \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) {}_1F_1(-1/2;1/2;4x^2) - x \Gamma(1/2) {}_1F_1(1/2;3/2;4x^2) \right\} \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ - \Gamma(-1/2) (1/x) \gamma(-1/2,4x^2) - \Gamma(1/2) (1/x) (\sqrt{\pi}/8) \erf(4x^2) \right\} .
-\end{align}
-
-And in $d=2$, since
-\begin{align}
-    G(-2/3,2/3,x) &= \Gamma(-2/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (-1/3)_m (1/6)_m }{ (1/3)_m (2/3)_m (1)_m }
-                    - x \Gamma(4/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (2/3)_m (7/6)_m }{ (2/3)_m (1)_m (4/3)_m }
-                        + \frac{1}{2} x^2 \Gamma(10/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (5/3)_m (13/6)_m }{ (2)_m (4/3)_m (5/3)_m } \\
-                &= \Gamma(-2/3) {}_2F_2(-1/3,1/6;1/3,2/3;-\frac{27x^2}{4})
-                    -x \Gamma(4/3) {}_1F_1(7/6;4/3;-\frac{27x^2}{4})
-                       -\frac{1}{2}x^2 \Gamma(10/3) {}_1F_1(13/6;4/3;-\frac{27x^2}{4}) .
-\end{align}
-
+\input{Spatial_adaptation/preamble}
 \section{Characteristic lengths}
@@
 \[
-  \omega_d \chi^d (\lambda_0 + \lambda v \chi) = 1 .
+  \omega_d \chi^d (\lambda_0 + \lambda \chi/v) = 1 .
 \]
@@
 \[
-    \chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v } .
+    \chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda/v } - 2 \lambda_0 }{ 4 \lambda/v } .
 \]
@@
 
-\section{General relationship between various quantities}
+\subsection{Mathematica Says}
 
-We know that there are $\lambda_0$ successful standing variants per unit area;
-suppose that there are $\nu_+$ successful new mutants per unit area.
-Therefore, $\lambda_0+\nu_+$ is the mean number of successful mutants per unit area (duh),
-and $1/(\lambda_0+\nu_+)$ is the mean area associated with a randomly chosen patch.
-Let $p_0$ be the proportion of space occupied by standing variants, 
-let $a_0$ be the mean area associated with a randomly chosen standing variant,
-and let $a_+$ be the mean area associated with a randomly chosen new variant,
-so that
+... (and Wikipedia too) that the solutions to $ax^2 + bx^3 = 1$ are:
+let 
 \[
-    p_0 = \frac{ \lambda_0 a_0 }{ \lambda_0 a_0 + \nu_+ a_+ } = \lambda_0 a_0 ,
+   u = \left( -2a^3 + 27b^2 + 3 \sqrt{3} \sqrt{ -4a^3b^2+27b^4} \right)
 \]
-since the denominator is equal to 1.
-
-We know how to calculate $p_0$, above, from thinking about the competing exponentials,
-which allows us to find $a_0$.
-
-Now, as in the first paper, the mean density of new patches is easy to calculate -- 
-$\lambda \nu_+ = \E[\tau]$, where $\tau$ is the mean adaptation time of a point;
-this is because 
-\begin{align*}
-    \P\{ \tau > t \} &= \P\{ \mbox{no points in $t$-cone} \} \\
-            &= (1/\lambda) \P\{ \mbox{new successful adaptation arises in non-occupied $(t+dt,x+dx)$} \} ;
-\end{align*}
-integrating over $t$ produces the result.
-
-Finally, the proportion of patches that come from new mutation is $\nu_+ / (\nu_+ + \lambda_0)$,
-and we know that, conditioning on the type of a randomly chosen patch,
+and $(1, r, \bar r)$ are the cube roots of $1$, with $r=(1+i\sqrt{3})/2$ then
 \[
-    \frac{1}{\lambda_0+\nu_+} = \frac{\lambda_0}{\lambda_0+\nu_+} a_0 + \frac{\nu_+}{\lambda_0+\nu_+} a_+ .
+    x = \begin{cases} 
+        -\frac{a}{3b} + \frac{ 2^{1/3} a^2 }{ 3 b u^{1/3} } + \frac{ u^{1/3} }{ 3 b 2^{1/3} } \\
+        -\frac{a}{3b} + r \frac{ 2^{1/3} a^2 }{ 3 b u^{1/3} } + \bar r \frac{ u^{1/3} }{ 3 b 2^{1/3} } \\
+        -\frac{a}{3b} + \bar r \frac{ 2^{1/3} a^2 }{ 3 b u^{1/3} } + r \frac{ u^{1/3} }{ 3 b 2^{1/3} }  .
+    \end{cases}
 \]
-So, given $\lambda_0$, $\nu_+$, and $a_0$, we can find $a_+$.
-
-\section{Poisson--Vorono\u\i\ facts}
-
-Fill in some facts about PV tesselations.
+Plugging in $a=\pi \lambda_0$ and $b=\pi \lambda/v$ we get
+$u = - 2 \pi^3 \lambda_0^3 + 3 \pi^2 \left( 9 v^{-2} \lambda^2 + \sqrt{ 81 b^4 \lambda^4 - 12 \pi v^{-2} \lambda^2 \lambda_0^3 } \right)$
+... and the whole expression is ugly.
 

### 2011-05-03 Spatial_adaptation/standing-variation-general.tex
WorkingWiki replaced project description

--- 2011-04-06 Spatial_adaptation/standing-variation.tex
@@
 \documentclass{article}
-\usepackage{amsmath,amssymb}
-
-\newcommand{\var}{\mathop{\mbox{Var}}}
-\renewcommand{\P}{\mathbb{P}}
-\newcommand{\E}{\mathbb{E}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\one}{\mathbf{1}}
-
-\begin{document}
-
-\section{Size of inflorescences}
-
-Let the local rate of growth at location $x$ and time $t$ be given by $f(x,t)$,
-and for each $x$ and $t$ let $(X^{(x,t)}_{s})_{s \ge t}$ have the distribution of a branching process begun with one individual at $x$ and time $t$,
-and let $\kappa^{(x,t)}_s(dz) = \P\{ |X^{(x,t)}_s| \in dz \}$  be the marginal distribution of the size of $X^{(x,t)}$.
-For a given mutation rate $\mu$ and population density $\rho(dx)$ we have a Poisson process of inflorescences of rate $\mu \rho$,
-each of which having distribution $\kappa$.  
-Then the set of inflorescences alive at time $t$ is Poisson, 
-and the mean measure of such inflorescences is
-\[ \nu(dx,A) = \mu \rho(dx) \int_{-\infty}^t \P\{ X^{(x,s)}_t \in A \} ds .\]
-The first thing we'd want to do is to ignore the spatial extent of $X$ and only look at the total number of individuals,
-so we need to know
-\[ \int_{-\infty}^t \kappa^{(x,s)}_t(dz) ds . \]
-
-Now assume everything is homogeneous, and let $Z_t$ be a continuous-time branching process begun at zero 
-and having branching rate $1$ and offspring distribution $\{p_k : k\ge 0\}$.
-Suppose that $Z$ has total lifetime $\tau$ and that $\tau$ is finite almost surely.
-If we treat each inflorescence as a point mass, then the set of (location, mass) coordinates $(x,z)$
-are Poisson with intensity $dx \otimes q(1,z) dz$ where $q(1,z)=q(z)$ is the Green function of $Z$, namely
-\[  
-  q(y,z) = \E^y \left[ \int_0^\tau \one(Z_t = z) dt \right] = \int_0^\infty P^t_{yz} dt .
-\]
-and $q$ solves $L q(y,z) = y \sum_k p_k (q(y+k-1,z)-q(y,z)) = \delta_{y=z}$.
-
-If we are interested in the density of ultimately successful influorescences,
-we need to futhermore weight by the probability that the offspring of any of the initial (assumed to be small) number of individuals in an inflorescence survive.
-If $p_s$ is the probability of survival in the new selective regime,
-since the probability that an influorescence of size $z$ survives in the new selective regime is $1 - (1-p_s)^z$,
-the mean measure for successful influorescences is
-\[
-   \mu dx \int_0^\infty (1-F(1-p_s,t)) dt ,
-\]
-where $F$ is the generating function for $Z_t$ (in the old selective regime),
-\[
-  F(a,t) = \E[a^{Z_t} \; | \; Z_0 = 1] .
-\]
-Note that 
-if $\phi$ be the generating function for the offspring distribution in the old selective regime,
-then $F$ solves
-\[
-   \partial_t F(a,t) = \phi(F(a,t)) - F(a,t).
-\]
-If $\phi_+$ be the generating function for the offspring distribution in the new selective regime,
-\[
-  \phi_+(a) = \sum_k p_k a^k .
-\]
-Now, $p_s$ is the minimal solution to $\phi_+(a) = a$.
-
-\section{Notes on branching process approximation}
-
-The branching process approximation will be good as long as the probability of "interactions" are small,
-i.e.\ we can couple the branching process with the actual density--regulated model.
-Let $Z_t(x)$ be the numbers of selected individuals in deme $x$.
-If we assume that $Z_t$ is smooth enough, 
-the numbers of nearby individuals to interact with is approximately $Z_t(x) \sigma^d$,
-and the total number of individuals is about $\rho \sigma^d$,
-so the rate of interactions is about $(Z_t(x) \sigma^d)^2/(\rho \sigma^d)$.
-Now if $Z$ is a spatial branching process beginning from a single individual at $x=0$ at $t=0$, 
-and growing at rate $r$, then 
-\[ 
-    Z_t(x) \approx \exp(rt - \|x\|^2/(2\sigma^2 t))/ (2 \pi \sigma^2)^{d/2} 
-        \le e^{rt} / ( \sigma^2 t)^{d/2}
-\]
-and so we need $t$ such that
-\[
-    e^{2rt} t^{-d} << \rho \sigma^{3d} .
-\]
-or equivalently,
-\[
-    e^{2rt/d} t^{-1} << \rho^{1/d} \sigma^3 .
-\]
-This will be true if $t$ is of smaller order than the logarithm of the effective population density, $\log \rho \sigma^d$.
-
-\subsection{On the random delay}
-
-Let $Y_t$ be the total number of offspring coming from a single individual
-(in the spatial, density-regulated model).
-Then we know that $(Y_t/t)^{1/d} \to v$, 
-but we saw from simulations that it looked like $Y_t^{1/d} \approx v(t-\tau)$,
-where $\tau$ is the (random) establishment time.
-(It would be useful to look at the coefficient of variation of $\tau$.)
-Well, we know that at first, $Y_t \approx W e^{rt}$, 
-where $W= \lim e^{-rt} Z_t$, and $Z_t$ is the associated branching process,
-and $r$ is the growth rate (er, $r=s$?).
-And eventually, $\partial_t Y_t^{1/d} \approx v$.
-If this switchover happened at a fixed size $C$,
-the size at which wavelike behavior took over,
-i.e.\ when $W e^{rt} = C$,
-then for $\tau = \frac{1}{r}\log(C/W)$ we'd have 
-\[
-    Y_t = \begin{cases}
-        W e^{rt} \quad & t \le \tau \\
-        (v (t-\tau))^d + C \quad & t > \tau ,
-\end{cases}
-\]
-which if $C$ is small relative to the scale we look on will look as described above.
-So the random delay is $(1/r)\log(C/W)$; and so the mean delay is something like $(1/r)\log C$
-and the randomness is determined by $(1/r)\log (1/W)$.
-
-What about how much variance is left in the branching process?
-Well, since $Z_{t+s} = \sum_{k=1}^{Z_t} Z^k_{s}$, where $Z^k_s$ are iid copies of $Z_s$,
-we have that $\var[Z_{t+s}|Z_t] = Z_t \var[Z_s]$,
-and so the coefficient of variation of $Z_{t+s}$ given $Z_t$ is
-\[
-   \mathop{CV}[Z_{t+s}|Z_t] &= \frac{\sqrt{Z_t \var[Z_s]}}{ Z_t \E[Z_s] } \\
-        &= e^{-rs} \sqrt{ \var[Z_s] } Z^{-1/2} \\
-        &\simeq e^{-rt/2} \sqrt{ \var[W] } .
-\]
-
-\section{Properties of the mixed model}
-
-Suppose that there are (point masses of) mutations present at time 0 as a Poisson process of rate $\lambda_0$ (calculated above),
-and that after time 0 mutations arise at rate $\lambda$.
-Suppose that the mutations spread distance $f(t)$ in time $t$, 
-and that the volume of the cone of height $t$ is given by $h(t) = \int_0^t \omega_d f(s)^d ds$.
-
-Let $\tau$ denote the time until the origin is adapted.
-Then
-\begin{align}
-   \P \{ \tau > t \} &= \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\}, \quad \mbox{and} \\
-   \E [ \tau ] &= \int_0^\infty \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} .
-\end{align}
-
-Let $z_0$ be the mean proportion of space covered by mutations present at time 0.
-Then, by the general result on competing inhomogeneous exponentials,
-\begin{align}
-    z_0 &= \P\{ \mbox{origin reached by mutations from $\lambda_0$ before $\lambda$} \} \\
-        &= \int_0^\infty \lambda_0 d \omega_d f'(t) f(t)^{d-1} \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} dt .
-\end{align}
-The mean density of new mutations that would have taken off but were prevented by standing variation is
-\begin{align}
-    \nu_* = \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt
-\end{align}
-and so the mean proportion of new mutations that would be successful in the absence of standing variation, but are prevented by standing variation,
-is 
-\begin{align}
-    z_* = \frac{ \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt }
-                { \int_0^\infty \lambda \exp\{ - \lambda h(t) \} dt } .
-\end{align}
-
-Let $X$ be the distance to the eventually enclosing mutation and $R=f(\tau)$ (see first paper).
-Then if $g(t) = \lambda_0 \omega_d f(t)^d + \lambda h(t)$ then $(X,R)$ has density
-\[
-    d \omega_d x^{d-1} \exp(-g(f^{-1}(r)) dx dr,
-\]
-so integrating over $r$,
-\[
-    \P\{ X \in dx \} = dx d \omega_d x^{d-1} \int_x^\infty \exp( - g(f^{-1}(r)) ) dr ,
-\]
-and
-\[
-    \E[ X^n ] = \frac{d\omega_d}{n+d} \int_0^\infty r^{n+d} \exp( - g(f^{-1}(r)) ) dr .
-\]
-
-If we rescale time by $\alpha$ and space by $\beta$, 
-then this transforms $(\lambda_0, \lambda, f(\cdot))$ into $(\lambda_0/\beta^d, \lambda/(\alpha \beta^d), f(\cdot/\alpha)/\beta)$.  
-Setting $\beta = \lambda_0^{1/d}$ and $\alpha = \lambda / \lambda_0$, this is mapped to $(1,1,f(\cdot \lambda_0/\lambda)/\lambda_0^{1/d})$.
-
-\subsection{Constant speed}
-
-Now let $f(t) = vt$, so $h(t) = \frac{ \omega_d v^d }{ d+1 } t^{d+1}$.
-The integrals that appear above are of the following form:
-\begin{align}
-  \int_0^\infty t^c \exp \left( - \alpha t^a - \beta t^b \right) dt 
-            &= \left( b^{-1} \beta^{ (1-c)/b } \right) \int_0^\infty u^{(c+1-b)/b} \exp\left( - \alpha \beta^{-a/b} u^{a/b} - u \right) du ,
-\end{align}
-so it suffices to learn something about the function
-\begin{equation}
-    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt .
-\end{equation}
-We will only be interested in this as a function of $x$; 
-the coefficients $a$ and $b$ will merely depend on the dimension and what quantity we are calculating.
-We can construct a power series for $G$ in $x$:
-\begin{align}
-    \partial_x^n G(a,b,x) &= \int_0^\infty (-1)^n t^{a+nb} \exp\left( -x t^b - t \right) dt \\
-            &= (-1)^n G(a+nb,b,x) , \quad \mbox{so} \\
-    \partial_x^n G(a,b,x) \vert_{x=0} &= \Gamma(a+nb+1) .
-\end{align}
-So, if everything works out, we will have that
-\begin{align}
-    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
-\end{align}
-Something simliar is noted at the Wikipedia page on the Gaussian integral, and claims such integrals show up in quantum field theory.
-
-A few facts, first.
-Using repeatedly the identity $\Gamma(x+1) = x\Gamma(x)$
-and the Pochammer symbol $(x)_n = x(x+1)\cdots(x+n-1)$,
-note that for $m \in \N$,
-\[
-    \Gamma(m+y) = (y)_m \Gamma(y).
-\]
-Also note that for $c$ and $m$ integers,
-\[
-    (y)_{cm} = c^{cm} \prod_{k=0}^{c-1} (\frac{y+k}{c})_{m} ,
-\]
-and in particular, that
-\[
-    (cm+\ell)! = \ell! (\ell+1)_{cm} = \ell! c^{cm} \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m .
-\]
-
-Now suppose that $b$ is rational (as will be the case), letting $b = b'/c$ with $b'$ and $c$ positive integers.
-Therefore, setting $n = mc+\ell$,
-\begin{align}
-    G(a,b,x) &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} \Gamma(a+\ell b'/c+mb'+1) \\
-            &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} (a+\ell b')_{mb'} \Gamma(a + \ell b') \\
-            &= \sum_{\ell=0}^{c-1} \frac{ (-x)^{\ell} }{ \ell! } \Gamma(a + \ell b') \sum_{m \ge 0} ( (-cx)^c / (b')^{b'} )^m 
-                    \frac{ \prod_{j=0}^{b'-1} (\frac{ a+b'\ell+j }{ b' } )_m }{ \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m } ,
-\end{align}
-which is the sum of $c$ (generalized) hypergeometric series.
-For these to converge, we need $b' \le c$, e.g.\ $b\le 1$.
-
-In particular, if we let $A=\lambda_0 \omega_d v^d $ and $B=\lambda h(1)$, then 
-\begin{align}
-    \E[\tau] &= \int_0^\infty e^{-A t^d - B t^{d+1}} dt \\
-            &= \frac{B^{1/(d+1)}}{d+1} \int_0^\infty u^{-d/(d+1))} \exp\left( -AB^{-d/(d+1))} u^{d/(d+1)} -u \right) du \\
-            &= \frac{B^{1/(d+1)}}{d+1} G \left(-d/(d+1), d/(d+1), A B^{-d/(d+1)} \right) .
-\end{align}
-So, if $d=1$, then we have {\bf (verify this by usual methods!)}, with $x = AB^{-1/2}$,
-\begin{align}
-    \E[\tau] &= \frac{\lambda^{1/(2)}}{2} G \left( -1/2, 1/2, x \right) \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) \sum_{m\ge0} (4x^2)^m \frac{ (-1/2)_m }{ (1/2)_m (1)_m } 
-                        - x \Gamma(1/2) \sum_{m\ge0} (4x^2)^m \frac{ (1/2)_m }{ (1)_m (3/2)_m }  \right\} \\
-            &=  \frac{\lambda^{1/(2)}}{2} \left\{ \Gamma(-1/2) {}_1F_1(-1/2;1/2;4x^2) - x \Gamma(1/2) {}_1F_1(1/2;3/2;4x^2) \right\} \\
-            &= \frac{\lambda^{1/(2)}}{2} \left\{ - \Gamma(-1/2) (1/x) \gamma(-1/2,4x^2) - \Gamma(1/2) (1/x) (\sqrt{\pi}/8) \erf(4x^2) \right\} .
-\end{align}
-
-And in $d=2$, since
-\begin{align}
-    G(-2/3,2/3,x) &= \Gamma(-2/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (-1/3)_m (1/6)_m }{ (1/3)_m (2/3)_m (1)_m }
-                    - x \Gamma(4/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (2/3)_m (7/6)_m }{ (2/3)_m (1)_m (4/3)_m }
-                        + \frac{1}{2} x^2 \Gamma(10/3) \sum_{m\ge0} (\frac{-27 x^3}{4})^m \frac{ (5/3)_m (13/6)_m }{ (2)_m (4/3)_m (5/3)_m } \\
-                &= \Gamma(-2/3) {}_2F_2(-1/3,1/6;1/3,2/3;-\frac{27x^2}{4})
-                    -x \Gamma(4/3) {}_1F_1(7/6;4/3;-\frac{27x^2}{4})
-                       -\frac{1}{2}x^2 \Gamma(10/3) {}_1F_1(13/6;4/3;-\frac{27x^2}{4}) .
-\end{align}
-
-\section{Characteristic lengths}
-
-How to pick appropriate rule-of-thumb parameters?
-Recall that before we had $\chi$ being the smallest solution to
-\[
-    f\left( \frac{1}{\lambda \chi^d \omega_d} \right) = \chi.
-\]
-Now, suppose we draw a cylinder in space-time with radius $\chi$ and height $t=f^{-1}(\chi)$;
-as before we choose $\chi$ so that the mean number of points in this cylinder is unity.
-Then
-\[
-    1 = \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d ,
-\]
-and so we choose $\chi$ to be the unique positive solution to 
-\[
-    f\left( \frac{1 - \lambda_0 \chi^d \omega_d}{\lambda \chi^d \omega_d} \right) = \chi.
-\]
-
-As for the proportion of mutations coming from standing variation,
-it seems a natural measure would be the mean number of such mutations in such a cylinder,
-which is using the definition of $\chi$,
-\begin{align*}
-    \pi_0 &= \frac{\lambda_0 \omega_d \chi^d }{ \lambda_0 \omega_d \chi^d + \lambda \omega_d f^{-1}(\chi) \chi^d} \\
-        & = \frac{ 1 }{ 1 + \lambda f^{-1}(\chi) / \lambda_0 } \\
-        &= \lambda_0 \omega_d \chi^d .
-\end{align*}
-
-Now, if $f(t) = vt$, then $f^{-1}(r) = r/v$,
-and so we have $\chi$ the positive root of
-\[
-  \omega_d \chi^d (\lambda_0 + \lambda v \chi) = 1 .
-\]
-If $d=1$ then this is
-\[
-    \chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v } .
-\]
-In $d=2$ ??
-
+\input{Spatial_adaptation/preamble}
 \section{General relationship between various quantities}
@@
 
-\section{Poisson--Vorono\u\i\ facts}
-
-Fill in some facts about PV tesselations.
-
 \end{document}

### 2011-05-03 standing-patches-paper.tex
fixed char len formula but need still fix discussion
fixed ch len discussion

@@
 the mutant type has fitness $1+s_b$,
-where $0<s_d<1$, and $s_b>0$ will usually be assumed to be small.
+where $s_b>0$ will usually be assumed to be small, and $0<s_d<1$.
 We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
@@
 \end{align*}
-Now for $s_b$ small, and using $p_s \approx 2s_b$,
-we know that $\zeta(1-2s_b,T) \approx 1-2s_b \E[Z_T] = 1-2s_b e^{-Ts_d}$,
+Now for $s_b$ small, using $p_s \approx 2s_b$ and that $\E[Z_T]=(1-s_d)^T$,
+we know that $\zeta(1-2s_b,T) \approx 1-2s_b \E[Z_T] = 1-2s_b (1-s_d)^T$,
 resulting in the approximation
 \begin{align}
-    \lambda_0 &= \mu \rho \int_0^\infty 2s_b e^{-s_d T} dT \\
-        &= \frac{ 2 \mu \rho s_b }{ s_d } .
+    \lambda_0 &= \mu \rho \int_0^\infty 2s_b (1-s_d)^{T} dT \\
+        &= \frac{ 2 \mu \rho s_b }{ -\log(1-s_d) } .
 \end{align}
@@
 \begin{equation} \label{eqn:defines_chi}
-%    s_d^{-1} (\sqrt{2s_b} \chi)^2 + \sigma (\sqrt{2s_b} \chi)^3 = (\pi \rho \mu)^{-1} .
-\left( \chi s_d \sigma \sqrt{2s_b} \right)^2 + \left( \chi s_d \sigma \sqrt{2s_b} \right)^3 
-        = \frac{ s_d^3 \sigma^2 }{ \pi \rho \mu } .
+%    \left( \frac{ - \chi \log(1-s_d) }{ 2 s_b \sigma } \right)^2 + \left( \frac{ - \chi \log(1-s_d) }{ 2 s_b \sigma } \right)^3 
+%        = \frac{ - \log(1-s_d)^3 }{ 4 s_b^2 \sigma^2 \pi \rho \mu } .
+   \frac{\sqrt{2s_b} }{-\log(1-s_d) } \chi^2 + \frac{1}{\sigma} \chi^3 = \frac{1}{\sqrt{2s_b} \rho\mu\pi}
 \end{equation}
-From this we see that $\chi$ increases linearly with $1/\sqrt{s_b}$,
-and decreases with $\rho$ and $\mu$.
-Furthermore, the relative sizes of $\sigma$ and $1/s_d$ determine the relative contributions of new and standing variants, respectively.
-For instance, if $\sigma^3 s_d^3/(\pi \rho \mu)$ is small, then $\chi \approx \sqrt{ s_d / (2 s_b \pi \rho \mu)} = 1/\sqrt{\pi \lambda_0}$,
-% + O( \gamma ) with $\gamma = \sigma^3 s_d^3/(\pi \rho \mu)$
-and nearly all mutants come from standing variation.
-On the other hand, if $\sigma^3 s_d^3/(\pi \rho \mu)$ is large, then $\chi \approx (2s_b)^{-1/2} \left( \sigma \pi \rho \mu)\right)^{-1/3}$,
-% + O( 1 ) .
-the characteristic length from \citet{ralphcoop2010},
-in which case standing variation is relatively unimportant.
-% More generally, as $\sigma s_d$ goes from $0$ to $\infty$, the positive solution to \eqref{eqn:defines_chi} moves continuously between the two.
+From this we see that $\chi$ decreases with $\rho$ and $\mu$.
+Furthermore, for large $\sigma$, the characteristic length is close to the value obtained just from standing variation:
+$\chi = \sqrt( -\log(s_d) / 2 s_b \rho \mu \pi ) + O(1/\sigma)$.
+On the other hand, if the mutant allele is highly deleterious before $t=0$,
+then the characteristic length is close to the value from \citet{ralphcoop2010}:
+$\chi = ( \sigma / \rho \mu \pi \sqrt{2 s_b} )^{1/3} + O(1/\log(1-s_d))$.
 

### 2011-05-03 Spatial_adaptation/standing-variation-fns.R
fixed sd -> log(1/(1-sd))

@@ -9,14 +9,14 @@
 # 
 # ... and
 # lambda <- 2 * mu * rho * sb    # is rate of new successful mutations
-# lambda0 <- lambda / sd         # is approximate rate of standing successful mutations
+# lambda0 <- lambda / log(1/(1-sd))         # is approximate rate of standing successful mutations
 # v <- sigma * sqrt(2*sb)        # speed of wave of advance
 
 meanTime <- function (mu, rho, sb, sd, sigma) {
     # mean time til adaptation of a point;
     # E[\tau] is int_0^\infty exp(-lambda0 pi v^2 t^2 - lambda pi v^2 t^3) dt
     f <- function (t) {
-        exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/sd + t) )
+        exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/log(1/(1-sd)) + t) )
     }
     return( integrate(f, 0, Inf) )
 }
@@ -33,16 +33,16 @@ standingProportionArea <- function (mu, rho, sb, sd, sigma) {
     # proportion of space arising from standing variation
     # is 2 * lambda0 * pi * v * t * exp( - 2 * lambda0 * pi * v * t - lambda * pi * v^2 * t^2 )
     f <- function (t) {
-        (1/sd) * 4 * sqrt(2) * mu * rho * sb^(3/2) * pi * t * exp( - 4 * mu * rho * sb^(3/2) * sigma * pi * t ( sqrt(2) / sd + sigma * t ) )
+        (1/log(1/(1-sd))) * 4 * sqrt(2) * mu * rho * sb^(3/2) * pi * t * exp( - 4 * mu * rho * sb^(3/2) * sigma * pi * t ( sqrt(2) / log(1/(1-sd)) + sigma * t ) )
     }
     return( integrate(f, 0, Inf) )
 }
 
 standingProportionNumbers <- function (mu, rho, sb, sd, sigma) {
     # proportion of patches arising from standing variation
-    #    is 1/(1+sd * E[\tau])
+    #    is 1/(1+log(1/(1-sd)) * E[\tau])
     x <- meanTime(mu, rho, sb, sd, sigma)
-    x$value <- 1 / (1 + sd * x$value)
+    x$value <- 1 / (1 + log(1/(1-sd)) * x$value)
     return( x )
 }
 
[…]

### 2011-05-05 Spatial_adaptation/standing-variation-plots.R
fixed mean proportion formula

@@ -44,7 +44,7 @@ sb <- .05
 sigmavals <- c(0.1,1,10,20,50,100)
 sigmacols <- rainbow(length(sigmavals)+5)[1:length(sigmavals)]
 sigmalabs <- c(expression(sigma==0.1), expression(sigma==1.0), expression(sigma==10), expression(sigma==20), expression(sigma==50), expression(sigma==100))
-sdvals <- (1:19)/200
+sdvals <- 1-exp(-(1:19)/20)
 otherlabs <- expression(paste( mu == 10^-8, "  ", s[b] == .05, "  ", rho == 25 ))
 
 # plot characteristic length as a function of sd

### 2011-05-08 standing-patches-paper.tex
more fixes
fixed mean proportion formula
minor fix

@@
 $\nu_+ = \lambda \E[\tau] $.
-Using that $\lambda_0 = \lambda / s_d$, this gives us that
+Using that $\lambda_0 = \lambda / \log(1/(1-s_d))$, this gives us that
 \begin{equation}
-    \nu_+ = (\mbox{mean density of new patches}) = \frac{1}{1+s_d \E[\tau]} .
+    \nu_+ = (\mbox{mean density of new patches}) = \frac{1}{1-\log(1-s_d) \E[\tau]} .
 \end{equation}
@@
 the probability it will be reached by time $t+dt$ 
-by a standing variant is approximately $2 \lambda_0 \pi vt dt$, 
+by a standing variant is approximately $2 \lambda_0 \pi v^2 t dt$, 
 and the probability it is reached by a new variant is $\lambda \pi v^2 t^2 dt$.
@@
 \begin{equation}
-    z_0 = \int_0^\infty {2 \lambda_0 \pi v t} \; \exp \left( - 2 \lambda_0 \pi v t - \lambda \pi v^2 t^2 \right) dt .
+    z_0 = \int_0^\infty {2 \lambda_0 \pi v^2 t} \; \exp \left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) dt .
 \end{equation}

### 2011-05-24 Spatial_adaptation/panmixia.tex
WorkingWiki importation

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}
\begin{document}

As I understood it, Dmitri's intuition (deriving from simulations) says that
in real populations, the plausibility of parallel adaptation 
is determined by the size of the mutational influx $2N\mu$,
and that additional opportunity for parallel adaptation caused by spatial structure is relatively unimportant,
at least as far as the plausibility (not necessarily the degree) goes.
This might be because: if $2N\mu \gg 1$,
then there are many successful new mutants {\em every generation},
and so even in the fairly brief time it takes a single mutation to fix locally,
new mutations will arise and fix locally.
In this case, we get parallel adaptation without invoking spatial structure.
Conversely, if $2N\mu \ll 1$,
then spatial spread allows more time for new mutants, 
but since fixation is still fast, the additional amount of time is still insignificant.

Clearly, there are parameter ranges where the above holds, and does not hold,
but the question is how restrictive are those conditions.
I am turning this into three questions:
\begin{enumerate}
  \item When will other new mutations arise in the amount of time it takes, say, the first mutation to fix, locally?
  \item When does spatial structure significantly increase the probability of parallel adaptation?  
  \item When does spatial structure significantly increase the degree of parallel adaptation? 
    (i.e.\ number of origins)
\end{enumerate}
Throughout I'll work in two dimensions.
Let $\rho$ be the population density, 
$\sigma$ be the dispersal distance, 
$\mu$ be the mutation rate,
$s$ be the selection coefficient,
$N$ be the total population size,
$A=N/\rho$ be the total area occupied,
and $r = \sqrt{A}/\sigma$.
I assume that $r$ is not small, since otherwise the problem is not spatial.

The first question is something we did not include explicitly in our first paper.
The area occupied by an advantageous mutation grows quadratically, 
but only after a (random) initial period of local fixation.
We can call ``local fixation'' 
the time when the growth of a panmictic and a spatially restricted family
begins to differ significantly,
which occurs when the number of individuals is $\sigma^2 \rho$.
This time is thus approximately
\[
  T_l = \frac{1}{s} \log (\sigma^2 \rho) .
\]
(The standard deviation of this time is of the same order; let's ignore the randomness.)
On the other hand, the time we have to wait to see a new, successful mutation is on the order of
\[
  T_n = \frac{1}{4N\mu s} .
\]
Therefore, we will see new mutations in the time it takes a new mutation to fix locally
(and hence, the time it takes a wave to get going)
if
\[
    T_l > T_n ,
\]
or since $\rho = N/A$,
\begin{equation} \label{eqn:tlggtn}
  2 \log(\sigma / \sqrt{A}) + \log(N) > \frac{1}{4\mu N} .
\end{equation}
A reasonable range for $r=\sqrt{A}/\sigma$ is $10 \le r \le 1000$,
over which range $4 < 2\log(r) < 14$, so $\log(N)$ will dominate the inequality.
In other words, to get new mutations in the time it takes to locally fix,
we'd need $1/(4 N \log N) < \mu$.
Setting $\mu=10^{-8}$ tells us that we get multiple single base mutations in this time 
if $N$ is larger than about a million.

For the second and third questions,
we need to compare the panmictic case to the spatial case.
In the panmictic case, the offspring of a family continues to grow logistically,
so the amount of time it takes to go from frequency $\sigma^2 \rho / N$ to frequency $1/2$ 
is approximately
\[
  T_p = \frac{1}{s} \log (N/\sigma^2 \rho) = \frac{1}{s} \log(A/\sigma^2) .
\]
On the other hand, the spatially restricted family will grow at speed $v=\sigma \sqrt{s}$, 
so the numbers grow like
\[
  \rho \sigma^2 s \left(t - s^{-1/2} \right)^2,
\]
with time measured in units after local fixation.
Therefore, the additional time it takes to reach frequency $1/2$ is 
\[
T_s = \sqrt{\frac{N}{2\sigma^2 s \rho}} + \sqrt{\frac{\rho}{s}} .
\]
This will give us significant extra time if $T_s \gg T_p$, i.e.\ if
\begin{equation} \label{eqn:tsggtp}
  % \sqrt{s}\left( 1 + \sqrt{\frac{A}{2\sigma^2}} \right) \gg \log(A/\sigma^2) . %exact
  \sqrt{\frac{s}{2}} \gg \frac{ \log(A/\sigma^2) }{ \sqrt{A/\sigma^2} }. %approx
\end{equation}
This extra time will be enough for some mutations if $T_s > T_n$, or
\begin{equation}
  \sqrt{\frac{sA}{2\sigma^2}} > \frac{1}{4N\mu} .
\end{equation}

Looking again at equation \eqref{eqn:tsggtp},
the spatial restriction will give much more time if
\[
  s \gg 8 \left( \frac{\log(r)}{r} \right)^2,
\]
where $r=\sqrt{A/\sigma^2}$ is the radius of the species range in units of dispersal distance.
Here is a plot of $s = 8 \left( \log(r)/r \right)^2$,
for $10 \le r \le 1000$, giving the lower limit for possible values of $s$.
At smaller species ranges, this is unlikely to happen, 
but for species ranges hundreds of times the dispersal distance, it starts to become reasonable.
Humans and Drosophila are on the edge -- if $\sigma=10$km,
then the spatial distribution is giving extra time; 
but if $\sigma=100$km, perhaps it is not.

% \begin{figure}[hht]
  \includegraphics[width=1.0\textwidth]{min-s-versus-r.pdf}
% \caption{Plot of $s= 8 \left( \log(r)/r \right)^2$ against $r=\sqrt{A/\sigma^2}$.} 
% \end{figure}

To answer the second question, 
to expect parallel adaptation in the spatial case but not in the panmictic case,
we would clearly need $T_l \ll T_n$, so no new mutations are arising in the time to local fixation,
since this part of the process is shared by both.
Recall that this effectively puts an upper bound on $N$.
We'd also need
$T_s > T_n \gg T_p$,
or
\begin{equation}
  \sqrt{\frac{sA}{2\sigma^2}} > \frac{1}{4N\mu} \gg \log(A/\sigma^2) ,
\end{equation}
or
\begin{equation}
  % \sqrt{s/2} > \frac{1}{4 N \mu r} \gg 2 \frac{ \log(r) }{ r } .
  s > \frac{1}{8 N^2 \mu^2 r^2} \gg 8 \left(\frac{ \log(r) }{ r }\right)^2 .
\end{equation}
This puts an upper and lower bound on the mutational influx $4N\mu$:
taking the middle values in the figure, 
this will be something like $1/20 < 4N\mu < 800$.

For the third question,
to get an idea we can use the results on standing variation,
and treat the mutations arising while the first is fixing, if such exist, as ``standing variation''.
This means setting 
\[
  \lambda_0 = 4 \rho \mu s T_l = 4 \rho \mu \log(\sigma^2 \rho) .
\]
Recall the characteristic length solves $\lambda_0 \pi \chi^2 + \lambda \pi \chi^3/v = 1$.
Set $\chi_0 = 1/\sqrt{\pi \lambda_0}$, 
the characteristic length only from the mutations arising while the first fixes locally;
we are interested in the numbers of mutations arising relative to these,
which is about $x=(\chi/\chi_0)^2$.
This proportion is between zero and one and satisfies
$x = 1-\gamma x^{3/2}$, 
where 
\[
  \gamma = \frac{ \sqrt{s} }{ \sigma \sqrt{4 \pi \rho \mu} \log^{3/2}(\sigma^2 \rho) } .
\]
The relationship between $\gamma$ and $f$ is shown in this figure:

\includegraphics[width=1.0\textwidth]{f-versus-gamma.pdf}

showing that if $\gamma>1$ then $f$ drops off quickly.
If we then fix $\mu=10^{-8}$ and $\rho=1$ (working in units of $1/\sqrt{\rho}$), 
then we get the following picture of $\log \gamma$:

\includegraphics[width=1.0\textwidth]{gamma-contour-1e-8.pdf}

showing that $\gamma$ is mostly affected by $\sigma$,
needing something like $\sigma<20/\sqrt{\rho}$.
If $\mu=10^{-6}$ this drops to something like $2/\sqrt{\rho}$.

\paragraph{In summary,}
the conclusions are:
\begin{enumerate}

  \item Multiple successful mutations in the time it takes a single mutation to fix locally
    are likely if $1/(4 N \log N) < \mu$; if $\mu=10^{-8}$ this means in populations larger than a million.

  \item Spatial structure can make parallel adaptation likely where it was unlikely under panmixia
    if the above does not happen and if
    $s > \frac{1}{8 N^2 \mu^2 r^2} \gg 8 \left(\frac{ \log(r) }{ r }\right)^2$;
    this latter restricts the range of mutational influxes, but not dramatically.

  \item Spatial structure can dramatically increase the number of parallel adaptations if $\sigma$ is small enough;
    the requirement is that $\sigma<C/\sqrt{\rho}$, for some $C$ around ten.

\end{enumerate}

It is easy to think of populations that satisfy any of these requirements;
some natural examples that come to mind seem to lie in the ambiguous region,
although others (e.g.\ {\it Drosophila}?) lie firmly on one side of the boundary or the other.

\end{document}

### 2011-05-24 panmixia.R
WorkingWiki removed source file ‘panmixia.R’

@@ -1,64 +1,0 @@
-panmictic <- function (t,s,N) {
-    # logistic growth beginning from 1/N
-    1/(1 + (N-1)*exp(-s*t))
-}
-
-panmicticinv <- function (x,s,N) {
-    # inverse of logistic function above:
-    # panmictic( panmicticinv(x,s,N), s, N) = x
-    log( N*x*(1-1/N)/(1-x) ) / s
-}
-
-quadratic <- function (t,s,N,sigma,rho) {
-    # as panmictic until rho*sigma^2/N;
-    # as pi*(v*t)^2 after that
-    # with v = sigma * sqrt(s)
-    t0 <- panmicticinv(rho*sigma^2/N,s,N)
-    # needs to be vectorized to be able to integrate
-    ans <- panmictic(t,s,N)
-    ans[t>t0] <- pmin( 1, (pi*rho*sigma^2*s/N) * (t-t0+1/sqrt(s))^2 )[t>t0]
-    return(ans)
-}
-
-N <- 1e8
-s <- .05
-sigma <- 20
-rho <- 3
-ttt <- 4*(1:100)*log(N)/(100*s)
-
-plot(ttt, sapply(ttt, function(t) panmictic(t,s,N)), type="l", ylim=c(0,1))
-lines(ttt, sapply(ttt, function(t) quadratic(t,s,N,sigma,rho)), col="red")
-
-integrate(function (t) {panmictic(t,s,N)-quadratic(t,s,N,sigma,rho)}, 0, Inf)
-
-ss <- sqrt(2)^(-(20:1)); 
-rr <- (1:1000)
-ssrr <- expand.grid(ss,rr)
-ddd <- apply(ssrr,1,function(x) { log( log(x[2])/(x[2]*sqrt(x[1]*2)) ) } )
-dim(ddd)<-c(length(ss),length(rr))
-contour(x=ss,y=rr,z=ddd,xlab="s",ylab="r")
-
-pdf(file="min-s-versus-r.pdf",width=6,height=4)
-rr <- 10:1000
-plot(rr,8*((log(rr)/rr)^2), type="l", xlab=expression(r==sqrt(A/sigma^2)), ylab=expression(s== 8 * ( log(r)/r )^2), log="xy")
-dev.off()
-
-NN <- exp(8:20)
-plot(NN, 1/(4*NN*log(NN)), type="l")
-
[…]

### 2011-05-24 Spatial_adaptation/panmixia.R
WorkingWiki replaced project description

@@ -1,0 +1,64 @@
+panmictic <- function (t,s,N) {
+    # logistic growth beginning from 1/N
+    1/(1 + (N-1)*exp(-s*t))
+}
+
+panmicticinv <- function (x,s,N) {
+    # inverse of logistic function above:
+    # panmictic( panmicticinv(x,s,N), s, N) = x
+    log( N*x*(1-1/N)/(1-x) ) / s
+}
+
+quadratic <- function (t,s,N,sigma,rho) {
+    # as panmictic until rho*sigma^2/N;
+    # as pi*(v*t)^2 after that
+    # with v = sigma * sqrt(s)
+    t0 <- panmicticinv(rho*sigma^2/N,s,N)
+    # needs to be vectorized to be able to integrate
+    ans <- panmictic(t,s,N)
+    ans[t>t0] <- pmin( 1, (pi*rho*sigma^2*s/N) * (t-t0+1/sqrt(s))^2 )[t>t0]
+    return(ans)
+}
+
+N <- 1e8
+s <- .05
+sigma <- 20
+rho <- 3
+ttt <- 4*(1:100)*log(N)/(100*s)
+
+plot(ttt, sapply(ttt, function(t) panmictic(t,s,N)), type="l", ylim=c(0,1))
+lines(ttt, sapply(ttt, function(t) quadratic(t,s,N,sigma,rho)), col="red")
+
+integrate(function (t) {panmictic(t,s,N)-quadratic(t,s,N,sigma,rho)}, 0, Inf)
+
+ss <- sqrt(2)^(-(20:1)); 
+rr <- (1:1000)
+ssrr <- expand.grid(ss,rr)
+ddd <- apply(ssrr,1,function(x) { log( log(x[2])/(x[2]*sqrt(x[1]*2)) ) } )
+dim(ddd)<-c(length(ss),length(rr))
+contour(x=ss,y=rr,z=ddd,xlab="s",ylab="r")
+
+pdf(file="min-s-versus-r.pdf",width=6,height=4)
+rr <- 10:1000
+plot(rr,8*((log(rr)/rr)^2), type="l", xlab=expression(r==sqrt(A/sigma^2)), ylab=expression(s== 8 * ( log(r)/r )^2), log="xy")
+dev.off()
+
+NN <- exp(8:20)
+plot(NN, 1/(4*NN*log(NN)), type="l")
+
[…]

### 2011-06-09 Spatial_adaptation/panmixia.R
pdf -> png

@@ -38,7 +38,8 @@ ddd <- apply(ssrr,1,function(x) { log( log(x[2])/(x[2]*sqrt(x[1]*2)) ) } )
 dim(ddd)<-c(length(ss),length(rr))
 contour(x=ss,y=rr,z=ddd,xlab="s",ylab="r")
 
-pdf(file="min-s-versus-r.pdf",width=6,height=4)
+png(file="min-s-versus-r.png",width=600,height=400)
+# pdf(file="min-s-versus-r.pdf",width=6,height=4)
 rr <- 10:1000
 plot(rr,8*((log(rr)/rr)^2), type="l", xlab=expression(r==sqrt(A/sigma^2)), ylab=expression(s== 8 * ( log(r)/r )^2), log="xy")
 dev.off()
@@ -46,7 +47,8 @@ dev.off()
 NN <- exp(8:20)
 plot(NN, 1/(4*NN*log(NN)), type="l")
 
-pdf(file="f-versus-gamma.pdf",width=6,height=4)
+png(file="f-versus-gamma.png",width=600,height=400)
+# pdf(file="f-versus-gamma.pdf",width=6,height=4)
 xx <- (1:999)/1000
 plot((1-xx^2)/xx^3, xx, type="l", log="xy", xlab=expression(gamma==(1-x)/x^(3/2)), ylab=expression(x==(chi/chi[0])^2) )
 dev.off()
@@ -58,7 +60,8 @@ sigsig <- (10:100)
 gg <- apply( expand.grid(ss,sigsig), 1, function (ssig) log( sqrt(ssig[1]) / ( ssig[2] * sqrt(4*pi*mu) * log(ssig[2]^2)^(3/2) ) ) )
 dim(gg) <- c(length(ss), length(sigsig))
 # pdf(file="gamma-contour-1e-8.pdf", width=6, height=4)
-pdf(file="gamma-contour-1e-8.pdf", width=6, height=4)
+png(file="gamma-contour-1e-8.png", width=600, height=400)
+# pdf(file="gamma-contour-1e-8.pdf", width=6, height=4)
 contour(x=ss,y=sigsig,z=gg, xlab="s", ylab=expression(sigma), main=expression(paste(log(gamma),", ", mu==10^-8)))
 dev.off()
 

### 2011-06-10 standing-patches-paper.tex
/* Paper */ added some migration text

@@
 
+\section{Local establishment}
+
+Include stuff from ``comparison to panmixia'' here.
+
 \section{Patchy selection}
@@
 
-patches where selection is $1+s_b$ separated by $1-s_d$ -- separation large relative to migration.   
-(or relative to $s_d/m$?)
+In this section, we suppose that instead of a uniform habitat,
+that the selective pressure is {\em patchy}, so that the new alleles are only advantageous in isolated locations, 
+and are disadvantageous in the intervening region.
+As before, we assume that the new allele has fitness $1+s_b$ relative to the unmutated type in the ``good'' patches,
+and assume that it has fitness $1+s_m$ in the intervening areas, with $s_m<0$.
+To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
+and that the disadvantage $s_m$ ensures that on the relevant timescale,
+the allele is very unlikely to fix in the regions where it is at a disadvantage.
+With these assumptions as justification, we will move to a model of discrete demes exchanging migrants,
+where ``migration'' occurs through the rare, transient passage of the new allele through regions where it is disadvantageous.
+
+equal deme sizes?  general migration rates?
 
@@
 
+\subsection{Mutational influx}
+
+Consider first a single, isolated patch of area $A$ in which no new allele has appeared.
+\citet{barton1987} (or \citet{slatkin1981}) has studied this situation (in one dimension)
+and found that:
+mutations appearing further than a few multiples of $\sigma/\sqrt{2s_m}$ away from the patch have vanishingly small probability of establishment;
+mutations appearing within the patch have probability strictly less but approaching that $2s_b$ of establishment;
+and if the patch width is less than $2 \tan^{-1} (\sqrt{s_m})$, then local fixation does not occur (rather, is not stable).
+Let $p(x)$ be the probability that a new allele arising at location $x$ fixes.
+The function $p(x)$ can be found in terms of Jacobi elliptic functions,
+XXX as shown in the Appendix,
+but the expressions are particularly unhelpful.
+The total successful mutational influx per generation is then given by $\int \rho \mu p(x) dx$,
+and depends in a complicated way on the relationship of patch width and selection coefficients,
+but still scales linearly with the population mutational influx $\rho \mu$.
+Therefore, we define $p_f = \int p(x) dx$, and in a pinch will approximate this by $2 s_b A$,
+which will be a good approximation if the width $\sqrt{A}$ is large relative to $\sigma/\sqrt{2s_m}$.
+
+XXX We are interested in the case where a single mutation only takes over? XXX
+
+\subsection{Migration rates}
+
+Now suppose that there are two patches of area $A$ of good environment separated by distance $R$,
+and that the allele has begun to fix in the first, but has not yet appeared in the second.
+As above, any copy of the new allele that appears at location $x$ sufficiently near to the second patch 
+has probability approximately $p(x)$ of fixing locally in this patch,
+whether the allele is a new mutation or is descended from the occupants of the first patch.
+There will be a small number of alleles descended from those of the first patch present in the regions where they are disadvantageous,
+held at mutation--selection balance;
+the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX),
+although the equations are again not particularly helpful.
+However, we do know that if $q(r)$ is the expected proportion present at mutation--selection balance at distance $r$ from the first patch
+in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|s_m r / \sigma|)$,
+where $C$ is a constant depending on the structure of the populations and the selection coefficients.
+
+Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
+supposing that patches are large enough for local fixation to be stable.
+If the alleles were not advantageous in the second patch,
+the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |s_m R/\sigma| )$;
+so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |s_m R/\sigma|)^{-1}$.
+If $T_m$ is large, then we would expect successful migrants to be rare,
+so that if the second patch adapts through migration, it is likely to come from only a single migrant.
+In the converse situation, XXX more than one migrant? XXX
+
 Get migration and mutation rates from a homogeneous continuous model:
-The migration rate from patch $i$ to a patch at distance $r$ is $C_i \exp(-\gamma r \sqrt{s} /\sigma)$,
+The migration rate from patch $i$ to a patch at distance $r$ is $C_i \exp(-\gamma r s /\sigma)$,
 where $C_i$ depends on the size of the patch and (to a lesser degree?) the local structural details of the selective difference.

### 2011-06-13 standing-patches-paper.tex
/* Paper */ put in some discrete results from previous paper

@@
 
-In this section, we suppose that instead of a uniform habitat,
-that the selective pressure is {\em patchy}, so that the new alleles are only advantageous in isolated locations, 
+In this section, we will relax the assumption that the uniform is homogenous.
+The model we analyze in section \ref{ss:discretedemes} is a model of discrete demes exchanging rare migrants,
+but since the object of this paper is to work in continuous space,
+we will first in sections \ref{ss:patchymutation} and \ref{ss:patchymigration} describe how a model with {\em patchy} selection can be approximated by such a discrete model.
+
+For this purpose, suppose that instead of a uniform habitat,
+the selective pressure is such that the new alleles are only advantageous in isolated locations, 
 and are disadvantageous in the intervening region.
@@
 and that the disadvantage $s_m$ ensures that on the relevant timescale,
-the allele is very unlikely to fix in the regions where it is at a disadvantage.
+the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
 With these assumptions as justification, we will move to a model of discrete demes exchanging migrants,
@@
 
-Describe simple discrete model, for general migration rates.
-
 \subsection{Mutational influx}
+\label{ss:patchymutation}
 
 Consider first a single, isolated patch of area $A$ in which no new allele has appeared.
-\citet{barton1987} (or \citet{slatkin1981}) has studied this situation (in one dimension)
+\citet{barton1987} (or \citet{slatkin1981}?) has studied this situation (in one dimension)
 and found that:
@@
 \subsection{Migration rates}
+\label{ss:patchymigration}
 
-Now suppose that there are two patches of area $A$ of good environment separated by distance $R$,
+Now suppose that there are two patches, each of area $A$, of good environment separated by distance $R$,
 and that the allele has begun to fix in the first, but has not yet appeared in the second.
-As above, any copy of the new allele that appears at location $x$ sufficiently near to the second patch 
+As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
 has probability approximately $p(x)$ of fixing locally in this patch,
 whether the allele is a new mutation or is descended from the occupants of the first patch.
-There will be a small number of alleles descended from those of the first patch present in the regions where they are disadvantageous,
-held at mutation--selection balance;
-the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX),
+Due to mutation--selection balance there will be a small number of alleles descended from those of the first patch (``migrants'') 
+present in the regions where they are disadvantageous;
+the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX, citeXXX),
 although the equations are again not particularly helpful.
@@
 so that if the second patch adapts through migration, it is likely to come from only a single migrant.
-In the converse situation, XXX more than one migrant? XXX
+We leave aside the converse situation, for which there is likely to be more than one migrant
+that founds the population mutant individuals in the new patch.
 
-Get migration and mutation rates from a homogeneous continuous model:
-The migration rate from patch $i$ to a patch at distance $r$ is $C_i \exp(-\gamma r s /\sigma)$,
-where $C_i$ depends on the size of the patch and (to a lesser degree?) the local structural details of the selective difference.
-The mutation rate is more or less (size of patch) times $\mu$; 
-more accurately we can integrate using Barton assuming step selection.
+\subsection{ss:discretedemes}
 
-\subsection{Results}
+Suppose now that we have a set of $L$ discrete populations of individuals that may reproduce or migrate,
+and that migration and mutation are rare, so that the waiting times until either event is approximately exponential.
+Suppose that for each pair of populations, labeled $i$ and $j$, the mean number of migrants that travel from $i$ to $j$ per generation
+is $M_{ij}$ and that the mean number of new mutations appearing in population $i$ per generation is $\mu_i$.
+
+If we suppose that fixation occurs on a faster time scale than mutation or migration,
+so that it is very unlikely that two different migrants or mutants begin to fix locally in the same population,
+then we can treat the process of extinction or local fixation as instantaneous 
+(a limit previous studied by \cite{Slatkin:81}).
+In this limit we may also ignore swamping effects of inmigration of nonmutant alleles.
+
+In this case, we have the following picture.
+The population at first starts out with no selected alleles. 
+At some later time point suppose that there are $T$ different alleles present in the population, 
+that the $k^\mathrm{th}$ allele is occupying demes $\{i^k_1, \ldots, i^k_{n_k}\}$, for $1\le k \le T$.
+Let $I$ denote the set of all adapted demes, 
+and let $J = \{j_1, \ldots, j_\ell\}$ denote the demes that have not yet adapted, with $\ell = L - \sum_k n_k$.
+Then we are assuming that two types of event can happen which could change the state:
+either at rate $\mu_j p_f$, an unadapted deme $j$ produces a new mutant, which fixes in this deme;
+or at rate $M_{ij} p_f$, an adapted deme $i$ sends a migrant to unadapted deme $j$, which fixes in that deme.
+Therefore, under these assumptions, the probability of fixation $p_f$ (and hence the selection coefficient)
+only enter as a time scaling: for instance,
+if we define $R = \sum_{j \in J} \left( \mu_j + \sum_{i \in I} M_{ij} \right)$,
+then the total rate of such events is $p_f R$.
+The probability that the next event results in, say, adaptation of deme $j$
+is 
+\[
+ \left( \mu_j + \sum_{i \in I} M_{ij} \right) / R,
+\]
+and the probability that deme $j$ adapts through migration from deme $i$, given that deme $j$ is the next to adapt, is
+$M_ij / R$,
+while the probability that it adapts through a new mutation is $\mu_j / R$.
+
+The main point here is that the final pattern,
+and hence the likely extent of parallel adaptation,
+is independent of the strength of selection.
+This is in contrast to the continuous case, however, this observation is dependent on the assumption of weak migration
+and fails to hold if different migrant lineages interact.
+This can be viewed as a continuous-time Markov process XXX.
+Explicit solutions are not available in the general case,
+but if we specialize to a completely symmetric model (i.e.\ the ``island'' model),
+then the mathematical picture is a pretty one.
+
+Indeed, suppose that migration and mutation are symmetric: $M_{ij}=m$ and $\mu_j = \mu$ for all $j$ and $i\neq j$.
+Suppose at some time the $k^\mathrm{th}$ allele is occupying $n_k$ demes, 
+with $n = \sum_k n_k < L$.
+Then at rate $(L-n) N \mu p_f + n (L-n) N m p_f$,
+either a migration or mutation event happens.
+With probability $\mu/(\mu + n m)$, it is a mutation,
+and a new deme is randomly chosen and assigned a new allele.
+With probability $n_k m / (\mu + n m)$,
+the $k^\mathrm{th}$ type sends a successful migrant to a randomly chosen new island,
+which is assigned the $k^\mathrm{th}$ allele.
+
+This process is a continuous-time version of the ``Chinese restaurant process''
+described in \citet{aldous1985exchangeability} and \citet{pitman1995partitions},
+and so the final partition of types across demes has the Ewens sampling distribution with parameter $\mu/m$.
+Again note that the selection coefficient, which is implicit in the probability $p_f$,
+does not enter into the final distribution.
+
+Now we can compute most properties we might want about the process.
+For instance, the expected number of distinct mutational origins is
+\begin{equation}
+    \E\left[ \mbox{ \# of independent mutations } \right] = 
+            1 + \frac{\mu}{m} \sum_{k=2}^L \frac{1}{k+(\mu/m)}. \label{discrete_expected}
+\end{equation}
+The probability that there is only a single successful mutation is
+\begin{equation}
+\P \left\{ \mbox{only one mutational origin} \right\} = 
+            \prod_{k=1}^{L-1} \frac{ k }{ k+(\mu/m)} .
+\end{equation}
+More generally, suppose we sample a single deme, and are interested in the total number of demes $S$
+(including the sampled one) that share the same fixed mutation as our sampled deme.
+Then $S$ has distribution
+\begin{equation} \label{eqn:Sdistrn}
+\P\{ S=s \} = \frac{ (s-1)! (\mu/m)^{L-s} }{ \prod_{k=1}^{L-1} (k+(\mu/m)) } .
+\end{equation}
+Furthermore, if $L$ is large, then $S/L$ has probability density approximately
+\begin{equation} \label{eqn:betadistrn}
+   (\mu/m) (1-x)^{(\mu/m)-1}
+\end{equation}
+namely, is a $\mathrm{Beta}((\mu/m), 1)$ distribution--- see \cite{donnelly-joyce} and \cite{permanPitmanYor92}.
+
+% In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
+% for a model with 20 interconnected demes, across different parameter values.  
+% This indicates that we need the population-scaled mutation and migration rates to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.
+If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
+
+The high connectedness of the discrete deme island model means that the expected number of distinct alleles, (Equation \eqref{discrete_expected}), 
+grows with the $\log$ of the number of demes.
+This strongly contrasts with the continuous spatial model where the local nature of dispersal means that doubling the species range will double the number of mutations expected.
+Furthermore, while in the continuous model the areas occupied by distinct mutations are similar in area, in this strong-selection discrete island model, 
+only a few types tend to dominate even as the number of demes (and thus, distinct types) grows large (see Equations \eqref{eqn:Sdistrn} and \eqref{eqn:betadistrn}).
+
+A more general model would allow migration only along edges of some graph connecting the demes.
+In the low migration limit such a model still produces a partition distribution independent of the selection coefficient,
+but it does not in general have the Ewens distribution.
+Another extension would be to include the time alleles need to achieve an intermediate frequency,
+along the lines of \citet{Navarro:03}, which would reintroduce dependence on the selection coefficient.
+
+\texttt{XXX omit this? XXX}
+We have treated the time during which the selected allele is at intermediate frequency in a deme (about $log(2N)/s$ generations) as negligible, 
+which ignores migration events that occur while the mutation is spreading through the population. 
+we now sketch of a more realistic approximation.
+For each occupied--unoccupied pair of demes, if the number of mutant alleles in the occupied deme is $n(t)>0$,
+then at time $t$, migrants from the occupied deme pass the selected allele to the unoccupied deme at rate $2 n(t) m p_f(s)$,
+while newly arisen mutants arise and fix in the unoccupied deme at constant rate $2 N \mu p_f(s)$.
+The selected type should grow approximately as $n(t)=\exp(s(t-t_0))$.
+Of course, to follow this model through, we will have to account for multiple types within a single deme and other complications,
+but we can at least make a few observations.
+One is that the final distribution of types is no longer independent of $s$, which enters through the dynamics $n(t)$.
+However, it is clear that increasing the selection coefficient $s$ will decrease the number of types (independent mutational origins),
+while on the other hand, increasing the deme size $N$ will increase the number of types,
+and that the case presented above is at the extreme (see also \cite{Navarro:03} for deterministic approximations).
+
+\subsection{Patchy Results} 
+\label{ss:discreteresults}
 
@@
 This occurs when the effective migration rate is much smaller than the mutation rate, namely...
+
+ \begin{figure}[ht]
+ \begin{center}
+ %\includegraphics{discrete-by-mutrate.eps}
+ %\includegraphics{discrete-by-migrprob.eps}
+ \includegraphics{discrete-by-ratio.eps}
+ \filenote{discrete-by-ratio.eps}
+ \caption{ %
+ The mean number of types and the mean number of demes occupied by a sampled type, in a discrete model with 20 demes,
+ plotted as a function of the ratio of migration to mutation probabilities.
+ Note the log scale on the horizontal axis.
+ \label{fig:discbyratio}
+ }
+ \end{center}
+ \end{figure}
+
+We have shown that in the model of discrete demes connected by weak migration, the final distribution of types depends only on the ratio of mutation to migration rates,
+and that the distribution itself has a Dirichlet distribution.  
+In particular, the number of independently arisen types follows the Ewens distribution, with parameter $\mu/m$.
+In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
+for a model with 20 interconnected demes, across different parameter values.  
+This indicates that we need the population-scaled mutation and migration rates to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.  
+If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
+
+Also note that the properties of the discrete model are independent of the {\em sizes} of the demes themselves--- we have assumed away all such dependence.
+This is complementary to the results of \cite{softsweepsII}, who showed that multiple mutations are likely to arise {\em within} a panmictic deme
+if the population-scaled mutation rate $2 N \mu$ is greater than $1$.
 

### 2011-06-14 standing-patches-paper.tex
/* Paper */ removed missing figure
/* Paper */ added some figures...

@@
 
-\subsection{Results}
+\subsection{Standing Variation -- Results}
 
@@
 \begin{center}
+\includegraphics[width=1.0\textwidth]{charlen-by-sd-limit}
+\caption{ %
+{\bf Characteristic length,} as a function of selective disadvantage, compared to the corresponding quantity without standing variation (e.g.\ from \cite{ralphcoop2010}) and to the quantity only considering standing variants (e.g.\ $\sqrt{\lambda_0}$).
+}
+\end{center}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
 \includegraphics[width=1.0\textwidth]{proportion-by-sigma.eps}
@@
 }
-\label{fig:cartoon}
 \end{center}
+\end{figure}
 
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=1.0\textwidth]{proportion-contour.png}
+\caption{ %
+{\bf Mean proportion of area occupied by standing variants,} as a function of dispersal distance $\sigma$ and selective disadvantage $s_d$.  XXX at what parameters. XXX
+}
+\end{center}
 \end{figure}
@@
 
-\subsection{ss:discretedemes}
+\subsection{Discrete populations}
+\label{ss:discretedemes}
 
@@
 
-\subsection{Patchy Results} 
+\subsection{Patchy Selection -- Results} 
 \label{ss:discreteresults}
@@
  %\includegraphics{discrete-by-migrprob.eps}
- \includegraphics{discrete-by-ratio.eps}
- \filenote{discrete-by-ratio.eps}
+ \includegraphics{discrete-by-ratio}
  \caption{ %

### 2011-06-15 standing-patches-paper.tex
/* Paper */ figure suffix
/* Paper */ figure
/* Paper */ added discussion of migration rates

@@
 \begin{center}
-\includegraphics[width=1.0\textwidth]{charlen-by-sd-limit}
+\includegraphics[width=1.0\textwidth]{charlen-by-sd-limit.eps}
 \caption{ %
@@
 \texttt{XXX omit this? XXX}
-We have treated the time during which the selected allele is at intermediate frequency in a deme (about $log(2N)/s$ generations) as negligible, 
+We have treated the time during which the selected allele is at intermediate frequency in a deme (about $\log(\sigma^2 \rho)/s_b$ generations) as negligible, 
 which ignores migration events that occur while the mutation is spreading through the population. 
@@
 For each occupied--unoccupied pair of demes, if the number of mutant alleles in the occupied deme is $n(t)>0$,
-then at time $t$, migrants from the occupied deme pass the selected allele to the unoccupied deme at rate $2 n(t) m p_f(s)$,
-while newly arisen mutants arise and fix in the unoccupied deme at constant rate $2 N \mu p_f(s)$.
+then at time $t$, migrants from the occupied deme pass the selected allele to the unoccupied deme at rate $2 n(t) m p_f(s_b)$,
+while newly arisen mutants arise and fix in the unoccupied deme at constant rate $2 N \mu p_f(s_b)$.
 The selected type should grow approximately as $n(t)=\exp(s(t-t_0))$.
@@
 but we can at least make a few observations.
-One is that the final distribution of types is no longer independent of $s$, which enters through the dynamics $n(t)$.
+One is that the final distribution of types is no longer independent of $s_b$, which enters through the dynamics $n(t)$.
 However, it is clear that increasing the selection coefficient $s$ will decrease the number of types (independent mutational origins),
@@
 
-Pull out results from other paper on discrete model:
-if migration rates are equal then we get Chinese Restaurant/Ewens;
-in any case is independent of selection.
+If environment is continuous, with separated patches where the new allele is advantageous,
+we can only get a rough sense of whether the new adaptation is likely to spread through mutation (i.e.\ in parallel)
+or through migration.
+Taking the expressions from sections \ref{ss:patchymutation} and \ref{ss:patchymigration} above,
+we see that if patches each have sufficiently large area $A$,
+then the mutational input in each is on the order of $2 \rho \mu A$,
+while the influx due to migration between patches separated by distance $R$ is on the order $2 \rho A \exp(- |s_m R/\sigma|)$. 
+If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|s_m R/\sigma| )$,
+then adaptation is expected to take place both by migration and mutation.
+If we take $\mu = 10^{-8}$, then this is satisfied if $R/\sigma \approx 18/s_m$.
+so if the disadvantage to the new mutation between patches is fairly strong (say, $s_m=.05$),
+then the distance between patches must be several hundreds of times the dispersal distance;
+while if $s_m$ is much smaller, say $s_m = .001$, 
+then the patches must be separated by tens of thousands of times the dispersal distance.
+The situation does not change much if $\mu$ is larger -- even if $\mu = 10^{-6}$, 
+the required separation between patches is only reduced by 60\%.
+If the separation between patches is much larger -- say, by a factor of ten -- 
+then patches are effectively isolated -- most adaptation will be independent between patches.
+On the other hand, if the separation is much smaller, then a single mutation is likely to fix everywhere.
+The general picture is that for species with spatial structure but without extreme barriers to dispersal, 
+patches will not be completely isolated, but some parallel adaptation is likely.
 
-when are patches effectively isolated?   
-This occurs when the effective migration rate is much smaller than the mutation rate, namely...
+Furthermore, in the model of discrete demes connected by symmetric weak migration, 
+the final distribution of types depends only on the ratio of mutation to migration rates,
+and that the number of independently arisen types follows the Ewens distribution, with parameter $\mu/m$.
+In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
+for a model with 20 interconnected demes, across different parameter values.  
+This indicates that we need the population-scaled mutation and migration rates 
+to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.  
+If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
 
@@
  \begin{center}
- %\includegraphics{discrete-by-mutrate.eps}
- %\includegraphics{discrete-by-migrprob.eps}
- \includegraphics{discrete-by-ratio}
+ \includegraphics{discrete-by-ratio.eps}
  \caption{ %
@@
 \end{figure}
-
-We have shown that in the model of discrete demes connected by weak migration, the final distribution of types depends only on the ratio of mutation to migration rates,
-and that the distribution itself has a Dirichlet distribution.  
-In particular, the number of independently arisen types follows the Ewens distribution, with parameter $\mu/m$.
-In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
-for a model with 20 interconnected demes, across different parameter values.  
-This indicates that we need the population-scaled mutation and migration rates to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.  
-If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
 

### 2011-06-15 Spatial_adaptation/local-adaptation.tex
/* Local Adaptation */ oops missing N
/* Local Adaptation */ filled in some about asymptotics

@@
 \usepackage{amsmath,amssymb}
+\usepackage{hyperref}
 
@@
 \begin{align}
-    \partial_t \phi(t,x) &\approx \sum_y s(x) \kappa(y-x) (1-\one_{x=y}) \phi(t,y)(1-\phi(t,x)) 
+    \partial_t \phi(t,x) &\approx \sum_y s(x) \kappa(y-x) (1-\frac{1}{N}\one_{x=y}) \phi(t,y)(1-\phi(t,x)) 
             + \sum_y \kappa(y-x) (1-\frac{1}{N}\one_{x=y}) (\phi(t,x)-\phi(t,y))  \\
@@
 Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
-so that (setting $\sigma^2=2$)
+so that 
 \[
-    \partial_t p(t,x) = \partial_x^2 p(t,x) + s(x) p(t,x) (1-p(t,x)) .
+    \partial_t p(t,x) = \sigma^2 \partial_x^2 p(t,x) + s(x) p(t,x) (1-p(t,x)) .
 \]
@@
 \begin{align} \label{eqn:definingphi}
-    \partial_x^2 \phi(x) = - s(x) \phi(x) (1-\phi(x)) ,
+    \partial_x^2 \phi(x) = - s(x) \phi(x) (1-\phi(x)) /sigma^2,
 \end{align}
 with appropriate boundary conditions.
+First rescale space by $\sigma$ so that it dissappears.
 If we now assume that $s(x)$ is piecewise constant,
@@
 Unfortunately, we are given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
-Monty deals with this by using symmetry, so he knows $\phi(0)$ and $\phi'(0)$.
+Monty (in "Gene flow and selection in a cline") deals with this by using symmetry, so he knows $\phi(0)$ and $\phi'(0)$.
 
@@
 This from ``NEQwiki, the nonlinear equations encyclopedia'', \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
+Recall that $V_i(\phi) = (1/2) s_i \phi^2(1-\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries.
+Note that if we rescale space by $\sqrt{s}/\sigma$,
+then $s$ and $\sigma$ drop from the equation.
 We want to integrate
@@
     \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
-        \sqrt{2}{s} \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } ,
+        \sqrt{ 2 } \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } ,
 \]
-with $C=2K/s$.
+with $C=2K$.
 Let $\phi^2(1-\phi/3)+C = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
-and change variables first to $y^2=(\alpha-\phi$, 
+and change variables first to $y^2=(\alpha-\phi)$, 
 and then to $x = y/\sqrt{\alpha-\beta}$, so that
@@
     \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } 
-        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-p^2) } } \\
+        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-y^2) } } \\
         = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
@@
 
+Presumably, this gets the same answer as Monty, who finds that in his solvable cases,
+for $x$ in the last interval (which ends in $\infty$), that
+\begin{align*}
+    \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{s} / \sigma + C ) \right) \\
+        &\simeq C' \exp\left( - \frac{ x \sqrt{s} }{ \sigma } \right) .
+\end{align*}
+Indeed, as $\phi \to 0$, we have approximately $\phi' = \frac{ \sqrt{s} }{ \sigma } \phi$.
+
 \section{Mutation rates}
@@
 Barton (Gen Res 87), has expressions for the probability of establishment as a function of spatial location in fixed landscape of advantageous pockets:
-The probability of establishment beginning at state p is <latex>F(p) = 1 - \exp(-4Ns\int \phi(x) p(dx))</latex>,
+The probability of establishment beginning at state p is $$F(p) = 1 - \exp(-4Ns\int \phi(x) p(dx))$$,
 where 

### 2011-06-15 standing-patches-paper.tex
/* Paper */ fixed some math

@@
 Consider first a single, isolated patch of area $A$ in which no new allele has appeared.
-\citet{barton1987} (or \citet{slatkin1981}?) has studied this situation (in one dimension)
+\citet{barton1987establishment} (or \citet{slatkin1973geneflow}?) has studied this situation (in one dimension)
 and found that:
@@
 However, we do know that if $q(r)$ is the expected proportion present at mutation--selection balance at distance $r$ from the first patch
-in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|s_m r / \sigma|)$,
+in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
 where $C$ is a constant depending on the structure of the populations and the selection coefficients.
@@
 If the alleles were not advantageous in the second patch,
-the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |s_m R/\sigma| )$;
-so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |s_m R/\sigma|)^{-1}$.
+the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |\sqrt{s_m} R/\sigma| )$;
+so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |\sqrt{s_m} R/\sigma|)^{-1}$.
 If $T_m$ is large, then we would expect successful migrants to be rare,
@@
 then the mutational input in each is on the order of $2 \rho \mu A$,
-while the influx due to migration between patches separated by distance $R$ is on the order $2 \rho A \exp(- |s_m R/\sigma|)$. 
-If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|s_m R/\sigma| )$,
+while the influx due to migration between patches separated by distance $R$ is on the order $2 \rho A \exp(- |\sqrt{s_m} R/\sigma|)$. 
+If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|\sqrt{s_m} R/\sigma| )$,
 then adaptation is expected to take place both by migration and mutation.
-If we take $\mu = 10^{-8}$, then this is satisfied if $R/\sigma \approx 18/s_m$.
+If we take $\mu = 10^{-8}$, then this is satisfied if $R/\sigma \approx 18/\sqrt{s_m}$.
 so if the disadvantage to the new mutation between patches is fairly strong (say, $s_m=.05$),
-then the distance between patches must be several hundreds of times the dispersal distance;
+then the distance between patches must be about a hundred times the dispersal distance (i.e.\ $R/\sigma \approx 100$);
 while if $s_m$ is much smaller, say $s_m = .001$, 
-then the patches must be separated by tens of thousands of times the dispersal distance.
+then the patches must be separated by many hundreds of times the dispersal distance (i.e.\ $R/\sigma \approx 500$).
 The situation does not change much if $\mu$ is larger -- even if $\mu = 10^{-6}$, 
@@
 On the other hand, if the separation is much smaller, then a single mutation is likely to fix everywhere.
-The general picture is that for species with spatial structure but without extreme barriers to dispersal, 
-patches will not be completely isolated, but some parallel adaptation is likely.
 

### 2011-06-15 Spatial_adaptation/panmixia.tex
Undo revision 3397 by [[Special:Contributions/Peter|Peter]] ([[User talk:Peter|talk]])
/* Comparison to panmixia */
/* Comparison to panmixia */  small fix

@@
 A reasonable range for $r=\sqrt{A}/\sigma$ is $10 \le r \le 1000$,
-over which range $4 < 2\log(r) < 14$, so $\log(N)$ will dominate the inequality.
+over which range $4 < 2\log(r) < 14$, while perhaps $10 \le \log(N) \le 20$.
 In other words, to get new mutations in the time it takes to locally fix,
-we'd need $1/(4 N \log N) < \mu$.
-Setting $\mu=10^{-8}$ tells us that we get multiple single base mutations in this time 
+we'd need $1/(4 N (\log N + 2 \log r)) < \mu$.
+Setting $\mu=10^{-8}$ and taking $r$ in this range
+tells us that we only get multiple single base mutations in this time 
 if $N$ is larger than about a million.
@@
 
-% \begin{figure}[hht]
-  \includegraphics[width=1.0\textwidth]{min-s-versus-r.pdf}
-% \caption{Plot of $s= 8 \left( \log(r)/r \right)^2$ against $r=\sqrt{A/\sigma^2}$.} 
-% \end{figure}
+ \begin{figure}[hht]
+  \includegraphics[width=0.8\textwidth]{min-s-versus-r.png}
+ \caption{Plot of $s= 8 \left( \log(r)/r \right)^2$ against $r=\sqrt{A/\sigma^2}$.} 
+ \end{figure}
 
@@
 
-\includegraphics[width=1.0\textwidth]{f-versus-gamma.pdf}
+\includegraphics[width=1.0\textwidth]{f-versus-gamma}
 
@@
 
-\includegraphics[width=1.0\textwidth]{gamma-contour-1e-8.pdf}
+\includegraphics[width=0.8\textwidth]{gamma-contour-1e-8.png}
 

### 2011-06-15 Spatial_adaptation/panmixia.R
*Empty MediaWiki Message*

@@ -36,10 +36,10 @@ rr <- (1:1000)
 ssrr <- expand.grid(ss,rr)
 ddd <- apply(ssrr,1,function(x) { log( log(x[2])/(x[2]*sqrt(x[1]*2)) ) } )
 dim(ddd)<-c(length(ss),length(rr))
-contour(x=ss,y=rr,z=ddd,xlab="s",ylab="r")
+# contour(x=ss,y=rr,z=ddd,xlab="s",ylab="r")
 
-png(file="min-s-versus-r.png",width=600,height=400)
-# pdf(file="min-s-versus-r.pdf",width=6,height=4)
+# png(file="min-s-versus-r.png",width=600,height=400)
+pdf(file="min-s-versus-r.pdf",width=6,height=4)
 rr <- 10:1000
 plot(rr,8*((log(rr)/rr)^2), type="l", xlab=expression(r==sqrt(A/sigma^2)), ylab=expression(s== 8 * ( log(r)/r )^2), log="xy")
 dev.off()
@@ -47,8 +47,8 @@ dev.off()
 NN <- exp(8:20)
 plot(NN, 1/(4*NN*log(NN)), type="l")
 
-png(file="f-versus-gamma.png",width=600,height=400)
-# pdf(file="f-versus-gamma.pdf",width=6,height=4)
+# png(file="f-versus-gamma.png",width=600,height=400)
+pdf(file="f-versus-gamma.pdf",width=6,height=4)
 xx <- (1:999)/1000
 plot((1-xx^2)/xx^3, xx, type="l", log="xy", xlab=expression(gamma==(1-x)/x^(3/2)), ylab=expression(x==(chi/chi[0])^2) )
 dev.off()
@@ -59,9 +59,8 @@ mu <- 1e-8
 sigsig <- (10:100)
 gg <- apply( expand.grid(ss,sigsig), 1, function (ssig) log( sqrt(ssig[1]) / ( ssig[2] * sqrt(4*pi*mu) * log(ssig[2]^2)^(3/2) ) ) )
 dim(gg) <- c(length(ss), length(sigsig))
-# pdf(file="gamma-contour-1e-8.pdf", width=6, height=4)
-png(file="gamma-contour-1e-8.png", width=600, height=400)
-# pdf(file="gamma-contour-1e-8.pdf", width=6, height=4)
+# png(file="gamma-contour-1e-8.png", width=600, height=400)
[…]

### 2011-06-15 Spatial_adaptation/panmixia.tex
/* Comparison to panmixia */ see figures?

@@
  \begin{figure}[hht]
-  \includegraphics[width=0.8\textwidth]{min-s-versus-r.png}
+  \includegraphics[width=0.8\textwidth]{min-s-versus-r.pdf}
  \caption{Plot of $s= 8 \left( \log(r)/r \right)^2$ against $r=\sqrt{A/\sigma^2}$.} 
@@
 taking the middle values in the figure, 
-this will be something like $1/20 < 4N\mu < 800$.
+spatial structure significantly increases the chance of parallel mutations if $1/20 < 4N\mu < 800$.
 
@@
 
-\includegraphics[width=1.0\textwidth]{f-versus-gamma}
+\includegraphics[width=1.0\textwidth]{f-versus-gamma.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth]{gamma-contour-1e-8.png}
+\includegraphics[width=0.8\textwidth]{gamma-contour-1e-8.pdf}
 
@@
 \paragraph{In summary,}
-the conclusions are:
+the conclusions are: (recall $r = \sqrt{A}/\sigma$)
 \begin{enumerate}
@@
     if the above does not happen and if
-    $s > \frac{1}{8 N^2 \mu^2 r^2} \gg 8 \left(\frac{ \log(r) }{ r }\right)^2$;
+    $s > \frac{1}{8 (N \mu r)^2} \gg 8 \left(\frac{ \log(r) }{ r }\right)^2$;
     this latter restricts the range of mutational influxes, but not dramatically.

### 2011-06-15 Spatial_adaptation/panmixia.R
*Empty MediaWiki Message*

@@ -39,7 +39,7 @@ dim(ddd)<-c(length(ss),length(rr))
 # contour(x=ss,y=rr,z=ddd,xlab="s",ylab="r")
 
 # png(file="min-s-versus-r.png",width=600,height=400)
-pdf(file="min-s-versus-r.pdf",width=6,height=4)
+pdf(file="Spatial_adaptation/min-s-versus-r.pdf",width=6,height=4)
 rr <- 10:1000
 plot(rr,8*((log(rr)/rr)^2), type="l", xlab=expression(r==sqrt(A/sigma^2)), ylab=expression(s== 8 * ( log(r)/r )^2), log="xy")
 dev.off()
@@ -48,7 +48,7 @@ NN <- exp(8:20)
 plot(NN, 1/(4*NN*log(NN)), type="l")
 
 # png(file="f-versus-gamma.png",width=600,height=400)
-pdf(file="f-versus-gamma.pdf",width=6,height=4)
+pdf(file="Spatial_adaptation/f-versus-gamma.pdf",width=6,height=4)
 xx <- (1:999)/1000
 plot((1-xx^2)/xx^3, xx, type="l", log="xy", xlab=expression(gamma==(1-x)/x^(3/2)), ylab=expression(x==(chi/chi[0])^2) )
 dev.off()
@@ -60,7 +60,7 @@ sigsig <- (10:100)
 gg <- apply( expand.grid(ss,sigsig), 1, function (ssig) log( sqrt(ssig[1]) / ( ssig[2] * sqrt(4*pi*mu) * log(ssig[2]^2)^(3/2) ) ) )
 dim(gg) <- c(length(ss), length(sigsig))
 # png(file="gamma-contour-1e-8.png", width=600, height=400)
-pdf(file="gamma-contour-1e-8.pdf", width=6, height=4)
+pdf(file="Spatial_adaptation/gamma-contour-1e-8.pdf", width=6, height=4)
 contour(x=ss,y=sigsig,z=gg, xlab="s", ylab=expression(sigma), main=expression(paste(log(gamma),", ", mu==10^-8)))
 dev.off()
 

### 2011-06-15 Spatial_adaptation/panmixia.R
Undo revision 3587 by [[Special:Contributions/Peter|Peter]] ([[User talk:Peter|talk]])

@@ -39,7 +39,7 @@ dim(ddd)<-c(length(ss),length(rr))
 # contour(x=ss,y=rr,z=ddd,xlab="s",ylab="r")
 
 # png(file="min-s-versus-r.png",width=600,height=400)
-pdf(file="Spatial_adaptation/min-s-versus-r.pdf",width=6,height=4)
+pdf(file="min-s-versus-r.pdf",width=6,height=4)
 rr <- 10:1000
 plot(rr,8*((log(rr)/rr)^2), type="l", xlab=expression(r==sqrt(A/sigma^2)), ylab=expression(s== 8 * ( log(r)/r )^2), log="xy")
 dev.off()
@@ -48,7 +48,7 @@ NN <- exp(8:20)
 plot(NN, 1/(4*NN*log(NN)), type="l")
 
 # png(file="f-versus-gamma.png",width=600,height=400)
-pdf(file="Spatial_adaptation/f-versus-gamma.pdf",width=6,height=4)
+pdf(file="f-versus-gamma.pdf",width=6,height=4)
 xx <- (1:999)/1000
 plot((1-xx^2)/xx^3, xx, type="l", log="xy", xlab=expression(gamma==(1-x)/x^(3/2)), ylab=expression(x==(chi/chi[0])^2) )
 dev.off()
@@ -60,7 +60,7 @@ sigsig <- (10:100)
 gg <- apply( expand.grid(ss,sigsig), 1, function (ssig) log( sqrt(ssig[1]) / ( ssig[2] * sqrt(4*pi*mu) * log(ssig[2]^2)^(3/2) ) ) )
 dim(gg) <- c(length(ss), length(sigsig))
 # png(file="gamma-contour-1e-8.png", width=600, height=400)
-pdf(file="Spatial_adaptation/gamma-contour-1e-8.pdf", width=6, height=4)
+pdf(file="gamma-contour-1e-8.pdf", width=6, height=4)
 contour(x=ss,y=sigsig,z=gg, xlab="s", ylab=expression(sigma), main=expression(paste(log(gamma),", ", mu==10^-8)))
 dev.off()
 

### 2011-06-15 Spatial_adaptation/panmixia.tex
/* Comparison to panmixia */ captions.

@@
 \]
-The relationship between $\gamma$ and $f$ is shown in this figure:
+The relationship between $\gamma$ and $x$ is shown in this figure:
 
+\begin{figure}[h!]
 \includegraphics[width=1.0\textwidth]{f-versus-gamma.pdf}
+\caption{ Proportion difference in numbers of mutations ($x$) versus $\gamma$, on a log-log scale. }
+\end{figure}
 
-showing that if $\gamma>1$ then $f$ drops off quickly.
+showing that if $\gamma>1$ then $x$ drops off quickly.
 If we then fix $\mu=10^{-8}$ and $\rho=1$ (working in units of $1/\sqrt{\rho}$), 
@@
 
+\begin{figure}[h!]
 \includegraphics[width=0.8\textwidth]{gamma-contour-1e-8.pdf}
+\caption{ Contour plot of $\log\gamma$ as a function of $s$ and $\sigma$. }
+\end{figure}
 

### 2011-06-15 standing-patches-paper.tex
/* Paper */ note about characteristic length
/* The Next Paper */

@@
 
+XXX could do calculations as in "panmixia" here, e.g. if $f = (\chi/\sqrt{\pi \lambda_0})^2$ then
+$f = 1-\gamma f^{3/2}$, where
+$\gamma = (4 \pi^{5/2} / \sqrt{2}) \mu^2 s_b^{3/2} \rho^2 / \sigma \log(1/(1-s_d))$.
+Then if $\gamma \ll 1$, $f \approx 1$, while if $\gamma \gg 1$, then $f \approx 0$.
+
 By the above calculation, we know that the relevant mutations occur about distance $\chi$ apart, 
@@
 
-Include stuff from ``comparison to panmixia'' here.
+Include stuff from ``comparison to panmixia'' here:
+
+Note that we are not considering the time it takes the mutation to fix locally.  
+How long is this time?  Mean, variance.
+When would we expect more mutations in this time?
+Roughly, if populations are larger than $C/\mu$ for some $C \approx 100$.
+
+Since we don't take this into account... we can ad-hoc it in by using the standing variation calculations -- add this to $\lambda_0$.
+This is equivalent to changing $s_d$ by... how much?
+
+Compare to panmixia: increases degree of parallel mutation if... ?
 

### 2011-06-20 standing-patches-paper.tex
/* The Next Paper */ rearranged the discrete section

@@
 \section{Patchy selection} 
+\label{ss:discretedemes}
 
-\subsection{Model}
-
-In this section, we will relax the assumption that the uniform is homogenous.
-The model we analyze in section \ref{ss:discretedemes} is a model of discrete demes exchanging rare migrants,
-but since the object of this paper is to work in continuous space,
-we will first in sections \ref{ss:patchymutation} and \ref{ss:patchymigration} describe how a model with {\em patchy} selection can be approximated by such a discrete model.
-
-For this purpose, suppose that instead of a uniform habitat,
-the selective pressure is such that the new alleles are only advantageous in isolated locations, 
-and are disadvantageous in the intervening region.
-As before, we assume that the new allele has fitness $1+s_b$ relative to the unmutated type in the ``good'' patches,
-and assume that it has fitness $1+s_m$ in the intervening areas, with $s_m<0$.
-To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
-and that the disadvantage $s_m$ ensures that on the relevant timescale,
-the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
-With these assumptions as justification, we will move to a model of discrete demes exchanging migrants,
-where ``migration'' occurs through the rare, transient passage of the new allele through regions where it is disadvantageous.
-
-equal deme sizes?  general migration rates?
-
-\subsection{Mutational influx}
-\label{ss:patchymutation}
-
-Consider first a single, isolated patch of area $A$ in which no new allele has appeared.
-\citet{barton1987establishment} (or \citet{slatkin1973geneflow}?) has studied this situation (in one dimension)
-and found that:
-mutations appearing further than a few multiples of $\sigma/\sqrt{2s_m}$ away from the patch have vanishingly small probability of establishment;
-mutations appearing within the patch have probability strictly less but approaching that $2s_b$ of establishment;
-and if the patch width is less than $2 \tan^{-1} (\sqrt{s_m})$, then local fixation does not occur (rather, is not stable).
-Let $p(x)$ be the probability that a new allele arising at location $x$ fixes.
-The function $p(x)$ can be found in terms of Jacobi elliptic functions,
-XXX as shown in the Appendix,
-but the expressions are particularly unhelpful.
-The total successful mutational influx per generation is then given by $\int \rho \mu p(x) dx$,
-and depends in a complicated way on the relationship of patch width and selection coefficients,
-but still scales linearly with the population mutational influx $\rho \mu$.
-Therefore, we define $p_f = \int p(x) dx$, and in a pinch will approximate this by $2 s_b A$,
-which will be a good approximation if the width $\sqrt{A}$ is large relative to $\sigma/\sqrt{2s_m}$.
-
-XXX We are interested in the case where a single mutation only takes over? XXX
-
-\subsection{Migration rates}
-\label{ss:patchymigration}
-
-Now suppose that there are two patches, each of area $A$, of good environment separated by distance $R$,
-and that the allele has begun to fix in the first, but has not yet appeared in the second.
-As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
-has probability approximately $p(x)$ of fixing locally in this patch,
-whether the allele is a new mutation or is descended from the occupants of the first patch.
-Due to mutation--selection balance there will be a small number of alleles descended from those of the first patch (``migrants'') 
-present in the regions where they are disadvantageous;
-the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX, citeXXX),
-although the equations are again not particularly helpful.
-However, we do know that if $q(r)$ is the expected proportion present at mutation--selection balance at distance $r$ from the first patch
-in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
-where $C$ is a constant depending on the structure of the populations and the selection coefficients.
-
-Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
-supposing that patches are large enough for local fixation to be stable.
-If the alleles were not advantageous in the second patch,
-the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |\sqrt{s_m} R/\sigma| )$;
-so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |\sqrt{s_m} R/\sigma|)^{-1}$.
-If $T_m$ is large, then we would expect successful migrants to be rare,
-so that if the second patch adapts through migration, it is likely to come from only a single migrant.
-We leave aside the converse situation, for which there is likely to be more than one migrant
-that founds the population mutant individuals in the new patch.
-
-\subsection{Discrete populations}
-\label{ss:discretedemes}
+We have been working in a continuous model of geographic space; 
+in this section we move to a discrete model of isolated populations exchanging rare migrants,
+but in section \ref{ss:patchyspace} discuss how a model with {\em patchy} selection can be approximated by such a discrete model.
 
@@
 then the mathematical picture is a pretty one.
+
+\subsection{Symmetric patches}
 
@@
 and that the case presented above is at the extreme (see also \cite{Navarro:03} for deterministic approximations).
+
+\subsection{Relation to a continuous model
+\label{ss:patchyspace}
+
+A similar situation of interest is the case when the species occupies a continuous range, but the selective pressure itself is patchy.
+If the areas where the new allele is advantageous are relatively isolated,
+we might hope to apply the results above to such a situation.
+For this purpose, suppose that instead of a uniform habitat,
+the selective pressure is such that the new alleles are only advantageous in isolated locations, 
+and are disadvantageous in the intervening region.
+As before, we assume that the new allele has fitness $1+s_b$ relative to the unmutated type in the ``good'' patches,
+and assume that it has fitness $1+s_m$ in the intervening areas, with $s_m<0$.
+To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
+and that the disadvantage $s_m$ ensures that on the relevant timescale,
+the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
+With these assumptions as justification, we will move to a model of discrete demes exchanging migrants,
+where ``migration'' occurs through the rare, transient passage of the new allele through regions where it is disadvantageous.
+
+equal deme sizes?  general migration rates?
+
+\subsubsection{Mutational influx}
+\label{ss:patchymutation}
+
+Consider first a single, isolated patch of area $A$ in which no new allele has appeared.
+\citet{barton1987establishment} (or \citet{slatkin1973geneflow}?) has studied this situation (in one dimension)
+and found that:
+mutations appearing further than a few multiples of $\sigma/\sqrt{2s_m}$ away from the patch have vanishingly small probability of establishment;
+mutations appearing within the patch have probability strictly less but approaching that $2s_b$ of establishment;
+and if the patch width is less than $2 \tan^{-1} (\sqrt{s_m})$, then local fixation does not occur (rather, is not stable).
+Let $p(x)$ be the probability that a new allele arising at location $x$ fixes.
+The function $p(x)$ can be found in terms of Jacobi elliptic functions,
+XXX as shown in the Appendix,
+but the expressions are particularly unhelpful.
+The total successful mutational influx per generation is then given by $\int \rho \mu p(x) dx$,
+and depends in a complicated way on the relationship of patch width and selection coefficients,
+but still scales linearly with the population mutational influx $\rho \mu$.
+Therefore, we define $p_f = \int p(x) dx$, and in a pinch will approximate this by $2 s_b A$,
+which will be a good approximation if the width $\sqrt{A}$ is large relative to $\sigma/\sqrt{2s_m}$.
+
+XXX We are interested in the case where a single mutation only takes over? XXX
+
+\subsubsection{Migration rates}
+\label{ss:patchymigration}
+
+Now suppose that there are two patches, each of area $A$, of good environment separated by distance $R$,
+and that the allele has begun to fix in the first, but has not yet appeared in the second.
+As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
+has probability approximately $p(x)$ of fixing locally in this patch,
+whether the allele is a new mutation or is descended from the occupants of the first patch.
+Due to mutation--selection balance there will be a small number of alleles descended from those of the first patch (``migrants'') 
+present in the regions where they are disadvantageous;
+the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX, citeXXX),
+although the equations are again not particularly helpful.
+However, we do know that if $q(r)$ is the expected proportion present at mutation--selection balance at distance $r$ from the first patch
+in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
+where $C$ is a constant depending on the structure of the populations and the selection coefficients.
+
+Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
+supposing that patches are large enough for local fixation to be stable.
+If the alleles were not advantageous in the second patch,
+the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |\sqrt{s_m} R/\sigma| )$;
+so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |\sqrt{s_m} R/\sigma|)^{-1}$.
+If $T_m$ is large, then we would expect successful migrants to be rare,
+so that if the second patch adapts through migration, it is likely to come from only a single migrant.
+We leave aside the converse situation, for which there is likely to be more than one migrant
+that founds the population mutant individuals in the new patch.
 

### 2011-06-21 standing-patches-paper.tex
/* The Next Paper */
updated with notes from train jun 20

@@
 \begin{center}
-\includegraphics[width=1.0\textwidth]{proportion-by-sigma.eps}
+\includegraphics[width=1.0\textwidth]{proportion-by-sd.eps}
 \caption{ %
@@
 \begin{center}
-\includegraphics[width=1.0\textwidth]{proportion-contour.png}
+\includegraphics[width=1.0\textwidth]{charlen-by-sd-sigma.png}
 \caption{ %
-{\bf Mean proportion of area occupied by standing variants,} as a function of dispersal distance $\sigma$ and selective disadvantage $s_d$.  XXX at what parameters. XXX
+{\bf Characteristic length,} as a function of dispersal distance $\sigma$ and selective disadvantage $s_d$.  XXX at what parameters. XXX
 }
@@
 influence of migration on things -- plot these versus migration
+
+{\tt Simulations:} compare mean density against $s_d$ at different values of $\sigma$; omit variability if needed for readability.
+Compare to panmictic results.
 
@@
 
-Since we don't take this into account... we can ad-hoc it in by using the standing variation calculations -- add this to $\lambda_0$.
-This is equivalent to changing $s_d$ by... how much?
+Both these results and the panmictic calculations are underestimates.  
+If the time to local fixation is enough to change things significantly, 
+we are in the realm where the panmictic results apply better.
+
+Graph: sigma against mean number of origins; different curves for spatial, panmictic (flat), and simulation;
+different values of theta as different line types?
+Mention Monty's empirical speeds paper.
 
 Compare to panmixia: increases degree of parallel mutation if... ?
+Check P and H papers and Orr and Betancourt.  (standing variation only)
 
@@
 is $M_{ij}$ and that the mean number of new mutations appearing in population $i$ per generation is $\mu_i$.
+
+{\tt Include cartoon here.}
 
@@
 
-Look at some human parameters?
+HbS
+
+lava bed peromyscus
+
+G6PD
 
@@
 discussion of applications.
+
+discuss mixing of types; 
+Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
 

### 2011-06-21 standing-patches-paper.tex
/* The Next Paper */ added some to "local fixation" section

@@
 
-\section{Local establishment}
+\section{Local establishment and comparison to panmixia}
 
-Include stuff from ``comparison to panmixia'' here:
+In the above (and in \citet{ralphcoop2010}) we have assumed that once the mutation appears,
+conditional on eventual fixation, it begins to spread spatially at speed $v$ instantly,
+effectively neglecting the time it must first spend fixing locally.
+In \citet{ralphcoop2010} we addressed this by noting that there would be no change at all in our results 
+if all mutations had to wait the same amount of time before fixing locally,
+and that this time was short relative to the time it took the wave to spread across the characteristic length;
+we then showed via simulation that this was reasonable in certain situations.
+In this section we examine this assumption in more detail, although mostly through heuristic arguments,
+and also compare our results above to those of \citet{softsweeps}.
 
-Note that we are not considering the time it takes the mutation to fix locally.  
-How long is this time?  Mean, variance.
-When would we expect more mutations in this time?
-Roughly, if populations are larger than $C/\mu$ for some $C \approx 100$.
+We are assuming that shortly after a new mutation appears, 
+it can be approximated by a branching process growing at rate $s$
+until the point that it grows large enough to ``feel'' spatial structure,
+at which point it begins to spread as a more--or--less deterministic wave.
+Although we are not aware of good analysis of this transition, 
+the relevant size when spatial structure becomes important
+should be something close to $\sigma^2 \rho$ 
+(i.e.\ Wright's ``local effective population size'' \citep{XXX}).
+Let $Z_t$ be a continuous-time branching process with $Z_0=1$ and $\E[Z_t] = e^{s_b t}$.
+Then we know that there exists a random variable $W$ such that $\lim_{t\to\infty} e^{-s_b t} Z_t = W$ almost surely,
+so that if $\tau$ is the time $Z_t$ reaches size $\sigma^2 \rho$, i.e.\ $\tau = \inf\{ t \ge 0: Z_t > \sigma^2 \rho$,
+then $\sigma^2 \rho = Z_\tau \approx e^{s_b \tau} W$.
+From this we know that $\tau \approx (1/s_b) (\log (\sigma^2 \rho) - \log W)$;
+although more detailed information is available (e.g.\ a CLT for $\tau$, \citet{XXX}),
+we will stick to the loose interpretation.
+\marginnote{check jaegers.  isn't there a CLT for $\tau$?}
+As $\sigma^2 \rho$ increases, the mean of $\tau$ increases, but the variance approaches a constant
+($\var[\tau] \to (1/s_b) \var[\log W]$).
 
-Both these results and the panmictic calculations are underestimates.  
-If the time to local fixation is enough to change things significantly, 
-we are in the realm where the panmictic results apply better.
+So, roughly speaking, we need to evaluate the importance of a delay of about $T = (1/s_b) \log (\sigma^2 \rho)$.
+New mutations will appear and fix during this time if $4N\mu s_b \ge 1/T$,
+i.e.\ if $4 N \mu \ge 1/\log (\sigma^2 \rho)$.
+Our model will still be a good approximation, however, 
+as long as $T$ is short relative to the time a wave takes to spread between nearby mutational origins.
+This can be worked out,
+but it is simpler to note that if the converse is true, 
+that the process is largely unaffected by spatial structure,
+and so the panmictic model is a good approximation for the true process.
+Since both our spatial model and the panmictic model underestimate the true degree of parallel adaptation,
+it suffices to compare results from the two models to avoid applying the wrong model.
 
-Graph: sigma against mean number of origins; different curves for spatial, panmictic (flat), and simulation;
-different values of theta as different line types?
-Mention Monty's empirical speeds paper.
+% When would we expect more mutations in this time?
+% Roughly, if populations are larger than $C/\mu$ for some $C \approx 100$.
+% Compare to panmixia: increases degree of parallel mutation if... ?
 
-Compare to panmixia: increases degree of parallel mutation if... ?
-Check P and H papers and Orr and Betancourt.  (standing variation only)
+\citet{softsweepsII}, it is shown that under a panmictic model with certain assumptions on the parameters,
+the number of independent origins due to both standing variation and new mutation seen in a sample of size $n$
+has approximately the Ewens distribution with parameters $n$ and $\theta = 4 N \mu$.
+As $n$ increases, the total number of types seen grows as $\log n$, 
+so the ``mean density of types'' corresponding to our $\lambda_0 + \nu_+$ in their model is infinite.
+\marginnote{Graph: sigma against mean number of origins; different curves for spatial, panmictic (flat), and simulation;
+different values of theta as different line types?  but how to do this?
+Mention Monty's empirical speeds paper.}
+We could compare the probability of parallel adaptation under the two models (see eq'n 18 in \citet{softsweepsII}),
+but this seems less useful than an assessment of the degree of parallel adaptation.
 

### 2011-06-21 standing-patches-paper.tex
/* The Next Paper */ intro notes

@@
 
-Reference on parallel adapation, standing variation, etc.:
+References on parallel adapation, standing variation, etc.:
   coop,
-  pennings and hermisson,
+  pennings and hermisson.
+Note pennings and hermisson found standing deleterious variation was very important.
+
+recall peromyscus examples with patchy environments
 
@@
 
-Stick to constant-speed waves.
-
 \section{General model}
@@
 outline here assumptions common to both models.
+
+Stick to constant-speed waves.
+Note monty's empircal speeds paper.
 

### 2011-06-29 Spatial_adaptation/patchy-selection-fns.R
WorkingWiki replaced project description

@@ -1,0 +1,26 @@
+#
+# Implement theory from patchy selection parts
+
+# mu is mutation rate
+# rho is population density
+# sigma is mean squared dispersal distance
+# sb >0 is advantage of mutation within patches
+# sd >0 is disadvantage of mutation between patches
+# R is distance between patches
+# 
+
+paramString <- function(ps=c("mu","rho","sb","sd","sigma","R"), pos=-1) {
+    # return a string for use in plots of the parameters.
+    # takes a vector or list of parameter names
+    if ( is.null(names(ps)) ) { names(ps) <- ps }
+    if ( mode(ps) != mode(1.0) ) { ps[1:length(ps)] <- 0; mode(ps) <- mode(1.0) }
+    for ( x in names(ps) ) { 
+        if (exists(x,where=pos,mode=mode(1.0))) {
+            ps[x] <- get(x, pos=pos, mode=mode(1.0)) 
+        } else {
+            ps[x] <- NA
+        }
+    }
+    paste( sapply( 1:length(ps), function (k) { sprintf( "%s=%.2g", names(ps)[k], ps[k] ) } ), collapse=", " )
+}
+

### 2011-06-29 Spatial_adaptation/patchy-selection-plots.R
WorkingWiki replaced project description

@@ -1,0 +1,15 @@
+
+source("Spatial_adaptation/patchy-selection-fns.R")
+
+# how big to make the figures
+figwidth <- 10
+figheight <- 8
+
+# playing around
+mu <- 10^(-8)
+rho <- 10
+sb <- .01
+sd <- .01
+sigmavals <- c(0.1,1,10,20,50)
+
+

### 2011-06-29 Spatial_adaptation/patchy-selection-fns.R
*Empty MediaWiki Message*

@@ -5,11 +5,41 @@
 # rho is population density
 # sigma is mean squared dispersal distance
 # sb >0 is advantage of mutation within patches
-# sd >0 is disadvantage of mutation between patches
+# sm <0 is disadvantage of mutation between patches
 # R is distance between patches
+# A is area of patches
 # 
+# rearrange so that
+# s = max(|sb|,|sm|)
+# sb = gb * s
+# sm = gm * s
 
-paramString <- function(ps=c("mu","rho","sb","sd","sigma","R"), pos=-1) {
+
+clineScale <- function(mu, rho, s, gb, gm, sigma, R, A) {
+    # sigma / sqrt(2 s) = the scale on which things happen between patches; 
+    # see slatkin; barton etc.
+    return( sigma / sqrt(2*sm) )
+}
+
+minimumPatch <- function(mu, rho, s, gb, gm, sigma, R, A) {
+    # the minimal patch size 
+    return( 2 * atanh( sqrt(gm/gb) ) )
+}
+
+mutationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
+    # approximation for mutational influx per patch
+    # XXX to-do: do numerics?
+    return( 2 * sb * A * rho * mu )
+}
+
+migrationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
+    # approximation for migrational influx per patch
+    # XXX to-do: do numerics?
+    return( 2 * sb * A * rho * exp( - sqrt(abs(s*gm)*R/sigma) ) )
+}
+
+
+paramString <- function(ps=c("mu","rho","s","gb","gm","sigma","R", "A"), pos=-1) {
     # return a string for use in plots of the parameters.
     # takes a vector or list of parameter names
     if ( is.null(names(ps)) ) { names(ps) <- ps }

### 2011-06-29 Spatial_adaptation/patchy-selection-plots.R

@@ -9,7 +9,7 @@ figheight <- 8
 mu <- 10^(-8)
 rho <- 10
 sb <- .01
-sd <- .01
+sm <- -.01
 sigmavals <- c(0.1,1,10,20,50)
 
 

### 2011-06-29 standing-patches-paper.tex
/* The Next Paper */ fixed latex oops
/* The Next Paper */ minor corrections

@@
 
-\subsection{Relation to a continuous model
+\subsection{Relation to a continuous model}
 \label{ss:patchyspace}
@@
 and found that:
-mutations appearing further than a few multiples of $\sigma/\sqrt{2s_m}$ away from the patch have vanishingly small probability of establishment;
+mutations appearing further than a few multiples of $\sigma/\sqrt{2|s_m|}$ away from the patch have vanishingly small probability of establishment;
 mutations appearing within the patch have probability strictly less but approaching that $2s_b$ of establishment;
-and if the patch width is less than $2 \tan^{-1} (\sqrt{s_m})$, then local fixation does not occur (rather, is not stable).
+and if the patch width is less than $2 \tan^{-1} (\sqrt{|s_m/s_b|})$, then local fixation does not occur (rather, is not stable).
 Let $p(x)$ be the probability that a new allele arising at location $x$ fixes.

### 2011-06-30 helianthus-ex.tex
WorkingWiki replaced project description

\documentclass{article} \usepackage[landscape]{geometry} \begin{document}
% latex table generated in R 2.13.0 by xtable 1.5-6 package
% Thu Jun 30 06:13:14 2011

\input{helianthus-ex-table.tex}

\paragraph{Parallel adaptation between patches?}
All values are in kilometers (squared km, etc).
The parameters: 
mu is the migration rate (per generation),
rho is the population density,
sb is the selection coefficient in the patches of dunes (chosen so the probability of fixation is $2\times sb$),
sm is the selection coefficient inbetween,
R is the distance between the patches,
and A is the area around a patch that might contribute adaptive mutations.
{\tt clScale} is the length over which the cline between the two types is expected to drop off;
{\tt minPatch} is the minimum area predicted to be able to locally adapt;
{\tt mutIn} is the mutational flux into an as-yet-colonized patch;
{\tt migIn} is the migrational flux from one patch to another.

\input{helianthus-standing-ex-table.tex}

\paragraph{Parallel adaptation within the main range?}
Here {\tt chi} is the characteristic scale of parallel adaptation;
and {\tt Etau} is the mean time it would take to adapt.

\end{document}

### 2011-06-30 patchy-selection-fns.R

@@ -1,0 +1,71 @@
+#
+# Implement theory from patchy selection parts
+
+# mu is mutation rate
+# rho is population density
+# sigma is mean squared dispersal distance
+# sb >0 is advantage of mutation within patches
+# sm <0 is disadvantage of mutation between patches
+# R is distance between patches
+# A is area of patches
+# 
+# rearrange so that
+# s = max(|sb|,|sm|)
+# sb = gb * s
+# sm = gm * s
+
+everything <- function(mu, rho, sb, sm, sigma, R, A) {
+    # compute all these things
+    s <- max(sb,sd)
+    gb <- sb/s
+    gm <- sm/s
+    mutI <- mutationInflux(mu, rho, s, gb, gm, sigma, R, A)
+    migI <- migrationInflux(mu, rho, s, gb, gm, sigma, R, A)
+    return( c( c( clScale=clineScale(mu, rho, s, gb, gm, sigma, R, A),
+        minPatch=minimumPatch(mu, rho, s, gb, gm, sigma, R, A),
+        mutIn=mutI,
+        migIn=migI,
+        ratioMutMig = mutI/migI
+        ), c(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A)
+        ) )
+}
+
+clineScale <- function(mu, rho, s, gb, gm, sigma, R, A) {
+    # sigma / sqrt(2 s) = the scale on which things happen between patches; 
+    # see slatkin; barton etc.
+    return( sigma / sqrt(2*sm) )
+}
+
+minimumPatch <- function(mu, rho, s, gb, gm, sigma, R, A) {
+    # the minimal patch area ( = width^2)
+    return( ( 2 * atanh( sqrt(gm/gb) ) * sigma / sqrt(2*sm) )^2 )
+}
+
+mutationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
+    # approximation for mutational influx per patch
+    # XXX to-do: do numerics?
+    return( 2 * sb * A * rho * mu )
+}
+
+migrationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
+    # approximation for migrational influx per patch
+    # XXX to-do: do numerics?
+    return( 2 * sb * A * rho * exp( - sqrt(abs(s*gm)*R/sigma) ) )
+}
+
+
+paramString <- function(ps=c("mu","rho","s","gb","gm","sigma","R", "A"), pos=-1) {
[…]

### 2011-06-30 patchy-selection-plots.R

@@ -1,0 +1,71 @@
+source("Spatial_adaptation/patchy-selection-fns.R")
+require(xtable)
+
+# Helianthus petiolaris/neglectus example:
+#  H.p. in southern Colorado; H.n. in Texas-ish
+
+# variance effective population density: total numbers / variance in offspring number
+rhovals <- c( 1e4 / 20, 1e5 / 10 )  # per km^2
+# dispersal distance: seeds, pollen
+sigmavals <- c(.020, .100, 1)  # km / gen
+# beneficial selection coefficient within the patches
+# --- set so that prob of establishment is 2*sb
+sb <- 0.2
+# deleterious selection coefficient outside of patches
+sm <- .001
+# mutation rate (note things will be linear in this
+muvals <- c(10^-8,10^-5)
+# size of patches is A <- 60 # km^2
+# ... but here we want the source area available for new mutations to come from
+#    if the dunes are unadapted, i.e. a ring around them of distance, say, 4*sigma
+A <- function(sigma) { 4*sigma*2*pi*10 }
+# distance between patches
+R <- 1000 # km
+
+###
+# Table up those values
+params <- expand.grid(rho=rhovals, sigma=sigmavals, mu=muvals)
+exvalues <- t( sapply(1:dim(params)[1], function (k) {
+        with( params[k,], everything(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A(sigma)) )
+        } ) )
+
+# write out the table in latex
+extable <- xtable(exvalues,digits=0)
+digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<1 | abs((x-floor(x))/x)>.01) }))] <- 3
+digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<.001) | any(abs(x)>1e5) }))] <- -2
+
+filename <- "helianthus-ex-table.tex"
+# write("\\documentclass{article} \\usepackage[landscape]{geometry} \\begin{document}", file=filename)
+print(extable, file=filename)#, append=TRUE)
+# write("\\end{document}", file=filename, append=TRUE)
+
+####
+# And look at parallel adaptation within the big species?
+source("Spatial_adaptation/standing-variation-fns.R")
+
[…]

### 2011-06-30 standing-variation-fns.R

@@ -1,0 +1,134 @@
+#
+# Implement theory from standing variation parts
+
+# mu is mutation rate
+# rho is population density
+# sigma is mean squared dispersal distance
+# sb >0 is advantage of mutation for t>0
+# sd >0 is disadvantage of mutation for t<0
+# 
+# ... and
+# lambda <- 2 * mu * rho * sb    # is rate of new successful mutations
+# lambda0 <- lambda / log(1/(1-sd))         # is approximate rate of standing successful mutations
+# v <- sigma * sqrt(2*sb)        # speed of wave of advance
+
+everything <- function (mu, rho, sb, sd, sigma) {
+    # bundle up everything
+        return( c( 
+            c( z0 = standingProportionArea(mu,rho,sb,sd,sigma)$value,
+                chi = charLength(mu,rho,sb,sd,sigma)$value,
+                Etau = meanTime(mu,rho,sb,sd,sigma)$value),
+            c( mu=mu, rho=rho, sb=sb, sd=sd, sigma=sigma )
+        ) )
+}
+
+meanTime <- function (mu, rho, sb, sd, sigma) {
+    # mean time til adaptation of a point;
+    # E[\tau] is int_0^\infty exp(-lambda0 pi v^2 t^2 - lambda pi v^2 t^3) dt
+    if (sd==0) {
+        return( list(value=0) )
+    } else if (sd==1) {
+        f <- function (t) {
+            exp( - 4 * mu * rho * sb^2 * sigma^2 * t^3 / 3) 
+        }
+    } else { 
+        f <- function (t) {
+            exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/log(1/(1-sd)) + t/3) )
+        }
+    }
+    xx <- 2^(-30:30)
+    yy <- sapply(xx, function(x) { f(x) } )
+    scale <- xx[ which.min( abs(10-yy*xx) ) ]
+    return( integrate(function(x) { f(x*scale)*scale }, 0, Inf) )
+}
+
+meanNewDensity <- function (mu, rho, sb, sd, sigma) {
+    # mean density of patches arising from new mutations
+    #  is lambda * E[\tau]
+    x <- meanTime(mu, rho, sb, sd, sigma)
+    x$value <- x$value * 2 * mu * rho * sb 
+    return( x )
+}
+
+standingProportionArea <- function (mu, rho, sb, sd, sigma) {
[…]

### 2011-06-30 patchy-selection-plots.R
WorkingWiki importation

@@ -43,12 +43,24 @@ print(extable, file=filename)#, append=TRUE)
 # And look at parallel adaptation within the big species?
 source("Spatial_adaptation/standing-variation-fns.R")
 
+# pre-environment-change deleterious selection coefficient
+sd <- .01
+
+params <- expand.grid(rho=rhovals, sigma=sigmavals, mu=muvals)
+
 exvalues <- t( sapply(1:dim(params)[1], function (k) {
-        with( params[k,], everything(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A(sigma)) )
+            with(params[k,], everything(mu=mu, rho=rho, sb=sb, sd=sd, sigma=sigma) )
         } ) )
-exvalues <- lapply(examples, function (x) {
-    with(x, everything(mu=mu, rho=rho, sb=sb, sd=sd, sigma=sigma) )
-        } )
+
+# write out the table in latex
+extable <- xtable(exvalues,digits=0)
+digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<1 | abs((x-floor(x))/x)>.01) }))] <- 3
+digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<.001) | any(abs(x)>1e5) }))] <- -2
+
+filename <- "helianthus-standing-ex-table.tex"
+# write("\\documentclass{article} \\usepackage[landscape]{geometry} \\begin{document}", file=filename)
+print(extable, file=filename)#, append=TRUE)
+# write("\\end{document}", file=filename, append=TRUE)
 
 
 #####

### 2011-07-01 Spatial_adaptation/patchy-selection-fns.R
WorkingWiki importation

@@ -14,6 +14,21 @@
 # sb = gb * s
 # sm = gm * s
 
+everything <- function(mu, rho, sb, sm, sigma, R, A) {
+    # compute all these things
+    s <- max(sb,sd)
+    gb <- sb/s
+    gm <- sm/s
+    mutI <- mutationInflux(mu, rho, s, gb, gm, sigma, R, A)
+    migI <- migrationInflux(mu, rho, s, gb, gm, sigma, R, A)
+    return( c( c( clScale=clineScale(mu, rho, s, gb, gm, sigma, R, A),
+        minPatch=minimumPatch(mu, rho, s, gb, gm, sigma, R, A),
+        mutIn=mutI,
+        migIn=migI,
+        ratioMutMig = mutI/migI
+        ), c(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A)
+        ) )
+}
 
 clineScale <- function(mu, rho, s, gb, gm, sigma, R, A) {
     # sigma / sqrt(2 s) = the scale on which things happen between patches; 
@@ -22,8 +37,8 @@ clineScale <- function(mu, rho, s, gb, gm, sigma, R, A) {
 }
 
 minimumPatch <- function(mu, rho, s, gb, gm, sigma, R, A) {
-    # the minimal patch size 
-    return( 2 * atanh( sqrt(gm/gb) ) )
+    # the minimal patch area ( = width^2)
+    return( ( 2 * atanh( sqrt(gm/gb) ) * sigma / sqrt(2*sm) )^2 )
 }
 
 mutationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {

### 2011-07-01 Spatial_adaptation/patchy-selection-plots.R

@@ -1,15 +1,83 @@
-
 source("Spatial_adaptation/patchy-selection-fns.R")
+require(xtable)
 
-# how big to make the figures
-figwidth <- 10
-figheight <- 8
+# Helianthus petiolaris/neglectus example:
+#  H.p. in southern Colorado; H.n. in Texas-ish
 
-# playing around
-mu <- 10^(-8)
-rho <- 10
-sb <- .01
-sm <- -.01
-sigmavals <- c(0.1,1,10,20,50)
+# variance effective population density: total numbers / variance in offspring number
+rhovals <- c( 1e4 / 20, 1e5 / 10 )  # per km^2
+# dispersal distance: seeds, pollen
+sigmavals <- c(.020, .100, 1)  # km / gen
+# beneficial selection coefficient within the patches
+# --- set so that prob of establishment is 2*sb
+sb <- 0.2
+# deleterious selection coefficient outside of patches
+sm <- .001
+# mutation rate (note things will be linear in this
+muvals <- c(10^-8,10^-5)
+# size of patches is A <- 60 # km^2
+# ... but here we want the source area available for new mutations to come from
+#    if the dunes are unadapted, i.e. a ring around them of distance, say, 4*sigma
+A <- function(sigma) { 4*sigma*2*pi*10 }
+# distance between patches
+R <- 1000 # km
+
+###
+# Table up those values
+params <- expand.grid(rho=rhovals, sigma=sigmavals, mu=muvals)
+exvalues <- t( sapply(1:dim(params)[1], function (k) {
+        with( params[k,], everything(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A(sigma)) )
+        } ) )
+
+# write out the table in latex
+extable <- xtable(exvalues,digits=0)
+digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<1 | abs((x-floor(x))/x)>.01) }))] <- 3
+digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<.001) | any(abs(x)>1e5) }))] <- -2
+
+filename <- "helianthus-ex-table.tex"
+# write("\\documentclass{article} \\usepackage[landscape]{geometry} \\begin{document}", file=filename)
+print(extable, file=filename)#, append=TRUE)
+# write("\\end{document}", file=filename, append=TRUE)
+
[…]

### 2011-07-01 Spatial_adaptation/standing-variation-fns.R

@@ -12,6 +12,16 @@
 # lambda0 <- lambda / log(1/(1-sd))         # is approximate rate of standing successful mutations
 # v <- sigma * sqrt(2*sb)        # speed of wave of advance
 
+everything <- function (mu, rho, sb, sd, sigma) {
+    # bundle up everything
+        return( c( 
+            c( z0 = standingProportionArea(mu,rho,sb,sd,sigma)$value,
+                chi = charLength(mu,rho,sb,sd,sigma)$value,
+                Etau = meanTime(mu,rho,sb,sd,sigma)$value),
+            c( mu=mu, rho=rho, sb=sb, sd=sd, sigma=sigma )
+        ) )
+}
+
 meanTime <- function (mu, rho, sb, sd, sigma) {
     # mean time til adaptation of a point;
     # E[\tau] is int_0^\infty exp(-lambda0 pi v^2 t^2 - lambda pi v^2 t^3) dt

### 2011-07-01 Spatial_adaptation/patchy-selection-plots.R
/* Plotting Code */

@@ -34,7 +34,7 @@ extable <- xtable(exvalues,digits=0)
 digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<1 | abs((x-floor(x))/x)>.01) }))] <- 3
 digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<.001) | any(abs(x)>1e5) }))] <- -2
 
-filename <- "helianthus-ex-table.tex"
+filename <- "Spatial_adaptation/helianthus-ex-table.tex"
 # write("\\documentclass{article} \\usepackage[landscape]{geometry} \\begin{document}", file=filename)
 print(extable, file=filename)#, append=TRUE)
 # write("\\end{document}", file=filename, append=TRUE)
@@ -57,7 +57,7 @@ extable <- xtable(exvalues,digits=0)
 digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<1 | abs((x-floor(x))/x)>.01) }))] <- 3
 digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<.001) | any(abs(x)>1e5) }))] <- -2
 
-filename <- "helianthus-standing-ex-table.tex"
+filename <- "Spatial_adaptation/helianthus-standing-ex-table.tex"
 # write("\\documentclass{article} \\usepackage[landscape]{geometry} \\begin{document}", file=filename)
 print(extable, file=filename)#, append=TRUE)
 # write("\\end{document}", file=filename, append=TRUE)

### 2011-07-01 patchy-selection-fns.R
*Empty MediaWiki Message*

@@ -1,71 +1,0 @@
-#
-# Implement theory from patchy selection parts
-
-# mu is mutation rate
-# rho is population density
-# sigma is mean squared dispersal distance
-# sb >0 is advantage of mutation within patches
-# sm <0 is disadvantage of mutation between patches
-# R is distance between patches
-# A is area of patches
-# 
-# rearrange so that
-# s = max(|sb|,|sm|)
-# sb = gb * s
-# sm = gm * s
-
-everything <- function(mu, rho, sb, sm, sigma, R, A) {
-    # compute all these things
-    s <- max(sb,sd)
-    gb <- sb/s
-    gm <- sm/s
-    mutI <- mutationInflux(mu, rho, s, gb, gm, sigma, R, A)
-    migI <- migrationInflux(mu, rho, s, gb, gm, sigma, R, A)
-    return( c( c( clScale=clineScale(mu, rho, s, gb, gm, sigma, R, A),
-        minPatch=minimumPatch(mu, rho, s, gb, gm, sigma, R, A),
-        mutIn=mutI,
-        migIn=migI,
-        ratioMutMig = mutI/migI
-        ), c(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A)
-        ) )
-}
-
-clineScale <- function(mu, rho, s, gb, gm, sigma, R, A) {
-    # sigma / sqrt(2 s) = the scale on which things happen between patches; 
-    # see slatkin; barton etc.
-    return( sigma / sqrt(2*sm) )
-}
-
-minimumPatch <- function(mu, rho, s, gb, gm, sigma, R, A) {
-    # the minimal patch area ( = width^2)
-    return( ( 2 * atanh( sqrt(gm/gb) ) * sigma / sqrt(2*sm) )^2 )
-}
-
-mutationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
-    # approximation for mutational influx per patch
-    # XXX to-do: do numerics?
-    return( 2 * sb * A * rho * mu )
-}
-
-migrationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
-    # approximation for migrational influx per patch
-    # XXX to-do: do numerics?
-    return( 2 * sb * A * rho * exp( - sqrt(abs(s*gm)*R/sigma) ) )
-}
-
-
-paramString <- function(ps=c("mu","rho","s","gb","gm","sigma","R", "A"), pos=-1) {
[…]

### 2011-07-01 patchy-selection-plots.R

@@ -1,83 +1,0 @@
-source("Spatial_adaptation/patchy-selection-fns.R")
-require(xtable)
-
-# Helianthus petiolaris/neglectus example:
-#  H.p. in southern Colorado; H.n. in Texas-ish
-
-# variance effective population density: total numbers / variance in offspring number
-rhovals <- c( 1e4 / 20, 1e5 / 10 )  # per km^2
-# dispersal distance: seeds, pollen
-sigmavals <- c(.020, .100, 1)  # km / gen
-# beneficial selection coefficient within the patches
-# --- set so that prob of establishment is 2*sb
-sb <- 0.2
-# deleterious selection coefficient outside of patches
-sm <- .001
-# mutation rate (note things will be linear in this
-muvals <- c(10^-8,10^-5)
-# size of patches is A <- 60 # km^2
-# ... but here we want the source area available for new mutations to come from
-#    if the dunes are unadapted, i.e. a ring around them of distance, say, 4*sigma
-A <- function(sigma) { 4*sigma*2*pi*10 }
-# distance between patches
-R <- 1000 # km
-
-###
-# Table up those values
-params <- expand.grid(rho=rhovals, sigma=sigmavals, mu=muvals)
-exvalues <- t( sapply(1:dim(params)[1], function (k) {
-        with( params[k,], everything(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A(sigma)) )
-        } ) )
-
-# write out the table in latex
-extable <- xtable(exvalues,digits=0)
-digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<1 | abs((x-floor(x))/x)>.01) }))] <- 3
-digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<.001) | any(abs(x)>1e5) }))] <- -2
-
-filename <- "helianthus-ex-table.tex"
-# write("\\documentclass{article} \\usepackage[landscape]{geometry} \\begin{document}", file=filename)
-print(extable, file=filename)#, append=TRUE)
-# write("\\end{document}", file=filename, append=TRUE)
-
-####
-# And look at parallel adaptation within the big species?
-source("Spatial_adaptation/standing-variation-fns.R")
-
-# pre-environment-change deleterious selection coefficient
[…]

### 2011-07-01 standing-variation-fns.R

@@ -1,134 +1,0 @@
-#
-# Implement theory from standing variation parts
-
-# mu is mutation rate
-# rho is population density
-# sigma is mean squared dispersal distance
-# sb >0 is advantage of mutation for t>0
-# sd >0 is disadvantage of mutation for t<0
-# 
-# ... and
-# lambda <- 2 * mu * rho * sb    # is rate of new successful mutations
-# lambda0 <- lambda / log(1/(1-sd))         # is approximate rate of standing successful mutations
-# v <- sigma * sqrt(2*sb)        # speed of wave of advance
-
-everything <- function (mu, rho, sb, sd, sigma) {
-    # bundle up everything
-        return( c( 
-            c( z0 = standingProportionArea(mu,rho,sb,sd,sigma)$value,
-                chi = charLength(mu,rho,sb,sd,sigma)$value,
-                Etau = meanTime(mu,rho,sb,sd,sigma)$value),
-            c( mu=mu, rho=rho, sb=sb, sd=sd, sigma=sigma )
-        ) )
-}
-
-meanTime <- function (mu, rho, sb, sd, sigma) {
-    # mean time til adaptation of a point;
-    # E[\tau] is int_0^\infty exp(-lambda0 pi v^2 t^2 - lambda pi v^2 t^3) dt
-    if (sd==0) {
-        return( list(value=0) )
-    } else if (sd==1) {
-        f <- function (t) {
-            exp( - 4 * mu * rho * sb^2 * sigma^2 * t^3 / 3) 
-        }
-    } else { 
-        f <- function (t) {
-            exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/log(1/(1-sd)) + t/3) )
-        }
-    }
-    xx <- 2^(-30:30)
-    yy <- sapply(xx, function(x) { f(x) } )
-    scale <- xx[ which.min( abs(10-yy*xx) ) ]
-    return( integrate(function(x) { f(x*scale)*scale }, 0, Inf) )
-}
-
-meanNewDensity <- function (mu, rho, sb, sd, sigma) {
-    # mean density of patches arising from new mutations
-    #  is lambda * E[\tau]
-    x <- meanTime(mu, rho, sb, sd, sigma)
-    x$value <- x$value * 2 * mu * rho * sb 
-    return( x )
-}
-
-standingProportionArea <- function (mu, rho, sb, sd, sigma) {
[…]

### 2012-12-03 maize-ex.tex
WorkingWiki replaced project description

--- 2011-06-30 helianthus-ex.tex
@@
 \documentclass{article} \usepackage[landscape]{geometry} \begin{document}
-% latex table generated in R 2.13.0 by xtable 1.5-6 package
-% Thu Jun 30 06:13:14 2011
 
@@
 
-\paragraph{Parallel adaptation between patches?}
-All values are in kilometers (squared km, etc).
-The parameters: 
-mu is the migration rate (per generation),
-rho is the population density,
-sb is the selection coefficient in the patches of dunes (chosen so the probability of fixation is $2\times sb$),
-sm is the selection coefficient inbetween,
-R is the distance between the patches,
-and A is the area around a patch that might contribute adaptive mutations.
-{\tt clScale} is the length over which the cline between the two types is expected to drop off;
-{\tt minPatch} is the minimum area predicted to be able to locally adapt;
-{\tt mutIn} is the mutational flux into an as-yet-colonized patch;
-{\tt migIn} is the migrational flux from one patch to another.
-
-\input{helianthus-standing-ex-table.tex}
-
-\paragraph{Parallel adaptation within the main range?}
-Here {\tt chi} is the characteristic scale of parallel adaptation;
-and {\tt Etau} is the mean time it would take to adapt.
-
 \end{document}

### 2012-12-03 patchy-selection-plots.R
/* Plotting Code */

@@ -81,3 +81,65 @@ print(extable, file=filename)#, append=TRUE)
 #     if (sigmaval==sigmavals[1]) { legend("topleft",col=rhocols,pch=1, legend=legtext) }
 # }
 # 
+
+
+######
+# Maize example:
+source("patchy-selection-fns.R")
+
+# variance effective population density: total numbers / variance in offspring number
+rhovals <- c( 1e4 / 100, 1e5 / 100 )  # per km^2
+# dispersal distance: seeds, pollen
+sigmavals <- c(.020, .100, 1)  # km / gen
+# beneficial selection coefficient within the patches
+# --- set so that prob of establishment is 2*sb
+sb <- 0.005
+# deleterious selection coefficient outside of patches
+sm <- .001
+# mutation rate (note things will be linear in this
+muvals <- c(10^-8,10^-5)
+# size of area of influx
+A <- 500 # km^2
+# distance between patches
+R <- 4000 # km
+
+###
+# Table up those values
+params <- expand.grid(rho=rhovals, sigma=sigmavals, mu=muvals)
+exvalues <- t( sapply(1:dim(params)[1], function (k) {
+        with( params[k,], everything(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A) )
+        } ) )
+
+# write out the table in latex
+extable <- xtable(exvalues,digits=0)
+digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<1 | abs((x-floor(x))/x)>.01) }))] <- 3
+digits(extable)[1+which(apply(exvalues, 2, function(x) { any(abs(x)<.001) | any(abs(x)>1e5) }))] <- -2
+
+filename <- "maize-ex-table.tex"
+# write("\\documentclass{article} \\usepackage[landscape]{geometry} \\begin{document}", file=filename)
+print(extable, file=filename)#, append=TRUE)
+# write("\\end{document}", file=filename, append=TRUE)
+
+####
+# And look at parallel adaptation within the big species?
+source("Spatial_adaptation/standing-variation-fns.R")
+
+# pre-environment-change deleterious selection coefficient
+sd <- .01
+
+params <- expand.grid(rho=rhovals, sigma=sigmavals, mu=muvals)
+
+exvalues <- t( sapply(1:dim(params)[1], function (k) {
[…]

### 2012-12-22 standing-patches-paper.tex
/* The Next Paper */

@@
 
+The covergent evolution of similar phenotypes in response to shared selection pressures is a testiment to the power of selection to repeatedly sculpt phenotypic variation.
+In some cases this convergence extends to the molecular level with dispate taxa converging to the same phenotype through parallel genetic changes in the same pathway, genes, or even by precisely the same genetic changes.
+Such precise convergence points to the conservation of function of the genes underlying adaptive traits, sometimes over deep time scales \citep{deephomologypapers}.
+Convergence on a molecular level also suggests that the path taken by adaptation can sometimes be relatively constained as few changes could have resulted in the adaptive phenotype, or that relatively few changes that can confer the advantagous phenotype of interest were sufficiently free of deleterious pleiotropic consequences to contribute to adaptation.
+
+%For highly polygenic traits
+
+Convergent adaptation can also occur within species, with individuals within a species 
+adapted to the same environment in parallel through different large effect changes. 
+There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
+All else being equal the evolution of convergent phenotypes within a species convergence is more likely with a higher mutational input, i.e.
+when the mutation target and population sizes are larger \citep{}.
+Convergence is also more likely when standing variation is present for the phenotype, 
+as multiple pre-existing alleles can spread from low frequency, 
+if our now adaptive alleles were previously not too deleterious \citep{Orr,Hermission}. 
+When multiple rare changes sweep up from low frequency in parallel within a population this has been termed a soft sweep.\\
+
+The geographic distribution of populations can also shape the probability of parallel mutation within a species.
+A geographically wide spread species is more likely to adapt by multiple mutations im parallel if dispersal 
+is geographically limited, as populations will adapt via new mutations rather than waiting for migration \citep{RalphCoop}. 
+Convergence is also more likely if geographically separated populations adapting to similar environments 
+via alleles which are maladapted in the environmentals separating them. 
+Such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{lenormand} 
+and so might encourage local adaptation by parallel mutation rather than migration.\\
+
+For example a variety of plant species have repeatedly adapted to patchs of heavy metal soils
+ (e.g. serpentine outcrops and mine tailings) within their ranges
+and  alleles that adapt the plants to heavy metals may be disfavored off of these patches. 
+Similarly across the American Southwest light colored rock pocket mice have adapted to outcrops of dark rock 
+through the evolution of darker cryptic pigmentation. These dark outcrops are geographically
+ separated by areas of light colored sand and rock, 
+and strong selection appears to favour the matching of crypsis to substrate. 
+Perhaps as a result of this strong selection against migrants, 
+at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops. 
+This situation is repeated in XXXX where darker colored mice
+ have repeated evolved lighter pigmentation to match lighter color sands on
+ beaches via different genetic routes.
+
+%http://www.nature.com/hdy/journal/v94/n2/full/6800600a.html
+
+In \citep{RalphCoop} we formulated a simple model of parallel adaptation in a geographic setting
+where the selection pressure was constant across the entire environment, 
+and that a single mutational change was sufficient to adapt populations to the novel environment. 
+We assumed that there was no standing variation for the adaptive allele, such that parallel mutation must be due to multiple mutations after the environmental shift.
+In that setting we found that there was a characteristic geographic scale, over which we would expect multiple instances of our adaptive allele to have arisen in parallel.
+This characteristic length could be expressed in terms of a simple compound parameter determined by our parameters of interest.
 Recall questions of origin of adaptations
@@
 \begin{center}
-\includegraphics[width=1.0\textwidth]{charlen-by-sd-limit.eps}
+  \includegraphics[width=1.0\textwidth]{charlen-by-sd-limit}
 \caption{ %
@@
 \begin{center}
-\includegraphics[width=1.0\textwidth]{proportion-by-sd.eps}
+  \includegraphics[width=1.0\textwidth]{proportion-by-sd}
 \caption{ %
@@
 \begin{center}
-\includegraphics[width=1.0\textwidth]{charlen-by-sd-sigma.png}
+  \includegraphics[width=1.0\textwidth]{charlen-by-sd-sigma}
 \caption{ %
@@
  \begin{center}
- \includegraphics{discrete-by-ratio.eps}
+   \includegraphics{discrete-by-ratio}
  \caption{ %

### 2013-01-21 standing-patches-paper.tex
/* The Next Paper */

@@
 Note monty's empircal speeds paper.
-
-\section{Standing variation}
-
-\subsection{The set-up}
-
-% selection pressure changes at $t=0$ from $1-s_d$ to $1+s_b$ -- homogeneous
-
-For now we ignore geographic hetergeneity, and assume that the species range
-is a homogenous, one- or two-dimensional region. 
-There are two selective classes -- the {\em neutral} type, and the {\em mutated} type.
-We assume that different mutations are distinguishable --
-either as selectively equivalent mutations, or by linked variation.
-For a long time in the past, the mutated type has been at a selective disadvantage (long enough to be at selection-mutation equilibrium),
-but at a certain time the selective regime changes, so that the mutated type has a selective advantage and quickly spreads to fixation.
-After fixation, alleles are either descended
-from families of mutants present as standing variation when the selective regime changed,
-or from new mutants arising since that time.
-
-For concreteness, suppose that before time $t=0$,
-the mutant type has fitness $1-s_d$ relative to the neutral type
-(i.e.\ it produces on average $1-s_d$ times the number of offspring per generation),
-and that after time $t=0$,
-the mutant type has fitness $1+s_b$,
-where $s_b>0$ will usually be assumed to be small, and $0<s_d<1$.
-We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
-The number of offspring has finite variance, which to keep the formulas simple we assume to be equal to one.
-(If the mutation is recessive, XXX more complicated things happen.)
-As for the other parameters,
-suppose that each offspring of a neutral parent is of the mutant type with probability $\mu$,
-and that the mean squared distance between parent and child is $\sigma^2$ (the {\em dispersal variance}).
-The species occupies an area with mean density
-$\rho$ haploid individuals (or chromosomes) per unit area.
-\marginnote{XXX or do diploid? below is haploid.}
-
-% Branching-process-sized point mass approx to local groups present at $t=0$
-% -- depends on local population density large enough relative to migration (to be "branching")
-% -- and migration not too large (since we approximate as a point mass).
-% Find mean density of clusters of successful standing mutants. 
-% Is greater-than-one size of clusters likely to make a difference?
-
-We make use of the commonly used approximation that neglects competition between relatives,
-treating the offspring of a new mutant that appears in an area not already occupied by the mutated type
-as a branching process.
-After $t=0$, the offspring of a single mutant form (approximately) a branching process with growth rate $s_b$,
-so each new mutant establishes locally with probability $p_s \approx 2s_b$.
-As in \cite{ralphcoop2010}, each new offspring has a very small probability of being a mutant and establishing locally,
-so the collections of times and locations at which mutants appear and establish locally 
-is well--approximated by a Poisson process in space and time.
-This the rate of this Poisson process, 
-the rate per unit area at which new mutants appear and establish locally after $t=0$ in areas not already occupied by the mutant type
-is approximately $\lambda = 2 \mu \rho s_b$.
-
-Before $t=0$, on the other hand, the mutation is deleterious, so the genetic descendents of each new mutation are (with high probability) doomed to extinction,
-but may persist for some time.
-Since the times and locations of mutants before $t=0$ also well-approximated by a Poisson process with rate $\mu \rho$,
-the locations of all mutant families extant at $t=0$ whose descendents are destined to fix locally is also,
-by the Mapping theorem, a Poisson process with rate we define to be $\lambda_0$.
-If we assume that the descendants of at most only a few members of any extant mutant family at $t=0$ will survive,
-and that these progenitors are near to each other in space, we can then treat each such family as a single mutant,
-equivalent to the mutants arising later.
-(The approximation will be good if the log of the size of an extant families is small relative to the establishment time,
-and the spatial distribution of the family is small relative to the spread between them.)
-To find $\lambda_0$, consider a mutation that arose at time $-T<0$, and let $Z_s$ be the number of its descendants at time $s-T$.
-At time $t=0$, there are $Z_T$ individuals present with the mutation,
-and each has probability $p_s \approx 2s_b$ of establishing, approximately independently.
-Therefore, the probability that some descendants of this mutation establish and fix locally is $1-(1-p_s)^{Z_T}$,
-so defining $\zeta(u,t) = \E[u^{Z_t} | Z_0=1 ]$ to be the generating function of $Z_T$,
-the rate is
-\begin{align*}
-    \mu \rho \int_0^\infty \left( 1- \zeta(1-p_s,T) \right) dT .
-\end{align*}
-Now for $s_b$ small, using $p_s \approx 2s_b$ and that $\E[Z_T]=(1-s_d)^T$,
-we know that $\zeta(1-2s_b,T) \approx 1-2s_b \E[Z_T] = 1-2s_b (1-s_d)^T$,
-resulting in the approximation
-\begin{align}
-    \lambda_0 &= \mu \rho \int_0^\infty 2s_b (1-s_d)^{T} dT \\
-        &= \frac{ 2 \mu \rho s_b }{ -\log(1-s_d) } .
-\end{align}
-
-XXX More discussion on whether this is a good approximation?  
-Compute typical size and radius of a standing cluster?
-
-recall PPP computations; draw cones of opportunity for new mutations
-
-\subsection{Characteristic parameters}
-
-% a characteristic length of regions and a proportion of haplotypes (or of space?) that are from standing variation
-
-In \citet{ralphcoop2010}, studying the model without standing variation,
-we defined a {\em characteristic length} which gave the spatial scale across with mutants with distinct origins would establish.
-This was proportional to the mean distance between neighboring established mutants,
-but had the advantage of being easier to calculate.
-Furthermore, the time scale over which adaptation occurred could be found by dividing the characteristic length 
-by the speed at which the mutants spread.
-We will define the characteristic length for this model,
-as well as a similar compound parameter describing the relative importance of standing variation to the process of adaptation.
-(Note that a similar approach, summarizing with a {\em characteristic length} was taken by \citep{slatkin1973geneflow} for a different problem.)
-
-Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
-As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
-The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
-$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
-(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
-% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
-Therefore, if we define $\chi$ to be the unique positive solution to
-\[
-    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
-\]
-then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
-before it is expected that one other successful mutation would have arisen in the area covered so far.
-The explicit formula for $\chi$ is cumbersome; here we omit it.
-(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
-This can be rewritten 
-\begin{equation} \label{eqn:defines_chi}
-%    \left( \frac{ - \chi \log(1-s_d) }{ 2 s_b \sigma } \right)^2 + \left( \frac{ - \chi \log(1-s_d) }{ 2 s_b \sigma } \right)^3 
-%        = \frac{ - \log(1-s_d)^3 }{ 4 s_b^2 \sigma^2 \pi \rho \mu } .
-   \frac{\sqrt{2s_b} }{-\log(1-s_d) } \chi^2 + \frac{1}{\sigma} \chi^3 = \frac{1}{\sqrt{2s_b} \rho\mu\pi}
-\end{equation}
-From this we see that $\chi$ decreases with $\rho$ and $\mu$.
-Furthermore, for large $\sigma$, the characteristic length is close to the value obtained just from standing variation:
-$\chi = \sqrt( -\log(s_d) / 2 s_b \rho \mu \pi ) + O(1/\sigma)$.
-On the other hand, if the mutant allele is highly deleterious before $t=0$,
-then the characteristic length is close to the value from \citet{ralphcoop2010}:
-$\chi = ( \sigma / \rho \mu \pi \sqrt{2 s_b} )^{1/3} + O(1/\log(1-s_d))$.
-
-XXX could do calculations as in "panmixia" here, e.g. if $f = (\chi/\sqrt{\pi \lambda_0})^2$ then
-$f = 1-\gamma f^{3/2}$, where
-$\gamma = (4 \pi^{5/2} / \sqrt{2}) \mu^2 s_b^{3/2} \rho^2 / \sigma \log(1/(1-s_d))$.
-Then if $\gamma \ll 1$, $f \approx 1$, while if $\gamma \gg 1$, then $f \approx 0$.
-
-By the above calculation, we know that the relevant mutations occur about distance $\chi$ apart, 
-and occur within the first $\chi/v$ generations.
-Said another way, if we look in a circular region of space of radius $\chi$ over $\chi/v$ generations,
-we expect to find one mutational origin;
-we denote the probability that this mutation is from standing variation by $\pi_0$.
-Therefore, $\pi_0$ is roughly the proportion of establishing mutations that come from standing variation,
-and it is an elementary consequence of the Poisson process that
-\begin{align} \label{eqn:pizero}
-    \pi_0 = \lambda_0 \pi \chi^2 .
-\end{align}
-\marginnote{put in value of $\chi$ computed above.}
-(In one dimension, $\pi_0 = 2 \lambda_0 \chi$.)
-
-\subsection{Standing Variation -- Results}
-
-The quantity that is perhaps easiest to compute is the mean time until adaptation.
-Fix some geographic location, and let $\tau\ge0$ be the time at which the point is reached by the advantageous mutations.
-Then, as shown in Figure XX,
-$\tau > t$ if and only if the cone with point at $(x,t)$ and slope $v$ extending back to $t=0$ is empty of successful mutations.
-Since we assume these are a Poisson process, 
-\[
-    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) ,
-\]
-and so we have
-\begin{align}
-    \E[\tau] % &= \int_0^\infty \P\{ \tau > t \} dt \\
-        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) dt.
-\end{align}
-For means of evaluting this integral, see Appendix \ref{apx:integrals}.
-
-% proportion of haplotypes (and of space) that are from standing variation
-
-We have defined $\lambda_0$ to be the mean density of standing variants that reach local fixation.
-Define $\nu_+$ to be the mean density of new mutants whose offspring fix locally.
-Since the probability that a mutant arising at $(x,t)$ is lucky enough to be born in a location not already occupied by mutants
-is $\P\{ \tau > t \}$,
-we can see  $\nu_+ = \int_0^\infty \lambda \P\{\tau>t\} dt$, and hence
-$\nu_+ = \lambda \E[\tau] $.
-Using that $\lambda_0 = \lambda / \log(1/(1-s_d))$, this gives us that
-\begin{equation}
-    \nu_+ = (\mbox{mean density of new patches}) = \frac{1}{1-\log(1-s_d) \E[\tau]} .
-\end{equation}
-Since the mean proportion of patches that come from standing variation is $\lambda_0 / (\lambda_0 + \nu_+)$,
-this gives us the exact quantity we approximated by $\pi_0$ in \eqref{eqn:pizero}.
-There are $\lambda_0 + \nu_+$ patches per unit area, so
-the typical patch (i.e.\ with distribution given by the Palm measure) occupies area $1/(\lambda_0 + \nu_+)$.
-
-We can also find the mean proportion of space covered by standing variants.
-If by time $t$ a geographic location has not yet been reached by the mutation,
-the probability it will be reached by time $t+dt$ 
-by a standing variant is approximately $2 \lambda_0 \pi v^2 t dt$, 
-and the probability it is reached by a new variant is $\lambda \pi v^2 t^2 dt$.
-Therefore, as is standard for competing exponentials,
-the probability a given location is reached first by a standing variant,
-and therefore the mean proportion of space covered  by standing variants,
-is
-\begin{equation}
-    z_0 = \int_0^\infty {2 \lambda_0 \pi v^2 t} \; \exp \left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) dt .
-\end{equation}
-\marginnote{put in this value in terms of Gaussian CDF?}
-To evaluate this integral, again see Appendix \ref{apx:integrals}.
-
-Furthermore, if we define $a_0$ to be the mean area occupied by a typical standing variant, 
-then $a_0$ is given by the proportion of the range occupied by standing variants over the mean density of unique standing variants,
-i.e.\ $a_0 = z_0 / \lambda_0$.
-We can compare this to the corresponding quantity $a_+$ for a new variant through the formula
-$a_0 / a_+ = z_0 / (1-z_0)$.
-
-\begin{figure}[ht]
-\begin{center}
-  \includegraphics[width=1.0\textwidth]{charlen-by-sd-limit}
-\caption{ %
-{\bf Characteristic length,} as a function of selective disadvantage, compared to the corresponding quantity without standing variation (e.g.\ from \cite{ralphcoop2010}) and to the quantity only considering standing variants (e.g.\ $\sqrt{\lambda_0}$).
-}
-\end{center}
-\end{figure}
-
-\begin{figure}[ht]
-\begin{center}
-  \includegraphics[width=1.0\textwidth]{proportion-by-sd}
-\caption{ %
-{\bf Mean proportion of patches arising from standing variation.} The paramters given are for the sickle-cell example.  XXX say more.
-}
-\end{center}
-\end{figure}
-
-\begin{figure}[ht]
-\begin{center}
-  \includegraphics[width=1.0\textwidth]{charlen-by-sd-sigma}
-\caption{ %
-{\bf Characteristic length,} as a function of dispersal distance $\sigma$ and selective disadvantage $s_d$.  XXX at what parameters. XXX
-}
-\end{center}
-\end{figure}
-
-\marginnote{do mean (and variance?) of distance between haplotypes?}
-
-influence of migration on things -- plot these versus migration
-
-{\tt Simulations:} compare mean density against $s_d$ at different values of $\sigma$; omit variability if needed for readability.
-Compare to panmictic results.
-
-\section{Local establishment and comparison to panmixia}
-
-In the above (and in \citet{ralphcoop2010}) we have assumed that once the mutation appears,
-conditional on eventual fixation, it begins to spread spatially at speed $v$ instantly,
-effectively neglecting the time it must first spend fixing locally.
-In \citet{ralphcoop2010} we addressed this by noting that there would be no change at all in our results 
-if all mutations had to wait the same amount of time before fixing locally,
-and that this time was short relative to the time it took the wave to spread across the characteristic length;
-we then showed via simulation that this was reasonable in certain situations.
-In this section we examine this assumption in more detail, although mostly through heuristic arguments,
-and also compare our results above to those of \citet{softsweeps}.
-
-We are assuming that shortly after a new mutation appears, 
-it can be approximated by a branching process growing at rate $s$
-until the point that it grows large enough to ``feel'' spatial structure,
-at which point it begins to spread as a more--or--less deterministic wave.
-Although we are not aware of good analysis of this transition, 
-the relevant size when spatial structure becomes important
-should be something close to $\sigma^2 \rho$ 
-(i.e.\ Wright's ``local effective population size'' \citep{XXX}).
-Let $Z_t$ be a continuous-time branching process with $Z_0=1$ and $\E[Z_t] = e^{s_b t}$.
-Then we know that there exists a random variable $W$ such that $\lim_{t\to\infty} e^{-s_b t} Z_t = W$ almost surely,
-so that if $\tau$ is the time $Z_t$ reaches size $\sigma^2 \rho$, i.e.\ $\tau = \inf\{ t \ge 0: Z_t > \sigma^2 \rho$,
-then $\sigma^2 \rho = Z_\tau \approx e^{s_b \tau} W$.
-From this we know that $\tau \approx (1/s_b) (\log (\sigma^2 \rho) - \log W)$;
-although more detailed information is available (e.g.\ a CLT for $\tau$, \citet{XXX}),
-we will stick to the loose interpretation.
-\marginnote{check jaegers.  isn't there a CLT for $\tau$?}
-As $\sigma^2 \rho$ increases, the mean of $\tau$ increases, but the variance approaches a constant
-($\var[\tau] \to (1/s_b) \var[\log W]$).
-
-So, roughly speaking, we need to evaluate the importance of a delay of about $T = (1/s_b) \log (\sigma^2 \rho)$.
-New mutations will appear and fix during this time if $4N\mu s_b \ge 1/T$,
-i.e.\ if $4 N \mu \ge 1/\log (\sigma^2 \rho)$.
-Our model will still be a good approximation, however, 
-as long as $T$ is short relative to the time a wave takes to spread between nearby mutational origins.
-This can be worked out,
-but it is simpler to note that if the converse is true, 
-that the process is largely unaffected by spatial structure,
-and so the panmictic model is a good approximation for the true process.
-Since both our spatial model and the panmictic model underestimate the true degree of parallel adaptation,
-it suffices to compare results from the two models to avoid applying the wrong model.
-
-% When would we expect more mutations in this time?
-% Roughly, if populations are larger than $C/\mu$ for some $C \approx 100$.
-% Compare to panmixia: increases degree of parallel mutation if... ?
-
-\citet{softsweepsII}, it is shown that under a panmictic model with certain assumptions on the parameters,
-the number of independent origins due to both standing variation and new mutation seen in a sample of size $n$
-has approximately the Ewens distribution with parameters $n$ and $\theta = 4 N \mu$.
-As $n$ increases, the total number of types seen grows as $\log n$, 
-so the ``mean density of types'' corresponding to our $\lambda_0 + \nu_+$ in their model is infinite.
-\marginnote{Graph: sigma against mean number of origins; different curves for spatial, panmictic (flat), and simulation;
-different values of theta as different line types?  but how to do this?
-Mention Monty's empirical speeds paper.}
-We could compare the probability of parallel adaptation under the two models (see eq'n 18 in \citet{softsweepsII}),
-but this seems less useful than an assessment of the degree of parallel adaptation.
 

### 2013-01-21 standing-variation-paper.tex
WorkingWiki replaced project description

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{natbib}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{latexml} % for \iflatexml

\newcommand{\var}{\mathop{\mbox{Var}}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}

\bibliographystyle{plain}

\begin{document}

\section{Introduction}

The covergent evolution of similar phenotypes in response to shared selection pressures is a testiment to the power of selection to repeatedly sculpt phenotypic variation. In some cases this convergence extends to the molecular level with dispate taxa converging to the same phenotype through parallel genetic changes in the same pathway, genes, or even by precisely the same genetic changes. 
Such precise convergence points to the conservation of function of the genes underlying adaptive traits, sometimes over deep time scales \citep{deephomologypapers}. Convergence on a molecular level also suggests that the path taken by adaptation can sometimes be relatively constained as few changes could have resulted in the adaptive phenotype, or that relatively few changes that can confer the advantagous phenotype of interest were sufficiently free of deleterious pleiotropic consequences to contribute to adaptation.

%For highly polygenic traits

Convergent adaptation can also occur within species, with individuals within a species 
adapted to the same environment in parallel through different large effect changes. 
There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
All else being equal the evolution of convergent phenotypes within a species convergence is more likely with a higher mutational input, i.e. when the mutation target and population sizes are larger \citep{}. 
Convergence is also more likely when standing variation is present for the phenotype, 
as multiple pre-existing alleles can spread from low frequency, 
if our now adaptive alleles were previously not too deleterious \citep{Orr,Hermission}. 
When multiple rare changes sweep up from low frequency in parallel within a population this has been termed a soft sweep.\\

The geographic distribution of populations can also shape the probability of parallel mutation within a species.
A geographically wide spread species is more likely to adapt by multiple mutations im parallel if dispersal 
is geographically limited, as populations will adapt via new mutations rather than waiting for migration \citep{RalphCoop}. 
Convergence is also more likely if geographically separated populations adapting to similar environments 
via alleles which are maladapted in the environmentals separating them. 
Such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{lenormand} 
and so might encourage local adaptation by parallel mutation rather than migration.\\

For example a variety of plant species have repeatedly adapted to patchs of heavy metal soils
 (e.g. serpentine outcrops and mine tailings) within their ranges
and  alleles that adapt the plants to heavy metals may be disfavored off of these patches. 
Similarly across the American Southwest light colored rock pocket mice have adapted to outcrops of dark rock 
through the evolution of darker cryptic pigmentation. These dark outcrops are geographically
 separated by areas of light colored sand and rock, 
and strong selection appears to favour the matching of crypsis to substrate. 
Perhaps as a result of this strong selection against migrants, 
at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops. 
This situation is repeated in XXXX where darker colored mice
 have repeated evolved lighter pigmentation to match lighter color sands on
 beaches via different genetic routes.

%http://www.nature.com/hdy/journal/v94/n2/full/6800600a.html

In \citep{RalphCoop} we formulated a simple model of parallel adaptation in a geographic setting
where the selection pressure was constant across the entire environment, 
and that a single mutational change was sufficient to adapt populations to the novel environment. 
We assumed that there was no standing variation for the adaptive allele, such that parallel mutation must be due to multiple mutations after the environmental shift. In that setting we found that there was a characteristic geographic scale, over which we would expect multiple instances of our adaptive allele to have arisen in parallel. This characteristic length could be expressed in terms of a simple compound parameter determined by our parameters of interest.
Recall questions of origin of adaptations

References on parallel adapation, standing variation, etc.:
  coop,
  pennings and hermisson.
Note pennings and hermisson found standing deleterious variation was very important.

recall peromyscus examples with patchy environments

Summarize previous paper

Question assumptions of no standing variation and homogenous selection

\section{General model}

outline here assumptions common to both models.

Stick to constant-speed waves.
Note monty's empircal speeds paper.

\section{Standing variation}

\subsection{The set-up}

% selection pressure changes at $t=0$ from $1-s_d$ to $1+s_b$ -- homogeneous

For now we ignore geographic hetergeneity, and assume that the species range
is a homogenous, one- or two-dimensional region. 
There are two selective classes -- the {\em neutral} type, and the {\em mutated} type.
We assume that different mutations are distinguishable --
either as selectively equivalent mutations, or by linked variation.
For a long time in the past, the mutated type has been at a selective disadvantage (long enough to be at selection-mutation equilibrium),
but at a certain time the selective regime changes, so that the mutated type has a selective advantage and quickly spreads to fixation.
After fixation, alleles are either descended
from families of mutants present as standing variation when the selective regime changed,
or from new mutants arising since that time.

For concreteness, suppose that before time $t=0$,
the mutant type has fitness $1-s_d$ relative to the neutral type
(i.e.\ it produces on average $1-s_d$ times the number of offspring per generation),
and that after time $t=0$,
the mutant type has fitness $1+s_b$,
where $s_b>0$ will usually be assumed to be small, and $0<s_d<1$.
We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
The number of offspring has finite variance, which to keep the formulas simple we assume to be equal to one.
(If the mutation is recessive, XXX more complicated things happen.)
As for the other parameters,
suppose that each offspring of a neutral parent is of the mutant type with probability $\mu$,
and that the mean squared distance between parent and child is $\sigma^2$ (the {\em dispersal variance}).
The species occupies an area with mean density
$\rho$ haploid individuals (or chromosomes) per unit area.
\marginnote{XXX or do diploid? below is haploid.}

% Branching-process-sized point mass approx to local groups present at $t=0$
% -- depends on local population density large enough relative to migration (to be "branching")
% -- and migration not too large (since we approximate as a point mass).
% Find mean density of clusters of successful standing mutants. 
% Is greater-than-one size of clusters likely to make a difference?

We make use of the commonly used approximation that neglects competition between relatives,
treating the offspring of a new mutant that appears in an area not already occupied by the mutated type
as a branching process.
After $t=0$, the offspring of a single mutant form (approximately) a branching process with growth rate $s_b$,
so each new mutant establishes locally with probability $p_s \approx 2s_b$.
As in \cite{ralphcoop2010}, each new offspring has a very small probability of being a mutant and establishing locally,
so the collections of times and locations at which mutants appear and establish locally 
is well--approximated by a Poisson process in space and time.
This the rate of this Poisson process, 
the rate per unit area at which new mutants appear and establish locally after $t=0$ in areas not already occupied by the mutant type
is approximately $\lambda = 2 \mu \rho s_b$.

Before $t=0$, on the other hand, the mutation is deleterious, so the genetic descendents of each new mutation are (with high probability) doomed to extinction,
but may persist for some time.
Since the times and locations of mutants before $t=0$ also well-approximated by a Poisson process with rate $\mu \rho$,
the locations of all mutant families extant at $t=0$ whose descendents are destined to fix locally is also,
by the Mapping theorem, a Poisson process with rate we define to be $\lambda_0$.
If we assume that the descendants of at most only a few members of any extant mutant family at $t=0$ will survive,
and that these progenitors are near to each other in space, we can then treat each such family as a single mutant,
equivalent to the mutants arising later.
(The approximation will be good if the log of the size of an extant families is small relative to the establishment time,
and the spatial distribution of the family is small relative to the spread between them.)
To find $\lambda_0$, consider a mutation that arose at time $-T<0$, and let $Z_s$ be the number of its descendants at time $s-T$.
At time $t=0$, there are $Z_T$ individuals present with the mutation,
and each has probability $p_s \approx 2s_b$ of establishing, approximately independently.
Therefore, the probability that some descendants of this mutation establish and fix locally is $1-(1-p_s)^{Z_T}$,
so defining $\zeta(u,t) = \E[u^{Z_t} | Z_0=1 ]$ to be the generating function of $Z_T$,
the rate is
\begin{align*}
    \mu \rho \int_0^\infty \left( 1- \zeta(1-p_s,T) \right) dT .
\end{align*}
Now for $s_b$ small, using $p_s \approx 2s_b$ and that $\E[Z_T]=(1-s_d)^T$,
we know that $\zeta(1-2s_b,T) \approx 1-2s_b \E[Z_T] = 1-2s_b (1-s_d)^T$,
resulting in the approximation
\begin{align}
    \lambda_0 &= \mu \rho \int_0^\infty 2s_b (1-s_d)^{T} dT \\
        &= \frac{ 2 \mu \rho s_b }{ -\log(1-s_d) } .
\end{align}

XXX More discussion on whether this is a good approximation?  
Compute typical size and radius of a standing cluster?

recall PPP computations; draw cones of opportunity for new mutations

\subsection{Characteristic parameters}

% a characteristic length of regions and a proportion of haplotypes (or of space?) that are from standing variation

In \citet{ralphcoop2010}, studying the model without standing variation,
we defined a {\em characteristic length} which gave the spatial scale across with mutants with distinct origins would establish.
This was proportional to the mean distance between neighboring established mutants,
but had the advantage of being easier to calculate.
Furthermore, the time scale over which adaptation occurred could be found by dividing the characteristic length 
by the speed at which the mutants spread.
We will define the characteristic length for this model,
as well as a similar compound parameter describing the relative importance of standing variation to the process of adaptation.
(Note that a similar approach, summarizing with a {\em characteristic length} was taken by \citep{slatkin1973geneflow} for a different problem.)

Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
$\lambda_0 \pi L^2 + \lambda \pi L^3 /v$ in two dimensions
(and $\lambda_0 2 L + \lambda \pi L^2 /v$ in one dimension).
% $\lambda_0 \omega_d L^d + \lambda \omega_d L^{d+1} /v$ in $d$ dimensions
Therefore, if we define $\chi$ to be the unique positive solution to
\[
    \lambda_0 \pi \chi^2 + \lambda \pi \chi^3 /v = 1,
\]
then $\chi$ gives the distance spread unobstructed by the descendants of a new mutant
before it is expected that one other successful mutation would have arisen in the area covered so far.
The explicit formula for $\chi$ is cumbersome; here we omit it.
(This is in two dimensions; in one dimension the characteristic length is $\chi = \frac{ \sqrt{ 4 \lambda_0^2 + 2\lambda v } - 2 \lambda_0 }{ 4 \lambda v }$.)
This can be rewritten 
\begin{equation} \label{eqn:defines_chi}
%    \left( \frac{ - \chi \log(1-s_d) }{ 2 s_b \sigma } \right)^2 + \left( \frac{ - \chi \log(1-s_d) }{ 2 s_b \sigma } \right)^3 
%        = \frac{ - \log(1-s_d)^3 }{ 4 s_b^2 \sigma^2 \pi \rho \mu } .
   \frac{\sqrt{2s_b} }{-\log(1-s_d) } \chi^2 + \frac{1}{\sigma} \chi^3 = \frac{1}{\sqrt{2s_b} \rho\mu\pi}
\end{equation}
From this we see that $\chi$ decreases with $\rho$ and $\mu$.
Furthermore, for large $\sigma$, the characteristic length is close to the value obtained just from standing variation:
$\chi = \sqrt( -\log(s_d) / 2 s_b \rho \mu \pi ) + O(1/\sigma)$.
On the other hand, if the mutant allele is highly deleterious before $t=0$,
then the characteristic length is close to the value from \citet{ralphcoop2010}:
$\chi = ( \sigma / \rho \mu \pi \sqrt{2 s_b} )^{1/3} + O(1/\log(1-s_d))$.

XXX could do calculations as in "panmixia" here, e.g. if $f = (\chi/\sqrt{\pi \lambda_0})^2$ then
$f = 1-\gamma f^{3/2}$, where
$\gamma = (4 \pi^{5/2} / \sqrt{2}) \mu^2 s_b^{3/2} \rho^2 / \sigma \log(1/(1-s_d))$.
Then if $\gamma \ll 1$, $f \approx 1$, while if $\gamma \gg 1$, then $f \approx 0$.

By the above calculation, we know that the relevant mutations occur about distance $\chi$ apart, 
and occur within the first $\chi/v$ generations.
Said another way, if we look in a circular region of space of radius $\chi$ over $\chi/v$ generations,
we expect to find one mutational origin;
we denote the probability that this mutation is from standing variation by $\pi_0$.
Therefore, $\pi_0$ is roughly the proportion of establishing mutations that come from standing variation,
and it is an elementary consequence of the Poisson process that
\begin{align} \label{eqn:pizero}
    \pi_0 = \lambda_0 \pi \chi^2 .
\end{align}
\marginnote{put in value of $\chi$ computed above.}
(In one dimension, $\pi_0 = 2 \lambda_0 \chi$.)

\subsection{Standing Variation -- Results}

The quantity that is perhaps easiest to compute is the mean time until adaptation.
Fix some geographic location, and let $\tau\ge0$ be the time at which the point is reached by the advantageous mutations.
Then, as shown in Figure XX,
$\tau > t$ if and only if the cone with point at $(x,t)$ and slope $v$ extending back to $t=0$ is empty of successful mutations.
Since we assume these are a Poisson process, 
\[
    \P\{ \tau > t \} = \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) ,
\]
and so we have
\begin{align}
    \E[\tau] % &= \int_0^\infty \P\{ \tau > t \} dt \\
        &= \int_0^\infty \exp\left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) dt.
\end{align}
For means of evaluting this integral, see Appendix \ref{apx:integrals}.

% proportion of haplotypes (and of space) that are from standing variation

We have defined $\lambda_0$ to be the mean density of standing variants that reach local fixation.
Define $\nu_+$ to be the mean density of new mutants whose offspring fix locally.
Since the probability that a mutant arising at $(x,t)$ is lucky enough to be born in a location not already occupied by mutants
is $\P\{ \tau > t \}$,
we can see  $\nu_+ = \int_0^\infty \lambda \P\{\tau>t\} dt$, and hence
$\nu_+ = \lambda \E[\tau] $.
Using that $\lambda_0 = \lambda / \log(1/(1-s_d))$, this gives us that
\begin{equation}
    \nu_+ = (\mbox{mean density of new patches}) = \frac{1}{1-\log(1-s_d) \E[\tau]} .
\end{equation}
Since the mean proportion of patches that come from standing variation is $\lambda_0 / (\lambda_0 + \nu_+)$,
this gives us the exact quantity we approximated by $\pi_0$ in \eqref{eqn:pizero}.
There are $\lambda_0 + \nu_+$ patches per unit area, so
the typical patch (i.e.\ with distribution given by the Palm measure) occupies area $1/(\lambda_0 + \nu_+)$.

We can also find the mean proportion of space covered by standing variants.
If by time $t$ a geographic location has not yet been reached by the mutation,
the probability it will be reached by time $t+dt$ 
by a standing variant is approximately $2 \lambda_0 \pi v^2 t dt$, 
and the probability it is reached by a new variant is $\lambda \pi v^2 t^2 dt$.
Therefore, as is standard for competing exponentials,
the probability a given location is reached first by a standing variant,
and therefore the mean proportion of space covered  by standing variants,
is
\begin{equation}
    z_0 = \int_0^\infty {2 \lambda_0 \pi v^2 t} \; \exp \left( - \lambda_0 \pi v^2 t^2 - \lambda \pi v^2 t^3 / 3 \right) dt .
\end{equation}
\marginnote{put in this value in terms of Gaussian CDF?}
To evaluate this integral, again see Appendix \ref{apx:integrals}.

Furthermore, if we define $a_0$ to be the mean area occupied by a typical standing variant, 
then $a_0$ is given by the proportion of the range occupied by standing variants over the mean density of unique standing variants,
i.e.\ $a_0 = z_0 / \lambda_0$.
We can compare this to the corresponding quantity $a_+$ for a new variant through the formula
$a_0 / a_+ = z_0 / (1-z_0)$.

\begin{figure}[ht]
\begin{center}
  \includegraphics[width=1.0\textwidth]{charlen-by-sd-limit}
\caption{ %
{\bf Characteristic length,} as a function of selective disadvantage, compared to the corresponding quantity without standing variation (e.g.\ from \cite{ralphcoop2010}) and to the quantity only considering standing variants (e.g.\ $\sqrt{\lambda_0}$).
}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
  \includegraphics[width=1.0\textwidth]{proportion-by-sd}
\caption{ %
{\bf Mean proportion of patches arising from standing variation.} The paramters given are for the sickle-cell example.  XXX say more.
}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
  \includegraphics[width=1.0\textwidth]{charlen-by-sd-sigma}
\caption{ %
{\bf Characteristic length,} as a function of dispersal distance $\sigma$ and selective disadvantage $s_d$.  XXX at what parameters. XXX
}
\end{center}
\end{figure}

\marginnote{do mean (and variance?) of distance between haplotypes?}

influence of migration on things -- plot these versus migration

{\tt Simulations:} compare mean density against $s_d$ at different values of $\sigma$; omit variability if needed for readability.
Compare to panmictic results.

\section{Local establishment and comparison to panmixia}

In the above (and in \citet{ralphcoop2010}) we have assumed that once the mutation appears,
conditional on eventual fixation, it begins to spread spatially at speed $v$ instantly,
effectively neglecting the time it must first spend fixing locally.
In \citet{ralphcoop2010} we addressed this by noting that there would be no change at all in our results 
if all mutations had to wait the same amount of time before fixing locally,
and that this time was short relative to the time it took the wave to spread across the characteristic length;
we then showed via simulation that this was reasonable in certain situations.
In this section we examine this assumption in more detail, although mostly through heuristic arguments,
and also compare our results above to those of \citet{softsweeps}.

We are assuming that shortly after a new mutation appears, 
it can be approximated by a branching process growing at rate $s$
until the point that it grows large enough to ``feel'' spatial structure,
at which point it begins to spread as a more--or--less deterministic wave.
Although we are not aware of good analysis of this transition, 
the relevant size when spatial structure becomes important
should be something close to $\sigma^2 \rho$ 
(i.e.\ Wright's ``local effective population size'' \citep{XXX}).
Let $Z_t$ be a continuous-time branching process with $Z_0=1$ and $\E[Z_t] = e^{s_b t}$.
Then we know that there exists a random variable $W$ such that $\lim_{t\to\infty} e^{-s_b t} Z_t = W$ almost surely,
so that if $\tau$ is the time $Z_t$ reaches size $\sigma^2 \rho$, i.e.\ $\tau = \inf\{ t \ge 0: Z_t > \sigma^2 \rho$,
then $\sigma^2 \rho = Z_\tau \approx e^{s_b \tau} W$.
From this we know that $\tau \approx (1/s_b) (\log (\sigma^2 \rho) - \log W)$;
although more detailed information is available (e.g.\ a CLT for $\tau$, \citet{XXX}),
we will stick to the loose interpretation.
\marginnote{check jaegers.  isn't there a CLT for $\tau$?}
As $\sigma^2 \rho$ increases, the mean of $\tau$ increases, but the variance approaches a constant
($\var[\tau] \to (1/s_b) \var[\log W]$).

So, roughly speaking, we need to evaluate the importance of a delay of about $T = (1/s_b) \log (\sigma^2 \rho)$.
New mutations will appear and fix during this time if $4N\mu s_b \ge 1/T$,
i.e.\ if $4 N \mu \ge 1/\log (\sigma^2 \rho)$.
Our model will still be a good approximation, however, 
as long as $T$ is short relative to the time a wave takes to spread between nearby mutational origins.
This can be worked out,
but it is simpler to note that if the converse is true, 
that the process is largely unaffected by spatial structure,
and so the panmictic model is a good approximation for the true process.
Since both our spatial model and the panmictic model underestimate the true degree of parallel adaptation,
it suffices to compare results from the two models to avoid applying the wrong model.

% When would we expect more mutations in this time?
% Roughly, if populations are larger than $C/\mu$ for some $C \approx 100$.
% Compare to panmixia: increases degree of parallel mutation if... ?

\citet{softsweepsII}, it is shown that under a panmictic model with certain assumptions on the parameters,
the number of independent origins due to both standing variation and new mutation seen in a sample of size $n$
has approximately the Ewens distribution with parameters $n$ and $\theta = 4 N \mu$.
As $n$ increases, the total number of types seen grows as $\log n$, 
so the ``mean density of types'' corresponding to our $\lambda_0 + \nu_+$ in their model is infinite.
\marginnote{Graph: sigma against mean number of origins; different curves for spatial, panmictic (flat), and simulation;
different values of theta as different line types?  but how to do this?
Mention Monty's empirical speeds paper.}
We could compare the probability of parallel adaptation under the two models (see eq'n 18 in \citet{softsweepsII}),
but this seems less useful than an assessment of the degree of parallel adaptation.

\end{document}

### 2013-01-24 patchy-selection-paper.tex
/* Patchy Selection paper */

--- 2013-01-21 standing-patches-paper.tex
@@
 
-\section{General model}
+\section{Methods}
+\subsection{Introduction to a patchily constructed model}
+\label{ss:patchyspace}
 
-outline here assumptions common to both models.
+Consider a population spread uniformly across a landscape. Within this environment there are patches of
+habitat that individuals are initially poorly adapted to, surrounded by large areas that our population is well adapted to.
 
-Stick to constant-speed waves.
-Note monty's empircal speeds paper.
+It takes only a single mutational change to create an allele ($B$) that adapts an individual to the
+poor habitat type. This mutational change occurs at a low rate of
+$\mu$ per chromosome per generation. We assume that this initially rare new
+allele has fitness $1+s_b$ relative to the unmutated type ($b$) in the ``poor'' habitat patches,
+and assume that it has fitness $1+s_m$ when rare in the intervening areas, with
+$s_m<0$. To simplify matters we assume that the disadvantage $s_m$ ensures that on the relevant timescale,
+the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
+
+We are interested in the establishment of mutations in the ``poor'' patches by either
+migration or mutation, and so we are only interested in whether the allele
+can escape initial loss by drift when rare. Thus we do not have to
+specify the the fitness of the homozygote, only that its
+fitness is such the dynamics of the  allele when it is rare are
+determined by the fitness of the heterozygote. This allows us to treat
+a diploid model as being essentially haploid for our purposes. \gc{TRUE?}
+
+For the sake of simplicity we will assume a constant haploid population density $\rho$, 
+across both types of habitat (we will return to discuss this equal
+density point later). We further assume that the
+variance in offspring number is $1$, and that the mean squared distance between parent and child is $\sigma^2$ (the dispersal variance).
+
+As only a single mutation is required to adapt individuals to the
+poor habitat patch, subsequent mutations that arise after an allele
+becomes established at its equilibrium frequency on the patch gain no selective
+benefit. Similarly an allele introduced into a patch by migration will
+not be favored by selection to spread, if the patch has already been
+colonized. Thus mutations selective exclude each other from
+patches, over short-time scales, and they will only slowly filter
+together over longer time scales by drift.
+
+\paragraph{Previous work} We will make use of a number previous
+results from the literature. \citet{slatkin1973geneflow}, working in
+one dimensional deterministic framework, showed that if the
+physical width of the patch is less than $2 \tan^{-1}
+(\sqrt{|s_m/s_b|})$, then $B$ can not 
+become stably established within the patch, as the local adaptation prevented by
+migrational swamping \citep[see also][ for a review]{Lenormand}.
+ \citet{barton1987establishment}, again working in one dimension,
+ showed that this critical patch size also held in a stochastic
+ framework \citep[see also the work of][]{Polk}. 
+For patches above this size, \citet{barton1987establishment}
+ founds that mutations appearing further than a few multiples of
+ $\sigma/\sqrt{2|s_m|}$ away from such a patch have vanishingly small
+ probability of becoming established within the patch. 
+Barton also showed that mutations appearing within the patch have probability 
+strictly less but approaching that $2s_b$ of establishment as the size
+of the patch increased. Thus for large enough patches the probability
+of establishment of a locally favorable allele approaches the standard
+establishment probability ($2s_b$) for a panamitic population
+\citep[][]{haldane, fisher}, a result that also (generally) holds for
+a geographically spread population experiencing a uniform selection
+pressure \citep{Maruyama, cherry}.
+
+%Maruyama: On the fixation probability of mutant genes in a subdivided population*
+
+%To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
+%and that the disadvantage $s_m$ ensures that on the relevant timescale,
+%the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
+
+%A similar situation of interest is the case when the species occupies a continuous range, but the selective pressure itself is patchy.
+I%f the areas where the new allele is advantageous are relatively isolated,
+%we might hope to apply the results above to such a situation.
+%For this purpose, suppose that instead of a uniform habitat,
+%the selective pressure is such that the new alleles are only advantageous in isolated locations, 
+%and are disadvantageous in the intervening region.
+%As before, we assume that the new allele has fitness $1+s_b$ relative to the unmutated type in the ``good'' patches,
+%and assume that it has fitness $1+s_m$ in the intervening areas, with $s_m<0$.
+%To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
+%and that the disadvantage $s_m$ ensures that on the relevant timescale,
+%the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
+%With these assumptions as justification, we will move to a model of discrete demes exchanging migrants,
+%where ``migration'' occurs through the rare, transient passage of the new allele through regions where it is disadvantageous.
+
+%equal deme sizes?  general migration rates?
+
+\subsubsection{Mutational influx}
+\label{ss:patchymutation}
+Consider first a single, isolated poor habitat patch of area $A$ in
+which no $B$ allele has yet become established. As
+we are interested in local adaptation that can occur we will assume
+that our patch is considerably larger than the cutoff for local
+establishment, i.e. $\sqrt{A} \gg 2 \tan^{-1} (\sqrt{|s_m/s_b|})$.
+
+Let $p(x)$ be the probability that a new allele arising at location
+$x$ away from the patch fixes within the poor habitat patch.
+The function $p(x)$ can be found in terms of Jacobi elliptic functions,
+XXX as shown in the Appendix, but the expressions are particularly unhelpful.
+The total successful mutational influx per generation is then given by $\int \rho \mu p(x) dx$,
+and depends in a complicated way on the relationship of patch width and selection coefficients,
+but still scales linearly with the population mutational influx $\rho \mu$.
+Therefore, we define $p_f = \int p(x) dx$.  If the width of the patch, $\sqrt{A}$, is large
+relative to $\sigma/\sqrt{2s_m}$ then a reasonable approximation is to
+set $p_f \approx 2 s_b A$. This follows from the fact that if the
+patch is large than the majority of the mutations that could become
+established within the patch, arose within the patch.
+
+The rate at which mutations arise and colonize the patch is on the order of $\rho \mu
+p_f$ generations, which will approximately be $2 s_b \rho \mu$. If
+this rate is low,  then the waiting time till a mutation arises that
+will become locally established within the patch is exponentially
+distributed with mean $2 s_b \rho \mu$.  Assuming that once a mutation
+becomes established
+ it quickly reaches its equilibrium frequency across the patch, the
+ expected time it takes for our patch to become colonized by the $B$
+ allele is $1/(2 s_b \rho \mu)$.
+
+\subsubsection{Migration rates}
+\label{ss:patchymigration}
+
+Now suppose that there are two patches, each of area $A$, of the poor
+habitat type separated by distance $R$. Consider the case where the
+allele has arisen and become established in the first patch, but has not yet appeared in the second.
+As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
+has probability approximately $p(x)$ of fixing locally in this patch,
+whether the allele is a new mutation or is descended from the occupants of the first patch.
+Due to mutation--selection balance there will be a small number of alleles descended from those of the first patch (``migrants'') 
+present in the regions where they are disadvantageous;
+the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX, citeXXX),
+although the equations are again not particularly helpful.
+However, we do know that if $q(r)$ is the expected proportion present at mutation--selection balance at distance $r$ from the first patch
+in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
+where $C$ is a constant depending on the structure of the populations and the selection coefficients.
+
+Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
+supposing that patches are large enough for local fixation to be
+stable. If the alleles were not advantageous in the second patch,
+the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |\sqrt{s_m} R/\sigma| )$;
+so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |\sqrt{s_m} R/\sigma|)^{-1}$.
+If $T_m$ is large, then we would expect successful migrants to be rare,
+so that if the second patch adapts through migration, it is likely to come from only a single migrant.
+We leave aside the converse situation, for which there is likely to be more than one migrant
+that founds the population mutant individuals in the new patch.
+
+\subsection{Patchy Selection -- Results} 
+\label{ss:discreteresults}
+
+We will assume that once a mutation escapes loss and becomes
+established in the poor habitat patch it reaches its equilbrium frequency very rapidly.
+
+If environment is continuous, with separated patches where the new allele is advantageous,
+we can only get a rough sense of whether the new adaptation is likely to spread through mutation (i.e.\ in parallel)
+or through migration.
+Taking the expressions from sections \ref{ss:patchymutation} and \ref{ss:patchymigration} above,
+we see that if patches each have sufficiently large area $A$,
+then the mutational input in each is on the order of $2 \rho \mu A$,
+while the influx due to migration between patches separated by distance $R$ is on the order $2 \rho A \exp(- |\sqrt{s_m} R/\sigma|)$. 
+If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|\sqrt{s_m} R/\sigma| )$,
+then adaptation is expected to take place both by migration and mutation.
+If we take $\mu = 10^{-8}$, then this is satisfied if $R/\sigma \approx 18/\sqrt{s_m}$.
+so if the disadvantage to the new mutation between patches is fairly strong (say, $s_m=.05$),
+then the distance between patches must be about a hundred times the dispersal distance (i.e.\ $R/\sigma \approx 100$);
+while if $s_m$ is much smaller, say $s_m = .001$, 
+then the patches must be separated by many hundreds of times the dispersal distance (i.e.\ $R/\sigma \approx 500$).
+The situation does not change much if $\mu$ is larger -- even if $\mu = 10^{-6}$, 
+the required separation between patches is only reduced by 60\%.
+If the separation between patches is much larger -- say, by a factor of ten -- 
+then patches are effectively isolated -- most adaptation will be independent between patches.
+On the other hand, if the separation is much smaller, then a single mutation is likely to fix everywhere.
+
+Furthermore, in the model of discrete demes connected by symmetric weak migration, 
+the final distribution of types depends only on the ratio of mutation to migration rates,
+and that the number of independently arisen types follows the Ewens distribution, with parameter $\mu/m$.
+In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
+for a model with 20 interconnected demes, across different parameter values.  
+This indicates that we need the population-scaled mutation and migration rates 
+to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.  
+If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
+
+\begin{figure}[ht]
+ \begin{center}
+   \includegraphics{discrete-by-ratio}
+ \caption{ %
+ The mean number of types and the mean number of demes occupied by a sampled type, in a discrete model with 20 demes,
+ plotted as a function of the ratio of migration to mutation probabilities.
+ Note the log scale on the horizontal axis.
+ \label{fig:discbyratio}
+ }
+ \end{center}
+\end{figure}
+
+Also note that the properties of the discrete model are independent of the {\em sizes} of the demes themselves--- we have assumed away all such dependence.
+This is complementary to the results of \cite{softsweepsII}, who showed that multiple mutations are likely to arise {\em within} a panmictic deme
+if the population-scaled mutation rate $2 N \mu$ is greater than $1$.
+
+\section{Applications}
+
+HbS
+
+lava bed peromyscus
+
+G6PD
+
+\section{Discussion}
+
+if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
+recall others' work.
+
+patchy selection helps this happen; 
+effective migration rate looks like what; 
+strength of (positive?) selection suprisingly does not affect results (?)
+
+standing variation will be important under these circumstances; 
+migration rate (and so geographic structure) will not affect numbers of types if this is true; 
+however migration always affcts spatial patterns.
+
+discussion of applications.
+
+discuss mixing of types; 
+Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
+
+\bibliography{standing_patches_refs}
+
+\appendix
+
+\section{Integrals appearing in the text}
+    \label{apx:integrals}
+
+In the text at several points appear integrals of the form
+\begin{align}
+  \int_0^\infty t^c \exp \left( - \alpha t^d - \beta t^{d+1} \right) dt 
+\end{align}
+where $c$ is a positive integer, $d$ is the dimension, and $\alpha$ and $\beta$ are positive real numbers.
+This could be evaluated through standard numerical methods; below we describe a power series expansion.
+Changing variables to $u = \beta t^{d+1}$, this becomes
+\begin{align}
+    \left( (d+1)^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
+\end{align}
+so it suffices to evaluate the function
+\begin{equation}
+    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
+\end{equation}
+in the case that $a=(c-d+1)/(d+1)$, $b=d/(d+1)$, and $x$ is a function of the demographic paramters.
+Since $a$ and $b$ only depend on the dimension and the quantity being computed,
+we are interested in $G$ as a function of $x$.
+At least in the case $b=d/(d+1)$ it is possible to express $G$ as a finite sum of gamma functions,
+but we proceed with a simpler method.
+Note that $\partial_x G(a,b,x) = -G(a+b,b,x)$,
+and that at $x=0$, the function $G$ is the gamma function $G(a,b,0) = \Gamma(a+1)$.
+Therefore, a Taylor series for $G$ would be
+\[
+    G(a,b,x) = \sum_{n \ge 0} = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
+\]
+It is easy to check using Stirling's formula that $\limsup_{n \to \infty} ( x^n \Gamma(a+nb+1)/n! )^{1/n} = 0$
+if $b<1$, so the sum converges.
 
@@
 
-\subsection{Relation to a continuous model}
-\label{ss:patchyspace}
-
-A similar situation of interest is the case when the species occupies a continuous range, but the selective pressure itself is patchy.
-If the areas where the new allele is advantageous are relatively isolated,
-we might hope to apply the results above to such a situation.
-For this purpose, suppose that instead of a uniform habitat,
-the selective pressure is such that the new alleles are only advantageous in isolated locations, 
-and are disadvantageous in the intervening region.
-As before, we assume that the new allele has fitness $1+s_b$ relative to the unmutated type in the ``good'' patches,
-and assume that it has fitness $1+s_m$ in the intervening areas, with $s_m<0$.
-To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
-and that the disadvantage $s_m$ ensures that on the relevant timescale,
-the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
-With these assumptions as justification, we will move to a model of discrete demes exchanging migrants,
-where ``migration'' occurs through the rare, transient passage of the new allele through regions where it is disadvantageous.
-
-equal deme sizes?  general migration rates?
-
-\subsubsection{Mutational influx}
-\label{ss:patchymutation}
-
-Consider first a single, isolated patch of area $A$ in which no new allele has appeared.
-\citet{barton1987establishment} (or \citet{slatkin1973geneflow}?) has studied this situation (in one dimension)
-and found that:
-mutations appearing further than a few multiples of $\sigma/\sqrt{2|s_m|}$ away from the patch have vanishingly small probability of establishment;
-mutations appearing within the patch have probability strictly less but approaching that $2s_b$ of establishment;
-and if the patch width is less than $2 \tan^{-1} (\sqrt{|s_m/s_b|})$, then local fixation does not occur (rather, is not stable).
-Let $p(x)$ be the probability that a new allele arising at location $x$ fixes.
-The function $p(x)$ can be found in terms of Jacobi elliptic functions,
-XXX as shown in the Appendix,
-but the expressions are particularly unhelpful.
-The total successful mutational influx per generation is then given by $\int \rho \mu p(x) dx$,
-and depends in a complicated way on the relationship of patch width and selection coefficients,
-but still scales linearly with the population mutational influx $\rho \mu$.
-Therefore, we define $p_f = \int p(x) dx$, and in a pinch will approximate this by $2 s_b A$,
-which will be a good approximation if the width $\sqrt{A}$ is large relative to $\sigma/\sqrt{2s_m}$.
-
-XXX We are interested in the case where a single mutation only takes over? XXX
-
-\subsubsection{Migration rates}
-\label{ss:patchymigration}
-
-Now suppose that there are two patches, each of area $A$, of good environment separated by distance $R$,
-and that the allele has begun to fix in the first, but has not yet appeared in the second.
-As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
-has probability approximately $p(x)$ of fixing locally in this patch,
-whether the allele is a new mutation or is descended from the occupants of the first patch.
-Due to mutation--selection balance there will be a small number of alleles descended from those of the first patch (``migrants'') 
-present in the regions where they are disadvantageous;
-the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX, citeXXX),
-although the equations are again not particularly helpful.
-However, we do know that if $q(r)$ is the expected proportion present at mutation--selection balance at distance $r$ from the first patch
-in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
-where $C$ is a constant depending on the structure of the populations and the selection coefficients.
-
-Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
-supposing that patches are large enough for local fixation to be stable.
-If the alleles were not advantageous in the second patch,
-the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |\sqrt{s_m} R/\sigma| )$;
-so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |\sqrt{s_m} R/\sigma|)^{-1}$.
-If $T_m$ is large, then we would expect successful migrants to be rare,
-so that if the second patch adapts through migration, it is likely to come from only a single migrant.
-We leave aside the converse situation, for which there is likely to be more than one migrant
-that founds the population mutant individuals in the new patch.
-
-\subsection{Patchy Selection -- Results} 
-\label{ss:discreteresults}
-
-If environment is continuous, with separated patches where the new allele is advantageous,
-we can only get a rough sense of whether the new adaptation is likely to spread through mutation (i.e.\ in parallel)
-or through migration.
-Taking the expressions from sections \ref{ss:patchymutation} and \ref{ss:patchymigration} above,
-we see that if patches each have sufficiently large area $A$,
-then the mutational input in each is on the order of $2 \rho \mu A$,
-while the influx due to migration between patches separated by distance $R$ is on the order $2 \rho A \exp(- |\sqrt{s_m} R/\sigma|)$. 
-If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|\sqrt{s_m} R/\sigma| )$,
-then adaptation is expected to take place both by migration and mutation.
-If we take $\mu = 10^{-8}$, then this is satisfied if $R/\sigma \approx 18/\sqrt{s_m}$.
-so if the disadvantage to the new mutation between patches is fairly strong (say, $s_m=.05$),
-then the distance between patches must be about a hundred times the dispersal distance (i.e.\ $R/\sigma \approx 100$);
-while if $s_m$ is much smaller, say $s_m = .001$, 
-then the patches must be separated by many hundreds of times the dispersal distance (i.e.\ $R/\sigma \approx 500$).
-The situation does not change much if $\mu$ is larger -- even if $\mu = 10^{-6}$, 
-the required separation between patches is only reduced by 60\%.
-If the separation between patches is much larger -- say, by a factor of ten -- 
-then patches are effectively isolated -- most adaptation will be independent between patches.
-On the other hand, if the separation is much smaller, then a single mutation is likely to fix everywhere.
-
-Furthermore, in the model of discrete demes connected by symmetric weak migration, 
-the final distribution of types depends only on the ratio of mutation to migration rates,
-and that the number of independently arisen types follows the Ewens distribution, with parameter $\mu/m$.
-In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
-for a model with 20 interconnected demes, across different parameter values.  
-This indicates that we need the population-scaled mutation and migration rates 
-to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.  
-If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
-
-\begin{figure}[ht]
- \begin{center}
-   \includegraphics{discrete-by-ratio}
- \caption{ %
- The mean number of types and the mean number of demes occupied by a sampled type, in a discrete model with 20 demes,
- plotted as a function of the ratio of migration to mutation probabilities.
- Note the log scale on the horizontal axis.
- \label{fig:discbyratio}
- }
- \end{center}
-\end{figure}
-
-Also note that the properties of the discrete model are independent of the {\em sizes} of the demes themselves--- we have assumed away all such dependence.
-This is complementary to the results of \cite{softsweepsII}, who showed that multiple mutations are likely to arise {\em within} a panmictic deme
-if the population-scaled mutation rate $2 N \mu$ is greater than $1$.
-
-\section{Applications}
-
-HbS
-
-lava bed peromyscus
-
-G6PD
-
-\section{Discussion}
-
-if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
-recall others' work.
-
-patchy selection helps this happen; 
-effective migration rate looks like what; 
-strength of (positive?) selection suprisingly does not affect results (?)
-
-standing variation will be important under these circumstances; 
-migration rate (and so geographic structure) will not affect numbers of types if this is true; 
-however migration always affcts spatial patterns.
-
-discussion of applications.
-
-discuss mixing of types; 
-Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
-
-\bibliography{standing_patches_refs}
-
-\appendix
-
-\section{Integrals appearing in the text}
-    \label{apx:integrals}
-
-In the text at several points appear integrals of the form
-\begin{align}
-  \int_0^\infty t^c \exp \left( - \alpha t^d - \beta t^{d+1} \right) dt 
-\end{align}
-where $c$ is a positive integer, $d$ is the dimension, and $\alpha$ and $\beta$ are positive real numbers.
-This could be evaluated through standard numerical methods; below we describe a power series expansion.
-Changing variables to $u = \beta t^{d+1}$, this becomes
-\begin{align}
-    \left( (d+1)^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
-\end{align}
-so it suffices to evaluate the function
-\begin{equation}
-    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
-\end{equation}
-in the case that $a=(c-d+1)/(d+1)$, $b=d/(d+1)$, and $x$ is a function of the demographic paramters.
-Since $a$ and $b$ only depend on the dimension and the quantity being computed,
-we are interested in $G$ as a function of $x$.
-At least in the case $b=d/(d+1)$ it is possible to express $G$ as a finite sum of gamma functions,
-but we proceed with a simpler method.
-Note that $\partial_x G(a,b,x) = -G(a+b,b,x)$,
-and that at $x=0$, the function $G$ is the gamma function $G(a,b,0) = \Gamma(a+1)$.
-Therefore, a Taylor series for $G$ would be
-\[
-    G(a,b,x) = \sum_{n \ge 0} = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
-\]
-It is easy to check using Stirling's formula that $\limsup_{n \to \infty} ( x^n \Gamma(a+nb+1)/n! )^{1/n} = 0$
-if $b<1$, so the sum converges.
-
 \end{document}

### 2013-01-24 patchy-selection-paper.tex
/* Patchy Selection paper */

@@
 \usepackage{graphicx}
+\usepackage{color}
 \usepackage{hyperref}
@@
 \newcommand{\R}{\mathbb{R}}
+
+\newcommand{\gc}[1]{{ \it \color{green} (#1) } }
 
@@
 There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
-All else being equal the evolution of convergent phenotypes within a species convergence is more likely with a higher mutational input, i.e.
-when the mutation target and population sizes are larger \citep{}.
-Convergence is also more likely when standing variation is present for the phenotype, 
+All else being equal the evolution of convergent phenotypes is more
+likely with a 
+higher mutational input, i.e. when the mutation target and population sizes are larger \citep{}. 
+Standing variation for a trait can also increase the probability of convergence, 
 as multiple pre-existing alleles can spread from low frequency, 
 if our now adaptive alleles were previously not too deleterious \citep{Orr,Hermission}. 
-When multiple rare changes sweep up from low frequency in parallel within a population this has been termed a soft sweep.\\
-
 The geographic distribution of populations can also shape the probability of parallel mutation within a species.
@@
 is geographically limited, as populations will adapt via new mutations rather than waiting for migration \citep{RalphCoop}.
-Convergence is also more likely if geographically separated populations adapting to similar environments 
-via alleles which are maladapted in the environmentals separating them. 
-Such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{lenormand} 
-and so might encourage local adaptation by parallel mutation rather than migration.\\
+
+%When multiple rare changes sweep up from low frequency in parallel within a population this has been termed a soft sweep.\\
+
+Intuitively, convergence is also more likely when geographically
+separated populations adapt to ecologically similar conditions. The
+probability of convergent adaptation rather than populations adapting
+through allelic variants shared by migration, will be enhanced if
+these adaptive alleles are maladapted in the environmentals separating
+the populations. Such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{slatkin} 
+and so might encourage local adaptation by parallel adaptation rather than migration.\\
 
@@
 at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops. 
-This situation is repeated in XXXX where darker colored mice
- have repeated evolved lighter pigmentation to match lighter color sands on
+This situation is repeated in XXXX where darker colored mice have repeated evolved lighter pigmentation to match lighter color sands on
  beaches via different genetic routes.
@@
 
-In \citep{RalphCoop} we formulated a simple model of parallel adaptation in a geographic setting
-where the selection pressure was constant across the entire environment, 
-and that a single mutational change was sufficient to adapt populations to the novel environment. 
-We assumed that there was no standing variation for the adaptive allele, such that parallel mutation must be due to multiple mutations after the environmental shift.
+%In \citep{RalphCoop} we formulated a simple model of parallel adaptation in a geographic setting
+%where a continuous population was subjected to a selection pressure
+%that was constant across the entire environment.
+
+%and that a single mutational change was sufficient to adapt populations to the novel environment. 
+%We assumed that there was no standing variation for the adaptive allele, such that parallel mutation must be due to multiple mutations after the environmental shift.
 In that setting we found that there was a characteristic geographic scale, over which we would expect multiple instances of our adaptive allele to have arisen in parallel.
 This characteristic length could be expressed in terms of a simple compound parameter determined by our parameters of interest.
-Recall questions of origin of adaptations
+%Recall questions of origin of adaptations
 
-References on parallel adapation, standing variation, etc.:
-  coop,
-  pennings and hermisson.
-Note pennings and hermisson found standing deleterious variation was very important.
+%References on parallel adapation, standing variation, etc.:
+ % coop,
+ % pennings and hermisson.
+%Note pennings and hermisson found standing deleterious variation was very important.
 
-recall peromyscus examples with patchy environments
+%recall peromyscus examples with patchy environments
 
-Summarize previous paper
+%Summarize previous paper
 
-Question assumptions of no standing variation and homogenous selection
+%Question assumptions of no standing variation and homogenous selection
 
@@
 $\mu$ per chromosome per generation. We assume that this initially rare new
-allele has fitness $1+s_b$ relative to the unmutated type ($b$) in the ``poor'' habitat patches,
+allele has fitness $1+s_B$ relative to the unmutated type ($b$) in the ``poor'' habitat patches,
 and assume that it has fitness $1+s_m$ when rare in the intervening areas, with
@@
 
-As only a single mutation is required to adapt individuals to the
-poor habitat patch, subsequent mutations that arise after an allele
-becomes established at its equilibrium frequency on the patch gain no selective
-benefit. Similarly an allele introduced into a patch by migration will
-not be favored by selection to spread, if the patch has already been
-colonized. Thus mutations selective exclude each other from
-patches, over short-time scales, and they will only slowly filter
-together over longer time scales by drift.
-
 \paragraph{Previous work} We will make use of a number previous
@@
 physical width of the patch is less than $2 \tan^{-1}
-(\sqrt{|s_m/s_b|})$, then $B$ can not 
+(\sqrt{|s_m/s_B|})$, then $B$ can not 
 become stably established within the patch, as the local adaptation prevented by
@@
 Barton also showed that mutations appearing within the patch have probability 
-strictly less but approaching that $2s_b$ of establishment as the size
+strictly less but approaching that $2s_B$ of establishment as the size
 of the patch increased. Thus for large enough patches the probability
 of establishment of a locally favorable allele approaches the standard
-establishment probability ($2s_b$) for a panamitic population
+establishment probability ($2s_B$) for a panamitic population
 \citep[][]{haldane, fisher}, a result that also (generally) holds for
@@
 %Maruyama: On the fixation probability of mutant genes in a subdivided population*
+%should we cite Cherry http://www.ncbi.nlm.nih.gov/pmc/articles/PMC1462507/
 
@@
 
-\subsubsection{Mutational influx}
+\subsubsection{Time to the establishment of a locally adaptive allele
+  due to mutational influx}
 \label{ss:patchymutation}
@@
 that our patch is considerably larger than the cutoff for local
-establishment, i.e. $\sqrt{A} \gg 2 \tan^{-1} (\sqrt{|s_m/s_b|})$.
+establishment, i.e. $\sqrt{A} \gg 2 \tan^{-1} (\sqrt{|s_m/s_B|})$.
 
@@
 relative to $\sigma/\sqrt{2s_m}$ then a reasonable approximation is to
-set $p_f \approx 2 s_b A$. This follows from the fact that if the
+set $p_f \approx 2 s_B A$. This follows from the fact that if the
 patch is large than the majority of the mutations that could become
@@
 
-The rate at which mutations arise and colonize the patch is on the order of $\rho \mu
-p_f$ generations, which will approximately be $2 s_b \rho \mu$. If
-this rate is low,  then the waiting time till a mutation arises that
+The rate at which mutations arise and colonize the patch is
+\begin{equation}
+\lambda_{Mut} = \rho \mu p_f \approx 2 s_B \rho \mu.
+\end{equation}
+
+If this rate is low,  then the waiting time till a mutation arises that
 will become locally established within the patch is exponentially
-distributed with mean $2 s_b \rho \mu$.  Assuming that once a mutation
+distributed with rate $\lambda_{Mut}$.  Assuming that once a mutation
 becomes established
@@
  expected time it takes for our patch to become colonized by the $B$
- allele is $1/(2 s_b \rho \mu)$.
+ allele is $T_{Mut} \approx 1/ \lambda_{Mut} $.
 
-\subsubsection{Migration rates}
+\subsubsection{Time to the establishment of a locally adaptive allele
+  due to migrational influx}
 \label{ss:patchymigration}
-
 Now suppose that there are two patches, each of area $A$, of the poor
 habitat type separated by distance $R$. Consider the case where the
-allele has arisen and become established in the first patch, but has not yet appeared in the second.
+$B$ allele has arisen and become established in the first patch, but has not yet appeared in the second.
 As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
@@
 whether the allele is a new mutation or is descended from the occupants of the first patch.
-Due to mutation--selection balance there will be a small number of alleles descended from those of the first patch (``migrants'') 
+Due to mutation--selection balance there will be a small number of $B$ alleles descended from those of the first patch (``migrants'') 
 present in the regions where they are disadvantageous;
@@
 although the equations are again not particularly helpful.
-However, we do know that if $q(r)$ is the expected proportion present at mutation--selection balance at distance $r$ from the first patch
+However, we do know that if $q(r)$ is the expected proportion present at \gc{migration}--selection balance at distance $r$ from the first patch
 in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
-where $C$ is a constant depending on the structure of the populations and the selection coefficients.
+where $C$ is a constant depending on the structure of the populations
+and the selection coefficients. \gc{comment on size of C? Should we
+  also use this approx. to get the approx. rate of establishing mutations
+  arising outside of the patch, above. Feel like in the mutation rate
+  part we could pull $s_B$ out front of all rates.}
+
+\gc{Below are we setting $C=1$?}
 
 Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
-supposing that patches are large enough for local fixation to be
-stable. If the alleles were not advantageous in the second patch,
-the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |\sqrt{s_m} R/\sigma| )$;
-so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |\sqrt{s_m} R/\sigma|)^{-1}$.
+supposing that patches are large enough for local fixation to be stable. If the alleles were not advantageous in the second patch,
+the average number of mutant alleles present in a small area $\Delta$
+near the second patch, due
+to migration from the first patch, would be around $\rho A \exp( -
+|\sqrt{s_m} R/\sigma| ) \Delta$. Assuming that the patch is
+sufficiently large to give an allele a $2s_B$ probability of becoming established, the rate at which migrants find
+themselves in the patch and become established  is approximately
+\begin{equation}
+\lamda_{Mig,R} = \rho A^2 s_B \exp( -|\sqrt{s_m} R/\sigma| ) \Delta
+\end{equation}
+. \gc{check my squared here. Do we want to use $q(R)$ here?|
+If the establishment rate is low, the time it takes
+a mutation to arise that will become established will be  $\sim Exp(\lambda_{mig,R})$
+
+If  the time it takes for a mutation that it destined to 
+become established in the second patch to become establish is short,
+then  the expected waiting time until some migrant allele fixes in the
+second patch will be $T_{Mig} = \lamda_{Mig,R} $.
+
+\gc{Note that we can set patches sizes to be different}
+
+\gc{think we've covered this next part by $q(r)$ being number of migs. such
+that $2sq(r)$ is (for small $s$ the right answer.}
 If $T_m$ is large, then we would expect successful migrants to be rare,
@@
 
-\subsection{Patchy Selection -- Results} 
-\label{ss:discreteresults}
+\subsubsection{The probability of parallel adaptation between patches} 
+\label{ss:probparallel}
 
-We will assume that once a mutation escapes loss and becomes
-established in the poor habitat patch it reaches its equilbrium frequency very rapidly.
+As only a single mutation is required to adapt individuals to the
+poor habitat patch, subsequent mutations that arise after an allele
+becomes established at its equilibrium frequency on the patch gain no selective
+benefit. Similarly an allele introduced into a patch by migration will
+not be favored by selection to spread, if the patch has already been
+colonized. Thus, mutations selective exclude each other from
+patches, over short-time scales, and they will only slowly filter
+together over longer time scales by drift.
+
+We will assume that once a introduced $B$ allele (from migration or mutation) becomes
+established in the poor habitat patch it reaches its equilbrium
+frequency very rapidly. Assuming that we being from our geographically 
+spread population containing no copies of the $B$ allele,
+
+Under this approximation, and our selective
+exclusion assumption, the second patch will become colonized by migration or
+mutation, but not by both. This means that the question of whether the
+second patch adapt in parallel, or through a shared migrant allele,
+simply comes down to whether the a new mutation or a migrant allele
+becomes established first.
 
@@
 If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|\sqrt{s_m} R/\sigma| )$,
-then adaptation is expected to take place both by migration and mutation.
+then adaptation is expected to take place both by migration and
+mutation.
+
 If we take $\mu = 10^{-8}$, then this is satisfied if $R/\sigma \approx 18/\sqrt{s_m}$.
@@
 
-\begin{figure}[ht]
- \begin{center}
-   \includegraphics{discrete-by-ratio}
- \caption{ %
- The mean number of types and the mean number of demes occupied by a sampled type, in a discrete model with 20 demes,
- plotted as a function of the ratio of migration to mutation probabilities.
- Note the log scale on the horizontal axis.
- \label{fig:discbyratio}
- }
- \end{center}
-\end{figure}
+%\begin{figure}[ht]
+% \begin{center}
+ %  \includegraphics{discrete-by-ratio}
+ %\caption{ %
+ %The mean number of types and the mean number of demes occupied by a sampled type, in a discrete model with 20 demes,
+ %plotted as a function of the ratio of migration to mutation probabilities.
+% Note the log scale on the horizontal axis.
+% \label{fig:discbyratio}
+ %}
+ %\end{center}
+%\end{figure}
 

### 2013-01-24 patchy-selection-paper.tex
*Empty MediaWiki Message*

@@
 \begin{equation}
-\lamda_{Mig,R} = \rho A^2 s_B \exp( -|\sqrt{s_m} R/\sigma| ) \Delta
+\lamda_{Mig,R} = \rho A^2 s_B \exp( -|\sqrt{s_m} R/\sigma| ).
 \end{equation}
-. \gc{check my squared here. Do we want to use $q(R)$ here?|
-If the establishment rate is low, the time it takes
+\gc{check my squared here. Do we want to use $q(R)$ here?|
+If this establishment rate is low, the time it takes
 a mutation to arise that will become established will be  $\sim Exp(\lambda_{mig,R})$

### 2013-01-24 patchy-selection-paper.tex
/* Patchy Selection paper */

@@
 (\sqrt{|s_m/s_B|})$, then $B$ can not 
-become stably established within the patch, as the local adaptation prevented by
+become \gc{fixed/at a stable equilibrium/?} within the patch, as the local adaptation prevented by
 migrational swamping \citep[see also][ for a review]{Lenormand}.
@@
  framework \citep[see also the work of][]{Polk}. 
-For patches above this size, \citet{barton1987establishment}
+For patches above this critical size, \citet{barton1987establishment}
  founds that mutations appearing further than a few multiples of
@@
 which no $B$ allele has yet become established. As
-we are interested in local adaptation that can occur we will assume
-that our patch is considerably larger than the cutoff for local
-establishment, i.e. $\sqrt{A} \gg 2 \tan^{-1} (\sqrt{|s_m/s_B|})$.
+we are interested in patchess where local adaptation can occur we will assume
+that our patch is larger than the cutoff for local
+establishment, i.e. $\sqrt{A} > 2 \tan^{-1} (\sqrt{|s_m/s_B|})$.
 
@@
 \end{equation}
-
 If this rate is low,  then the waiting time till a mutation arises that
@@
 distributed with rate $\lambda_{Mut}$.  Assuming that once a mutation
-becomes established
- it quickly reaches its equilibrium frequency across the patch, the
+becomes established it quickly reaches its equilibrium frequency across the patch, the
  expected time it takes for our patch to become colonized by the $B$
@@
   arising outside of the patch, above. Feel like in the mutation rate
-  part we could pull $s_B$ out front of all rates.}
+  part we could pull $s_B$ out front of the rate as well.}
 
@@
 near the second patch, due
-to migration from the first patch, would be around $\rho A \exp( -
+to migration from the first patch, would be $\approx \rho A \exp( -
 |\sqrt{s_m} R/\sigma| ) \Delta$. Assuming that the patch is
@@
 \end{equation}
-\gc{check my squared here. Do we want to use $q(R)$ here?|
+\gc{Is this $A^2$ right here?. Do we want to use $q(R)$ here?}
 If this establishment rate is low, the time it takes
 a mutation to arise that will become established will be  $\sim Exp(\lambda_{mig,R})$
-
-If  the time it takes for a mutation that it destined to 
-become established in the second patch to become establish is short,
+Assuming that the time it takes for a allele from the first patch to
+become established in the second patch is short,
 then  the expected waiting time until some migrant allele fixes in the
-second patch will be $T_{Mig} = \lamda_{Mig,R} $.
+second patch will be $T_{Mig} = 1/\lamda_{Mig,R} $.
 
 \gc{Note that we can set patches sizes to be different}
+
+\gc{Put time it takes for a migrant allele to move between patches
+  here? Or below }
 
@@
 
-\subsubsection{The probability of parallel adaptation between patches} 
+\subsubsection{The probability of parallel adaptation between patches.} 
 \label{ss:probparallel}
+We now turn to the question of whether our populations adapt in
+parallel genetic changes or by the spread of a migrant allele between the
+patches. These two scenerios are assumed to be distinguishable, either
+because in the case of parallelism the mutations occur at different
+positions or if the mutations occur at the same position they can be
+distinguished due to their haplotypic background.
+
+Assuming that we begin with our geographically 
+spread population containing no copies of the $B$ allele, one or other
+of our poor habitat patches will be the first to adapt via the
+establishment of a $B$ allele. The question
 
@@
 established in the poor habitat patch it reaches its equilbrium
-frequency very rapidly. Assuming that we being from our geographically 
-spread population containing no copies of the $B$ allele,
+frequency very rapidly. Under this approximation, and our selective
+exclusion assumption, after one of the patches becomes colonized by a
+single $B$ allele, the other patch will become colonized by migration or
+mutation, but not by both. As such, the question of whether the
+population adapts to the second poor habitat patch in parallel, simply comes down to whether the a new mutation or a migrant allele
+is the first to become established in the second patch.
 
-Under this approximation, and our selective
-exclusion assumption, the second patch will become colonized by migration or
-mutation, but not by both. This means that the question of whether the
-second patch adapt in parallel, or through a shared migrant allele,
-simply comes down to whether the a new mutation or a migrant allele
-becomes established first.
-
-If environment is continuous, with separated patches where the new allele is advantageous,
-we can only get a rough sense of whether the new adaptation is likely to spread through mutation (i.e.\ in parallel)
-or through migration.
 Taking the expressions from sections \ref{ss:patchymutation} and \ref{ss:patchymigration} above,
 we see that if patches each have sufficiently large area $A$,
-then the mutational input in each is on the order of $2 \rho \mu A$,
-while the influx due to migration between patches separated by distance $R$ is on the order $2 \rho A \exp(- |\sqrt{s_m} R/\sigma|)$. 
+then the mutational input in each is on the order of $2 s_B \rho \mu A$,
+while the influx due to migration between patches separated by
+distance $R$ is on the order $2 s_B \rho A \exp(- |\sqrt{s_m}
+R/\sigma|)$ \gc{check $A^2$}. 
 If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|\sqrt{s_m} R/\sigma| )$,
@@
 then patches are effectively isolated -- most adaptation will be independent between patches.
-On the other hand, if the separation is much smaller, then a single mutation is likely to fix everywhere.
+On the other hand, if the separation is much smaller, then a single
+mutation is likely to fix in both patches.
 
-Furthermore, in the model of discrete demes connected by symmetric weak migration, 
-the final distribution of types depends only on the ratio of mutation to migration rates,
-and that the number of independently arisen types follows the Ewens distribution, with parameter $\mu/m$.
-In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
-for a model with 20 interconnected demes, across different parameter values.  
-This indicates that we need the population-scaled mutation and migration rates 
-to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.  
-If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
+If we are willing to take these approximations at face
+value, and assuming both rates are small, the probability of parallel
+adaptation is 
+\begin{equation}
+\frac{\mu}{\mu + \exp(- |\sqrt{s_m} R/\sigma|}
+\end{equation}
+so that the probability of parallel mutation should increase
+approximately logistically with the distance between the patches at a
+rate $\sqrt{s_m} /\sigma$. However, we caution that these calculations
+are rough and so should be taken wi
+
+%Furthermore, in the model of discrete demes connected by symmetric weak migration, 
+%the final distribution of types depends only on the ratio of mutation to migration rates,
+%and that the number of independently arisen types follows the Ewens distribution, with parameter $\mu/m$.
+%In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
+%for a model with 20 interconnected demes, across different parameter values.  
+%This indicates that we need the population-scaled mutation and migration rates 
+%to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.  
+%If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
 

### 2013-01-25 patchy-selection-paper.tex
/* Patchy Selection paper */

@@
 
+\gc{work in a parameter that the the migrant outflux of a patch, and
+  explain that depends on geometry }
+
 Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
@@
 near the second patch, due
-to migration from the first patch, would be $\approx \rho A \exp( -
+to migration from the first patch, would be $\approx \rho E \exp( -
 |\sqrt{s_m} R/\sigma| ) \Delta$. Assuming that the patch is
@@
 \begin{equation}
-\lamda_{Mig,R} = \rho A^2 s_B \exp( -|\sqrt{s_m} R/\sigma| ).
+\lambda_{Mig,R} = \rho E s_B \exp( -|\sqrt{s_m} R/\sigma| ).
 \end{equation}
@@
 while the influx due to migration between patches separated by
-distance $R$ is on the order $2 s_B \rho A \exp(- |\sqrt{s_m}
-R/\sigma|)$ \gc{check $A^2$}. 
+distance $R$ is on the order $2 s_B \rho E \exp(- |\sqrt{s_m}
+R/\sigma|)$. \gc{what to assume about A and $E$ here?}
 If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|\sqrt{s_m} R/\sigma| )$,
@@
 If we are willing to take these approximations at face
-value, and assuming both rates are small, the probability of parallel
+value, and assuming both rates are small, the the time till the first
+patch is colonized will be approximately exponentially
+distributed with parameter $2 \lambda_{Mut}$. Then time till the
+second patch is subsequently colonized by the $B$
+allele (via either migration or new mutation) will be exponentially
+distributed with parameter $\lambda_{Mut}+\lambda_{Mig,R}$. Both of
+these times scale linearly with $1/(\rho s_{B})$, so that
+
+then probability of parallel
 adaptation is 
 \begin{equation}
-\frac{\mu}{\mu + \exp(- |\sqrt{s_m} R/\sigma|}
+\frac{\lambda_{Mut}}{\lambda_{Mut} + \lambda_{Mig,R}} = \frac{\mu (A/E)}{\mu (A/E) + \exp(- |\sqrt{s_m} R/\sigma|}
 \end{equation}
@@
 rate $\sqrt{s_m} /\sigma$. However, we caution that these calculations
-are rough and so should be taken wi
+are rough and so should be treated mainly as a rule of thumb.
+
+We can extend this reasoning to multiple patches ($>2$) within our population's range
+where our $B$ allele can facilitate local adaptation. Under our
+assumptions of mutational exclusion and rapid establishment, each one
+of these patches will adapt through a single copy of the $B$
+alllele. The occupation of our different patches by the $B$ allele will form a
+discrete state markov chain, where the states of the chain are which patches have
+been colonized so far by $B$, and which colonized patches have shared
+versus distinct copy of $B$.
+
+For simplicity we will assume that each of these patches has
+the same size and geometry (i.e. same $E$). A particular patch adapts in parallel through an independent
+instance of the $B$ mutation at rate $2 s_B \mu \rho A$.
+While, if patches $1$ through $k$ are already occupied by
+established copies of the $B$ allele, and these patches are at
+distances $R_1,~\cdots,~R_k$ away from our patch, the rate of
+adaptation through migration in our patch is 
+\begin{equation}
+\sum_{i=1}^{k} 2 s_B \rho E \exp(- |\sqrt{s_m} R_k/\sigma|}.
+\end{equation}
+Once again the compound parameter $2s_B \rho$ will function to
+scale the times between events (parallel adaptation or the spread of a
+migrant), and so is critical to the time-scale over which
+adaptation occurs. However, this compound parameter again has no effect on
+the final configuration of the pattern of which patches have adapted
+in parallel versus have shared genetic basis to their adaptive
+response to the poor environment.
+
+The states of this
 

### 2013-01-25 patchy-selection-paper.tex
/* Patchy Selection paper */

@@
 
-The covergent evolution of similar phenotypes in response to shared selection pressures is a testiment to the power of selection to repeatedly sculpt phenotypic variation.
-In some cases this convergence extends to the molecular level with dispate taxa converging to the same phenotype through parallel genetic changes in the same pathway, genes, or even by precisely the same genetic changes.
-Such precise convergence points to the conservation of function of the genes underlying adaptive traits, sometimes over deep time scales \citep{deephomologypapers}.
+The covergent evolution of similar phenotypes in response to shared
+selection pressures is a testiment to the power of selection to
+repeatedly sculpt phenotypic variation. 
+In some cases this convergence extends to the molecular level with
+dispate taxa converging to the same phenotype through parallel genetic
+changes in the same pathway, genes, 
+or even by precisely the same genetic changes. Such precise
+convergence points to the conservation of function of the genes
+underlying adaptive traits, 
+sometimes over deep time scales \citep{deephomologypapers}.
 Convergence on a molecular level also suggests that the path taken by adaptation can sometimes be relatively constained as few changes could have resulted in the adaptive phenotype, or that relatively few changes that can confer the advantagous phenotype of interest were sufficiently free of deleterious pleiotropic consequences to contribute to adaptation.
@@
 has probability approximately $p(x)$ of fixing locally in this patch,
-whether the allele is a new mutation or is descended from the occupants of the first patch.
-Due to mutation--selection balance there will be a small number of $B$ alleles descended from those of the first patch (``migrants'') 
+whether the allele is a new mutation or is descended from the
+occupants of the first patch.
+
+For the moment we will ignore the
+possibility of mutation generating a $B$ allele can concentrate on the
+flow of migrants from patch $1$ to patch $2$.
+Due to \gc{migration}--selection balance there will be a small number of $B$ alleles descended from those of the first patch (``migrants'') 
 present in the regions where they are disadvantageous;
@@
 then  the expected waiting time until some migrant allele fixes in the
-second patch will be $T_{Mig} = 1/\lamda_{Mig,R} $.
+second patch will be $T_{Mig} = 1/ \lambda_{Mig,R} $.
 

### 2013-01-31 branching-transit.tex
WorkingWiki replaced project description

\documentclass{article}
\usepackage{fullpage} 
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\var}{\mathop{\mbox{Var}}}

\begin{document}

Consider a largeish occupied patch of habitat in which an allele is beneficial ($s_b$),
surrounded by habitat where it is deleterious ($s_d$).
The allele is present in surrounding areas at low frequency due to migration-selection balance.
We want the first time that another patch at distance $x$ is successfully colonized.

The general picture:
\begin{itemize}

\item The excursions sufficiently far from the patch are like subcritical branching processes,
with mean $1-s_d$ and variance $\xi^2$.

\item An excursion that makes it to the new patch in $k$ copies establishes with probability $1-(1-p_s)^k$,
where $p_s \approx 2s_b/\xi^2$, so establishment has approximate probability $2 k s_b / \xi^2$.

\item The probability that the branching process lives for $t$ generations is asymptotically a constant times $(1-s_d)^t$.

\item Conditioned on living for time $t$, the process can be represented by a single lineage whose offspring distribution
is the original distribution, conditioned to be nonzero, minus one.
(Also, it is known that conditional on being alive at $t\to\infty$, the allele is present in $K$ copies; where $K$ has calculable mean and variance.)

\item Spatial motion is independent of branching;
so we can take th emotion of the trunk lineage to be Brownian motion;
the chance that the motion first hits $x$ at time $t$ is (Borodin \& Salminen 2.02)
\[
\frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) .
\]

\item Therefore, the probability that an excursion makes it to $x$ at time $t$ is
\[
  (1-s_d)^t \frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) ,
\]
or the chance that it is still alive then time the chance the trunk lineage hits $x$ at that time.

\item This induces a distribution on $T$, the time that a successful excursion took to cross,
and hence the length of the hitchhiking segment.

\end{itemize}

From Jagers, section 2.6,
the mean and variance of $K$ are as follows.
Let the probability of $k$ offpsring be $p_k$ and let
\[
 f(u) =  \sum_k p_k u^k,
\]
and $f_n$ the $n^\mathrm{th}$ composition of $f$.
Note that the mean number of offspring is $1-s_b = f'(1)$, and the variance is $\xi^2 = f''(1) + f'(1)(1-f'(1)) = f''(1) + s_d(1-s_d)$.
If $\sum_k p_k k \log k < \infty$, then 
\[
\phi = \lim_{n \to \infty} \frac{ 1-f_n(0) }{ (1-s_d)^n } > 0 .
\]
Furthermore,
\begin{gather*}
\E[K] = 1/\phi \qquad \mbox{and} \\
\E[K(K-1)] = \frac{ f''(1) }{ \phi s_d (1-s_d) } \qquad \mbox{so} \\
\var[K] = \frac{1}{\phi} \left( \frac{ \xi^2 }{ s_d (1-s_d) } + \frac{1}{1-\phi} - 1 \right) .
\end{gather*}

\end{document}

### 2013-02-02 branching-transit.tex
WorkingWiki importation

@@
 
-\item The probability that the branching process lives for $t$ generations is asymptotically a constant times $(1-s_d)^t$.
+\item The probability that the branching process lives for $t$ generations is asymptotically a constant ($\phi$, below) times $(1-s_d)^t$.
 
@@
 is the original distribution, conditioned to be nonzero, minus one.
-(Also, it is known that conditional on being alive at $t\to\infty$, the allele is present in $K$ copies; where $K$ has calculable mean and variance.)
+(Also, it is known that conditional on being alive at $t\to\infty$, the allele is present in $K$ copies; where $K$ has mean $1/\phi$ and variance calculated below.)
 
 \item Spatial motion is independent of branching;
-so we can take th emotion of the trunk lineage to be Brownian motion;
+so we can take the motion of the trunk lineage to be Brownian;
 the chance that the motion first hits $x$ at time $t$ is (Borodin \& Salminen 2.02)
@@
 
-\item Therefore, the probability that an excursion makes it to $x$ at time $t$ is
+\item Therefore, the probability that an excursion makes it to $x$ at time $t$ is asymptotically
 \[
-  (1-s_d)^t \frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) ,
+  (1-s_d)^t \phi \frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) ,
 \]
@@
 
-\item This induces a distribution on $T$, the time that a successful excursion took to cross,
-and hence the length of the hitchhiking segment.
+\item This induces a distribution on $T$, the time that a successful excursion took to cross:
+assuming, as is asymptotically correct, that the number of arriving lineages is independent of how long it took to get there,
+if $\tau_x$ is the random time of arrival of a successful lineage,
+\[
+ \P\{\tau_x \in dt\} \propto \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) ,
+\]
+and so
+
+\item Hence the length of the hitchhiking segment on \emph{one side} of the selected site $L$ has distribution
+(using Gradshteyn \& Ryzhik 3.471.9; note $\ell > 0 > \log(1-s_d)$):
+\begin{align*}
+ \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] \\
+ & \propto \int_0^\infty \frac{1}{t^{3/2}} \exp\left(-\frac{x^2}{2t} + t(\log(1-s_d) -\ell) \right)  dt \\
+ &= \frac{\sqrt{2\pi}}{x} e^{-x\sqrt{2(\ell-\log(1-s_d))}} ;
+\end{align*}
+hence, defining $\alpha = - \log(1-s_d) > 0$,
+\begin{align*}
+ \P\{ L > \ell \} &= \exp\left\{{-x\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} \qquad \mbox{for}\; \ell>0.
+\end{align*}
+This is a shifted Weibull distribution, or equivalently, $L=X^2-\alpha$, where $X$ is a Exponential($x\sqrt{2}$) conditioned to be at least $\sqrt{\alpha}$.
+Note that the total length of a hitchhiking segment is the sum of two independent copies of this.
 
 \end{itemize}
+
+\subsection{$K$ and $\phi$}
 
@@
 and $f_n$ the $n^\mathrm{th}$ composition of $f$.
-Note that the mean number of offspring is $1-s_b = f'(1)$, and the variance is $\xi^2 = f''(1) + f'(1)(1-f'(1)) = f''(1) + s_d(1-s_d)$.
+Note that the mean number of offspring is $1-s_d = f'(1)$, and the variance is $\xi^2 = f''(1) + f'(1)(1-f'(1)) = f''(1) + s_d(1-s_d)$.
 If $\sum_k p_k k \log k < \infty$, then 
 \[
-\phi = \lim_{n \to \infty} \frac{ 1-f_n(0) }{ (1-s_d)^n } > 0 .
+\phi = \lim_{n \to \infty} \frac{ \P\{ \mbox{survival to }n \}  }{ (1-s_d)^n } = \lim_{n \to \infty} \frac{ 1-f_n(0) }{ (1-s_d)^n } > 0 .
 \]
@@
 \end{gather*}
+Note additionally that 
+\[
+    1 - \frac{1-f_n(u)}{1-f_n(0)} \searrow g(u) = \E[u^K] .
+\]
 

### 2013-02-03 branching-transit.tex
WorkingWiki importation

@@
 \newcommand{\var}{\mathop{\mbox{Var}}}
+\newcommand{\Exp}{\mathop{\mbox{Exp}}}
 
@@
 \[
- \P\{\tau_x \in dt\} \propto \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) ,
+ \P\{\tau_x \in dt\} \propto \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) .
 \]
-and so
 
-\item Hence the length of the hitchhiking segment on \emph{one side} of the selected site $L$ has distribution
+\item The length of the hitchhiking segment on \emph{one side} of the selected site $L$ is exponential with rate depending on $T$;
+integrating over $T$ we get
 (using Gradshteyn \& Ryzhik 3.471.9; note $\ell > 0 > \log(1-s_d)$):
@@
 \begin{align*}
- \P\{ L > \ell \} &= \exp\left\{{-x\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} \qquad \mbox{for}\; \ell>0.
+ \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] = \exp\left\{{-x\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} \qquad \mbox{for}\; \ell>0.
 \end{align*}
-This is a shifted Weibull distribution, or equivalently, $L=X^2-\alpha$, where $X$ is a Exponential($x\sqrt{2}$) conditioned to be at least $\sqrt{\alpha}$.
+This is a shifted Weibull distribution, or equivalently, $L=X^2-\alpha$, where $X$ is a Exponential with rate $x\sqrt{2}$ conditioned to be at least $\sqrt{\alpha}$.
 Note that the total length of a hitchhiking segment is the sum of two independent copies of this.
+
+\item We could rescale with $x$ and get a limiting distribution for $\tau_x$:
+\begin{align*}
+  \E\left[\exp\left( -\ell \frac{\tau_x}{x} \right)\right]
+        &= \exp\left\{ - x\sqrt{2}\left(\sqrt{\frac{\ell}{x}+\alpha} - \sqrt{\alpha}\right) \right\} \\
+        &\xrightarrow{x\to\infty} \exp\left\{ - \frac{\ell}{\sqrt{2\alpha}} \right\} ,
+\end{align*}
+implying that $\tau_x/x \xrightarrow{a.s.} 1/\sqrt{2\alpha}$.
+
+\item Ok, but how about the CLT?  Differentiating, we see that $\E[\tau_x]=x/\sqrt{2\alpha}$ and $\var[\tau_x]=x/2^{3/2}\alpha^{3/2}$,
+which suggests looking at
+\begin{align*}
+\E\left[\exp\left( -\ell \left(\frac{\tau_x-x/\sqrt{2\alpha}}{\sqrt{x}} \right) \right)\right]
+    &= \exp\left\{ - x\sqrt{2}\left(\sqrt{\frac{\ell}{\sqrt{x}}+\alpha} - \sqrt{\alpha}\right) + \frac{\ell}{\sqrt{2\alpha x}} \right\} \\
+    &\simeq \exp\left\{ - x\sqrt{2}\left( \frac{\ell}{2\sqrt{\alpha x}} - \frac{\ell^2}{8\alpha^{3/2}x}  \right) + \frac{\ell}{\sqrt{2\alpha x}} \right\} \\
+    &= \exp\left\{ \frac{\ell^2}{2^{5/2}\alpha^{3/2}} \right\} ,
+\end{align*}
+since $\sqrt{\alpha+\epsilon}=\sqrt{\alpha} +\epsilon/2\sqrt{\alpha}-\epsilon^2/8\alpha^{3/2} + O(\epsilon^3)$.
+The Laplace transform of a $N(0,\sigma^2)$ is $\exp(\ell^2\sigma^2/2)$, so
+\[
+\frac{\tau_x - x/\sqrt{2\alpha}}{\sqrt{x}} \xrightarrow{d} N(0,2^{-3/2}\alpha^{-3/2}) \qquad \mbox{as}\;x\to\infty.
+\]
+
+\item Translating this limit law back to $L$, we have that
+\[
+  \tau \simeq \frac{x}{2\alpha} + 2\sqrt{2\alpha^3x} B,
+\]
+with $B \sim N(0,1)$; since $L \sim \Exp(\tau)$,
+\[
+  \frac{x}{\sqrt{2\alpha}} L \simeq \Exp\left(1+\frac{4\alpha^2}{\sqrt{x}}B\right) \xrightarrow{x\to\infty} \Exp(1),
+\]
+i.e.\ $L$ is approximately a Exponential with rate $\sqrt{2\alpha}/x$.
+
+\item Note: if we wanted to kill the motion at some negative location (returns to original patch)
+we'd use the result for BM, on p.296 of Feller.
 

### 2013-02-06 Spatial_adaptation/local-adaptation.tex
/* Local Adaptation */

@@
 \[
-    \partial_t p(t,x) = \sigma^2 \partial_x^2 p(t,x) + s(x) p(t,x) (1-p(t,x)) .
+\partial_t p(t,x) = \frac{1}{2} \sigma^2 \partial_x^2 p(t,x) + s(x) p(t,x) (1-p(t,x)) .
 \]
@@
 \begin{align} \label{eqn:definingphi}
-    \partial_x^2 \phi(x) = - s(x) \phi(x) (1-\phi(x)) /sigma^2,
+    \partial_x^2 \phi(x) = - 2 s(x) \phi(x) (1-\phi(x)) /\sigma^2,
 \end{align}
 with appropriate boundary conditions.
-First rescale space by $\sigma$ so that it dissappears.
+First rescale space by $\sigma/\sqrt{2}$ so the $\sigma^2/2$ term dissappears.
 If we now assume that $s(x)$ is piecewise constant,
 $s(x) = s_i$ for $x \in [x_i,x_{i+1})$, with $x_0=-\infty$ and $x_{n+2}=\infty$,
-then the equation is integrable: if we multiply through by $\partial_x \phi(x)$ and integrate, then we get that
+then the equation is integrable: if we multiply through by $2\partial_x \phi(x)$ and integrate, then we get that
 \begin{align} \label{eqn:conservation}
-    ( \partial_x \phi(x) )^2  &= - \int^{x} s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
-        &= - \frac{1}{2} s_i \phi(x)^2 \left( 1 - \frac{1}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) \\
+    ( \partial_x \phi(x) )^2  &= - \int^{x} 2 s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
+        &= - s_i \phi(x)^2 \left( 1 - \frac{2}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) \\
         &:= - V_i(\phi(x)) \quad \mbox{for } x \in [x_i,x_{i+1}) ,
@@
 \[
-        V_i(\phi) = \frac{1}{2} s_i \phi^2 \left( 1 - \frac{1}{3} \phi \right) + K_i .
+        V_i(\phi) = s_i \phi^2 \left( 1 - \frac{2}{3} \phi \right) + K_i .
 \]
-Note that $V_i'(0)=V_i'(1)=0$, that $V_i(0)=K_i$ and $V_i(1) = K_i-s_i/6$.  
+Note that $V_i'(0)=V_i'(1)=0$, that $V_i(0)=K_i$ and $V_i(1) = K_i+s_i/3$.  
 We will always have that $V(\phi) \le 0$.
@@
 \[
-    K_{i+1} - K_i = (s_i - s_{i+1}) \frac{1}{2} \phi(x_{i+1})^2 (1-\phi(x_{i+1})/3) .
+    K_{i+1} - K_i = (s_i - s_{i+1}) \phi(x_{i+1})^2 (1-2\phi(x_{i+1})/3) .
 \]
@@
 This from ``NEQwiki, the nonlinear equations encyclopedia'', \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
-Recall that $V_i(\phi) = (1/2) s_i \phi^2(1-\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries.
+Recall that $V_i(\phi) =  s_i \phi^2(1-2\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries.
 Note that if we rescale space by $\sqrt{s}/\sigma$,
@@
     \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
-        \sqrt{ 2 } \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } ,
+         \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } .
 \]
-with $C=2K$.
-Let $\phi^2(1-\phi/3)+C = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
+Let $\phi^2(1-2\phi/3)+K = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
 and change variables first to $y^2=(\alpha-\phi)$, 
@@
 \begin{align*}
-    \frac{ d\phi }{ \sqrt{ \phi^2 (1-\phi/3) + C } } 
+    \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } 
         = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-y^2) } } \\
@@
 \begin{align*}
-    \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{s} / \sigma + C ) \right) \\
-        &\simeq C' \exp\left( - \frac{ x \sqrt{s} }{ \sigma } \right) .
+    \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / \sigma + C ) \right) \\
+        &\simeq C' \exp\left( - \frac{ 2 x \sqrt{2s} }{ \sigma } \right) .
 \end{align*}
-Indeed, as $\phi \to 0$, we have approximately $\phi' = \frac{ \sqrt{s} }{ \sigma } \phi$.
+% Indeed, as $\phi \to 0$, we have approximately $\phi' = \frac{ \sqrt{s} }{ \sigma } \phi$.  % huh????
 

### 2013-02-07 branching-transit.tex
WorkingWiki importation

@@
 the chance that the motion first hits $x$ at time $t$ is (Borodin \& Salminen 2.02)
-\[
+\begin{align}
 \frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) .
-\]
+\end{align}
 
 \item Therefore, the probability that an excursion makes it to $x$ at time $t$ is asymptotically
-\[
+\begin{align}
   (1-s_d)^t \phi \frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) ,
-\]
+\end{align}
 or the chance that it is still alive then time the chance the trunk lineage hits $x$ at that time.
@@
 if $\tau_x$ is the random time of arrival of a successful lineage,
-\[
- \P\{\tau_x \in dt\} \propto \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) .
-\]
+\begin{align}
+ \P\{\tau_x \in dt\} = C_{s_d,x} \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) ,
+\end{align}
+where $C_{s_d,x}$ is a constant depending only on $s_d$ and $x$.
 
@@
 (using Gradshteyn \& Ryzhik 3.471.9; note $\ell > 0 > \log(1-s_d)$):
-\begin{align*}
+\begin{align}
  \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] \\
- & \propto \int_0^\infty \frac{1}{t^{3/2}} \exp\left(-\frac{x^2}{2t} + t(\log(1-s_d) -\ell) \right)  dt \\
- &= \frac{\sqrt{2\pi}}{x} e^{-x\sqrt{2(\ell-\log(1-s_d))}} ;
-\end{align*}
-hence, defining $\alpha = - \log(1-s_d) > 0$,
-\begin{align*}
+ & = C_{s_d,x} \int_0^\infty \frac{1}{t^{3/2}} \exp\left(-\frac{x^2}{2t} + t(\log(1-s_d) -\ell) \right)  dt \\
+ &= C_{s_d,x} \frac{\sqrt{2\pi}}{x} e^{-x\sqrt{2(\ell-\log(1-s_d))}} ; \label{eqn:first_integral}
+\end{align}
+hence, defining $\alpha = - \log(1-s_d) \approx s_d$ (note $\alpha>0$),
+\begin{align}
  \P\{ L > \ell \} &= \E\left[ e^{-\ell \tau_x} \right] = \exp\left\{{-x\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} \qquad \mbox{for}\; \ell>0.
-\end{align*}
+\end{align}
 This is a shifted Weibull distribution, or equivalently, $L=X^2-\alpha$, where $X$ is a Exponential with rate $x\sqrt{2}$ conditioned to be at least $\sqrt{\alpha}$.
@@
 \item We could rescale with $x$ and get a limiting distribution for $\tau_x$:
-\begin{align*}
+\begin{align}
   \E\left[\exp\left( -\ell \frac{\tau_x}{x} \right)\right]
@@
         &\xrightarrow{x\to\infty} \exp\left\{ - \frac{\ell}{\sqrt{2\alpha}} \right\} ,
-\end{align*}
+\end{align}
 implying that $\tau_x/x \xrightarrow{a.s.} 1/\sqrt{2\alpha}$.
@@
 which suggests looking at
-\begin{align*}
+\begin{align}
 \E\left[\exp\left( -\ell \left(\frac{\tau_x-x/\sqrt{2\alpha}}{\sqrt{x}} \right) \right)\right]
@@
     &= \exp\left\{ \frac{\ell^2}{2^{5/2}\alpha^{3/2}} \right\} ,
-\end{align*}
+\end{align}
 since $\sqrt{\alpha+\epsilon}=\sqrt{\alpha} +\epsilon/2\sqrt{\alpha}-\epsilon^2/8\alpha^{3/2} + O(\epsilon^3)$.
 The Laplace transform of a $N(0,\sigma^2)$ is $\exp(\ell^2\sigma^2/2)$, so
-\[
+\begin{align}
 \frac{\tau_x - x/\sqrt{2\alpha}}{\sqrt{x}} \xrightarrow{d} N(0,2^{-3/2}\alpha^{-3/2}) \qquad \mbox{as}\;x\to\infty.
-\]
+\end{align}
 
 \item Translating this limit law back to $L$, we have that
-\[
-  \tau \simeq \frac{x}{2\alpha} + 2\sqrt{2\alpha^3x} B,
-\]
+\begin{align}
+  \tau \simeq \frac{x}{2\alpha} + \sqrt{\frac{x}{8\alpha^3}} B,
+\end{align}
 with $B \sim N(0,1)$; since $L \sim \Exp(\tau)$,
-\[
+\begin{align}
   \frac{x}{\sqrt{2\alpha}} L \simeq \Exp\left(1+\frac{4\alpha^2}{\sqrt{x}}B\right) \xrightarrow{x\to\infty} \Exp(1),
-\]
+\end{align}
 i.e.\ $L$ is approximately a Exponential with rate $\sqrt{2\alpha}/x$.
@@
 \item Note: if we wanted to kill the motion at some negative location (returns to original patch)
-we'd use the result for BM, on p.296 of Feller.
+we'd use the result for BM killed at $-y$, say (Feller vol.2, XIV.5):
+First note that since we kill at rate $\alpha$, the Laplace transform of $\tau$ is proportional to the Laplace transform of the first passage time,
+shifted by $\alpha$, since $\E[e^{-\ell \tau}] \propto \int_0^\infty e^{-\alpha t} e^{-\ell t} p_x(t) dt$.
+The Laplace transform of the first passage time from $0$ to $x$, killed on hitting $-y$, is
+\begin{align}
+ \frac{ e^{y\sqrt{2\ell}} - e^{-y\sqrt{2\ell}} } { e^{(x+y)\sqrt{2\ell}} - e^{-(x+y)\sqrt{2\ell}} } .
+\end{align}
+Note this reduces to the answer above ($e^{-x\sqrt{2\ell}}$) as $y \to \infty$.
+
+\end{itemize}
+
+\subsection{Two dimensions}
+
+\begin{itemize}
+
+\item In two dimensions, if everything is radially symmetric, we need to replace simple Brownian motion by a two-dimensional Bessel process.
+Conditioning it to survive until $x$ is equivalent to giving it drift (cite?),
+so this reduces to finding the Laplace transform of the hitting time of a diffusion.
+I think this is in some Pitman \& Yor paper.
+
+\item Radial symmetry implies a big annulus of habitat around the central patch.
+If instead there are two smallish patches in two dimensions, 
+we need only multiply the one-dimensional case by the probability that the $y$-coordinate is within the required range,
+which is something involving the error function, or for large $x$,
+is the width of the patch multiplied by $1/(2 \pi \sigma^2 t)$.  
+This replaces the $t^{3/2}$ by a $t^{5/2}$, and back to G\&R, we have that the Laplace transform is in terms of $K_{3/2}$ instead of $K_{1/2}$;
+and this is found in G\&R 8.468, $K_{3/2}(z) = \sqrt{\pi/2z} e^{-z} (1+1/z)$.
+The formula corresponding to equation \eqref{eqn:first_integral} above is then
+\begin{align}
+  \E[e^{-\ell \tau_x}] \propto \sqrt{\pi} \frac{2\sqrt{\ell+\alpha}}{x^2} e^{-x\sqrt{2(\ell+\alpha)}} \left( 1 + \frac{1}{x\sqrt{2(\ell+\alpha)}} \right) ,
+\end{align}
+and so
+\begin{align}
+  \E[e^{-\ell \tau_x}] &= \frac{\sqrt{\ell+\alpha}}{\sqrt{\alpha}} e^{-x\sqrt{2}\left(\sqrt{\ell+\alpha}-\sqrt{\alpha}\right)} \left( 1 + \frac{1}{x\sqrt{2(\ell+\alpha)}} \right) \left( 1 + \frac{1}{x\sqrt{2(\alpha)}} \right)^{-1} \\
+  &= \frac{x\sqrt{2(\ell+\alpha)}+1}{x\sqrt{2\alpha}+1} e^{-x\sqrt{2}\left(\sqrt{\ell+\alpha}-\sqrt{\alpha}\right)} .
+\end{align}
+
+\item Differentiating the above, we get that $\E[\tau_x] = x^2 / (x\sqrt{2\alpha}+1) \sim x/\sqrt{2\alpha}$ and $\var[\tau_x] = x^3 / \sqrt{2\alpha}(x\sqrt{2\alpha}+1)^2 \sim x/(2\alpha)^{3/2}$; in fact, the same scaling produces the same limit law, since the $\ell/\sqrt{x}$ term dissappears.
 
@@
 Let the probability of $k$ offpsring be $p_k$ and let
-\[
+\begin{align}
  f(u) =  \sum_k p_k u^k,
-\]
+\end{align}
 and $f_n$ the $n^\mathrm{th}$ composition of $f$.
@@
 If $\sum_k p_k k \log k < \infty$, then 
-\[
+\begin{align}
 \phi = \lim_{n \to \infty} \frac{ \P\{ \mbox{survival to }n \}  }{ (1-s_d)^n } = \lim_{n \to \infty} \frac{ 1-f_n(0) }{ (1-s_d)^n } > 0 .
-\]
+\end{align}
 Furthermore,
@@
 Note additionally that 
-\[
+\begin{align}
     1 - \frac{1-f_n(u)}{1-f_n(0)} \searrow g(u) = \E[u^K] .
-\]
+\end{align}
 

### 2013-02-08 patchy-selection-plots.R
WorkingWiki importation

@@ -1,5 +1,145 @@
 source("patchy-selection-fns.R")
+source("plotting-fns.R")
 require(xtable)
+
+#######
+# Phase diagram:
+#  time to migration and mutations as a function of migration rate, mutation rate, and/or deleterious selection coef
+
+# Plot level sets of time to mut and mig
+#  first against characteristic dist and pop density
+#  then agains mutation rate and pop density
+rhovals <- logseq(1,100000,length.out=100)
+clvals <- seq(1,20,length.out=100)
+muvals <- logseq(1e-8,1e-6,length.out=100)
+A <- 10000 # km^2
+mu <- 1e-8
+sb <- .1
+sd <- .01
+const <- .5
+cl <- 5  # cline length
+clplot <- expand.grid(cl=clvals,rho=rhovals)
+clplot$mut <- with(clplot, 1/(A*mu*2*sb*rho) )
+clplot$mig <- with(clplot, 1/(const*2*sd*rho*exp(-cl)) )
+muplot <- expand.grid(mu=muvals,rho=rhovals)
+muplot$mut <- with(muplot, 1/(A*mu*2*sb*rho) )
+muplot$mig <- with(muplot, 1/(const*2*sd*rho*exp(-cl)) )
+f <- function (x) { matrix(x,nrow=(sqrt(length(x)))) }
+
+pdf(file="phase-diagram-log.pdf",width=6,height=3,pointsize=10)
+layout(t(1:2))
+par(mar=c(4,3.5,1,1)+.1)
+levelsets <- 10^c(0,2,4,6)  # which lines to draw
+labelsets <- levelsets
+# plot(0,type='n',xlab='',ylab='',xlim=range(clvals),ylim=range(rhovals),log='y')
+# contour( clvals, rhovals, f(clplot$mut), col='red', levels=levelsets, labels=labelsets, method="edge", lwd=2, add=TRUE )
+logcontour( clvals, rhovals, f(clplot$mut), col='red', levels=levelsets, labels=labelsets, method="edge", log='y' )
+logcontour( clvals, rhovals, f(clplot$mig), col='blue', add=TRUE, labels=labelsets, levels=levelsets, method="edge", log='y' )
+for (lev in levelsets) { # note rho is log10-scale.
+    rholev <- 1/(A*mu*2*sb*lev)  # mutation time determines rho
[…]

### 2013-02-08 plotting-fns.R

@@ -1,0 +1,29 @@
+# misc plotting fns
+
+logcontour <- function(x,y,z,log="",add=FALSE,...) {
+    # contour plot with log scale axes
+    logx <- grepl("x",log)
+    logy <- grepl("y",log)
+    xlims <- range( if (logx) { log10(x) } else { x } )
+    ylims <- range( if (logy) { log10(y) } else { y } )
+    if (!add) { 
+        plot(0,type='n',xlab='',ylab='',xlim=xlims, ylim=ylims, xaxt='n', yaxt='n') 
+        if (logx) {
+            xat <- unique(floor(pretty(c(ceiling(xlims[1]-diff(ylims)/1e5),floor(xlims[2]+diff(ylims)/1e5)))))
+            # xat <- axisTicks(range(x),log=logx)
+            axis(1, at=xat, labels=10^xat)
+        } else {
+            axis(1)
+        }
+        if (logy) {
+            yat <- unique(floor(pretty(c(ceiling(ylims[1]-diff(ylims)/1e5),floor(ylims[2]+diff(ylims)/1e5)))))
+            # yat <- axisTicks(ylims,log=logy)
+            axis(2, at=yat, labels=10^yat)
+        } else {
+            axis(2)
+        }
+    }
+    contour( if(logx){log10(x)}else{x}, if(logy){log10(y)}else{y}, z, add=TRUE, ... )
+}
+
+

### 2013-02-08 maize-ex.tex
/* Calculatated values */
WorkingWiki importation

@@
-\documentclass{article} \usepackage[landscape]{geometry} \begin{document}
+\documentclass{article} 
+\usepackage{fullpage} 
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\newcommand{\E}{\mathbb{E}}
+\renewcommand{\P}{\mathbb{P}}
+\newcommand{\var}{\mathop{\mbox{Var}}}
+\newcommand{\mutrate}{\lambda_{\mbox{mut}}}
+\begin{document}
 
-\input{helianthus-ex-table.tex}
+Maize is cultivated in several highland regions of North and South America,
+ranging in size and hundreds to thousands of kilometers apart;
+first in modern-day central Mexico, and some thousands of years later in the Andes.
+The (relevant) highland-adapted traits are thought to be deleterious in the intervening lowlands.
+The question is then whether the Andes acquired these traits through independent evolution (i.e.~new mutation)
+or by migration from Mexico.
+As many of the relevant parameter values are only vaguely guessable --
+e.g.~population size, mutational target size, migration rates --
+we consider the ease of adaptation by migration or mutation across a range of parameter values.
+A set of parameters that deems adaptation by either means highly unlikely
+can be discarded, since we know adaptation has occurred.
+On the other hand, it may well be the case that both methods are likely, so that some adaptations have arisen {\it de novo}
+and others have moved by migration.
+
+In the case of either migration or mutation, we can estimate the rate of adaptation by the expected time to adaptation.
+If this is sufficiently small (hundreds of years) then it is almost certain to have occurred within the three thousand years available;
+if it is large (tens of thousands of years) it is very unlikely.
+Both models assume a continuous distribution of populations across geography.
+To evaluate the chance of adaptation by mutation,
+we suppose that there is a large patch of highland habitat, but no highland-adapted alleles yet;
+the chance that a new mutation beneficial at high altitude achieves local fixation depends on the distance to the highland.
+As for migration, we compute the expected time before a rare excursion of migrants from a patch of highland habitat already fixed for the highland alleles
+travels a distance comparable to the distance between highlands.
+
+\section{General modeling}
+
+Throughout we use in many ways the {\em branching process approximation} --
+if an allele is locally rare, then for at least a few generations,
+the fates of each offspring are nearly independent.
+So, if the allele is locally deleterious, the total numbers of that allele behave as a subcritical branching process,
+destined for ultimate extinction.
+On the other hand, if the allele is advantageous,
+it will either die out or become locally common, with its fate determined in the first few generations.
+If the number of offspring of an individual with this allele is the random variable $X$, 
+with mean $\E[X] = 1+s$ (selective advantage $s>0$), variance $\var[X]=\xi^2$, and $\P\{X=0\}>0$ (some chance of leaving no offspring),
+then the probability of local nonextinction $p_*$ is approximately $p_* \approx 2s/\xi^2$ to second order in $s$.
+The precise value can be found by defining the generating function $\Phi(u) = \E[u^X]$; 
+the probability of local nonextinction $p_*$ is the minimal solution to $\Phi(1-u) = 1-u$.
+(This can be seen because: $1-p_*$ is the probability that an individual's family dies out;
+this is equal to the probability that the families of all that individuals' children die out;
+since each child's family behaves independently, if the individual has $x$ offspring, this is equal to $(1-p_*)^x$;
+and $\Phi(1-p_*) = \E[(1-p_*)^X]$.)
+
+If the selective advantage ($s$) depends on geographic location, 
+a similar fact holds: index spatial location by $i \in 1, \ldots, n$,
+and for $u = (u_1,u_2,\ldots,u_n)$ define the functions $\Phi_i(u) = \E[ \prod_j u_j^{X_{ij}} ]$,
+where $X_{ij}$ is the (random) number of offspring that an individual at $i$ produces at location $j$.
+Then $p_* = (p_{*1}, \ldots, p_{*n})$, the vector of probabilities that a new mutation at each location eventually fixes,
+is the minimal solution to $\Phi(1-p_*) = 1-p_*$,
+i.e.\ $\Phi_i(1-p_*) = 1-p_{*i}$.
+
+% # maize
+% N <- 103359  # plants per field
+% Nf <- 561    # ears to make next generation
+% e <- 0.068   # prob of total seed stock replacement
+% mg <- .0083  # pollen migration
+% pm <- .02    # prob of partial stock addition
+% m <- 0.2     # proportion of stock replaced
+% eqvals <- c(1, iterroot( function (p) { 
+%             p.dead <- e + (1/2)*(1-e)*(1-Nf/N)
+%             return( p.dead * 1 + 
+%                     (1/2)*(1-e)*exp(((1+sb)/p.dead)*( p - 1 )) + 
+%                     ((1/2)*(1-e)*Nf/N)*exp(((N/Nf)*(1+sb)/(1-e))*( p - 1 )) 
+%                 )
+%         } ) )
+% genfn <- function (f,selx,migr=.5) { 
+%         (e + (1/2)*(1-e)*(1-Nf/N))* 1 + #  no offspring
+%         (1/2)*(1-e)*exp(((1+selx)/(1-e))*( f + migr*diff(c(eqvals[1],f,eqvals[2]),differences=2) - 1 )) + # pollen, Poisson mean 1+s(x)
+%         (1/2)*(1-e)*(Nf/N)*exp(((N/Nf)*(1+selx/(1-e)))*( f + migr*diff(c(eqvals[1],f,eqvals[2]),differences=2) - 1 )) # seed, Poisson mean (1+s(x))*N/Nf
+%     }
+
+Here we consider a linear habitat, so that the selection coefficient at location $\ell_i$ is $s_i = \min( s_b, \max( - s_d, \alpha \ell_i ) )$.
+There does not seem to be a nice analytic expression for $p_*$ in this case,
+but since $1-p_*$ is a fixed point of $\Phi$, the solution can be found by iteration:
+$1-p_* = \lim_{n \to \infty} \Phi^n(u)$ for an appropriate starting point $u$.
+
+\subsection{Maize model}
+
+Specifically, the migration and reproduction dynamics we use are as follows.
+On a large scale,
+fields of $N$ plants are replanted each year from $N_f$ ears,
+either from completely new stock (with probability $p_e$),
+from partially new stock (a proportion $r_m$ with probability $p_m$),
+or entirely from the same field.
+Plants have an average of $\mu_E$ ears per plant, and ears have an average of $N/N_f$ kernels;
+so a plant has on average $\mu_E N/N_f$ kernels, and a field has on average $\mu_E N$ ears and $\mu_E N^2/N_f$ kernels.
+We suppose that a plant with the selected allele is pollen parent to $(1+s) \mu_E N/N_f$ kernels,
+and also seed parent to $(1+s)\mu_E N/N_f$ kernels, still in $\mu_E$ ears.
+The number of offspring a plant has depends on how many of its offspring kernels get replanted.
+Some proportion $m_g$ of the pollen-parent kernels are in other fields, and may be replanted;
+but with probability $p_e$ no other kernels (i.e.~those in the same field) are replanted.
+Otherwise, with probability $1-p_m$ the farmer chooses $N_f$ of the ears from this field to replant
+(or, $(1-r_m) N_f$ of them, with probability $p_m$);
+this results in a mean number $N_f/N$ (or, $(1-r_m)N_f/N$) of the plant's ears of seed children being chosen,
+and a mean number $1+s$ of the plant's pollen children kernels being chosen.
+Furthermore, the field is used to completely (or partially) replant another's field with chance $p_e/(1-p_e)$ (or $p_m$);
+resulting in another $N_f/N$ (or $r_m N_f/N$) ears and $1+s$ (or $r_m (1+s)$) pollen children being replanted elsewhere.
+Here we have assumed that pollen is well-mixed within a field,
+and that the selected allele is locally rare.
+Finally, we must divide all these offspring numbers by 2,
+since we look at the offspring carrying a particular haplotype, not of the diploid plant's genome.
+
+The above gives mean values; to get a probability model we assume that every count in sight is Poisson.
+In other words, we suppose that the number of pollen children is Poisson with random mean $\lambda_P$,
+and the number of seed children is a mixture of $K$ independent Poissons with mean $(1+s)N/N_f$ each,
+where $K$ is the random number of ears chosen to replant, which is itself Poisson with mean $\mu_K$.
+By Poisson additivity, the numbers of local and migrant offspring are Poisson,
+with means $\lambda_P = \lambda_{PL} + \lambda_{PM}$ and $\mu_K = \mu_{KL} + \mu_{KM}$ respectively.
+With probability $p_e$, $\lambda_{PM} = m_g(1+s)$ and $\mu_K = \lambda_{PL} = 0$.
+Otherwise, with probability $(1-p_e)(1-p_m)$, $\mu_{KL} = N_f/N$ and $\lambda_{PL} = (1+s)(1-m_g)$;
+and with probability $(1-p_e)p_m$, $\mu_{KL} = (1-r_m) N_f/N$ and $\lambda_{PL} = (1-r_m)(1+s)(1-m_g)$.
+The migrant means are,
+with probability $(1-p_e) p_e/(1-p_e) = p_e$, $\mu_{KM} = N_f/N$ and $\lambda_{PM} = 1+s$;
+while with probability $(1-p_e) p_m$, $\mu_{KM} = r_m N_f/N$ and $\lambda_{PM} = (1+s)(r_m(1-m_g) + m_g)$,
+and otherwise $\mu_{KM} = 0$ and $\lambda_{PM} = m_g(1+s)$.
+
+\begin{table}[htb!!]
+  \begin{center}
+  \begin{tabular}{|crl|}
+    \hline
+    complete seed stock replacement prob & $p_e$     & 0.068 \\
+    pollen migration rate               & $m_g$      & 0.0083 \\
+    number of plants per field          & $N$        & $10^5$ \\
+    number of ears used to replant      & $N_f$      & 561 \\
+    mean ears per plant                 & $\mu_E$    & 3 \\
+    partial stock replacement prob      & $p_m$      & 0.02 \\
+    mean proportion stock replaced      & $r_m$      & 0.2 \\
+    pollen migration distance           & $\sigma_p$ & 0 km \\
+    seed replacement distance           & $\sigma_s$ & 50 km \\
+    distance between demes              & $a$        & 15 km \\
+    width of altitudinal cline          & $w$        & 62km \\
+    deleterious selection coefficient   & $s_d$      & varies \\
+    beneficial selection coefficient    & $s_b$      & varies \\
+    slope of selection gradient         & $\alpha$   & $(s_d+s_b)/w$ \\
+    variance in offspring number        & $\xi^2$    & varies \\
+    maize population density            & $\rho$     & $5 \times 10^3$ \\
+    \hline 
+  \end{tabular}
+\end{center}
+  \caption{Parameter estimates used in calculations, and other notation.
+  \label{tab:parameters}
+  }
+\end{table}
+
+\subsection{Math}
+
+The generating function of a Poisson with mean $\lambda$ is $\phi(u;\lambda)=\exp(\lambda(u-1))$,
+and the generating function of a Poisson($\mu$) sum of Poisson($\lambda$) values is $\phi(\phi(u;\lambda);\mu)$.
+Therefore, the generating function for the diploid process, ignoring spatial structure,
+is
+\begin{align}
+ \Phi(u) &= \begin{aligned}[t]
+   & p_e \phi(u;m_g(1+s)) \\
+   & \quad {} + \left\{ (1-p_e) (1-p_m) \phi(u;(1+s)(1-m_g))\phi(\phi(u;(1+s)N/N_f);N_f/N) \right. \\
+   & \quad \qquad \left. {} + (1-p_e) p_m \phi(u;(1+s)(1-r_m)(1-m_g)) \phi(\phi(u;(1+s)N/N_f);(1-r_m)N_f/N) \right\} \\
+   & \quad \quad {} \times \left\{ p_e/(1-p_e) \phi(u;1+s) \phi(\phi(u;(1+s)N_f/N);N_f/N) \right. \\
+   & \quad \qquad \left. {} + p_m \phi(u;(1+s)(r_m(1-p_e)(1-m_g)+m_g)) \right. \\ 
+   & \quad \qquad \quad \left. {}\times \phi(\phi(u;(1+s)N/N_f);r_m N_f/N)\right. \\
+   & \quad \qquad \left. {} + (1 - p_e/(1-p_e)-p_m) \phi(u;m_g(1+s)) \right\}
+ \end{aligned} \label{eqn:genfn} \\
+ &= \begin{aligned}[t]
+   & \phi(u;m_g(1+s)) \big( \; p_e \big.\\
+   & \quad {} + \left\{ (1-p_e) (1-p_m) \phi(u;(1+s)(1-m_g))\phi(\phi(u;(1+s)N/N_f);N_f/N) \right. \\
+   & \quad \qquad \left. {} + (1-p_e) p_m \phi(u;(1+s)(1-r_m)(1-m_g)) \phi(\phi(u;(1+s)N/N_f);(1-r_m)N_f/N) \right\} \\
+   & \quad \quad {} \times \left\{ p_e/(1-p_e) \phi(u;(1+s)(1-m_g)) \phi(\phi(u;(1+s)N_f/N);N_f/N) \right. \\
+   & \quad \qquad {} + p_m \phi(u;(1+s)r_m(1-m_g)) \\
+   & \quad \qquad \quad {} \times \phi(\phi(u;(1+s)N/N_f);r_m N_f/N) \\
+   & \quad \qquad \left. {} + (1-p_e/(1-p_e)-p_m) \right\}  \big)
+ \end{aligned}
+\end{align}
+To get the generating function for a haploid, replace every instance of $1+s$ by $(1+s)/2$.
+
+As a quick check,
+% \begin{align}
+%   \Phi(0) &= p_e + \left\{ (1-p_e)(1-p_m)+(1-p_e)p_m \right\}
+%     \times \left\{ p_e/(1-p_e) + p_m + (1-p_e/(1-p_e)-p_m) \right\}
+%   &= p_e + (1-p_e)
+%   &= 1 .
+% \end{align}
+% And, 
+the mean total number of offspring of a diploid is
+\begin{align}
+  & \begin{aligned}
+(1+s) \big(
+    m_g
+    + & (1-p_e)\left\{ (1-p_m) ( (1-m_g) + 1 ) + p_m ( (1-r_m)(1-m_g) + (1-r_m) ) \right\} \\
+    & \qquad + \left\{ p_e ( (1-m_g) + 1 ) + p_m (1-p_e)( r_m(1-m_g) + r_m ) \right\} 
+    \big) 
+  \end{aligned}  \\
+& \qquad = 
+    (1+s) \big( m_g + 
+      (1-p_e)(2-m_g)(1 - p_m r_m)
+      + ( p_e (2-m_g) + p_m r_m (1-p_e)(2-m_g) )
+    \big) \\
+& \qquad = (1+s) \big( 
+    m_g + (2-m_g) ( (1-p_e) (1 - p_m r_m) + p_e + p_m r_m(1-p_e) )
+  \big) \\
+& \qquad = (1+s) \big( 
+    m_g + (2-m_g)
+  \big) \\
+& \qquad = 2(1+s) .
+\end{align}
+Check!
+
+We show numerically later that the probability of establishment is very close to $2s$ over the variance in reproductive number (as expected).
+It is possible to write down an expression for the variance, but it's a big, ugly one that doesn't lend itself to intuition.
+
+\subsection{Migration and spatial structure}
+
+To incorporate spatial structure, suppose that the locations $\ell_k$ are arranged in a regular grid, so that $\ell_k = a k$. 
+Recall that $s_k$ is the selection coefficient at location $k$.
+If the total number of offspring produced by an individual at $\ell_i$ is Poisson($\lambda_i$), with each offspring independently migrating to location $j$
+with probability $m_{ij}$,
+then the number of offspring at $j$ is Poisson($m_{ij}\lambda_i$),
+and so the generating function is
+\begin{align}
+  \phi(u;\lambda,m) &= \prod_j \exp( \lambda_i m_{ij} ( u_j - 1 ) ) \\
+  &= \exp\left\{ \lambda_i \left(\left(\sum_j m_{ij} u_j\right) - 1\right) \right\} .
+\end{align}
+We can then substitute this expression into equation \eqref{eqn:genfn},
+with appropriate migration kernels for pollen and seed dispersal.
+
+For migration, we need migration rates and migration distances for both wind-blown pollen
+and for farmer seed exchange.
+The rates are parameterized as above;
+we need the typical dispersal distances, however.
+One option is to say that the typical distance between villages is $d_v$,
+and that villages are discrete demes,
+so that pollen stays within the deme (pollen migration distance 0)
+and seed is exchanged with others from nearby villages;
+on average $\sigma_s$ distance away in a random direction.
+The number of villages away the seed comes from could be geometric 
+(including the possibility of coming from the same village).
+
+\subsection{Results}
+
+Iterating the generating function above finds the probability of establishment as a function of distance along the cline.
+This is shown in figure \ref{fig:prob_estab}.
+Note that the approximation $2s$ divided by the variance in offspring number is pretty darn close.
+
+\begin{figure}[ht!!]
+  \begin{center}
+    \includegraphics{prob-estab-62}
+    \includegraphics{prob-estab-500}
+  \end{center}
+  \caption{
+  Probability of establishment, as a function of distance along and around an altitudinal cline, whise boundaries are marked by the green lines.
+  {\bf (A)} The parameters above; with cline width 62km; {\bf (B)} the same, except with cline width 500km.
+  \label{fig:prob_estab}
+  }
+\end{figure}
+
+%%%%%%%
+\section{Adaptation by mutation}
+
+First, we'd like to compute how difficult is it for the beneficial adaptation to arise by new mutation.
+The rate of appearance of mutant alleles is a Poisson process,
+and we can assume that each is successful or not independently,
+so the time until the new mutant appears and fixes is exponentially distributed,
+with rate equal to the mutation rate multiplied by the probability of establishment integrated over the population.
+Referring to figure~\ref{fig:prob_estab},
+we see that this is pretty close to ( (area of high altitude) $+$ ($1/2$ area of altitudial gradient) ) $\times$ (population density) $\times$ (prob of establishment at high altitude).
+
+Let $A$ denote (area of high altitude) plus ($1/2$ area of altitudial gradient).
+The population density $\rho$ is roughly
+0.5--5 people per km$^2$
+$\times$ (0.5 ha field/person)
+$\times$ (2$\times10^4$ plants per field ha)
+$=$ (5000--50000 plants per km$^2$).
+As a check, the other set of numbers was
+``one village per 15 km''; i.e.\ per square with 15km on a side,
+which is 0.444 people per km$^2$.
+
+Since the probability of establishment at high altitude is approximately $2 s_b / \xi^2$, 
+with $\xi^2$ the variance in offspring number,
+the rate of appearance is just 
+\begin{align*}
+  \mutrate = 2 \rho A s_b / \xi^2 .
+\end{align*}
+
+%%%%%%%
+\section{Adaptation by migration}
+
+% \input{maize-ex-table.tex}
+
+% \input{maize-standing-ex-table.tex}
 

### 2013-02-08 patchy-selection-fns.R

@@ -50,7 +50,8 @@ mutationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
 migrationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
     # approximation for migrational influx per patch
     # XXX to-do: do numerics?
-    return( 2 * s*gb * A * rho * exp( - sqrt(abs(s*gm))*R/sigma ) )
+    # return( 2 * s*gb * A * rho * exp( - sqrt(abs(s*gm))*R/sigma ) )
+    return( 2 * s*gb * rho * exp( - sqrt(abs(s*gm))*R/sigma ) )
 }
 
 
@@ -67,5 +68,29 @@ paramString <- function(ps=c("mu","rho","s","gb","gm","sigma","R", "A"), pos=-1)
         }
     }
     paste( sapply( 1:length(ps), function (k) { sprintf( "%s=%.2g", names(ps)[k], ps[k] ) } ), collapse=", " )
+}
+
+logseq <- function(from,to,...) { 
+    # returns a sequence whose logarithms are uniformly spaced,
+    #  i.e. a multiplicative sequence
+    exp( seq( log(from), log(to), ... ) ) 
+}
+expseq <- function(from,to,...) { 
+    # returns a sequence that when exponentiated is uniformly spaced
+    log( seq( exp(from), exp(to), ... ) ) 
+}
+
+require(colorspace)
+hcolor <- function (z,alpha=.75,cols=adjustcolor(heat_hcl(64,h=c(40,360),l=70,c.=c(70,100)),alpha),nc=length(cols),...) {
+    # coloring function for hplot
+    cols[as.numeric(cut(pmin(1,pmax(-1,z)),breaks=seq(-1,1,length.out=nc+1),include.lowest=TRUE))]
+}
+hplot <- function (z,x=as.vector(col(z)),y=as.vector(row(z)),scale=1,max.cex=4,labs,xlabs=labs,ylabs=labs,alpha=.75,...) {
+    # Like heatmap but with circles...
+    plot( x=x, y=y, pch=20, cex=max.cex*pmin(1,sqrt(abs(z)/scale)), col=hcolor(z/scale,alpha=alpha), xaxt='n', yaxt='n', xlab="", ylab="", ... )
+    if (!missing(labs) | !missing(xlabs)) 
+        axis(1, at=1:length(xlabs), labels=xlabs, las=2) 
[…]

### 2013-02-08 patchy-selection-plots.R
WorkingWiki importation

@@ -109,12 +109,12 @@ get.prob.estab <- function (clinewidth) {
     ff <- matrix(NA,ncol=niter,nrow=length(xx))
     ff[,1] <- f(xx)
     for (k in 2:niter) { ff[,k] <- genfn(ff[,k-1],selx) }
-    return( list( prob.estab=ff[,niter], selx=selx, eqvals=eqvals ) )
+    return( list( prob.estab=ff[,niter], selx=selx, eqvals=eqvals, m=m ) )
 }
 
 pdf(file="prob-establishment.pdf",width=6,height=3,pointsize=10)
 layout(t(1:2))
-par(mar=c(4,4,1,1)+.1,mgp=c(2,1,0))
+par(mar=c(4,4,1,1)+.1,mgp=c(2.5,1,0))
 with( get.prob.estab(50), {
         plot(xx,1-prob.estab,type='n',lwd=2,ylim=c(0,max(1-prob.estab)),xlab="distance (km)", ylab="prob of establishment")
         polygon(c(xx,rev(xx)),c(1-prob.estab,rep(0,length(xx))),col=adjustcolor("blue",.5))
@@ -123,7 +123,8 @@ with( get.prob.estab(50), {
         polygon(xx[c(min(which(selx>-sd)),max(which(selx<sb)))][c(1,1,2,2)], c(0,par("usr")[4])[c(1,2,2,1)],density=5,col=grey(.5))
         text( xx[min(which(selx>-sd))], max(1-prob.estab), labels=as.expression(substitute(s[d]==sd,list(sd=-sd))), adj=c(1.2,1) )
         text( xx[max(which(selx<sb))], max(1-prob.estab), labels=as.expression(substitute(s[b]==sb,list(sb=sb))), adj=c(-0.2,1) )
-        arrows( x0=min(xx), y0=mean(1-prob.estab), x1=sqrt(sd/m), angle=90 )
+        arrows( x0=min(xx) + 5*dx, y0=mean(1-prob.estab), x1=min(xx)+2*dx*m/sqrt(sd) + 5*dx, angle=90, code=3, length=.05 )
+        text( min(xx) + 5*dx + (2/2)*dx*m/sqrt(sd), mean(1-prob.estab), pos=3, labels=expression(2*sigma/sqrt(s)) )
     } )
 with( get.prob.estab(500), {
[…]

### 2013-02-09 patchy-selection-paper.tex
/* Patchy Selection paper */

@@
 \begin{equation}
-\sum_{i=1}^{k} 2 s_B \rho E \exp(- |\sqrt{s_m} R_k/\sigma|}.
+\sum_{i=1}^{k} 2 s_B \rho E \exp(- |\sqrt{s_m} R_k/\sigma|).
 \end{equation}
@@
 \section{Applications} 
+Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX.
+These mice live in the desert?
+through the American SouthWest and into Mexico.
+In the majority of their range they live on the light colored rocks, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}.
+However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation.
+On one of these flows (Pinacate) an allele at the Mc1r locus is responsible for much of the change to a dark pelage \citep{Nachman:03}.
+This dark allele differs from the light allele by four amino acid changes - and has a dominant or partially dominant effect depending on the measure of coat color assessed.
+These lava flows are separated from each other by a range of distances, with some being separated from any other flow by hundreds of kilometers of light colored substrate.
+The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel.
+However, there is some evidence \citep{Hoekstra:05} that some close populations may have adapted by the spread of another dark allele by migration.
 
-HbS
+As can be seen from above a key parameter is $\sigma/\sqrt{s}$.
+\citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites.
+On the lava flow the dark MC1R allele is at $86\%$ frequency.
+At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency.
+At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass.
+While at the light substrate O'Neill site, which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.
 
-lava bed peromyscus
+I estimate that a characteristic length of 2.8km gives the frequency dropping from 50\% to ~1\% in this distance.
+So this doesn't seem so off.
+A characteristic length that is 1.4x gives 2.8\% from 50\% in 12km, which may be a little more accurate.
 
-G6PD
+2) The frequency at O'Neill [another site off the lava, on opposite side from Tule]  is 9km from the lava and has a frequency of 34\% of the dark allele.
+I estimate that we need a sigma/sqrt(s) of ~28km (i.e.
+ten times the characteristic length) to give a drop from 50\% to 34\%.
+
+[Obviously the cline isn't decaying from 50\%, the frequency of the light allele on the lava to be ~14%, we could put this in more carefully]
+So it seems to me that a sigma/sqrt(s) of 3-30km seems to encompass a range of interesting parameters, and can be justified without too much trouble.
+
+Using the dispersal distance of 0.28km [from marker recapture] and an s=0.01 [from paper] gives us sigma/sqrt(s)=2.8km
 

### 2013-02-11 patchy-selection-paper.tex
WorkingWiki importation

--- 2013-01-25 patchy-selection-paper.tex
@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\dconv}{\xrightarrow{d}}
+\newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
 
-\newcommand{\gc}[1]{{ \it \color{green} (#1) } }
+\newcommand{\gc}[1]{{\it\color{green}(#1)} }
+\newcommand{\plr}[1]{{\it\color{blue}(#1)}}
 
@@
 
+\subsection{Length of the hitchhiking haplotype}
+
+If a patch adapts through new mutation, the genomic region surrounding the selected site will hitchhike \citep{hitchhiking} along with it,
+so that at least initially, all adapted individuals share a fairly long stretch of haplotype.
+This association gets slowly whittled down by recombination during interbreeding at the edge of the patch;
+but there will always be longer LD nearby to the selected site.
+\plr{Could work out on what time scale it gets whittled down and the shared length at stationarity.}
+
+On the other hand, if a an already adapted patch colonizes another through migration,
+then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
+The length of this haplotype is inversely proportional to the number of generations the allele spends ``in transit'' --
+this is because, following back the successful migrant lineage, 
+each reproductive event in locations where the allele is at low frequency
+whittles down the hitchhiking haplotype.
+(In this case, the haplotype is literally hitchhiking across geography!)
+
+Here is a simplified picture of the process of adaptation by migration.
+There is a constant outflux of migrants from the patch where the allele is already established; 
+any descendants of these carrying the allele are locally maladapted, so are mostly doomed to extinction.
+However, eventually the descendants of one lucky migrant will reach the new patch and colonize it.
+(If the new patch is at some distance, it will almost certainly \plr{a.c.} be colonized first by only one family.)
+We will estimate the total outflux of migrants, and assume that their fates are independent of each other,
+so that we need only calculate the probability that offspring carrying the focal allele of a given migrant colonize the new patch,
+and the distribution of the transit time, given that this occurs.
+In doing so, we only follow the offspring of the migrant that carry the focal allele,
+and approximate these as a branching process, 
+i.e.\ assume that the number of offspring each produces is independent of each other.
+
+Since we are concerned with the rare migrant families that transit quickly between patches, 
+we should remove any backmigrants into the already-established patch;
+we do this by assuming rather that the allele is deleterious \emph{everywhere} --
+backmigrants will contribute little, since they are going the wrong direction,
+and we deal with arrival at the new patch later.
+Spatial motion is now independent of reproduction,
+so the chance that offspring of the migrant first arrive in the new patch after $\tau$ generations 
+is the probability of survival of the family until time $\tau$
+multiplied by the probability that some of the family are in the new patch, given survival.
+The first probability comes from well-known properties of branching processes,
+and the second from Brownian motion.
+Then, the chance that the migrant family establishes in the new patch 
+depends on the size of the family that arrives in the patch,
+and the probability that each establishes.
+
+If each individual has on average $(1-s_d)$ offspring, then after $t$ generations,
+the family has on average $(1-s_d)^t$ members.
+Let $p_e(t)$ be the chance of extinction by time $t$, and let $K_t$ be the (random) family size conditioned on nonextinction,
+so that $(1-s_d)^t = (1-p_e(t))\E[K_t]$.
+It turns out that by assuming the fate of each offspring is independent (i.e.\ that this forms a branching process),
+and that the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jaegers1976branching},
+that $K_t$ has a limiting distribution: $K_t \dconv K$ 
+and that $(1-s_d)^{t}/(1-p_e(t)) \to \E[K]$ as $t \to \infty$.
+In other words, the chance of survival to time $t$ is a constant multiplied by $(1-s_d)^t$,
+and conditional on survival, the family size is fixed.
+
+We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary}.
+When we condition on survival, we condition on existence of at least one lineage from time $0$ to time $t$.
+Once given this lineage, nothing distinguishes any of the other reproduction events --
+then each individual in this lineage must give birth to at least one offspring,
+and all other reproduction events are unconditioned.
+In other words, the genealogy of a family that survives until $t$
+can be decomposed into a ``trunk'' that lasts until $t$
+anmd independent ``branches'' growing from each intermediate time $0\le s \le t$ 
+that may or may not survive until $t$.
+This process is depicted in Figure~XA.
+Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$
+(so $\P\{Z_t = 0\} = p_e(t)$),
+then
+\begin{align}
+  Z_t \; \vert \; Z_t>0 \deq \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
+\end{align}
+where each $Z^{(s,k)}$ is an independent copy of $Z$,
+and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.
+
+\begin{figure}[ht!!]
+  \begin{center}
+  \input{branching-concept.pdf_tex}
+  \end{center}
+\caption{Depiction of the decomposition described in the text.
+In red is the central lineage (whose spatial motion we follow);
+in black are the surviving branches off of this lineage, 
+and in grey are offspring of the central lineage that did not survive until $t$.
+\plr{Label $W_s$ in this figure?}
+\plr{Add as part B a simulation figure.}
+\label{fig:branching_decomp}
+}
+\end{figure}
+
+Now we can add in the spatial motion, by following only the motion of the ``trunk''
+(i.e.\ the red line in figure \ref{fig:branching_decomp}).
+We will first work in one dimension.
+If the variance of the parent-offspring distance is $\sigma^2$ (and the distribution is not too heavy-tailed),
+then the motion of this is approximately Brownian, with variance $\sigma^2$ per generation.
+Suppose that the new patch is at distance $x$.
+The probability density for the first time such a Brownian motion $B_s$ hits $x$ is
+\citep[XXX]{feller}
+% note $\sigma B_t \deq B_{\sigma^2 t}$ so $\P\{ \sup_{s\le t} \sigma B_s \le x \} = \P\{ \sup_{s \le \sigma^2 t} B_s \le x \}
+\begin{align}
+  \P\{ B_t=x \;\mbox{and} \; B_u<x \;\mbox{for all}\; u<t\} =  \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right) .
+\end{align}
+
+We are interested in two things: the probability that the family of a given migrant will eventually establish in the new patch,
+and the transit time conditioned on establishment.
+
+Write $\tau$ for the first time the trunk lineage hits $x$, i.e.\ reaches the new patch, 
+conditioned on this occurrence.
+Let $Z_t$ denote the size of the family at time $t$,
+and $B_t$ the location of the trunk lineage,
+so that
+\begin{align}
+  \P\{\tau \in dt\} = \P\{ Z_t>0 \} \P\{ 
+\end{align}
+If we define $\mu_K$ to be the limiting mean family size conditioned on survival
+(i.e.\ $\mu_K = \E[K]$), then combining the results above,
+\begin{align}
+  \P\{ \tau \in dt \} \propto 
+  \frac{(1-s_d)^t}{\mu_K} \frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) ,
+\end{align}
+
+%%%%%%%%%%
 \section{Applications}

### 2013-02-11 patchy-selection-paper.tex
WorkingWiki importation

@@
 
+\section{Applications} 
+Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX.
+These mice live in the desert?
+through the American SouthWest and into Mexico.
+In the majority of their range they live on the light colored rocks, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}.
+However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation.
+On one of these flows (Pinacate) an allele at the Mc1r locus is responsible for much of the change to a dark pelage \citep{Nachman:03}.
+This dark allele differs from the light allele by four amino acid changes - and has a dominant or partially dominant effect depending on the measure of coat color assessed.
+These lava flows are separated from each other by a range of distances, with some being separated from any other flow by hundreds of kilometers of light colored substrate.
+The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel.
+However, there is some evidence \citep{Hoekstra:05} that some close populations may have adapted by the spread of another dark allele by migration.
+
+As can be seen from above a key parameter is $\sigma/\sqrt{s}$.
+\citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites.
+On the lava flow the dark MC1R allele is at $86\%$ frequency.
+At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency.
+At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass.
+While at the light substrate O'Neill site, which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.
+
+I estimate that a characteristic length of 2.8km gives the frequency dropping from 50\% to ~1\% in this distance.
+So this doesn't seem so off.
+A characteristic length that is 1.4x gives 2.8\% from 50\% in 12km, which may be a little more accurate.
+
+2) The frequency at O'Neill [another site off the lava, on opposite side from Tule]  is 9km from the lava and has a frequency of 34\% of the dark allele.
+I estimate that we need a sigma/sqrt(s) of ~28km (i.e.
+ten times the characteristic length) to give a drop from 50\% to 34\%.
+
+[Obviously the cline isn't decaying from 50\%, the frequency of the light allele on the lava to be ~14%, we could put this in more carefully]
+So it seems to me that a sigma/sqrt(s) of 3-30km seems to encompass a range of interesting parameters, and can be justified without too much trouble.
+
+Using the dispersal distance of 0.28km [from marker recapture] and an s=0.01 [from paper] gives us sigma/sqrt(s)=2.8km
+
+%%%%%%%%%
 \subsection{Length of the hitchhiking haplotype}
@@
 any descendants of these carrying the allele are locally maladapted, so are mostly doomed to extinction.
+(We do not require that migration be blind to the environment --
+although adaptive migration would reduce the outflux of migrants --
+but do assume that the focal allele does not affect migration once outside the patch.)
+\plr{or maybe we should discuss this later?}
 However, eventually the descendants of one lucky migrant will reach the new patch and colonize it.

### 2013-02-12 patchy-selection-paper.tex
WorkingWiki importation

@@
 
-\newcommand{\var}{\mathop{\mbox{Var}}}
+% \newcommand{\var}{\mathop{\mbox{Var}}}
+% \newcommand{\Exp}{\mathop{\mbox{Exp}}}
+\DeclareMathOperator{\var}{Var}
+\DeclareMathOperator{\Exp}{Exp}
 \renewcommand{\P}{\mathbb{P}}
@@
 \newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
+\newcommand{\migrate}{\lambda_\text{mig}}
 
@@
 If this establishment rate is low, the time it takes
-a mutation to arise that will become established will be  $\sim Exp(\lambda_{mig,R})$
+a mutation to arise that will become established will be  $\sim \Exp(\lambda_{mig,R})$
 Assuming that the time it takes for a allele from the first patch to
@@
 
-\section{Applications} 
-Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX.
-These mice live in the desert?
-through the American SouthWest and into Mexico.
-In the majority of their range they live on the light colored rocks, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}.
-However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation.
-On one of these flows (Pinacate) an allele at the Mc1r locus is responsible for much of the change to a dark pelage \citep{Nachman:03}.
-This dark allele differs from the light allele by four amino acid changes - and has a dominant or partially dominant effect depending on the measure of coat color assessed.
-These lava flows are separated from each other by a range of distances, with some being separated from any other flow by hundreds of kilometers of light colored substrate.
-The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel.
-However, there is some evidence \citep{Hoekstra:05} that some close populations may have adapted by the spread of another dark allele by migration.
-
-As can be seen from above a key parameter is $\sigma/\sqrt{s}$.
-\citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites.
-On the lava flow the dark MC1R allele is at $86\%$ frequency.
-At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency.
-At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass.
-While at the light substrate O'Neill site, which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.
-
-I estimate that a characteristic length of 2.8km gives the frequency dropping from 50\% to ~1\% in this distance.
-So this doesn't seem so off.
-A characteristic length that is 1.4x gives 2.8\% from 50\% in 12km, which may be a little more accurate.
-
-2) The frequency at O'Neill [another site off the lava, on opposite side from Tule]  is 9km from the lava and has a frequency of 34\% of the dark allele.
-I estimate that we need a sigma/sqrt(s) of ~28km (i.e.
-ten times the characteristic length) to give a drop from 50\% to 34\%.
-
-[Obviously the cline isn't decaying from 50\%, the frequency of the light allele on the lava to be ~14%, we could put this in more carefully]
-So it seems to me that a sigma/sqrt(s) of 3-30km seems to encompass a range of interesting parameters, and can be justified without too much trouble.
-
-Using the dispersal distance of 0.28km [from marker recapture] and an s=0.01 [from paper] gives us sigma/sqrt(s)=2.8km
-
 %%%%%%%%%
-\subsection{Length of the hitchhiking haplotype}
+\section{Length of the hitchhiking haplotype}
 
@@
 
-If each individual has on average $(1-s_d)$ offspring, then after $t$ generations,
-the family has on average $(1-s_d)^t$ members.
+If each individual has on average $(1-s_m)$ offspring, then after $t$ generations,
+the family has on average $(1-s_m)^t$ members.
 Let $p_e(t)$ be the chance of extinction by time $t$, and let $K_t$ be the (random) family size conditioned on nonextinction,
-so that $(1-s_d)^t = (1-p_e(t))\E[K_t]$.
+so that $(1-s_m)^t = (1-p_e(t))\E[K_t]$.
 It turns out that by assuming the fate of each offspring is independent (i.e.\ that this forms a branching process),
@@
 that $K_t$ has a limiting distribution: $K_t \dconv K$ 
-and that $(1-s_d)^{t}/(1-p_e(t)) \to \E[K]$ as $t \to \infty$.
-In other words, the chance of survival to time $t$ is a constant multiplied by $(1-s_d)^t$,
+and that $(1-s_m)^{t}/(1-p_e(t)) \to \E[K]$ as $t \to \infty$.
+In other words, the chance of survival to time $t$ is a constant multiplied by $(1-s_m)^t$,
 and conditional on survival, the family size is fixed.
@@
 % note $\sigma B_t \deq B_{\sigma^2 t}$ so $\P\{ \sup_{s\le t} \sigma B_s \le x \} = \P\{ \sup_{s \le \sigma^2 t} B_s \le x \}
-\begin{align}
+\begin{align} \label{eqn:hitting_dist}
   \P\{ B_t=x \;\mbox{and} \; B_u<x \;\mbox{for all}\; u<t\} =  \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right) .
 \end{align}
+The chance that the descendants of a given migrant arrives at the new patch at time $t$
+is now just equation~\eqref{eqn:hitting_dist} multiplied by the chance that 
+the migrant family is still alive at time $t$.
+For convenience, we define $\alpha = - \log(1-s_m) \approx s_m$,
+so that $1-p_e(t) \simeq (1/\mu_K) \exp(-\alpha t)$,
+and define $\mu_K$ to be the limiting mean family size conditioned on survival.
+By integrating over $t$, we get the chance that descendants of the migrant ever arrive at the new patch is
+\begin{align} \label{eqn:estab_prob}
+  & \int_0^\infty \frac{(1-s_m)^t}{\mu_K} \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right)  dt \\
+  & \qquad = \frac{1}{\mu_K} \exp\left( - \frac{ x \sqrt{\alpha}}{\sigma} \right) .
+\end{align}
+(Note that this is, up to the constant $\mu_K$, the chance that a Brownian motion killed at rate $\alpha$
+ever hits $x$ \citep{feller}.
+This approximation holds if $x$ is large enough that the approximation $1-p_e(t) \approx (1-s_m)^t/\mu_K$ holds.)
 
-We are interested in two things: the probability that the family of a given migrant will eventually establish in the new patch,
+We now have the tools to compute the two quantities of interest:
+the probability that the family of a given migrant will eventually establish in the new patch,
 and the transit time conditioned on establishment.
 
-Write $\tau$ for the first time the trunk lineage hits $x$, i.e.\ reaches the new patch, 
-conditioned on this occurrence.
-Let $Z_t$ denote the size of the family at time $t$,
-and $B_t$ the location of the trunk lineage,
-so that
+The probability of establishment of a single migrant in a patch distance $x$ away
+is equal to the probability that the migrant's family makes it to the patch at all,
+multiplied but the chance that it establishes locally, given that it makes it there.
+We have computed the first term in equation~\eqref{eqn:estab_prob},
+and we make the small assumption that the second term is just the probability of local fixation
+of a beneficial allele present in $K$ copies,
+i.e.~$\E[1-(1-p_f(s_b))^K] \approx 2 s_b \mu_K/\xi^2$,
+where $p_f(s_b) \approx 2s_b/\xi^2$ is the probability of fixation of a single mutant with additive benefit $s_b$.
+Therefore, the probability of establishment is
 \begin{align}
-  \P\{\tau \in dt\} = \P\{ Z_t>0 \} \P\{ 
+  F(x) &\approx \frac{2s_b}{\xi^2} \exp\left( - \frac{\sqrt{\alpha} x}{\sigma} \right) ,
 \end{align}
-If we define $\mu_K$ to be the limiting mean family size conditioned on survival
-(i.e.\ $\mu_K = \E[K]$), then combining the results above,
+which verifies the less intuitive calculation of $\migrate$ above.
+
+Now, write $\tau$ for the first time the trunk lineage hits $x$, i.e.\ reaches the new patch,
+with the convention that $\tau=\infty$ if it never does.
+By the above arguments, and equation~\eqref{eqn:estab_prob}, we have that the conditional probability density is
 \begin{align}
-  \P\{ \tau \in dt \} \propto 
-  \frac{(1-s_d)^t}{\mu_K} \frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) ,
+  \P\{\tau\in dt | \tau<\infty \} &= \frac{\P\{\tau \in dt\}}{\P\{\tau<\infty\}} \\
+  &= \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2} -\alpha t + \sqrt{\alpha} \frac{x}{\sigma} \right) .
 \end{align}
+Although $\tau$ is interesting, we are more interested in the length of the hitchhiking segment,
+since that is where our signal would come from.
+Since each generation spent in transit provides an opportunity for recombination,
+if recombination is Poisson, the length of the hitchhiking segment (in Morgans) on each side of the site will be exponentially distributed
+with mean $\tau$, so that if $L$ is the length on, say, the right, then
+\begin{align} \label{eqn:haplen_cdf}
+\P\{L>\ell\} &= \E[e^{-\ell \tau}|\tau<\infty] \\
+&= \exp\left\{{-\frac{x}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
+\end{align}
+This is also the Laplace transform of $\tau$;
+we can differentiate it to find that $\E[\tau] = x/(\sigma\sqrt{2\alpha})$
+and $\var[\tau] = x/( 2^{3/2}\alpha^{3/2}\sigma)$.
+Integrating~\eqref{eqn:haplen_cdf} gives that 
+$\E[L] = \sigma \sqrt{2\alpha}/x + \sigma^2/x^2$ and $\var[L] = 2\alpha\sigma^2/x^2 + 4 \sigma^3 \sqrt{2\alpha}/x^3 + 5 \sigma^4 / x^4$;
+an easier way to see this is to note that if $Y$ is an exponential random variable with rate $x\sqrt{2}/\sigma$,
+then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$.
+
+%%%
+\subsection{Limiting distributions}
+
+We are already assuming that $x$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
+by taking the large-$x$ limit above we can obtain more intuitive descriptions of the process
+via a central limit for $\tau$.
+Define $\eta_x$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
+so that $\tau = x/\sigma(2\alpha)^{-1/2} + \eta_x \sqrt{x/\sigma} (2\alpha)^{-3/4}$.
+Plugging this into equation~\eqref{eqn:haplen_cdf},
+after some algebra (and using that $\sqrt{x+\epsilon} = \sqrt{x} + \frac{\epsilon}{2\sqrt{x}} - \frac{\epsilon^2}{2\sqrt{x}^3} + O(\epsilon^3)$) we obtain that
+\begin{align}
+  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{x}\right) \right\}.
+\end{align}
+By the properties of Laplace transforms \citep{durrett},
+this implies that $\eta_x$ converges in distribution to a standard Gaussian random variable as $x \to \infty$.
+In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
+Since the standard deviation of $\tau$ is of lower order than its mean,
+a similar calculation shows that for large $x$, the haplotype halflength $L$ is approximately exponentially distributed, 
+with mean $\sqrt{2\alpha} \sigma/x$.
+
+% \begin{align}
+%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{x}{\sigma\sqrt{2\alpha}}\right)\frac{(2\alpha)^{3/4}\sqrt{\sigma}}{\sqrt{x}} \right\} \right] \\
+%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2\alpha)^{3/4} \sqrt{\sigma}}{\sqrt{x}}\right) \tau \right\} \right] \exp\left( (2\alpha)^{1/4} \sqrt{\frac{x}{\sigma}} \right) \\
+% \end{align}
 
@@
 \section{Applications} 
+Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX.
+These mice live in the desert?
+through the American SouthWest and into Mexico.
+In the majority of their range they live on the light colored rocks, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}.
+However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation.
+On one of these flows (Pinacate) an allele at the Mc1r locus is responsible for much of the change to a dark pelage \citep{Nachman:03}.
+This dark allele differs from the light allele by four amino acid changes - and has a dominant or partially dominant effect depending on the measure of coat color assessed.
+These lava flows are separated from each other by a range of distances, with some being separated from any other flow by hundreds of kilometers of light colored substrate.
+The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel.
+However, there is some evidence \citep{Hoekstra:05} that some close populations may have adapted by the spread of another dark allele by migration.
+
+As can be seen from above a key parameter is $\sigma/\sqrt{s}$.
+\citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites.
+On the lava flow the dark MC1R allele is at $86\%$ frequency.
+At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency.
+At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass.
+While at the light substrate O'Neill site, which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.
+
+I estimate that a characteristic length of 2.8km gives the frequency dropping from 50\% to $\sim$1\% in this distance.
+So this doesn't seem so off.
+A characteristic length that is 1.4x gives 2.8\% from 50\% in 12km, which may be a little more accurate.
+
+2) The frequency at O'Neill [another site off the lava, on opposite side from Tule]  is 9km from the lava and has a frequency of 34\% of the dark allele.
+I estimate that we need a sigma/sqrt(s) of $\sim$28km (i.e.
+ten times the characteristic length) to give a drop from 50\% to 34\%.
+
+[Obviously the cline isn't decaying from 50\%, the frequency of the light allele on the lava to be $\sim$14\%, we could put this in more carefully]
+So it seems to me that a sigma/sqrt(s) of 3-30km seems to encompass a range of interesting parameters, and can be justified without too much trouble.
+
+Using the dispersal distance of 0.28km [from marker recapture] and an s=0.01 [from paper] gives us sigma/sqrt(s)=2.8km
 
 HbS
-
-lava bed peromyscus
 
@@
 \section{Discussion}
+
+\plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
 

### 2013-02-14 patchy-selection-paper.tex
WorkingWiki importation

@@
 \newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
+
 \newcommand{\migrate}{\lambda_\text{mig}}
+\newcommand{\mutrate}{\lambda_\text{mut}}
 
@@
 likely with a 
-higher mutational input, i.e. when the mutation target and population sizes are larger \citep{}. 
+higher mutational input, i.e.\ when the mutational target and population sizes are larger \citep{}. 
 Standing variation for a trait can also increase the probability of convergence, 
 as multiple pre-existing alleles can spread from low frequency, 
-if our now adaptive alleles were previously not too deleterious \citep{Orr,Hermission}. 
+if the newly beneficial alleles were previously not too deleterious \citep{Orr,Hermission}. 
 The geographic distribution of populations can also shape the probability of parallel mutation within a species.
-A geographically wide spread species is more likely to adapt by multiple mutations im parallel if dispersal 
-is geographically limited, as populations will adapt via new mutations rather than waiting for migration \citep{RalphCoop}.
+A widespread species is more likely to adapt by multiple mutations in parallel if dispersal 
+is geographically limited, as subpopulations will adapt via new mutations rather than waiting for migration \citep{RalphCoop}.
 
@@
 
-Intuitively, convergence is also more likely when geographically
-separated populations adapt to ecologically similar conditions. The
-probability of convergent adaptation rather than populations adapting
-through allelic variants shared by migration, will be enhanced if
-these adaptive alleles are maladapted in the environmentals separating
-the populations. Such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{slatkin} 
-and so might encourage local adaptation by parallel adaptation rather than migration.\\
+Intuitively, convergence is also more likely 
+when geographically separated populations adapt to ecologically similar conditions. 
+The probability that convergent adaptations arise independently
+before adaptations spread between the populations by migration
+will be enhanced if these adaptive alleles are maladapted in intervening environments,
+since such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{slatkin}.
 
-For example a variety of plant species have repeatedly adapted to patchs of heavy metal soils
- (e.g. serpentine outcrops and mine tailings) within their ranges
-and alleles that adapt the plants to heavy metals may be disfavored off of these patches. 
-Similarly across the American Southwest light colored rock pocket mice have adapted to outcrops of dark rock 
-through the evolution of darker cryptic pigmentation. These dark outcrops are geographically
- separated by areas of light colored sand and rock, 
-and strong selection appears to favour the matching of crypsis to substrate. 
-Perhaps as a result of this strong selection against migrants, 
-at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops. 
-This situation is repeated in XXXX where darker colored mice have repeated evolved lighter pigmentation to match lighter color sands on
- beaches via different genetic routes.
+For example, a variety of plant species have repeatedly adapted to patches of soil with high concentrations of heavy metals
+(e.g.\ serpentine outcrops and mine tailings) \citep{turnerarabidopsis,mimulus};
+the alleles conferring heavy metal tolerance are thought to be disfavored off of these patches. 
+Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration,
+to particular substrates, e.g.\ dark rock outcrops or white sand dunes \citep{benson}.
+One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
+which is primarily dark on a number of black lava flows separated by much lighter soil \citep{dice}.
+Strong selection appears to favour such crypsis \citep{owls},
+so perhaps as a result of this strong selection against migrants, 
+at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{nachman}. 
+Similar situations have been demonstrated in other mouse systems \citep{beachmice,peromyscus}.
 
@@
 
+%%%%%%%%%%%%%
 \section{Methods}
-\subsection{Introduction to a patchily constructed model}
+
+%%%
+\subsection{Introduction to a patchy model}
 \label{ss:patchyspace}
@@
 Consider a population spread uniformly across a landscape. Within this environment there are patches of
-habitat that individuals are initially poorly adapted to, surrounded by large areas that our population is well adapted to.
+habitat that individuals are initially poorly adapted to, surrounded by large areas that the population is well adapted to.
 
-It takes only a single mutational change to create an allele ($B$) that adapts an individual to the
+Suppose it takes only a single mutational change to create an allele ($B$) that adapts an individual to the
 poor habitat type. This mutational change occurs at a low rate of
@@
 allele has fitness $1+s_B$ relative to the unmutated type ($b$) in the ``poor'' habitat patches,
-and assume that it has fitness $1+s_m$ when rare in the intervening areas, with
-$s_m<0$. To simplify matters we assume that the disadvantage $s_m$ ensures that on the relevant timescale,
+and assume that it has fitness $1-s_m$ when rare in the intervening areas, with
+$s_m>0$. To simplify matters we assume that the disadvantage $s_m$ 
+is sufficiently large that, on the relevant timescale,
 the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
+(In particular, we require $s_m\neq 0$.)
+\plr{NOTE: changed over to $s_m>0$.}
 
@@
 fitness is such the dynamics of the  allele when it is rare are
-determined by the fitness of the heterozygote. This allows us to treat
-a diploid model as being essentially haploid for our purposes. \gc{TRUE?}
+determined by the fitness of the heterozygote. 
+More general dominance will only make small corrections,
+with the exception of the recessive case, which we will omit.
+Mostly, we follow the literature in treating the diploid model as essentially haploid.
+% This allows us to treat a diploid model as being essentially haploid for our purposes. \gc{TRUE?}  \plr{would change e.g. freqs at migration-selection balance}
 
@@
 density point later). We further assume that the
-variance in offspring number is $1$, and that the mean squared distance between parent and child is $\sigma^2$ (the dispersal variance).
+variance in offspring number is $\xi^2$, 
+and that the mean squared distance between parent and child is $\sigma^2$ (the dispersal variance).
 
-\paragraph{Previous work} We will make use of a number previous
-results from the literature. \citet{slatkin1973geneflow}, working in
-one dimensional deterministic framework, showed that if the
-physical width of the patch is less than $2 \tan^{-1}
-(\sqrt{|s_m/s_B|})$, then $B$ can not 
-become \gc{fixed/at a stable equilibrium/?} within the patch, as the local adaptation prevented by
-migrational swamping \citep[see also][ for a review]{Lenormand}.
+\paragraph{Previous work} 
+We will make use of a number previous results from the literature. 
+\citet{slatkin1973geneflow}, working in one dimensional deterministic framework, 
+showed that if the physical width of the patch is less than $2 \tan^{-1} (\sqrt{s_m/s_B})$, 
+then there is not a stable equilibrium with both $b$ and $B$ present,
+so that $B$ cannot establish within the patch due to migrational swamping \citep[see also][ for a review]{Lenormand}.
 \citet{barton1987establishment}, again working in one dimension,
- showed that this critical patch size also held in a stochastic
- framework \citep[see also the work of][]{Polk}. 
+showed that this critical patch size also held in a stochastic framework \citep[see also the work of][]{Polk}. 
 For patches above this critical size, \citet{barton1987establishment}
- founds that mutations appearing further than a few multiples of
- $\sigma/\sqrt{2|s_m|}$ away from such a patch have vanishingly small
- probability of becoming established within the patch. 
-Barton also showed that mutations appearing within the patch have probability 
-strictly less but approaching that $2s_B$ of establishment as the size
-of the patch increased. Thus for large enough patches the probability
-of establishment of a locally favorable allele approaches the standard
-establishment probability ($2s_B$) for a panamitic population
-\citep[][]{haldane, fisher}, a result that also (generally) holds for
-a geographically spread population experiencing a uniform selection
-pressure \citep{Maruyama, cherry}.
+found that the probability of establishment of a new mutant that appears at distance $x$ from the patch
+decays exponentially with distance from the patch, with the scale given by $\sigma/\sqrt{s_m}$.
+% mutations appearing further than a few multiples of $\sigma/\sqrt{2s_m}$ away from such a patch 
+% have vanishingly small probability of becoming established within the patch. 
+\citet{barton1987establishment} also showed that mutations appearing within the patch have probability of establishment
+less than the probability for a panimictic population 
+\citep[which we denote $p_e$, and often approximate by $2 s_B / \xi^2$][]{haldane,fisher};
+and that the probability approaches $p_e$ as the size of the patch increases.
+This result holds quite generally for a geographically spread population that experiences a uniform selection
+pressure \citep{Maruyama,cherry}. 
+\plr{wait, what result? that fixation is hindered by geographic structure?}
 
@@
 
-%A similar situation of interest is the case when the species occupies a continuous range, but the selective pressure itself is patchy.
-I%f the areas where the new allele is advantageous are relatively isolated,
-%we might hope to apply the results above to such a situation.
-%For this purpose, suppose that instead of a uniform habitat,
-%the selective pressure is such that the new alleles are only advantageous in isolated locations, 
-%and are disadvantageous in the intervening region.
-%As before, we assume that the new allele has fitness $1+s_b$ relative to the unmutated type in the ``good'' patches,
-%and assume that it has fitness $1+s_m$ in the intervening areas, with $s_m<0$.
-%To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
-%and that the disadvantage $s_m$ ensures that on the relevant timescale,
-%the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
-%With these assumptions as justification, we will move to a model of discrete demes exchanging migrants,
-%where ``migration'' occurs through the rare, transient passage of the new allele through regions where it is disadvantageous.
-
 %equal deme sizes?  general migration rates?
 
-\subsubsection{Time to the establishment of a locally adaptive allele
+%%%%%
+\subsection{Time to the establishment of a locally adaptive allele
   due to mutational influx}
 \label{ss:patchymutation}
-Consider first a single, isolated poor habitat patch of area $A$ in
-which no $B$ allele has yet become established. As
-we are interested in patchess where local adaptation can occur we will assume
-that our patch is larger than the cutoff for local
-establishment, i.e. $\sqrt{A} > 2 \tan^{-1} (\sqrt{|s_m/s_B|})$.
 
-Let $p(x)$ be the probability that a new allele arising at location
-$x$ away from the patch fixes within the poor habitat patch.
-The function $p(x)$ can be found in terms of Jacobi elliptic functions,
-XXX as shown in the Appendix, but the expressions are particularly unhelpful.
-The total successful mutational influx per generation is then given by $\int \rho \mu p(x) dx$,
+Consider first a single, isolated poor habitat patch of area $A$ in which no $B$ allele has yet become established. 
+As we are interested in patches where local adaptation can occur,
+we will assume that our patch is larger than the cutoff for local establishment, 
+i.e.\ wider than $2 \tan^{-1} (\sqrt{s_m/s_B})$.
+
+Let $p(x)$ be the probability that a new mutant $B$ allele arising at location $x$ 
+relative to the patch fixes within the poor habitat patch.
+The function $p(x)$ can be found in one dimension in terms of Jacobi elliptic functions,
+XXX as shown in the Appendix, but the expressions are not particularly helpful.
+The total successful mutational influx per generation is then given by the integral of $p(x)$ over the entire species range $\int \rho \mu p(x) dx$,
 and depends in a complicated way on the relationship of patch width and selection coefficients,
-but still scales linearly with the population mutational influx $\rho \mu$.
-Therefore, we define $p_f = \int p(x) dx$.  If the width of the patch, $\sqrt{A}$, is large
-relative to $\sigma/\sqrt{2s_m}$ then a reasonable approximation is to
-set $p_f \approx 2 s_B A$. This follows from the fact that if the
-patch is large than the majority of the mutations that could become
-established within the patch, arose within the patch.
+but still scales linearly with the mutational influx density $\rho \mu$.
+If the width of the patch is large relative to $\sigma/\sqrt{2s_m}$, 
+then a reasonable approximation is to assume that $p(x) = p_e$ within the patch, and $p(x) = 0$ otherwise,
+for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_B A / \xi^2$. 
+This holds for large patches, in which the number of mutations arising within the patch is larger 
+than those arising sufficiently nearby;
+we examine this more generally via simulation in Appendix XXX.
+\plr{In Appendix: we simulated fixation prob in a patch with gradual transition.  We have convincing figures.}
 
-The rate at which mutations arise and colonize the patch is
-\begin{equation}
-\lambda_{Mut} = \rho \mu p_f \approx 2 s_B \rho \mu.
-\end{equation}
-If this rate is low,  then the waiting time till a mutation arises that
-will become locally established within the patch is exponentially
-distributed with rate $\lambda_{Mut}$.  Assuming that once a mutation
-becomes established it quickly reaches its equilibrium frequency across the patch, the
- expected time it takes for our patch to become colonized by the $B$
- allele is $T_{Mut} \approx 1/ \lambda_{Mut} $.
+The rate at which mutations arise and colonize a patch of area $A$ is therefore
+\begin{align} \label{eqn:mutrate}
+  \mutrate = \rho \mu \int p(x) dx  \approx 2 s_B \rho A \mu / \xi^2.
+\end{align}
+If this rate is low,  then the number of generations until a mutation arises that
+will become locally established within the patch is exponentially distributed with rate $\mutrate$.  
+Assuming that once a mutation becomes established it quickly reaches its equilibrium frequency across the patch, 
+the expected time it takes for our patch to become colonized by the $B$ allele is approximately $1/\mutrate$.
 
-\subsubsection{Time to the establishment of a locally adaptive allele
-  due to migrational influx}
+%%%%
+\subsection{Time to the establishment of a locally adaptive allele due to migrational influx}
 \label{ss:patchymigration}
 Now suppose that there are two patches, each of area $A$, of the poor
-habitat type separated by distance $R$. Consider the case where the
-$B$ allele has arisen and become established in the first patch, but has not yet appeared in the second.
-As above, any copy of the new allele that appears at location $x$ relative to the second, unadapted patch 
+habitat type separated by distance $R$. 
+We will consider the case where the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
+for the purpose of comparison ignoring the possibility that a $B$ allele arises independently by mutation in the first patch.
+As above, any copy of the new allele appears at location $x$ relative to the second, unadapted patch 
 has probability approximately $p(x)$ of fixing locally in this patch,
-whether the allele is a new mutation or is descended from the
-occupants of the first patch.
+whether the allele is a new mutation or is descended from the occupants of the first patch.
 
-For the moment we will ignore the
-possibility of mutation generating a $B$ allele can concentrate on the
-flow of migrants from patch $1$ to patch $2$.
-Due to \gc{migration}--selection balance there will be a small number of $B$ alleles descended from those of the first patch (``migrants'') 
+Due to migration--selection balance, 
+there will be a small number of $B$ alleles descended from those of the first patch
 present in the regions where they are disadvantageous;
-the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX, citeXXX),
+the expected frequency at distance $r$ may again be expressed in terms of Jacobi elliptic functions \citep[Appendix XXX][]{barton},
 although the equations are again not particularly helpful.
-However, we do know that if $q(r)$ is the expected proportion present at \gc{migration}--selection balance at distance $r$ from the first patch
-in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|\sqrt{s_m} r / \sigma|)$,
-where $C$ is a constant depending on the structure of the populations
-and the selection coefficients. \gc{comment on size of C? Should we
-  also use this approx. to get the approx. rate of establishing mutations
+This expected frequency is the \emph{time-averaged} occupation frequency,
+i.e.\ the average frequency of the allele at each location over a long time period,
+below we study the fluctuations about this mean.
+More helpful than the exact expressions,
+we know that if $q(r)$ is the expected proportion present at migration--selection balance at distance $r$ from the first patch
+in the absence of the second patch, 
+then for large $r$ this proportion decays exponentially, 
+\begin{align} \label{eqn:eqfreq}
+  q(r) \approx C \exp( -\sqrt{s_m} r / \sigma),
+\end{align}
+where $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
+(Probably $C=1/2$ is close enough for applications; see Appendix XXX.)
+\plr{same appendix with sims as above}
+\gc{Should we also use this approx. to get the approx. rate of establishing mutations
 arising outside of the patch, above. Feel like in the mutation rate
 part we could pull $s_B$ out front of the rate as well.} 
+\plr{huh?}
 
-\gc{Below are we setting $C=1$?}
+This allows us to get a good sense of the time scale of adaptation by migration between separated patches, as follows.
+First imagine that the new allele is not advantageous in the second patch,
+so that the expected number of alleles present in a small area $\Delta$ at distance $r$ from the already established patch
+is $\Delta \rho q(r)$.
+As these copies are disadvantageous, they form a subcritical branching process, which die out fairly soon.
+Now add to this local selective advantage by stipulating that any copy of the allele in the new patch
+has ``extra'' offspring that are marked in some way, on average $(1+s_B)-(1-s_m)=s_B+s_m$ additional offspring per generation.
+If the probability of establishment \emph{in the new patch} is $p(x)$ as above,
+then each of these new, marked copies appearing at location $x$ has probability $p(x)$ of establishing.
+(This general relation is stated formally in Appendix XXX.)
+Therefore, the total rate of establishment 
+(i.e.\ the mean number of successfully establishing migrant lineages per generation)
+is $\rho (s_B+s_m) \int q(|x|) p(x) dx$.
+Using a similar approximation as in section \ref{ss:patchymutation},
+we can approximate this, for a patch of area $A$ at distance $R$ from an already occupied patch, as
+\begin{align}
+  \migrate(R) = \rho A \frac{s_B}{\xi^2} (s_B+s_m) \exp\left( -\frac{ \sqrt{s_m} R}{\sigma} \right),
+\end{align}
+where we have set $C=1/2$ in equation \eqref{eqn:eqfreq}.
+Using different methods, we arrive at the same functional form below in section \ref{ss:haplotype_length}.
+(Note that if the size of the new patch is large relative to $\sqrt{s_m}/\sigma$,
+then the integral of $q(x)$ over the patch will not be well approximated by $A \times q(R)$,
+but the adjustment is easy to make.)
 
-\gc{work in a parameter that the the migrant outflux of a patch, and
-  explain that depends on geometry }
-
-Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
-supposing that patches are large enough for local fixation to be stable. If the alleles were not advantageous in the second patch,
-the average number of mutant alleles present in a small area $\Delta$
-near the second patch, due
-to migration from the first patch, would be $\approx \rho E \exp( -
-|\sqrt{s_m} R/\sigma| ) \Delta$. Assuming that the patch is
-sufficiently large to give an allele a $2s_B$ probability of becoming established, the rate at which migrants find
-themselves in the patch and become established  is approximately
-\begin{equation}
-\lambda_{Mig,R} = \rho E s_B \exp( -|\sqrt{s_m} R/\sigma| ).
-\end{equation}
-\gc{Is this $A^2$ right here?. Do we want to use $q(R)$ here?}
 If this establishment rate is low, the time it takes
-a mutation to arise that will become established will be  $\sim \Exp(\lambda_{mig,R})$
+a mutation to arise that will become established 
+will be approximately exponentially distributed with rate $\migrate$.
 Assuming that the time it takes for a allele from the first patch to
@@
 then  the expected waiting time until some migrant allele fixes in the
-second patch will be $T_{Mig} = 1/ \lambda_{Mig,R} $.
+second patch will be $1/\migrate$.
 
-\gc{Note that we can set patches sizes to be different}
+%%%
+\subsection{The probability of parallel adaptation between patches.} 
+\label{ss:probparallel}
 
-\gc{Put time it takes for a migrant allele to move between patches
-  here? Or below }
+We now turn to the question of whether the separated populations adapt by parallel genetic changes 
+or by the spread of a migrant allele between the patches. 
+% These two scenerios are assumed to be distinguishable, 
+% either because in the case of parallelism the mutations occur at different positions 
+% or if the mutations occur at the same position they can be distinguished due to their haplotypic background. 
+% Assuming that the geographically spread population initally contains no copies of the $B$ allele, 
+% one or other of the patches of different habitat 
+% will be the first to adapt via the establishment of a $B$ allele;
+% so the question is then the relative timescale of new mutation and migration to the other patch.
+As only a single mutation is required to adapt individuals to the
+poor habitat patch, subsequent mutations that arise after an allele becomes established on the patch gain no selective benefit. 
+Similarly, an allele introduced into a patch by migration will not be favored by selection to spread, 
+if the patch has already been colonized. 
+Therefore, mutations selectively exclude each other from patches, over short time scales, 
+and will only slowly mix together over longer time scales by drift.
 
-\gc{think we've covered this next part by $q(r)$ being number of migs. such
-that $2sq(r)$ is (for small $s$ the right answer.}
-If $T_m$ is large, then we would expect successful migrants to be rare,
-so that if the second patch adapts through migration, it is likely to come from only a single migrant.
-We leave aside the converse situation, for which there is likely to be more than one migrant
-that founds the population mutant individuals in the new patch.
+We will assume that once a $B$ allele is introduced (by migration or mutation), 
+it becomes established in the poor habitat patch rapidly. 
+Under this approximation, and our selective exclusion assumption, 
+after one of the patches becomes colonized by a single $B$ allele, 
+the other patch will become colonized by migration or mutation, but not by both. 
+As such, the question of whether the
+population adapts to the second poor habitat patch in parallel, 
+simply comes down to whether the a new mutation or a migrant allele is the first to become established in the second patch.
 
-\subsubsection{The probability of parallel adaptation between patches.} 
-\label{ss:probparallel}
-We now turn to the question of whether our populations adapt in
-parallel genetic changes or by the spread of a migrant allele between the
-patches. These two scenerios are assumed to be distinguishable, either
-because in the case of parallelism the mutations occur at different
-positions or if the mutations occur at the same position they can be
-distinguished due to their haplotypic background.
+Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
+we see that if the patches each have sufficiently large area,
+the rate of migration and mutation will be on the same order if 
+$\mu \approx ((s_B+s_m)/2) \exp(-\sqrt{s_m} R / \sigma)$,
+where $R$ is the distance between the patches.
+(Note that in both cases, $A$ is the area of the not-yet-adapted patch.)
 
-Assuming that we begin with our geographically 
-spread population containing no copies of the $B$ allele, one or other
-of our poor habitat patches will be the first to adapt via the
-establishment of a $B$ allele. The question
-
-As only a single mutation is required to adapt individuals to the
-poor habitat patch, subsequent mutations that arise after an allele
-becomes established at its equilibrium frequency on the patch gain no selective
-benefit. Similarly an allele introduced into a patch by migration will
-not be favored by selection to spread, if the patch has already been
-colonized. Thus, mutations selective exclude each other from
-patches, over short-time scales, and they will only slowly filter
-together over longer time scales by drift.
-
-We will assume that once a introduced $B$ allele (from migration or mutation) becomes
-established in the poor habitat patch it reaches its equilbrium
-frequency very rapidly. Under this approximation, and our selective
-exclusion assumption, after one of the patches becomes colonized by a
-single $B$ allele, the other patch will become colonized by migration or
-mutation, but not by both. As such, the question of whether the
-population adapts to the second poor habitat patch in parallel, simply comes down to whether the a new mutation or a migrant allele
-is the first to become established in the second patch.
-
-Taking the expressions from sections \ref{ss:patchymutation} and \ref{ss:patchymigration} above,
-we see that if patches each have sufficiently large area $A$,
-then the mutational input in each is on the order of $2 s_B \rho \mu A$,
-while the influx due to migration between patches separated by
-distance $R$ is on the order $2 s_B \rho E \exp(- |\sqrt{s_m}
-R/\sigma|)$. \gc{what to assume about A and $E$ here?}
-If the input from mutation and migration are comparable, i.e.\ if $\mu \approx \exp( -|\sqrt{s_m} R/\sigma| )$,
-then adaptation is expected to take place both by migration and
-mutation.
-
-If we take $\mu = 10^{-8}$, then this is satisfied if $R/\sigma \approx 18/\sqrt{s_m}$.
-so if the disadvantage to the new mutation between patches is fairly strong (say, $s_m=.05$),
-then the distance between patches must be about a hundred times the dispersal distance (i.e.\ $R/\sigma \approx 100$);
-while if $s_m$ is much smaller, say $s_m = .001$, 
-then the patches must be separated by many hundreds of times the dispersal distance (i.e.\ $R/\sigma \approx 500$).
+If we take $\mu = 10^{-8}$, and $s_B=s_m=s$, then this is satisfied if $R/\sigma \approx (18+\log(s))/\sqrt{s}$,
+so if the selective pressure is fairly strong (say, $s=.05$),
+then the distance between patches must be about 70 times the dispersal distance (i.e.\ $R/\sigma \approx 68$);
+while if $s$ is much smaller, say $s = .001$, 
+then the patches must be separated by many hundreds of times the dispersal distance (i.e.\ $R/\sigma \approx 368$).
 The situation does not change much if $\mu$ is larger -- even if $\mu = 10^{-6}$, 
-the required separation between patches is only reduced by 60\%.
-If the separation between patches is much larger -- say, by a factor of ten -- 
-then patches are effectively isolated -- most adaptation will be independent between patches.
+the required separation between patches is only reduced by $5/\sqrt{s_m}$.
+If the separation between patches is much larger than this value -- certainly, by a factor of ten -- 
+then patches are effectively isolated -- most adaptation will occur independent between patches.
 On the other hand, if the separation is much smaller, then a single
@@
 
-If we are willing to take these approximations at face
-value, and assuming both rates are small, the the time till the first
-patch is colonized will be approximately exponentially
-distributed with parameter $2 \lambda_{Mut}$. Then time till the
-second patch is subsequently colonized by the $B$
-allele (via either migration or new mutation) will be exponentially
-distributed with parameter $\lambda_{Mut}+\lambda_{Mig,R}$. Both of
-these times scale linearly with $1/(\rho s_{B})$, so that
-
-then probability of parallel
-adaptation is 
+If we are willing to take these approximations at face value, 
+and assuming both rates are small, 
+the time till the first patch is colonized by the $B$ allele will be approximately exponentially distributed with rate $2 \mutrate$.
+Following this, the time till the second patch is subsequently colonized 
+(via either migration or new mutation) 
+will be exponentially distributed with rate $\mutrate+\migrate$.
+Both these times scale linearly with $\xi^2/(\rho A s_{B})$, so that 
+the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
+and decrease with patch population size and benefical seleciton coefficient.
+Furthermore, the probability that the second adaptation is a new mutation,
+i.e.\ the probability of parallel adapatation, is
 \begin{equation}
-\frac{\lambda_{Mut}}{\lambda_{Mut} + \lambda_{Mig,R}} = \frac{\mu (A/E)}{\mu (A/E) + \exp(- |\sqrt{s_m} R/\sigma|}
+  \frac{\mutrate}{\mutrate+\migrate} = \frac{2\mu/(s_B+s_m)}{2\mu/(s_B+s_m) + \exp\left(- \sqrt{s_m} R / \sigma \right) },
 \end{equation}
 so that the probability of parallel mutation should increase
-approximately logistically with the distance between the patches at a
-rate $\sqrt{s_m} /\sigma$. However, we caution that these calculations
-are rough and so should be treated mainly as a rule of thumb.
+approximately logistically with the distance between the patches at rate $\sqrt{s_m} /\sigma$.
+
+%%%%
+\subsection{Multiple patches}
+
+\plr{Aim here is to discuss transition to discrete model?}
 
@@
 For simplicity we will assume that each of these patches has
-the same size and geometry (i.e. same $E$). A particular patch adapts in parallel through an independent
+the same size and geometry (i.e.\ same $E$). A particular patch adapts in parallel through an independent
 instance of the $B$ mutation at rate $2 s_B \mu \rho A$.
@@
 \begin{equation}
-\sum_{i=1}^{k} 2 s_B \rho E \exp(- |\sqrt{s_m} R_k/\sigma|).
+\sum_{i=1}^{k} 2 s_B \rho E \exp(- \sqrt{s_m} R_k/\sigma).
 \end{equation}
@@
 \section{Length of the hitchhiking haplotype}
+\label{ss:haplotype_length}
 
@@
 2) The frequency at O'Neill [another site off the lava, on opposite side from Tule]  is 9km from the lava and has a frequency of 34\% of the dark allele.
-I estimate that we need a sigma/sqrt(s) of $\sim$28km (i.e.
-ten times the characteristic length) to give a drop from 50\% to 34\%.
+I estimate that we need a sigma/sqrt(s) of $\sim$28km (i.e.\ ten times the characteristic length) to give a drop from 50\% to 34\%.
 

### 2013-02-14 patchy-selection-paper.tex
WorkingWiki importation

@@
 
-\plr{Aim here is to discuss transition to discrete model?}
+We can extend this reasoning to multiple patches of unadapted habitat 
+in which the $B$ allele can facilitate local adaptation. 
+Under our assumptions of mutational exclusion and rapid establishment, 
+each patch will adapt through establishment of a single copy of the $B$ allele, 
+either by migration or mutation.
+Since we assume that the times between successive establishments is the result of many nearly-independent attempts
+with small probability of success,
+the process of occupation is well approximated by a discrete-state Markov chain,
+whose state is the labeling of patches according to the colonizing allele
+(or none, for those that haven't yet adapted).
 
-We can extend this reasoning to multiple patches ($>2$) within our population's range
-where our $B$ allele can facilitate local adaptation. Under our
-assumptions of mutational exclusion and rapid establishment, each one
-of these patches will adapt through a single copy of the $B$
-alllele. The occupation of our different patches by the $B$ allele will form a
-discrete state markov chain, where the states of the chain are which patches have
-been colonized so far by $B$, and which colonized patches have shared
-versus distinct copy of $B$.
+If not yet adapted,
+a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_B \mu \rho A/\xi^2$.
+Without loss of generality, assume that patches $1, \ldots, k$ are already adapted,
+and these patches are at distances $R_1, \ldots, R_k$ away from this unadapted patch, 
+then the total rate of adaptation through migration is
+\begin{equation}
+  \frac{ s_B (s_B+s_m) \rho A}{\xi^2} \sum_{i=1}^{k} \exp\left(- \sqrt{s_m} R_i/\sigma\right),
+\end{equation}
+and the probability that the adaptation is from patch $i$ is $\exp\left(- \sqrt{s_m} R_i/\sigma\right)/\sum_{j=1}^{k} \exp\left(- \sqrt{s_m} R_j/\sigma\right)$.
+These specify the transition rates of the process.
 
-For simplicity we will assume that each of these patches has
-the same size and geometry (i.e.\ same $E$). A particular patch adapts in parallel through an independent
-instance of the $B$ mutation at rate $2 s_B \mu \rho A$.
-While, if patches $1$ through $k$ are already occupied by
-established copies of the $B$ allele, and these patches are at
-distances $R_1,~\cdots,~R_k$ away from our patch, the rate of
-adaptation through migration in our patch is 
-\begin{equation}
-\sum_{i=1}^{k} 2 s_B \rho E \exp(- \sqrt{s_m} R_k/\sigma).
-\end{equation}
-Once again the compound parameter $2s_B \rho$ will function to
-scale the times between events (parallel adaptation or the spread of a
-migrant), and so is critical to the time-scale over which
-adaptation occurs. However, this compound parameter again has no effect on
-the final configuration of the pattern of which patches have adapted
-in parallel versus have shared genetic basis to their adaptive
-response to the poor environment.
+Since the compound parameter $s_B \rho / \xi^2$ is common to both rates,
+it functions only as a time scaling of the process, 
+and therefore has no effect on the final configuration of the pattern of which patches have adapted
+in parallel versus have shared genetic basis to their adaptive response to the poor environment.
 
-The states of this
-
-%Furthermore, in the model of discrete demes connected by symmetric weak migration, 
-%the final distribution of types depends only on the ratio of mutation to migration rates,
-%and that the number of independently arisen types follows the Ewens distribution, with parameter $\mu/m$.
-%In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
-%for a model with 20 interconnected demes, across different parameter values.  
-%This indicates that we need the population-scaled mutation and migration rates 
-%to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.  
-%If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
-
-%\begin{figure}[ht]
-% \begin{center}
- %  \includegraphics{discrete-by-ratio}
- %\caption{ %
- %The mean number of types and the mean number of demes occupied by a sampled type, in a discrete model with 20 demes,
- %plotted as a function of the ratio of migration to mutation probabilities.
-% Note the log scale on the horizontal axis.
-% \label{fig:discbyratio}
- %}
- %\end{center}
-%\end{figure}
-
-Also note that the properties of the discrete model are independent of the {\em sizes} of the demes themselves--- we have assumed away all such dependence.
-This is complementary to the results of \cite{softsweepsII}, who showed that multiple mutations are likely to arise {\em within} a panmictic deme
+Also note that we can rescale time by the typical patch size, and introduce a parameter, say $a_k = A_k/\bar A$,
+making the properties (other that the time scaling) of the discrete model independent of the \emph{numerical sizes} of the demes themselves.
+This is complementary to the results of \cite{softsweepsII}, who showed that multiple mutations are likely to arise \emph{within} a panmictic deme
 if the population-scaled mutation rate $2 N \mu$ is greater than $1$.
 
+\plr{conclusion?}
+
 %%%%%%%%%
-\section{Length of the hitchhiking haplotype}
+\subsection{Length of the hitchhiking haplotype}
 \label{ss:haplotype_length}
@@
 %%%
-\subsection{Limiting distributions}
+\subsubsection{Limiting distributions}
 

### 2013-02-15 patchy-selection-paper.tex
/* Patchy Selection paper */

@@
 Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX.
-These mice live in the desert?
-through the American SouthWest and into Mexico.
-In the majority of their range they live on the light colored rocks, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}.
+These mice live throughout the American SouthWest and into Mexico, and mostly live on the light colored rock outcrops, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}.
 However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation.
@@
 The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel.
-However, there is some evidence \citep{Hoekstra:05} that some close populations may have adapted by the spread of another dark allele by migration.
+However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, some nearby populations may share a dark phenotype whose genetic basis has been spread by migration.
 
-As can be seen from above a key parameter is $\sigma/\sqrt{s}$.
+As can be seen from above a key parameter is the characteristic lengths $\sigma/\sqrt{s_m}$.
 \citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites.
-On the lava flow the dark MC1R allele is at $86\%$ frequency.
-At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency.
+ On the lava flow the dark MC1R allele is at $86\%$ frequency. At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency. 
 At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass. 
-While at the light substrate O'Neill site, which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.
+While at the other light substrate site (O'Neill), which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.
+On the basis of these numbers we can gain a sense of a plausible range of characteristic lengths.
+Assuming the allele is at $50\%$ frequency at the edge of the lava flow, 
+we can fit the frequency found at each of the light substrate sites that are 
+polymorphic for the dark allele under a simple allele frequency cline model \citep{XXX}.
+Doing this for the Tule we obtain $\sigma/\sqrt{s_m} \approx 4$km and for the O'Neill site
+$\sigma/\sqrt{s_m} \approx 30$km. \\
 
-I estimate that a characteristic length of 2.8km gives the frequency dropping from 50\% to $\sim$1\% in this distance.
-So this doesn't seem so off.
-A characteristic length that is 1.4x gives 2.8\% from 50\% in 12km, which may be a little more accurate.
-
-2) The frequency at O'Neill [another site off the lava, on opposite side from Tule]  is 9km from the lava and has a frequency of 34\% of the dark allele.
-I estimate that we need a sigma/sqrt(s) of $\sim$28km (i.e.\ ten times the characteristic length) to give a drop from 50\% to 34\%.
-
-[Obviously the cline isn't decaying from 50\%, the frequency of the light allele on the lava to be $\sim$14\%, we could put this in more carefully]
-So it seems to me that a sigma/sqrt(s) of 3-30km seems to encompass a range of interesting parameters, and can be justified without too much trouble.
-
-Using the dispersal distance of 0.28km [from marker recapture] and an s=0.01 [from paper] gives us sigma/sqrt(s)=2.8km
-
-HbS
-
-G6PD
+The mutational target size $\mu$ for the trait is unclear. 
+While the Pinacate dark haplotype differs from the light haplotype at three amino-acid residues,
+it is likely that not all of these changes are needed for a population to begin to  
+adapt. Also there a number of genes, beyond MC1R, at which adaptative changes affecting 
+pigmentation have been identified in closely related species and more
+broadly across vertibrates \citep{}.To span a range of plausible
+values, we use a low mutation rate 
+We use a low mutation rate $\mu= 10^{-8}$, where the adaptive
+mtation has to occur at a single base pair.
+An intermediate mutation rate $\mu= 10^{-5}$, equivalent to
+the mutation being able to arise at any of a thousand bases pairs,
+e.g. any of the coding bases of a typical length gene. And a high
+mutation rate $\mu = 10^{3}$, corresponding to the adaptive mutation
+being able to arise in any of $100$ genes, e.g. anywhere in a large pathway.
 

### 2013-02-25 patchy-selection-paper.tex
WorkingWiki importation

@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\calE}{\mathcal{E}}
+\newcommand{\calC}{\mathcal{C}}
 \newcommand{\dconv}{\xrightarrow{d}}
@@
 
+%%%%%%%%%%
+\section{The probability of establishment}
+\label{ss:prob_estab}
+
+In the main text we have made use of two decompositons of the spatial branching process
+to approximate in two different ways the rate of establishment of migrant alleles.
+Here we make the relationship more formal.
+The decompositions use a similar idea to that found in \citep[section D.12]{athreyaney},
+where a supercritical branching process is decomposed into individuals whose families eventually die out
+and those having infinite lines of descent.
+We also note that the approximations fall generally in the realm covered by \citet{aldouspoissonclumping}.
+
+Consider a discrete-time branching process $Z_t$ with spatial motion and offspring distribution depending on spatial location.
+Suppose that the offspring distribution for an individual at location $x$ is distributed as $X_x$.
+We can record the state of the process simply as the list of locations of individuals:
+$Z_t = (x_1, \ldots, x_{N(t)})$ means that there are $N(t)$ individuals alive at $t$, at locations $x_1, \ldots, x_{N(t)}$.
+We can construct the process iteratively: given $Z_t$,
+we sample first the number of offspring of each, and then the locations the offspring migrate to.
+Concretely, given $Z_{t} = (x_1, \ldots, x_{N(t)})$,
+first let $X_1(t),\ldots,X_{N(t)(t)}$ be independent random draws from the offspring distribution,
+with $X_k(t)$ having distribution $X_{x_k}$.
+The number of new offspring is $N(t+1) = \sum_{k=1}^{N(t)} X_k(t)$,
+and we can sample their locations independently given the location of their parent to obtain $Z_{t+1}$.
+Note that the genealogy is implicitly recorded:
+the parent of the $k^\mathrm{th}$ individual at time $t$ is
+the $a(k,t) = \min\{ \ell : \sum_{j=1}^\ell X_j(t-1) \ge k \}$-th individual at time $t-1$.
+It will be useful to extend this to more distant ancestors:
+let $a_1(k,t) = a(k,t)$ and recursively define $a_m(k,t) = a(a_{m-1}(k,t),t-m+1)$, as long as $m\le t$.
+
+We want to find (approximations to) $p(x)$, the probability that a branching process beginning with a single individual at $x$ does not die out. 
+(In other words, $p(x) = \P\{ N(t) = 0 \; \mbox{for some}\; t>0 \; \vert \; Z_0 = (x) \}$.)
+We assume that the mean number of offspring $\E[X_x]$ is only greater than one for $x \in A$ (``the patch''),
+and so the process can only establish if some offspring eventually get to the patch (with probability one).
+
+The second decomposition we use is simpler, only conditioning on the number of individuals who ever reach the patch.  
+Let $\calE = \{(t_i,k_i)\}$ be the times and indices of individuals who are in the patch but none of whose ancestors have been --
+i.e.\ $(t,k) \in \calE$ iff $x_k(t) \in A$ and $a_m(k,t) \notin A$ for all $1\le m \le t$.
+If we remove these individuals (and their offspring), the resulting branching process has mean number of offspring less than or equal to 1 everywhere,
+so there are a finite number of such individuals with probability one (the process dies out).
+Therefore, the probability that the process dies out is then the probability that the families descending from each individual in $\calE$ dies out.
+By independence of different offspring, this is
+\begin{align}
+  1-p(x) = \E\left[ \prod_{(t,k) \in \calE} (1-p(x_k(t))) \; \vert \; Z_0 = (x) \right] .
+\end{align}
+Above, we make the approximation that $p(x_k(t)) \approx p_e$,
+the probability of establishment ignoring the possibility of migrating outside the patch.
+In this case,
+\begin{align}
+  p(x) \approx \E\left[ p_e^{|\calE|} \; \vert \; Z_0 = (x) \right] .
+\end{align}
+Here $|\calE|$ is the ``size of the family that reaches the patch'';
+in section \ref{ss:haplotype_length} we estimated the probability that any of the family reaches the patch
+and argued that the distribution of $|\calE|$ is given by $K$,
+the size of the subcritical branching process conditioned on still being alive.
+
+The first approximation uses a similar trick, but rather than treating all individuals migrating into the patch
+as an opportunity for establishment,
+we couple the branching process to a strictly subcritical branching process
+that produces extra offspring, each of which provides an opportunity for establishment.
+To this end, suppose that the offspring distribution off the patch is given by $X'$ (so $X_x \deq X'$ for $x \notin A$),
+and the offspring distribution within the patch stochastically dominates $X$,
+i.e.\ $\P\{X_x\ge n\} \ge \P\{X'\ge n\}$ for all $x\in A$ and $n \ge 0$.
+Stochastic domination implies that we can \emph{couple} any instance of $X_x$ with one of $X'$ and a remainder --
+i.e.\ jointly construct an $X'$ and a $Y_x$ such that $X_x = X' + Y_x$,
+and that $Y_x$ also takes values in the nonnegative integers.
+We can therefore modify the construction of the branching process above,
+so that $X_k(t) = X_k'(t) + Y_k(t)$, where $X_k'(t) \deq X$ and $Y_k(t) \deq Y_{x_k(t)}$.
+We can define a \emph{core} branching process $Z'$ using the $X'$, which will be finite (since $\E[X']<1$).
+Formally, say that the first $X_k'(t)$ offspring of the $(t,k)^\mathrm{th}$ individual are ``core'' offspring,
+denoted by $\calC_t = \{ (t,k) \; : \; 0 \le k - \sum_{j<\ell} X_j(t) \le X_\ell'(t) \; \mbox{for some}\; 1 \le \ell \le N(t) \}$;
+and define $Z'$ as the core individuals whose ancestors are all ``core'' as well,
+i.e.\ $Z'_t = \{ x_k \in Z_t \; : \;  a_m(k,t) \in \calC_{t-m} \; \mbox{for all}\; 0 \le m \le t\}$.
+The boundary analogous to above in the first computation is then the non-core offspring of core individuals;
+denote the times and indices of these by $\partial \calC$,
+so that $(t,k) \in \partial \calC$ iff $(t,k) \notin \calC_t$ and $a(k,t) \in \calC_{t-1}$.
+We can now use this as above to compute the probability of establishment beginning from a single individual:
+\begin{align}
+  1-p(x) = \E\left[ \prod_{(t,k) \in \partial\calC} (1-p(x_k(t))) \; \vert \; Z_0 = (x) \right] ,
+\end{align}
+but this is less helpful, as the distribution of $\calC$ is harder to get at.
+
+However, recall the actual problem at hand.
+Roughly speaking, we approximate the process of colonization of the new patch by migrants from an established patch
+by imagining that the established patch sends out migrants, and that each of these acts as an independent branching process,
+at least in the region of the new patch.
+We can suppose that migrant families appear at some rate, say, midway between the patches, and far enough from the original one that we can ignore it,
+and leave the precise distribution of their appearance unspecified
+We can decompose each of these branching processes as above into ``core'' and ``extra'',
+so that the point in time of ``establishment'' is well-defined:
+the time $\tau$ of first appearance of an ``extra'' offspring with infinite line of descent.
+Given that establishment has not yet occurred by $t$, 
+and the combined state of the core process $Z'_t$,
+then the mean number of establishment events occurring at time $t$ is $\sum_{x \in Z'_t} p(x) \E[Y_x]$.
+
+If establishment events were independent and sufficiently rare,
+the influx of branching processes begain long enough ago that $Z'$ was at stationarity
+with $z_A := \E[ \# \{ (x,t) \in Z'_t : x \in A \} ]$,
+and we again approximate $p(x) \approx p_e$,
+then we would have that 
+\begin{align}
+  \P\{\tau>t\} \approx \exp\left\{ - t z_A \E[Y_x] p_e \right\} ;
+\end{align}
+in our case, $\E[Y_x] \approx (s_B+s_m)$ and $p_e \approx 2 s_B/\xi^2$,
+so $\tau$ is approximately exponentially distributed with rate $ 2 z_A (s_B+s_m) s_B/\xi^2 $.
+We can get $z_A$, the occupation density without the new patch, 
+directly by integrating the expression~\eqref{eqn:eqfreq} over the patch.
+
+In obtaining an exponential distribution, we have assumed effectively that establishment events
+are independent of each other.
+This will not be strictly true, since an establishment implies that a family of migrants are present in the patch,
+and hence others may produce offspring with infinite lines of descent at a similar time.
+The mean total family size of an allele with fitness $1-s_m$ (i.e.\ a subcritical branching process) is $\sum_{t \ge 0} (1-s_m)^t = 1/s_m$;
+so the mean number of ``extra'' offspring is $1+s_B/s_m$.
+Establishment events are correlated to the extent that single such families give rise to multiple establishment events,
+
+% We will record the state of the branching process as a finite point measure,
+% so that if there are $n$ individuals alive at $t$, at locations $x_1, \ldots, x_n$,
+% then $Z_t = \sum_{k=1}^n \delta_{x_k}$, and we write $Z_t(A) = \sum_{k=1}^n \bone_{x_k \in A}$ for the number of individuals in a set $A$.
+
+%%%%%%%%%%%
 \section{Patchy selection} 

### 2013-02-25 patchy-selection-paper.tex
WorkingWiki importation

@@
 By integrating over $t$, we get the chance that descendants of the migrant ever arrive at the new patch is
-\begin{align} \label{eqn:estab_prob}
+\begin{align} 
   & \int_0^\infty \frac{(1-s_m)^t}{\mu_K} \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right)  dt \\
-  & \qquad = \frac{1}{\mu_K} \exp\left( - \frac{ x \sqrt{\alpha}}{\sigma} \right) .
+  & \qquad = \frac{1}{\mu_K} \exp\left( - \frac{ x \sqrt{\alpha}}{\sigma} \right) .\label{eqn:estab_prob}
 \end{align}
@@
 Therefore, the probability of establishment is
-\begin{align}
+\begin{align} \label{eqn:bp_prob_estab}
   F(x) &\approx \frac{2s_b}{\xi^2} \exp\left( - \frac{\sqrt{\alpha} x}{\sigma} \right) ,
@@
 \label{ss:prob_estab}
+
+\plr{Notes for clarification.}
+We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
+\begin{align}
+  (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \\
+  &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
+  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \\
+  &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \\
+  & \qquad \qquad \times (\mbox{prob of estab in the patch}) .
+\end{align}
+The probability of establishment of a branching process is in equation~\eqref{eqn:bp_prob_estab};
+the subcritical occupation density is in equation~\eqref{eqn:eqfreq}, up to a constant $C$ depending on geometry;
+the number of (extra) offspring per subcritical individual is $s_B+s_m$;
+and the probability of establishment in the patch is $2s_B/\xi^2$.
+The ``outflux of branching processes'' we have not computed;
+but we know that the probability a branching process arrives at the patch is given by equation~\eqref{eqn:estab_prob};
+if it does it is present in $\mu_K$ copies, each of which leave a total of $1/s_m$ offspring; 
+so equating ($\rho$ times equation~\eqref{eqn:eqfreq}) and (outflux times $\mu_K/s_m$ times equation~\eqref{eqn:estab_prob})
+we have that the effective outflux (at the edge of the already occupied patch!) is $C \rho s_m$,
+rather than $C\rho$ which we might have guessed.
+The number of establishment events per BP is as follows, in averages: 
+with prob $1/\mu_K \exp(-\sqrt{\alpha} x/ \sigma)$ a total of $\mu_K$ indivs arrive at the new patch; 
+each produce $1/s_m$ total core offspring;
+each of these produce $s_B+s_m$ extra offspring which each have chance $2s_B/\xi^2$ of establishing.
+This squares with the previous argument.
+\textbf{Now} note that what we \textbf{actually} want to compute is not this rate, but rather 
+\begin{align}
+  \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a BP}) \\
+  &= \frac{\rho C}{s_m} \times \frac{2 s_b}{\xi^2} \exp\left(- \frac{\sqrt{\alpha}x}{\sigma} \right) .
+\end{align}
+\plr{end of notes}
 

### 2013-02-28 patchy-selection-paper.tex
WorkingWiki importation

--- 2013-02-25 patchy-selection-paper.tex
@@
 \label{ss:prob_estab}
+
+\plr{Notes for clarification.}
+We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
+\begin{align}
+  (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \\
+    \begin{split} &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
+  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split} \\
+  \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \\
+    & \qquad \qquad \times (\mbox{prob of estab in the patch}) . \end{split} 
+\end{align}
+The probability of establishment of a branching process is in equation~\eqref{eqn:bp_prob_estab};
+the subcritical occupation density is in equation~\eqref{eqn:eqfreq}, up to a constant $C$ depending on geometry;
+the number of (extra) offspring per subcritical individual is $s_B+s_m$;
+and the probability of establishment in the patch is $2s_B/\xi^2$.
+The ``outflux of branching processes'' we have not computed;
+but we know that the probability a branching process arrives at the patch is given by equation~\eqref{eqn:estab_prob};
+if it does it is present in $\mu_K$ copies, each of which leave a total of $1/s_m$ offspring; 
+so equating ($\rho$ times equation~\eqref{eqn:eqfreq}) and (outflux times $\mu_K/s_m$ times equation~\eqref{eqn:estab_prob})
+we have that the effective outflux (at the edge of the already occupied patch!) is 
+\begin{align}
+  (\mbox{outflux of BPs from old patch}) = C \rho s_m ,
+\end{align}
+rather than $C\rho$ which we might have guessed.
+The number of establishment events per BP is as follows, in averages: 
+with prob $1/\mu_K \exp(-\sqrt{\alpha} x/ \sigma)$ a total of $\mu_K$ indivs arrive at the new patch; 
+each produce $1/s_m$ total core offspring;
+each of these produce $s_B+s_m$ extra offspring which each have chance $2s_B/\xi^2$ of establishing,
+i.e.
+\begin{align}
+  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{\alpha} x}{\sigma}\right) .
+\end{align}
+\textbf{Now} note that what we \textbf{actually} want to compute is not this rate, but rather 
+\begin{align}
+  \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
+  &= C \rho s_m \times \frac{2 s_b}{\xi^2} \exp\left(- \frac{\sqrt{\alpha}x}{\sigma} \right) .
+\end{align}
+Finally, combining these, we can also compute that
+\begin{align}
+  (\mbox{\# estab events per successful BP}) = \frac{s_B+s_m}{s_m} .
+\end{align}
+\plr{end of notes}
 

### 2013-03-01 branching-transit.tex
WorkingWiki importation

@@
 \newcommand{\Exp}{\mathop{\mbox{Exp}}}
+\newcommand{\migrate}{\lambda_\text{mig}}
 
@@
 
+\subsection{Occupation density}
+
+Suppose we have branching processes beginning at the origin at some constant rate $\lambda$ in time.
+Then the occupation density at distance $x$, i.e.\ the expected number of particles found at location $x$ at any given time, at stationarity,
+is $\lambda \int_0^\infty \E[Z_t(dx)] dt$, where $Z_t$ is the branching process.
+Since spatial motion is independent of branching,
+\begin{align}
+  \E[Z_t(dx)] = e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) ,
+\end{align}
+and so using G\&R 3.471.9 and 8.468 again,
+\begin{align} \label{eqn:bp_occupation}
+  \int_0^\infty \E[Z_t(dx)] dt  &= \int_0^\infty e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) dt \\
+  &= \begin{cases}
+    \frac{ 2^{1/4} x^{1/2} }{ \sqrt{\pi} \sigma^{3/2} \alpha^{1/4} } K_{1/2} \left(\frac{x \sqrt{2\alpha}}{\sigma}\right) =
+    \frac{1}{\sigma \sqrt{2 \alpha}} \exp\left( - \frac{ \sqrt{2\alpha} x }{ \sigma } \right) \qquad & d=1 \\
+    \frac{1}{\sigma^2 \pi} K_0\left(\frac{x \sqrt{2\alpha}}{\sigma}\right) \sim \frac{ e^{-\frac{x\sqrt{2\alpha}{\sigma}}} }{ \sqrt{ \sigma^3  2\pi x \qquad & d=2  ,
+\end{cases} 
+\end{align}
+where $K_0(z)$ is the modified Bessel function of the second kind of order 0.
+Note that this is the Green function for a Brownian motion killed at rate $\alpha$.
+
+A more back-of-the-envelope calculation of this would be:
+we know that the probability a branching process arrives at the patch is given by $(1/\mu_K)\exp(-x\sqrt{2 \alpha}/\sigma)$;
+if it does it is present in $\mu_K$ copies, each of which leave a total of $1/s_m$ offspring.
+To translate this to the above we need the typical area covered by these (in some sense?), 
+which must be around $\sigma \sqrt{2 \alpha}$.
+
+\subsection{An integral}
+
+For reference, by G\&R 3.471.9 and 8.468,
+\begin{align}
+  \int_0^\infty x^{n+1/2} e^{-\frac{\beta}{x}-\gamma x} dx &= 2 \left( \frac{\beta}{\gamma} \right)^{\frac{2n+1}{4}} K_{n+3/2}(2\sqrt{\beta\gamma}) \quad \mbox{(for general $n$)} \\
+  \mbox{($n$ an integer)} \quad &= \sqrt{\pi} \beta^{\frac{n}{2}} \gamma^{-\frac{n+1}{2}} e^{-2\sqrt{\beta\gamma}} \sum_{k=1}^n \frac{ (n+k)! }{ k! (n-k)! (4\sqrt{\beta \gamma})^k } .
+\end{align}
+
+Note that (8.451.6) $K_\nu(z) = \sqrt{\frac{\pi}{2z}} e^{-z} (1 + O(1/z))$.
+
+\section{The big picture}
+
+Here are notes to clarify what's going on for writing the main paper.
+
+We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
+\begin{align}
+  (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \\
+    \begin{split} &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
+  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split} \\
+  \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \\
+    & \qquad \qquad \times (\mbox{prob of estab in the patch}) . \end{split} 
+\end{align}
+The probability of establishment of a branching process is in equation~\eqref{eqn:bp_prob_estab};
+the subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
+the number of (extra) offspring per subcritical individual is $s_B+s_m$;
+and the probability of establishment in the patch is $2s_B/\xi^2$.
+The ``outflux of branching processes'' we have not computed;
+but we have compute the occupation density from this branching process approximation above,
+so equating ($\rho$ times $C\exp(-\sqrt{2 s_m}x/\sigma)$) and (outflux times equation~\eqref{eqn:bp_occupation})
+we have that the effective outflux (from the edge of the already occupied patch!) is 
+\begin{align}
+  (\mbox{outflux of BPs from old patch}) = \frac{ C \rho \sqrt{\alpha} }{ \sqrt{\pi} } ,
+\end{align}
+rather than $C\rho$ which we might have guessed.
+The number of establishment events per BP is as follows, in averages: 
+with prob $1/\mu_K \exp(-\sqrt{\alpha} x/ \sigma)$ a total of $\mu_K$ indivs arrive at the new patch; 
+each produce $1/s_m$ total core offspring;
+each of these produce $s_B+s_m$ extra offspring which each have chance $2s_B/\xi^2$ of establishing,
+i.e.
+\begin{align}
+  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{\alpha} x}{\sigma}\right) .
+\end{align}
+\textbf{Now} note that what we \textbf{actually} want to compute is not this rate, but rather 
+\begin{align}
+  \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
+  &= C \rho \sqrt{\alpha} \times \frac{2 s_b}{\sqrt{\pi} \xi^2} \exp\left(- \frac{\sqrt{\alpha}x}{\sigma} \right) .
+\end{align}
+Finally, combining these, we can also compute that
+\begin{align}
+  (\mbox{\# estab events per successful BP}) = \frac{s_B+s_m}{ \sqrt{s_m} } .
+\end{align}
+
 \end{document}

### 2013-03-01 patchy-selection-paper.tex
WorkingWiki importation

--- 2013-02-25 patchy-selection-paper.tex
@@
 \begin{align} \label{eqn:eqfreq}
-  q(r) \approx C \exp( -\sqrt{s_m} r / \sigma),
+  q(r) \approx C \exp( -\sqrt{2 s_m} r / \sigma),
 \end{align}
@@
 \begin{align}
-  \migrate(R) = \rho A \frac{s_B}{\xi^2} (s_B+s_m) \exp\left( -\frac{ \sqrt{s_m} R}{\sigma} \right),
+  \migrate(R) = \rho A \frac{s_B}{\xi^2} (s_B+s_m) \exp\left( -\frac{ \sqrt{2 s_m} R}{\sigma} \right),
 \end{align}
@@
 the rate of migration and mutation will be on the same order if 
-$\mu \approx ((s_B+s_m)/2) \exp(-\sqrt{s_m} R / \sigma)$,
+$\mu \approx ((s_B+s_m)/2) \exp(-\sqrt{2 s_m} R / \sigma)$,
 where $R$ is the distance between the patches.
@@
 \begin{equation}
-  \frac{\mutrate}{\mutrate+\migrate} = \frac{2\mu/(s_B+s_m)}{2\mu/(s_B+s_m) + \exp\left(- \sqrt{s_m} R / \sigma \right) },
+  \frac{\mutrate}{\mutrate+\migrate} = \frac{2\mu/(s_B+s_m)}{2\mu/(s_B+s_m) + \exp\left(- \sqrt{2 s_m} R / \sigma \right) },
 \end{equation}
@@
 \begin{equation}
-  \frac{ s_B (s_B+s_m) \rho A}{\xi^2} \sum_{i=1}^{k} \exp\left(- \sqrt{s_m} R_i/\sigma\right),
+  \frac{ s_B (s_B+s_m) \rho A}{\xi^2} \sum_{i=1}^{k} \exp\left(- \sqrt{2 s_m} R_i/\sigma\right),
 \end{equation}
-and the probability that the adaptation is from patch $i$ is $\exp\left(- \sqrt{s_m} R_i/\sigma\right)/\sum_{j=1}^{k} \exp\left(- \sqrt{s_m} R_j/\sigma\right)$.
+and the probability that the adaptation is from patch $i$ is $\exp\left(- \sqrt{2 s_m} R_i/\sigma\right)/\sum_{j=1}^{k} \exp\left(- \sqrt{2 s_m} R_j/\sigma\right)$.
 These specify the transition rates of the process.
@@
 By integrating over $t$, we get the chance that descendants of the migrant ever arrive at the new patch is
-\begin{align} \label{eqn:estab_prob}
+\begin{align} 
   & \int_0^\infty \frac{(1-s_m)^t}{\mu_K} \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right)  dt \\
-  & \qquad = \frac{1}{\mu_K} \exp\left( - \frac{ x \sqrt{\alpha}}{\sigma} \right) .
+  & \qquad = \frac{1}{\mu_K} \exp\left( - \frac{ x \sqrt{2 \alpha}}{\sigma} \right) .\label{eqn:estab_prob}
 \end{align}
@@
 Therefore, the probability of establishment is
-\begin{align}
-  F(x) &\approx \frac{2s_b}{\xi^2} \exp\left( - \frac{\sqrt{\alpha} x}{\sigma} \right) ,
+\begin{align} \label{eqn:bp_prob_estab}
+  F(x) &\approx \frac{2s_b}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} x}{\sigma} \right) ,
 \end{align}
@@
 \[
-    G(a,b,x) = \sum_{n \ge 0} = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
+    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
 \]

### 2013-03-01 branching-transit.tex
WorkingWiki importation

@@
   &= \begin{cases}
-    \frac{ 2^{1/4} x^{1/2} }{ \sqrt{\pi} \sigma^{3/2} \alpha^{1/4} } K_{1/2} \left(\frac{x \sqrt{2\alpha}}{\sigma}\right) =
-    \frac{1}{\sigma \sqrt{2 \alpha}} \exp\left( - \frac{ \sqrt{2\alpha} x }{ \sigma } \right) \qquad & d=1 \\
-    \frac{1}{\sigma^2 \pi} K_0\left(\frac{x \sqrt{2\alpha}}{\sigma}\right) \sim \frac{ e^{-\frac{x\sqrt{2\alpha}{\sigma}}} }{ \sqrt{ \sigma^3  2\pi x \qquad & d=2  ,
+    \sqrt{ \frac{ 2 x }{ \pi \sigma^{3} \sqrt{2\alpha} } }  K_{1/2} \left(\frac{x \sqrt{2\alpha}}{\sigma}\right) =
+    \frac{1}{\sigma \sqrt{2 \alpha}} \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) \qquad & d=1 \\
+    \frac{1}{\sigma^2 \pi} K_0\left(\frac{x \sqrt{2\alpha}}{\sigma}\right) 
+    \sim \frac{ 1 }{ \sqrt{ \sigma^3  2\pi x \sqrt{2\alpha} } } \exp\left(-\frac{x\sqrt{2\alpha}}{\sigma}\right) \qquad & d=2  ,
 \end{cases} 
@@
 if it does it is present in $\mu_K$ copies, each of which leave a total of $1/s_m$ offspring.
-To translate this to the above we need the typical area covered by these (in some sense?), 
-which must be around $\sigma \sqrt{2 \alpha}$.
+To translate this to the above we would need the typical area covered by these (in some sense?).
+% which must be around $\sigma \sqrt{2 \alpha}$.
 
@@
 \begin{align}
-  \int_0^\infty x^{n+1/2} e^{-\frac{\beta}{x}-\gamma x} dx &= 2 \left( \frac{\beta}{\gamma} \right)^{\frac{2n+1}{4}} K_{n+3/2}(2\sqrt{\beta\gamma}) \quad \mbox{(for general $n$)} \\
+  \int_0^\infty x^{n-1/2} e^{-\frac{\beta}{x}-\gamma x} dx &= 2 \left( \frac{\beta}{\gamma} \right)^{\frac{2n+1}{4}} K_{n+1/2}(2\sqrt{\beta\gamma}) \quad \mbox{(for general $n$)} \\
   \mbox{($n$ an integer)} \quad &= \sqrt{\pi} \beta^{\frac{n}{2}} \gamma^{-\frac{n+1}{2}} e^{-2\sqrt{\beta\gamma}} \sum_{k=1}^n \frac{ (n+k)! }{ k! (n-k)! (4\sqrt{\beta \gamma})^k } .
@@
 
-Note that (8.451.6) $K_\nu(z) = \sqrt{\frac{\pi}{2z}} e^{-z} (1 + O(1/z))$.
+Note that (8.451.6) $K_\nu(z) = \sqrt{\frac{\pi}{2z}} e^{-z} (1 + O(1/z))$ for all $\nu$, and $K_{-\nu}(z)=K_\nu(z)$.
 
+We have $n-1/2 = - d/2$, $\beta=x^2/2\sigma^2$, and $\gamma = \alpha$, leading to
+\begin{align}
+\frac{1}{(2\pi\sigma^2 )^{d/2}} \int_0^\infty x^{n-1/2} e^{-\frac{\beta}{x}-\gamma x} dx \sim
+\frac{ 1 }{ (2\pi)^\frac{d-1}{2} \sigma^d } \left( \frac{\sigma}{x} \right)^{\frac{d-1}{2}} ({2\alpha})^\frac{3-d}{4} \exp\left( - \frac{ x \sqrt{2 \alpha} }{ \sigma } \right) \quad \mbox{as}\; \frac{x\sqrt{2\alpha}}{\sigma} \to \infty .
+\end{align}
+
+%%%%%%%%%
 \section{The big picture}
@@
 The ``outflux of branching processes'' we have not computed;
-but we have compute the occupation density from this branching process approximation above,
+but we have computed the occupation density from this branching process approximation above,
 so equating ($\rho$ times $C\exp(-\sqrt{2 s_m}x/\sigma)$) and (outflux times equation~\eqref{eqn:bp_occupation})
@@
 \begin{align}
-  (\mbox{outflux of BPs from old patch}) = \frac{ C \rho \sqrt{\alpha} }{ \sqrt{\pi} } ,
+  (\mbox{outflux of BPs from old patch}) = C \rho \sigma \sqrt{2 \alpha} ,
 \end{align}

### 2013-05-31 patchy-selection-paper.tex
WorkingWiki importation

@@
 
+\section{The equilibrium frequency}
+
+One route to the ``equlibrium frequency'' of the allele outside the range where it is advantageous is as follows;
+see \citet{slatkin} or \citet{barton,pollack} for other arguments in this case, or \citet{etheridge} or \citet{dawson} for general theory
+Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
+and that selection (for or against) the allele is given by the function $s(x)$, with $x$ denoting the spatial location.
+Then each individual at location $x$ reproduces after a random lifetime,
+dying and giving birth in the same deme to a random number of offspring with distribution given by $X$;
+who all migrate to a new location chosen as $x+R$
+and replace randomly chosen individuals there.
+If $x+R$ is outside of the range, then they perish.
+Each individual's reproduction time is exponentially distributed, 
+either with rate 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
+Suppose that $R$ has mean zero and variance $\sigma^2$, and that $X$ has mean $\mu$ and variance $\xi^2$.
+We also assume that the distribution of $X$ makes the resulting Markov chain irreducible and aperiodic.
+
+Let $\Phi^N_t(x)$ be the number of mutant alleles present at location $x$ at time $t$,
+and denote by $\delta_x$ a single unit at location $x$, so that e.g.~$\Phi^N_t + \delta_x/N$
+is the configuration after a mutant allele has been added to location $x$.
+For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the number of mutant alleles added if $X$ new offspring carrying mutant alleles
+replace randomly chosen individuals in a deme where the mutant allele is at frequncy $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
+similarly, $\tilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
+(We like to think of $\Phi^N_t$ as a measure, but it does not hurt to think of $\Phi^N$ as a vector;
+we aren't providing the rigorous justification here.)
+Then we know that for any sufficiently nice function $f(\phi)$ that
+\begin{align} \label{eqn:discrete_generator}
+  \begin{split} \frac{\partial}{\partial t} \E\left[ f(\Phi^N_t) \right] 
+  &= N \sum_x \left\{ \E\left[ (1+s(x+R)) \Phi^N_t(x+R) \left( f\left(\Phi^N_t + \frac{\bar X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right. \\
+     & \qquad  \qquad \left.
+{} + \E\left[ \left(1-\Phi^N_t(x+R)\right) \left( f\left(\Phi^N_t - \frac{\tilde X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right\}  \end{split} \\
+     &= \mu \sum_x \E\left[ \left(\partial_{\phi(x)} f(\Phi_t) \right) \left\{ \Phi_t(x+R) - \Phi_t(x) + s(x+R) \Phi_t(x+R) (1-\Phi_t(x)) \right\} \right] + O\left(\frac{1}{N}\right).
+\end{align}
+This follows in part from $\E[\bar X_\phi] = \phi \mu$ and $\E[\tilde X_\phi] = (1-\phi)\mu$.
+We can see two things from this:
+First, since this is a first-order differential operator, the limiting of the stochastic process
+is deterministic (check by applying to $f(\phi) = \phi(x)^2$ to find the variance).
+Second, if we want to rescale space as well to get the usual differential equation, 
+we need to choose $\var[R]=\sigma^2$ and $s(x)$ to be of the same, small, order, another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale.
+More concretely, suppose that the grid size is $\epsilon \to 0$, that $\var[R] = \sigma^2 \epsilon$, and that $s(x)/\epsilon \to \gamma(x)$,
+and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
+then the previous equation converges to the familiar form:
+\begin{align}
+  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \partial_x^2 \xi(t,x) + \gamma(x) \xi(t,x) (1-\xi(t,x)) \right) .
+\end{align}
+
+The process $\Phi_t$ is an irreducible, aperiodic finite-state Markov chain with absorbing states at 0 and 1;
+therefore, the inevitable outcome for finite $N$ is extinction of one type or another.
+Setting the left-hand side to zero in \eqref{eqn:discrete_generator} gives us equations for the moments of the stationary distribution,
+for instance, if $\phi(x) = \lim_{t \to \infty} \E[ \Phi_t(x) ]
+
 \end{document}

### 2013-05-31 branching-transit.tex
WorkingWiki importation

@@
 if $\tau_x$ is the random time of arrival of a successful lineage,
-\begin{align}
+\begin{align} \label{eqn:arrival_time_density}
  \P\{\tau_x \in dt\} = C_{s_d,x} \frac{(1-s_d)^t}{t^{3/2}} \exp\left(-\frac{x^2}{2t}\right) ,
@@
 which is something involving the error function, or for large $x$,
-is the width of the patch multiplied by $1/(2 \pi \sigma^2 t)$.  
-This replaces the $t^{3/2}$ by a $t^{5/2}$, and back to G\&R, we have that the Laplace transform is in terms of $K_{3/2}$ instead of $K_{1/2}$;
-and this is found in G\&R 8.468, $K_{3/2}(z) = \sqrt{\pi/2z} e^{-z} (1+1/z)$.
-The formula corresponding to equation \eqref{eqn:first_integral} above is then
+is the width of the patch multiplied by $1/\sqrt{2 \pi \sigma^2 t}$.  
+This replaces the $t^{3/2}$ in \eqref{eqn:arrival_time_density} by $t^2$, and back to G\&R, we have that the Laplace transform is in terms of $K_{1}$ instead of $K_{1/2}$;
+in this case, we have that
 \begin{align}
-  \E[e^{-\ell \tau_x}] \propto \sqrt{\pi} \frac{2\sqrt{\ell+\alpha}}{x^2} e^{-x\sqrt{2(\ell+\alpha)}} \left( 1 + \frac{1}{x\sqrt{2(\ell+\alpha)}} \right) ,
+    \P\{ \tau < \infty \} &= \frac{x dy}{2\pi \sigma^4} \int_0^\infty t^{-2} \exp\left\{-\frac{x^2}{2\sigma^2 t} - \alpha t \right\} dt \\
+                          &= \frac{x dy}{2\pi \sigma^4} 2 \left( \frac{ x^2 }{ 2 \alpha \sigma^2 } \right)^{-1/2} K_1\left(2 \sqrt\frac{x^2 \alpha}{2\sigma^2} \right) \\
+                          &= \frac{dy\sqrt{2\alpha}}{\pi \sigma^3} K_1\left(\frac{x \sqrt{2\alpha}}{\sigma} \right)  \qquad \text{in } d=2
 \end{align}
-and so
-\begin{align}
-  \E[e^{-\ell \tau_x}] &= \frac{\sqrt{\ell+\alpha}}{\sqrt{\alpha}} e^{-x\sqrt{2}\left(\sqrt{\ell+\alpha}-\sqrt{\alpha}\right)} \left( 1 + \frac{1}{x\sqrt{2(\ell+\alpha)}} \right) \left( 1 + \frac{1}{x\sqrt{2(\alpha)}} \right)^{-1} \\
-  &= \frac{x\sqrt{2(\ell+\alpha)}+1}{x\sqrt{2\alpha}+1} e^{-x\sqrt{2}\left(\sqrt{\ell+\alpha}-\sqrt{\alpha}\right)} .
-\end{align}
+and we get $\E[e^{-\ell \tau_x}]$ by replacing $\alpha$ with $\alpha+\ell$.
 
-\item Differentiating the above, we get that $\E[\tau_x] = x^2 / (x\sqrt{2\alpha}+1) \sim x/\sqrt{2\alpha}$ and $\var[\tau_x] = x^3 / \sqrt{2\alpha}(x\sqrt{2\alpha}+1)^2 \sim x/(2\alpha)^{3/2}$; in fact, the same scaling produces the same limit law, since the $\ell/\sqrt{x}$ term dissappears.
-
+% WRONGLY BASED ON 1/t INSTEAD OF 1/\sqrt{t}:
+% and this is found in G\&R 8.468, $K_{3/2}(z) = \sqrt{\pi/2z} e^{-z} (1+1/z)$.
+% The formula corresponding to equation \eqref{eqn:first_integral} above is then
+% \begin{align}
+%   \E[e^{-\ell \tau_x}] \propto \sqrt{\pi} \frac{2\sqrt{\ell+\alpha}}{x^2} e^{-x\sqrt{2(\ell+\alpha)}} \left( 1 + \frac{1}{x\sqrt{2(\ell+\alpha)}} \right) ,
+% \end{align}
+% and so
+% \begin{align}
+%   \E[e^{-\ell \tau_x}] &= \frac{\sqrt{\ell+\alpha}}{\sqrt{\alpha}} e^{-x\sqrt{2}\left(\sqrt{\ell+\alpha}-\sqrt{\alpha}\right)} \left( 1 + \frac{1}{x\sqrt{2(\ell+\alpha)}} \right) \left( 1 + \frac{1}{x\sqrt{2(\alpha)}} \right)^{-1} \\
+%   &= \frac{x\sqrt{2(\ell+\alpha)}+1}{x\sqrt{2\alpha}+1} e^{-x\sqrt{2}\left(\sqrt{\ell+\alpha}-\sqrt{\alpha}\right)} .
+% \end{align}
+% 
+% \item Differentiating the above, we get that $\E[\tau_x] = x^2 / (x\sqrt{2\alpha}+1) \sim x/\sqrt{2\alpha}$ and $\var[\tau_x] = x^3 / \sqrt{2\alpha}(x\sqrt{2\alpha}+1)^2 \sim x/(2\alpha)^{3/2}$; in fact, the same scaling produces the same limit law, since the $\ell/\sqrt{x}$ term dissappears.
+% 
 \end{itemize}
@@
 
-A more back-of-the-envelope calculation of this would be:
+We can combine this with the formula \eqref{eqn:eqfreq} for the equilibrium frequency $q(r) \approx C \exp( -\sqrt{2 \alpha} r / \sigma)$
+to estimate $\lambda$, the effective outflux from the occupied patch: equating $\rho q(r)$ with $\lambda$ multiplied by equation \eqref{eqn:bp_occupation}, we get
+\begin{align} \label{eqn:outflux_from_patch}
+    \lambda = \text{(outflux from old patch)} = \begin{cases}
+        C \rho \sigma \sqrt{2\alpha} \quad & d=1 \\
+    C \rho \sigma^{3/2} (2\alpha)^{1/4} \sqrt{2 \pi x} \quad & d=2 
+    \end{cases}
+\end{align}
+Note that this differs from the value $C\rho$ which we might have guessed.
+
+\emph{The following is not making sense to me at the moment.} A more back-of-the-envelope calculation of this would be:
 we know that the probability a branching process arrives at the patch is given by $(1/\mu_K)\exp(-x\sqrt{2 \alpha}/\sigma)$;
@@
 
+From the main paper, we have that
+\begin{align} \label{eqn:bp_prob_estab}
+    \text{(prob of estab of a distant BP)} &\approx \frac{2s_b}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} x}{\sigma} \right) ,
+\end{align}
+and the subcritical occupation density per unit area at distance $r$ is
+\begin{align} \label{eqn:eqfreq}
+    q(r) \approx C \exp( -\sqrt{2 \alpha} r / \sigma).
+\end{align}
+
 We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
 \begin{align}
-  (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \\
-    \begin{split} &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
-  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split} \\
-  \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \\
+    (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \label{eqn:heuristic1a} \\
+    \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \label{eqn:heuristic1b} \\
     & \qquad \qquad \times (\mbox{prob of estab in the patch}) . \end{split} 
 \end{align}
-The probability of establishment of a branching process is in equation~\eqref{eqn:bp_prob_estab};
-the subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
+The subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
 the number of (extra) offspring per subcritical individual is $s_B+s_m$;
-and the probability of establishment in the patch is $2s_B/\xi^2$.
-The ``outflux of branching processes'' we have not computed;
-but we have computed the occupation density from this branching process approximation above,
-so equating ($\rho$ times $C\exp(-\sqrt{2 s_m}x/\sigma)$) and (outflux times equation~\eqref{eqn:bp_occupation})
-we have that the effective outflux (from the edge of the already occupied patch!) is 
-\begin{align}
-  (\mbox{outflux of BPs from old patch}) = C \rho \sigma \sqrt{2 \alpha} ,
-\end{align}
-rather than $C\rho$ which we might have guessed.
+and the probability of establishment of each such extra offspring in the patch is $2s_B/\xi^2$.
+The effective outflux of branching processes is computed in equation \eqref{eqn:outflux_from_patch}.
 The number of establishment events per BP is as follows, in averages: 
@@
 \begin{align}
-  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{\alpha} x}{\sigma}\right) .
+  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{2\alpha} x}{\sigma}\right) .
 \end{align}
-\textbf{Now} note that what we \textbf{actually} want to compute is not this rate, but rather 
+
+However, this formulation counts separately each of possible multiple establishment events by a single transiting family.
+Writing
+\begin{align} \label{eqn:heuristic2}
+  \begin{split} (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
+  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split},
+\end{align}
+for $\migrate$ we actually want to omit the last term.
+Using the probability of establishment of a distant branching process is in equation~\eqref{eqn:bp_prob_estab}, this is
 \begin{align}
   \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
-  &= C \rho \sqrt{\alpha} \times \frac{2 s_b}{\sqrt{\pi} \xi^2} \exp\left(- \frac{\sqrt{\alpha}x}{\sigma} \right) .
+           &= \begin{cases}
+    C \rho \sqrt{2 \alpha} \times \frac{2 s_b}{\sqrt{\pi} \xi^2} \exp\left(- \frac{\sqrt{2\alpha}x}{\sigma} \right)  \qquad & d=1 \\
+    C \rho \sqrt{2 \alpha} \times \frac{ 2 s_b \sqrt{2 \alpha A} }{ \xi^2 \pi \sigma^3} K_1\left(\frac{x\sqrt{2\alpha}}{\sigma}\right)  \qquad & d=2 .
+\end{cases}
 \end{align}
-Finally, combining these, we can also compute that
+Finally, combining \eqref{eqn:heuristic1b} and \eqref{eqn:heuristic2}, we can also compute that
 \begin{align}

### 2013-05-31 patchy-selection-paper.tex
WorkingWiki importation

@@
 \plr{same appendix with sims as above}
-\gc{Should we also use this approx. to get the approx. rate of establishing mutations
-arising outside of the patch, above. Feel like in the mutation rate
-part we could pull $s_B$ out front of the rate as well.} 
+\gc{Should we also use this approx. to get the approx. rate of establishing mutations arising outside of the patch, above.} 
+\plr{Difficulties with this... and I'm not sure how interesting that is?  Convince me otherwise?}
+\gc{Feel like in the mutation rate part we could pull $s_B$ out front of the rate as well.} 
 \plr{huh?}
 
-This allows us to get a good sense of the time scale of adaptation by migration between separated patches, as follows.
-First imagine that the new allele is not advantageous in the second patch,
-so that the expected number of alleles present in a small area $\Delta$ at distance $r$ from the already established patch
-is $\Delta \rho q(r)$.
-As these copies are disadvantageous, they form a subcritical branching process, which die out fairly soon.
-Now add to this local selective advantage by stipulating that any copy of the allele in the new patch
-has ``extra'' offspring that are marked in some way, on average $(1+s_B)-(1-s_m)=s_B+s_m$ additional offspring per generation.
-If the probability of establishment \emph{in the new patch} is $p(x)$ as above,
-then each of these new, marked copies appearing at location $x$ has probability $p(x)$ of establishing.
-(This general relation is stated formally in Appendix XXX.)
-Therefore, the total rate of establishment 
-(i.e.\ the mean number of successfully establishing migrant lineages per generation)
-is $\rho (s_B+s_m) \int q(|x|) p(x) dx$.
-Using a similar approximation as in section \ref{ss:patchymutation},
-we can approximate this, for a patch of area $A$ at distance $R$ from an already occupied patch, as
-\begin{align}
-  \migrate(R) = \rho A \frac{s_B}{\xi^2} (s_B+s_m) \exp\left( -\frac{ \sqrt{2 s_m} R}{\sigma} \right),
+As we show below in section \ref{ss:haplotype_length}, the rate of establishment, for a patch of area $A$ at distance $R$ from an already occupied patch, is approximately
+\begin{align} \label{eqn:migrate}
+    \migrate(R) = \rho \frac{s_B\sqrt{2s_m} }{\xi^2\sqrt{\pi}}  \exp\left( -\frac{ \sqrt{2 s_m} R}{\sigma} \right).
 \end{align}
-where we have set $C=1/2$ in equation \eqref{eqn:eqfreq}.
-Using different methods, we arrive at the same functional form below in section \ref{ss:haplotype_length}.
-(Note that if the size of the new patch is large relative to $\sqrt{s_m}/\sigma$,
-then the integral of $q(x)$ over the patch will not be well approximated by $A \times q(R)$,
-but the adjustment is easy to make.)
+The exponential term is expected based on \eqref{eqn:eqfreq} -- 
+since establishment requires rare long distant migrants to be present, and \eqref{eqn:eqfreq} gives the density of such --
+but the constant requires other arguments.
+
+% This allows us to get a good sense of the time scale of adaptation by migration between separated patches, as follows.
+% First imagine that the new allele is not advantageous in the second patch,
+% so that the expected number of alleles present in a small area $\Delta$ at distance $r$ from the already established patch
+% is $\Delta \rho q(r)$.
+% As these copies are disadvantageous, they form a subcritical branching process, which die out fairly soon.
+% Now add to this local selective advantage by stipulating that any copy of the allele in the new patch
+% has ``extra'' offspring that are marked in some way, on average $(1+s_B)-(1-s_m)=s_B+s_m$ additional offspring per generation.
+% If the probability of establishment \emph{in the new patch} is $p(x)$ as above,
+% then each of these new, marked copies appearing at location $x$ has probability $p(x)$ of establishing.
+% (This general relation is stated formally in Appendix XXX.)
+% Therefore, the total rate of establishment 
+% (i.e.\ the mean number of successfully establishing migrant lineages per generation)
+% is $\rho (s_B+s_m) \int q(|x|) p(x) dx$.
+% Using a similar approximation as in section \ref{ss:patchymutation},
+% we can approximate this, for a patch of area $A$ at distance $R$ from an already occupied patch, as
+% \begin{align}
+%   \migrate(R) = \rho A \frac{s_B}{\xi^2} (s_B+s_m) \exp\left( -\frac{ \sqrt{2 s_m} R}{\sigma} \right),
+% \end{align}
+% where we have set $C=1/2$ in equation \eqref{eqn:eqfreq}.
+% Using different methods, we arrive at the same functional form below in section \ref{ss:haplotype_length}.
+% (Note that if the size of the new patch is large relative to $\sqrt{s_m}/\sigma$,
+% then the integral of $q(x)$ over the patch will not be well approximated by $A \times q(R)$,
+% but the adjustment is easy to make.)
 
@@
 Setting the left-hand side to zero in \eqref{eqn:discrete_generator} gives us equations for the moments of the stationary distribution,
-for instance, if $\phi(x) = \lim_{t \to \infty} \E[ \Phi_t(x) ]
+for instance, if $\phi(x) = \lim_{t \to \infty} \E[ \Phi_t(x) ]$
 

### 2013-06-01 maize-ex.tex
WorkingWiki importation

@@
 \newcommand{\mutrate}{\lambda_{\mbox{mut}}}
+\newcommand{\migrate}{\lambda_{\mbox{mig}}}
 \begin{document}
@@
   \begin{center}
-  \begin{tabular}{|crl|}
+  \begin{tabular}{|cll|}
     \hline
@@
     maize population density             & $\rho$     & $5 \times 10^3$ \\
+    area of highland habitat             & $A$        & 500 km$^2$ \\
+    mean dispersal distance              & $\sigma$   & 1.8 km \\
     \hline 
@@
 
+\subsection{Dispersal distance}
+
+The dispersal distance
+-- the mean distance between parent and offspring --
+is the average of the pollen and seed mean dispersal distances.
+With the above assumptions, the pollen dispersal distance is zero,
+and the seed dispersal distance is the chance of inter-village movement
+multiplied by the mean distance moved.
+This is
+% pe <- .068; pm <- .02; rm <- .2; sigma.s <- 50
+% (1/2) * (pe + (1-pe)*pm*rm) * sigma.s
+\begin{align}
+  \sigma = \frac{1}{2} (p_e + (1-p_e) p_m r_m ) \sigma_s = 1.7932 \text{km}
+\end{align}
+at the parameter values above.
+
 \subsection{Results}
@@
 \begin{align*}
-  \mutrate = 2 \rho A s_b / \xi^2 .
+  \mutrate = 2 \rho A s_b \mu / \xi^2 .
 \end{align*}
+% rho <- 5000; sb <- 10^(-(1:3)); xisq <- 50; A <- 500
+% sapply( c(1e-8,1e-5), function (mu) 2 * rho * A * sb *mu / xisq )
+At the values above, with $.1 \le s_b \le .001$, the factor $2 \rho A s_b / \xi^2$ multiplying the mutation rate
+varies between $10^2$ and $10^5$,
+implying that a single-base mutation with $\mu=10^{-8}$ would have to wait between $10^4$ and $10^6$ generations to fix,
+but a mutation with a larger target, say $\mu=10^{-5}$, would fix in tens to thousands of generations, depending on the selection coefficient.
 
@@
 \section{Adaptation by migration}
+
+As we show in the theory paper, 
+the rate of adaptation by diffusive migration is roughly
+\begin{align*}
+  \migrate = \rho \frac{s_b \sqrt{2 s_m}}{2\xi^2} \exp\left(- \frac{\sqrt{2 s_m}R}{\sigma} \right) .
+\end{align*}
+We can talk about this more and give some simulations.
+But for now, let's interpret.
+
+First note that for $10^{-1} \le s_m \le 10^{-4}$, the value $1/\sqrt{2s_m}$ is between 2 and 70 --
+so the exponential decay of the chance of migration fallws off on a scale of between 2 and 70 times the dispersal distance.
+Above we have estimated the dispersal distance to be $\sigma \approx 2$ km,
+and far below the mean distance $\sigma_s$ to the field that a farmer replants seed from, when this happens,
+which we have as $\sigma_s = 50$ km.
+Taking $\sigma=2$ km, we have that $4 \le \sigma/\sqrt{2s_m} \le 150$ km.
+A very conservative upper bound might be $\sigma \le \sigma_s/20$ (if farmers replaced 10\% of their seed with long-distance seed every year).
+At this upper bound, we would have $5 \le \sigma/\sqrt{2s_m} \le 175$ km,
+which is not very different.
+This makes the exponential term very small since $R$ is on the order of 1,000 km.
+
+Taking $\sigma=2$ km, we then compute that 
+% rho <- 5000; sb <- .01; sm <- 10^(-(1:4)); xisq <- 50; sigma <- 2
+% sapply( 1000*(1:4), function (R) rho * sb * sqrt(2*sm) / (2 *  xisq) * exp(- sqrt(2*sm)*R/sigma ) )
+if $s_m = 10^{-4}$ (very weak selection in the lowlands), then for $R=1,000$ km, the migration rate is $\migrate \le 10^{-5}$,
+i.e.\ it would take on the order of 100,000 generations (years) to get a successful migrant only 1,000 km away,
+under this model of undirected, diffusive dispersal.
+For larger $s_m$, the migration rate is much smaller.
+
+%%%%%%%%%%%%
+\section{Conclusion}
+
+It seems unlikely that any alleles that are adaptive in the highlands and deleterious at all in the lowlands
+would have transited central America by undirected (diffusive) sharing of seed.
+The conclusions could change if we drastically underestimate the rate of very long distance sharing of seed,
+e.g.\ if sharing across hundreds of kilometers was common at some point.
+
+Both calculations are very pessimistic about the chance of shared single-base changes through either migration or independent mutation.
+However, independent mutations could be expected in kilobase-size targets,
+suggesting there might be signal for genes that share adaptive changes.
 

### 2013-06-01 patchy-selection-fns.R

@@ -51,7 +51,7 @@ migrationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
     # approximation for migrational influx per patch
     # XXX to-do: do numerics?
     # return( 2 * s*gb * A * rho * exp( - sqrt(abs(s*gm))*R/sigma ) )
-    return( 2 * s*gb * rho * exp( - sqrt(abs(s*gm))*R/sigma ) )
+    return( s*gb * sqrt(s*gm/pi) * rho * exp( - sqrt(abs(s*gm))*R/sigma ) )
 }
 
 

### 2013-06-01 patchy-selection-plots.R

@@ -253,12 +253,14 @@ muvals <- c(10^-8,10^-5)
 A <- 500 # km^2
 # distance between patches
 R <- 4000 # km
+# variance in reproductive number -- sb above is scaled by this.
+xisq <- 50
 
 ###
 # Table up those values
 params <- expand.grid(rho=rhovals, sigma=sigmavals, mu=muvals)
 exvalues <- t( sapply(1:dim(params)[1], function (k) {
-        with( params[k,], everything(mu=mu, rho=rho, sb=sb, sm=sm, sigma=sigma, R=R, A=A) )
+        with( params[k,], everything(mu=mu, rho=rho, sb=sb*xisq, sm=sm, sigma=sigma, R=R, A=A) )
         } ) )
 
 # write out the table in latex

### 2013-06-01 sim-patchy-selection-fns.R

@@ -179,10 +179,10 @@ colorRampTrans <- function (basecol, n=12) {
 }
 
 
-plotpophist <- function(pophist,...) {
+plotpophist <- function(pophist,timeslice=TRUE,...) {
     # Plot one-dimensional population history with time on the x-axis
     if (!any(dim(pophist$pophist)[1:2]==1)) { stop("Population is not one-dimensional.") } 
-    zimg <- plotpop( aperm(pophist$pophist[,,,,drop=TRUE],c(3,1,2)), params=pophist$pop$params, xlab="time", ylab="space" )
+    zimg <- plotpop( aperm(pophist$pophist[,,,timeslice,drop=TRUE],c(3,1,2)), params=pophist$pop$params, xlab="time", ylab="space", ... )
     axis(1); axis(2)
     return( invisible(zimg) )
 }
@@ -231,21 +231,16 @@ plotpop <- function (n, params, maxtimes=200, thresh=.05, cols, coltrans, ...) {
     # subsample if too many columns
     if ((dim(n)[1]>maxtimes)) { n <- n[(1:(dim(n)[1]))%%floor(dim(n)[1]/maxtimes)==0,,] }
     if ((dim(n)[2]>maxtimes)) { n <- n[,(1:(dim(n)[2]))%%floor(dim(n)[2]/maxtimes)==0,] }
-
-    if (missing(cols)) {
-        cols <- c(list(c("#FFFFFF")), lapply(rainbow(dim(n)[3]), colorRampTrans, n=12) )
-    }
-    if (missing(coltrans)) {
-        coltrans <- function (x,cols=cols,ncols=length(cols)) { cols[ceiling((ncols-1)*x/params$N+0.5)] }
-    }
+    # color picking
+    if (missing(cols)) { cols <- c(list(c("#FFFFFF")), lapply(rainbow(dim(n)[3]), colorRampTrans, n=12) ) }
+    if (missing(coltrans)) { coltrans <- function (x,colors,ncols=length(colors)) { colors[ceiling((ncols-1)*x/params$N+1)] } }
 
     # Bulk plotting
     img <- array(1,dim=c(3,dim(n)[1:2]))  # has r,g,b layers above population
     for (k in 2:dim(n)[3]) {
         # image(1-log(1+n[,,k])/log(N),zlim=c(0,1), col=cols[[k]], new=k>1)
         # image(n[,,k]/N,zlim=c(0,1), col=cols[[k]], new=k>1)
[…]

### 2013-06-01 sim-patchy-selection.R
WorkingWiki importation

@@ -9,10 +9,31 @@ rm(pophist); rm(lins)
 run.name <- sample(run.list,1)
 run.id <- gsub( "([^-]*)-.*","\\1",run.name )
 load(run.name)
-print(pophist$pop$params$range)
+print(c("run",run.id,"range",pophist$pop$params$range))
 
 plotpophist(pophist)
 
+timeslice <- 1000+1:1000
+plotpophist(pophist, timeslice=timeslice, maxtimes=length(timeslice))
+for (k in 1:dim(lins$lineagelocs)[3]) {
+    ltmp <- t(lins$lineagelocs[,timeslice,k])
+    usethese <- ( !is.na(ltmp[,2]) &  pophist$pop$params$patchdist[ltmp] > 15 )
+    if (any(usethese)) { lines( seq_along(timeslice)[usethese], subset(ltmp,usethese)[,2], col=rainbow(60)[k] ) }
+}
+points( lins$T - min(timeslice), lins$linit[,2], pch=20, cex=.25, col=adjustcolor("black",.25) )
+
+pdf(file="verylong.pdf",width=100,height=8,pointsize=10)
+timeslice <- 1:dim(lins$lineagelocs)[2]
+plotpophist(pophist, timeslice=timeslice, maxtimes=length(timeslice))
+for (k in 1:dim(lins$lineagelocs)[3]) {
+    ltmp <- t(lins$lineagelocs[,timeslice,k])
+    usethese <- ( !is.na(ltmp[,2]) &  pophist$pop$params$patchdist[ltmp] > 15 )
+    if (any(usethese)) { lines( seq_along(timeslice)[usethese], subset(ltmp,usethese)[,2], col=rainbow(60)[k] ) }
+}
+points( lins$T - min(timeslice), lins$linit[,2], pch=20, cex=.25, col=adjustcolor("black",.25) )
+dev.off()
+
+# 1-D
 animpophist(pophist)
 
 # 2-D
@@ -21,9 +42,9 @@ with(pophist$pop, contour( x=1:nrow(params$patchdist), y=1:ncol(params$patchdist
 plotlins(lins,add=TRUE,subset=1:100)
 
 # Exponential decay in mean occupation density
-plot( pophist$pop$params$patchdist, rowMeans(pophist$pophist[,,2,200:dim(pophist$pophist)[4],drop=FALSE],dims=2)/pophist$pop$params$N, log='y' )
[…]

### 2013-06-01 generate-patchy-run.R
WorkingWiki importation

@@ -3,6 +3,11 @@
 source("sim-patchy-selection-fns.R")
 source("lineages.R")
 
+nsteps <- 10000
+stepsize <- 1
+run.id <- floor(runif(1)*10000)
+nlins <- 500
+
 # Parameters
 params <- list(
         mu = 0,           # mutation rate
@@ -30,18 +35,18 @@ if (min(params$range)>1) { # 2D
             lapply( 1:5, function (n) { c( 2^(-(n+2)), 0, n ) } ), 
             lapply( 1:5, function (n) { c( 2^(-(n+2)), 0, -n ) } ) )
 } else if (params$range[1]==1) {
-    params$migrsteps = c( lapply( 1:5, function (n) { c( 2^(-(n+2)), n, 0 ) } ),  ## 1D
-            lapply( 1:5, function (n) { c( 2^(-(n+2)), -n, 0 ) } ) )
+    params$migrsteps = c( lapply( 1:5, function (n) { c( 2^(-(n+2)), n, 0 ) } ),  ## 1D, y
+                          lapply( 1:5, function (n) { c( 2^(-(n+2)), -n, 0 ) } ) )
 } else {
-    params$migrsteps = c( lapply( 1:5, function (n) { c( 2^(-(n+2)), 0, n ) } ),  ## 1D
-            lapply( 1:5, function (n) { c( 2^(-(n+2)), 0, -n ) } ) )
+    params$migrsteps = c( lapply( 1:5, function (n) { c( 2^(-(n+2)), 0, n ) } ),  ## 1D, x
+                          lapply( 1:5, function (n) { c( 2^(-(n+2)), 0, -n ) } ) )
 }
 params$migrsteps <- { msum <- sum( sapply(params$migrsteps,"[",1) ); lapply( params$migrsteps, function (x) { c(x[1]/msum,x[-1]) } ) }
 # define patch
 params$s <- with(params, matrix( sm, nrow=range[1], ncol=range[2] ) )
 params$s[ (row(params$s)-params$range[1]/2)^2 + (col(params$s)-params$range[2]/2)^2 <= (params$patchsize/2)^2 ] <- params$sb  # a central circle
 params$sigma <- getsigma(params)
-# compute distance to patch
+# compute distance to center of patch
 params$patchdist <- sqrt( (row(params$s)-params$range[1]/2)^2 + (col(params$s)-params$range[2]/2)^2 ) 
 
[…]

### 2013-06-01 lineages.R

@@ -68,7 +68,26 @@ lineages <- function ( pophist, nlin, migrsteps, T=dim(pophist)[4], m, linit=NUL
             }
         }
     }
-    return( list( linit=linit, lineagelocs=LL, localsizes=LS, coalevents=coalevents ) )
+    return( list( linit=linit, T=T, lineagelocs=LL, localsizes=LS, coalevents=coalevents ) )
+}
+
+getfams <- function (eps=1, lins, params, LL=lins$lineagelocs, coalevents=lins$coalevents) {
+    # pull out all lineages coalescing before hitting the patch
+    coal <- subset(coalevents,params$patchdist[as.matrix(coalevents[,c("x","y")])]>params$patchsize+eps*params$sigma)
+    fams <- lapply( sort(setdiff(coal$coalto,coal$from)), function (k) {
+            thisLL <- t(LL[,,k]) 
+            exittime <- max( c( min(which(!is.na(thisLL[,1]))), which( params$patchdist[thisLL]<=params$patchsize ) ) ) # last exit from patch
+            coaltime <- min( coal$t[coal$coalto==k] )
+            usetimes <- min(coaltime,exittime):max(which(!is.na(thisLL[,1])))
+            return( data.frame( k=k, t=usetimes, thisLL[ usetimes, ] ) )
+        } )
+    names(fams) <- sort(setdiff(coal$coalto,coal$from))
+    for (j in 1:nrow(coal)) {
+        fams[[paste(coal$coalto[j])]] <- rbind( fams[[paste(coal$coalto[j])]], 
+                data.frame( k=coal$orig[j], t=which(!is.na(LL[1,,coal$orig[j]])), na.omit(t(LL[,,coal$orig[j]])) )
+            )
+    }
+    return(fams)
 }
 
 plotlins <- function( lineages, coalevents, linit, cols, jit=TRUE, add=FALSE, subset=TRUE, ... ) {

### 2013-07-23 branching-transit.tex
WorkingWiki importation

@@
 
-\item This induces a distribution on $T$, the time that a successful excursion took to cross:
+\item This induces a distribution on $\tau_x$, the time that a successful excursion took to cross:
 assuming, as is asymptotically correct, that the number of arriving lineages is independent of how long it took to get there,
@@
 
-\item The length of the hitchhiking segment on \emph{one side} of the selected site $L$ is exponential with rate depending on $T$;
-integrating over $T$ we get
+\item The length of the hitchhiking segment on \emph{one side} of the selected site $L$ is exponential with rate depending on $\tau_x$;
+integrating over $\tau_x$ we get
 (using Gradshteyn \& Ryzhik 3.471.9; note $\ell > 0 > \log(1-s_d)$):
@@
 \begin{align}
-  \E[Z_t(dx)] = e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) ,
+  \E[Z_t(dx)] = e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) dx ,
 \end{align}
@@
 Note that this differs from the value $C\rho$ which we might have guessed.
+NOT RIGHT in $d=2$: SHOULD NOT DEPEND ON $x$.
 
@@
 The subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
-the number of (extra) offspring per subcritical individual is $s_B+s_m$;
-and the probability of establishment of each such extra offspring in the patch is $2s_B/\xi^2$.
+the number of (extra) offspring per subcritical individual is $s_p+s_m$;
+and the probability of establishment of each such extra offspring in the patch is $2s_p/\xi^2$.
 The effective outflux of branching processes is computed in equation \eqref{eqn:outflux_from_patch}.
@@
 each produce $1/s_m$ total core offspring;
-each of these produce $s_B+s_m$ extra offspring which each have chance $2s_B/\xi^2$ of establishing,
+each of these produce $s_p+s_m$ extra offspring which each have chance $2s_p/\xi^2$ of establishing,
 i.e.
 \begin{align}
-  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{2\alpha} x}{\sigma}\right) .
+  (\mbox{\# estab events per BP}) = \frac{(s_p+s_m) 2 s_p }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{2\alpha} x}{\sigma}\right) .
 \end{align}
@@
 \begin{align}
-  (\mbox{\# estab events per successful BP}) = \frac{s_B+s_m}{ \sqrt{s_m} } .
+  (\mbox{\# estab events per successful BP}) = \frac{s_p+s_m}{ \sqrt{s_m} } .
 \end{align}

### 2013-07-23 patchy-selection-paper.tex
WorkingWiki replaced project description

--- 2013-05-31 patchy-selection-paper.tex
@@
 \DeclareMathOperator{\Exp}{Exp}
+\DeclareMathOperator{\sn}{sn}
 \renewcommand{\P}{\mathbb{P}}
@@
 changes in the same pathway, genes, 
-or even by precisely the same genetic changes. Such precise
-convergence points to the conservation of function of the genes
+or even by precisely the same genetic changes \citep{zhen2012parallel}. 
+Such precise convergence points to the conservation of function of the genes
 underlying adaptive traits, 
 sometimes over deep time scales \citep{deephomologypapers}. 
-Convergence on a molecular level also suggests that the path taken by adaptation can sometimes be relatively constained as few changes could have resulted in the adaptive phenotype, or that relatively few changes that can confer the advantagous phenotype of interest were sufficiently free of deleterious pleiotropic consequences to contribute to adaptation.
+Convergence on a molecular level also suggests that the path 
+accessible to adaptation is sometimes relatively constrained, 
+as would be the case if only a few changes confer the advantagous phenotype 
+and are sufficiently free of deleterious pleiotropic consequences.
 
@@
 
-Convergent adaptation can also occur within species, with individuals within a species 
-adapted to the same environment in parallel through different large effect changes. 
+Convergent adaptation can also occur within species, if individuals within a species 
+adapt to the same environment, in parallel, through different genotypic changes. 
 There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
-All else being equal the evolution of convergent phenotypes is more
-likely with a 
-higher mutational input, i.e.\ when the mutational target and population sizes are larger \citep{}. 
+All else being equal, such evolution of convergent phenotypes is more likely with a 
+higher mutational input, that is, when the mutational target and population sizes are larger \citep{}. 
+The geographic distribution of populations can also affect the probability of parallel mutation within a species:
+a widespread species is more likely to adapt by multiple, parallel mutations if dispersal is geographically limited, 
+since subpopulations will adapt via new mutations before the adaptive allele arrives via migration \citep{ralph2010parallel}. 
 Standing variation for a trait can also increase the probability of convergence, 
-as multiple pre-existing alleles can spread from low frequency, 
-if the newly beneficial alleles were previously not too deleterious \citep{Orr,Hermission}. 
-The geographic distribution of populations can also shape the probability of parallel mutation within a species.
-A widespread species is more likely to adapt by multiple mutations in parallel if dispersal 
-is geographically limited, as subpopulations will adapt via new mutations rather than waiting for migration \citep{RalphCoop}.
-
-%When multiple rare changes sweep up from low frequency in parallel within a population this has been termed a soft sweep.\\
+and standing variation for an allele increases the probability that the selective sweep will be \emph{soft}
+(beginning from a base of multiple copies),
+leading to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}.
 
@@
 before adaptations spread between the populations by migration
-will be enhanced if these adaptive alleles are maladapted in intervening environments,
-since such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{slatkin}.
-
-For example, a variety of plant species have repeatedly adapted to patches of soil with high concentrations of heavy metals
-(e.g.\ serpentine outcrops and mine tailings) \citep{turnerarabidopsis,mimulus};
-the alleles conferring heavy metal tolerance are thought to be disfavored off of these patches. 
-Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration,
-to particular substrates, e.g.\ dark rock outcrops or white sand dunes \citep{benson}.
+will be larger if these adaptive alleles are maladapted in intervening environments,
+since such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{slatkin1973geneflow,}.
+
+One elegant set of such examples is provided by the assortment of plant species 
+that have repeatedly adapted to patches of soil with high concentrations of heavy metals
+(e.g.\ serpentine outcrops and mine tailings) \citep{turner2010serpentine,mimulus};
+the alleles conferring heavy metal tolerance are thought to be locally restricted 
+because they incur a cost off of these patches. 
+Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration
+to particular substrates, e.g.\ dark rock outcrops or white sand dunes \citep{benson1933concealing}.
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
-which is primarily dark on a number of black lava flows separated by much lighter soil \citep{dice}.
-Strong selection appears to favour such crypsis \citep{owls},
-so perhaps as a result of this strong selection against migrants, 
-at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{nachman}. 
-Similar situations have been demonstrated in other mouse systems \citep{beachmice,peromyscus}.
+which is primarily dark on a number of black lava flows separated by much lighter soil \citep{dice1940ecologic}.
+Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
+and, perhaps as a result of this strong selection against migrants, 
+at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{nachman2003different}. 
+Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}.
+
+In this paper, we study this situation, namely,
+when a set of alleles provide an adaptive benefit in geographically localized patches, 
+separated by (inhabited) regions where the alleles are deleterious.
+\plr{or neutral?}
+The main questions are:
+Under what conditions is it likely that each patch adapts in parallel, through new mutation,
+and when is it likely that migration carries these alleles between patches?
+How could we tell, after the fact, which has occurred?
+We work in a model of continuous geography,
+using a variety of known results and new methods.
+In section XX we \dots QUICK OUTLINE
 
 %http://www.nature.com/hdy/journal/v94/n2/full/6800600a.html
-
-%In \citep{RalphCoop} we formulated a simple model of parallel adaptation in a geographic setting
-%where a continuous population was subjected to a selection pressure
-%that was constant across the entire environment.
 
@@
 %%%
-\subsection{Introduction to a patchy model}
+\subsection{Model of a patchy landscape}
 \label{ss:patchyspace}
 
-Consider a population spread uniformly across a landscape. Within this environment there are patches of
-habitat that individuals are initially poorly adapted to, surrounded by large areas that the population is well adapted to.
-
-Suppose it takes only a single mutational change to create an allele ($B$) that adapts an individual to the
-poor habitat type. This mutational change occurs at a low rate of
-$\mu$ per chromosome per generation. We assume that this initially rare new
-allele has fitness $1+s_B$ relative to the unmutated type ($b$) in the ``poor'' habitat patches,
-and assume that it has fitness $1-s_m$ when rare in the intervening areas, with
-$s_m>0$. To simplify matters we assume that the disadvantage $s_m$ 
+Consider a population spread across a landscaper, 
+within which are patches of habitat to which individuals are (initially) poorly adapted,
+surrounded by large areas to which the population is well adapted.
+(When we refer to ``patches'' it is to these pieces of habitat.)
+Suppose it takes only a single mutational change to create an allele ($B$) that adapts an individual to the poor habitat type, 
+a change that occurs at a (low) rate of $\mu$ per chromosome per generation. 
+We assume that this initially rare new
+allele has fitness $1+s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches,
+and assume that it has fitness $1-s_m$ when rare in the intervening areas, with $s_p, s_m > 0$.
+To simplify matters, we assume that the disadvantage $s_m$ 
 is sufficiently large that, on the relevant timescale,
 the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
-(In particular, we require $s_m\neq 0$.)
-\plr{NOTE: changed over to $s_m>0$.}
+(The case where $s_m=0$ requires a different approach, which we do not treat here.)
 
 We are interested in the establishment of mutations in the ``poor'' patches by either
-migration or mutation, and so we are only interested in whether the allele
-can escape initial loss by drift when rare. Thus we do not have to
-specify the the fitness of the homozygote, only that its
-fitness is such the dynamics of the  allele when it is rare are
-determined by the fitness of the heterozygote. 
+migration or mutation, and so are mainly interested in whether the allele
+can escape initial loss by drift when rare. 
+Therefore, we do not have to specify the the fitness of the homozygote; 
+only that the dynamics of the allele when rare 
+are determined by the fitness of the heterozygote. 
 More general dominance will only make small corrections,
-with the exception of the recessive case, which we will omit.
-Mostly, we follow the literature in treating the diploid model as essentially haploid.
-% This allows us to treat a diploid model as being essentially haploid for our purposes. \gc{TRUE?}  \plr{would change e.g. freqs at migration-selection balance}
-
-For the sake of simplicity we will assume a constant haploid population density $\rho$, 
-across both types of habitat (we will return to discuss this equal
-density point later). We further assume that the
-variance in offspring number is $\xi^2$, 
-and that the mean squared distance between parent and child is $\sigma^2$ (the dispersal variance).
+with the exception of the recessive case, which we omit.
+For the most part, 
+we follow the literature in treating the diploid model as essentially haploid.
+
+For the sake of simplicity, we assume a constant haploid population density $\rho$, 
+across both types of habitat (we return to discuss this equal density point later). 
+We further assume that the variance in offspring number is $\xi^2$, 
+and that the mean squared distance between parent and child is $\sigma^2$
+(i.e.\ $\sigma$ is the dispersal distance).
+We also aim to describe the scenario where it is sensible to think of large, discrete patches;
+the case of a fine-scale hetergeneous environment is treated \citep{elsewhere}.
 
 \paragraph{Previous work} 
-We will make use of a number previous results from the literature. 
-\citet{slatkin1973geneflow}, working in one dimensional deterministic framework, 
-showed that if the physical width of the patch is less than $2 \tan^{-1} (\sqrt{s_m/s_B})$, 
-then there is not a stable equilibrium with both $b$ and $B$ present,
-so that $B$ cannot establish within the patch due to migrational swamping \citep[see also][ for a review]{Lenormand}.
-\citet{barton1987establishment}, again working in one dimension,
-showed that this critical patch size also held in a stochastic framework \citep[see also the work of][]{Polk}. 
-For patches above this critical size, \citet{barton1987establishment}
-found that the probability of establishment of a new mutant that appears at distance $x$ from the patch
-decays exponentially with distance from the patch, with the scale given by $\sigma/\sqrt{s_m}$.
-% mutations appearing further than a few multiples of $\sigma/\sqrt{2s_m}$ away from such a patch 
-% have vanishingly small probability of becoming established within the patch. 
-\citet{barton1987establishment} also showed that mutations appearing within the patch have probability of establishment
-less than the probability for a panimictic population 
-\citep[which we denote $p_e$, and often approximate by $2 s_B / \xi^2$][]{haldane,fisher};
-and that the probability approaches $p_e$ as the size of the patch increases.
+We will make use of several previous results from the literature. 
+\citet{slatkin1973geneflow}, working in a one dimensional, deterministic framework, 
+showed that if the physical width of the patch is less than $2 \tan^{-1} (\sqrt{s_m/s_p})$, 
+then there is no stable equilibrium with both $b$ and $B$ present,
+so that migrational swamping prevents $B$ from establishing \citep[see also][ for a review]{lenormand2002limits}.
+\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed (again in one dimension) 
+that this critical patch size also holds in a stochastic framework.
+\citet{barton1987establishment} also showed that
+for patches above this critical size,
+the probability of establishment of a new mutant that appears at distance $r$ from the patch
+decays exponentially with distance from the patch, with the scale given by $\sigma/\sqrt{s_m}$,
+and that mutations appearing within the patch have probability of establishment
+less than the probability for a panimictic population,
+but approaching this as the size of the patch increases.
+(This latter panmictic probability of establishment we denote $p_e$,
+and often approximate by $2 s_p / \xi^2$ \citep{haldane,fisher}.)
 This result holds quite generally for a geographically spread population that experiences a uniform selection
-pressure \citep{Maruyama,cherry}. 
-\plr{wait, what result? that fixation is hindered by geographic structure?}
-
-%Maruyama: On the fixation probability of mutant genes in a subdivided population*
-%should we cite Cherry http://www.ncbi.nlm.nih.gov/pmc/articles/PMC1462507/
-
-%To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
-%and that the disadvantage $s_m$ ensures that on the relevant timescale,
-%the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
-
-%equal deme sizes?  general migration rates?
+pressure \citep{maruyama1970fixation,cherry2003diffusion}.
 
 %%%%%
-\subsection{Time to the establishment of a locally adaptive allele
-  due to mutational influx}
+\subsection{Establishment of a locally adaptive allele due to mutational influx}
 \label{ss:patchymutation}
@@
 Consider first a single, isolated poor habitat patch of area $A$ in which no $B$ allele has yet become established. 
+We first compute the time scale on which new $B$ mutations appear and fix in such a patch.
 As we are interested in patches where local adaptation can occur,
 we will assume that our patch is larger than the cutoff for local establishment, 
-i.e.\ wider than $2 \tan^{-1} (\sqrt{s_m/s_B})$.
+i.e.\ wider than $2 \tan^{-1} (\sqrt{s_m/s_p})$ \citep{slatkin1973geneflow}.
 
@@
 relative to the patch fixes within the poor habitat patch.
-The function $p(x)$ can be found in one dimension in terms of Jacobi elliptic functions,
-XXX as shown in the Appendix, but the expressions are not particularly helpful.
-The total successful mutational influx per generation is then given by the integral of $p(x)$ over the entire species range $\int \rho \mu p(x) dx$,
-and depends in a complicated way on the relationship of patch width and selection coefficients,
+As shown in Appendix \ref{apx:elliptic_integrals},
+the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
+but the precise expressions are not particularly helpful.
+The total successful mutational influx per generation is then given by the integral of $p(x)$ over the entire species range,
+% $\int \rho \mu p(x) dx$,
+and hence depends in a complicated way on the relationship of patch width and selection coefficients,
 but still scales linearly with the mutational influx density $\rho \mu$.
@@
 then a reasonable approximation is to assume that $p(x) = p_e$ within the patch, and $p(x) = 0$ otherwise,
-for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_B A / \xi^2$. 
-This holds for large patches, in which the number of mutations arising within the patch is larger 
-than those arising sufficiently nearby;
-we examine this more generally via simulation in Appendix XXX.
-\plr{In Appendix: we simulated fixation prob in a patch with gradual transition.  We have convincing figures.}
+for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_p A / \xi^2$. 
+This holds for large patches, in which the rate at which mutations arise within the patch is larger 
+than the rate at which they appear nearby;
+we examine this more generally via simulation in Appendix \ref{apx:establishment_sims}.
 
@@
 \begin{align} \label{eqn:mutrate}
-  \mutrate = \rho \mu \int p(x) dx  \approx 2 s_B \rho A \mu / \xi^2.
-\end{align}
-If this rate is low,  then the number of generations until a mutation arises that
-will become locally established within the patch is exponentially distributed with rate $\mutrate$.  
+  \mutrate = \rho \mu \int p(x) dx  \approx 2 s_p \rho A \mu / \xi^2.
+\end{align}
+If this rate is low, then the time (in generations) until a mutation arises that
+will become locally established within the patch is exponentially distributed with mean $1/\mutrate$.  
 Assuming that once a mutation becomes established it quickly reaches its equilibrium frequency across the patch, 
-the expected time it takes for our patch to become colonized by the $B$ allele is approximately $1/\mutrate$.
+the time scale on which new patches become colonized by the $B$ allele from new mutation is therefore approximately $1/\mutrate$.
+
+%%%%%%%
+\subsection{Establishment of a locally adaptive allele due to migrational influx}
+\label{ss:patchymigration}
+
+Now suppose that there are two patches, each of area $A$ and separated by distance $r$. 
+If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
+we would like to know the time scale on which copies of $B$ move between patches by migration
+(so, ignoring the possibility that a $B$ allele arises independently by mutation in the first patch).
+To determine this, we dissect the process by which an allele transits between the patches,
+in the process obtaining other useful information about the genealogy.
+
+Migration--selection balance ensures that 
+there will be some number of $B$ alleles present in the regions where they are disadvantageous,
+but only very few, and rarely, far away from the patch where $B$ is established.
+Denote the expected frequency of allele $B$ at distance $r$ from the patch by $q(r)$.
+Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(r)$ is the solution,
+which can then be solved to give $q(r)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
+Asymptotics of these implies that for large $r$ the mean occupation frequency decays exponentially, 
+\begin{align} \label{eqn:eqfreq}
+    q(r) \approx C \exp( -\sqrt{2 s_m} r / \sigma) \qquad \text{for large $r$},
+\end{align}
+where $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
+In Appendix \ref{apx:patchy_sims} we argue that $C=1/2$ is about right for most parameter values,
+and also that these asymptotics are correct in both one and two dimensions
+(the differential equation is only analytically solvable in one dimension).
+
+This expected frequency $q(r)$ is the \emph{time-averaged} occupation frequency,
+or in other words, the total number of $B$ alleles found at distance $r$ across $T$ generations, divided by $T$.
+Therefore, if we look at a small area of size $\epsilon$, 
+then $q(r) \epsilon$ is equal to the probability of finding any $B$ alleles there,
+multiplied by the mean number of such, if there are any.
+This points to an important fact: 
+close to the original patch, the ``equilibrium frequency'' $q(r)$ describes well the density of the $B$ allele at most times,
+but far away from the patch, there are rare excursions of a family of $B$ alleles, interspersed by long periods of absence.
+To compute the rate of establishment by such migration,
+we use properties of subcritical branching processes and Brownian motion,
+combined with \eqref{eqn:eqfreq},
+to obtain an approximation that we argue is valid if the distance $r$ is large relative to $\sigma/\sqrt{s_p}$.
+
+The arguments depends on the following decomposition of the genealogy of $B$ alleles.
+First, fix some intermediate distance $r_0$ from the established patch,
+and classify all $B$ alleles present at any time outside $r_0$ into families
+if they share a common ancestor living outside $r_0$.
+If $r_0$ is large enough that there are not too many families
+and interactions between family members are weak,
+then these families can be approximated by a collection of independent spatial branching processes.
+(This can be made formal in the limit of large population density.)
+We can extend these families by 
+including any individual outside the original patch (but perhaps inside $r_0$) 
+that has a parent outside $r_0$,
+so that these branching processes are rooted at $r_0$, and are killed upon reaching the original patch.
+This decomposition is illustrated in figure \ref{fig:sim_excursions}.
+In terms of these families, the equilibrium frequency above is then
+\begin{align}
+    \label{eqn:gestalt_q}
+    \begin{split}
+        q(r) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $r$ ) } ,
+\end{split}
+\end{align}
+and the quantity we wish to compute,
+the effective rate of at which \emph{families of} migrant $B$ alleles establish in a new patch at distance $r$, 
+is
+\begin{align}
+    \label{eqn:gestalt_migrate}
+    \begin{split}
+        \migrate(r) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in patch at $r$ ) }
+    \end{split}
+\end{align}
+All terms except the ``outflux of families'' are simple calculations with spatial branching processes;
+and we can solve for this outflux using the established form for \eqref{eqn:gestalt_q}.
+Here $\migrate$ reflects the mean number of establishing migrant families per unit time,
+rather than individuals,
+because family establishment events may include more than one individual,
+so individual establishment events are clustered in time,
+and using their overall density to obtain a typical waiting time until the next event
+will underestimate that time.
 
 %%%%
-\subsection{Time to the establishment of a locally adaptive allele due to migrational influx}
-\label{ss:patchymigration}
-Now suppose that there are two patches, each of area $A$, of the poor
-habitat type separated by distance $R$. 
-We will consider the case where the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
-for the purpose of comparison ignoring the possibility that a $B$ allele arises independently by mutation in the first patch.
-As above, any copy of the new allele appears at location $x$ relative to the second, unadapted patch 
-has probability approximately $p(x)$ of fixing locally in this patch,
-whether the allele is a new mutation or is descended from the occupants of the first patch.
-
-Due to migration--selection balance, 
-there will be a small number of $B$ alleles descended from those of the first patch
-present in the regions where they are disadvantageous;
-the expected frequency at distance $r$ may again be expressed in terms of Jacobi elliptic functions \citep[Appendix XXX][]{barton},
-although the equations are again not particularly helpful.
-This expected frequency is the \emph{time-averaged} occupation frequency,
-i.e.\ the average frequency of the allele at each location over a long time period,
-below we study the fluctuations about this mean.
-More helpful than the exact expressions,
-we know that if $q(r)$ is the expected proportion present at migration--selection balance at distance $r$ from the first patch
-in the absence of the second patch, 
-then for large $r$ this proportion decays exponentially, 
-\begin{align} \label{eqn:eqfreq}
-  q(r) \approx C \exp( -\sqrt{2 s_m} r / \sigma),
-\end{align}
-where $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
-(Probably $C=1/2$ is close enough for applications; see Appendix XXX.)
-\plr{same appendix with sims as above}
-\gc{Should we also use this approx. to get the approx. rate of establishing mutations
-arising outside of the patch, above. Feel like in the mutation rate
-part we could pull $s_B$ out front of the rate as well.} 
-\plr{huh?}
-
-This allows us to get a good sense of the time scale of adaptation by migration between separated patches, as follows.
-First imagine that the new allele is not advantageous in the second patch,
-so that the expected number of alleles present in a small area $\Delta$ at distance $r$ from the already established patch
-is $\Delta \rho q(r)$.
-As these copies are disadvantageous, they form a subcritical branching process, which die out fairly soon.
-Now add to this local selective advantage by stipulating that any copy of the allele in the new patch
-has ``extra'' offspring that are marked in some way, on average $(1+s_B)-(1-s_m)=s_B+s_m$ additional offspring per generation.
-If the probability of establishment \emph{in the new patch} is $p(x)$ as above,
-then each of these new, marked copies appearing at location $x$ has probability $p(x)$ of establishing.
-(This general relation is stated formally in Appendix XXX.)
-Therefore, the total rate of establishment 
-(i.e.\ the mean number of successfully establishing migrant lineages per generation)
-is $\rho (s_B+s_m) \int q(|x|) p(x) dx$.
-Using a similar approximation as in section \ref{ss:patchymutation},
-we can approximate this, for a patch of area $A$ at distance $R$ from an already occupied patch, as
-\begin{align}
-  \migrate(R) = \rho A \frac{s_B}{\xi^2} (s_B+s_m) \exp\left( -\frac{ \sqrt{2 s_m} R}{\sigma} \right),
-\end{align}
-where we have set $C=1/2$ in equation \eqref{eqn:eqfreq}.
-Using different methods, we arrive at the same functional form below in section \ref{ss:haplotype_length}.
-(Note that if the size of the new patch is large relative to $\sqrt{s_m}/\sigma$,
-then the integral of $q(x)$ over the patch will not be well approximated by $A \times q(R)$,
-but the adjustment is easy to make.)
-
-If this establishment rate is low, the time it takes
-a mutation to arise that will become established 
-will be approximately exponentially distributed with rate $\migrate$.
-Assuming that the time it takes for a allele from the first patch to
-become established in the second patch is short,
-then  the expected waiting time until some migrant allele fixes in the
-second patch will be $1/\migrate$.
-
-%%%
-\subsection{The probability of parallel adaptation between patches.} 
-\label{ss:probparallel}
-
-We now turn to the question of whether the separated populations adapt by parallel genetic changes 
-or by the spread of a migrant allele between the patches. 
-% These two scenerios are assumed to be distinguishable, 
-% either because in the case of parallelism the mutations occur at different positions 
-% or if the mutations occur at the same position they can be distinguished due to their haplotypic background. 
-% Assuming that the geographically spread population initally contains no copies of the $B$ allele, 
-% one or other of the patches of different habitat 
-% will be the first to adapt via the establishment of a $B$ allele;
-% so the question is then the relative timescale of new mutation and migration to the other patch.
-As only a single mutation is required to adapt individuals to the
-poor habitat patch, subsequent mutations that arise after an allele becomes established on the patch gain no selective benefit. 
-Similarly, an allele introduced into a patch by migration will not be favored by selection to spread, 
-if the patch has already been colonized. 
-Therefore, mutations selectively exclude each other from patches, over short time scales, 
-and will only slowly mix together over longer time scales by drift.
-
-We will assume that once a $B$ allele is introduced (by migration or mutation), 
-it becomes established in the poor habitat patch rapidly. 
-Under this approximation, and our selective exclusion assumption, 
-after one of the patches becomes colonized by a single $B$ allele, 
-the other patch will become colonized by migration or mutation, but not by both. 
-As such, the question of whether the
-population adapts to the second poor habitat patch in parallel, 
-simply comes down to whether the a new mutation or a migrant allele is the first to become established in the second patch.
-
-Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
-we see that if the patches each have sufficiently large area,
-the rate of migration and mutation will be on the same order if 
-$\mu \approx ((s_B+s_m)/2) \exp(-\sqrt{2 s_m} R / \sigma)$,
-where $R$ is the distance between the patches.
-(Note that in both cases, $A$ is the area of the not-yet-adapted patch.)
-
-If we take $\mu = 10^{-8}$, and $s_B=s_m=s$, then this is satisfied if $R/\sigma \approx (18+\log(s))/\sqrt{s}$,
-so if the selective pressure is fairly strong (say, $s=.05$),
-then the distance between patches must be about 70 times the dispersal distance (i.e.\ $R/\sigma \approx 68$);
-while if $s$ is much smaller, say $s = .001$, 
-then the patches must be separated by many hundreds of times the dispersal distance (i.e.\ $R/\sigma \approx 368$).
-The situation does not change much if $\mu$ is larger -- even if $\mu = 10^{-6}$, 
-the required separation between patches is only reduced by $5/\sqrt{s_m}$.
-If the separation between patches is much larger than this value -- certainly, by a factor of ten -- 
-then patches are effectively isolated -- most adaptation will occur independent between patches.
-On the other hand, if the separation is much smaller, then a single
-mutation is likely to fix in both patches.
-
-If we are willing to take these approximations at face value, 
-and assuming both rates are small, 
-the time till the first patch is colonized by the $B$ allele will be approximately exponentially distributed with rate $2 \mutrate$.
-Following this, the time till the second patch is subsequently colonized 
-(via either migration or new mutation) 
-will be exponentially distributed with rate $\mutrate+\migrate$.
-Both these times scale linearly with $\xi^2/(\rho A s_{B})$, so that 
-the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
-and decrease with patch population size and benefical seleciton coefficient.
-Furthermore, the probability that the second adaptation is a new mutation,
-i.e.\ the probability of parallel adapatation, is
-\begin{equation}
-  \frac{\mutrate}{\mutrate+\migrate} = \frac{2\mu/(s_B+s_m)}{2\mu/(s_B+s_m) + \exp\left(- \sqrt{2 s_m} R / \sigma \right) },
-\end{equation}
-so that the probability of parallel mutation should increase
-approximately logistically with the distance between the patches at rate $\sqrt{s_m} /\sigma$.
-
-%%%%
-\subsection{Multiple patches}
-
-We can extend this reasoning to multiple patches of unadapted habitat 
-in which the $B$ allele can facilitate local adaptation. 
-Under our assumptions of mutational exclusion and rapid establishment, 
-each patch will adapt through establishment of a single copy of the $B$ allele, 
-either by migration or mutation.
-Since we assume that the times between successive establishments is the result of many nearly-independent attempts
-with small probability of success,
-the process of occupation is well approximated by a discrete-state Markov chain,
-whose state is the labeling of patches according to the colonizing allele
-(or none, for those that haven't yet adapted).
-
-If not yet adapted,
-a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_B \mu \rho A/\xi^2$.
-Without loss of generality, assume that patches $1, \ldots, k$ are already adapted,
-and these patches are at distances $R_1, \ldots, R_k$ away from this unadapted patch, 
-then the total rate of adaptation through migration is
-\begin{equation}
-  \frac{ s_B (s_B+s_m) \rho A}{\xi^2} \sum_{i=1}^{k} \exp\left(- \sqrt{2 s_m} R_i/\sigma\right),
-\end{equation}
-and the probability that the adaptation is from patch $i$ is $\exp\left(- \sqrt{2 s_m} R_i/\sigma\right)/\sum_{j=1}^{k} \exp\left(- \sqrt{2 s_m} R_j/\sigma\right)$.
-These specify the transition rates of the process.
-
-Since the compound parameter $s_B \rho / \xi^2$ is common to both rates,
-it functions only as a time scaling of the process, 
-and therefore has no effect on the final configuration of the pattern of which patches have adapted
-in parallel versus have shared genetic basis to their adaptive response to the poor environment.
-
-Also note that we can rescale time by the typical patch size, and introduce a parameter, say $a_k = A_k/\bar A$,
-making the properties (other that the time scaling) of the discrete model independent of the \emph{numerical sizes} of the demes themselves.
-This is complementary to the results of \cite{softsweepsII}, who showed that multiple mutations are likely to arise \emph{within} a panmictic deme
-if the population-scaled mutation rate $2 N \mu$ is greater than $1$.
-
-\plr{conclusion?}
-
-%%%%%%%%%
-\subsection{Length of the hitchhiking haplotype}
-\label{ss:haplotype_length}
-
-If a patch adapts through new mutation, the genomic region surrounding the selected site will hitchhike \citep{hitchhiking} along with it,
-so that at least initially, all adapted individuals share a fairly long stretch of haplotype.
-This association gets slowly whittled down by recombination during interbreeding at the edge of the patch;
-but there will always be longer LD nearby to the selected site.
-\plr{Could work out on what time scale it gets whittled down and the shared length at stationarity.}
-
-On the other hand, if a an already adapted patch colonizes another through migration,
-then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
-The length of this haplotype is inversely proportional to the number of generations the allele spends ``in transit'' --
-this is because, following back the successful migrant lineage, 
-each reproductive event in locations where the allele is at low frequency
-whittles down the hitchhiking haplotype.
-(In this case, the haplotype is literally hitchhiking across geography!)
-
-Here is a simplified picture of the process of adaptation by migration.
-There is a constant outflux of migrants from the patch where the allele is already established; 
-any descendants of these carrying the allele are locally maladapted, so are mostly doomed to extinction.
-(We do not require that migration be blind to the environment --
-although adaptive migration would reduce the outflux of migrants --
-but do assume that the focal allele does not affect migration once outside the patch.)
-\plr{or maybe we should discuss this later?}
-However, eventually the descendants of one lucky migrant will reach the new patch and colonize it.
-(If the new patch is at some distance, it will almost certainly \plr{a.c.} be colonized first by only one family.)
-We will estimate the total outflux of migrants, and assume that their fates are independent of each other,
-so that we need only calculate the probability that offspring carrying the focal allele of a given migrant colonize the new patch,
-and the distribution of the transit time, given that this occurs.
-In doing so, we only follow the offspring of the migrant that carry the focal allele,
-and approximate these as a branching process, 
-i.e.\ assume that the number of offspring each produces is independent of each other.
-
+\subsubsection{The genealogy of migrant families}
+
+To understand the process of adaptation by migration,
+we therefore need to look more closely at the rare families of $B$ the allele far from the patches where it is established.
+Consider one such family.
+As noted above, we can model this by assuming that each individual
+migrates, then reproduces, independently of all other family members.
+The mean squared migration distance is $\sigma^2$, takes the migrant in a random direction,
+and if migration takes the individual into a patch where the $B$ allele is favored, 
+we drop the individual from our family.
+Since the $B$ allele is uniformly deleterious elsewhere,
+we can decouple reproduction and migration
+by first forming a tree in which each individual has a random number of offspring
+(with mean $1-s_m$ and variance $\xi^2$),
+then assigning spatial locations to each individual in the tree,
+and then pruning any branches that happen to wander into a patch.
+This pruning removes both branches leading back into the already established patch (which are irrelevant)
+and branches leading into the new patch (we will need to count these).
 Since we are concerned with the rare migrant families that transit quickly between patches, 
-we should remove any backmigrants into the already-established patch;
-we do this by assuming rather that the allele is deleterious \emph{everywhere} --
-backmigrants will contribute little, since they are going the wrong direction,
-and we deal with arrival at the new patch later.
-Spatial motion is now independent of reproduction,
-so the chance that offspring of the migrant first arrive in the new patch after $\tau$ generations 
-is the probability of survival of the family until time $\tau$
-multiplied by the probability that some of the family are in the new patch, given survival.
-The first probability comes from well-known properties of branching processes,
-and the second from Brownian motion.
-Then, the chance that the migrant family establishes in the new patch 
-depends on the size of the family that arrives in the patch,
-and the probability that each establishes.
-
-If each individual has on average $(1-s_m)$ offspring, then after $t$ generations,
-the family has on average $(1-s_m)^t$ members.
+backmigrants to the already established patch will contribute little, since they are going the wrong direction;
+so to simplify the analysis we ignore this detail.  % but could add this detail in an appendix
+
+Since each individual has on average $(1-s_m)$ offspring,
+a family has on average $(1-s_m)^t$ members in generation $t$.
 Let $p_e(t)$ be the chance of extinction by time $t$, and let $K_t$ be the (random) family size conditioned on nonextinction,
@@
 It turns out that by assuming the fate of each offspring is independent (i.e.\ that this forms a branching process),
-and that the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jaegers1976branching},
+and that the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jagers1975branching},
 that $K_t$ has a limiting distribution: $K_t \dconv K$ 
@@
 We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary}.
-When we condition on survival, we condition on existence of at least one lineage from time $0$ to time $t$.
+When we condition on survival until $t$, we condition on existence of at least one lineage from time $0$ to time $t$.
 Once given this lineage, nothing distinguishes any of the other reproduction events --
-then each individual in this lineage must give birth to at least one offspring,
+each individual in this lineage must give birth to at least one offspring,
 and all other reproduction events are unconditioned.
@@
 can be decomposed into a ``trunk'' that lasts until $t$
-anmd independent ``branches'' growing from each intermediate time $0\le s \le t$ 
-that may or may not survive until $t$.
-This process is depicted in Figure~XA.
+sprouting independent ``branches'' that may or may not survive until $t$.
+This process is depicted in Figure~\ref{fig:branching_decomp}A.
 Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$
@@
 \begin{align}
-  Z_t \; \vert \; Z_t>0 \deq \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
+  \left( Z_t \; \vert \; Z_t>0 \right) \deq \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
 \end{align}
@@
 
-Now we can add in the spatial motion, by following only the motion of the ``trunk''
-(i.e.\ the red line in figure \ref{fig:branching_decomp}).
-We will first work in one dimension.
-If the variance of the parent-offspring distance is $\sigma^2$ (and the distribution is not too heavy-tailed),
-then the motion of this is approximately Brownian, with variance $\sigma^2$ per generation.
-Suppose that the new patch is at distance $x$.
-The probability density for the first time such a Brownian motion $B_s$ hits $x$ is
+This suggests that we add in the spatial motion by following only the motion of the ``trunk''
+(i.e.\ the red line in figure \ref{fig:branching_decomp}),
+whose motion is is approximately Brownian, with variance $\sigma^2$ per generation.
+(Recall $\sigma$ is the dispersal distance.)
+Suppose that the new patch is at distance $r$.
+The probability density for the first time such a Brownian motion $B_s$ hits $r$ is
 \citep[XXX]{feller}
-% note $\sigma B_t \deq B_{\sigma^2 t}$ so $\P\{ \sup_{s\le t} \sigma B_s \le x \} = \P\{ \sup_{s \le \sigma^2 t} B_s \le x \}
+\plr{work in $d=2$}
+% note $\sigma B_t \deq B_{\sigma^2 t}$ so $\P\{ \sup_{s\le t} \sigma B_s \le r \} = \P\{ \sup_{s \le \sigma^2 t} B_s \le r \}
 \begin{align} \label{eqn:hitting_dist}
-  \P\{ B_t=x \;\mbox{and} \; B_u<x \;\mbox{for all}\; u<t\} =  \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right) .
-\end{align}
-The chance that the descendants of a given migrant arrives at the new patch at time $t$
-is now just equation~\eqref{eqn:hitting_dist} multiplied by the chance that 
-the migrant family is still alive at time $t$.
-For convenience, we define $\alpha = - \log(1-s_m) \approx s_m$,
-so that $1-p_e(t) \simeq (1/\mu_K) \exp(-\alpha t)$,
-and define $\mu_K$ to be the limiting mean family size conditioned on survival.
-By integrating over $t$, we get the chance that descendants of the migrant ever arrive at the new patch is
-\begin{align} 
-  & \int_0^\infty \frac{(1-s_m)^t}{\mu_K} \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right)  dt \\
-  & \qquad = \frac{1}{\mu_K} \exp\left( - \frac{ x \sqrt{2 \alpha}}{\sigma} \right) .\label{eqn:estab_prob}
-\end{align}
-(Note that this is, up to the constant $\mu_K$, the chance that a Brownian motion killed at rate $\alpha$
-ever hits $x$ \citep{feller}.
-This approximation holds if $x$ is large enough that the approximation $1-p_e(t) \approx (1-s_m)^t/\mu_K$ holds.)
-
-We now have the tools to compute the two quantities of interest:
-the probability that the family of a given migrant will eventually establish in the new patch,
-and the transit time conditioned on establishment.
-
-The probability of establishment of a single migrant in a patch distance $x$ away
-is equal to the probability that the migrant's family makes it to the patch at all,
-multiplied but the chance that it establishes locally, given that it makes it there.
-We have computed the first term in equation~\eqref{eqn:estab_prob},
-and we make the small assumption that the second term is just the probability of local fixation
-of a beneficial allele present in $K$ copies,
-i.e.~$\E[1-(1-p_f(s_b))^K] \approx 2 s_b \mu_K/\xi^2$,
-where $p_f(s_b) \approx 2s_b/\xi^2$ is the probability of fixation of a single mutant with additive benefit $s_b$.
-Therefore, the probability of establishment is
+  \P\{ B_t=r \;\mbox{and} \; B_u<r \;\mbox{for all}\; u<t\} =  \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2}\right) .
+\end{align}
+Let $\tau$ be this random time at which a given migrant family hits the new patch,
+with the convention that $\tau =\infty$ if the family never gets here.
+We approximate the chance that a given family arrives at the new patch at time $t$, i.e.\ that $\tau=t$,
+by the product of equation~\eqref{eqn:hitting_dist} and the probability that the migrant family survives until time $t$.
+For convenience, define $\alpha := - \log(1-s_m)$ to be the rate of decrease between patches (note that $\alpha \approx s_m$),
+and $\mu_K = \E[K]$ to be the limiting mean family size conditioned on survival,
+so that $1-p_e(t) \simeq (1/\mu_K) \exp(-\alpha t)$.
+Then, the probability distribution of $\tau$ is given by
+\begin{align}
+    \P\{ \tau \in dt \} =  \frac{1}{\mu_K} \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2} - \alpha t \right) dt  .
+\end{align}
+
+By integrating over~$t$, the chance that this family ever arrives at the new patch is
+\begin{align} 
+\label{eqn:estab_prob}
+  \begin{split}
+      \P\{ \tau < \infty \} &= \int_0^\infty \P\{ \tau \in dt \} = \frac{1}{\mu_K} \exp\left( - \frac{ r \sqrt{2 \alpha}}{\sigma} \right) .
+  \end{split}
+\end{align}
+This is, up to the constant $\mu_K$, the chance that a Brownian motion killed at rate $\alpha$
+ever hits $r$ \citep{feller}.
+This approximation depends on $r$ being large enough so that the approximation $1-p_e(t) \approx (1-s_m)^t/\mu_K$ holds.
+
+Now, each member of a family that successfully arrives at the new patch
+(where the $B$ allele is again beneficial)
+has a chance $p_e$ of establishing.
+The precise distribution of the number of members of a family that enter the patch
+given the trunk arrives there
+depends on the geometry of the patch,
+but since family members should not be too far from the trunk,
+we can approximate this by the total size of the family.
+If the family has $K$ individuals when the trunk arrives at the new patch,
+the chance that any establish is approximately $1-(1-p_e)^K$;
+and so that chance that the family establishes, given it arrives at the new patch, 
+is 
+\begin{align}
+    \label{eqn:prob_estab_arrival}
+    \E[1-(1-p_e)^K] \approx 2 s_p \mu_K/\xi^2 .
+\end{align}
+
+The probability of establishment of an outmigrant from the already-established patch is 
+equal to the probably the trunk arrives (equation~\eqref{eqn:estab_prob}) multiplied by the probability it establishes (equation~\eqref{eqn:prob_estab_arrival}),
+namely
 \begin{align} \label{eqn:bp_prob_estab}
-  F(x) &\approx \frac{2s_b}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} x}{\sigma} \right) ,
-\end{align}
-which verifies the less intuitive calculation of $\migrate$ above.
-
-Now, write $\tau$ for the first time the trunk lineage hits $x$, i.e.\ reaches the new patch,
-with the convention that $\tau=\infty$ if it never does.
+    F(r) &:= \text{( prob family establishes at $r$ )} = \frac{2s_p}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} r}{\sigma} \right) ,
+\end{align}
+
+%%%%%
+\subsubsection{The rate of outflux of migrant families}
+
+Now, we can obtain the effective ``outflux of families'' from information about the migrant families, combined with \eqref{eqn:gestalt_q}.
+Define $\gamma$ to be this outflux; 
+i.e.\ the mean number of individuals per unit time that are born carrying the $B$ allele closer than $r_0$ to the original patch
+but migrate to a larger distance.
+Then for $r>r_0$, the mean occupation density $q(r)$ is equal to $\gamma$ 
+multiplied by the mean occupation density of the family descending from such a migrant,
+which we denote $u(r)$.
+Denote by $Z_t$ the number of individuals such a family has after time $t$,
+and the locations of these by $X_1(t), \ldots, X_{Z_t}(t)$.
+Then this mean occupation density, at $x$, is the expected number of individuals ever present at $x$. 
+Since the expected number of individuals present at time $t$ is $e^{\alpha t}$
+and the location of each of these individuals is marginally distributed as a  Gaussian with variance $\sigma^2 t$,
+we can write this expectation as
+\begin{align}
+    u(x) &= \int_0^t \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
+    &= \int_0^t e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
+    &= \frac{ 1 }{ \sigma \sqrt{2 \alpha} } \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) 
+\end{align}
+\plr{check that}
+The integral is evaluated in Appendix~\ref{apx:some_integrals}.
+\plr{Also examine non-dependence on $r_0$.}
+
+Combining this with the solution for $q(r)$ (equation~\eqref{eqn:eqfreq}) as in \eqref{eqn:q_gestalt},
+we then have that (with $C=1/2$)
+\begin{align} \label{eqn:outflux}
+    \gamma = \text{( outflux of families )} \approx (1/2) \rho \sigma \sqrt{2\alpha} .
+\end{align}
+
+Combining this with \eqref{eqn:bp_prob_estab}, as in \eqref{eqn:gestalt_q}, 
+the rate of establishment, for a patch of area $A$ at distance $r$ from an already occupied patch, is approximately
+\begin{align} \label{eqn:migrate}
+    \migrate(r) = \gamma F(r) = \frac{\rho \sigma s_p\sqrt{2s_m} }{\xi^2}  \exp\left( -\frac{ \sqrt{2 s_m} r}{\sigma} \right).
+\end{align}
+Note that the exponential term is expected based on \eqref{eqn:eqfreq} -- 
+since establishment requires rare long distant migrants to be present, and \eqref{eqn:eqfreq} gives the density of such --
+but the first term required other arguments.
+
+If this establishment rate is low, 
+the time (in generations) before migrant copies of the allele successfully establish through migration in the second patch
+is approximately exponentially distributed with mean $1/\migrate$.
+Neglecting the time for such a migrant allele to establish once present in the second patch,
+the time scale on which migrant alleles transit and fix between patchs is $1/\migrate$.
+
+%%%%%%%%%
+\subsection{The probability of parallel adaptation between patches.} 
+\label{ss:probparallel}
+
+We now turn to the question of whether the separated populations adapt by parallel genetic changes 
+or by the spread of a migrant allele between the patches
+(or both, or neither).
+As only a single mutation is required for individuals to adapt to the
+poor habitat patch, subsequent mutations that arise after an allele becomes established in the patch gain no selective benefit. 
+Similarly, an allele introduced into a patch by migration will not be favored by selection to spread, 
+if the patch has already been colonized. 
+Therefore, mutations selectively exclude each other from patches, over short time scales, 
+and will only slowly mix together over longer time scales by drift.
+
+We will assume that once a $B$ allele is introduced (by migration or mutation), 
+it becomes established in the poor habitat patch rapidly. 
+Under this approximation, and the ``selective exclusion'' just discussed,
+after one of the patches becomes colonized by a single $B$ allele, 
+the other patch will become colonized by migration or mutation, but not by both. 
+As such, the question of how the the second patch adapts
+simply comes down to whether the a new mutation or a migrant allele is the first to become established in the second patch.
+
+Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
+we see that if the patches each have sufficiently large area,
+the rate of migration and mutation will be on the same order if 
+$\mu \approx ((s_p+s_m)/2) \exp(-\sqrt{2 s_m} r / \sigma)$,
+\plr{check}
+where $r$ is the distance between the patches.
+(Note that in both cases, $A$ is the area of the not-yet-adapted patch.)
+
+If we take $\mu = 10^{-8}$, and $s_p=s_m=s$, then this is satisfied if $r/\sigma \approx (18+\log(s))/\sqrt{s}$,
+so if the selective pressure is fairly strong (say, $s=.05$),
+then the distance between patches must be about 70 times the dispersal distance (i.e.\ $r/\sigma \approx 68$);
+while if $s$ is much smaller, say $s = .001$, 
+then the patches must be separated by many hundreds of times the dispersal distance (i.e.\ $r/\sigma \approx 368$).
+The situation does not change much if $\mu$ is larger -- even if $\mu = 10^{-6}$, 
+the required separation between patches is only reduced by $5/\sqrt{s_m}$.
+If the separation between patches is much larger than this value -- certainly, by a factor of ten -- 
+then patches are effectively isolated -- most adaptation will occur independent between patches.
+On the other hand, if the separation is much smaller, then a single
+mutation is likely to fix in both patches.
+
+If we are willing to take these approximations at face value, 
+and assuming both rates are small, 
+the time until the first of the two patches is colonized by the $B$ allele will be approximately exponentially distributed with mean $1/(2 \mutrate)$.
+Following this, the time till the second patch is subsequently colonized 
+(via either migration or new mutation) 
+will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
+Both these times scale linearly with $\xi^2/(\rho A s_{B})$, \plr{check}
+so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
+and decrease with patch population size and benefical selection coefficient.
+Furthermore, the probability that the second adaptation is a new mutation,
+i.e.\ the probability of parallel adapatation, is
+\begin{equation}
+  \frac{\mutrate}{\mutrate+\migrate} = \frac{2\mu/(s_p+s_m)}{2\mu/(s_p+s_m) + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
+\end{equation}
+\plr{check}
+so that the probability of parallel mutation should increase
+approximately logistically with the distance between the patches, at rate $\sqrt{s_m} /\sigma$.
+
+%%%%%%%%%%%%
+\subsection{Multiple patches}
+
+We can extend this reasoning to multiple patches of unadapted habitat 
+in which the $B$ allele can facilitate local adaptation. 
+Under our assumptions of mutational exclusion and rapid establishment, 
+each patch will adapt through establishment of a single copy of the $B$ allele, 
+either by migration or mutation.
+If we imagine that the times between successive establishments is the result of many nearly-independent attempts
+with small probability of success,
+the process of occupation is well approximated by a discrete-state Markov chain,
+whose state is the labeling of patches according to the colonizing allele
+(or none, for those that haven't yet adapted).
+
+If not yet adapted,
+a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$.
+Without loss of generality, assume that patches $1, \ldots, k$ are already adapted,
+and these patches are at distances $r_1, \ldots, r_k$ away from this unadapted patch, 
+then the total rate of adaptation through migration is
+\begin{equation}
+  \frac{ s_p (s_p+s_m) \rho A}{\xi^2} \sum_{i=1}^{k} \exp\left(- \sqrt{2 s_m} r_i/\sigma\right),
+\end{equation}
+and the probability that the migrant, adapted allele 
+comes from patch $i$ is $\exp\left(- \sqrt{2 s_m} r_i/\sigma\right)/\sum_{j=1}^{k} \exp\left(- \sqrt{2 s_m} r_j/\sigma\right)$.
+These specify the transition rates of the process.
+
+Since the compound parameter $s_p \rho / \xi^2$ is common to both rates,
+it functions only as a time scaling of the process, 
+and therefore has no effect on the final configuration, namely, 
+the sets of patches sharing the genetic basis of their adaptive response.
+
+Also note that we can rescale time by the typical patch size, and introduce a parameter, say $a_k = A_k/\bar A$,
+making the properties (other that the time scaling) of the discrete model independent of the \emph{numerical sizes} of the demes themselves.
+This is complementary to the results of \cite{softsweepsII}, who showed that multiple mutations are likely to arise \emph{within} a panmictic deme
+if the population-scaled mutation rate $2 N \mu$ is greater than $1$.
+
+%%%%%%%%%
+\subsection{Length of the hitchhiking haplotype}
+\label{ss:haplotype_length}
+
+If a patch adapts through new mutation, the genomic region surrounding the selected site will hitchhike \citep{maynardsmith1974hitchhiking,kaplan1989hitchhiking} along with it,
+so that at least initially, all adapted individuals share a fairly long stretch of haplotype.
+This association gets slowly whittled down by recombination during interbreeding at the edge of the patch;
+but there will always be longer LD nearby to the selected site.
+\plr{Could work out on what time scale it gets whittled down and the shared length at stationarity.}
+
+On the other hand, if an already adapted patch colonizes another through migration,
+then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
+The length of this haplotype is roughly inversely proportional to the number of generations the allele spends ``in transit'',
+because each reproductive event in locations where the allele is at low frequency
+recombines haplotypes of the $b$ allele closer to the adaptive allele.
+(In this case, the haplotype is literally hitchhiking across geography!)
+
+As noted above, the length of the hitchhiking segment depends on the time
+spent ``in transit'' between the patches.
+Above, we approximated this by the time $\tau$ it takes the trunk lineage to hit the new patch,
+except that studying a successful migrant family means we need to condition on the establishment occurring, i.e.\ on $\tau < \infty$.
 By the above arguments, and equation~\eqref{eqn:estab_prob}, we have that the conditional probability density is
@@
   \P\{\tau\in dt | \tau<\infty \} &= \frac{\P\{\tau \in dt\}}{\P\{\tau<\infty\}} \\
-  &= \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2} -\alpha t + \sqrt{\alpha} \frac{x}{\sigma} \right) .
-\end{align}
-Although $\tau$ is interesting, we are more interested in the length of the hitchhiking segment,
-since that is where our signal would come from.
+  &= \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2} -\alpha t + \sqrt{\alpha} \frac{r}{\sigma} \right) .
+\end{align}
+\plr{more dimensions}
 Since each generation spent in transit provides an opportunity for recombination,
@@
 \P\{L>\ell\} &= \E[e^{-\ell \tau}|\tau<\infty] \\
-  &= \exp\left\{{-\frac{x}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
+  &= \exp\left\{{-\frac{r}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
 \end{align}
 This is also the Laplace transform of $\tau$;
-we can differentiate it to find that $\E[\tau] = x/(\sigma\sqrt{2\alpha})$
-and $\var[\tau] = x/( 2^{3/2}\alpha^{3/2}\sigma)$.
-Integrating~\eqref{eqn:haplen_cdf} gives that 
-$\E[L] = \sigma \sqrt{2\alpha}/x + \sigma^2/x^2$ and $\var[L] = 2\alpha\sigma^2/x^2 + 4 \sigma^3 \sqrt{2\alpha}/x^3 + 5 \sigma^4 / x^4$;
-an easier way to see this is to note that if $Y$ is an exponential random variable with rate $x\sqrt{2}/\sigma$,
-then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$.
+we can differentiate it to find that $\E[\tau] = r/(\sigma\sqrt{2\alpha})$
+and $\var[\tau] = r/( 2^{3/2}\alpha^{3/2}\sigma)$.
+Furthermore, this implies that if $Y$ is an exponential random variable with rate $r\sqrt{2}/\sigma$,
+then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$,
+from which we can obtain that
+$\E[L] = \sigma \sqrt{2\alpha}/r + \sigma^2/r^2$ and $\var[L] = 2\alpha\sigma^2/r^2 + 4 \sigma^3 \sqrt{2\alpha}/r^3 + 5 \sigma^4 / r^4$.
 
 %%%
-\subsubsection{Limiting distributions}
-
-We are already assuming that $x$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
-by taking the large-$x$ limit above we can obtain more intuitive descriptions of the process
+\subsubsection{Limiting distribution of the transit time and hitchhiking haplotype}
+
+In the above we already assumed that the inter-patch distance $r$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
+by taking the large-$r$ limit above we can obtain a more intuitive description of the process
 via a central limit for $\tau$.
-Define $\eta_x$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
-so that $\tau = x/\sigma(2\alpha)^{-1/2} + \eta_x \sqrt{x/\sigma} (2\alpha)^{-3/4}$.
+Define $\eta_r$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
+so that $\tau = r/\sigma(2\alpha)^{-1/2} + \eta_r \sqrt{r/\sigma} (2\alpha)^{-3/4}$.
 Plugging this into equation~\eqref{eqn:haplen_cdf},
-after some algebra (and using that $\sqrt{x+\epsilon} = \sqrt{x} + \frac{\epsilon}{2\sqrt{x}} - \frac{\epsilon^2}{2\sqrt{x}^3} + O(\epsilon^3)$) we obtain that
-\begin{align}
-  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{x}\right) \right\}.
+after some algebra (and using that $\sqrt{r+\epsilon} = \sqrt{r} + \frac{\epsilon}{2\sqrt{r}} - \frac{\epsilon^2}{2\sqrt{r}^3} + O(\epsilon^3)$) we obtain that
+\begin{align}
+  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{r}\right) \right\}.
 \end{align}
 By the properties of Laplace transforms \citep{durrett},
-this implies that $\eta_x$ converges in distribution to a standard Gaussian random variable as $x \to \infty$.
+this implies that $\eta_r$ converges in distribution to a standard Gaussian random variable as $r \to \infty$.
 In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
 Since the standard deviation of $\tau$ is of lower order than its mean,
-a similar calculation shows that for large $x$, the haplotype halflength $L$ is approximately exponentially distributed, 
-with mean $\sqrt{2\alpha} \sigma/x$.
+a similar calculation shows that for large $r$, the haplotype halflength $L$ is approximately exponentially distributed, 
+with mean $\sqrt{2\alpha} \sigma/r$.
 
 % \begin{align}
-%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{x}{\sigma\sqrt{2\alpha}}\right)\frac{(2\alpha)^{3/4}\sqrt{\sigma}}{\sqrt{x}} \right\} \right] \\
-%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2\alpha)^{3/4} \sqrt{\sigma}}{\sqrt{x}}\right) \tau \right\} \right] \exp\left( (2\alpha)^{1/4} \sqrt{\frac{x}{\sigma}} \right) \\
+%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{r}{\sigma\sqrt{2\alpha}}\right)\frac{(2\alpha)^{3/4}\sqrt{\sigma}}{\sqrt{r}} \right\} \right] \\
+%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2\alpha)^{3/4} \sqrt{\sigma}}{\sqrt{r}}\right) \tau \right\} \right] \exp\left( (2\alpha)^{1/4} \sqrt{\frac{r}{\sigma}} \right) \\
 % \end{align}
@@
 \section{Applications} 
-Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by XXX.
-These mice live throughout the American SouthWest and into Mexico, and mostly live on the light colored rock outcrops, and have light pigmentation \citep[presumably an adaptation to visual-based predators][]{}.
-However, in some regions these mice live on dark colored substrates, including old lava flows, and have adapted to these patches in part through changes to a dark pigmentation.
-On one of these flows (Pinacate) an allele at the Mc1r locus is responsible for much of the change to a dark pelage \citep{Nachman:03}.
-This dark allele differs from the light allele by four amino acid changes - and has a dominant or partially dominant effect depending on the measure of coat color assessed.
-These lava flows are separated from each other by a range of distances, with some being separated from any other flow by hundreds of kilometers of light colored substrate.
-The Pinacate allele is not present in a number of other dark colored rock pocket mouse populations suggesting these populations have adapted in parallel.
-However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, some nearby populations may share a dark phenotype whose genetic basis has been spread by migration.
-
-As can be seen from above a key parameter is the characteristic lengths $\sigma/\sqrt{s_m}$.
-\citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby light substrate sites.
- On the lava flow the dark MC1R allele is at $86\%$ frequency. At the Tule light substrate site, which is $\sim 12km$ from the flow, the dark MC1R allele was at $\sim 3\%$ frequency. 
-At a greater distance from the lava ($XXX$km) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass. 
-While at the other light substrate site (O'Neill), which is $10$km from the flow on opposite side from Tule, the dark MC1R allele has a frequency of $\sim 34\%$.
+Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by \citet{dice1940ecologic,benson1933concealing}.
+These mice live on rock outcrops throughout the southwestern United States and northern Mexico,
+and generally have light pigmentation, presumably to match the color of the rock, avoiding visual-based predators. 
+However, in some regions these mice live on darker substrates such as old lava flows, 
+and in these patches have much darker pigmentation. 
+\citet{nachman2003different} demonstrated that on one of these flows (Pinacate), an allele at the MC1R locus is responsible for much of the change to a dark pelage.
+This dark allele differs from the light allele by four amino acid changes, 
+and has a dominant or partially dominant effect depending on the measure of coat color assessed. 
+These lava flows are separated from each other by a range of distances, 
+with some being separated from any other by hundreds of kilometers of light colored substrate. 
+The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
+suggesting these populations have adapted in parallel. 
+However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, 
+multiple dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
+despite intervening light habitat.
+
+As can be seen from the above, a key parameter is the characteristic length $\sigma/\sqrt{s_m}$. 
+\citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby sites in light-colored rock.
+On the lava flow the dark MC1R allele is at 86\% frequency. 
+At the Tule light substrate site, which is $\sim$ 12km from the flow, the dark MC1R allele was at $\sim$ 3\% frequency. 
+At a greater distance from the lava (XXXkm) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass. 
+While at the other light substrate site (O'Neill), which is 10km from the flow on opposite side from Tule, 
+the dark MC1R allele has a frequency of $\sim$ 34\%. 
 On the basis of these numbers we can gain a sense of a plausible range of characteristic lengths.
-Assuming the allele is at $50\%$ frequency at the edge of the lava flow, 
+Assuming the allele is at 50\% frequency at the edge of the lava flow, 
 we can fit the frequency found at each of the light substrate sites that are 
@@
 Doing this for the Tule we obtain $\sigma/\sqrt{s_m} \approx 4$km and for the O'Neill site
-$\sigma/\sqrt{s_m} \approx 30$km. \\
+$\sigma/\sqrt{s_m} \approx 30$km.
 
@@
 
+unequal density: Lenormand, gene flow and limits of\dots
+
+small-scale fluctuations
+
 \bibliography{standing_patches_refs}
@@
 
+%%%%%
+\section{Numerical calculation of the probability of establishment}
+\label{apx:establishment_sims}
+
+To investigate the impact of the various approximations surrounding the mutational influx,
+we carried out numerical calculations of a fine-grained, discrete model.
+These are shown, and described, in figure~\ref{fig:prob_estab_calcs}.
+
+Here (and at other parameter choices) we see that the probability of establishment $p(x)$ goes to the equilibrium value
+(approximately $p_e = 2s/\xi^2$) within the patch;
+the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
+This lends credence to our approximation that the integral of $p(x)$ over the entire range
+is close to $p_e$ multiplied by the area of the patch.
+
+\begin{figure}[ht!]
+    \begin{center}
+        % made in patchy-selection-plots.R
+        \includegraphics{prob-establishment}
+    \end{center}
+    \label{fig:prob_estab_calcs}
+    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
+    in an abrupt transition (left) and a gradual transition (right).
+    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp .05$ respecively);
+    and the selection coefficient interpolates linearly between these in the central shaded region.
+    The number of offspring is Poisson.
+    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
+    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
+    }
+\end{figure}
+
+%%%%%%%%%
+\section{Simulations of a patchy environment}
+\label{apx:patchy_sims}
+
+Describe sims.
+
+Argue that $C=0.5$ is a reasonable number.
+
+\begin{figure}[ht!]
+    \begin{center}
+        % made in patchy-sim-writeup-plots.R
+        \includegraphics{example-equilibrium}
+    \end{center}
+    \label{fig:prob_estab_calcs}
+    \caption{
+    Snapshots of a simulation of allelic frequency dynamics around a patch where the allele is beneficial:
+    the focal allele is benefical in the middle region and deleterious outside.
+    The dynamics were simulated for 10000 generations across demes with 1000 individuals in each
+    and random dispersal to nearby demes with a mean dispersal distance of $\sigma \approx 2$ demes.
+    Colored lines show allele frequencies across space in 50 uniformly spaced generations;
+    the solid line shows the mean across all generations.
+    Below is the same plot, except on a log scale.
+    }
+\end{figure}
+
+%%%%%%%%%%%%
+\section{The equilibrium frequency}
+
+One route to the ``equlibrium frequency'' of the allele outside the range where it is advantageous is as follows;
+see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} for other arguments in this case, 
+or \citet{etheridge} or \citet{dawson} for general theory used below.
+Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
+and that selection (for or against) the allele is given by the function $s(x)$, with $x$ denoting the spatial location.
+Then each individual at location $x$ reproduces after a random lifetime,
+dying and giving birth in the same deme to a random number of offspring with distribution given by $X$;
+who all migrate to a new location chosen as $x+R$
+and replace randomly chosen individuals there.
+If $x+R$ is outside of the range, then they perish.
+Each individual's reproduction time is exponentially distributed, 
+either with rate 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
+Suppose that $R$ has mean zero and variance $\sigma^2$, and that $X$ has mean $\mu$ and variance $\xi^2$.
+We also assume that the distribution of $X$ makes the resulting Markov chain irreducible and aperiodic.
+
+Let $\Phi^N_t(x)$ be the number of mutant alleles present at location $x$ at time $t$,
+and denote by $\delta_x$ a single unit at location $x$, so that e.g.~$\Phi^N_t + \delta_x/N$
+is the configuration after a mutant allele has been added to location $x$.
+For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the number of mutant alleles added if $X$ new offspring carrying mutant alleles
+replace randomly chosen individuals in a deme where the mutant allele is at frequncy $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
+similarly, $\tilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
+(We like to think of $\Phi^N_t$ as a measure, but it does not hurt to think of $\Phi^N$ as a vector;
+we aren't providing the rigorous justification here.)
+Then we know that for any sufficiently nice function $f(\phi)$ that
+\begin{align} \label{eqn:discrete_generator}
+  \begin{split} \frac{\partial}{\partial t} \E\left[ f(\Phi^N_t) \right] 
+  &= N \sum_x \left\{ \E\left[ (1+s(x+R)) \Phi^N_t(x+R) \left( f\left(\Phi^N_t + \frac{\bar X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right. \\
+     & \qquad  \qquad \left.
+{} + \E\left[ \left(1-\Phi^N_t(x+R)\right) \left( f\left(\Phi^N_t - \frac{\tilde X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right\}  \end{split} \\
+     &= \mu \sum_x \E\left[ \left(\partial_{\phi(x)} f(\Phi_t) \right) \left\{ \Phi_t(x+R) - \Phi_t(x) + s(x+R) \Phi_t(x+R) (1-\Phi_t(x)) \right\} \right] + O\left(\frac{1}{N}\right).
+\end{align}
+This follows in part from $\E[\bar X_\phi] = \phi \mu$ and $\E[\tilde X_\phi] = (1-\phi)\mu$.
+We can see two things from this:
+First, since this is a first-order differential operator, the limiting stochastic process
+is deterministic (check by applying to $f(\phi) = \phi(x)^2$ to find the variance).
+Second, if we want to rescale space as well to get the usual differential equation, 
+we need to choose $\var[R]=\sigma^2$ and $s(x)$ to be of the same, small, order; 
+this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale.
+More concretely, suppose that the grid size is $\epsilon \to 0$, that $\var[R] = \sigma^2 \epsilon$, and that $s(x)/\epsilon \to \gamma(x)$,
+and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
+then the previous equation converges to the familiar form:
+\begin{align}
+  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \partial_x^2 \xi(t,x) + \gamma(x) \xi(t,x) (1-\xi(t,x)) \right) .
+\end{align}
+
+The process $\Phi_t$ is an irreducible, aperiodic finite-state Markov chain with absorbing states at 0 and 1;
+therefore, the inevitable outcome for finite $N$ is extinction of one type or another.
+Setting the left-hand side to zero in \eqref{eqn:discrete_generator} gives us equations for the moments of the stationary distribution,
+for instance, if $\phi(x) = \lim_{t \to \infty} \E[ \Phi_t(x) ]$
+
+%%%%%%%
+\section{The exact solution to the equilibrium frequency}
+\label{apx:elliptic_integrals},
+
+Or, in other words, solving the Fisher-KPP equation with varying selection.
+Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
+so that 
+\[
+\partial_t p(t,x) = \frac{1}{2} \sigma^2 \partial_x^2 p(t,x) + s(x) p(t,x) (1-p(t,x)) .
+\]
+The stable distribution $\phi(x) = \lim_{t\to\infty} p(t,x)$ then solves
+\begin{align} \label{eqn:definingphi}
+    \partial_x^2 \phi(x) = - 2 s(x) \phi(x) (1-\phi(x)) /\sigma^2,
+\end{align}
+with appropriate boundary conditions.
+First rescale space by $\sigma/\sqrt{2}$ so the $\sigma^2/2$ term dissappears.
+If we now assume that $s(x)$ is piecewise constant,
+$s(x) = s_i$ for $x \in [x_i,x_{i+1})$, with $x_0=-\infty$ and $x_{n+2}=\infty$,
+then the equation is integrable: if we multiply through by $2\partial_x \phi(x)$ and integrate, then we get that
+\begin{align} \label{eqn:conservation}
+    ( \partial_x \phi(x) )^2  &= - \int^{x} 2 s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
+        &= - s_i \phi(x)^2 \left( 1 - \frac{2}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) \\
+        &:= - V_i(\phi(x)) \quad \mbox{for } x \in [x_i,x_{i+1}) ,
+\end{align}
+where $K_i$ is a constant of integration and we define
+\[
+        V_i(\phi) = s_i \phi^2 \left( 1 - \frac{2}{3} \phi \right) + K_i .
+\]
+Note that $V_i'(0)=V_i'(1)=0$, that $V_i(0)=K_i$ and $V_i(1) = K_i+s_i/3$.  
+We will always have that $V(\phi) \le 0$.
+(We have then that $\phi'' = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
+Also note that this implies ``conservation of energy'', i.e.\ $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
+Rearranging, we get that $dx = d\phi / \sqrt{-V(\phi)}$, so where $\phi(x)$ is monotone, the inverse is
+\[
+    x(\phi) = x(\phi_0) \pm \int_{\phi_0}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } .
+\]
+For each $i$ then define the elliptic function
+\[
+    F_i(\phi) = \int_{\phi_i^*}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } ,
+\]
+where take the positive branch of the square root, and $\phi_i^*$ will be chosen later.
+Then we have that
+$x(\phi) - x(\phi_0) = \pm( F(\phi) - F(\phi_0))$,
+or for an appropriate $x_0$,
+\[
+    \phi(x) = F^{-1}\left( F(\phi(x_0)) \pm (x - x_0) \right).
+\]
+
+We clearly want $\lim_{x \to \infty} \partial_x \phi(\pm x) = 0$, 
+and $\lim_{x \to \infty} \phi(\pm x)$ to be zero or one depending on the sign of $s_0$ and $s_{n+1}$.
+Since $(\partial_x \phi(x))^2 = -V(x)$, this implies that if $s_0<0$, then $K_0 = 0$,
+while if $s_0>0$ then $K_0 = s_0/6$; and likewise for $K_{n+1}$.
+
+We also require that $\phi(x)$ and $\phi'(x)$ are continuous.
+Continuity of $\phi'(x)$ is equivalent to $V_i(x_{i+1}) = V_{i+1}(x_{i+1})$,
+which we can rearrange to find an equation for $K_{i+1}$ in terms of $K_i$ and $\phi(x_{i+1})$:
+\[
+    K_{i+1} - K_i = (s_i - s_{i+1}) \phi(x_{i+1})^2 (1-2\phi(x_{i+1})/3) .
+\]
+What about $\phi(x_i)$?  Well, if $\phi(x)$ is monotone on $[x_i,x_{i+1})$ then we can wolog take $\phi_i^*=0$ or $1$ depending on the sign of $s_i$.
+Otherwise, let $\phi_i^*$ be the (unique) root of $V_i$ in $[x_i,x_{i+1})$, so $V_i(\phi_i^*)=0$.
+In this case, $\phi_i^*$ is the maximum or minimum of $\phi$ in the interval:
+if $s>0$, then $\phi_i = \max\{ \phi(x) : x \in [x_i,x_{i+1})\}$. 
+Recall we defined $F_i$ using $\phi_i^*$; now using the fact that $\phi$ is monotone with the opposite sign on either side of $\phi_i^*$,
+$x_{i+1} - x_i = F(\phi(x_{i+1})) - F(\phi_i^*) + F(\phi(x_{i})) - F(\phi_i^*)$,
+and that $F(\phi_i^*) = 0$,
+we know that the length of the $i$th stretch is
+\[
+    x_{i+1} - x_i = \pm \left( F_i(\phi(x_i)) + F_i(\phi(x_{i+1})) \right).
+\]
+
+Note that all the $\pm{}$'s are easily relatable to the signs of $s_i$.
+If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be all set.
+Unfortunately, we are given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
+Monty (in "Gene flow and selection in a cline") deals with this by using symmetry, so he knows $\phi(0)$ and $\phi'(0)$.
+
+We are interested in the following case:
+\[
+    s(x) = \begin{cases}
+        +s_b \qquad \mbox{for } |x|<\epsilon \\
+        -s_d \qquad \mbox{for } x<-\epsilon \; \mbox{or}\; \epsilon \le x \le R \\
+        -s_* \qquad \mbox{for } R < x < R+2\epsilon \\
+        -s_d \qquad \mbox{for } x>R+2\epsilon .
+    \end{cases}
+\]
+We can solve the case with $s_* = s_d$, since $\phi'(0) = 0$ (as can Monty, although he doesn't say what it is).
+We should be able to argue that if $R$ is large then the value of $s_*$ won't make much difference.
+
+\subsection{Doing the integrals}
+
+This from ``NEQwiki, the nonlinear equations encyclopedia'', \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
+Recall that $V_i(\phi) =  s_i \phi^2(1-2\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries.
+Note that if we rescale space by $\sqrt{s}/\sigma$,
+then $s$ and $\sigma$ drop from the equation.
+We want to integrate
+\[
+    \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
+         \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } .
+\]
+Let $\phi^2(1-2\phi/3)+K = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
+and change variables first to $y^2=(\alpha-\phi)$, 
+and then to $x = y/\sqrt{\alpha-\beta}$, so that
+\begin{align*}
+    \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } 
+        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-y^2) } } \\
+        = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
+\end{align*}
+with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
+Now Jacobi's ``incomplete elliptic integral of the first kind'' is
+\[
+    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) }} ,
+\]
+and the Jacobian elliptic function $\sn(x;k)$ is the inverse: $F(\sn(x;k);k) = x$.
+
+As $k \to 0$, $\sn(x;k) \to \sin(x)$, while as $k \to 1$, $\sn(x;k) \to \sinh(x)$.
+
+Presumably, this gets the same answer as Monty, who finds that in his solvable cases,
+for $x$ in the last interval (which ends in $\infty$), that
+\begin{align*}
+    \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / \sigma + C ) \right) \\
+        &\simeq C' \exp\left( - \frac{ 2 x \sqrt{2s} }{ \sigma } \right) .
+\end{align*}
+% Indeed, as $\phi \to 0$, we have approximately $\phi' = \frac{ \sqrt{s} }{ \sigma } \phi$.  % huh????
+
+%%%%%%%%%%%%
 \section{Integrals appearing in the text}
@@
 \end{align}
-in our case, $\E[Y_x] \approx (s_B+s_m)$ and $p_e \approx 2 s_B/\xi^2$,
-so $\tau$ is approximately exponentially distributed with rate $ 2 z_A (s_B+s_m) s_B/\xi^2 $.
+in our case, $\E[Y_x] \approx (s_p+s_m)$ and $p_e \approx 2 s_p/\xi^2$,
+so $\tau$ is approximately exponentially distributed with rate $ 2 z_A (s_p+s_m) s_p/\xi^2 $.
 We can get $z_A$, the occupation density without the new patch, 
@@
 The mean total family size of an allele with fitness $1-s_m$ (i.e.\ a subcritical branching process) is $\sum_{t \ge 0} (1-s_m)^t = 1/s_m$;
-so the mean number of ``extra'' offspring is $1+s_B/s_m$.
+so the mean number of ``extra'' offspring is $1+s_p/s_m$.
 Establishment events are correlated to the extent that single such families give rise to multiple establishment events,
@@
 \texttt{XXX omit this? XXX}
-We have treated the time during which the selected allele is at intermediate frequency in a deme (about $\log(\sigma^2 \rho)/s_b$ generations) as negligible, 
+We have treated the time during which the selected allele is at intermediate frequency in a deme (about $\log(\sigma^2 \rho)/s_p$ generations) as negligible, 
 which ignores migration events that occur while the mutation is spreading through the population. 
@@
 For each occupied--unoccupied pair of demes, if the number of mutant alleles in the occupied deme is $n(t)>0$,
-then at time $t$, migrants from the occupied deme pass the selected allele to the unoccupied deme at rate $2 n(t) m p_f(s_b)$,
-while newly arisen mutants arise and fix in the unoccupied deme at constant rate $2 N \mu p_f(s_b)$.
+then at time $t$, migrants from the occupied deme pass the selected allele to the unoccupied deme at rate $2 n(t) m p_f(s_p)$,
+while newly arisen mutants arise and fix in the unoccupied deme at constant rate $2 N \mu p_f(s_p)$.
 The selected type should grow approximately as $n(t)=\exp(s(t-t_0))$.
@@
 but we can at least make a few observations.
-One is that the final distribution of types is no longer independent of $s_b$, which enters through the dynamics $n(t)$.
+One is that the final distribution of types is no longer independent of $s_p$, which enters through the dynamics $n(t)$.
 However, it is clear that increasing the selection coefficient $s$ will decrease the number of types (independent mutational origins),
@@
 
-\section{The equilibrium frequency}
-
-One route to the ``equlibrium frequency'' of the allele outside the range where it is advantageous is as follows;
-see \citet{slatkin} or \citet{barton,pollack} for other arguments in this case, or \citet{etheridge} or \citet{dawson} for general theory
-Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
-and that selection (for or against) the allele is given by the function $s(x)$, with $x$ denoting the spatial location.
-Then each individual at location $x$ reproduces after a random lifetime,
-dying and giving birth in the same deme to a random number of offspring with distribution given by $X$;
-who all migrate to a new location chosen as $x+R$
-and replace randomly chosen individuals there.
-If $x+R$ is outside of the range, then they perish.
-Each individual's reproduction time is exponentially distributed, 
-either with rate 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
-Suppose that $R$ has mean zero and variance $\sigma^2$, and that $X$ has mean $\mu$ and variance $\xi^2$.
-We also assume that the distribution of $X$ makes the resulting Markov chain irreducible and aperiodic.
-
-Let $\Phi^N_t(x)$ be the number of mutant alleles present at location $x$ at time $t$,
-and denote by $\delta_x$ a single unit at location $x$, so that e.g.~$\Phi^N_t + \delta_x/N$
-is the configuration after a mutant allele has been added to location $x$.
-For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the number of mutant alleles added if $X$ new offspring carrying mutant alleles
-replace randomly chosen individuals in a deme where the mutant allele is at frequncy $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
-similarly, $\tilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
-(We like to think of $\Phi^N_t$ as a measure, but it does not hurt to think of $\Phi^N$ as a vector;
-we aren't providing the rigorous justification here.)
-Then we know that for any sufficiently nice function $f(\phi)$ that
-\begin{align} \label{eqn:discrete_generator}
-  \begin{split} \frac{\partial}{\partial t} \E\left[ f(\Phi^N_t) \right] 
-  &= N \sum_x \left\{ \E\left[ (1+s(x+R)) \Phi^N_t(x+R) \left( f\left(\Phi^N_t + \frac{\bar X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right. \\
-     & \qquad  \qquad \left.
-{} + \E\left[ \left(1-\Phi^N_t(x+R)\right) \left( f\left(\Phi^N_t - \frac{\tilde X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right\}  \end{split} \\
-     &= \mu \sum_x \E\left[ \left(\partial_{\phi(x)} f(\Phi_t) \right) \left\{ \Phi_t(x+R) - \Phi_t(x) + s(x+R) \Phi_t(x+R) (1-\Phi_t(x)) \right\} \right] + O\left(\frac{1}{N}\right).
-\end{align}
-This follows in part from $\E[\bar X_\phi] = \phi \mu$ and $\E[\tilde X_\phi] = (1-\phi)\mu$.
-We can see two things from this:
-First, since this is a first-order differential operator, the limiting of the stochastic process
-is deterministic (check by applying to $f(\phi) = \phi(x)^2$ to find the variance).
-Second, if we want to rescale space as well to get the usual differential equation, 
-we need to choose $\var[R]=\sigma^2$ and $s(x)$ to be of the same, small, order, another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale.
-More concretely, suppose that the grid size is $\epsilon \to 0$, that $\var[R] = \sigma^2 \epsilon$, and that $s(x)/\epsilon \to \gamma(x)$,
-and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
-then the previous equation converges to the familiar form:
-\begin{align}
-  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \partial_x^2 \xi(t,x) + \gamma(x) \xi(t,x) (1-\xi(t,x)) \right) .
-\end{align}
-
-The process $\Phi_t$ is an irreducible, aperiodic finite-state Markov chain with absorbing states at 0 and 1;
-therefore, the inevitable outcome for finite $N$ is extinction of one type or another.
-Setting the left-hand side to zero in \eqref{eqn:discrete_generator} gives us equations for the moments of the stationary distribution,
-for instance, if $\phi(x) = \lim_{t \to \infty} \E[ \Phi_t(x) ]
-
 \end{document}

### 2013-07-23 patchy-sim-writeup-plots.R

@@ -1,0 +1,79 @@
+#!/usr/bin/R
+source("sim-patchy-selection-fns.R")
+source("lineages.R")
+source("sim-patchy-selection-fns.R")
+
+
+run.names <- c( '3975-1-10000-pophistory-run.Rdata', '5473-1-10000-pophistory-run.Rdata' )
+
+run.id <- gsub( "([^-]*)-.*","\\1",run.names[1] )
+load(run.name)
+print(c("run",run.id,"range",pophist$pop$params$range))
+
+# time slices of the process
+pdf(file='example-equilibrium.pdf', width=6, height=6, pointsize=10)
+par(mar=c(5,4,1,1)+.1)
+layout(1:2)
+timeslice <- floor( seq( 2000, 10000, length.out=50 ) )
+
+matplot( pophist$pophist[,,2,timeslice]/pophist$pop$params$N, type='l', xlab='deme number (space)', ylab='allele frequency' )
+abline(v=0.5 + which(diff(as.vector(pophist$pop$params$s))!=0), lty=2 )
+text( mean(which(as.vector(pophist$pop$params$s)>0)), .100, labels=as.expression(substitute(s==sb,list(sb=max(pophist$pop$params$s)))) )
+text( c(.05,.95)*dim(pophist$pophist)[2], .900, labels=as.expression(substitute(s==sb,list(sb=min(pophist$pop$params$s)))) )
+lines( rowMeans(pophist$pophist[,,2,2000:10000])/pophist$pop$params$N, lwd=2 )
+
+matplot( pophist$pophist[,,2,timeslice]/pophist$pop$params$N, type='l', xlab='deme number (space)', ylab='allele frequency', log='y' )
+abline(v=0.5 + which(diff(as.vector(pophist$pop$params$s))!=0), lty=2 )
+lines( rowMeans(pophist$pophist[,,2,2000:10000])/pophist$pop$params$N, lwd=2 )
+
+dev.off()
+
+# the exp'l decay
+patchloc <- pmax( min(which(pophist$pop$params$s>0))-seq_along(pophist$pop$params$s), seq_along(pophist$pop$params$s)-max(which(pophist$pop$params$s>0)) )
+
+plot( sort(unique(patchloc)), tapply(rowMeans(pophist$pophist[,,2,2000:10000])/pophist$pop$params$N,patchloc,mean), log='y' )
[…]

### 2013-07-23 sim-patchy-selection.R

@@ -1,8 +1,15 @@
 #!/usr/bin/R
 source("sim-patchy-selection-fns.R")
 source("lineages.R")
+source("sim-patchy-selection-fns.R")
 
 run.list <- list.files(".","*-pophistory-run.Rdata")
+if (!file.exists("run-info.csv")) {
+    rundims <- sapply( run.list, function (x) { tmp <- load(x); if ('pophist' %in% tmp) { dim(pophist$pophist) } else { NA } } )
+    write.table(rundims,file="run-info.csv",sep=',')
+} else {
+    rundims <- read.csv("run-info.csv")
+}
 
 # load random run
 rm(pophist); rm(lins)
@@ -43,17 +50,20 @@ plotlins(lins,add=TRUE,subset=1:100)
 
 # Exponential decay in mean occupation density
 plot( pophist$pop$params$patchdist, rowMeans(pophist$pophist[,,2,200:dim(pophist$pophist)[4],drop=FALSE],dims=2)/pophist$pop$params$N, log='y', col=adjustcolor("black",.25), pch=20 )
-with(pophist$pop$params, lines( seq(0,max(pophist$pop$params$patchdist),length.out=100), 1.5*exp(-(sqrt(2*abs(sm))/getsigma(pophist$pop$params))*seq(0,max(pophist$pop$params$patchdist),length.out=100)) ) )
-with(pophist$pop$params, lines( seq(0,max(pophist$pop$params$patchdist),length.out=100), 1.5*exp(-2*(sqrt(2*abs(sm))/getsigma(pophist$pop$params))*seq(0,max(pophist$pop$params$patchdist),length.out=100)), lty=2 ) )  # fudge factor of 2
-
+with(pophist$pop$params, 
+    lines( seq(0,max(pophist$pop$params$patchdist),length.out=100), 
+        1.5*exp(-(sqrt(2*abs(sm))/getsigma(pophist$pop$params))*seq(0,max(pophist$pop$params$patchdist),length.out=100)) ) )
+with(pophist$pop$params, 
+    lines( seq(0,max(pophist$pop$params$patchdist),length.out=100), 
+        1.5*exp(-2*(sqrt(2*abs(sm))/getsigma(pophist$pop$params))*seq(0,max(pophist$pop$params$patchdist),length.out=100)), lty=2 ) )  # fudge factor of 2 
 
 # Look at occupation times of farout stuff
[…]

### 2013-07-28 patchy-selection-paper.tex
more plots

@@
 
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics{sim-snapshots}
+  \end{center}
+  \label{fig:sim_snapshots}
+  \caption{
+  Results from individual-based simulation, on a one-dimensional grid:
+  the focal allele is beneficial in the central 20 demes, with $s_b=0.1$,
+  and deleterious elsewhere, with $s_m=-0.02$.
+  The dynamics were simulated for 10000 generations; demes had 1000 individuals in each;
+  dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
+  \textbf{(A)} Allele frequencies across space: colored lines are snapshots at an evenly spaced set of times,
+  and the solid lines shows the cumulative mean occupation frequency across all generations.
+  \textbf{(B)} A segment of the temporal dynamics: space is again shown across the horizontal axis;
+  with time on the vertical axis; demes are colored darker red the greater then allele frequency at the corresponding time.
+  The 50\% and 5\% frequency contour lines are in grey,
+  and the genealogy of 50 individuals in a rare long-distance excursion is traced back with black lines.
+  }
+\end{figure}
+
 Migration--selection balance ensures that 
@@
 to obtain an approximation that we argue is valid if the distance $r$ is large relative to $\sigma/\sqrt{s_p}$.
+
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics{sim-occupation-freqs}
+  \end{center}
+  \label{fig:sim_occupation_freqs}
+  \caption{
+  Mean occupation frequencies from simulation; details are as in figure~\ref{fig:sim_snapshots} except as otherwise noted.
+  Each panel shows mean occupation frequencies of individual demes (grey dots)
+  and averaged over demes at similar distances (black circles), on
+  \textbf{(A)} a one-dimensional grid of 150 demes, and
+  \textbf{(B)} a two-dimensional grid of $50\times 50$ demes.
+  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq} (with $C=1/2$).
+  The two distinct lines of grey points in (A) are to the left and right of the patch, respectively.
+  }
+\end{figure}
 

### 2013-07-28 lineages.R

@@ -15,6 +15,7 @@ lineages <- function ( pophist, nlin, migrsteps, T=dim(pophist)[4], m, linit=NUL
     #    with types either 0 (nonselected) or 1 (selected)
     # returns:
     #  LL = location of lineages through time (2 x T x nlin)
+    if (missing(migrsteps)) { migrsteps <- pophist$pop$params$migrsteps }
     if (missing(m)) { m <- pophist$pop$params$m }
     if ("pophist" %in% names(pophist)) { pophist <- pophist$pophist }
 
@@ -26,7 +27,7 @@ lineages <- function ( pophist, nlin, migrsteps, T=dim(pophist)[4], m, linit=NUL
         initpop <- pophist[,,2,T]
         dim(initpop) <- dim(pophist)[1:2]
         locs <- sample( nx*ny, nlin, replace=TRUE, prob=initpop )
-        linit <- cbind( row(initpop)[locs], col(initpop)[locs] )
+        linit <- cbind( col(initpop)[locs], row(initpop)[locs], T )
     }
 
     # add a zero step and reweight migrsteps
@@ -36,10 +37,11 @@ lineages <- function ( pophist, nlin, migrsteps, T=dim(pophist)[4], m, linit=NUL
     # the Lineage Locations
     #   and the Local Sizes (migration-weighted)
     #   and coalescent events
+    T <- linit[,3]
     LL <- array(NA, c(2,max(T),nrow(linit)))  # dimensions are: (x,y) , (time ago) , (lineage number)
     LS <- array(NA, c(1,max(T),nrow(linit)))  # dimensions are: (size) , (time ago) , (lineage number)
-    LL[cbind(1:2,rep(T,each=2),rep(1:nrow(linit),each=2))] <- t(linit)
-    LS[cbind(1,T,1:nrow(linit))] <- pophist[cbind(linit,2,T)]  # ?'[': in x[i], 'i' can be a matrix with as many columns as there are dimensions of ‘x’...
+    LL[cbind(1:2,rep(T,each=2),rep(1:nrow(linit),each=2))] <- t(linit[,2:1])
+    LS[cbind(1,T,1:nrow(linit))] <- pophist[cbind(linit[,2:1],2,T)]  # ?'[': in x[i], 'i' can be a matrix with as many columns as there are dimensions of ‘x’...
     dimnames(LL) <- list( c("y","x"), dimnames(pophist)[[4]][1:max(T)], NULL )  
[…]

### 2013-07-28 patchy-sim-writeup-plots.R

@@ -5,90 +5,188 @@ source("lineages.R")
 run.list <- list.files(".","*-pophistory-run.Rdata")
 rundims <- read.csv("run-info.csv")
 
-run.names <- c( '3975-1-10000-pophistory-run.Rdata', '5473-1-10000-pophistory-run.Rdata' )
+# run.name <- "3975-1-10000-pophistory-run.Rdata"
 
+for (run.name in run.list) {
+
+    run.id <- gsub( "([^-]*)-.*","\\1",run.name)
+    load(run.name)
+    print(c("run",run.id,"range",pophist$pop$params$range))
+
+    dimension <- sum(dim(pophist$pophist)[1:2]>1)
+    theory.decay <- dimension * sqrt(2*abs(getgrowth(pophist$pop$params)$gm) / getsigma(pophist$pop$params) )
+
+    # time slices of the process
+    timeslice <- floor( seq( .05*dim(pophist$pophist)[4], dim(pophist$pophist)[4], length.out=50 ) )
+
+    # the exp'l decay
+    patchloc <- with(pophist$pop$params, (-1)^(s>0) * sqrt( abs( ( row(s) - range[1]/2 )^2 + ( col(s) - range[2]/2 )^2 - (patchsize/2)^2 ) ) )
+    tmpdists <- seq(min(patchloc),max(patchloc),length.out=27)
+    tmplocs <- tmpdists[-1] - diff(tmpdists)/2
+    patchdist <- cut( as.vector(patchloc), breaks=tmpdists, include.lowest=TRUE, ordered_result=TRUE )
+    # equil <- getequilibrium(pophist,init=rowMeans(pophist$pophist[,,2,2000:10000,drop=FALSE])/pophist$pop$params$N) #,keep.convergence=TRUE)
+    if (is.null(pophist$occupation)) { pophist$occupation <- rowSums(pophist$pophist,dim=3) }
+
+    if (dimension==1) {
+        sliced <- pophist$pophist[,,2,timeslice]/pophist$pop$params$N
+        # if (length(dim(sliced))>2) { dim(sliced) <- c( prod(dim(sliced)[1:2]), dim(sliced)[3] ) }
+
+        pdf(file=paste(run.id,'example-equilibrium.pdf',sep='-'), width=6, height=6, pointsize=10)
+        par(mar=c(5,4,1,1)+.1)
+        layout(1:2)
+
[…]

### 2013-07-28 sim-patchy-selection-fns.R

@@ -243,10 +243,16 @@ colorRampTrans <- function (basecol, n=12) {
 }
 
 
-plotpophist <- function(pophist,timeslice=TRUE,...) {
+plotpophist <- function(pophist,timeslice=1:dim(pophist$pophist)[4],transposed=FALSE,...) {
     # Plot one-dimensional population history with time on the x-axis
     if (!any(dim(pophist$pophist)[1:2]==1)) { stop("Population is not one-dimensional.") } 
-    zimg <- plotpop( aperm(pophist$pophist[,,,timeslice,drop=TRUE],c(3,1,2)), params=pophist$pop$params, xlab="time", ylab="space", ... )
+    if (!transposed) {
+        popn <- aperm(pophist$pophist[,,,,drop=TRUE],c(3,1,2))
+        zimg <- plotpop( popn, params=pophist$pop$params, xlab="time (generations)", ylab="space (demes)", userows=timeslice, ... )
+    } else {
+        popn <- aperm(pophist$pophist[,,,,drop=TRUE],c(1,3,2))
+        zimg <- plotpop( popn, params=pophist$pop$params, xlab="space (demes)", ylab="time (generations)", usecols=timeslice, ... )
+    }
     axis(1); axis(2)
     return( invisible(zimg) )
 }
@@ -284,22 +290,28 @@ plotlinearpop <- function(n, thresh=.05, cols, ...) {
     }
 }
 
-plotpop <- function (n, params, maxtimes=200, thresh=.05, cols, coltrans, ...) {
+plotpop <- function (n, params, maxtimes=200, thresh=.05, cols, coltrans, userows=1:dim(n)[1], usecols=1:dim(n)[2], ...) {
     # Layer the colors on top of eachother (multiplicatively) 
     # for the different types
 
     if (missing(params)) { params <- n$params }
     if ("n" %in% names(n)) { n <- n$n }
     if (length(dim(n))==4 & dim(n)[4]==1) { dim(n) <- dim(n)[1:3] }
+    if (is.logical(userows)) { userows <- which(userows) }
+    if (is.logical(usecols)) { usecols <- which(usecols) }
 
     # subsample if too many columns
-    if ((dim(n)[1]>maxtimes)) { n <- n[(1:(dim(n)[1]))%%floor(dim(n)[1]/maxtimes)==0,,] }
[…]

### 2013-07-28 patchy-selection-paper.tex
.

@@
 \begin{align} \label{eqn:eqfreq}
-    q(r) \approx C \exp( -\sqrt{2 s_m} r / \sigma) \qquad \text{for large $r$},
-\end{align}
-where $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
+    q(r) \approx C \exp( - r \sqrt{2 s_m} / \sigma) \qquad \text{for large $r$},
+\end{align}
+where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
 In Appendix \ref{apx:patchy_sims} we argue that $C=1/2$ is about right for most parameter values,
@@
 so that these branching processes are rooted at $r_0$, and are killed upon reaching the original patch.
-This decomposition is illustrated in figure \ref{fig:sim_excursions}.
+This decomposition is illustrated in~\ref{fig:sim_snapshots}, and later in figure~\ref{fig:branching_decomp}.
 In terms of these families, the equilibrium frequency above is then
@@
   \begin{center}
-  \input{branching-concept.pdf_tex}
+    \includegraphics{branching-concept}
   \end{center}
-\caption{Depiction of the decomposition described in the text.
-In red is the central lineage (whose spatial motion we follow);
-in black are the surviving branches off of this lineage, 
-and in grey are offspring of the central lineage that did not survive until $t$.
-\plr{Label $W_s$ in this figure?}
-\plr{Add as part B a simulation figure.}
+\caption{Cartoon of the decomposition described in the text.
+Migrants families are emerging from the already-established patch (grey circle);
+each is composed of a ``trunk'' lineage (in red), whose spatial motion we follow,
+embellished by transient branches.
+For the depicted long-lived family, these are further distinguished by those still alive at time $t$ (in black)
+and those that did not survive until $t$ (in grey).
+Figure~\ref{fig:sim_snapshots} has a depiction of such a family, from a simulation.
 \label{fig:branching_decomp}
@@
     \end{center}
-    \label{fig:prob_estab_calcs}
+    \label{fig:example_equil}
     \caption{
@@
 either with rate 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
-Suppose that $R$ has mean zero and variance $\sigma^2$, and that $X$ has mean $\mu$ and variance $\xi^2$.
+Suppose that $X$ has mean $\mu$; the variance of $X$ will not enter into the formula.
+Also suppose that $R$ has mean zero and variance $\sigma^2$;
+if we are in more than one dimension, we mean that the components of the dispersal distance are uncorrelated
+and each have variance $\sigma^2$.
 We also assume that the distribution of $X$ makes the resulting Markov chain irreducible and aperiodic.
@@
 \begin{align}
-  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \partial_x^2 \xi(t,x) + \gamma(x) \xi(t,x) (1-\xi(t,x)) \right) .
-\end{align}
+  \label{eqn:derived_fkpp}
+  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \xi(t,x) + \gamma(x) \xi(t,x) (1-\xi(t,x)) \right) .
+\end{align}
+
+In one dimension, we are done (see section~\ref{apx:elliptic_integrals});
+in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
+If we take a radially symmetric patch centered at the origin (so $\gamma$ only depends on $r$), 
+and let $\xi(t,r)$ denote the mean occupation frequency at distance $r$,
+then the polar form of the Laplacian in $d$ dimensions gives us that \eqref{eqn:derived_fkpp} is
+\begin{align}
+  \label{eqn:radial_fkpp}
+  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + \gamma(r) \xi(t,r) (1-\xi(t,r)) \right) .
+\end{align}
+However, the second term goes to zero as $r \to \infty$,
+leading to the same asymptotic behavior (CITE SOMETHING).
 

### 2013-07-28 patchy-sim-writeup-plots.R

@@ -98,7 +98,9 @@ LL <- plotlins(lins,plotit=FALSE)
 Ltimes <- as.numeric(dimnames(LL)[[2]])
 useLtimes <- (Ltimes>min(timeslice)) & (Ltimes<max(timeslice))
 LL <- LL[,useLtimes,] 
-plotpophist(pophist,timeslice=timeslice,maxtimes=300,plotlegend=FALSE, cols=cols,transposed=TRUE)
+# plotpophist(pophist,timeslice=timeslice,maxtimes=300,plotlegend=FALSE, cols=cols,transposed=TRUE)
+image( x=1:dim(pophist$pophist)[2], y=seq(min(timeslice),max(timeslice),length.out=200), subph, col=colorRampTrans('red',n=32)[c(1,7:32)], xlab='space (demes)', ylab='time (generations)' )
+image( x=1:dim(pophist$pophist)[2], y=seq(min(timeslice),max(timeslice),length.out=200), subph>0, col=c(NA,adjustcolor("red",.1)), add=TRUE )
 contour( x=1:dim(pophist$pophist)[2], y=seq(min(timeslice),max(timeslice),length.out=200), subph, levels=c(.5,.05), add=TRUE, col=grey(.4) )
 invisible( lapply( 1:dim(LL)[3], function (k) { lines( LL[2,,k], Ltimes[useLtimes] ) } ) )
 

### 2013-08-28 patchy-selection-paper.tex
added phase diagram

@@
 \newcommand{\mutrate}{\lambda_\text{mut}}
+\newcommand{\Tmig}{T_\text{mig}}
+\newcommand{\Tmut}{T_\text{mut}}
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 
@@
 mutation is likely to fix in both patches.
+
+One way to visualize this is by a phase diagram.
+In figure \ref{fig:phase_diagram},
+we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
+as a function of population density, mutation rate, and distance between patches (scaled by $\sqrt{s_m}/\sigma$).
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{phase-diagram-log}
+  \end{center}
+  \caption{
+  Mean time until adaptation in a new patch occurs by migration ($\Tmig$, blue) or by mutation ($\Tmut$, red),
+  as a function of population density ($\rho$), mutation rate ($\mu$), 
+  and distance to an already adapted patch ($R$), scaled by $\sqrt{s_m}/\sigma$.
+  Note that population density and mutation rate are shown on a log scale.
+  \plr{fix axis labels}
+  \label{fig:phase_diagram}
+  }
+\end{figure}
 

### 2013-08-28 patchy-selection-paper.tex
.

@@
 
-%%%%%%%%%%%%
-\section{Integrals appearing in the text}
-    \label{apx:integrals}
-
-In the text at several points appear integrals of the form
-\begin{align}
-  \int_0^\infty t^c \exp \left( - \alpha t^d - \beta t^{d+1} \right) dt 
-\end{align}
-where $c$ is a positive integer, $d$ is the dimension, and $\alpha$ and $\beta$ are positive real numbers.
-This could be evaluated through standard numerical methods; below we describe a power series expansion.
-Changing variables to $u = \beta t^{d+1}$, this becomes
-\begin{align}
-    \left( (d+1)^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
-\end{align}
-so it suffices to evaluate the function
-\begin{equation}
-    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
-\end{equation}
-in the case that $a=(c-d+1)/(d+1)$, $b=d/(d+1)$, and $x$ is a function of the demographic paramters.
-Since $a$ and $b$ only depend on the dimension and the quantity being computed,
-we are interested in $G$ as a function of $x$.
-At least in the case $b=d/(d+1)$ it is possible to express $G$ as a finite sum of gamma functions,
-but we proceed with a simpler method.
-Note that $\partial_x G(a,b,x) = -G(a+b,b,x)$,
-and that at $x=0$, the function $G$ is the gamma function $G(a,b,0) = \Gamma(a+1)$.
-Therefore, a Taylor series for $G$ would be
-\[
-    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
-\]
-It is easy to check using Stirling's formula that $\limsup_{n \to \infty} ( x^n \Gamma(a+nb+1)/n! )^{1/n} = 0$
-if $b<1$, so the sum converges.
-
 %%%%%%%%%%

### 2013-09-12 patchy-selection-paper.tex
merged in changes

@@
 will be larger if these adaptive alleles are maladapted in intervening environments,
-since such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{slatkin1973geneflow,}.
+since such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{slatkin1973geneflow}.
 
@@
 relative to the patch fixes within the poor habitat patch.
-As shown in Appendix \ref{apx:elliptic_integrals},
+As shown in Appendix \ref{apx:eqfreq},
 the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
@@
 \section{The equilibrium frequency}
+\label{apx:eqfreq}
 
@@
 see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} for other arguments in this case, 
-or \citet{etheridge} or \citet{dawson} for general theory used below.
+or \citet{etheridge2000introduction} or \citet{dawson1993measurevalued} for general theory used below.
 Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
 and that selection (for or against) the allele is given by the function $s(x)$, with $x$ denoting the spatial location.
-Then each individual at location $x$ reproduces after a random lifetime,
+Each individual at location $x$ reproduces after a random lifetime,
 dying and giving birth in the same deme to a random number of offspring with distribution given by $X$;
-who all migrate to a new location chosen as $x+R$
-and replace randomly chosen individuals there.
+each offspring migrates independently to a new location chosen randomly from the distribution given by $x+R$,
+where they replaces a randomly chosen individual.
 If $x+R$ is outside of the range, then they perish.
 Each individual's reproduction time is exponentially distributed, 
-either with rate 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
-Suppose that $X$ has mean $\mu$; the variance of $X$ will not enter into the formula.
-Also suppose that $R$ has mean zero and variance $\sigma^2$;
+the reproduction rate is 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
+Suppose that the number of offspring $X$ has mean $\mu$; the variance of $X$ will not enter into the formula.
+Also suppose that the migration displacement $R$ has mean zero and variance $\sigma^2$;
 if we are in more than one dimension, we mean that the components of the dispersal distance are uncorrelated
 and each have variance $\sigma^2$.
-We also assume that the distribution of $X$ makes the resulting Markov chain irreducible and aperiodic.
-
-Let $\Phi^N_t(x)$ be the number of mutant alleles present at location $x$ at time $t$,
-and denote by $\delta_x$ a single unit at location $x$, so that e.g.~$\Phi^N_t + \delta_x/N$
+We also assume that the distribution of $X$ makes the associated random walk irreducible and aperiodic.
+
+Let $\Phi^N_t(x)$ be the proportion of mutant alleles present at location $x$ at time $t$,
+and $\Phi_t(x)$ the process obtained by taking $N \to \infty$ (which we assume exists).
+Denote by $\delta_x$ a single unit at location $x$, so that e.g.~$\Phi^N_t + \delta_x/N$
 is the configuration after a mutant allele has been added to location $x$.
-For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the number of mutant alleles added if $X$ new offspring carrying mutant alleles
+For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the random number of mutant alleles added if $X$ new offspring carrying mutant alleles
 replace randomly chosen individuals in a deme where the mutant allele is at frequncy $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
-similarly, $\tilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
+similarly, $\widetilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
 (We like to think of $\Phi^N_t$ as a measure, but it does not hurt to think of $\Phi^N$ as a vector;
 we aren't providing the rigorous justification here.)
-Then we know that for any sufficiently nice function $f(\phi)$ that
+Then we know that for any sufficiently nice function $f(\Phi)$ that
 \begin{align} \label{eqn:discrete_generator}
@@
      & \qquad  \qquad \left.
-{} + \E\left[ \left(1-\Phi^N_t(x+R)\right) \left( f\left(\Phi^N_t - \frac{\tilde X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right\}  \end{split} \\
+{} + \E\left[ \left(1-\Phi^N_t(x+R)\right) \left( f\left(\Phi^N_t - \frac{\widetilde X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right\}  \end{split} \\
      &= \mu \sum_x \E\left[ \left(\partial_{\phi(x)} f(\Phi_t) \right) \left\{ \Phi_t(x+R) - \Phi_t(x) + s(x+R) \Phi_t(x+R) (1-\Phi_t(x)) \right\} \right] + O\left(\frac{1}{N}\right).
 \end{align}
-This follows in part from $\E[\bar X_\phi] = \phi \mu$ and $\E[\tilde X_\phi] = (1-\phi)\mu$.
+In the final expectation, $R$ and $\Phi$ are independent.
+This follows by taking first-order terms in $1/N$ in the Taylor series for $f$, 
+and the fact that $\E[\bar X_\phi] = \phi \mu$ and $\E[\widetilde X_\phi] = (1-\phi)\mu$.
 We can see two things from this:
-First, since this is a first-order differential operator, the limiting stochastic process
-is deterministic (check by applying to $f(\phi) = \phi(x)^2$ to find the variance).
+First, since this is a first-order differential operator, the limiting stochastic process $\Phi$
+is in fact deterministic (check by applying to $f(\Phi) = \Phi(x)^2$ to find the variance).
 Second, if we want to rescale space as well to get the usual differential equation, 
@@
 this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale.
-More concretely, suppose that the grid size is $\epsilon \to 0$, that $\var[R] = \sigma^2 \epsilon$, and that $s(x)/\epsilon \to \gamma(x)$,
+More concretely, suppose that the grid size is $\epsilon \to 0$, 
+that $\var[R] = (\sigma \epsilon)^2$, and that $s(x)/\epsilon \to \gamma(x)$,
 and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
-then the previous equation converges to the familiar form:
+then the previous equation with $f(\Phi) = \Phi(x)$ converges to the familiar form:
 \begin{align}
@@
 \end{align}
+Here we have taken first the population size $N \to \infty$ and then the grid size $\epsilon \to 0$;
+we could alternatively take both limits together, but not if $\epsilon$ goes to zero too much faster than $N$ grows.
+One reason for this is that at finite $N$,
+the process $\Phi_t$ is an irreducible, aperiodic finite-state Markov chain with absorbing states at 0 and 1;
+therefore, the inevitable outcome is extinction of one type or another,
+which is not the regime we want to study.
 
@@
 However, the second term goes to zero as $r \to \infty$,
-leading to the same asymptotic behavior (CITE SOMETHING).
-
-The process $\Phi_t$ is an irreducible, aperiodic finite-state Markov chain with absorbing states at 0 and 1;
-therefore, the inevitable outcome for finite $N$ is extinction of one type or another.
-Setting the left-hand side to zero in \eqref{eqn:discrete_generator} gives us equations for the moments of the stationary distribution,
-for instance, if $\phi(x) = \lim_{t \to \infty} \E[ \Phi_t(x) ]$
+leading to the same asymptotic behavior for the tail \plr{CITE SOMETHING}.
 
@@
 \section{The exact solution to the equilibrium frequency}
-\label{apx:elliptic_integrals},
-
-Or, in other words, solving the Fisher-KPP equation with varying selection.
+\label{apx:elliptic_integrals}
+
+Here we describe how to solve \eqref{eqn:derived_fkpp} in one dimension,
+i.e.\ the Fisher-KPP equation with varying selection.
 Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
@@
 \[
-\partial_t p(t,x) = \frac{1}{2} \sigma^2 \partial_x^2 p(t,x) + s(x) p(t,x) (1-p(t,x)) .
+\partial_t p(t,x) = \frac{\mu}{2} \sigma^2 \partial_x^2 p(t,x) + \mu s(x) p(t,x) (1-p(t,x)) .
 \]
@@
     ( \partial_x \phi(x) )^2  &= - \int^{x} 2 s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
-        &= - s_i \phi(x)^2 \left( 1 - \frac{2}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) \\
-        &:= - V_i(\phi(x)) \quad \mbox{for } x \in [x_i,x_{i+1}) ,
-\end{align}
-where $K_i$ is a constant of integration and we define
+        &= - s_i \phi(x)^2 \left( 1 - \frac{2}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) ,
+\end{align}
+if we define
+\begin{align} 
+  K_i = - s_i \phi(x_i)^2 \left( 1 - \frac{2}{3} \phi(x_i) \right) - ( \partial_x \phi(x_i) )^2  .
+\end{align}
+For ease of reference, we define
 \[
@@
 We will always have that $V(\phi) \le 0$.
-(We have then that $\phi'' = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
+(We have then that $\partial_x^2 \phi = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
 Also note that this implies ``conservation of energy'', i.e.\ $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
@@
 \[
-    x(\phi) = x(\phi_0) \pm \int_{\phi_0}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } .
+    x(\phi) = x(\phi^*) \pm \int_{\phi^*}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } .
 \]
 For each $i$ then define the elliptic function
-\[
-    F_i(\phi) = \int_{\phi_i^*}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } ,
-\]
+\begin{align}  \label{eqn:elliptic_function}
+    F_i(\phi) = \int_{\phi_i^*}^\phi \frac{ d\psi }{ \sqrt{ -V_i(\psi) } } ,
+\end{align}
 where take the positive branch of the square root, and $\phi_i^*$ will be chosen later.
@@
 \]
-What about $\phi(x_i)$?  Well, if $\phi(x)$ is monotone on $[x_i,x_{i+1})$ then we can wolog take $\phi_i^*=0$ or $1$ depending on the sign of $s_i$.
+What about the frequency at the points the selection changes, $\phi(x_i)$?  
+Well, if $\phi(x)$ is monotone on $[x_i,x_{i+1})$ then we can without loss of generality take $\phi_i^*=0$ or $1$ depending on the sign of $s_i$.
 Otherwise, let $\phi_i^*$ be the (unique) root of $V_i$ in $[x_i,x_{i+1})$, so $V_i(\phi_i^*)=0$.
@@
 Recall we defined $F_i$ using $\phi_i^*$; now using the fact that $\phi$ is monotone with the opposite sign on either side of $\phi_i^*$,
-$x_{i+1} - x_i = F(\phi(x_{i+1})) - F(\phi_i^*) + F(\phi(x_{i})) - F(\phi_i^*)$,
+$x_{i+1} - x_i = F_i(\phi(x_{i+1})) - F_i(\phi_i^*) + F_i(\phi(x_{i})) - F_i(\phi_i^*)$,
 and that $F(\phi_i^*) = 0$,
@@
 \]
-
 Note that all the $\pm{}$'s are easily relatable to the signs of $s_i$.
-If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be all set.
-Unfortunately, we are given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
-Monty (in "Gene flow and selection in a cline") deals with this by using symmetry, so he knows $\phi(0)$ and $\phi'(0)$.
+If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be able to solve the equations for $\phi(x_i)$ and $K_i$ recursively upwards.
+In some cases, such as \citet{slatkin1973geneflow}, we can infer $\phi(0)$ and $\partial_x \phi(0)$ by spatial symmetry.
+In other cases, we are only given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
 
@@
 \]
-We can solve the case with $s_* = s_d$, since $\phi'(0) = 0$ (as can Monty, although he doesn't say what it is).
+We can solve the case with $s_* = s_d$, since $\phi'(0) = 0$ (as done implicitly in \citet{slatkin1973geneflow}).
 We should be able to argue that if $R$ is large then the value of $s_*$ won't make much difference.
@@
 
-This from ``NEQwiki, the nonlinear equations encyclopedia'', \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
-Recall that $V_i(\phi) =  s_i \phi^2(1-2\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries.
-Note that if we rescale space by $\sqrt{s}/\sigma$,
-then $s$ and $\sigma$ drop from the equation.
-We want to integrate
+An important ingredient in the above method is the integral \eqref{eqn:elliptic_function},
+to which a method for solving the Korteweg-de Vries equation can be applied \citep{NEQwiki}.
+Recall that $V_i(\phi) =  s_i \phi^2(1-2\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries;
+since we're just working within an interval with $s$ constant, we can rescale space by $\sqrt{2 s_i}/\sigma$,
+so that $s$ and $\sigma$ drop from the equation.
+We then want to integrate
 \[
     \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
-         \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } .
+         \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } ,
 \]
+over a domain where $V$ is always negative.
 Let $\phi^2(1-2\phi/3)+K = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
@@
 with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
-Now Jacobi's ``incomplete elliptic integral of the first kind'' is
+Now Jacobi's incomplete elliptic integral of the first kind is defined by
 \[
@@
 and the Jacobian elliptic function $\sn(x;k)$ is the inverse: $F(\sn(x;k);k) = x$.
-
 As $k \to 0$, $\sn(x;k) \to \sin(x)$, while as $k \to 1$, $\sn(x;k) \to \sinh(x)$.
@@
 \begin{align*}
-    \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / \sigma + C ) \right) \\
-        &\simeq C' \exp\left( - \frac{ 2 x \sqrt{2s} }{ \sigma } \right) .
+    \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / 2 \sigma + C ) \right) \\
+        &\simeq C' \exp\left( - \frac{ x \sqrt{2s} }{ \sigma } \right) .
 \end{align*}

### 2013-09-12 standing-variation-paper.tex
merged in changes

@@
 
+\bibliography{standing_patches_refs}
+
+\appendix
+
+%%%%%%%%%%%%
+\section{Integrals appearing in the text}
+    \label{apx:integrals}
+
+In the text at several points appear integrals of the form
+\begin{align}
+  \int_0^\infty t^c \exp \left( - \alpha t^d - \beta t^{d+1} \right) dt 
+\end{align}
+where $c$ is a positive integer, $d$ is the dimension, and $\alpha$ and $\beta$ are positive real numbers.
+This could be evaluated through standard numerical methods; below we describe a power series expansion.
+Changing variables to $u = \beta t^{d+1}$, this becomes
+\begin{align}
+    \left( (d+1)^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
+\end{align}
+so it suffices to evaluate the function
+\begin{equation}
+    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
+\end{equation}
+in the case that $a=(c-d+1)/(d+1)$, $b=d/(d+1)$, and $x$ is a function of the demographic paramters.
+Since $a$ and $b$ only depend on the dimension and the quantity being computed,
+we are interested in $G$ as a function of $x$.
+At least in the case $b=d/(d+1)$ it is possible to express $G$ as a finite sum of gamma functions,
+but we proceed with a simpler method.
+Note that $\partial_x G(a,b,x) = -G(a+b,b,x)$,
+and that at $x=0$, the function $G$ is the gamma function $G(a,b,0) = \Gamma(a+1)$.
+Therefore, a Taylor series for $G$ would be
+\[
+    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
+\]
+It is easy to check using Stirling's formula that $\limsup_{n \to \infty} ( x^n \Gamma(a+nb+1)/n! )^{1/n} = 0$
+if $b<1$, so the sum converges.
+
 \end{document}

### 2013-11-20 patchy-selection-paper.tex
.

@@
 
-Consider a population spread across a landscaper, 
+Consider a population spread across a landscape, 
 within which are patches of habitat to which individuals are (initially) poorly adapted,
@@
 We will make use of several previous results from the literature. 
-\citet{slatkin1973geneflow}, working in a one dimensional, deterministic framework, 
+\citet{haldane1948theory}, working with the one dimensional partial differential equation,
+showed that with large patches, the spatial scale over which the two types intergrade is given by $\sigma/\sqrt{s}$,
+with $s$ either $s_m$ or $s_p$ as appropriate.
+\citet{slatkin1973geneflow}, in the same setting,
 showed that if the physical width of the patch is less than $2 \tan^{-1} (\sqrt{s_m/s_p})$, 
@@
   \end{center}
-  \label{fig:sim_snapshots}
   \caption{
@@
   and the genealogy of 50 individuals in a rare long-distance excursion is traced back with black lines.
+  \label{fig:sim_snapshots}
   }
@@
 and the quantity we wish to compute,
-the effective rate of at which \emph{families of} migrant $B$ alleles establish in a new patch at distance $r$, 
+the effective rate of at which \emph{families} of migrant $B$ alleles establish in a new patch at distance $r$, 
 is
@@
 All terms except the ``outflux of families'' are simple calculations with spatial branching processes;
-and we can solve for this outflux using the established form for \eqref{eqn:gestalt_q}.
+and we can solve for this outflux using equation \eqref{eqn:eqfreq}.
 Here $\migrate$ reflects the mean number of establishing migrant families per unit time,
@@
 and then pruning any branches that happen to wander into a patch.
-This pruning removes both branches leading back into the already established patch (which are irrelevant)
-and branches leading into the new patch (we will need to count these).
-Since we are concerned with the rare migrant families that transit quickly between patches, 
-backmigrants to the already established patch will contribute little, since they are going the wrong direction;
-so to simplify the analysis we ignore this detail.  % but could add this detail in an appendix
+This pruning removes both 
+branches leading into the new patch (we will need to count these),
+and branches leading back into the already established patch 
+(which contribute little, since they are going the wrong direction;
+so we ignore these).  % but could add detail in appendix
 
@@
 In other words, the chance of survival to time $t$ is a constant multiplied by $(1-s_m)^t$,
-and conditional on survival, the family size is fixed.
+and conditional on survival, the family size is of fixed magnitude.
 
@@
 In other words, the genealogy of a family that survives until $t$
-can be decomposed into a ``trunk'' that lasts until $t$
-sprouting independent ``branches'' that may or may not survive until $t$.
+can be decomposed into a ``trunk'' that 
+sprouts independent ``branches'' that may or may not survive until $t$.
 This process is depicted in Figure~\ref{fig:branching_decomp}A.
@@
 
-This suggests that we add in the spatial motion by following only the motion of the ``trunk''
+This suggests that we will lose little by following only the motion of the ``trunk''
 (i.e.\ the red line in figure \ref{fig:branching_decomp}),
-whose motion is is approximately Brownian, with variance $\sigma^2$ per generation.
-(Recall $\sigma$ is the dispersal distance.)
+and adding the other $K-1$ surviving family members as a cluster about this trunk.
+The motion is is approximately Brownian, with variance $\sigma^2$ per generation,
+with $\sigma$ is the dispersal distance.
 Suppose that the new patch is at distance $r$.
@@
 
-The probability of establishment of an outmigrant from the already-established patch is 
+In summary, the probability that an outmigrant from the already-established patch establishes in a new patch at distance $r$ away is 
 equal to the probably the trunk arrives (equation~\eqref{eqn:estab_prob}) multiplied by the probability it establishes (equation~\eqref{eqn:prob_estab_arrival}),
@@
 \begin{align} \label{eqn:bp_prob_estab}
-    F(r) &:= \text{( prob family establishes at $r$ )} = \frac{2s_p}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} r}{\sigma} \right) ,
+    F(r) &:= \text{( prob family establishes at $r$ )} = \frac{2s_p}{\xi^2} \exp\left( - \frac{r \sqrt{2 \alpha}}{\sigma} \right) ,
 \end{align}
@@
 Define $\gamma$ to be this outflux; 
-i.e.\ the mean number of individuals per unit time that are born carrying the $B$ allele closer than $r_0$ to the original patch
+i.e.\ the mean number of individuals per unit time whose ancestors carrying the $B$ allele all lived closer than $r_0$ to the original patch
 but migrate to a larger distance.
@@
 \begin{align}
-    u(x) &= \int_0^t \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
-    &= \int_0^t e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
+    u(x) &= \int_0^\infty \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
+    &= \int_0^\infty e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
     &= \frac{ 1 }{ \sigma \sqrt{2 \alpha} } \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) 

### 2013-11-20 Spatial_adaptation/maize_calcs.R

@@ -1,3 +1,150 @@
+#######
+# Find prob of survival with nonconstant selection
+
+iterroot <- function (genfn,tol=sqrt(.Machine$double.eps),maxiter=1000,...) {
+    p0 <- genfn(1/2,...)
+    for (k in 1:maxiter) {
+        p <- genfn(p0,...)
+        if (max(abs(p0-p))<tol) { break } else { p0 <- p }
+    }
+    attr(p,"niter") <- k
+    attr(p,"converged") <- (k<maxiter)
+    return(p)
+}
+
+dx <- 15            # distance between demes (km)
+rangesize <- 1000   # half-length of range (km)
+xx <- seq(-rangesize,rangesize,by=dx)  # deme locations (km)
+sb <- .1
+sd <- .00001
+elevends <- c(500,2500)     # endpoints of elevation (m)
+clinewidth <- 500           # a much wider cline
+clinewidth <- 26            # altitudinal cline width (km): S.Am
+clinewidth <- 62            # altitudinal cline width (km): Mexico
+alpha <- (sb+sd)/clinewidth      # slope of selection with altitude (units of s per km)
+selfn <- function (x) { pmin( sb, pmax( -sd, alpha*x ) ) }
+selx <- selfn(xx)
+
+# binary branching
+m <- .5
+eqvals <- c( 1, iterroot( function (p) (1/2)*(1-sb) + (1/2)*(1+sb)*p^2 ) )
+genfn <- function (f,selx,migr=m) { (1/2)*(1-selx) + (1/2)*(1+selx)*f*( f + migr*diff(c(eqvals[1],f,eqvals[2]),differences=2) ) }
+# Poisson
+m <- .5
+eqvals <- c( 1, iterroot( function (p) exp((1+sb)*(p-1)) ) )
+genfn <- function (f,selx,migr=m) { exp((1+selx)*(f+migr*diff(c(eqvals[1],f,eqvals[2]),differences=2)-1)) }
+
+# maize
+N <- 103359  # plants per field
+Nf <- 561    # ears to make next generation
+pe <- 0.068  # prob of total seed stock replacement
+mg <- .0083  # pollen migration
+psig <- 0    # pollen migration distance
[…]

### 2013-11-25 patchy-selection-paper.tex
.

@@
 
-In one dimension, we are done (see section~\ref{apx:elliptic_integrals});
-in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
-If we take a radially symmetric patch centered at the origin (so $\gamma$ only depends on $r$), 
-and let $\xi(t,r)$ denote the mean occupation frequency at distance $r$,
-then the polar form of the Laplacian in $d$ dimensions gives us that \eqref{eqn:derived_fkpp} is
-\begin{align}
-  \label{eqn:radial_fkpp}
-  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + \gamma(r) \xi(t,r) (1-\xi(t,r)) \right) .
-\end{align}
-However, the second term goes to zero as $r \to \infty$,
-leading to the same asymptotic behavior for the tail \plr{CITE SOMETHING}.
+We are interested in the equilibrium frequency,
+i.e.\ the mean frequency of the selected allele at migration-selection equilibrium.
+The frequency $\phi(x)$ at location $x$ solves
+\begin{align}
+  \label{eqn:general_phi}
+  \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \phi(x) + \gamma(x) \phi(x) (1-\phi(x)) = 0 
+\end{align}
 
@@
 %%%%%%%%%%
+\section{Asymptotics for the equilibrium frequency}
+\label{ss:asymptotic_freq}
+
+We provide the above discussion on exact solutions mostly for completeness;
+in practice, we use asymptotics of the solutions.
+These are furthermore available for $d=2$, where we have no analytic solutions.
+In $d=2$ (and higher) dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
+Suppose therefore that the patch is radially symmetric,
+and that the allele is uniformly deleterious outside of some radius $r_0$
+(in the notation above, $\gamma(x) = -s_m$ for $|x|>r_0$).
+Then, denoting by $\phi(r)$ the mean occupation frequency at distance $r$ at equilibrium,
+the polar form of the Laplacian in $d$ dimensions translates equation \eqref{eqn:general_phi} to
+\begin{align}
+  \label{eqn:radial_phi}
+  \frac{\sigma^2}{2} \partial_{r}^2 \phi(r) + \sigma^2\frac{d-1}{2r} \partial_r \phi(r) &= s_m \phi(r) (1-\phi(r)) \quad r>r_0 , \\
+  \phi(r_0) = \phi_0 , \qquad \phi(r) \ge 0 .
+\end{align}
+We wish to find the asymptotic behavior of $\phi$ for large $r$.
+The argument can likely be extended to patches that are not radially symmetric.
+
+The solution with $d=2$ is $C Y_0(ir) \sim C e^{r}/\sqrt{2 \pi r}$ (G\&R 8.451.2).
+
+%%%%%%%%%%
 \section{The probability of establishment}

### 2014-01-31 patchy-selection-paper.tex
additional citation
.

@@
 We will make use of several previous results from the literature. 
+XXX cite nagylaki 1974: does 2d clines XXX
 \citet{haldane1948theory}, working with the one dimensional partial differential equation,
@@
 \citet{slatkin1973geneflow}, in the same setting,
-showed that if the physical width of the patch is less than $2 \tan^{-1} (\sqrt{s_m/s_p})$, 
+showed that if the physical width of the patch is less than $2 (\sigma/\sqrt{s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
 then there is no stable equilibrium with both $b$ and $B$ present,

### 2014-01-31 patchy-selection-paper.tex
.

@@
 \citet{slatkin1973geneflow}, in the same setting,
-showed that if the physical width of the patch is less than $2 (\sigma/\sqrt{s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
+showed that if the physical width of the patch is less than $2 (\sigma/\sqrt{s_m}) \tan^{-1} (\sqrt{s_m/s_p})$, 
 then there is no stable equilibrium with both $b$ and $B$ present,

### 2014-02-02 patchy-selection-paper.tex
.

@@
 we will assume that our patch is larger than the cutoff for local establishment, 
-i.e.\ wider than $2 \tan^{-1} (\sqrt{s_m/s_p})$ \citep{slatkin1973geneflow}.
+i.e.\ wider than XXX $\sigma/\sqrt{s_m}$ times ? XXX $2 \tan^{-1} (\sqrt{s_m/s_p})$ \citep{slatkin1973geneflow}.
 
@@
   \textbf{(B)} A segment of the temporal dynamics: space is again shown across the horizontal axis;
-  with time on the vertical axis; demes are colored darker red the greater then allele frequency at the corresponding time.
-  The 50\% and 5\% frequency contour lines are in grey,
+  with time on the vertical axis; demes colored darker have larger allele frequency.
+  The 50\% and 5\% frequency contour lines for the allele are in grey,
   and the genealogy of 50 individuals in a rare long-distance excursion is traced back with black lines.
@@
 Denote the expected frequency of allele $B$ at distance $r$ from the patch by $q(r)$.
-Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(r)$ is the solution,
-which can then be solved to give $q(r)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
-Asymptotics of these implies that for large $r$ the mean occupation frequency decays exponentially, 
+Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(r)$ is the solution.
+In one dimension, this equation can be solved to give $q(r)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}),
+but in a fairly cumbersome form.
+We will only be concerned with the large $r$ asymptotics,
+which can be obtained by linearization of the differential equation in any dimension.
+The result, discussed in Appendix \ref{apx:asymptotic_freq}, is that the expected frequency decays exponentially,
+with a polynomial correction in $d=2$:
 \begin{align} \label{eqn:eqfreq}
-    q(r) \approx C \exp( - r \sqrt{2 s_m} / \sigma) \qquad \text{for large $r$},
-\end{align}
-where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
-In Appendix \ref{apx:patchy_sims} we argue that $C=1/2$ is about right for most parameter values,
-and also that these asymptotics are correct in both one and two dimensions
-(the differential equation is only analytically solvable in one dimension).
+  q(r) \approx C \left( \frac{ r \sqrt{2 s_m} }{ \sigma } \right)^{-(d-1)/2} \exp\left( - \frac{ r \sqrt{2 s_m} }{ \sigma } \right) + O(1/r)  ,
+\end{align}
+where $d$ is the dimension, equal to 1 or 2, 
+and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
+In Appendix \ref{apx:patchy_sims} we argue that $C=1/2$ is about right for most parameter values.
 
@@
   Note that population density and mutation rate are shown on a log scale.
-  \plr{fix axis labels}
+  \plr{fix axis labels; add $\Tmig=\Tmut$ line}
   \label{fig:phase_diagram}
@@
 the reproduction rate is 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
-Suppose that the number of offspring $X$ has mean $\mu$; the variance of $X$ will not enter into the formula.
+Suppose that the number of offspring $X$ has mean $\eta$; the variance of $X$ will not enter into the formula.
 Also suppose that the migration displacement $R$ has mean zero and variance $\sigma^2$;
@@
 {} + \E\left[ \left(1-\Phi^N_t(x+R)\right) \left( f\left(\Phi^N_t - \frac{\widetilde X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right\}  \end{split} \\
-     &= \mu \sum_x \E\left[ \left(\partial_{\phi(x)} f(\Phi_t) \right) \left\{ \Phi_t(x+R) - \Phi_t(x) + s(x+R) \Phi_t(x+R) (1-\Phi_t(x)) \right\} \right] + O\left(\frac{1}{N}\right).
+     &= \eta \sum_x \E\left[ \left(\partial_{\phi(x)} f(\Phi_t) \right) \left\{ \Phi_t(x+R) - \Phi_t(x) + s(x+R) \Phi_t(x+R) (1-\Phi_t(x)) \right\} \right] + O\left(\frac{1}{N}\right).
 \end{align}
@@
 This follows by taking first-order terms in $1/N$ in the Taylor series for $f$, 
-and the fact that $\E[\bar X_\phi] = \phi \mu$ and $\E[\widetilde X_\phi] = (1-\phi)\mu$.
+and the fact that $\E[\bar X_\phi] = \phi \eta$ and $\E[\widetilde X_\phi] = (1-\phi)\eta$.
 We can see two things from this:
@@
 this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale.
-More concretely, suppose that the grid size is $\epsilon \to 0$, 
-that $\var[R] = (\sigma \epsilon)^2$, and that $s(x)/\epsilon \to \gamma(x)$,
-and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
-then the previous equation with $f(\Phi) = \Phi(x)$ converges to the familiar form:
+Somewhat informally, if we assume we can Taylor expand $\Phi$ and $s$ in $x$, then
+setting $f(\Phi) = \Phi(x)$,
 \begin{align} 
   \label{eqn:derived_fkpp}
-  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \xi(t,x) + \gamma(x) \xi(t,x) (1-\xi(t,x)) \right) .
+  \frac{\partial}{\partial t} \Phi_t(x)
+  &= \eta \left( \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \Phi_t(x) + s(x) \Phi_t(x) \left(1-\Phi_t(x)\right) \right) + O(\E[R^4]) .
+\end{align} 
+More formally, suppose that the grid size is $\epsilon \to 0$, 
+that $\var[R]/\epsilon^2 \to \sigma^2$, and that $s(x)/\epsilon \to \gamma(x)$,
+that $\Phi_t(x)$ is sufficiently differentiable,
+and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
+then equation \eqref{eqn:discrete_generator} converges to the familiar form:
+\begin{align}
+  \partial_t \xi(t,x) = \eta \left( \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \xi(t,x) + \gamma(x) \xi(t,x) (1-\xi(t,x)) \right) .
 \end{align}
@@
 i.e.\ the mean frequency of the selected allele at migration-selection equilibrium.
-The frequency $\phi(x)$ at location $x$ solves
+Setting the left side of \eqref{eqn:derived_fkpp} to zero, the frequency $\phi(x)$ at location $x$ solves
 \begin{align}
   \label{eqn:general_phi}
-  \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \phi(x) + \gamma(x) \phi(x) (1-\phi(x)) = 0 
-\end{align}
+  \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \phi(x) + s(x) \phi(x) (1-\phi(x)) = 0 ,
+\end{align}
+which we study in subsequent sections.
 
 %%%%%%%
-\section{The exact solution to the equilibrium frequency}
+\section{The exact solution to the equilibrium frequency in one dimension}
 \label{apx:elliptic_integrals}
@@
 \[
-\partial_t p(t,x) = \frac{\mu}{2} \sigma^2 \partial_x^2 p(t,x) + \mu s(x) p(t,x) (1-p(t,x)) .
+\partial_t p(t,x) = \frac{\eta}{2} \sigma^2 \partial_x^2 p(t,x) + \eta s(x) p(t,x) (1-p(t,x)) .
 \]
@@
 \section{Asymptotics for the equilibrium frequency}
-\label{ss:asymptotic_freq}
+\label{apx:asymptotic_freq}
+
+REFER TO WASOW
 
@@
 \label{ss:discretedemes}
+
+\plr{NOTE:} could do the random walk on a lattice with characteristic spacing
 

### 2014-03-07 patchy-selection-paper.tex
.

--- 2014-01-31 patchy-selection-paper.tex
@@
 
-Consider a population spread across a landscape, 
+Consider a population spread across a landscaper, 
 within which are patches of habitat to which individuals are (initially) poorly adapted,
@@
 We will make use of several previous results from the literature. 
-XXX cite nagylaki 1974: does 2d clines XXX
-\citet{haldane1948theory}, working with the one dimensional partial differential equation,
-showed that with large patches, the spatial scale over which the two types intergrade is given by $\sigma/\sqrt{s}$,
-with $s$ either $s_m$ or $s_p$ as appropriate.
-\citet{slatkin1973geneflow}, in the same setting,
-showed that if the physical width of the patch is less than $2 (\sigma/\sqrt{s_m}) \tan^{-1} (\sqrt{s_m/s_p})$, 
-then there is no stable equilibrium with both $b$ and $B$ present,
+Haldane, Fisher, and later \citet{slatkin1973geneflow} first analyzed models of this sort.
+\citet{nagylaki1975conditions} and \citet{conley1975application} showed that in one dimension, if the physical width of the patch is less than $(\sigma/\sqrt{2s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
+then there is no stable equilibrium with both $b$ and $B$ present,  % see p.603 in Nagylaki
 so that migrational swamping prevents $B$ from establishing \citep[see also][ for a review]{lenormand2002limits}.
-\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed (again in one dimension) 
-that this critical patch size also holds in a stochastic framework.
+\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed  
+that patches must be at least this critical size so that a new mutation has positive probability of establishment
+(in the large population density limit).
 \citet{barton1987establishment} also showed that
-for patches above this critical size,
-the probability of establishment of a new mutant that appears at distance $r$ from the patch
-decays exponentially with distance from the patch, with the scale given by $\sigma/\sqrt{s_m}$,
+this probability of establishment of a new mutation decays exponentially with distance from the patch, 
+with the scale given by $\sigma/\sqrt{s_m}$,
 and that mutations appearing within the patch have probability of establishment
-less than the probability for a panimictic population,
+strictly less than the probability for a panimictic population,
 but approaching this as the size of the patch increases.
@@
   \end{center}
+  \label{fig:sim_snapshots}
   \caption{
@@
   and the genealogy of 50 individuals in a rare long-distance excursion is traced back with black lines.
-  \label{fig:sim_snapshots}
   }
@@
 and the quantity we wish to compute,
-the effective rate of at which \emph{families} of migrant $B$ alleles establish in a new patch at distance $r$, 
+the effective rate of at which \emph{families of} migrant $B$ alleles establish in a new patch at distance $r$, 
 is
@@
 All terms except the ``outflux of families'' are simple calculations with spatial branching processes;
-and we can solve for this outflux using equation \eqref{eqn:eqfreq}.
+and we can solve for this outflux using the established form for \eqref{eqn:gestalt_q}.
 Here $\migrate$ reflects the mean number of establishing migrant families per unit time,
@@
 and then pruning any branches that happen to wander into a patch.
-This pruning removes both 
-branches leading into the new patch (we will need to count these),
-and branches leading back into the already established patch 
-(which contribute little, since they are going the wrong direction;
-so we ignore these).  % but could add detail in appendix
+This pruning removes both branches leading back into the already established patch (which are irrelevant)
+and branches leading into the new patch (we will need to count these).
+Since we are concerned with the rare migrant families that transit quickly between patches, 
+backmigrants to the already established patch will contribute little, since they are going the wrong direction;
+so to simplify the analysis we ignore this detail.  % but could add this detail in an appendix
 
@@
 In other words, the chance of survival to time $t$ is a constant multiplied by $(1-s_m)^t$,
-and conditional on survival, the family size is of fixed magnitude.
+and conditional on survival, the family size is fixed.
 
@@
 In other words, the genealogy of a family that survives until $t$
-can be decomposed into a ``trunk'' that 
-sprouts independent ``branches'' that may or may not survive until $t$.
+can be decomposed into a ``trunk'' that lasts until $t$
+sprouting independent ``branches'' that may or may not survive until $t$.
 This process is depicted in Figure~\ref{fig:branching_decomp}A.
@@
 
-This suggests that we will lose little by following only the motion of the ``trunk''
+This suggests that we add in the spatial motion by following only the motion of the ``trunk''
 (i.e.\ the red line in figure \ref{fig:branching_decomp}),
-and adding the other $K-1$ surviving family members as a cluster about this trunk.
-The motion is is approximately Brownian, with variance $\sigma^2$ per generation,
-with $\sigma$ is the dispersal distance.
+whose motion is is approximately Brownian, with variance $\sigma^2$ per generation.
+(Recall $\sigma$ is the dispersal distance.)
 Suppose that the new patch is at distance $r$.
@@
 
-In summary, the probability that an outmigrant from the already-established patch establishes in a new patch at distance $r$ away is 
+The probability of establishment of an outmigrant from the already-established patch is 
 equal to the probably the trunk arrives (equation~\eqref{eqn:estab_prob}) multiplied by the probability it establishes (equation~\eqref{eqn:prob_estab_arrival}),
@@
 \begin{align} \label{eqn:bp_prob_estab}
-    F(r) &:= \text{( prob family establishes at $r$ )} = \frac{2s_p}{\xi^2} \exp\left( - \frac{r \sqrt{2 \alpha}}{\sigma} \right) ,
+    F(r) &:= \text{( prob family establishes at $r$ )} = \frac{2s_p}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} r}{\sigma} \right) ,
 \end{align}
@@
 Define $\gamma$ to be this outflux; 
-i.e.\ the mean number of individuals per unit time whose ancestors carrying the $B$ allele all lived closer than $r_0$ to the original patch
+i.e.\ the mean number of individuals per unit time that are born carrying the $B$ allele closer than $r_0$ to the original patch
 but migrate to a larger distance.
@@
 \begin{align}
-    u(x) &= \int_0^\infty \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
-    &= \int_0^\infty e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
+    u(x) &= \int_0^t \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
+    &= \int_0^t e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
     &= \frac{ 1 }{ \sigma \sqrt{2 \alpha} } \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) 
@@
 
-We are interested in the equilibrium frequency,
-i.e.\ the mean frequency of the selected allele at migration-selection equilibrium.
-The frequency $\phi(x)$ at location $x$ solves
-\begin{align}
-  \label{eqn:general_phi}
-  \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \phi(x) + \gamma(x) \phi(x) (1-\phi(x)) = 0 
-\end{align}
+In one dimension, we are done (see section~\ref{apx:elliptic_integrals});
+in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
+If we take a radially symmetric patch centered at the origin (so $\gamma$ only depends on $r$), 
+and let $\xi(t,r)$ denote the mean occupation frequency at distance $r$,
+then the polar form of the Laplacian in $d$ dimensions gives us that \eqref{eqn:derived_fkpp} is
+\begin{align}
+  \label{eqn:radial_fkpp}
+  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + \gamma(r) \xi(t,r) (1-\xi(t,r)) \right) .
+\end{align}
+However, the second term goes to zero as $r \to \infty$,
+leading to the same asymptotic behavior for the tail \plr{CITE SOMETHING}.
 
@@
 %%%%%%%%%%
-\section{Asymptotics for the equilibrium frequency}
-\label{ss:asymptotic_freq}
-
-We provide the above discussion on exact solutions mostly for completeness;
-in practice, we use asymptotics of the solutions.
-These are furthermore available for $d=2$, where we have no analytic solutions.
-In $d=2$ (and higher) dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
-Suppose therefore that the patch is radially symmetric,
-and that the allele is uniformly deleterious outside of some radius $r_0$
-(in the notation above, $\gamma(x) = -s_m$ for $|x|>r_0$).
-Then, denoting by $\phi(r)$ the mean occupation frequency at distance $r$ at equilibrium,
-the polar form of the Laplacian in $d$ dimensions translates equation \eqref{eqn:general_phi} to
-\begin{align}
-  \label{eqn:radial_phi}
-  \frac{\sigma^2}{2} \partial_{r}^2 \phi(r) + \sigma^2\frac{d-1}{2r} \partial_r \phi(r) &= s_m \phi(r) (1-\phi(r)) \quad r>r_0 , \\
-  \phi(r_0) = \phi_0 , \qquad \phi(r) \ge 0 .
-\end{align}
-We wish to find the asymptotic behavior of $\phi$ for large $r$.
-The argument can likely be extended to patches that are not radially symmetric.
-
-The solution with $d=2$ is $C Y_0(ir) \sim C e^{r}/\sqrt{2 \pi r}$ (G\&R 8.451.2).
-
-%%%%%%%%%%
 \section{The probability of establishment}

### 2014-03-07 spatial-equation-soluations.tex
.

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{fullpage}

\begin{document}

Slatkin (1973) argues that if selection at $x$ is $s\gamma(x)$,
the variance in dispersal distance is $\sigma^2$,
and $A(x) = 2p(x)-1 = p(x)-q(x)$, then
\begin{equation}
    \partial_x^2 A(x) = - \frac{s \gamma(x)}{\sigma^2} \left( 1- A^2(x) \right).
\end{equation}
(note the erroneous minus sign in equation (7).
Rescaling to $\xi = x \sqrt{s}/\sigma$, this is
\begin{equation} \label{eqn:reaction_diffusion}
    \partial_\xi^2 A(\xi) = - \gamma(x) \left( 1- A^2(\xi) \right).
\end{equation}

Note that if $f(x) = 2 - 3 \tanh^2(x/2+x_0)$ then since $\partial_x \tanh(x) = (1-\tanh^2(x))$,
\begin{align*}
    2 &\ge f(x) \ge -1 \\
    \partial_x f(x) &= - 3 \tanh(x/2+x_0) \left( 1- \tanh^2(x/2+x_0) \right) \\
    \partial^2_x f(x) &= - \frac{3}{2} \left( 1- \tanh^2(x/2+x_0) \right) 
    + \frac{9}{2} \tanh^2(x/2+x_0) \left(1-\tanh^2(x/2+x_0)\right) \\
        &= \frac{1}{2} \left\{ - 3 + 12 \tanh^2(x/2+x_0) - 9 \tanh^4(x/2+x_0) \right\} \\
        &= \frac{1}{2} \left(1-f(x)^2 \right)
\end{align*}
and a solution to equation \eqref{eqn:reaction_diffusion}, if $\gamma(x)=\gamma$ is constant,
is obtained by choosing $x_0$ to match the boundary conditions.

Note that since $-1\le A(x) \le 1$, we certainly can't apply the solution $f(x)$
for $|x/2+x_0| \le \tanh^{-1}(1/\sqrt{3})$.

\end{document}

### 2014-03-11 patchy-selection-paper.tex
.

@@
 (in the large population density limit).
-\citet{barton1987establishment} also showed that
+The same paper also showed that
 this probability of establishment of a new mutation decays exponentially with distance from the patch, 
@@
 (This latter panmictic probability of establishment we denote $p_e$,
-and often approximate by $2 s_p / \xi^2$ \citep{haldane,fisher}.)
+and often approximate by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.)
 This result holds quite generally for a geographically spread population that experiences a uniform selection
@@
 As we are interested in patches where local adaptation can occur,
-we will assume that our patch is larger than the cutoff for local establishment, 
-i.e.\ wider than $2 \tan^{-1} (\sqrt{s_m/s_p})$ \citep{slatkin1973geneflow}.
+we will assume that our patch is larger than the cutoff for local establishment 
+mentioned above.
 
@@
 relative to the patch fixes within the poor habitat patch.
-As shown in Appendix \ref{apx:eqfreq},
+As shown in Appendix \ref{apx:elliptic_integrals} XXX,
 the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
@@
   Results from individual-based simulation, on a one-dimensional grid:
-  the focal allele is beneficial in the central 20 demes, with $s_b=0.1$,
+  the focal allele is beneficial in the central 20 demes, with $s_p=0.1$,
   and deleterious elsewhere, with $s_m=-0.02$.
@@
 but only very few, and rarely, far away from the patch where $B$ is established.
-Denote the expected frequency of allele $B$ at distance $r$ from the patch by $q(r)$.
+Denote the expected frequency of allele $B$ at displacement $r$ relative to the patch by $q(r)$.
 Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(r)$ is the solution,
-which can then be solved to give $q(r)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
-Asymptotics of these implies that for large $r$ the mean occupation frequency decays exponentially, 
+which in one dimension can be solved to give $q(r)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
+More general, asymptotics of the solution implies that for large $r$ the mean occupation frequency decays exponentially, with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}):
 \begin{align} \label{eqn:eqfreq}
-    q(r) \approx C \exp( - r \sqrt{2 s_m} / \sigma) \qquad \text{for large $r$},
+    q(r) \approx C |r|^{-(d-1)/2} \exp( - |r| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|r|$},
 \end{align}
 where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
-In Appendix \ref{apx:patchy_sims} we argue that $C=1/2$ is about right for most parameter values,
-and also that these asymptotics are correct in both one and two dimensions
-(the differential equation is only analytically solvable in one dimension).
+In Appendix \ref{apx:patchy_sims} we argue that $C=1/2$ is about right for most parameter values.
 
@@
 close to the original patch, the ``equilibrium frequency'' $q(r)$ describes well the density of the $B$ allele at most times,
-but far away from the patch, there are rare excursions of a family of $B$ alleles, interspersed by long periods of absence.
+but far away from the patch, 
+the equilibrium is composed of rare excursions of a family of $B$ alleles, interspersed by long periods of absence.
 To compute the rate of establishment by such migration,
 we use properties of subcritical branching processes and Brownian motion,
-combined with \eqref{eqn:eqfreq},
-to obtain an approximation that we argue is valid if the distance $r$ is large relative to $\sigma/\sqrt{s_p}$.
+combined with equation \eqref{eqn:eqfreq},
+to obtain an approximation that is valid if the distance $r$ is large (at least five times $\sigma/\sqrt{s_m}$, say).
 
@@
   \textbf{(B)} a two-dimensional grid of $50\times 50$ demes.
-  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq} (with $C=1/2$).
+  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq}.
   The two distinct lines of grey points in (A) are to the left and right of the patch, respectively.
@@
 \label{apx:elliptic_integrals}
+
+XXX add equation that probability of establishment solves also XXX
 
@@
 \section{Patchy selection} 
+\label{apx:asymptotics}
+
+The general equation XXX is not analytically solvable,
+but the theory in \citet{wasow1965asymptotic}
+lets us conclude that XXX
+
+%%%%%%%%%%%
+\section{Patchy selection} 
 \label{ss:discretedemes}

### 2014-03-11 generate-patchy-run.R

@@ -1,4 +1,6 @@
 #!/usr/bin/R
+# example:
+# Rscript generate-patchy-run.R 'nsteps=1e3' 'stepsize=1e3' 'range=c(201,201)'
 
 source("sim-patchy-selection-fns.R")
 source("lineages.R")
@@ -80,7 +82,7 @@ initpop$n[,,2][params$s>0] <- params$N
 
 
 # Generate and save runs
-pophist <- pophistory( pop=initpop, nsteps=nsteps, step=stepsize, progress=stepsize )
+pophist <- pophistory( pop=initpop, nsteps=nsteps, step=stepsize, progress=max(100,stepsize) )
 
 if (do.lineages) {
     ## trace lineages back

### 2014-03-11 patchy-sim-writeup-plots.R

@@ -113,16 +113,21 @@ par(mar=c(5,4,3,1)+.1)
 for (run.name in c("5038-1-10000-pophistory-run.Rdata","352-1-10000-pophistory-run.Rdata")) {
     load(run.name)
     dimension <- sum(dim(pophist$pophist)[1:2]>1)
-    theory.decay <- dimension * sqrt(2*abs(getgrowth(pophist$pop$params)$gm) / getsigma(pophist$pop$params) )
+    theory.decay <- dimension* sqrt(2*abs(getgrowth(pophist$pop$params)$gm) / getsigma(pophist$pop$params) )
     # the exp'l decay
     patchloc <- with(pophist$pop$params, (-1)^(s>0) * sqrt( abs( ( row(s) - range[1]/2 )^2 + ( col(s) - range[2]/2 )^2 - (patchsize/2)^2 ) ) )
     tmpdists <- seq(min(patchloc),max(patchloc),length.out=27)
     tmplocs <- tmpdists[-1] - diff(tmpdists)/2
     patchdist <- cut( as.vector(patchloc), breaks=tmpdists, include.lowest=TRUE, ordered_result=TRUE )
     if (is.null(pophist$occupation)) { pophist$occupation <- rowSums(pophist$pophist,dim=3) }
+    # NOTE: pophist$occupation records *every step*; while pophist$pophist only records every 'stepsize' steps
     plot( patchloc, rowMeans(pophist$pophist[,,2,2000:10000,drop=FALSE],dim=2)/pophist$pop$params$N, log='y', xlab='deme number (space)', ylab='allele frequency', pch=20, cex=.5, col=grey(.7), main=substitute(d==thisdim,list(thisdim=dimension)), ylim=c(1e-4,1) )
     points( tmplocs, tapply(pophist$occupation[,,2]/(pophist$pop$gen*pophist$pop$params$N),patchdist,mean), lwd=2 )
-    lines( tmplocs, (1/2) * exp( - tmplocs * theory.decay ), lwd=2 )
+    if (dimension==1) {
+        lines( tmplocs, (1/2) * exp( - tmplocs * theory.decay ), lwd=2 )
+    } else {
+        lines( tmplocs, 2 * tmplocs^(-1/2) * exp( - tmplocs * theory.decay ), lwd=2 )
+    }
     abline(v=0,lty=2)
 }
 dev.off()

### 2014-03-13 patchy-selection-paper.tex
.

@@
 
-The covergent evolution of similar phenotypes in response to shared
+The convergent evolution of similar phenotypes in response to shared
 selection pressures is a testiment to the power of selection to
@@
 
-The arguments depends on the following decomposition of the genealogy of $B$ alleles.
-First, fix some intermediate distance $r_0$ from the established patch,
-and classify all $B$ alleles present at any time outside $r_0$ into families
-if they share a common ancestor living outside $r_0$.
+The arguments depend on the following decomposition.
+First, consider the genealogy of all $B$ alleles that were alive at any time outside of the original patch.  
+This is a collection of trees, each rooted at an allele living outside the patch whose parent was born inside the patch.
+Next, fix some intermediate distance $r_0$ from the established patch,
+and erase from the genealogy every allele that has no ancestors living further away than $r_0$ to the patch.
+This fragments the original set of trees into a collection of smaller trees that relate to each other all $B$ alleles living outside of $r_0$ at any time,
+and some living inside of $r_0$;
+we refer to these trees as ``families''.
 If $r_0$ is large enough that there are not too many families
 and interactions between family members are weak,
-then these families can be approximated by a collection of independent spatial branching processes.
-(This can be made formal in the limit of large population density.)
-We can extend these families by 
-including any individual outside the original patch (but perhaps inside $r_0$) 
-that has a parent outside $r_0$,
-so that these branching processes are rooted at $r_0$, and are killed upon reaching the original patch.
+then these families can be approximated by a collection of independent spatial branching processes
+that are killed upon reaching the original patch.
+(This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
 This decomposition is illustrated in~\ref{fig:sim_snapshots}, and later in figure~\ref{fig:branching_decomp}.
@@
 \end{align}
+These quantities should not depend on $r_0$, of course.
 All terms except the ``outflux of families'' are simple calculations with spatial branching processes;
@@
 To understand the process of adaptation by migration,
-we therefore need to look more closely at the rare families of $B$ the allele far from the patches where it is established.
-Consider one such family.
-As noted above, we can model this by assuming that each individual
+we therefore need to look more closely at the rare families of $B$ alleles far from the patch where it is established.
+Consider one such family,
+that lives far from the original patch.
+As noted above, we will model this by assuming that each individual
 migrates, then reproduces, independently of all other family members.
@@
 we drop the individual from our family.
-Since the $B$ allele is uniformly deleterious elsewhere,
-we can decouple reproduction and migration
+We assume the $B$ allele is uniformly deleterious elsewhere,
+and so can decouple reproduction and migration
 by first forming a tree in which each individual has a random number of offspring
 (with mean $1-s_m$ and variance $\xi^2$),
-then assigning spatial locations to each individual in the tree,
-and then pruning any branches that happen to wander into a patch.
+then recursively assigning spatial locations to each individual in the tree,
+and finally pruning any branches that happen to wander into a patch.
 This pruning removes both branches leading back into the already established patch (which are irrelevant)
@@
 
-Since each individual has on average $(1-s_m)$ offspring,
-a family has on average $(1-s_m)^t$ members in generation $t$.
 Let $p_e(t)$ be the chance of extinction by time $t$, and let $K_t$ be the (random) family size conditioned on nonextinction,
@@
 In other words, the chance of survival to time $t$ is a constant multiplied by $(1-s_m)^t$,
-and conditional on survival, the family size is fixed.
+and conditional on survival, the family size is given by a fixed distribution.
 
@@
 can be decomposed into a ``trunk'' that lasts until $t$
-sprouting independent ``branches'' that may or may not survive until $t$.
+which sprouts independent ``branches'' that may or may not survive until $t$.
 This process is depicted in Figure~\ref{fig:branching_decomp}A.
@@
 \caption{Cartoon of the decomposition described in the text.
-Migrants families are emerging from the already-established patch (grey circle);
+Migrants families that get further than $r_0$ from the already-established patch (grey circle);
 each is composed of a ``trunk'' lineage (in red), whose spatial motion we follow,
@@
 (Recall $\sigma$ is the dispersal distance.)
+We can use these two ingredients, Brownian motion and a branching process, 
+to compute the terms of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
+
+\subsubsection*{Mean occupation density}
+
+First, we compute the second term in equation \eqref{eqn:gestalt_q}, ``mean occupation density of a family at $r$'',
+by which we mean the expected total number of offspring of a single family that ever live at distance $r$ from the patch.
+Concretely, suppose that $Z$ is such a branching process, started with a single individual at $r_0$,
+defined so that $Z_t(A)$ is the number of individuals alive at time $t$ living in the set $A$.
+The mean occupation density of $Z$ is then defined to be the density of $\E[\int_0^t Z_t dt]$,
+i.e.\ the smooth function $u(x)$ such that
+\begin{align} \label{eqn:occupation_density_defn}
+    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x) dx .
+\end{align}
+Since each individual has on average $(1-s_m)$ offspring,
+a family has on average $(1-s_m)^t$ members in generation $t$;
+and the marginal distribution of any one surviving individual is Gaussian with variance $t \sigma^2$,
+the density $u(x)$ is equal to the Green's function for Brownian motion,
+\begin{align} \label{eqn:occupation_density_solution}
+    u(x) = \int_0^\infty (1-s_m)^t p_{\sigma^2 t}(x_0,x) dt
+\end{align}
+where $p_t$ is the transition density for standard Brownian motion.
+
+\subsubsection*{Probability of establishment}
+
 Suppose that the new patch is at distance $r$.

### 2014-03-13 generate-patchy-run.R

@@ -2,8 +2,9 @@
 # example:
 # Rscript generate-patchy-run.R 'nsteps=1e3' 'stepsize=1e3' 'range=c(201,201)'
 
-source("sim-patchy-selection-fns.R")
-source("lineages.R")
+# SOURCE THESE BELOW:
+# source("sim-patchy-selection-fns.R")
+# source("lineages.R")
 
 ngens <- 1e5
 nsteps <- 1000  # record at this many steps
@@ -12,6 +13,7 @@ stepsize <- max(1,floor(ngens/nsteps))
 run.id <- floor(runif(1)*10000)
 nlins <- 500
 do.lineages <- FALSE
+burnin <- 5e3
 
 # Parameters
 params <- list(
@@ -82,7 +84,7 @@ initpop$n[,,2][params$s>0] <- params$N
 
 
 # Generate and save runs
-pophist <- pophistory( pop=initpop, nsteps=nsteps, step=stepsize, progress=max(100,stepsize) )
+pophist <- pophistory( pop=initpop, nsteps=nsteps, step=stepsize, progress=max(100,stepsize), burnin=burnin )
 
 if (do.lineages) {
     ## trace lineages back

### 2014-03-13 patchy-sim-writeup-plots.R

@@ -119,10 +119,11 @@ for (run.name in c("5038-1-10000-pophistory-run.Rdata","352-1-10000-pophistory-r
     tmpdists <- seq(min(patchloc),max(patchloc),length.out=27)
     tmplocs <- tmpdists[-1] - diff(tmpdists)/2
     patchdist <- cut( as.vector(patchloc), breaks=tmpdists, include.lowest=TRUE, ordered_result=TRUE )
-    if (is.null(pophist$occupation)) { pophist$occupation <- rowSums(pophist$pophist,dim=3) }
-    # NOTE: pophist$occupation records *every step*; while pophist$pophist only records every 'stepsize' steps
-    plot( patchloc, rowMeans(pophist$pophist[,,2,2000:10000,drop=FALSE],dim=2)/pophist$pop$params$N, log='y', xlab='deme number (space)', ylab='allele frequency', pch=20, cex=.5, col=grey(.7), main=substitute(d==thisdim,list(thisdim=dimension)), ylim=c(1e-4,1) )
-    points( tmplocs, tapply(pophist$occupation[,,2]/(pophist$pop$gen*pophist$pop$params$N),patchdist,mean), lwd=2 )
+    # NOTE: if present, pophist$occupation records *every step* after 'burnin'; while pophist$pophist only records every 'stepsize' steps
+    # if (is.null(pophist$occupation)) { warning("occupation density not recorded."); pophist$occupation <- rowSums(pophist$pophist,dim=3) }
+    # plot( patchloc, rowMeans(pophist$pophist[,,2,2000:10000,drop=FALSE],dim=2)/pophist$pop$params$N, log='y', xlab='deme number (space)', ylab='allele frequency', pch=20, cex=.5, col=grey(.7), main=substitute(d==thisdim,list(thisdim=dimension)), ylim=c(1e-4,1) )
+    plot( patchloc, pophist$occupation[,,2]/((pophist$pop$gen-pophist$burnin)*pophist$pop$params$N), log='y', xlab='deme number (space)', ylab='allele frequency', pch=20, cex=.5, col=grey(.7), main=substitute(d==thisdim,list(thisdim=dimension)), ylim=c(1e-4,1) )
+    points( tmplocs, tapply(pophist$occupation[,,2]/((pophist$pop$gen-pophist$burnin)*pophist$pop$params$N),patchdist,mean), lwd=2 )
     if (dimension==1) {
         lines( tmplocs, (1/2) * exp( - tmplocs * theory.decay ), lwd=2 )
     } else {

### 2014-03-28 patchy-selection-paper.tex
plotting

@@
 but only very few, and rarely, far away from the patch where $B$ is established.
-Denote the expected frequency of allele $B$ at displacement $r$ relative to the patch by $q(r)$.
-Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(r)$ is the solution,
-which in one dimension can be solved to give $q(r)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
+Denote the expected frequency of allele $B$ at displacement $x$ relative to the patch by $q(x)$.
+Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(x)$ is the solution,
+which in one dimension can be solved to give $q(x)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
 More general, asymptotics of the solution implies that for large $r$ the mean occupation frequency decays exponentially, with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}):
 \begin{align} \label{eqn:eqfreq}
-    q(r) \approx C |r|^{-(d-1)/2} \exp( - |r| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|r|$},
+    q(x) \approx C |x|^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
 \end{align}
 where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
-In Appendix \ref{apx:patchy_sims} we argue that $C=1/2$ is about right for most parameter values.
+In practice we take $C=1/2$,
+since from simulations, $C$ does not seem to vary from this by more than a factor of 2 or so (see Appendix \ref{apx:patchy_sims}).
 
@@
 the equilibrium is composed of rare excursions of a family of $B$ alleles, interspersed by long periods of absence.
+The relevant time scale on which $B$ alleles migrate between patches is given by the rate of appearance of such families.
 To compute the rate of establishment by such migration,
@@
 
-The arguments depend on the following decomposition.
+To make the argument, we decompose the genealogy of $B$ alleles into families in the following way.
 First, consider the genealogy of all $B$ alleles that were alive at any time outside of the original patch.  
@@
 then these families can be approximated by a collection of independent spatial branching processes
-that are killed upon reaching the original patch.
+whose members are killed (actually, ignored) if they enter the original patch.
 (This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
@@
     \begin{split}
-        q(r) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $r$ ) } ,
+        q(x) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $x$ ) } ,
 \end{split}
@@
     \begin{split}
-        \migrate(r) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in patch at $r$ ) }
+        \migrate(x) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in patch at $x$ ) }
     \end{split}
@@
 These quantities should not depend on $r_0$, of course.
-All terms except the ``outflux of families'' are simple calculations with spatial branching processes;
-and we can solve for this outflux using the established form for \eqref{eqn:gestalt_q}.
-Here $\migrate$ reflects the mean number of establishing migrant families per unit time,
-rather than individuals,
-because family establishment events may include more than one individual,
-so individual establishment events are clustered in time,
-and using their overall density to obtain a typical waiting time until the next event
-will underestimate that time.
+All terms except the ``outflux of families'' are calculations with spatial branching processes;
+and we can solve for this outflux using the established form for $q(r)$ in \eqref{eqn:gestalt_q}.
+Note that $\migrate$ reflects the mean number of establishing migrant families per unit time,  rather than individuals.
 
@@
 we drop the individual from our family.
-We assume the $B$ allele is uniformly deleterious elsewhere,
-and so can decouple reproduction and migration
+Since the $B$ allele is uniformly deleterious elsewhere,
+we can decouple reproduction and migration
 by first forming a tree in which each individual has a random number of offspring
@@
 
-Let $p_e(t)$ be the chance of extinction by time $t$, and let $K_t$ be the (random) family size conditioned on nonextinction,
-so that $(1-s_m)^t = (1-p_e(t))\E[K_t]$.
-It turns out that by assuming the fate of each offspring is independent (i.e.\ that this forms a branching process),
-and that the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jagers1975branching},
-that $K_t$ has a limiting distribution: $K_t \dconv K$ 
+Since each individual has on average $(1-s_m)$ offspring,
+the expected family size after time $t$ is $(1-s_m)^t$.
+Therefore, if we define $p_e(t)$ to be the chance of extinction by time $t$, 
+and let $K_t$ be the (random) family size conditioned on nonextinction,
+then by Bayes' rule, $(1-s_m)^t = (1-p_e(t))\E[K_t]$.
+It turns out that if the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jagers1975branching},
+then $K_t$ has a limiting distribution: $K_t \dconv K$ 
 and that $(1-s_m)^{t}/(1-p_e(t)) \to \E[K]$ as $t \to \infty$.
@@
 We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary}.
-When we condition on survival until $t$, we condition on existence of at least one lineage from time $0$ to time $t$.
+When we condition on survival until $t$, we condition on existence of at least one lineage surviving from time $0$ to time $t$.
 Once given this lineage, nothing distinguishes any of the other reproduction events --
@@
 This process is depicted in Figure~\ref{fig:branching_decomp}A.
-Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$
-(so $\P\{Z_t = 0\} = p_e(t)$),
-then
+Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
+% (so $\P\{Z_t = 0\} = p_e(t)$),
+then this gives the following decomposition:
 \begin{align}
@@
 
+%%%%%%% %%%%%%%%%%%
 \subsubsection*{Mean occupation density}
 
-First, we compute the second term in equation \eqref{eqn:gestalt_q}, ``mean occupation density of a family at $r$'',
+First, we compute the second term in equation \eqref{eqn:gestalt_q}, (mean occupation density of a family at $x$),
 by which we mean the expected total number of offspring of a single family that ever live at distance $r$ from the patch.
-Concretely, suppose that $Z$ is such a branching process, started with a single individual at $r_0$,
-defined so that $Z_t(A)$ is the number of individuals alive at time $t$ living in the set $A$.
+In two dimensions, this depends on where the family began, 
+and right-hand side of equation \eqref{eqn:gestalt_q} should be integrated over this starting point.
+Here we'll write $x_0$ for the starting point, and $u(x_0,x)$ for the mean occupation density at $x$.
+Concretely, suppose that $Z$ is a spatial branching process as above, started with a single individual at $x_0$,
+defined now so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the set $A$.
 The mean occupation density of $Z$ is then defined to be the density of $\E[\int_0^t Z_t dt]$,
-i.e.\ the smooth function $u(x)$ such that
+i.e.\ the smooth function $u(x_0,x)$ that, integrated over a set $A$, 
+gives the expected total number of individuals ever living in $A$:
 \begin{align} \label{eqn:occupation_density_defn}
-    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x) dx .
-\end{align}
-Since each individual has on average $(1-s_m)$ offspring,
-a family has on average $(1-s_m)^t$ members in generation $t$;
-and the marginal distribution of any one surviving individual is Gaussian with variance $t \sigma^2$,
-the density $u(x)$ is equal to the Green's function for Brownian motion,
+    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x_0,x) dx .
+\end{align}
+Since a family has on average $(1-s_m)^t$ members in generation $t$,
+and the marginal distribution of any one surviving individual is approximately Gaussian with variance $t \sigma^2$,
+exchanging the expectation and the integral shows that
+the density $u(x_0,x)$ is equal to the Green's function for Brownian motion killed at rate $1-s_m$,
 \begin{align} \label{eqn:occupation_density_solution}
-    u(x) = \int_0^\infty (1-s_m)^t p_{\sigma^2 t}(x_0,x) dt
-\end{align}
-where $p_t$ is the transition density for standard Brownian motion.
+    u(x_0,x) = \int_0^\infty (1-s_m)^t p_{\sigma^2 t}(x_0,x) dt
+\end{align}
+where $p_t(x,y) = (2 \pi t)^{-d/2} \exp(-\|x-y\|^2/2t)$ is the transition density for standard Brownian motion.
+This is known exactly; the solution is
+
+%%%%%%% %%%%%%%%%%%
+\subsubsection*{Outflux of families}
+
+Now, we can obtain the first term in equation \eqref{eqn:gestalt_q}, the (outflux of families).
+Define $\gamma$ to be this outflux; 
+i.e.\ the mean number of individuals per unit time that are born carrying the $B$ allele closer than $r_0$ to the original patch
+but migrate to a larger distance.
+Then for $r>r_0$, the mean occupation density $q(r)$ is equal to $\gamma$ 
+multiplied by the mean occupation density of the family descending from such a migrant,
+which we denote $u(r)$.
+Denote by $Z_t$ the number of individuals such a family has after time $t$,
+and the locations of these by $X_1(t), \ldots, X_{Z_t}(t)$.
+Then this mean occupation density, at $x$, is the expected number of individuals ever present at $x$. 
+Since the expected number of individuals present at time $t$ is $e^{\alpha t}$
+and the location of each of these individuals is marginally distributed as a  Gaussian with variance $\sigma^2 t$,
+we can write this expectation as
+\begin{align}
+    u(x) &= \int_0^t \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
+    &= \int_0^t e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
+    &= \frac{ 1 }{ \sigma \sqrt{2 \alpha} } \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) 
+\end{align}
+\plr{check that}
+The integral is evaluated in Appendix~\ref{apx:some_integrals}.
+\plr{Also examine non-dependence on $r_0$.}
+
+Combining this with the solution for $q(r)$ (equation~\eqref{eqn:eqfreq}) as in \eqref{eqn:q_gestalt},
+we then have that (with $C=1/2$)
+\begin{align} \label{eqn:outflux}
+    \gamma = \text{( outflux of families )} \approx (1/2) \rho \sigma \sqrt{2\alpha} .
+\end{align}
 
@@
 
-%%%%%
-\subsubsection{The rate of outflux of migrant families}
-
-Now, we can obtain the effective ``outflux of families'' from information about the migrant families, combined with \eqref{eqn:gestalt_q}.
-Define $\gamma$ to be this outflux; 
-i.e.\ the mean number of individuals per unit time that are born carrying the $B$ allele closer than $r_0$ to the original patch
-but migrate to a larger distance.
-Then for $r>r_0$, the mean occupation density $q(r)$ is equal to $\gamma$ 
-multiplied by the mean occupation density of the family descending from such a migrant,
-which we denote $u(r)$.
-Denote by $Z_t$ the number of individuals such a family has after time $t$,
-and the locations of these by $X_1(t), \ldots, X_{Z_t}(t)$.
-Then this mean occupation density, at $x$, is the expected number of individuals ever present at $x$. 
-Since the expected number of individuals present at time $t$ is $e^{\alpha t}$
-and the location of each of these individuals is marginally distributed as a  Gaussian with variance $\sigma^2 t$,
-we can write this expectation as
-\begin{align}
-    u(x) &= \int_0^t \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
-    &= \int_0^t e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
-    &= \frac{ 1 }{ \sigma \sqrt{2 \alpha} } \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) 
-\end{align}
-\plr{check that}
-The integral is evaluated in Appendix~\ref{apx:some_integrals}.
-\plr{Also examine non-dependence on $r_0$.}
-
-Combining this with the solution for $q(r)$ (equation~\eqref{eqn:eqfreq}) as in \eqref{eqn:q_gestalt},
-we then have that (with $C=1/2$)
-\begin{align} \label{eqn:outflux}
-    \gamma = \text{( outflux of families )} \approx (1/2) \rho \sigma \sqrt{2\alpha} .
-\end{align}
+\subsubsection*{Putting it all together}
 

### 2014-03-28 patchy-sim-writeup-plots.R

@@ -9,53 +9,6 @@ rundims <- read.csv("run-info.csv")
 
 for (run.name in run.list) {
 
-    run.id <- gsub( "([^-]*)-.*","\\1",run.name)
-    load(run.name)
-    print(c("run",run.id,"range",pophist$pop$params$range))
-
-    dimension <- sum(dim(pophist$pophist)[1:2]>1)
-    theory.decay <- dimension * sqrt(2*abs(getgrowth(pophist$pop$params)$gm) / getsigma(pophist$pop$params) )
-
-    # time slices of the process
-    timeslice <- floor( seq( .05*dim(pophist$pophist)[4], dim(pophist$pophist)[4], length.out=50 ) )
-
-    # the exp'l decay
-    patchloc <- with(pophist$pop$params, (-1)^(s>0) * sqrt( abs( ( row(s) - range[1]/2 )^2 + ( col(s) - range[2]/2 )^2 - (patchsize/2)^2 ) ) )
-    tmpdists <- seq(min(patchloc),max(patchloc),length.out=27)
-    tmplocs <- tmpdists[-1] - diff(tmpdists)/2
-    patchdist <- cut( as.vector(patchloc), breaks=tmpdists, include.lowest=TRUE, ordered_result=TRUE )
-    # equil <- getequilibrium(pophist,init=rowMeans(pophist$pophist[,,2,2000:10000,drop=FALSE])/pophist$pop$params$N) #,keep.convergence=TRUE)
-    if (is.null(pophist$occupation)) { pophist$occupation <- rowSums(pophist$pophist,dim=3) }
-
-    if (dimension==1) {
-        sliced <- pophist$pophist[,,2,timeslice]/pophist$pop$params$N
-        # if (length(dim(sliced))>2) { dim(sliced) <- c( prod(dim(sliced)[1:2]), dim(sliced)[3] ) }
-
-        pdf(file=paste(run.id,'example-equilibrium.pdf',sep='-'), width=6, height=6, pointsize=10)
-        par(mar=c(5,4,1,1)+.1)
-        layout(1:2)
-
-        matplot( sliced, type='l', xlab='deme number (space)', ylab='allele frequency' )
-        abline(v=0.5 + which(diff(as.vector(pophist$pop$params$s))!=0), lty=2 )
[…]

### 2014-03-28 plot-patchy-sim.R

@@ -1,0 +1,56 @@
+#!/usr/bin/R
+source("sim-patchy-selection-fns.R")
+source("lineages.R")
+
+run.name <- commandArgs(TRUE)[1]
+
+run.id <- gsub( "([^-]*)-.*","\\1",run.name)
+load(run.name)
+print(c("run",run.id,"range",pophist$pop$params$range))
+
+dimension <- sum(dim(pophist$pophist)[1:2]>1)
+theory.decay <- sqrt(2*abs(getgrowth(pophist$pop$params)$gm) / getsigma(pophist$pop$params) )
+
+# time slices of the process
+timeslice <- floor( seq( .05*dim(pophist$pophist)[4], dim(pophist$pophist)[4], length.out=50 ) )
+
+# the exp'l decay
+patchloc <- with(pophist$pop$params, (-1)^(s>0) * sqrt( abs( ( row(s) - range[1]/2 )^2 + ( col(s) - range[2]/2 )^2 - (patchsize/2)^2 ) ) )
+obs.freqs <- pophist$occupation[,,2]/(pophist$pop$params$N*(pophist$pop$gen-pophist$burnin))
+pred.freqs <-  ( if (dimension==2) {patchloc^(-1/2)} else {1} )  * exp( - patchloc * theory.decay )
+fit.const <- exp( mean( log(obs.freqs/pred.freqs)[obs.freqs<.01], na.rm=TRUE ) )
+tmpdists <- seq(min(patchloc),max(patchloc),length.out=27)
+tmplocs <- tmpdists[-1] - diff(tmpdists)/2
+patchdist <- cut( as.vector(patchloc), breaks=tmpdists, include.lowest=TRUE, ordered_result=TRUE )
+# equil <- getequilibrium(pophist,init=rowMeans(pophist$pophist[,,2,2000:10000,drop=FALSE])/pophist$pop$params$N) #,keep.convergence=TRUE)
+if (is.null(pophist$occupation)) { pophist$occupation <- rowSums(pophist$pophist,dim=3); pophist$burnin <- 0 }
+
+if (dimension==1) {
+    sliced <- pophist$pophist[,,2,timeslice]/pophist$pop$params$N
+    # if (length(dim(sliced))>2) { dim(sliced) <- c( prod(dim(sliced)[1:2]), dim(sliced)[3] ) }
+
+    pdf(file=paste(run.id,'example-equilibrium.pdf',sep='-'), width=6, height=6, pointsize=10)
+    par(mar=c(5,4,1,1)+.1)
[…]

### 2014-03-28 patchy-selection-paper.tex
.

@@
 \begin{align} \label{eqn:eqfreq}
-    q(x) \approx C |x|^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
-\end{align}
-where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
-In practice we take $C=1/2$,
-since from simulations, $C$ does not seem to vary from this by more than a factor of 2 or so (see Appendix \ref{apx:patchy_sims}).
+    q(x) \approx C (|x|-r_0)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
+\end{align}
+where $d$ is the dimension ($d=1$ or $2$), and $C$ and $r_0$ are constants depending on the geographic shape of the populations and the selection coefficients
+(but note that $r_0$ is only necessary in $d=2$).
+In applications we fit $C$ and $r_0$ to the data;
+to get concrete numbers from theory we take $C=1$ and $r_0=0$ if necessary.
+From simulations, $C$ seems to be close to 1 in practice (see Appendix \ref{apx:patchy_sims})
+and XXX $r_0$.
 
@@
 \end{align}
-These quantities should not depend on $r_0$, of course.
+The quantities on the right-hand side depend implicitly on $r_0$.
 All terms except the ``outflux of families'' are calculations with spatial branching processes;
@@
 where $p_t(x,y) = (2 \pi t)^{-d/2} \exp(-\|x-y\|^2/2t)$ is the transition density for standard Brownian motion.
-This is known exactly; the solution is
+The integral can be expressed in terms of modified Bessel functions:
+using \citet{gradshteyn2007table} (3.471.9 and 8.451.6), the mean occupation density is,
+with $\alpha = -\log(1-s_m)$ and $d$ the dimension,
+\begin{align}
+    u(x_0,x) &= \frac{1}{\sigma^d \pi}\left(\sqrt{ \frac{\|x-x_0\|^2}{2\alpha\sigma^2} }\right)^{(d+2)/2} K_{(d+2)/2}\left( \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right) \\
+             &= \frac{1}{\sigma^d \sqrt{2\pi}} \left(\sqrt{ \frac{\|x-x_0\|^2}{2\alpha\sigma^2} }\right)^{d/2} \exp\left(- \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right)+ O(\sigma/\|x-x_0\|) .
+\label{eqn:occupation_asymptotics}
+\end{align}
+Since we only have asymptotics for the equilibrium frequency, 
+we will only use the latter (asymptotic) form.
 
@@
 Now, we can obtain the first term in equation \eqref{eqn:gestalt_q}, the (outflux of families).
-Define $\gamma$ to be this outflux; 
+Define $\gamma(r_0)$ to be this outflux; 
 i.e.\ the mean number of individuals per unit time that are born carrying the $B$ allele closer than $r_0$ to the original patch
 but migrate to a larger distance.
-Then for $r>r_0$, the mean occupation density $q(r)$ is equal to $\gamma$ 
-multiplied by the mean occupation density of the family descending from such a migrant,
-which we denote $u(r)$.
+Then for any point $x$ further than $r_0$ from the patch, 
+the mean occupation density $q(x)$ is equal to the integral of $\gamma$ over all locations $x_0$ at distance $r_0$ from the patch,
+multiplied by the mean occupation density at $x$ for migrant families beginning at at location.
+This is a more precise statement of the decomposition of the equilibrium frequency from equation \eqref{eqn:gestalt_q}.
+However, this integral is not tractable; if we assume that $r_0$ is small relative to $\|x\|$, then we get 
+\begin{align}
+    q(x) &= \int_{B(r_0)} u(y,x) \gamma(|y|) dy \\
+         &\simeq \gamma(r_0) 2 (\pi r_0)^{d-1} u(0,x) .
+\end{align}
+Now, using equations \eqref{eqn:eqfreq} and \eqref{eqn:occupation_asymptotics}, we get that
+\begin{align}
+    \gamma(r_0) \simeq 
+\end{align}
+
 Denote by $Z_t$ the number of individuals such a family has after time $t$,
@@
 \end{align}
-However, the second term goes to zero as $r \to \infty$,
-leading to the same asymptotic behavior for the tail \plr{CITE SOMETHING}.
+
+%%%%%%%%%%%
+\subsection{Asymptotic solution for the equilibrium frequency} 
+\label{apx:asymptotics}
+
+If $q(r)$ is a radially symmetric equilibrium frequency and $\xi(t,x)=q(|x|)$,
+then $\partial_t \xi(t,x) = 0$.
+So, using \eqref{eqn:radial_fkpp}, a radially symmetric solution 
+with $\gamma(r) = -s < 0$ for all $r>r_0$ 
+solves, for $r_0 < r < \infty$,
+\begin{align}
+    \partial_r^2 q(r) + \frac{d-1}{r} \partial_r q(r) - \frac{2s}{\sigma^2} q(r) (1-q(r)) = 0 \\
+    \lim_{r \to \infty} q(r) = \lim_{r \to \infty} \partial_r q(r) = 0 \qquad 
+    0 < q(r) < 1 
+\end{align}
+Since $q(r) \to 0$ as $r \to \infty$, so $q(r) (1-q(r)) \approx q(r)$,
+XXX theory in \citet{wasow1965asymptotic} XXX
+implies that the true equilibrium frequency $q$ is close, for large $r$, to the solution to
+\begin{align}
+    \partial_r^2 u(r) + \frac{d-1}{r} \partial_r u(r) - \frac{2s}{\sigma^2} u(r) = 0  \label{eqn:bessel} \\
+    \lim_{r \to \infty} u(r) = \lim_{r \to \infty} \partial_r u(r) = 0 \qquad  .
+\end{align}
+This has general solution given by a modified Bessel function:
+using \citet{gradshteyn2007table} 8.494.8,
+the general solution is
+\begin{align}
+    u(r) = C' (r-r_0)^{d/2} K_{d/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
+\end{align}
+where $C'$ and $r_0$ are chosen to match boundary conditions.
+Asymptotics of Bessel functions \citep[][8.451.6]{gradshteyn2007table} then imply that
+\begin{align}
+    q(r) \simeq C (r-r_0)^{(d-1)/2} \exp \left( -r \sqrt{2s}/\sigma \right) ,
+\end{align}
+where $C$ is a different constant.
 
 %%%%%%%
-\section{The exact solution to the equilibrium frequency}
+\subsection{The exact solution to the equilibrium frequency}
 \label{apx:elliptic_integrals}
@@
 \section{Patchy selection} 
-\label{apx:asymptotics}
-
-The general equation XXX is not analytically solvable,
-but the theory in \citet{wasow1965asymptotic}
-lets us conclude that XXX
-
-%%%%%%%%%%%
-\section{Patchy selection} 
 \label{ss:discretedemes}

### 2014-03-28 generate-patchy-run.R

@@ -39,6 +39,18 @@ for (x in commandArgs(TRUE)) { eval(parse(text=x)) }
 for (x in gsub("^([^ <=]*[ <=])","params$\\1",commandArgs(TRUE))) { eval(parse(text=x)) }
 # print(params)
 
+# load restarting parameters (but then make command-line modifications again)
+if (!is.null(restart)) {
+    stopifnot(file.exists(restart))
+    renv <- new.env()
+    load(restart,envir=renv)
+    old.params <- params
+    params <- with(renv,pophist$pop$params)
+    for (x in gsub("^([^ <=]*[ <=])","params$\\1",commandArgs(TRUE))) { eval(parse(text=x)) }
+    initpop <- with(renv,pophist$pop)
+    initpop$params <- params
+}
+
 filename <- paste(run.id,"-r",paste(params$range,collapse="-"),"-sb",params$sb,"-sm",params$sm,"-pophistory-run.Rdata",sep="")
 run.list <- list.files(".","*-pophistory-run.Rdata")
 while (filename %in% run.list) {  # make sure don't overwrite something
@@ -50,18 +62,6 @@ if (!interactive()) {
     logcon <- file(logfile,open="wt")
     sink(file=logcon, type="message") 
     sink(file=logcon, type="output")   # send both to log file
-}
-
-# load restarting parameters (but then make command-line modifications again)
-if (!is.null(restart)) {
-    stopifnot(file.exists(restart))
-    renv <- new.env()
-    load(restart,envir=renv)
-    old.params <- params
-    params <- with(renv,pophist$pop$params)
-    for (x in gsub("^([^ <=]*[ <=])","params$\\1",commandArgs(TRUE))) { eval(parse(text=x)) }
-    initpop <- with(renv,pophist$pop)
-    initpop$params <- params
 }
 
 # parameters all set up now

### 2014-03-28 patchy-sim-writeup-plots.R

@@ -7,9 +7,6 @@ rundims <- read.csv("run-info.csv")
 
 # run.name <- "3975-1-10000-pophistory-run.Rdata"
 
-for (run.name in run.list) {
-
-}
 
 ### snapshots
 run.name <- "3975-1-10000-pophistory-run.Rdata"

### 2014-03-28 plot-patchy-sim.R

@@ -18,7 +18,7 @@ timeslice <- floor( seq( .05*dim(pophist$pophist)[4], dim(pophist$pophist)[4], l
 patchloc <- with(pophist$pop$params, (-1)^(s>0) * sqrt( abs( ( row(s) - range[1]/2 )^2 + ( col(s) - range[2]/2 )^2 - (patchsize/2)^2 ) ) )
 obs.freqs <- pophist$occupation[,,2]/(pophist$pop$params$N*(pophist$pop$gen-pophist$burnin))
 pred.freqs <-  ( if (dimension==2) {patchloc^(-1/2)} else {1} )  * exp( - patchloc * theory.decay )
-fit.const <- exp( mean( log(obs.freqs/pred.freqs)[obs.freqs<.01], na.rm=TRUE ) )
+fit.const <- exp( mean( log(obs.freqs/pred.freqs)[(obs.freqs<.01)&(obs.freqs>1e-6)], na.rm=TRUE ) )
 tmpdists <- seq(min(patchloc),max(patchloc),length.out=27)
 tmplocs <- tmpdists[-1] - diff(tmpdists)/2
 patchdist <- cut( as.vector(patchloc), breaks=tmpdists, include.lowest=TRUE, ordered_result=TRUE )

### 2014-03-28 restart.sh
.

@@ -10,7 +10,7 @@ do
     ARGS="$ARGS '"$1"'"
     shift
 done
-if [ $(echo $ARGS | grep -v burnin) ]
+if  $(echo $ARGS | grep -v -q burnin) 
 then
     ARGS="$ARGS 'burnin=0'"
 fi

### 2014-04-09 patchy-selection-paper.tex
debugged exponents!

@@
 which in one dimension can be solved to give $q(x)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
-More general, asymptotics of the solution implies that for large $r$ the mean occupation frequency decays exponentially, with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}):
+More general, asymptotics of the solution implies that for large $|x|$ the mean occupation frequency decays exponentially, with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}):
 \begin{align} \label{eqn:eqfreq}
-    q(x) \approx C (|x|-r_0)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
+    q(x) \approx C |x|^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
 \end{align}
@@
 to get concrete numbers from theory we take $C=1$ and $r_0=0$ if necessary.
-From simulations, $C$ seems to be close to 1 in practice (see Appendix \ref{apx:patchy_sims})
-and XXX $r_0$.
+From simulations, $C$ seems to be not too far from 1 in practice (see Appendix \ref{apx:patchy_sims}).
 
@@
   and averaged over demes at similar distances (black circles), on
-  \textbf{(A)} a one-dimensional grid of 150 demes, and
-  \textbf{(B)} a two-dimensional grid of $50\times 50$ demes.
-  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq}.
-  The two distinct lines of grey points in (A) are to the left and right of the patch, respectively.
+  \textbf{(A)} a one-dimensional grid of 201 demes, and
+  \textbf{(B)} a two-dimensional grid of $101\times 101$ demes.
+  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq},
+  with $C$ chosen arbitrarily to give a reasonable fit.
   }
@@
 \begin{align}
-    u(x_0,x) &= \frac{1}{\sigma^d \pi}\left(\sqrt{ \frac{\|x-x_0\|^2}{2\alpha\sigma^2} }\right)^{(d+2)/2} K_{(d+2)/2}\left( \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right) \\
-             &= \frac{1}{\sigma^d \sqrt{2\pi}} \left(\sqrt{ \frac{\|x-x_0\|^2}{2\alpha\sigma^2} }\right)^{d/2} \exp\left(- \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right)+ O(\sigma/\|x-x_0\|) .
+u(x_0,x) &= \frac{\alpha^{(d-2)/2}}{\sigma^d \pi^{d/2}}\left(\frac{ \sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(2-d)/2} K_{(2-d)/2}\left( \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right) \\
+         &= \frac{\alpha^{(d-2)/2}}{\sigma^d 2 \pi^{(d-1)/2}} \left( \frac{\sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(1-d)/2} \exp\left(- \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right)+ O(\sigma/\|x-x_0\|) .
 \label{eqn:occupation_asymptotics}
@@
 This has general solution given by a modified Bessel function:
-using \citet{gradshteyn2007table} 8.494.8,
+using \citet{gradshteyn2007table} 8.494.9,
 the general solution is
 \begin{align}
-    u(r) = C' (r-r_0)^{d/2} K_{d/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
+    u(r) = C' (r-r_0)^{(2-d)/2} K_{(2-d)/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
 \end{align}
@@
 \begin{align}
-    q(r) \simeq C (r-r_0)^{(d-1)/2} \exp \left( -r \sqrt{2s}/\sigma \right) ,
+    q(r) \approx C r^{(1-d)/2} \exp \left( -r \sqrt{2s}/\sigma \right) + O(1/r),
 \end{align}

### 2014-04-09 patchy-sim-writeup-plots.R

@@ -60,28 +60,45 @@ dev.off()
 pdf(file="sim-occupation-freqs.pdf",width=6,height=3,pointsize=10)
 layout(t(1:2))
 par(mar=c(5,4,3,1)+.1)
-for (run.name in c("5038-1-10000-pophistory-run.Rdata","352-1-10000-pophistory-run.Rdata")) {
+# for (run.name in c("2611-r1-201-sb0.05-sm-0.005-pophistory-run.Rdata","7705124-r101-101-sb0.05-sm-0.005-pophistory-run.Rdata")) {
+for (run.name in c("3994-1000-1000-pophistory-run.Rdata","7705124-r101-101-sb0.05-sm-0.005-pophistory-run.Rdata")) {
     # see also plot-patchy-sim.R
     load(run.name)
     dimension <- sum(dim(pophist$pophist)[1:2]>1)
-    theory.decay <- sqrt(2*abs(getgrowth(pophist$pop$params)$gm)) / getsigma(pophist$pop$params)
+    theory.decay <- sqrt(2*abs(getgrowth(pophist$pop$params)$gm)) / getsigma(pophist$pop$params) 
+    # yes, the parentheses differ in the exponents for x:
+    expl.approx <- function (x,d) { sqrt(pi/2) * x^((1-d)/2) * exp(-x) }
+    bessel.approx <- function (x,d) { x^(1-(d/2)) * bessel_Knu( nu=1-d/2, x=x ) } 
     # the exp'l decay
     patchloc <- with(pophist$pop$params, (-1)^(s>0) * sqrt( abs( ( row(s) - range[1]/2 )^2 + ( col(s) - range[2]/2 )^2 - (patchsize/2)^2 ) ) )
     obs.freqs <- pophist$occupation[,,2]/(pophist$pop$params$N*(pophist$pop$gen-pophist$burnin))
-    pred.freqs <-  ( if (dimension==2) {patchloc^(-1/2)} else {1} )  * exp( - patchloc * theory.decay )
-    fit.const <- exp( mean( log(obs.freqs/pred.freqs)[obs.freqs<.01], na.rm=TRUE ) )
+    expl.freqs <-  expl.approx( (patchloc*theory.decay), d=dimension )
+
+    # also allow a shift?
+    if (FALSE) {
+        f <- function (x0) { 
+            goodones <- (obs.freqs>3e-3) & (patchloc > -x0)
[…]

### 2014-04-09 plot-patchy-sim.R

@@ -16,13 +16,16 @@ if (is.null(pophist$occupation) | max(pophist$occupation)==0 ) { stop("occupatio
 
 dimension <- sum(dim(pophist$pophist)[1:2]>1)
 theory.decay <- sqrt(2*abs(getgrowth(pophist$pop$params)$gm)) / getsigma(pophist$pop$params) 
+# yes, the parentheses differ in the exponents for x:
+expl.approx <- function (x,d) { sqrt(pi/2) * x^((1-d)/2) * exp(-x) }
+bessel.approx <- function (x,d) { x^(1-(d/2)) * bessel_Knu( nu=1-d/2, x=x ) } 
 
 # the exp'l decay
 patchloc <- with(pophist$pop$params, (-1)^(s>0) * sqrt( abs( ( row(s) - range[1]/2 )^2 + ( col(s) - range[2]/2 )^2 - (patchsize/2)^2 ) ) )
 obs.freqs <- pophist$occupation[,,2]/(pophist$pop$params$N*(pophist$pop$gen-pophist$burnin))
-expl.freqs <-  (patchloc*theory.decay)^((dimension-1)/2)  * exp( - patchloc * theory.decay )
+expl.freqs <-  expl.approx( (patchloc*theory.decay), d=dimension )
 # and the bessel function
-bessel.freqs <- (patchloc*theory.decay)^(dimension/2) * bessel_Knu( nu=dimension/2, x=(patchloc * theory.decay) )
+bessel.freqs <-  bessel.approx( (patchloc*theory.decay), d=dimension )
 expl.const <- exp( mean( log(obs.freqs/expl.freqs)[(log(obs.freqs)<quantile(log(obs.freqs),.5))&(log(obs.freqs)>quantile(log(obs.freqs),.2))], na.rm=TRUE ) )
 bessel.const <- exp( mean( log(obs.freqs/bessel.freqs)[(log(obs.freqs)<quantile(log(obs.freqs),.9))&(log(obs.freqs)>quantile(log(obs.freqs),.4))], na.rm=TRUE ) )
 tmpdists <- seq(min(patchloc),max(patchloc),length.out=27)
@@ -42,8 +45,8 @@ pdf(file=paste(run.id,'expl-decay.pdf',sep='-'), width=6, height=6, pointsize=10
 plot( patchloc, obs.freqs, log='y', xlab='deme number (space)', ylab='allele frequency', pch=20, cex=.5, col=grey(.7) )
[…]

### 2014-04-10 patchy-selection-paper.tex
.

@@
 multiplied by the mean occupation density at $x$ for migrant families beginning at at location.
-This is a more precise statement of the decomposition of the equilibrium frequency from equation \eqref{eqn:gestalt_q}.
-However, this integral is not tractable; if we assume that $r_0$ is small relative to $\|x\|$, then we get 
-\begin{align}
-    q(x) &= \int_{B(r_0)} u(y,x) \gamma(|y|) dy \\
-         &\simeq \gamma(r_0) 2 (\pi r_0)^{d-1} u(0,x) .
-\end{align}
+This is a more precise statement of the decomposition of the equilibrium frequency from equation \eqref{eqn:gestalt_q}:
+\begin{align}
+    q(x) &= \gamma(r_0) \int_{B(r_0)} u(y,x) dy,
+\end{align}
+so that $\gamma(r_0) = q(x) / \int_{B(r_0)} u(y,x) dy$, for any choice of $x$.
+XXX Green function for BM killed on disk XXX
 Now, using equations \eqref{eqn:eqfreq} and \eqref{eqn:occupation_asymptotics}, we get that

### 2014-04-11 patchy-selection-paper.tex
ref
hitting times

@@
 %%%%%%% %%%%%%%%%%%
+\subsubsection*{Putting it together}
+
+We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
+as the sum of independent migrant families, each of whose dynamics are given by a spatial branching process.
+Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\delta B(r_0)$ its boundary.
+Suppose that there is a constant outflux of migrant families of rate $\gamma(x)$ from each point $x \in \delta B(r_0)$,
+e.g.\ with locations and times of appearances as the points of a Poisson process.
+Roughly speaking, equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
+differ only by substituting the average number of members of a migrant family in the new patch
+for the probability that a family has any members in the new patch;
+by the discussion of the previous section, these differ only by a constant factor.
+
+First we will make this idea more formal.
+Suppose that $Z$ is one such spatial branching process as above, started with a single individual at $x$,
+defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the set $A$.
+The \emph{mean occupation density} of $Z$ is then defined to be the density of $\E[\int_0^t Z_t dt]$,
+i.e.\ the smooth function $u(x,y)$ that, integrated over a set $A$, 
+gives the expected total number of individuals ever living in $A$:
+\begin{align} \label{eqn:occupation_density_defn}
+    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x,y) dy .
+\end{align}
+We can think of $u(x,y)$ as the expected total number of offspring of a family beginning at $x$
+that ever live at $y$.
+By independence of reproduction and spatial motion,
+$u(x,y)$ is the sum over all times of the expected number of offspring
+multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
+i.e.\
+\begin{align}
+    u(x,y) = \int_0^t e^{-\alpha t} p_t(x,y) dt .
+\end{align}
+Now, \eqref{eqn:gestalt_q} is
+\begin{align}
+    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
+         &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-\alpha t} p_t(x,y) dt dx .
+\end{align}
+
+Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
+which we assume has relatively small area $A$ and is located near position $y$.
+The probability that this family reaches $A$ by time $t$ is $(1-p_e(t))$
+multiplied by the probability that the family reaches the patch, given survival.
+If $t$ is large then the previous section argues that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
+and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
+will have hit the patch by then.
+If $w = A^{1/d}$ is the width of the patch, then
+
+%%%%%%% %%%%%%%%%%%
 \subsubsection*{Mean occupation density}
+
+XXX OMIT FROM HERE? XXX
 
@@
 u(x_0,x) &= \frac{\alpha^{(d-2)/2}}{\sigma^d \pi^{d/2}}\left(\frac{ \sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(2-d)/2} K_{(2-d)/2}\left( \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right) \\
-         &= \frac{\alpha^{(d-2)/2}}{\sigma^d 2 \pi^{(d-1)/2}} \left( \frac{\sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(1-d)/2} \exp\left(- \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right)+ O(\sigma/\|x-x_0\|) .
-\label{eqn:occupation_asymptotics}
+         &= 
+    \exp\left(- \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right)
+    \left\{
+    \frac{\alpha^{(d-2)/2}}{\sigma^d 2 \pi^{(d-1)/2}} \left( \frac{\sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(1-d)/2} 
++ O\left( \left(\frac{\sqrt{2\alpha}\|x-x_0\|}{\sigma}\right)^{-(d+1)/2} \right) 
+    \right\}
+    . \label{eqn:occupation_asymptotics}
 \end{align}
@@
 so that $\gamma(r_0) = q(x) / \int_{B(r_0)} u(y,x) dy$, for any choice of $x$.
-XXX Green function for BM killed on disk XXX
-Now, using equations \eqref{eqn:eqfreq} and \eqref{eqn:occupation_asymptotics}, we get that
+Now, using equations \eqref{eqn:eqfreq} and \eqref{eqn:occupation_asymptotics}, 
+we can obtain that (see Appendix \ref{apx:lesbegue})
+the outflux of families 
 \begin{align}
@@
 
+\subsection{Temporary work: outflux of families}
+\label{apx:lesbegue}
+
+We have that
+\[
+    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
+\]
+for any $y$.  
+Take $y=(r,0)$.
+Note that after scaling by $\sqrt{2\alpha}/\sigma$,
+\begin{align}
+    \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\|} }{ \|y\|^{-(d-1)/2} e^{-\|y\|} } \\
+        &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r (\|\hat x - x/r\| - 1)} .
+\end{align}
+Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
+so that the expression in the exponent is to, first order in $1/r$,
+\begin{align}
+    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
+            &= - 2 r_0 \cos \theta + O(1/r) .
+\end{align}
+This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
+\begin{align}
+    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} ,
+\end{align}
+and hence, by bounded convergence, then \citet{gradshteyn2007table} (3.339),
+\begin{align}
+    \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
+                                         &= 2 \pi C' I_0(-2r_0),
+\end{align}
+where $I_0$ is a modified Bessel function of order zero of the first kind.
+The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2\alpha} r_0 / \sigma ) $.
+
+\subsection{Temporary work: hitting probs}
+
+Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
+which we assume has relatively small area $A$ and is located near position $y$.
+The probability that this family reaches $A$ by time $t$ is $(1-p_e(t))$
+multiplied by the probability that the family reaches the patch, given survival.
+If $t$ is large then the previous section argues that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
+and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
+will have hit the patch by then.
+If $w = A^{1/d}$ is the width of the patch, then 
+this is the chance that a $d$-dimensional Bessel process begun at $\|y-x\|$ hits $w$ by time $t$.
+Let $\tau_w$ be the hitting time of $w$,
+so, integrating by parts, and using Borodin \& Salminen (2.2.0.1 and 4.2.0.1)
+\begin{align}
+    \int_0^\infty (1-p_e(t)) \P^x\{ \tau_w \le t \} dt  
+    &\approx \int_0^\infty e^{-\alpha t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\
+    &= \frac{1}{\alpha \E[K]} \int_0^\infty e^{-\alpha t} \P^x\{ \tau_w \in dt \} \\
+    &= \frac{1}{\alpha \E[K]} \E^x[e^{-\alpha \tau_w}] \\
+    &= 
+    \frac{1}{\alpha \E[K]} \times
+    \begin{cases}
+        e^{-(x-w) \sqrt{2\alpha}} \qquad & d=1 \\
+        \frac{ K_0(x\sqrt{2\alpha}) }{ K_0(w\sqrt{2\alpha}) } \qquad & d=2 
+    \end{cases}
+\end{align}
+Note that the $d=2$ case
+\begin{align}
+    \frac{ K_0(x\sqrt{2\alpha}) }{ K_0(w\sqrt{2\alpha}) } = \sqrt{\frac{w}{x}} e^{-(x-w)\sqrt{2\alpha}} ( 1 + O(1/w) ).
+\end{align}
+
 \end{document}

### 2014-04-13 patchy-selection-paper.tex
.

@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\one}{\mathbb{1}}
 \newcommand{\calE}{\mathcal{E}}
@@
 
-We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary}.
+We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} \citep[see also][]{chauvin1991growing}.
 When we condition on survival until $t$, we condition on existence of at least one lineage surviving from time $0$ to time $t$.
@@
 
+Now note that if $\tau$ is Exponential($\alpha$), then
+\begin{align}
+  q(y) &= \frac{1}{\alpha} \int_{\partial B} \E^x[ \int_0^\tau \one_A(B_t) dt ] dx 
+\end{align}
+and that if $\mu(x,A)$ is the hitting distribution of $\partial A$ by $B_{t \min \tau}$,
+\begin{align}
+  \E^x[ \int_0^\tau \one_A(B_t) dt ] &= \P^x\{ \tau_A < \tau \}  \E^{\mu(x,A)}[ \int_0^\tau \one_A(B_t) dt ] .
+\end{align}
+If $A$ is circular, then the latter term is the integral of the Green function, so using Borodin \& Salminen (Appendix 1),
+in $d=1$,
+\begin{align}
+  \E^w[ \int_0^\tau \one_A(B_t) dt ] &= \alpha \int_0^{2w} G_\alpha(0,y) dy \\
+  &= \alpha \int_0^{2w} \frac{ e^{- y \sqrt{2\alpha} }}{2\sqrt{2\alpha}} dy
+\end{align}
+and in $d=2$,
+\begin{align}
+  \E^w[ \int_0^\tau \one_A(B_t) dt ] &= \alpha \int_0^{2w} G_\alpha(0,y) dy \\
+  &= \alpha \int_0^{2w} K_0(y \sqrt{2\alpha}) dy 
+\end{align}
+
 \end{document}

### 2014-04-14 patchy-selection-paper.tex
.

--- 2014-04-11 patchy-selection-paper.tex
@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\one}{\mathbf{1}}
 \newcommand{\calE}{\mathcal{E}}
@@
 
+\subsection{Temporary}
+
+Consider a single migrant family $Z$ beginning at $x$.
+Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
+and $\tau_\dagger$ an independent Exponential($\alpha$) time.
+The mean occupation time of $Z$ spent in a region $A$ is 
+\begin{align}
+    u(x,y) &= \E[\int_0^\infty Z_t(A) dt ] \\
+           &= \int_0^\infty \E[Z_t] p_t(x,A) dt \\
+        &= \int_0^\infty e^{-\alpha t} p_t(x,A) dt \\
+        &= \int_0^\infty \P^x\{ \tau_\dagger < t \; \& \; B_t \in A \} dt \\
+        &= \E[ \int_0^{\tau_\dagger} \one_A(B_t) dt ]  .
+\end{align}
+If we define
+$\tau_A$ to be the hitting time of $A$ by $B$,
+and $\mu(x,A)$ to be the hitting distribution of $\partial A$ by $B_{\min(t,\tau)}$,
+by the strong Markov property this is equal to
+\begin{align}
+    u(x,y) &= \E^x[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] \\
+           &= \P^x\{ \tau_A < \tau_\dagger \}  \E^{\mu(x,A)}[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] .
+\end{align}
+If $A$ is circular, the latter expectation will not depend on $x$;
+call this $h(A)$.
+
+Now consider the chance that $Z$ ever hits $A$.
+If we make the approximation that $Z$ hits $A$ iff the trunk of $Z$ does,
+then this is equal to
+\begin{align}
+    F(x,y) &= \int_0^\infty (1-p_e(t)) \P\{ \tau_A \in dt \} \\
+           &= \int_0^\infty e^{\alpha t} (1-p_e(t)) e^{- \alpha t } \P\{ \tau_A < t \} dt \\
+    &=\int_0^\infty e^{\alpha t} (1-p_e(t)) \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &= \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &= \frac{1}{\E[K]} \P^x[ \tau_A < \tau_\dagger ] .
+\end{align}
+Therefore, we have that
+\begin{align}
+    F(x,y) \E[K] 
+    \approx \P^x[ \tau_A < \tau_\dagger ] 
+    = u(x,y) / h(A) .
+\end{align}
+
+Now, 
+\begin{align}
+    q(y)  &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
+          &= h(A) \int_{\partial B(r_0)} \gamma(x) F(x,y) dx \\
+          &= \frac{ h(A) }{ \E[K] } \migrate(y) .
+\end{align}
+The function $h(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
+is expected to spend inside the disk before $\tau_\dagger$.
+This is integral of the Green function for the Bessel process of the appropriate order, 
+so using Borodin \& Salminen (Appendix 1),
+and letting $w$ be the width of the patch,
+in $d=1$,
+\begin{align}
+    h_1(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+  % &= \int_0^{2w} G_\alpha(0,y) dy \\
+  &= \int_0^{2w} \frac{ e^{- y \sqrt{2\alpha} }}{\sqrt{2\alpha}} dy \\
+  &= (1-e^{-2w\sqrt{2\alpha}})/(2\alpha)
+\end{align}
+and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
+\begin{align}
+    h_2(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_\alpha(0,y) dy \\
+  &= \int_0^{2w} 2 y K_0(y \sqrt{2\alpha}) dy \\
+  &= \frac{1}{\alpha}\int_0^{2w\sqrt{2\alpha}} y K_0(y) dy \\
+  &= \frac{1}{2\alpha}\left(1- 2w\sqrt{2\alpha} K_1(2w\sqrt{2\alpha}) \right)  .
+\end{align}
+
 \end{document}

### 2014-04-19 patchy-selection-paper.tex
.

@@
 as the sum of independent migrant families, each of whose dynamics are given by a spatial branching process.
-Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\delta B(r_0)$ its boundary.
-Suppose that there is a constant outflux of migrant families of rate $\gamma(x)$ from each point $x \in \delta B(r_0)$,
-e.g.\ with locations and times of appearances as the points of a Poisson process.
-Roughly speaking, equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
-differ only by substituting the average number of members of a migrant family in the new patch
-for the probability that a family has any members in the new patch;
-by the discussion of the previous section, these differ only by a constant factor.
-
-First we will make this idea more formal.
+Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
+Suppose that there is a constant outflux of migrant families of rate $\gamma(x)$ from each point $x \in \partial B(r_0)$.
+To make expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} more concrete,
+we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
+of the outflux $\gamma$ at that point, mulitplied by the relevant second quantity.
+By examining the constituent terms,
+it will turn out that \eqref{eqn:gestalt_migrate} is a constant multiple of \eqref{eqn:gestalt_q}.
+
+First consider \eqref{eqn:gestalt_q}.
 Suppose that $Z$ is one such spatial branching process as above, started with a single individual at $x$,
 defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the set $A$.
-The \emph{mean occupation density} of $Z$ is then defined to be the density of $\E[\int_0^t Z_t dt]$,
+The \emph{mean occupation density} of $Z$ can be thought of 
+as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
+Concretely, it is the density of $\E[\int_0^t Z_t dt]$,
 i.e.\ the smooth function $u(x,y)$ that, integrated over a set $A$, 
@@
 \end{align}
-We can think of $u(x,y)$ as the expected total number of offspring of a family beginning at $x$
-that ever live at $y$.
-By independence of reproduction and spatial motion,
-$u(x,y)$ is the sum over all times of the expected number of offspring
+Since reproduction and spatial motion are independent,
+$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
 multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
@@
 \end{align}
-Now, \eqref{eqn:gestalt_q} is
-\begin{align}
-    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
-         &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-\alpha t} p_t(x,y) dt dx .
-\end{align}
-
-Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
-which we assume has relatively small area $A$ and is located near position $y$.
-The probability that this family reaches $A$ by time $t$ is $(1-p_e(t))$
-multiplied by the probability that the family reaches the patch, given survival.
-If $t$ is large then the previous section argues that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
-and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
-will have hit the patch by then.
-If $w = A^{1/d}$ is the width of the patch, then
+The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
+is the sum of mean occupation densities of a constant outflux of branching processes
+from $\partial B(r_0)$:
+\begin{align} \label{eqn:occupation_integral}
+    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  % \\
+         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-\alpha t} p_t(x,y) dt dx .
+\end{align}
+
+Now we turn to the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
+Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
+which we assume is circular, has relatively small area $A$ and is located near position $y$.
+% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-p_e(t))$,
+% multiplied by the probability that the family reaches the patch, given survival.
+% If $t$ is large then as in the previous section, $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$.
+It would be possible to analyze directly the probability that a given migrant family establishes in $A$,
+as in \citet{barton1987probability,branchingbrownian};
+but we make a simpler approximation, approximating this by the chance that the ``trunk'' hits the new patch,
+multiplied by the chance that at least member of the family escapes demographic stochasticity to establish.
+Write $h(x,y)$ for the probability that the Brownian trunk
+hits the new patch centered at $y$ before the family dies out,
+and $f(A)$ for the chance that the family manages to establish in the new patch,
+given that it successfully arrives.
+As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is an integral of $h(x,y)$:
+\begin{align}\label{eqn:migrate_integral}
+  \migrate(y) &= \int_{\partial B(r_0)} \gamma(x) h(x,y) f(A) dx .
+\end{align}
+
+Intuitively, $h(x,y)$ counts each family once if it hits the patch,
+while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
+Since the expected total time spent in the patch does not depend on the distance at which the family started
+if we know that it survived to hit the patch,
+it is intuitively reasonable that $u(x,y)$ and $h(x,y)$ are simply related.
+Using this fact,
+and that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
+it turns out (see Appendix~\ref{apx:hitting_occupation}) that
+\begin{align} 
+  h(x,y) = \frac{ s(A) }{ \E[K] } u(x,y),
+\end{align}
+where $s(A)$ is the total expected occupation time of the patch given the family arrives,
+and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
+Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a constant factor
+(at least asymptotically for large distances),
+we now know that the outflux of successful migrant families is equal to the mean occupation frequency
+multiplied by $s(A) f(A) / \E[K]$.
+The probability $f(A)$ that the family does not die out once it arrives at the new patch
+is approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$.
+% If we let $w$ be the width of the patch, then the remaining term is
+% \begin{align}
+%   s(w) \simeq \frac{1}{2\alpha} \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
+% \end{align}
+Combining these terms with the expression for $s(A)$ from Appendix~\ref{apx:hitting_occupation},
+\begin{align}
+  \migrate(y) \approx q(y) \times \frac{ s_p  }{ \alpha \xi^2 } 
+  \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
+\end{align}
 

### 2014-04-20 patchy-selection-paper.tex
.

@@
 Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
-Suppose that there is a constant outflux of migrant families of rate $\gamma(x)$ from each point $x \in \partial B(r_0)$.
+Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
+such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ that initiate new migrant families
+is $\gamma(x)$.
 To make expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} more concrete,
@@
 \begin{align}\label{eqn:migrate_integral}
-  \migrate(y) &= \int_{\partial B(r_0)} \gamma(x) h(x,y) f(A) dx .
+  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h(x,y) f(A) dx .
 \end{align}
@@
 \begin{align}
-  \migrate(y) \approx q(y) \times \frac{ s_p  }{ \alpha \xi^2 } 
-  \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
-\end{align}
+  \label{eqn:migrate}
+  \migrate(y) \approx q(y) \times \rho \frac{ s_p  }{ \alpha \xi^2 } 
+  % \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
+\end{align}
+For patches that are small relative to $\sigma/\sqrt{2s_m}$, there is an additional correction term (see Appendix~\ref{apx:hitting_occupation}),
+but at this point we would be violating various other assumptions above, so we omit it.
 
@@
 the rate of establishment, for a patch of area $A$ at distance $r$ from an already occupied patch, is approximately
-\begin{align} \label{eqn:migrate}
+\begin{align} % \label{eqn:migrate}
     \migrate(r) = \gamma F(r) = \frac{\rho \sigma s_p\sqrt{2s_m} }{\xi^2}  \exp\left( -\frac{ \sqrt{2 s_m} r}{\sigma} \right).
@@
 As such, the question of how the the second patch adapts
-simply comes down to whether the a new mutation or a migrant allele is the first to become established in the second patch.
+simply comes down to whether a new mutation or a migrant allele is the first to become established in the second patch.
 
@@
 the rate of migration and mutation will be on the same order if 
-$\mu \approx ((s_p+s_m)/2) \exp(-\sqrt{2 s_m} r / \sigma)$,
-\plr{check}
-where $r$ is the distance between the patches.
-(Note that in both cases, $A$ is the area of the not-yet-adapted patch.)
-
-If we take $\mu = 10^{-8}$, and $s_p=s_m=s$, then this is satisfied if $r/\sigma \approx (18+\log(s))/\sqrt{s}$,
-so if the selective pressure is fairly strong (say, $s=.05$),
-then the distance between patches must be about 70 times the dispersal distance (i.e.\ $r/\sigma \approx 68$);
-while if $s$ is much smaller, say $s = .001$, 
-then the patches must be separated by many hundreds of times the dispersal distance (i.e.\ $r/\sigma \approx 368$).
-The situation does not change much if $\mu$ is larger -- even if $\mu = 10^{-6}$, 
-the required separation between patches is only reduced by $5/\sqrt{s_m}$.
-If the separation between patches is much larger than this value -- certainly, by a factor of ten -- 
+$A \mu \approx (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
+where $R$ is the distance between the patches,
+and recall that $A$ is the area of the not-yet-adapted patch.
+In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
+is $\rho (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
+which depends fairly strongly on the selective disadvantage $s_m$ between patches.
+Equivalently, migration will occur much quicker than mutation 
+if $R/\sigma > (\log(1/s_m) - \log(A\mu))/\sqrt{2s_m}$.
+
+If we take the mutational influx in a patch to be one per 1,000 generations ($A \mu = .001$),
+and the selection against transiting alleles $s_m = .01$,
+adaptation procedes independently for distances $R/\sigma > \log(s_m)/\sqrt{s_m}$,
+so if the selective pressure is fairly strong (say, $s_m=.05$),
+then the distance between patches must be about 30 times the dispersal distance (i.e.\ $R/\sigma > 30$),
+while if $s_m$ is much smaller, say $s_m = .001$, 
+then the patches must be separated by several hundred times the dispersal distance (i.e.\ $R/\sigma > 300$).
+The situation does not change much if the influx is quite a bit higher -- even if $A \mu = 1$,
+the required separation between patches is only reduced by $7/\sqrt{s_m}$,
+to $R/\sigma > 10$ with $s_m=.05$ and $R/\sigma>150$ for $s_m=.001$.
+If the separation between patches is much larger than this transitional value -- certainly, by a factor of ten -- 
 then patches are effectively isolated -- most adaptation will occur independent between patches.
@@
 we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
-as a function of population density, mutation rate, and distance between patches (scaled by $\sqrt{s_m}/\sigma$).
+as a function of various parameters.
 
@@
   \caption{
-  Mean time until adaptation in a new patch occurs by migration ($\Tmig$, blue) or by mutation ($\Tmut$, red),
-  as a function of population density ($\rho$), mutation rate ($\mu$), 
-  and distance to an already adapted patch ($R$), scaled by $\sqrt{s_m}/\sigma$.
+  Shaded contour plots of
+  mean time until adaptation in a new patch occurs by migration ($\Tmig$, blue) or by mutation ($\Tmut$, red),
+  as a function of population density ($\rho$), 
+  distance to an already adapted patch in units of dispersal distance ($R/\sigma$),
+  strength of deleterious selection ($s_m$),
+  and mutation rate scaled by patch area ($\mu A$).
+  The line at which $\Tmig=\Tmut$ is shown in black.
+  Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=50$, and $\xi^2=1$.
   Note that population density and mutation rate are shown on a log scale.
-  \plr{fix axis labels}
   \label{fig:phase_diagram}
@@
 will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
-Both these times scale linearly with $\xi^2/(\rho A s_{B})$, \plr{check}
+Both these times scale linearly with $\xi^2/(\rho s_{B})$, 
 so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
-and decrease with patch population size and benefical selection coefficient.
+and decrease with population density and benefical selection coefficient.
 Furthermore, the probability that the second adaptation is a new mutation,
-i.e.\ the probability of parallel adapatation, is
+i.e.\ the probability of parallel adapatation, is in $d=1$ ($d=2$ only differs slightly)
 \begin{equation}
-  \frac{\mutrate}{\mutrate+\migrate} = \frac{2\mu/(s_p+s_m)}{2\mu/(s_p+s_m) + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
+  \frac{\mutrate}{\mutrate+\migrate} = \frac{2 A \mu s_m}{2 A \mu s_m + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
 \end{equation}
-\plr{check}
 so that the probability of parallel mutation should increase
@@
 and these patches are at distances $r_1, \ldots, r_k$ away from this unadapted patch, 
-then the total rate of adaptation through migration is
+then the total rate of adaptation through migration is approximately
 \begin{equation}
-  \frac{ s_p (s_p+s_m) \rho A}{\xi^2} \sum_{i=1}^{k} \exp\left(- \sqrt{2 s_m} r_i/\sigma\right),
+  \frac{ s_p \rho }{s_m \xi^2} \sum_{i=1}^{k} \exp\left(- r_i \sqrt{2 s_m} /\sigma\right),
 \end{equation}
 and the probability that the migrant, adapted allele 
-comes from patch $i$ is $\exp\left(- \sqrt{2 s_m} r_i/\sigma\right)/\sum_{j=1}^{k} \exp\left(- \sqrt{2 s_m} r_j/\sigma\right)$.
+comes from patch $i$ is 
+\begin{align}
+  \exp\left(- r_i \sqrt{2 s_m} /\sigma\right)/\sum_{j=1}^{k} \exp\left(- r_j \sqrt{2 s_m} /\sigma\right) .
+\end{align}
 These specify the transition rates of the process.
@@
 
-\subsection{Temporary}
+\subsection{Hitting and occupation}
+\label{apx:hitting_occupation}
 

### 2014-04-20 patchy-selection-fns.R

@@ -43,15 +43,12 @@ minimumPatch <- function(mu, rho, s, gb, gm, sigma, R, A) {
 
 mutationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
     # approximation for mutational influx per patch
-    # XXX to-do: do numerics?
     return( 2 * s*gb * A * rho * mu )
 }
 
 migrationInflux <- function(mu, rho, s, gb, gm, sigma, R, A) {
     # approximation for migrational influx per patch
-    # XXX to-do: do numerics?
-    # return( 2 * s*gb * A * rho * exp( - sqrt(abs(s*gm))*R/sigma ) )
-    return( s*gb * sqrt(s*gm/pi) * rho * exp( - sqrt(abs(s*gm))*R/sigma ) )
+    return( (1/2) * (s*gb / s*gm) * rho * (1/sqrt(sqrt(abs(s*gm))*R/sigma)) * exp( - sqrt(abs(s*gm))*R/sigma ) )
 }
 
 

### 2014-04-20 patchy-selection-plots.R

@@ -9,21 +9,26 @@ require(xtable)
 # Plot level sets of time to mut and mig
 #  first against characteristic dist and pop density
 #  then agains mutation rate and pop density
-rhovals <- logseq(1,100000,length.out=100)
+rhovals <- logseq(1,1e5,length.out=100) # per km^2
 clvals <- seq(1,20,length.out=100)
-muvals <- logseq(1e-8,1e-6,length.out=100)
-A <- 10000 # km^2
-mu <- 1e-8
+muvals <- logseq(1e-8,1e-6,length.out=100) 
+smvals <- logseq(1e-6,1e-1,length.out=100)
+A <- 100 # per km^2
+rho <- 1e2  # 
+mu <- 1e-8 
 sb <- .1
 sd <- .01
 const <- .5
 cl <- 5  # cline length
 clplot <- expand.grid(cl=clvals,rho=rhovals)
 clplot$mut <- with(clplot, 1/(A*mu*2*sb*rho) )
-clplot$mig <- with(clplot, 1/(const*2*sd*(sd+sb)*rho*exp(-cl)) )
+clplot$mig <- with(clplot, 1/(const*(sb/sd)*rho*(1/sqrt(cl))*exp(-cl)) )
 muplot <- expand.grid(mu=muvals,rho=rhovals)
 muplot$mut <- with(muplot, 1/(A*mu*2*sb*rho) )
-muplot$mig <- with(muplot, 1/(const*2*sd*(sd+sb)*rho*exp(-cl)) )
+muplot$mig <- with(muplot, 1/(const*(sb/sd)*rho*(1/sqrt(cl))*exp(-cl)) )
+smplot <- expand.grid(mu=muvals,sm=smvals)
+smplot$mut <- with(muplot, 1/(A*mu*2*sb*rho) )
+smplot$mig <- with(muplot, 1/(const*(sb/sd)*rho*(1/sqrt(cl))*exp(-cl)) )
 f <- function (x) { matrix(x,nrow=(sqrt(length(x)))) }
 pminmax <- function (x,lower,upper) { pmin(upper,pmax(lower,x)) }
 
@@ -44,9 +49,13 @@ for (lev in levelsets) { # note rho is log10-scale.
     miglevs <- 1/(const*2*sd*(sb+sd)*exp(-clvals)*lev)  # migration:
[…]

### 2014-04-20 patchy-selection-plots.R
.

@@ -9,26 +9,28 @@ require(xtable)
 # Plot level sets of time to mut and mig
 #  first against characteristic dist and pop density
 #  then agains mutation rate and pop density
+muttime <- function (env) {with(env,1/(A*mu*2*sb*rho))}
+migtime <- function (env) {with(env,1/(const*(sb/sd)*rho*(1/sqrt(cl*sqrt(sd)))*exp(-cl*sqrt(sd))))}
 rhovals <- logseq(1,1e5,length.out=100) # per km^2
-clvals <- seq(1,20,length.out=100)
-muvals <- logseq(1e-8,1e-6,length.out=100) 
-smvals <- logseq(1e-6,1e-1,length.out=100)
-A <- 100 # per km^2
-rho <- 1e2  # 
+clvals <- seq(1,200,length.out=100)
+muvals <- logseq(1e-9,1e-4,length.out=100) 
+sdvals <- logseq(1e-4,1e-1,length.out=100)
+A <- 100 # km^2
+rho <- 10  # per km^2
 mu <- 1e-8 
 sb <- .1
 sd <- .01
 const <- .5
-cl <- 5  # cline length
+cl <- 50  # R/sigma
 clplot <- expand.grid(cl=clvals,rho=rhovals)
-clplot$mut <- with(clplot, 1/(A*mu*2*sb*rho) )
-clplot$mig <- with(clplot, 1/(const*(sb/sd)*rho*(1/sqrt(cl))*exp(-cl)) )
+clplot$mut <- muttime(clplot)
+clplot$mig <- migtime(clplot)
 muplot <- expand.grid(mu=muvals,rho=rhovals)
-muplot$mut <- with(muplot, 1/(A*mu*2*sb*rho) )
-muplot$mig <- with(muplot, 1/(const*(sb/sd)*rho*(1/sqrt(cl))*exp(-cl)) )
-smplot <- expand.grid(mu=muvals,sm=smvals)
-smplot$mut <- with(muplot, 1/(A*mu*2*sb*rho) )
-smplot$mig <- with(muplot, 1/(const*(sb/sd)*rho*(1/sqrt(cl))*exp(-cl)) )
+muplot$mut <- muttime(muplot)
+muplot$mig <- migtime(muplot)
+sdplot <- expand.grid(mu=muvals,sd=sdvals)
+sdplot$mut <- muttime(sdplot)
+sdplot$mig <- migtime(sdplot)
[…]

### 2014-04-21 patchy-selection-paper.tex
draft! cleaned up!

@@
 relative to the patch fixes within the poor habitat patch.
-As shown in Appendix \ref{apx:elliptic_integrals} XXX,
-the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
-but the precise expressions are not particularly helpful.
-The total successful mutational influx per generation is then given by the integral of $p(x)$ over the entire species range,
+Under various assumptions, precise expressions can be found for $p(x)$ \citep{barton1987establishment},
+% the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
+but these are not particularly helpful.
+The total successful mutational influx per generation is given by the integral of $p(x)$ over the entire species range,
 % $\int \rho \mu p(x) dx$,
@@
 which in one dimension can be solved to give $q(x)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
-More general, asymptotics of the solution implies that for large $|x|$ the mean occupation frequency decays exponentially, with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}):
+More general, asymptotics of the solution implies that for large $|x|$ the mean occupation frequency decays exponentially, 
+with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}): 
 \begin{align} \label{eqn:eqfreq}
-    q(x) \approx C |x|^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
-\end{align}
-where $d$ is the dimension ($d=1$ or $2$), and $C$ and $r_0$ are constants depending on the geographic shape of the populations and the selection coefficients
-(but note that $r_0$ is only necessary in $d=2$).
-In applications we fit $C$ and $r_0$ to the data;
-to get concrete numbers from theory we take $C=1$ and $r_0=0$ if necessary.
+  q(x) \approx C \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
+\end{align}
+where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients.
+In applications we fit $C$ to the data;
+to get concrete numbers from theory we take $C=1$ if necessary.
 From simulations, $C$ seems to be not too far from 1 in practice (see Appendix \ref{apx:patchy_sims}).
@@
 or in other words, the total number of $B$ alleles found at distance $r$ across $T$ generations, divided by $T$.
-Therefore, if we look at a small area of size $\epsilon$, 
-then $q(r) \epsilon$ is equal to the probability of finding any $B$ alleles there,
+Therefore, if we look at a small area of size $dx$, 
+then $q(r) dx$ is equal to the probability of finding any $B$ alleles there,
 multiplied by the mean number of such, if there are any.
@@
 It would be possible to analyze directly the probability that a given migrant family establishes in $A$,
-as in \citet{barton1987probability,branchingbrownian};
+as in \citet{barton1987establishment,mckean1975KPP};
 but we make a simpler approximation, approximating this by the chance that the ``trunk'' hits the new patch,
@@
 The probability $f(A)$ that the family does not die out once it arrives at the new patch
-is approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$.
+is if each member arriving has chance $p_e$ of establishing, independently, $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$.
 % If we let $w$ be the width of the patch, then the remaining term is
@@
 For patches that are small relative to $\sigma/\sqrt{2s_m}$, there is an additional correction term (see Appendix~\ref{apx:hitting_occupation}),
-but at this point we would be violating various other assumptions above, so we omit it.
-
-%%%%%%% %%%%%%%%%%%
-\subsubsection*{Mean occupation density}
-
-XXX OMIT FROM HERE? XXX
-
-First, we compute the second term in equation \eqref{eqn:gestalt_q}, (mean occupation density of a family at $x$),
-by which we mean the expected total number of offspring of a single family that ever live at distance $r$ from the patch.
-In two dimensions, this depends on where the family began, 
-and right-hand side of equation \eqref{eqn:gestalt_q} should be integrated over this starting point.
-Here we'll write $x_0$ for the starting point, and $u(x_0,x)$ for the mean occupation density at $x$.
-Concretely, suppose that $Z$ is a spatial branching process as above, started with a single individual at $x_0$,
-defined now so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the set $A$.
-The mean occupation density of $Z$ is then defined to be the density of $\E[\int_0^t Z_t dt]$,
-i.e.\ the smooth function $u(x_0,x)$ that, integrated over a set $A$, 
-gives the expected total number of individuals ever living in $A$:
-\begin{align} \label{eqn:occupation_density_defn}
-    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x_0,x) dx .
-\end{align}
-Since a family has on average $(1-s_m)^t$ members in generation $t$,
-and the marginal distribution of any one surviving individual is approximately Gaussian with variance $t \sigma^2$,
-exchanging the expectation and the integral shows that
-the density $u(x_0,x)$ is equal to the Green's function for Brownian motion killed at rate $1-s_m$,
-\begin{align} \label{eqn:occupation_density_solution}
-    u(x_0,x) = \int_0^\infty (1-s_m)^t p_{\sigma^2 t}(x_0,x) dt
-\end{align}
-where $p_t(x,y) = (2 \pi t)^{-d/2} \exp(-\|x-y\|^2/2t)$ is the transition density for standard Brownian motion.
-The integral can be expressed in terms of modified Bessel functions:
-using \citet{gradshteyn2007table} (3.471.9 and 8.451.6), the mean occupation density is,
-with $\alpha = -\log(1-s_m)$ and $d$ the dimension,
-\begin{align}
-u(x_0,x) &= \frac{\alpha^{(d-2)/2}}{\sigma^d \pi^{d/2}}\left(\frac{ \sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(2-d)/2} K_{(2-d)/2}\left( \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right) \\
-         &= 
-    \exp\left(- \sqrt{2\alpha} \frac{\|x-x_0\|}{\sigma} \right)
-    \left\{
-    \frac{\alpha^{(d-2)/2}}{\sigma^d 2 \pi^{(d-1)/2}} \left( \frac{\sqrt{2 \alpha} \|x-x_0\|}{\sigma} \right)^{(1-d)/2} 
-+ O\left( \left(\frac{\sqrt{2\alpha}\|x-x_0\|}{\sigma}\right)^{-(d+1)/2} \right) 
-    \right\}
-    . \label{eqn:occupation_asymptotics}
-\end{align}
-Since we only have asymptotics for the equilibrium frequency, 
-we will only use the latter (asymptotic) form.
-
-%%%%%%% %%%%%%%%%%%
-\subsubsection*{Outflux of families}
-
-Now, we can obtain the first term in equation \eqref{eqn:gestalt_q}, the (outflux of families).
-Define $\gamma(r_0)$ to be this outflux; 
-i.e.\ the mean number of individuals per unit time that are born carrying the $B$ allele closer than $r_0$ to the original patch
-but migrate to a larger distance.
-Then for any point $x$ further than $r_0$ from the patch, 
-the mean occupation density $q(x)$ is equal to the integral of $\gamma$ over all locations $x_0$ at distance $r_0$ from the patch,
-multiplied by the mean occupation density at $x$ for migrant families beginning at at location.
-This is a more precise statement of the decomposition of the equilibrium frequency from equation \eqref{eqn:gestalt_q}:
-\begin{align}
-    q(x) &= \gamma(r_0) \int_{B(r_0)} u(y,x) dy,
-\end{align}
-so that $\gamma(r_0) = q(x) / \int_{B(r_0)} u(y,x) dy$, for any choice of $x$.
-Now, using equations \eqref{eqn:eqfreq} and \eqref{eqn:occupation_asymptotics}, 
-we can obtain that (see Appendix \ref{apx:lesbegue})
-the outflux of families 
-\begin{align}
-    \gamma(r_0) \simeq 
-\end{align}
-
-Denote by $Z_t$ the number of individuals such a family has after time $t$,
-and the locations of these by $X_1(t), \ldots, X_{Z_t}(t)$.
-Then this mean occupation density, at $x$, is the expected number of individuals ever present at $x$. 
-Since the expected number of individuals present at time $t$ is $e^{\alpha t}$
-and the location of each of these individuals is marginally distributed as a  Gaussian with variance $\sigma^2 t$,
-we can write this expectation as
-\begin{align}
-    u(x) &= \int_0^t \E[ \#\{ k : \; X_k(t) \in dx \} ] dt \\
-    &= \int_0^t e^{\alpha t} \frac{1}{\sqrt{2 \pi t \sigma^2}} e^{-\frac{x^2}{2t\sigma^2}} dt \\
-    &= \frac{ 1 }{ \sigma \sqrt{2 \alpha} } \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) 
-\end{align}
-\plr{check that}
-The integral is evaluated in Appendix~\ref{apx:some_integrals}.
-\plr{Also examine non-dependence on $r_0$.}
-
-Combining this with the solution for $q(r)$ (equation~\eqref{eqn:eqfreq}) as in \eqref{eqn:q_gestalt},
-we then have that (with $C=1/2$)
-\begin{align} \label{eqn:outflux}
-    \gamma = \text{( outflux of families )} \approx (1/2) \rho \sigma \sqrt{2\alpha} .
-\end{align}
-
-\subsubsection*{Probability of establishment}
-
-Suppose that the new patch is at distance $r$.
-The probability density for the first time such a Brownian motion $B_s$ hits $r$ is
-\citep[XXX]{feller}
-\plr{work in $d=2$}
-% note $\sigma B_t \deq B_{\sigma^2 t}$ so $\P\{ \sup_{s\le t} \sigma B_s \le r \} = \P\{ \sup_{s \le \sigma^2 t} B_s \le r \}
-\begin{align} \label{eqn:hitting_dist}
-  \P\{ B_t=r \;\mbox{and} \; B_u<r \;\mbox{for all}\; u<t\} =  \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2}\right) .
-\end{align}
-Let $\tau$ be this random time at which a given migrant family hits the new patch,
-with the convention that $\tau =\infty$ if the family never gets here.
-We approximate the chance that a given family arrives at the new patch at time $t$, i.e.\ that $\tau=t$,
-by the product of equation~\eqref{eqn:hitting_dist} and the probability that the migrant family survives until time $t$.
-For convenience, define $\alpha := - \log(1-s_m)$ to be the rate of decrease between patches (note that $\alpha \approx s_m$),
-and $\mu_K = \E[K]$ to be the limiting mean family size conditioned on survival,
-so that $1-p_e(t) \simeq (1/\mu_K) \exp(-\alpha t)$.
-Then, the probability distribution of $\tau$ is given by
-\begin{align}
-    \P\{ \tau \in dt \} =  \frac{1}{\mu_K} \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2} - \alpha t \right) dt  .
-\end{align}
-
-By integrating over~$t$, the chance that this family ever arrives at the new patch is
-\begin{align} 
-\label{eqn:estab_prob}
-  \begin{split}
-      \P\{ \tau < \infty \} &= \int_0^\infty \P\{ \tau \in dt \} = \frac{1}{\mu_K} \exp\left( - \frac{ r \sqrt{2 \alpha}}{\sigma} \right) .
-  \end{split}
-\end{align}
-This is, up to the constant $\mu_K$, the chance that a Brownian motion killed at rate $\alpha$
-ever hits $r$ \citep{feller}.
-This approximation depends on $r$ being large enough so that the approximation $1-p_e(t) \approx (1-s_m)^t/\mu_K$ holds.
-
-Now, each member of a family that successfully arrives at the new patch
-(where the $B$ allele is again beneficial)
-has a chance $p_e$ of establishing.
-The precise distribution of the number of members of a family that enter the patch
-given the trunk arrives there
-depends on the geometry of the patch,
-but since family members should not be too far from the trunk,
-we can approximate this by the total size of the family.
-If the family has $K$ individuals when the trunk arrives at the new patch,
-the chance that any establish is approximately $1-(1-p_e)^K$;
-and so that chance that the family establishes, given it arrives at the new patch, 
-is 
-\begin{align}
-    \label{eqn:prob_estab_arrival}
-    \E[1-(1-p_e)^K] \approx 2 s_p \mu_K/\xi^2 .
-\end{align}
-
-The probability of establishment of an outmigrant from the already-established patch is 
-equal to the probably the trunk arrives (equation~\eqref{eqn:estab_prob}) multiplied by the probability it establishes (equation~\eqref{eqn:prob_estab_arrival}),
-namely
-\begin{align} \label{eqn:bp_prob_estab}
-    F(r) &:= \text{( prob family establishes at $r$ )} = \frac{2s_p}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} r}{\sigma} \right) ,
-\end{align}
-
-\subsubsection*{Putting it all together}
-
-Combining this with \eqref{eqn:bp_prob_estab}, as in \eqref{eqn:gestalt_q}, 
-the rate of establishment, for a patch of area $A$ at distance $r$ from an already occupied patch, is approximately
-\begin{align} % \label{eqn:migrate}
-    \migrate(r) = \gamma F(r) = \frac{\rho \sigma s_p\sqrt{2s_m} }{\xi^2}  \exp\left( -\frac{ \sqrt{2 s_m} r}{\sigma} \right).
-\end{align}
-Note that the exponential term is expected based on \eqref{eqn:eqfreq} -- 
-since establishment requires rare long distant migrants to be present, and \eqref{eqn:eqfreq} gives the density of such --
-but the first term required other arguments.
-
-If this establishment rate is low, 
-the time (in generations) before migrant copies of the allele successfully establish through migration in the second patch
-is approximately exponentially distributed with mean $1/\migrate$.
-Neglecting the time for such a migrant allele to establish once present in the second patch,
-the time scale on which migrant alleles transit and fix between patchs is $1/\migrate$.
+but small patches violate other assumptions above, so we omit it.
 
@@
 spent ``in transit'' between the patches.
-Above, we approximated this by the time $\tau$ it takes the trunk lineage to hit the new patch,
-except that studying a successful migrant family means we need to condition on the establishment occurring, i.e.\ on $\tau < \infty$.
-By the above arguments, and equation~\eqref{eqn:estab_prob}, we have that the conditional probability density is
-\begin{align}
-  \P\{\tau\in dt | \tau<\infty \} &= \frac{\P\{\tau \in dt\}}{\P\{\tau<\infty\}} \\
-  &= \frac{r}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{r^2}{2t\sigma^2} -\alpha t + \sqrt{\alpha} \frac{r}{\sigma} \right) .
-\end{align}
-\plr{more dimensions}
+Above we argued that a good model for this transit is a single Brownian trunk lineage
+surrounded by a cloud of close relatives of random size $K$
+whose chance of surviving until $t$ generations is $p_e(t) \simeq e^{-\alpha t} / \E[K]$.
+Consider a single such family, and let $\tau$ be the (random) time it first hits the new patch,
+with $\tau = \infty$ if it never does.
+By pulling out the constant $1/\E[K]$, we can use standard results on hitting times of $d$-dimensional Brownian motion
+that is killed at rate $\alpha$.
+In particular, we know that if the patch is circular with radius $w$, then
+using \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook}
+\begin{align}
+  \E[e^{-\ell \tau}] &=
+    \begin{cases}
+      e^{- R \sqrt{2(\alpha+\ell)}/\sigma} \qquad & d=1 \\
+      \frac{ K_0( (R+w)\sqrt{2(\alpha+\ell)}/\sigma) }{ K_0( w\sqrt{2(\alpha+\ell)}/\sigma) } \qquad & d=2 
+    \end{cases} 
+    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(\alpha+\ell)}/\sigma}  (1+O(1/w)) .
+\end{align}
+In particular, we can find 
+$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E[\exp(-\ell \tau)]$. % \simeq (R/w+1)^{-(d-1)/2} \exp(- R \sqrt{2\alpha}/\sigma)$.
+
+% Above, we approximated this by the time $\tau$ it takes the trunk lineage to hit the new patch;
+% now that we are studying a successful migrant family means we need to condition on the establishment occurring, i.e.\ on $\tau < \infty$.
+% By the above arguments, and equation~\eqref{eqn:estab_prob}, we have that the conditional probability density is
+% \begin{align}
+%   \P\{\tau\in dt | \tau<\infty \} &= \frac{\P\{\tau \in dt\}}{\P\{\tau<\infty\}} \\
+%   &= \frac{R}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{R^2}{2t\sigma^2} -\alpha t + \sqrt{\alpha} \frac{R}{\sigma} \right) .
+% \end{align}
+% \plr{more dimensions}
+
 Since each generation spent in transit provides an opportunity for recombination,
 if recombination is Poisson, the length of the hitchhiking segment (in Morgans) on each side of the site will be exponentially distributed
-with mean $\tau$, so that if $L$ is the length on, say, the right, then
+with mean $\tau$, so that if $L$ is the length on, say, the right, then in one dimension
 \begin{align} \label{eqn:haplen_cdf}
 \P\{L>\ell\} &= \E[e^{-\ell \tau}|\tau<\infty] \\
-  &= \exp\left\{{-\frac{r}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
+  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
 \end{align}
 This is also the Laplace transform of $\tau$;
-we can differentiate it to find that $\E[\tau] = r/(\sigma\sqrt{2\alpha})$
-and $\var[\tau] = r/( 2^{3/2}\alpha^{3/2}\sigma)$.
-Furthermore, this implies that if $Y$ is an exponential random variable with rate $r\sqrt{2}/\sigma$,
+we can differentiate it to find that $\E[\tau] = (R\sqrt{2\alpha}/\sigma)\times 1/(2\alpha)$
+and $\var[\tau] = (R\sqrt{2\alpha}/\sigma) \times 1/(2\alpha)^2$.
+Furthermore, this implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
 then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$,
 from which we can obtain that
-$\E[L] = \sigma \sqrt{2\alpha}/r + \sigma^2/r^2$ and $\var[L] = 2\alpha\sigma^2/r^2 + 4 \sigma^3 \sqrt{2\alpha}/r^3 + 5 \sigma^4 / r^4$.
+$\E[L] = \sigma \sqrt{2\alpha}/R + \sigma^2/R^2$ and $\var[L] = 2\alpha\sigma^2/R^2 + 4 \sigma^3 \sqrt{2\alpha}/R^3 + 5 \sigma^4 / R^4$.
+For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
+but how other properties might change is unclear.
 
@@
 
-In the above we already assumed that the inter-patch distance $r$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
-by taking the large-$r$ limit above we can obtain a more intuitive description of the process
+In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
+by taking the large-$R$ limit above we can obtain a more intuitive description of the process
 via a central limit for $\tau$.
-Define $\eta_r$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
-so that $\tau = r/\sigma(2\alpha)^{-1/2} + \eta_r \sqrt{r/\sigma} (2\alpha)^{-3/4}$.
+Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
+so that $\tau = R/\sigma(2\alpha)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2\alpha)^{-3/4}$.
 Plugging this into equation~\eqref{eqn:haplen_cdf},
-after some algebra (and using that $\sqrt{r+\epsilon} = \sqrt{r} + \frac{\epsilon}{2\sqrt{r}} - \frac{\epsilon^2}{2\sqrt{r}^3} + O(\epsilon^3)$) we obtain that
-\begin{align}
-  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{r}\right) \right\}.
-\end{align}
-By the properties of Laplace transforms \citep{durrett},
-this implies that $\eta_r$ converges in distribution to a standard Gaussian random variable as $r \to \infty$.
+after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
+\begin{align}
+  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
+\end{align}
+Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
+this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
 In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
 Since the standard deviation of $\tau$ is of lower order than its mean,
-a similar calculation shows that for large $r$, the haplotype halflength $L$ is approximately exponentially distributed, 
-with mean $\sqrt{2\alpha} \sigma/r$.
+a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
+with mean $\sqrt{2\alpha} \sigma/R$.
 
 % \begin{align}
-%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{r}{\sigma\sqrt{2\alpha}}\right)\frac{(2\alpha)^{3/4}\sqrt{\sigma}}{\sqrt{r}} \right\} \right] \\
-%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2\alpha)^{3/4} \sqrt{\sigma}}{\sqrt{r}}\right) \tau \right\} \right] \exp\left( (2\alpha)^{1/4} \sqrt{\frac{r}{\sigma}} \right) \\
+%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{R}{\sigma\sqrt{2\alpha}}\right)\frac{(2\alpha)^{3/4}\sqrt{\sigma}}{\sqrt{R}} \right\} \right] \\
+%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2\alpha)^{3/4} \sqrt{\sigma}}{\sqrt{R}}\right) \tau \right\} \right] \exp\left( (2\alpha)^{1/4} \sqrt{\frac{R}{\sigma}} \right) \\
 % \end{align}
+
+XXX Check what happens in two dimensions. XXX
 
@@
 
-%%%%%
-\section{Numerical calculation of the probability of establishment}
-\label{apx:establishment_sims}
-
-To investigate the impact of the various approximations surrounding the mutational influx,
-we carried out numerical calculations of a fine-grained, discrete model.
-These are shown, and described, in figure~\ref{fig:prob_estab_calcs}.
-
-Here (and at other parameter choices) we see that the probability of establishment $p(x)$ goes to the equilibrium value
-(approximately $p_e = 2s/\xi^2$) within the patch;
-the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
-This lends credence to our approximation that the integral of $p(x)$ over the entire range
-is close to $p_e$ multiplied by the area of the patch.
-
-\begin{figure}[ht!]
-    \begin{center}
-        % made in patchy-selection-plots.R
-        \includegraphics{prob-establishment}
-    \end{center}
-    \label{fig:prob_estab_calcs}
-    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
-    in an abrupt transition (left) and a gradual transition (right).
-    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp .05$ respecively);
-    and the selection coefficient interpolates linearly between these in the central shaded region.
-    The number of offspring is Poisson.
-    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
-    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
-    }
-\end{figure}
-
-%%%%%%%%%
-\section{Simulations of a patchy environment}
-\label{apx:patchy_sims}
-
-Describe sims.
-
-Argue that $C=0.5$ is a reasonable number.
-
-\begin{figure}[ht!]
-    \begin{center}
-        % made in patchy-sim-writeup-plots.R
-        \includegraphics{example-equilibrium}
-    \end{center}
-    \label{fig:example_equil}
-    \caption{
-    Snapshots of a simulation of allelic frequency dynamics around a patch where the allele is beneficial:
-    the focal allele is benefical in the middle region and deleterious outside.
-    The dynamics were simulated for 10000 generations across demes with 1000 individuals in each
-    and random dispersal to nearby demes with a mean dispersal distance of $\sigma \approx 2$ demes.
-    Colored lines show allele frequencies across space in 50 uniformly spaced generations;
-    the solid line shows the mean across all generations.
-    Below is the same plot, except on a log scale.
-    }
-\end{figure}
-
 %%%%%%%%%%%%
@@
 Since $q(r) \to 0$ as $r \to \infty$, so $q(r) (1-q(r)) \approx q(r)$,
-XXX theory in \citet{wasow1965asymptotic} XXX
-implies that the true equilibrium frequency $q$ is close, for large $r$, to the solution to
+it can be shown 
+% theory in \citet{wasow1965asymptotic} implies 
+that the true equilibrium frequency $q$ is close, for large $r$, to the solution to
 \begin{align}
@@
 \label{apx:elliptic_integrals}
-
-XXX add equation that probability of establishment solves also XXX
 
@@
 \subsection{Doing the integrals}
+\label{apx:some_integrals}
 
@@
 
-%%%%%%%%%%
-\section{The probability of establishment}
-\label{ss:prob_estab}
-
-In the main text we have made use of two decompositons of the spatial branching process
-to approximate in two different ways the rate of establishment of migrant alleles.
-Here we make the relationship more formal.
-The decompositions use a similar idea to that found in \citep[section D.12]{athreyaney},
-where a supercritical branching process is decomposed into individuals whose families eventually die out
-and those having infinite lines of descent.
-We also note that the approximations fall generally in the realm covered by \citet{aldouspoissonclumping}.
-
-Consider a discrete-time branching process $Z_t$ with spatial motion and offspring distribution depending on spatial location.
-Suppose that the offspring distribution for an individual at location $x$ is distributed as $X_x$.
-We can record the state of the process simply as the list of locations of individuals:
-$Z_t = (x_1, \ldots, x_{N(t)})$ means that there are $N(t)$ individuals alive at $t$, at locations $x_1, \ldots, x_{N(t)}$.
-We can construct the process iteratively: given $Z_t$,
-we sample first the number of offspring of each, and then the locations the offspring migrate to.
-Concretely, given $Z_{t} = (x_1, \ldots, x_{N(t)})$,
-first let $X_1(t),\ldots,X_{N(t)(t)}$ be independent random draws from the offspring distribution,
-with $X_k(t)$ having distribution $X_{x_k}$.
-The number of new offspring is $N(t+1) = \sum_{k=1}^{N(t)} X_k(t)$,
-and we can sample their locations independently given the location of their parent to obtain $Z_{t+1}$.
-Note that the genealogy is implicitly recorded:
-the parent of the $k^\mathrm{th}$ individual at time $t$ is
-the $a(k,t) = \min\{ \ell : \sum_{j=1}^\ell X_j(t-1) \ge k \}$-th individual at time $t-1$.
-It will be useful to extend this to more distant ancestors:
-let $a_1(k,t) = a(k,t)$ and recursively define $a_m(k,t) = a(a_{m-1}(k,t),t-m+1)$, as long as $m\le t$.
-
-We want to find (approximations to) $p(x)$, the probability that a branching process beginning with a single individual at $x$ does not die out. 
-(In other words, $p(x) = \P\{ N(t) = 0 \; \mbox{for some}\; t>0 \; \vert \; Z_0 = (x) \}$.)
-We assume that the mean number of offspring $\E[X_x]$ is only greater than one for $x \in A$ (``the patch''),
-and so the process can only establish if some offspring eventually get to the patch (with probability one).
-
-The second decomposition we use is simpler, only conditioning on the number of individuals who ever reach the patch.  
-Let $\calE = \{(t_i,k_i)\}$ be the times and indices of individuals who are in the patch but none of whose ancestors have been --
-i.e.\ $(t,k) \in \calE$ iff $x_k(t) \in A$ and $a_m(k,t) \notin A$ for all $1\le m \le t$.
-If we remove these individuals (and their offspring), the resulting branching process has mean number of offspring less than or equal to 1 everywhere,
-so there are a finite number of such individuals with probability one (the process dies out).
-Therefore, the probability that the process dies out is then the probability that the families descending from each individual in $\calE$ dies out.
-By independence of different offspring, this is
-\begin{align}
-  1-p(x) = \E\left[ \prod_{(t,k) \in \calE} (1-p(x_k(t))) \; \vert \; Z_0 = (x) \right] .
-\end{align}
-Above, we make the approximation that $p(x_k(t)) \approx p_e$,
-the probability of establishment ignoring the possibility of migrating outside the patch.
-In this case,
-\begin{align}
-  p(x) \approx \E\left[ p_e^{|\calE|} \; \vert \; Z_0 = (x) \right] .
-\end{align}
-Here $|\calE|$ is the ``size of the family that reaches the patch'';
-in section \ref{ss:haplotype_length} we estimated the probability that any of the family reaches the patch
-and argued that the distribution of $|\calE|$ is given by $K$,
-the size of the subcritical branching process conditioned on still being alive.
-
-The first approximation uses a similar trick, but rather than treating all individuals migrating into the patch
-as an opportunity for establishment,
-we couple the branching process to a strictly subcritical branching process
-that produces extra offspring, each of which provides an opportunity for establishment.
-To this end, suppose that the offspring distribution off the patch is given by $X'$ (so $X_x \deq X'$ for $x \notin A$),
-and the offspring distribution within the patch stochastically dominates $X$,
-i.e.\ $\P\{X_x\ge n\} \ge \P\{X'\ge n\}$ for all $x\in A$ and $n \ge 0$.
-Stochastic domination implies that we can \emph{couple} any instance of $X_x$ with one of $X'$ and a remainder --
-i.e.\ jointly construct an $X'$ and a $Y_x$ such that $X_x = X' + Y_x$,
-and that $Y_x$ also takes values in the nonnegative integers.
-We can therefore modify the construction of the branching process above,
-so that $X_k(t) = X_k'(t) + Y_k(t)$, where $X_k'(t) \deq X$ and $Y_k(t) \deq Y_{x_k(t)}$.
-We can define a \emph{core} branching process $Z'$ using the $X'$, which will be finite (since $\E[X']<1$).
-Formally, say that the first $X_k'(t)$ offspring of the $(t,k)^\mathrm{th}$ individual are ``core'' offspring,
-denoted by $\calC_t = \{ (t,k) \; : \; 0 \le k - \sum_{j<\ell} X_j(t) \le X_\ell'(t) \; \mbox{for some}\; 1 \le \ell \le N(t) \}$;
-and define $Z'$ as the core individuals whose ancestors are all ``core'' as well,
-i.e.\ $Z'_t = \{ x_k \in Z_t \; : \;  a_m(k,t) \in \calC_{t-m} \; \mbox{for all}\; 0 \le m \le t\}$.
-The boundary analogous to above in the first computation is then the non-core offspring of core individuals;
-denote the times and indices of these by $\partial \calC$,
-so that $(t,k) \in \partial \calC$ iff $(t,k) \notin \calC_t$ and $a(k,t) \in \calC_{t-1}$.
-We can now use this as above to compute the probability of establishment beginning from a single individual:
-\begin{align}
-  1-p(x) = \E\left[ \prod_{(t,k) \in \partial\calC} (1-p(x_k(t))) \; \vert \; Z_0 = (x) \right] ,
-\end{align}
-but this is less helpful, as the distribution of $\calC$ is harder to get at.
-
-However, recall the actual problem at hand.
-Roughly speaking, we approximate the process of colonization of the new patch by migrants from an established patch
-by imagining that the established patch sends out migrants, and that each of these acts as an independent branching process,
-at least in the region of the new patch.
-We can suppose that migrant families appear at some rate, say, midway between the patches, and far enough from the original one that we can ignore it,
-and leave the precise distribution of their appearance unspecified
-We can decompose each of these branching processes as above into ``core'' and ``extra'',
-so that the point in time of ``establishment'' is well-defined:
-the time $\tau$ of first appearance of an ``extra'' offspring with infinite line of descent.
-Given that establishment has not yet occurred by $t$, 
-and the combined state of the core process $Z'_t$,
-then the mean number of establishment events occurring at time $t$ is $\sum_{x \in Z'_t} p(x) \E[Y_x]$.
-
-If establishment events were independent and sufficiently rare,
-the influx of branching processes begain long enough ago that $Z'$ was at stationarity
-with $z_A := \E[ \# \{ (x,t) \in Z'_t : x \in A \} ]$,
-and we again approximate $p(x) \approx p_e$,
-then we would have that 
-\begin{align}
-  \P\{\tau>t\} \approx \exp\left\{ - t z_A \E[Y_x] p_e \right\} ;
-\end{align}
-in our case, $\E[Y_x] \approx (s_p+s_m)$ and $p_e \approx 2 s_p/\xi^2$,
-so $\tau$ is approximately exponentially distributed with rate $ 2 z_A (s_p+s_m) s_p/\xi^2 $.
-We can get $z_A$, the occupation density without the new patch, 
-directly by integrating the expression~\eqref{eqn:eqfreq} over the patch.
-
-In obtaining an exponential distribution, we have assumed effectively that establishment events
-are independent of each other.
-This will not be strictly true, since an establishment implies that a family of migrants are present in the patch,
-and hence others may produce offspring with infinite lines of descent at a similar time.
-The mean total family size of an allele with fitness $1-s_m$ (i.e.\ a subcritical branching process) is $\sum_{t \ge 0} (1-s_m)^t = 1/s_m$;
-so the mean number of ``extra'' offspring is $1+s_p/s_m$.
-Establishment events are correlated to the extent that single such families give rise to multiple establishment events,
-
-% We will record the state of the branching process as a finite point measure,
-% so that if there are $n$ individuals alive at $t$, at locations $x_1, \ldots, x_n$,
-% then $Z_t = \sum_{k=1}^n \delta_{x_k}$, and we write $Z_t(A) = \sum_{k=1}^n \bone_{x_k \in A}$ for the number of individuals in a set $A$.
+%%%%%
+\section{Numerical calculation of the probability of establishment}
+\label{apx:establishment_sims}
+
+To investigate the impact of the various approximations surrounding the mutational influx,
+we carried out numerical calculations of a fine-grained, discrete model.
+These are shown, and described, in figure~\ref{fig:prob_estab_calcs}.
+
+Here (and at other parameter choices) we see that the probability of establishment $p(x)$ goes to the equilibrium value
+(approximately $p_e = 2s/\xi^2$) within the patch;
+the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
+This lends credence to our approximation that the integral of $p(x)$ over the entire range
+is close to $p_e$ multiplied by the area of the patch.
+
+\begin{figure}[ht!]
+    \begin{center}
+        % made in patchy-selection-plots.R
+        \includegraphics{prob-establishment}
+    \end{center}
+    \label{fig:prob_estab_calcs}
+    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
+    in an abrupt transition (left) and a gradual transition (right).
+    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp .05$ respecively);
+    and the selection coefficient interpolates linearly between these in the central shaded region.
+    The number of offspring is Poisson.
+    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
+    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
+    }
+\end{figure}
+
+%%%%%%%%%
+\section{Simulations of a patchy environment}
+\label{apx:patchy_sims}
+
+Describe sims.
+
+Argue that $C=0.5$ is a reasonable number.
+
+\begin{figure}[ht!]
+    \begin{center}
+        % made in patchy-sim-writeup-plots.R
+        \includegraphics{example-equilibrium}
+    \end{center}
+    \label{fig:example_equil}
+    \caption{
+    Snapshots of a simulation of allelic frequency dynamics around a patch where the allele is beneficial:
+    the focal allele is benefical in the middle region and deleterious outside.
+    The dynamics were simulated for 10000 generations across demes with 1000 individuals in each
+    and random dispersal to nearby demes with a mean dispersal distance of $\sigma \approx 2$ demes.
+    Colored lines show allele frequencies across space in 50 uniformly spaced generations;
+    the solid line shows the mean across all generations.
+    Below is the same plot, except on a log scale.
+    }
+\end{figure}
 
@@
 \end{equation}
-namely, is a $\mathrm{Beta}((\mu/m), 1)$ distribution--- see \cite{donnelly-joyce} and \cite{permanPitmanYor92}.
+namely, is a $\mathrm{Beta}((\mu/m), 1)$ distribution--- see \cite{donnelly1989continuity} and \cite{perman1992sizebiased}.
 
@@
 
-\subsection{Temporary work: outflux of families}
-\label{apx:lesbegue}
-
-We have that
+%%%%%%%%% %%%%%%%%%%%%%%
+\section{Hitting and occupation}
+\label{apx:hitting_occupation}
+
+Consider a single migrant family $Z$ beginning at $x$.
+Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
+and $\tau_\dagger$ an independent Exponential($\alpha$) time.
+The mean occupation time of $Z$ spent in a region $A$ is 
+\begin{align}
+    u(x,y) &= \E[\int_0^\infty Z_t(A) dt ] \\
+           &= \int_0^\infty \E[Z_t] p_t(x,A) dt \\
+        &= \int_0^\infty e^{-\alpha t} p_t(x,A) dt \\
+        &= \int_0^\infty \P^x\{ \tau_\dagger < t \; \& \; B_t \in A \} dt \\
+        &= \E[ \int_0^{\tau_\dagger} \one_A(B_t) dt ]  .
+\end{align}
+If we define
+$\tau_A$ to be the hitting time of $A$ by $B$,
+and $\mu(x,A)$ to be the hitting distribution of $\partial A$ by $B_{\min(t,\tau)}$,
+by the strong Markov property this is equal to
+\begin{align}
+    u(x,y) &= \E^x[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] \\
+           &= \P^x\{ \tau_A < \tau_\dagger \}  \E^{\mu(x,A)}[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] .
+\end{align}
+If $A$ is circular, the latter expectation will not depend on $x$;
+call this $h(A)$.
+
+Now consider the chance that $Z$ ever hits $A$.
+If we make the approximation that $Z$ hits $A$ iff the trunk of $Z$ does,
+then this is equal to
+\begin{align}
+    F(x,y) &= \int_0^\infty (1-p_e(t)) \P\{ \tau_A \in dt \} \\
+           &= \int_0^\infty e^{\alpha t} (1-p_e(t)) e^{- \alpha t } \P\{ \tau_A < t \} dt \\
+    &=\int_0^\infty e^{\alpha t} (1-p_e(t)) \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &= \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &= \frac{1}{\E[K]} \P^x[ \tau_A < \tau_\dagger ] .
+\end{align}
+Therefore, we have that
+\begin{align}
+    F(x,y) \E[K] 
+    \approx \P^x[ \tau_A < \tau_\dagger ] 
+    = u(x,y) / h(A) .
+\end{align}
+
+Now, 
+\begin{align}
+    q(y)  &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
+          &= h(A) \int_{\partial B(r_0)} \gamma(x) F(x,y) dx \\
+          &= \frac{ h(A) }{ \E[K] } \migrate(y) .
+\end{align}
+The function $h(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
+is expected to spend inside the disk before $\tau_\dagger$.
+This is integral of the Green function for the Bessel process of the appropriate order, 
+so using Borodin \& Salminen (Appendix 1),
+and letting $w$ be the width of the patch,
+in $d=1$,
+\begin{align}
+    h_1(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+  % &= \int_0^{2w} G_\alpha(0,y) dy \\
+  &= \int_0^{2w} \frac{ e^{- y \sqrt{2\alpha} }}{\sqrt{2\alpha}} dy \\
+  &= (1-e^{-2w\sqrt{2\alpha}})/(2\alpha)
+\end{align}
+and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
+\begin{align}
+    h_2(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_\alpha(0,y) dy \\
+  &= \int_0^{2w} 2 y K_0(y \sqrt{2\alpha}) dy \\
+  &= \frac{1}{\alpha}\int_0^{2w\sqrt{2\alpha}} y K_0(y) dy \\
+  &= \frac{1}{2\alpha}\left(1- 2w\sqrt{2\alpha} K_1(2w\sqrt{2\alpha}) \right)  .
+\end{align}
+
+%%%%%%% %%%%%%%%%%%%
+\section{The outflux of families}
+\label{apx:outflux}
+
+It is of interest to have an expression for the ``outflux of families'',
+$\gamma(r_0)$, 
+defined to be the mean proportion of $B$ alleles at distance $r_0$
+whose lineage traces back to the original patch without venturing further away than $r_0$.
+By expression \eqref{eqn:occupation_integral},
 \[
@@
 \begin{align}
-    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} ,
-\end{align}
-and hence, by bounded convergence, then \citet{gradshteyn2007table} (3.339),
-\begin{align}
-    \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
+    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} .
+\end{align}
+In one dimension we are done.
+In two dimensions,
+by bounded convergence and \citet{gradshteyn2007table} (3.339),
+\begin{align}
+  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
+            &= C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
                          &= 2 \pi C' I_0(-2r_0),
@@
 
-\subsection{Temporary work: hitting probs}
+\end{document}
+
+\section{Temporary work: hitting probs}
 
@@
 \end{align}
-
-\subsection{Hitting and occupation}
-\label{apx:hitting_occupation}
-
-Consider a single migrant family $Z$ beginning at $x$.
-Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
-and $\tau_\dagger$ an independent Exponential($\alpha$) time.
-The mean occupation time of $Z$ spent in a region $A$ is 
-\begin{align}
-    u(x,y) &= \E[\int_0^\infty Z_t(A) dt ] \\
-           &= \int_0^\infty \E[Z_t] p_t(x,A) dt \\
-        &= \int_0^\infty e^{-\alpha t} p_t(x,A) dt \\
-        &= \int_0^\infty \P^x\{ \tau_\dagger < t \; \& \; B_t \in A \} dt \\
-        &= \E[ \int_0^{\tau_\dagger} \one_A(B_t) dt ]  .
-\end{align}
-If we define
-$\tau_A$ to be the hitting time of $A$ by $B$,
-and $\mu(x,A)$ to be the hitting distribution of $\partial A$ by $B_{\min(t,\tau)}$,
-by the strong Markov property this is equal to
-\begin{align}
-    u(x,y) &= \E^x[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] \\
-           &= \P^x\{ \tau_A < \tau_\dagger \}  \E^{\mu(x,A)}[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] .
-\end{align}
-If $A$ is circular, the latter expectation will not depend on $x$;
-call this $h(A)$.
-
-Now consider the chance that $Z$ ever hits $A$.
-If we make the approximation that $Z$ hits $A$ iff the trunk of $Z$ does,
-then this is equal to
-\begin{align}
-    F(x,y) &= \int_0^\infty (1-p_e(t)) \P\{ \tau_A \in dt \} \\
-           &= \int_0^\infty e^{\alpha t} (1-p_e(t)) e^{- \alpha t } \P\{ \tau_A < t \} dt \\
-    &=\int_0^\infty e^{\alpha t} (1-p_e(t)) \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &= \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &= \frac{1}{\E[K]} \P^x[ \tau_A < \tau_\dagger ] .
-\end{align}
-Therefore, we have that
-\begin{align}
-    F(x,y) \E[K] 
-    \approx \P^x[ \tau_A < \tau_\dagger ] 
-    = u(x,y) / h(A) .
-\end{align}
-
-Now, 
-\begin{align}
-    q(y)  &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
-          &= h(A) \int_{\partial B(r_0)} \gamma(x) F(x,y) dx \\
-          &= \frac{ h(A) }{ \E[K] } \migrate(y) .
-\end{align}
-The function $h(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
-is expected to spend inside the disk before $\tau_\dagger$.
-This is integral of the Green function for the Bessel process of the appropriate order, 
-so using Borodin \& Salminen (Appendix 1),
-and letting $w$ be the width of the patch,
-in $d=1$,
-\begin{align}
-    h_1(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-  % &= \int_0^{2w} G_\alpha(0,y) dy \\
-  &= \int_0^{2w} \frac{ e^{- y \sqrt{2\alpha} }}{\sqrt{2\alpha}} dy \\
-  &= (1-e^{-2w\sqrt{2\alpha}})/(2\alpha)
-\end{align}
-and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
-\begin{align}
-    h_2(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-    % &= \int_0^{2w} G_\alpha(0,y) dy \\
-  &= \int_0^{2w} 2 y K_0(y \sqrt{2\alpha}) dy \\
-  &= \frac{1}{\alpha}\int_0^{2w\sqrt{2\alpha}} y K_0(y) dy \\
-  &= \frac{1}{2\alpha}\left(1- 2w\sqrt{2\alpha} K_1(2w\sqrt{2\alpha}) \right)  .
-\end{align}
-
-\end{document}

### 2014-04-21 spatial-equation-solutions.tex
draft! cleaned up!

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{fullpage}

\begin{document}

Slatkin (1973) argues that if selection at $x$ is $s\gamma(x)$,
the variance in dispersal distance is $\sigma^2$,
and $A(x) = 2p(x)-1 = p(x)-q(x)$, then
\begin{equation}
    \partial_x^2 A(x) = - \frac{s \gamma(x)}{\sigma^2} \left( 1- A^2(x) \right).
\end{equation}
(note the erroneous minus sign in equation (7).
Rescaling to $\xi = x \sqrt{s}/\sigma$, this is
\begin{equation} \label{eqn:reaction_diffusion}
    \partial_\xi^2 A(\xi) = - \gamma(x) \left( 1- A^2(\xi) \right).
\end{equation}

Note that if $f(x) = 2 - 3 \tanh^2(x/2+x_0)$ then since $\partial_x \tanh(x) = (1-\tanh^2(x))$,
\begin{align*}
    2 &\ge f(x) \ge -1 \\
    \partial_x f(x) &= - 3 \tanh(x/2+x_0) \left( 1- \tanh^2(x/2+x_0) \right) \\
    \partial^2_x f(x) &= - \frac{3}{2} \left( 1- \tanh^2(x/2+x_0) \right) 
    + \frac{9}{2} \tanh^2(x/2+x_0) \left(1-\tanh^2(x/2+x_0)\right) \\
        &= \frac{1}{2} \left\{ - 3 + 12 \tanh^2(x/2+x_0) - 9 \tanh^4(x/2+x_0) \right\} \\
        &= \frac{1}{2} \left(1-f(x)^2 \right)
\end{align*}
and a solution to equation \eqref{eqn:reaction_diffusion}, if $\gamma(x)=\gamma$ is constant,
is obtained by choosing $x_0$ to match the boundary conditions.

Note that since $-1\le A(x) \le 1$, we certainly can't apply the solution $f(x)$
for $|x/2+x_0| \le \tanh^{-1}(1/\sqrt{3})$.

%%%%%%%%%%
\section{The probability of establishment}
\label{ss:prob_estab}

In the main text we have made use of two decompositons of the spatial branching process
to approximate in two different ways the rate of establishment of migrant alleles.
Here we make the relationship more formal.
The decompositions use a similar idea to that found in \citep[section D.12]{athreya2004branching},
where a supercritical branching process is decomposed into individuals whose families eventually die out
and those having infinite lines of descent.
We also note that the approximations fall generally in the realm covered by \citet{aldous1989poisson}.

Consider a discrete-time branching process $Z_t$ with spatial motion and offspring distribution depending on spatial location.
Suppose that the offspring distribution for an individual at location $x$ is distributed as $X_x$.
We can record the state of the process simply as the list of locations of individuals:
$Z_t = (x_1, \ldots, x_{N(t)})$ means that there are $N(t)$ individuals alive at $t$, at locations $x_1, \ldots, x_{N(t)}$.
We can construct the process iteratively: given $Z_t$,
we sample first the number of offspring of each, and then the locations the offspring migrate to.
Concretely, given $Z_{t} = (x_1, \ldots, x_{N(t)})$,
first let $X_1(t),\ldots,X_{N(t)(t)}$ be independent random draws from the offspring distribution,
with $X_k(t)$ having distribution $X_{x_k}$.
The number of new offspring is $N(t+1) = \sum_{k=1}^{N(t)} X_k(t)$,
and we can sample their locations independently given the location of their parent to obtain $Z_{t+1}$.
Note that the genealogy is implicitly recorded:
the parent of the $k^\mathrm{th}$ individual at time $t$ is
the $a(k,t) = \min\{ \ell : \sum_{j=1}^\ell X_j(t-1) \ge k \}$-th individual at time $t-1$.
It will be useful to extend this to more distant ancestors:
let $a_1(k,t) = a(k,t)$ and recursively define $a_m(k,t) = a(a_{m-1}(k,t),t-m+1)$, as long as $m\le t$.

We want to find (approximations to) $p(x)$, the probability that a branching process beginning with a single individual at $x$ does not die out. 
(In other words, $p(x) = \P\{ N(t) = 0 \; \mbox{for some}\; t>0 \; \vert \; Z_0 = (x) \}$.)
We assume that the mean number of offspring $\E[X_x]$ is only greater than one for $x \in A$ (``the patch''),
and so the process can only establish if some offspring eventually get to the patch (with probability one).

The second decomposition we use is simpler, only conditioning on the number of individuals who ever reach the patch.  
Let $\calE = \{(t_i,k_i)\}$ be the times and indices of individuals who are in the patch but none of whose ancestors have been --
i.e.\ $(t,k) \in \calE$ iff $x_k(t) \in A$ and $a_m(k,t) \notin A$ for all $1\le m \le t$.
If we remove these individuals (and their offspring), the resulting branching process has mean number of offspring less than or equal to 1 everywhere,
so there are a finite number of such individuals with probability one (the process dies out).
Therefore, the probability that the process dies out is then the probability that the families descending from each individual in $\calE$ dies out.
By independence of different offspring, this is
\begin{align}
  1-p(x) = \E\left[ \prod_{(t,k) \in \calE} (1-p(x_k(t))) \; \vert \; Z_0 = (x) \right] .
\end{align}
Above, we make the approximation that $p(x_k(t)) \approx p_e$,
the probability of establishment ignoring the possibility of migrating outside the patch.
In this case,
\begin{align}
  p(x) \approx \E\left[ p_e^{|\calE|} \; \vert \; Z_0 = (x) \right] .
\end{align}
Here $|\calE|$ is the ``size of the family that reaches the patch'';
in section \ref{ss:haplotype_length} we estimated the probability that any of the family reaches the patch
and argued that the distribution of $|\calE|$ is given by $K$,
the size of the subcritical branching process conditioned on still being alive.

The first approximation uses a similar trick, but rather than treating all individuals migrating into the patch
as an opportunity for establishment,
we couple the branching process to a strictly subcritical branching process
that produces extra offspring, each of which provides an opportunity for establishment.
To this end, suppose that the offspring distribution off the patch is given by $X'$ (so $X_x \deq X'$ for $x \notin A$),
and the offspring distribution within the patch stochastically dominates $X$,
i.e.\ $\P\{X_x\ge n\} \ge \P\{X'\ge n\}$ for all $x\in A$ and $n \ge 0$.
Stochastic domination implies that we can \emph{couple} any instance of $X_x$ with one of $X'$ and a remainder --
i.e.\ jointly construct an $X'$ and a $Y_x$ such that $X_x = X' + Y_x$,
and that $Y_x$ also takes values in the nonnegative integers.
We can therefore modify the construction of the branching process above,
so that $X_k(t) = X_k'(t) + Y_k(t)$, where $X_k'(t) \deq X$ and $Y_k(t) \deq Y_{x_k(t)}$.
We can define a \emph{core} branching process $Z'$ using the $X'$, which will be finite (since $\E[X']<1$).
Formally, say that the first $X_k'(t)$ offspring of the $(t,k)^\mathrm{th}$ individual are ``core'' offspring,
denoted by $\calC_t = \{ (t,k) \; : \; 0 \le k - \sum_{j<\ell} X_j(t) \le X_\ell'(t) \; \mbox{for some}\; 1 \le \ell \le N(t) \}$;
and define $Z'$ as the core individuals whose ancestors are all ``core'' as well,
i.e.\ $Z'_t = \{ x_k \in Z_t \; : \;  a_m(k,t) \in \calC_{t-m} \; \mbox{for all}\; 0 \le m \le t\}$.
The boundary analogous to above in the first computation is then the non-core offspring of core individuals;
denote the times and indices of these by $\partial \calC$,
so that $(t,k) \in \partial \calC$ iff $(t,k) \notin \calC_t$ and $a(k,t) \in \calC_{t-1}$.
We can now use this as above to compute the probability of establishment beginning from a single individual:
\begin{align}
  1-p(x) = \E\left[ \prod_{(t,k) \in \partial\calC} (1-p(x_k(t))) \; \vert \; Z_0 = (x) \right] ,
\end{align}
but this is less helpful, as the distribution of $\calC$ is harder to get at.

However, recall the actual problem at hand.
Roughly speaking, we approximate the process of colonization of the new patch by migrants from an established patch
by imagining that the established patch sends out migrants, and that each of these acts as an independent branching process,
at least in the region of the new patch.
We can suppose that migrant families appear at some rate, say, midway between the patches, and far enough from the original one that we can ignore it,
and leave the precise distribution of their appearance unspecified
We can decompose each of these branching processes as above into ``core'' and ``extra'',
so that the point in time of ``establishment'' is well-defined:
the time $\tau$ of first appearance of an ``extra'' offspring with infinite line of descent.
Given that establishment has not yet occurred by $t$, 
and the combined state of the core process $Z'_t$,
then the mean number of establishment events occurring at time $t$ is $\sum_{x \in Z'_t} p(x) \E[Y_x]$.

If establishment events were independent and sufficiently rare,
the influx of branching processes begain long enough ago that $Z'$ was at stationarity
with $z_A := \E[ \# \{ (x,t) \in Z'_t : x \in A \} ]$,
and we again approximate $p(x) \approx p_e$,
then we would have that 
\begin{align}
  \P\{\tau>t\} \approx \exp\left\{ - t z_A \E[Y_x] p_e \right\} ;
\end{align}
in our case, $\E[Y_x] \approx (s_p+s_m)$ and $p_e \approx 2 s_p/\xi^2$,
so $\tau$ is approximately exponentially distributed with rate $ 2 z_A (s_p+s_m) s_p/\xi^2 $.
We can get $z_A$, the occupation density without the new patch, 
directly by integrating the expression~\eqref{eqn:eqfreq} over the patch.

In obtaining an exponential distribution, we have assumed effectively that establishment events
are independent of each other.
This will not be strictly true, since an establishment implies that a family of migrants are present in the patch,
and hence others may produce offspring with infinite lines of descent at a similar time.
The mean total family size of an allele with fitness $1-s_m$ (i.e.\ a subcritical branching process) is $\sum_{t \ge 0} (1-s_m)^t = 1/s_m$;
so the mean number of ``extra'' offspring is $1+s_p/s_m$.
Establishment events are correlated to the extent that single such families give rise to multiple establishment events,

% We will record the state of the branching process as a finite point measure,
% so that if there are $n$ individuals alive at $t$, at locations $x_1, \ldots, x_n$,
% then $Z_t = \sum_{k=1}^n \delta_{x_k}$, and we write $Z_t(A) = \sum_{k=1}^n \bone_{x_k \in A}$ for the number of individuals in a set $A$.

\end{document}

### 2014-05-12 patchy-selection-paper.tex
edited text through the probability of parallel mig

@@
 \usepackage{hyperref}
-\usepackage{latexml} % for \iflatexml
+%\usepackage{latexml} % for \iflatexml
 
@@
 
+\gc{Lets also give the numbers of key equations in this road map so
+  folks can understand the biologically important results and then work
+  their way back through math, once they know where we are going}
+
 %http://www.nature.com/hdy/journal/v94/n2/full/6800600a.html
@@
 
-Consider a population spread across a landscaper, 
+Consider a population spread across a landscape, 
 within which are patches of habitat to which individuals are (initially) poorly adapted,
 surrounded by large areas to which the population is well adapted.
-(When we refer to ``patches'' it is to these pieces of habitat.)
-Suppose it takes only a single mutational change to create an allele ($B$) that adapts an individual to the poor habitat type, 
-a change that occurs at a (low) rate of $\mu$ per chromosome per generation. 
+(When we refer to ``patches'' it is to these pieces of poor habitat.)
+Suppose it takes only a single mutational change to create an allele
+($B$) that adapts an individual to the poor habitat type.
+This change could be as specific as a single base change, a knock out
+of a gene, or a mutation anywhere in a set of genes. We assume that
+this mutational change occurs at a (low) rate of $\mu$ per chromosome per generation. 
 We assume that this initially rare new
-allele has fitness $1+s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches,
-and assume that it has fitness $1-s_m$ when rare in the intervening areas, with $s_p, s_m > 0$.
+allele has fitness $1+s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches.
+Outside of these patches we assume that $B$ has a relative fitness
+$1-s_m$  when rare in the intervening areas, with $s_p, s_m > 0$.
 To simplify matters, we assume that the disadvantage $s_m$ 
@@
 \paragraph{Previous work} 
-We will make use of several previous results from the literature. 
-Haldane, Fisher, and later \citet{slatkin1973geneflow} first analyzed models of this sort.
+We will make use of several previous results from the literature on
+migration in spatially varying selection models. \gc{cite Haldane and zfisher} 
+Haldane, Fisher, and later \citet{slatkin1973geneflow} first analyzed
+deterministic models of of this sort.
 \citet{nagylaki1975conditions} and \citet{conley1975application} showed that in one dimension, if the physical width of the patch is less than $(\sigma/\sqrt{2s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
@@
 this probability of establishment of a new mutation decays exponentially with distance from the patch, 
-with the scale given by $\sigma/\sqrt{s_m}$,
-and that mutations appearing within the patch have probability of establishment
+with the scale given by $\sigma/\sqrt{s_m}$. Mutations appearing within the patch have probability of establishment
 strictly less than the probability for a panimictic population,
 but approaching this as the size of the patch increases.
-(This latter panmictic probability of establishment we denote $p_e$,
-and often approximate by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.)
+This latter panmictic probability of establishment we denote $p_e$,
+and often approximate by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
 This result holds quite generally for a geographically spread population that experiences a uniform selection
@@
 
-Let $p(x)$ be the probability that a new mutant $B$ allele arising at location $x$ 
+Let $p(x)$ be the probability that a new mutant $B$ allele arising at
+location $x$
 relative to the patch fixes within the poor habitat patch.
@@
 but these are not particularly helpful.
-The total successful mutational influx per generation is given by the integral of $p(x)$ over the entire species range,
+The total successful mutational influx per generation is given by the
+integral of $p(x)$ over the entire species range:
+\begin{equation}
+  \mutrate = \rho \mu \int p(x) dx \label{eqn:exactmutrate}
+\end{equation}
 % $\int \rho \mu p(x) dx$,
@@
 but still scales linearly with the mutational influx density $\rho \mu$.
-If the width of the patch is large relative to $\sigma/\sqrt{2s_m}$, 
-then a reasonable approximation is to assume that $p(x) = p_e$ within the patch, and $p(x) = 0$ otherwise,
-for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_p A / \xi^2$. 
+If we consider a patch of area $A$, whose width of the patch is large relative to $\sigma/\sqrt{2s_m}$, 
+then a reasonable approximation is to assume that $p(x) = p_e$ within the patch, and $p(x) = 0$ otherwise.
+%for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_p A / \xi^2$. 
 This holds for large patches, in which the rate at which mutations arise within the patch is larger 
@@
 \label{ss:patchymigration}
-
+\gc{There's a bunch of switching between x and r in this section}
 Now suppose that there are two patches, each of area $A$ and separated by distance $r$. 
@@
 
+\gc{I'd suggest making Figure 2 and Fig 1A all part of same figure
+  they are showing the same basic thing good fit of equation of
+  cline. The s's in 1A should be labeled sm and sb. I'd suggest that
+  the second part of 1 become its own figure or be paired with fig.3
+  as they are both about rare excursions. Could you make the CI or
+  sample path in Figure 1B be different colors, the grey and black are
+too close to each other.}
+
 \begin{figure}[ht!]
@@
   and the genealogy of 50 individuals in a rare long-distance excursion is traced back with black lines.
-  }
+  }   \label{fig:sim_snapshots}
 \end{figure}
@@
 but far away from the patch, 
-the equilibrium is composed of rare excursions of a family of $B$ alleles, interspersed by long periods of absence.
+the equilibrium is composed of rare excursions of a family of $B$
+alleles, interspersed by long periods of absence. An example of one of these rare
+excursions of a migrant families is shown in Figure \gc{XXX} in \gc{XXXcolour}.
 The relevant time scale on which $B$ alleles migrate between patches is given by the rate of appearance of such families.
@@
 to obtain an approximation that is valid if the distance $r$ is large (at least five times $\sigma/\sqrt{s_m}$, say).
+\gc{We now work through the steps in developing this approximation,
+  and give our resulting approximation for $\migrate(y)$ in eqn. \eqref{eqn:migrate}. }
 
@@
   \end{center}
-  \label{fig:sim_occupation_freqs}
+
   \caption{
@@
   with $C$ chosen arbitrarily to give a reasonable fit.
-  }
+  }   \label{fig:sim_occupation_freqs}
 \end{figure}
@@
 Since we are concerned with the rare migrant families that transit quickly between patches, 
-backmigrants to the already established patch will contribute little, since they are going the wrong direction;
+back migrants to the already established patch will contribute little, since they are going the wrong direction;
 so to simplify the analysis we ignore this detail.  % but could add this detail in an appendix
@@
 the expected family size after time $t$ is $(1-s_m)^t$.
-Therefore, if we define $p_e(t)$ to be the chance of extinction by time $t$, 
+Therefore, if we define $p_e(t)$ to be the chance of a \gc{family}
+going extinct by time $t$, 
 and let $K_t$ be the (random) family size conditioned on nonextinction,
@@
 
+\gc{We exploit this idea to consider the survival of a central ``trunk'',
+  and then conditional on that consider the family size of side
+  ``branches'' off of this trunk. This process is depicted in Figure~\ref{fig:branching_decomp}.}
 We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} \citep[see also][]{chauvin1991growing}.
-When we condition on survival until $t$, we condition on existence of at least one lineage surviving from time $0$ to time $t$.
-Once given this lineage, nothing distinguishes any of the other reproduction events --
-each individual in this lineage must give birth to at least one offspring,
+When we condition on survival until $t$, we condition on existence of
+at least one lineage surviving from time $0$ to time $t$ (the long red
+lineage in Figure \ref{fig:branching_decomp}).
+\gc{Given this lineage, nothing distinguishes any of the other reproduction events --
+each individual along this lineage} must give birth to at least one offspring,
 and all other reproduction events are unconditioned.
@@
 which sprouts independent ``branches'' that may or may not survive until $t$.
-This process is depicted in Figure~\ref{fig:branching_decomp}A.
 Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
@@
 \end{align}
-where each $Z^{(s,k)}$ is an independent copy of $Z$,
+where each $Z^{(s,k)}$ is an independent copy of $Z$ \gc{that arose
+  from the main branch in generation $s$},
 and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.
+\gc{Is there no 1+ in the equation above for the main branch? }
 
@@
 and those that did not survive until $t$ (in grey).
-Figure~\ref{fig:sim_snapshots} has a depiction of such a family, from a simulation.
-\label{fig:branching_decomp}
-}
+Figure~\ref{fig:sim_snapshots} has a depiction of such a family, from
+a simulation. \gc{In reality the spatial paths taken by lineages are
+  much more wiggly but they are shown as straight lines here for simplicity.}
+} \label{fig:branching_decomp}
 \end{figure}
@@
 We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
-as the sum of independent migrant families, each of whose dynamics are given by a spatial branching process.
+as the sum of independent migrant families, each of whose dynamics are
+given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
 Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
@@
 By examining the constituent terms,
-it will turn out that \eqref{eqn:gestalt_migrate} is a constant multiple of \eqref{eqn:gestalt_q}.
-
-First consider \eqref{eqn:gestalt_q}.
+it will turn out that eqn. \eqref{eqn:gestalt_migrate} is a constant multiple of \eqref{eqn:gestalt_q}.
+
+First consider eqn. \eqref{eqn:gestalt_q}.
 Suppose that $Z$ is one such spatial branching process as above, started with a single individual at $x$,
-defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the set $A$.
+defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the \gc{area} $A$.
 The \emph{mean occupation density} of $Z$ can be thought of 
@@
 \end{align}
+\gc{do we need to talk about u a little more it }
 Since reproduction and spatial motion are independent,
@@
 \end{align}
+\gc{$\alpha$ introduced here? Switch from $(1-s_m)^t$ }
 The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
@@
 \end{align}
+\gc{define $\gamma(x)$ presumably the outflux of branches leaving ball}
 
@@
 Therefore, mutations selectively exclude each other from patches, over short time scales, 
-and will only slowly mix together over longer time scales by drift.
+and will only slowly mix together over longer time scales by drift \gc{and
+migration. We made use of this selective exclusion of alleles in our
+previous paper \citep{RalphCoop}. In reality different mutations may
+not be truly selective equivalent in which case the mixing will occur
+over a time-scale dictated by the difference in selection
+coefficients. }
 
 We will assume that once a $B$ allele is introduced (by migration or mutation), 
-it becomes established in the poor habitat patch rapidly. 
+\gc{condition on it escaping loss,} it becomes established in the poor habitat patch rapidly. 
 Under this approximation, and the ``selective exclusion'' just discussed,
@@
 As such, the question of how the the second patch adapts
-simply comes down to whether a new mutation or a migrant allele is the first to become established in the second patch.
-
-Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
+simply comes down to whether a new mutation or a migrant allele is the first to become established in the second patch. Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
 we see that if the patches each have sufficiently large area,
@@
 if $R/\sigma > (\log(1/s_m) - \log(A\mu))/\sqrt{2s_m}$.
-
-If we take the mutational influx in a patch to be one per 1,000 generations ($A \mu = .001$),
+\gc{the two paragraphs either side of this will need some reworking.}
+
+If we take the mutational influx in a patch to be one per 1,000
+generations ($A \mu = .001$), \gc{don't think this is mutational
+  influx as it doesn't have $\rho in it$}
 and the selection against transiting alleles $s_m = .01$,
@@
 we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
-as a function of various parameters.
+as a function of various parameters. \gc{Parallel mutation will occur
+where $\Tmut \gg \Tmig $ (reder areas), while adpatation by migration will occur
+where $\Tmig \gg \Tmut $ (bluer areas).}
 
@@
 and assuming both rates are small, 
-the time until the first of the two patches is colonized by the $B$ allele will be approximately exponentially distributed with mean $1/(2 \mutrate)$.
+the time until the first of the two patches is colonized by the $B$
+allele will be approximately exponentially distributed with mean $1/(2
+\mutrate)$.
 Following this, the time till the second patch is subsequently colonized 
@@
 \begin{equation}
-  \frac{\mutrate}{\mutrate+\migrate} = \frac{2 A \mu s_m}{2 A \mu s_m + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
+  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A \mu \alpha +
+    \left(r \sqrt{2 s_m} r / \sigma \right)^{(d-1)/2} \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
 \end{equation}

### 2014-05-12 standing-variation-paper.tex
edited text through the probability of parallel mig

@@
 \usepackage{hyperref}
-\usepackage{latexml} % for \iflatexml
+%\usepackage{latexml} % for \iflatexml
+\usepackage{color}
+\newcommand{\gc}[1]{{\it\color{green}(#1)} }
+\newcommand{\plr}[1]{{\it\color{blue}(#1)}}
 
@@
 
-For now we ignore geographic hetergeneity, and assume that the species range
+We assume that the species range
 is a homogenous, one- or two-dimensional region. 
@@
 where $s_b>0$ will usually be assumed to be small, and $0<s_d<1$.
-We assume that diploid fitness is additive, or at least that the important dynamics are determined by the haploid fitness.
+We assume that diploid fitness is additive, or at least that the
+important early dynamics are determined by the haploid fitness
+\gc{heterozygous fintess, with no reference to the fitness of the
+  homozygote for the mutant alleles}.
 The number of offspring has finite variance, which to keep the formulas simple we assume to be equal to one.
@@
 
-XXX More discussion on whether this is a good approximation?  
+\plr{XXX More discussion on whether this is a good approximation?  
 Compute typical size and radius of a standing cluster?
 
-recall PPP computations; draw cones of opportunity for new mutations
+recall PPP computations; draw cones of opportunity for new mutations }
+
+\gc{Point out connection to standard mutation selection balance formula}
 
@@
 as well as a similar compound parameter describing the relative importance of standing variation to the process of adaptation.
-(Note that a similar approach, summarizing with a {\em characteristic length} was taken by \citep{slatkin1973geneflow} for a different problem.)
+%(Note that a similar approach, summarizing with a {\em characteristic length} was taken by \citep{slatkin1973geneflow} for a different problem.)
 
@@
 
+\section{Applications}
+
+% G6PD frequency across world http://www.google.com/imgres?imgurl=http%3A%2F%2Fwww.plosmedicine.org%2Farticle%2Finfo%3Adoi%2F10.1371%2Fjournal.pmed.1001339.g002%2Flargerimage&imgrefurl=http%3A%2F%2Fwww.plosmedicine.org%2Farticle%2Finfo%253Adoi%252F10.1371%252Fjournal.pmed.1001339&h=2938&w=2381&tbnid=oD_tX3ea8IbLfM%3A&zoom=1&docid=k37axqAZDS7ljM&ei=5rdOU5TVDMqj8QGntICYBg&tbm=isch&ved=0CF8QMygLMAs&iact=rc&uact=3&dur=1447&page=1&start=0&ndsp=15
+
+% Spatial distribution of G6PD alleles:  http://www.malariajournal.com/content/12/1/418
+
 \bibliography{standing_patches_refs}

### 2014-05-14 lineage-movement.tex
lineage movement.

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\deriv}[1]{\frac{d}{d#1}}
\newcommand{\dderiv}[1]{\frac{d^2}{d^2#1}}
\newcommand{\given}{\;\vert\;}

\begin{document}

\subsection{Discrete model}

Consider a discrete model in which the total number of individuals at location $x$ is $N(x)$,
each individual is of type either A or B at a selected locus,
and the number of individuals of type A and location $x$ is $n(x)$.
Suppose that type A individuals at location $x$ reproduce at rate $1+s_A(x)$, 
and likewise type B at rate $1+s_B(x)$.
At reproduction, individuals recombine with others in the same location,
and the offspring choose a new location according to the kernel $m(x,y)$.
The population dynamics are random, but suppose that $N(x)$ is sufficiently large that these do not vary with time;
instead we want to follow a lineage backwards in time,
at a locus separated from the selected locus by recombination fraction $r$.

Suppose this is a Moran model.
There are four things that can happen:
\begin{enumerate}
    \item[$x\xrightarrow{AA}y$] One type A individual at location $x$ reproduces, 
        either does not recombine or recombines with another type A,
        and sends the offspring to $y$.
    \item[$x\xrightarrow{AB}y$] An individual at location $x$ reproduces, 
        recombines with the other type,
        and sends to $y$ an offspring
        who inherits at the selected locus from the type B parent 
        and the at the neutral locus from the type A parent.
    \item[$x\xrightarrow{BA}y$] An individual at location $x$ reproduces, 
        recombines with the other type,
        and sends to $y$ an offspring
        who inherits at the selected locus from the type A parent 
        and the at the neutral locus from the type B parent.
    \item[$x\xrightarrow{BB}y$] One type B individual at location $x$ reproduces, 
        either does not recombine or recombines with another type B,
        and sends the offspring to $y$.
\end{enumerate}
Let $p(x) = n(x)/N(x)$.
These four things happen at rates:
\begin{align}
    & x\xrightarrow{AA}y & \qquad w_{AA}(x,y) &= p(y) (1+s_A(y)) \left(1 - r (1-p(y)) \right)  N(y) m(y,x) \\
    & x\xrightarrow{AB}y & \qquad w_{AB}(x,y) &= r p(y) (1-p(y)) \left(1+\frac{s_A(y)+s_B(y)}{2}\right) N(y) m(y,x) \\
    & x\xrightarrow{BA}y & \qquad w_{BA}(x,y) &= r p(y) (1-p(y)) \left(1+\frac{s_A(y)+s_B(y)}{2}\right) N(y) m(y,x) \\
    & x\xrightarrow{BB}y & \qquad w_{BB}(x,y) &= (1-p(y)) (1+s_B(y)) \left(1 - r p(y) \right) N(y) m(y,x) 
\end{align}

These rates tell us precisely the rates at which a lineage will move, backwards in time.
For instance, the rate at which a lineage currently in a type A individual at location $x$
jumps to another type A individual at location $y$ is just $w_{AA}(y,x)$.
A lineage at the selected locus (with $r=0$) in a type A individual moves according to these jump rates.
If we follow a linked locus until the first time it is found in a type B individual,
it moves according to these jump rates, and at location $x$ recombines at rate
\begin{align}
    \gamma_A(x) = \sum_y w_{BA}(y,x) .
\end{align}
Suppose that we kill the lineage when it recombines onto type B.
Let $X_t$ denote the position of the lineage at time $t$ in the past,
with $X_t = \dagger$ if it has recombined onto B,
and let $f$ be test function with $f(\dagger)=0$.
If we let $M_r(y) = p(y) (1+s_A(y)) \left(1 - r (1-p(y)) \right)  N(y)$, then
\begin{align}
    \deriv{t} \E[f(X_t) \given X_0=x ] &= \sum_y w_{AA}(y,x) ( f(y)-f(x) ) - \gamma_A(x) f(x) \\
                                       &= \sum_y  M_r(y) m(y,x) ( f(y) - f(x)  ) - \gamma_A(x) f(x) .
\end{align}

\section{Diffusion limit}

Now suppose that $m(x,y)$ is symmetric, and depends on a parameter $\sigma$ so that as $\sigma \to 0$,
the associate random walk converges to Brownian motion, so that
\begin{align}
    \lim_{\sigma \to 0} \sum_y \frac{ m(x,y) ( f(y) - f(x) ) }{\sigma^2} = \frac{1}{2} \dderiv{x} f(x) .
\end{align}
Write $f'(x) = \deriv{x}f(x)$, and note that
\begin{align}
    \frac{1}{\sigma^2} \sum_y g(y) m(y,x) (f(y)-f(x)) 
    &= \frac{1}{\sigma^2} \sum_y m(y,x) \left( g(y) f(y) - g(x) f(x) + (g(x)-g(y)) f(x) \right) \\
    &= \frac{1}{\sigma^2} \sum_y m(y,x) \left( g(y) f(y) - g(x) f(x) \right) \\
    & \qquad - f(x) \frac{1}{\sigma^2} \sum_y m(y,x) (g(y)-g(x)) \\
    &\xrightarrow{\sigma \to 0} \frac{1}{2} \dderiv{x}\left( g(x)f(x) \right) - \frac{1}{2} f(x) \dderiv{x} g(x) \\
    &= \frac{1}{2} \left( g(x) f'(x) + 2 g'(x) f'(x) + f(x) g''(x) - f(x) g''(x) \right) \\
    &= \frac{1}{2} g(x) f''(x) + g'(x) f'(x) .
\end{align}
If we also rescale recombination, so that $\rho = r/\sigma^2$,
then 
\begin{align}
    \frac{1}{\sigma^2} \gamma_A(x) \xrightarrow{\sigma \to 0} \rho p(x) (1-p(x)) \left( 1 + \frac{s_A(x)+s_B(x)}{2} \right) N(x) ,
\end{align}
which we can define to be equal to $\rho k(x)$.
In this case, note that since $r = \rho \sigma^2$,
\begin{align}
    M_r(x) \xrightarrow{\sigma \to 0} M(x) = p(x) (1+s_A(x)) N(x) .
\end{align}
Under these assumptions,
\begin{align}
    \deriv{t} \E[f(X_{t/\sigma^2}) \given X_0=x ] &\xrightarrow{\sigma \to 0} 
    \frac{1}{2} M(x) f''(x) + M'(x) f'(x) - \rho k(x) f(x) ,
\end{align}
i.e.\ $X_{t/\sigma^2}$ converges to a diffusion with drift $M'(x)$ and killed at rate $\rho k(x)$.

\paragraph{Stationary distribution}
If $X$ has a stationary distribution then its density $\pi(x)$ solves the forward Kolmogorov equation,
\begin{align}
    0 &= - \deriv{x}\left( M'(x) \pi(x) \right) + \frac{1}{2} \dderiv{x} \left( M(x) \pi(x) \right) \\
      &= - M''(x) \pi(x) - M'(x) \pi'(x) + \frac{1}{2} M''(x) \pi(x) + M'(x) \pi'(x) + \frac{1}{2} M(x) \pi''(x)  \\
      &= \frac{1}{2} \left( M(x) \pi''(x) - M''(x) \pi(x) \right) .
\end{align}
It is immediate that $\pi(x) = M(x)/Z$,
where
\begin{align}
    Z = \int p(x) (1+s_A(x)) N(x) dx .
\end{align}

\paragraph{Recombination time}
Let $\tau_\dagger$ be the first time  the lineage recombines onto type B.
If $\rho$ is much smaller than the mixing time of a type A lineage,
then the time to recombination is roughly Exponential,
with mean
\begin{align}
    \E[\tau_\dagger] \approx \frac{1}{ \rho \int \pi(x) k(x) dx } .
\end{align}
The product $\pi(x) k(x)$ is:
\begin{align}
    \pi(x) k(x) = \frac{1}{Z} N(x)^2 p(x)^2 (1-p(x)) (1+s_A(x)) (1+(s_A(x)+s_B(x))/2) .
\end{align}

More generally, if we define
\begin{align}
    h(x) = \E[\tau_\dagger \given X_0 = x]
\end{align}
then $h$ solves the equation
\begin{align}
    0 = M'(x) h'(x) + \frac{1}{2} M(x) h''(x) - \rho k(x) h(x) + 1 .
\end{align}

\end{document}

### 2014-05-14 lineage-movement.tex
.

@@
 Let $X_t$ denote the position of the lineage at time $t$ in the past,
-with $X_t = \dagger$ if it has recombined onto B,
-and let $f$ be test function with $f(\dagger)=0$.
+with $X_t = \rho$ if it has recombined onto B,
+and let $f$ be test function with $f(\rho)=0$.
 If we let $M_r(y) = p(y) (1+s_A(y)) \left(1 - r (1-p(y)) \right)  N(y)$, then
@@
 \paragraph{Recombination time}
-Let $\tau_\dagger$ be the first time  the lineage recombines onto type B.
+Let $\tau_\rho$ be the first time  the lineage recombines onto type B.
 If $\rho$ is much smaller than the mixing time of a type A lineage,
@@
 \begin{align}
-    \E[\tau_\dagger] \approx \frac{1}{ \rho \int \pi(x) k(x) dx } .
+    \E[\tau_\rho] \approx \frac{1}{ \rho \int \pi(x) k(x) dx } .
 \end{align}
@@
 \begin{align}
-    h(x) = \E[\tau_\dagger \given X_0 = x]
+    h(x) = \E[\tau_\rho \given X_0 = x]
 \end{align}
@@
 
+The probability that a segment of length $r = \rho \sigma^2$ to one side of the selected locus
+does not recombine across time $t$ is $\P\{\tau_\rho > t\}$.
+Therefore, if $L_t$ is the length of the nonrecombined segment to one side of the selected locus
+after time $t$, then
+\begin{align}
+    \E[ L_t \given X_0=x ] &= \sigma^2 \int_0^\infty \P\{ \tau_\rho > t \} d\rho \\
+\end{align}
+The appropriate thing to do is to look at $\E[L_T]$, where $T$ is the coalescence time.
+
 \end{document}

### 2014-05-15 patchy-selection-paper.tex

@@
 \subsection{Multiple patches}
-
 We can extend this reasoning to multiple patches of unadapted habitat 
@@
 
-If not yet adapted,
-a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$.
+If not yet adapted, a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$.
 Without loss of generality, assume that patches $1, \ldots, k$ are already adapted,
@@
 \begin{equation}
-  \frac{ s_p \rho }{s_m \xi^2} \sum_{i=1}^{k} \exp\left(- r_i \sqrt{2 s_m} /\sigma\right),
+  \frac{ s_p \rho }{\alpha \xi^2} \sum_{i=1}^{k} \left(r_i \sqrt{2 s_m} /\sigma \right) \exp\left(- r_i \sqrt{2 s_m} /\sigma\right),
 \end{equation}
-and the probability that the migrant, adapted allele 
+and conditional on coming from migration the probability the adapted allele 
 comes from patch $i$ is 
 \begin{align}
-  \exp\left(- r_i \sqrt{2 s_m} /\sigma\right)/\sum_{j=1}^{k} \exp\left(- r_j \sqrt{2 s_m} /\sigma\right) .
+  \frac{\left(r_i \sqrt{2 s_m} /\sigma \right) \exp\left(- r_i \sqrt{2 s_m}
+    /\sigma\right)} {\sum_{j=1}^{k}  \left(r_i \sqrt{2 s_m} /\sigma \right) \exp\left(- r_i \sqrt{2 s_m}
+    /\sigma\right) } .
 \end{align}
@@
 
-If a patch adapts through new mutation, the genomic region surrounding the selected site will hitchhike \citep{maynardsmith1974hitchhiking,kaplan1989hitchhiking} along with it,
-so that at least initially, all adapted individuals share a fairly long stretch of haplotype.
+If a patch adapts through new mutation or a rare migrant lineage, the
+genomic region surrounding the selected site will hitchhike along with
+it \citep{maynardsmith1974hitchhiking},
+so that at least initially, all adapted individuals within a patch
+share a fairly long stretch of haplotype. \gc{Pairs of individuals
+  within a patch will share a haplotype of mean genetic length of
+  $\approx \approx \log(\rho A s_p)/s_p$ around the selected site,
+as long as the patch is reasonably well mixed by dispersal
+\citep[otherwise see][]{BartonEtheridge13spatialHH}. }
 This association gets slowly whittled down by recombination during interbreeding at the edge of the patch;
@@
 
-On the other hand, if an already adapted patch colonizes another through migration,
+When an already adapted patch colonizes another through migration,
 then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
-The length of this haplotype is roughly inversely proportional to the number of generations the allele spends ``in transit'',
-because each reproductive event in locations where the allele is at low frequency
-recombines haplotypes of the $b$ allele closer to the adaptive allele.
+The genetic length of this haplotype is roughly inversely proportional to the
+number of generations the allele spends ``in transit'', 
+because each reproductive event in locations where the allele is at low frequency recombines haplotypes of the $b$ allele closer to the adaptive allele.
 (In this case, the haplotype is literally hitchhiking across geography!)
 
-As noted above, the length of the hitchhiking segment depends on the time
-spent ``in transit'' between the patches.
 Above we argued that a good model for this transit is a single Brownian trunk lineage
@@
 Consider a single such family, and let $\tau$ be the (random) time it first hits the new patch,
-with $\tau = \infty$ if it never does.
+with $\tau = \infty$ if it never does. The length of the hitchhiking segment depends on the time
+spent by this trunk lineage ``in transit'' between the patches, i.e. $\tau$.
+
 By pulling out the constant $1/\E[K]$, we can use standard results on hitting times of $d$-dimensional Brownian motion
-that is killed at rate $\alpha$.
-In particular, we know that if the patch is circular with radius $w$, then
-using \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook}
+that is killed at rate $\alpha$ \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook} 
+In particular, if the patch is circular with radius $w$ and a distance
+$R$ from the already adapted patch, then we have the useful result
+about the Laplace transform of $\tau$ our transit time
 \begin{align}
@@
       \frac{ K_0( (R+w)\sqrt{2(\alpha+\ell)}/\sigma) }{ K_0( w\sqrt{2(\alpha+\ell)}/\sigma) } \qquad & d=2 
-    \end{cases} 
+    \end{cases} \label{borodinresult}
     % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(\alpha+\ell)}/\sigma}  (1+O(1/w)) .
 \end{align}
-In particular, we can find 
-$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E[\exp(-\ell \tau)]$. % \simeq (R/w+1)^{-(d-1)/2} \exp(- R \sqrt{2\alpha}/\sigma)$.
+\gc{define $K_0$, bessel? We've used K for family size}
+We can find that the probability that a particular
+trunk lineageis killed before it reaches the new patch is
+$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau)
+\right]$. We are interested in lineages that
+manage to establish on the patch i.e. $\tau< \infty$, and so in one dimension
+\begin{equation} \label{eqn:haplen_cdf}
+\E[e^{-\ell \tau}|\tau<\infty]  = \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
+\end{equation}
+We can differentiate this to find that our expected transit time is
+\begin{equation}
+\E[\tau] = (R\sqrt{2\alpha}/\sigma)\times 1/(2\alpha)
+\end{equation}
+and $\var[\tau] = (R\sqrt{2\alpha}/\sigma) \times 1/(2\alpha)^2$ in
+one dimension.
+
+% \simeq (R/w+1)^{-(d-1)/2} \exp(- R \sqrt{2\alpha}/\sigma)$.
 
@@
 Since each generation spent in transit provides an opportunity for recombination,
-if recombination is Poisson, the length of the hitchhiking segment (in Morgans) on each side of the site will be exponentially distributed
-with mean $\tau$, so that if $L$ is the length on, say, the right, then in one dimension
-\begin{align} \label{eqn:haplen_cdf}
-\P\{L>\ell\} &= \E[e^{-\ell \tau}|\tau<\infty] \\
-  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
-\end{align}
-This is also the Laplace transform of $\tau$;
-we can differentiate it to find that $\E[\tau] = (R\sqrt{2\alpha}/\sigma)\times 1/(2\alpha)$
-and $\var[\tau] = (R\sqrt{2\alpha}/\sigma) \times 1/(2\alpha)^2$.
-Furthermore, this implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
+if recombination is Poisson, the length of the haplotype (in Morgans)
+initially shared between populations on each side of the site will be exponentially distributed
+with mean $\tau$, so that if $L$ is the length of hitchhiking segment on, say, the right of
+the selected locus, then
+\begin{equation} \label{eqn:haplen_cdf2}
+\P\{L>\ell\} = \E[e^{-\ell \tau}|\tau<\infty] 
+%  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
+\end{equation}
+which is our Laplace transform from above that in one dimension
+gives us eqn. \eqref{eqn:haplen_cdf}. This implies that in one dimension, if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
 then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$,
-from which we can obtain that
-$\E[L] = \sigma \sqrt{2\alpha}/R + \sigma^2/R^2$ and $\var[L] = 2\alpha\sigma^2/R^2 + 4 \sigma^3 \sqrt{2\alpha}/R^3 + 5 \sigma^4 / R^4$.
+from which we can obtain that the expected length of shared hitchiking
+haplotype is
+\begin{equation} \label{eqn:haplotype_length}
+\E[L] = \sigma \sqrt{2\alpha}/R + \sigma^2/R^2
+\end{equation}
+and thar $\var[L] = 2\alpha\sigma^2/R^2 + 4 \sigma^3 \sqrt{2\alpha}/R^3 + 5 \sigma^4 / R^4$.
 For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
-but how other properties might change is unclear.
+but how other properties might change is unclear. \gc{add some
+  sentences of intuition of how parameters play into haplotype length
+  here. is it really $\sigma \sqrt{\alpha}$ in \eqref{eqn:haplotype_length} and not $\sqrt{\alpha}/sigma$???}
+
+\gc{Perhaps we should separate the above bit into things that are
+  useful about $\tau$ and put those up by  \eqref{borodinresult} and
+  things that are useful about $L$ here? We should also turn some of
+  this into numbered equations. I tried to do this above, and have the
+commented out original below}
+
+% Since each generation spent in transit provides an opportunity for recombination,
+% if recombination is Poisson, the length of the hitchhiking segment (in Morgans) on each side of the site will be exponentially distributed
+% with mean $\tau$, so that if $L$ is the length of hitchhiking segment on, say, the right of
+% the selected locus, then in one dimension
+% \begin{align} \label{eqn:haplen_cdf}
+% \P\{L>\ell\} &= \E[e^{-\ell \tau}|\tau<\infty] \\
+%   &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
+% \end{align}
+% found by using eqn. \eqref{borodinresult}. This is also the Laplace transform of $\tau$;
+% we can differentiate it to find that $\E[\tau] = (R\sqrt{2\alpha}/\sigma)\times 1/(2\alpha)$
+% and $\var[\tau] = (R\sqrt{2\alpha}/\sigma) \times 1/(2\alpha)^2$.
+% Furthermore, this implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
+% then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$,
+% from which we can obtain that
+% $\E[L] = \sigma \sqrt{2\alpha}/R + \sigma^2/R^2$ and $\var[L] = 2\alpha\sigma^2/R^2 + 4 \sigma^3 \sqrt{2\alpha}/R^3 + 5 \sigma^4 / R^4$.
+% For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
+% but how other properties might change is unclear.
 

### 2014-05-15 lineage-movement.tex
lineage movement

@@
 \begin{align}
-    0 = M'(x) h'(x) + \frac{1}{2} M(x) h''(x) - \rho k(x) h(x) + 1 .
+    -1 = M'(x) h'(x) + \frac{1}{2} M(x) h''(x) - \rho k(x) h(x) .
 \end{align}
@@
 \end{align}
+
 The appropriate thing to do is to look at $\E[L_T]$, where $T$ is the coalescence time.
+Generally, then
+\begin{align}
+  \P\{ L_t > \rho \} = \P\{ \tau_\rho > T \} .
+\end{align}
+Do do this properly, we need to run two diffusions.
+However, if we let $T$ instead be independent Exponential($\lambda$), then 
+\begin{align}
+  \ell(x) = \P\{ \tau_\rho > T \given X_0 = x\} 
+\end{align}
+solves the equation
+\begin{align}
+  M'(x) \ell'(x) + \frac{1}{2} M(x) \ell''(x) - (\lambda + \rho k(x) ) \ell(x) = \lambda .
+\end{align}
+
+\paragraph{Questions:} 
+\begin{enumerate}
+
+    \item Can we make any progress analytically with any of these quantities,
+      using the fact that
+      \[ \frac{1}{2} p''(x) + s(x) p(x)(1-p(x)) = 0 \quad ?\]
+
+    \item Do we get the same system using, say, a Wright--Fisher model?  (I assume so?)
+
+    \item Can we prove convergence of the finite model to this one, under suitable assumptions?
+A main difficulty here would be showing that the lineage spends all its time in the part of the range where $p(x)$ does not fluctuate significantly, i.e.\ is not too small\ldots but if this is not true, it would be even more interesting.
+
+\end{enumerate}
 

### 2014-05-15 lineages.R

@@ -4,7 +4,6 @@ require(colorspace)
 #  and record statistics about their path
 
 lineages <- function ( pophist, nlin, migrsteps, T=dim(pophist)[4], m, linit=NULL, coal=TRUE ) {
-
     # simulate the movement of nlin selected lineages back through pophist,
     # beginning at locations in linit 
     #   which is a nlin x 3 matrix recording (x,y,t)

### 2014-05-17 patchy-selection-paper.tex
rejiggered plots

@@
 repeatedly sculpt phenotypic variation. 
-In some cases this convergence extends to the molecular level with
-dispate taxa converging to the same phenotype through parallel genetic
-changes in the same pathway, genes, 
+In some cases this convergence extends to the molecular level, 
+with disparate taxa converging to the same phenotype through parallel 
+genetic changes in the same pathway, genes, 
 or even by precisely the same genetic changes \citep{zhen2012parallel}. 
-Such precise convergence points to the conservation of function of the genes
-underlying adaptive traits, 
+Such precise convergence points to the conservation of function of the genes underlying adaptive traits, 
 sometimes over deep time scales \citep{deephomologypapers}. 
@@
 There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
-All else being equal, such evolution of convergent phenotypes is more likely with a 
-higher mutational input, that is, when the mutational target and population sizes are larger \citep{}. 
+Such evolution of convergent phenotypes is favored by higher mutational input, 
+i.e.\ larger mutational target and/or population size \citep{dmitri}.
 The geographic distribution of populations can also affect the probability of parallel mutation within a species:
@@
 at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{nachman2003different}. 
-Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}.
+Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
+and in lizards \citep{whitesands}.
 
@@
 
-%http://www.nature.com/hdy/journal/v94/n2/full/6800600a.html
+  new mutation: \eqref{eqn:mutrate};
+
+% http://www.nature.com/hdy/journal/v94/n2/full/6800600a.html   -- this is \citep{hoekstra2004local}
 
@@
 ($B$) that adapts an individual to the poor habitat type.
-This change could be as specific as a single base change, a knock out
-of a gene, or a mutation anywhere in a set of genes. We assume that
-this mutational change occurs at a (low) rate of $\mu$ per chromosome per generation. 
-We assume that this initially rare new
-allele has fitness $1+s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches.
-Outside of these patches we assume that $B$ has a relative fitness
-$1-s_m$  when rare in the intervening areas, with $s_p, s_m > 0$.
+The required change could be as specific as a single base change, 
+or as general as a knock out of a gene or one of a set of genes.
+We assume that this mutational change occurs at a (low) rate of $\mu$ per chromosome per generation,
+and that this initially rare new allele has fitness $1+s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches.
+Outside of these patches we assume that the new allele has a relative fitness
+$1-s_m$  when rare in the intervening areas, with $s_p$ and $s_m$ both positive.
 To simplify matters, we assume that the disadvantage $s_m$ 
@@
 are determined by the fitness of the heterozygote. 
-More general dominance will only make small corrections,
+More general dominance will only make small corrections to the the dynamics until initial fixation,
 with the exception of the recessive case, which we omit.
@@
 For the sake of simplicity, we assume a constant haploid population density $\rho$, 
-across both types of habitat (we return to discuss this equal density point later). 
+across both types of habitat (but return to this point later),
 We further assume that the variance in offspring number is $\xi^2$, 
@@
 We also aim to describe the scenario where it is sensible to think of large, discrete patches;
-the case of a fine-scale hetergeneous environment is treated \citep{elsewhere}.
-
-\paragraph{Previous work} 
+the case of a fine-scale heterogeneous environment is treated \citep{elsewhere}.
+
+\paragraph{Past work on spatial models}
 We will make use of several previous results from the literature on
-migration in spatially varying selection models. \gc{cite Haldane and zfisher} 
-Haldane, Fisher, and later \citet{slatkin1973geneflow} first analyzed
+migration in spatially varying selection models. 
+\citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
 deterministic models of of this sort.
@@
 
-Let $p(x)$ be the probability that a new mutant $B$ allele arising at
-location $x$
+Let $p(x)$ be the probability that a new mutant $B$ allele arising at location $x$
 relative to the patch fixes within the poor habitat patch.
@@
 % the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
-but these are not particularly helpful.
-The total successful mutational influx per generation is given by the
+but results will be more interpretable if we proceed with a simple approxmation.
+The total influx per generation of mutations that successfully establish is given by the
 integral of $p(x)$ over the entire species range:
@@
 % $\int \rho \mu p(x) dx$,
-and hence depends in a complicated way on the relationship of patch width and selection coefficients,
+and hence depends in a complicated way on the patch geometry and selection coefficients,
 but still scales linearly with the mutational influx density $\rho \mu$.
-If we consider a patch of area $A$, whose width of the patch is large relative to $\sigma/\sqrt{2s_m}$, 
-then a reasonable approximation is to assume that $p(x) = p_e$ within the patch, and $p(x) = 0$ otherwise.
+If we consider a patch of area $A$, whose width is large relative to $\sigma/\sqrt{2s_m}$, 
+then a reasonable approximation is to assume that $p(x) = p_e \approx 2 s_p / \xi^2$ within the patch, and $p(x) = 0$ otherwise.
 %for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_p A / \xi^2$. 
-This holds for large patches, in which the rate at which mutations arise within the patch is larger 
-than the rate at which they appear nearby;
-we examine this more generally via simulation in Appendix \ref{apx:establishment_sims}.
-
-The rate at which mutations arise and colonize a patch of area $A$ is therefore
+We would then approximate the integrand $p(x)$ in \eqref{eqn:exactmutrate} by a step function,
+which will be a good approximation if the patch is large relative the scale over which $p(x)$ goes from 0 to $p_e$,
+or if $p(x)$ is symmetric about the edge of the patch.
+We examine this more generally via exact calculation of $p(x)$ in Appendix \ref{apx:establishment_sims}.
+
+The rate at which mutations arise and colonize a patch of area $A$ is then
 \begin{align} \label{eqn:mutrate}
-  \mutrate = \rho \mu \int p(x) dx  \approx 2 s_p \rho A \mu / \xi^2.
-\end{align}
+  \mutrate % = \rho \mu \int p(x) dx  
+  \approx 2 s_p \rho A \mu / \xi^2,
+\end{align}
+i.e.\ the product of mutational influx in the patch ($A\rho\mu$) and the probability of establishment ($2s_p/\xi^2$).
 If this rate is low, then the time (in generations) until a mutation arises that
@@
 \gc{There's a bunch of switching between x and r in this section}
-Now suppose that there are two patches, each of area $A$ and separated by distance $r$. 
+Now suppose that there are two patches, each of area $A$ and separated by distance $R$. 
 If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
@@
 To determine this, we dissect the process by which an allele transits between the patches,
-in the process obtaining other useful information about the genealogy.
+in the process obtaining other useful information about the genealogy of B alleles.
 
@@
 too close to each other.}
+
+Migration--selection balance ensures that 
+there will be some number of $B$ alleles present in the regions where they are disadvantageous,
+but only very few, and rarely, far away from the patch where $B$ is established.
+Denote the expected frequency of allele $B$ at location $x$ relative to the patch by $q(x)$,
+and assume that the shape of the patch is at least roughly circular.
+Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(x)$ is the solution,
+and show that $q(x)$ decays exponentially for large $|x|$,
+with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}): 
+\begin{align} \label{eqn:eqfreq}
+  q(x) \approx C \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
+\end{align}
+where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients.
+These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
+(In one dimension, the equation can be solved to give $q(x)$ in terms of Jacobi elliptic functions; see Appendix \ref{apx:elliptic_integrals}.)
+In applications we fit $C$ to the data;
+to get concrete numbers from theory we take $C=1$ if necessary.
+From simulations, $C$ seems to be not too far from 1 in practice (see Appendix \ref{apx:patchy_sims}).
+
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics{sim-occupation-freqs}
+  \end{center}
+  \caption{
+  Mean occupation frequencies from simulation; details are as in figure~\ref{fig:sim_snapshots} except as otherwise noted.
+  Each panel shows mean occupation frequencies of individual demes (grey dots)
+  and averaged over demes at similar distances (black circles), on
+  \textbf{(A)} a one-dimensional grid of 201 demes, and
+  \textbf{(B)} a two-dimensional grid of $101\times 101$ demes.
+  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq},
+  with $C$ chosen arbitrarily to give a reasonable fit.
+  }   \label{fig:sim_occupation_freqs}
+\end{figure}
+
+This expected frequency $q(x)$ is the \emph{time-averaged} occupation frequency,
+or in other words, the total number of $B$ alleles found across $T$ generations near location $x$, per unit area, divided by $T$.
+If, for instance, $q(x)=.01$ and the population density is $\rho=10$ individuals per unit area, 
+then in a survey tract of unit area around $x$ we expect to find one individual every 10 generations
+-- or, perhaps more likely, 10 individuals every 100 generations.
+This points to an important fact: 
+close to the original patch, the ``equilibrium frequency'' $q(x)$ describes well the density of the $B$ allele at most times,
+but far away from the patch, 
+the equilibrium is composed of rare excursions of families of $B$
+alleles, interspersed by long periods of absence. 
+An example of one of these rare excursions is shown in Figure~\ref{fig:sim_snapshots}.
+The relevant time scale on which $B$ alleles migrate between patches is given by the rate of appearance of such families.
+To compute the rate of establishment by such migration,
+we argue that these families (suitably defined) can be approximated by
+independent branching processes whose spatial motion is described by Brownian motion.
+Properties of these, combined with equation \eqref{eqn:eqfreq},
+leads us to the expression \eqref{eqn:migrate} for the rate of effective migration,
+an approximation that is valid if the distance $R$ is large (at least five times $\sigma/\sqrt{2s_m}$, say).
 
@@
   }   \label{fig:sim_snapshots}
-\end{figure}
-
-Migration--selection balance ensures that 
-there will be some number of $B$ alleles present in the regions where they are disadvantageous,
-but only very few, and rarely, far away from the patch where $B$ is established.
-Denote the expected frequency of allele $B$ at displacement $x$ relative to the patch by $q(x)$.
-Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(x)$ is the solution,
-which in one dimension can be solved to give $q(x)$ in terms of Jacobi elliptic functions (see Appendix \ref{apx:elliptic_integrals}).
-More general, asymptotics of the solution implies that for large $|x|$ the mean occupation frequency decays exponentially, 
-with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}): 
-\begin{align} \label{eqn:eqfreq}
-  q(x) \approx C \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) \qquad \text{for large $|x|$},
-\end{align}
-where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients.
-In applications we fit $C$ to the data;
-to get concrete numbers from theory we take $C=1$ if necessary.
-From simulations, $C$ seems to be not too far from 1 in practice (see Appendix \ref{apx:patchy_sims}).
-
-This expected frequency $q(r)$ is the \emph{time-averaged} occupation frequency,
-or in other words, the total number of $B$ alleles found at distance $r$ across $T$ generations, divided by $T$.
-Therefore, if we look at a small area of size $dx$, 
-then $q(r) dx$ is equal to the probability of finding any $B$ alleles there,
-multiplied by the mean number of such, if there are any.
-This points to an important fact: 
-close to the original patch, the ``equilibrium frequency'' $q(r)$ describes well the density of the $B$ allele at most times,
-but far away from the patch, 
-the equilibrium is composed of rare excursions of a family of $B$
-alleles, interspersed by long periods of absence. An example of one of these rare
-excursions of a migrant families is shown in Figure \gc{XXX} in \gc{XXXcolour}.
-The relevant time scale on which $B$ alleles migrate between patches is given by the rate of appearance of such families.
-To compute the rate of establishment by such migration,
-we use properties of subcritical branching processes and Brownian motion,
-combined with equation \eqref{eqn:eqfreq},
-to obtain an approximation that is valid if the distance $r$ is large (at least five times $\sigma/\sqrt{s_m}$, say).
-\gc{We now work through the steps in developing this approximation,
-  and give our resulting approximation for $\migrate(y)$ in eqn. \eqref{eqn:migrate}. }
-
-\begin{figure}[ht!]
-  \begin{center}
-    \includegraphics{sim-occupation-freqs}
-  \end{center}
-
-  \caption{
-  Mean occupation frequencies from simulation; details are as in figure~\ref{fig:sim_snapshots} except as otherwise noted.
-  Each panel shows mean occupation frequencies of individual demes (grey dots)
-  and averaged over demes at similar distances (black circles), on
-  \textbf{(A)} a one-dimensional grid of 201 demes, and
-  \textbf{(B)} a two-dimensional grid of $101\times 101$ demes.
-  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq},
-  with $C$ chosen arbitrarily to give a reasonable fit.
-  }   \label{fig:sim_occupation_freqs}
 \end{figure}

### 2014-05-17 patchy-sim-writeup-plots.R

@@ -1,13 +1,82 @@
 #!/usr/bin/R
 source("sim-patchy-selection-fns.R")
 source("lineages.R")
+require(gsl)
 
 run.list <- list.files(".","*-pophistory-run.Rdata")
 rundims <- read.csv("run-info.csv")
 
 # run.name <- "3975-1-10000-pophistory-run.Rdata"
 
+#######
+### occupation frequencies
+plotdecay <- function (pophist, shift=FALSE) {
+    # return averaged freqs and theoretical prediction
+    dimension <- sum(dim(pophist$pophist)[1:2]>1)
+    theory.decay <- sqrt(2*abs(getgrowth(pophist$pop$params)$gm)) / getsigma(pophist$pop$params) 
+    # yes, the parentheses differ in the exponents for x:
+    expl.approx <- function (x,d) { sqrt(pi/2) * x^((1-d)/2) * exp(-x) }
+    bessel.approx <- function (x,d) { x^(1-(d/2)) * bessel_Knu( nu=1-d/2, x=x ) } 
+    # the exp'l decay
+    patchloc <- with(pophist$pop$params, (-1)^(s>0) * sqrt( abs( ( row(s) - range[1]/2 )^2 + ( col(s) - range[2]/2 )^2 - (patchsize/2)^2 ) ) )
+    obs.freqs <- pophist$occupation[,,2]/(pophist$pop$params$N*(pophist$pop$gen-pophist$burnin))
+    expl.freqs <-  expl.approx( (patchloc*theory.decay), d=dimension )
+    # also estimate the shift?
+    if (shift) {
+        f <- function (x0) { 
+            goodones <- (obs.freqs>3e-3) & (patchloc > -x0)
+            bfreqs <- (bessel.approx( (patchloc+x0)*theory.decay, d=dimension ))
+            mean( ((log(obs.freqs)-log(bfreqs))[goodones] - mean(log(obs.freqs)[goodones]) + mean(log(bfreqs)[goodones]) )^2 ) }
+        bessel.shift <- optimize( f, interval=c(0,50) )
+        bessel.shift.const <- exp( mean( log(obs.freqs/bessel.approx( (patchloc+bessel.shift$minimum)*theory.decay, d=dimension ))[(log(obs.freqs)<quantile(log(obs.freqs),.9))&(log(obs.freqs)>quantile(log(obs.freqs),.4))], na.rm=TRUE ) )
[…]

### 2014-05-17 patchy-selection-paper.tex
edited text up to 2.4

@@
 too close to each other.}
+\plr{how's that?  I'm not sure what to do about the grey; the real problem is that the CIs are too small. If I ran them for less time they would be\dots}
 
@@
   \caption{
-  Mean occupation frequencies from simulation; details are as in figure~\ref{fig:sim_snapshots} except as otherwise noted.
-  Each panel shows mean occupation frequencies of individual demes (grey dots)
+  \textbf{(A)} 
+  Mean occupation frequencies of individual demes (grey dots)
   and averaged over demes at similar distances (black circles), on
-  \textbf{(A)} a one-dimensional grid of 201 demes, and
-  \textbf{(B)} a two-dimensional grid of $101\times 101$ demes.
+  a one-dimensional grid of 201 demes, and
+  a two-dimensional grid of $101\times 101$ demes.
   Superimposed in black is the decay predicted by \eqref{eqn:eqfreq},
   with $C$ chosen arbitrarily to give a reasonable fit.
+  \textbf{(B)} 
+  Fluctuations: colored lines are snapshots of allele frequencies at evenly spaced times
+  from a one-dimensional grid of 151 demes;
+  the solid lines shows the cumulative mean occupation frequency across all generations.
+  For both plots, the simulation was run for 1000 generations;
+  the focal allele is beneficial in the central 20 demes, with $s_p=0.1$,
+  and deleterious elsewhere, with $s_m=-0.02$; 
+  demes had 1000 individuals in each;
+  dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
   }   \label{fig:sim_occupation_freqs}
@@
   \caption{
-  Results from individual-based simulation, on a one-dimensional grid:
-  the focal allele is beneficial in the central 20 demes, with $s_p=0.1$,
-  and deleterious elsewhere, with $s_m=-0.02$.
-  The dynamics were simulated for 10000 generations; demes had 1000 individuals in each;
-  dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
-  \textbf{(A)} Allele frequencies across space: colored lines are snapshots at an evenly spaced set of times,
-  and the solid lines shows the cumulative mean occupation frequency across all generations.
-  \textbf{(B)} A segment of the temporal dynamics: space is again shown across the horizontal axis;
+  The temporal dynamics of the same simulation as Figure~\ref{fig:sim_occupation_freqs}B.
+  A segment of the temporal dynamics: space is again shown across the horizontal axis;
   with time on the vertical axis; demes are colored darker red the greater then allele frequency at the corresponding time.
@@
 (This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
-This decomposition is illustrated in~\ref{fig:sim_snapshots}, and later in figure~\ref{fig:branching_decomp}.
+This decomposition is illustrated in figure~\ref{fig:branching_decomp}.
 In terms of these families, the equilibrium frequency above is then
@@
 the expected family size after time $t$ is $(1-s_m)^t$.
-Therefore, if we define $p_e(t)$ to be the chance of a \gc{family}
-going extinct by time $t$, 
+Therefore, if we define $k_e(t)$ to be the chance that the family
+has gone extinct by time $t$, 
+\plr{switched from $p_e(t)$ to $k_e(t)$ to not conflict with prob of estab.  better suggestion?}
 and let $K_t$ be the (random) family size conditioned on nonextinction,
-then by Bayes' rule, $(1-s_m)^t = (1-p_e(t))\E[K_t]$.
+then by Bayes' rule, $(1-s_m)^t = (1-k_e(t))\E[K_t]$.
 It turns out that if the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jagers1975branching},
 then $K_t$ has a limiting distribution: $K_t \dconv K$ 
-and that $(1-s_m)^{t}/(1-p_e(t)) \to \E[K]$ as $t \to \infty$.
+and that $(1-s_m)^{t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
 In other words, the chance of survival to time $t$ is a constant multiplied by $(1-s_m)^t$,
@@
 
-\gc{We exploit this idea to consider the survival of a central ``trunk'',
-  and then conditional on that consider the family size of side
-  ``branches'' off of this trunk. This process is depicted in Figure~\ref{fig:branching_decomp}.}
+This suggests that families can be well-described by motion of a central ``trunk'',
+dies at a random time whose distribution is given by $k_e(t)$,
+and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
+as depicted in Figure~\ref{fig:branching_decomp}.
 We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} \citep[see also][]{chauvin1991growing}.
 When we condition on survival until $t$, we condition on existence of
-at least one lineage surviving from time $0$ to time $t$ (the long red
+at least one ``trunk'' lineage surviving from time $0$ to time $t$ (the long red
 lineage in Figure \ref{fig:branching_decomp}).
-\gc{Given this lineage, nothing distinguishes any of the other reproduction events --
-each individual along this lineage} must give birth to at least one offspring,
+Given this lineage, nothing distinguishes any of the other reproduction events --
+each individual along the trunk lineage must give birth to at least one offspring,
 and all other reproduction events are unconditioned.
@@
 Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
-% (so $\P\{Z_t = 0\} = p_e(t)$),
+% (so $\P\{Z_t = 0\} = k_e(t)$),
 then this gives the following decomposition:
 \begin{align}
-  \left( Z_t \; \vert \; Z_t>0 \right) \deq \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
-\end{align}
-where each $Z^{(s,k)}$ is an independent copy of $Z$ \gc{that arose
-  from the main branch in generation $s$},
+  \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
+\end{align}
+where each $Z^{(s,k)}$ is an independent copy of $Z$ 
+that arose from the main branch in generation $s$,
 and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.
-\gc{Is there no 1+ in the equation above for the main branch? }
+\gc{Is there no 1+ in the equation above for the main branch? } \plr{oops; yep.}
 
@@
 and those that did not survive until $t$ (in grey).
-Figure~\ref{fig:sim_snapshots} has a depiction of such a family, from
-a simulation. \gc{In reality the spatial paths taken by lineages are
-  much more wiggly but they are shown as straight lines here for simplicity.}
+Figure~\ref{fig:sim_snapshots} has a depiction of such a family, from a simulation. 
+Note that the spatial motions of lineages are independent random walks
+(and hence should be much wigglier than in this cartoon).
 } \label{fig:branching_decomp}
@@
 By examining the constituent terms,
-it will turn out that eqn. \eqref{eqn:gestalt_migrate} is a constant multiple of \eqref{eqn:gestalt_q}.
-
-First consider eqn. \eqref{eqn:gestalt_q}.
-Suppose that $Z$ is one such spatial branching process as above, started with a single individual at $x$,
-defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the \gc{area} $A$.
+it will turn out that equation \eqref{eqn:gestalt_migrate} is a constant multiple of \eqref{eqn:gestalt_q}.
+
+First consider equation \eqref{eqn:gestalt_q}.
+Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
+defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
 The \emph{mean occupation density} of $Z$ can be thought of 
@@
 \gc{do we need to talk about u a little more it }
+\plr{not sure what to say}
 Since reproduction and spatial motion are independent,
@@
 \begin{align}
-    u(x,y) = \int_0^t e^{-\alpha t} p_t(x,y) dt .
-\end{align}
-\gc{$\alpha$ introduced here? Switch from $(1-s_m)^t$ }
+    u(x,y) = \int_0^t (1-s_m)^t p_t(x,y) dt .
+\end{align}
 The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
@@
 \begin{align} \label{eqn:occupation_integral}
-    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  % \\
-         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-\alpha t} p_t(x,y) dt dx .
+    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
+         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t (1-s_m)^t p_t(x,y) dt dx .
 \end{align}
 \gc{define $\gamma(x)$ presumably the outflux of branches leaving ball}
-
-Now we turn to the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
+\plr{it was defined at the start of this paragraph.}
+
+This form we can now compare to 
+the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
 Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
 which we assume is circular, has relatively small area $A$ and is located near position $y$.
-% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-p_e(t))$,
+% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
 % multiplied by the probability that the family reaches the patch, given survival.
-% If $t$ is large then as in the previous section, $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$.
+% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-\alpha t}/\E[K]$.
 It would be possible to analyze directly the probability that a given migrant family establishes in $A$,
-as in \citet{barton1987establishment,mckean1975KPP};
-but we make a simpler approximation, approximating this by the chance that the ``trunk'' hits the new patch,
-multiplied by the chance that at least member of the family escapes demographic stochasticity to establish.
+as in \citet{barton1987establishment};
+but we make a simpler approximation, approximating this probability by the chance that the ``trunk'' hits the new patch,
+multiplied by the chance that at least one member of the family escapes demographic stochasticity,
+successfulling establishing in the new patch.
 Write $h(x,y)$ for the probability that the Brownian trunk
@@
 given that it successfully arrives.
-As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is an integral of $h(x,y)$:
+As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral of $h(x,y)$:
 \begin{align}\label{eqn:migrate_integral}
@@
 it is intuitively reasonable that $u(x,y)$ and $h(x,y)$ are simply related.
-Using this fact,
-and that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
+Following this intuition,
+and that $(1-k_e(t)) \approx (1-s_m)^t/\E[K]$,
 it turns out (see Appendix~\ref{apx:hitting_occupation}) that
 \begin{align} 
-  h(x,y) = \frac{ s(A) }{ \E[K] } u(x,y),
-\end{align}
-where $s(A)$ is the total expected occupation time of the patch given the family arrives,
+  h(x,y) = \frac{ g(A) }{ \E[K] } u(x,y),
+\end{align}
+where $g(A)$ is the total expected occupation time of the patch given the family arrives,
 and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
 Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a constant factor
-(at least asymptotically for large distances),
+(at least asymptotically as $x$ and $y$ are far apart),
 we now know that the outflux of successful migrant families is equal to the mean occupation frequency
-multiplied by $s(A) f(A) / \E[K]$.
-The probability $f(A)$ that the family does not die out once it arrives at the new patch
-is if each member arriving has chance $p_e$ of establishing, independently, $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$.
+multiplied by $g(A) f(A) / \E[K]$.
+The probability $f(A)$ that the family does not die out once it arrives at the new patch is
+$\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
+if each member arriving has chance $p_e$ of establishing, independently.
 % If we let $w$ be the width of the patch, then the remaining term is
@@
 % \end{align}
-Combining these terms with the expression for $s(A)$ from Appendix~\ref{apx:hitting_occupation},
+Combining these terms with the expression for $g(A)$ from Appendix~\ref{apx:hitting_occupation},
 \begin{align}
   \label{eqn:migrate}
-  \migrate(y) \approx q(y) \times \rho \frac{ s_p  }{ \alpha \xi^2 } 
+  \migrate(y) \approx q(y) \times \frac{ \rho s_p  }{ \xi^2 \log(1/(1-s_m)) } 
   % \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
@@
 Therefore, mutations selectively exclude each other from patches, over short time scales, 
-and will only slowly mix together over longer time scales by drift \gc{and
-migration. We made use of this selective exclusion of alleles in our
-previous paper \citep{RalphCoop}. In reality different mutations may
-not be truly selective equivalent in which case the mixing will occur
-over a time-scale dictated by the difference in selection
-coefficients. }
-
-We will assume that once a $B$ allele is introduced (by migration or mutation), 
-\gc{condition on it escaping loss,} it becomes established in the poor habitat patch rapidly. 
+and will only slowly mix together over longer time scales by drift and migration,
+an assumption we also made in \citep{ralph2010parallel}. 
+In reality, different mutations will likely not be truly selective equivalent,
+in which case this mixing occurs over a time-scale dictated by the difference in selection coefficients.
+
+We will assume that once a $B$ allele is introduced (by migration or mutation) and successfully begins to establish,
+it becomes established in the poor habitat patch rapidly. 
 Under this approximation, and the ``selective exclusion'' just discussed,
@@
 As such, the question of how the the second patch adapts
-simply comes down to whether a new mutation or a migrant allele is the first to become established in the second patch. Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
+simply comes down to whether a new mutation or a migrant allele is the first to become established in the second patch. 
+Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
 we see that if the patches each have sufficiently large area,
@@
 \gc{the two paragraphs either side of this will need some reworking.}
+\plr{want to do that? or be a bit more specific? what works/doesn't?}
 
@@
 surrounded by a cloud of close relatives of random size $K$
-whose chance of surviving until $t$ generations is $p_e(t) \simeq e^{-\alpha t} / \E[K]$.
+whose chance of surviving until $t$ generations is $k_e(t) \simeq e^{-\alpha t} / \E[K]$.
 Consider a single such family, and let $\tau$ be the (random) time it first hits the new patch,
@@
 If $A$ is circular, the latter expectation will not depend on $x$;
-call this $h(A)$.
+call this $g(A)$.
 
@@
 \begin{align}
-    F(x,y) &= \int_0^\infty (1-p_e(t)) \P\{ \tau_A \in dt \} \\
-           &= \int_0^\infty e^{\alpha t} (1-p_e(t)) e^{- \alpha t } \P\{ \tau_A < t \} dt \\
-    &=\int_0^\infty e^{\alpha t} (1-p_e(t)) \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    F(x,y) &= \int_0^\infty (1-k_e(t)) \P\{ \tau_A \in dt \} \\
+           &= \int_0^\infty e^{\alpha t} (1-k_e(t)) e^{- \alpha t } \P\{ \tau_A < t \} dt \\
+    &=\int_0^\infty e^{\alpha t} (1-k_e(t)) \P\{ \tau_A < t < \tau_\dagger \} dt \\
     &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
@@
     \approx \P^x[ \tau_A < \tau_\dagger ] 
-    = u(x,y) / h(A) .
+    = u(x,y) / g(A) .
 \end{align}
@@
     q(y)  &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
-          &= h(A) \int_{\partial B(r_0)} \gamma(x) F(x,y) dx \\
-          &= \frac{ h(A) }{ \E[K] } \migrate(y) .
-\end{align}
-The function $h(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
+          &= g(A) \int_{\partial B(r_0)} \gamma(x) F(x,y) dx \\
+          &= \frac{ g(A) }{ \E[K] } \migrate(y) .
+\end{align}
+The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
 is expected to spend inside the disk before $\tau_\dagger$.
@@
 which we assume has relatively small area $A$ and is located near position $y$.
-The probability that this family reaches $A$ by time $t$ is $(1-p_e(t))$
+The probability that this family reaches $A$ by time $t$ is $(1-k_e(t))$
 multiplied by the probability that the family reaches the patch, given survival.
-If $t$ is large then the previous section argues that $(1-p_e(t)) \approx e^{-\alpha t}/\E[K]$,
+If $t$ is large then the previous section argues that $(1-k_e(t)) \approx e^{-\alpha t}/\E[K]$,
 and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
@@
 \begin{align}
-    \int_0^\infty (1-p_e(t)) \P^x\{ \tau_w \le t \} dt  
+    \int_0^\infty (1-k_e(t)) \P^x\{ \tau_w \le t \} dt  
     &\approx \int_0^\infty e^{-\alpha t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\

### 2014-05-18 patchy-selection-paper.tex
edited 2.4; slight figure modifications

@@
 $1-s_m$  when rare in the intervening areas, with $s_p$ and $s_m$ both positive.
+(Here ``fitness'' we take to be the intrinsic growth rate of the type when rare.)
 To simplify matters, we assume that the disadvantage $s_m$ 
@@
 
-\gc{I'd suggest making Figure 2 and Fig 1A all part of same figure
-  they are showing the same basic thing good fit of equation of
-  cline. The s's in 1A should be labeled sm and sb. I'd suggest that
-  the second part of 1 become its own figure or be paired with fig.3
-  as they are both about rare excursions. Could you make the CI or
-  sample path in Figure 1B be different colors, the grey and black are
-too close to each other.}
+\gc{Could you make the CI or sample path in Figure 1B be different colors, the grey and black are too close to each other.}
 \plr{how's that?  I'm not sure what to do about the grey; the real problem is that the CIs are too small. If I ran them for less time they would be\dots}
@@
   with $C$ chosen arbitrarily to give a reasonable fit.
+  (The uptick at the right end is due to boundary effects near the edge of the range.)
   \textbf{(B)} 
@@
 by first forming a tree in which each individual has a random number of offspring
-(with mean $1-s_m$ and variance $\xi^2$),
 then recursively assigning spatial locations to each individual in the tree,
 and finally pruning any branches that happen to wander into a patch.
+By our definition of fitness, the mean number of offspring per individual is $\exp(-s_m)\approx 1-s_m$,
+and the variance is $\xi^2$.
+\plr{I changed this since equation \eqref{eqn:eqfreq} is in the small-$s_m$ limit, 
+and therefore doesn't differentiate between $s_m$ and $\log(1/(1-s_m))$;
+doing it this way means we don't have to define another parameter.}
 This pruning removes both branches leading back into the already established patch (which are irrelevant)
@@
 
-Since each individual has on average $(1-s_m)$ offspring,
-the expected family size after time $t$ is $(1-s_m)^t$.
+Since each individual has on average $e^{-s_m}$ offspring,
+the expected family size after time $t$ is $e^{t s_m}$.
 Therefore, if we define $k_e(t)$ to be the chance that the family
@@
 and let $K_t$ be the (random) family size conditioned on nonextinction,
-then by Bayes' rule, $(1-s_m)^t = (1-k_e(t))\E[K_t]$.
+then by Bayes' rule, $e^{-s_m t} = (1-k_e(t))\E[K_t]$.
 It turns out that if the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jagers1975branching},
 then $K_t$ has a limiting distribution: $K_t \dconv K$ 
-and that $(1-s_m)^{t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
-In other words, the chance of survival to time $t$ is a constant multiplied by $(1-s_m)^t$,
+and that $e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
+In other words, the chance of survival to time $t$ is a constant multiplied by $e^{-s_m t}$,
 and conditional on survival, the family size is given by a fixed distribution.
@@
 \begin{align}
-    u(x,y) = \int_0^t (1-s_m)^t p_t(x,y) dt .
+    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
 \end{align}
@@
     q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
-         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t (1-s_m)^t p_t(x,y) dt dx .
+         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
 \end{align}
@@
 Following this intuition,
-and that $(1-k_e(t)) \approx (1-s_m)^t/\E[K]$,
+and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
 it turns out (see Appendix~\ref{apx:hitting_occupation}) that
@@
 % \end{align}
-Combining these terms with the expression for $g(A)$ from Appendix~\ref{apx:hitting_occupation},
+Combining the terms with expressions for $g(A)$ in Appendix~\ref{apx:hitting_occupation},
+we obtain that
 \begin{align}
   \label{eqn:migrate}
-  \migrate(y) \approx q(y) \times \frac{ \rho s_p  }{ \xi^2 \log(1/(1-s_m)) } 
+  \migrate(y) &\approx q(y) \times \frac{ \rho s_p  }{ \xi^2 s_m }  \\
   % \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
-\end{align}
-For patches that are small relative to $\sigma/\sqrt{2s_m}$, there is an additional correction term (see Appendix~\ref{apx:hitting_occupation}),
-but small patches violate other assumptions above, so we omit it.
+  &\approx C \frac{ \rho s_p  }{ \xi^2 s_m } \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) .
+\end{align}
+This does not depend on $A$, because although
+$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
+the dependence is very weak:
+for instance, the width of a circular patch only very weakly affects the chance 
+of establishment of a new migrant that appears on its edge,
+as long as the patch is large enough.
+For more discussion, and correction terms, see Appendix~\ref{apx:hitting_occupation}.
 
@@
 simply comes down to whether a new mutation or a migrant allele is the first to become established in the second patch. 
-Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
-we see that if the patches each have sufficiently large area,
-the rate of migration and mutation will be on the same order if 
+To work out how adaptation is likely to proceed, 
+we can compare expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above
+for the effective rates of adaptation by new mutation and by migration.
+
+The first thing to notice is that effective migration and mutation rates will be on the same order if 
 $A \mu \approx (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
-where $R$ is the distance between the patches,
-and recall that $A$ is the area of the not-yet-adapted patch.
+where recall that $R$ is the distance between the patches,
+and $A$ is the area of the not-yet-adapted patch.
 In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
@@
 which depends fairly strongly on the selective disadvantage $s_m$ between patches.
-Equivalently, migration will occur much quicker than mutation 
-if $R/\sigma > (\log(1/s_m) - \log(A\mu))/\sqrt{2s_m}$.
-\gc{the two paragraphs either side of this will need some reworking.}
-\plr{want to do that? or be a bit more specific? what works/doesn't?}
-
-If we take the mutational influx in a patch to be one per 1,000
-generations ($A \mu = .001$), \gc{don't think this is mutational
-  influx as it doesn't have $\rho in it$}
-and the selection against transiting alleles $s_m = .01$,
-adaptation procedes independently for distances $R/\sigma > \log(s_m)/\sqrt{s_m}$,
-so if the selective pressure is fairly strong (say, $s_m=.05$),
-then the distance between patches must be about 30 times the dispersal distance (i.e.\ $R/\sigma > 30$),
-while if $s_m$ is much smaller, say $s_m = .001$, 
+Equivalently, the rates are roughly equal if
+$R/\sigma = \log(1/(s_m A \mu))/\sqrt{2s_m}$,
+which gives the critical gap size past which adaptation will be mostly parallel
+in terms of selection coefficients, patch size, and mutation rate.
+
+% ss <- 10^(-(1:6))
+% log(1/ss)
+% 1/sqrt(2*ss)
+% sm <- .05; (-log(c(1e-3,1))-log(sm))/sqrt(2*sm)
+% sm <- .001; (-log(c(1e-3,1))-log(sm))/sqrt(2*sm)
+% (7+log(1/ss))/sqrt(2*(ss))
+% (7+log(1/(.5*ss)))/sqrt(2*(.5*ss))
+If we take the mutational influx in a patch to be one per 1,000 generations ($\rho A \mu = .001$)
+and the population density to be 1 ($\rho =1$, so $\log(A\mu)\approx 7$)
+then adaptation is mostly parallel for patches separated by gaps larger than $R/\sigma > (7+\log(s_m))/\sqrt{s_m}$.
+If the selective pressure is fairly strong (say, $s_m=.05$),
+then for migration to be slower than mutation,
+the distance between patches must be at least 30 times the dispersal distance (i.e.\ $R/\sigma > 30$).
+If $s_m$ is much smaller, say $s_m = .001$, 
 then the patches must be separated by several hundred times the dispersal distance (i.e.\ $R/\sigma > 300$).
-The situation does not change much if the influx is quite a bit higher -- even if $A \mu = 1$,
-the required separation between patches is only reduced by $7/\sqrt{s_m}$,
-to $R/\sigma > 10$ with $s_m=.05$ and $R/\sigma>150$ for $s_m=.001$.
-If the separation between patches is much larger than this transitional value -- certainly, by a factor of ten -- 
-then patches are effectively isolated -- most adaptation will occur independent between patches.
-On the other hand, if the separation is much smaller, then a single
-mutation is likely to fix in both patches.
+The situation does not change much if the influx is quite a bit higher -- even if $A \mu = 1$ (so $\log(A\mu) \approx 2.3$),
+the required separation between patches is only reduced to $R/\sigma > 10$ with $s_m=.05$ and $R/\sigma>150$ for $s_m=.001$.
+On the other hand, if the separation is much smaller than this transitional value, 
+then adaptations are likely to be shared between patches.
 
@@
 we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
-as a function of various parameters. \gc{Parallel mutation will occur
-where $\Tmut \gg \Tmig $ (reder areas), while adpatation by migration will occur
-where $\Tmig \gg \Tmut $ (bluer areas).}
+as a function of various parameters. 
+This gives us an idea of not only the parameter space where parallel adaptation is likely ($\Tmig \ll \Tmut$),
+but also where adaptation by any means takes a very long time ($\Tmig$ and $\Tmut$ both very large),
+and important check in applications.
 
@@
   and mutation rate scaled by patch area ($\mu A$).
-  The line at which $\Tmig=\Tmut$ is shown in black.
+  The black, dotted line shows values where $\Tmig=\Tmut$.
   Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=50$, and $\xi^2=1$.
@@
 
-If we are willing to take these approximations at face value, 
+Taking these approximations at face value, 
 and assuming both rates are small, 
@@
 \mutrate)$.
-Following this, the time till the second patch is subsequently colonized 
+Following this, the time until the second patch is subsequently colonized 
 (via either migration or new mutation) 
 will be exponentially distributed with mean $1/(\mutrate+\migrate)$.
-Both these times scale linearly with $\xi^2/(\rho s_{B})$, 
+Both these rates scale linearly with population density ($\rho$) 
+and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$),
 so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
@@
 \begin{equation}
-  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A \mu \alpha +
-    \left(r \sqrt{2 s_m} r / \sigma \right)^{(d-1)/2} \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
+  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A \mu \alpha + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
 \end{equation}

### 2014-05-19 patchy-selection-paper.tex
edited up until applications

@@
 \subsection{Multiple patches}
-We can extend this reasoning to multiple patches of unadapted habitat 
+Above we worked with two patches separated by distance $R$;
+but these calculations apply as well to multiple patches
 in which the $B$ allele can facilitate local adaptation. 
@@
 with small probability of success,
-the process of occupation is well approximated by a discrete-state Markov chain,
-whose state is the labeling of patches according to the colonizing allele
+the process of occupation is well approximated by a Markov chain,
+whose state records the labeling of patches according to the colonizing allele
 (or none, for those that haven't yet adapted).
 
-If not yet adapted, a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$.
-Without loss of generality, assume that patches $1, \ldots, k$ are already adapted,
-and these patches are at distances $r_1, \ldots, r_k$ away from this unadapted patch, 
-then the total rate of adaptation through migration is approximately
+If not yet adapted, a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$ (equation \eqref{eqn:mutrate}).
+Suppose that patches $1, \ldots, k$ are already adapted,
+and that these patches are at distances $r_1, \ldots, r_k$ away from this unadapted patch.
+The total rate of adaptation through migration, from equation \eqref{eqn:migrate}, is
 \begin{equation}
-  \frac{ s_p \rho }{\alpha \xi^2} \sum_{i=1}^{k} \left(r_i \sqrt{2 s_m} /\sigma \right) \exp\left(- r_i \sqrt{2 s_m} /\sigma\right),
+  \migrate(r_i) = C \frac{ \rho s_p }{\xi^2 s_m} \sum_{i=1}^{k} \left(r_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m} /\sigma\right),
 \end{equation}
-and conditional on coming from migration the probability the adapted allele 
+and if the patch adapts through migration, the probability the adapted allele 
 comes from patch $i$ is 
 \begin{align}
-  \frac{\left(r_i \sqrt{2 s_m} /\sigma \right) \exp\left(- r_i \sqrt{2 s_m}
-    /\sigma\right)} {\sum_{j=1}^{k}  \left(r_i \sqrt{2 s_m} /\sigma \right) \exp\left(- r_i \sqrt{2 s_m}
+  \frac{\left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
+  /\sigma\right)} {\sum_{j=1}^{k}  \left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
     /\sigma\right) } .
@@
 so that at least initially, all adapted individuals within a patch
-share a fairly long stretch of haplotype. \gc{Pairs of individuals
-  within a patch will share a haplotype of mean genetic length of
-  $\approx \approx \log(\rho A s_p)/s_p$ around the selected site,
+share a fairly long stretch of haplotype. 
+Pairs of individuals within a patch will share a haplotype of mean genetic length of
+$\approx \log(\rho A s_p)/s_p$ around the selected site, 
 as long as the patch is reasonably well mixed by dispersal
-\citep[otherwise see][]{BartonEtheridge13spatialHH}. }
+\citep[otherwise see][]{barton2013genetic}.
 This association gets slowly whittled down by recombination during interbreeding at the edge of the patch;
@@
 surrounded by a cloud of close relatives of random size $K$
-whose chance of surviving until $t$ generations is $k_e(t) \simeq e^{-\alpha t} / \E[K]$.
-Consider a single such family, and let $\tau$ be the (random) time it first hits the new patch,
-with $\tau = \infty$ if it never does. The length of the hitchhiking segment depends on the time
-spent by this trunk lineage ``in transit'' between the patches, i.e. $\tau$.
-
-By pulling out the constant $1/\E[K]$, we can use standard results on hitting times of $d$-dimensional Brownian motion
-that is killed at rate $\alpha$ \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook} 
+whose chance of surviving until $t$ generations is $k_e(t)$.
+Consider a single such family, and let $\tau$ be the (random) time at which it first hits the new patch,
+with $\tau = \infty$ if it never does. 
+We assume that the length of the hitchhiking segment depends only on 
+the time spent by the trunk lineage of the first successful migrant ``in transit'' between the patches,
+i.e.\ $\tau$ conditioned on $\tau < \infty$.
+In so doing, we neglect 
+recombination between B alleles (since they are at low density in transit, we assume this is rare),
+and the possibility that more than one migrant family is in transit at once
+(so that faster migrants would be more likely to have arrived first);
+but these factors should be minor if the rest of our framework fits.
+
+Our first obstacle is that we do not know $k_e(t)$, 
+the distribution of the time until a family dies out, exactly.
+However, we do know that $k_e(t) \simeq e^{-s_m t} / \E[K]$,
+i.e.\ that it has exponential tails.
+Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
+replacing this random time by an exponential with rate $s_m$.
+Once we do this, 
+we can use standard results on hitting times of $d$-dimensional Brownian motion
+that is killed at rate $s_m$ \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook} 
 In particular, if the patch is circular with radius $w$ and a distance
-$R$ from the already adapted patch, then we have the useful result
-about the Laplace transform of $\tau$ our transit time
+$R$ from the already adapted patch, then 
+we know that the Laplace transform of the transit time $\tau$ is
 \begin{align}
@@
     \begin{cases}
-      e^{- R \sqrt{2(\alpha+\ell)}/\sigma} \qquad & d=1 \\
-      \frac{ K_0( (R+w)\sqrt{2(\alpha+\ell)}/\sigma) }{ K_0( w\sqrt{2(\alpha+\ell)}/\sigma) } \qquad & d=2 
-    \end{cases} \label{borodinresult}
-    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(\alpha+\ell)}/\sigma}  (1+O(1/w)) .
-\end{align}
-\gc{define $K_0$, bessel? We've used K for family size}
-We can find that the probability that a particular
-trunk lineageis killed before it reaches the new patch is
-$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau)
-\right]$. We are interested in lineages that
-manage to establish on the patch i.e. $\tau< \infty$, and so in one dimension
+      e^{- R \sqrt{2(s_m+\ell)}/\sigma} \qquad & d=1 \\
+      \frac{ K_0( (R+w)\sqrt{2(s_m+\ell)}/\sigma) }{ K_0( w\sqrt{2(s_m+\ell)}/\sigma) } \qquad & d=2  ,
+    \end{cases} \label{eqn:borodinresult}
+    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(s_m+\ell)}/\sigma}  (1+O(1/w)) .
+\end{align}
+where $K_0$ is a modified Bessel function of the second kind.
+We are interested in lineages that manage to reach the patch before being killed,
+i.e.\ having $\tau < \infty$,
+which occurs with probability
+$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau) \right]$.
+
+To keep the expressions simple, in the remainder of this section we only compute quantities for one dimension.
+By Bayes' rule, the Laplace transform of $\tau$ for a successful lineage is
 \begin{equation} \label{eqn:haplen_cdf}
-\E[e^{-\ell \tau}|\tau<\infty]  = \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
+\E[e^{-\ell \tau}|\tau<\infty]  = \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
 \end{equation}
-We can differentiate this to find that our expected transit time is
+This can be differentiated to find that the expected transit time is
 \begin{equation} 
-\E[\tau] = (R\sqrt{2\alpha}/\sigma)\times 1/(2\alpha)
+  \E[\tau] = (R\sqrt{2s_m}/\sigma)\times 1/(2s_m) \label{eqn:mean_tau}
 \end{equation}
-and $\var[\tau] = (R\sqrt{2\alpha}/\sigma) \times 1/(2\alpha)^2$ in
-one dimension.
-
-% \simeq (R/w+1)^{-(d-1)/2} \exp(- R \sqrt{2\alpha}/\sigma)$.
+and that $\var[\tau] = (R\sqrt{2s_m}/\sigma) \times 1/(2s_m)^2$.
+
+% \simeq (R/w+1)^{-(d-1)/2} \exp(- R \sqrt{2s_m}/\sigma)$.
 
@@
 %   \P\{\tau\in dt | \tau<\infty \} &= \frac{\P\{\tau \in dt\}}{\P\{\tau<\infty\}} \\
-%   &= \frac{R}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{R^2}{2t\sigma^2} -\alpha t + \sqrt{\alpha} \frac{R}{\sigma} \right) .
+%   &= \frac{R}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{R^2}{2t\sigma^2} -s_m t + \sqrt{s_m} \frac{R}{\sigma} \right) .
 % \end{align}
@@
 if recombination is Poisson, the length of the haplotype (in Morgans)
-initially shared between populations on each side of the site will be exponentially distributed
+initially shared between populations on each side of the selected locus is exponentially distributed
 with mean $\tau$, so that if $L$ is the length of hitchhiking segment on, say, the right of
@@
 \begin{equation} \label{eqn:haplen_cdf2}
-\P\{L>\ell\} = \E[e^{-\ell \tau}|\tau<\infty] 
-%  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
+\P\{L>\ell\} = \E[e^{-\ell \tau}|\tau<\infty] ,
+%  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
 \end{equation}
-which is our Laplace transform from above that in one dimension
-gives us eqn. \eqref{eqn:haplen_cdf}. This implies that in one dimension, if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
-then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$,
-from which we can obtain that the expected length of shared hitchiking
-haplotype is
+and so (one minus) equation~\eqref{eqn:haplen_cdf} in fact gives the cumulative distribution function of $L$.
+The form of this distribution function implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
+then $L$ has the same distribution as $(Y + \sqrt{s_m})^2 - s_m$.
+Furthermore, the expected length of shared hitchiking haplotype is
 \begin{equation} \label{eqn:haplotype_length}
-\E[L] = \sigma \sqrt{2\alpha}/R + \sigma^2/R^2
+\E[L] = \sigma \sqrt{2s_m}/R + \sigma^2/R^2
 \end{equation}
-and thar $\var[L] = 2\alpha\sigma^2/R^2 + 4 \sigma^3 \sqrt{2\alpha}/R^3 + 5 \sigma^4 / R^4$.
+and $\var[L] = 2s_m\sigma^2/R^2 + 4 \sigma^3 \sqrt{2s_m}/R^3 + 5 \sigma^4 / R^4$.
 For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
-but how other properties might change is unclear. \gc{add some
+but how other properties might change is unclear.
+
+As a general rule of thumb, equation~\eqref{eqn:mean_tau} means that families who successfully establish
+move at speed $\sqrt{s_m} \sigma$ towards the new patch 
+-- if $s_m$, the strength of the selection against them, is smaller, the need to move quickly is less imperative.
+Then, equation~\eqref{eqn:haplotype_length} means that the haplotype length
+is roughly the length one would expect given the mean transit time.
+Note that we have become used to seeing $\sigma$ divided by $\sqrt{s_m}$, rather than multiplied,
+the $\sigma/\sqrt{s_m}$ is a length scale, and to convert it to a time this is divided by $1/s_m$.
+
+\gc{add some
   sentences of intuition of how parameters play into haplotype length
-  here. is it really $\sigma \sqrt{\alpha}$ in \eqref{eqn:haplotype_length} and not $\sqrt{\alpha}/sigma$???}
+  here. is it really $\sigma \sqrt{s_m}$ in \eqref{eqn:haplotype_length} and not $\sqrt{s_m}/sigma$???}
 
 \gc{Perhaps we should separate the above bit into things that are
-  useful about $\tau$ and put those up by  \eqref{borodinresult} and
+  useful about $\tau$ and put those up by  \eqref{eqn:borodinresult} and
   things that are useful about $L$ here? We should also turn some of
@@
 
-% Since each generation spent in transit provides an opportunity for recombination,
-% if recombination is Poisson, the length of the hitchhiking segment (in Morgans) on each side of the site will be exponentially distributed
-% with mean $\tau$, so that if $L$ is the length of hitchhiking segment on, say, the right of
-% the selected locus, then in one dimension
-% \begin{align} \label{eqn:haplen_cdf}
-% \P\{L>\ell\} &= \E[e^{-\ell \tau}|\tau<\infty] \\
-%   &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+\alpha)} - \sqrt{2\alpha}\right)}\right\} .
-% \end{align}
-% found by using eqn. \eqref{borodinresult}. This is also the Laplace transform of $\tau$;
-% we can differentiate it to find that $\E[\tau] = (R\sqrt{2\alpha}/\sigma)\times 1/(2\alpha)$
-% and $\var[\tau] = (R\sqrt{2\alpha}/\sigma) \times 1/(2\alpha)^2$.
-% Furthermore, this implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
-% then $L$ has the same distribution as $(Y + \sqrt{\alpha})^2 - \alpha$,
-% from which we can obtain that
-% $\E[L] = \sigma \sqrt{2\alpha}/R + \sigma^2/R^2$ and $\var[L] = 2\alpha\sigma^2/R^2 + 4 \sigma^3 \sqrt{2\alpha}/R^3 + 5 \sigma^4 / R^4$.
-% For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
-% but how other properties might change is unclear.
-
 %%%
@@
 by taking the large-$R$ limit above we can obtain a more intuitive description of the process
-via a central limit for $\tau$.
+via a central limit theorem for $\tau$.
 Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
-so that $\tau = R/\sigma(2\alpha)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2\alpha)^{-3/4}$.
+so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
 Plugging this into equation~\eqref{eqn:haplen_cdf},
@@
 a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
-with mean $\sqrt{2\alpha} \sigma/R$.
+with mean $\sqrt{2s_m} \sigma/R$.
 
 % \begin{align}
-%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{R}{\sigma\sqrt{2\alpha}}\right)\frac{(2\alpha)^{3/4}\sqrt{\sigma}}{\sqrt{R}} \right\} \right] \\
-%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2\alpha)^{3/4} \sqrt{\sigma}}{\sqrt{R}}\right) \tau \right\} \right] \exp\left( (2\alpha)^{1/4} \sqrt{\frac{R}{\sigma}} \right) \\
+%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{R}{\sigma\sqrt{2s_m}}\right)\frac{(2s_m)^{3/4}\sqrt{\sigma}}{\sqrt{R}} \right\} \right] \\
+%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2s_m)^{3/4} \sqrt{\sigma}}{\sqrt{R}}\right) \tau \right\} \right] \exp\left( (2s_m)^{1/4} \sqrt{\frac{R}{\sigma}} \right) \\
 % \end{align}
 
-XXX Check what happens in two dimensions. XXX
+\plr{let's leave two dimensions for the revisions. or, as an exercise? =)}
 
@@
 small-scale fluctuations
+
+when landscape is not made of discrete (circular) patches
 

### 2014-05-20 lineage-movement.tex
.

@@
 \documentclass{article}
+
 \usepackage{fullpage}
@@
 
-\subsection{Discrete model}
+\section{Discrete model}
 
 Consider a discrete model in which the total number of individuals at location $x$ is $N(x)$,
-each individual is of type either A or B at a selected locus,
-and the number of individuals of type A and location $x$ is $n(x)$.
-Suppose that type A individuals at location $x$ reproduce at rate $1+s_A(x)$, 
-and likewise type B at rate $1+s_B(x)$.
+each individual is of type either B or b at a selected locus,
+and the number of individuals of type B and location $x$ is $n(x)$.
+Suppose that type B individuals at location $x$ reproduce at rate $1+s_B(x)$, 
+and likewise type b at rate $1+s_b(x)$.
 At reproduction, individuals recombine with others in the same location,
@@
 \begin{enumerate}
-    \item[$x\xrightarrow{AA}y$] One type A individual at location $x$ reproduces, 
-        either does not recombine or recombines with another type A,
+    \item[$x\xrightarrow{AA}y$] One type B individual at location $x$ reproduces, 
+        either does not recombine or recombines with another type B,
         and sends the offspring to $y$.
@@
         and sends to $y$ an offspring
-        who inherits at the selected locus from the type B parent 
-        and the at the neutral locus from the type A parent.
+        who inherits at the selected locus from the type b parent 
+        and the at the neutral locus from the type B parent.
     \item[$x\xrightarrow{BA}y$] An individual at location $x$ reproduces, 
@@
         and sends to $y$ an offspring
-        who inherits at the selected locus from the type A parent 
-        and the at the neutral locus from the type B parent.
-    \item[$x\xrightarrow{BB}y$] One type B individual at location $x$ reproduces, 
-        either does not recombine or recombines with another type B,
+        who inherits at the selected locus from the type B parent 
+        and the at the neutral locus from the type b parent.
+    \item[$x\xrightarrow{BB}y$] One type b individual at location $x$ reproduces, 
+        either does not recombine or recombines with another type b,
         and sends the offspring to $y$.
@@
 \begin{align}
-    & x\xrightarrow{AA}y & \qquad w_{AA}(x,y) &= p(y) (1+s_A(y)) \left(1 - r (1-p(y)) \right)  N(y) m(y,x) \\
-    & x\xrightarrow{AB}y & \qquad w_{AB}(x,y) &= r p(y) (1-p(y)) \left(1+\frac{s_A(y)+s_B(y)}{2}\right) N(y) m(y,x) \\
-    & x\xrightarrow{BA}y & \qquad w_{BA}(x,y) &= r p(y) (1-p(y)) \left(1+\frac{s_A(y)+s_B(y)}{2}\right) N(y) m(y,x) \\
-    & x\xrightarrow{BB}y & \qquad w_{BB}(x,y) &= (1-p(y)) (1+s_B(y)) \left(1 - r p(y) \right) N(y) m(y,x) 
+    & x\xrightarrow{AA}y & \qquad w_{AA}(x,y) &= p(y) (1+s_B(y)) \left(1 - r (1-p(y)) \right)  N(y) m(y,x) \\
+    & x\xrightarrow{AB}y & \qquad w_{AB}(x,y) &= r p(y) (1-p(y)) \left(1+\frac{s_B(y)+s_b(y)}{2}\right) N(y) m(y,x) \\
+    & x\xrightarrow{BA}y & \qquad w_{BA}(x,y) &= r p(y) (1-p(y)) \left(1+\frac{s_B(y)+s_b(y)}{2}\right) N(y) m(y,x) \\
+    & x\xrightarrow{BB}y & \qquad w_{BB}(x,y) &= (1-p(y)) (1+s_b(y)) \left(1 - r p(y) \right) N(y) m(y,x) 
 \end{align}
 
-These rates tell us precisely the rates at which a lineage will move, backwards in time.
-For instance, the rate at which a lineage currently in a type A individual at location $x$
-jumps to another type A individual at location $y$ is just $w_{AA}(y,x)$.
-A lineage at the selected locus (with $r=0$) in a type A individual moves according to these jump rates.
-If we follow a linked locus until the first time it is found in a type B individual,
+Note that at equilibrium, we require
+\begin{align}
+  0 = \sum_y N(y) m(y,x) \left\{ p(y)(1+s_B(y))(1-p(x)) - (1-p(y))(1+s_b(y))p(x) \right\} .
+\end{align}
+
+\subsection{Lineage movement}
+
+These rates tell us the rates at which a lineage will move, backwards in time.
+For instance, the rate at which a lineage at the selected locus
+currently in a type B individual at location $x$
+jumps to another type B individual at location $y$ is equal to the rate of influx of migrants from $y$
+divided by the number of B alleles at $x$,
+or
+\begin{align}
+  r_A(x,y) = \frac{w_AA(y,x)}{p(x)N(x)} = N(y) p(y)  (1+s_B(y)) m(y,x) \frac{ 1 }{ N(x) p(x) }
+\end{align}
+
+A lineage at the selected locus (with $r=0$) in a type B individual moves according to these jump rates.
+If we follow a linked locus until the first time it is found in a type b individual,
 it moves according to these jump rates, and at location $x$ recombines at rate
 \begin{align}
-    \gamma_A(x) = \sum_y w_{BA}(y,x) .
+    \gamma_B(x) = \sum_y w_{BA}(y,x) .
 \end{align}
-Suppose that we kill the lineage when it recombines onto type B.
+Suppose that we kill the lineage when it recombines onto type b.
 Let $X_t$ denote the position of the lineage at time $t$ in the past,
-with $X_t = \rho$ if it has recombined onto B,
+with $X_t = \rho$ if it has recombined onto b,
 and let $f$ be test function with $f(\rho)=0$.
-If we let $M_r(y) = p(y) (1+s_A(y)) \left(1 - r (1-p(y)) \right)  N(y)$, then
+If we let $M_r(y) = p(y) (1+s_B(y)) \left(1 - r (1-p(y)) \right)  N(y)$, then
 \begin{align}
-    \deriv{t} \E[f(X_t) \given X_0=x ] &= \sum_y w_{AA}(y,x) ( f(y)-f(x) ) - \gamma_A(x) f(x) \\
-                                       &= \sum_y  M_r(y) m(y,x) ( f(y) - f(x)  ) - \gamma_A(x) f(x) .
+    \deriv{t} \E[f(X_t) \given X_0=x ] &= \sum_y w_{AA}(y,x) ( f(y)-f(x) ) - \gamma_B(x) f(x) \\
+                                       &= \sum_y  M_r(y) m(y,x) ( f(y) - f(x)  ) - \gamma_B(x) f(x) .
 \end{align}
@@
 \begin{align}
-    \frac{1}{\sigma^2} \gamma_A(x) \xrightarrow{\sigma \to 0} \rho p(x) (1-p(x)) \left( 1 + \frac{s_A(x)+s_B(x)}{2} \right) N(x) ,
+    \frac{1}{\sigma^2} \gamma_B(x) \xrightarrow{\sigma \to 0} \rho p(x) (1-p(x)) \left( 1 + \frac{s_B(x)+s_b(x)}{2} \right) N(x) ,
 \end{align}
@@
 \begin{align}
-    M_r(x) \xrightarrow{\sigma \to 0} M(x) = p(x) (1+s_A(x)) N(x) .
+    M_r(x) \xrightarrow{\sigma \to 0} M(x) = p(x) (1+s_B(x)) N(x) .
 \end{align}
@@
 \begin{align}
-    Z = \int p(x) (1+s_A(x)) N(x) dx .
+    Z = \int p(x) (1+s_B(x)) N(x) dx .
 \end{align}
@@
 \paragraph{Recombination time}
-Let $\tau_\rho$ be the first time  the lineage recombines onto type B.
-If $\rho$ is much smaller than the mixing time of a type A lineage,
+Let $\tau_\rho$ be the first time  the lineage recombines onto type b.
+If $\rho$ is much smaller than the mixing time of a type B lineage,
 then the time to recombination is roughly Exponential,
@@
 \begin{align}
-    \pi(x) k(x) = \frac{1}{Z} N(x)^2 p(x)^2 (1-p(x)) (1+s_A(x)) (1+(s_A(x)+s_B(x))/2) .
+    \pi(x) k(x) = \frac{1}{Z} N(x)^2 p(x)^2 (1-p(x)) (1+s_B(x)) (1+(s_B(x)+s_b(x))/2) .
 \end{align}

### 2014-05-20 patchy-selection-paper.tex
did sometihng
.

@@
 \plr{let's leave two dimensions for the revisions. or, as an exercise? =)}
+
+%%%%%%% %%%%%%%%%%
+\subsection{Length of the shared haplotype within a patch}
+
+To really make use of the expressions in the previous section for the length of a haplotype shared
+between patches that are coadapted through migration,
+we need to know the length of the haplotype shared by individuals \emph{within} a patch.
+A detailed treatment of this is well beyond the scope of this paper,
+but we can get a good idea of the scale through heuristic arguments.
+To do this, we compare the rates of coalescence to the rate of recombination between B and b haplotypes.
+
+Sharing is ultimately determined by the coalescence time within a patch at the B allele.
+We could make arguments about coalescence between lineages moving in the patch \citep[as in][]{hallatschek},
+but given the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis},
+it seems reasonable to treat the patch as panmictic as far as coalescence is concerned.
+In other words, we suppose that all adapted individuals in a given patch have inherited their B allele from a single common ancestor living in the patch,
+and that the time back to the common ancestor at the selected locus between two randomly sampled individuals
+is exponentially distributed, with mean inversely proportional to the patch size, $1/(A\rho)$.
+If this is sufficiently shorter than the typical time to a common ancestor with a b allele,
+then the signal of a shared haplotype will be noticible.
+
+Consider a site at some distance to the selected locus, in two chromosomes carrying the B allele.
+The most recent common ancestor at this site may not be the same as that at the selected locus,
+but the two chromosomes will still share a haplotype at this site
+if coalescence at this site occurs before recombination onto the b background.
+In other words, if all ancestral chromosomes along the lineages connecting the two chromosomes to their most recent common ancestor
+carried the B allele,
+they share a B haplotype at that site.
+Therefore, we need to understand how often the ancestral lineage at a site linked to the selected locus
+moves between b and B backgrounds, as it moves across geography.
+\plr{alternate setup:}
+Consider the genotypes of two B-carrying chromosomes at a site linked to the selected site.
+The main thing we need to know is which of the following lived more recently:
+their most recent common ancestor at that site,
+or an ancestor of either of them at that site that carried the b allele.
+
+First consider the lineage of a single B allele, moving back through time.
+Suppose that the B allele is at mutation-selection balance, 
+with the local proportion of B alleles near $x$ equal to $p(x)$,
+and that we are close enough to the patch that fluctuations are small (unlike in previous sections).
+If we suppose that type B individuals at location $y$ reproduce at rate $1+s(y)$,
+and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
+then the total rate of influx of B alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
+This means that in time $dt$, a total of $ dt \, \rho p(y) (1+s(y)) m(y,x) \, dy \, dx $ new, type B migrants 
+will move from an area near $y$ of size $dy$ to an area near $x$ of size $dx$.
+Looking backwards in time, this is the number of $B$ alleles living at $x$
+who were produced by a parent from near $y$ in time $dt$.
+At equilibrium, there are $ \rho p(x) dx$ type B alleles in this area,
+so the probability that the lineage of a B allele at $x$ moves to near $y$ in time $dt$
+is 
+\begin{align}
+  r(x,y) \, dy = \frac{p(y)}{p(x)} (1+s(y)) m(y,x) \, dy .
+\end{align}
 

### 2014-05-21 lineage-movement.tex

@@
 and the number of individuals of type B and location $x$ is $n(x)$.
-Suppose that type B individuals at location $x$ reproduce at rate $1+s_B(x)$, 
-and likewise type b at rate $1+s_b(x)$.
+Suppose that type B individuals at location $x$ reproduce at rate $s_B(x)$, 
+and likewise type b at rate $s_b(x)$.
 At reproduction, individuals recombine with others in the same location,
@@
 \begin{enumerate}
-    \item[$x\xrightarrow{AA}y$] One type B individual at location $x$ reproduces, 
+    \item[$x\xrightarrow{BB}y$] One type B individual at location $x$ reproduces, 
         either does not recombine or recombines with another type B,
         and sends the offspring to $y$.
-    \item[$x\xrightarrow{AB}y$] An individual at location $x$ reproduces, 
+    \item[$x\xrightarrow{Bb}y$] An individual at location $x$ reproduces, 
         recombines with the other type,
@@
         and the at the neutral locus from the type B parent.
-    \item[$x\xrightarrow{BA}y$] An individual at location $x$ reproduces, 
+    \item[$x\xrightarrow{bB}y$] An individual at location $x$ reproduces, 
         recombines with the other type,
@@
         and the at the neutral locus from the type b parent.
-    \item[$x\xrightarrow{BB}y$] One type b individual at location $x$ reproduces, 
+    \item[$x\xrightarrow{bb}y$] One type b individual at location $x$ reproduces, 
         either does not recombine or recombines with another type b,
@@
 \begin{align}
-    & x\xrightarrow{AA}y & \qquad w_{AA}(x,y) &= p(y) (1+s_B(y)) \left(1 - r (1-p(y)) \right)  N(y) m(y,x) \\
-    & x\xrightarrow{AB}y & \qquad w_{AB}(x,y) &= r p(y) (1-p(y)) \left(1+\frac{s_B(y)+s_b(y)}{2}\right) N(y) m(y,x) \\
-    & x\xrightarrow{BA}y & \qquad w_{BA}(x,y) &= r p(y) (1-p(y)) \left(1+\frac{s_B(y)+s_b(y)}{2}\right) N(y) m(y,x) \\
-    & x\xrightarrow{BB}y & \qquad w_{BB}(x,y) &= (1-p(y)) (1+s_b(y)) \left(1 - r p(y) \right) N(y) m(y,x) 
+    & x\xrightarrow{BB}y & \qquad w_{BB}(x,y) &= p(x) s_B(x) \left(1 - r (1-p(x)) \right)  N(x) m(x,y) \\
+    & x\xrightarrow{Bb}y & \qquad w_{Bb}(x,y) &= r p(x) (1-p(x)) \frac{s_B(x)+s_b(x)}{2} N(x) m(x,y) \\
+    & x\xrightarrow{bB}y & \qquad w_{bB}(x,y) &= r p(x) (1-p(x)) \frac{s_B(x)+s_b(x)}{2} N(x) m(x,y) \\
+    & x\xrightarrow{bb}y & \qquad w_{bb}(x,y) &= (1-p(x)) s_b(x) \left(1 - r p(x) \right) N(x) m(x,y) 
 \end{align}
@@
 \begin{align}
-  0 = \sum_y N(y) m(y,x) \left\{ p(y)(1+s_B(y))(1-p(x)) - (1-p(y))(1+s_b(y))p(x) \right\} .
+  0 = \sum_y N(y) m(y,x) \left\{ p(y)s_B(y)(1-p(x)) - (1-p(y))s_b(y)p(x) \right\} .
 \end{align}
@@
 \begin{align}
-  r_A(x,y) = \frac{w_AA(y,x)}{p(x)N(x)} = N(y) p(y)  (1+s_B(y)) m(y,x) \frac{ 1 }{ N(x) p(x) }
+  r_A(x,y) &= \frac{w_{BB}(y,x)}{p(x)N(x)} \\
+  &= N(y) p(y)  s_B(y) m(y,x) \frac{ 1 }{ N(x) p(x) }
 \end{align}
 
-A lineage at the selected locus (with $r=0$) in a type B individual moves according to these jump rates.
 If we follow a linked locus until the first time it is found in a type b individual,
@@
 \begin{align}
-    \gamma_B(x) = \sum_y w_{BA}(y,x) .
+  \gamma_B(x) &= \frac{1}{N(x)p(x)}\sum_y w_{bB}(y,x)  \\
+  &= r p(y) (1-p(y)) \frac{s_B(y)+s_b(y)}{2} N(y) m(y,x) \frac{1}{N(x)p(x)}.
 \end{align}
@@
 and let $f$ be test function with $f(\rho)=0$.
-If we let $M_r(y) = p(y) (1+s_B(y)) \left(1 - r (1-p(y)) \right)  N(y)$, then
+If we let $M_r(y) = s_B(y) \left(1 - r (1-p(y)) \right)$, then
 \begin{align}
-    \deriv{t} \E[f(X_t) \given X_0=x ] &= \sum_y w_{AA}(y,x) ( f(y)-f(x) ) - \gamma_B(x) f(x) \\
-                                       &= \sum_y  M_r(y) m(y,x) ( f(y) - f(x)  ) - \gamma_B(x) f(x) .
+    \deriv{t} \E[f(X_t) \given X_0=x ] &= \sum_y r_{B}(y,x) ( f(y)-f(x) ) - \gamma_B(x) f(x) \\
+    &= \frac{1}{N(x)p(x)} \sum_y  N(y) p(y) M_r(y) m(y,x) ( f(y) - f(x)  ) - \gamma_B(x) f(x) . \label{eqn:discrete_generator}
 \end{align}
@@
     &= \frac{1}{2} \left( g(x) f'(x) + 2 g'(x) f'(x) + f(x) g''(x) - f(x) g''(x) \right) \\
-    &= \frac{1}{2} g(x) f''(x) + g'(x) f'(x) .
+    &= \frac{1}{2} g(x) f''(x) + g'(x) f'(x) . \label{eqn:deriv_limit}
 \end{align}
+
 If we also rescale recombination, so that $\rho = r/\sigma^2$,
@@
 \begin{align}
-    \frac{1}{\sigma^2} \gamma_B(x) \xrightarrow{\sigma \to 0} \rho p(x) (1-p(x)) \left( 1 + \frac{s_B(x)+s_b(x)}{2} \right) N(x) ,
+    \frac{1}{\sigma^2} \gamma_B(x) \xrightarrow{\sigma \to 0} \rho (1-p(x)) \frac{s_B(x)+s_b(x)}{2} ,
 \end{align}
@@
 \begin{align}
-    M_r(x) \xrightarrow{\sigma \to 0} M(x) = p(x) (1+s_B(x)) N(x) .
+    M_r(x) \xrightarrow{\sigma \to 0} M(x) = s_B(x) .
 \end{align}
-Under these assumptions,
+Under these assumptions, combining \eqref{eqn:discrete_generator} and \eqref{eqn:deriv_limit},
 \begin{align}
     \deriv{t} \E[f(X_{t/\sigma^2}) \given X_0=x ] &\xrightarrow{\sigma \to 0} 
-    \frac{1}{2} M(x) f''(x) + M'(x) f'(x) - \rho k(x) f(x) ,
+    \frac{1}{2} s_B(x) \dderiv{x} f(x) + \frac{1}{N(x)p(x)} \deriv{x} \left\{ N(x)p(x) s_B(x) \right\}  \deriv{x} f(x) - \rho k(x) f(x) \\
+    &= \frac{1}{2} s_B(x) f''(x) + \left( s_B'(x) + \log(N(x)p(x))' s_B(x) \right) f'(x)  - \rho k(x) f(x) ,
 \end{align}
-i.e.\ $X_{t/\sigma^2}$ converges to a diffusion with drift $M'(x)$ and killed at rate $\rho k(x)$.
+i.e.\ $X_{t/\sigma^2}$ converges to a diffusion with drift $\frac{1}{ N(x)p(x)} \deriv{x} ( N(x) p(x) s_B(x) )$ and killed at rate $\rho k(x)$.
 
@@
 If $X$ has a stationary distribution then its density $\pi(x)$ solves the forward Kolmogorov equation,
+which is here (letting $n(x) = N(x) p(x)$ for convenience)
 \begin{align}
-    0 &= - \deriv{x}\left( M'(x) \pi(x) \right) + \frac{1}{2} \dderiv{x} \left( M(x) \pi(x) \right) \\
-      &= - M''(x) \pi(x) - M'(x) \pi'(x) + \frac{1}{2} M''(x) \pi(x) + M'(x) \pi'(x) + \frac{1}{2} M(x) \pi''(x)  \\
-      &= \frac{1}{2} \left( M(x) \pi''(x) - M''(x) \pi(x) \right) .
+  0 &= \frac{1}{2} \dderiv{x} \left( s_B(x) \pi(x) \right) - \deriv{x} \left( \frac{\pi(x)}{n(x)} \deriv{x}(n(x) s_B(x)) \right) .
 \end{align}
-It is immediate that $\pi(x) = M(x)/Z$,
-where
+It is easy to check that $\pi(x) \propto n(x)^2 s(x) $ is a solution,
+and since it is nonnegative, the stationary distribution,
+i.e.
 \begin{align}
-    Z = \int p(x) (1+s_B(x)) N(x) dx .
+  \pi(x) = \frac{ \left( N(x) p(x) \right)^2 s_B(x) }{ \int_y \left( N(y) p(y) \right)^2 s_B(y) dy } ,
 \end{align}
+if the denominator (below denoted by $Z$) is finite.
+
+In other words, the stationary distribution is proportional to the local rate of production of $B$ alleles, \emph{squared}.
+
+\paragraph{Speed and scale}
+It will be useful to think about the Dubins-Schwartz construction of this diffusion.
+Note that if we define
+\begin{align}
+  z(x) = \int_{0}^x \frac{ dy }{ n(y)^2 s_B(y)^2 },
+\end{align}
+where $n(x) = N(x) p(x)$ as before,
+then
+\begin{align}
+  \deriv{t} \E^x[z(X_t)] &= \frac{1}{2} s_B(x) \deriv{x} \frac{1}{ n(x)^2 s_B(x)^2 } + \frac{1}{n(x)} \frac{ \deriv{x}(n(x)s_B(x)) }{  n(x)^2 s_B(x)^2 } \\
+  &= - s_B(x) \frac{ \deriv{x} (n(x)s_B(x)) }{ n(x)^3 s_B(x)^3 } + \frac{1}{n(x)} \frac{ \deriv{x}(n(x)s_B(x)) }{  n(x)^2 s_B(x)^2 } \\
+  &= 0 ,
+\end{align}
+that $Z_t = z(X_t)$ is a martingale.
+
+Since
+\begin{align}
+  dX_t = \frac{\deriv{x}(n(X_t)s_B(X_t))}{n(X_t)} dt + \sqrt{s_B(X_t)} dB_t,
+\end{align}
+then if $s_B(x)$ is bounded below, the lineage won't get stuck,
+and will always come back to the patch.
+Since $z'(x) = 1/(n(x)s_B(x))^2$,
+applying It\^o's formula,
+\begin{align}
+  dZ_t &=  z'(X_t) dX_t + \frac{1}{2} z''(X_t) s_B(X_t) dt \\
+  &= \frac{ dB_t }{ n^2(X_t) s_B^{3/2}(X_t) } ,
+\end{align}
+i.e.\ $Z_t$ ``looks like Brownian motion moving at speed $1/(n^4(X_t) s_B^3(X_t))$''.
+This happens because the transformation $z(x)$ squooshes up locations far away from the patch,
+where the lineage moves at a more-or-less constant speed;
+so that $Z_t$ has to move faster in those squooshed-up regions.
 
@@
 \begin{align}
-    \pi(x) k(x) = \frac{1}{Z} N(x)^2 p(x)^2 (1-p(x)) (1+s_B(x)) (1+(s_B(x)+s_b(x))/2) .
+    \pi(x) k(x) = \frac{1}{Z} N(x)^2 p(x)^2 s_B(x) (1-p(x)) (s_B(x)+s_b(x))/2 .
 \end{align}
@@
 \begin{align}
-    -1 = M'(x) h'(x) + \frac{1}{2} M(x) h''(x) - \rho k(x) h(x) .
+  -1 = \frac{1}{n(x)}(n(x)s_B(x))' h'(x) + \frac{1}{2} s_B(x) h''(x) - \rho k(x) h(x) .
 \end{align}
@@
 \begin{align}
-  M'(x) \ell'(x) + \frac{1}{2} M(x) \ell''(x) - (\lambda + \rho k(x) ) \ell(x) = \lambda .
+  \frac{1}{n(x)}(n(x)s_B(x))' h'(x) + \frac{1}{2} s_B(x) h''(x) - (\lambda + \rho k(x)) h(x) = \lambda .
 \end{align}

### 2014-05-21 patchy-selection-paper.tex
.

@@
   dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
-  }   \label{fig:sim_occupation_freqs}
+ \gc{The approximation and simulations disagree at long distances from
+   the patch, but this is due to edge effects in the simulations.} }   \label{fig:sim_occupation_freqs}
 \end{figure}
@@
 who were produced by a parent from near $y$ in time $dt$.
-At equilibrium, there are $ \rho p(x) dx$ type B alleles in this area,
-so the probability that the lineage of a B allele at $x$ moves to near $y$ in time $dt$
-is 
-\begin{align}
-  r(x,y) \, dy = \frac{p(y)}{p(x)} (1+s(y)) m(y,x) \, dy .
-\end{align}
+
+XXX
 

### 2014-05-22 lineage-movement.tex

@@
 \begin{align}
-    M_r(x) \xrightarrow{\sigma \to 0} M(x) = s_B(x) .
+    M_r(x) \xrightarrow{\sigma \to 0} s_B(x) .
 \end{align}
@@
 
+\paragraph{Rescaling selection out}
+Suppose for the moment that $s_b(x)=1$. 
+It hasn't entered here much anyhow.
+First note that if we change spatial variables to
+\begin{align}
+  s(x) = \int_0^x \sqrt{s_B(y)} dy
+\end{align}
+then $ds = \sqrt{s(x)} dx$, and the generator for the process becomes
+\begin{align}
+  f \mapsto 
+    \frac{1}{2} f''(s) + \log\left( N(s) p(s) s_B(s) \right)' f'(s) - \rho k(s) f(s) .
+\end{align}
+At first sight this looks like the usual scaling of $\sigma/\sqrt{s}$,
+but it isn't, because in the notation I'm using there, $s_B$ is around 1,
+i.e.\ $s_B = 1+s$ if $s$ is the usual selection coefficient.
+
 \paragraph{Recombination time}
@@
 
+%%%%%%%%%
+\section{Integrals against $p$}
+
+Recall that if $N(x)$ is constant, the equilibrium frequency $p(x)$ solves
+\[ \frac{1}{2} p''(x) + s(x) p(x)(1-p(x)) = 0 \quad ,\]
+where $s(x) = s_B(x)-s_b(x)$,
+and that if $s(x) = s$ is constant on a region, the \emph{inverse} solves
+\begin{align}
+  \frac{dx}{dp} = \frac{1}{\sqrt{sp^2(2p/3-1) + K}},
+\end{align}
+where $K$ is a constant to make the boundary conditions come out right.
+
+If this is the case, then we can change variables:
+\begin{align}
+  \int_a^b f(p(x)) dx = \int_{p(a)}^{p(b)} \frac{f(p) dp }{\sqrt{sp^2(2p/3-1) + K}} .
+\end{align}
+so for instance, 
+the denominator for the stationary distribution (dropping the $N$'s)
+\begin{align}
+  Z &= \int p(y)^2 s_B(y) dy \\
+  &= \int \frac{ p (1+s) dp }{ \sqrt{ s (2p/3-1) + K } } .
+\end{align}
+Note, however, that the first expression makes it clear that if we make the patch bigger, that $Z$ will increase roughly proportionally
+-- since this increases the area over which $p(y)\approx 1$.
+To make this occur in the integral over $p$ on the second line, 
+it must be the case that $K$ is approximately $s/3$ plus a term inversely proportional to the area of the patch.
+
+The integral of $\pi(x) k(x)$ is
+\begin{align}
+  \frac{1}{Z} \int_a^b p(x)^2(1-p(x))s_B(x)(s_b(x)+s_B(x)) dx = \frac{1}{Z} \int_{p(a)}^{p(b)} \frac{p (1-p) (1+s)(1+s/2) dp }{\sqrt{s(2p/3-1) + K}} .
+\end{align}
+I have the feeling Barton does similar things in one of his papers,
+but haven't been able to track it down.
+
 \paragraph{Questions:} 
 \begin{enumerate}
-
-    \item Can we make any progress analytically with any of these quantities,
-      using the fact that
-      \[ \frac{1}{2} p''(x) + s(x) p(x)(1-p(x)) = 0 \quad ?\]
 

### 2014-05-22 patchy-selection-paper.tex
added Lava_flow_mice_app.R Lava_flow_mice_prob_parallel.pdf

@@
 On the basis of these numbers we can gain a sense of a plausible range of characteristic lengths.
-Assuming the allele is at 50\% frequency at the edge of the lava flow, 
+Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
 we can fit the frequency found at each of the light substrate sites that are 
 polymorphic for the dark allele under a simple allele frequency cline model \citep{XXX}.
-Doing this for the Tule we obtain $\sigma/\sqrt{s_m} \approx 4$km and for the O'Neill site
-$\sigma/\sqrt{s_m} \approx 30$km.
+Doing this for the Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km and for the O'Neill site
+$\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. We also need estimate of $s_m$,
+using  $\sigma=1km$ \gc{REF} our cline widths imply $s_m=3^{-1/2}$ and
+$30^{-1/2}$ are reasonable values.
 
@@
 pigmentation have been identified in closely related species and more
-broadly across vertibrates \citep{}.To span a range of plausible
-values, we use a low mutation rate 
-We use a low mutation rate $\mu= 10^{-8}$, where the adaptive
-mtation has to occur at a single base pair.
-An intermediate mutation rate $\mu= 10^{-5}$, equivalent to
-the mutation being able to arise at any of a thousand bases pairs,
-e.g. any of the coding bases of a typical length gene. And a high
-mutation rate $\mu = 10^{3}$, corresponding to the adaptive mutation
-being able to arise in any of $100$ genes, e.g. anywhere in a large pathway.
+broadly across vertebrates \citep{REF}.To span a range of plausible
+values we use a low mutation rate of $\mu= 10^{-8}$, where the adaptive
+mtation has to occur at a single base pair; and a high mutation
+rate $\mu= 10^{-5}$, equivalent to the mutation being able to arise at any of a thousand base pairs
+e.g. any of the coding bases of a typical length gene.  We set $A=100km^2$ reflecting a patch with a width of $\sim 10km$.
 

### 2014-05-22 Lava_flow_mice_app.R

@@ -1,0 +1,29 @@
+char.length= c(3,30)
+
+
+A=100
+
+#pdf(file="~/Dropbox/Talks_posters/coopibarra_patchy_2013/figures/prob_parallel_hoekstra_NEW.pdf")
+R = seq(5,600,length=10000) ##distance in Km
+#layout(t(1:2))
+mu = 1e-8
+prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
+plot(R, prob.parallel,type="l",lwd=3, xlab="R, km",ylab="Prob. parallel mutation.",cex.lab=1.5,cex.axis=1.5,log="x") 
+mu=1e-5
+prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
+
+lines(R, prob.parallel,lwd=3,col="red")
+
+
+R = seq(5,600,length=1000) ##distance in Km
+mu = 1e-8
+prob.parallel <- mu*A/(mu*A + (1/char.length[2])* exp(-R/char.length[2]))  ##assuming sigma=1
+lines(R, prob.parallel,lty=2,lwd=3)
+
+mu=1e-5
+prob.parallel <- mu*A/(mu*A + (1/char.length[2])* exp(-R/char.length[2]))  ##assuming sigma=1
+
+lines(R, prob.parallel,lwd=3,col="red",lty=2)
+
+legend("topleft", legend=expression(mu == 10^-8, mu == 10^-5, sigma/sqrt(s[m]) == 3,sigma/sqrt(s[m]) == 30),col=c("black","red","grey","grey"),lty=c(1,1,1,2),lwd=2)
+#dev.off()

### 2014-05-24 lineage-movement.tex
poorly worded meanderings about haplotypes within the patch

@@
 
-\paragraph{Questions:} 
+%%%%%%
+\section{Haplotype lengths}
+
+Some statements above about ``haplotype length'' aren't quite right.
+Consider the haplotype shared about two B alleles;
+by the time these coalesce, their ``shared haplotype'' in the above sense may be broken up into a number of blocks --
+all coalescing linked to other B alleles.
+More properly, the length of the contiguous haplotype
+is determined by branching diffusions.
+The single-lineage version is thus:
+begin a diffusion with mass equal to genetic length (of something, rescaled);
+then, any remaining particles split in two at rate proportional to their length,
+always labeling the particle corresponding to the selected allele;
+when this happens any particles other than the labeled particle may die
+or may continue as an independent diffusion.
+
+\section{Questions:} 
 \begin{enumerate}

### 2014-05-24 patchy-selection-paper.tex
poorly worded meanderings about haplotypes within the patch

--- 2014-05-21 patchy-selection-paper.tex
@@
   dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
- \gc{The approximation and simulations disagree at long distances from
-   the patch, but this is due to edge effects in the simulations.} }   \label{fig:sim_occupation_freqs}
+  }   \label{fig:sim_occupation_freqs}
 \end{figure}
@@
 
+\plr{how's that?}
+
 %%%
@@
 then the total rate of influx of B alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
-This means that in time $dt$, a total of $ dt \, \rho p(y) (1+s(y)) m(y,x) \, dy \, dx $ new, type B migrants 
-will move from an area near $y$ of size $dy$ to an area near $x$ of size $dx$.
-Looking backwards in time, this is the number of $B$ alleles living at $x$
-who were produced by a parent from near $y$ in time $dt$.
-
-XXX
+As argued by \citet{hudson},
+in a discrete model the rate at which a B lineage moves from $x$ to $y$
+is this influx divided by the local abundance of $B$ alleles:
+\begin{align}
+  r(x,y) = \frac{ p(y) (1+s(y)) m(y,x) }{ p(x) } .
+\end{align}
+After the same large-population, continuous-space limit that underlies the rest of this paper,
+this motion should converge to a diffusion process \citep[that is similar to that studied by][]{hallatschek},
+notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
+In It\^o differential notation, 
+the location $X_t$ of a B lineage $t$ generations ago satisfies
+\begin{align}
+  d X_t &= (1+s(X_t)) D \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,\\
+    % &=  ( \log(p(X_t))' (1+s(X_t)) + s'(X_t) ) dt + \sqrt{1+s(X_t)} dB_t,\\
+\end{align}
+where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
+$B_t$ is a standard Brownian motion, and $D$ denotes the gradient operator.
+The main effect is the drift term, that pulls the lineage towards regions where $p(x)$ is larger;
+since $s$ is generally small, the speed of the Brownian noise term has little effect.
+Furthermore, it is easy to verify that the stationary distribution of this process is
+\begin{align}
+  \pi(x) = \frac{ p(x)^2 (1+s(x)) }{ \int_y p(y)^2 (1+s(y)) dy } .
+\end{align}
+
+If we are instead following a lineage at a locus at genetic distance $r$ from the selected locus,
+and begin linked to a B allele,
+then the lineage will move as a B lineage up until the most recent time
+it is found linked to a b allele.
+If the lineage is at $x$, the probability that a recombination that moves it from a B background
+to a b background occurs in time $dt$ is $dt (1-p(x)) (1+s(x)/2)$.
+The Feynman-Kac formula could then be used to write down differential equations
+solved by various moments of the haplotype length as a function of the sampling location of the lineage.
+
+More simply, consider: within the patch, where $p(x) \approx 1$, 
+a linked lineage moves roughly as Brownian motion, and is unlikely to recombine onto a b background.
+The rate of recombination between B and b backgrounds depends on the proportion of the patch in which $p(x)$ is intermediate.
+This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
+As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
+Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
+(and, therefore, the rate of recombination)
+is roughly $r (\sigma/\sqrt{s})A^{-1/d}$.
+This will be equal to the coalescence rate if
+$1/(A\rho) = r (\sigma/\sqrt{s}) A^{-1/d}$, 
+and hence the length of linkage shared within the patch is on the order of
+\begin{align}
+  \frac{1}{r} = \frac{ \sigma A^{(d-1)/d} }{\sqrt{s}} ,
+\end{align}
+which does not increase with $A$ in one dimension,
+but increases with $\sqrt{A}$ in two dimensions.
 

### 2014-05-24 patchy-selection-paper.tex
added applications

--- 2014-05-22 patchy-selection-paper.tex
@@
   new mutation: \eqref{eqn:mutrate}; 
+ via migration  \eqref{eqn:migrate}
+
+Expected transit time \label{eqn:mean_tau}
+Expected haplotype length \label{eqn:haplotype_length}
 
@@
 i.e.\ the probability of parallel adapatation, is in $d=1$ ($d=2$ only differs slightly)
-\begin{equation}
-  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A \mu \alpha + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },
+\begin{equation} \label{eqn:parallel_prob}
+  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A
+    \mu \alpha + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
 \end{equation}
@@
 comes from patch $i$ is 
-\begin{align}
+\begin{align}  \label{eqn:prob-parallel}
   \frac{\left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
@@
 
-\gc{add some
-  sentences of intuition of how parameters play into haplotype length
-  here. is it really $\sigma \sqrt{s_m}$ in \eqref{eqn:haplotype_length} and not $\sqrt{s_m}/sigma$???}
-
-\gc{Perhaps we should separate the above bit into things that are
-  useful about $\tau$ and put those up by  \eqref{eqn:borodinresult} and
-  things that are useful about $L$ here? We should also turn some of
-  this into numbered equations. I tried to do this above, and have the
-commented out original below}
-
 %%%
@@
 multiple dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
-despite intervening light habitat.
+despite intervening light habitat. We use our framework to investigate
+how the probability of parallel mutation depends on the distance
+between lava flow patches.
 
@@
 rate $\mu= 10^{-5}$, equivalent to the mutation being able to arise at any of a thousand base pairs
-e.g. any of the coding bases of a typical length gene.  We set $A=100km^2$ reflecting a patch with a width of $\sim 10km$.
+e.g. any of the coding bases of a typical length gene.  Finally we set
+$A=100km^2$ reflecting a patch with a width of $\sim 10km$. In Figure
+\ref{fig:mice_prob_parallel} we show the dependance of the probability
+of parallel mutation on the distance between lava flow patches using
+these parameters. As can be seen parallel mutation should become
+likely over a scale of tens to a few hundred kilometers between patches.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{Lava_flow_mice_prob_parallel}
+  \end{center}
+  \caption{
+The probability of parallel mutation as a function of the distance
+between environmental patches for two different cline widths
+and two different mutation rates (using eqn. \eqref{eqn:prob-parallel}). The parameters were chosen to match  
+  \label{fig:mice_prob_parallel}
+  }
+\end{figure}
 

### 2014-05-25 patchy-selection-paper.tex
edited application

@@
 \section{Applications} 
-Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, \emph{Chaetodipus intermedius}, building on classic work by \citet{dice1940ecologic,benson1933concealing}.
-These mice live on rock outcrops throughout the southwestern United States and northern Mexico, 
-and generally have light pigmentation, presumably to match the color of the rock, avoiding visual-based predators. 
-However, in some regions these mice live on darker substrates such as old lava flows, 
-and in these patches have much darker pigmentation. 
-\citet{nachman2003different} demonstrated that on one of these flows (Pinacate), an allele at the MC1R locus is responsible for much of the change to a dark pelage.
+Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, 
+\emph{Chaetodipus intermedius}, building on classic work by \citet{benson1933concealing} and \citet{dice1940ecologic}.
+These mice live on rocky outcrops throughout the southwestern United States and northern Mexico, 
+and generally have a light pelage similar to the predominant rock color.
+However, in some regions these mice live on darker substrates (e.g.\ old lava flows),
+and in these patches have much darker pigmentation, 
+presumably to avoid visual predators.
+\citet{nachman2003different} demonstrated that on one of these flows (Pinacate), 
+much of the change in coloration is due to an allele at the MC1R locus.
 This dark allele differs from the light allele by four amino acid changes, 
 and has a dominant or partially dominant effect depending on the measure of coat color assessed. 
-These lava flows are separated from each other by a range of distances, 
+These areas of dark rock are separated from each other by a range of distances, 
 with some being separated from any other by hundreds of kilometers of light colored substrate. 
 The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
-suggesting these populations have adapted in parallel. 
+suggesting these populations have adapted in parallel \citep{xxx}.
 However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, 
 multiple dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
-despite intervening light habitat. We use our framework to investigate
-how the probability of parallel mutation depends on the distance
-between lava flow patches.
-
-As can be seen from the above, a key parameter is the characteristic length $\sigma/\sqrt{s_m}$. 
-\citep{Hoekstra:04} studied the frequency of the dark MC1R allele, and coat color phenotypes, at sites on the Pinacate lava flow and at two nearby sites in light-colored rock.
-On the lava flow the dark MC1R allele is at 86\% frequency. 
-At the Tule light substrate site, which is $\sim$ 12km from the flow, the dark MC1R allele was at $\sim$ 3\% frequency. 
-At a greater distance from the lava (XXXkm) the light colored substrate the dark MC1R was absent from a sample at the Christmas pass. 
-While at the other light substrate site (O'Neill), which is 10km from the flow on opposite side from Tule, 
-the dark MC1R allele has a frequency of $\sim$ 34\%. 
-On the basis of these numbers we can gain a sense of a plausible range of characteristic lengths.
+despite intervening light habitat. 
+We apply our theory to this example,
+in particular investigating how the probability of parallel mutation depends on 
+the distance between lava flow patches.
+
+A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. 
+\citep{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
+at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
+On the lava flow the dark MC1R allele is at 86\% frequency,
+while at a site with light substrate 12km to the west (Tule), the frequency is 3\%.
+The dark allele was entirely absent from Christmas pass, a site with light substrate 7km north of Tule, and 3km further from the lava flow.
+In the other direction, the dark MC1R allele was at 34\% at a site with light substrate 10km to the east of the flow (O'Neill).
+These numbers give us a sense of a plausible range of characteristic lengths.
 Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
-we can fit the frequency found at each of the light substrate sites that are 
-polymorphic for the dark allele under a simple allele frequency cline model \citep{XXX}.
-Doing this for the Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km and for the O'Neill site
-$\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. We also need estimate of $s_m$,
-using  $\sigma=1km$ \gc{REF} our cline widths imply $s_m=3^{-1/2}$ and
-$30^{-1/2}$ are reasonable values.
+we can fit formula \eqref{eqn:eqfreq} to these frequencies.
+Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
+and for O'Neill, $\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. 
+We also need estimate of $s_m$. 
+Using $\sigma=1km$ \gc{REF}, these cline widths imply $s_m=1/\sqrt{3}$ and
+$1/\sqrt{30}$ are reasonable values.
+\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
+which has dispersal distance of 0.26km.  Shall we ask Nachman?}
 
@@
 While the Pinacate dark haplotype differs from the light haplotype at three amino-acid residues,
-it is likely that not all of these changes are needed for a population to begin to  
-adapt. Also there a number of genes, beyond MC1R, at which adaptative changes affecting 
-pigmentation have been identified in closely related species and more
-broadly across vertebrates \citep{REF}.To span a range of plausible
-values we use a low mutation rate of $\mu= 10^{-8}$, where the adaptive
-mtation has to occur at a single base pair; and a high mutation
-rate $\mu= 10^{-5}$, equivalent to the mutation being able to arise at any of a thousand base pairs
-e.g. any of the coding bases of a typical length gene.  Finally we set
-$A=100km^2$ reflecting a patch with a width of $\sim 10km$. In Figure
-\ref{fig:mice_prob_parallel} we show the dependance of the probability
-of parallel mutation on the distance between lava flow patches using
-these parameters. As can be seen parallel mutation should become
+it is likely that not all of these changes are needed for a population to begin to  adapt. 
+Also there a number of genes besides MC1R at which adaptative changes affecting pigmentation 
+have been identified in closely related species and more broadly across vertebrates \citep{REF}.
+To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
+where the adaptive mtation has to occur at a single base pair; 
+and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation being able to arise at any of a thousand base pairs,
+e.g.\ any of the coding bases of a typical gene.  
+Finally, we set $A=100\text{km}^2$.
+Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
+of parallel mutation on the distance between lava flow patches using these parameters. 
+As can be seen parallel mutation should become
 likely over a scale of tens to a few hundred kilometers between patches.
@@
 between environmental patches for two different cline widths
-and two different mutation rates (using eqn. \eqref{eqn:prob-parallel}). The parameters were chosen to match  
+and two different mutation rates (using equation \eqref{eqn:prob-parallel}). 
+The parameters were chosen to match  
   \label{fig:mice_prob_parallel}

### 2014-05-28 patchy-selection-paper.tex
missing rho
haplotype within a patch section rewrite

@@
 \subsection{Length of the shared haplotype within a patch}
-
-To really make use of the expressions in the previous section for the length of a haplotype shared
-between patches that are coadapted through migration,
+\label{ss:patch_haplotype}
+
+To really make use of the expressions in the previous section,
 we need to know the length of the haplotype shared by individuals \emph{within} a patch.
@@
 but we can get a good idea of the scale through heuristic arguments.
-To do this, we compare the rates of coalescence to the rate of recombination between B and b haplotypes.
-
-Sharing is ultimately determined by the coalescence time within a patch at the B allele.
-We could make arguments about coalescence between lineages moving in the patch \citep[as in][]{hallatschek},
-but given the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis},
-it seems reasonable to treat the patch as panmictic as far as coalescence is concerned.
-In other words, we suppose that all adapted individuals in a given patch have inherited their B allele from a single common ancestor living in the patch,
-and that the time back to the common ancestor at the selected locus between two randomly sampled individuals
-is exponentially distributed, with mean inversely proportional to the patch size, $1/(A\rho)$.
-If this is sufficiently shorter than the typical time to a common ancestor with a b allele,
-then the signal of a shared haplotype will be noticible.
-
-Consider a site at some distance to the selected locus, in two chromosomes carrying the B allele.
-The most recent common ancestor at this site may not be the same as that at the selected locus,
-but the two chromosomes will still share a haplotype at this site
-if coalescence at this site occurs before recombination onto the b background.
-In other words, if all ancestral chromosomes along the lineages connecting the two chromosomes to their most recent common ancestor
-carried the B allele,
-they share a B haplotype at that site.
-Therefore, we need to understand how often the ancestral lineage at a site linked to the selected locus
-moves between b and B backgrounds, as it moves across geography.
-\plr{alternate setup:}
-Consider the genotypes of two B-carrying chromosomes at a site linked to the selected site.
-The main thing we need to know is which of the following lived more recently:
-their most recent common ancestor at that site,
-or an ancestor of either of them at that site that carried the b allele.
-
-First consider the lineage of a single B allele, moving back through time.
-Suppose that the B allele is at mutation-selection balance, 
-with the local proportion of B alleles near $x$ equal to $p(x)$,
-and that we are close enough to the patch that fluctuations are small (unlike in previous sections).
+To do this, we compare the rate of recombination between B and b haplotypes
+to the rate of coalescence of B haplotypes within a patch,
+which we take to be $1/(A\rho)$ 
+(bypassing the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis}).
+
+Consider the lineage of a single B allele, moving back through time,
+in a patch at mutation-selection balance, 
+i.e.\ with the local proportion of B alleles near $x$ equal to $p(x)$,
+and that the lineage stays close enough to the patch 
+that fluctuations about $p(x)$ are small (unlike in previous sections).
 If we suppose that type B individuals at location $y$ reproduce at rate $1+s(y)$,
@@
 then the total rate of influx of B alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
-As argued by \citet{hudson},
-in a discrete model the rate at which a B lineage moves from $x$ to $y$
-is this influx divided by the local abundance of $B$ alleles:
-\begin{align}
-  r(x,y) = \frac{ p(y) (1+s(y)) m(y,x) }{ p(x) } .
-\end{align}
+As argued by \citet{hudson}, the rate at which a B lineage moves from $x$ to $y$
+is this influx divided by the local abundance of $B$ alleles,
+or $\frac{ p(y) (1+s(y)) m(y,x) }{ p(x) }$.
 After the same large-population, continuous-space limit that underlies the rest of this paper,
-this motion should converge to a diffusion process \citep[that is similar to that studied by][]{hallatschek},
+this motion should converge to a diffusion process \citep[as assumed by][]{hallatschek},
 notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
@@
 \begin{align}
-  d X_t &= (1+s(X_t)) D \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,\\
+    d X_t &= (1+s(X_t)) \frac{d}{d\,x} \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,\\
     % &=  ( \log(p(X_t))' (1+s(X_t)) + s'(X_t) ) dt + \sqrt{1+s(X_t)} dB_t,\\
@@
 where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
-$B_t$ is a standard Brownian motion, and $D$ denotes the gradient operator.
-The main effect is the drift term, that pulls the lineage towards regions where $p(x)$ is larger;
-since $s$ is generally small, the speed of the Brownian noise term has little effect.
+and $B_t$ is a standard Brownian motion.
+This is a diffusion that, when at $x$, 
+is pulled toward regions with more B alleles
+by the drift term $(1+s(x))\frac{d}{d\,x} \log( p(x) (1+s(x)) )$,
+and moves at nearly constant speed (since $\sqrt{1+s(x)}$ is close to 1).
 Furthermore, it is easy to verify that the stationary distribution of this process is
@@
 
-If we are instead following a lineage at a locus at genetic distance $r$ from the selected locus,
-and begin linked to a B allele,
-then the lineage will move as a B lineage up until the most recent time
-it is found linked to a b allele.
-If the lineage is at $x$, the probability that a recombination that moves it from a B background
-to a b background occurs in time $dt$ is $dt (1-p(x)) (1+s(x)/2)$.
+A lineage at a locus at genetic distance $r$ that begins linked to a B allele
+moves as a B lineage up until the most recent time
+it recombined from a b background to a B background,
+which occurs when at $x$ with rate $r (1-p(x)) (1+s(x)/2)$.
 The Feynman-Kac formula could then be used to write down differential equations
@@
 
-More simply, consider: within the patch, where $p(x) \approx 1$, 
-a linked lineage moves roughly as Brownian motion, and is unlikely to recombine onto a b background.
-The rate of recombination between B and b backgrounds depends on the proportion of the patch in which $p(x)$ is intermediate.
+Or, more simply, if recombination is on a slower time scale than mixing within the patch,
+then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
+against the stationary distribution.
+Within the patch, $p(x) \approx 1$,
+so recombination between backgrounds is unlikely to occur,
+and so the rate of recombination depends to a first approximation 
+on the proportion of the patch in which $p(x)$ is intermediate.
 This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
@@
 Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
-(and, therefore, the rate of recombination)
-is roughly $r (\sigma/\sqrt{s})A^{-1/d}$.
+is roughly $2(\sigma/\sqrt{s})A^{-1/d}$,
+and so the rate of recombination of a linked lineage to the b background
+is, taking $p(x) \approx 1/2$ in this region, roughly $r (\sigma/\sqrt{s})A^{-1/d}$,
 This will be equal to the coalescence rate if
@@
 \begin{align}
-  \frac{1}{r} = \frac{ \sigma A^{(d-1)/d} }{\sqrt{s}} ,
+  \frac{1}{r} = \frac{ \sigma \rho A^{(d-1)/d} }{\sqrt{s}} ,
 \end{align}

### 2014-05-30 patchy-selection-paper.tex
edited the haplotype sharing section

@@
 
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{sim-transit.pdf}
+  \end{center}
+  \caption{
+  BLAH BLAH WIGGLY LINE
+  \label{fig:lineagesmotion}
+  }
+\end{figure}
+
 %%%%%%%%%
@@
 but there will always be longer LD nearby to the selected site.
-\plr{Could work out on what time scale it gets whittled down and the shared length at stationarity.}
 
@@
 number of generations the allele spends ``in transit'', 
-because each reproductive event in locations where the allele is at low frequency recombines haplotypes of the $b$ allele closer to the adaptive allele.
-(In this case, the haplotype is literally hitchhiking across geography!)
+because each reproductive event in locations where the allele is at
+low frequency recombines haplotypes of the $b$ allele closer to the
+adaptive allele. In this case, the haplotype is literally hitchhiking across geography!
+\gc{In Figure \ref{fig:lineagesmotion} we show the lineage that founds a
+mutant population on a second patch. The entire time that the lineage
+is outside the dark red zone (where the B allele is common) the
+haplotype that accompanies it is being broken down. After the lineage
+establishes on the patch the rate of decay of the haplotype is
+significant slowed as it is again surrounded by other B
+gneotypes with the similar haplotypic backgrounds. We first discuss
+the transit of the lineage, and its accompanying haplotype, and return
+to this point about the whittling down of haplotypes within
+estalbished patches in the next section.}
 
@@
 the $\sigma/\sqrt{s_m}$ is a length scale, and to convert it to a time this is divided by $1/s_m$.
-
+\gc{In the Appendix Section \ref{apx:limiting_dist_of_HH_haplotype} we
+  give some expressions for the limiting distribution of
+  the length of shared haplotype.}
+
+\gc{Graham pasted this to the appendix as it felt like too much information to drop on
+  people}
 %%%
-\subsubsection{Limiting distribution of the transit time and hitchhiking haplotype}
-
-In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
-by taking the large-$R$ limit above we can obtain a more intuitive description of the process
-via a central limit theorem for $\tau$.
-Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
-so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
-Plugging this into equation~\eqref{eqn:haplen_cdf},
-after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
-\begin{align}
-  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
-\end{align}
-Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
-this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
-In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
-Since the standard deviation of $\tau$ is of lower order than its mean,
-a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
-with mean $\sqrt{2s_m} \sigma/R$.
 
@@
 
-\plr{let's leave two dimensions for the revisions. or, as an exercise? =)}
-
 %%%%%%% %%%%%%%%%%
@@
 
-To really make use of the expressions in the previous section,
-we need to know the length of the haplotype shared by individuals \emph{within} a patch.
+The other key issue in thinking about the population genetic signal of
+adaptation to spatial patches is the process by which haplotypes are
+whittled down within patches, which plays into a number of aspects of
+haplotype sharing. Thee initial haplotype of the sweep within a patch will
+diserpate over time as recombination brokes up the haplotype; likewise the haplotype that is shared 
+between patches that have
+adapted through a shared migrant allele
+(eqn. \ref{eqn:haplotype_length}) will also breakdown. 
+We first develop a simple approximation to the rate of
+recombination controling the breakdown of haplotypes within patches. 
+If we wait long enough after the initial sweep,
+either due to migration or mutation, the initial pattern will have
+decayed due to this process of recombination. However, within patches
+we should expect to find a larger extent of haplotype sharing close to
+the selected locus, than in surrounding geographic regions as
+selection against migrant lineage slows the diserpation of
+genetic drift accumulated within the patch \citep{bartonpaper,charlesworths}.
+
+We refer the reader again to Figure \ref{fig:lineagesmotion}. When a lineage in
+established patch is in the dark red area, where the B allele is common, it is very unlikely to
+recombine off the B background. However, when it wanders out of the
+patch into the paler red and white areas when it recombines it will
+usually be on to the b allele background.
+
 A detailed treatment of this is well beyond the scope of this paper,
-but we can get a good idea of the scale through heuristic arguments.
-To do this, we compare the rate of recombination between B and b haplotypes
-to the rate of coalescence of B haplotypes within a patch,
-which we take to be $1/(A\rho)$ 
-(bypassing the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis}).
-
+but we can get a good idea of the genomic scale of these effects through heuristic arguments.
+To understand the probability of a recombination moving our lineage
+off of the B background we need to understand what fraction of the
+time our lineage spends in geographic regions where $b$ in common. 
 Consider the lineage of a single B allele, moving back through time,
-in a patch at mutation-selection balance, 
+in a patch at migration-selection balance, 
 i.e.\ with the local proportion of B alleles near $x$ equal to $p(x)$,
 and that the lineage stays close enough to the patch 
-that fluctuations about $p(x)$ are small (unlike in previous sections).
+that fluctuations about $p(x)$ are small. (Note that this assumption differs from
+previous sections, but here our focus is on a typical lineage in the
+patch rather than the rare lineages that make it far from the patch).
 If we suppose that type B individuals at location $y$ reproduce at rate $1+s(y)$,
@@
 and $B_t$ is a standard Brownian motion.
-This is a diffusion that, when at $x$, 
+This is a diffusion describing the movement of our lineage backward in
+timethat, when at $x$, 
 is pulled toward regions with more B alleles
@@
 \begin{align}
-  \pi(x) = \frac{ p(x)^2 (1+s(x)) }{ \int_y p(y)^2 (1+s(y)) dy } .
-\end{align}
-
+  \pi(x) = \frac{ p(x)^2 (1+s(x)) }{ \int_y p(y)^2 (1+s(y)) dy } . \label{eqn:lineagestatdist}
+\end{align}
 A lineage at a locus at genetic distance $r$ that begins linked to a B allele
@@
 
-Or, more simply, if recombination is on a slower time scale than mixing within the patch,
+As our goal here to describe the rough genomic length scale, 
+based on the above, we view 
+the lineage is simply bouncing around in the patch backward in time and we propose a
+simple heurstic to describe the probability of recombination. 
+If recombination is on a slower time scale than mixing within the patch,
 then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
-against the stationary distribution.
+against the stationary distribution (eqn. \ref{eqn:lineagestatdist}).
 Within the patch, $p(x) \approx 1$,
-so recombination between backgrounds is unlikely to occur,
+so recombination off the $B$ background is unlikely to occur,
 and so the rate of recombination depends to a first approximation 
-on the proportion of the patch in which $p(x)$ is intermediate.
+on the proportion of the patch in which $p(x)$ is intermediate,
+i.e. were both $B$ and $b$ are present.
 This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
@@
 Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
-is roughly $2(\sigma/\sqrt{s})A^{-1/d}$,
-and so the rate of recombination of a linked lineage to the b background
-is, taking $p(x) \approx 1/2$ in this region, roughly $r (\sigma/\sqrt{s})A^{-1/d}$,
-This will be equal to the coalescence rate if
-$1/(A\rho) = r (\sigma/\sqrt{s}) A^{-1/d}$, 
-and hence the length of linkage shared within the patch is on the order of
-\begin{align}
-  \frac{1}{r} = \frac{ \sigma \rho A^{(d-1)/d} }{\sqrt{s}} ,
+is roughly $2(\sigma/\sqrt{s})A^{-1/d}$, and in these regions $p(x)
+\approx 1/2$. Under these approximations the per generation rate of
+recombination of a linked lineage to the b background is, taking in this region, roughly 
+\begin{equation}
+r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
+\end{equation}
+We can use this to first of all think about the length of haplotype
+associated with the $B$ allele after a given length of time $\tau$,
+this could the the time since the sweep within a patch due to a
+mutation or migration of a beneficial allele. If the recombination
+rate per base is $r_{BP}$ then the length of haplotype associated with
+the $B$ allele is
+\begin{equation}
+\frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
+\end{equation} 
+note that the length of haplotype decays at a rate proportional to the
+cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
+opportunity for recombination to move our lineage off the $B$
+background.
+
+Finally we can consider the expected length of haplotype shared within
+a patch. To do this, we compare the rate of recombination between B
+and b haplotypes (eqn. \ref{eqn:rate-rec-in-patch})
+to the rate of coalescence of B haplotypes within a patch. We will
+assume that mixing within the patch is fast compared to the rate of
+drift, so that the rate of coalescence of B alleles within a patch is 
+$1/(A\rho)$ 
+(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis}).
+The rate of recombination and coalescence will be on the same order if
+$\xi /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$, 
+and hence the genetic length of haplotype shared between pairs of
+individuals within the patches is on the order of
+\begin{align}
+  r = \frac{\sqrt{s} \xi}{\sigma \rho A^{(d-1)/d}} 
 \end{align}
 which does not increase with $A$ in one dimension,
-but increases with $\sqrt{A}$ in two dimensions.
+but decreases with $\sqrt{A}$ in two dimensions .
 
@@
 
+\section{Limiting distribution of the transit time and hitchhiking haplotype}
+\label{apx:limiting_dist_of_HH_haplotype}
+
+In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
+by taking the large-$R$ limit above we can obtain a more intuitive description of the process
+via a central limit theorem for $\tau$.
+Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
+so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
+Plugging this into equation~\eqref{eqn:haplen_cdf},
+after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
+\begin{align}
+  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
+\end{align}
+Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
+this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
+In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
+Since the standard deviation of $\tau$ is of lower order than its mean,
+a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
+with mean $\sqrt{2s_m} \sigma/R$.
+
 %%%%%%%%%%%

### 2014-05-30 Lava_flow_mice_app.R

@@ -1,14 +1,18 @@
-char.length= c(3,30)
+char.length= c(3,30)  ## sigma /sqrt(s)
 
 
 A=100
 
-#pdf(file="~/Dropbox/Talks_posters/coopibarra_patchy_2013/figures/prob_parallel_hoekstra_NEW.pdf")
+#pdf(file="~/Dropbox/postdocs/Peter/Spatial_adaptation/Lava_flow_mice_prob_parallel.pdf")
+layout((1:2))
+par(mar=c(3,3,0.2,1))
 R = seq(5,600,length=10000) ##distance in Km
+
 #layout(t(1:2))
 mu = 1e-8
 prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
-plot(R, prob.parallel,type="l",lwd=3, xlab="R, km",ylab="Prob. parallel mutation.",cex.lab=1.5,cex.axis=1.5,log="x") 
+plot(R, prob.parallel,type="l",lwd=3, xlab="R, km",ylab="",cex.lab=1.3,cex.axis=1.3,log="x") 
+mtext(side=2,line=2,"Prob. parallel mutation",cex=1.3)
 mu=1e-5
 prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
 
@@ -27,3 +31,17 @@ lines(R, prob.parallel,lwd=3,col="red",lty=2)
 
 legend("topleft", legend=expression(mu == 10^-8, mu == 10^-5, sigma/sqrt(s[m]) == 3,sigma/sqrt(s[m]) == 30),col=c("black","red","grey","grey"),lty=c(1,1,1,2),lwd=2)
 #dev.off()
+
+
+####haplotype lengths
+##assuming sigma=1
+length.hap.1<-sqrt(2)*1/(char.length[1])/R +1/R^2
+length.hap.2<-sqrt(2)*1/(char.length[2])/R +1/R^2
+plot(R,length.hap.1*100,type="l",log="x",ylab="", xlab="",lty=1,lwd=2,cex.lab=1.3,cex.axis=1.3)
+mtext(side=2,line=2,"Shared Haplotype Length, cM",cex=1.3)
+mtext(side=1,line=2,"Distance between patches, R km",cex=1.3)
+lines(R,length.hap.2*100,lty=2,lwd=2)
+
[…]

### 2014-06-01 patchy-selection-paper.tex
note
edited applications and altered scaled rec rate to reflect ratio of areas

@@
 Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
-is roughly $2(\sigma/\sqrt{s})A^{-1/d}$, and in these regions $p(x)
-\approx 1/2$. Under these approximations the per generation rate of
+multiplied by $p(x)$ in these intermediate regions, where we assume
+$p(x) \approx 1/2$. Under these approximations the per generation rate of
 recombination of a linked lineage to the b background is, taking in this region, roughly 
-\begin{equation}
-r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
-\end{equation}
+\begin{align}
+&=
+    \begin{cases}
+    r \frac{\frac{\sigma}{\sqrt{s}} }{\frac{\sigma}{\sqrt{s}} + A }    \qquad & d=1 \\
+   r \frac{\pi  \frac{\sigma}{\sqrt{s}} \left( \frac{\sigma}{\sqrt{s}} +
+  \sqrt{\frac{A}{\pi}} \right)}{A+\pi  \frac{\sigma}{\sqrt{s}} \left( \frac{\sigma}{\sqrt{s}} +
+  \sqrt{\frac{A}{\pi}} \right) }    \qquad & d=2 \\  
+\end{cases}
+\label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
+\end{align}
+\gc{Altered the above.}
 We can use this to first of all think about the length of haplotype
@@
 To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
-where the adaptive mtation has to occur at a single base pair; 
-and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation being able to arise at any of a thousand base pairs,
-e.g.\ any of the coding bases of a typical gene.  
+where the adaptive mutation has to occur at a single base pair; 
+and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation
+being able to arise at any of a thousand base pairs.  
 Finally, we set $A=100\text{km}^2$.
@@
 likely over a scale of tens to a few hundred kilometers between patches.
+
+Givn the large selection coefficient associated with the dark allele
+on the dark substrate we expect the initial sweep associated with
+either a new mutation or migrant allele to be large ($s/\log(\rho A
+s)$). We can use eqn \eqref{eqn:haplotype_length} to calculate how long a
+founding haplotype we initially expect to find shared between populations
+this in Figure \ref{fig:mice_prob_parallel}. The initial length can be
+quite long between geographically close patches (tens of km). However,
+for our wider cline width ($\sigma/\sqrt{s_m} =30km$) adaptation by
+migration can still be likely for patches 100km apart but its shared
+basis may be hard to detect as the length of shared haplotype can be
+quite short.
+
+Given the relatively small size of the dark substrate patches where
+$A$ is often $<10km$ we should expect the haplotype to be rapidly
+whittled down within patches (eqn. \ref{})
 
@@
 
+recessive alleles will have different dynamics. Actually not really
+obvious what will happen here, cline width will be much wider but
+prob. of taking off will go as the square of the occupancy density.
+
 discussion of applications.
@@
 small-scale fluctuations
-
 when landscape is not made of discrete (circular) patches

### 2014-06-02 patchy-selection-paper.tex

@@
 recombination of a linked lineage to the b background is, taking in this region, roughly 
+% \begin{equation}
+% r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
+% \end{equation}
 \begin{align}
@@
 \end{align}
-\gc{Altered the above.}
 We can use this to first of all think about the length of haplotype

### 2014-06-02 patchy-selection-paper.tex
added shared lineage

@@
 \label{ss:patchymigration}
-\gc{There's a bunch of switching between x and r in this section}
+
 Now suppose that there are two patches, each of area $A$ and separated by distance $R$. 
@@
 in the process obtaining other useful information about the genealogy of B alleles.
-
-\gc{Could you make the CI or sample path in Figure 1B be different colors, the grey and black are too close to each other.}
-\plr{how's that?  I'm not sure what to do about the grey; the real problem is that the CIs are too small. If I ran them for less time they would be\dots}
 
@@
 and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.
-\gc{Is there no 1+ in the equation above for the main branch? } \plr{oops; yep.}
 
@@
 \end{align}
-\gc{define $\gamma(x)$ presumably the outflux of branches leaving ball}
-\plr{it was defined at the start of this paragraph.}
 
@@
   BLAH BLAH WIGGLY LINE
+  \plr{can we please keep that caption? ok, fine\dots}
+  \textbf{Adaptation by migration:}
+  An initially unadapted patch (10 demes at the left end of the range) is colonized around generation 5000;
+  a representative lineage from the new patch is shown transiting between the two and coalescing with a lineage
+  in the unadapted patch.
+  As in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
+  other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
   \label{fig:lineagesmotion}

### 2014-06-02 patchy-sim-writeup-plots.R

@@ -123,6 +123,7 @@ dimension <- sum(dim(pophist$pophist)[1:2]>1)
 
 
 pdf(file="sim-transit.pdf",width=5,height=3,pointsize=10)
+layout(t(1:2),width=c(3,1))
 par(mar=c(5,4,1,1)+.1)
 # make a copy of the slice we want
 tmphist <- pophist
@@ -151,6 +152,28 @@ LL <- plotlins(lins,plotit=FALSE)
 ltimes <- seq(from=1,to=dim(LL)[2],length.out=250)
 lcols <- rep("black",length(linit)) # terrain_hcl(length(linit))
 invisible( lapply( 1:dim(LL)[3], function (k) { lines( LL[2,ltimes,k], ltimes, col=lcols[k]) } ) )
+# simulate shared haplotype length
+localprops <- lins$localsizes[,,,drop=TRUE] / tmphist$pop$params$N
+recombs <- list( matrix( rexp(prod(dim(localprops))), nrow=nrow(localprops) ),
+        matrix( rexp(prod(dim(localprops))), nrow=nrow(localprops) ) )
+coaltime <- lins$coalevents$t[2]
+visible <- list( matrix( rbinom(prod(dim(localprops)),size=1, prob=ifelse(seq_along(localprops)>coaltime,localprops[,1],0) ), nrow=nrow(localprops) ),
+        matrix( rbinom(prod(dim(localprops)),size=1, prob=ifelse(seq_along(localprops)>coaltime,localprops[,2],0) ), nrow=nrow(localprops) ) )
+for (k in 1:2) {
+    recombs[[k]][!visible[[k]]] <- Inf
+    recombs[[k]] <- cummin( recombs[[k]] )
+}
+par(mar=c(5,1,1,1)+.1)
+changepoints <- lapply( recombs, function (x) { z <- sort(unique(c(1,outer(which(diff(x)!=0),c(1,0),"+")),length(x))); z[z>0] } )
+x1 <- (-1)*recombs[[1]][changepoints[[1]]]
+x2 <- rev(recombs[[2]][changepoints[[2]]])
+xx <- c( -1, x1, max(x1), min(x2), x2, 1 )
+y1 <- changepoints[[1]]
+y2 <- rev(changepoints[[2]])
+yy <- c(1, y1, length(recombs[[1]]), length(recombs[[1]]), y2 , 1)
[…]

### 2014-06-02 patchy-selection-paper.tex
edited haplotype length section; added haplotypes to figure

@@
   in the unadapted patch.
-  As in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
+  In the left panel, as in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
   other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
+  The right panel shows (in red) the length of the shared haplotype between the two lineages
+  around the selected locus,
+  which decreases as time progresses.
   \label{fig:lineagesmotion}
@@
 If a patch adapts through new mutation or a rare migrant lineage, the
-genomic region surrounding the selected site will hitchhike along with
-it \citep{maynardsmith1974hitchhiking},
+genomic region surrounding the selected site will hitchhike along with it \citep{maynardsmith1974hitchhiking},
 so that at least initially, all adapted individuals within a patch
@@
 Pairs of individuals within a patch will share a haplotype of mean genetic length of
-$\approx \log(\rho A s_p)/s_p$ around the selected site, 
+about $\log(\rho A s_p)/s_p$ around the selected site, 
 as long as the patch is reasonably well mixed by dispersal
@@
 because each reproductive event in locations where the allele is at
-low frequency recombines haplotypes of the $b$ allele closer to the
-adaptive allele. In this case, the haplotype is literally hitchhiking across geography!
-\gc{In Figure \ref{fig:lineagesmotion} we show the lineage that founds a
-mutant population on a second patch. The entire time that the lineage
-is outside the dark red zone (where the B allele is common) the
-haplotype that accompanies it is being broken down. After the lineage
-establishes on the patch the rate of decay of the haplotype is
-significant slowed as it is again surrounded by other B
-gneotypes with the similar haplotypic backgrounds. We first discuss
-the transit of the lineage, and its accompanying haplotype, and return
-to this point about the whittling down of haplotypes within
-estalbished patches in the next section.}
+low frequency will likely recombine haplotypes of the $b$ allele closer to the adaptive allele. 
+In this case, the haplotype is literally hitchhiking across geography!
+Figure~\ref{fig:lineagesmotion} shows a simulation of such an event,
+including the lineage that founds an adaptive population on a second patch,
+and the length of the haplotype shared between this lineage and one in the original patch.
+The entire time that the lineage is outside the region where the B allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
+the haplotype that accompanies it is broken down rapidly. 
+After the lineage establishes on the patch, the rate of decay of the haplotype is slowed significantly, 
+since most others with which it recombines have similar haplotypic backgrounds. 
+We first discuss the transit of the lineage, and its accompanying haplotype, and return
+to how haplotypes are whittled down within estalbished patches in the next section.
 
@@
 and $\var[L] = 2s_m\sigma^2/R^2 + 4 \sigma^3 \sqrt{2s_m}/R^3 + 5 \sigma^4 / R^4$.
-For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
+For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior;
 but how other properties might change is unclear.
@@
 is roughly the length one would expect given the mean transit time.
-Note that we have become used to seeing $\sigma$ divided by $\sqrt{s_m}$, rather than multiplied,
-the $\sigma/\sqrt{s_m}$ is a length scale, and to convert it to a time this is divided by $1/s_m$.
-\gc{In the Appendix Section \ref{apx:limiting_dist_of_HH_haplotype} we
-  give some expressions for the limiting distribution of
-  the length of shared haplotype.}
-
-\gc{Graham pasted this to the appendix as it felt like too much information to drop on
-  people}
+Note that we have become used to seeing $\sigma$ divided by $\sqrt{s_m}$, rather than multiplied;
+it appears here because $\sigma/\sqrt{s_m}$ is a length scale, 
+and to convert it to a time this is divided by $1/s_m$.
+This intuitive picture can be expanded by a central limit theorem:
+for large $R$, the deviation of $\tau$ about its mean is asymptotically Gaussian,
+and the haplotype halflength $L$ is exponential.
+
+% In Appendix~\ref{apx:limiting_dist_of_HH_haplotype}, 
+% we also give some expressions for the limiting distribution of the length of shared haplotype.
+
+\gc{Graham pasted this to the appendix as it felt like too much information to drop on people}
+\plr{I am sympathetic, but it's only a paragraph\dots how about that last sentence, instad?}
 %%%
+% 
+% \section{Limiting distribution of the transit time and hitchhiking haplotype}
+% \label{apx:limiting_dist_of_HH_haplotype}
+% 
+% In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
+% also taking the large-$R$ limit, obtains a more intuitive description of the process
+% via a central limit theorem for $\tau$.
+% Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
+% so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
+% Plugging this into equation~\eqref{eqn:haplen_cdf},
+% after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
+% \begin{align}
+%   \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
+% \end{align}
+% Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
+% this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
+% In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
+% Since the standard deviation of $\tau$ is of lower order than its mean,
+% a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
+% with mean $\sqrt{2s_m} \sigma/R$.
 
@@
 
-The other key issue in thinking about the population genetic signal of
-adaptation to spatial patches is the process by which haplotypes are
-whittled down within patches, which plays into a number of aspects of
-haplotype sharing. Thee initial haplotype of the sweep within a patch will
-diserpate over time as recombination brokes up the haplotype; likewise the haplotype that is shared 
-between patches that have
-adapted through a shared migrant allele
-(eqn. \ref{eqn:haplotype_length}) will also breakdown. 
-We first develop a simple approximation to the rate of
-recombination controling the breakdown of haplotypes within patches. 
-If we wait long enough after the initial sweep,
-either due to migration or mutation, the initial pattern will have
-decayed due to this process of recombination. However, within patches
-we should expect to find a larger extent of haplotype sharing close to
-the selected locus, than in surrounding geographic regions as
-selection against migrant lineage slows the diserpation of
-genetic drift accumulated within the patch \citep{bartonpaper,charlesworths}.
-
-We refer the reader again to Figure \ref{fig:lineagesmotion}. When a lineage in
-established patch is in the dark red area, where the B allele is common, it is very unlikely to
-recombine off the B background. However, when it wanders out of the
-patch into the paler red and white areas when it recombines it will
-usually be on to the b allele background.
+The other key issue in thinking about the population genetic signal of adaptation in spatial patches 
+is the process by which haplotypes are whittled down within patches, 
+which plays into a number of aspects of haplotype sharing. 
+The initial haplotype that sweeps within a patch 
+will be dispersed over time by recombination;
+likewise, the haplotype that is shared between patches coadapted by migration
+(equation~\ref{eqn:haplotype_length}) will also break down. 
+We first develop a simple approximation to the rate 
+at which recombination breaks down haplotypes within patches. 
+If we wait long enough after the initial adaptation of a patch,
+the haplotype that swept initially will have decayed due to this process of recombination. 
+However, we still expect to find individuals within the patch sharing long haplotypes around the selected locus
+than with individuals elsewhere,
+since selection against migrants 
+decreases mean coalescence times within the patch near the selected locus.
+\citep{bartonpaper,charlesworths}.
+
+Refer again to Figure~\ref{fig:lineagesmotion}. 
+When a lineage is in a region with high frequency of B alleles (dark red), 
+it is very unlikely to recombine off the B background. 
+However, when it wanders out of the patch,
+into regions with lower frequency of B alleles,
+when it recombines it will
+usually be on to the background of the b allele.
 
@@
 and that the lineage stays close enough to the patch 
-that fluctuations about $p(x)$ are small. (Note that this assumption differs from
+that fluctuations about $p(x)$ are small. 
+(Note that this assumption differs from
 previous sections, but here our focus is on a typical lineage in the
@@
 The Feynman-Kac formula could then be used to write down differential equations
-solved by various moments of the haplotype length as a function of the sampling location of the lineage.
-
-As our goal here to describe the rough genomic length scale, 
-based on the above, we view 
-the lineage is simply bouncing around in the patch backward in time and we propose a
-simple heurstic to describe the probability of recombination. 
+solved by various moments of the haplotype length as a function of the sampling location of the lineage;
+instead, we only use the above as a guide for rough approximation.
+
+Based on this picture of
+the lineage bouncing around in the patch,
+we propose a simple heurstic to describe the probability of recombination. 
 If recombination is on a slower time scale than mixing within the patch,
 then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
-against the stationary distribution (eqn. \ref{eqn:lineagestatdist}).
+against the stationary distribution (equation~\ref{eqn:lineagestatdist}).
 Within the patch, $p(x) \approx 1$,
@@
 on the proportion of the patch in which $p(x)$ is intermediate,
-i.e. were both $B$ and $b$ are present.
+i.e.\ where both $B$ and $b$ are present.
 This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
@@
 $p(x) \approx 1/2$. Under these approximations the per generation rate of
-recombination of a linked lineage to the b background is, taking in this region, roughly 
+recombination of a linked lineage to the b background is
+reduced by a factor roughly equal to the proportion of the patch lying within distance $\sigma/\sqrt{s}$ of the boundary,
+i.e.
 % \begin{equation}
@@
 % \end{equation}
-\begin{align}
-&=
+\plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
+and you shrink it to have radius $r_0$,
+then the resulting object has area $A \times (r_0/r)^d$;
+here we have $r_0 = r - \sigma/\sqrt{s}$,
+in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
+\begin{align}
+  r_\text{eff} &=
     \begin{cases}
-    r \frac{\frac{\sigma}{\sqrt{s}} }{\frac{\sigma}{\sqrt{s}} + A }    \qquad & d=1 \\
-   r \frac{\pi  \frac{\sigma}{\sqrt{s}} \left( \frac{\sigma}{\sqrt{s}} +
-  \sqrt{\frac{A}{\pi}} \right)}{A+\pi  \frac{\sigma}{\sqrt{s}} \left( \frac{\sigma}{\sqrt{s}} +
-  \sqrt{\frac{A}{\pi}} \right) }    \qquad & d=2 \\  
-\end{cases}
+      r \frac{2\sigma}{A\sqrt{s}}    \qquad & d=1 \\
+      r \frac{\sqrt{\pi} \sigma }{ \sqrt{ s A } } \qquad & d=2 \\  
+\end{cases} ,
 \label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
 \end{align}
-We can use this to first of all think about the length of haplotype
-associated with the $B$ allele after a given length of time $\tau$,
-this could the the time since the sweep within a patch due to a
-mutation or migration of a beneficial allele. If the recombination
-rate per base is $r_{BP}$ then the length of haplotype associated with
-the $B$ allele is
+i.e.\ the reduction in recombination rate is proportional to $(\sigma/\sqrt{s})A^{-1/d}$.
+We can use this to think about the length of the haplotype
+associated with the $B$ allele after a given length of time $T$,
+for instance, the time since the $B$ allele joined the patch.
+\plr{changed this to genetic lengths so as not to introduce another parameter, $r_{BP}$}
+% If the recombination rate per base is $r_{BP}$, 
+% then the length of the haplotype associated with the $B$ allele is
+% \begin{equation}
+% \frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
+% \end{equation} 
+In units of genetic length (Morgans),
+the length of the haplotype associated with the $B$ allele is, up to a constant,
 \begin{equation}
-\frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
+  \frac{A^{1/d}\sqrt{s}}{\sigma T}
 \end{equation} 
-note that the length of haplotype decays at a rate proportional to the
+Note that the length of haplotype decays at a rate proportional to the
 cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
-opportunity for recombination to move our lineage off the $B$
-background.
-
-Finally we can consider the expected length of haplotype shared within
-a patch. To do this, we compare the rate of recombination between B
-and b haplotypes (eqn. \ref{eqn:rate-rec-in-patch})
-to the rate of coalescence of B haplotypes within a patch. We will
-assume that mixing within the patch is fast compared to the rate of
-drift, so that the rate of coalescence of B alleles within a patch is 
-$1/(A\rho)$ 
+opportunity for recombination to move our lineage off the $B$ background.
+
+Finally we can consider the expected length of haplotype shared within a patch. 
+To do this, we compare the rate of recombination between B and b haplotypes (equation~\ref{eqn:rate-rec-in-patch})
+to the rate of coalescence of B haplotypes within a patch. 
+We will assume that mixing within the patch is fast compared to the rate of drift, 
+so that the rate of coalescence of B alleles within a patch is 
+$\xi^2/(A\rho)$,
+the ``variance effective population size'' of the patch
 (this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis}).
 The rate of recombination and coalescence will be on the same order if
-$\xi /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$, 
+$\xi^2 /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$, 
 and hence the genetic length of haplotype shared between pairs of
@@
 \begin{align}
-  r = \frac{\sqrt{s} \xi}{\sigma \rho A^{(d-1)/d}} 
-\end{align}
-which does not increase with $A$ in one dimension,
-but decreases with $\sqrt{A}$ in two dimensions .
+  r = \frac{\sqrt{s} \xi^2}{\sigma \rho A^{(d-1)/d}} 
+\end{align}
+which does not depend on $A$ in one dimension,
+but decreases with $\sqrt{A}$ in two dimensions.
 
@@
 
-\section{Limiting distribution of the transit time and hitchhiking haplotype}
-\label{apx:limiting_dist_of_HH_haplotype}
-
-In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
-by taking the large-$R$ limit above we can obtain a more intuitive description of the process
-via a central limit theorem for $\tau$.
-Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
-so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
-Plugging this into equation~\eqref{eqn:haplen_cdf},
-after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
-\begin{align}
-  \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
-\end{align}
-Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
-this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
-In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
-Since the standard deviation of $\tau$ is of lower order than its mean,
-a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
-with mean $\sqrt{2s_m} \sigma/R$.
-
 %%%%%%%%%%%

### 2014-06-02 patchy-sim-writeup-plots.R

@@ -153,7 +153,7 @@ ltimes <- seq(from=1,to=dim(LL)[2],length.out=250)
 lcols <- rep("black",length(linit)) # terrain_hcl(length(linit))
 invisible( lapply( 1:dim(LL)[3], function (k) { lines( LL[2,ltimes,k], ltimes, col=lcols[k]) } ) )
 # simulate shared haplotype length
-localprops <- lins$localsizes[,,,drop=TRUE] / tmphist$pop$params$N
+localprops <- 1-lins$localsizes[,,,drop=TRUE] / tmphist$pop$params$N
 recombs <- list( matrix( rexp(prod(dim(localprops))), nrow=nrow(localprops) ),
         matrix( rexp(prod(dim(localprops))), nrow=nrow(localprops) ) )
 coaltime <- lins$coalevents$t[2]

### 2014-06-27 patchy-selection-paper.tex
changed back to my version of the decrease in rec
wrote chunk of discussion

@@
 
-\gc{Graham pasted this to the appendix as it felt like too much information to drop on people}
-\plr{I am sympathetic, but it's only a paragraph\dots how about that last sentence, instad?}
-%%%
 % 
@@
 \begin{align}
-  r_\text{eff} &=
+r_{eff} &=
     \begin{cases}
-      r \frac{2\sigma}{A\sqrt{s}}    \qquad & d=1 \\
-      r \frac{\sqrt{\pi} \sigma }{ \sqrt{ s A } } \qquad & d=2 \\  
-\end{cases} ,
+    r \frac{ \sigma /\sqrt{s} }{ \sigma/\sqrt{s} + A }    \qquad & d=1 \\
+   r \frac{\pi  \sigma /\sqrt{s}  \left( \sigma /\sqrt{s}+
+  \sqrt{A/\pi} \right)}{A+\pi \sigma/\sqrt{s} \left( \sigma/\sqrt{s} +
+  \sqrt{A/\pi} \right) }    \qquad & d=2 \\  
+\end{cases}
 \label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
 \end{align}
-i.e.\ the reduction in recombination rate is proportional to $(\sigma/\sqrt{s})A^{-1/d}$.
+\gc{Altered the above.}
+which scales the recombination rate roughly with $(\sigma/\sqrt{s})A^{-1/d}$.
 We can use this to think about the length of the haplotype
@@
 \section{Discussion}
+
+We have developed a simple set of approximations to describe the rate
+at which organisms in a patchy environment adapt through novel
+mutations or migration from other nearby patches (eqn.
+\eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
+probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
+We have also determined the characteristic genomic length scales over which
+the various signals of hitchhiking should be visable
+(eqn. \eqref{eqn:haplotype_length} and ).
+
+These results show that the probability of parallel adaptation is
+independent of both the population density and the beneficial
+selection coefficient. This follows from the fact that these
+parameters both increase the rate of adaptation via both migration and
+mutation equally. Thus they are important for the time-scale over
+which adaptation occurs but not the mode of adaptation. This
+constrasts with models of convergent adaptation within panmictic
+populations \citep{} and geographically spread populations adapting to
+homogeneous populations \citep{RalphCoop}, where the number of
+adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
+coefficient is absent or weak, due to the fact that selection both
+aids establiishment of new alleles and the spread of existing
+alleles.
+
+The primary determinant of the probability of parallel adaptation is
+$\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
+allele frequency decay off the patch. We find that patches that are close together on
+this scale will adapt via migration, but there will be a sharp
+transition to convergent adaptation as we move to patches that are further apart.
+
+In coming to this conclusion, we have ignored the possibility of very long distance migration,
+instead focusing local dispersal (as parameterized by a Gaussian
+dispersal model). However, dispersal distributions can be very heavy
+tailed, with a small fraction of individuals moving very long distances indeed, in
+addition over long time-scales very rare chance events (mice sucked up in
+hurricanes etc) could play a role in spreading migrant alleles. Such
+tail events could greatly increase the probability of parallel
+adaptation above that predicted by our model. Furthermore, if
+adaptive alleles do move between distant patches via rare long distance
+migration they will be associated with a much longer shared haplotype than
+predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
+results as a null model by which the contribution of long distribution
+migrants to adaptation could be empirically judged.
+
+We have also assumed that the alleles have a fitness effect in
+heterozygotes both on and off the patch. However, 
+ recessive alleles could also contribute to adaptation to patchy
+ environments. \gc{should we cite Nachman eg here}.  Our
+ simple branching process framework does not easily accomodate recessive
+ alleles as it ignores the selection interaction of alleles, and so a
+ different approach will be needed. At first sight it
+ seems like recessive alleles may be able to contribute to adaptation
+ via migration over longer distances, as the migrant alleles will be at low
+ frequency between patches and so usually invisible to
+ selection. However, this fact also means that they are unlikely to
+ establish on the beneficial patch upon reaching it via migration. As
+ such it is not clear what our intuition should be about the
+ contribution of recessive alleles to adaptation via migration.
+
+\gc{will write a couple more broader paragraphs}
+
+The falling cost of sequencing over the next decade offers the
+opportunity to investigate the genomics of parallel adaptation over a
+large s
+Over the coming decade population genomic data will be generated
 

### 2014-06-27 Lava_flow_mice_app.R

@@ -30,6 +30,8 @@ prob.parallel <- mu*A/(mu*A + (1/char.length[2])* exp(-R/char.length[2]))  ##ass
 lines(R, prob.parallel,lwd=3,col="red",lty=2)
 
 legend("topleft", legend=expression(mu == 10^-8, mu == 10^-5, sigma/sqrt(s[m]) == 3,sigma/sqrt(s[m]) == 30),col=c("black","red","grey","grey"),lty=c(1,1,1,2),lwd=2)
+
+mtext(side=1,line=2,"Distance between patches, r km",cex=1.3)
 #dev.off()
 
 
@@ -45,3 +47,52 @@ lines(R,length.hap.2*100,lty=2,lwd=2)
 legend("topright", legend=expression( sigma/sqrt(s[m]) == 3,sigma/sqrt(s[m]) == 30),col=c("grey","grey"),lty=c(1,2),lwd=2)
 
 dev.copy2pdf(file="~/Dropbox/postdocs/Peter/Spatial_adaptation/Lava_flow_mice_prob_parallel.pdf")
+
+##
+
+
+
+
+######version of graphs for SMBE
+
+R = seq(5,600,length=10000) ##distance in Km
+
+#layout(t(1:2))
+mu = 1e-8
+prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
+plot(R, prob.parallel,type="l",lwd=3, xlab="R, km",ylab="",cex.lab=1.3,cex.axis=1.3,log="x") 
+mtext(side=2,line=2,"Prob. parallel mutation",cex=1.3)
+mu=1e-5
+prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
+
+lines(R, prob.parallel,lwd=3,col="red")
+
+
+R = seq(5,600,length=1000) ##distance in Km
+mu = 1e-8
+prob.parallel <- mu*A/(mu*A + (1/char.length[2])* exp(-R/char.length[2]))  ##assuming sigma=1
+lines(R, prob.parallel,lty=2,lwd=3)
+
+mu=1e-5
+prob.parallel <- mu*A/(mu*A + (1/char.length[2])* exp(-R/char.length[2]))  ##assuming sigma=1
+
+lines(R, prob.parallel,lwd=3,col="red",lty=2)
+
[…]

### 2014-06-28 patchy-selection-paper.tex
rearranged appendices; began discussion

--- 2014-06-02 patchy-selection-paper.tex
@@
 
-Expected transit time \label{eqn:mean_tau}
-Expected haplotype length \label{eqn:haplotype_length}
+Expected transit time \eqref{eqn:mean_tau}
+Expected haplotype length \eqref{eqn:haplotype_length}
 
@@
 \end{align}
-i.e.\ the product of mutational influx in the patch ($A\rho\mu$) and the probability of establishment ($2s_p/\xi^2$).
+i.e.\ the product of mutational influx in the patch ($\rho A\mu$) and the probability of establishment ($2s_p/\xi^2$).
 If this rate is low, then the time (in generations) until a mutation arises that
@@
 we would like to know the time scale on which copies of $B$ move between patches by migration
-(so, ignoring the possibility that a $B$ allele arises independently by mutation in the first patch).
+(supposing that no $B$ allele arises independently by mutation in the meantime).
 To determine this, we dissect the process by which an allele transits between the patches,
@@
 and show that $q(x)$ decays exponentially for large $|x|$,
-with a square-root correction in two dimensions (see Appendix \ref{apx:asymptotics}): 
+with a square-root correction in two dimensions:
 \begin{align} \label{eqn:eqfreq}
@@
 \end{align}
-where $d$ is the dimension ($d=1$ or $2$), and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients.
+where $d$ is the dimension ($d=1$ or $2$), 
+and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients.
 These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
-(In one dimension, the equation can be solved to give $q(x)$ in terms of Jacobi elliptic functions; see Appendix \ref{apx:elliptic_integrals}.)
+To be clear about the assumptions (especially regarding scaling) implicit here,
+we provide a derivation in Appendix~\ref{apx:eqfreq}; 
+as well as justification for the asymptotics in Appendix \ref{apx:asymptotics}.
+In one dimension, the equation can be solved to give $q(x)$ 
+in terms of Jacobi elliptic functions; see Appendix \ref{apx:elliptic_integrals}.
 In applications we fit $C$ to the data;
 to get concrete numbers from theory we take $C=1$ if necessary.
-From simulations, $C$ seems to be not too far from 1 in practice (see Appendix \ref{apx:patchy_sims}).
+From simulations, $C$ seems to be not too far from 1 in practice (see below).
 
@@
   with $C$ chosen arbitrarily to give a reasonable fit.
-  (The uptick at the right end is due to boundary effects near the edge of the range.)
+  (The disagreement at long distances is due to boundary effects near the edge of the range.)
   \textbf{(B)} 
@@
   dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
- \gc{The approximation and simulations disagree at long distances from
-   the patch, but this is due to edge effects in the simulations.} }   \label{fig:sim_occupation_freqs}
+   }   \label{fig:sim_occupation_freqs}
 \end{figure}
@@
 Properties of these, combined with equation \eqref{eqn:eqfreq},
-leads us to the expression \eqref{eqn:migrate} for the rate of effective migration,
+will lead us eventually to the expression \eqref{eqn:migrate} for the rate of effective migration,
 an approximation that is valid if the distance $R$ is large (at least five times $\sigma/\sqrt{2s_m}$, say).
@@
 and the variance is $\xi^2$.
-\plr{I changed this since equation \eqref{eqn:eqfreq} is in the small-$s_m$ limit, 
-and therefore doesn't differentiate between $s_m$ and $\log(1/(1-s_m))$;
-doing it this way means we don't have to define another parameter.}
 This pruning removes both branches leading back into the already established patch (which are irrelevant)
@@
 \end{align}
-\gc{do we need to talk about u a little more it }
-\plr{not sure what to say}
+Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
 Since reproduction and spatial motion are independent,
@@
 
+\paragraph{Island model}
+The resulting discrete process is quite general,
+but if we oversimplify back to the island model, it is tractable:
+suppose the probability of adaptaton by mutation is the same for each patch
+and by migration is the same for each pair of patches.
+The resulting process is a continuous-time version of the ``Chinese restaurant process''
+described in \citet{aldous1985exchangeability} and \citet{pitman1995partitions},
+and so the final partition of types across demes has the Ewens sampling distribution with parameter $\mu/m$.
+
+Now we can compute most properties we might want about the process.
+For instance, the proportion of demes that shares the same origin as a randomly sampled deme
+is approximately Beta distributed -- see \cite{donnelly1989continuity} and \cite{perman1992sizebiased}.
+The high connectedness of the discrete deme island model means that the expected number of distinct alleles
+grows with the $\log$ of the number of demes, 
+which strongly contrasts with the continuous spatial model 
+where the local nature of dispersal means that doubling the species range will double the number of mutations expected. 
+Furthermore, while in the continuous model the areas occupied by distinct mutations are similar in area, 
+in this strong-selection discrete island model, 
+only a few types tend to dominate even as the number of demes (and thus, distinct types) grows large.
+
 \begin{figure}[ht]
@@
   \caption{
-  BLAH BLAH WIGGLY LINE
-  \plr{can we please keep that caption? ok, fine\dots}
   \textbf{Adaptation by migration:}
@@
 
+%%%%%%% %%%%%%% %%%%%%%
 \section{Discussion} 
-
-\plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
+\label{ss:discussion}
+
+This paper is an investigation into the basic, ill-formed question: 
+What is the spatial resolution of adaptation?
+In other words, what scale of environmental patchiness can a species develop and sustain local adaptations,
+and over what spatial scale will evolution develop independent solutions to evolutionary problems?
+The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
+the dispersal distance divided by the square root of the selection against the new allele between patches,
+and much more weakly on the selective benefit within the patches or (perhaps suprisingly) the population density.
+Weaker selection against alleles will make sharing of adaptations between patches easier,
+but will also make it harder to spot,
+since the lucky, transiting alleles move slower and carry a shorter haplotype.
+
+Observations of similar adaptations occuring within the same species in widely separated patches
+is strongly suggestive of parallel, independent adaptations.
+Such observations are very interesting 
+as they tell us about the degree of constraint faced by evolution --
+how many different methods are there typically available to a species
+with which to solve an evolutionary problem?
+
+It is already well-appreciated that larger populations can lead to more parallelism \citep{softsweepsI}
+\plr{fisher probably said this too?}
+and that spatial structure contributes to this \citep{ralph2010parallel}.
+Naturally, then, local adaptation to well-separated patches of some environment is even more likely to occur in parallel;
+we have studied the case where patches are separated by long distances.
+This is complementary to the work of \citet{XXX}, who studied local adaptation in rapidly varying environments.
+Intermediate cases where habitat types are distributed in some more complex geometry
+-- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance --
+remains to be studied.
+
+To complement this theory, we need more empirical work,
+in which the parameters (migration rate, sensitivity to environmental factors, etc)
+are reasonably well-understood.
+However, learning about these aspects is difficult in the field,
+and it is still not understood how much of the intraspecific variation we observe
+is due to adaptation to local conditions, or even how to tell.
+Notable steps forward include \citet{turner2010serpentine}, Mimulus, Amity, others?
+
+\paragraph{Standing variation} 
+We have tended to focus on relative rates of adaptation,
+since we are thinking of applications in which we know that adaptation has occurred,
+and the question is whether or not adaptations in distinct patches
+have adapted indpendently or not.
+However, especially if rates of adaptation by new mutation are very low,
+any adaptation that does occur may have to make use of standing variation.
+The unstructured case was studied by \citet{softsweepsI}
+and we study the continuous, spatial case in \citet{toappear}.
+If local, parallel adaptation is from standing variation,
+then the dynamics of adaptation should not depend strongly on migration patterns
+(but note that the spatial distribution of standing variation will be).
+
+\paragraph{Dominance}
+We have mostly swept the issue of dominance under the rug
+by dealing with essentially haploid models,
+and appealing to the fact that the dynamics we study occur where the mutation is rare,
+and hence mostly present only in heterozygotes.
+If, however, the mutation is recessive, then it is essentially neutral where rare,
+and so could have a much easier time spreading between patches.
+This is counteracted, however, by the increased difficulty with which the mutation would establish
+once arriving in a new patch.
+On the other hand, a dominant mutation would not differ greatly from the theory presented here.
+In fact, MC1r is ?dominant? and agouti is ?recessive?
+Other schemes could be obtained.
+
+optimal $s_m$ for migration makes migration harder to spot
 
@@
 \appendix
+
+%%%%%
+\section{Numerical calculation of the probability of establishment}
+\label{apx:establishment_sims}
+
+To investigate the impact of the various approximations surrounding the mutational influx,
+we carried out numerical calculations of a fine-grained, discrete model.
+These are shown, and described, in figure~\ref{fig:prob_estab_calcs}.
+
+Here (and at other parameter choices) we see that the probability of establishment $p(x)$ goes to the equilibrium value
+(approximately $p_e = 2s/\xi^2$) within the patch;
+the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
+This lends credence to our approximation that the integral of $p(x)$ over the entire range
+is close to $p_e$ multiplied by the area of the patch.
+
+\begin{figure}[ht!]
+    \begin{center}
+        % made in patchy-selection-plots.R
+        \includegraphics{prob-establishment}
+    \end{center}
+    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
+    in an abrupt transition (left) and a gradual transition (right).
+    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respecively);
+    and the selection coefficient interpolates linearly between these in the central shaded region.
+    The number of offspring is Poisson.
+    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
+    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
+    \label{fig:prob_estab_calcs}
+    }
+\end{figure}
 
@@
 
-%%%%%
-\section{Numerical calculation of the probability of establishment}
-\label{apx:establishment_sims}
-
-To investigate the impact of the various approximations surrounding the mutational influx,
-we carried out numerical calculations of a fine-grained, discrete model.
-These are shown, and described, in figure~\ref{fig:prob_estab_calcs}.
-
-Here (and at other parameter choices) we see that the probability of establishment $p(x)$ goes to the equilibrium value
-(approximately $p_e = 2s/\xi^2$) within the patch;
-the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
-This lends credence to our approximation that the integral of $p(x)$ over the entire range
-is close to $p_e$ multiplied by the area of the patch.
-
-\begin{figure}[ht!]
-    \begin{center}
-        % made in patchy-selection-plots.R
-        \includegraphics{prob-establishment}
-    \end{center}
-    \label{fig:prob_estab_calcs}
-    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
-    in an abrupt transition (left) and a gradual transition (right).
-    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp .05$ respecively);
-    and the selection coefficient interpolates linearly between these in the central shaded region.
-    The number of offspring is Poisson.
-    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
-    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
-    }
-\end{figure}
+%%%%%%%%% %%%%%%%%%%%%%%
+\section{Hitting and occupation}
+\label{apx:hitting_occupation}
+
+Consider a single migrant family $Z$ beginning at $x$.
+Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
+and $\tau_\dagger$ an independent Exponential($\alpha$) time.
+The mean occupation time of $Z$ spent in a region $A$ is 
+\begin{align}
+    u(x,y) &= \E[\int_0^\infty Z_t(A) dt ] \\
+           &= \int_0^\infty \E[Z_t] p_t(x,A) dt \\
+        &= \int_0^\infty e^{-\alpha t} p_t(x,A) dt \\
+        &= \int_0^\infty \P^x\{ \tau_\dagger < t \; \& \; B_t \in A \} dt \\
+        &= \E[ \int_0^{\tau_\dagger} \one_A(B_t) dt ]  .
+\end{align}
+If we define
+$\tau_A$ to be the hitting time of $A$ by $B$,
+and $\mu(x,A)$ to be the hitting distribution of $\partial A$ by $B_{\min(t,\tau)}$,
+by the strong Markov property this is equal to
+\begin{align}
+    u(x,y) &= \E^x[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] \\
+           &= \P^x\{ \tau_A < \tau_\dagger \}  \E^{\mu(x,A)}[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] .
+\end{align}
+If $A$ is circular, the latter expectation will not depend on $x$;
+call this $g(A)$.
+
+Now consider the chance that $Z$ ever hits $A$.
+If we make the approximation that $Z$ hits $A$ iff the trunk of $Z$ does,
+then this is equal to
+\begin{align}
+    F(x,y) &= \int_0^\infty (1-k_e(t)) \P\{ \tau_A \in dt \} \\
+           &= \int_0^\infty e^{\alpha t} (1-k_e(t)) e^{- \alpha t } \P\{ \tau_A < t \} dt \\
+    &=\int_0^\infty e^{\alpha t} (1-k_e(t)) \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &= \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
+    &= \frac{1}{\E[K]} \P^x[ \tau_A < \tau_\dagger ] .
+\end{align}
+Therefore, we have that
+\begin{align}
+    F(x,y) \E[K] 
+    \approx \P^x[ \tau_A < \tau_\dagger ] 
+    = u(x,y) / g(A) .
+\end{align}
+
+Now, 
+\begin{align}
+    q(y)  &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
+          &= g(A) \int_{\partial B(r_0)} \gamma(x) F(x,y) dx \\
+          &= \frac{ g(A) }{ \E[K] } \migrate(y) .
+\end{align}
+The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
+is expected to spend inside the disk before $\tau_\dagger$.
+This is integral of the Green function for the Bessel process of the appropriate order, 
+so using Borodin \& Salminen (Appendix 1),
+and letting $w$ be the width of the patch,
+in $d=1$,
+\begin{align}
+    h_1(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+  % &= \int_0^{2w} G_\alpha(0,y) dy \\
+  &= \int_0^{2w} \frac{ e^{- y \sqrt{2\alpha} }}{\sqrt{2\alpha}} dy \\
+  &= (1-e^{-2w\sqrt{2\alpha}})/(2\alpha)
+\end{align}
+and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
+\begin{align}
+    h_2(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_\alpha(0,y) dy \\
+  &= \int_0^{2w} 2 y K_0(y \sqrt{2\alpha}) dy \\
+  &= \frac{1}{\alpha}\int_0^{2w\sqrt{2\alpha}} y K_0(y) dy \\
+  &= \frac{1}{2\alpha}\left(1- 2w\sqrt{2\alpha} K_1(2w\sqrt{2\alpha}) \right)  .
+\end{align}
+
+%%%%%%% %%%%%%%%%%%%
+\section{The outflux of families}
+\label{apx:outflux}
+
+It is of interest to have an expression for the ``outflux of families'',
+$\gamma(r_0)$, 
+defined to be the mean proportion of $B$ alleles at distance $r_0$
+whose lineage traces back to the original patch without venturing further away than $r_0$.
+By expression \eqref{eqn:occupation_integral},
+\[
+    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
+\]
+for any $y$.  
+Take $y=(r,0)$.
+Note that after scaling by $\sqrt{2\alpha}/\sigma$,
+\begin{align}
+    \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\|} }{ \|y\|^{-(d-1)/2} e^{-\|y\|} } \\
+        &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r (\|\hat x - x/r\| - 1)} .
+\end{align}
+Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
+so that the expression in the exponent is to, first order in $1/r$,
+\begin{align}
+    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
+            &= - 2 r_0 \cos \theta + O(1/r) .
+\end{align}
+This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
+\begin{align}
+    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} .
+\end{align}
+In one dimension we are done.
+In two dimensions,
+by bounded convergence and \citet{gradshteyn2007table} (3.339),
+\begin{align}
+  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
+            &= C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
+                         &= 2 \pi C' I_0(-2r_0),
+\end{align}
+where $I_0$ is a modified Bessel function of order zero of the first kind.
+The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2\alpha} r_0 / \sigma ) $.
+
+\section{UNUSED AFTER HERE?}
 
@@
 
-%%%%%%%%% %%%%%%%%%%%%%%
-\section{Hitting and occupation}
-\label{apx:hitting_occupation}
-
-Consider a single migrant family $Z$ beginning at $x$.
-Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
-and $\tau_\dagger$ an independent Exponential($\alpha$) time.
-The mean occupation time of $Z$ spent in a region $A$ is 
-\begin{align}
-    u(x,y) &= \E[\int_0^\infty Z_t(A) dt ] \\
-           &= \int_0^\infty \E[Z_t] p_t(x,A) dt \\
-        &= \int_0^\infty e^{-\alpha t} p_t(x,A) dt \\
-        &= \int_0^\infty \P^x\{ \tau_\dagger < t \; \& \; B_t \in A \} dt \\
-        &= \E[ \int_0^{\tau_\dagger} \one_A(B_t) dt ]  .
-\end{align}
-If we define
-$\tau_A$ to be the hitting time of $A$ by $B$,
-and $\mu(x,A)$ to be the hitting distribution of $\partial A$ by $B_{\min(t,\tau)}$,
-by the strong Markov property this is equal to
-\begin{align}
-    u(x,y) &= \E^x[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] \\
-           &= \P^x\{ \tau_A < \tau_\dagger \}  \E^{\mu(x,A)}[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] .
-\end{align}
-If $A$ is circular, the latter expectation will not depend on $x$;
-call this $g(A)$.
-
-Now consider the chance that $Z$ ever hits $A$.
-If we make the approximation that $Z$ hits $A$ iff the trunk of $Z$ does,
-then this is equal to
-\begin{align}
-    F(x,y) &= \int_0^\infty (1-k_e(t)) \P\{ \tau_A \in dt \} \\
-           &= \int_0^\infty e^{\alpha t} (1-k_e(t)) e^{- \alpha t } \P\{ \tau_A < t \} dt \\
-    &=\int_0^\infty e^{\alpha t} (1-k_e(t)) \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &= \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &= \frac{1}{\E[K]} \P^x[ \tau_A < \tau_\dagger ] .
-\end{align}
-Therefore, we have that
-\begin{align}
-    F(x,y) \E[K] 
-    \approx \P^x[ \tau_A < \tau_\dagger ] 
-    = u(x,y) / g(A) .
-\end{align}
-
-Now, 
-\begin{align}
-    q(y)  &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
-          &= g(A) \int_{\partial B(r_0)} \gamma(x) F(x,y) dx \\
-          &= \frac{ g(A) }{ \E[K] } \migrate(y) .
-\end{align}
-The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
-is expected to spend inside the disk before $\tau_\dagger$.
-This is integral of the Green function for the Bessel process of the appropriate order, 
-so using Borodin \& Salminen (Appendix 1),
-and letting $w$ be the width of the patch,
-in $d=1$,
-\begin{align}
-    h_1(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-  % &= \int_0^{2w} G_\alpha(0,y) dy \\
-  &= \int_0^{2w} \frac{ e^{- y \sqrt{2\alpha} }}{\sqrt{2\alpha}} dy \\
-  &= (1-e^{-2w\sqrt{2\alpha}})/(2\alpha)
-\end{align}
-and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
-\begin{align}
-    h_2(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-    % &= \int_0^{2w} G_\alpha(0,y) dy \\
-  &= \int_0^{2w} 2 y K_0(y \sqrt{2\alpha}) dy \\
-  &= \frac{1}{\alpha}\int_0^{2w\sqrt{2\alpha}} y K_0(y) dy \\
-  &= \frac{1}{2\alpha}\left(1- 2w\sqrt{2\alpha} K_1(2w\sqrt{2\alpha}) \right)  .
-\end{align}
-
-%%%%%%% %%%%%%%%%%%%
-\section{The outflux of families}
-\label{apx:outflux}
-
-It is of interest to have an expression for the ``outflux of families'',
-$\gamma(r_0)$, 
-defined to be the mean proportion of $B$ alleles at distance $r_0$
-whose lineage traces back to the original patch without venturing further away than $r_0$.
-By expression \eqref{eqn:occupation_integral},
-\[
-    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
-\]
-for any $y$.  
-Take $y=(r,0)$.
-Note that after scaling by $\sqrt{2\alpha}/\sigma$,
-\begin{align}
-    \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\|} }{ \|y\|^{-(d-1)/2} e^{-\|y\|} } \\
-        &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r (\|\hat x - x/r\| - 1)} .
-\end{align}
-Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
-so that the expression in the exponent is to, first order in $1/r$,
-\begin{align}
-    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
-            &= - 2 r_0 \cos \theta + O(1/r) .
-\end{align}
-This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
-\begin{align}
-    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} .
-\end{align}
-In one dimension we are done.
-In two dimensions,
-by bounded convergence and \citet{gradshteyn2007table} (3.339),
-\begin{align}
-  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-            &= C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
-                         &= 2 \pi C' I_0(-2r_0),
-\end{align}
-where $I_0$ is a modified Bessel function of order zero of the first kind.
-The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2\alpha} r_0 / \sigma ) $.
-
 \end{document}

### 2014-06-28 patchy-selection-paper.tex

@@
 
+\subsection{PETER'S VERSION}
+
 This paper is an investigation into the basic, ill-formed question: 
@@
 optimal $s_m$ for migration makes migration harder to spot
+
+\subsection{GRAHAM'S VERSION}
+
+We have developed a simple set of approximations to describe the rate
+at which organisms in a patchy environment adapt through novel
+mutations or migration from other nearby patches (eqn.
+\eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
+probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
+We have also determined the characteristic genomic length scales over which
+the various signals of hitchhiking should be visable
+(eqn. \eqref{eqn:haplotype_length} and ).
+
+These results show that the probability of parallel adaptation is
+independent of both the population density and the beneficial
+selection coefficient. This follows from the fact that these
+parameters both increase the rate of adaptation via both migration and
+mutation equally. Thus they are important for the time-scale over
+which adaptation occurs but not the mode of adaptation. This
+constrasts with models of convergent adaptation within panmictic
+populations \citep{} and geographically spread populations adapting to
+homogeneous populations \citep{RalphCoop}, where the number of
+adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
+coefficient is absent or weak, due to the fact that selection both
+aids establiishment of new alleles and the spread of existing
+alleles.
+
+The primary determinant of the probability of parallel adaptation is
+$\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
+allele frequency decay off the patch. We find that patches that are close together on
+this scale will adapt via migration, but there will be a sharp
+transition to convergent adaptation as we move to patches that are further apart.
+
+In coming to this conclusion, we have ignored the possibility of very long distance migration,
+instead focusing local dispersal (as parameterized by a Gaussian
+dispersal model). However, dispersal distributions can be very heavy
+tailed, with a small fraction of individuals moving very long distances indeed, in
+addition over long time-scales very rare chance events (mice sucked up in
+hurricanes etc) could play a role in spreading migrant alleles. Such
+tail events could greatly increase the probability of parallel
+adaptation above that predicted by our model. Furthermore, if
+adaptive alleles do move between distant patches via rare long distance
+migration they will be associated with a much longer shared haplotype than
+predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
+results as a null model by which the contribution of long distribution
+migrants to adaptation could be empirically judged.
+
+We have also assumed that the alleles have a fitness effect in
+heterozygotes both on and off the patch. However, 
+ recessive alleles could also contribute to adaptation to patchy
+ environments. \gc{should we cite Nachman eg here}.  Our
+ simple branching process framework does not easily accomodate recessive
+ alleles as it ignores the selection interaction of alleles, and so a
+ different approach will be needed. At first sight it
+ seems like recessive alleles may be able to contribute to adaptation
+ via migration over longer distances, as the migrant alleles will be at low
+ frequency between patches and so usually invisible to
+ selection. However, this fact also means that they are unlikely to
+ establish on the beneficial patch upon reaching it via migration. As
+ such it is not clear what our intuition should be about the
+ contribution of recessive alleles to adaptation via migration.
+
+\gc{will write a couple more broader paragraphs}
+
+The falling cost of sequencing over the next decade offers the
+opportunity to investigate the genomics of parallel adaptation over a
+large s
+Over the coming decade population genomic data will be generated
+
+\plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
 

### 2014-06-29 patchy-selection-paper.tex
notes on lit

@@
 \end{align}
+
+\section{Notes: for intro/discussion}
+
+- leimu2008metaanalysis: local adaptation not ubiquitous but common; mostly in large population sizes.
+- robertson1960theory: dependence of Ne on the degree of adaptation
+- lots of local adaptation stuff on plants and conservation genetics; ties into plasticity and range limits.
+- but also: Neurospora (ellison2011) and salmonids (fraser2011)
+- swamping in Mayr 1963, Animal Species and Evolution?
+- hanson1966effects: original observation of minimum patch size, in simulation?
+- macnair1991: larger effects more common in e.g. heavy metal sites
+- schat1996: silene parallel adaptation to mine tailings
+- kaweki: a review.
+- fourneir-level2011: map of arabidopsis local adaptation
+- reviews for local adapt in marine (sanford2011), trees (savolainen2007)

### 2014-06-30 patchy-selection-paper.tex
added refs, minor edits to appendices

@@
 Convergence on a molecular level also suggests that the path 
-accessible to adaptation is sometimes relatively constrained, 
+accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
 as would be the case if only a few changes confer the advantagous phenotype 
@@
 that have repeatedly adapted to patches of soil with high concentrations of heavy metals
-(e.g.\ serpentine outcrops and mine tailings) \citep{turner2010serpentine,mimulus};
+(e.g.\ serpentine outcrops and mine tailings) \citep{macnair1991evolution,schat1996identical,turner2010serpentine,mimulus};
 the alleles conferring heavy metal tolerance are thought to be locally restricted 
@@
 To investigate the impact of the various approximations surrounding the mutational influx,
-we carried out numerical calculations of a fine-grained, discrete model.
-These are shown, and described, in figure~\ref{fig:prob_estab_calcs}.
+we calculated the probability of establishment for an spatial branching process moving across discrete demes.
+The solution was obtained by iterating the multivariate generating function until convergence
+(since the function which gives the probability of extinction as a function of location of the progenitor
+is an attractive fixed point of the generating function; see \citet{jagers1975branching}).
+These solutions are shown, and parameters described, in figure~\ref{fig:prob_estab_calcs}.
 
@@
     The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respecively);
-    and the selection coefficient interpolates linearly between these in the central shaded region.
+    and the selection coefficient interpolates linearly between these in the central hatched region.
     The number of offspring is Poisson.
@@
 
+For completeness, and clarity as to the scalings on the relevant parameters,
+here we provide a derivation of the differential equations referred to above equation~\eqref{eqn:eqfreq},
+and establish the asymptotics given in that equation.
 One route to the ``equlibrium frequency'' of the allele outside the range where it is advantageous is as follows;
 see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} for other arguments in this case, 
-or \citet{etheridge2000introduction} or \citet{dawson1993measurevalued} for general theory used below.
+or \citet{etheridge2000introduction} and/or \citet{dawson1993measurevalued} for general theory used below.
+Following is a description of a discrete model (which is nearly the one we simulate from).
 Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
@@
 Each individual at location $x$ reproduces after a random lifetime,
-dying and giving birth in the same deme to a random number of offspring with distribution given by $X$;
-each offspring migrates independently to a new location chosen randomly from the distribution given by $x+R$,
-where they replaces a randomly chosen individual.
-If $x+R$ is outside of the range, then they perish.
-Each individual's reproduction time is exponentially distributed, 
+dying, and then leaving behind in the same deme a random number of offspring with distribution given by $X$;
+each offspring then migrates independently to a new location chosen randomly from the distribution given by $x+R$,
+where she replaces a randomly chosen individual.
+If $x+R$ is outside of the range, then she perishes.
+Each individual's time until reproduction is exponentially distributed:
 the reproduction rate is 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
@@
 and each have variance $\sigma^2$.
-We also assume that the distribution of $X$ makes the associated random walk irreducible and aperiodic.
+We also assume that the distribution of $R$ makes the associated random walk irreducible and aperiodic.
 
@@
 we need to choose $\var[R]=\sigma^2$ and $s(x)$ to be of the same, small, order; 
-this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale.
+this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale \citep[as noted by][]{slatkin1973geneflow}.
 More concretely, suppose that the grid size is $\epsilon \to 0$, 
@@
 We can solve the case with $s_* = s_d$, since $\phi'(0) = 0$ (as done implicitly in \citet{slatkin1973geneflow}).
-We should be able to argue that if $R$ is large then the value of $s_*$ won't make much difference.
+It should be possible to argue that if $R$ is large then the value of $s_*$ won't make much difference.
 
@@
 
-Presumably, this gets the same answer as Monty, who finds that in his solvable cases,
-for $x$ in the last interval (which ends in $\infty$), that
-\begin{align*}
-    \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / 2 \sigma + C ) \right) \\
-        &\simeq C' \exp\left( - \frac{ x \sqrt{2s} }{ \sigma } \right) .
-\end{align*}
-% Indeed, as $\phi \to 0$, we have approximately $\phi' = \frac{ \sqrt{s} }{ \sigma } \phi$.  % huh????
+% Presumably, this gets the same answer as \citep{slatkin1973geneflow}, who finds that in the solvable cases,
+% for $x$ in the last interval (which ends in $\infty$), that
+% \begin{align*}
+%     \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / 2 \sigma + C ) \right) \\
+%         &\simeq C' \exp\left( - \frac{ x \sqrt{2s} }{ \sigma } \right) .
+% \end{align*}
 
@@
 - leimu2008metaanalysis: local adaptation not ubiquitous but common; mostly in large population sizes.
-- robertson1960theory: dependence of Ne on the degree of adaptation
+- robertson1960theory: dependence of the degree of adaptation on Ne 
 - lots of local adaptation stuff on plants and conservation genetics; ties into plasticity and range limits.
@@
 - hanson1966effects: original observation of minimum patch size, in simulation?
-- macnair1991: larger effects more common in e.g. heavy metal sites
-- schat1996: silene parallel adaptation to mine tailings
-- kaweki: a review.
 - fourneir-level2011: map of arabidopsis local adaptation

### 2014-06-30 Lava_flow_mice_app.R

@@ -3,16 +3,16 @@ char.length= c(3,30)  ## sigma /sqrt(s)
 
 A=100
 
-#pdf(file="~/Dropbox/postdocs/Peter/Spatial_adaptation/Lava_flow_mice_prob_parallel.pdf")
+pdf(file="Lava_flow_mice_prob_parallel.pdf",pointsize=10,width=5,height=5)
 layout((1:2))
-par(mar=c(3,3,0.2,1))
+par(mar=c(4,3.5,1,1)+.1)
 R = seq(5,600,length=10000) ##distance in Km
 
 #layout(t(1:2))
 mu = 1e-8
 prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
-plot(R, prob.parallel,type="l",lwd=3, xlab="R, km",ylab="",cex.lab=1.3,cex.axis=1.3,log="x") 
-mtext(side=2,line=2,"Prob. parallel mutation",cex=1.3)
+plot(R, prob.parallel,type="l",lwd=3, xlab="", ylab="",cex.lab=1.3,cex.axis=1.3,log="x") 
+mtext(side=2,line=2.5,"prob parallel mutation",cex=1.3)
 mu=1e-5
 prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
 
@@ -31,8 +31,7 @@ lines(R, prob.parallel,lwd=3,col="red",lty=2)
 
 legend("topleft", legend=expression(mu == 10^-8, mu == 10^-5, sigma/sqrt(s[m]) == 3,sigma/sqrt(s[m]) == 30),col=c("black","red","grey","grey"),lty=c(1,1,1,2),lwd=2)
 
-mtext(side=1,line=2,"Distance between patches, r km",cex=1.3)
-#dev.off()
+mtext(side=1,line=2.5,"Distance between patches (km)",cex=1.3)
 
 
 ####haplotype lengths
@@ -40,15 +39,16 @@ mtext(side=1,line=2,"Distance between patches, r km",cex=1.3)
 length.hap.1<-sqrt(2)*1/(char.length[1])/R +1/R^2
 length.hap.2<-sqrt(2)*1/(char.length[2])/R +1/R^2
[…]

### 2014-06-30 patchy-selection-paper.tex
put in dispersal refs

@@
 which has dispersal distance of 0.26km.  Shall we ask Nachman?}
+\plr{He referred to \citet{french1968dispersal,allred1963range}.}
 
@@
 
-\subsection{PETER'S VERSION}
-
-This paper is an investigation into the basic, ill-formed question: 
-What is the spatial resolution of adaptation?
-In other words, what scale of environmental patchiness can a species develop and sustain local adaptations,
-and over what spatial scale will evolution develop independent solutions to evolutionary problems?
+This paper is an investigation into the basic question: 
+What is the spatial resolution of convergent local adaptation?
+In other words, what spatial scale of environmental patchiness will evolution develop independent solutions to evolutionary problems?
 The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
-the dispersal distance divided by the square root of the selection against the new allele between patches,
-and much more weakly on the selective benefit within the patches or (perhaps suprisingly) the population density.
-Weaker selection against alleles will make sharing of adaptations between patches easier,
-but will also make it harder to spot,
-since the lucky, transiting alleles move slower and carry a shorter haplotype.
-
-Observations of similar adaptations occuring within the same species in widely separated patches
-is strongly suggestive of parallel, independent adaptations.
-Such observations are very interesting 
-as they tell us about the degree of constraint faced by evolution --
-how many different methods are there typically available to a species
-with which to solve an evolutionary problem?
-
-It is already well-appreciated that larger populations can lead to more parallelism \citep{softsweepsI}
-\plr{fisher probably said this too?}
-and that spatial structure contributes to this \citep{ralph2010parallel}.
-Naturally, then, local adaptation to well-separated patches of some environment is even more likely to occur in parallel;
-we have studied the case where patches are separated by long distances.
-This is complementary to the work of \citet{XXX}, who studied local adaptation in rapidly varying environments.
-Intermediate cases where habitat types are distributed in some more complex geometry
--- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance --
-remains to be studied.
-
-To complement this theory, we need more empirical work,
-in which the parameters (migration rate, sensitivity to environmental factors, etc)
-are reasonably well-understood.
-However, learning about these aspects is difficult in the field,
-and it is still not understood how much of the intraspecific variation we observe
-is due to adaptation to local conditions, or even how to tell.
-Notable steps forward include \citet{turner2010serpentine}, Mimulus, Amity, others?
+the dispersal distance divided by the square root of the selection
+against the new allele between patches. It depends much more weakly on the
+selective benefit within the patches or (perhaps suprisingly) the
+population density. Although these two factors will determine the
+time-scale over which adaptation will occur. This is contrast with models of panmictic populations
+\citep{softsweepsI} and geographically spread populations adapting to homogeneous
+selection pressures \citep{RalphCoop}, where the probability of
+multiple adaptive alleles increases with the population
+size/density. However, in all of these models the dependence on the beneficial selection
+coefficient is absent or weak, due to the fact that selection both
+aids establishment of new alleles and the spread of existing
+alleles.
+
+While weaker selection against alleles will make sharing of
+adaptations between patches easier, 
+it will also makes it harder to spot the sharing,
+since the lucky, transiting alleles move slower and so carry a shorter
+shared haplotype.
+This issue is amplified by the fact that the length of haplotype
+shared within patches decays over time, potentially making the
+identification of shared adaptations to old selection pressures difficult.
+
+% Observations of similar adaptations occuring within the same species in widely separated patches
+% is strongly suggestive of parallel, independent adaptations.
+% Such observations are very interesting 
+% as they tell us about the degree of constraint faced by evolution --
+% how many different methods are there typically available to a species
+% with which to solve an evolutionary problem?
+
+% It is already well-appreciated that larger populations can lead to more parallelism \citep{softsweepsI}
+% \plr{fisher probably said this too?}
+% and that spatial structure contributes to this \citep{ralph2010parallel}.
+
+% This is complementary to the work of \citet{XXX}, who studied local adaptation in rapidly varying environments.
+% Intermediate cases where habitat types are distributed in some more complex geometry
+% -- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance --
+% remains to be studied.
+
+% To complement this theory, we need more empirical work,
+% in which the parameters (migration rate, sensitivity to environmental factors, etc)
+% are reasonably well-understood.
+% However, learning about these aspects is difficult in the field,
+% and it is still not understood how much of the intraspecific variation we observe
+% is due to adaptation to local conditions, or even how to tell.
+% Notable steps forward include \citet{turner2010serpentine}, Mimulus, Amity, others?
 
@@
 and hence mostly present only in heterozygotes. 
+Our results should hold as a good approximation to dominant and
+partially dominant alleles.
 If, however, the mutation is recessive, then it is essentially neutral where rare,
@@
 This is counteracted, however, by the increased difficulty with which the mutation would establish
-once arriving in a new patch.
-On the other hand, a dominant mutation would not differ greatly from the theory presented here.
-In fact, MC1r is ?dominant? and agouti is ?recessive?
-Other schemes could be obtained.
-
-optimal $s_m$ for migration makes migration harder to spot
-
-\subsection{GRAHAM'S VERSION}
-
-We have developed a simple set of approximations to describe the rate
-at which organisms in a patchy environment adapt through novel
-mutations or migration from other nearby patches (eqn.
-\eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
-probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
-We have also determined the characteristic genomic length scales over which
-the various signals of hitchhiking should be visable
-(eqn. \eqref{eqn:haplotype_length} and ).
-
-These results show that the probability of parallel adaptation is
-independent of both the population density and the beneficial
-selection coefficient. This follows from the fact that these
-parameters both increase the rate of adaptation via both migration and
-mutation equally. Thus they are important for the time-scale over
-which adaptation occurs but not the mode of adaptation. This
-constrasts with models of convergent adaptation within panmictic
-populations \citep{} and geographically spread populations adapting to
-homogeneous populations \citep{RalphCoop}, where the number of
-adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
-coefficient is absent or weak, due to the fact that selection both
-aids establiishment of new alleles and the spread of existing
-alleles.
-
-The primary determinant of the probability of parallel adaptation is
-$\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
-allele frequency decay off the patch. We find that patches that are close together on
-this scale will adapt via migration, but there will be a sharp
-transition to convergent adaptation as we move to patches that are further apart.
-
-In coming to this conclusion, we have ignored the possibility of very long distance migration,
+once arriving in a new patch. As
+ such it is not clear what our intuition should be about the
+ contribution of recessive alleles to adaptation via migration. 
+Further work is needed to understand how recessive alleles may
+contribute to convergent local adaptation.
+
+\paragraph{Long distance migration}
+We have also ignored the possibility of very long distance migration,
 instead focusing local dispersal (as parameterized by a Gaussian
@@
 
-We have also assumed that the alleles have a fitness effect in
-heterozygotes both on and off the patch. However, 
- recessive alleles could also contribute to adaptation to patchy
- environments. \gc{should we cite Nachman eg here}.  Our
- simple branching process framework does not easily accomodate recessive
- alleles as it ignores the selection interaction of alleles, and so a
- different approach will be needed. At first sight it
- seems like recessive alleles may be able to contribute to adaptation
- via migration over longer distances, as the migrant alleles will be at low
- frequency between patches and so usually invisible to
- selection. However, this fact also means that they are unlikely to
- establish on the beneficial patch upon reaching it via migration. As
- such it is not clear what our intuition should be about the
- contribution of recessive alleles to adaptation via migration.
-
-\gc{will write a couple more broader paragraphs}
-
-The falling cost of sequencing over the next decade offers the
-opportunity to investigate the genomics of parallel adaptation over a
-large s
-Over the coming decade population genomic data will be generated
-
-\plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
-
-if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
-recall others' work.
-
-patchy selection helps this happen; 
-effective migration rate looks like what; 
-strength of (positive?) selection suprisingly does not affect results (?)
-
-standing variation will be important under these circumstances; 
-migration rate (and so geographic structure) will not affect numbers of types if this is true; 
-however migration always affcts spatial patterns.
-
-recessive alleles will have different dynamics. Actually not really
-obvious what will happen here, cline width will be much wider but
-prob. of taking off will go as the square of the occupancy density.
-
-discussion of applications.
-
-discuss mixing of types; 
-Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
-
-unequal density: Lenormand, gene flow and limits of\dots
-
-small-scale fluctuations
-when landscape is not made of discrete (circular) patches
+The falling cost of population genomic sequencing means that we will soon have the
+opportunity to study the effects of geography and ecology on
+adaptation across many populations within a species. 
+Our work suggests that even quite geographically close populations may
+be forced to adapt convergently when migration is geographically limited. 
+Thus, systems where populations have repeated to strong local
+selection pressures limited may offer the opportunity to study
+highly replicated convergent adaptation within a similar genetic
+background.  
+Such empricial work will also strongly inform our understanding of the role of gene flow in
+keeping ecological similar populations evolving in concert \citep{sexton}.
+Our results suggest that adaptation to shared environments is
+certainly no guarentee of a shared genetic basis to adaptation, suggesting that
+rapid adaptation to a share environment could potentially drive
+speciation under some scenerios.
+
+% \subsection{GRAHAM'S VERSION}
+
+% We have developed a simple set of approximations to describe the rate
+% at which organisms in a patchy environment adapt through novel
+% mutations or migration from other nearby patches (eqn.
+% \eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
+% probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
+% We have also determined the characteristic genomic length scales over which
+% the various signals of hitchhiking should be visable
+% (eqn. \eqref{eqn:haplotype_length} and ).
+
+% These results show that the probability of parallel adaptation is
+% independent of both the population density and the beneficial
+% selection coefficient. This follows from the fact that these
+% parameters both increase the rate of adaptation via both migration and
+% mutation equally. Thus they are important for the time-scale over
+% which adaptation occurs but not the mode of adaptation. This
+% constrasts with models of convergent adaptation within panmictic
+% populations \citep{} and geographically spread populations adapting to
+% homogeneous populations \citep{RalphCoop}, where the number of
+% adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
+% coefficient is absent or weak, due to the fact that selection both
+% aids establiishment of new alleles and the spread of existing
+% alleles.
+
+% The primary determinant of the probability of parallel adaptation is
+% $\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
+% allele frequency decay off the patch. We find that patches that are close together on
+% this scale will adapt via migration, but there will be a sharp
+% transition to convergent adaptation as we move to patches that are further apart.
+
+% We have also assumed that the alleles have a fitness effect in
+% heterozygotes both on and off the patch. However, 
+%  recessive alleles could also contribute to adaptation to patchy
+%  environments. \gc{should we cite Nachman eg here}.  Our
+%  simple branching process framework does not easily accomodate recessive
+%  alleles as it ignores the selection interaction of alleles, and so a
+%  different approach will be needed. At first sight it
+%  seems like recessive alleles may be able to contribute to adaptation
+%  via migration over longer distances, as the migrant alleles will be at low
+%  frequency between patches and so usually invisible to
+%  selection. However, this fact also means that they are unlikely to
+%  establish on the beneficial patch upon reaching it via migration. As
+%  such it is not clear what our intuition should be about the
+%  contribution of recessive alleles to adaptation via migration.
+
+% \gc{will write a couple more broader paragraphs}
+
+% The falling cost of sequencing over the next decade offers the
+% opportunity to investigate the genomics of parallel adaptation over a
+% large s
+% Over the coming decade population genomic data will be generated
+
+% \plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
+
+% if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
+% recall others' work.
+
+% patchy selection helps this happen; 
+% effective migration rate looks like what; 
+% strength of (positive?) selection suprisingly does not affect results (?)
+
+% standing variation will be important under these circumstances; 
+% migration rate (and so geographic structure) will not affect numbers of types if this is true; 
+% however migration always affcts spatial patterns.
+
+% recessive alleles will have different dynamics. Actually not really
+% obvious what will happen here, cline width will be much wider but
+% prob. of taking off will go as the square of the occupancy density.
+
+% discussion of applications.
+
+% discuss mixing of types; 
+% Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
+
+% unequal density: Lenormand, gene flow and limits of\dots
+
+% small-scale fluctuations
+% when landscape is not made of discrete (circular) patches
 

### 2014-06-30 patchy-selection-paper.tex
put in dispersal refs

@@
 
+\title{
+Convergent Adaptation while adapting 
+}
+
+\author{Peter L. Ralph$^1$ \\ email: pralph@usc.edu  \and Graham Coop$^2$ \\ email: gmcoop@ucdavis.edu }
+
 \begin{document}
@@
 \section{Introduction}
+
+% Such precise convergence points to the conservation of function of the genes underlying adaptive traits, 
+% sometimes over deep time scales \citep{deephomologypapers}. 
+% Convergence on a molecular level also suggests that the path 
+% accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
+% as would be the case if only a few changes confer the advantagous phenotype 
+% and are sufficiently free of deleterious pleiotropic consequences.
 
@@
 genetic changes in the same pathway, genes, 
-or even by precisely the same genetic changes \citep{zhen2012parallel}. 
-Such precise convergence points to the conservation of function of the genes underlying adaptive traits, 
-sometimes over deep time scales \citep{deephomologypapers}. 
-Convergence on a molecular level also suggests that the path 
-accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
-as would be the case if only a few changes confer the advantagous phenotype 
-and are sufficiently free of deleterious pleiotropic consequences.
-
-%For highly polygenic traits
-
+or even by precisely the same genetic changes \citep{zhen2012parallel,,martin2013repeated,deephomologypapers}. 
 Convergent adaptation can also occur within species, if individuals within a species 
-adapt to the same environment, in parallel, through different genotypic changes. 
+adapt to the same environment, in parallel, through different genetics changes. 
 There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
 Such evolution of convergent phenotypes is favored by higher mutational input, 
-i.e.\ larger mutational target and/or population size \citep{dmitri}.
+i.e.\ larger mutational target and/or population size \citep{softsweepsII}.
 The geographic distribution of populations can also affect the probability of parallel mutation within a species:
@@
 that have repeatedly adapted to patches of soil with high concentrations of heavy metals
-(e.g.\ serpentine outcrops and mine tailings) \citep{macnair1991evolution,schat1996identical,turner2010serpentine,mimulus};
+(e.g.\ serpentine outcrops and mine tailings) \citep{macnair1991evolution,schat1996identical,turner2010serpentine};
 the alleles conferring heavy metal tolerance are thought to be locally restricted 
-because they incur a cost off of these patches. 
+because they incur a cost off of these patches \citep{wright2013indirect}.
 Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration
 to particular substrates, e.g.\ dark rock outcrops or white sand dunes \citep{benson1933concealing}.
+\plr{we are missing what Nachman said was the original ref for this... I think you took a picture?}
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
@@
 and, perhaps as a result of this strong selection against migrants, 
-at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{nachman2003different}. 
+at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{hoekstra2003different}. 
 Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
-and in lizards \citep{whitesands}.
+and in lizards \citep{rosenblum2010molecular}.
 
@@
 separated by (inhabited) regions where the alleles are deleterious.
-\plr{or neutral?}
 The main questions are:
@@
 and when is it likely that migration carries these alleles between patches?
+
 How could we tell, after the fact, which has occurred?
+
 We work in a model of continuous geography,
 using a variety of known results and new methods.
-In section XX we \dots QUICK OUTLINE
+In section \ref{ss:patchymutation} we develop a simple approximation,
+eqn. \eqref{eqn:mutrate}, for the rate at which patches become
+established by new mutations. Then in section \ref{ss:patchymigration} we
+develop an approximation, eqn. \eqref{eqn:migrate}, to the rate which alleles move from an
+existing patch to colonize a novel patch (this represents a big chunk
+of the results of the paper). We combine these two results in section
+\ref{ss:probparallel} to develop an approximation to the probability of
+parallel adaptation, eqn. \eqref{}.
+To understand the genomic signal left by adaptations shared by
+migration in section \ref{ss:haplotype_length} we approximate the time
+it will take for an allele to transit between patches,
+eqn. \eqref{eqn:mean_tau}, and use this to approximate the length of
+haplotype that hitchhikes with it,
+eqn. \eqref{eqn:haplotype_length}. We then quantify how quickly this
+haplotype decays within patches eqn. \eqref{} in section \ref{ss:patch_haplotype}.
+
+\dots QUICK OUTLINE
 
@@
 pressure \citep{maruyama1970fixation,cherry2003diffusion}. 
+More general work on the mathematically equivalent problem of density-dependent population growth 
+has obtained much more general criteria
+under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1989diffusive},
+and determined the habitat configuration that maximizes the probability of establishment \citep{lou2006minimization}.
 
@@
 To determine this, we dissect the process by which an allele transits between the patches,
-in the process obtaining other useful information about the genealogy of B alleles.
+in the process obtaining other useful information about the genealogy of $B$ alleles.
 
@@
 we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
-of the outflux $\gamma$ at that point, mulitplied by the relevant second quantity.
+of the outflux $\gamma$ at that point, multiplied by the relevant second quantity.
 By examining the constituent terms,
-it will turn out that equation \eqref{eqn:gestalt_migrate} is a constant multiple of \eqref{eqn:gestalt_q}.
-
-First consider equation \eqref{eqn:gestalt_q}.
+it will turn out that equation \eqref{eqn:gestalt_migrate} 
+is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.
+
+First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
 Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
@@
 as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
-Concretely, it is the density of $\E[\int_0^t Z_t dt]$,
-i.e.\ the smooth function $u(x,y)$ that, integrated over a set $A$, 
-gives the expected total number of individuals ever living in $A$:
+Concretely, it is the density of $\E[\int_0^t Z_t(\cdot) dt]$,
+i.e.\ the smooth function $u(x,y)$ that, integrated over a set $S$, 
+gives the expected total number of individuals ever living in $S$:
 \begin{align} \label{eqn:occupation_density_defn}
-    \E\left[ \int_0^\infty Z_t(A) dt \right] = \int_A u(x,y) dy .
-\end{align}
-Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
+    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
+\end{align}
+% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
 Since reproduction and spatial motion are independent,
@@
 Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
-which we assume is circular, has relatively small area $A$ and is located near position $y$.
+which we assume is circular, has relatively small area $A$ and is centered at position $y$.
 % The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
 % multiplied by the probability that the family reaches the patch, given survival.
-% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-\alpha t}/\E[K]$.
-It would be possible to analyze directly the probability that a given migrant family establishes in $A$,
+% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
+It would be possible to analyze directly the probability that a given migrant family establishes in the patch,
 as in \citet{barton1987establishment};
@@
 multiplied by the chance that at least one member of the family escapes demographic stochasticity,
-successfulling establishing in the new patch.
-Write $h(x,y)$ for the probability that the Brownian trunk
+successfully establishing in the new patch.
+Write $h_A(x,y)$ for the probability that the Brownian trunk
 hits the new patch centered at $y$ before the family dies out,
@@
 given that it successfully arrives.
-As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral of $h(x,y)$:
+As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral of $h_A(x,y)$:
 \begin{align}\label{eqn:migrate_integral}
-  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h(x,y) f(A) dx .
-\end{align}
-
-Intuitively, $h(x,y)$ counts each family once if it hits the patch,
+  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
+\end{align}
+
+Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
 while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
@@
 if we know that it survived to hit the patch,
-it is intuitively reasonable that $u(x,y)$ and $h(x,y)$ are simply related.
+it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
 Following this intuition,
@@
 \begin{align} 
-  h(x,y) = \frac{ g(A) }{ \E[K] } u(x,y),
+  h_A(x,y) \approx \frac{ g(A) }{ \E[K] } u(x,y),
 \end{align}
@@
 and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
-Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a constant factor
-(at least asymptotically as $x$ and $y$ are far apart),
-we now know that the outflux of successful migrant families is equal to the mean occupation frequency
-multiplied by $g(A) f(A) / \E[K]$.
+Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
+that (at least asymptotically) does not depend on the distance between $x$ and $y$,
+we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
 The probability $f(A)$ that the family does not die out once it arrives at the new patch is
@@
 % \begin{align}
-%   s(w) \simeq \frac{1}{2\alpha} \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
+%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
 % \end{align}
-Combining the terms with expressions for $g(A)$ in Appendix~\ref{apx:hitting_occupation},
+Combining with expressions for $g(A)$ in Appendix~\ref{apx:hitting_occupation},
 we obtain that
@@
   \migrate(y) &\approx q(y) \times \frac{ \rho s_p  }{ \xi^2 s_m }  \\
-  % \left( 1 - \left( \frac{\pi w \sqrt{2 \alpha}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2\alpha}/\sigma} \right).
+  % \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
   &\approx C \frac{ \rho s_p  }{ \xi^2 s_m } \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) .
@@
 as long as the patch is large enough.
-For more discussion, and correction terms, see Appendix~\ref{apx:hitting_occupation}.
+For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see Appendix~\ref{apx:hitting_occupation}.
 
@@
 \begin{equation} \label{eqn:parallel_prob}
-  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu \alpha }{2 A
-    \mu \alpha + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
+  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu s_m }{2 A \mu s_m + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
 \end{equation}
@@
 %%%%%%%%%%%%
-\subsection{Multiple patches}
+\subsubsection{Multiple patches}
 Above we worked with two patches separated by distance $R$;
@@
 and the length of the haplotype shared between this lineage and one in the original patch.
-The entire time that the lineage is outside the region where the B allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
+The entire time that the lineage is outside the region where the $B$ allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
 the haplotype that accompanies it is broken down rapidly. 
@@
 surrounded by a cloud of close relatives of random size $K$
-whose chance of surviving until $t$ generations is $k_e(t)$.
+whose chance of surviving until $t$ generations is $1-k_e(t)$.
 Consider a single such family, and let $\tau$ be the (random) time at which it first hits the new patch,
@@
 In so doing, we neglect 
-recombination between B alleles (since they are at low density in transit, we assume this is rare),
+recombination between $B$ alleles (since they are at low density in transit, we assume this is rare),
 and the possibility that more than one migrant family is in transit at once
@@
 
-Our first obstacle is that we do not know $k_e(t)$, 
-the distribution of the time until a family dies out, exactly.
-However, we do know that $k_e(t) \simeq e^{-s_m t} / \E[K]$,
+Our first obstacle is that we do not know $1-k_e(t)$, 
+the probability a family survives until $t$, exactly.
+However, we do know that $1-k_e(t) \simeq e^{-s_m t} / \E[K]$,
 i.e.\ that it has exponential tails.
@@
 Refer again to Figure~\ref{fig:lineagesmotion}. 
-When a lineage is in a region with high frequency of B alleles (dark red), 
-it is very unlikely to recombine off the B background. 
+When a lineage is in a region with high frequency of $B$ alleles (dark red), 
+it is very unlikely to recombine off the $B$ background. 
 However, when it wanders out of the patch,
-into regions with lower frequency of B alleles,
+into regions with lower frequency of $B$ alleles,
 when it recombines it will
-usually be on to the background of the b allele.
+usually be on to the background of the $b$ allele.
 
@@
 off of the B background we need to understand what fraction of the
-time our lineage spends in geographic regions where $b$ in common. 
-Consider the lineage of a single B allele, moving back through time,
+time our lineage spends in geographic regions where $b$ is common. 
+Consider the lineage of a single $B$ allele, moving back through time,
 in a patch at migration-selection balance, 
-i.e.\ with the local proportion of B alleles near $x$ equal to $p(x)$,
-and that the lineage stays close enough to the patch 
+i.e.\ with the local proportion of $B$ alleles near $x$ equal to $p(x)$,
+and suppose that the lineage stays close enough to the patch 
 that fluctuations about $p(x)$ are small. 
@@
 patch rather than the rare lineages that make it far from the patch).
-If we suppose that type B individuals at location $y$ reproduce at rate $1+s(y)$,
+If we suppose that type $B$ individuals at location $y$ reproduce at rate $1+s(y)$,
 and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
-then the total rate of influx of B alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
-As argued by \citet{hudson}, the rate at which a B lineage moves from $x$ to $y$
+then the total rate of influx of $B$ alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
+As argued by \citet{hudson1988coalescent}, the rate at which a $B$ lineage moves from $x$ to $y$
 is this influx divided by the local abundance of $B$ alleles,
-or $\frac{ p(y) (1+s(y)) m(y,x) }{ p(x) }$.
-After the same large-population, continuous-space limit that underlies the rest of this paper,
-this motion should converge to a diffusion process \citep[as assumed by][]{hallatschek},
+or $p(y) (1+s(y)) m(y,x) / p(x) $.
+After the same large-population, continuous-space limit that underlies the rest of this paper (Appendix~\ref{apx:eqfreq}),
+this motion should converge to a diffusion process \citep[as assumed by e.g.][]{hallatschek2008surfing},
 notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
+This diffusion, when at $x$, 
+is pulled toward regions with more $B$ alleles
+by a drift term $(1+s(x))\frac{d}{d\,x} \log( p(x) (1+s(x)) )$,
+where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
+and moves at speed $\sqrt{1+s(x)}$ (which is nearly constant, since $s(x)$ is small).
 In It\^o differential notation, 
-the location $X_t$ of a B lineage $t$ generations ago satisfies
-\begin{align}
-    d X_t &= (1+s(X_t)) \frac{d}{d\,x} \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,\\
+the location $X_t$ of a $B$ lineage $t$ generations ago satisfies
+\begin{align}
+    d X_t &= (1+s(X_t)) \frac{d}{dx} \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,
     % &=  ( \log(p(X_t))' (1+s(X_t)) + s'(X_t) ) dt + \sqrt{1+s(X_t)} dB_t,\\
 \end{align}
-where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
-and $B_t$ is a standard Brownian motion.
-This is a diffusion describing the movement of our lineage backward in
-timethat, when at $x$, 
-is pulled toward regions with more B alleles
-by the drift term $(1+s(x))\frac{d}{d\,x} \log( p(x) (1+s(x)) )$,
-and moves at nearly constant speed (since $\sqrt{1+s(x)}$ is close to 1).
-Furthermore, it is easy to verify that the stationary distribution of this process is
+where $B_t$ is a standard Brownian motion.
+Furthermore, it is easy to verify that the stationary distribution of this process has density
 \begin{align}
@@
 \end{align}
-A lineage at a locus at genetic distance $r$ that begins linked to a B allele
-moves as a B lineage up until the most recent time
-it recombined from a b background to a B background,
+A lineage at a locus at genetic distance $r$ that begins linked to a $B$ allele
+moves as a $B$ lineage up until the most recent time
+it recombined from a $b$ background to a $B$ background,
 which occurs when at $x$ with rate $r (1-p(x)) (1+s(x)/2)$.
@@
 $p(x) \approx 1/2$. Under these approximations the per generation rate of
-recombination of a linked lineage to the b background is
+recombination of a linked lineage to the $b$ background is
 reduced by a factor roughly equal to the proportion of the patch lying within distance $\sigma/\sqrt{s}$ of the boundary,
-i.e.
-% \begin{equation}
-% r (\sigma/\sqrt{s})A^{-1/d} \label{eqn:rate-rec-in-patch}
-% \end{equation}
-\plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
-and you shrink it to have radius $r_0$,
-then the resulting object has area $A \times (r_0/r)^d$;
-here we have $r_0 = r - \sigma/\sqrt{s}$,
-in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
-\begin{align}
-r_{eff} &=
+so the effective recombination rate is
+% \plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
+% and you shrink it to have radius $r_0$,
+% then the resulting object has area $A \times (r_0/r)^d$;
+% here we have $r_0 = r - \sigma/\sqrt{s}$,
+% in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
+\begin{align}
+r_\text{eff} &=
     \begin{cases}
@@
   \sqrt{A/\pi} \right)}{A+\pi \sigma/\sqrt{s} \left( \sigma/\sqrt{s} +
-  \sqrt{A/\pi} \right) }    \qquad & d=2 \\  
+  \sqrt{A/\pi} \right) }    \qquad & d=2 ,  
 \end{cases}
@@
 \end{align}
-\gc{Altered the above.}
-which scales the recombination rate roughly with $(\sigma/\sqrt{s})A^{-1/d}$.
-We can use this to think about the length of the haplotype
-associated with the $B$ allele after a given length of time $T$,
-for instance, the time since the $B$ allele joined the patch.
-\plr{changed this to genetic lengths so as not to introduce another parameter, $r_{BP}$}
-% If the recombination rate per base is $r_{BP}$, 
-% then the length of the haplotype associated with the $B$ allele is
-% \begin{equation}
-% \frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
-% \end{equation} 
-In units of genetic length (Morgans),
-the length of the haplotype associated with the $B$ allele is, up to a constant,
-\begin{equation}
-  \frac{A^{1/d}\sqrt{s}}{\sigma T}
-\end{equation} 
-Note that the length of haplotype decays at a rate proportional to the
-cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
+so that $T$ generations after the founding of a patch we expect a
+haplotype of genetic length $1/(r_\text{eff}T)$ Morgans to remain linked to
+the $B$ allele in a typical lineage. 
+Note that the length of haplotype
+decays at a rate roughly proportional to $\sqrt{s}/ \sigma$, as the narrower the cline the less
 opportunity for recombination to move our lineage off the $B$ background.
@@
 Finally we can consider the expected length of haplotype shared within a patch. 
-To do this, we compare the rate of recombination between B and b haplotypes (equation~\ref{eqn:rate-rec-in-patch})
-to the rate of coalescence of B haplotypes within a patch. 
+To do this, we compare the rate of recombination between $B$ and $b$ haplotypes (equation~\ref{eqn:rate-rec-in-patch})
+to the rate of coalescence of $B$ haplotypes within a patch. 
 We will assume that mixing within the patch is fast compared to the rate of drift, 
-so that the rate of coalescence of B alleles within a patch is 
+so that the rate of coalescence of $B$ alleles within a patch is 
 $\xi^2/(A\rho)$,
 the ``variance effective population size'' of the patch
-(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis}).
+(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1975pain,barton2002neutral}).
 The rate of recombination and coalescence will be on the same order if
@@
 and hence the genetic length of haplotype shared between pairs of
-individuals within the patches is on the order of
+individuals within the patches at equilibrium is on the order of
 \begin{align}
@@
 but decreases with $\sqrt{A}$ in two dimensions.
+In applications, this can be compared to equation~\eqref{eqn:haplotype_length}
+for the length of shared haplotype between patches.
 
@@
 presumably to avoid visual predators.
-\citet{nachman2003different} demonstrated that on one of these flows (Pinacate), 
+\citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
 much of the change in coloration is due to an allele at the MC1R locus.
@@
 The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
-suggesting these populations have adapted in parallel \citep{xxx}.
+suggesting these populations have adapted in parallel \citep{hoekstra2003different}.
 However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, 
@@
 A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. 
-\citep{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes,
+\citet{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
 at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
@@
 Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
-we can fit formula \eqref{eqn:eqfreq} to these frequencies.
+we can fit formula \eqref{eqn:eqfreq} to these frequencies
+\citep[similar to][]{hoekstra2004ecological}.
 Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
@@
 We also need estimate of $s_m$. 
-Using $\sigma=1km$ \gc{REF}, these cline widths imply $s_m=1/\sqrt{3}$ and
+Using $\sigma \approx 1km$ \citep{french1968dispersal,allred1963range}, these cline widths imply $s_m=1/\sqrt{3}$ and
 $1/\sqrt{30}$ are reasonable values.
-\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
-which has dispersal distance of 0.26km.  Shall we ask Nachman?}
-\plr{He referred to \citet{french1968dispersal,allred1963range}.}
 
@@
 Also there a number of genes besides MC1R at which adaptative changes affecting pigmentation 
-have been identified in closely related species and more broadly across vertebrates \citep{REF}.
+have been identified in closely related species and more broadly across vertebrates \citep{Hoekstra:06}.
 To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
@@
 being able to arise at any of a thousand base pairs.  
-Finally, we set $A=100\text{km}^2$.
+Finally, we set $A=100\text{km}^2$ (roughly the size of the Pinacate patch).
 Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
-of parallel mutation on the distance between lava flow patches using these parameters. 
-As can be seen parallel mutation should become
+of parallel mutation on the distance between lava flow patches using these parameters, 
+showing that parallel mutation should become
 likely over a scale of tens to a few hundred kilometers between patches.
 
-Givn the large selection coefficient associated with the dark allele
-on the dark substrate we expect the initial sweep associated with
-either a new mutation or migrant allele to be large ($s/\log(\rho A
-s)$). We can use eqn \eqref{eqn:haplotype_length} to calculate how long a
-founding haplotype we initially expect to find shared between populations
-this in Figure \ref{fig:mice_prob_parallel}. The initial length can be
-quite long between geographically close patches (tens of km). However,
-for our wider cline width ($\sigma/\sqrt{s_m} =30km$) adaptation by
-migration can still be likely for patches 100km apart but its shared
-basis may be hard to detect as the length of shared haplotype can be
-quite short.
-
-Given the relatively small size of the dark substrate patches where
-$A$ is often $<10km$ we should expect the haplotype to be rapidly
-whittled down within patches (eqn. \ref{})
+Given the large selection coefficient associated with the dark allele
+on the dark substrate, we expect the initial haplotype associated with
+either a new mutation or migrant allele to be large. 
+We can use equation \eqref{eqn:haplotype_length} to calculate how long the founding
+haplotype shared between populations is expected to be,
+as shown in Figure~\ref{fig:mice_prob_parallel}. 
+The initial length can be quite long between geographically close patches (tens of kilometers). 
+However,
+for the wider cline width ($\sigma/\sqrt{s_m} = 30$km), 
+adaptation by migration can still be likely for patches 100km apart, 
+but the shared basis may be hard to detect, as the length of shared haplotype can be quite short. 
+Furthermore, given the relatively small radius of the patches of dark substrate (often $<10km$),
+we should expect the haplotype to be rapidly whittled down within patches after establishment (equation \ref{}).
 
@@
 and two different mutation rates (using equation \eqref{eqn:prob-parallel}). 
-The parameters were chosen to match  
+The parameters were chosen to match  XXX
   \label{fig:mice_prob_parallel}
@@
 What is the spatial resolution of convergent local adaptation?
-In other words, what spatial scale of environmental patchiness will evolution develop independent solutions to evolutionary problems?
+In other words, 
+over what spatial scale of environmental patchiness will evolution develop independent solutions to evolutionary problems?
 The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
-the dispersal distance divided by the square root of the selection
+the dispersal distance divided by the square root of the strength of selection
 against the new allele between patches. It depends much more weakly on the
-selective benefit within the patches or (perhaps suprisingly) the
-population density. Although these two factors will determine the
-time-scale over which adaptation will occur. This is contrast with models of panmictic populations
-\citep{softsweepsI} and geographically spread populations adapting to homogeneous
-selection pressures \citep{RalphCoop}, where the probability of
-multiple adaptive alleles increases with the population
-size/density. However, in all of these models the dependence on the beneficial selection
+selective benefit within the patches or (perhaps suprisingly) the population density, 
+although these two factors will determine the time-scale over which adaptation will occur. 
+This is in contrast with models of panmictic populations \citep{softsweepsI} 
+and geographically spread populations adapting to homogeneous selection pressures \citep{ralph2010parallel}, 
+where the probability of multiple, independently arising adaptive alleles increases with the population size/density. 
+However, in all of these models the dependence on the beneficial selection
 coefficient is absent or weak, due to the fact that selection both
@@
 
-While weaker selection against alleles will make sharing of
-adaptations between patches easier, 
+We have also shown that
+while weaker selection against alleles will make sharing 
+of adaptations between patches easier, 
 it will also makes it harder to spot the sharing,
-since the lucky, transiting alleles move slower and so carry a shorter
-shared haplotype.
+since the lucky alleles that manage to colonize new patches move slower,
+and thus carry a shorter shared haplotype.
 This issue is amplified by the fact that the length of haplotype
@@
 
-% Observations of similar adaptations occuring within the same species in widely separated patches
-% is strongly suggestive of parallel, independent adaptations.
-% Such observations are very interesting 
-% as they tell us about the degree of constraint faced by evolution --
-% how many different methods are there typically available to a species
-% with which to solve an evolutionary problem?
-
-% It is already well-appreciated that larger populations can lead to more parallelism \citep{softsweepsI}
-% \plr{fisher probably said this too?}
-% and that spatial structure contributes to this \citep{ralph2010parallel}.
-
-% This is complementary to the work of \citet{XXX}, who studied local adaptation in rapidly varying environments.
-% Intermediate cases where habitat types are distributed in some more complex geometry
-% -- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance --
-% remains to be studied.
-
-% To complement this theory, we need more empirical work,
-% in which the parameters (migration rate, sensitivity to environmental factors, etc)
-% are reasonably well-understood.
-% However, learning about these aspects is difficult in the field,
-% and it is still not understood how much of the intraspecific variation we observe
-% is due to adaptation to local conditions, or even how to tell.
-% Notable steps forward include \citet{turner2010serpentine}, Mimulus, Amity, others?
+In this work, we have tried to emphasize an intuitive, individual-centered understanding of the process,
+determining, for instance, 
+the speed at which successful migrants traverse regions where they are maladapted,
+and the effective rate at which new families of migrant alleles appear
+at long distances from a patch.
+These quantities may be as useful for applied work
+as more abstract ones such as the probability that adaptation occurs by migration versus new mutation.
 
@@
 and the question is whether or not adaptations in distinct patches
-have adapted indpendently or not.
+have appeared indpendently or not.
 However, especially if rates of adaptation by new mutation are very low,
 any adaptation that does occur may have to make use of standing variation.
-The unstructured case was studied by \citet{softsweepsI}
-and we study the continuous, spatial case in \citet{toappear}.
-If local, parallel adaptation is from standing variation,
+The unstructured case was studied by \citet{softsweepsI};
+we study the continuous, spatial case in an upcoming paper.
+If parallelism in local adaptation of the sort we study here is due to standing variation
+rather than new mutation,
 then the dynamics of adaptation should not depend strongly on migration patterns
-(but note that the spatial distribution of standing variation will be).
+(but note that the spatial distribution of standing variation may be).
 
@@
 and hence mostly present only in heterozygotes. 
-Our results should hold as a good approximation to dominant and
-partially dominant alleles.
+Our results should hold as a good approximation to dominant and partially dominant alleles
+(with $s_m$ the selection against heterozygotes).
 If, however, the mutation is recessive, then it is essentially neutral where rare,
-and so could have a much easier time spreading between patches.
+and so would encounter much less resistance to spreading between patches.
 This is counteracted, however, by the increased difficulty with which the mutation would establish
-once arriving in a new patch. As
- such it is not clear what our intuition should be about the
- contribution of recessive alleles to adaptation via migration. 
-Further work is needed to understand how recessive alleles may
-contribute to convergent local adaptation.
+once arriving in a new patch. 
+As such it is not clear what our intuition should be 
+about the contribution of recessive alleles to adaptation via migration. 
+Further work is needed to put empirical observations of local adaptation by recessive alleles 
+\plr{cite nachman? but is this published?}
+in a theoretical context.
 
@@
 We have also ignored the possibility of very long distance migration,
-instead focusing local dispersal (as parameterized by a Gaussian
-dispersal model). However, dispersal distributions can be very heavy
-tailed, with a small fraction of individuals moving very long distances indeed, in
-addition over long time-scales very rare chance events (mice sucked up in
-hurricanes etc) could play a role in spreading migrant alleles. Such
-tail events could greatly increase the probability of parallel
-adaptation above that predicted by our model. Furthermore, if
-adaptive alleles do move between distant patches via rare long distance
-migration they will be associated with a much longer shared haplotype than
-predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
-results as a null model by which the contribution of long distribution
-migrants to adaptation could be empirically judged.
-
+instead focusing on local dispersal (and using a Gaussian model).
+However, dispersal distributions can be very heavy tailed, 
+with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
+In addition, over long time-scales, very rare chance events (mice sucked up in hurricanes and the like) 
+could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
+Such tail events could greatly increase the probability of parallel adaptation above that predicted by our model. 
+Furthermore, if adaptive alleles do move between distant patches via rare, long distance migration 
+then they will be associated with a much longer shared haplotype than predicted by equation \eqref{eqn:haplotype_length}. 
+As such, we view our results as a null model by which the contribution of long distribution migrants to adaptation could be empirically judged.
+
+\paragraph{More realistic geometries}
+Here we have studied circular patches of habitat at long distances from each other,
+while real habitat is often much more complex
+-- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance.
+The work of \citet{cantrell1991diffusive} comes closest to a general theory of balanced polymorphisms in such habitats,
+and it is possible that our techniques could be applied in their much more general setting,
+as both are based, fundamentally, on branching process approximations.
+
+\paragraph{Future work}
 The falling cost of population genomic sequencing means that we will soon have the
-opportunity to study the effects of geography and ecology on
-adaptation across many populations within a species. 
-Our work suggests that even quite geographically close populations may
-be forced to adapt convergently when migration is geographically limited. 
-Thus, systems where populations have repeated to strong local
-selection pressures limited may offer the opportunity to study
-highly replicated convergent adaptation within a similar genetic
-background.  
-Such empricial work will also strongly inform our understanding of the role of gene flow in
-keeping ecological similar populations evolving in concert \citep{sexton}.
+opportunity to study the interplay of adaptation with geography and ecology 
+across many populations within a species. 
+Our work suggests that even quite geographically close populations 
+may be forced to adapt convergently when migration is geographically limited. 
+Thus, systems where populations have been repeatedly subject to 
+strong local selection pressures may offer the opportunity 
+to study highly replicated convergent adaptation within a similar genetic background.  
+Such empricial work will also strongly inform our understanding of how gene flow 
+may work to keep ecologically similar populations evolving in concert \citep{sexton2013genetic}.
 Our results suggest that adaptation to shared environments is
-certainly no guarentee of a shared genetic basis to adaptation, suggesting that
-rapid adaptation to a share environment could potentially drive
+certainly no guarantee of a shared genetic basis to adaptation, 
+suggesting that rapid adaptation to a shared environment could potentially drive
 speciation under some scenerios.
-
-% \subsection{GRAHAM'S VERSION}
-
-% We have developed a simple set of approximations to describe the rate
-% at which organisms in a patchy environment adapt through novel
-% mutations or migration from other nearby patches (eqn.
-% \eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
-% probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
-% We have also determined the characteristic genomic length scales over which
-% the various signals of hitchhiking should be visable
-% (eqn. \eqref{eqn:haplotype_length} and ).
-
-% These results show that the probability of parallel adaptation is
-% independent of both the population density and the beneficial
-% selection coefficient. This follows from the fact that these
-% parameters both increase the rate of adaptation via both migration and
-% mutation equally. Thus they are important for the time-scale over
-% which adaptation occurs but not the mode of adaptation. This
-% constrasts with models of convergent adaptation within panmictic
-% populations \citep{} and geographically spread populations adapting to
-% homogeneous populations \citep{RalphCoop}, where the number of
-% adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
-% coefficient is absent or weak, due to the fact that selection both
-% aids establiishment of new alleles and the spread of existing
-% alleles.
-
-% The primary determinant of the probability of parallel adaptation is
-% $\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
-% allele frequency decay off the patch. We find that patches that are close together on
-% this scale will adapt via migration, but there will be a sharp
-% transition to convergent adaptation as we move to patches that are further apart.
-
-% We have also assumed that the alleles have a fitness effect in
-% heterozygotes both on and off the patch. However, 
-%  recessive alleles could also contribute to adaptation to patchy
-%  environments. \gc{should we cite Nachman eg here}.  Our
-%  simple branching process framework does not easily accomodate recessive
-%  alleles as it ignores the selection interaction of alleles, and so a
-%  different approach will be needed. At first sight it
-%  seems like recessive alleles may be able to contribute to adaptation
-%  via migration over longer distances, as the migrant alleles will be at low
-%  frequency between patches and so usually invisible to
-%  selection. However, this fact also means that they are unlikely to
-%  establish on the beneficial patch upon reaching it via migration. As
-%  such it is not clear what our intuition should be about the
-%  contribution of recessive alleles to adaptation via migration.
-
-% \gc{will write a couple more broader paragraphs}
-
-% The falling cost of sequencing over the next decade offers the
-% opportunity to investigate the genomics of parallel adaptation over a
-% large s
-% Over the coming decade population genomic data will be generated
-
-% \plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
-
-% if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
-% recall others' work.
-
-% patchy selection helps this happen; 
-% effective migration rate looks like what; 
-% strength of (positive?) selection suprisingly does not affect results (?)
-
-% standing variation will be important under these circumstances; 
-% migration rate (and so geographic structure) will not affect numbers of types if this is true; 
-% however migration always affcts spatial patterns.
-
-% recessive alleles will have different dynamics. Actually not really
-% obvious what will happen here, cline width will be much wider but
-% prob. of taking off will go as the square of the occupancy density.
 
@@
 % unequal density: Lenormand, gene flow and limits of\dots
-
-% small-scale fluctuations
-% when landscape is not made of discrete (circular) patches
 
@@
 In other cases, we are only given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
-
-We are interested in the following case:
-\[
-    s(x) = \begin{cases}
-        +s_b \qquad \mbox{for } |x|<\epsilon \\
-        -s_d \qquad \mbox{for } x<-\epsilon \; \mbox{or}\; \epsilon \le x \le R \\
-        -s_* \qquad \mbox{for } R < x < R+2\epsilon \\
-        -s_d \qquad \mbox{for } x>R+2\epsilon .
-    \end{cases}
-\]
-We can solve the case with $s_* = s_d$, since $\phi'(0) = 0$ (as done implicitly in \citet{slatkin1973geneflow}).
-It should be possible to argue that if $R$ is large then the value of $s_*$ won't make much difference.
 
@@
 
-Consider a single migrant family $Z$ beginning at $x$.
+Consider a single migrant family $Z$ beginning at $x$,
+and write $S$ for the new patch, which has area $A$ and is centered at $y$.
 Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
-and $\tau_\dagger$ an independent Exponential($\alpha$) time.
-The mean occupation time of $Z$ spent in a region $A$ is 
-\begin{align}
-    u(x,y) &= \E[\int_0^\infty Z_t(A) dt ] \\
-           &= \int_0^\infty \E[Z_t] p_t(x,A) dt \\
-        &= \int_0^\infty e^{-\alpha t} p_t(x,A) dt \\
-        &= \int_0^\infty \P^x\{ \tau_\dagger < t \; \& \; B_t \in A \} dt \\
-        &= \E[ \int_0^{\tau_\dagger} \one_A(B_t) dt ]  .
+and $\tau_\dagger$ an independent Exponential($s_m$) time.
+The mean occupation time of $Z$ spent in a region $S$ is,
+since the marginal distribution of a single lineage is Brownian,
+\begin{align}
+    u(x,S) &= \E[\int_0^\infty Z_t(S) dt ] \\
+           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
+        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
+        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
+        &= \E\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  .
 \end{align}
 If we define
-$\tau_A$ to be the hitting time of $A$ by $B$,
-and $\mu(x,A)$ to be the hitting distribution of $\partial A$ by $B_{\min(t,\tau)}$,
+$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
+and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\min(t,\tau_\dagger)}$,
 by the strong Markov property this is equal to
 \begin{align}
-    u(x,y) &= \E^x[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] \\
-           &= \P^x\{ \tau_A < \tau_\dagger \}  \E^{\mu(x,A)}[ \int_0^{\tau_\dagger} \one_A(B_t) dt ] .
-\end{align}
-If $A$ is circular, the latter expectation will not depend on $x$;
+    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
+           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] ,
+\end{align}
+provided $S$ is not pathological.
+If $S$ is circular with area $A$, the latter expectation will not depend on $x$;
 call this $g(A)$.
 
-Now consider the chance that $Z$ ever hits $A$.
-If we make the approximation that $Z$ hits $A$ iff the trunk of $Z$ does,
+Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
+(In the main body, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
+If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
+and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
 then this is equal to
 \begin{align}
-    F(x,y) &= \int_0^\infty (1-k_e(t)) \P\{ \tau_A \in dt \} \\
-           &= \int_0^\infty e^{\alpha t} (1-k_e(t)) e^{- \alpha t } \P\{ \tau_A < t \} dt \\
-    &=\int_0^\infty e^{\alpha t} (1-k_e(t)) \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &= \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_A < t < \tau_\dagger \} dt \\
-    &= \frac{1}{\E[K]} \P^x[ \tau_A < \tau_\dagger ] .
+    h(x,S) &= \int_0^\infty (1-k_e(t)) \P\{ \tau_S \in dt \} \\
+           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P\{ \tau_S < t \} dt \\
+    &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P\{ \tau_S < t < \tau_\dagger \} dt \\
+    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_S < t < \tau_\dagger \} dt \\
+    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
 \end{align}
@@
 \begin{align}
-    F(x,y) \E[K] 
-    \approx \P^x[ \tau_A < \tau_\dagger ] 
-    = u(x,y) / g(A) .
-\end{align}
-
-Now, 
-\begin{align}
-    q(y)  &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  \\
-          &= g(A) \int_{\partial B(r_0)} \gamma(x) F(x,y) dx \\
-          &= \frac{ g(A) }{ \E[K] } \migrate(y) .
+    h(x,S) \E[K] 
+    \approx \P^x[ \tau_S < \tau_\dagger ] 
+    = u(x,S) / g(A) .
+\end{align}
+
+If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
+then by equation~\eqref{eqn:migrate_integral},
+\begin{align}
+    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
+          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
+          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) .
 \end{align}
@@
 This is integral of the Green function for the Bessel process of the appropriate order, 
-so using Borodin \& Salminen (Appendix 1),
+so using \citet{borodin2002handbook}, % (Appendix 1),
 and letting $w$ be the width of the patch,
@@
 \begin{align}
-    h_1(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-  % &= \int_0^{2w} G_\alpha(0,y) dy \\
-  &= \int_0^{2w} \frac{ e^{- y \sqrt{2\alpha} }}{\sqrt{2\alpha}} dy \\
-  &= (1-e^{-2w\sqrt{2\alpha}})/(2\alpha)
+    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
+  &= \int_0^{2w} \frac{ e^{- y \sqrt{2s_m} }}{\sqrt{2s_m}} dy \\
+  &= (1-e^{-2w\sqrt{2s_m}})/(2s_m)
 \end{align}
@@
 \begin{align}
-    h_2(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-    % &= \int_0^{2w} G_\alpha(0,y) dy \\
-  &= \int_0^{2w} 2 y K_0(y \sqrt{2\alpha}) dy \\
-  &= \frac{1}{\alpha}\int_0^{2w\sqrt{2\alpha}} y K_0(y) dy \\
-  &= \frac{1}{2\alpha}\left(1- 2w\sqrt{2\alpha} K_1(2w\sqrt{2\alpha}) \right)  .
-\end{align}
+    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
+  &= \int_0^{2w} 2 y K_0(y \sqrt{2s_m}) dy \\
+  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}} y K_0(y) dy \\
+  &= \frac{1}{2s_m}\left(1- 2w\sqrt{2s_m} K_1(2w\sqrt{2s_m}) \right)  .
+\end{align}
+In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.
 
 %%%%%%% %%%%%%%%%%%%
-\section{The outflux of families}
+\subsection{The outflux of families}
 \label{apx:outflux}
@@
 Take $y=(r,0)$.
-Note that after scaling by $\sqrt{2\alpha}/\sigma$,
+Note that after scaling by $\sqrt{2s_m}/\sigma$,
 \begin{align}
@@
 where $I_0$ is a modified Bessel function of order zero of the first kind.
-The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2\alpha} r_0 / \sigma ) $.
+The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2s_m} r_0 / \sigma ) $.
 
@@
 Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
-which we assume has relatively small area $A$ and is located near position $y$.
+which we assume has relatively small area $A$ and is centered at position $y$.
 The probability that this family reaches $A$ by time $t$ is $(1-k_e(t))$
 multiplied by the probability that the family reaches the patch, given survival.
-If $t$ is large then the previous section argues that $(1-k_e(t)) \approx e^{-\alpha t}/\E[K]$,
+If $t$ is large then the previous section argues that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
 and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
@@
     \int_0^\infty (1-k_e(t)) \P^x\{ \tau_w \le t \} dt  
-    &\approx \int_0^\infty e^{-\alpha t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\
-    &= \frac{1}{\alpha \E[K]} \int_0^\infty e^{-\alpha t} \P^x\{ \tau_w \in dt \} \\
-    &= \frac{1}{\alpha \E[K]} \E^x[e^{-\alpha \tau_w}] \\
+    &\approx \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\
+    &= \frac{1}{s_m \E[K]} \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \in dt \} \\
+    &= \frac{1}{s_m \E[K]} \E^x[e^{-s_m \tau_w}] \\
     &= 
-    \frac{1}{\alpha \E[K]} \times
+    \frac{1}{s_m \E[K]} \times
     \begin{cases}
-        e^{-(x-w) \sqrt{2\alpha}} \qquad & d=1 \\
-        \frac{ K_0(x\sqrt{2\alpha}) }{ K_0(w\sqrt{2\alpha}) } \qquad & d=2 
+        e^{-(x-w) \sqrt{2s_m}} \qquad & d=1 \\
+        \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } \qquad & d=2 
     \end{cases}
@@
 \begin{align}
-    \frac{ K_0(x\sqrt{2\alpha}) }{ K_0(w\sqrt{2\alpha}) } = \sqrt{\frac{w}{x}} e^{-(x-w)\sqrt{2\alpha}} ( 1 + O(1/w) ).
+    \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } = \sqrt{\frac{w}{x}} e^{-(x-w)\sqrt{2s_m}} ( 1 + O(1/w) ).
 \end{align}

### 2014-06-30 patchy-selection-paper.tex
put in dispersal refs

--- 2014-06-30 patchy-selection-paper.tex
@@
 Convergence on a molecular level also suggests that the path 
-accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
+accessible to adaptation is sometimes relatively constrained, 
 as would be the case if only a few changes confer the advantagous phenotype 
@@
 that have repeatedly adapted to patches of soil with high concentrations of heavy metals
-(e.g.\ serpentine outcrops and mine tailings) \citep{macnair1991evolution,schat1996identical,turner2010serpentine,mimulus};
+(e.g.\ serpentine outcrops and mine tailings) \citep{turner2010serpentine,mimulus};
 the alleles conferring heavy metal tolerance are thought to be locally restricted 
@@
 
+\gc{Graham pasted this to the appendix as it felt like too much information to drop on people}
+\plr{I am sympathetic, but it's only a paragraph\dots how about that last sentence, instad?}
+%%%
 % 
@@
 \begin{align}
-r_{eff} &=
+  r_\text{eff} &=
     \begin{cases}
-    r \frac{ \sigma /\sqrt{s} }{ \sigma/\sqrt{s} + A }    \qquad & d=1 \\
-   r \frac{\pi  \sigma /\sqrt{s}  \left( \sigma /\sqrt{s}+
-  \sqrt{A/\pi} \right)}{A+\pi \sigma/\sqrt{s} \left( \sigma/\sqrt{s} +
-  \sqrt{A/\pi} \right) }    \qquad & d=2 \\  
-\end{cases}
+      r \frac{2\sigma}{A\sqrt{s}}    \qquad & d=1 \\
+      r \frac{\sqrt{\pi} \sigma }{ \sqrt{ s A } } \qquad & d=2 \\  
+\end{cases} ,
 \label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
 \end{align}
-\gc{Altered the above.}
-which scales the recombination rate roughly with $(\sigma/\sqrt{s})A^{-1/d}$.
+i.e.\ the reduction in recombination rate is proportional to $(\sigma/\sqrt{s})A^{-1/d}$.
 We can use this to think about the length of the haplotype
@@
 
-\subsection{PETER'S VERSION}
-
 This paper is an investigation into the basic, ill-formed question: 
@@
 
-\subsection{GRAHAM'S VERSION}
-
-We have developed a simple set of approximations to describe the rate
-at which organisms in a patchy environment adapt through novel
-mutations or migration from other nearby patches (eqn.
-\eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
-probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
-We have also determined the characteristic genomic length scales over which
-the various signals of hitchhiking should be visable
-(eqn. \eqref{eqn:haplotype_length} and ).
-
-These results show that the probability of parallel adaptation is
-independent of both the population density and the beneficial
-selection coefficient. This follows from the fact that these
-parameters both increase the rate of adaptation via both migration and
-mutation equally. Thus they are important for the time-scale over
-which adaptation occurs but not the mode of adaptation. This
-constrasts with models of convergent adaptation within panmictic
-populations \citep{} and geographically spread populations adapting to
-homogeneous populations \citep{RalphCoop}, where the number of
-adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
-coefficient is absent or weak, due to the fact that selection both
-aids establiishment of new alleles and the spread of existing
-alleles.
-
-The primary determinant of the probability of parallel adaptation is
-$\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
-allele frequency decay off the patch. We find that patches that are close together on
-this scale will adapt via migration, but there will be a sharp
-transition to convergent adaptation as we move to patches that are further apart.
-
-In coming to this conclusion, we have ignored the possibility of very long distance migration,
-instead focusing local dispersal (as parameterized by a Gaussian
-dispersal model). However, dispersal distributions can be very heavy
-tailed, with a small fraction of individuals moving very long distances indeed, in
-addition over long time-scales very rare chance events (mice sucked up in
-hurricanes etc) could play a role in spreading migrant alleles. Such
-tail events could greatly increase the probability of parallel
-adaptation above that predicted by our model. Furthermore, if
-adaptive alleles do move between distant patches via rare long distance
-migration they will be associated with a much longer shared haplotype than
-predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
-results as a null model by which the contribution of long distribution
-migrants to adaptation could be empirically judged.
-
-We have also assumed that the alleles have a fitness effect in
-heterozygotes both on and off the patch. However, 
- recessive alleles could also contribute to adaptation to patchy
- environments. \gc{should we cite Nachman eg here}.  Our
- simple branching process framework does not easily accomodate recessive
- alleles as it ignores the selection interaction of alleles, and so a
- different approach will be needed. At first sight it
- seems like recessive alleles may be able to contribute to adaptation
- via migration over longer distances, as the migrant alleles will be at low
- frequency between patches and so usually invisible to
- selection. However, this fact also means that they are unlikely to
- establish on the beneficial patch upon reaching it via migration. As
- such it is not clear what our intuition should be about the
- contribution of recessive alleles to adaptation via migration.
-
-\gc{will write a couple more broader paragraphs}
-
-The falling cost of sequencing over the next decade offers the
-opportunity to investigate the genomics of parallel adaptation over a
-large s
-Over the coming decade population genomic data will be generated
-
-\plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
-
 if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
@@
 To investigate the impact of the various approximations surrounding the mutational influx,
-we calculated the probability of establishment for an spatial branching process moving across discrete demes.
-The solution was obtained by iterating the multivariate generating function until convergence
-(since the function which gives the probability of extinction as a function of location of the progenitor
-is an attractive fixed point of the generating function; see \citet{jagers1975branching}).
-These solutions are shown, and parameters described, in figure~\ref{fig:prob_estab_calcs}.
+we carried out numerical calculations of a fine-grained, discrete model.
+These are shown, and described, in figure~\ref{fig:prob_estab_calcs}.
 
@@
     The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respecively);
-    and the selection coefficient interpolates linearly between these in the central hatched region.
+    and the selection coefficient interpolates linearly between these in the central shaded region.
     The number of offspring is Poisson.
@@
 
-For completeness, and clarity as to the scalings on the relevant parameters,
-here we provide a derivation of the differential equations referred to above equation~\eqref{eqn:eqfreq},
-and establish the asymptotics given in that equation.
 One route to the ``equlibrium frequency'' of the allele outside the range where it is advantageous is as follows;
 see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} for other arguments in this case, 
-or \citet{etheridge2000introduction} and/or \citet{dawson1993measurevalued} for general theory used below.
-Following is a description of a discrete model (which is nearly the one we simulate from).
+or \citet{etheridge2000introduction} or \citet{dawson1993measurevalued} for general theory used below.
 Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
@@
 Each individual at location $x$ reproduces after a random lifetime,
-dying, and then leaving behind in the same deme a random number of offspring with distribution given by $X$;
-each offspring then migrates independently to a new location chosen randomly from the distribution given by $x+R$,
-where she replaces a randomly chosen individual.
-If $x+R$ is outside of the range, then she perishes.
-Each individual's time until reproduction is exponentially distributed:
+dying and giving birth in the same deme to a random number of offspring with distribution given by $X$;
+each offspring migrates independently to a new location chosen randomly from the distribution given by $x+R$,
+where they replaces a randomly chosen individual.
+If $x+R$ is outside of the range, then they perish.
+Each individual's reproduction time is exponentially distributed, 
 the reproduction rate is 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
@@
 and each have variance $\sigma^2$.
-We also assume that the distribution of $R$ makes the associated random walk irreducible and aperiodic.
+We also assume that the distribution of $X$ makes the associated random walk irreducible and aperiodic.
 
@@
 we need to choose $\var[R]=\sigma^2$ and $s(x)$ to be of the same, small, order; 
-this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale \citep[as noted by][]{slatkin1973geneflow}.
+this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale.
 More concretely, suppose that the grid size is $\epsilon \to 0$, 
@@
 We can solve the case with $s_* = s_d$, since $\phi'(0) = 0$ (as done implicitly in \citet{slatkin1973geneflow}).
-It should be possible to argue that if $R$ is large then the value of $s_*$ won't make much difference.
+We should be able to argue that if $R$ is large then the value of $s_*$ won't make much difference.
 
@@
 
-% Presumably, this gets the same answer as \citep{slatkin1973geneflow}, who finds that in the solvable cases,
-% for $x$ in the last interval (which ends in $\infty$), that
-% \begin{align*}
-%     \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / 2 \sigma + C ) \right) \\
-%         &\simeq C' \exp\left( - \frac{ x \sqrt{2s} }{ \sigma } \right) .
-% \end{align*}
+Presumably, this gets the same answer as Monty, who finds that in his solvable cases,
+for $x$ in the last interval (which ends in $\infty$), that
+\begin{align*}
+    \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / 2 \sigma + C ) \right) \\
+        &\simeq C' \exp\left( - \frac{ x \sqrt{2s} }{ \sigma } \right) .
+\end{align*}
+% Indeed, as $\phi \to 0$, we have approximately $\phi' = \frac{ \sqrt{s} }{ \sigma } \phi$.  % huh????
 
@@
 \end{align}
-
-\section{Notes: for intro/discussion}
-
-- leimu2008metaanalysis: local adaptation not ubiquitous but common; mostly in large population sizes.
-- robertson1960theory: dependence of the degree of adaptation on Ne 
-- lots of local adaptation stuff on plants and conservation genetics; ties into plasticity and range limits.
-- but also: Neurospora (ellison2011) and salmonids (fraser2011)
-- swamping in Mayr 1963, Animal Species and Evolution?
-- hanson1966effects: original observation of minimum patch size, in simulation?
-- fourneir-level2011: map of arabidopsis local adaptation
-- reviews for local adapt in marine (sanford2011), trees (savolainen2007)

### 2014-07-01 patchy-selection-paper.tex

@@
 
+\subsection{GRAHAM'S VERSION}
+
+We have developed a simple set of approximations to describe the rate
+at which organisms in a patchy environment adapt through novel
+mutations or migration from other nearby patches (eqn.
+\eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
+probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
+We have also determined the characteristic genomic length scales over which
+the various signals of hitchhiking should be visable
+(eqn. \eqref{eqn:haplotype_length} and ).
+
+These results show that the probability of parallel adaptation is
+independent of both the population density and the beneficial
+selection coefficient. This follows from the fact that these
+parameters both increase the rate of adaptation via both migration and
+mutation equally. Thus they are important for the time-scale over
+which adaptation occurs but not the mode of adaptation. This
+constrasts with models of convergent adaptation within panmictic
+populations \citep{} and geographically spread populations adapting to
+homogeneous populations \citep{RalphCoop}, where the number of
+adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
+coefficient is absent or weak, due to the fact that selection both
+aids establiishment of new alleles and the spread of existing
+alleles.
+
+The primary determinant of the probability of parallel adaptation is
+$\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
+allele frequency decay off the patch. We find that patches that are close together on
+this scale will adapt via migration, but there will be a sharp
+transition to convergent adaptation as we move to patches that are further apart.
+
+In coming to this conclusion, we have ignored the possibility of very long distance migration,
+instead focusing local dispersal (as parameterized by a Gaussian
+dispersal model). However, dispersal distributions can be very heavy
+tailed, with a small fraction of individuals moving very long distances indeed, in
+addition over long time-scales very rare chance events (mice sucked up in
+hurricanes etc) could play a role in spreading migrant alleles. Such
+tail events could greatly increase the probability of parallel
+adaptation above that predicted by our model. Furthermore, if
+adaptive alleles do move between distant patches via rare long distance
+migration they will be associated with a much longer shared haplotype than
+predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
+results as a null model by which the contribution of long distribution
+migrants to adaptation could be empirically judged.
+
+We have also assumed that the alleles have a fitness effect in
+heterozygotes both on and off the patch. However, 
+ recessive alleles could also contribute to adaptation to patchy
+ environments. \gc{should we cite Nachman eg here}.  Our
+ simple branching process framework does not easily accomodate recessive
+ alleles as it ignores the selection interaction of alleles, and so a
+ different approach will be needed. At first sight it
+ seems like recessive alleles may be able to contribute to adaptation
+ via migration over longer distances, as the migrant alleles will be at low
+ frequency between patches and so usually invisible to
+ selection. However, this fact also means that they are unlikely to
+ establish on the beneficial patch upon reaching it via migration. As
+ such it is not clear what our intuition should be about the
+ contribution of recessive alleles to adaptation via migration.
+
+\gc{will write a couple more broader paragraphs}
+
+The falling cost of sequencing over the next decade offers the
+opportunity to investigate the genomics of parallel adaptation over a
+large s
+Over the coming decade population genomic data will be generated
+
+\plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
+
 if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
@@
 \end{align}
+
+\section{Notes: for intro/discussion}
+
+- leimu2008metaanalysis: local adaptation not ubiquitous but common; mostly in large population sizes.
+- robertson1960theory: dependence of the degree of adaptation on Ne 
+- lots of local adaptation stuff on plants and conservation genetics; ties into plasticity and range limits.
+- but also: Neurospora (ellison2011) and salmonids (fraser2011)
+- swamping in Mayr 1963, Animal Species and Evolution?
+- hanson1966effects: original observation of minimum patch size, in simulation?
+- fourneir-level2011: map of arabidopsis local adaptation
+- reviews for local adapt in marine (sanford2011), trees (savolainen2007)

### 2014-07-01 patchy-selection-paper.tex
Graham merged discussions and added a bit more

--- 2014-06-30 patchy-selection-paper.tex
@@
 which has dispersal distance of 0.26km.  Shall we ask Nachman?}
-\plr{He referred to \citet{french1968dispersal,allred1963range}.}
 
@@
 
-This paper is an investigation into the basic, ill-formed question: 
-What is the spatial resolution of adaptation?
-In other words, what scale of environmental patchiness can a species develop and sustain local adaptations,
-and over what spatial scale will evolution develop independent solutions to evolutionary problems?
+This paper is an investigation into the basic question: 
+What is the spatial resolution of convergent local adaptation?
+In other words, what spatial scale of environmental patchiness will evolution develop independent solutions to evolutionary problems?
 The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
-the dispersal distance divided by the square root of the selection against the new allele between patches,
-and much more weakly on the selective benefit within the patches or (perhaps suprisingly) the population density.
-Weaker selection against alleles will make sharing of adaptations between patches easier,
-but will also make it harder to spot,
-since the lucky, transiting alleles move slower and carry a shorter haplotype.
-
-Observations of similar adaptations occuring within the same species in widely separated patches
-is strongly suggestive of parallel, independent adaptations.
-Such observations are very interesting 
-as they tell us about the degree of constraint faced by evolution --
-how many different methods are there typically available to a species
-with which to solve an evolutionary problem?
-
-It is already well-appreciated that larger populations can lead to more parallelism \citep{softsweepsI}
-\plr{fisher probably said this too?}
-and that spatial structure contributes to this \citep{ralph2010parallel}.
-Naturally, then, local adaptation to well-separated patches of some environment is even more likely to occur in parallel;
-we have studied the case where patches are separated by long distances.
-This is complementary to the work of \citet{XXX}, who studied local adaptation in rapidly varying environments.
-Intermediate cases where habitat types are distributed in some more complex geometry
--- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance --
-remains to be studied.
-
-To complement this theory, we need more empirical work,
-in which the parameters (migration rate, sensitivity to environmental factors, etc)
-are reasonably well-understood.
-However, learning about these aspects is difficult in the field,
-and it is still not understood how much of the intraspecific variation we observe
-is due to adaptation to local conditions, or even how to tell.
-Notable steps forward include \citet{turner2010serpentine}, Mimulus, Amity, others?
+the dispersal distance divided by the square root of the selection
+against the new allele between patches. It depends much more weakly on the
+selective benefit within the patches or (perhaps suprisingly) the
+population density. Although these two factors will determine the
+time-scale over which adaptation will occur. This is contrast with models of panmictic populations
+\citep{softsweepsI} and geographically spread populations adapting to homogeneous
+selection pressures \citep{RalphCoop}, where the probability of
+multiple adaptive alleles increases with the population
+size/density. However, in all of these models the dependence on the beneficial selection
+coefficient is absent or weak, due to the fact that selection both
+aids establishment of new alleles and the spread of existing
+alleles.
+
+While weaker selection against alleles will make sharing of
+adaptations between patches easier, 
+it will also makes it harder to spot the sharing,
+since the lucky, transiting alleles move slower and so carry a shorter
+shared haplotype.
+This issue is amplified by the fact that the length of haplotype
+shared within patches decays over time, potentially making the
+identification of shared adaptations to old selection pressures difficult.
+
+% Observations of similar adaptations occuring within the same species in widely separated patches
+% is strongly suggestive of parallel, independent adaptations.
+% Such observations are very interesting 
+% as they tell us about the degree of constraint faced by evolution --
+% how many different methods are there typically available to a species
+% with which to solve an evolutionary problem?
+
+% It is already well-appreciated that larger populations can lead to more parallelism \citep{softsweepsI}
+% \plr{fisher probably said this too?}
+% and that spatial structure contributes to this \citep{ralph2010parallel}.
+
+% This is complementary to the work of \citet{XXX}, who studied local adaptation in rapidly varying environments.
+% Intermediate cases where habitat types are distributed in some more complex geometry
+% -- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance --
+% remains to be studied.
+
+% To complement this theory, we need more empirical work,
+% in which the parameters (migration rate, sensitivity to environmental factors, etc)
+% are reasonably well-understood.
+% However, learning about these aspects is difficult in the field,
+% and it is still not understood how much of the intraspecific variation we observe
+% is due to adaptation to local conditions, or even how to tell.
+% Notable steps forward include \citet{turner2010serpentine}, Mimulus, Amity, others?
 
@@
 This is counteracted, however, by the increased difficulty with which the mutation would establish
-once arriving in a new patch.
-On the other hand, a dominant mutation would not differ greatly from the theory presented here.
-In fact, MC1r is ?dominant? and agouti is ?recessive?
-Other schemes could be obtained.
-
-optimal $s_m$ for migration makes migration harder to spot
-
-if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
-recall others' work.
-
-patchy selection helps this happen; 
-effective migration rate looks like what; 
-strength of (positive?) selection suprisingly does not affect results (?)
-
-standing variation will be important under these circumstances; 
-migration rate (and so geographic structure) will not affect numbers of types if this is true; 
-however migration always affcts spatial patterns.
-
-recessive alleles will have different dynamics. Actually not really
-obvious what will happen here, cline width will be much wider but
-prob. of taking off will go as the square of the occupancy density.
-
-discussion of applications.
-
-discuss mixing of types; 
-Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
-
-unequal density: Lenormand, gene flow and limits of\dots
-
-small-scale fluctuations
-when landscape is not made of discrete (circular) patches
+once arriving in a new patch. As
+ such it is not clear what our intuition should be about the
+ contribution of recessive alleles to adaptation via migration. 
+Further work is needed to understand how recessive alleles may
+contribute to convergent local adaptation.
+
+\paragraph{Long distance migration}
+We have also ignored the possibility of very long distance migration,
+instead focusing local dispersal (as parameterized by a Gaussian
+dispersal model). However, dispersal distributions can be very heavy
+tailed, with a small fraction of individuals moving very long distances indeed, in
+addition over long time-scales very rare chance events (mice sucked up in
+hurricanes etc) could play a role in spreading migrant alleles. Such
+tail events could greatly increase the probability of parallel
+adaptation above that predicted by our model. Furthermore, if
+adaptive alleles do move between distant patches via rare long distance
+migration they will be associated with a much longer shared haplotype than
+predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
+results as a null model by which the contribution of long distribution
+migrants to adaptation could be empirically judged.
+
+The falling cost of population genomic sequencing means that we will soon have the
+opportunity to study the effects of geography and ecology on
+adaptation across many populations within a species. 
+Our work suggests that even quite geographically close populations may
+be forced to adapt convergently when migration is geographically limited. 
+Thus, systems where populations have repeated to strong local
+selection pressures limited may offer the opportunity to study
+highly replicated convergent adaptation within a similar genetic
+background.  
+Such empricial work will also strongly inform our understanding of the role of gene flow in
+keeping ecological similar populations evolving in concert \citep{sexton}.
+Our results suggest that adaptation to shared environments is
+certainly no guarentee of a shared genetic basis to adaptation, suggesting that
+rapid adaptation to a share environment could potentially drive
+speciation under some scenerios.
+
+% \subsection{GRAHAM'S VERSION}
+
+% We have developed a simple set of approximations to describe the rate
+% at which organisms in a patchy environment adapt through novel
+% mutations or migration from other nearby patches (eqn.
+% \eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
+% probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
+% We have also determined the characteristic genomic length scales over which
+% the various signals of hitchhiking should be visable
+% (eqn. \eqref{eqn:haplotype_length} and ).
+
+% These results show that the probability of parallel adaptation is
+% independent of both the population density and the beneficial
+% selection coefficient. This follows from the fact that these
+% parameters both increase the rate of adaptation via both migration and
+% mutation equally. Thus they are important for the time-scale over
+% which adaptation occurs but not the mode of adaptation. This
+% constrasts with models of convergent adaptation within panmictic
+% populations \citep{} and geographically spread populations adapting to
+% homogeneous populations \citep{RalphCoop}, where the number of
+% adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
+% coefficient is absent or weak, due to the fact that selection both
+% aids establiishment of new alleles and the spread of existing
+% alleles.
+
+% The primary determinant of the probability of parallel adaptation is
+% $\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
+% allele frequency decay off the patch. We find that patches that are close together on
+% this scale will adapt via migration, but there will be a sharp
+% transition to convergent adaptation as we move to patches that are further apart.
+
+% We have also assumed that the alleles have a fitness effect in
+% heterozygotes both on and off the patch. However, 
+%  recessive alleles could also contribute to adaptation to patchy
+%  environments. \gc{should we cite Nachman eg here}.  Our
+%  simple branching process framework does not easily accomodate recessive
+%  alleles as it ignores the selection interaction of alleles, and so a
+%  different approach will be needed. At first sight it
+%  seems like recessive alleles may be able to contribute to adaptation
+%  via migration over longer distances, as the migrant alleles will be at low
+%  frequency between patches and so usually invisible to
+%  selection. However, this fact also means that they are unlikely to
+%  establish on the beneficial patch upon reaching it via migration. As
+%  such it is not clear what our intuition should be about the
+%  contribution of recessive alleles to adaptation via migration.
+
+% \gc{will write a couple more broader paragraphs}
+
+% The falling cost of sequencing over the next decade offers the
+% opportunity to investigate the genomics of parallel adaptation over a
+% large s
+% Over the coming decade population genomic data will be generated
+
+% \plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
+
+% if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
+% recall others' work.
+
+% patchy selection helps this happen; 
+% effective migration rate looks like what; 
+% strength of (positive?) selection suprisingly does not affect results (?)
+
+% standing variation will be important under these circumstances; 
+% migration rate (and so geographic structure) will not affect numbers of types if this is true; 
+% however migration always affcts spatial patterns.
+
+% recessive alleles will have different dynamics. Actually not really
+% obvious what will happen here, cline width will be much wider but
+% prob. of taking off will go as the square of the occupancy density.
+
+% discussion of applications.
+
+% discuss mixing of types; 
+% Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
+
+% unequal density: Lenormand, gene flow and limits of\dots
+
+% small-scale fluctuations
+% when landscape is not made of discrete (circular) patches
 

### 2014-07-02 patchy-selection-paper.tex

@@
 Such precise convergence points to the conservation of function of the genes underlying adaptive traits, 
-sometimes over deep time scales \citep{deephomologypapers}. 
+sometimes over deep time scales \citep{zhen2012parallel,martin2013repeated,deephomologypapers}. 
 Convergence on a molecular level also suggests that the path 
@@
 Such evolution of convergent phenotypes is favored by higher mutational input, 
-i.e.\ larger mutational target and/or population size \citep{dmitri}.
+i.e.\ larger mutational target and/or population size \citep{softsweepsII}.
 The geographic distribution of populations can also affect the probability of parallel mutation within a species:
@@
 Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
-and in lizards \citep{whitesands}.
+and in lizards \citep{rosenblum2010molecular}.
 
@@
 pressure \citep{maruyama1970fixation,cherry2003diffusion}. 
+More general work on the mathematically equivalent problem of density-dependent population growth 
+has obtained much more general criteria
+under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1991diffusive},
+and determined the habitat configuration that maximizes the probability of establishment \citep{lou2006minimization}.
 
@@
 the ``variance effective population size'' of the patch
-(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1978pain,bartonetheridgedepaulis}).
+(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1975pain,barton2002neutral}).
 The rate of recombination and coalescence will be on the same order if

### 2014-07-02 patchy-selection-paper.tex
changes to intro and apps

--- 2014-07-01 patchy-selection-paper.tex
@@
 or even by precisely the same genetic changes \citep{zhen2012parallel}. 
-Such precise convergence points to the conservation of function of the genes underlying adaptive traits, 
-sometimes over deep time scales \citep{deephomologypapers}. 
-Convergence on a molecular level also suggests that the path 
-accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
-as would be the case if only a few changes confer the advantagous phenotype 
-and are sufficiently free of deleterious pleiotropic consequences.
-
-%For highly polygenic traits
-
 Convergent adaptation can also occur within species, if individuals within a species 
-adapt to the same environment, in parallel, through different genotypic changes. 
+adapt to the same environment, in parallel, through different genetics changes. 
 There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
@@
 and when is it likely that migration carries these alleles between patches?
+
 How could we tell, after the fact, which has occurred?
+
 We work in a model of continuous geography,
 using a variety of known results and new methods.
-In section XX we \dots QUICK OUTLINE
+In section \ref{ss:patchymutation} we develop a simple approximation,
+eqn. \eqref{eqn:mutrate}, for the rate at which patches become
+established by new mutations. Then in section \ref{ss:patchymigration} we
+develop an approximation, eqn. \eqref{eqn:migrate}, to the rate which alleles move from an
+existing patch to colonize a novel patch (this represents a big chunk
+of the results of the paper). We combine these two results in section
+\ref{ss:probparallel} to develop an approximation to the probability of
+parallel adaptation, eqn. \eqref{}.
+To understand the genomic signal left by adaptations shared by
+migration in section \ref{ss:haplotype_length} we approximate the time
+it will take for an allele to transit between patches,
+eqn. \eqref{eqn:mean_tau}, and use this to approximate the length of
+haplotype that hitchhikes with it,
+eqn. \eqref{eqn:haplotype_length}. We then quantify how quickly this
+haplotype decays within patches eqn. \eqref{} in section \ref{ss:patch_haplotype}.
+
+ {ss:patch_haplotype}
+
+{ss:haplotype_length}
+l{ss:probparallel}
+\dots QUICK OUTLINE
 
@@
 % \end{equation}
-\plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
-and you shrink it to have radius $r_0$,
-then the resulting object has area $A \times (r_0/r)^d$;
-here we have $r_0 = r - \sigma/\sqrt{s}$,
-in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
+%\plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
+%and you shrink it to have radius $r_0$,
+%then the resulting object has area $A \times (r_0/r)^d$;
+%here we have $r_0 = r - \sigma/\sqrt{s}$,
+%in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
 \begin{align}
@@
 \end{align}
-\gc{Altered the above.}
-which scales the recombination rate roughly with $(\sigma/\sqrt{s})A^{-1/d}$.
-We can use this to think about the length of the haplotype
-associated with the $B$ allele after a given length of time $T$,
-for instance, the time since the $B$ allele joined the patch.
-\plr{changed this to genetic lengths so as not to introduce another parameter, $r_{BP}$}
-% If the recombination rate per base is $r_{BP}$, 
-% then the length of the haplotype associated with the $B$ allele is
+so that $T$ generations after the founding of a patch we expect a
+haplotype of genetic length $1/(r_{eff}T)$ Morgans to remain linked to
+the $B$ for a typical lineage. Note that the length of haplotype
+decays at a rate roughly proportional to $\sqrt{s}/ \sigma$, as the narrower the cline the less
+opportunity for recombination to move our lineage off the $B$ background.
+
+% which scales the recombination rate roughly with $(\sigma/\sqrt{s})A^{-1/d}$.
+% We can use this to think about the length of the haplotype
+% associated with the $B$ allele after a given length of time $T$,
+% for instance, the time since the $B$ allele joined the patch.
+% \plr{changed this to genetic lengths so as not to introduce another parameter, $r_{BP}$}
+% % If the recombination rate per base is $r_{BP}$, 
+% % then the length of the haplotype associated with the $B$ allele is
+% % \begin{equation}
+% % \frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
+% % \end{equation} 
+% In units of genetic length (Morgans),
+% the length of the haplotype associated with the $B$ allele is, up to a constant,
 % \begin{equation}
-% \frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
+%   \frac{A^{1/d}\sqrt{s}}{\sigma T}
 % \end{equation}
-In units of genetic length (Morgans),
-the length of the haplotype associated with the $B$ allele is, up to a constant,
-\begin{equation}
-  \frac{A^{1/d}\sqrt{s}}{\sigma T}
-\end{equation} 
-Note that the length of haplotype decays at a rate proportional to the
-cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
-opportunity for recombination to move our lineage off the $B$ background.
 
@@
 %%%%%%%%%%
-\section{Applications} 
+\section{Application} 
 Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, 
@@
 The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
-suggesting these populations have adapted in parallel \citep{xxx}.
+suggesting these populations have adapted in parallel \citep{nachman2003different}.
 However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, 
@@
 We also need estimate of $s_m$. 
-Using $\sigma=1km$ \gc{REF}, these cline widths imply $s_m=1/\sqrt{3}$ and
+Using $\sigma \approx 1km$ \gc{REF}, these cline widths imply $s_m=1/\sqrt{3}$ and
 $1/\sqrt{30}$ are reasonable values.
-\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
-which has dispersal distance of 0.26km.  Shall we ask Nachman?}
+%\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
+%which has dispersal distance of 0.26km.  Shall we ask Nachman?}
 
@@
 being able to arise at any of a thousand base pairs.  
-Finally, we set $A=100\text{km}^2$.
+Finally, we set $A=100\text{km}^2$ (roughly thr size of the Pinacate patch).
 Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
@@
 on the dark substrate we expect the initial sweep associated with
-either a new mutation or migrant allele to be large ($s/\log(\rho A
-s)$). We can use eqn \eqref{eqn:haplotype_length} to calculate how long a
-founding haplotype we initially expect to find shared between populations
-this in Figure \ref{fig:mice_prob_parallel}. The initial length can be
+either a new mutation or migrant allele to be large. We can use eqn
+\eqref{eqn:haplotype_length} to calculate how long a founding
+haplotype we initially expect to find shared between populations this in Figure \ref{fig:mice_prob_parallel}. The initial length can be
 quite long between geographically close patches (tens of km). However,
@@
 basis may be hard to detect as the length of shared haplotype can be
-quite short.
-
-Given the relatively small size of the dark substrate patches where
-$A$ is often $<10km$ we should expect the haplotype to be rapidly
-whittled down within patches (eqn. \ref{})
+quite short. Furthermore, given the relatively small radius of the dark substrate
+patches (often $<10km$) we should expect the haplotype to be rapidly
+whittled down within patches after establishment (eqn. \ref{})
 
@@
 Such empricial work will also strongly inform our understanding of the role of gene flow in
-keeping ecological similar populations evolving in concert \citep{sexton}.
+keeping ecological similar populations evolving in concert \citep{Sexton:14}.
 Our results suggest that adaptation to shared environments is

### 2014-07-02 patchy-selection-paper.tex
peter edited through discussion

--- 2014-07-02 patchy-selection-paper.tex
@@
 and, perhaps as a result of this strong selection against migrants, 
-at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{nachman2003different}. 
+at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{hoekstra2003different}. 
 Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
@@
 has obtained much more general criteria
-under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1991diffusive},
+under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1989diffusive},
 and determined the habitat configuration that maximizes the probability of establishment \citep{lou2006minimization}.
@@
 surrounded by a cloud of close relatives of random size $K$
-whose chance of surviving until $t$ generations is $k_e(t)$.
+whose chance of dying out by $t$ generations is $k_e(t)$.
 Consider a single such family, and let $\tau$ be the (random) time at which it first hits the new patch,
@@
 \end{align}
-\gc{Altered the above.}
 which scales the recombination rate roughly with $(\sigma/\sqrt{s})A^{-1/d}$.
@@
 for instance, the time since the $B$ allele joined the patch.
-\plr{changed this to genetic lengths so as not to introduce another parameter, $r_{BP}$}
 % If the recombination rate per base is $r_{BP}$, 
@@
 Finally we can consider the expected length of haplotype shared within a patch. 
-To do this, we compare the rate of recombination between B and b haplotypes (equation~\ref{eqn:rate-rec-in-patch})
-to the rate of coalescence of B haplotypes within a patch. 
+To do this, we compare the rate of recombination between $B$ and $b$ haplotypes (equation~\ref{eqn:rate-rec-in-patch})
+to the rate of coalescence of $B$ haplotypes within a patch. 
 We will assume that mixing within the patch is fast compared to the rate of drift, 
-so that the rate of coalescence of B alleles within a patch is 
+so that the rate of coalescence of $B$ alleles within a patch is 
 $\xi^2/(A\rho)$,
@@
 Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
-of parallel mutation on the distance between lava flow patches using these parameters. 
-As can be seen parallel mutation should become
+of parallel mutation on the distance between lava flow patches using these parameters, 
+showing that parallel mutation should become
 likely over a scale of tens to a few hundred kilometers between patches.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{Lava_flow_mice_prob_parallel}
+  \end{center}
+  \caption{
+The probability of parallel mutation as a function of the distance
+between environmental patches for two different cline widths
+and two different mutation rates (using equation \eqref{eqn:prob-parallel}). 
+The parameters were chosen to match  XXX
+  \label{fig:mice_prob_parallel}
+  }
+\end{figure}
 
@@
 
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics{Lava_flow_mice_prob_parallel}
-  \end{center}
-  \caption{
-The probability of parallel mutation as a function of the distance
-between environmental patches for two different cline widths
-and two different mutation rates (using equation \eqref{eqn:prob-parallel}). 
-The parameters were chosen to match  
-  \label{fig:mice_prob_parallel}
-  }
-\end{figure}
-
 %%%%%%% %%%%%%% %%%%%%%
@@
 We have also ignored the possibility of very long distance migration,
-instead focusing local dispersal (as parameterized by a Gaussian
-dispersal model). However, dispersal distributions can be very heavy
-tailed, with a small fraction of individuals moving very long distances indeed, in
-addition over long time-scales very rare chance events (mice sucked up in
-hurricanes etc) could play a role in spreading migrant alleles. Such
-tail events could greatly increase the probability of parallel
-adaptation above that predicted by our model. Furthermore, if
-adaptive alleles do move between distant patches via rare long distance
-migration they will be associated with a much longer shared haplotype than
-predicted by eqn. \eqref{eqn:haplotype_length}. As such we view our
-results as a null model by which the contribution of long distribution
-migrants to adaptation could be empirically judged.
-
+instead focusing on local dispersal (and using a Gaussian model).
+However, dispersal distributions can be very heavy tailed, 
+with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
+In addition, over long time-scales, very rare chance events (mice sucked up in hurricanes and the like) 
+could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
+Such tail events could greatly increase the probability of parallel adaptation above that predicted by our model. 
+Furthermore, if adaptive alleles do move between distant patches via rare, long distance migration 
+then they will be associated with a much longer shared haplotype than predicted by equation \eqref{eqn:haplotype_length}. 
+As such, we view our results as a null model by which the contribution of long distribution migrants to adaptation could be empirically judged.
+
+\paragraph{More realistic geometries}
+Here we have studied circular patches of habitat at long distances from each other,
+while real habitat is often much more complex
+-- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance.
+The work of \citet{cantrell1991diffusive} comes closest to a general theory of balanced polymorphisms in such habitats,
+and it is possible that our techniques could be applied in their much more general setting,
+as both are based, fundamentally, on branching process approximations.
+
+\paragraph{Future work}
 The falling cost of population genomic sequencing means that we will soon have the
-opportunity to study the effects of geography and ecology on
-adaptation across many populations within a species. 
-Our work suggests that even quite geographically close populations may
-be forced to adapt convergently when migration is geographically limited. 
-Thus, systems where populations have repeated to strong local
-selection pressures limited may offer the opportunity to study
-highly replicated convergent adaptation within a similar genetic
-background.  
-Such empricial work will also strongly inform our understanding of the role of gene flow in
-keeping ecological similar populations evolving in concert \citep{sexton}.
+opportunity to study the interplay of adaptation with geography and ecology 
+across many populations within a species. 
+Our work suggests that even quite geographically close populations 
+may be forced to adapt convergently when migration is geographically limited. 
+Thus, systems where populations have been repeatedly subject to 
+strong local selection pressures may offer the opportunity 
+to study highly replicated convergent adaptation within a similar genetic background.  
+Such empricial work will also strongly inform our understanding of how gene flow 
+may work to keep ecologically similar populations evolving in concert \citep{sexton2013genetic}.
 Our results suggest that adaptation to shared environments is
-certainly no guarentee of a shared genetic basis to adaptation, suggesting that
-rapid adaptation to a share environment could potentially drive
+certainly no guarantee of a shared genetic basis to adaptation, 
+suggesting that rapid adaptation to a shared environment could potentially drive
 speciation under some scenerios.
-
-% \subsection{GRAHAM'S VERSION}
-
-% We have developed a simple set of approximations to describe the rate
-% at which organisms in a patchy environment adapt through novel
-% mutations or migration from other nearby patches (eqn.
-% \eqref{eqn:mutrate} and \eqref{eqn:migrate}), and the relative
-% probability of each of these outcomes (eqn. \eqref{eqn:parallel_prob}). 
-% We have also determined the characteristic genomic length scales over which
-% the various signals of hitchhiking should be visable
-% (eqn. \eqref{eqn:haplotype_length} and ).
-
-% These results show that the probability of parallel adaptation is
-% independent of both the population density and the beneficial
-% selection coefficient. This follows from the fact that these
-% parameters both increase the rate of adaptation via both migration and
-% mutation equally. Thus they are important for the time-scale over
-% which adaptation occurs but not the mode of adaptation. This
-% constrasts with models of convergent adaptation within panmictic
-% populations \citep{} and geographically spread populations adapting to
-% homogeneous populations \citep{RalphCoop}, where the number of
-% adaptive alleles increases with the population size/density. In all of these models the dependence on the beneficial selection
-% coefficient is absent or weak, due to the fact that selection both
-% aids establiishment of new alleles and the spread of existing
-% alleles.
-
-% The primary determinant of the probability of parallel adaptation is
-% $\sigma/\sqrt{s_d}$, the characteristic geographic distance of the
-% allele frequency decay off the patch. We find that patches that are close together on
-% this scale will adapt via migration, but there will be a sharp
-% transition to convergent adaptation as we move to patches that are further apart.
-
-% We have also assumed that the alleles have a fitness effect in
-% heterozygotes both on and off the patch. However, 
-%  recessive alleles could also contribute to adaptation to patchy
-%  environments. \gc{should we cite Nachman eg here}.  Our
-%  simple branching process framework does not easily accomodate recessive
-%  alleles as it ignores the selection interaction of alleles, and so a
-%  different approach will be needed. At first sight it
-%  seems like recessive alleles may be able to contribute to adaptation
-%  via migration over longer distances, as the migrant alleles will be at low
-%  frequency between patches and so usually invisible to
-%  selection. However, this fact also means that they are unlikely to
-%  establish on the beneficial patch upon reaching it via migration. As
-%  such it is not clear what our intuition should be about the
-%  contribution of recessive alleles to adaptation via migration.
-
-% \gc{will write a couple more broader paragraphs}
-
-% The falling cost of sequencing over the next decade offers the
-% opportunity to investigate the genomics of parallel adaptation over a
-% large s
-% Over the coming decade population genomic data will be generated
-
-% \plr{Note that optimal $s_m$ for migration makes migration harder to spot.}
-
-% if populations and mutation rates are large enough, parallel adaptation is likely, aided by geography. 
-% recall others' work.
 

### 2014-07-02 Lava_flow_mice_app.R

@@ -11,8 +11,8 @@ R = seq(5,600,length=10000) ##distance in Km
 #layout(t(1:2))
 mu = 1e-8
 prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
-plot(R, prob.parallel,type="l",lwd=3, xlab="", ylab="",cex.lab=1.3,cex.axis=1.3,log="x") 
-mtext(side=2,line=2.5,"prob parallel mutation",cex=1.3)
+plot(R, prob.parallel,type="l",lwd=3, xlab="", ylab="",log="x") 
+mtext(side=2,line=2.5,"prob parallel mutation")
 mu=1e-5
 prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
 
@@ -31,16 +31,16 @@ lines(R, prob.parallel,lwd=3,col="red",lty=2)
 
 legend("topleft", legend=expression(mu == 10^-8, mu == 10^-5, sigma/sqrt(s[m]) == 3,sigma/sqrt(s[m]) == 30),col=c("black","red","grey","grey"),lty=c(1,1,1,2),lwd=2)
 
-mtext(side=1,line=2.5,"Distance between patches (km)",cex=1.3)
+mtext(side=1,line=2.5,"Distance between patches (km)")
 
 
 ####haplotype lengths
 ##assuming sigma=1
 length.hap.1<-sqrt(2)*1/(char.length[1])/R +1/R^2
 length.hap.2<-sqrt(2)*1/(char.length[2])/R +1/R^2
-plot(R,length.hap.1*100,type="l",log="x",ylab="", xlab="",lty=1,lwd=2,cex.lab=1.3,cex.axis=1.3)
-mtext(side=2,line=2.5,"Shared Haplotype Length (cM)",cex=1.3)
-mtext(side=1,line=2.5,"Distance between patches (km)",cex=1.3)
+plot(R,length.hap.1*100,type="l",log="x",ylab="", xlab="",lty=1,lwd=2)
+mtext(side=2,line=2.5,"shared haplotype length (cM)")
+mtext(side=1,line=2.5,"distance between patches (km)")
 lines(R,length.hap.2*100,lty=2,lwd=2)
 
 legend("topright", legend=expression( sigma/sqrt(s[m]) == 3,sigma/sqrt(s[m]) == 30),col=c("grey","grey"),lty=c(1,2),lwd=2)

### 2014-07-02 patchy-selection-paper.tex

--- 2014-07-02 patchy-selection-paper.tex
@@
 genetic changes in the same pathway, genes, 
-or even by precisely the same genetic changes \citep{zhen2012parallel}. 
+or even by precisely the same genetic changes \citep{zhen2012parallel,,martin2013repeated,deephomologypapers}. 
 Convergent adaptation can also occur within species, if individuals within a species 
@@
 $1/\sqrt{30}$ are reasonable values.
-%\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
-%which has dispersal distance of 0.26km.  Shall we ask Nachman?}
+
+\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
+which has dispersal distance of 0.26km.  Shall we ask Nachman?}
+\plr{He referred to \citet{french1968dispersal,allred1963range}.}
 

### 2014-07-02 patchy-selection-paper.tex
addition to applications

--- 2014-07-02 patchy-selection-paper.tex
@@
 genetic changes in the same pathway, genes, 
-or even by precisely the same genetic changes \citep{zhen2012parallel}. 
-Such precise convergence points to the conservation of function of the genes underlying adaptive traits, 
-sometimes over deep time scales \citep{zhen2012parallel,martin2013repeated,deephomologypapers}. 
-Convergence on a molecular level also suggests that the path 
-accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
-as would be the case if only a few changes confer the advantagous phenotype 
-and are sufficiently free of deleterious pleiotropic consequences.
-
-%For highly polygenic traits
-
+or even by precisely the same genetic changes \citep{zhen2012parallel,,martin2013repeated,deephomologypapers}. 
 Convergent adaptation can also occur within species, if individuals within a species 
-adapt to the same environment, in parallel, through different genotypic changes. 
+adapt to the same environment, in parallel, through different genetics changes. 
 There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
@@
 comes from patch $i$ is 
-\begin{align}  \label{eqn:prob-parallel}
+\begin{align}  
   \frac{\left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
@@
 surrounded by a cloud of close relatives of random size $K$
-whose chance of dying out by $t$ generations is $k_e(t)$.
+whose chance of surviving until $t$ generations is $1-k_e(t)$.
 Consider a single such family, and let $\tau$ be the (random) time at which it first hits the new patch,
@@
   \sqrt{A/\pi} \right)}{A+\pi \sigma/\sqrt{s} \left( \sigma/\sqrt{s} +
-  \sqrt{A/\pi} \right) }    \qquad & d=2 \\  
+  \sqrt{A/\pi} \right) }    \qquad & d=2 ,  
 \end{cases}
-\label{eqn:rate-rec-in-patch}   %(\sigma/\sqrt{s})A^{-1/d}
-\end{align}
-which scales the recombination rate roughly with $(\sigma/\sqrt{s})A^{-1/d}$.
-We can use this to think about the length of the haplotype
-associated with the $B$ allele after a given length of time $T$,
-for instance, the time since the $B$ allele joined the patch.
-% If the recombination rate per base is $r_{BP}$, 
-% then the length of the haplotype associated with the $B$ allele is
-% \begin{equation}
-% \frac{\sqrt{s}}{\sigma A^{-1/d}}\frac{1}{r_{BP}}
-% \end{equation} 
-In units of genetic length (Morgans),
-the length of the haplotype associated with the $B$ allele is, up to a constant,
-\begin{equation}
-  \frac{A^{1/d}\sqrt{s}}{\sigma T}
-\end{equation} 
-Note that the length of haplotype decays at a rate proportional to the
-cline width $\sqrt{s} / \sigma$, as the narrower the cline the less
+\label{eqn:rec_rate_within}   %(\sigma/\sqrt{s})A^{-1/d}
+\end{align}
+so that $T$ generations after the founding of a patch we expect a
+haplotype of genetic length $1/(r_\text{eff}T)$ Morgans to remain linked to
+the $B$ allele in a typical lineage. 
+Note that the length of haplotype
+decays at a rate roughly proportional to $\sqrt{s}/ \sigma$, as the narrower the cline the less
 opportunity for recombination to move our lineage off the $B$ background.
@@
 Finally we can consider the expected length of haplotype shared within a patch. 
-To do this, we compare the rate of recombination between $B$ and $b$ haplotypes (equation~\ref{eqn:rate-rec-in-patch})
+To do this, we compare the rate of recombination between $B$ and $b$ haplotypes (equation~\ref{eqn:rec_rate_within})
 to the rate of coalescence of $B$ haplotypes within a patch. 
@@
 and hence the genetic length of haplotype shared between pairs of
-individuals within the patches is on the order of
-\begin{align}
+individuals within the patches at equilibrium is on the order of
+\begin{align} \label{eqn:within_haplotype_length}
   r = \frac{\sqrt{s} \xi^2}{\sigma \rho A^{(d-1)/d}} 
@@
 \section{Applications}
+
+% FOR REFERENCE:
+% On Wed, May 28, 2014 at 10:20 AM, Michael Nachman <mnachman@berkeley.edu> wrote:
+% > Graham:
+% >
+% > The density is probably in the range of 10-40 animals per hectare (10,000
+% > m^s).  We built enclosures of 0.25 hectare and found that they could support
+% > about 10 animals with food provided ad libitum, so I think this is an upper
+% > density (i.e. 40/hectare).  As for dispersal distance, I'm less sure.
+% > Studies on closely related species (genus Perognathus) in similar habitats
+% > suggest distances of 100's of feet.  Attached are some old papers that tried
+% > to estimate dispersal distances in Perognathus.
+% > best,
+% > Michael
+
 Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, 
@@
 A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. 
-\citep{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
+\citet{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
 at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
@@
 We also need estimate of $s_m$. 
-Using $\sigma=1km$ \gc{REF}, these cline widths imply $s_m=1/\sqrt{3}$ and
-$1/\sqrt{30}$ are reasonable values.
-\plr{Going back to previous conversations, we came upon \citet{mullen2008natural}, 
-which has dispersal distance of 0.26km.  Shall we ask Nachman?}
-\plr{He referred to \citet{french1968dispersal,allred1963range}.}
-
-The mutational target size $\mu$ for the trait is unclear. 
-While the Pinacate dark haplotype differs from the light haplotype at three amino-acid residues,
-it is likely that not all of these changes are needed for a population to begin to adapt. 
-Also there a number of genes besides MC1R at which adaptative changes affecting pigmentation 
-have been identified in closely related species and more broadly across vertebrates \citep{REF}.
-To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
-where the adaptive mutation has to occur at a single base pair; 
-and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation
-being able to arise at any of a thousand base pairs.  
-Finally, we set $A=100\text{km}^2$.
-Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
-of parallel mutation on the distance between lava flow patches using these parameters, 
-showing that parallel mutation should become
-likely over a scale of tens to a few hundred kilometers between patches.
+Using $\sigma \approx 1$km \citep{french1968dispersal,allred1963range}, 
+these cline widths imply that $s_m=1/\sqrt{3}$ and $1/\sqrt{30}$ are reasonable values.
 
@@
 between environmental patches for two different cline widths
-and two different mutation rates (using equation \eqref{eqn:prob-parallel}). 
-The parameters were chosen to match  XXX
+and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
+The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
+see text.
   \label{fig:mice_prob_parallel}
@@
 
-Givn the large selection coefficient associated with the dark allele
-on the dark substrate we expect the initial sweep associated with
-either a new mutation or migrant allele to be large ($s/\log(\rho A
-s)$). We can use eqn \eqref{eqn:haplotype_length} to calculate how long a
-founding haplotype we initially expect to find shared between populations
-this in Figure \ref{fig:mice_prob_parallel}. The initial length can be
-quite long between geographically close patches (tens of km). However,
-for our wider cline width ($\sigma/\sqrt{s_m} =30km$) adaptation by
-migration can still be likely for patches 100km apart but its shared
-basis may be hard to detect as the length of shared haplotype can be
-quite short.
-
-Given the relatively small size of the dark substrate patches where
-$A$ is often $<10km$ we should expect the haplotype to be rapidly
-whittled down within patches (eqn. \ref{})
+The mutational target size $\mu$ for the trait is unclear. 
+While the Pinacate dark haplotype differs from the light haplotype at three amino acid residues,
+it is likely that not all of these changes are needed for a population to begin to adapt. 
+Also there a number of genes besides MC1R at which adaptative changes affecting pigmentation 
+have been identified in closely related species and more broadly across vertebrates \citep{Hoekstra:06}.
+To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
+where the adaptive mutation has to occur at a single base pair; 
+and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation
+being able to arise at any of a thousand base pairs.  
+Finally, we set $A=100\text{km}^2$ (roughly the size of the Pinacate patch).
+Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
+of parallel mutation on the distance between lava flow patches using these parameters, 
+showing that parallel mutation should become
+likely over a scale of tens to a few hundred kilometers between patches.
+
+Given the large selection coefficient associated with the dark allele
+on the dark substrate, we expect the initial haplotype associated with
+either a new mutation or migrant allele to be large. 
+We can use equation \eqref{eqn:haplotype_length} to calculate how long the founding
+haplotype shared between populations is expected to be,
+as shown in Figure~\ref{fig:mice_prob_parallel}. 
+The initial length can be quite long between geographically close patches (tens of kilometers). 
+However, for the wider cline width ($\sigma/\sqrt{s_m} = 30$km), 
+adaptation by migration can still be likely for patches 100km apart, 
+but the shared basis may be hard to detect, as the length of shared haplotype can be quite short.
+
+Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
+we should expect the haplotype to be rapidly whittled down within patches after establishment (equation \ref{eqn:rec_rate_within}).
+Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare (M.~Nachman, personal communication),
+and the variance in offspring number to be $\xi^2 = 3$ (mostly to get round numbers),
+then the shared haplotype length within patches from equation \eqref{eqn:within_haplotype_length}
+is only expected to be on the order of hundreds or thousands of bases
+($10^{-3}$--$10^{-4}$cM).
+This suggests that unless the migration has happened fairly recently \plr{age of the lava flows?}
+then it will be difficult to identify shared haplotypes
+(how recently depends on the background level of LD in the population).
 
@@
 
-% patchy selection helps this happen; 
-% effective migration rate looks like what; 
-% strength of (positive?) selection suprisingly does not affect results (?)
-
-% standing variation will be important under these circumstances; 
-% migration rate (and so geographic structure) will not affect numbers of types if this is true; 
-% however migration always affcts spatial patterns.
-
-% recessive alleles will have different dynamics. Actually not really
-% obvious what will happen here, cline width will be much wider but
-% prob. of taking off will go as the square of the occupancy density.
-
 % discussion of applications.

### 2014-07-02 patchy-selection-paper.tex
small changes throughout

@@
 \title{
-Convergent Adaptation while adapting 
+The Probability of Convergent Adaptation to Environmental Patches. 
 }
@@
 Standing variation for a trait can also increase the probability of convergence, 
-and standing variation for an allele increases the probability that the selective sweep will be \emph{soft}
+as standing variation for an allele increases the probability that the selective sweep will be \emph{soft}
 (beginning from a base of multiple copies),
@@
 Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration
-to particular substrates, e.g.\ dark rock outcrops or white sand dunes \citep{benson1933concealing}.
-\plr{we are missing what Nachman said was the original ref for this... I think you took a picture?}
+to particular substrates, e.g.\ dark rock outcrops or white sand dunes
+\citep{benson1933concealing, DiceBlossom1937}.
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
-which is primarily dark on a number of black lava flows separated by much lighter soil \citep{dice1940ecologic}.
+which is primarily dark on a number of black lava flows separated by
+much lighter soil \citep{dice1940ecologic, DiceBlossom1937}.
 Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
@@
 and when is it likely that migration carries these alleles between patches?
-
-How could we tell, after the fact, which has occurred?
+How easy will it be to detect adaptive alleles that are shared by
+migration, i.e. over what genomic scale will a population genetic
+signal be visible?
 
@@
 eqn. \eqref{eqn:haplotype_length}. We then quantify how quickly this
-haplotype decays within patches eqn. \eqref{} in section \ref{ss:patch_haplotype}.
-
-\dots QUICK OUTLINE
-
-\gc{Lets also give the numbers of key equations in this road map so
-  folks can understand the biologically important results and then work
-  their way back through math, once they know where we are going}
-
-  new mutation: \eqref{eqn:mutrate}; 
- via migration  \eqref{eqn:migrate}
-
-Expected transit time \eqref{eqn:mean_tau}
-Expected haplotype length \eqref{eqn:haplotype_length}
+haplotype decays within patches eqn. \eqref{eqn:rec_rate_within } and
+use this to obtain properties about the equilibrium distribution of
+haplotype lengths
+\eqref{eqn:within_haplotype_length}  in section \ref{ss:patch_haplotype}. 
+Finally in section \ref{ss:applications} we apply this work to understand the geographic scale of
+convergent evolution in \textit{Chaetodipus intermedius}.
 
@@
 has gone extinct by time $t$, 
-\plr{switched from $p_e(t)$ to $k_e(t)$ to not conflict with prob of estab.  better suggestion?}
+%\plr{switched from $p_e(t)$ to $k_e(t)$ to not conflict with prob of estab.  better suggestion?}
 and let $K_t$ be the (random) family size conditioned on nonextinction,
@@
 Furthermore, the probability that the second adaptation is a new mutation,
-i.e.\ the probability of parallel adapatation, is in $d=1$ ($d=2$ only differs slightly)
+i.e.\ the probability of parallel adapatation, is 
 \begin{equation} \label{eqn:parallel_prob}
-  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu s_m }{2 A \mu s_m + \exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
+  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu s_m }{2 A \mu s_m + \left(r \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
 \end{equation}
@@
 
-\paragraph{Island model}
 The resulting discrete process is quite general,
 but if we oversimplify back to the island model, it is tractable:
-suppose the probability of adaptaton by mutation is the same for each patch
-and by migration is the same for each pair of patches.
+suppose the probability of adaptation by mutation is the same for each patch
+and by distance is the same for each pair of patches.
 The resulting process is a continuous-time version of the ``Chinese restaurant process''
 described in \citet{aldous1985exchangeability} and \citet{pitman1995partitions},
-and so the final partition of types across demes has the Ewens sampling distribution with parameter $\mu/m$.
-
+and so the final partition of types across demes has the Ewens
+sampling distribution with parameter $A \mu s_m/ \exp (-sqrt{2 s_m}r/\sigma)$.
 Now we can compute most properties we might want about the process.
@@
 where the local nature of dispersal means that doubling the species range will double the number of mutations expected.
-Furthermore, while in the continuous model the areas occupied by distinct mutations are similar in area, 
-in this strong-selection discrete island model, 
-only a few types tend to dominate even as the number of demes (and thus, distinct types) grows large.
 
@@
 %%%%%%%%%%
-\section{Applications}
-
+\subsection{Applications} 
+\label{ss:applications}
 % FOR REFERENCE:
@@
 Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, 
-\emph{Chaetodipus intermedius}, building on classic work by \citet{benson1933concealing} and \citet{dice1940ecologic}.
+\emph{Chaetodipus intermedius}, building on classic work by
+\citet{benson1933concealing}, \citet{DiceBlossom1937} and \citet{dice1940ecologic}.
 These mice live on rocky outcrops throughout the southwestern United States and northern Mexico, 
@@
 Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
-we should expect the haplotype to be rapidly whittled down within patches after establishment (equation \ref{eqn:rec_rate_within}).
+we should expect the haplotype to be rapidly whittled down within
+patches after establishment (equation \ref{eqn:rec_rate_within}). We
+estimate that the effective recombination rate is at most only 50 \% lower
+than the actual recombination rate.
 Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare (M.~Nachman, personal communication),
@@
 ($10^{-3}$--$10^{-4}$cM).
-This suggests that unless the migration has happened fairly recently \plr{age of the lava flows?}
+This suggests that unless the migration has happened fairly recently
 then it will be difficult to identify shared haplotypes
@@
 then the dynamics of adaptation should not depend strongly on migration patterns
-(but note that the spatial distribution of standing variation may be).
+(but note that the initial spatial distribution of standing variation may be).
 

### 2014-07-02 patchy-selection-paper.tex
small changes throughout

@@
 The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
-suggesting these populations have adapted in parallel \citep{hoekstra2003different}.
+suggesting these populations have adapted in parallel \citep{nachman2003genetic ,hoekstra2003different}.
 However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, 
@@
 
-\paragraph{More realistic geometries}
+\paragraph{Other patches geometries}
 Here we have studied circular patches of habitat at long distances from each other,
-while real habitat is often much more complex
+while real habitat is can be much more complex
 -- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance.
@@
 
-\paragraph{Future work}
+\paragraph{Concluding thoughts}
 The falling cost of population genomic sequencing means that we will soon have the

### 2014-07-02 patchy-selection-paper.tex
began abstract

@@
 \usepackage{color}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 %\usepackage{latexml} % for \iflatexml
@@
 
+% The Spatial Resolution of Convergent Adaptation
+% Convergent Evolution in Patchy Landscapes
+% Parallel Evolution in Patchy Landscapes
+
 \title{
-The Probability of Convergent Adaptation to Environmental Patches. 
+The Probability of Convergent Evolution During Local Adaptation. 
 }
@@
 \begin{document}
+
+\maketitle
+\begin{center}
+$^1$ Computational Biology and Bioinformatics \\ University of
+Southern California, Los Angeles, CA
+$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
+\end{center}
+
+\abstract{
+Habitats are often heterogeneous,
+so that the selective pressures encountered by a species across its range
+are usually not uniform.
+Large enough patches of habitat with divergenct selective pressures
+can lead to local adaptation,
+i.e.\ a balanced polymorphism at loci responding to the selective pressure
+-- a number of well-studied examples concern cryptic coloration in small mammals
+whose range extends across several colors of substrate.
+If there is a similar selective pressure in multiple patches,
+then 
+}
 
@@
 
-The convergent evolution of similar phenotypes in response to shared
-selection pressures is a testiment to the power of selection to
-repeatedly sculpt phenotypic variation. 
+The convergent evolution of similar phenotypes 
+in response to shared selection pressures is a testiment to the power of selection 
+to repeatedly sculpt phenotypic variation. 
 In some cases this convergence extends to the molecular level, 
@@
 genetic changes in the same pathway, genes, 
-or even by precisely the same genetic changes \citep{zhen2012parallel,,martin2013repeated,deephomologypapers}. 
+or even by precisely the same genetic changes
+\citep{stern2013genetic}. 
+\plr{want to also cite \citep{zhen2012parallel,martin2013repeated}?}
 Convergent adaptation can also occur within species, if individuals within a species 
 adapt to the same environment, in parallel, through different genetics changes. 
-There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
+There are a growing number of examples of this in a range of well studied organisms and phenotypes \cite{Arendt:08}.
 Such evolution of convergent phenotypes is favored by higher mutational input, 
@@
 
+%\citep{wright2013indirect}
 One elegant set of such examples is provided by the assortment of plant species 
@@
 the alleles conferring heavy metal tolerance are thought to be locally restricted 
-because they incur a cost off of these patches \citep{wright2013indirect}.
+because they incur a cost off of these patches.
 Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration
 to particular substrates, e.g.\ dark rock outcrops or white sand dunes
-\citep{benson1933concealing, DiceBlossom1937}.
+\citep{benson1933concealing}.
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
-which is primarily dark on a number of black lava flows separated by
-much lighter soil \citep{dice1940ecologic, DiceBlossom1937}.
+which is on a number of black lava flows has much darker pelage than on intervening,
+much lighter soil \citep{DiceBlossom1937,dice1940ecologic}.
 Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
 and, perhaps as a result of this strong selection against migrants, 
-at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{hoekstra2003different}. 
+at least two distinct genetic changes are responsible for the dark pigmentation found on different outcrops \citep{hoekstra2003different}. 
 Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
@@
 \ref{ss:probparallel} to develop an approximation to the probability of
-parallel adaptation, eqn. \eqref{}.
+parallel adaptation, eqn. \eqref{eqn:parallel_prob}.
 To understand the genomic signal left by adaptations shared by
@@
 eqn. \eqref{eqn:haplotype_length}. We then quantify how quickly this
-haplotype decays within patches eqn. \eqref{eqn:rec_rate_within } and
+haplotype decays within patches eqn. \eqref{eqn:rec_rate_within} and
 use this to obtain properties about the equilibrium distribution of
@@
 For the sake of simplicity, we assume a constant haploid population density $\rho$, 
-across both types of habitat (but return to this point later),
+across both types of habitat,
 We further assume that the variance in offspring number is $\xi^2$, 
 and that the mean squared distance between parent and child is $\sigma^2$
-(i.e.\ $\sigma$ is the dispersal distance).
+(i.e.\ $\sigma$ is the dispersal distance). Differences in population
+densities on and off the patches may arise if selection is hard, in
+that case these differences can be incorporated into the density \citep[see][]{lenormand2002limits}. 
 We also aim to describe the scenario where it is sensible to think of large, discrete patches;
-the case of a fine-scale heterogeneous environment is treated \citep{elsewhere}.
+the case of a fine-scale heterogeneous environment is treated
+\citep{elsewhere}. \gc{IS THIS LAST SENTENCE NEEDED IF SO GIVE CITATIOn}
 
@@
 % (7+log(1/(.5*ss)))/sqrt(2*(.5*ss))
+
+\gc{NOT SURE IF ANYONE HAS LOOKED AT THI IN LONG TIME?}
 If we take the mutational influx in a patch to be one per 1,000 generations ($\rho A \mu = .001$)
@@
 the haplotype that swept initially will have decayed due to this process of recombination. 
-However, we still expect to find individuals within the patch sharing long haplotypes around the selected locus
+However, we may still expect to find individuals within the patch sharing long haplotypes around the selected locus
 than with individuals elsewhere,
@@
 decreases mean coalescence times within the patch near the selected locus.
-\citep{bartonpaper,charlesworths}.
+%\citep{bartonpaper,charlesworths}.
 
@@
 The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
-suggesting these populations have adapted in parallel \citep{nachman2003genetic ,hoekstra2003different}.
-However, there is some evidence \citep{Hoekstra:05} that, elsewhere in the range, 
+suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
+However, there is some evidence \citep{hoekstra2004local} that, elsewhere in the range, 
 multiple dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
@@
 Further work is needed to put empirical observations of local adaptation by recessive alleles 
-\plr{cite nachman? but is this published?}
 in a theoretical context.
@@
 strong local selection pressures may offer the opportunity 
-to study highly replicated convergent adaptation within a similar genetic background.  
+to study highly replicated convergent adaptation within a similar
+genetic background \citep{stern2013genetic}.  
 Such empricial work will also strongly inform our understanding of how gene flow 
@@
 suggesting that rapid adaptation to a shared environment could potentially drive
-speciation under some scenerios.
+speciation under some scenerios \citep[in a manner similar to ][]{Kondrashov:03}.
 

### 2014-07-03 patchy-selection-paper.tex
small changes throughout

@@
 \usepackage{color}
-\usepackage[hidelinks]{hyperref}
+\usepackage{hyperref}
 %\usepackage{latexml} % for \iflatexml
@@
 
-% The Spatial Resolution of Convergent Adaptation
-% Convergent Evolution in Patchy Landscapes
-% Parallel Evolution in Patchy Landscapes
-
 \title{
@@
 \begin{document}
-
-\maketitle
-\begin{center}
-$^1$ Computational Biology and Bioinformatics \\ University of
-Southern California, Los Angeles, CA
-$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
-\end{center}
-
-\abstract{
-Habitats are often heterogeneous,
-so that the selective pressures encountered by a species across its range
-are usually not uniform.
-Large enough patches of habitat with divergenct selective pressures
-can lead to local adaptation,
-i.e.\ a balanced polymorphism at loci responding to the selective pressure
--- a number of well-studied examples concern cryptic coloration in small mammals
-whose range extends across several colors of substrate.
-If there is a similar selective pressure in multiple patches,
-then 
-}
 
@@
 
-The convergent evolution of similar phenotypes 
-in response to shared selection pressures is a testiment to the power of selection 
-to repeatedly sculpt phenotypic variation. 
+The convergent evolution of similar phenotypes in response to shared
+selection pressures is a testiment to the power of selection to
+repeatedly sculpt phenotypic variation. 
 In some cases this convergence extends to the molecular level, 
@@
 or even by precisely the same genetic changes
-\citep{stern2013genetic}. 
-\plr{want to also cite \citep{zhen2012parallel,martin2013repeated}?}
+\citep{Stern:13}. 
 Convergent adaptation can also occur within species, if individuals within a species 
@@
 to particular substrates, e.g.\ dark rock outcrops or white sand dunes
-\citep{benson1933concealing}.
+\citep{benson1933concealing, DiceBlossom1937}.
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
-which is on a number of black lava flows has much darker pelage than on intervening,
-much lighter soil \citep{DiceBlossom1937,dice1940ecologic}.
+which is primarily dark on a number of black lava flows separated by
+much lighter soil \citep{dice1940ecologic, DiceBlossom1937}.
 Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
 and, perhaps as a result of this strong selection against migrants, 
-at least two distinct genetic changes are responsible for the dark pigmentation found on different outcrops \citep{hoekstra2003different}. 
+at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{hoekstra2003different}. 
 Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
@@
 The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
-suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
+suggesting these populations have adapted in parallel \citep{nachman2003genetic ,hoekstra2003different}.
 However, there is some evidence \citep{hoekstra2004local} that, elsewhere in the range, 
@@
 to study highly replicated convergent adaptation within a similar
-genetic background \citep{stern2013genetic}.  
+genetic background \citep{Stern:13}.  
 Such empricial work will also strongly inform our understanding of how gene flow 
@@
 suggesting that rapid adaptation to a shared environment could potentially drive
-speciation under some scenerios \citep[in a manner similar to ][]{Kondrashov:03}.
+speciation under some scenerios \citep[in a manner similar to ][]{}.
 

### 2014-07-03 patchy-selection-paper.tex
small changes throughout

@@
 suggesting that rapid adaptation to a shared environment could potentially drive
-speciation under some scenerios \citep[in a manner similar to ][]{}.
+speciation under some scenerios \citep[in a manner similar to ][]{Kondrashov:03}.
 

### 2014-07-03 patchy-selection-paper.tex
small changes throughout

@@
 \begin{document}
+
+\maketitle
+\begin{center}
+$^1$ Computational Biology and Bioinformatics \\ University of
+Southern California, Los Angeles, CA
+$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
+\end{center}
 
@@
 The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
-suggesting these populations have adapted in parallel \citep{nachman2003genetic ,hoekstra2003different}.
+suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
 However, there is some evidence \citep{hoekstra2004local} that, elsewhere in the range, 

### 2014-07-03 patchy-selection-paper.tex
abstract

@@
 
+% The Spatial Resolution of Convergent Adaptation
+% Convergent Evolution in Patchy Landscapes
+% Parallel Evolution in Patchy Landscapes
+
 \title{
@@
 \begin{center}
-$^1$ Computational Biology and Bioinformatics \\ University of
-Southern California, Los Angeles, CA
-$^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
+$^1$ Computational Biology and Bioinformatics \\ 
+University of Southern California, Los Angeles, CA \\
+$^2$Center for Population Biology \& Department of Evolution and Ecology \\ 
+University of California -- Davis, Davis, CA, 95616
 \end{center}
 
+%%%%%%%%%%%
+\abstract{
+When the same phenotype is found to be responding to selection in parallel in isolated parts of a species range,
+it provides strong evidence as to the degree of constraint imposed on evolution by the underlying genetic architecture.
+Migration can counfound such inferences by synchronizing adaptation across the range;
+but the importance of such sharing of local adaptations is unclear.
+This paper studies the balance between these two in continuous environments with patchy selection pressures,
+such as when an animal's coloration is selected to match its background,
+but the species range contains large areas of differently colored substrate.
+The main questions answered are: 
+How long does it take for a new, locally adapted allele to appear in a patch of habitat through new mutation?
+Or, through migration from another, already adapted patch?
+Which is more likely to occur, as a function of patch separation?
+How can we tell which has occurred?
+These are answered by decomposing the migration--selection equilibrium
+surrounding an already adapted patch
+into families of migrant alleles,
+in particular treating those rare families that reach new patches
+as spatial branching processes.
+The speed at which successful migrant alleles traverse regions where they are maladapted
+determines the length of the haplotype shared by around shared, adaptive alleles,
+which is compared to the length of the haplotype shared within a patch.
+Cryptic coloration in the rock pocket mouse, \textit{Chaetodipus intermedius}, is discussed as an empirical example.
+
+}
+
+%%%%%%%%%%% %%%%%%%%%%% 
 \section{Introduction}
@@
 or even by precisely the same genetic changes
-\citep{Stern:13}. 
+\citep{stern2013genetic}. 
+\plr{want to also cite \citep{zhen2012parallel,martin2013repeated}?}
 Convergent adaptation can also occur within species, if individuals within a species 
@@
 to particular substrates, e.g.\ dark rock outcrops or white sand dunes
-\citep{benson1933concealing, DiceBlossom1937}.
+\citep{benson1933concealing}.
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
-which is primarily dark on a number of black lava flows separated by
-much lighter soil \citep{dice1940ecologic, DiceBlossom1937}.
+which is on a number of black lava flows has much darker pelage than on intervening,
+much lighter soil \citep{DiceBlossom1937,dice1940ecologic}.
 Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
 and, perhaps as a result of this strong selection against migrants, 
-at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops \citep{hoekstra2003different}. 
+at least two distinct genetic changes are responsible for the dark pigmentation found on different outcrops \citep{hoekstra2003different}. 
 Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
@@
 to study highly replicated convergent adaptation within a similar
-genetic background \citep{Stern:13}.  
+genetic background \citep{stern2013genetic}.  
 Such empricial work will also strongly inform our understanding of how gene flow 

### 2014-07-03 patchy-selection-paper.tex
small changes throughout

--- 2014-07-03 patchy-selection-paper.tex
@@
 $^1$ Computational Biology and Bioinformatics \\ University of
-Southern California, Los Angeles, CA
+Southern California, Los Angeles, CA\\
 $^2$Center for Population Biology \& Department of Evolution and Ecology \\ University of California -- Davis. Davis, CA, 95616
@@
 %%%%%%%%%%%
+
+\end{document}
+
+\section{Temporary work: hitting probs}
+
+Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
+which we assume has relatively small area $A$ and is centered at position $y$.
+The probability that this family reaches $A$ by time $t$ is $(1-k_e(t))$
+multiplied by the probability that the family reaches the patch, given survival.
+If $t$ is large then the previous section argues that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
+and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
+will have hit the patch by then.
+If $w = A^{1/d}$ is the width of the patch, then 
+this is the chance that a $d$-dimensional Bessel process begun at $\|y-x\|$ hits $w$ by time $t$.
+Let $\tau_w$ be the hitting time of $w$,
+so, integrating by parts, and using Borodin \& Salminen (2.2.0.1 and 4.2.0.1)
+\begin{align}
+    \int_0^\infty (1-k_e(t)) \P^x\{ \tau_w \le t \} dt  
+    &\approx \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\
+    &= \frac{1}{s_m \E[K]} \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \in dt \} \\
+    &= \frac{1}{s_m \E[K]} \E^x[e^{-s_m \tau_w}] \\
+    &= 
+    \frac{1}{s_m \E[K]} \times
+    \begin{cases}
+        e^{-(x-w) \sqrt{2s_m}} \qquad & d=1 \\
+        \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } \qquad & d=2 
+    \end{cases}
+\end{align}
+Note that the $d=2$ case
+\begin{align}
+    \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } = \sqrt{\frac{w}{x}} e^{-(x-w)\sqrt{2s_m}} ( 1 + O(1/w) ).
+\end{align}
+
+\section{Notes: for intro/discussion}
+
+- leimu2008metaanalysis: local adaptation not ubiquitous but common; mostly in large population sizes.
+- robertson1960theory: dependence of the degree of adaptation on Ne 
+- lots of local adaptation stuff on plants and conservation genetics; ties into plasticity and range limits.
+- but also: Neurospora (ellison2011) and salmonids (fraser2011)
+- swamping in Mayr 1963, Animal Species and Evolution?
+- hanson1966effects: original observation of minimum patch size, in simulation?
+- fourneir-level2011: map of arabidopsis local adaptation
+- reviews for local adapt in marine (sanford2011), trees (savolainen2007)
+
 \section{Patchy selection} 
@@
 and that the case presented above is at the extreme (see also \cite{Navarro:03} for deterministic approximations).
-
-\end{document}
-
-\section{Temporary work: hitting probs}
-
-Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
-which we assume has relatively small area $A$ and is centered at position $y$.
-The probability that this family reaches $A$ by time $t$ is $(1-k_e(t))$
-multiplied by the probability that the family reaches the patch, given survival.
-If $t$ is large then the previous section argues that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
-will have hit the patch by then.
-If $w = A^{1/d}$ is the width of the patch, then 
-this is the chance that a $d$-dimensional Bessel process begun at $\|y-x\|$ hits $w$ by time $t$.
-Let $\tau_w$ be the hitting time of $w$,
-so, integrating by parts, and using Borodin \& Salminen (2.2.0.1 and 4.2.0.1)
-\begin{align}
-    \int_0^\infty (1-k_e(t)) \P^x\{ \tau_w \le t \} dt  
-    &\approx \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\
-    &= \frac{1}{s_m \E[K]} \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \in dt \} \\
-    &= \frac{1}{s_m \E[K]} \E^x[e^{-s_m \tau_w}] \\
-    &= 
-    \frac{1}{s_m \E[K]} \times
-    \begin{cases}
-        e^{-(x-w) \sqrt{2s_m}} \qquad & d=1 \\
-        \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } \qquad & d=2 
-    \end{cases}
-\end{align}
-Note that the $d=2$ case
-\begin{align}
-    \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } = \sqrt{\frac{w}{x}} e^{-(x-w)\sqrt{2s_m}} ( 1 + O(1/w) ).
-\end{align}
-
-\section{Notes: for intro/discussion}
-
-- leimu2008metaanalysis: local adaptation not ubiquitous but common; mostly in large population sizes.
-- robertson1960theory: dependence of the degree of adaptation on Ne 
-- lots of local adaptation stuff on plants and conservation genetics; ties into plasticity and range limits.
-- but also: Neurospora (ellison2011) and salmonids (fraser2011)
-- swamping in Mayr 1963, Animal Species and Evolution?
-- hanson1966effects: original observation of minimum patch size, in simulation?
-- fourneir-level2011: map of arabidopsis local adaptation
-- reviews for local adapt in marine (sanford2011), trees (savolainen2007)

### 2014-07-03 patchy-selection-paper.tex

--- 2014-07-03 patchy-selection-paper.tex
@@
 \title{
-The Probability of Convergent Evolution During Local Adaptation. 
+Convergent Evolution During Local Adaptation to Patchy Landscapes
 }
@@
 \begin{center}
+
 $^1$ Computational Biology and Bioinformatics \\ 
@@
 \abstract{
-When the same phenotype is found to be responding to selection in parallel in isolated parts of a species range,
-it provides strong evidence as to the degree of constraint imposed on evolution by the underlying genetic architecture.
-Migration can counfound such inferences by synchronizing adaptation across the range;
-but the importance of such sharing of local adaptations is unclear.
-This paper studies the balance between these two in continuous environments with patchy selection pressures,
-such as when an animal's coloration is selected to match its background,
-but the species range contains large areas of differently colored substrate.
-The main questions answered are: 
-How long does it take for a new, locally adapted allele to appear in a patch of habitat through new mutation?
+Populations within a species often repeatedly locally adapt to similar environmental conditions
+across their range. Such adaptation can occur through convergent
+evolution as different alleles arise and spread in different parts of
+the range. Or it can occur through the spread of alleles by migration
+acting to synchronizing adaptation across the species.
+The tension between the two reflects the degree of 
+constraint imposed on evolution by the underlying genetic
+architecture versus how how effectively selection acts to inhibt the geographic spread of local
+adapted alleles.
+This paper studies a model of the the balance between these two routes to adaptation in continuous environments with patchy selection pressures.
+We address the following questions: 
+How long does it take for a new, locally adapted allele to appear in a
+patch of favorable habitat through new mutation?
 Or, through migration from another, already adapted patch?
 Which is more likely to occur, as a function of patch separation?
-How can we tell which has occurred?
-These are answered by decomposing the migration--selection equilibrium
+How can we tell which has occurred, i.e. what population genetic
+signal is left by the spread of migrant alleles and how long does this
+signal persist for?
+To answer these questions we decompose the migration--selection equilibrium
 surrounding an already adapted patch
@@
 as spatial branching processes.
-The speed at which successful migrant alleles traverse regions where they are maladapted
-determines the length of the haplotype shared by around shared, adaptive alleles,
-which is compared to the length of the haplotype shared within a patch.
-Cryptic coloration in the rock pocket mouse, \textit{Chaetodipus intermedius}, is discussed as an empirical example.
+This provides a set of tractable approximations to understand the role
+of geographic separation between favorable patches in promoting
+convergent adaptation and the genomic signals left.
+We illustrate these ideas using the convergent evolution of cryptic coloration in the rock pocket mouse, \textit{Chaetodipus intermedius}, as an empirical example.
 
@@
 or even by precisely the same genetic changes
-\citep{stern2013genetic}. 
-\plr{want to also cite \citep{zhen2012parallel,martin2013repeated}?}
+\citep{stern2013genetic,zhen2012parallel,martin2013repeated}. 
 Convergent adaptation can also occur within species, if individuals within a species 
@@
 that case these differences can be incorporated into the density \citep[see][]{lenormand2002limits}. 
-We also aim to describe the scenario where it is sensible to think of large, discrete patches;
-the case of a fine-scale heterogeneous environment is treated
-\citep{elsewhere}. \gc{IS THIS LAST SENTENCE NEEDED IF SO GIVE CITATIOn}
+%We also aim to describe the scenario where it is sensible to think of large, discrete patches;
+%the case of a fine-scale heterogeneous environment is treated
+%\citep{elsewhere}. \gc{IS THIS LAST SENTENCE NEEDED IF SO GIVE CITATIOn}
 
@@
 
-\gc{NOT SURE IF ANYONE HAS LOOKED AT THI IN LONG TIME?}
+\gc{NOT SURE IF ANYONE HAS LOOKED AT THIS NEXT PARAGRAPH IN LONG TIME?}
 If we take the mutational influx in a patch to be one per 1,000 generations ($\rho A \mu = .001$)
@@
 
-In this work, we have tried to emphasize an intuitive, individual-centered understanding of the process,
-determining, for instance, 
-the speed at which successful migrants traverse regions where they are maladapted,
-and the effective rate at which new families of migrant alleles appear
-at long distances from a patch.
-These quantities may be as useful for applied work
-as more abstract ones such as the probability that adaptation occurs by migration versus new mutation.
+%commented out as wasn't quite sure what purpose this paragraph served.
+% In this work, we have tried to emphasize an intuitive, individual-centered understanding of the process,
+% determining, for instance, 
+% the speed at which successful migrants traverse regions where they are maladapted,
+% and the effective rate at which new families of migrant alleles appear
+% at long distances from a patch.
+% These quantities may be as useful for applied work
+% as more abstract ones such as the probability that adaptation occurs by migration versus new mutation.
+
+In developing the set of approximations in the paper we have ignored a
+number of complicating factors. We now briefly discuss these.
 
@@
 Our work suggests that even quite geographically close populations 
-may be forced to adapt convergently when migration is geographically limited. 
+may be forced to locally adapt convergently when migration is geographically limited. 
 Thus, systems where populations have been repeatedly subject to 
@@
 genetic background \citep{stern2013genetic}.  
-Such empricial work will also strongly inform our understanding of how gene flow 
-may work to keep ecologically similar populations evolving in concert \citep{sexton2013genetic}.
+Such empricial work will also strongly inform our understanding of the
+ability of gene flow to keep ecologically similar populations evolving
+in concert \citep{sexton2013genetic}.
 Our results suggest that adaptation to shared environments is
@@
 suggesting that rapid adaptation to a shared environment could potentially drive
-speciation under some scenerios \citep[in a manner similar to ][]{Kondrashov:03}.
+speciation if the alleles that spread in each population fail to work
+well together \citep{Kondrashov:03}.
 

### 2014-07-03 patchy-selection-paper.tex
figure legend

@@
 adapt to the same environment, in parallel, through different genetics changes. 
-There are a growing number of examples of this in a range of well studied organisms and phenotypes \cite{Arendt:08}.
+There are a growing number of examples of this in a range of well studied organisms and phenotypes \citep{Arendt:08}.
 Such evolution of convergent phenotypes is favored by higher mutational input, 
@@
   \caption{
-The probability of parallel mutation as a function of the distance
+Top panel: The probability of parallel mutation as a function of the distance
 between environmental patches for two different cline widths
@@
 The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
-see text.
+see text. Bottom panel: The initial genetic length of the haplotype
+shared between patches due to adaptation via migration as a function
+of the distance between environmental patches. 
   \label{fig:mice_prob_parallel}

### 2014-07-03 patchy-selection-paper.tex
fixed typo

@@
 
-\paragraph{Other patches geometries}
+\paragraph{Other patch geometries}
 Here we have studied circular patches of habitat at long distances from each other,

### 2014-07-04 patchy-selection-paper.tex
edited in light of Nachman comments

@@
 to particular substrates, e.g.\ dark rock outcrops or white sand dunes
-\citep{benson1933concealing}.
+\citep{benson1933concealing, dice1940ecologic}.
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
 which is on a number of black lava flows has much darker pelage than on intervening,
-much lighter soil \citep{DiceBlossom1937,dice1940ecologic}.
+much lighter soil \citep{DiceBlossom1937}.
 Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
@@
 
-Hoekstra, Nachman, and colleagues have made an extensive study of coat color in the rock pocket mouse, 
-\emph{Chaetodipus intermedius}, building on classic work by
-\citet{benson1933concealing}, \citet{DiceBlossom1937} and \citet{dice1940ecologic}.
+Coat color in the rock pocket mouse
+\emph{Chaetodipus intermedius} is a classic example of local
+adaptation 
+\citep{benson1933concealing,DiceBlossom1937}.
 These mice live on rocky outcrops throughout the southwestern United States and northern Mexico, 
@@
 suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
-However, there is some evidence \citep{hoekstra2004local} that, elsewhere in the range, 
-multiple dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
+However, \citet{hoekstra2004local} reasoned that, elsewhere in the range, 
+multiple close dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
 despite intervening light habitat. 
@@
 in particular investigating how the probability of parallel mutation depends on 
-the distance between lava flow patches.
+the distance between dark outcrops.
 
@@
 The mutational target size $\mu$ for the trait is unclear. 
-While the Pinacate dark haplotype differs from the light haplotype at three amino acid residues,
+While the Pinacate dark haplotype differs from the light haplotype at four amino acid residues,
 it is likely that not all of these changes are needed for a population to begin to adapt. 

### 2014-07-04 standing-variation-paper.tex
edited in light of Nachman comments

@@
 \section{Introduction}
-
-The covergent evolution of similar phenotypes in response to shared selection pressures is a testiment to the power of selection to repeatedly sculpt phenotypic variation.
-In some cases this convergence extends to the molecular level with dispate taxa converging to the same phenotype through parallel genetic changes in the same pathway, genes, or even by precisely the same genetic changes.
-Such precise convergence points to the conservation of function of the genes underlying adaptive traits, sometimes over deep time scales \citep{deephomologypapers}.
-Convergence on a molecular level also suggests that the path taken by adaptation can sometimes be relatively constained as few changes could have resulted in the adaptive phenotype, or that relatively few changes that can confer the advantagous phenotype of interest were sufficiently free of deleterious pleiotropic consequences to contribute to adaptation.
-
-%For highly polygenic traits
-
-Convergent adaptation can also occur within species, with individuals within a species 
-adapted to the same environment in parallel through different large effect changes. 
-There are a growing number of examples of this in a range of well studied organisms and phenotypes REFs.
-All else being equal the evolution of convergent phenotypes within a species convergence is more likely with a higher mutational input, i.e.
-when the mutation target and population sizes are larger \citep{}.
-Convergence is also more likely when standing variation is present for the phenotype, 
-as multiple pre-existing alleles can spread from low frequency, 
-if our now adaptive alleles were previously not too deleterious \citep{Orr,Hermission}. 
-When multiple rare changes sweep up from low frequency in parallel within a population this has been termed a soft sweep.\\
-
-The geographic distribution of populations can also shape the probability of parallel mutation within a species.
-A geographically wide spread species is more likely to adapt by multiple mutations im parallel if dispersal 
-is geographically limited, as populations will adapt via new mutations rather than waiting for migration \citep{RalphCoop}. 
-Convergence is also more likely if geographically separated populations adapting to similar environments 
-via alleles which are maladapted in the environmentals separating them. 
-Such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{lenormand} 
-and so might encourage local adaptation by parallel mutation rather than migration.\\
-
-For example a variety of plant species have repeatedly adapted to patchs of heavy metal soils
- (e.g. serpentine outcrops and mine tailings) within their ranges
-and  alleles that adapt the plants to heavy metals may be disfavored off of these patches. 
-Similarly across the American Southwest light colored rock pocket mice have adapted to outcrops of dark rock 
-through the evolution of darker cryptic pigmentation. These dark outcrops are geographically
- separated by areas of light colored sand and rock, 
-and strong selection appears to favour the matching of crypsis to substrate. 
-Perhaps as a result of this strong selection against migrants, 
-at least two distinct genetic changes are responsible from the dark pigmentation adaptation on different outcrops. 
-This situation is repeated in XXXX where darker colored mice
- have repeated evolved lighter pigmentation to match lighter color sands on
- beaches via different genetic routes.
-
-%http://www.nature.com/hdy/journal/v94/n2/full/6800600a.html
 
@@
 
+\paragraph{Rates of origination of standing and new mutations}
 We make use of the commonly used approximation that neglects competition between relatives,
@@
 
-Before $t=0$, on the other hand, the mutation is deleterious, so the genetic descendents of each new mutation are (with high probability) doomed to extinction,
+Before $t=0$, on the other hand, the mutation is deleterious, so the
+genetic descendents of each new mutation are (with high probability)
+doomed (doomed I say!) to extinction,
 but may persist for some time.
@@
 \plr{XXX More discussion on whether this is a good approximation?  
-Compute typical size and radius of a standing cluster?
+Compute typical size and radius of a standing cluster?}
+\gc{Point out connection to standard mutation selection balance formula}
 
-recall PPP computations; draw cones of opportunity for new mutations }
+\paragraph{Geographic spread of alleles} 
+Once an allele has become local established it can begin to spatially spread. 
+We assume that the allele, having become established, quickly settles
+down to spread as a traveling wave of constant speed. The behavior of
+this wave of advance of a beneficial allele was first described by 
+\citet{fisher1937wave} and \citet*{KPP1937}.
 
-\gc{Point out connection to standard mutation selection balance formula}
+If the mean squared distance between parent and child is $\sigma^2$
+(i.e.\ $\sigma$ is the dispersal distance) then the speed of advance
+of this wave will be $v = \sigma \sqrt{2s}$.
+
+\plr{recall PPP computations; draw cones of opportunity for new mutations }
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{spreading_alleles_trimmed}
+  \end{center}
+  \caption{
+Cartoon of the geographic spread of standing and new mutations.
+
+  \label{fig:cartoon}
+  }
+\end{figure}
 
@@
 by the speed at which the mutants spread.
-We will define the characteristic length for this model,
+We will define the characteristic length for this new model,
 as well as a similar compound parameter describing the relative importance of standing variation to the process of adaptation.
@@
 Suppose we fix our attention on a particular new mutation that happens to be the first to occur in some region.
-As it spreads, it will cover a distance $L$ in time $L/v$, where $v = \sigma \sqrt{2s_b}$ is the speed of the wave of advance.
+As it spreads, it will cover a distance $L$ in time $L/v$.
 The number of other mutations appearing in the circle it has covered up until this time is Poisson with mean
@@
 
+Thoughts
+species coherence.
+The Biological Reality of Species: Gene Flow, Selection, and Collective Evolution
+
 \bibliography{standing_patches_refs}

### 2014-07-05 patchy-selection-paper.tex
peter edited through entirely

@@
 \abstract{
-Populations within a species often repeatedly locally adapt to similar environmental conditions
-across their range. Such adaptation can occur through convergent
-evolution as different alleles arise and spread in different parts of
-the range. Or it can occur through the spread of alleles by migration
-acting to synchronizing adaptation across the species.
+Species often encounter, and adapt to, many patches of locally similar environmental conditions
+across their range. 
+Such adaptation can occur through convergent evolution as different alleles arise and spread in different parts of the range, 
+or through the spread of alleles by migration
+acting to synchronize adaptation across the species.
 The tension between the two reflects the degree of 
 constraint imposed on evolution by the underlying genetic
-architecture versus how how effectively selection acts to inhibt the geographic spread of local
+architecture versus how how effectively selection acts to inhibit the geographic spread of locally
 adapted alleles.
-This paper studies a model of the the balance between these two routes to adaptation in continuous environments with patchy selection pressures.
+This paper studies a model of the balance between these two routes to adaptation in continuous environments with patchy selection pressures.
 We address the following questions: 
 How long does it take for a new, locally adapted allele to appear in a
-patch of favorable habitat through new mutation?
+patch of habitat where it is favored through new mutation?
 Or, through migration from another, already adapted patch?
-Which is more likely to occur, as a function of patch separation?
-How can we tell which has occurred, i.e. what population genetic
+Which is more likely to occur, as a function of distance between the patches?
+How can we tell which has occurred, i.e.\ what population genetic
 signal is left by the spread of migrant alleles and how long does this
@@
 as spatial branching processes.
-This provides a set of tractable approximations to understand the role
-of geographic separation between favorable patches in promoting
-convergent adaptation and the genomic signals left.
+This provides a way to understand the role of geographic separation between patches 
+in promoting convergent adaptation and the genomic signals it leaves behind.
 We illustrate these ideas using the convergent evolution of cryptic coloration in the rock pocket mouse, \textit{Chaetodipus intermedius}, as an empirical example.
@@
 % accessible to adaptation is sometimes relatively constrained \citep{stern2013genetic}, 
-% as would be the case if only a few changes confer the advantagous phenotype 
+% as would be the case if only a few changes confer the advantageous phenotype 
 % and are sufficiently free of deleterious pleiotropic consequences.
@@
 The convergent evolution of similar phenotypes 
-in response to shared selection pressures is a testiment to the power of selection 
+in response to shared selection pressures is a testament to the power of selection 
 to repeatedly sculpt phenotypic variation. 
@@
 Convergent adaptation can also occur within species, if individuals within a species 
-adapt to the same environment, in parallel, through different genetics changes. 
+adapt to the same environment, in parallel, through different genetic changes. 
 There are a growing number of examples of this in a range of well studied organisms and phenotypes \citep{Arendt:08}.
@@
 that have repeatedly adapted to patches of soil with high concentrations of heavy metals
-(e.g.\ serpentine outcrops and mine tailings) \citep{macnair1991evolution,schat1996identical,turner2010serpentine};
+(e.g.\ serpentine outcrops and mine tailings) \citep{kruckeberg1951intraspecific,turner2010serpentine,macnair1991evolution,schat1996identical};
 the alleles conferring heavy metal tolerance are thought to be locally restricted 
@@
 to particular substrates, e.g.\ dark rock outcrops or white sand dunes
-\citep{benson1933concealing, dice1940ecologic}.
+\citep{benson1933concealing}.
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
@@
 at least two distinct genetic changes are responsible for the dark pigmentation found on different outcrops \citep{hoekstra2003different}. 
-Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism}
+Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism,dice1940ecologic}
 and in lizards \citep{rosenblum2010molecular}.
@@
 How easy will it be to detect adaptive alleles that are shared by
-migration, i.e. over what genomic scale will a population genetic
+migration, i.e.\ over what genomic scale will a population genetic
 signal be visible?
@@
 In section \ref{ss:patchymutation} we develop a simple approximation,
-eqn. \eqref{eqn:mutrate}, for the rate at which patches become
-established by new mutations. Then in section \ref{ss:patchymigration} we
-develop an approximation, eqn. \eqref{eqn:migrate}, to the rate which alleles move from an
-existing patch to colonize a novel patch (this represents a big chunk
-of the results of the paper). We combine these two results in section
-\ref{ss:probparallel} to develop an approximation to the probability of
-parallel adaptation, eqn. \eqref{eqn:parallel_prob}.
+equation \eqref{eqn:mutrate}, for the rate at which patches become established by new mutations. 
+The most important conceptual work of the paper occurs in section \ref{ss:patchymigration},
+where we develop an understanding of the process by which alleles move from an existing patch 
+to colonize a novel patch, culminating in equation \eqref{eqn:migrate} for the rate at which this happens.
+We combine these two results in section \ref{ss:probparallel} 
+to discuss to the probability of parallel adaptation, equation \eqref{eqn:parallel_prob}.
 To understand the genomic signal left by adaptations shared by
@@
 it will take for an allele to transit between patches,
-eqn. \eqref{eqn:mean_tau}, and use this to approximate the length of
+equation \eqref{eqn:mean_tau}, and use this to approximate the length of
 haplotype that hitchhikes with it,
-eqn. \eqref{eqn:haplotype_length}. We then quantify how quickly this
-haplotype decays within patches eqn. \eqref{eqn:rec_rate_within} and
+equation \eqref{eqn:haplotype_length}. We then quantify how quickly this
+haplotype decays within patches equation \eqref{eqn:rec_rate_within} and
 use this to obtain properties about the equilibrium distribution of
@@
 \eqref{eqn:within_haplotype_length}  in section \ref{ss:patch_haplotype}. 
-Finally in section \ref{ss:applications} we apply this work to understand the geographic scale of
+Finally, in section \ref{ss:applications} we apply this work to understand the geographic scale of
 convergent evolution in \textit{Chaetodipus intermedius}.
-
-% http://www.nature.com/hdy/journal/v94/n2/full/6800600a.html   -- this is \citep{hoekstra2004local}
-
-%and that a single mutational change was sufficient to adapt populations to the novel environment. 
-%We assumed that there was no standing variation for the adaptive allele, such that parallel mutation must be due to multiple mutations after the environmental shift.
-In that setting we found that there was a characteristic geographic scale, over which we would expect multiple instances of our adaptive allele to have arisen in parallel.
-This characteristic length could be expressed in terms of a simple compound parameter determined by our parameters of interest.
-%Recall questions of origin of adaptations
-
-%References on parallel adapation, standing variation, etc.:
- % coop,
- % pennings and hermisson.
-%Note pennings and hermisson found standing deleterious variation was very important.
-
-%recall peromyscus examples with patchy environments
-
-%Summarize previous paper
-
-%Question assumptions of no standing variation and homogenous selection
 
@@
 can escape initial loss by drift when rare. 
-Therefore, we do not have to specify the the fitness of the homozygote; 
+Therefore, we do not have to specify the fitness of the homozygote; 
 only that the dynamics of the allele when rare 
 are determined by the fitness of the heterozygote. 
-More general dominance will only make small corrections to the the dynamics until initial fixation,
+More general dominance will only make small corrections to the dynamics until initial fixation,
 with the exception of the recessive case, which we omit.
-For the most part, 
+In other words,
 we follow the literature in treating the diploid model as essentially haploid.
 
-For the sake of simplicity, we assume a constant haploid population density $\rho$, 
-across both types of habitat,
+Also, to avoid introducing more parameters,
+we assume population density $\rho$ is equal across the range (even in the ``poor'' patches).
+If this is not the case, in most equations below, the population density that should be used is the density
+on the poor patches; see \citet{lenormand2002limits} for a more thorough approach.
 We further assume that the variance in offspring number is $\xi^2$, 
 and that the mean squared distance between parent and child is $\sigma^2$
-(i.e.\ $\sigma$ is the dispersal distance). Differences in population
-densities on and off the patches may arise if selection is hard, in
-that case these differences can be incorporated into the density \citep[see][]{lenormand2002limits}. 
+(i.e.\ $\sigma$ is the dispersal distance). 
 %We also aim to describe the scenario where it is sensible to think of large, discrete patches;
 %the case of a fine-scale heterogeneous environment is treated
-%\citep{elsewhere}. \gc{IS THIS LAST SENTENCE NEEDED IF SO GIVE CITATIOn}
+%\citep{elsewhere}. \gc{IS THIS LAST SENTENCE NEEDED IF SO GIVE CITATION} \plr{haven't found it yet}
 
@@
 with the scale given by $\sigma/\sqrt{s_m}$. Mutations appearing within the patch have probability of establishment
-strictly less than the probability for a panimictic population,
+strictly less than the probability for a panmictic population,
 but approaching this as the size of the patch increases.
@@
 
-Consider first a single, isolated poor habitat patch of area $A$ in which no $B$ allele has yet become established. 
-We first compute the time scale on which new $B$ mutations appear and fix in such a patch.
+Consider first a single, isolated poor habitat patch in which no $B$ allele has yet become established. 
+We first compute the time scale on which a new $B$ mutations will appear and fix in such a patch,
+if the rate of mutation from $b$ to $B$ is $\mu$ per individual per generation.
 As we are interested in patches where local adaptation can occur,
@@
 % the function $p(x)$ can be found (in one dimension) in terms of Jacobi elliptic functions,
-but results will be more interpretable if we proceed with a simple approxmation.
+but results will be more interpretable if we proceed with a simple approximation.
 The total influx per generation of mutations that successfully establish is given by the
@@
 If we consider a patch of area $A$, whose width is large relative to $\sigma/\sqrt{2s_m}$, 
-then a reasonable approximation is to assume that $p(x) = p_e \approx 2 s_p / \xi^2$ within the patch, and $p(x) = 0$ otherwise.
+then a reasonable approximation is to ignore migration, 
+so that $p(x) = p_e \approx 2 s_p / \xi^2$ within the patch, and $p(x) = 0$ otherwise.
 %for a patch of area $A$ approximating $\int p(x) dx \approx 2 s_p A / \xi^2$. 
-We would then approximate the integrand $p(x)$ in \eqref{eqn:exactmutrate} by a step function,
+This approximates the integrand $p(x)$ in \eqref{eqn:exactmutrate} by a step function,
 which will be a good approximation if the patch is large relative the scale over which $p(x)$ goes from 0 to $p_e$,
@@
 %%%%%%%
+\subsection{Simulation methods}
+\label{ss:simulations}
+
+First we briefly describe the simulations used below for illustration and validation.
+We simulated forward-time dynamics of the number of alleles of each type in a regular grid of demes with fixed size $N$.
+Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
+locally beneficial alleles were more likely to reproduce.
+Each then produced a random, Poisson distributed number of offspring that each independently migrated to demes no more than six steps away
+according to a symmetric migration kernel.
+Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
+(Although we described the simulation in terms of individuals,
+we kept track only of total numbers in an equivalent way.)
+
+The reproduction rate in simulations for type $b$ alleles was 30\%; 
+this was then multiplied by $1+s$ to get the probability of reproduction for type $B$.
+These $s$ are the values reported in the figures, and do not depend on the basic rate of reproduction.
+However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
+we computed the rate of intrinsic growth, 
+i.e.\ the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
+in the absence of migration.
+(The resulting values are close to the first notion of $s$,
+but give better agreement with theory,
+which uses the second definition.)
+
+To sample lineages, we first simulated forwards in time the population dynamics,
+then sampled lineages moving back through time
+by, in each generation,
+moving each lineage to a new deme with probability proportional to the reverse migration probability
+weighted by the number of $B$ alleles in that deme in the previous time step.
+If more than one lineage was found in a deme with $n$ $B$ alleles,
+then each lineage picked a label uniformly from $1 \ldots n$,
+and those picking the same label coalesced.
+Since reproduction is Poisson, this correctly samples from the distribution of lineages given the population dynamics.
+
+%%%%%%%
 \subsection{Establishment of a locally adaptive allele due to migrational influx}
@@
 To determine this, we dissect the process by which an allele transits between the patches,
-in the process obtaining other useful information about the genealogy of $B$ alleles.
+along the way obtaining other useful information about the genealogy of $B$ alleles.
 
@@
 and assume that the shape of the patch is at least roughly circular.
-Following \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(x)$ is the solution,
+Following \citet{haldane1948theory} or \citet{slatkin1973geneflow}, one can write down a differential equation to which $q(x)$ is the solution,
 and show that $q(x)$ decays exponentially for large $|x|$,
@@
   a two-dimensional grid of $101\times 101$ demes.
-  Superimposed in black is the decay predicted by \eqref{eqn:eqfreq},
+  Superimposed in color is the decay predicted by \eqref{eqn:eqfreq},
   with $C$ chosen arbitrarily to give a reasonable fit.
@@
   from a one-dimensional grid of 151 demes;
-  the solid lines shows the cumulative mean occupation frequency across all generations.
+  the solid line shows the cumulative mean occupation frequency across all generations.
   For both plots, the simulation was run for 1000 generations;
-  the focal allele is beneficial in the central 20 demes, with $s_p=0.1$,
+  the focal allele is beneficial in the central 10 demes, with $s_p=0.1$,
   and deleterious elsewhere, with $s_m=-0.02$; 
@@
 This expected frequency $q(x)$ is the \emph{time-averaged} occupation frequency,
-or in other words, the total number of $B$ alleles found across $T$ generations near location $x$, per unit area, divided by $T$.
+i.e.\ the total number of $B$ alleles found across $T$ generations near location $x$, per unit area, divided by $T$.
 If, for instance, $q(x)=.01$ and the population density is $\rho=10$ individuals per unit area, 
@@
   \caption{
-  The temporal dynamics of the same simulation as Figure~\ref{fig:sim_occupation_freqs}B.
-  A segment of the temporal dynamics: space is again shown across the horizontal axis;
-  with time on the vertical axis; demes are colored darker red the greater then allele frequency at the corresponding time.
+  A segment of the temporal dynamics of the same simulation as Figure~\ref{fig:sim_occupation_freqs}B.
+  Space is again shown across the horizontal axis, with time on the vertical axis; 
+  demes are colored darker red the greater their allele frequency at the corresponding time.
   The 50\% and 5\% frequency contour lines are in grey,
@@
 then these families can be approximated by a collection of independent spatial branching processes
-whose members are killed (actually, ignored) if they enter the original patch.
+whose members are ignored if they enter the original patch.
 (This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
@@
 and the quantity we wish to compute,
-the effective rate of at which \emph{families of} migrant $B$ alleles establish in a new patch at distance $r$, 
+the effective rate of at which \emph{families of} migrant $B$ alleles establish in a new patch at displacement $x$ from the original,
 is
@@
 and we can solve for this outflux using the established form for $q(r)$ in \eqref{eqn:gestalt_q}.
-Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals.
-
-%%%%
+Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
+-- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.
+
+%%%% %%%%
 \subsubsection{The genealogy of migrant families}
@@
 has gone extinct by time $t$, 
-%\plr{switched from $p_e(t)$ to $k_e(t)$ to not conflict with prob of estab.  better suggestion?}
 and let $K_t$ be the (random) family size conditioned on nonextinction,
@@
 then $K_t$ has a limiting distribution: $K_t \dconv K$ 
-and that $e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
+and $e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
 In other words, the chance of survival to time $t$ is a constant multiplied by $e^{-s_m t}$,
@@
 \caption{Cartoon of the decomposition described in the text. 
-Migrants families that get further than $r_0$ from the already-established patch (grey circle);
-each is composed of a ``trunk'' lineage (in red), whose spatial motion we follow,
+Migrants families that get further than $r_0$ from the already-established patch (grey circle)
+are each composed of a ``trunk'' lineage (in red), whose spatial motion we follow,
 embellished by transient branches.
@@
 Note that the spatial motions of lineages are independent random walks
-(and hence should be much wigglier than in this cartoon).
+(and hence should be much wigglier than in this cartoon, which confounds space and time).
 } \label{fig:branching_decomp}
@@
 %%%%%%% %%%%%%%%%%%
-\subsubsection*{Putting it together}
+\subsubsection*{The outflux of migrant alleles}
 
@@
 Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
-defined so that $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
+and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
 The \emph{mean occupation density} of $Z$ can be thought of 
@@
 given that it successfully arrives.
-As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral of $h_A(x,y)$:
+As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
 \begin{align}\label{eqn:migrate_integral}
@@
 The probability $f(A)$ that the family does not die out once it arrives at the new patch is
-$\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
+approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
 if each member arriving has chance $p_e$ of establishing, independently.
@@
 
-We now turn to the question of whether the separated populations adapt by parallel genetic changes 
+We now turn to the question of whether the separated patches adapt by parallel genetic changes 
 or by the spread of a migrant allele between the patches
@@
 the other patch will become colonized by migration or mutation, but not by both. 
-As such, the question of how the the second patch adapts
+As such, the question of how the second patch adapts
 simply comes down to whether a new mutation or a migrant allele is the first to become established in the second patch. 
@@
 for the effective rates of adaptation by new mutation and by migration.
+We work in one dimension, as the square root term appearing for two dimensions is relatively unimportant (see figure~\ref{fig:sim_occupation_freqs}).
 
@@
 
-\gc{NOT SURE IF ANYONE HAS LOOKED AT THIS NEXT PARAGRAPH IN LONG TIME?}
 If we take the mutational influx in a patch to be one per 1,000 generations ($\rho A \mu = .001$)
@@
 but also where adaptation by any means takes a very long time ($\Tmig$ and $\Tmut$ both very large),
-and important check in applications.
+an important check in applications.
 
@@
   Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=50$, and $\xi^2=1$.
-  Note that population density and mutation rate are shown on a log scale.
+  Note that population density $\rho$ and mutation influx $A\mu$ are shown on a log scale.
   \label{fig:phase_diagram}
@@
 so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
-and decrease with population density and benefical selection coefficient.
+and decrease with population density and beneficial selection coefficient.
 Furthermore, the probability that the second adaptation is a new mutation,
-i.e.\ the probability of parallel adapatation, is 
+i.e.\ the probability of parallel adaptation, is 
 \begin{equation} \label{eqn:parallel_prob}
-  \frac{\mutrate}{\mutrate+\migrate(r)} = \frac{ A \mu s_m }{2 A \mu s_m + \left(r \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} r / \sigma \right) },  
+  \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu s_m }{2 A \mu s_m + \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
 \end{equation}
@@
 For instance, the proportion of demes that shares the same origin as a randomly sampled deme
-is approximately Beta distributed -- see \cite{donnelly1989continuity} and \cite{perman1992sizebiased}.
+is approximately Beta distributed -- see \cite{donnelly1989continuity}.
 The high connectedness of the discrete deme island model means that the expected number of distinct alleles
@@
   An initially unadapted patch (10 demes at the left end of the range) is colonized around generation 5000;
-  a representative lineage from the new patch is shown transiting between the two and coalescing with a lineage
-  in the unadapted patch.
+  a representative lineage is shown in the new patch, which follows the original migrant ``trunk'' lineage
+  between the patches and coalesces with a lineage in the unadapted patch.
   In the left panel, as in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
   other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
-  The right panel shows (in red) the length of the shared haplotype between the two lineages
-  around the selected locus,
+  The right panel shows (in red) the extent of the shared haplotype between the two lineages around the selected locus (at position 0),
   which decreases as time progresses. 
@@
 We first discuss the transit of the lineage, and its accompanying haplotype, and return
-to how haplotypes are whittled down within estalbished patches in the next section.
+to how haplotypes are whittled down within established patches in the next section.
 
@@
 recombination between $B$ alleles (since they are at low density in transit, we assume this is rare),
-and the possibility that more than one migrant family is in transit at once
+and the possibility that more than one successful migrant family is in transit at once
 (so that faster migrants would be more likely to have arrived first);
@@
 we can use standard results on hitting times of $d$-dimensional Brownian motion
-that is killed at rate $s_m$ \citet[][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook} 
-In particular, if the patch is circular with radius $w$ and a distance
+that is killed at rate $s_m$ -- \citet[see][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook}.
+In particular, if the patch is circular with radius $w$ and lies at distance
 $R$ from the already adapted patch, then 
@@
 then $L$ has the same distribution as $(Y + \sqrt{s_m})^2 - s_m$.
-Furthermore, the expected length of shared hitchiking haplotype is
+Furthermore, the expected length of shared hitchhiking haplotype is
 \begin{equation} \label{eqn:haplotype_length}
@@
 it appears here because $\sigma/\sqrt{s_m}$ is a length scale, 
-and to convert it to a time this is divided by $1/s_m$.
+and to convert it to a speed this is divided by $1/s_m$.
 This intuitive picture can be expanded by a central limit theorem:
@@
 (equation~\ref{eqn:haplotype_length}) will also break down. 
+However, a long time after the initial sweep,
+we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
+than with individuals elsewhere,
+since selection against migrants 
+decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
 We first develop a simple approximation to the rate 
 at which recombination breaks down haplotypes within patches.
-If we wait long enough after the initial adaptation of a patch,
-the haplotype that swept initially will have decayed due to this process of recombination. 
-However, we may still expect to find individuals within the patch sharing long haplotypes around the selected locus
-than with individuals elsewhere,
-since selection against migrants 
-decreases mean coalescence times within the patch near the selected locus.
-%\citep{bartonpaper,charlesworths}.
 
@@
 After the same large-population, continuous-space limit that underlies the rest of this paper (Appendix~\ref{apx:eqfreq}),
+if the assumption that fluctuations in $p(x)$ are small,
 this motion should converge to a diffusion process \citep[as assumed by e.g.][]{hallatschek2008surfing},
@@
     d X_t &= (1+s(X_t)) \frac{d}{dx} \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,
-    % &=  ( \log(p(X_t))' (1+s(X_t)) + s'(X_t) ) dt + \sqrt{1+s(X_t)} dB_t,\\
 \end{align}
@@
 the lineage bouncing around in the patch,
-we propose a simple heurstic to describe the probability of recombination. 
+we propose a simple heuristic to describe the probability of recombination. 
 If recombination is on a slower time scale than mixing within the patch,
@@
 As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
-Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
-multiplied by $p(x)$ in these intermediate regions, where we assume
-$p(x) \approx 1/2$. Under these approximations the per generation rate of
-recombination of a linked lineage to the $b$ background is
+% Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
+% multiplied by $p(x)$ in these intermediate regions, where we assume $p(x) \approx 1/2$. 
+The per generation rate of recombination of a locus linked to the $b$ background is thus
 reduced by a factor roughly equal to the proportion of the patch lying within distance $\sigma/\sqrt{s}$ of the boundary,
@@
 $\xi^2 /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$, 
-and hence the genetic length of haplotype shared between pairs of
-individuals within the patches at equilibrium is on the order of
+and hence the genetic length of the haplotype shared between pairs of
+individuals within the same patches at equilibrium is on the order of
 \begin{align} \label{eqn:within_haplotype_length} 
-  r = \frac{\sqrt{s} \xi^2}{\sigma \rho A^{(d-1)/d}} 
-\end{align}
-which does not depend on $A$ in one dimension,
+  r = \frac{\sqrt{s} \xi^2}{\sigma \rho A^{(d-1)/d}}  .
+\end{align}
+This does not depend on $A$ in one dimension,
 but decreases with $\sqrt{A}$ in two dimensions.
@@
 % > about 10 animals with food provided ad libitum, so I think this is an upper
-% > density (i.e. 40/hectare).  As for dispersal distance, I'm less sure.
+% > density (i.e.\ 40/hectare).  As for dispersal distance, I'm less sure.
 % > Studies on closely related species (genus Perognathus) in similar habitats
@@
 it is likely that not all of these changes are needed for a population to begin to adapt. 
-Also there a number of genes besides MC1R at which adaptative changes affecting pigmentation 
+Also there a number of genes besides MC1R at which adaptive changes affecting pigmentation 
 have been identified in closely related species and more broadly across vertebrates \citep{Hoekstra:06}.
@@
 Finally, we set $A=100\text{km}^2$ (roughly the size of the Pinacate patch).
-Figure~\ref{fig:mice_prob_parallel} shows the dependance of the probability
+Figure~\ref{fig:mice_prob_parallel} shows the dependence of the probability
 of parallel mutation on the distance between lava flow patches using these parameters, 
@@
 patches after establishment (equation \ref{eqn:rec_rate_within}). We
-estimate that the effective recombination rate is at most only 50 \% lower
-than the actual recombination rate.
-Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare (M.~Nachman, personal communication),
+estimate that the effective recombination rate is no less than 50\% of the actual recombination rate.
+Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare 
+(M.~Nachman, personal communication),
 and the variance in offspring number to be $\xi^2 = 3$ (mostly to get round numbers),
@@
 The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
-the dispersal distance divided by the square root of the strength of selection
-against the new allele between patches. It depends much more weakly on the
-selective benefit within the patches or (perhaps suprisingly) the population density, 
+the dispersal distance divided by the square root of the strength of selection against the allele between patches. 
+It depends much more weakly on the
+selective benefit within the patches or (perhaps surprisingly) the population density, 
 although these two factors will determine the time-scale over which adaptation will occur. 
@@
 
+% ADD SOMETHING ABOUT:
+% In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
+% is $\rho (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
+% which depends fairly strongly on the selective disadvantage $s_m$ between patches.
+% Equivalently, the rates are roughly equal if
+% $R/\sigma = \log(1/(s_m A \mu))/\sqrt{2s_m}$,
+% which gives the critical gap size past which adaptation will be mostly parallel
+% in terms of selection coefficients, patch size, and mutation rate.
+
 %commented out as wasn't quite sure what purpose this paragraph served.
+%% the point was that most other work deals with bulk behavior, differential equations,
+%% not tracking individuals as you have to do here.
 % In this work, we have tried to emphasize an intuitive, individual-centered understanding of the process,
@@
 and the question is whether or not adaptations in distinct patches
-have appeared indpendently or not.
-However, especially if rates of adaptation by new mutation are very low,
+have appeared independently or not.
+However, especially absolute if rates of adaptation by new mutation are very low,
 any adaptation that does occur may have to make use of standing variation.
@@
 and so would encounter much less resistance to spreading between patches.
+The shape of the cline obtained is given by \citet{haldane1948theory}.
 This is counteracted, however, by the increased difficulty with which the mutation would establish
@@
 Our work suggests that even quite geographically close populations 
-may be forced to locally adapt convergently when migration is geographically limited. 
+may be forced to locally adapt convergently when migration is geographically limited
+and selective pressures are divergent.
 Thus, systems where populations have been repeatedly subject to 
@@
 genetic background \citep{stern2013genetic}.  
-Such empricial work will also strongly inform our understanding of the
+Such empirical work will also strongly inform our understanding of the
 ability of gene flow to keep ecologically similar populations evolving
@@
 
-% unequal density: Lenormand, gene flow and limits of\dots
-
 \bibliography{standing_patches_refs}
@@
 
-To investigate the impact of the various approximations surrounding the mutational influx,
-we calculated the probability of establishment for an spatial branching process moving across discrete demes.
-The solution was obtained by iterating the multivariate generating function until convergence
-(since the function which gives the probability of extinction as a function of location of the progenitor
-is an attractive fixed point of the generating function; see \citet{jagers1975branching}).
+When rare, under at least some demographic models,
+copies of a new mutant allele are approximately independent and experience a uniform selective benefit;
+and can therefore be treated as a branching process (a spatial branching process in this case).
+Furthermore, whether or not a new, beneficial mutation establishes or is lost to demographic stochasticity
+is determined by this phase where it is rare.
+Fortunately, the probability that a branching process does not establish
+(i.e.\ dies out)
+can be found as a fixed point of the generating function of the process \citep{jagers1975branching}.
+Therefore, we calculated explicitly the generating function for a spatial branching process
+with nearest-neighbor migration on a one-dimensional lattice
+and a Poisson number of offspring with mean $1+s$,
+where $s$ could vary by location,
+and iterated this forward to convergence to obtain $1-p(x)$, the probability a single mutation appearing at $x$ would fail to establish.
+We considered two situations:
+where $s$ is a step function,
+and where it has a linear transition.
 These solutions are shown, and parameters described, in figure~\ref{fig:prob_estab_calcs}.
@@
 the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
+The fit is equally good for other parameter values,
+even if migration can move further than one deme and offspring numbers are not Poisson (results not shown).
 This lends credence to our approximation that the integral of $p(x)$ over the entire range
@@
     in an abrupt transition (left) and a gradual transition (right).
-    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respecively);
+    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respectively);
     and the selection coefficient interpolates linearly between these in the central hatched region.
@@
 and establish the asymptotics given in that equation.
-One route to the ``equlibrium frequency'' of the allele outside the range where it is advantageous is as follows;
-see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} for other arguments in this case, 
+One route to the ``equilibrium frequency'' of the allele outside the range where it is advantageous is as follows;
+see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} (or \citet{KPP1937book} or \citet{fisher1937wave} or \citet{haldane1948theory})
+for other arguments in equivalent models,
 or \citet{etheridge2000introduction} and/or \citet{dawson1993measurevalued} for general theory used below.
@@
 the reproduction rate is 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
-Suppose that the number of offspring $X$ has mean $\mu$; the variance of $X$ will not enter into the formula.
+Suppose that the number of offspring $X$ has mean $\mu$; the variance of $X$ will not enter into the formula
+(but assume $X$ is well-behaved).
 Also suppose that the migration displacement $R$ has mean zero and variance $\sigma^2$;
@@
 For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the random number of mutant alleles added if $X$ new offspring carrying mutant alleles
-replace randomly chosen individuals in a deme where the mutant allele is at frequncy $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
+replace randomly chosen individuals in a deme where the mutant allele is at frequency $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
 similarly, $\widetilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
@@
 we aren't providing the rigorous justification here.)
-Then we know that for any sufficiently nice function $f(\Phi)$ that
+Then the above description implies that for any sufficiently nice function $f(\Phi)$ that
 \begin{align} \label{eqn:discrete_generator}
@@
 We can see two things from this:
-First, since this is a first-order differential operator, the limiting stochastic process $\Phi$
+First, since this is a first-order differential operator, the limiting stochastic process $\Phi$ obtained as $N \to \infty$
 is in fact deterministic (check by applying to $f(\Phi) = \Phi(x)^2$ to find the variance).
@@
 More concretely, suppose that the grid size is $\epsilon \to 0$, 
-that $\var[R] = (\sigma \epsilon)^2$, and that $s(x)/\epsilon \to \gamma(x)$,
+that $\var[R] = (\sigma \epsilon)^2$, and that the strength of selection is $s(x) \epsilon$,
 and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
@@
   \label{eqn:derived_fkpp}
-  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \xi(t,x) + \gamma(x) \xi(t,x) (1-\xi(t,x)) \right) .
+  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \xi(t,x) + s(x) \xi(t,x) (1-\xi(t,x)) \right) .
 \end{align}
@@
 in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
-If we take a radially symmetric patch centered at the origin (so $\gamma$ only depends on $r$), 
+If we take a radially symmetric patch centered at the origin (so $s$ only depends on $r$), 
 and let $\xi(t,r)$ denote the mean occupation frequency at distance $r$,
@@
   \label{eqn:radial_fkpp}
-  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + \gamma(r) \xi(t,r) (1-\xi(t,r)) \right) .
+  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + s(r) \xi(t,r) (1-\xi(t,r)) \right) .
 \end{align}
@@
 So, using \eqref{eqn:radial_fkpp}, a radially symmetric solution 
-with $\gamma(r) = -s < 0$ for all $r>r_0$ 
+with $s(r) = -s < 0$ for all $r>r_0$ 
 solves, for $r_0 < r < \infty$,
-\begin{align}
+\begin{align} \label{eqn:radial_eqfreq_diffeq}
     \partial_r^2 q(r) + \frac{d-1}{r} \partial_r q(r) - \frac{2s}{\sigma^2} q(r) (1-q(r)) = 0 \\
@@
 
-Here we describe how to solve \eqref{eqn:derived_fkpp} in one dimension,
-i.e.\ the Fisher-KPP equation with varying selection.
+Here we describe how to solve \eqref{eqn:radial_eqfreq_diffeq} exactly in one dimension,
+i.e.\ the Fisher-KPP equation with piecewise constant selection.
 Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
@@
 with appropriate boundary conditions.
-First rescale space by $\sigma/\sqrt{2}$ so the $\sigma^2/2$ term dissappears.
+First rescale space by $\sigma/\sqrt{2}$ so the $\sigma^2/2$ term disappears.
 If we now assume that $s(x)$ is piecewise constant,
@@
 \begin{align}
-    u(x,S) &= \E[\int_0^\infty Z_t(S) dt ] \\
+    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
            &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
@@
         &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
-        &= \E\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  .
-\end{align}
+        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
+\end{align}
+where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
+and likewise $\E^x$.
 If we define
@@
 \end{align}
+where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
 provided $S$ is not pathological.
@@
 \begin{align}
-    h(x,S) &= \int_0^\infty (1-k_e(t)) \P\{ \tau_S \in dt \} \\
-           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P\{ \tau_S < t \} dt \\
-    &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P\{ \tau_S < t < \tau_\dagger \} dt \\
-    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_S < t < \tau_\dagger \} dt \\
+    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P\{ \tau_S \in dt \} \\
+    &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P\{ \tau_S \in dt \} \\
+    &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
     &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
@@
 If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
-then by equation~\eqref{eqn:migrate_integral},
+then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
 \begin{align}
@@
 
-\section{UNUSED AFTER HERE?}
-
-%%%%%%%%%
-\section{Simulations of a patchy environment}
-\label{apx:patchy_sims}
-
-Describe sims.
-
-Argue that $C=0.5$ is a reasonable number.
+%%%%%%%%%%%
+
+\end{document}
+
+%%%%%%%% END %%%%%%%%%%%%%%%
+%%%%%%%% %%% %%%%%%%%%%%%%%%
 
@@
 \end{figure}
-
-%%%%%%%%%%%
-
-\end{document}
 

### 2014-07-05 standing-variation-paper.tex
added refs for standinf paper

@@
 \plr{recall PPP computations; draw cones of opportunity for new mutations }
+
+\paragraph{}
+So we can put these ingredients together for a simple model of the
+geographic spread of alleles. A cartoon example of this process is
+shown in Figure \ref{fig:cartoon} Initially, when the selection
+pressures change at $t=0$, a set of standing variants can start to
+spread having escaped loss through drift. These variants are shown as
+little lightning bolts, and occur at a density $\lambda_0$ across
+space. 
+They spread at velocity $v$ carving out cones in space-time. As these
+alleles proceed in their geographic spread other new alleles can arise
+and become established in parallel, these are shown as little
+stars. These new mutations arise and become established at rate $\lambda$.
 
@@
 
+\paragraph{G6PD}
+34 variants across 4; 000; 000 km2
+15 common variants
+
+515 codons
+
+Selection coefficient (s) 	A- 0.044 (0.019–0.048)  Med	0.034
+(0.014–0.049)  ``The mean age of the A− allele in these runs was 6357 years, with a 95\% credibility interval extending from 3840 to 11,760 years.
+For the Med alleles, the mean age was 3330 years with a 95\% credibility interval of 1600 to 6640 years"
+\citet{tishkoff-g6pd}
+
+s 0.24 and t= 40 generations \citep{Slatkin-age-est}
+
+The estimated age is ~100 generations with an upper bound of <150
+generations A- 0.1 < s < 0.2 \citep{Saunders-g6pd}
+
+46 to 58\% reduction in risk of severe malaria for both female heterozygotes and male hemizygotes estimates of selection pressure \citep{Ruwende-g6pd}
+
+Mahidol487A mutation started to increase at ~1500 YBP, with a
+selection intensity of ~0.23 \citep{Louicharoen-g6pd}
+Show's effect in het and homozy females
+
 Thoughts

### 2014-07-05 patchy-selection-paper.tex

--- 2014-07-05 patchy-selection-paper.tex
@@
 
-% ADD SOMETHING ABOUT:
-% In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
-% is $\rho (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
-% which depends fairly strongly on the selective disadvantage $s_m$ between patches.
-% Equivalently, the rates are roughly equal if
-% $R/\sigma = \log(1/(s_m A \mu))/\sqrt{2s_m}$,
-% which gives the critical gap size past which adaptation will be mostly parallel
-% in terms of selection coefficients, patch size, and mutation rate.
-
-%commented out as wasn't quite sure what purpose this paragraph served.
-%% the point was that most other work deals with bulk behavior, differential equations,
-%% not tracking individuals as you have to do here.
-% In this work, we have tried to emphasize an intuitive, individual-centered understanding of the process,
-% determining, for instance, 
-% the speed at which successful migrants traverse regions where they are maladapted,
-% and the effective rate at which new families of migrant alleles appear
-% at long distances from a patch.
-% These quantities may be as useful for applied work
-% as more abstract ones such as the probability that adaptation occurs by migration versus new mutation.
+Perhaps the most useful rule-of-thumb quantities we found were the following.
+The effective rate of migration into an as-yet-unadapted patch from an already-adapted patch distance $R$ away 
+-- the analogue of the mutational influx $\mu \rho A$ --
+is $\rho (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$.
+Equivalently, the critical gap size between patches past which adaptation is likely independent
+is $R = (\sigma/\sqrt{2s_m}) \log(1/(s_m A \mu))$.
+Finally, successfully transiting migrant lineages move at rate $\sigma \sqrt{2s_m}$,
+and so shared haplotype lengths between patches will be of order $R/\sigma \sqrt{2s_m}$.
 
@@
     the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
+    Note that the approximation used in the text is here $p_e \approx 2 s_p / \xi^2 = 0.1$,
+    agreeing with the asymptotic value on the right.
     \label{fig:prob_estab_calcs}

### 2014-07-07 patchy-selection-paper.tex

@@
 across their range. 
-Such adaptation can occur through convergent evolution as different alleles arise and spread in different parts of the range, 
+Such adaptation can occur through convergent evolution as different alleles arise and spread in different patches, 
 or through the spread of alleles by migration
@@
 use this to obtain properties about the equilibrium distribution of
-haplotype lengths
+haplotype lengths, equation 
 \eqref{eqn:within_haplotype_length}  in section \ref{ss:patch_haplotype}. 
@@
 The required change could be as specific as a single base change, 
-or as general as a knock out of a gene or one of a set of genes.
+or as general as a knockout of a gene or one of a set of genes.
 We assume that this mutational change occurs at a (low) rate of $\mu$ per chromosome per generation,
@@
 $1-s_m$  when rare in the intervening areas, with $s_p$ and $s_m$ both positive.
-(Here ``fitness'' we take to be the intrinsic growth rate of the type when rare.)
+(Here we take ``fitness'' to be the intrinsic growth rate of the type when rare.)
 To simplify matters, we assume that the disadvantage $s_m$ 
@@
 This suggests that families can be well-described by motion of a central ``trunk'',
-dies at a random time whose distribution is given by $k_e(t)$,
+which dies at a random time whose distribution is given by $k_e(t)$,
 and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
@@
 an assumption we also made in \citep{ralph2010parallel}. 
-In reality, different mutations will likely not be truly selective equivalent,
+In reality, different mutations will likely not be truly selectively equivalent,
 in which case this mixing occurs over a time-scale dictated by the difference in selection coefficients.
@@
 then for migration to be slower than mutation,
-the distance between patches must be at least 30 times the dispersal distance (i.e.\ $R/\sigma > 30$).
+the distance between patches must be at least 30 times the dispersal
+distance (i.e.\ $R/\sigma > 30$) for convergence to occur.
 If $s_m$ is much smaller, say $s_m = .001$, 
@@
 but also where adaptation by any means takes a very long time ($\Tmig$ and $\Tmut$ both very large),
-an important check in applications.
+an important check in applications of these results.
 
@@
 either by migration or mutation.
-If we imagine that the times between successive establishments is the result of many nearly-independent attempts
+If we imagine that the times between successive establishments are the result of many nearly-independent attempts
 with small probability of success,
@@
 When an already adapted patch colonizes another through migration,
-then the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
+ the newly colonized patch will inherit a long piece of haplotype around the selected site from the originating patch.
 The genetic length of this haplotype is roughly inversely proportional to the
@@
 being able to arise at any of a thousand base pairs.  
-Finally, we set $A=100\text{km}^2$ (roughly the size of the Pinacate patch).
-Figure~\ref{fig:mice_prob_parallel} shows the dependence of the probability
+Finally, we set $A=100\text{km}^2$ (roughly the size of the Pinacate
+patch). In
+Figure~\ref{fig:mice_prob_parallel} we show the dependence of the probability
 of parallel mutation on the distance between lava flow patches using these parameters, 
@@
 In other words, 
-over what spatial scale of environmental patchiness will evolution develop independent solutions to evolutionary problems?
+over what spatial scale of environmental patchiness will the process
+of adaptation develop independent solutions to evolutionary problems?
 The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
@@
 Finally, successfully transiting migrant lineages move at rate $\sigma \sqrt{2s_m}$,
-and so shared haplotype lengths between patches will be of order $R/\sigma \sqrt{2s_m}$.
+and so shared haplotype lengths between patches will be of order $R/(\sigma \sqrt{2s_m})$.
 

### 2014-07-07 patchy-selection-cover-letter.tex
cover letter

\documentclass{article}
\usepackage{natbib}
\usepackage{fullpage}
\usepackage{hyperref}

\begin{document}

% \section{Cover letter}

\begin{minipage}[b]{2.5in}
  Submission Cover Letter \\
  {\it PLoS Genetics}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
  Departments of Evolution and Ecology\\
  University of California at Davis\\
  \textbf{and} \\
  Molecular Biology and Bioinformatics \\
  University of Southern California\\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em
Please find enclosed our manuscript, entitled 
``Convergent Evolution During Local Adaptation to Patchy Landscapes'',
for consideration for publication in {\it PLoS Genetics}.

\vskip 1em

In the paper, we study the process of adaptation on a continuous landscape 
with patches of habitat in which a selective pressure is contrary to that experienced in the rest of the range,
determining the quantities that describe how adaptation will proceed.
The work is theoretical, with an empirical example.
We strove to present the substantial mathematical results in a clear manner accessible to the readership of {\it PLoS Genetics},
focusing instead on the biological intuition and interpretations.
The paper is in the same vein as our well-read paper 
``Parallel adaptation: one or many waves of advance of an advantageous allele?'' (Genetics 2010),
and fills a gap in practical workers' toolboxes and intuitions surrounding the geographic aspects of adaptation.

\vskip 1em

The work is motivated by the classical examples of intraspecific convergent evolution across patchy environments,
such as heavy metal tolerance in plants, or locally adaptive melanism in small mammals.
Such examples are of considerable interest because they 
imply a certain degree of constraint imposed on evolution by the genomic architecture underlying the trait.
Furthermore, as landscape genomics datasets become available, 
it is important to know (but almost entirely unknown) 
what proportion of the genome may be responding to patchy selection in a similar manner but on a non-obvious phenotype;
here we contribute to building this intuition.

\vskip 1em

We suggest as suitable reviewers: 
Joachim Hermisson (\href{mailto:joachim.hermisson@univie.ac.at}{joachim.hermisson@univie.ac.at}), 
Pleuni Pennings (\href{mailto:pleuni@dds.nl}{pleuni@dds.nl}),
% Magnus Nordborg (\href{mailto:magnus@usc.edu}{magnus@usc.edu}),
Phillip Messer (\href{mailto:messer@cornell.edu}{messer@cornell.edu}),
Mark Kirkpatrick (\href{mailto:kirkp@austin.utexas.edu}{kirkp@austin.utexas.edu}),
and Nick Barton (\href{mailto:nick.barton@ist.ac.at}{nick.barton@ist.ac.at}).

\vskip 2em

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf Peter Ralph, and\\
Graham Coop.}\\
\end{minipage}

\end{document}

### 2014-07-08 patchy-selection-cover-letter.tex

@@
 ``Parallel adaptation: one or many waves of advance of an advantageous allele?'' (Genetics 2010),
-and fills a gap in practical workers' toolboxes and intuitions surrounding the geographic aspects of adaptation.
+and fills a gap in empirical workers' toolboxes and intuitions surrounding the geographic aspects of adaptation.
 
@@
 imply a certain degree of constraint imposed on evolution by the genomic architecture underlying the trait.
-Furthermore, as landscape genomics datasets become available, 
-it is important to know (but almost entirely unknown) 
-what proportion of the genome may be responding to patchy selection in a similar manner but on a non-obvious phenotype;
-here we contribute to building this intuition.
+Over the coming decade a large number of population genomics datasets
+with fine-grain geographic sampling will be developed, these will likely highlight many cases of geographical convergent
+evolution. Our model helps build intuition about the scale at which we
+should expect such convergent evolution among environmental patches,
+and so may provide an important reference point in guiding the design of
+these studies.
 
@@
 Mark Kirkpatrick (\href{mailto:kirkp@austin.utexas.edu}{kirkp@austin.utexas.edu}),
-and Nick Barton (\href{mailto:nick.barton@ist.ac.at}{nick.barton@ist.ac.at}).
+and Nick Barton
+(\href{mailto:nick.barton@ist.ac.at}{nick.barton@ist.ac.at}). Any of
+these reviewers may also be a suitable external editor (as Coop is an
+associate editor).
 

### 2014-07-08 patchy-selection-paper.tex

@@
 although these two factors will determine the time-scale over which adaptation will occur. 
-This is in contrast with models of panmictic populations \citep{softsweepsI} 
+This is in contrast with models of panmictic populations
+\citep{softsweepsI, MesserPetrov, Wilson-softsweep} 
 and geographically spread populations adapting to homogeneous selection pressures \citep{ralph2010parallel}, 
@@
 aids establishment of new alleles and the spread of existing
-alleles.
+alleles \citep[but see][for the complications of varying population sizes]{Wilson-softsweep}.
 

### 2014-07-08 patchy-selection-paper.tex
added author summary

@@
 
-% The Spatial Resolution of Convergent Adaptation
-% Convergent Evolution in Patchy Landscapes
-% Parallel Evolution in Patchy Landscapes
-
-\title{
+% short title:  Convergent Evolution on Patchy Landscapes
+\title 
+{
 Convergent Evolution During Local Adaptation to Patchy Landscapes
@@
 \end{center}
+
+% Keywords
+% convergent evolution; parallel evolution; MC1r; local adaptation; landscape genetics; branching processes; reaction-diffusion equations
 
@@
 
+\textbf{Author Summary:}
+Often, a large species range will include patches where the species differs because it has adapted to locally differing conditions.
+For instance, rock pocket mice are often found with a coat color that matches the rocks they live in, 
+and the difference in coloration is known to be controlled genetically.
+Sometimes, similar genetic changes have occurred independently in different patches,
+suggesting that there were few accessible ways to evolve the locally adaptive form.
+However, the genetic basis could also be shared if
+rare migrants may carry the locally beneficial genotypes between nearby patches,
+despite being at a disadvantage between the patches.
+We use a mathematical model of random migration to determine how quickly adaptation is expected to occur through new mutation or migration from other patches,
+and study in more detail what we would expect successful migrations between patches to look like.
+The results are useful for determining whether similar adaptations in different locations are likely to have the same genetic basis or not,
+and more generally in understanding how species adapt to patchy, heterogeneous landscapes.
 }

### 2014-07-09 patchy-selection-paper.tex
plos section changes

@@
 %%%%%%%%%%%%%
-\section{Methods}
+\section{Results}
 
@@
 or if $p(x)$ is symmetric about the edge of the patch.
-We examine this more generally via exact calculation of $p(x)$ in Appendix \ref{apx:establishment_sims}.
+We examine this more generally via exact calculation of $p(x)$ in Section \ref{apx:establishment_sims}.
 
@@
 
-\bibliography{standing_patches_refs}
-
-\appendix
-
 %%%%%
-\section{Numerical calculation of the probability of establishment}
+\section{Materials \& Methods}
+
+\subsection{Numerical calculation of the probability of establishment}
 \label{apx:establishment_sims}
@@
 \end{figure}
+
+\bibliography{standing_patches_refs}
+
+\appendix
 

### 2014-07-09 patchy-selection-paper.tex
plos publications format hoops
began simon's edits

@@
 \usepackage{amsmath,amssymb}
-\usepackage{natbib}
 \usepackage{marginnote}
@@
 \usepackage{color}
+\usepackage{natbib}
 \usepackage[hidelinks]{hyperref}
-%\usepackage{latexml} % for \iflatexml
+
+% \usepackage{cite}
+
+% Use doublespacing - comment out for single spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm 
+\textheight 21cm
+
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
+
+% turn this on to doublespace and put all figures at the very end.
+\newif\ifsubmissionversion
+\submissionversionfalse
+\newif\ifincludesupplement
+\includesupplementfalse
+
+\ifsubmissionversion
+\usepackage{endfloat}
+\fi
+
+\ifsubmissionversion
+% \linespread{1.6}
+\doublespacing
+\fi
+
+% Use the PLoS provided bibtex style
+% OR NOT \bibliographystyle{plos2009}
+\bibliographystyle{abbrvnat}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+\pagestyle{myheadings}
+%% ** EDIT HERE **
 
@@
 }
+% Simon: I think the title only partly reflects the scope of the paper (adaptation by migration being underrepresented).
+What about "Relative importance of migration and convergent evolution for local adaptation in patchy landscapes" or similar?
 
@@
 constraint imposed on evolution by the underlying genetic
-architecture versus how how effectively selection acts to inhibit the geographic spread of locally
+architecture versus how effectively selection acts to inhibit the geographic spread of locally
 adapted alleles.
@@
 We address the following questions: 
-How long does it take for a new, locally adapted allele to appear in a
-patch of habitat where it is favored through new mutation?
+How long does it take for a novel, locally adapted allele to appear in a
+patch of habitat where it is favored through mutation?
 Or, through migration from another, already adapted patch?
@@
 Such evolution of convergent phenotypes is favored by higher mutational input, 
-i.e.\ larger mutational target and/or population size \citep{softsweepsII}.
+i.e.\ higher total mutational rate and/or population size \citep{softsweepsII}.
 The geographic distribution of populations can also affect the probability of parallel mutation within a species:
@@
 to discuss to the probability of parallel adaptation, equation \eqref{eqn:parallel_prob}.
-To understand the genomic signal left by adaptations shared by
-migration in section \ref{ss:haplotype_length} we approximate the time
-it will take for an allele to transit between patches,
-equation \eqref{eqn:mean_tau}, and use this to approximate the length of
-haplotype that hitchhikes with it,
-equation \eqref{eqn:haplotype_length}. We then quantify how quickly this
-haplotype decays within patches equation \eqref{eqn:rec_rate_within} and
-use this to obtain properties about the equilibrium distribution of
-haplotype lengths, equation 
-\eqref{eqn:within_haplotype_length}  in section \ref{ss:patch_haplotype}. 
+To understand the genomic signal left by adaptations shared by migration, 
+in section \ref{ss:haplotype_length}, 
+we approximate the time it will take for an allele to transit between patches,
+(equation \eqref{eqn:mean_tau}),
+and use this to approximate the length of haplotype that hitchhikes with it (equation \eqref{eqn:haplotype_length}). 
+We then quantify how quickly this haplotype decays within patches (equation \eqref{eqn:rec_rate_within}) 
+and use this to obtain properties about the equilibrium distribution of haplotype lengths 
+(equation \eqref{eqn:within_haplotype_length}), in section \ref{ss:patch_haplotype}. 
 Finally, in section \ref{ss:applications} we apply this work to understand the geographic scale of
@@
 
+\ifsubmissionversion
+\processdelayedfloats
+\fi
+
 \end{document}

### 2014-07-10 patchy-selection-paper.tex
acknowledgements
put in simon's changes and fixed up bibliography

@@
 
-\bibliographystyle{plainnat}
-
 % short title:  Convergent Evolution on Patchy Landscapes
@@
 }
-% Simon: I think the title only partly reflects the scope of the paper (adaptation by migration being underrepresented).
-What about "Relative importance of migration and convergent evolution for local adaptation in patchy landscapes" or similar?
 
@@
 will be larger if these adaptive alleles are maladapted in intervening environments,
-since such adverse conditions can strongly spatially restrict the spread of locally adapted alleles \citep{slatkin1973geneflow}.
+since such adverse conditions can strongly restrict the spread of locally adapted alleles \citep{slatkin1973geneflow}.
 
@@
 
-Consider a population spread across a landscape, 
-within which are patches of habitat to which individuals are (initially) poorly adapted,
-surrounded by large areas to which the population is well adapted.
+Consider a population spread across a landscape to which it is generally well adapted,
+but within which are patches of habitat to which individuals are (initially) poorly adapted.
 (When we refer to ``patches'' it is to these pieces of poor habitat.)
@@
 $1-s_m$  when rare in the intervening areas, with $s_p$ and $s_m$ both positive.
-(Here we take ``fitness'' to be the intrinsic growth rate of the type when rare.)
+(Here we define ``fitness'' to be the intrinsic growth rate of the type when rare.)
 To simplify matters, we assume that the disadvantage $s_m$ 
 is sufficiently large that, on the relevant timescale,
-the allele is very unlikely to fix locally in the regions where it is at a disadvantage.
+the allele is very unlikely to fix locally in these intervening regions.
 (The case where $s_m=0$ requires a different approach, which we do not treat here.)
@@
 We are interested in the establishment of mutations in the ``poor'' patches by either
-migration or mutation, and so are mainly interested in whether the allele
+migration or mutation, and so will focus on whether the allele
 can escape initial loss by drift when rare. 
-Therefore, we do not have to specify the fitness of the homozygote; 
+Therefore, we need not specify the fitness of the homozygote; 
 only that the dynamics of the allele when rare 
@@
 We will make use of several previous results from the literature on
-migration in spatially varying selection models. 
+migration in models of spatially varying selection.
 \citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
@@
 this probability of establishment of a new mutation decays exponentially with distance from the patch, 
-with the scale given by $\sigma/\sqrt{s_m}$. Mutations appearing within the patch have probability of establishment
+with the scale given by $\sigma/\sqrt{s_m}$. 
+Mutations appearing within the patch have probability of establishment
 strictly less than the probability for a panmictic population,
@@
 This latter panmictic probability of establishment we denote $p_e$,
-and often approximate by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
+and often approximate $p_e$ by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
 This result holds quite generally for a geographically spread population that experiences a uniform selection
@@
 
-Let $p(x)$ be the probability that a new mutant $B$ allele arising at location $x$
-relative to the patch fixes within the poor habitat patch.
+Let $p(x)$ be the probability that a new mutant $B$ allele that arises at location $x$
+relative to the center of the patch fixes within the poor habitat patch.
 Under various assumptions, precise expressions can be found for $p(x)$ \citep{barton1987establishment},
@@
 (supposing that no $B$ allele arises independently by mutation in the meantime).
-To determine this, we dissect the process by which an allele transits between the patches,
-along the way obtaining other useful information about the genealogy of $B$ alleles.
+To determine this, 
+study the fine-scale genealogy of alleles that transit between patches,
+obtaining along the way other useful information about the genealogy of $B$ alleles.
 
@@
 To be clear about the assumptions (especially regarding scaling) implicit here,
-we provide a derivation in Appendix~\ref{apx:eqfreq}; 
+we provide a derivation in Appendix~\ref{apx:eqfreq}, 
 as well as justification for the asymptotics in Appendix \ref{apx:asymptotics}.
@@
 and interactions between family members are weak,
-then these families can be approximated by a collection of independent spatial branching processes
+then these families can be approximated 
+by a collection of independent spatial branching processes
 whose members are ignored if they enter the original patch.
@@
 and the quantity we wish to compute,
-the effective rate of at which \emph{families of} migrant $B$ alleles establish in a new patch at displacement $x$ from the original,
+the effective rate at which \emph{families of} migrant $B$ alleles establish in a new patch at displacement $x$ from the original,
 is
@@
 All terms except the ``outflux of families'' are calculations with spatial branching processes;
-and we can solve for this outflux using the established form for $q(r)$ in \eqref{eqn:gestalt_q}.
-Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
--- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.
+and we can solve for this outflux using the established form for $q(x)$ in \eqref{eqn:gestalt_q}.
+% Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
+% -- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.
 
@@
 To understand the process of adaptation by migration,
-we therefore need to look more closely at the rare families of $B$ alleles far from the patch where it is established.
+we therefore need to look more closely at the rare families of $B$ alleles 
+far from the original patch where it is already established.
 Consider one such family,
 that lives far from the original patch.
-As noted above, we will model this by assuming that each individual
+As noted above, we will model this as a branching process,
+i.e.\ by assuming that each individual
 migrates, then reproduces, independently of all other family members.
-The mean squared migration distance is $\sigma^2$, takes the migrant in a random direction,
-and if migration takes the individual into a patch where the $B$ allele is favored, 
+Migration takes the individual in a random direction
+for a random distance whose mean square is $\sigma^2$,
+and if the individual arrives to a patch where the $B$ allele is favored, 
 we drop the individual from our family.
@@
 we can decouple reproduction and migration
-by first forming a tree in which each individual has a random number of offspring
+by first forming a tree in which each individual has a random number of offspring,
 then recursively assigning spatial locations to each individual in the tree,
 and finally pruning any branches that happen to wander into a patch.
-By our definition of fitness, the mean number of offspring per individual is $\exp(-s_m)\approx 1-s_m$,
+By our definition of fitness, the mean number of offspring per individual is $e^{-s_m}\approx 1-s_m$,
 and the variance is $\xi^2$.
@@
 Since each individual has on average $e^{-s_m}$ offspring,
-the expected family size after time $t$ is $e^{t s_m}$.
+the expected family size after time $t$ is $e^{- t s_m}$.
 Therefore, if we define $k_e(t)$ to be the chance that the family
@@
 It turns out that if the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jagers1975branching},
-then $K_t$ has a limiting distribution: $K_t \dconv K$ 
+then $K_t$ has a limiting distribution: $K_t \dconv K$,
 and $e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
@@
 When we condition on survival until $t$, we condition on existence of
-at least one ``trunk'' lineage surviving from time $0$ to time $t$ (the long red
+at least one trunk lineage surviving from time $0$ to time $t$ (the long red
 lineage in Figure \ref{fig:branching_decomp}).
@@
 In other words, the genealogy of a family that survives until $t$
-can be decomposed into a ``trunk'' that lasts until $t$
-which sprouts independent ``branches'' that may or may not survive until $t$.
-Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
+can be decomposed into a trunk that lasts until $t$
+and that sprouts independent branches that may or may not survive until $t$.
+Specifically, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
 % (so $\P\{Z_t = 0\} = k_e(t)$),
@@
 This suggests that we add in the spatial motion by following only the motion of the ``trunk''
-(i.e.\ the red line in figure \ref{fig:branching_decomp}),
-whose motion is is approximately Brownian, with variance $\sigma^2$ per generation.
+(i.e.\ the red line in figure \ref{fig:branching_decomp}).
+Its motion is approximately Brownian, with variance $\sigma^2$ per generation.
 (Recall $\sigma$ is the dispersal distance.)
@@
 we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
-of the outflux $\gamma$ at that point, multiplied by the relevant second quantity.
+of the outflux $\gamma$ at that point.
 By examining the constituent terms,
@@
 and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
-The \emph{mean occupation density} of $Z$ can be thought of 
+The \emph{mean occupation density} of $Z$ at location $y$
+can be thought of 
 as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
-Concretely, it is the density of $\E[\int_0^t Z_t(\cdot) dt]$,
-i.e.\ the smooth function $u(x,y)$ that, integrated over a set $S$, 
+It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
+i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
 gives the expected total number of individuals ever living in $S$:
@@
 the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
-Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch,
-which we assume is circular, has relatively small area $A$ and is centered at position $y$.
+Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
+We assume this new patch is circular, 
+has relatively small area $A$ and is centered at position $y$.
 % The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
@@
 % If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
-It would be possible to analyze directly the probability that a given migrant family establishes in the patch,
+It would be possible to analyze this probability directly,
 as in \citet{barton1987establishment};
-but we make a simpler approximation, approximating this probability by the chance that the ``trunk'' hits the new patch,
-multiplied by the chance that at least one member of the family escapes demographic stochasticity,
-successfully establishing in the new patch.
+but here we take a simpler route, 
+approximating this by the chance that the trunk hits the new patch,
+multiplied by the chance that at least one member of the family escapes demographic stochasticity
+and successfully establishes in the new patch.
 Write $h_A(x,y)$ for the probability that the Brownian trunk
@@
 \end{align}
-where $g(A)$ is the total expected occupation time of the patch given the family arrives,
-and recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
+where $g(A)$ is the total expected occupation time of the patch given the family arrives.
+Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
 Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
@@
   % \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
-  &\approx C \frac{ \rho s_p  }{ \xi^2 s_m } \left( |x| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |x| \sqrt{2 s_m} / \sigma) .
+  &\approx C \frac{ \rho s_p  }{ \xi^2 s_m } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
 \end{align}
@@
 $f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
-the dependence is very weak:
-for instance, the width of a circular patch only very weakly affects the chance 
+the dependence is very weak.
+For instance, the width of a circular patch only very weakly affects the chance 
 of establishment of a new migrant that appears on its edge,
@@
 We now turn to the question of whether the separated patches adapt by parallel genetic changes 
-or by the spread of a migrant allele between the patches
-(or both, or neither).
+or by the spread of a migrant allele between the patches.
 As only a single mutation is required for individuals to adapt to the
@@
 
-We will assume that once a $B$ allele is introduced (by migration or mutation) and successfully begins to establish,
-it becomes established in the poor habitat patch rapidly. 
+We will assume that once a $B$ allele is introduced (by migration or mutation) 
+it becomes established in the poor habitat patch rapidly
+if it escapes loss by demographic stochasticity.
 Under this approximation, and the ``selective exclusion'' just discussed,
@@
 
-The first thing to notice is that effective migration and mutation rates will be on the same order if 
+An important point to notice is that effective migration and mutation rates will be on the same order if 
 $A \mu \approx (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
-where recall that $R$ is the distance between the patches,
+where $R$ is the distance between the patches,
 and $A$ is the area of the not-yet-adapted patch.
@@
 If we take the mutational influx in a patch to be one per 1,000 generations ($\rho A \mu = .001$)
-and the population density to be 1 ($\rho =1$, so $\log(A\mu)\approx 7$)
-then adaptation is mostly parallel for patches separated by gaps larger than $R/\sigma > (7+\log(s_m))/\sqrt{s_m}$.
+and the population density to be 1 ($\rho =1$, so $-\log(A\mu)\approx 7$)
+then adaptation is mostly parallel for patches separated by gaps larger than $R/\sigma > (7-\log(s_m))/\sqrt{2s_m}$.
 If the selective pressure is fairly strong (say, $s_m=.05$),
 then for migration to be slower than mutation,
-the distance between patches must be at least 30 times the dispersal
-distance (i.e.\ $R/\sigma > 30$) for convergence to occur.
+the distance between patches must be at least 31.6 times the dispersal
+distance (i.e.\ $R/\sigma > 31.6$) for convergence to occur.
 If $s_m$ is much smaller, say $s_m = .001$, 
-then the patches must be separated by several hundred times the dispersal distance (i.e.\ $R/\sigma > 300$).
-The situation does not change much if the influx is quite a bit higher -- even if $A \mu = 1$ (so $\log(A\mu) \approx 2.3$),
-the required separation between patches is only reduced to $R/\sigma > 10$ with $s_m=.05$ and $R/\sigma>150$ for $s_m=.001$.
+then the patches must be separated by several hundred times the dispersal distance (i.e.\ $R/\sigma > 308$).
+The situation does not change much if the influx is quite a bit higher -- even if $A \mu = 1$,
+the required separation between patches is only reduced to $R/\sigma > 9.5$ with $s_m=.05$ and $R/\sigma>154$ for $s_m=.001$.
 On the other hand, if the separation is much smaller than this transitional value, 
@@
 in which the $B$ allele can facilitate local adaptation. 
-Under our assumptions of mutational exclusion and rapid establishment, 
+Under our assumptions of selective exclusion and rapid establishment, 
 each patch will adapt through establishment of a single copy of the $B$ allele, 
@@
 
-If not yet adapted, a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$ (equation \eqref{eqn:mutrate}).
+If not yet adapted, a focal patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_p \mu \rho A/\xi^2$ (equation \eqref{eqn:mutrate}).
 Suppose that patches $1, \ldots, k$ are already adapted,
-and that these patches are at distances $r_1, \ldots, r_k$ away from this unadapted patch.
+and that these patches are at distances $R_1, \ldots, R_k$ away from this unadapted patch.
 The total rate of adaptation through migration, from equation \eqref{eqn:migrate}, is
-\begin{equation}
-  \migrate(r_i) = C \frac{ \rho s_p }{\xi^2 s_m} \sum_{i=1}^{k} \left(r_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m} /\sigma\right),
+\begin{equation} \label{eqn:total_migrate}
+  \sum_{i=1}^k \migrate(R_i) = C \frac{ \rho s_p }{\xi^2 s_m} \sum_{i=1}^{k} \left(R_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m} /\sigma\right).
 \end{equation}
-and if the patch adapts through migration, the probability the adapted allele 
+If the patch adapts through migration, the probability the adapted allele 
 comes from patch $i$ is 
-\begin{align}  
-  \frac{\left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
-  /\sigma\right)} {\sum_{j=1}^{k}  \left(r_i \right)^{-(d-1)/2} \exp\left(- r_i \sqrt{2 s_m}
+\begin{align}   \label{eqn:migprob}
+  \frac{\left(R_i \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m}
+  /\sigma\right)} {\sum_{j=1}^{k}  \left(R_i \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m}
     /\sigma\right) } .
 \end{align}
-These specify the transition rates of the process.
-
-Since the compound parameter $s_p \rho / \xi^2$ is common to both rates,
+Equations \eqref{eqn:total_migrate} and \eqref{eqn:migprob} specify the transition rates of the process.
+Since the compound parameter $s_p \rho / \xi^2$ is common to both migration and mutation rates,
 it functions only as a time scaling of the process, 
@@
 Also note that we can rescale time by the typical patch size, and introduce a parameter, say $a_k = A_k/\bar A$,
-making the properties (other that the time scaling) of the discrete model independent of the \emph{numerical sizes} of the demes themselves.
+making the properties (other than the time scaling) of the discrete model independent of the \emph{numerical sizes} of the demes themselves.
 This is complementary to the results of \cite{softsweepsII}, who showed that multiple mutations are likely to arise \emph{within} a panmictic deme
@@
 
-The resulting discrete process is quite general,
+The resulting spatially discrete process is quite general,
 but if we oversimplify back to the island model, it is tractable:
 suppose the probability of adaptation by mutation is the same for each patch
-and by distance is the same for each pair of patches.
+and the probability of migration is the same between each pair of patches.
 The resulting process is a continuous-time version of the ``Chinese restaurant process''
@@
 and so the final partition of types across demes has the Ewens
-sampling distribution with parameter $A \mu s_m/ \exp (-sqrt{2 s_m}r/\sigma)$.
+sampling distribution with parameter $A \mu s_m/ \exp (-\sqrt{2 s_m}R/\sigma)$.
 Now we can compute most properties we might want about the process.
@@
 The high connectedness of the discrete deme island model means that the expected number of distinct alleles
-grows with the $\log$ of the number of demes, 
-which strongly contrasts with the continuous spatial model 
+grows with the $\log$ of the number of demes.
+This strongly contrasts with the continuous spatial model 
 where the local nature of dispersal means that doubling the species range will double the number of mutations expected.
@@
 share a fairly long stretch of haplotype. 
-Pairs of individuals within a patch will share a haplotype of mean genetic length of
+Pairs of adapted individuals within a patch will share a haplotype of mean genetic length of
 about $\log(\rho A s_p)/s_p$ around the selected site, 
@@
 \citep[otherwise see][]{barton2013genetic}.
-This association gets slowly whittled down by recombination during interbreeding at the edge of the patch;
+This association gets slowly whittled down by recombination during interbreeding at the edge of the patch,
 but there will always be longer LD nearby to the selected site.
@@
 number of generations the allele spends ``in transit'', 
-because each reproductive event in locations where the allele is at
-low frequency will likely recombine haplotypes of the $b$ allele closer to the adaptive allele. 
-In this case, the haplotype is literally hitchhiking across geography!
+because while very rare, the allele will be mostly in heterozygotes,
+and so each generation provides another opportunity for a different haplotype to recombine closer to the transiting $B$ allele.
+A large, linked haplotype may still arrive and fix in the new patch,
+in which case the haplotype has literally hitchhiked across geography!
 Figure~\ref{fig:lineagesmotion} shows a simulation of such an event,
@@
 In so doing, we neglect 
-recombination between $B$ alleles (since they are at low density in transit, we assume this is rare),
+recombination between $B$ alleles (since they are at low density in transit),
 and the possibility that more than one successful migrant family is in transit at once
-(so that faster migrants would be more likely to have arrived first);
-but these factors should be minor if the rest of our framework fits.
+(so that faster migrants would be more likely to have arrived first).
+These factors should have a minor impact if the rest of our framework fits.
 
@@
 Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
-replacing this random time by an exponential with rate $s_m$.
+supposing that the random lifetime of a family is exponentially distributed with mean $1/s_m$.
 Once we do this, 
 we can use standard results on hitting times of $d$-dimensional Brownian motion
-that is killed at rate $s_m$ -- \citet[see][(2.2.0.1 and 4.2.0.1)]{borodin2002handbook}.
+that is killed at rate $s_m$ \citep[see][2.2.0.1 and 4.2.0.1]{borodin2002handbook}.
 In particular, if the patch is circular with radius $w$ and lies at distance
@@
 and $\var[L] = 2s_m\sigma^2/R^2 + 4 \sigma^3 \sqrt{2s_m}/R^3 + 5 \sigma^4 / R^4$.
-For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior;
+For two dimensions, asymptotics of Bessel functions show that $L$ has the same tail behavior,
 but how other properties might change is unclear.
@@
 is roughly the length one would expect given the mean transit time.
-Note that we have become used to seeing $\sigma$ divided by $\sqrt{s_m}$, rather than multiplied;
-it appears here because $\sigma/\sqrt{s_m}$ is a length scale, 
-and to convert it to a speed this is divided by $1/s_m$.
+Note that we have become used to seeing $\sigma$ divided by $\sqrt{2s_m}$, rather than multiplied;
+it appears here because $\sigma/\sqrt{2s_m}$ is a length scale, 
+and to convert it to a speed this is divided by $1/2s_m$.
 This intuitive picture can be expanded by a central limit theorem:
@@
 The initial haplotype that sweeps within a patch 
-will be dispersed over time by recombination;
-likewise, the haplotype that is shared between patches coadapted by migration
+will be dispersed over time by recombination.
+Likewise, the haplotype that is shared between patches coadapted by migration
 (equation~\ref{eqn:haplotype_length}) will also break down. 
@@
 time our lineage spends in geographic regions where $b$ is common. 
-Consider the lineage of a single $B$ allele, moving back through time,
+We will follow the lineage of a single $B$ allele
 in a patch at migration-selection balance, 
-i.e.\ with the local proportion of $B$ alleles near $x$ equal to $p(x)$,
+i.e.\ with the local proportion of $B$ alleles near $x$ equal to $q(x)$,
 and suppose that the lineage stays close enough to the patch 
-that fluctuations about $p(x)$ are small. 
+that fluctuations about $q(x)$ are small. 
 (Note that this assumption differs from
-previous sections, but here our focus is on a typical lineage in the
-patch rather than the rare lineages that make it far from the patch).
+previous sections, but here our focus is on a typical lineage 
+in the patch rather than the rare lineages that make it far from the patch).
 If we suppose that type $B$ individuals at location $y$ reproduce at rate $1+s(y)$,
 and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
-then the total rate of influx of $B$ alleles from $y$ into $x$ is $\rho p(y) (1+s(y)) m(y,x)$.
-As argued by \citet{hudson1988coalescent}, the rate at which a $B$ lineage moves from $x$ to $y$
-is this influx divided by the local abundance of $B$ alleles,
-or $p(y) (1+s(y)) m(y,x) / p(x) $.
+then the total rate of influx of $B$ alleles from $y$ into $x$ is $\rho q(y) (1+s(y)) m(y,x)$.
+As argued by \citet{hudson1988coalescent}, 
+this implies that the chance that a $B$ allele at $x$
+has been inherited from a parent at $y$ in a time interval $dt$
+is equal to this influx divided by the local abundance of $B$ alleles,
+i.e.\ $dt \times q(y) (1+s(y)) m(y,x) / q(x)$.
+This therefore describes the movement of a lineage backwards in time.
 After the same large-population, continuous-space limit that underlies the rest of this paper (Appendix~\ref{apx:eqfreq}),
-if the assumption that fluctuations in $p(x)$ are small,
+if the assumption that fluctuations in $q(x)$ are small,
 this motion should converge to a diffusion process \citep[as assumed by e.g.][]{hallatschek2008surfing},
@@
 is pulled toward regions with more $B$ alleles
-by a drift term $(1+s(x))\frac{d}{d\,x} \log( p(x) (1+s(x)) )$,
+by a mean displacement term $(1+s(x))\frac{d}{d\,x} \log( q(x) (1+s(x)) )$,
 where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
 and moves at speed $\sqrt{1+s(x)}$ (which is nearly constant, since $s(x)$ is small).
-In It\^o differential notation, 
+In It\=o differential notation, 
 the location $X_t$ of a $B$ lineage $t$ generations ago satisfies
 \begin{align}
-    d X_t &= (1+s(X_t)) \frac{d}{dx} \log( p(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,
+    d X_t &= (1+s(X_t)) \frac{d}{dx} \log( q(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,
 \end{align}
@@
 \begin{align}
-  \pi(x) = \frac{ p(x)^2 (1+s(x)) }{ \int_y p(y)^2 (1+s(y)) dy } . \label{eqn:lineagestatdist}
-\end{align}
-A lineage at a locus at genetic distance $r$ that begins linked to a $B$ allele
-moves as a $B$ lineage up until the most recent time
-it recombined from a $b$ background to a $B$ background,
-which occurs when at $x$ with rate $r (1-p(x)) (1+s(x)/2)$.
+  \pi(x) = \frac{ q(x)^2 (1+s(x)) }{ \int_y q(y)^2 (1+s(y)) dy } . \label{eqn:lineagestatdist}
+\end{align}
+A lineage at a locus at genetic distance $R$ from a $B$ allele
+moves as a $B$ lineage until it recombines onto a $b$ background,
+which occurs when at $x$ with rate $r (1-q(x)) (1+s(x)/2)$.
 The Feynman-Kac formula could then be used to write down differential equations
@@
 If recombination is on a slower time scale than mixing within the patch,
-then the time until the last recombination is the integral of $r (1-p(x)) (1+s(x)/2)$ 
+then the time until the last recombination is the integral of $r (1-q(x)) (1+s(x)/2)$ 
 against the stationary distribution (equation~\ref{eqn:lineagestatdist}).
-Within the patch, $p(x) \approx 1$,
+Within the patch, $q(x) \approx 1$,
 so recombination off the $B$ background is unlikely to occur,
 and so the rate of recombination depends to a first approximation 
-on the proportion of the patch in which $p(x)$ is intermediate,
+on the proportion of the patch in which $q(x)$ is intermediate,
 i.e.\ where both $B$ and $b$ are present.
-This region where $p(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
-As we increase the size of a patch, the area on which $p(x)$ is intermediate scales with the perimeter of the patch.
+This region where $q(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
+As we increase the size of a patch, the area on which $q(x)$ is intermediate scales with the perimeter of the patch.
 % Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
-% multiplied by $p(x)$ in these intermediate regions, where we assume $p(x) \approx 1/2$. 
+% multiplied by $q(x)$ in these intermediate regions, where we assume $q(x) \approx 1/2$. 
 The per generation rate of recombination of a locus linked to the $b$ background is thus
@@
   \sqrt{A/\pi} \right)}{A+\pi \sigma/\sqrt{s} \left( \sigma/\sqrt{s} +
-  \sqrt{A/\pi} \right) }    \qquad & d=2 ,  
+  \sqrt{A/\pi} \right) }    \qquad & d=2 .
 \end{cases}
@@
 \end{align}
-so that $T$ generations after the founding of a patch we expect a
+It follows that $T$ generations after the founding of a patch we expect a
 haplotype of genetic length $1/(r_\text{eff}T)$ Morgans to remain linked to
@@
 Note that the length of haplotype
-decays at a rate roughly proportional to $\sqrt{s}/ \sigma$, as the narrower the cline the less
-opportunity for recombination to move our lineage off the $B$ background.
+decays at a rate roughly proportional to $\sqrt{s}/ \sigma$, as the narrower the cline 
+the less opportunity for recombination to move our lineage off the $B$ background.
 
@@
 The rate of recombination and coalescence will be on the same order if
-$\xi^2 /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$, 
-and hence the genetic length of the haplotype shared between pairs of
+$\xi^2 /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$.
+Hence the genetic length of the haplotype shared between pairs of
 individuals within the same patches at equilibrium is on the order of
@@
 This dark allele differs from the light allele by four amino acid changes, 
-and has a dominant or partially dominant effect depending on the measure of coat color assessed. 
+and has a dominant or partially dominant effect depending on the measure of coat color.
 These areas of dark rock are separated from each other by a range of distances, 
@@
 Using $\sigma \approx 1$km \citep{french1968dispersal,allred1963range}, 
-these cline widths imply that $s_m=1/\sqrt{3}$ and $1/\sqrt{30}$ are reasonable values.
+these cline widths imply that $s_m=1/9$ and $1/900$ are reasonable values.
 
@@
 \paragraph{Standing variation} 
-We have tended to focus on relative rates of adaptation,
+We have focused on relative rates of adaptation,
 since we are thinking of applications in which we know that adaptation has occurred,
@@
 any adaptation that does occur may have to make use of standing variation.
-The unstructured case was studied by \citet{softsweepsI};
-we study the continuous, spatial case in an upcoming paper.
+The case of a panmictic population was studied by \citet{softsweepsI},
+and we study the case of a continuous, spatial population in an upcoming paper.
 If parallelism in local adaptation of the sort we study here is due to standing variation
@@
 then the dynamics of adaptation should not depend strongly on migration patterns
-(but note that the initial spatial distribution of standing variation may be).
+(but note that the initial spatial distribution of standing variation may).
 
 \paragraph{Dominance}
-We have mostly swept the issue of dominance under the rug
+We have mostly ignored the issue of dominance 
 by dealing with essentially haploid models,
@@
 in a theoretical context.
+It is simliarly unclear how the model should extend to a polygenic trait.
 
@@
 with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
-In addition, over long time-scales, very rare chance events (mice sucked up in hurricanes and the like) 
+In addition, over long time-scales, very rare chance events \citep[mice carried off by hurricanes and the like][]{censky1998overwater,nathan2008mechanisms}
 could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
@@
 \paragraph{Other patch geometries}
-Here we have studied circular patches of habitat at long distances from each other,
-while real habitat is can be much more complex
--- archipelagos of patches of varying sizes, or connected by long, skinny corridors, for instance.
-The work of \citet{cantrell1991diffusive} comes closest to a general theory of balanced polymorphisms in such habitats,
-and it is possible that our techniques could be applied in their much more general setting,
+We have studied circular patches of habitat at long distances from each other.
+Real habitat geometry can be much more complex,
+e.g.\ with archipelagos of patches of varying sizes, 
+or patches connected by long, skinny corridors, for instance.
+The work of \citet{cantrell1991diffusive} comes closest to a general theory of balanced polymorphisms in such habitats.
+It is possible that our techniques could be applied in their much more general setting,
 as both are based, fundamentally, on branching process approximations.
@@
 Our work suggests that even quite geographically close populations 
-may be forced to locally adapt convergently when migration is geographically limited
+may be forced to locally adapt by repeated, convergent, \textit{de novo} mutation 
+when migration is geographically limited
 and selective pressures are divergent.
@@
 % Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
+
+\subsection*{Acknowledgements}
+We would like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
+as well as Gideon Bradburd and the rest of the Coop lab for helpful comments on the manuscript.
 

### 2014-07-10 patchy-selection-paper.tex
factor of 2

@@
 this probability of establishment of a new mutation decays exponentially with distance from the patch, 
-with the scale given by $\sigma/\sqrt{s_m}$. 
+with the scale given by $\sigma/\sqrt{2s_m}$. 
 Mutations appearing within the patch have probability of establishment

### 2014-07-10 patchy-selection-paper.tex
small change to discussion/ack

@@
 
-\paragraph{Other patch geometries}
+\paragraph{Other geographic models}
 We have studied circular patches of habitat at long distances from each other.
@@
 as both are based, fundamentally, on branching process approximations.
+It is also interesting to think about the probability of convergent
+adaptation to continuously varying environments, e.g. replicated
+environmental clines.
 
@@
 We would like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
-as well as Gideon Bradburd and the rest of the Coop lab for helpful comments on the manuscript.
+as well as Gideon, and Simon Aeschbacher and the rest of the Coop lab
+for helpful comments and discussion of the manuscript.
 

### 2014-07-10 patchy-selection-paper.tex
moved appendices to methods and made sections non-numbered

@@
 
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
+% % Text layout
+% \topmargin 0.0cm
+% \oddsidemargin 0.5cm
+% \evensidemargin 0.5cm
+% \textwidth 16cm 
+% \textheight 21cm
 
@@
 %%%%%%%%%%% %%%%%%%%%%% 
-\section{Introduction}
+\section*{Introduction}
 
@@
 using a variety of known results and new methods.
-In section \ref{ss:patchymutation} we develop a simple approximation,
+In the section \nameref{ss:patchymutation} we develop a simple approximation,
 equation \eqref{eqn:mutrate}, for the rate at which patches become established by new mutations. 
-The most important conceptual work of the paper occurs in section \ref{ss:patchymigration},
+The most important conceptual work of the paper occurs in the section \nameref{ss:patchymigration},
 where we develop an understanding of the process by which alleles move from an existing patch 
 to colonize a novel patch, culminating in equation \eqref{eqn:migrate} for the rate at which this happens.
-We combine these two results in section \ref{ss:probparallel} 
+We combine these two results in the section \nameref{ss:probparallel} 
 to discuss to the probability of parallel adaptation, equation \eqref{eqn:parallel_prob}.
 To understand the genomic signal left by adaptations shared by migration, 
-in section \ref{ss:haplotype_length}, 
+in the section \nameref{ss:haplotype_length}, 
 we approximate the time it will take for an allele to transit between patches,
@@
 and use this to obtain properties about the equilibrium distribution of haplotype lengths 
-(equation \eqref{eqn:within_haplotype_length}), in section \ref{ss:patch_haplotype}. 
-Finally, in section \ref{ss:applications} we apply this work to understand the geographic scale of
+(equation \eqref{eqn:within_haplotype_length}), in the section \nameref{ss:patch_haplotype}. 
+Finally, in the section \nameref{ss:applications} we apply this work to understand the geographic scale of
 convergent evolution in \textit{Chaetodipus intermedius}.
@@
 %%%%%%%%%%%%%
-\section{Results}
+\section*{Results}
 
 %%%
-\subsection{Model of a patchy landscape}
+\subsection*{Model of a patchy landscape}
 \label{ss:patchyspace}
@@
 %%%%%
-\subsection{Establishment of a locally adaptive allele due to mutational influx}
+\subsection[Establishment by Mutation]{Establishment of a locally adaptive allele due to mutational influx}
 \label{ss:patchymutation}
@@
 or if $p(x)$ is symmetric about the edge of the patch.
-We examine this more generally via exact calculation of $p(x)$ in Section \ref{apx:establishment_sims}.
+We examine this more generally via exact calculation of $p(x)$ in the section \nameref{apx:establishment_sims}.
 
@@
 %%%%%%%
-\subsection{Simulation methods}
+\subsection*{Simulation methods}
 \label{ss:simulations}
@@
 %%%%%%%
-\subsection{Establishment of a locally adaptive allele due to migrational influx}
+\subsection[Establishment by Migration]{Establishment of a locally adaptive allele due to migrational influx}
 \label{ss:patchymigration}
@@
 To be clear about the assumptions (especially regarding scaling) implicit here,
-we provide a derivation in Appendix~\ref{apx:eqfreq}, 
-as well as justification for the asymptotics in Appendix \ref{apx:asymptotics}.
+we provide a derivation in section~\nameref{apx:eqfreq}, 
+as well as justification for the asymptotics in section~\nameref{apx:asymptotics}.
 In one dimension, the equation can be solved to give $q(x)$ 
-in terms of Jacobi elliptic functions; see Appendix \ref{apx:elliptic_integrals}.
+in terms of Jacobi elliptic functions; see appendix \ref{apx:elliptic_integrals}.
 In applications we fit $C$ to the data;
@@
 %%%% %%%%
-\subsubsection{The genealogy of migrant families}
+\subsubsection*{The genealogy of migrant families}
 
@@
 and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-it turns out (see Appendix~\ref{apx:hitting_occupation}) that
+it turns out (see section~\nameref{apx:hitting_occupation}) that
 \begin{align} 
@@
 % \end{align}
-Combining with expressions for $g(A)$ in Appendix~\ref{apx:hitting_occupation},
+Combining with expressions for $g(A)$ in section~\nameref{apx:hitting_occupation},
 we obtain that
@@
 as long as the patch is large enough.
-For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see Appendix~\ref{apx:hitting_occupation}.
+For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{apx:hitting_occupation}.
 
 %%%%%%%%%
-\subsection{The probability of parallel adaptation between patches.} 
+\subsection[Probability of Parallel Adaptation]{The probability of parallel adaptation between patches.} 
 \label{ss:probparallel}
@@
 %%%%%%%%%%%%
-\subsubsection{Multiple patches}
+\subsubsection*{Multiple patches}
 Above we worked with two patches separated by distance $R$;
@@
 %%%%%%%%%
-\subsection{Length of the hitchhiking haplotype}
+\subsection[Haplotypes Shared Between Patches]{Length of the hitchhiking haplotype}
 \label{ss:haplotype_length}
@@
 %%%%%%% %%%%%%%%%%
-\subsection{Length of the shared haplotype within a patch}
+\subsection[Haplotypes Shared Within a Patch]{Length of the shared haplotype within a patch}
 \label{ss:patch_haplotype}
@@
 This therefore describes the movement of a lineage backwards in time.
-After the same large-population, continuous-space limit that underlies the rest of this paper (Appendix~\ref{apx:eqfreq}),
+After the same large-population, continuous-space limit that underlies the rest of this paper (section~\nameref{apx:eqfreq}),
 if the assumption that fluctuations in $q(x)$ are small,
@@
 %%%%%%%%%%
-\subsection{Applications} 
+\subsection*{Applications} 
 \label{ss:applications}
@@
 %%%%%%% %%%%%%% %%%%%%%
-\section{Discussion} 
+\section*{Discussion} 
 \label{ss:discussion}
@@
 We would like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
-as well as Gideon, and Simon Aeschbacher and the rest of the Coop lab
+as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
 for helpful comments and discussion of the manuscript.
@@
 %%%%%
-\section{Materials \& Methods}
-
-\subsection{Numerical calculation of the probability of establishment}
+\section*{Materials \& Methods}
+
+\subsection*[Probability of Establishment]{Numerical calculation of the probability of establishment}
 \label{apx:establishment_sims}
@@
 
-\bibliography{standing_patches_refs}
-
-\appendix
-
 %%%%%%%%%%%%
-\section{The equilibrium frequency}
+\subsection*[Equilibrium Frequency]{The equilibrium frequency}
 \label{apx:eqfreq}
@@
 %%%%%%%%%%%
-\subsection{Asymptotic solution for the equilibrium frequency} 
+\subsection*[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
 \label{apx:asymptotics}
@@
 
+%%%%%%%%% %%%%%%%%%%%%%%
+\subsection*[Hitting and Occupation]{Hitting and occupation}
+\label{apx:hitting_occupation}
+
+Consider a single migrant family $Z$ beginning at $x$,
+and write $S$ for the new patch, which has area $A$ and is centered at $y$.
+Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
+and $\tau_\dagger$ an independent Exponential($s_m$) time.
+The mean occupation time of $Z$ spent in a region $S$ is,
+since the marginal distribution of a single lineage is Brownian,
+\begin{align}
+    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
+           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
+        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
+        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
+        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
+\end{align}
+where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
+and likewise $\E^x$.
+If we define
+$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
+and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\min(t,\tau_\dagger)}$,
+by the strong Markov property this is equal to
+\begin{align}
+    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
+           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] ,
+\end{align}
+where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
+provided $S$ is not pathological.
+If $S$ is circular with area $A$, the latter expectation will not depend on $x$;
+call this $g(A)$.
+
+Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
+(In the main body, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
+If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
+and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
+then this is equal to
+\begin{align}
+    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P\{ \tau_S \in dt \} \\
+    &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P\{ \tau_S \in dt \} \\
+    &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
+\end{align}
+Therefore, we have that
+\begin{align}
+    h(x,S) \E[K] 
+    \approx \P^x[ \tau_S < \tau_\dagger ] 
+    = u(x,S) / g(A) .
+\end{align}
+
+If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
+then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
+\begin{align}
+    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
+          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
+          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) .
+\end{align}
+The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
+is expected to spend inside the disk before $\tau_\dagger$.
+This is integral of the Green function for the Bessel process of the appropriate order, 
+so using \citet{borodin2002handbook}, % (Appendix 1),
+and letting $w$ be the width of the patch,
+in $d=1$,
+\begin{align}
+    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
+  &= \int_0^{2w} \frac{ e^{- y \sqrt{2s_m} }}{\sqrt{2s_m}} dy \\
+  &= (1-e^{-2w\sqrt{2s_m}})/(2s_m)
+\end{align}
+and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
+\begin{align}
+    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
+  &= \int_0^{2w} 2 y K_0(y \sqrt{2s_m}) dy \\
+  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}} y K_0(y) dy \\
+  &= \frac{1}{2s_m}\left(1- 2w\sqrt{2s_m} K_1(2w\sqrt{2s_m}) \right)  .
+\end{align}
+In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.
+
+%%%%%%% %%%%%%%%%%%%
+\subsection*[Outflux of Families]{The outflux of families}
+\label{apx:outflux}
+
+It is of interest to have an expression for the ``outflux of families'',
+$\gamma(r_0)$, 
+defined to be the mean proportion of $B$ alleles at distance $r_0$
+whose lineage traces back to the original patch without venturing further away than $r_0$.
+By expression \eqref{eqn:occupation_integral},
+\[
+    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
+\]
+for any $y$.  
+Take $y=(r,0)$.
+Note that after scaling by $\sqrt{2s_m}/\sigma$,
+\begin{align}
+    \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\|} }{ \|y\|^{-(d-1)/2} e^{-\|y\|} } \\
+        &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r (\|\hat x - x/r\| - 1)} .
+\end{align}
+Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
+so that the expression in the exponent is to, first order in $1/r$,
+\begin{align}
+    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
+            &= - 2 r_0 \cos \theta + O(1/r) .
+\end{align}
+This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
+\begin{align}
+    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} .
+\end{align}
+In one dimension we are done.
+In two dimensions,
+by bounded convergence and \citet{gradshteyn2007table} (3.339),
+\begin{align}
+  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
+            &= C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
+                         &= 2 \pi C' I_0(-2r_0),
+\end{align}
+where $I_0$ is a modified Bessel function of order zero of the first kind.
+The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2s_m} r_0 / \sigma ) $.
+
+%%%%%%%%%%%
+
+\bibliography{standing_patches_refs}
+
+\appendix
+
 %%%%%%%
-\subsection{The exact solution to the equilibrium frequency}
+\section{The exact solution to the equilibrium frequency}
 \label{apx:elliptic_integrals}
@@
 
-%%%%%%%%% %%%%%%%%%%%%%%
-\section{Hitting and occupation}
-\label{apx:hitting_occupation}
-
-Consider a single migrant family $Z$ beginning at $x$,
-and write $S$ for the new patch, which has area $A$ and is centered at $y$.
-Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
-and $\tau_\dagger$ an independent Exponential($s_m$) time.
-The mean occupation time of $Z$ spent in a region $S$ is,
-since the marginal distribution of a single lineage is Brownian,
-\begin{align}
-    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
-           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
-        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
-        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
-        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
-\end{align}
-where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
-and likewise $\E^x$.
-If we define
-$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
-and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\min(t,\tau_\dagger)}$,
-by the strong Markov property this is equal to
-\begin{align}
-    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
-           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] ,
-\end{align}
-where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
-provided $S$ is not pathological.
-If $S$ is circular with area $A$, the latter expectation will not depend on $x$;
-call this $g(A)$.
-
-Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
-(In the main body, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
-If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
-and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
-then this is equal to
-\begin{align}
-    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P\{ \tau_S \in dt \} \\
-    &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P\{ \tau_S \in dt \} \\
-    &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
-    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
-    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
-\end{align}
-Therefore, we have that
-\begin{align}
-    h(x,S) \E[K] 
-    \approx \P^x[ \tau_S < \tau_\dagger ] 
-    = u(x,S) / g(A) .
-\end{align}
-
-If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
-then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
-\begin{align}
-    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
-          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
-          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) .
-\end{align}
-The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
-is expected to spend inside the disk before $\tau_\dagger$.
-This is integral of the Green function for the Bessel process of the appropriate order, 
-so using \citet{borodin2002handbook}, % (Appendix 1),
-and letting $w$ be the width of the patch,
-in $d=1$,
-\begin{align}
-    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
-  &= \int_0^{2w} \frac{ e^{- y \sqrt{2s_m} }}{\sqrt{2s_m}} dy \\
-  &= (1-e^{-2w\sqrt{2s_m}})/(2s_m)
-\end{align}
-and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
-\begin{align}
-    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
-  &= \int_0^{2w} 2 y K_0(y \sqrt{2s_m}) dy \\
-  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}} y K_0(y) dy \\
-  &= \frac{1}{2s_m}\left(1- 2w\sqrt{2s_m} K_1(2w\sqrt{2s_m}) \right)  .
-\end{align}
-In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.
-
-%%%%%%% %%%%%%%%%%%%
-\subsection{The outflux of families}
-\label{apx:outflux}
-
-It is of interest to have an expression for the ``outflux of families'',
-$\gamma(r_0)$, 
-defined to be the mean proportion of $B$ alleles at distance $r_0$
-whose lineage traces back to the original patch without venturing further away than $r_0$.
-By expression \eqref{eqn:occupation_integral},
-\[
-    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
-\]
-for any $y$.  
-Take $y=(r,0)$.
-Note that after scaling by $\sqrt{2s_m}/\sigma$,
-\begin{align}
-    \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\|} }{ \|y\|^{-(d-1)/2} e^{-\|y\|} } \\
-        &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r (\|\hat x - x/r\| - 1)} .
-\end{align}
-Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
-so that the expression in the exponent is to, first order in $1/r$,
-\begin{align}
-    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
-            &= - 2 r_0 \cos \theta + O(1/r) .
-\end{align}
-This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
-\begin{align}
-    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} .
-\end{align}
-In one dimension we are done.
-In two dimensions,
-by bounded convergence and \citet{gradshteyn2007table} (3.339),
-\begin{align}
-  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-            &= C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
-                         &= 2 \pi C' I_0(-2r_0),
-\end{align}
-where $I_0$ is a modified Bessel function of order zero of the first kind.
-The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2s_m} r_0 / \sigma ) $.
-
-%%%%%%%%%%%
-
 \ifsubmissionversion
@@
 \end{align}
-
-\section{Notes: for intro/discussion}
-
-- leimu2008metaanalysis: local adaptation not ubiquitous but common; mostly in large population sizes.
-- robertson1960theory: dependence of the degree of adaptation on Ne 
-- lots of local adaptation stuff on plants and conservation genetics; ties into plasticity and range limits.
-- but also: Neurospora (ellison2011) and salmonids (fraser2011)
-- swamping in Mayr 1963, Animal Species and Evolution?
-- hanson1966effects: original observation of minimum patch size, in simulation?
-- fourneir-level2011: map of arabidopsis local adaptation
-- reviews for local adapt in marine (sanford2011), trees (savolainen2007)
 

### 2014-07-10 patchy-selection-paper.tex
made sections non-numbered CORRECTLY now

@@
 \usepackage[hidelinks]{hyperref}
+
+\setcounter{secnumdepth}{0}
 
@@
 %%%
-\subsection*{Model of a patchy landscape}
+\subsection{Model of a patchy landscape}
 \label{ss:patchyspace}
@@
 %%%%%%%
-\subsection*{Simulation methods}
+\subsection{Simulation methods}
 \label{ss:simulations}
@@
 %%%% %%%%
-\subsubsection*{The genealogy of migrant families}
+\subsubsection{The genealogy of migrant families}
 
@@
 %%%%%%% %%%%%%%%%%%
-\subsubsection*{The outflux of migrant alleles}
+\subsubsection{The outflux of migrant alleles}
 
@@
 %%%%%%%%%%%%
-\subsubsection*{Multiple patches}
+\subsubsection{Multiple patches}
 Above we worked with two patches separated by distance $R$;
@@
 %%%%%%%%%%
-\subsection*{Applications} 
+\subsection{Applications} 
 \label{ss:applications}
@@
 
-\subsection*[Probability of Establishment]{Numerical calculation of the probability of establishment}
+\subsection[Probability of Establishment]{Numerical calculation of the probability of establishment}
 \label{apx:establishment_sims}
@@
 %%%%%%%%%%%%
-\subsection*[Equilibrium Frequency]{The equilibrium frequency}
+\subsection[Equilibrium Frequency]{The equilibrium frequency}
 \label{apx:eqfreq}
@@
 %%%%%%%%%%%
-\subsection*[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
+\subsection[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
 \label{apx:asymptotics}
@@
 %%%%%%%%% %%%%%%%%%%%%%%
-\subsection*[Hitting and Occupation]{Hitting and occupation}
+\subsection[Hitting and Occupation]{Hitting and occupation}
 \label{apx:hitting_occupation}
@@
 %%%%%%% %%%%%%%%%%%%
-\subsection*[Outflux of Families]{The outflux of families}
+\subsection[Outflux of Families]{The outflux of families}
 \label{apx:outflux}
@@
 \appendix
+
+\setcounter{secnumdepth}{2}
 

### 2014-07-11 patchy-selection-paper.tex
plos reference style

@@
 \usepackage{color}
-\usepackage{natbib}
 \usepackage[hidelinks]{hyperref}
@@
 \setcounter{secnumdepth}{0}
-
-% \usepackage{cite}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
 
@@
 \newif\ifsubmissionversion
+% \submissionversiontrue
 \submissionversionfalse
-\newif\ifincludesupplement
-\includesupplementfalse
 
@@
     \usepackage{endfloat}
+    \usepackage{cite}
+    % Use the PLoS provided bibtex style
+    \bibliographystyle{plos2009}
+    \newcommand{\citep}[1]{\cite{#1}}
+    \newcommand{\citet}[1]{\cite{#1}}
+\else
+    \usepackage{natbib}
+    \bibliographystyle{abbrvnat}
 \fi
@@
 % \linespread{1.6}
-\doublespacing
+%\usepackage{setspace} 
+%\doublespacing
 \fi
-
-% Use the PLoS provided bibtex style
-% OR NOT \bibliographystyle{plos2009}
-\bibliographystyle{abbrvnat}
 
@@
 then there is no stable equilibrium with both $b$ and $B$ present,  % see p.603 in Nagylaki
-so that migrational swamping prevents $B$ from establishing \citep[see also][ for a review]{lenormand2002limits}.
+so that migrational swamping prevents $B$ from establishing (see also \citet{lenormand2002limits} for a review).
 \citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed  
@@
 then by Bayes' rule, $e^{-s_m t} = (1-k_e(t))\E[K_t]$.
-It turns out that if the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jagers1975branching},
+It turns out that if the distribution of offspring numbers is not too heavy-tailed (see \citet{jagers1975branching} for details),
 then $K_t$ has a limiting distribution: $K_t \dconv K$,
@@
 as depicted in Figure~\ref{fig:branching_decomp}.
-We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} \citep[see also][]{chauvin1991growing}.
+We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} (see also \citet{chauvin1991growing}).
 When we condition on survival until $t$, we condition on existence of
@@
   \begin{center}
-    \includegraphics{sim-transit.pdf}
+      \includegraphics{sim-transit}
   \end{center}
@@
 as long as the patch is reasonably well mixed by dispersal
-\citep[otherwise see][]{barton2013genetic}.
+(otherwise see \citet{barton2013genetic}).
 This association gets slowly whittled down by recombination during interbreeding at the edge of the patch,
@@
 we can use standard results on hitting times of $d$-dimensional Brownian motion
-that is killed at rate $s_m$ \citep[see][2.2.0.1 and 4.2.0.1]{borodin2002handbook}.
+that is killed at rate $s_m$ (see \citet{borodin2002handbook} 2.2.0.1 and 4.2.0.1).
 In particular, if the patch is circular with radius $w$ and lies at distance
@@
 if the assumption that fluctuations in $q(x)$ are small,
-this motion should converge to a diffusion process \citep[as assumed by e.g.][]{hallatschek2008surfing},
+this motion should converge to a diffusion process (as assumed by e.g.\ \citet{hallatschek2008surfing}),
 notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
@@
 we can fit formula \eqref{eqn:eqfreq} to these frequencies
-\citep[similar to][]{hoekstra2004ecological}.
+(similar to \citet{hoekstra2004ecological}).
 Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
@@
 aids establishment of new alleles and the spread of existing
-alleles \citep[but see][for the complications of varying population sizes]{Wilson-softsweep}.
+alleles (but see \citet{Wilson-softsweep} for the complications of varying population sizes).
 
@@
 with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
-In addition, over long time-scales, very rare chance events \citep[mice carried off by hurricanes and the like][]{censky1998overwater,nathan2008mechanisms}
+In addition, over long time-scales, very rare chance events (mice carried off by hurricanes and the like \citet{censky1998overwater,nathan2008mechanisms})
 could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
@@
 we need to choose $\var[R]=\sigma^2$ and $s(x)$ to be of the same, small, order; 
-this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale \citep[as noted by][]{slatkin1973geneflow}.
+this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale (as noted by \citet{slatkin1973geneflow}).
 More concretely, suppose that the grid size is $\epsilon \to 0$, 
@@
 where $C'$ and $r_0$ are chosen to match boundary conditions.
-Asymptotics of Bessel functions \citep[][8.451.6]{gradshteyn2007table} then imply that
+Asymptotics of Bessel functions (\citet{gradshteyn2007table}, 8.451.6) then imply that
 \begin{align}

### 2014-08-15 patchy-selection-paper.tex
fixed bug joachim found (1/sm -> sm in migrate)

@@
 \begin{align} 
-  h_A(x,y) \approx \frac{ g(A) }{ \E[K] } u(x,y),
+    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
 \end{align}
@@
 % \end{align}
-Combining with expressions for $g(A)$ in section~\nameref{apx:hitting_occupation},
+Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{apx:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
 we obtain that
@@
   \label{eqn:migrate} 
-  \migrate(y) &\approx q(y) \times \frac{ \rho s_p  }{ \xi^2 s_m }  \\
-  % \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
-  &\approx C \frac{ \rho s_p  }{ \xi^2 s_m } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
+  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
+  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
 \end{align}
@@
 An important point to notice is that effective migration and mutation rates will be on the same order if 
-$A \mu \approx (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
+$A \mu \approx 2 s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
 where $R$ is the distance between the patches,
@@
 In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
-is $\rho (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$,
+is $2 \rho s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
 which depends fairly strongly on the selective disadvantage $s_m$ between patches.
+% Note in particular that if $2 s_m < \mu A$ then migration is always slower than new mutation.
 Equivalently, the rates are roughly equal if
-$R/\sigma = \log(1/(s_m A \mu))/\sqrt{2s_m}$,
+$R/\sigma = \log(2 s_m/(A \mu))/\sqrt{2s_m}$,
 which gives the critical gap size past which adaptation will be mostly parallel
@@
 
-% ss <- 10^(-(1:6))
-% log(1/ss)
-% 1/sqrt(2*ss)
-% sm <- .05; (-log(c(1e-3,1))-log(sm))/sqrt(2*sm)
-% sm <- .001; (-log(c(1e-3,1))-log(sm))/sqrt(2*sm)
-% (7+log(1/ss))/sqrt(2*(ss))
-% (7+log(1/(.5*ss)))/sqrt(2*(.5*ss))
-
-If we take the mutational influx in a patch to be one per 1,000 generations ($\rho A \mu = .001$)
-and the population density to be 1 ($\rho =1$, so $-\log(A\mu)\approx 7$)
-then adaptation is mostly parallel for patches separated by gaps larger than $R/\sigma > (7-\log(s_m))/\sqrt{2s_m}$.
+% cdist <- function (sm,z) { (-log(z)+log(2*sm))/sqrt(2*sm) }
+% cdist(c(.05,.001),z=1e-4)
+% cdist(c(.05,.001),z=1e-2)
+% cdist(c(.05,.001),z=1)
+% ss <- exp(seq(log(1e-6),log(1e-2),length.out=100))
+% plot( ss, cdist(sm=.001,z=ss), type='l', log='x', xlab='mutation influx', ylab='R/sigma' )
+% lines( ss, cdist(sm=.05,z=ss), col='red')
+% abline(h=0,lty=2)
+
+If we take the mutational influx in a patch to be one per 10,000 generations ($\rho A \mu = .0001$)
+and the population density to be 1 ($\rho =1$, so $-\log(A\mu)\approx 9.2$)
+then adaptation is mostly parallel for patches separated by gaps larger than $R/\sigma > (9.2+\log(2s_m))/\sqrt{2s_m}$.
 If the selective pressure is fairly strong (say, $s_m=.05$),
-then for migration to be slower than mutation,
-the distance between patches must be at least 31.6 times the dispersal
-distance (i.e.\ $R/\sigma > 31.6$) for convergence to occur.
+then for convergence to be likely,
+the distance between patches must be at least 21.8 times the dispersal distance (i.e.\ $R/\sigma > 21.8$).
 If $s_m$ is much smaller, say $s_m = .001$, 
-then the patches must be separated by several hundred times the dispersal distance (i.e.\ $R/\sigma > 308$).
-The situation does not change much if the influx is quite a bit higher -- even if $A \mu = 1$,
-the required separation between patches is only reduced to $R/\sigma > 9.5$ with $s_m=.05$ and $R/\sigma>154$ for $s_m=.001$.
+the required distance increases to $R/\sigma > 67$.
 On the other hand, if the separation is much smaller than this transitional value, 
 then adaptations are likely to be shared between patches.
+
+If the influx is higher -- say, $\rho A \mu = .01$,
+the required separation between patches is only reduced to $R/\sigma > 7.3$ with $s_m=.05$.
+If $\rho A \mu = .01$ and $s_m=.001$, then the prediction is that mutation will \emph{always} be faster than migration
+-- however, with $s_m=.001$, for the assumption of well-separated patches to hold we would need $R/\sigma > 1/\sqrt{2s_m} = 22.4$;
+therefore, at spatial scales smaller than about 20 dispersal distances the model breaks down.
 
@@
   and mutation rate scaled by patch area ($\mu A$).
+  The right-hand plot shows that $\migrate$ is relatively insensitive to $s_m$.
   The black, dotted line shows values where $\Tmig=\Tmut$.
-  Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=50$, and $\xi^2=1$.
+  Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=20$, and $\xi^2=1$.
   Note that population density $\rho$ and mutation influx $A\mu$ are shown on a log scale.
@@
 \begin{equation} \label{eqn:parallel_prob}
-  \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu s_m }{2 A \mu s_m + \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
+    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
 \end{equation}
@@
 \begin{equation} \label{eqn:total_migrate}
-  \sum_{i=1}^k \migrate(R_i) = C \frac{ \rho s_p }{\xi^2 s_m} \sum_{i=1}^{k} \left(R_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m} /\sigma\right).
+  \sum_{i=1}^k \migrate(R_i) = C \frac{ 4 \rho s_p s_m }{\xi^2} \sum_{i=1}^{k} \left(R_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m} /\sigma\right).
 \end{equation}
@@
 and so the final partition of types across demes has the Ewens
-sampling distribution with parameter $A \mu s_m/ \exp (-\sqrt{2 s_m}R/\sigma)$.
+sampling distribution with parameter $A \mu / (4 s_m \exp (-\sqrt{2 s_m}R/\sigma))$.
 Now we can compute most properties we might want about the process.
@@
 -- the analogue of the mutational influx $\mu \rho A$ --
-is $\rho (1/s_m) \exp(- R \sqrt{2 s_m} / \sigma)$.
+is $2 \rho s_m \exp(- R \sqrt{2 s_m} / \sigma)$.
 Equivalently, the critical gap size between patches past which adaptation is likely independent
-is $R = (\sigma/\sqrt{2s_m}) \log(1/(s_m A \mu))$.
+is $R = (\sigma/\sqrt{2s_m}) \log(2s_m/(A \mu))$.
 Finally, successfully transiting migrant lineages move at rate $\sigma \sqrt{2s_m}$,
@@
           &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
-          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) .
+          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) . \label{eqn:q_migrate_relation}
 \end{align}

### 2014-08-15 patchy-selection-plots.R

@@ -10,18 +10,18 @@ require(xtable)
 #  first against characteristic dist and pop density
 #  then agains mutation rate and pop density
 muttime <- function (env) {with(env,1/(A*mu*2*sb*rho))}
-migtime <- function (env) {with(env,1/(const*(sb/sd)*rho*(1/sqrt(cl*sqrt(sd)))*exp(-cl*sqrt(sd))))}
+migtime <- function (env) {with(env,1/(const*(sb*4*sd)*rho*(1/sqrt(cl*sqrt(sd)))*exp(-cl*sqrt(sd))))}
 rhovals <- logseq(1,1e5,length.out=100) # per km^2
-clvals <- seq(1,200,length.out=100)
+clvals <- seq(1,120,length.out=100)
 muvals <- logseq(1e-9,1e-4,length.out=100) 
-sdvals <- logseq(1e-4,1e-1,length.out=100)
+sdvals <- logseq(1e-6,1e-2,length.out=100)
 A <- 100 # km^2
 rho <- 10  # per km^2
 mu <- 1e-8 
 sb <- .1
 sd <- .01
 const <- .5
-cl <- 50  # R/sigma
+cl <- 20  # R/sigma
 clplot <- expand.grid(cl=clvals,rho=rhovals)
 clplot$mut <- muttime(clplot)
 clplot$mig <- migtime(clplot)
@@ -37,7 +37,7 @@ pminmax <- function (x,lower,upper) { pmin(upper,pmax(lower,x)) }
 pdf(file="phase-diagram-log.pdf",width=6,height=3,pointsize=10)
 layout(t(1:2))
 par(mar=c(4,3.5,1,1)+.1)
-levelsets <- 10^c(0,2,4,6)  # which lines to draw
+levelsets <- 10^rev(c(0,1,2,3,4,5,6))  # which lines to draw
 labelsets <- levelsets
 # plot(0,type='n',xlab='',ylab='',xlim=range(clvals),ylim=range(rhovals),log='y')
 # contour( clvals, rhovals, f(clplot$mut), col='red', levels=levelsets, labels=labelsets, method="edge", lwd=2, add=TRUE )
@@ -52,7 +52,7 @@ for (lev in levelsets) { # note rho is log10-scale.
     if (rholev>min(rhovals) & rholev<max(rhovals)) {
[…]

### 2014-08-15 patchy-selection-paper.tex
minor changes

@@
 
-If the influx is higher -- say, $\rho A \mu = .01$,
+If the mutational influx is higher -- say, $\rho A \mu = .01$ --
 the required separation between patches is only reduced to $R/\sigma > 7.3$ with $s_m=.05$.
-If $\rho A \mu = .01$ and $s_m=.001$, then the prediction is that mutation will \emph{always} be faster than migration
--- however, with $s_m=.001$, for the assumption of well-separated patches to hold we would need $R/\sigma > 1/\sqrt{2s_m} = 22.4$;
-therefore, at spatial scales smaller than about 20 dispersal distances the model breaks down.
+If $s_m=.001$, then with this higher mutational influx this calculation predicts that mutation will \emph{always} be faster than migration
+-- however, this is misleading, 
+since with $s_m=.001$, for the assumption of well-separated patches to hold we would need $R/\sigma > 1/\sqrt{2s_m} = 22.4$.
+The correct conclusion is that convergent adaptation is much more likely than shared adaptation
+between patches separated by at least 22.4$\sigma$;
+for shorter separations the model breaks down.
 
@@
   and mutation rate scaled by patch area ($\mu A$).
-  The right-hand plot shows that $\migrate$ is relatively insensitive to $s_m$.
+  The right-hand plot shows that $\migrate$ is relatively insensitive to $s_m$ (thanks to the square root).
   The black, dotted line shows values where $\Tmig=\Tmut$.

### 2014-08-18 patchy-selection-paper.tex
thanked joachim

@@
 \subsection*{Acknowledgements}
-We would like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
+We would like to heartily thank Joachim Hermisson for a careful reading of the paper,
+and catching the algebra error present in the first version of this paper on the bioRxiv,
+and also to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
 as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab

### 2014-11-22 analyze-sims.R
figures comparing sims and predicted

@@ -2,6 +2,8 @@
 
 source("sim-patchy-selection-fns.R")
 
+maxtime <- 25000  # number of gens these were run for
+
 mutsims <- read.table("mutation-sims-info.tsv",sep='\t',header=TRUE)
 
 names(mutsims)[names(mutsims)=="N"] <- "N.name"
@@ -14,6 +16,9 @@ names(mutsims)[names(mutsims)=="mu"] <- "mu.name"
 mutsims$mu.name <- factor( mutsims$mu.name, levels=levels(mutsims$mu.name)[order(as.numeric(gsub("mu","",levels(mutsims$mu.name))))])
 names(mutsims)[names(mutsims)=="mu.1"] <- "mu"
 mutsims$gb <- getgrowth(mutsims)$gb
+
+# create factor for parameter combination
+mutsims$paramstring <- with(mutsims, factor( paste(mu.name,sm.name,N.name,dims,sep='_') ) )
 
 mutsims$adjust <- 0  # with( mutsims, log(N)/gb + size1/2 / ( sigma * sqrt(gb) ) )
 mutsims$muttime <- with( mutsims, 1/(2 * mu * gb * size1 * N) + adjust )
@@ -132,3 +137,61 @@ pairs( migsims[c("N", "gb", "gm", "R", "sigma", "migtime")],
 pairs( migsims[c("final1", "time1", "hit100.1", "final2", "time2", "hit100.2","migtime")],
     upper.panel=function (x,y,...) { points(x,y,...,col=migsims$sm.name) }, 
     lower.panel=function (x,y,...) { points(x,y,...,col=migsims$N.name) } )
+
+####
+# for paper
+
+require(colorspace)
+
+pdf( file="times-predicted-observed.pdf", width=6.5, height=3.5, pointsize=10 )
+layout(t(1:2))
+par(mar=c(4,3,1,1)+.1)
+# MUTATION
+    # remove simulation configurations where a substantial portion haven't yet adapted
+    usethese <- with( mutsims, adapted & ( muttime < maxtime/10 ) ) 
+    sm.vals <- (levels(droplevels(mutsims[usethese,]$sm.name)))
+    sm.cols <- sequential_hcl(length(sm.vals))[ match( levels(mutsims$sm.name)[tapply(mutsims$sm.name,mutsims$paramstring,"[",1)], sm.vals ) ]
+    N.vals <- (levels(droplevels(mutsims[usethese,]$N.name)))
[…]

### 2014-11-22 patchy-selection-paper.tex
new fig

@@
 For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{apx:hitting_occupation}.
+
+%%%%%%%%%
+\subsection{Comparison to simulation}
+\label{ss:comparison_to_sims}
+
+We used the simulations describe above to test the theory,
+etcetera.
+
+\begin{figure}[ht!]
+  \begin{center}
+      \includegraphics{times-predicted-observed}
+  \end{center}
+  \caption{
+      \textbf{(A)} Time to adaptation by new mutation 
+      in 3,128 simulations,
+      compared to the value predicted by equation X.
+      For each of 52 different parameter combinations (Supp Table XX),
+      we performed between 40 and 100 independent simulations;
+      shown are the median (points) and interquartile range (lines)
+      of the time until the patch had at least 100 adapted individuals,
+      for each unique parameter combination (example runs in Supp Figure XX).
+      \textbf{(A)} Similar, for time to adaptation by migration
+      in 4,158 simulations.
+      TODO: put equations on axis labels, describe more, make supp figs.
+  }   \label{fig:sim_times}
+\end{figure}
 

### 2014-11-24 analyze-sims.R
figs

@@ -91,6 +91,8 @@ migsims$R <- as.numeric(gsub("R-","",migsims$R.name))
 migsims$gb <- getgrowth(migsims)$gb
 migsims$gm <- getgrowth(migsims)$gm
 
+migsims$paramstring <- with(migsims, factor( paste(sm.name,R.name,N.name,dims,sep='_') ) )
+
 migsims$adjust <- 0  # with( migsims, log(N)/gb + size1/2 / ( sigma * sqrt(gb) ) )
 migsims$migtime <- with( migsims, 0.5/( 4 * N * gb * abs(sm) * exp(- R * sqrt(2*abs(sm))/sigma ) ) + adjust )
 migsims$adapted <- ( 
@@ -102,7 +104,8 @@ pdf(file="migration-time-predicted.pdf",width=9, height=5, pointsize=10)
 layout(t(1:2))
 with( subset(migsims,adapted), {
         plot( migtime, hit100.2, col=sm.name, pch=as.numeric(N.name), log='xy', xlim=c(10,1e8), ylab="time to hit 100 in second patch", xlab="mean migration time") 
-        points( sort(unique(migtime)), tapply( hit100.2, factor(migtime,levels=sort(unique(migtime))), mean, na.rm=TRUE ), pch=20, cex=2 )
+        # points( sort(unique(migtime)), tapply( hit100.2, factor(migtime,levels=sort(unique(migtime))), mean, na.rm=TRUE ), pch=20, col=adjustcolor("red",.8), cex=2 )
+        points( sort(unique(migtime)), tapply( hit100.2, factor(migtime,levels=sort(unique(migtime))), median, na.rm=TRUE ), pch=20, cex=2 )
         abline(0,1,untf=TRUE)
         legend("bottomright",legend=c(levels(sm.name),levels(N.name)), col=c(1:nlevels(sm.name),rep("black",nlevels(N.name))), pch=c(rep(1,nlevels(sm.name)),1:nlevels(N.name)) )
     } )
@@ -143,21 +146,21 @@ pairs( migsims[c("final1", "time1", "hit100.1", "final2", "time2", "hit100.2","m
 
 require(colorspace)
 
-pdf( file="times-predicted-observed.pdf", width=6.5, height=3.5, pointsize=10 )
+pdf( file="times-predicted-observed.pdf", width=6.5, height=3.3, pointsize=10 )
 layout(t(1:2))
-par(mar=c(4,3,1,1)+.1)
+par(mar=c(4,4,1,1)+.1)
 # MUTATION
[…]

### 2014-11-25 patchy-review-responses.tex
review responses begun

% \documentclass{article}
% \usepackage{fullpage}
% \input{review-response-commands}
% 
% \begin{document}

\section*{From the editor}
% From: "Catriona J. MacCallum" <cmaccallum@plos.org>
% To: Peter Ralph <pralph@usc.edu>
% Date: Fri, 15 Feb 2013 04:01:05 -0500
% Subject: PLOS Biology Decision [PBIOLOGY-D-12-02842R1]
% Dear Dr. Ralph,

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
  {\it PLoS Biology}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
  Department of Evolution and Ecology\\
  University of California at Davis\\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

\begin{quote}
As can you see from the reviews, two general points were raised by all
reviewers and I feel that thoroughly addressing both of these points poses a
prerequisite for further consideration of the manuscript:

(i) There needs to be some sort of numerical verification (e.g. with
individual-based simulations) to test the main conclusions, check their
robustness, and identify the parameter ranges for which they hold.

(ii) The presentation of the main result needs to be improved. In particular,
all reviewers feel that the main result can be derived in a simpler and more
intuitive way. I agree with them that at present the manuscript is overly
technical and would benefit strongly from a more intuitive derivation of the
results and discussion of the assumptions under which they hold. The reviewers
provide detailed suggestions on how this can be achieved.
\end{quote}

%%%%%%%%%%%%%%
\reviewersection

%%%%%%
\point{}{
    Am I right that the basic argument is the
    following:

    A migrant family in unfavorable territory will be typically of size $n \sim 1/s$
    ($s=s_m$), stick around for a time $\tau \sim 1/s$ and be spread out over $\ell \sim \sigma/\sqrt(s)$. 
    Given an equilibrium density $\rho*q(x)$, we need an arrival rate
    of migrant families of
    \[ k ~ \rho q(x) l/n/tau ~ \rho q(x) l s^2 \]
}

\reply
The factor of $\ell$ shouldn't be there:
suppose that $X$ is a Poisson process of ``centers'' of rate 1, 
and that around each center of $X$ we distribute $n$ points uniformly within a circle of radius $\ell$;
the expected number of points in a set of area 1 is $n$, regardless of $\ell$.
The right thing to do with the factor of $\tau$ is harder to intuit:
it is clear that $\rho q(x)$ should be $E[K]$ multiplied by the number of families present;
and it might be argued that the number of families present should be the influx of families
multiplied by the time they stick around for,
but given that the typical size of a family decreases once it arrives,
the correct way forward becomes non-intuitive, to us, at least.

%%%%%%%%
\point{}{ 
    The probability of establishment of at least one member of family is then
    \[ P_{est} = 1 - exp(-p n) ~ 1 - exp(-p_e/s) \]
    where $p_e$ is the establishment probability of a single individual. The result
    for $\migrate$ will then depend on whether $p_e/s$ is large or small.
    For $p_e/s$ large, $\migrate = \rho q(x) s^2 \ell$,
    otherwise,
    $\migrate = \rho q(x) s \ell p_e$.

    The latter is the case discussed in the manuscript. The other is important too
    and should be discussed. 
}

%%%%%%%%
\point{}{ 
    I don't seem to be able to identify the equivalent of
    l in the author's formulae, but I believe it should be there. After all, rho
    q(x) is a density and needs to be multiplied by a length (or area in d=2) to
    give a meaningful rate.
}

\reply
    See above.

%%%%%%%%
\point{}{ 
    To assess the accuracy of the various approximations made, the process should
    be simulated and the measured rates explicitly compared to analytical
    predictions.
}

\reply

%%%%%%%%
\point{long range migration}{All results depend exponentially on the distance, which
      is a consequence of the model only allowing for short range dispersal. Rare
    long range dispersal will alter all these rates dramatically, see for example
    recent work by Hallatschek and Fisher.
}

\point{Fig 4:}{ Why not use natural units for the axis of these plots,
  $R\sqrt(s)/\sigma$, $\rho \mu A/s$.
  }

%%%%%%%%
\point{in the discussion on page 10 "probability of parallel adaptation.."}{
    it should be noted that many of the scenarios at high mutation influx are quite
    hypothetical. While the colonization of migrants might dominate over local
    adaptation, this argument assumes that the migrants are at equilibrium
    frequency. The transient to reach this equilibrium might be much longer than
    the time required for de novo local adaptation.
}

%%%%%%%%%%%%%%
\reviewersection

\point{1.}{
    I think that the main result (12) can be derived much more simply, and
    also extended. By writing that the probability that a family that arrives
    at the patch is successful is approximately $ 2s_p \E [K] /\xi^2$, the authors have
    implicitly assumed that this quantity is $<\ll 1$. But this implies that the
    fluctuations in the number of B individuals arriving at the patch can be
    ignored, since even a large fluctuation does not saturate the probability
    of success. This is valid for $s_m \gg s_p$. Given that the fluctuations can be
    ignored, (12) can be derived immediately from $q(y)$ -- simply apply Fick’s
    Law to $q(y)\rho$ to get that the flux of new B individuals is $\sigma^2 |q'(y)|\rho/2$, and
    then multiply by the success probability $2s_p/\xi^2$. (Actually, the result is
    slightly different from (12) -- see X below.)
}

\reply
This argument falls in the trap of treating each migrant individual as independent,
ignoring the correlations of arrival times induced by families;
see the discussion in which we now set this up as a straw man in XXX.

\point{2.}{
    However, I think we should also care about the opposite limit, $s_m \ll s p$. In
    this case, the fluctuations do matter -- arrivals typically occur in clumps
    (families) large enough so that success is almost certain conditioned on
    arrival. Since on average each family consists of $\E[K]$ individuals, the
    family flux is just the individual flux divided by $\E[K]$. $\E[K] = \xi^2 /(2s_m)$,
    so the result is the same as above, except with $s_p$ replaced by an additional
    factor of $s_m$. (This could be made more formal, but I don’t think that’s
    what the authors are going for.) In the intermediate case $s_m \approx s_p$ which
    might be the most relevant for the mice, the two limiting regimes cross
    over smoothly. So for the first limit ($s_m \gg s_p$), all you need is $q$, and
    for the second, you only need to add basic facts about branching random
    walks (ie, limiting number distribution and localization at long times).
}

\reply
In this argument, the reviewer is taking into account clumpiness for establishment of migrants,
but still not for the time interval
between arrivals of families.

%%%%%%%%%%%%%%
\reviewersection

\point{1a.}{
        What I miss is a clearer discussion of the scope and limits of the
    main results. Take Eq. 13: in its revised form, this is a non-monotonic
    function of the selection coefficient $s_m$. I am almost sure that this is
    an artefact and that the parameter range of the theory needs to be
    restricted to the part where the transition probability declines with $s_m$.
    But it is not easy to see how this needs to be done in a prober way. We
    certainly need a minimal distance among patches. On top of that, we may
    need a lower bound for $s_m$. Both these restrictions (or any other) should
    not be just ad-hoc, but motivated from within the theory (see my remarks at eq. 13).
}

\point{1b.}{
    Linked to this: Figure 4, right panel, shows results in a parameter
    range where (I think) the theory does not apply.
}

\point{1c.}{
    Also linked to this: I believe your theory is sound, but it is really
    tricky to clearly define its range of application and also to see how good
    it is quantitatively. Bottom line: we need individual-based simulations to
    test at least some of the main conclusions. E.g.: do simulations for 2
    patches in 1d to test the predicted demarcation line between the
    "mutation" and the "migration" regime. Do this for several parameter
    choices to test the limits of the theory.
}

\point{2.}{
    Section "outflux of migrant alleles": I found this section much less
    clear than others and have made suggestions for a clearer presentation
    (see ms).
}

\point{3.}{
    Section "length of shared haplotype": This is very handwavy and
    (without simulation backup) I find it hard to believe that these results
    are accurate. Since it is also not central to the paper, the section
    should maybe just be dropped.
}

%%%%%%
\point{[beginning comments in the pdf]}{
``Do you assume a Gaussian kernel? Or is this irrelevant? Which properties of the kernel are required? Only existence of a finite $sigma^2$?''
}

\reply
We've tried to make it more clear here 
\nameref{rr:longtailed}
exactly what we sweep under the rug later.

%%%%%%
\point{}{``This is not sufficient: the patch also needs to be large enough that p_e is (almost) reached at its center. Generally, the approximation should over-estimate the true value.''}

\reply
Good point; we have adjusted this slightly. \nameref{rr:p_estab}

%%%%%%
\point{[more complete information about simulation model]}{}

\reply
We've added more description of how we did the simulations
\nameref{rr:sim_details},
and don't think the choice of a bounded kernel is particularly salient
(given that it's not long-tailed, of course).

%%%%%%
\point{[comments near equation \eqref{eqn:eqfreq}]}{}

\reply
Both comments are excellent. 
We have added discussion of the intuition and the restrictions on $|x|$, as suggested
\nameref{rr:q_form}.

%%%%%%
\point{The outflux of migrant alleles}{
    ``This section is less clear than the previous one. You try to be intuitive, but also give most of the formulas. It seems to me that there should be either more or (rather:) fewer equations. On the formal level, I was only able to follow after reading the "Methods" section. And on the intuitive level, I was hampered by trying to understand the equations from the information in the main text (and failing to do so).

I think you could devote the formal derivation to the "Methods" and argue on an intuitive level in the main text:

First, you should clearly spell out what the approximation is. You use the trunc + outgrowth picture of the branching process. Then you assume that the mutant family will only be able to cross to the other patch if the trunc hits this new patch. And when the trunc hits, you assume that all the outgrowth branches also enter it. I.e.: Eventually, either the whole family is "in" and then stays in (as in Fig 3) or no part ever enters. This approximation should be fine if the size of the new patch is sufficiently large.

As a consequence of this, we can just follow the trunc, which moves by simple Brownian motion with an exponential life span. This defines h(x,y) as probability that the trunc arrives at the new patch.

Now, how can we relate h(x,y) to u(x,y)? - As follows: Once the trunc reaches the new patch (and assuming s_m selection everywhere) it will live there for an exponential time with mean 1/2s_m. Since it "lives or dies" with its entire outgrowth family (of expected size E[K]) it thus contributes to the density like u(x,y) = h(x,y) E[K] / (2s_m). This motivates Eq. 11. [no need to introduce g(A) in the main text]
}

% \end{document}

### 2014-11-25 patchy-selection-paper.tex
figs
review responses begun

@@
 \setcounter{secnumdepth}{0}
+
+%%%%%
+% Responses to reviews:
+\input{review-response-commands}
+% set this to show line numbers and include responses to reviews or not
+\newif\ifreviewresponses
+% \reviewresponsestrue  % include them
+\ifreviewresponses\else\renewcommand{\linelabel}[1]{}\fi
+\reviewresponsesfalse  % don't include them
+\iflatexml\reviewresponsesfalse\fi  % they don't work so well in html
+% \newcommand{\responsefile}{pbio-reviews-19sept12-responses.tex}  % name of the review reponses file
+\newcommand{\responsefile}{patchy-review-responses.tex}  % name of the review reponses file
 
@@
 (i.e.\ $\sigma$ is the dispersal distance). 
+We will deal with migration by immediately appealing to the central limit theorem,
+treating the total distance traveled across $t$  dispersal events as Gaussian with variance $t \sigma^2$,
+and do not discuss the (interesting) cases where rare, long-distance dispersal events
+are more important \citep[see][for discussion]{longtailed,ralphcoop2010,hallatschek??}.  \linelabel{rr:longtailed}
 %We also aim to describe the scenario where it is sensible to think of large, discrete patches;
@@
 which will be a good approximation if the patch is large relative the scale over which $p(x)$ goes from 0 to $p_e$,
-or if $p(x)$ is symmetric about the edge of the patch.
+or if $p(x)$ is close to $p_e$ at some point and is symmetric about the edge of the patch. \lineref{rr:p_estab}
 We examine this more generally via exact calculation of $p(x)$ in the section \nameref{apx:establishment_sims}.
@@
 
-First we briefly describe the simulations used below for illustration and validation.
-We simulated forward-time dynamics of the number of alleles of each type in a regular grid of demes with fixed size $N$.
+First we briefly describe the simulations used below for illustration and validation
+(and provide the \texttt{R} code used as a supplement).
+We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
 Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
 locally beneficial alleles were more likely to reproduce.
-Each then produced a random, Poisson distributed number of offspring that each independently migrated to demes no more than six steps away
-according to a symmetric migration kernel.
+Each then produced a random, Poisson distributed number of offspring that each, independently, 
+either remained in the same location or
+migrated a random number of steps in a uniformly chosen cardinal direction;
+the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
+\linelabel{rr:sim_details}
 Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
@@
 The reproduction rate in simulations for type $b$ alleles was 30\%; 
-this was then multiplied by $1+s$ to get the probability of reproduction for type $B$.
-These $s$ are the values reported in the figures, and do not depend on the basic rate of reproduction.
+this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
+where the value of $s$ is either $s_m$ or $s_p$ depending on the individual's location.
+This determines the values of $s_m$ and $s_p$ reported in the figures, 
+and do not depend on the basic rate of reproduction.
 However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
@@
 and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients.
+In applications we fit $C$ to the data;
+to get concrete numbers from theory we take $C=1$ if necessary.
+
+From simulations, $C$ seems to be not too far from 1 in practice (see below).
+As J.\ Hermisson pointed out in an earlier draft,
+this functional form has a simple intuition:
+local migration leads to the exponential decay,
+since if each migrant at distance $x$ 
+produces an average of $c$ descendants that make it to distance $x+1$, 
+then in one dimension, the number of migrants must decay as $\exp(-cx)$;
+and the square-root term in two dimensions 
+enters because the available area grows with $x$.  
+The ``renewal'' aspect of this same argument suggests that for equation~\eqref{eqn:eqfreq} to hold,
+$|x|$ must be larger than only 5--10 multiples of $\sigma$.
+\linelabel{rr:q_form}
+
 These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
 To be clear about the assumptions (especially regarding scaling) implicit here,
-we provide a derivation in section~\nameref{apx:eqfreq}, 
-as well as justification for the asymptotics in section~\nameref{apx:asymptotics}.
+we provide a derivation below in \nameref{apx:eqfreq}, 
+as well as justification for the asymptotics below in \nameref{apx:asymptotics}.
 In one dimension, the equation can be solved to give $q(x)$ 
 in terms of Jacobi elliptic functions; see appendix \ref{apx:elliptic_integrals}.
-In applications we fit $C$ to the data;
-to get concrete numbers from theory we take $C=1$ if necessary.
-From simulations, $C$ seems to be not too far from 1 in practice (see below).
 
@@
   with $C$ chosen arbitrarily to give a reasonable fit.
+  \plr{what is $C$?}
   (The disagreement at long distances is due to boundary effects near the edge of the range.)
@@
 \end{align}
-The quantities on the right-hand side depend implicitly on $r_0$.
+The quantities on the right-hand side depend implicitly on $r_0$,
+but their product does not.
 All terms except the ``outflux of families'' are calculations with spatial branching processes;
-and we can solve for this outflux using the established form for $q(x)$ in \eqref{eqn:gestalt_q}.
+and we can solve for this outflux using the established form for $q(x)$.
 % Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
 % -- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.
+
+%%%% %%%%
+\subsubsection{Heuristics}
+
+At first it is tempting to argue
+that since the probability of establishment of a beneficial allele is $p_e$,
+and in a patch of area $A$ there are $\rho A q(x)$ beneficial alleles present at equilibrium,
+the mean rate of establishment is $p_e \rho A q(x)$.
+This is wrong, since it ignores the family structure of migrants.
+
+However, this wrong start suggests a correct one:
+notice that the mean density $q(x)$ counts up each family's total size,
+while $\migrate(x)$ counts the probability the family establishes.
+We argue below that
 
@@
 then $K_t$ has a limiting distribution: $K_t \dconv K$,
-and $e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
+and 
+\begin{align} \label{eqn:EK_limit}
+    e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty .
+\end{align}
 In other words, the chance of survival to time $t$ is a constant multiplied by $e^{-s_m t}$,
@@
 if each member arriving has chance $p_e$ of establishing, independently.
+\plr{And if $p_e$ is small?}
 % If we let $w$ be the width of the patch, then the remaining term is
@@
   \begin{center}
+      % made in analyze-sims.R
       \includegraphics{times-predicted-observed}
@@
   }   \label{fig:sim_times}
+\end{figure}
+
+\begin{figure}[ht!]
+  \begin{center}
+      % made in analyze-sims.R
+      \includegraphics{prob-mutation-compared}
+  \end{center}
+  \caption{
+      Probability that two patches separated by distance $R$
+      evolve by independent mutations (i.e.\ in parallel),
+      as predicted by theory (colors, contours)
+      and in simulations (points).
+      \textbf{With $C=5$.}
+      XXX: more info
+  }   \label{fig:sim_probs}
 \end{figure}

### 2014-11-25 patchy-review-responses.tex
reviews...

@@
 
-%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%  Reviewer 2
 \reviewersection
@@
 
-%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%  Reviewer 3
 \reviewersection
@@
 We've tried to make it more clear here 
-\nameref{rr:longtailed}
+\llname{rr:longtailed}
 exactly what we sweep under the rug later.
@@
 \reply
-Good point; we have adjusted this slightly. \nameref{rr:p_estab}
+Good point; we have adjusted this slightly. \llname{rr:p_estab}
 
@@
 We've added more description of how we did the simulations
-\nameref{rr:sim_details},
+\llname{rr:sim_details},
 and don't think the choice of a bounded kernel is particularly salient
@@
 We have added discussion of the intuition and the restrictions on $|x|$, as suggested
-\nameref{rr:q_form}.
+\llname{rr:q_form}.
 
@@
 \point{The outflux of migrant alleles}{
-    ``This section is less clear than the previous one.
-You try to be intuitive, but also give most of the formulas.
-It seems to me that there should be either more or (rather:) fewer equations.
-On the formal level, I was only able to follow after reading the "Methods" section.
-And on the intuitive level, I was hampered by trying to understand the equations from the information in the main text (and failing to do so).
-
+    \begin{quote}
+        This section is less clear than the previous one. [...]
     I think you could devote the formal derivation to the "Methods" and argue on an intuitive level in the main text:
@@
     First, you should clearly spell out what the approximation is.
-You use the trunc + outgrowth picture of the branching process.
-Then you assume that the mutant family will only be able to cross to the other patch if the trunc hits this new patch.
-And when the trunc hits, you assume that all the outgrowth branches also enter it.
+You use the trunk + outgrowth picture of the branching process.
+Then you assume that the mutant family will only be able to cross to the other patch if the trunk hits this new patch.
+And when the trunk hits, you assume that all the outgrowth branches also enter it.
 I.e.: Eventually, either the whole family is "in" and then stays in (as in Fig 3) or no part ever enters.
@@
 
-As a consequence of this, we can just follow the trunc, which moves by simple Brownian motion with an exponential life span. This defines h(x,y) as probability that the trunc arrives at the new patch.
+    As a consequence of this, we can just follow the trunk, which moves by simple Brownian motion with an exponential life span.
+This defines h(x,y) as probability that the trunk arrives at the new patch.
 
-Now, how can we relate h(x,y) to u(x,y)?
-- As follows: Once the trunc reaches the new patch (and assuming s_m selection everywhere) it will live there for an exponential time with mean 1/2s_m.
-Since it "lives or dies" with its entire outgrowth family (of expected size E[K]) it thus contributes to the density like u(x,y) = h(x,y) E[K] / (2s_m).
+    Now, how can we relate $h(x,y)$ to $u(x,y)$?
+- As follows: Once the trunk reaches the new patch (and assuming $s_m$ selection everywhere) it will live there for an exponential time with mean $1/2s_m$.
+Since it "lives or dies" with its entire outgrowth family (of expected size $\E[K]$) it thus contributes to the density like $u(x,y) = h(x,y) \E[K] / (2s_m)$.
 This motivates Eq.
 11.
-[no need to introduce g(A) in the main text]
+(no need to introduce $g(A)$ in the main text)
+    \end{quote}
 }
 
+\reply
+   Good suggestion; as discussed above,
+   we've taken what we think is a similar approach, 
+   providing a more intuitive arguemnt (in \nameref{ss:heuristics})
+   and merged the section referred to here with the Methods \llname{rr:moved_argument}.
+
+%%%%%%%%
+\point{[awkward sentence]}{}
+
+\reply
+Fixed? \llname{rr:awkward}
+
+%%%%%%%%
+\point{[expected total number]}{if y is a point, this number would be 0 (its a density)}
+
+\reply
+We have added the word ``informally''. \llname{rr:density}
+
+%%%%%%%%
+\point{}{ 
+    \begin{quote}
+    This function is non-monotonic in $s_m$ and I think this needs to be discussed: can it really be that $\migrate$ increases with $s_m$ or is this an artifact?
+
+    In general, we need a clearer idea about the limits of this theory. There are several things to consider:
+    \begin{itemize}
+        \item equation (3) for the expected mutant density is only an asymptotic. Therefore, we probably need to require $|y| \gg \sigma / \sqrt(2s_m)$.
+
+        \item the approximation $g(A) \sim 1/2s_m$ in the appendix requires that the patch size is much larger than $1/s_m$
+
+        \item in your arguments, you use the outflow of families from a ball with (large) radius $r_0$ [although I do not see how this figures in]
+
+        \item over short ranges, results must depend on details of the dispersal kernel anyway.
+[Depending on the shape of the distribution - heavy tail - this even holds for large distances].
+As mentioned above: spell out assumptions on the kernel.
+    \end{itemize}
+
+    So: what is the minimal distance $|y|$ for the result to be valid?
+And: does this minimal distance eliminate the anomaly of an increasing $\migrate$ with $s_m$ for small $s_m?$ Alternatively: is there also a lower limit for $s_m$?
+I have noted that you do assume above that $s_m$ is large enough to avoid local fixation - but I'm not sure that this comes down to the same thing.
+\end{quote}
+}
+
+\reply
+
+%%%%%%%%
+\point{[expressio nfor the outflux of migrants]}{do we need this for anything?}
+
+\reply
+We don't need it for anything else, 
+but we think it adds sufficiently to the intuition about the process
+that we'd like to keep it in.
+In practice it could provide a good reality check.
+\plr{alternatively, compute this for the mice example?}
+
+%%%%%% p.12
+\point{[justify $\E[K]\approx 1/s_m$]}{}
+
+\reply
+See \eqref{eqn:mean_K}, and modified discussion around \llname{rr:EK}.
+
 % \end{document}

### 2014-11-25 patchy-selection-paper.tex
reviews...

@@
 %%%% %%%%
-\subsubsection{Heuristics}
-
-At first it is tempting to argue
-that since the probability of establishment of a beneficial allele is $p_e$,
-and in a patch of area $A$ there are $\rho A q(x)$ beneficial alleles present at equilibrium,
-the mean rate of establishment is $p_e \rho A q(x)$.
-This is wrong, since it ignores the family structure of migrants.
-
-However, this wrong start suggests a correct one:
-notice that the mean density $q(x)$ counts up each family's total size,
-while $\migrate(x)$ counts the probability the family establishes.
-We argue below that
-
-%%%% %%%%
 \subsubsection{The genealogy of migrant families}
+\label{ss:migrant_math}
 
@@
 then by Bayes' rule, $e^{-s_m t} = (1-k_e(t))\E[K_t]$.
-It turns out that if the distribution of offspring numbers is not too heavy-tailed (see \citet{jagers1975branching} for details),
+It turns out that if the distribution of offspring numbers is not too heavy-tailed 
+(see \citet{jagers1975branching} for details),
 then $K_t$ has a limiting distribution: $K_t \dconv K$,
@@
 then this gives the following decomposition:
-\begin{align}
+\begin{align} \label{eqn:Zt_decomposition}
   \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
@@
 that arose from the main branch in generation $s$,
-and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.
+and each $W_s$ is an independent copy of the offspring distribution $W$, conditioned on $W_s \ge 1$.
+Since each subfamily $Z^{(s,k)}$ is a branching process with mean offspring number less than one,
+most of these die out quickly,
+and so $Z_t$ is composed of a cluster of recently split subfamilies.
+If we take the expectation of \eqref{eqn:Zt_decomposition},
+we get the limiting mean family size conditioned on nonextinction:
+\begin{align} \label{eqn:mean_K}
+    \E[K] = 1 + \E[W \given W \ge 1] /(e^{s_m}-1) \approx \E[W \given W \ge 1] / s_m .
+\end{align}
 
@@
 
-This suggests that we add in the spatial motion by following only the motion of the ``trunk''
-(i.e.\ the red line in figure \ref{fig:branching_decomp}).
-Its motion is approximately Brownian, with variance $\sigma^2$ per generation.
-(Recall $\sigma$ is the dispersal distance.)
-We can use these two ingredients, Brownian motion and a branching process, 
-to compute the terms of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
-
-%%%%%%% %%%%%%%%%%%
-\subsubsection{The outflux of migrant alleles}
-
-We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
-as the sum of independent migrant families, each of whose dynamics are
-given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
-Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
-Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
-such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ that initiate new migrant families
-is $\gamma(x)$.
-To make expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} more concrete,
-we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
-of the outflux $\gamma$ at that point.
-By examining the constituent terms,
-it will turn out that equation \eqref{eqn:gestalt_migrate} 
-is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.
-
-First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
-Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
-and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
-The \emph{mean occupation density} of $Z$ at location $y$
-can be thought of 
-as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
-It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
-i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
-gives the expected total number of individuals ever living in $S$:
-\begin{align} \label{eqn:occupation_density_defn}
-    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
-\end{align}
-% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
-Since reproduction and spatial motion are independent,
-$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
-multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
-i.e.\
-\begin{align}
-    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
-\end{align}
-The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
-is the sum of mean occupation densities of a constant outflux of branching processes
-from $\partial B(r_0)$:
-\begin{align} \label{eqn:occupation_integral}
-    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
-         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
-\end{align}
-
-This form we can now compare to 
-the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
-Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
-We assume this new patch is circular, 
-has relatively small area $A$ and is centered at position $y$.
-% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
-% multiplied by the probability that the family reaches the patch, given survival.
-% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
-It would be possible to analyze this probability directly,
-as in \citet{barton1987establishment};
-but here we take a simpler route, 
-approximating this by the chance that the trunk hits the new patch,
-multiplied by the chance that at least one member of the family escapes demographic stochasticity
-and successfully establishes in the new patch.
-Write $h_A(x,y)$ for the probability that the Brownian trunk
-hits the new patch centered at $y$ before the family dies out,
-and $f(A)$ for the chance that the family manages to establish in the new patch,
-given that it successfully arrives.
-As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
-\begin{align}\label{eqn:migrate_integral}
-  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
-\end{align}
-
-Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
-while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
-Since the expected total time spent in the patch does not depend on the distance at which the family started
-if we know that it survived to hit the patch,
-it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
-Following this intuition,
-and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-it turns out (see section~\nameref{apx:hitting_occupation}) that
-\begin{align} 
-    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
-\end{align}
-where $g(A)$ is the total expected occupation time of the patch given the family arrives.
-Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
-Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
-that (at least asymptotically) does not depend on the distance between $x$ and $y$,
-we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
-The probability $f(A)$ that the family does not die out once it arrives at the new patch is
-approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
-if each member arriving has chance $p_e$ of establishing, independently.
-\plr{And if $p_e$ is small?}
-% If we let $w$ be the width of the patch, then the remaining term is
-% \begin{align}
-%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
-% \end{align}
-Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{apx:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
-we obtain that
-\begin{align}
-  \label{eqn:migrate} 
-  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
-  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
-\end{align}
-This does not depend on $A$, because although
-$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
-the dependence is very weak.
-For instance, the width of a circular patch only very weakly affects the chance 
-of establishment of a new migrant that appears on its edge,
-as long as the patch is large enough.
-For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{apx:hitting_occupation}.
+%%%% %%%%
+\subsubsection{Heuristics}
+\label{ss:heuristics}
+
+The parallel form suggested by \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
+suggests that we can find $\migrate(x)$ by multiplying the mean density $q(x)$
+by the ratio of the second terms in the two equations,
+i.e.
+\begin{align} \label{eqn:ratio_gestalt}
+  \migrate(x) 
+    &= \rho q(x) \times 
+    \frac{
+        \text{ ( probability a family establishes in patch at $x$ ) }
+    }{
+        \text{ ( mean occupation density of a family at $x$ ) }
+    } .
+\end{align}
+To make this precise we need to understand what we mean by ``outflux of families'';
+we do this below in \nameref{ss:hitting_occupation},
+and provide a heuristic argument here.
+
+Both numerator and denominator in \eqref{eqn:ratio_gestalt} 
+count only families that have arrived at the new patch,
+so we can consider a family conditioned on having successfully transited between the patches.
+To obtain simply interpretable expressions,
+we assume that once any member of the family arrives at the patch,
+all other members are close by, and so have roughly equal opportunity
+to establish in the new patch.
+
+Suppose the family arrives with $K$ individuals,
+i.e.\ there are $K$ individuals that manage to establish
+The probability that at least one individual leaves a lineage that establishes in the patch
+is $1-(1-p_e)^K$.
+On the other hand, in the absence of the patch of new habitat,
+each of the $K$ newly arrived individuals would leave behind a lineage with 
+mean total size $\int_0^\infty e^{-t s_m} dt = 1/s_m$,
+roughly half of which would stay in the new patch.
+Therefore, the expected contribution of the family to the occupation density is $K/2s_m$.
+(Readers who feel uncomfortable about the geometry of the patch,
+or the distinction between occupation density and number of individuals
+are encouraged to take a detour through \nameref{ss:hitting_occupation}, below.)
+
+This suggests that 
+\begin{align} \label{eqn:ratio_K}
+  \migrate(x) 
+    &= \rho q(x) \times 
+    \frac{
+        \E[1-(1-p_e)^K]
+    }{
+        \E[K]/2s_m
+    } .
+\end{align}
+There are now two extremes to consider.
+If $\P\{K>1/p_e}$ is small, then each family has a small chance of establishment,
+and the numerator is approximately $p_e \E[K]$.
+On the other hand, if $p_e$ is not too small,
+then the numerator would be close to 1, as any family that arrives will likely establish.
+Combining these, and using that $\E[K] \approx 1/s_m$, 
+\begin{align} \label{eqn:twocase_migrate}
+    \migrate(x) &\approx 2 \rho s_m \min(s_m,p_e) \times q(x) .
+%     \begin{cases}
+%         q(x) \times 4 \rho s_m s_p / \xi^2  \qquad & \text{ if } s_p < s_m \\
+%         q(x) \times 2 \rho s_m^2 \qquad & \text{ if } s_p > s_m 
+%     \end{cases} .
+\end{align}
+When we use this, we set $p_e = 2 s_p / \xi^2$.
 
@@
 i.e.\ that it has exponential tails.
-Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
-supposing that the random lifetime of a family is exponentially distributed with mean $1/s_m$.
+This suggests one further approximation, 
+that the random lifetime of a family is exponentially distributed with mean $1/s_m$;
+equation \eqref{eqn:mean_K} shows this introduces a constant into the equations below. \linelabel{rr:EK}
 Once we do this, 
@@
 \subsection[Hitting and Occupation]{Hitting and occupation}
-\label{apx:hitting_occupation}
+\label{ss:hitting_occupation}
+
+Here we make a more precise argument to back up expression \eqref{eqn:twocase_migrate} for the migration rate.
+The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
+since it relies only on a decomposition of the migrant families upon hitting the new patch;
+but it is also imprecise in a number of subtle ways
+that are difficult to formalize.
+The argument we make here therefore takes a somewhat different tack,
+supposing that it suffices to model the spatial movement of a migrant family
+by following only the motion of the ``trunk''
+(i.e.\ the red line in figure \ref{fig:branching_decomp}),
+and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
+(Recall $\sigma$ is the dispersal distance.)
+We then use facts about Brownian motion and branching processes, 
+to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
+\linelabel{rr:moved_argument}
+
+We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
+as the sum of independent migrant families, each of whose dynamics are
+given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
+Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
+Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
+such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ 
+that initiate new migrant families is $\gamma(x)$.
+Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``ouflux of families'',
+when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
+By examining the constituent terms,
+it will turn out that equation \eqref{eqn:gestalt_migrate} 
+is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.
+
+First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
+Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
+and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
+The \emph{mean occupation density} of $Z$ at location $y$
+can be thought of  informally \linelabel{rr:density}
+as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
+It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
+i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
+gives the expected total number of individuals ever living in $S$:
+\begin{align} \label{eqn:occupation_density_defn}
+    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
+\end{align}
+% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
+Since reproduction and spatial motion are independent,
+$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
+multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
+i.e.\
+\begin{align}
+    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
+\end{align}
+The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
+is the sum of mean occupation densities of a constant outflux of branching processes
+from $\partial B(r_0)$:
+\begin{align} \label{eqn:occupation_integral}
+    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
+         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
+\end{align}
+
+This form we can now compare to 
+the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
+Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
+We assume this new patch is circular, 
+has relatively small area $A$ and is centered at position $y$.
+% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
+% multiplied by the probability that the family reaches the patch, given survival.
+% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
+It would be possible to analyze this probability directly,
+as in \citet{barton1987establishment};
+but here we take a simpler route, 
+approximating this by the chance that the trunk hits the new patch,
+multiplied by the chance that at least one member of the family escapes demographic stochasticity
+and successfully establishes in the new patch.
+Write $h_A(x,y)$ for the probability that the Brownian trunk
+hits the new patch centered at $y$ before the family dies out,
+and $f(A)$ for the chance that the family manages to establish in the new patch,
+given that it successfully arrives.
+As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
+\begin{align}\label{eqn:migrate_integral}
+  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
+\end{align}
+Recall that $\rho$ appears here but not in \eqref{eqn:occupation_integral}
+because $q(x)$ is the local \emph{proportion} of $B$ alleles.
+
+Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
+while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
+Since the expected total time spent in the patch does not depend on the distance at which the family started
+if we know that it survived to hit the patch,
+it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
+Following this intuition,
+and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
+it turns out (see section~\nameref{ss:hitting_occupation}) that
+\begin{align} 
+    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
+\end{align}
+where $g(A)$ is the total expected occupation time of the patch given the family arrives.
+Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
+Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
+that (at least asymptotically) does not depend on the distance between $x$ and $y$,
+we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
+The probability $f(A)$ that the family does not die out once it arrives at the new patch is
+approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
+if each member arriving has chance $p_e$ of establishing, independently.
+\plr{And if $p_e$ is small?}
+% If we let $w$ be the width of the patch, then the remaining term is
+% \begin{align}
+%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
+% \end{align}
+Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{ss:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
+we obtain that
+\begin{align}
+  \label{eqn:migrate} 
+  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
+  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
+\end{align}
+This does not depend on $A$, because although
+$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
+the dependence is very weak.
+For instance, the width of a circular patch only very weakly affects the chance 
+of establishment of a new migrant that appears on its edge,
+as long as the patch is large enough.
+For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{ss:hitting_occupation}.
 

### 2014-11-26 patchy-review-responses.tex
more responses...

@@
   Resubmission Cover Letter \\
-  {\it PLoS Biology}
+  {\it PLoS Genetics}
 \end{minipage}
@@
 \begin{minipage}[b]{2.5in}
-  Department of Evolution and Ecology\\
+    Department of Molecular and Computational Biology\\
+    University of Southern California, \\
+    \emph{and} Department of Evolution and Ecology\\
     University of California at Davis\\
@@
     of migrant families of
-    \[ k ~ \rho q(x) l/n/tau ~ \rho q(x) l s^2 \]
+    \[ k ~ \rho q(x) l/n/\tau \sim \rho q(x) l s^2 \]
 }
@@
     The probability of establishment of at least one member of family is then
-    \[ P_{est} = 1 - exp(-p n) ~ 1 - exp(-p_e/s) \]
+    \[ P_{est} = 1 - \exp(-p n) ~ 1 - \exp(-p_e/s) \]
     where $p_e$ is the establishment probability of a single individual. The result
@@
     I don't seem to be able to identify the equivalent of
-    l in the author's formulae, but I believe it should be there. After all, rho
-    q(x) is a density and needs to be multiplied by a length (or area in d=2) to
+    $\ell$ in the author's formulae, but I believe it should be there. After all, 
+    $\rho q(x)$ is a density and needs to be multiplied by a length (or area in d=2) to
     give a meaningful rate.
@@
 
+%%%%%%%%
 \point{Fig 4:}{ Why not use natural units for the axis of these plots,
@@
 
-\point{1.}{
+\begin{quote}
+Briefly, I think that the derivation of the main results (12), (21), and (25) can
+be made much shorter and easier to understand (for both empiricists and the-
+orists), and that there need to be simulations that quantitatively confirm these
+results. Overly-detailed comments are below.
+\end{quote}
+
+\point{}{
     I think that the main result (12) can be derived much more simply, and
@@
 
-\point{2.}{
+\point{}{
     However, I think we should also care about the opposite limit, $s_m \ll s p$. In
@@
 
+%%%%%%%%
+\point{}{ 
+    As it stands, $\migrate(y)$ is the rate of successful migrants at an infinitesimal
+    area at distance $y$, so it needs to be integrated over an area to get the rate
+    for a whole patch. The authors don’t do this, and as a result, expressions
+    like (14) are dimensionally inconsistent. It’s a little annoying to define the
+    right area of integration, but it’s straightforward to argue that up to a
+    constant the answer
+     should be given by integrating $\migrate(y)$ over a region
+     of thickness $\sigma/\sqrt{2sm}$ around the boundary of the patch. I think it’s easier
+    to work with fluxes and just integrate over the boundary of the patch, with
+    no worries about thickness, so that’s why I used them above. (Of course,
+    you get the same answer either way, except for small or weirdly-shaped
+    patches, where either or both can break down.)
+}
+
+%%%%%%%%
+\point{}{ 
+    I think that the authors need to show simulation results confirming the
+    analytic approximations. Spatial simulations are a pain, but it’s important
+    to have a (somewhat) independent check on the math and assumptions. At
+    a minimum, I think that there needs to be a figure showing that they have
+    the right expression for the rate of successful migrants (12) and migration
+    sojourn times (17) over a range of parameter values. If it’s not too much
+    work, I think it would be really nice to have a figure that doubles as a kind
+    of graphical summary, showing a scattering of patches in two dimensions,
+    with one starting as the homeland of the B allele and the rest colored
+    according to their probability in simulations of being colonized by that
+    original allele vs new mutations, and a curve marking the region predicted
+    as being colonized by the analytical approximations, or something like
+    that. But I understand if the authors don’t want to spend their time on
+    that.
+}
+
+%%%%%%%%
+\point{}{ 
+    It would be nice to have a table of symbol definitions.
+}
+
+\reply
+We've tried to say in words what each symbol is when we use it as much as possible,
+so we hope this isn't necessary.
+
+%%%%%%%%
+\point{}{ 
+    And speaking of symbols, I think it’s a bit nicer to write things in terms
+    of the characteristic patch length $a \sim A^{1/d}$, rather than the area. This
+    avoids having to talk about one-dimensional areas \ldots
+}
+
+\reply
+Good point, but $A$ is what shows up naturally in $\mutrate$, so we've left it as-is.
+
+%%%%%%%%
+\point{}{ 
+     It would be nice to have a clear, concise summary of the quantitative
+     assumptions needed for the results (all the stuff like $a \gg \sigma/\sqrt{2s_m}$ or
+     whatever).
+ }
+
+\reply
+See text added in \nameref{ss:assumptions}.
+
+%%%%%%%%
+\point{"Length of the hitchhiking haplotype"}{
+    I would focus more on the asymptotic expressions for $R \gg w  \gg \sigma/ \sqrt{2s_m}$, since that’s the regime
+     where the analysis makes the most sense. Then the expressions become
+     simpler and you can treat d = 2 as well. If you want, you could even
+     lead with the intuitive explanation at the end of the section, and save the
+     Bessel functions for an appendix.
+}
+
+ \reply
+
+%%%%%%%%
+\point{}{ 
+     I think (24) follows directly from Barton, Heredity 1979. (If you look at
+     the unnumbered equation above Fig. 1 in that paper and plug in $x \sim a$,
+     you find the time scale $r_{eff}$.) I think that the displayed equation should just
+     be $r_{eff} \approx r\sigma/\sqrt{s}/a$. (I feel like the extra correction terms and geometric
+      factor go against the spirit of the paper, and in any case are probably
+      small compared to the inaccuracies introduced just by applying the model
+      to any real system.) I think you can cut essentially everything before
+      the simple explanation just above (24); if people want the mathematical
+      details, they can read Barton. If I am misunderstanding things and there
+    is a new result here, rather than just a good new exposition of an existing
+    result, there should be a figure showing simulations confirming it.
+}
+
+\reply
+
+%%%%%%%%
+\point{In “Applications”,}{the authors spend a lot of time introducing the mice
+    and calculating the values for Fig. 6, but then they don’t show where the
+    actual data lie on those plots. Is there enough information to do some
+    kind of comparison? How close are the nearest samples to each other?
+    Have no examples of alleles spreading by migration been found? If not,
+    how much of a bound does this put on the parameter values? Is there any
+    data on the genetic map length of the region of reduced diversity around
+    the selected locus?
+}
+
+\reply
+
+%%%%%%%%
+\point{}{ 
+    A few sections or passages seem out of place:
+    \begin{itemize}
+        \item[A.]  I think “Simulation methods” should go in the “Materials and meth-
+            ods” section, or at least not right between “Establishment of a locally
+            adaptive allele due to mutational influx” and “Establishment of a lo-
+            cally adaptive allele due to mutational influx”.
+
+        \item[B.] The sentence on p7 beginning “By our definition of fitness. . . ”.
+
+        \item[C.]  The paragraph starting “Also note that. . . ” on p11.
+    \end{itemize}
+}
+
+\reply
+
+%%%%%%%%
+\point{Fig. 4:}{$R/\sigma$ should have dimensions of $\sqrt{\text{time}}$. In the caption, I think it’s
+    misleading to say that $\migrate$ is ``relatively insensitive to $s_m$''. It’s exponen-
+    tially sensitive! I don’t understand the blue numbers in the right-hand
+    panel. It might be easier to interpret these if they were contour plots
+    of the total time to adaptation, colored according to the value of (14),
+    although I’m not sure. Also, they aren’t really ``phase diagrams''.
+}
+
+\reply
+
+%%%%%%%%
+\point{Top of p11 (just above “Multiple patches”):}{should be $\sqrt{2s_m}/\sigma$.
+}
+
+\reply
+
+%%%%%%%%
+\point{Between (49) and (51):}{should be “Green’s”.
+}
+
+\reply
+
+%%%%%%%%
+\point{(51)-(55)}{are missing factors of $\sigma$.
+}
+
+\reply
+
+%%%%%%%%
+\point{}{ 
+ In (56) and following, it would be clearer to use different symbols for scaled
+and unscaled distances.
+
+}
+
+\reply
+
 %%%%%%%%%%%%%%  Reviewer 3
@@
     restricted to the part where the transition probability declines with $s_m$.
-    But it is not easy to see how this needs to be done in a prober way. We
+    But it is not easy to see how this needs to be done in a proper way. We
     certainly need a minimal distance among patches. On top of that, we may
@@
 \point{[beginning comments in the pdf]}{
-``Do you assume a Gaussian kernel? Or is this irrelevant? Which properties of the kernel are required? Only existence of a finite $sigma^2$?''
+``Do you assume a Gaussian kernel? Or is this irrelevant? Which properties of the kernel are required? Only existence of a finite $\sigma^2$?''
 }
@@
 %%%%%%
-\point{}{``This is not sufficient: the patch also needs to be large enough that p_e is (almost) reached at its center. Generally, the approximation should over-estimate the true value.''}
+\point{}{``This is not sufficient: the patch also needs to be large enough that $p_e$ is (almost) reached at its center. Generally, the approximation should over-estimate the true value.''}
 
@@
 %%%%%%%%
-\point{}{ 
+\point{[equation \eqref{eqn:migrate}]}{ 
     \begin{quote}
@@
     \begin{itemize}
-        \item equation (3) for the expected mutant density is only an asymptotic. Therefore, we probably need to require $|y| \gg \sigma / \sqrt(2s_m)$.
+        \item equation (3) for the expected mutant density is only asymptotic. Therefore, we probably need to require $|y| \gg \sigma / \sqrt(2s_m)$.
 
@@
 %%%%%%%%
-\point{[expressio nfor the outflux of migrants]}{do we need this for anything?}
+\point{[expression for the outflux of migrants]}{do we need this for anything?}
 
@@
 
+%%%%%%%%
+\point{[figure 1 showing unimportance of square root term]}{}
+
+\reply
+Clarified. \llname{rr:sqrt_term}
+
+%%%%%%%%
+\point{``the assumption of well-separated patches''}{ 
+}
+
+\reply
+We have changed this. \llname{rr:separated_patches}
+
+%%%%%%%%
+\point{[figure 4, the contour plots]}{
+}
+
+\reply
+Good point; we didn't notice that after correcting $\migrate$ the parameter regime on the right didn't make sense.
+We have substituted XXX for these.
+
+%%%%%%%%
+\point{[\eqref{eqn:total_migrate} assumes no interference}{ 
+}
+
+\reply
+Good point, but one we don't think we need to get into; we've added a small caveat.
+\llname{rr:interference}
+
+%%%%%%%%
+\point{}{ 
+    One message here seems to be that "migration vs mutation" is independent of the density.
+As far as the model goes, this is correct.
+However, in Nature selection will typically be density dependent, usually with stronger deleterious selection (higher $s_m$) at higher densities.
+One could therefore suspect that higher densities will often favor mutation relative to migration.
+}
+
+\reply
+Interesting point,
+but we think its applicability depends on what densities are being compared.
+If it is in two different species, then $s_m$ must clearly be measured separately in the two,
+so this is not a worry.
+If it is two populations of the same species,
+then one might also argue that lower densities implies worse habitat,
+and so stronger deleterious selection.
+We don't see anything straightforward we can say here, 
+other than adding this caveat later. \llname{rr:popdensity}
+
 %%%%%% p.12
@@
 
+%%%%%%%%
+\point{}{ 
+    There is a lot of handwaving in this section.
+Although each single step has some plausability, the final result seems to depend on a lot of assumptions.
+There is also an important difference to the other sections.
+Most of the other results are valid for large distances, where we can safely use asymptotic forms of the distribution (which arguably are quite universal).
+In contrast, this section relies on short-range dispersal (limited excursions of B-mutant lines to the surrounding area, where b is common).
+This is far less universal and should depend on details of dispersal.
+
+    Also, it seems to me that another route to break down LD is by b-immigrants to the patch, who will life there for an exponential life span and can recombine before they die out.
+Note that, in the pocket mouse example below, the frequency of dark mice on the lava flows is only 86\%.
+This should result in considerable recombination "on site".
+
+    So, in summary: I am sceptical concerning the results of this section.
+They are preliminary at best and need further study.
+I therefore suggest to cut them from the paper (or reduce it to some brief remarks).
+}
+
+\reply
+
+%%%%%%%%
+\point{Dominance.}{ 
+    the deleterious effect could be recessive and the beneficial one still (at least partially) domninant.
+}
+
+\reply
+Added that note. \llname{rr:dominance}
+
 % \end{document}

### 2014-11-26 patchy-selection-paper.tex
more responses...

@@
 \newif\ifreviewresponses
-% \reviewresponsestrue  % include them
+\reviewresponsestrue  % include them
+% \reviewresponsesfalse  % don't include them
 \ifreviewresponses\else\renewcommand{\linelabel}[1]{}\fi
-\reviewresponsesfalse  % don't include them
 \iflatexml\reviewresponsesfalse\fi  % they don't work so well in html
@@
 \newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
+\newcommand{\given}{\,\vert\,}
 
@@
 which will be a good approximation if the patch is large relative the scale over which $p(x)$ goes from 0 to $p_e$,
-or if $p(x)$ is close to $p_e$ at some point and is symmetric about the edge of the patch. \lineref{rr:p_estab}
+or if $p(x)$ is close to $p_e$ at some point and is symmetric about the edge of the patch. \linelabel{rr:p_estab}
 We examine this more generally via exact calculation of $p(x)$ in the section \nameref{apx:establishment_sims}.
@@
 \begin{align} \label{eqn:EK_limit}
-    e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty .
+    e^{-s_m t}/(1-k_e(t)) \to \E[K] \quad \text{as} \quad t \to \infty .
 \end{align}
@@
 There are now two extremes to consider.
-If $\P\{K>1/p_e}$ is small, then each family has a small chance of establishment,
+If $\P\{K>1/p_e\}$ is small, then each family has a small chance of establishment,
 and the numerator is approximately $p_e \E[K]$.
@@
 then the numerator would be close to 1, as any family that arrives will likely establish.
-Combining these, and using that $\E[K] \approx 1/s_m$, 
+Combining these, and absorbing the term $\E[W|W>1]$ from equation \eqref{eqn:mean_K} 
+into the constant $C$,
 \begin{align} \label{eqn:twocase_migrate}
-    \migrate(x) &\approx 2 \rho s_m \min(s_m,p_e) \times q(x) .
+    \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
 %     \begin{cases}
@@
 When we use this, we set $p_e = 2 s_p / \xi^2$.
+
+%%%%%%%%%
+\subsection{What are we assuming?}
+\label{ss:assumptions}
+
+A quick review of the parameter regime we are studying here:
+First, we assume that 
+\textbf{(a)} each patch is large enough to support a stable population of $B$ alleles
+($A^{1/d} > (\sigma/\sqrt{2s_p}) \tan^{-1}(\sqrt{s_m/s_p})$).
+Next, in using equation \eqref{eqn:eqfreq}, we assume that 
+\textbf{(b)} the inter-patch distance is at least several multiples of the dispersal distance
+($R \gg \sigma$),
+and that 
+\textbf{(c)} local drift is not so strong that the equilibirum frequency is at least approximately attained
+($s_m \gg 1/(\rho \sigma^d)$).
+We also neglect the time for migration-selection equilibrium to be reached,
+which has the effect of over-estimating $\migrate$.
+
+We also assume that migration to a new patch takes place over a number of generations;
+if there are sufficient rare, long-distance migration events
+that would move between patches in a single hop,
+this would require a different analysis.
 
@@
 for the effective rates of adaptation by new mutation and by migration.
-We work in one dimension, as the square root term appearing for two dimensions is relatively unimportant (see figure~\ref{fig:sim_occupation_freqs}).
+We work in one dimension, as the square root term appearing for two dimensions is relatively unimportant 
+(as seen in figure~\ref{fig:sim_occupation_freqs} by the fact that the $d=2$ curve is nearly straight).
+\linelabel{rr:sqrt_term}
 
@@
 If $s_m=.001$, then with this higher mutational influx this calculation predicts that mutation will \emph{always} be faster than migration
--- however, this is misleading, 
-since with $s_m=.001$, for the assumption of well-separated patches to hold we would need $R/\sigma > 1/\sqrt{2s_m} = 22.4$.
-The correct conclusion is that convergent adaptation is much more likely than shared adaptation
-between patches separated by at least 22.4$\sigma$;
-for shorter separations the model breaks down.
+-- however, this should be taken very cautiously,
+since as discussed above, this model does not hold if $R$ is of the same order as $\sigma$
+or if $s_m$ is small enough that local drift is more important.
+\linelabel{rr:separated_patches}
 
@@
 
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics{phase-diagram-log}
-  \end{center}
-  \caption{
-  Shaded contour plots of
-  mean time until adaptation in a new patch occurs by migration ($\Tmig$, blue) or by mutation ($\Tmut$, red),
-  as a function of population density ($\rho$), 
-  distance to an already adapted patch in units of dispersal distance ($R/\sigma$),
-  strength of deleterious selection ($s_m$),
-  and mutation rate scaled by patch area ($\mu A$).
-  The right-hand plot shows that $\migrate$ is relatively insensitive to $s_m$ (thanks to the square root).
-  The black, dotted line shows values where $\Tmig=\Tmut$.
-  Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=20$, and $\xi^2=1$.
-  Note that population density $\rho$ and mutation influx $A\mu$ are shown on a log scale.
-  \label{fig:phase_diagram}
-  }
-\end{figure}
+% \begin{figure}[ht]
+%   \begin{center}
+%     \includegraphics{phase-diagram-log}
+%   \end{center}
+%   \caption{
+%   Shaded contour plots of
+%   mean time until adaptation in a new patch occurs by migration ($\Tmig$, blue) or by mutation ($\Tmut$, red),
+%   as a function of population density ($\rho$), 
+%   distance to an already adapted patch in units of dispersal distance ($R/\sigma$),
+%   strength of deleterious selection ($s_m$),
+%   and mutation rate scaled by patch area ($\mu A$).
+%   The right-hand plot shows that $\migrate$ is relatively insensitive to $s_m$ (thanks to the square root).
+%   The black, dotted line shows values where $\Tmig=\Tmut$.
+%   Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=20$, and $\xi^2=1$.
+%   Note that population density $\rho$ and mutation influx $A\mu$ are shown on a log scale.
+%   \label{fig:phase_diagram}
+%   }
+% \end{figure}
 
@@
 and that these patches are at distances $R_1, \ldots, R_k$ away from this unadapted patch.
-The total rate of adaptation through migration, from equation \eqref{eqn:migrate}, is
+The total rate of adaptation through migration, from equation \eqref{eqn:migrate}
+if patches do not interfere with each other, \linelabel{rr:interference}
+is
 \begin{equation} \label{eqn:total_migrate}
@@
 selective benefit within the patches or (perhaps surprisingly) the population density, 
-although these two factors will determine the time-scale over which adaptation will occur. 
+although these two factors will determine the time-scale over which adaptation will occur
+(and note that population density could affect the selection coefficients). \linelabel{rr:popdensity}
 This is in contrast with models of panmictic populations
@@
 This is counteracted, however, by the increased difficulty with which the mutation would establish
-once arriving in a new patch. 
+once arriving in a new patch,
+if the beneficial effect is also recessive. \linelabel{rr:dominance}
 As such it is not clear what our intuition should be 
@@
 We have also ignored the possibility of very long distance migration,
-instead focusing on local dispersal (and using a Gaussian model).
+instead focusing on local dispersal (and using Gaussians via the central limit theorem).
 However, dispersal distributions can be very heavy tailed, 
@@
 \begin{align}
-    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P\{ \tau_S \in dt \} \\
-    &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P\{ \tau_S \in dt \} \\
-    &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
-    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
+           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
+           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
     &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
@@
 
+\includereviews
+
 \end{document}

### 2014-11-26 analyze-sims.R

@@ -94,7 +94,8 @@ migsims$gm <- getgrowth(migsims)$gm
 migsims$paramstring <- with(migsims, factor( paste(sm.name,R.name,N.name,dims,sep='_') ) )
 
 migsims$adjust <- 0  # with( migsims, log(N)/gb + size1/2 / ( sigma * sqrt(gb) ) )
-migsims$migtime <- with( migsims, 0.5/( 4 * N * gb * abs(sm) * exp(- R * sqrt(2*abs(sm))/sigma ) ) + adjust )
+# migsims$migtime <- with( migsims, 0.5/( 4 * N * gb * abs(sm) * exp(- R * sqrt(2*abs(sm))/sigma ) ) + adjust )
+migsims$migtime <- with( migsims, 1/( 5 * ((1+gb)/(1-exp(-(1+gb)))) * N * pmin(2*gb/(1+gb),abs(sm)) * abs(sm) * exp(- R * sqrt(2*abs(sm))/sigma ) ) + adjust )
 migsims$adapted <- ( 
     ( migsims$dims == "1D" ) &
     ( migsims$hit100.2 < 24000 ) &
@@ -145,6 +146,7 @@ pairs( migsims[c("final1", "time1", "hit100.1", "final2", "time2", "hit100.2","m
 # for paper
 
 require(colorspace)
+require(calibrate)
 
 pdf( file="times-predicted-observed.pdf", width=6.5, height=3.3, pointsize=10 )
 layout(t(1:2))
@@ -219,12 +221,50 @@ dev.off()
 ##
 # phase space picture of probability of migration versus mutation
 
-pmut <- function(sm,mu,R,A,sigma,C=1/2) { ( A * mu ) / ( A*mu + 2*C*sm*exp(-R*sqrt(2*sm)/sigma^2) ) }
-paramgrid <- expand.grid( N=seq(25,1600,length.out=100), sm=10^(seq(-4,-1,length.out=100)), mu=1e-6, R=seq(20,160,length.out=100), A=99, sigma=0.953770108230929 )
-
+pmut <- function(sm,mu,R,A,sigma,C=10/2) { ( A * mu ) / ( A*mu + 2*C*abs(sm)*exp(-R*sqrt(2*abs(sm))/sigma^2) ) }
+muval <- 1e-5
+Aval <- 99
+sigmaval <- median(migsims$sigma)  # i.e. most common one, here.
[…]

### 2014-11-26 patchy-review-responses.tex
more sims

@@
 \section*{From the editor}
-% From: "Catriona J. MacCallum" <cmaccallum@plos.org>
-% To: Peter Ralph <pralph@usc.edu>
-% Date: Fri, 15 Feb 2013 04:01:05 -0500
-% Subject: PLOS Biology Decision [PBIOLOGY-D-12-02842R1]
-% Dear Dr. Ralph,
 
@@
 
+Changes of note are:
+
+\begin{itemize}
+
+  \item We have done a large number of simulations, and compared the results to the theory;
+    see figures~\ref{fig:sim_probs} and \ref{fig:sim_times}.
+
+  \item Several reviewers pointed out that in approximating $\E[1-(1-p_e)^K] \approx \E[K] p_e$
+    we were neglecting the case that $s_m \ll s_p$.
+    This was pur laziness on our part,
+    which we have rectified; see the new equation~\eqref{eqn:twocase_migrate}.
+
+\end{itemize}
+
 %%%%%%%%%%%%%%
@@
     The probability of establishment of at least one member of family is then
-    \[ P_{est} = 1 - \exp(-p n) ~ 1 - \exp(-p_e/s) \]
+    \[ P_{est} = 1 - \exp(-p n) \sim 1 - \exp(-p_e/s) \]
     where $p_e$ is the establishment probability of a single individual. The result
@@
 
+\reply
+As discussed above, we have incorporated this good point; see around equation~\eqref{eqn:twocase_migrate}.
+
 %%%%%%%%
@@
 \reply
-    See above.
+    See the response to point 1.1 above.
 
@@
 \reply
+  We've done this; see above.
 
@@
 
+\reply
+This was already discussed in the discussion, but we've added a note about it earlier on, citing Hallatschek \& Fisher.
+\llname{rr:longtailed}
+
 %%%%%%%%
@@
   }
+
+\reply
+Good idea, but we've decided to cut this figure.
 
@@
 
+\reply
+Very good points.  
+We apologize for not catching the nonmonotonicity earlier.
+We've added a section, \nameref{ss:assumptions}, discussing this point,
+as well as a note regarding the nonmonotonicity wrt $s_m$
+\llname{rr:sm_condition}.
+
 \point{1b.}{

### 2014-11-26 patchy-selection-paper.tex
more sims

@@
 and do not discuss the (interesting) cases where rare, long-distance dispersal events
-are more important \citep[see][for discussion]{longtailed,ralphcoop2010,hallatschek??}.  \linelabel{rr:longtailed}
+are more important \citep[see e.g.][for discussion]{levin2003ecology,ralph2010parallel,hallatschek2014acceleration}.  \linelabel{rr:longtailed}
 %We also aim to describe the scenario where it is sensible to think of large, discrete patches;
@@
 \begin{align} \label{eqn:twocase_migrate}
-    \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
+  \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m} 
 %     \begin{cases}
@@
 \end{align}
-When we use this, we set $p_e = 2 s_p / \xi^2$.
+When we use this below, we set $p_e = 2 s_p / \xi^2$.
+Note that the function increases for small $s_m$, 
+but decreases if $s_m > |y|^2 / 2 \sigma^2$, 
+i.e.\ where the condition holds.
 
@@
 
-One way to visualize this is by a phase diagram.
-In figure \ref{fig:phase_diagram},
-we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
-as a function of various parameters. 
-This gives us an idea of not only the parameter space where parallel adaptation is likely ($\Tmig \ll \Tmut$),
-but also where adaptation by any means takes a very long time ($\Tmig$ and $\Tmut$ both very large),
-an important check in applications of these results.
-
+% One way to visualize this is by a phase diagram.
+% In figure \ref{fig:phase_diagram},
+% we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
+% as a function of various parameters. 
+% This gives us an idea of not only the parameter space where parallel adaptation is likely ($\Tmig \ll \Tmut$),
+% but also where adaptation by any means takes a very long time ($\Tmig$ and $\Tmut$ both very large),
+% an important check in applications of these results.
+% 
 % \begin{figure}[ht]
@@
 with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
-In addition, over long time-scales, very rare chance events (mice carried off by hurricanes and the like \citet{censky1998overwater,nathan2008mechanisms})
+In addition, over long time-scales, very rare chance events (mice carried off by hurricanes and the like; \citet{censky1998overwater,nathan2008mechanisms})
 could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.

### 2014-11-26 analyze-sims.R

@@ -155,15 +155,12 @@ par(mar=c(4,4,1,1)+.1)
     # remove simulation configurations where a substantial portion haven't yet adapted
     usethese <- with( mutsims, adapted & ( muttime < maxtime/10 ) ) 
     with( droplevels(subset(mutsims,usethese)), {
-            sm.vals <- (levels(droplevels(sm.name)))
-            sm.col.pal <- sequential_hcl(length(sm.vals))
+            sm.vals <- (levels(droplevels(sm.name))); sm.col.pal <- sequential_hcl(length(sm.vals))
             sm.pch <- match( levels(sm.name)[tapply(sm.name,paramstring,"[",1)], sm.vals )
             sm.cols <- sm.col.pal[ sm.pch ]
-            N.vals <- (levels(droplevels(N.name)))
-            N.col.pal <- terrain_hcl(length(N.vals))
+            N.vals <- (levels(droplevels(N.name))); N.col.pal <- terrain_hcl(length(N.vals))
             N.cols <- N.col.pal[ match( levels(N.name)[tapply(N.name,paramstring,"[",1)], N.vals ) ]
-            mu.vals <- (levels(droplevels(mu.name)))
-            mu.col.pal <- sequential_hcl(length(mu.vals))
+            mu.vals <- (levels(droplevels(mu.name))); mu.col.pal <- sequential_hcl(length(mu.vals))
             mu.pch <- match( levels(mu.name)[tapply(mu.name,paramstring,"[",1)], mu.vals )
             mu.cols <- mu.col.pal[ mu.pch ]
             xx <- exp(jitter(log(tapply(muttime,paramstring,mean,na.rm=TRUE))))
@@ -189,23 +186,20 @@ par(mar=c(4,4,1,1)+.1)
                 pt.bg=c(NA,sm.col.pal)[c( rep(1,nlevels(sm.name)), rep(1,nlevels(N.name)), 1+(1:nlevels(mu.name)) ) ] 
             )
         } )
+
 # MIGRATION
     usethese <- with( migsims, adapted  & ( migtime < maxtime ) ) 
     with( droplevels(subset(migsims,usethese)), {
-            sm.vals <- (levels(droplevels(sm.name)))
-            sm.col.pal <- sequential_hcl(length(sm.vals))
+            sm.vals <- (levels(droplevels(sm.name))); sm.col.pal <- sequential_hcl(length(sm.vals))
[…]

### 2014-11-26 sim-mutation-params-newR.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+# Simulate time to mutation in a new patch
+SB=".01"
+SM="-.0001"
+
+for SM in "-.0001" "-.001" "-.01" "-.1"
+do
+    for N in 100 1000
+    do
+        for MU in "1e-5" "1e-6"
+        do
+            THISDIR="mutation/sm${SM}/N-${N}/1D"
+            mkdir -p $THISDIR
+            Rscript generate-patchy-run.R \
+                "outdir=\"${THISDIR}\"" \
+                "N=${N}" "range=c(1,501)" "sb=${SB}" "sm=${SM}" "nsteps=1000" "stepsize=25" "ntypes=2" "mu=${MU}"\
+                's=matrix(ifelse((seq.int(1,range[2])>200)&(seq.int(1,range[2])<300),sb,sm),nrow=1,ncol=range[2])' \
+                'initdist=array(c(rep(N,prod(range)),rep(0,(ntypes-1)*prod(range))),c(range,ntypes))' &
+        done
+    done
+done
+

### 2014-11-26 patchy-review-responses.tex
varous

@@
 \reply
+We moved \nameref{ss:simulations} to Materials and Methods,
+modified the second sentence,
+and rearranged the structure around the third paragraph.
 
 %%%%%%%%
-\point{Fig. 4:}{$R/\sigma$ should have dimensions of $\sqrt{\text{time}}$. In the caption, I think it’s
-    misleading to say that $\migrate$ is ``relatively insensitive to $s_m$''. It’s exponen-
-    tially sensitive! I don’t understand the blue numbers in the right-hand
+\point{Fig. 4:}{
+    In the caption, I think it’s misleading to say that $\migrate$ is ``relatively insensitive to $s_m$''. 
+    It’s exponentially sensitive! I don’t understand the blue numbers in the right-hand
     panel. It might be easier to interpret these if they were contour plots
@@
 \reply
+Good points; we meant that it was less sensitive to $s_m$ than we'd otherwise think;
+we've replaced this figure with a different one anyhow.
 
@@
 \reply
+Rephrased. \revref
 
 %%%%%%%%
-\point{Between (49) and (51):}{should be “Green’s”.
+\point{Between (49) and (51):}{should be ``Green's''.
 }
@@
 \reply
+Always a contentious choice; we're sticking by our position,
+and can cite \citet{borodin2002handbook} as backup. ;)
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%  Reviewer 3

### 2014-11-26 patchy-selection-paper.tex
varous

@@
 %%%%%%%
-\subsection{Simulation methods}
-\label{ss:simulations}
-
-First we briefly describe the simulations used below for illustration and validation
-(and provide the \texttt{R} code used as a supplement).
-We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
-Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
-locally beneficial alleles were more likely to reproduce.
-Each then produced a random, Poisson distributed number of offspring that each, independently, 
-either remained in the same location or
-migrated a random number of steps in a uniformly chosen cardinal direction;
-the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
-\linelabel{rr:sim_details}
-Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
-(Although we described the simulation in terms of individuals,
-we kept track only of total numbers in an equivalent way.)
-
-The reproduction rate in simulations for type $b$ alleles was 30\%; 
-this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
-where the value of $s$ is either $s_m$ or $s_p$ depending on the individual's location.
-This determines the values of $s_m$ and $s_p$ reported in the figures, 
-and do not depend on the basic rate of reproduction.
-However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
-we computed the rate of intrinsic growth, 
-i.e.\ the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
-in the absence of migration.
-(The resulting values are close to the first notion of $s$,
-but give better agreement with theory,
-which uses the second definition.)
-
-To sample lineages, we first simulated forwards in time the population dynamics,
-then sampled lineages moving back through time
-by, in each generation,
-moving each lineage to a new deme with probability proportional to the reverse migration probability
-weighted by the number of $B$ alleles in that deme in the previous time step.
-If more than one lineage was found in a deme with $n$ $B$ alleles,
-then each lineage picked a label uniformly from $1 \ldots n$,
-and those picking the same label coalesced.
-Since reproduction is Poisson, this correctly samples from the distribution of lineages given the population dynamics.
-
-%%%%%%%
 \subsection[Establishment by Migration]{Establishment of a locally adaptive allele due to migrational influx}
@@
 and finally pruning any branches that happen to wander into a patch.
-By our definition of fitness, the mean number of offspring per individual is $e^{-s_m}\approx 1-s_m$,
-and the variance is $\xi^2$.
+By our definition of fitness, 
+in this tree the mean number of offspring per individual is $e^{-s_m} < 1$.
 This pruning removes both branches leading back into the already established patch (which are irrelevant)
@@
 into the constant $C$,
-\begin{align} \label{eqn:twocase_migrate}
+\begin{align} \label{eqn:migrate}
   \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m} 
@@
 i.e.\ where the condition holds.
+\linelabel{rr:sm_condition}
 
@@
 
-%%%%%%%%%
-\subsection{Comparison to simulation}
-\label{ss:comparison_to_sims}
-
-We used the simulations describe above to test the theory,
-etcetera.
+To test the robustness of our results to the various approximations we used,
+we used individual-based simulations on a one-dimensional lattice of 501 demes,
+run for 25,000 generations.
+More details are given below in \nameref{ss:simulations},
+and the number of simulations used and parameter values are given in supplementary tables \ref{stab:mutation_params} and \ref{stab:migration_params}.
+To estimate the mean time until adaptation by mutation,
+we used one centrally located patch of 99 demes
+and a mutation rate of $\mu=10^{-5}$,
+and to estimate the mean time until adaptation by migration,
+we used two centrally located patches of 99 demes
+separated by a varying number of demes,
+with one patch initialized to have the $B$ allele at frequency 0.8.
+In each case, we measured the ``time to adaptation'' as the amount of time until at least 100 $B$ alleles
+were present in the previously unadapted patch.
+Figure~\ref{fig:sim_times}
+summarizes how the results compare to theory,
+excluding parameter combinations where a majority of simulations did not adapt by 25,000 generations
+or the patch was not large enough to support a stable population of $B$ alleles.
 
@@
   \caption{
-      \textbf{(A)} Time to adaptation by new mutation 
-      in 3,128 simulations,
-      compared to the value predicted by equation X.
-      For each of 52 different parameter combinations (Supp Table XX),
-      we performed between 40 and 100 independent simulations;
+      \textbf{(A)} 
+      Time until adaptation by mutation in an isolated patch for different parameter combinations,
+      as observed in simulations (vertical axis),
+      and as predicted from equation~\eqref{eqn:mutrate} (as $1/\mutrate$, horizontal axis).
+      \textbf{(B)} 
+      Time until adaptation by migration from an already occupied patch to a new patch,
+      as observed in simulations (vertical axis),
+      and as predicted from equation~\eqref{eqn:migrate} (as $1/\migrate$, horizontal axis).
+      For each,
+      parameter combinations and numbers of simulations are in supplemental tables \ref{stab:migration_params} and \ref{stab:mutation_params};
       shown are the median (points) and interquartile range (lines)
       of the time until the patch had at least 100 adapted individuals,
-      for each unique parameter combination (example runs in Supp Figure XX).
-      \textbf{(A)} Similar, for time to adaptation by migration
-      in 4,158 simulations.
-      TODO: put equations on axis labels, describe more, make supp figs.
+      for each unique parameter combination.
   }   \label{fig:sim_times}
 \end{figure}
+
+%%%%%%%%%
+\subsection{Comparison to simulation}
+\label{ss:comparison_to_sims}
+
+We used the simulations describe above to test the theory,
+etcetera.
 
@@
 so that the probability of parallel mutation should increase
-approximately logistically with the distance between the patches, at rate $\sqrt{s_m} /\sigma$.
+approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{s_m}$. \revpoint{2}{13}
 
@@
 Equations \eqref{eqn:total_migrate} and \eqref{eqn:migprob} specify the transition rates of the process.
+
 Since the compound parameter $s_p \rho / \xi^2$ is common to both migration and mutation rates,
@@
 the sets of patches sharing the genetic basis of their adaptive response.
-
 Also note that we can rescale time by the typical patch size, and introduce a parameter, say $a_k = A_k/\bar A$,
@@
 
+%%%%%%%
+\subsection{Simulation methods}
+\label{ss:simulations}
+
+First we briefly describe the simulations we used for illustration and validation
+(the \texttt{R} code used is provided as a supplement).
+We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
+Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
+locally beneficial alleles were more likely to reproduce.
+Each then produced a random, Poisson distributed number of offspring that each, independently, 
+either remained in the same location with probabilty $0.8$ or
+migrated a random number of steps in a uniformly chosen cardinal direction;
+the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
+\linelabel{rr:sim_details}
+Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
+(Although we described the simulation in terms of individuals,
+we kept track only of total numbers in an equivalent way.)
+
+The reproduction rate in simulations for type $b$ alleles was 30\%; 
+this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
+where the value of $s$ is either $s_m$ or $s_p$ depending on the individual's location.
+This determines the values of $s_m$ and $s_p$ reported in the figures, 
+and do not depend on the basic rate of reproduction.
+However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
+we computed the rate of intrinsic growth, 
+i.e.\ the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
+in the absence of migration.
+(The resulting values are close to the first notion of $s$,
+but give better agreement with theory,
+which uses the second definition.)
+
+To sample lineages, we first simulated forwards in time the population dynamics,
+then sampled lineages moving back through time
+by, in each generation,
+moving each lineage to a new deme with probability proportional to the reverse migration probability
+weighted by the number of $B$ alleles in that deme in the previous time step.
+If more than one lineage was found in a deme with $n$ $B$ alleles,
+then each lineage picked a label uniformly from $1 \ldots n$,
+and those picking the same label coalesced.
+Since reproduction is Poisson, this correctly samples from the distribution of lineages given the population dynamics.
+
 \subsection[Probability of Establishment]{Numerical calculation of the probability of establishment}
@@
 Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{ss:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
-we obtain that
-\begin{align}
-  \label{eqn:migrate} 
+we obtain that XXX FIX
+\begin{align}
+  \label{eqn:orig_migrate} 
   \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
@@
 
+%%%%%%%
+\section{Parameter values for simulations}
+\label{apx:parameter_tables}
+
+\input{mutation-parameters-table.tex}
+
+\input{migration-parameters-table.tex}
+
 \ifsubmissionversion

### 2014-11-26 analyze-sims.R

@@ -28,20 +28,6 @@ mutsims$adapted <- (
     ( mutsims$time1 >= mutsims$hit100.1 ) )
 
 mutsims$adapttime <- mutsims$hit100.1
-
-# look at residuals
-with( subset(mutsims,adapted), {
-        x <- log(N)
-        plot( x, adapttime-muttime, log='x' ) 
-        points( sort(unique(x)), tapply(adapttime-muttime,factor(x,levels=sort(unique(x))),mean,na.rm=TRUE), col='red', cex=2 )
-        resid.lm <- lm( adapttime-muttime ~ x )
-        abline(coef(resid.lm),untf=TRUE)
-        coef(resid.lm)
-    } )
-
-muttime.lm <- with( subset(mutsims,adapted), lm( adapttime ~ muttime ) )
-muttime.lm
-
 pdf(file="mutation-times-predicted.pdf", width=5, height=5, pointsize=10)
 with( subset(mutsims,adapted), 
         plot( muttime, adapttime, log='xy', xlab="mean time to adaptation by mutation", ylab="time to hit 100 in patch",
@@ -61,17 +47,33 @@ legend("bottomright", legend=c(levels(mutsims$sm.name),levels(mutsims$N.name)), 
 dev.off()
 
 
-pairs( mutsims[c("time1", "hit100.1", "final1", "muttime" )  ], 
-    upper.panel=function (x,y,...) { points(x,y,...,col=mutsims$sm.name) }, 
-    lower.panel=function (x,y,...) { points(x,y,...,col=mutsims$N.name) } )
 
-layout(t(1:2))
-with( mutsims, plot( final1, adapttime, col=sm.name, pch=1+!adapted ) )
-with( subset(mutsims,adapted&N>100), abline(coef(lm( adapttime ~ final1 ) ) ) )
-legend("bottomleft",legend=levels(mutsims$sm.name),col=1:nlevels(mutsims$sm.name),pch=1)
-with( mutsims, plot( final1, adapttime, col=N.name, pch=1+!adapted ) )
-with( subset(mutsims,adapted&N>100), abline(coef(lm( adapttime ~ final1 ) ) ) )
-legend("bottomleft",legend=levels(mutsims$N.name),col=1:nlevels(mutsims$N.name),pch=1)
+# look at residuals
+if (interactive()) {
+    with( subset(mutsims,adapted), {
+        x <- log(N)
+        plot( x, adapttime-muttime, log='x' ) 
[…]

### 2014-11-26 sim-mutation-params-newR.sh

@@ -10,7 +10,7 @@ do
     do
         for MU in "1e-5" "1e-6"
         do
-            THISDIR="mutation/sm${SM}/N-${N}/1D"
+            THISDIR="mutation/mu${MU}/sm${SM}/N-${N}/1D"
             mkdir -p $THISDIR
             Rscript generate-patchy-run.R \
                 "outdir=\"${THISDIR}\"" \

### 2014-11-26 patchy-selection-paper.tex
figs

@@
 or the patch was not large enough to support a stable population of $B$ alleles.
+Supplementary figures \ref{sfig:sim_mutation_times} and \ref{sfig:sim_migration_times}
+show all times, 
+and depictions of typical simulation runs are shown in supplementary figures
+\ref{sfig:} .
+
+The agreement is reasonable for both
 
@@
   \caption{
-      \textbf{(A)} 
+      \textbf{(left)} 
       Time until adaptation by mutation in an isolated patch for different parameter combinations,
       as observed in simulations (vertical axis),
-      and as predicted from equation~\eqref{eqn:mutrate} (as $1/\mutrate$, horizontal axis).
-      \textbf{(B)} 
+      and as predicted from equation~\eqref{eqn:mutrate} (horizontal axis).
+      \textbf{(right)} 
       Time until adaptation by migration from an already occupied patch to a new patch,
       as observed in simulations (vertical axis),
-      and as predicted from equation~\eqref{eqn:migrate} (as $1/\migrate$, horizontal axis).
+      and as predicted from equation~\eqref{eqn:migrate} (horizontal axis).
       For each,
@@
 
+Here we provide additional details regarding the simulations we used
+to validate the theory in figures \ref{fig:sim_times} and \ref{fig:sim_probs}.
+
+% table of parameters, produced by analyze-sims.R
 \input{mutation-parameters-table.tex}
 
+% table of parameters, produced by analyze-sims.R
 \input{migration-parameters-table.tex}
+
+% example trace plots
 

### 2014-11-26 analyze-sims.R

@@ -165,11 +165,11 @@ par(mar=c(4,4,1,1)+.1)
             sm.labs <- parse( text=gsub("sm-","s[m]==",levels(sm.name)) )
             N.vals <- (levels(droplevels(N.name))); N.col.pal <- terrain_hcl(length(N.vals))
             N.cols <- N.col.pal[ match( levels(N.name)[tapply(N.name,paramstring,"[",1)], N.vals ) ]
-            N.labs <- parse( text=gsub("N-","N==",levels(N.name)) )
+            N.labs <- parse( text=gsub("N-","rho ==",levels(N.name)) )
             mu.vals <- (levels(droplevels(mu.name))); mu.col.pal <- sequential_hcl(length(mu.vals))
             mu.pch <- match( levels(mu.name)[tapply(mu.name,paramstring,"[",1)], mu.vals )
             mu.cols <- mu.col.pal[ mu.pch ]
-            mu.labs <- parse( text=gsub("mu-","mu==",levels(mu.name)) )
+            mu.labs <- parse( text=gsub("mu","mu == ",levels(mu.name)) )
             xx <- exp(jitter(log(tapply(muttime,paramstring,mean,na.rm=TRUE))))
             plot( 0, type='n', 
                     log='xy', xlim=c(100,10000),ylim=c(100,10000),
@@ -208,7 +208,7 @@ par(mar=c(4,4,1,1)+.1)
             N.vals <- (levels(droplevels(N.name))); N.col.pal <- terrain_hcl(length(N.vals))
             N.pch <- match( levels(N.name)[tapply(N.name,paramstring,"[",1)], N.vals )
             N.cols <- N.col.pal[ N.pch ]
-            N.labs <- parse( text=gsub("N-","N==",levels(N.name)) )
+            N.labs <- parse( text=gsub("N-","rho ==",levels(N.name)) )
             plot( 0, type='n', log='xy', xlim=c(10,3.5e4), ylim=c(10,3.5e4), 
                     ylab="time to hit 100 in second patch", 
                     xlab=expression(1/lambda['mig']) )
@@ -393,3 +393,29 @@ print.xtable(
 sink(NULL)
 
 
+####
+#  make example trace plots
+
+plotsims <- tapply( 1:nrow(mutsims), mutsims$paramstring, sample, 1 )
+
+for (k in plotsims) {
+    with( mutsims[k,],  {
+        fcon <- pipe(paste("find ../patchy-sim -name", filename))
[…]

### 2014-11-26 analyze-sims.R
.

@@ -396,26 +396,30 @@ sink(NULL)
 ####
 #  make example trace plots
 
-plotsims <- tapply( 1:nrow(mutsims), mutsims$paramstring, sample, 1 )
+plotsims <- with( subset(mutsims,dims=="1D"), tapply( seq_along(paramstring), droplevels(paramstring), sample, 1 ) )
 
 for (k in plotsims) {
+
     with( mutsims[k,],  {
         fcon <- pipe(paste("find ../patchy-sim -name", filename))
         fname <- scan(fcon,what='char')
         close(fcon)
-        oname <- gsub(".Rdata",".pdf",filename)
         cat("reading from ", fname, "\n")
         load(fname)
-        cat("writing to ", oname, "\n")
-        pdf(file=oname,width=7,height=3,pointsize=10)
+        pdf(file=gsub(".Rdata",".pdf",filename),width=7,height=3,pointsize=10)
+        # png(file=gsub(".Rdata",".png",filename),width=7*144,height=3*144,pointsize=10,res=144)
         layout(t(1:2))
-        plotpophist(pophist,plotlegend=FALSE)
+        plotpophist(pophist,plotlegend=FALSE,maxtimes=100,
+            main=with(pophist$pop$params,parse(text=sprintf("list(rho == %d, mu == %0.2e)",N,mu))) )
         freqs <- pophist$pophist[,,2,]/pophist$pop$params$N
-        matplot(freqs[,floor(seq(1,ncol(freqs),length.out=25))],type='l', 
+        xlocs <- floor(seq(1,nrow(freqs),length.out=200))
+        matplot(xlocs, freqs[xlocs,floor(seq(1,ncol(freqs),length.out=25))],type='l', 
             col=heat.colors(25), ylim=c(0,1),
-            ylab="frequency of B", xlab="space (demes)" )
+            ylab="frequency of B", xlab="space (demes)",
+            main=with(pophist$pop,parse(text=sprintf("list(s[m] == %0.2e, s[p] == %0.2e)",abs(params$sm),getgrowth(params)$gb))) )
         ss <- as.vector(pophist$pop$params$s)
-        lines( (ss-min(ss))/diff(range(ss)), col='black', lty=2, lwd=2 )
+        lines( xlocs, ((ss-min(ss))/diff(range(ss)))[xlocs], col='black', lty=2, lwd=2 )
         dev.off()
     } )
+
 }

### 2014-11-26 analyze-sims.R
.

@@ -400,7 +400,7 @@ plotsims <- with( subset(mutsims,dims=="1D"), tapply( seq_along(paramstring), dr
 
 for (k in plotsims) {
 
-    with( mutsims[k,],  {
+    with( subset(mutsims,dims=="1D")[k,],  {
         fcon <- pipe(paste("find ../patchy-sim -name", filename))
         fname <- scan(fcon,what='char')
         close(fcon)

### 2014-11-26 patchy-selection-paper.tex
example sim plots

@@
 
-The agreement is reasonable for both
+The agreement is reasonable for both.
+The theoretical value $1/\mutrate$ underestimates the mean time to mutation
+by a factor of around 2 that depends increases with $s_m$.
+This is to be expected for two reasons:
+First, we compute the time to reach 100 $B$ alleles, while theory predicts
+the time until the the progenitor of those 100 $B$ alleles arose.
+Second, the expression for $\mutrate$ neglects effects near the boundary of the patch,
+and the larger $s_m$ is, 
+the harder it is for mutations that arise near the edge of the patch to establish.
+The theoretical values $1/\migrate$ again has a margin of error of a factor of about 2,
+with the exception of the weakest disadvantage off the patch, $s_m=10^{-4}$,
+which adapt much more quickly than predicted by theory.
+However, these simulations fall outside the region where we expect the model to do well,
+since $R < \sigma/\sqrt{s_m} \approx 100$.
+A glance at SUPP FIGURE XXX reveals more:
+since selection is weaker than Wright's effective neighborhood size ($s_m < 1/2 \rho \sigma \le 1/8000$), 
+local drift is stronger than selection,
+and local migration-selection balance looks very little like the deterministic equilibirium.
 
@@
       as observed in simulations (vertical axis),
-      and as predicted from equation~\eqref{eqn:migrate} (horizontal axis).
+      and as predicted from equation~\eqref{eqn:migrate}, with $C=5$ (horizontal axis).
       For each,
@@
   }   \label{fig:sim_times}
-\end{figure}
-
-%%%%%%%%%
-\subsection{Comparison to simulation}
-\label{ss:comparison_to_sims}
-
-We used the simulations describe above to test the theory,
-etcetera.
-
-\begin{figure}[ht!]
-  \begin{center}
-      % made in analyze-sims.R
-      \includegraphics{prob-mutation-compared}
-  \end{center}
-  \caption{
-      Probability that two patches separated by distance $R$
-      evolve by independent mutations (i.e.\ in parallel),
-      as predicted by theory (colors, contours)
-      and in simulations (points).
-      \textbf{With $C=5$.}
-      XXX: more info
-  }   \label{fig:sim_probs}
 \end{figure}
@@
 approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{s_m}$. \revpoint{2}{13}
+
+\begin{figure}[ht!]
+  \begin{center}
+      % made in analyze-sims.R
+      \includegraphics{prob-mutation-compared}
+  \end{center}
+  \caption{
+      Probability that two patches separated by distance $R$
+      evolve by independent mutations (i.e.\ in parallel),
+      as predicted by theory (colors, contours)
+      and in simulations (points).
+      \textbf{With $C=5$.}
+      XXX: more info
+  }   \label{fig:sim_probs}
+\end{figure}
 
@@
 
+% sm=0.1
+\begin{figure}
+  \begin{center}
+    \includegraphics{{example-sims/18885-r1-501-sb0.01-sm-0.1-N1200-pophistory-run}.pdf}
+    \includegraphics{{example-sims/56325-r1-501-sb0.01-sm-0.1-N600-pophistory-run}.pdf}
+    \includegraphics{{example-sims/28432-r1-501-sb0.01-sm-0.1-N50-pophistory-run}.pdf}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.1$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left).
+  } \label{sfig:sims_1}
+\end{figure}
+
+% sm=0.01
+\begin{figure}
+  \begin{center}
+    \includegraphics{{example-sims/43099-r1-501-sb0.01-sm-0.01-N1200-pophistory-run}.pdf}
+    \includegraphics{{example-sims/69787-r1-501-sb0.01-sm-0.01-N600-pophistory-run}.pdf}
+    \includegraphics{{example-sims/11821-r1-501-sb0.01-sm-0.01-N50-pophistory-run}.pdf}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.01$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left).
+  } \label{sfig:sims_1}
+\end{figure}
+
+% sm=0.001
+\begin{figure}
+  \begin{center}
+    \includegraphics{{example-sims/59611-r1-501-sb0.01-sm-0.001-N1200-pophistory-run}.pdf}
+    \includegraphics{{example-sims/93713-r1-501-sb0.01-sm-0.001-N600-pophistory-run}.pdf}
+    \includegraphics{{example-sims/29850-r1-501-sb0.01-sm-0.001-N50-pophistory-run}.pdf}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left).
+  } \label{sfig:sims_1}
+\end{figure}
+
+% sm=0.0001
+\begin{figure}
+  \begin{center}
+    \includegraphics{{example-sims/15449-r1-501-sb0.01-sm-1e-04-N1600-pophistory-run}.pdf}
+    \includegraphics{{example-sims/5582-r1-501-sb0.01-sm-1e-04-N400-pophistory-run}.pdf}
+    \includegraphics{{example-sims/24639-r1-501-sb0.01-sm-1e-04-N25-pophistory-run}.pdf}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left).
+  } \label{sfig:sims_1}
+\end{figure}
+
 \ifsubmissionversion

### 2014-11-26 analyze-sims.R

@@ -205,7 +205,7 @@ par(mar=c(4,4,1,1)+.1)
             R.pch <- match( levels(R.name)[tapply(R.name,paramstring,"[",1)], R.vals )
             R.cols <- R.col.pal[ R.pch ]
             R.labs <- parse( text=gsub("R-","R==",levels(R.name)) )
-            N.vals <- (levels(droplevels(N.name))); N.col.pal <- terrain_hcl(length(N.vals))
+            N.vals <- (levels(droplevels(N.name))); N.col.pal <- sequential_hcl(length(N.vals))
             N.pch <- match( levels(N.name)[tapply(N.name,paramstring,"[",1)], N.vals )
             N.cols <- N.col.pal[ N.pch ]
             N.labs <- parse( text=gsub("N-","rho ==",levels(N.name)) )
@@ -396,17 +396,16 @@ sink(NULL)
 ####
 #  make example trace plots
 
+# mutation
 plotsims <- with( subset(mutsims,dims=="1D"), tapply( seq_along(paramstring), droplevels(paramstring), sample, 1 ) )
-
 for (k in plotsims) {
-
     with( subset(mutsims,dims=="1D")[k,],  {
         fcon <- pipe(paste("find ../patchy-sim -name", filename))
         fname <- scan(fcon,what='char')
         close(fcon)
         cat("reading from ", fname, "\n")
         load(fname)
-        pdf(file=gsub(".Rdata",".pdf",filename),width=7,height=3,pointsize=10)
+        pdf(file=file.path("example-mutation-sims",gsub("_Rdata",".pdf",gsub(".","_",filename,fixed=TRUE))), width=7,height=3,pointsize=10)
         # png(file=gsub(".Rdata",".png",filename),width=7*144,height=3*144,pointsize=10,res=144)
         layout(t(1:2))
         plotpophist(pophist,plotlegend=FALSE,maxtimes=100,
@@ -421,5 +420,30 @@ for (k in plotsims) {
         lines( xlocs, ((ss-min(ss))/diff(range(ss)))[xlocs], col='black', lty=2, lwd=2 )
         dev.off()
     } )
+}
 
+# migration
+plotsims <- with( subset(migsims,dims=="1D"), tapply( seq_along(paramstring), droplevels(paramstring), sample, 1 ) )
+for (k in plotsims) {
+    with( subset(mutsims,dims=="1D")[k,],  {
[…]

### 2014-11-26 patchy-selection-paper.tex
.

@@
   \begin{center}
-    \includegraphics{{example-sims/18885-r1-501-sb0.01-sm-0.1-N1200-pophistory-run}.pdf}
-    \includegraphics{{example-sims/56325-r1-501-sb0.01-sm-0.1-N600-pophistory-run}.pdf}
-    \includegraphics{{example-sims/28432-r1-501-sb0.01-sm-0.1-N50-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/18885-r1-501-sb0.01-sm-0.1-N1200-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/56325-r1-501-sb0.01-sm-0.1-N600-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/28432-r1-501-sb0.01-sm-0.1-N50-pophistory-run}.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics{{example-sims/43099-r1-501-sb0.01-sm-0.01-N1200-pophistory-run}.pdf}
-    \includegraphics{{example-sims/69787-r1-501-sb0.01-sm-0.01-N600-pophistory-run}.pdf}
-    \includegraphics{{example-sims/11821-r1-501-sb0.01-sm-0.01-N50-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/43099-r1-501-sb0.01-sm-0.01-N1200-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/69787-r1-501-sb0.01-sm-0.01-N600-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/11821-r1-501-sb0.01-sm-0.01-N50-pophistory-run}.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics{{example-sims/59611-r1-501-sb0.01-sm-0.001-N1200-pophistory-run}.pdf}
-    \includegraphics{{example-sims/93713-r1-501-sb0.01-sm-0.001-N600-pophistory-run}.pdf}
-    \includegraphics{{example-sims/29850-r1-501-sb0.01-sm-0.001-N50-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/59611-r1-501-sb0.01-sm-0.001-N1200-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/93713-r1-501-sb0.01-sm-0.001-N600-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/29850-r1-501-sb0.01-sm-0.001-N50-pophistory-run}.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics{{example-sims/15449-r1-501-sb0.01-sm-1e-04-N1600-pophistory-run}.pdf}
-    \includegraphics{{example-sims/5582-r1-501-sb0.01-sm-1e-04-N400-pophistory-run}.pdf}
-    \includegraphics{{example-sims/24639-r1-501-sb0.01-sm-1e-04-N25-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/15449-r1-501-sb0.01-sm-1e-04-N1600-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/5582-r1-501-sb0.01-sm-1e-04-N400-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/24639-r1-501-sb0.01-sm-1e-04-N25-pophistory-run}.pdf}
   \end{center}

### 2014-11-26 analyze-sims.R
.

@@ -397,53 +397,60 @@ sink(NULL)
 #  make example trace plots
 
 # mutation
-plotsims <- with( subset(mutsims,dims=="1D"), tapply( seq_along(paramstring), droplevels(paramstring), sample, 1 ) )
-for (k in plotsims) {
-    with( subset(mutsims,dims=="1D")[k,],  {
-        fcon <- pipe(paste("find ../patchy-sim -name", filename))
-        fname <- scan(fcon,what='char')
-        close(fcon)
-        cat("reading from ", fname, "\n")
-        load(fname)
-        pdf(file=file.path("example-mutation-sims",gsub("_Rdata",".pdf",gsub(".","_",filename,fixed=TRUE))), width=7,height=3,pointsize=10)
-        # png(file=gsub(".Rdata",".png",filename),width=7*144,height=3*144,pointsize=10,res=144)
-        layout(t(1:2))
-        plotpophist(pophist,plotlegend=FALSE,maxtimes=100,
-            main=with(pophist$pop$params,parse(text=sprintf("list(rho == %d, mu == %0.2e)",N,mu))) )
-        freqs <- pophist$pophist[,,2,]/pophist$pop$params$N
-        xlocs <- floor(seq(1,nrow(freqs),length.out=200))
-        matplot(xlocs, freqs[xlocs,floor(seq(1,ncol(freqs),length.out=25))],type='l', 
-            col=heat.colors(25), ylim=c(0,1),
-            ylab="frequency of B", xlab="space (demes)",
-            main=with(pophist$pop,parse(text=sprintf("list(s[m] == %0.2e, s[p] == %0.2e)",abs(params$sm),getgrowth(params)$gb))) )
-        ss <- as.vector(pophist$pop$params$s)
-        lines( xlocs, ((ss-min(ss))/diff(range(ss)))[xlocs], col='black', lty=2, lwd=2 )
-        dev.off()
-    } )
+if (!file.exists("example-mutation-sims")) {
+    dir.create("example-mutation-sims")
+    plotsims <- with( subset(mutsims,dims=="1D"), tapply( seq_along(paramstring), droplevels(paramstring), sample, 1 ) )
+    for (k in plotsims) {
+        with( subset(mutsims,dims=="1D")[k,],  {
+            fcon <- pipe(paste("find ../patchy-sim -name", filename))
+            fname <- scan(fcon,what='char')
+            close(fcon)
+            cat("reading from ", fname, "\n")
+            load(fname)
[…]

### 2014-11-26 patchy-selection-paper.tex
.

@@
   \begin{center}
-    \includegraphics{{example-mutation-sims/43099-r1-501-sb0.01-sm-0.01-N1200-pophistory-run}.pdf}
-    \includegraphics{{example-mutation-sims/69787-r1-501-sb0.01-sm-0.01-N600-pophistory-run}.pdf}
-    \includegraphics{{example-mutation-sims/11821-r1-501-sb0.01-sm-0.01-N50-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/43099-r1-501-sb0_01-sm-0_01-N1200-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/69787-r1-501-sb0_01-sm-0_01-N600-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/11821-r1-501-sb0_01-sm-0_01-N50-pophistory-run}.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics{{example-mutation-sims/59611-r1-501-sb0.01-sm-0.001-N1200-pophistory-run}.pdf}
-    \includegraphics{{example-mutation-sims/93713-r1-501-sb0.01-sm-0.001-N600-pophistory-run}.pdf}
-    \includegraphics{{example-mutation-sims/29850-r1-501-sb0.01-sm-0.001-N50-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/59611-r1-501-sb0_01-sm-0_001-N1200-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/93713-r1-501-sb0_01-sm-0_001-N600-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/29850-r1-501-sb0_01-sm-0_001-N50-pophistory-run}.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics{{example-mutation-sims/15449-r1-501-sb0.01-sm-1e-04-N1600-pophistory-run}.pdf}
-    \includegraphics{{example-mutation-sims/5582-r1-501-sb0.01-sm-1e-04-N400-pophistory-run}.pdf}
-    \includegraphics{{example-mutation-sims/24639-r1-501-sb0.01-sm-1e-04-N25-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/15449-r1-501-sb0_01-sm-1e-04-N1600-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/5582-r1-501-sb0_01-sm-1e-04-N400-pophistory-run}.pdf}
+    \includegraphics{{example-mutation-sims/24639-r1-501-sb0_01-sm-1e-04-N25-pophistory-run}.pdf}
   \end{center}

### 2014-11-26 analyze-sims.R

@@ -440,7 +440,7 @@ if (!file.exists("example-migration-sims")) {
             # png(file=gsub(".Rdata",".png",filename),width=7*144,height=3*144,pointsize=10,res=144)
             layout(t(1:2))
             plotpophist(pophist,plotlegend=FALSE,maxtimes=100,
-                main=with(pophist$pop$params,parse(text=sprintf("list(rho == %d, mu == %0.2e)",N,mu))) )
+                main=with(pophist$pop$params,parse(text=sprintf("list(rho == %d, R == %0.2d)",N,R))) )
             freqs <- pophist$pophist[,,2,]/pophist$pop$params$N
             xlocs <- floor(seq(1,nrow(freqs),length.out=200))
             matplot(xlocs, freqs[xlocs,floor(seq(1,ncol(freqs),length.out=25))],type='l', 

### 2014-11-27 patchy-review-responses.tex
more sims

@@
 \begin{minipage}[b]{2.5in}
-    Department of Molecular and Computational Biology\\
+    Molecular and Computational Biology\\
     University of Southern California, \\
-    \emph{and} Department of Evolution and Ecology\\
+    \emph{and} Evolution and Ecology\\
     University of California at Davis\\
@@
   robustness, and identify the parameter ranges for which they hold.
+\end{quote}
 
+We have done a large number of simulations, and compared the results to the theory;
+the main comparison occurs in \ref{fig:sim_times} and figures~\ref{fig:sim_probs};
+we have also added in supplementary section \ref{apx:parameter_tables} 
+visualizations of the simulations at various parameter values
+that we think will help some readers in verifying the assumptions.
+
+\begin{quote}
   (ii) The presentation of the main result needs to be improved. In particular,
@@
 
-Changes of note are:
+We sympathize entirely: we felt the same way ourselves,
+and went around in similar circles for quite some time
+before settling on the argument given in the first draft.
+As suggested by the third reviewer, 
+we have opted to move the most difficult part of the argument to the Materials \& Methods,
+replacing it in the Results with a more \textit{ad hoc} argument
+along the lines of that suggested by the reviewers.
+(We do think the more formal argument is more solid, however,
+and therefore necessary; as discussed below, various seemingly reasonable arguments
+give different answers for subtle reasons.)
 
+Other changes of note are:
 \begin{itemize}
-
-  \item We have done a large number of simulations, and compared the results to the theory;
-    see figures~\ref{fig:sim_probs} and \ref{fig:sim_times}.
 
@@
     we were neglecting the case that $s_m \ll s_p$.
-    This was pur laziness on our part,
+    This omission was pure laziness on our part,
     which we have rectified; see the new equation~\eqref{eqn:twocase_migrate}.
+
+  \item We have also improved our discussion of the limits to the theory.
 

### 2014-11-27 patchy-selection-paper.tex
more sims

--- 2014-11-26 patchy-selection-paper.tex
@@
 %%%%%%%%%
-\subsection{What are we assuming?}
+\subsection{A review, and a test, of the assumptions}
 \label{ss:assumptions}
@@
 Next, in using equation \eqref{eqn:eqfreq}, we assume that 
-\textbf{(b)} the inter-patch distance is at least several multiples of the dispersal distance
-($R \gg \sigma$),
+\textbf{(b)} the inter-patch distance is at least the characteristic length
+($R > \sigma/\sqrt{s_m}$),
 and that
 \textbf{(c)} local drift is not so strong that the equilibirum frequency is at least approximately attained
-($s_m \gg 1/(\rho \sigma^d)$).
+(using Wright's effective neighborhood size, $s_m \gg 1/(\rho \sigma^d)$).
+The last requirement is necessary because if the $B$ allele fixes in large neighborhoods where it is deleterious,
+we cannot approximate its dynamics via a branching process.
 We also neglect the time for migration-selection equilibrium to be reached,
-which has the effect of over-estimating $\migrate$.
-
-We also assume that migration to a new patch takes place over a number of generations;
+which has the effect of over-estimating $\migrate$,
+but don't know a good way to estimate this contribution.
+As discussed above, we also assume that migration to a new patch takes place over a number of generations;
 if there are sufficient rare, long-distance migration events
@@
 or the patch was not large enough to support a stable population of $B$ alleles.
-Supplementary figures \ref{sfig:sim_mutation_times} and \ref{sfig:sim_migration_times}
+Supplementary figures \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times} 
 show all times, 
 and depictions of typical simulation runs are shown in supplementary figures
-\ref{sfig:} .
+\ref{sfig:sims_1}, \ref{sfig:sims_2}, \ref{sfig:sims_3}, \ref{sfig:sims_4}, \ref{sfig:sims_5}, \ref{sfig:sims_6}, \ref{sfig:sims_7}, and \ref{sfig:sims_8}.
 
@@
 since $R < \sigma/\sqrt{s_m} \approx 100$.
-A glance at SUPP FIGURE XXX reveals more:
-since selection is weaker than Wright's effective neighborhood size ($s_m < 1/2 \rho \sigma \le 1/8000$), 
-local drift is stronger than selection,
-and local migration-selection balance looks very little like the deterministic equilibirium.
+A glance at supplementary figure \ref{sfig:sims_8} shows more:
+since $s_m$ is so small, the second patch is reached by migrants before the migration-selection eqilibrium is reached,
+and at migration-selection equilibrium, the two patches act as a single meta-patch.
 
@@
 
+Let $\gamma = \min(1,s_m/p_e)$.
 An important point to notice is that effective migration and mutation rates will be on the same order if 
-$A \mu \approx 2 s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
+$A \mu \approx 2 \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
 where $R$ is the distance between the patches,
@@
 In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
-is $2 \rho s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
+is $2 \rho \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
 which depends fairly strongly on the selective disadvantage $s_m$ between patches.
@@
 Equivalently, the rates are roughly equal if
-$R/\sigma = \log(2 s_m/(A \mu))/\sqrt{2s_m}$,
+$R/\sigma = \log(2 \gamma s_m/(A \mu))/\sqrt{2s_m}$,
 which gives the critical gap size past which adaptation will be mostly parallel
@@
 
-If we take the mutational influx in a patch to be one per 10,000 generations ($\rho A \mu = .0001$)
+If we take the mutational influx in a patch to be one per 10,000 generations ($\rho A \mu = .0001$),
+$s_m>p_e$ so that $\gamma=1$,
 and the population density to be 1 ($\rho =1$, so $-\log(A\mu)\approx 9.2$)
@@
 Both these rates scale linearly with population density ($\rho$) 
-and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$),
+and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$, for $p_e>s_m$),
 so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
@@
 Furthermore, the probability that the second adaptation is a new mutation,
-i.e.\ the probability of parallel adaptation, is 
+i.e.\ the probability of parallel adaptation, 
+with $\gamma = \min(1,s_m/p_e)$,
+is 
 \begin{equation} \label{eqn:parallel_prob}
-    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
+    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
 \end{equation}
@@
 \subsection*{Acknowledgements}
-We would like to heartily thank Joachim Hermisson for a careful reading of the paper,
+We would like to heartily thank Joachim Hermisson for a number of insightful suggestions,
 and catching the algebra error present in the first version of this paper on the bioRxiv,
-and also to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
+as well as the two anonymous reviewers.
+We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
 as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
@@
 %%%%%%%
-\section{Parameter values for simulations}
+\section{More information about the simulations}
 \label{apx:parameter_tables}
@@
 \input{migration-parameters-table.tex}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{mutation-times-predicted}
+  \end{center}
+  \caption{
+    The same data shown in the left panel of figure~\ref{fig:sim_times},
+    but all times shown (not just the interquartile ranges),
+    and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
+  } \label{sfig:sim_migration_times}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{migration-time-predicted}
+  \end{center}
+  \caption{
+    The same data shown in the right panel of figure~\ref{fig:sim_times},
+    but all times shown (not just the interquartile ranges),
+    and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
+    The left panel has the predicted time to adaptation on the horizontal axis as in figure~\ref{fig:sim_times},
+    and the right panel has, for comparison, the raw distance between patches 
+    (which predicts time to adaptation, but not as well).
+  } \label{sfig:sim_mutation_times}
+\end{figure}
 
@@
   \begin{center}
-    \includegraphics{{example-sims/18885-r1-501-sb0.01-sm-0.1-N1200-pophistory-run}.pdf}
-    \includegraphics{{example-sims/56325-r1-501-sb0.01-sm-0.1-N600-pophistory-run}.pdf}
-    \includegraphics{{example-sims/28432-r1-501-sb0.01-sm-0.1-N50-pophistory-run}.pdf}
+    \includegraphics{example-mutation-sims/18885-r1-501-sb0_01-sm-0_1-N1200-pophistory-run}
+    \includegraphics{example-mutation-sims/56325-r1-501-sb0_01-sm-0_1-N600-pophistory-run}
+    \includegraphics{example-mutation-sims/28432-r1-501-sb0_01-sm-0_1-N50-pophistory-run}
   \end{center}
@@
     and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left).
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
   } \label{sfig:sims_1}
@@
   \begin{center}
-    \includegraphics{{example-sims/43099-r1-501-sb0.01-sm-0.01-N1200-pophistory-run}.pdf}
-    \includegraphics{{example-sims/69787-r1-501-sb0.01-sm-0.01-N600-pophistory-run}.pdf}
-    \includegraphics{{example-sims/11821-r1-501-sb0.01-sm-0.01-N50-pophistory-run}.pdf}
+    \includegraphics{example-mutation-sims/43099-r1-501-sb0_01-sm-0_01-N1200-pophistory-run}
+    \includegraphics{example-mutation-sims/69787-r1-501-sb0_01-sm-0_01-N600-pophistory-run}
+    \includegraphics{example-mutation-sims/11821-r1-501-sb0_01-sm-0_01-N50-pophistory-run}
   \end{center}
@@
     and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left).
-  } \label{sfig:sims_1}
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_2}
 \end{figure}
@@
   \begin{center}
-    \includegraphics{{example-sims/59611-r1-501-sb0.01-sm-0.001-N1200-pophistory-run}.pdf}
-    \includegraphics{{example-sims/93713-r1-501-sb0.01-sm-0.001-N600-pophistory-run}.pdf}
-    \includegraphics{{example-sims/29850-r1-501-sb0.01-sm-0.001-N50-pophistory-run}.pdf}
+    \includegraphics{example-mutation-sims/59611-r1-501-sb0_01-sm-0_001-N1200-pophistory-run}
+    \includegraphics{example-mutation-sims/93713-r1-501-sb0_01-sm-0_001-N600-pophistory-run}
+    \includegraphics{example-mutation-sims/29850-r1-501-sb0_01-sm-0_001-N50-pophistory-run}
   \end{center}
@@
     and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left).
-  } \label{sfig:sims_1}
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_3}
 \end{figure}
@@
   \begin{center}
-    \includegraphics{{example-sims/15449-r1-501-sb0.01-sm-1e-04-N1600-pophistory-run}.pdf}
-    \includegraphics{{example-sims/5582-r1-501-sb0.01-sm-1e-04-N400-pophistory-run}.pdf}
-    \includegraphics{{example-sims/24639-r1-501-sb0.01-sm-1e-04-N25-pophistory-run}.pdf}
+    \includegraphics{example-mutation-sims/15449-r1-501-sb0_01-sm-1e-04-N1600-pophistory-run}
+    \includegraphics{example-mutation-sims/5582-r1-501-sb0_01-sm-1e-04-N400-pophistory-run}
+    \includegraphics{example-mutation-sims/24639-r1-501-sb0_01-sm-1e-04-N25-pophistory-run}
   \end{center}
@@
     and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left).
-  } \label{sfig:sims_1}
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_4}
+\end{figure}
+
+% sm=0.1, R=80
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-migration-sims/89826-r1-501-sb0_01-sm-0_1-N4000-pophistory-run}
+    \includegraphics{example-migration-sims/76178-r1-501-sb0_01-sm-0_1-N1000-pophistory-run}
+    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.1$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_5}
+\end{figure}
+
+% sm=0.01, R=80
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-migration-sims/42080-r1-501-sb0_01-sm-0_01-N4000-pophistory-run}
+    \includegraphics{example-migration-sims/82698-r1-501-sb0_01-sm-0_01-N1000-pophistory-run}
+    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.01$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_6}
+\end{figure}
+
+% sm=0.001, R=80
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-migration-sims/88079-r1-501-sb0_01-sm-0_001-N4000-pophistory-run}
+    \includegraphics{example-migration-sims/3464-r1-501-sb0_01-sm-0_001-N1000-pophistory-run}
+    \includegraphics{example-migration-sims/37774-r1-501-sb0_01-sm-0_001-N100-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_7}
+\end{figure}
+
+% sm=0.0001, R=80
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-migration-sims/26044-r1-501-sb0_01-sm-1e-04-N4000-pophistory-run}
+    \includegraphics{example-migration-sims/85750-r1-501-sb0_01-sm-1e-04-N1000-pophistory-run}
+    \includegraphics{example-migration-sims/4111-r1-501-sb0_01-sm-1e-04-N100-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.0001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_8}
 \end{figure}
@@
 
+%%%% Reviews
 \includereviews

### 2014-11-27 analyze-sims.R

@@ -268,6 +268,11 @@ simgrid <- cbind( simgrid, t( sapply( 1:nrow(simgrid), function (k) { params <- 
             }
             c( migtime=median(migtimes), muttime=median(muttimes), pmut=pmut )
         } ) ) )
+simgrid$valid <- with(simgrid, 
+        ( R*sqrt(abs(sm))/sigma > 1 ) & 
+        ( A > (sigma/sqrt(abs(sm)))*atan(sqrt(abs(sm/gb))) ) &
+        ( abs(sm) * 2 * sigma * N > 1 )
+    )
 
 
 # look at results

### 2014-11-27 sim-migration-params-newsm.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+# Simulate time to migration in a new patch
+SB=".01"
+
+for SM in "-.003" "-.03"
+do
+    for N in 100 1000
+    do
+        for R in 60 100 120 # distance between patches
+        do
+            THISDIR="migration/sm${SM}/R-${R}/N-${N}/1D"
+            mkdir -p $THISDIR
+            Rscript generate-patchy-run.R \
+                "outdir=\"${THISDIR}\"" \
+                "N=${N}" "range=c(1,501)" "sb=${SB}" "sm=${SM}" "nsteps=1000" "stepsize=25" "ntypes=2" \
+                "s=matrix(ifelse(abs(abs(seq.int(1,range[2])-range[2]/2)-(${R}/2+20))<20,sb,sm),nrow=1,ncol=range[2])" \
+                'initdist=array(c(rep(N,prod(range)),rep(0,(ntypes-1)*prod(range))),c(range,ntypes));initdist[(seq.int(1,range[2])<range[2]/2)&s>0]=floor(.8*N)' &
+        done
+    done
+done
+

### 2014-11-27 sim-mutation-params-newsm.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+# Simulate time to mutation in a new patch
+SB=".01"
+SM="-.0001"
+
+for SM in "-.003" "-.03"
+do
+    for N in 100 1000
+    do
+        for MU in "1e-5" "1e-6"
+        do
+            THISDIR="mutation/mu${MU}/sm${SM}/N-${N}/1D"
+            mkdir -p $THISDIR
+            Rscript generate-patchy-run.R \
+                "outdir=\"${THISDIR}\"" \
+                "N=${N}" "range=c(1,501)" "sb=${SB}" "sm=${SM}" "nsteps=1000" "stepsize=25" "ntypes=2" "mu=${MU}"\
+                's=matrix(ifelse((seq.int(1,range[2])>200)&(seq.int(1,range[2])<300),sb,sm),nrow=1,ncol=range[2])' \
+                'initdist=array(c(rep(N,prod(range)),rep(0,(ntypes-1)*prod(range))),c(range,ntypes))' &
+        done
+    done
+done
+

### 2014-11-27 patchy-selection-paper.tex
hwew

@@
 \begin{equation} \label{eqn:parallel_prob}
-    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
+    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + C \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
 \end{equation}
@@
 approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{s_m}$. \revpoint{2}{13}
+
+We tested this using the same simulations as figure \ref{fig:sim_times},
+by using the empirical distributions of the respective times to adaptation
+to estimate the probability,
+for each parameter combination,
+that a new, successful mutation appears 
+before a successful migrant arrives from another, already adapted patch.
+The results are compared to \eqref{eqn:parallel_prob}
+in figure~\ref{fig:sim_probs}.
 
@@
   \caption{
-      Probability that two patches separated by distance $R$
+      Probability that two patches 
       evolve by independent mutations (i.e.\ in parallel),
-      as predicted by theory (colors, contours)
-      and in simulations (points).
-      \textbf{With $C=5$.}
-      XXX: more info
+      as a function of their geographic separation ($R$)
+      and the strength of selection between patches ($s_m$).
+      The background shades of grey and contours
+      give the probability predicted by theory (equation~\eqref{eqn:parallel_prob}, with $C=5$).
+      Points correspond to parameter combinations of the simulations in figure \ref{fig:sim_times},
+      with the shade of grey and the red numeric label indicating the same probability
+      estimated from these simulations (see text for details).
+      The two densities, $\rho=100$ and $\rho=1000$ would lie directly on top of each other, so the $\rho=1000$ points have been shifted vertically.
   }   \label{fig:sim_probs}

### 2014-11-28 patchy-review-responses.tex
.

@@
 As suggested by the third reviewer, 
-we have opted to move the most difficult part of the argument to the Materials \& Methods,
+we have moved the most difficult part of the argument to the Materials \& Methods,
 replacing it in the Results with a more \textit{ad hoc} argument
@@
 (We do think the more formal argument is more solid, however,
-and therefore necessary; as discussed below, various seemingly reasonable arguments
+and therefore necessary to keep; 
+as discussed below, various seemingly reasonable arguments
 give different answers for subtle reasons.)
@@
 
-  \item We have also improved our discussion of the limits to the theory.
+  \item We have also improved our discussion of the limits to the theory,
+    mostly in the (short) section \nameref{ss:assumptions}.
+    The main point of note was that, indeed as Reviewer 3 pointed out,
+    we were running tests in regions where $s_m$ was too small for the theory to apply,
+    resulting in $\migrate$ that was not monotonic with $s_m$.
+    Happily, the natural restriction of $R \sqrt{s_m}/\sigma > 1$ coincides with the condition
+    for $\migrate$ to decrease with $s_m$;
+    we don't have good intuition for why the formula should increase in the region where it is inapplicable,
+    but this seems beside the point.
 
@@
 \point{}{
-    Am I right that the basic argument is the
-    following:
+    Am I right that the basic argument is the following:
 
@@
 \reply
-    See the response to point 1.1 above.
+    The correspondence in \eqref{eqn:ratio_K} boils down to \eqref{eqn:boils_downto};
+    but rearranging this to $u(x,S) \approx h(x,S) \E[K] g(A)$,
+    this says that the mean occupation density in $S$ due to a family beginning at $x$
+    is the mean number of lineages hitting $S$ ($h(x,S) \E[K]$)
+    multiplied by the mean occupation density of a single lineage starting on the edge of the patch.
+    So, it seems that the missing length (or area) enters through $g(A)$,
+    which in our case is $1/(2s_m)$.
+    We agree, it is not obvious how the units work out here.
 
@@
 \reply
-This was already discussed in the discussion, but we've added a note about it earlier on, citing Hallatschek \& Fisher.
+This was already discussed in the discussion, but we've added a note about it earlier on, in particular adding a citation to the nice paper of Hallatschek \& Fisher.
 \llname{rr:longtailed}
@@
 \point{in the discussion on page 10 "probability of parallel adaptation.."}{
-    it should be noted that many of the scenarios at high mutation influx are quite
+    It should be noted that many of the scenarios at high mutation influx are quite
     hypothetical. While the colonization of migrants might dominate over local
@@
 }
+
+\reply
+We don't know of even a poor approximation to the time to reach migration-selection equilibrium,
+unfortunately.
+However, we have pointed out in \nameref{ss:assumptions} that we're missing this point,
+and discuss there how the time lag to reach this equilbrium contributes to discrepancies between theory and simulations.
 
@@
 \reply
-This argument falls in the trap of treating each migrant individual as independent,
-ignoring the correlations of arrival times induced by families;
-see the discussion in which we now set this up as a straw man in XXX.
+This argument implicitly treats each migrant individual as independent,
+ignoring the correlations of arrival times induced by families.
+This is the reason for explicitly defining families to be those who share an ancestor living outside of $r_0$
+more recently than one living inside the patch;
+many of the particles that Fick's law would pick up leaving $r_0$ would be already counted as part of a family.
+However, we think our simplified argument in \nameref{ss:heuristics} gets the same spirit.
 
@@
 \reply
-In this argument, the reviewer is taking into account clumpiness for establishment of migrants,
-but still not for the time interval
-between arrivals of families.
+Good point about $s_m \ll s_p$ (also made by reviewer \#1);
+we've modified the key equation \eqref{eqn:migrate}
+and surrounding discussion along these lines.
 
@@
 }
+
+\reply
+See the response to reviwer 1's point 3 above.
 
@@
 
+\reply
+As discussed above, we've done this.
+
 %%%%%%%%
@@
 We've tried to say in words what each symbol is when we use it as much as possible,
-so we hope this isn't necessary.
+so we hope this isn't still necessary.
 
@@
      where the analysis makes the most sense. Then the expressions become
-     simpler and you can treat d = 2 as well. If you want, you could even
+     simpler and you can treat $d = 2$ as well. If you want, you could even
      lead with the intuitive explanation at the end of the section, and save the
@@
      I think (24) follows directly from Barton, Heredity 1979. (If you look at
-     the unnumbered equation above Fig. 1 in that paper and plug in $x \sim a$,
+     the unnumbered equation above Fig.\ 1 in that paper and plug in $x \sim a$,
      you find the time scale $r_{eff}$.) I think that the displayed equation should just
@@
 \reply
+This is a very interesting point,
+and the basic tool of a diffusion approximation seems to be the same in both.
+However, something is not matching up,
+because \citet{barton1979geneflow} finds that the length of the region of differentiated genome 
+decays with $1/\sqrt{T}$,
+whereas with our approximation the length of the differentiated haplotype decays as $1/T$.
+The difference may be that Barton considers LD and we consider a haplotype unbroken by recombination.
+On the balance, and considering the other reviewer's wise comments,
+we've cut most of this from \nameref{ss:patch_haplotype}.
 

### 2014-11-28 patchy-selection-paper.tex
.

@@
   Superimposed in color is the decay predicted by \eqref{eqn:eqfreq},
-  with $C$ chosen arbitrarily to give a reasonable fit.
-  \plr{what is $C$?}
+  with $C$ chosen arbitrarily to give a reasonable fit: 0.85 and 0.11 respectively.
   (The disagreement at long distances is due to boundary effects near the edge of the range.)
@@
   demes had 1000 individuals in each;
-  dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
+  dispersal was to nearby demes with mean dispersal distances of 
+  $\sigma = 0.95$ and 0.74 times the deme spacing, respectively.
    }   \label{fig:sim_occupation_freqs}
@@
     \begin{split}
-        q(x) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $x$ ) } ,
+        \rho q(x) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $x$ ) } ,
 \end{split}
@@
 
-A quick review of the parameter regime we are studying here:
-First, we assume that 
+Now that we have expressions for the mean rates of adaptation by new mutation, \eqref{eqn:mutrate},
+and by migration from an already colonized patch, \eqref{eqn:migrate},
+it seems helpful to step back and review the assumptions we have made and the parameter regime
+we expect the results to apply to.
+For the mutation rate,
+our results will hold best if $s_m \approx s_p$,
+if the patch diameter is large relative to $\sigma/\sqrt{\min(s_m,s_p)}$,
+the patch is not too far from circular.
+If $s_m \ll s_p$ then \eqref{eqn:mutrate} will underestimate the rate,
+and conversely if $s_m \gg s_p$.
+As for the migration rate,
+we assume that 
 \textbf{(a)} each patch is large enough to support a stable population of $B$ alleles
@@
 We also neglect the time for migration-selection equilibrium to be reached,
-which has the effect of over-estimating $\migrate$,
-but don't know a good way to estimate this contribution.
+which has the effect of over-estimating $\migrate$.
 As discussed above, we also assume that migration to a new patch takes place over a number of generations;
@@
       For each,
-      parameter combinations and numbers of simulations are in supplemental tables \ref{stab:migration_params} and \ref{stab:mutation_params};
+      parameter combinations and numbers of simulations are in supplemental tables \ref{stab:mutation_params} and \ref{stab:migration_params};
       shown are the median (points) and interquartile range (lines)
@@
 % \end{align}
+
+%%%%%%% %%%%%%%%%%
+\subsection[Haplotypes Shared Within a Patch]{Length of the shared haplotype within a patch}
+\label{ss:patch_haplotype}
+
+The other key issue in thinking about the population genetic signal of adaptation in spatial patches 
+is the process by which haplotypes are whittled down within patches, 
+which plays into a number of aspects of haplotype sharing. 
+The initial haplotype that sweeps within a patch 
+will be dispersed over time by recombination.
+Likewise, the haplotype that is shared between patches coadapted by migration
+(equation~\ref{eqn:haplotype_length}) will also break down. 
+However, a long time after the initial sweep,
+we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
+than with individuals elsewhere,
+since selection against migrants 
+decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}.
+
+Finally we can consider the expected length of haplotype shared within a patch. 
+To do this, we compare the rate of recombination between $B$ and $b$ haplotypes (equation~\ref{eqn:rec_rate_within})
+to the rate of coalescence of $B$ haplotypes within a patch. 
+We will assume that mixing within the patch is fast compared to the rate of drift, 
+so that the rate of coalescence of $B$ alleles within a patch is 
+$\xi^2/(A\rho)$,
+the ``variance effective population size'' of the patch
+(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1975pain,barton2002neutral}).
+The rate of recombination and coalescence will be on the same order if
+$\xi^2 /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$.
+Hence the genetic length of the haplotype shared between pairs of
+individuals within the same patches at equilibrium is on the order of
+\begin{align} \label{eqn:within_haplotype_length} 
+  r = \frac{\sqrt{s} \xi^2}{\sigma \rho A^{(d-1)/d}}  .
+\end{align}
+This does not depend on $A$ in one dimension,
+but decreases with $\sqrt{A}$ in two dimensions.
+In applications, this can be compared to equation~\eqref{eqn:haplotype_length}
+for the length of shared haplotype between patches.
+
+%%%%%%%%%%
+\subsection{Applications} 
+\label{ss:applications}
+% FOR REFERENCE:
+% On Wed, May 28, 2014 at 10:20 AM, Michael Nachman <mnachman@berkeley.edu> wrote:
+% > Graham:
+% >
+% > The density is probably in the range of 10-40 animals per hectare (10,000
+% > m^s).  We built enclosures of 0.25 hectare and found that they could support
+% > about 10 animals with food provided ad libitum, so I think this is an upper
+% > density (i.e.\ 40/hectare).  As for dispersal distance, I'm less sure.
+% > Studies on closely related species (genus Perognathus) in similar habitats
+% > suggest distances of 100's of feet.  Attached are some old papers that tried
+% > to estimate dispersal distances in Perognathus.
+% > best,
+% > Michael
+
+Coat color in the rock pocket mouse
+\emph{Chaetodipus intermedius} is a classic example of local
+adaptation 
+\citep{benson1933concealing,DiceBlossom1937}.
+These mice live on rocky outcrops throughout the southwestern United States and northern Mexico, 
+and generally have a light pelage similar to the predominant rock color.
+However, in some regions these mice live on darker substrates (e.g.\ old lava flows),
+and in these patches have much darker pigmentation, 
+presumably to avoid visual predators.
+\citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
+much of the change in coloration is due to an allele at the MC1R locus.
+This dark allele differs from the light allele by four amino acid changes, 
+and has a dominant or partially dominant effect depending on the measure of coat color.
+These areas of dark rock are separated from each other by a range of distances, 
+with some being separated from any other by hundreds of kilometers of light colored substrate. 
+The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
+suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
+However, \citet{hoekstra2004local} reasoned that, elsewhere in the range, 
+multiple close dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
+despite intervening light habitat. 
+We apply our theory to this example,
+in particular investigating how the probability of parallel mutation depends on 
+the distance between dark outcrops.
+
+A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. 
+\citet{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
+at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
+On the lava flow the dark MC1R allele is at 86\% frequency,
+while at a site with light substrate 12km to the west (Tule), the frequency is 3\%.
+The dark allele was entirely absent from Christmas pass, a site with light substrate 7km north of Tule, and 3km further from the lava flow.
+In the other direction, the dark MC1R allele was at 34\% at a site with light substrate 10km to the east of the flow (O'Neill).
+These numbers give us a sense of a plausible range of characteristic lengths.
+Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
+we can fit formula \eqref{eqn:eqfreq} to these frequencies
+(similar to \citet{hoekstra2004ecological}).
+Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
+and for O'Neill, $\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. 
+We also need estimate of $s_m$. 
+Using $\sigma \approx 1$km \citep{french1968dispersal,allred1963range}, 
+these cline widths imply that $s_m=1/9$ and $1/900$ are reasonable values.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{Lava_flow_mice_prob_parallel}
+  \end{center}
+  \caption{
+Top panel: The probability of parallel mutation as a function of the distance
+between environmental patches for two different cline widths
+and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
+The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
+see text. Bottom panel: The initial genetic length of the haplotype
+shared between patches due to adaptation via migration as a function
+of the distance between environmental patches. 
+  \label{fig:mice_prob_parallel}
+  }
+\end{figure}
+
+The mutational target size $\mu$ for the trait is unclear. 
+While the Pinacate dark haplotype differs from the light haplotype at four amino acid residues,
+it is likely that not all of these changes are needed for a population to begin to adapt. 
+Also there a number of genes besides MC1R at which adaptive changes affecting pigmentation 
+have been identified in closely related species and more broadly across vertebrates \citep{Hoekstra:06}.
+To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
+where the adaptive mutation has to occur at a single base pair; 
+and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation
+being able to arise at any of a thousand base pairs.  
+Finally, we set $A=100\text{km}^2$ (roughly the size of the Pinacate
+patch). In
+Figure~\ref{fig:mice_prob_parallel} we show the dependence of the probability
+of parallel mutation on the distance between lava flow patches using these parameters, 
+showing that parallel mutation should become
+likely over a scale of tens to a few hundred kilometers between patches.
+
+Given the large selection coefficient associated with the dark allele
+on the dark substrate, we expect the initial haplotype associated with
+either a new mutation or migrant allele to be large. 
+We can use equation \eqref{eqn:haplotype_length} to calculate how long the founding
+haplotype shared between populations is expected to be,
+as shown in Figure~\ref{fig:mice_prob_parallel}. 
+The initial length can be quite long between geographically close patches (tens of kilometers). 
+However, for the wider cline width ($\sigma/\sqrt{s_m} = 30$km), 
+adaptation by migration can still be likely for patches 100km apart, 
+but the shared basis may be hard to detect, as the length of shared haplotype can be quite short.
+
+Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
+we should expect the haplotype to be rapidly whittled down within
+patches after establishment (equation \ref{eqn:rec_rate_within}). We
+estimate that the effective recombination rate is no less than 50\% of the actual recombination rate.
+Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare 
+(M.~Nachman, personal communication),
+and the variance in offspring number to be $\xi^2 = 3$ (mostly to get round numbers),
+then the shared haplotype length within patches from equation \eqref{eqn:within_haplotype_length}
+is only expected to be on the order of hundreds or thousands of bases
+($10^{-3}$--$10^{-4}$cM).
+This suggests that unless the migration has happened fairly recently
+then it will be difficult to identify shared haplotypes
+(how recently depends on the background level of LD in the population).
+
+%%%%%%% %%%%%%% %%%%%%%
+\section*{Discussion} 
+\label{ss:discussion}
+
+This paper is an investigation into the basic question: 
+What is the spatial resolution of convergent local adaptation?
+In other words, 
+over what spatial scale of environmental patchiness will the process
+of adaptation develop independent solutions to evolutionary problems?
+The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
+the dispersal distance divided by the square root of the strength of selection against the allele between patches. 
+It depends much more weakly on the
+selective benefit within the patches or (perhaps surprisingly) the population density, 
+although these two factors will determine the time-scale over which adaptation will occur
+(and note that population density could affect the selection coefficients). \linelabel{rr:popdensity}
+This is in contrast with models of panmictic populations
+\citep{softsweepsI, MesserPetrov, Wilson-softsweep} 
+and geographically spread populations adapting to homogeneous selection pressures \citep{ralph2010parallel}, 
+where the probability of multiple, independently arising adaptive alleles increases with the population size/density. 
+However, in all of these models the dependence on the beneficial selection
+coefficient is absent or weak, due to the fact that selection both
+aids establishment of new alleles and the spread of existing
+alleles (but see \citet{Wilson-softsweep} for the complications of varying population sizes).
+
+We have also shown that
+while weaker selection against alleles will make sharing 
+of adaptations between patches easier, 
+it will also makes it harder to spot the sharing,
+since the lucky alleles that manage to colonize new patches move slower,
+and thus carry a shorter shared haplotype.
+This issue is amplified by the fact that the length of haplotype
+shared within patches decays over time, potentially making the
+identification of shared adaptations to old selection pressures difficult.
+
+Perhaps the most useful rule-of-thumb quantities we found were the following.
+The effective rate of migration into an as-yet-unadapted patch from an already-adapted patch distance $R$ away 
+-- the analogue of the mutational influx $\mu \rho A$ --
+is $2 \rho s_m \exp(- R \sqrt{2 s_m} / \sigma)$.
+Equivalently, the critical gap size between patches past which adaptation is likely independent
+is $R = (\sigma/\sqrt{2s_m}) \log(2s_m/(A \mu))$.
+Finally, successfully transiting migrant lineages move at rate $\sigma \sqrt{2s_m}$,
+and so shared haplotype lengths between patches will be of order $R/(\sigma \sqrt{2s_m})$.
+
+In developing the set of approximations in the paper we have ignored a
+number of complicating factors. We now briefly discuss these.
+
+\paragraph{Standing variation} 
+We have focused on relative rates of adaptation,
+since we are thinking of applications in which we know that adaptation has occurred,
+and the question is whether or not adaptations in distinct patches
+have appeared independently or not.
+However, especially absolute if rates of adaptation by new mutation are very low,
+any adaptation that does occur may have to make use of standing variation.
+The case of a panmictic population was studied by \citet{softsweepsI},
+and we study the case of a continuous, spatial population in an upcoming paper.
+If parallelism in local adaptation of the sort we study here is due to standing variation
+rather than new mutation,
+then the dynamics of adaptation should not depend strongly on migration patterns
+(but note that the initial spatial distribution of standing variation may).
+
+\paragraph{Dominance}
+We have mostly ignored the issue of dominance 
+by dealing with essentially haploid models,
+and appealing to the fact that the dynamics we study occur where the mutation is rare,
+and hence mostly present only in heterozygotes. 
+Our results should hold as a good approximation to dominant and partially dominant alleles
+(with $s_m$ the selection against heterozygotes).
+If, however, the mutation is recessive, then it is essentially neutral where rare,
+and so would encounter much less resistance to spreading between patches.
+The shape of the cline obtained is given by \citet{haldane1948theory}.
+This is counteracted, however, by the increased difficulty with which the mutation would establish
+once arriving in a new patch,
+if the beneficial effect is also recessive. \linelabel{rr:dominance}
+As such it is not clear what our intuition should be 
+about the contribution of recessive alleles to adaptation via migration. 
+Further work is needed to put empirical observations of local adaptation by recessive alleles 
+in a theoretical context.
+It is simliarly unclear how the model should extend to a polygenic trait.
+
+\paragraph{Long distance migration}
+We have also ignored the possibility of very long distance migration,
+instead focusing on local dispersal (and using Gaussians via the central limit theorem).
+However, dispersal distributions can be very heavy tailed, 
+with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
+In addition, over long time-scales, very rare chance events (mice carried off by hurricanes and the like; \citet{censky1998overwater,nathan2008mechanisms})
+could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
+Such tail events could greatly increase the probability of parallel adaptation above that predicted by our model. 
+Furthermore, if adaptive alleles do move between distant patches via rare, long distance migration 
+then they will be associated with a much longer shared haplotype than predicted by equation \eqref{eqn:haplotype_length}. 
+As such, we view our results as a null model by which the contribution of long distribution migrants to adaptation could be empirically judged.
+
+\paragraph{Other geographic models}
+We have studied circular patches of habitat at long distances from each other.
+Real habitat geometry can be much more complex,
+e.g.\ with archipelagos of patches of varying sizes, 
+or patches connected by long, skinny corridors, for instance.
+The work of \citet{cantrell1991diffusive} comes closest to a general theory of balanced polymorphisms in such habitats.
+It is possible that our techniques could be applied in their much more general setting,
+as both are based, fundamentally, on branching process approximations.
+It is also interesting to think about the probability of convergent
+adaptation to continuously varying environments, e.g. replicated
+environmental clines.
+
+\paragraph{Concluding thoughts}
+The falling cost of population genomic sequencing means that we will soon have the
+opportunity to study the interplay of adaptation with geography and ecology 
+across many populations within a species. 
+Our work suggests that even quite geographically close populations 
+may be forced to locally adapt by repeated, convergent, \textit{de novo} mutation 
+when migration is geographically limited
+and selective pressures are divergent.
+Thus, systems where populations have been repeatedly subject to 
+strong local selection pressures may offer the opportunity 
+to study highly replicated convergent adaptation within a similar
+genetic background \citep{stern2013genetic}.  
+Such empirical work will also strongly inform our understanding of the
+ability of gene flow to keep ecologically similar populations evolving
+in concert \citep{sexton2013genetic}.
+Our results suggest that adaptation to shared environments is
+certainly no guarantee of a shared genetic basis to adaptation, 
+suggesting that rapid adaptation to a shared environment could potentially drive
+speciation if the alleles that spread in each population fail to work
+well together \citep{Kondrashov:03}.
+
+% discussion of applications.
+
+% discuss mixing of types; 
+% Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
+
+\subsection*{Acknowledgements}
+We would like to heartily thank Joachim Hermisson for a number of insightful suggestions,
+and catching the algebra error present in the first version of this paper on the bioRxiv,
+as well as the two anonymous reviewers.
+We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
+as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
+for helpful comments and discussion of the manuscript.
+
+%%%%%
+\section*{Materials \& Methods}
+
+%%%%%%%
+\subsection{Simulation methods}
+\label{ss:simulations}
+
+First we briefly describe the simulations we used for illustration and validation
+(the \texttt{R} code used is provided as a supplement).
+We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
+Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
+locally beneficial alleles were more likely to reproduce.
+Each then produced a random, Poisson distributed number of offspring that each, independently, 
+either remained in the same location with probabilty $1-m$ or
+migrated a random number of steps with probability $m$
+in a uniformly chosen cardinal direction;
+for most simulations $m=0.2$ and
+the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
+\linelabel{rr:sim_details}
+Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
+(Although we described the simulation in terms of individuals,
+we kept track only of total numbers in an equivalent way.)
+
+The reproduction rate in simulations for type $b$ alleles was 30\%; 
+this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
+where the value of $s$ is either $s_m$ or $s_p$ depending on the individual's location.
+This determines the values of $s_m$ and $s_p$ reported in the figures, 
+and do not depend on the basic rate of reproduction.
+However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
+we computed the rate of intrinsic growth, 
+i.e.\ the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
+in the absence of migration.
+(The resulting values are close to the first notion of $s$,
+but give better agreement with theory,
+which uses the second definition.)
+
+To sample lineages, we first simulated forwards in time the population dynamics,
+then sampled lineages moving back through time
+by, in each generation,
+moving each lineage to a new deme with probability proportional to the reverse migration probability
+weighted by the number of $B$ alleles in that deme in the previous time step.
+If more than one lineage was found in a deme with $n$ $B$ alleles,
+then each lineage picked a label uniformly from $1 \ldots n$,
+and those picking the same label coalesced.
+Since reproduction is Poisson, this correctly samples from the distribution of lineages given the population dynamics.
+
+\subsection[Probability of Establishment]{Numerical calculation of the probability of establishment}
+\label{apx:establishment_sims}
+
+When rare, under at least some demographic models,
+copies of a new mutant allele are approximately independent and experience a uniform selective benefit;
+and can therefore be treated as a branching process (a spatial branching process in this case).
+Furthermore, whether or not a new, beneficial mutation establishes or is lost to demographic stochasticity
+is determined by this phase where it is rare.
+Fortunately, the probability that a branching process does not establish
+(i.e.\ dies out)
+can be found as a fixed point of the generating function of the process \citep{jagers1975branching}.
+Therefore, we calculated explicitly the generating function for a spatial branching process
+with nearest-neighbor migration on a one-dimensional lattice
+and a Poisson number of offspring with mean $1+s$,
+where $s$ could vary by location,
+and iterated this forward to convergence to obtain $1-p(x)$, the probability a single mutation appearing at $x$ would fail to establish.
+We considered two situations:
+where $s$ is a step function,
+and where it has a linear transition.
+These solutions are shown, and parameters described, in figure~\ref{fig:prob_estab_calcs}.
+
+Here (and at other parameter choices) we see that the probability of establishment $p(x)$ goes to the equilibrium value
+(approximately $p_e = 2s/\xi^2$) within the patch;
+the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
+The fit is equally good for other parameter values,
+even if migration can move further than one deme and offspring numbers are not Poisson (results not shown).
+This lends credence to our approximation that the integral of $p(x)$ over the entire range
+is close to $p_e$ multiplied by the area of the patch.
+
+\begin{figure}[ht!]
+    \begin{center}
+        % made in patchy-selection-plots.R
+        \includegraphics{prob-establishment}
+    \end{center}
+    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
+    in an abrupt transition (left) and a gradual transition (right).
+    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respectively);
+    and the selection coefficient interpolates linearly between these in the central hatched region.
+    The number of offspring is Poisson.
+    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
+    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
+    Note that the approximation used in the text is here $p_e \approx 2 s_p / \xi^2 = 0.1$,
+    agreeing with the asymptotic value on the right.
+    \label{fig:prob_estab_calcs}
+    }
+\end{figure}
+
+%%%%%%%%%%%%
+\subsection[Equilibrium Frequency]{The equilibrium frequency}
+\label{apx:eqfreq}
+
+For completeness, and clarity as to the scalings on the relevant parameters,
+here we provide a derivation of the differential equations referred to above equation~\eqref{eqn:eqfreq},
+and establish the asymptotics given in that equation.
+One route to the ``equilibrium frequency'' of the allele outside the range where it is advantageous is as follows;
+see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} (or \citet{KPP1937book} or \citet{fisher1937wave} or \citet{haldane1948theory})
+for other arguments in equivalent models,
+or \citet{etheridge2000introduction} and/or \citet{dawson1993measurevalued} for general theory used below.
+Following is a description of a discrete model (which is nearly the one we simulate from).
+Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
+and that selection (for or against) the allele is given by the function $s(x)$, with $x$ denoting the spatial location.
+Each individual at location $x$ reproduces after a random lifetime,
+dying, and then leaving behind in the same deme a random number of offspring with distribution given by $X$;
+each offspring then migrates independently to a new location chosen randomly from the distribution given by $x+R$,
+where she replaces a randomly chosen individual.
+If $x+R$ is outside of the range, then she perishes.
+Each individual's time until reproduction is exponentially distributed:
+the reproduction rate is 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
+Suppose that the number of offspring $X$ has mean $\mu$; the variance of $X$ will not enter into the formula
+(but assume $X$ is well-behaved).
+Also suppose that the migration displacement $R$ has mean zero and variance $\sigma^2$;
+if we are in more than one dimension, we mean that the components of the dispersal distance are uncorrelated
+and each have variance $\sigma^2$.
+We also assume that the distribution of $R$ makes the associated random walk irreducible and aperiodic.
+
+Let $\Phi^N_t(x)$ be the proportion of mutant alleles present at location $x$ at time $t$,
+and $\Phi_t(x)$ the process obtained by taking $N \to \infty$ (which we assume exists).
+Denote by $\delta_x$ a single unit at location $x$, so that e.g.~$\Phi^N_t + \delta_x/N$
+is the configuration after a mutant allele has been added to location $x$.
+For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the random number of mutant alleles added if $X$ new offspring carrying mutant alleles
+replace randomly chosen individuals in a deme where the mutant allele is at frequency $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
+similarly, $\widetilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
+(We like to think of $\Phi^N_t$ as a measure, but it does not hurt to think of $\Phi^N$ as a vector;
+we aren't providing the rigorous justification here.)
+Then the above description implies that for any sufficiently nice function $f(\Phi)$ that
+\begin{align} \label{eqn:discrete_generator}
+  \begin{split} \frac{\partial}{\partial t} \E\left[ f(\Phi^N_t) \right] 
+  &= N \sum_x \left\{ \E\left[ (1+s(x+R)) \Phi^N_t(x+R) \left( f\left(\Phi^N_t + \frac{\bar X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right. \\
+     & \qquad  \qquad \left.
+{} + \E\left[ \left(1-\Phi^N_t(x+R)\right) \left( f\left(\Phi^N_t - \frac{\widetilde X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right\}  \end{split} \\
+     &= \mu \sum_x \E\left[ \left(\partial_{\phi(x)} f(\Phi_t) \right) \left\{ \Phi_t(x+R) - \Phi_t(x) + s(x+R) \Phi_t(x+R) (1-\Phi_t(x)) \right\} \right] + O\left(\frac{1}{N}\right).
+\end{align}
+In the final expectation, $R$ and $\Phi$ are independent.
+This follows by taking first-order terms in $1/N$ in the Taylor series for $f$, 
+and the fact that $\E[\bar X_\phi] = \phi \mu$ and $\E[\widetilde X_\phi] = (1-\phi)\mu$.
+We can see two things from this:
+First, since this is a first-order differential operator, the limiting stochastic process $\Phi$ obtained as $N \to \infty$
+is in fact deterministic (check by applying to $f(\Phi) = \Phi(x)^2$ to find the variance).
+Second, if we want to rescale space as well to get the usual differential equation, 
+we need to choose $\var[R]=\sigma^2$ and $s(x)$ to be of the same, small, order; 
+this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale (as noted by \citet{slatkin1973geneflow}).
+More concretely, suppose that the grid size is $\epsilon \to 0$, 
+that $\var[R] = (\sigma \epsilon)^2$, and that the strength of selection is $s(x) \epsilon$,
+and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
+then the previous equation with $f(\Phi) = \Phi(x)$ converges to the familiar form:
+\begin{align}
+  \label{eqn:derived_fkpp}
+  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \xi(t,x) + s(x) \xi(t,x) (1-\xi(t,x)) \right) .
+\end{align}
+Here we have taken first the population size $N \to \infty$ and then the grid size $\epsilon \to 0$;
+we could alternatively take both limits together, but not if $\epsilon$ goes to zero too much faster than $N$ grows.
+One reason for this is that at finite $N$,
+the process $\Phi_t$ is an irreducible, aperiodic finite-state Markov chain with absorbing states at 0 and 1;
+therefore, the inevitable outcome is extinction of one type or another,
+which is not the regime we want to study.
+
+In one dimension, we are done (see section~\ref{apx:elliptic_integrals});
+in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
+If we take a radially symmetric patch centered at the origin (so $s$ only depends on $r$), 
+and let $\xi(t,r)$ denote the mean occupation frequency at distance $r$,
+then the polar form of the Laplacian in $d$ dimensions gives us that \eqref{eqn:derived_fkpp} is
+\begin{align}
+  \label{eqn:radial_fkpp}
+  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + s(r) \xi(t,r) (1-\xi(t,r)) \right) .
+\end{align}
+
+%%%%%%%%%%%
+\subsection[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
+\label{apx:asymptotics}
+
+If $q(r)$ is a radially symmetric equilibrium frequency and $\xi(t,x)=q(|x|)$,
+then $\partial_t \xi(t,x) = 0$.
+So, using \eqref{eqn:radial_fkpp}, a radially symmetric solution 
+with $s(r) = -s < 0$ for all $r>r_0$ 
+solves, for $r_0 < r < \infty$,
+\begin{align} \label{eqn:radial_eqfreq_diffeq}
+    \partial_r^2 q(r) + \frac{d-1}{r} \partial_r q(r) - \frac{2s}{\sigma^2} q(r) (1-q(r)) = 0 \\
+    \lim_{r \to \infty} q(r) = \lim_{r \to \infty} \partial_r q(r) = 0 \qquad 
+    0 < q(r) < 1 
+\end{align}
+Since $q(r) \to 0$ as $r \to \infty$, so $q(r) (1-q(r)) \approx q(r)$,
+it can be shown 
+% theory in \citet{wasow1965asymptotic} implies 
+that the true equilibrium frequency $q$ is close, for large $r$, to the solution to
+\begin{align}
+    \partial_r^2 u(r) + \frac{d-1}{r} \partial_r u(r) - \frac{2s}{\sigma^2} u(r) = 0  \label{eqn:bessel} \\
+    \lim_{r \to \infty} u(r) = \lim_{r \to \infty} \partial_r u(r) = 0 \qquad  .
+\end{align}
+This has general solution given by a modified Bessel function:
+using \citet{gradshteyn2007table} 8.494.9,
+the general solution is
+\begin{align}
+    u(r) = C' (r-r_0)^{(2-d)/2} K_{(2-d)/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
+\end{align}
+where $C'$ and $r_0$ are chosen to match boundary conditions.
+Asymptotics of Bessel functions (\citet{gradshteyn2007table}, 8.451.6) then imply that
+\begin{align}
+    q(r) \approx C r^{(1-d)/2} \exp \left( -r \sqrt{2s}/\sigma \right) + O(1/r),
+\end{align}
+where $C$ is a different constant.
+
+%%%%%%%%% %%%%%%%%%%%%%%
+\subsection[Hitting and Occupation]{Hitting and occupation}
+\label{ss:hitting_occupation}
+
+Here we make a more precise argument to back up expression \eqref{eqn:twocase_migrate} for the migration rate.
+The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
+since it relies only on a decomposition of the migrant families upon hitting the new patch;
+but it is also imprecise in a number of subtle ways
+that are difficult to formalize.
+The argument we make here therefore takes a somewhat different tack,
+supposing that it suffices to model the spatial movement of a migrant family
+by following only the motion of the ``trunk''
+(i.e.\ the red line in figure \ref{fig:branching_decomp}),
+and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
+(Recall $\sigma$ is the dispersal distance.)
+We then use facts about Brownian motion and branching processes, 
+to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
+\linelabel{rr:moved_argument}
+
+We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
+as the sum of independent migrant families, each of whose dynamics are
+given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
+Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
+Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
+such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ 
+that initiate new migrant families is $\gamma(x)$.
+Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``ouflux of families'',
+when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
+By examining the constituent terms,
+it will turn out that equation \eqref{eqn:gestalt_migrate} 
+is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.
+
+First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
+Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
+and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
+The \emph{mean occupation density} of $Z$ at location $y$
+can be thought of  informally \linelabel{rr:density}
+as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
+It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
+i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
+gives the expected total number of individuals ever living in $S$:
+\begin{align} \label{eqn:occupation_density_defn}
+    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
+\end{align}
+% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
+Since reproduction and spatial motion are independent,
+$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
+multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
+i.e.\
+\begin{align}
+    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
+\end{align}
+The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
+is the sum of mean occupation densities of a constant outflux of branching processes
+from $\partial B(r_0)$:
+\begin{align} \label{eqn:occupation_integral}
+    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
+         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
+\end{align}
+
+This form we can now compare to 
+the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
+Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
+We assume this new patch is circular, 
+has relatively small area $A$ and is centered at position $y$.
+% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
+% multiplied by the probability that the family reaches the patch, given survival.
+% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
+It would be possible to analyze this probability directly,
+as in \citet{barton1987establishment};
+but here we take a simpler route, 
+approximating this by the chance that the trunk hits the new patch,
+multiplied by the chance that at least one member of the family escapes demographic stochasticity
+and successfully establishes in the new patch.
+Write $h_A(x,y)$ for the probability that the Brownian trunk
+hits the new patch centered at $y$ before the family dies out,
+and $f(A)$ for the chance that the family manages to establish in the new patch,
+given that it successfully arrives.
+As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
+\begin{align}\label{eqn:migrate_integral}
+  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
+\end{align}
+Recall that $\rho$ appears here but not in \eqref{eqn:occupation_integral}
+because $q(x)$ is the local \emph{proportion} of $B$ alleles.
+
+Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
+while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
+Since the expected total time spent in the patch does not depend on the distance at which the family started
+if we know that it survived to hit the patch,
+it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
+Following this intuition,
+and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
+it turns out (see section~\nameref{ss:hitting_occupation}) that
+\begin{align} 
+    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
+\end{align}
+where $g(A)$ is the total expected occupation time of the patch given the family arrives.
+Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
+Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
+that (at least asymptotically) does not depend on the distance between $x$ and $y$,
+we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
+The probability $f(A)$ that the family does not die out once it arrives at the new patch is
+approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
+if each member arriving has chance $p_e$ of establishing, independently.
+\plr{And if $p_e$ is small?}
+% If we let $w$ be the width of the patch, then the remaining term is
+% \begin{align}
+%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
+% \end{align}
+Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{ss:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
+we obtain that XXX FIX
+\begin{align}
+  \label{eqn:orig_migrate} 
+  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
+  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
+\end{align}
+This does not depend on $A$, because although
+$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
+the dependence is very weak.
+For instance, the width of a circular patch only very weakly affects the chance 
+of establishment of a new migrant that appears on its edge,
+as long as the patch is large enough.
+For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{ss:hitting_occupation}.
+
+Consider a single migrant family $Z$ beginning at $x$,
+and write $S$ for the new patch, which has area $A$ and is centered at $y$.
+Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
+and $\tau_\dagger$ an independent Exponential($s_m$) time.
+The mean occupation time of $Z$ spent in a region $S$ is,
+since the marginal distribution of a single lineage is Brownian,
+\begin{align}
+    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
+           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
+        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
+        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
+        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
+\end{align}
+where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
+and likewise $\E^x$.
+If we define
+$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
+and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\min(t,\tau_\dagger)}$,
+by the strong Markov property this is equal to
+\begin{align}
+    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
+           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] ,
+\end{align}
+where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
+provided $S$ is not pathological.
+If $S$ is circular with area $A$, the latter expectation will not depend on $x$;
+call this $g(A)$.
+
+Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
+(In the main body, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
+If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
+and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
+then this is equal to
+\begin{align}
+    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
+           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
+           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
+\end{align}
+Therefore, we have that
+\begin{align} \label{eqn:boils_downto}
+    h(x,S) \E[K] 
+    \approx \P^x[ \tau_S < \tau_\dagger ] 
+    = u(x,S) / g(A) .
+\end{align}
+
+If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
+then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
+\begin{align}
+    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
+          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
+          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) . \label{eqn:q_migrate_relation}
+\end{align}
+The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
+is expected to spend inside the disk before $\tau_\dagger$.
+This is integral of the Green function for the Bessel process of the appropriate order, 
+so using \citet{borodin2002handbook}, % (Appendix 1),
+and letting $w$ be the width of the patch,
+in $d=1$,
+\begin{align}
+    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
+  &= \int_0^{2w} \frac{ e^{- y \sqrt{2s_m} }}{\sqrt{2s_m}} dy \\
+  &= (1-e^{-2w\sqrt{2s_m}})/(2s_m)
+\end{align}
+and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
+\begin{align}
+    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
+    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
+  &= \int_0^{2w} 2 y K_0(y \sqrt{2s_m}) dy \\
+  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}} y K_0(y) dy \\
+  &= \frac{1}{2s_m}\left(1- 2w\sqrt{2s_m} K_1(2w\sqrt{2s_m}) \right)  .
+\end{align}
+In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.
+
+%%%%%%% %%%%%%%%%%%%
+\subsection[Outflux of Families]{The outflux of families}
+\label{apx:outflux}
+
+It is of interest to have an expression for the ``outflux of families'',
+$\gamma(r_0)$, 
+defined to be the mean proportion of $B$ alleles at distance $r_0$
+whose lineage traces back to the original patch without venturing further away than $r_0$.
+By expression \eqref{eqn:occupation_integral},
+\[
+    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
+\]
+for any $y$.  
+Take $y=(r,0)$.
+Note that after scaling by $\sqrt{2s_m}/\sigma$,
+\begin{align}
+    \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\|} }{ \|y\|^{-(d-1)/2} e^{-\|y\|} } \\
+        &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r (\|\hat x - x/r\| - 1)} .
+\end{align}
+Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
+so that the expression in the exponent is to, first order in $1/r$,
+\begin{align}
+    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
+            &= - 2 r_0 \cos \theta + O(1/r) .
+\end{align}
+This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
+\begin{align}
+    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} .
+\end{align}
+In one dimension we are done.
+In two dimensions,
+by bounded convergence and \citet{gradshteyn2007table} (3.339),
+\begin{align}
+  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
+            &= C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
+                         &= 2 \pi C' I_0(-2r_0),
+\end{align}
+where $I_0$ is a modified Bessel function of order zero of the first kind.
+The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2s_m} r_0 / \sigma ) $.
+
+%%%%%%%%%%%
+
+\bibliography{standing_patches_refs}
+
+\appendix
+
+\setcounter{secnumdepth}{2}
+
+%%%%%%%
+\section{The exact solution to the equilibrium frequency}
+\label{apx:elliptic_integrals}
+
+Here we describe how to solve \eqref{eqn:radial_eqfreq_diffeq} exactly in one dimension,
+i.e.\ the Fisher-KPP equation with piecewise constant selection.
+Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
+so that 
+\[
+\partial_t p(t,x) = \frac{\mu}{2} \sigma^2 \partial_x^2 p(t,x) + \mu s(x) p(t,x) (1-p(t,x)) .
+\]
+The stable distribution $\phi(x) = \lim_{t\to\infty} p(t,x)$ then solves
+\begin{align} \label{eqn:definingphi}
+    \partial_x^2 \phi(x) = - 2 s(x) \phi(x) (1-\phi(x)) /\sigma^2,
+\end{align}
+with appropriate boundary conditions.
+First rescale space by $\sigma/\sqrt{2}$ so the $\sigma^2/2$ term disappears.
+If we now assume that $s(x)$ is piecewise constant,
+$s(x) = s_i$ for $x \in [x_i,x_{i+1})$, with $x_0=-\infty$ and $x_{n+2}=\infty$,
+then the equation is integrable: if we multiply through by $2\partial_x \phi(x)$ and integrate, then we get that
+\begin{align} \label{eqn:conservation}
+    ( \partial_x \phi(x) )^2  &= - \int^{x} 2 s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
+        &= - s_i \phi(x)^2 \left( 1 - \frac{2}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) ,
+\end{align}
+if we define
+\begin{align} 
+  K_i = - s_i \phi(x_i)^2 \left( 1 - \frac{2}{3} \phi(x_i) \right) - ( \partial_x \phi(x_i) )^2  .
+\end{align}
+For ease of reference, we define
+\[
+        V_i(\phi) = s_i \phi^2 \left( 1 - \frac{2}{3} \phi \right) + K_i .
+\]
+Note that $V_i'(0)=V_i'(1)=0$, that $V_i(0)=K_i$ and $V_i(1) = K_i+s_i/3$.  
+We will always have that $V(\phi) \le 0$.
+(We have then that $\partial_x^2 \phi = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
+Also note that this implies ``conservation of energy'', i.e.\ $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
+Rearranging, we get that $dx = d\phi / \sqrt{-V(\phi)}$, so where $\phi(x)$ is monotone, the inverse is
+\[
+    x(\phi) = x(\phi^*) \pm \int_{\phi^*}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } .
+\]
+For each $i$ then define the elliptic function
+\begin{align}  \label{eqn:elliptic_function}
+    F_i(\phi) = \int_{\phi_i^*}^\phi \frac{ d\psi }{ \sqrt{ -V_i(\psi) } } ,
+\end{align}
+where take the positive branch of the square root, and $\phi_i^*$ will be chosen later.
+Then we have that
+$x(\phi) - x(\phi_0) = \pm( F(\phi) - F(\phi_0))$,
+or for an appropriate $x_0$,
+\[
+    \phi(x) = F^{-1}\left( F(\phi(x_0)) \pm (x - x_0) \right).
+\]
+
+We clearly want $\lim_{x \to \infty} \partial_x \phi(\pm x) = 0$, 
+and $\lim_{x \to \infty} \phi(\pm x)$ to be zero or one depending on the sign of $s_0$ and $s_{n+1}$.
+Since $(\partial_x \phi(x))^2 = -V(x)$, this implies that if $s_0<0$, then $K_0 = 0$,
+while if $s_0>0$ then $K_0 = s_0/6$; and likewise for $K_{n+1}$.
+
+We also require that $\phi(x)$ and $\phi'(x)$ are continuous.
+Continuity of $\phi'(x)$ is equivalent to $V_i(x_{i+1}) = V_{i+1}(x_{i+1})$,
+which we can rearrange to find an equation for $K_{i+1}$ in terms of $K_i$ and $\phi(x_{i+1})$:
+\[
+    K_{i+1} - K_i = (s_i - s_{i+1}) \phi(x_{i+1})^2 (1-2\phi(x_{i+1})/3) .
+\]
+What about the frequency at the points the selection changes, $\phi(x_i)$?  
+Well, if $\phi(x)$ is monotone on $[x_i,x_{i+1})$ then we can without loss of generality take $\phi_i^*=0$ or $1$ depending on the sign of $s_i$.
+Otherwise, let $\phi_i^*$ be the (unique) root of $V_i$ in $[x_i,x_{i+1})$, so $V_i(\phi_i^*)=0$.
+In this case, $\phi_i^*$ is the maximum or minimum of $\phi$ in the interval:
+if $s>0$, then $\phi_i = \max\{ \phi(x) : x \in [x_i,x_{i+1})\}$. 
+Recall we defined $F_i$ using $\phi_i^*$; now using the fact that $\phi$ is monotone with the opposite sign on either side of $\phi_i^*$,
+$x_{i+1} - x_i = F_i(\phi(x_{i+1})) - F_i(\phi_i^*) + F_i(\phi(x_{i})) - F_i(\phi_i^*)$,
+and that $F(\phi_i^*) = 0$,
+we know that the length of the $i$th stretch is
+\[
+    x_{i+1} - x_i = \pm \left( F_i(\phi(x_i)) + F_i(\phi(x_{i+1})) \right).
+\]
+Note that all the $\pm{}$'s are easily relatable to the signs of $s_i$.
+If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be able to solve the equations for $\phi(x_i)$ and $K_i$ recursively upwards.
+In some cases, such as \citet{slatkin1973geneflow}, we can infer $\phi(0)$ and $\partial_x \phi(0)$ by spatial symmetry.
+In other cases, we are only given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
+
+\subsection{Doing the integrals}
+\label{apx:some_integrals}
+
+An important ingredient in the above method is the integral \eqref{eqn:elliptic_function},
+to which a method for solving the Korteweg-de Vries equation can be applied \citep{NEQwiki}.
+Recall that $V_i(\phi) =  s_i \phi^2(1-2\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries;
+since we're just working within an interval with $s$ constant, we can rescale space by $\sqrt{2 s_i}/\sigma$,
+so that $s$ and $\sigma$ drop from the equation.
+We then want to integrate
+\[
+    \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
+         \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } ,
+\]
+over a domain where $V$ is always negative.
+Let $\phi^2(1-2\phi/3)+K = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
+and change variables first to $y^2=(\alpha-\phi)$, 
+and then to $x = y/\sqrt{\alpha-\beta}$, so that
+\begin{align*}
+    \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } 
+        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-y^2) } } \\
+        = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
+\end{align*}
+with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
+Now Jacobi's incomplete elliptic integral of the first kind is defined by
+\[
+    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) }} ,
+\]
+and the Jacobian elliptic function $\sn(x;k)$ is the inverse: $F(\sn(x;k);k) = x$.
+As $k \to 0$, $\sn(x;k) \to \sin(x)$, while as $k \to 1$, $\sn(x;k) \to \sinh(x)$.
+
+% Presumably, this gets the same answer as \citep{slatkin1973geneflow}, who finds that in the solvable cases,
+% for $x$ in the last interval (which ends in $\infty$), that
+% \begin{align*}
+%     \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / 2 \sigma + C ) \right) \\
+%         &\simeq C' \exp\left( - \frac{ x \sqrt{2s} }{ \sigma } \right) .
+% \end{align*}
+
+%%%%%%%
+\section{More information about the simulations}
+\label{apx:parameter_tables}
+
+Here we provide additional details regarding the simulations we used
+to validate the theory in figures \ref{fig:sim_times} and \ref{fig:sim_probs}.
+
+% table of parameters, produced by analyze-sims.R
+\input{mutation-parameters-table.tex}
+
+% table of parameters, produced by analyze-sims.R
+\input{migration-parameters-table.tex}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{mutation-times-predicted}
+  \end{center}
+  \caption{
+    The same data shown in the left panel of figure~\ref{fig:sim_times},
+    but all times shown (not just the interquartile ranges),
+    and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
+  } \label{sfig:sim_migration_times}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics{migration-time-predicted}
+  \end{center}
+  \caption{
+    The same data shown in the right panel of figure~\ref{fig:sim_times},
+    but all times shown (not just the interquartile ranges),
+    and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
+    The left panel has the predicted time to adaptation on the horizontal axis as in figure~\ref{fig:sim_times},
+    and the right panel has, for comparison, the raw distance between patches 
+    (which predicts time to adaptation, but not as well).
+  } \label{sfig:sim_mutation_times}
+\end{figure}
+
+% example trace plots
+
+% sm=0.1
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-mutation-sims/18885-r1-501-sb0_01-sm-0_1-N1200-pophistory-run}
+    \includegraphics{example-mutation-sims/56325-r1-501-sb0_01-sm-0_1-N600-pophistory-run}
+    \includegraphics{example-mutation-sims/28432-r1-501-sb0_01-sm-0_1-N50-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.1$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_1}
+\end{figure}
+
+% sm=0.01
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-mutation-sims/43099-r1-501-sb0_01-sm-0_01-N1200-pophistory-run}
+    \includegraphics{example-mutation-sims/69787-r1-501-sb0_01-sm-0_01-N600-pophistory-run}
+    \includegraphics{example-mutation-sims/11821-r1-501-sb0_01-sm-0_01-N50-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.01$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_2}
+\end{figure}
+
+% sm=0.001
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-mutation-sims/59611-r1-501-sb0_01-sm-0_001-N1200-pophistory-run}
+    \includegraphics{example-mutation-sims/93713-r1-501-sb0_01-sm-0_001-N600-pophistory-run}
+    \includegraphics{example-mutation-sims/29850-r1-501-sb0_01-sm-0_001-N50-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_3}
+\end{figure}
+
+% sm=0.0001
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-mutation-sims/15449-r1-501-sb0_01-sm-1e-04-N1600-pophistory-run}
+    \includegraphics{example-mutation-sims/5582-r1-501-sb0_01-sm-1e-04-N400-pophistory-run}
+    \includegraphics{example-mutation-sims/24639-r1-501-sb0_01-sm-1e-04-N25-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_4}
+\end{figure}
+
+% sm=0.1, R=80
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-migration-sims/89826-r1-501-sb0_01-sm-0_1-N4000-pophistory-run}
+    \includegraphics{example-migration-sims/76178-r1-501-sb0_01-sm-0_1-N1000-pophistory-run}
+    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.1$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_5}
+\end{figure}
+
+% sm=0.01, R=80
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-migration-sims/42080-r1-501-sb0_01-sm-0_01-N4000-pophistory-run}
+    \includegraphics{example-migration-sims/82698-r1-501-sb0_01-sm-0_01-N1000-pophistory-run}
+    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.01$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_6}
+\end{figure}
+
+% sm=0.001, R=80
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-migration-sims/88079-r1-501-sb0_01-sm-0_001-N4000-pophistory-run}
+    \includegraphics{example-migration-sims/3464-r1-501-sb0_01-sm-0_001-N1000-pophistory-run}
+    \includegraphics{example-migration-sims/37774-r1-501-sb0_01-sm-0_001-N100-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_7}
+\end{figure}
+
+% sm=0.0001, R=80
+\begin{figure}
+  \begin{center}
+    \includegraphics{example-migration-sims/26044-r1-501-sb0_01-sm-1e-04-N4000-pophistory-run}
+    \includegraphics{example-migration-sims/85750-r1-501-sb0_01-sm-1e-04-N1000-pophistory-run}
+    \includegraphics{example-migration-sims/4111-r1-501-sb0_01-sm-1e-04-N100-pophistory-run}
+  \end{center}
+  \caption{
+    Randomly chosen simulations with $s_m=0.0001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
+    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
+    (i.e.\ each line represents a vertical slice through the plot on the left);
+    dotted black lines indicate the patches where $B$ is advantageous.
+  } \label{sfig:sims_8}
+\end{figure}
+
+\ifsubmissionversion
+\processdelayedfloats
+\fi
+
+%%%% Reviews
+\includereviews
+
+\end{document}
+
+%%%%%%%% END %%%%%%%%%%%%%%%
+%%%%%%%% %%% %%%%%%%%%%%%%%%
+
+\begin{figure}[ht!]
+    \begin{center}
+        % made in patchy-sim-writeup-plots.R
+        \includegraphics{example-equilibrium}
+    \end{center}
+    \label{fig:example_equil}
+    \caption{
+    Snapshots of a simulation of allelic frequency dynamics around a patch where the allele is beneficial:
+    the focal allele is benefical in the middle region and deleterious outside.
+    The dynamics were simulated for 10000 generations across demes with 1000 individuals in each
+    and random dispersal to nearby demes with a mean dispersal distance of $\sigma \approx 2$ demes.
+    Colored lines show allele frequencies across space in 50 uniformly spaced generations;
+    the solid line shows the mean across all generations.
+    Below is the same plot, except on a log scale.
+    }
+\end{figure}
+
+\section{Temporary work: hitting probs}
+
+Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
+which we assume has relatively small area $A$ and is centered at position $y$.
+The probability that this family reaches $A$ by time $t$ is $(1-k_e(t))$
+multiplied by the probability that the family reaches the patch, given survival.
+If $t$ is large then the previous section argues that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
+and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
+will have hit the patch by then.
+If $w = A^{1/d}$ is the width of the patch, then 
+this is the chance that a $d$-dimensional Bessel process begun at $\|y-x\|$ hits $w$ by time $t$.
+Let $\tau_w$ be the hitting time of $w$,
+so, integrating by parts, and using Borodin \& Salminen (2.2.0.1 and 4.2.0.1)
+\begin{align}
+    \int_0^\infty (1-k_e(t)) \P^x\{ \tau_w \le t \} dt  
+    &\approx \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\
+    &= \frac{1}{s_m \E[K]} \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \in dt \} \\
+    &= \frac{1}{s_m \E[K]} \E^x[e^{-s_m \tau_w}] \\
+    &= 
+    \frac{1}{s_m \E[K]} \times
+    \begin{cases}
+        e^{-(x-w) \sqrt{2s_m}} \qquad & d=1 \\
+        \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } \qquad & d=2 
+    \end{cases}
+\end{align}
+Note that the $d=2$ case
+\begin{align}
+    \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } = \sqrt{\frac{w}{x}} e^{-(x-w)\sqrt{2s_m}} ( 1 + O(1/w) ).
+\end{align}
+
+\section{Patchy selection} 
+\label{ss:discretedemes}
+
+We have been working in a continuous model of geographic space; 
+in this section we move to a discrete model of isolated populations exchanging rare migrants,
+but in section \ref{ss:patchyspace} discuss how a model with {\em patchy} selection can be approximated by such a discrete model.
+
+Suppose now that we have a set of $L$ discrete populations of individuals that may reproduce or migrate,
+and that migration and mutation are rare, so that the waiting times until either event is approximately exponential.
+Suppose that for each pair of populations, labeled $i$ and $j$, the mean number of migrants that travel from $i$ to $j$ per generation
+is $M_{ij}$ and that the mean number of new mutations appearing in population $i$ per generation is $\mu_i$.
+
+{\tt Include cartoon here.}
+
+If we suppose that fixation occurs on a faster time scale than mutation or migration,
+so that it is very unlikely that two different migrants or mutants begin to fix locally in the same population,
+then we can treat the process of extinction or local fixation as instantaneous 
+(a limit previous studied by \cite{Slatkin:81}).
+In this limit we may also ignore swamping effects of inmigration of nonmutant alleles.
+
+In this case, we have the following picture.
+The population at first starts out with no selected alleles. 
+At some later time point suppose that there are $T$ different alleles present in the population, 
+that the $k^\mathrm{th}$ allele is occupying demes $\{i^k_1, \ldots, i^k_{n_k}\}$, for $1\le k \le T$.
+Let $I$ denote the set of all adapted demes, 
+and let $J = \{j_1, \ldots, j_\ell\}$ denote the demes that have not yet adapted, with $\ell = L - \sum_k n_k$.
+Then we are assuming that two types of event can happen which could change the state:
+either at rate $\mu_j p_f$, an unadapted deme $j$ produces a new mutant, which fixes in this deme;
+or at rate $M_{ij} p_f$, an adapted deme $i$ sends a migrant to unadapted deme $j$, which fixes in that deme.
+Therefore, under these assumptions, the probability of fixation $p_f$ (and hence the selection coefficient)
+only enter as a time scaling: for instance,
+if we define $R = \sum_{j \in J} \left( \mu_j + \sum_{i \in I} M_{ij} \right)$,
+then the total rate of such events is $p_f R$.
+The probability that the next event results in, say, adaptation of deme $j$
+is 
+\[
+ \left( \mu_j + \sum_{i \in I} M_{ij} \right) / R,
+\]
+and the probability that deme $j$ adapts through migration from deme $i$, given that deme $j$ is the next to adapt, is
+$M_ij / R$,
+while the probability that it adapts through a new mutation is $\mu_j / R$.
+
+The main point here is that the final pattern,
+and hence the likely extent of parallel adaptation,
+is independent of the strength of selection.
+This is in contrast to the continuous case, however, this observation is dependent on the assumption of weak migration
+and fails to hold if different migrant lineages interact.
+This can be viewed as a continuous-time Markov process XXX.
+Explicit solutions are not available in the general case,
+but if we specialize to a completely symmetric model (i.e.\ the ``island'' model),
+then the mathematical picture is a pretty one.
+
+\subsection{Symmetric patches}
+
+Indeed, suppose that migration and mutation are symmetric: $M_{ij}=m$ and $\mu_j = \mu$ for all $j$ and $i\neq j$.
+Suppose at some time the $k^\mathrm{th}$ allele is occupying $n_k$ demes, 
+with $n = \sum_k n_k < L$.
+Then at rate $(L-n) N \mu p_f + n (L-n) N m p_f$,
+either a migration or mutation event happens.
+With probability $\mu/(\mu + n m)$, it is a mutation,
+and a new deme is randomly chosen and assigned a new allele.
+With probability $n_k m / (\mu + n m)$,
+the $k^\mathrm{th}$ type sends a successful migrant to a randomly chosen new island,
+which is assigned the $k^\mathrm{th}$ allele.
+
+This process is a continuous-time version of the ``Chinese restaurant process''
+described in \citet{aldous1985exchangeability} and \citet{pitman1995partitions},
+and so the final partition of types across demes has the Ewens sampling distribution with parameter $\mu/m$.
+Again note that the selection coefficient, which is implicit in the probability $p_f$,
+does not enter into the final distribution.
+
+Now we can compute most properties we might want about the process.
+For instance, the expected number of distinct mutational origins is
+\begin{equation}
+    \E\left[ \mbox{ \# of independent mutations } \right] = 
+            1 + \frac{\mu}{m} \sum_{k=2}^L \frac{1}{k+(\mu/m)}. \label{discrete_expected}
+\end{equation}
+The probability that there is only a single successful mutation is
+\begin{equation}
+\P \left\{ \mbox{only one mutational origin} \right\} = 
+            \prod_{k=1}^{L-1} \frac{ k }{ k+(\mu/m)} .
+\end{equation}
+More generally, suppose we sample a single deme, and are interested in the total number of demes $S$
+(including the sampled one) that share the same fixed mutation as our sampled deme.
+Then $S$ has distribution
+\begin{equation} \label{eqn:Sdistrn}
+\P\{ S=s \} = \frac{ (s-1)! (\mu/m)^{L-s} }{ \prod_{k=1}^{L-1} (k+(\mu/m)) } .
+\end{equation}
+Furthermore, if $L$ is large, then $S/L$ has probability density approximately
+\begin{equation} \label{eqn:betadistrn}
+   (\mu/m) (1-x)^{(\mu/m)-1}
+\end{equation}
+namely, is a $\mathrm{Beta}((\mu/m), 1)$ distribution--- see \cite{donnelly1989continuity} and \cite{perman1992sizebiased}.
+
+% In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
+% for a model with 20 interconnected demes, across different parameter values.  
+% This indicates that we need the population-scaled mutation and migration rates to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.
+If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
+
+The high connectedness of the discrete deme island model means that the expected number of distinct alleles, (Equation \eqref{discrete_expected}), 
+grows with the $\log$ of the number of demes.
+This strongly contrasts with the continuous spatial model where the local nature of dispersal means that doubling the species range will double the number of mutations expected.
+Furthermore, while in the continuous model the areas occupied by distinct mutations are similar in area, in this strong-selection discrete island model, 
+only a few types tend to dominate even as the number of demes (and thus, distinct types) grows large (see Equations \eqref{eqn:Sdistrn} and \eqref{eqn:betadistrn}).
+
+A more general model would allow migration only along edges of some graph connecting the demes.
+In the low migration limit such a model still produces a partition distribution independent of the selection coefficient,
+but it does not in general have the Ewens distribution.
+Another extension would be to include the time alleles need to achieve an intermediate frequency,
+along the lines of \citet{Navarro:03}, which would reintroduce dependence on the selection coefficient.
+
+\texttt{XXX omit this? XXX}
+We have treated the time during which the selected allele is at intermediate frequency in a deme (about $\log(\sigma^2 \rho)/s_p$ generations) as negligible, 
+which ignores migration events that occur while the mutation is spreading through the population. 
+we now sketch of a more realistic approximation.
+For each occupied--unoccupied pair of demes, if the number of mutant alleles in the occupied deme is $n(t)>0$,
+then at time $t$, migrants from the occupied deme pass the selected allele to the unoccupied deme at rate $2 n(t) m p_f(s_p)$,
+while newly arisen mutants arise and fix in the unoccupied deme at constant rate $2 N \mu p_f(s_p)$.
+The selected type should grow approximately as $n(t)=\exp(s(t-t_0))$.
+Of course, to follow this model through, we will have to account for multiple types within a single deme and other complications,
+but we can at least make a few observations.
+One is that the final distribution of types is no longer independent of $s_p$, which enters through the dynamics $n(t)$.
+However, it is clear that increasing the selection coefficient $s$ will decrease the number of types (independent mutational origins),
+while on the other hand, increasing the deme size $N$ will increase the number of types,
+and that the case presented above is at the extreme (see also \cite{Navarro:03} for deterministic approximations).
 
@@
 for the length of shared haplotype between patches.
-
-%%%%%%%%%%
-\subsection{Applications} 
-\label{ss:applications}
-% FOR REFERENCE:
-% On Wed, May 28, 2014 at 10:20 AM, Michael Nachman <mnachman@berkeley.edu> wrote:
-% > Graham:
-% >
-% > The density is probably in the range of 10-40 animals per hectare (10,000
-% > m^s).  We built enclosures of 0.25 hectare and found that they could support
-% > about 10 animals with food provided ad libitum, so I think this is an upper
-% > density (i.e.\ 40/hectare).  As for dispersal distance, I'm less sure.
-% > Studies on closely related species (genus Perognathus) in similar habitats
-% > suggest distances of 100's of feet.  Attached are some old papers that tried
-% > to estimate dispersal distances in Perognathus.
-% > best,
-% > Michael
-
-Coat color in the rock pocket mouse
-\emph{Chaetodipus intermedius} is a classic example of local
-adaptation 
-\citep{benson1933concealing,DiceBlossom1937}.
-These mice live on rocky outcrops throughout the southwestern United States and northern Mexico, 
-and generally have a light pelage similar to the predominant rock color.
-However, in some regions these mice live on darker substrates (e.g.\ old lava flows),
-and in these patches have much darker pigmentation, 
-presumably to avoid visual predators.
-\citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
-much of the change in coloration is due to an allele at the MC1R locus.
-This dark allele differs from the light allele by four amino acid changes, 
-and has a dominant or partially dominant effect depending on the measure of coat color.
-These areas of dark rock are separated from each other by a range of distances, 
-with some being separated from any other by hundreds of kilometers of light colored substrate. 
-The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
-suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
-However, \citet{hoekstra2004local} reasoned that, elsewhere in the range, 
-multiple close dark outcrops may share a dark phenotype whose genetic basis has been spread by migration
-despite intervening light habitat. 
-We apply our theory to this example,
-in particular investigating how the probability of parallel mutation depends on 
-the distance between dark outcrops.
-
-A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. 
-\citet{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
-at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
-On the lava flow the dark MC1R allele is at 86\% frequency,
-while at a site with light substrate 12km to the west (Tule), the frequency is 3\%.
-The dark allele was entirely absent from Christmas pass, a site with light substrate 7km north of Tule, and 3km further from the lava flow.
-In the other direction, the dark MC1R allele was at 34\% at a site with light substrate 10km to the east of the flow (O'Neill).
-These numbers give us a sense of a plausible range of characteristic lengths.
-Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
-we can fit formula \eqref{eqn:eqfreq} to these frequencies
-(similar to \citet{hoekstra2004ecological}).
-Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
-and for O'Neill, $\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. 
-We also need estimate of $s_m$. 
-Using $\sigma \approx 1$km \citep{french1968dispersal,allred1963range}, 
-these cline widths imply that $s_m=1/9$ and $1/900$ are reasonable values.
-
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics{Lava_flow_mice_prob_parallel}
-  \end{center}
-  \caption{
-Top panel: The probability of parallel mutation as a function of the distance
-between environmental patches for two different cline widths
-and two different mutation rates (using equation \eqref{eqn:parallel_prob}). 
-The parameters were chosen to match the example of \textit{Chaetodipus intermedius};
-see text. Bottom panel: The initial genetic length of the haplotype
-shared between patches due to adaptation via migration as a function
-of the distance between environmental patches. 
-  \label{fig:mice_prob_parallel}
-  }
-\end{figure}
-
-The mutational target size $\mu$ for the trait is unclear. 
-While the Pinacate dark haplotype differs from the light haplotype at four amino acid residues,
-it is likely that not all of these changes are needed for a population to begin to adapt. 
-Also there a number of genes besides MC1R at which adaptive changes affecting pigmentation 
-have been identified in closely related species and more broadly across vertebrates \citep{Hoekstra:06}.
-To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
-where the adaptive mutation has to occur at a single base pair; 
-and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation
-being able to arise at any of a thousand base pairs.  
-Finally, we set $A=100\text{km}^2$ (roughly the size of the Pinacate
-patch). In
-Figure~\ref{fig:mice_prob_parallel} we show the dependence of the probability
-of parallel mutation on the distance between lava flow patches using these parameters, 
-showing that parallel mutation should become
-likely over a scale of tens to a few hundred kilometers between patches.
-
-Given the large selection coefficient associated with the dark allele
-on the dark substrate, we expect the initial haplotype associated with
-either a new mutation or migrant allele to be large. 
-We can use equation \eqref{eqn:haplotype_length} to calculate how long the founding
-haplotype shared between populations is expected to be,
-as shown in Figure~\ref{fig:mice_prob_parallel}. 
-The initial length can be quite long between geographically close patches (tens of kilometers). 
-However, for the wider cline width ($\sigma/\sqrt{s_m} = 30$km), 
-adaptation by migration can still be likely for patches 100km apart, 
-but the shared basis may be hard to detect, as the length of shared haplotype can be quite short.
-
-Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
-we should expect the haplotype to be rapidly whittled down within
-patches after establishment (equation \ref{eqn:rec_rate_within}). We
-estimate that the effective recombination rate is no less than 50\% of the actual recombination rate.
-Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare 
-(M.~Nachman, personal communication),
-and the variance in offspring number to be $\xi^2 = 3$ (mostly to get round numbers),
-then the shared haplotype length within patches from equation \eqref{eqn:within_haplotype_length}
-is only expected to be on the order of hundreds or thousands of bases
-($10^{-3}$--$10^{-4}$cM).
-This suggests that unless the migration has happened fairly recently
-then it will be difficult to identify shared haplotypes
-(how recently depends on the background level of LD in the population).
-
-%%%%%%% %%%%%%% %%%%%%%
-\section*{Discussion} 
-\label{ss:discussion}
-
-This paper is an investigation into the basic question: 
-What is the spatial resolution of convergent local adaptation?
-In other words, 
-over what spatial scale of environmental patchiness will the process
-of adaptation develop independent solutions to evolutionary problems?
-The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
-the dispersal distance divided by the square root of the strength of selection against the allele between patches. 
-It depends much more weakly on the
-selective benefit within the patches or (perhaps surprisingly) the population density, 
-although these two factors will determine the time-scale over which adaptation will occur
-(and note that population density could affect the selection coefficients). \linelabel{rr:popdensity}
-This is in contrast with models of panmictic populations
-\citep{softsweepsI, MesserPetrov, Wilson-softsweep} 
-and geographically spread populations adapting to homogeneous selection pressures \citep{ralph2010parallel}, 
-where the probability of multiple, independently arising adaptive alleles increases with the population size/density. 
-However, in all of these models the dependence on the beneficial selection
-coefficient is absent or weak, due to the fact that selection both
-aids establishment of new alleles and the spread of existing
-alleles (but see \citet{Wilson-softsweep} for the complications of varying population sizes).
-
-We have also shown that
-while weaker selection against alleles will make sharing 
-of adaptations between patches easier, 
-it will also makes it harder to spot the sharing,
-since the lucky alleles that manage to colonize new patches move slower,
-and thus carry a shorter shared haplotype.
-This issue is amplified by the fact that the length of haplotype
-shared within patches decays over time, potentially making the
-identification of shared adaptations to old selection pressures difficult.
-
-Perhaps the most useful rule-of-thumb quantities we found were the following.
-The effective rate of migration into an as-yet-unadapted patch from an already-adapted patch distance $R$ away 
--- the analogue of the mutational influx $\mu \rho A$ --
-is $2 \rho s_m \exp(- R \sqrt{2 s_m} / \sigma)$.
-Equivalently, the critical gap size between patches past which adaptation is likely independent
-is $R = (\sigma/\sqrt{2s_m}) \log(2s_m/(A \mu))$.
-Finally, successfully transiting migrant lineages move at rate $\sigma \sqrt{2s_m}$,
-and so shared haplotype lengths between patches will be of order $R/(\sigma \sqrt{2s_m})$.
-
-In developing the set of approximations in the paper we have ignored a
-number of complicating factors. We now briefly discuss these.
-
-\paragraph{Standing variation} 
-We have focused on relative rates of adaptation,
-since we are thinking of applications in which we know that adaptation has occurred,
-and the question is whether or not adaptations in distinct patches
-have appeared independently or not.
-However, especially absolute if rates of adaptation by new mutation are very low,
-any adaptation that does occur may have to make use of standing variation.
-The case of a panmictic population was studied by \citet{softsweepsI},
-and we study the case of a continuous, spatial population in an upcoming paper.
-If parallelism in local adaptation of the sort we study here is due to standing variation
-rather than new mutation,
-then the dynamics of adaptation should not depend strongly on migration patterns
-(but note that the initial spatial distribution of standing variation may).
-
-\paragraph{Dominance}
-We have mostly ignored the issue of dominance 
-by dealing with essentially haploid models,
-and appealing to the fact that the dynamics we study occur where the mutation is rare,
-and hence mostly present only in heterozygotes. 
-Our results should hold as a good approximation to dominant and partially dominant alleles
-(with $s_m$ the selection against heterozygotes).
-If, however, the mutation is recessive, then it is essentially neutral where rare,
-and so would encounter much less resistance to spreading between patches.
-The shape of the cline obtained is given by \citet{haldane1948theory}.
-This is counteracted, however, by the increased difficulty with which the mutation would establish
-once arriving in a new patch,
-if the beneficial effect is also recessive. \linelabel{rr:dominance}
-As such it is not clear what our intuition should be 
-about the contribution of recessive alleles to adaptation via migration. 
-Further work is needed to put empirical observations of local adaptation by recessive alleles 
-in a theoretical context.
-It is simliarly unclear how the model should extend to a polygenic trait.
-
-\paragraph{Long distance migration}
-We have also ignored the possibility of very long distance migration,
-instead focusing on local dispersal (and using Gaussians via the central limit theorem).
-However, dispersal distributions can be very heavy tailed, 
-with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
-In addition, over long time-scales, very rare chance events (mice carried off by hurricanes and the like; \citet{censky1998overwater,nathan2008mechanisms})
-could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
-Such tail events could greatly increase the probability of parallel adaptation above that predicted by our model. 
-Furthermore, if adaptive alleles do move between distant patches via rare, long distance migration 
-then they will be associated with a much longer shared haplotype than predicted by equation \eqref{eqn:haplotype_length}. 
-As such, we view our results as a null model by which the contribution of long distribution migrants to adaptation could be empirically judged.
-
-\paragraph{Other geographic models}
-We have studied circular patches of habitat at long distances from each other.
-Real habitat geometry can be much more complex,
-e.g.\ with archipelagos of patches of varying sizes, 
-or patches connected by long, skinny corridors, for instance.
-The work of \citet{cantrell1991diffusive} comes closest to a general theory of balanced polymorphisms in such habitats.
-It is possible that our techniques could be applied in their much more general setting,
-as both are based, fundamentally, on branching process approximations.
-It is also interesting to think about the probability of convergent
-adaptation to continuously varying environments, e.g. replicated
-environmental clines.
-
-\paragraph{Concluding thoughts}
-The falling cost of population genomic sequencing means that we will soon have the
-opportunity to study the interplay of adaptation with geography and ecology 
-across many populations within a species. 
-Our work suggests that even quite geographically close populations 
-may be forced to locally adapt by repeated, convergent, \textit{de novo} mutation 
-when migration is geographically limited
-and selective pressures are divergent.
-Thus, systems where populations have been repeatedly subject to 
-strong local selection pressures may offer the opportunity 
-to study highly replicated convergent adaptation within a similar
-genetic background \citep{stern2013genetic}.  
-Such empirical work will also strongly inform our understanding of the
-ability of gene flow to keep ecologically similar populations evolving
-in concert \citep{sexton2013genetic}.
-Our results suggest that adaptation to shared environments is
-certainly no guarantee of a shared genetic basis to adaptation, 
-suggesting that rapid adaptation to a shared environment could potentially drive
-speciation if the alleles that spread in each population fail to work
-well together \citep{Kondrashov:03}.
-
-% discussion of applications.
-
-% discuss mixing of types; 
-% Graph: from simulation showing local proportions in a single patch showing initial fixation of a single type and later mixing converging toward more than one type.  (and eventual loss of one?)
-
-\subsection*{Acknowledgements}
-We would like to heartily thank Joachim Hermisson for a number of insightful suggestions,
-and catching the algebra error present in the first version of this paper on the bioRxiv,
-as well as the two anonymous reviewers.
-We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
-as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
-for helpful comments and discussion of the manuscript.
-
-%%%%%
-\section*{Materials \& Methods}
-
-%%%%%%%
-\subsection{Simulation methods}
-\label{ss:simulations}
-
-First we briefly describe the simulations we used for illustration and validation
-(the \texttt{R} code used is provided as a supplement).
-We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
-Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
-locally beneficial alleles were more likely to reproduce.
-Each then produced a random, Poisson distributed number of offspring that each, independently, 
-either remained in the same location with probabilty $0.8$ or
-migrated a random number of steps in a uniformly chosen cardinal direction;
-the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
-\linelabel{rr:sim_details}
-Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
-(Although we described the simulation in terms of individuals,
-we kept track only of total numbers in an equivalent way.)
-
-The reproduction rate in simulations for type $b$ alleles was 30\%; 
-this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
-where the value of $s$ is either $s_m$ or $s_p$ depending on the individual's location.
-This determines the values of $s_m$ and $s_p$ reported in the figures, 
-and do not depend on the basic rate of reproduction.
-However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
-we computed the rate of intrinsic growth, 
-i.e.\ the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
-in the absence of migration.
-(The resulting values are close to the first notion of $s$,
-but give better agreement with theory,
-which uses the second definition.)
-
-To sample lineages, we first simulated forwards in time the population dynamics,
-then sampled lineages moving back through time
-by, in each generation,
-moving each lineage to a new deme with probability proportional to the reverse migration probability
-weighted by the number of $B$ alleles in that deme in the previous time step.
-If more than one lineage was found in a deme with $n$ $B$ alleles,
-then each lineage picked a label uniformly from $1 \ldots n$,
-and those picking the same label coalesced.
-Since reproduction is Poisson, this correctly samples from the distribution of lineages given the population dynamics.
-
-\subsection[Probability of Establishment]{Numerical calculation of the probability of establishment}
-\label{apx:establishment_sims}
-
-When rare, under at least some demographic models,
-copies of a new mutant allele are approximately independent and experience a uniform selective benefit;
-and can therefore be treated as a branching process (a spatial branching process in this case).
-Furthermore, whether or not a new, beneficial mutation establishes or is lost to demographic stochasticity
-is determined by this phase where it is rare.
-Fortunately, the probability that a branching process does not establish
-(i.e.\ dies out)
-can be found as a fixed point of the generating function of the process \citep{jagers1975branching}.
-Therefore, we calculated explicitly the generating function for a spatial branching process
-with nearest-neighbor migration on a one-dimensional lattice
-and a Poisson number of offspring with mean $1+s$,
-where $s$ could vary by location,
-and iterated this forward to convergence to obtain $1-p(x)$, the probability a single mutation appearing at $x$ would fail to establish.
-We considered two situations:
-where $s$ is a step function,
-and where it has a linear transition.
-These solutions are shown, and parameters described, in figure~\ref{fig:prob_estab_calcs}.
-
-Here (and at other parameter choices) we see that the probability of establishment $p(x)$ goes to the equilibrium value
-(approximately $p_e = 2s/\xi^2$) within the patch;
-the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
-The fit is equally good for other parameter values,
-even if migration can move further than one deme and offspring numbers are not Poisson (results not shown).
-This lends credence to our approximation that the integral of $p(x)$ over the entire range
-is close to $p_e$ multiplied by the area of the patch.
-
-\begin{figure}[ht!]
-    \begin{center}
-        % made in patchy-selection-plots.R
-        \includegraphics{prob-establishment}
-    \end{center}
-    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
-    in an abrupt transition (left) and a gradual transition (right).
-    The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respectively);
-    and the selection coefficient interpolates linearly between these in the central hatched region.
-    The number of offspring is Poisson.
-    The underlying landscape has demes every 15km and a migration rate of 0.5 between neighboring demes;
-    the probability was found by numerically solving the equation for the probability of establishment of a multi-type branching process.
-    Note that the approximation used in the text is here $p_e \approx 2 s_p / \xi^2 = 0.1$,
-    agreeing with the asymptotic value on the right.
-    \label{fig:prob_estab_calcs}
-    }
-\end{figure}
-
-%%%%%%%%%%%%
-\subsection[Equilibrium Frequency]{The equilibrium frequency}
-\label{apx:eqfreq}
-
-For completeness, and clarity as to the scalings on the relevant parameters,
-here we provide a derivation of the differential equations referred to above equation~\eqref{eqn:eqfreq},
-and establish the asymptotics given in that equation.
-One route to the ``equilibrium frequency'' of the allele outside the range where it is advantageous is as follows;
-see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} (or \citet{KPP1937book} or \citet{fisher1937wave} or \citet{haldane1948theory})
-for other arguments in equivalent models,
-or \citet{etheridge2000introduction} and/or \citet{dawson1993measurevalued} for general theory used below.
-Following is a description of a discrete model (which is nearly the one we simulate from).
-Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
-and that selection (for or against) the allele is given by the function $s(x)$, with $x$ denoting the spatial location.
-Each individual at location $x$ reproduces after a random lifetime,
-dying, and then leaving behind in the same deme a random number of offspring with distribution given by $X$;
-each offspring then migrates independently to a new location chosen randomly from the distribution given by $x+R$,
-where she replaces a randomly chosen individual.
-If $x+R$ is outside of the range, then she perishes.
-Each individual's time until reproduction is exponentially distributed:
-the reproduction rate is 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
-Suppose that the number of offspring $X$ has mean $\mu$; the variance of $X$ will not enter into the formula
-(but assume $X$ is well-behaved).
-Also suppose that the migration displacement $R$ has mean zero and variance $\sigma^2$;
-if we are in more than one dimension, we mean that the components of the dispersal distance are uncorrelated
-and each have variance $\sigma^2$.
-We also assume that the distribution of $R$ makes the associated random walk irreducible and aperiodic.
-
-Let $\Phi^N_t(x)$ be the proportion of mutant alleles present at location $x$ at time $t$,
-and $\Phi_t(x)$ the process obtained by taking $N \to \infty$ (which we assume exists).
-Denote by $\delta_x$ a single unit at location $x$, so that e.g.~$\Phi^N_t + \delta_x/N$
-is the configuration after a mutant allele has been added to location $x$.
-For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the random number of mutant alleles added if $X$ new offspring carrying mutant alleles
-replace randomly chosen individuals in a deme where the mutant allele is at frequency $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
-similarly, $\widetilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
-(We like to think of $\Phi^N_t$ as a measure, but it does not hurt to think of $\Phi^N$ as a vector;
-we aren't providing the rigorous justification here.)
-Then the above description implies that for any sufficiently nice function $f(\Phi)$ that
-\begin{align} \label{eqn:discrete_generator}
-  \begin{split} \frac{\partial}{\partial t} \E\left[ f(\Phi^N_t) \right] 
-  &= N \sum_x \left\{ \E\left[ (1+s(x+R)) \Phi^N_t(x+R) \left( f\left(\Phi^N_t + \frac{\bar X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right. \\
-     & \qquad  \qquad \left.
-{} + \E\left[ \left(1-\Phi^N_t(x+R)\right) \left( f\left(\Phi^N_t - \frac{\widetilde X_{\Phi_t(x+R)}}{N}\delta_{x}\right) - f(\Phi^N_t) \right) \right] \right\}  \end{split} \\
-     &= \mu \sum_x \E\left[ \left(\partial_{\phi(x)} f(\Phi_t) \right) \left\{ \Phi_t(x+R) - \Phi_t(x) + s(x+R) \Phi_t(x+R) (1-\Phi_t(x)) \right\} \right] + O\left(\frac{1}{N}\right).
-\end{align}
-In the final expectation, $R$ and $\Phi$ are independent.
-This follows by taking first-order terms in $1/N$ in the Taylor series for $f$, 
-and the fact that $\E[\bar X_\phi] = \phi \mu$ and $\E[\widetilde X_\phi] = (1-\phi)\mu$.
-We can see two things from this:
-First, since this is a first-order differential operator, the limiting stochastic process $\Phi$ obtained as $N \to \infty$
-is in fact deterministic (check by applying to $f(\Phi) = \Phi(x)^2$ to find the variance).
-Second, if we want to rescale space as well to get the usual differential equation, 
-we need to choose $\var[R]=\sigma^2$ and $s(x)$ to be of the same, small, order; 
-this is another way of seeing that $\sigma/\sqrt{s}$ is the relevant length scale (as noted by \citet{slatkin1973geneflow}).
-More concretely, suppose that the grid size is $\epsilon \to 0$, 
-that $\var[R] = (\sigma \epsilon)^2$, and that the strength of selection is $s(x) \epsilon$,
-and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
-then the previous equation with $f(\Phi) = \Phi(x)$ converges to the familiar form:
-\begin{align}
-  \label{eqn:derived_fkpp}
-  \partial_t \xi(t,x) = \mu \left( \frac{\sigma^2}{2} \sum_{k=1}^d \partial_{x_k}^2 \xi(t,x) + s(x) \xi(t,x) (1-\xi(t,x)) \right) .
-\end{align}
-Here we have taken first the population size $N \to \infty$ and then the grid size $\epsilon \to 0$;
-we could alternatively take both limits together, but not if $\epsilon$ goes to zero too much faster than $N$ grows.
-One reason for this is that at finite $N$,
-the process $\Phi_t$ is an irreducible, aperiodic finite-state Markov chain with absorbing states at 0 and 1;
-therefore, the inevitable outcome is extinction of one type or another,
-which is not the regime we want to study.
-
-In one dimension, we are done (see section~\ref{apx:elliptic_integrals});
-in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
-If we take a radially symmetric patch centered at the origin (so $s$ only depends on $r$), 
-and let $\xi(t,r)$ denote the mean occupation frequency at distance $r$,
-then the polar form of the Laplacian in $d$ dimensions gives us that \eqref{eqn:derived_fkpp} is
-\begin{align}
-  \label{eqn:radial_fkpp}
-  \partial_t \xi(t,r) = \mu \left( \frac{\sigma^2}{2} \partial_{r}^2 \xi(t,r) + \sigma^2\frac{d-1}{2r} \partial_r \xi(t,r) + s(r) \xi(t,r) (1-\xi(t,r)) \right) .
-\end{align}
-
-%%%%%%%%%%%
-\subsection[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
-\label{apx:asymptotics}
-
-If $q(r)$ is a radially symmetric equilibrium frequency and $\xi(t,x)=q(|x|)$,
-then $\partial_t \xi(t,x) = 0$.
-So, using \eqref{eqn:radial_fkpp}, a radially symmetric solution 
-with $s(r) = -s < 0$ for all $r>r_0$ 
-solves, for $r_0 < r < \infty$,
-\begin{align} \label{eqn:radial_eqfreq_diffeq}
-    \partial_r^2 q(r) + \frac{d-1}{r} \partial_r q(r) - \frac{2s}{\sigma^2} q(r) (1-q(r)) = 0 \\
-    \lim_{r \to \infty} q(r) = \lim_{r \to \infty} \partial_r q(r) = 0 \qquad 
-    0 < q(r) < 1 
-\end{align}
-Since $q(r) \to 0$ as $r \to \infty$, so $q(r) (1-q(r)) \approx q(r)$,
-it can be shown 
-% theory in \citet{wasow1965asymptotic} implies 
-that the true equilibrium frequency $q$ is close, for large $r$, to the solution to
-\begin{align}
-    \partial_r^2 u(r) + \frac{d-1}{r} \partial_r u(r) - \frac{2s}{\sigma^2} u(r) = 0  \label{eqn:bessel} \\
-    \lim_{r \to \infty} u(r) = \lim_{r \to \infty} \partial_r u(r) = 0 \qquad  .
-\end{align}
-This has general solution given by a modified Bessel function:
-using \citet{gradshteyn2007table} 8.494.9,
-the general solution is
-\begin{align}
-    u(r) = C' (r-r_0)^{(2-d)/2} K_{(2-d)/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
-\end{align}
-where $C'$ and $r_0$ are chosen to match boundary conditions.
-Asymptotics of Bessel functions (\citet{gradshteyn2007table}, 8.451.6) then imply that
-\begin{align}
-    q(r) \approx C r^{(1-d)/2} \exp \left( -r \sqrt{2s}/\sigma \right) + O(1/r),
-\end{align}
-where $C$ is a different constant.
-
-%%%%%%%%% %%%%%%%%%%%%%%
-\subsection[Hitting and Occupation]{Hitting and occupation}
-\label{ss:hitting_occupation}
-
-Here we make a more precise argument to back up expression \eqref{eqn:twocase_migrate} for the migration rate.
-The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
-since it relies only on a decomposition of the migrant families upon hitting the new patch;
-but it is also imprecise in a number of subtle ways
-that are difficult to formalize.
-The argument we make here therefore takes a somewhat different tack,
-supposing that it suffices to model the spatial movement of a migrant family
-by following only the motion of the ``trunk''
-(i.e.\ the red line in figure \ref{fig:branching_decomp}),
-and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
-(Recall $\sigma$ is the dispersal distance.)
-We then use facts about Brownian motion and branching processes, 
-to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
-\linelabel{rr:moved_argument}
-
-We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
-as the sum of independent migrant families, each of whose dynamics are
-given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
-Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
-Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
-such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ 
-that initiate new migrant families is $\gamma(x)$.
-Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``ouflux of families'',
-when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
-By examining the constituent terms,
-it will turn out that equation \eqref{eqn:gestalt_migrate} 
-is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.
-
-First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
-Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
-and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
-The \emph{mean occupation density} of $Z$ at location $y$
-can be thought of  informally \linelabel{rr:density}
-as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
-It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
-i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
-gives the expected total number of individuals ever living in $S$:
-\begin{align} \label{eqn:occupation_density_defn}
-    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
-\end{align}
-% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
-Since reproduction and spatial motion are independent,
-$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
-multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
-i.e.\
-\begin{align}
-    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
-\end{align}
-The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
-is the sum of mean occupation densities of a constant outflux of branching processes
-from $\partial B(r_0)$:
-\begin{align} \label{eqn:occupation_integral}
-    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
-         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
-\end{align}
-
-This form we can now compare to 
-the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
-Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
-We assume this new patch is circular, 
-has relatively small area $A$ and is centered at position $y$.
-% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
-% multiplied by the probability that the family reaches the patch, given survival.
-% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
-It would be possible to analyze this probability directly,
-as in \citet{barton1987establishment};
-but here we take a simpler route, 
-approximating this by the chance that the trunk hits the new patch,
-multiplied by the chance that at least one member of the family escapes demographic stochasticity
-and successfully establishes in the new patch.
-Write $h_A(x,y)$ for the probability that the Brownian trunk
-hits the new patch centered at $y$ before the family dies out,
-and $f(A)$ for the chance that the family manages to establish in the new patch,
-given that it successfully arrives.
-As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
-\begin{align}\label{eqn:migrate_integral}
-  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
-\end{align}
-Recall that $\rho$ appears here but not in \eqref{eqn:occupation_integral}
-because $q(x)$ is the local \emph{proportion} of $B$ alleles.
-
-Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
-while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
-Since the expected total time spent in the patch does not depend on the distance at which the family started
-if we know that it survived to hit the patch,
-it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
-Following this intuition,
-and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-it turns out (see section~\nameref{ss:hitting_occupation}) that
-\begin{align} 
-    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
-\end{align}
-where $g(A)$ is the total expected occupation time of the patch given the family arrives.
-Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
-Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
-that (at least asymptotically) does not depend on the distance between $x$ and $y$,
-we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
-The probability $f(A)$ that the family does not die out once it arrives at the new patch is
-approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
-if each member arriving has chance $p_e$ of establishing, independently.
-\plr{And if $p_e$ is small?}
-% If we let $w$ be the width of the patch, then the remaining term is
-% \begin{align}
-%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
-% \end{align}
-Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{ss:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
-we obtain that XXX FIX
-\begin{align}
-  \label{eqn:orig_migrate} 
-  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
-  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
-\end{align}
-This does not depend on $A$, because although
-$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
-the dependence is very weak.
-For instance, the width of a circular patch only very weakly affects the chance 
-of establishment of a new migrant that appears on its edge,
-as long as the patch is large enough.
-For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{ss:hitting_occupation}.
-
-Consider a single migrant family $Z$ beginning at $x$,
-and write $S$ for the new patch, which has area $A$ and is centered at $y$.
-Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
-and $\tau_\dagger$ an independent Exponential($s_m$) time.
-The mean occupation time of $Z$ spent in a region $S$ is,
-since the marginal distribution of a single lineage is Brownian,
-\begin{align}
-    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
-           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
-        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
-        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
-        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
-\end{align}
-where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
-and likewise $\E^x$.
-If we define
-$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
-and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\min(t,\tau_\dagger)}$,
-by the strong Markov property this is equal to
-\begin{align}
-    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
-           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] ,
-\end{align}
-where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
-provided $S$ is not pathological.
-If $S$ is circular with area $A$, the latter expectation will not depend on $x$;
-call this $g(A)$.
-
-Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
-(In the main body, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
-If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
-and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
-then this is equal to
-\begin{align}
-    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
-           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
-           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
-           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
-    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
-\end{align}
-Therefore, we have that
-\begin{align}
-    h(x,S) \E[K] 
-    \approx \P^x[ \tau_S < \tau_\dagger ] 
-    = u(x,S) / g(A) .
-\end{align}
-
-If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
-then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
-\begin{align}
-    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
-          &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
-          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) . \label{eqn:q_migrate_relation}
-\end{align}
-The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$
-is expected to spend inside the disk before $\tau_\dagger$.
-This is integral of the Green function for the Bessel process of the appropriate order, 
-so using \citet{borodin2002handbook}, % (Appendix 1),
-and letting $w$ be the width of the patch,
-in $d=1$,
-\begin{align}
-    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
-  &= \int_0^{2w} \frac{ e^{- y \sqrt{2s_m} }}{\sqrt{2s_m}} dy \\
-  &= (1-e^{-2w\sqrt{2s_m}})/(2s_m)
-\end{align}
-and in $d=2$, by \citet{gradshteyn2007table} 5.56.2,
-\begin{align}
-    g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
-    % &= \int_0^{2w} G_{s_m}(0,y) dy \\
-  &= \int_0^{2w} 2 y K_0(y \sqrt{2s_m}) dy \\
-  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}} y K_0(y) dy \\
-  &= \frac{1}{2s_m}\left(1- 2w\sqrt{2s_m} K_1(2w\sqrt{2s_m}) \right)  .
-\end{align}
-In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.
-
-%%%%%%% %%%%%%%%%%%%
-\subsection[Outflux of Families]{The outflux of families}
-\label{apx:outflux}
-
-It is of interest to have an expression for the ``outflux of families'',
-$\gamma(r_0)$, 
-defined to be the mean proportion of $B$ alleles at distance $r_0$
-whose lineage traces back to the original patch without venturing further away than $r_0$.
-By expression \eqref{eqn:occupation_integral},
-\[
-    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
-\]
-for any $y$.  
-Take $y=(r,0)$.
-Note that after scaling by $\sqrt{2s_m}/\sigma$,
-\begin{align}
-    \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\|} }{ \|y\|^{-(d-1)/2} e^{-\|y\|} } \\
-        &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r (\|\hat x - x/r\| - 1)} .
-\end{align}
-Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
-so that the expression in the exponent is to, first order in $1/r$,
-\begin{align}
-    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
-            &= - 2 r_0 \cos \theta + O(1/r) .
-\end{align}
-This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
-\begin{align}
-    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} .
-\end{align}
-In one dimension we are done.
-In two dimensions,
-by bounded convergence and \citet{gradshteyn2007table} (3.339),
-\begin{align}
-  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-            &= C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
-                         &= 2 \pi C' I_0(-2r_0),
-\end{align}
-where $I_0$ is a modified Bessel function of order zero of the first kind.
-The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2s_m} r_0 / \sigma ) $.
-
-%%%%%%%%%%%
-
-\bibliography{standing_patches_refs}
-
-\appendix
-
-\setcounter{secnumdepth}{2}
-
-%%%%%%%
-\section{The exact solution to the equilibrium frequency}
-\label{apx:elliptic_integrals}
-
-Here we describe how to solve \eqref{eqn:radial_eqfreq_diffeq} exactly in one dimension,
-i.e.\ the Fisher-KPP equation with piecewise constant selection.
-Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
-so that 
-\[
-\partial_t p(t,x) = \frac{\mu}{2} \sigma^2 \partial_x^2 p(t,x) + \mu s(x) p(t,x) (1-p(t,x)) .
-\]
-The stable distribution $\phi(x) = \lim_{t\to\infty} p(t,x)$ then solves
-\begin{align} \label{eqn:definingphi}
-    \partial_x^2 \phi(x) = - 2 s(x) \phi(x) (1-\phi(x)) /\sigma^2,
-\end{align}
-with appropriate boundary conditions.
-First rescale space by $\sigma/\sqrt{2}$ so the $\sigma^2/2$ term disappears.
-If we now assume that $s(x)$ is piecewise constant,
-$s(x) = s_i$ for $x \in [x_i,x_{i+1})$, with $x_0=-\infty$ and $x_{n+2}=\infty$,
-then the equation is integrable: if we multiply through by $2\partial_x \phi(x)$ and integrate, then we get that
-\begin{align} \label{eqn:conservation}
-    ( \partial_x \phi(x) )^2  &= - \int^{x} 2 s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
-        &= - s_i \phi(x)^2 \left( 1 - \frac{2}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) ,
-\end{align}
-if we define
-\begin{align} 
-  K_i = - s_i \phi(x_i)^2 \left( 1 - \frac{2}{3} \phi(x_i) \right) - ( \partial_x \phi(x_i) )^2  .
-\end{align}
-For ease of reference, we define
-\[
-        V_i(\phi) = s_i \phi^2 \left( 1 - \frac{2}{3} \phi \right) + K_i .
-\]
-Note that $V_i'(0)=V_i'(1)=0$, that $V_i(0)=K_i$ and $V_i(1) = K_i+s_i/3$.  
-We will always have that $V(\phi) \le 0$.
-(We have then that $\partial_x^2 \phi = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
-Also note that this implies ``conservation of energy'', i.e.\ $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
-Rearranging, we get that $dx = d\phi / \sqrt{-V(\phi)}$, so where $\phi(x)$ is monotone, the inverse is
-\[
-    x(\phi) = x(\phi^*) \pm \int_{\phi^*}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } .
-\]
-For each $i$ then define the elliptic function
-\begin{align}  \label{eqn:elliptic_function}
-    F_i(\phi) = \int_{\phi_i^*}^\phi \frac{ d\psi }{ \sqrt{ -V_i(\psi) } } ,
-\end{align}
-where take the positive branch of the square root, and $\phi_i^*$ will be chosen later.
-Then we have that
-$x(\phi) - x(\phi_0) = \pm( F(\phi) - F(\phi_0))$,
-or for an appropriate $x_0$,
-\[
-    \phi(x) = F^{-1}\left( F(\phi(x_0)) \pm (x - x_0) \right).
-\]
-
-We clearly want $\lim_{x \to \infty} \partial_x \phi(\pm x) = 0$, 
-and $\lim_{x \to \infty} \phi(\pm x)$ to be zero or one depending on the sign of $s_0$ and $s_{n+1}$.
-Since $(\partial_x \phi(x))^2 = -V(x)$, this implies that if $s_0<0$, then $K_0 = 0$,
-while if $s_0>0$ then $K_0 = s_0/6$; and likewise for $K_{n+1}$.
-
-We also require that $\phi(x)$ and $\phi'(x)$ are continuous.
-Continuity of $\phi'(x)$ is equivalent to $V_i(x_{i+1}) = V_{i+1}(x_{i+1})$,
-which we can rearrange to find an equation for $K_{i+1}$ in terms of $K_i$ and $\phi(x_{i+1})$:
-\[
-    K_{i+1} - K_i = (s_i - s_{i+1}) \phi(x_{i+1})^2 (1-2\phi(x_{i+1})/3) .
-\]
-What about the frequency at the points the selection changes, $\phi(x_i)$?  
-Well, if $\phi(x)$ is monotone on $[x_i,x_{i+1})$ then we can without loss of generality take $\phi_i^*=0$ or $1$ depending on the sign of $s_i$.
-Otherwise, let $\phi_i^*$ be the (unique) root of $V_i$ in $[x_i,x_{i+1})$, so $V_i(\phi_i^*)=0$.
-In this case, $\phi_i^*$ is the maximum or minimum of $\phi$ in the interval:
-if $s>0$, then $\phi_i = \max\{ \phi(x) : x \in [x_i,x_{i+1})\}$. 
-Recall we defined $F_i$ using $\phi_i^*$; now using the fact that $\phi$ is monotone with the opposite sign on either side of $\phi_i^*$,
-$x_{i+1} - x_i = F_i(\phi(x_{i+1})) - F_i(\phi_i^*) + F_i(\phi(x_{i})) - F_i(\phi_i^*)$,
-and that $F(\phi_i^*) = 0$,
-we know that the length of the $i$th stretch is
-\[
-    x_{i+1} - x_i = \pm \left( F_i(\phi(x_i)) + F_i(\phi(x_{i+1})) \right).
-\]
-Note that all the $\pm{}$'s are easily relatable to the signs of $s_i$.
-If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be able to solve the equations for $\phi(x_i)$ and $K_i$ recursively upwards.
-In some cases, such as \citet{slatkin1973geneflow}, we can infer $\phi(0)$ and $\partial_x \phi(0)$ by spatial symmetry.
-In other cases, we are only given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
-
-\subsection{Doing the integrals}
-\label{apx:some_integrals}
-
-An important ingredient in the above method is the integral \eqref{eqn:elliptic_function},
-to which a method for solving the Korteweg-de Vries equation can be applied \citep{NEQwiki}.
-Recall that $V_i(\phi) =  s_i \phi^2(1-2\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries;
-since we're just working within an interval with $s$ constant, we can rescale space by $\sqrt{2 s_i}/\sigma$,
-so that $s$ and $\sigma$ drop from the equation.
-We then want to integrate
-\[
-    \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
-         \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } ,
-\]
-over a domain where $V$ is always negative.
-Let $\phi^2(1-2\phi/3)+K = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
-and change variables first to $y^2=(\alpha-\phi)$, 
-and then to $x = y/\sqrt{\alpha-\beta}$, so that
-\begin{align*}
-    \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } 
-        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-y^2) } } \\
-        = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
-\end{align*}
-with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
-Now Jacobi's incomplete elliptic integral of the first kind is defined by
-\[
-    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) }} ,
-\]
-and the Jacobian elliptic function $\sn(x;k)$ is the inverse: $F(\sn(x;k);k) = x$.
-As $k \to 0$, $\sn(x;k) \to \sin(x)$, while as $k \to 1$, $\sn(x;k) \to \sinh(x)$.
-
-% Presumably, this gets the same answer as \citep{slatkin1973geneflow}, who finds that in the solvable cases,
-% for $x$ in the last interval (which ends in $\infty$), that
-% \begin{align*}
-%     \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / 2 \sigma + C ) \right) \\
-%         &\simeq C' \exp\left( - \frac{ x \sqrt{2s} }{ \sigma } \right) .
-% \end{align*}
-
-%%%%%%%
-\section{More information about the simulations}
-\label{apx:parameter_tables}
-
-Here we provide additional details regarding the simulations we used
-to validate the theory in figures \ref{fig:sim_times} and \ref{fig:sim_probs}.
-
-% table of parameters, produced by analyze-sims.R
-\input{mutation-parameters-table.tex}
-
-% table of parameters, produced by analyze-sims.R
-\input{migration-parameters-table.tex}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics{mutation-times-predicted}
-  \end{center}
-  \caption{
-    The same data shown in the left panel of figure~\ref{fig:sim_times},
-    but all times shown (not just the interquartile ranges),
-    and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
-  } \label{sfig:sim_migration_times}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics{migration-time-predicted}
-  \end{center}
-  \caption{
-    The same data shown in the right panel of figure~\ref{fig:sim_times},
-    but all times shown (not just the interquartile ranges),
-    and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
-    The left panel has the predicted time to adaptation on the horizontal axis as in figure~\ref{fig:sim_times},
-    and the right panel has, for comparison, the raw distance between patches 
-    (which predicts time to adaptation, but not as well).
-  } \label{sfig:sim_mutation_times}
-\end{figure}
-
-% example trace plots
-
-% sm=0.1
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/18885-r1-501-sb0_01-sm-0_1-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/56325-r1-501-sb0_01-sm-0_1-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/28432-r1-501-sb0_01-sm-0_1-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.1$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_1}
-\end{figure}
-
-% sm=0.01
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/43099-r1-501-sb0_01-sm-0_01-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/69787-r1-501-sb0_01-sm-0_01-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/11821-r1-501-sb0_01-sm-0_01-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.01$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_2}
-\end{figure}
-
-% sm=0.001
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/59611-r1-501-sb0_01-sm-0_001-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/93713-r1-501-sb0_01-sm-0_001-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/29850-r1-501-sb0_01-sm-0_001-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_3}
-\end{figure}
-
-% sm=0.0001
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/15449-r1-501-sb0_01-sm-1e-04-N1600-pophistory-run}
-    \includegraphics{example-mutation-sims/5582-r1-501-sb0_01-sm-1e-04-N400-pophistory-run}
-    \includegraphics{example-mutation-sims/24639-r1-501-sb0_01-sm-1e-04-N25-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_4}
-\end{figure}
-
-% sm=0.1, R=80
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-migration-sims/89826-r1-501-sb0_01-sm-0_1-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/76178-r1-501-sb0_01-sm-0_1-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.1$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_5}
-\end{figure}
-
-% sm=0.01, R=80
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-migration-sims/42080-r1-501-sb0_01-sm-0_01-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/82698-r1-501-sb0_01-sm-0_01-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.01$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_6}
-\end{figure}
-
-% sm=0.001, R=80
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-migration-sims/88079-r1-501-sb0_01-sm-0_001-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/3464-r1-501-sb0_01-sm-0_001-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/37774-r1-501-sb0_01-sm-0_001-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_7}
-\end{figure}
-
-% sm=0.0001, R=80
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-migration-sims/26044-r1-501-sb0_01-sm-1e-04-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/85750-r1-501-sb0_01-sm-1e-04-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/4111-r1-501-sb0_01-sm-1e-04-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.0001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_8}
-\end{figure}
-
-\ifsubmissionversion
-\processdelayedfloats
-\fi
-
-%%%% Reviews
-\includereviews
-
-\end{document}
-
-%%%%%%%% END %%%%%%%%%%%%%%%
-%%%%%%%% %%% %%%%%%%%%%%%%%%
-
-\begin{figure}[ht!]
-    \begin{center}
-        % made in patchy-sim-writeup-plots.R
-        \includegraphics{example-equilibrium}
-    \end{center}
-    \label{fig:example_equil}
-    \caption{
-    Snapshots of a simulation of allelic frequency dynamics around a patch where the allele is beneficial:
-    the focal allele is benefical in the middle region and deleterious outside.
-    The dynamics were simulated for 10000 generations across demes with 1000 individuals in each
-    and random dispersal to nearby demes with a mean dispersal distance of $\sigma \approx 2$ demes.
-    Colored lines show allele frequencies across space in 50 uniformly spaced generations;
-    the solid line shows the mean across all generations.
-    Below is the same plot, except on a log scale.
-    }
-\end{figure}
-
-\section{Temporary work: hitting probs}
-
-Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
-which we assume has relatively small area $A$ and is centered at position $y$.
-The probability that this family reaches $A$ by time $t$ is $(1-k_e(t))$
-multiplied by the probability that the family reaches the patch, given survival.
-If $t$ is large then the previous section argues that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
-will have hit the patch by then.
-If $w = A^{1/d}$ is the width of the patch, then 
-this is the chance that a $d$-dimensional Bessel process begun at $\|y-x\|$ hits $w$ by time $t$.
-Let $\tau_w$ be the hitting time of $w$,
-so, integrating by parts, and using Borodin \& Salminen (2.2.0.1 and 4.2.0.1)
-\begin{align}
-    \int_0^\infty (1-k_e(t)) \P^x\{ \tau_w \le t \} dt  
-    &\approx \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\
-    &= \frac{1}{s_m \E[K]} \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \in dt \} \\
-    &= \frac{1}{s_m \E[K]} \E^x[e^{-s_m \tau_w}] \\
-    &= 
-    \frac{1}{s_m \E[K]} \times
-    \begin{cases}
-        e^{-(x-w) \sqrt{2s_m}} \qquad & d=1 \\
-        \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } \qquad & d=2 
-    \end{cases}
-\end{align}
-Note that the $d=2$ case
-\begin{align}
-    \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } = \sqrt{\frac{w}{x}} e^{-(x-w)\sqrt{2s_m}} ( 1 + O(1/w) ).
-\end{align}
-
-\section{Patchy selection} 
-\label{ss:discretedemes}
-
-We have been working in a continuous model of geographic space; 
-in this section we move to a discrete model of isolated populations exchanging rare migrants,
-but in section \ref{ss:patchyspace} discuss how a model with {\em patchy} selection can be approximated by such a discrete model.
-
-Suppose now that we have a set of $L$ discrete populations of individuals that may reproduce or migrate,
-and that migration and mutation are rare, so that the waiting times until either event is approximately exponential.
-Suppose that for each pair of populations, labeled $i$ and $j$, the mean number of migrants that travel from $i$ to $j$ per generation
-is $M_{ij}$ and that the mean number of new mutations appearing in population $i$ per generation is $\mu_i$.
-
-{\tt Include cartoon here.}
-
-If we suppose that fixation occurs on a faster time scale than mutation or migration,
-so that it is very unlikely that two different migrants or mutants begin to fix locally in the same population,
-then we can treat the process of extinction or local fixation as instantaneous 
-(a limit previous studied by \cite{Slatkin:81}).
-In this limit we may also ignore swamping effects of inmigration of nonmutant alleles.
-
-In this case, we have the following picture.
-The population at first starts out with no selected alleles. 
-At some later time point suppose that there are $T$ different alleles present in the population, 
-that the $k^\mathrm{th}$ allele is occupying demes $\{i^k_1, \ldots, i^k_{n_k}\}$, for $1\le k \le T$.
-Let $I$ denote the set of all adapted demes, 
-and let $J = \{j_1, \ldots, j_\ell\}$ denote the demes that have not yet adapted, with $\ell = L - \sum_k n_k$.
-Then we are assuming that two types of event can happen which could change the state:
-either at rate $\mu_j p_f$, an unadapted deme $j$ produces a new mutant, which fixes in this deme;
-or at rate $M_{ij} p_f$, an adapted deme $i$ sends a migrant to unadapted deme $j$, which fixes in that deme.
-Therefore, under these assumptions, the probability of fixation $p_f$ (and hence the selection coefficient)
-only enter as a time scaling: for instance,
-if we define $R = \sum_{j \in J} \left( \mu_j + \sum_{i \in I} M_{ij} \right)$,
-then the total rate of such events is $p_f R$.
-The probability that the next event results in, say, adaptation of deme $j$
-is 
-\[
- \left( \mu_j + \sum_{i \in I} M_{ij} \right) / R,
-\]
-and the probability that deme $j$ adapts through migration from deme $i$, given that deme $j$ is the next to adapt, is
-$M_ij / R$,
-while the probability that it adapts through a new mutation is $\mu_j / R$.
-
-The main point here is that the final pattern,
-and hence the likely extent of parallel adaptation,
-is independent of the strength of selection.
-This is in contrast to the continuous case, however, this observation is dependent on the assumption of weak migration
-and fails to hold if different migrant lineages interact.
-This can be viewed as a continuous-time Markov process XXX.
-Explicit solutions are not available in the general case,
-but if we specialize to a completely symmetric model (i.e.\ the ``island'' model),
-then the mathematical picture is a pretty one.
-
-\subsection{Symmetric patches}
-
-Indeed, suppose that migration and mutation are symmetric: $M_{ij}=m$ and $\mu_j = \mu$ for all $j$ and $i\neq j$.
-Suppose at some time the $k^\mathrm{th}$ allele is occupying $n_k$ demes, 
-with $n = \sum_k n_k < L$.
-Then at rate $(L-n) N \mu p_f + n (L-n) N m p_f$,
-either a migration or mutation event happens.
-With probability $\mu/(\mu + n m)$, it is a mutation,
-and a new deme is randomly chosen and assigned a new allele.
-With probability $n_k m / (\mu + n m)$,
-the $k^\mathrm{th}$ type sends a successful migrant to a randomly chosen new island,
-which is assigned the $k^\mathrm{th}$ allele.
-
-This process is a continuous-time version of the ``Chinese restaurant process''
-described in \citet{aldous1985exchangeability} and \citet{pitman1995partitions},
-and so the final partition of types across demes has the Ewens sampling distribution with parameter $\mu/m$.
-Again note that the selection coefficient, which is implicit in the probability $p_f$,
-does not enter into the final distribution.
-
-Now we can compute most properties we might want about the process.
-For instance, the expected number of distinct mutational origins is
-\begin{equation}
-    \E\left[ \mbox{ \# of independent mutations } \right] = 
-            1 + \frac{\mu}{m} \sum_{k=2}^L \frac{1}{k+(\mu/m)}. \label{discrete_expected}
-\end{equation}
-The probability that there is only a single successful mutation is
-\begin{equation}
-\P \left\{ \mbox{only one mutational origin} \right\} = 
-            \prod_{k=1}^{L-1} \frac{ k }{ k+(\mu/m)} .
-\end{equation}
-More generally, suppose we sample a single deme, and are interested in the total number of demes $S$
-(including the sampled one) that share the same fixed mutation as our sampled deme.
-Then $S$ has distribution
-\begin{equation} \label{eqn:Sdistrn}
-\P\{ S=s \} = \frac{ (s-1)! (\mu/m)^{L-s} }{ \prod_{k=1}^{L-1} (k+(\mu/m)) } .
-\end{equation}
-Furthermore, if $L$ is large, then $S/L$ has probability density approximately
-\begin{equation} \label{eqn:betadistrn}
-   (\mu/m) (1-x)^{(\mu/m)-1}
-\end{equation}
-namely, is a $\mathrm{Beta}((\mu/m), 1)$ distribution--- see \cite{donnelly1989continuity} and \cite{perman1992sizebiased}.
-
-% In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
-% for a model with 20 interconnected demes, across different parameter values.  
-% This indicates that we need the population-scaled mutation and migration rates to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.
-If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
-
-The high connectedness of the discrete deme island model means that the expected number of distinct alleles, (Equation \eqref{discrete_expected}), 
-grows with the $\log$ of the number of demes.
-This strongly contrasts with the continuous spatial model where the local nature of dispersal means that doubling the species range will double the number of mutations expected.
-Furthermore, while in the continuous model the areas occupied by distinct mutations are similar in area, in this strong-selection discrete island model, 
-only a few types tend to dominate even as the number of demes (and thus, distinct types) grows large (see Equations \eqref{eqn:Sdistrn} and \eqref{eqn:betadistrn}).
-
-A more general model would allow migration only along edges of some graph connecting the demes.
-In the low migration limit such a model still produces a partition distribution independent of the selection coefficient,
-but it does not in general have the Ewens distribution.
-Another extension would be to include the time alleles need to achieve an intermediate frequency,
-along the lines of \citet{Navarro:03}, which would reintroduce dependence on the selection coefficient.
-
-\texttt{XXX omit this? XXX}
-We have treated the time during which the selected allele is at intermediate frequency in a deme (about $\log(\sigma^2 \rho)/s_p$ generations) as negligible, 
-which ignores migration events that occur while the mutation is spreading through the population. 
-we now sketch of a more realistic approximation.
-For each occupied--unoccupied pair of demes, if the number of mutant alleles in the occupied deme is $n(t)>0$,
-then at time $t$, migrants from the occupied deme pass the selected allele to the unoccupied deme at rate $2 n(t) m p_f(s_p)$,
-while newly arisen mutants arise and fix in the unoccupied deme at constant rate $2 N \mu p_f(s_p)$.
-The selected type should grow approximately as $n(t)=\exp(s(t-t_0))$.
-Of course, to follow this model through, we will have to account for multiple types within a single deme and other complications,
-but we can at least make a few observations.
-One is that the final distribution of types is no longer independent of $s_p$, which enters through the dynamics $n(t)$.
-However, it is clear that increasing the selection coefficient $s$ will decrease the number of types (independent mutational origins),
-while on the other hand, increasing the deme size $N$ will increase the number of types,
-and that the case presented above is at the extreme (see also \cite{Navarro:03} for deterministic approximations).

### 2014-11-29 patchy-selection-paper.tex
coop changes

--- 2014-11-27 patchy-selection-paper.tex
@@
 
-%%%%%
-% Responses to reviews:
-\input{review-response-commands}
-% set this to show line numbers and include responses to reviews or not
-\newif\ifreviewresponses
-\reviewresponsestrue  % include them
-% \reviewresponsesfalse  % don't include them
-\ifreviewresponses\else\renewcommand{\linelabel}[1]{}\fi
-\iflatexml\reviewresponsesfalse\fi  % they don't work so well in html
-% \newcommand{\responsefile}{pbio-reviews-19sept12-responses.tex}  % name of the review reponses file
-\newcommand{\responsefile}{patchy-review-responses.tex}  % name of the review reponses file
+%We should cite Papers by Slatkin and Wiehe and this, on spatial hitchhiking http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3176130/
 
@@
 \newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
-\newcommand{\given}{\,\vert\,}
 
@@
 (i.e.\ $\sigma$ is the dispersal distance). 
-We will deal with migration by immediately appealing to the central limit theorem,
-treating the total distance traveled across $t$  dispersal events as Gaussian with variance $t \sigma^2$,
-and do not discuss the (interesting) cases where rare, long-distance dispersal events
-are more important \citep[see e.g.][for discussion]{levin2003ecology,ralph2010parallel,hallatschek2014acceleration}.  \linelabel{rr:longtailed}
 %We also aim to describe the scenario where it is sensible to think of large, discrete patches;
@@
 which will be a good approximation if the patch is large relative the scale over which $p(x)$ goes from 0 to $p_e$,
-or if $p(x)$ is close to $p_e$ at some point and is symmetric about the edge of the patch. \linelabel{rr:p_estab}
+or if $p(x)$ is symmetric about the edge of the patch.
 We examine this more generally via exact calculation of $p(x)$ in the section \nameref{apx:establishment_sims}.
@@
 the time scale on which new patches become colonized by the $B$ allele from new mutation is therefore approximately $1/\mutrate$.
+
+%%%%%%%
+\subsection{Simulation methods}
+\label{ss:simulations}
+
+First we briefly describe the simulations used below for illustration and validation.
+We simulated forward-time dynamics of the number of alleles of each type in a regular grid of demes with fixed size $N$.
+Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
+locally beneficial alleles were more likely to reproduce.
+Each then produced a random, Poisson distributed number of offspring that each independently migrated to demes no more than six steps away
+according to a symmetric migration kernel.
+Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
+(Although we described the simulation in terms of individuals,
+we kept track only of total numbers in an equivalent way.)
+
+The reproduction rate in simulations for type $b$ alleles was 30\%; 
+this was then multiplied by $1+s$ to get the probability of reproduction for type $B$.
+These $s$ are the values reported in the figures, and do not depend on the basic rate of reproduction.
+However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
+we computed the rate of intrinsic growth, 
+i.e.\ the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
+in the absence of migration.
+(The resulting values are close to the first notion of $s$,
+but give better agreement with theory,
+which uses the second definition.)
+
+To sample lineages, we first simulated forwards in time the population dynamics,
+then sampled lineages moving back through time
+by, in each generation,
+moving each lineage to a new deme with probability proportional to the reverse migration probability
+weighted by the number of $B$ alleles in that deme in the previous time step.
+If more than one lineage was found in a deme with $n$ $B$ alleles,
+then each lineage picked a label uniformly from $1 \ldots n$,
+and those picking the same label coalesced.
+Since reproduction is Poisson, this correctly samples from the distribution of lineages given the population dynamics.
 
@@
 and $C$ is a constant depending on the geographic shape of the populations and the selection coefficients.
+These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
+To be clear about the assumptions (especially regarding scaling) implicit here,
+we provide a derivation in section~\nameref{apx:eqfreq}, 
+as well as justification for the asymptotics in section~\nameref{apx:asymptotics}.
+In one dimension, the equation can be solved to give $q(x)$ 
+in terms of Jacobi elliptic functions; see appendix \ref{apx:elliptic_integrals}.
 In applications we fit $C$ to the data;
 to get concrete numbers from theory we take $C=1$ if necessary.
-
 From simulations, $C$ seems to be not too far from 1 in practice (see below).
-As J.\ Hermisson pointed out in an earlier draft,
-this functional form has a simple intuition:
-local migration leads to the exponential decay,
-since if each migrant at distance $x$ 
-produces an average of $c$ descendants that make it to distance $x+1$, 
-then in one dimension, the number of migrants must decay as $\exp(-cx)$;
-and the square-root term in two dimensions 
-enters because the available area grows with $x$.  
-The ``renewal'' aspect of this same argument suggests that for equation~\eqref{eqn:eqfreq} to hold,
-$|x|$ must be larger than only 5--10 multiples of $\sigma$.
-\linelabel{rr:q_form}
-
-These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
-To be clear about the assumptions (especially regarding scaling) implicit here,
-we provide a derivation below in \nameref{apx:eqfreq}, 
-as well as justification for the asymptotics below in \nameref{apx:asymptotics}.
-In one dimension, the equation can be solved to give $q(x)$ 
-in terms of Jacobi elliptic functions; see appendix \ref{apx:elliptic_integrals}.
 
@@
   with $C$ chosen arbitrarily to give a reasonable fit.
-  \plr{what is $C$?}
   (The disagreement at long distances is due to boundary effects near the edge of the range.)
@@
 \end{align}
-The quantities on the right-hand side depend implicitly on $r_0$,
-but their product does not.
+The quantities on the right-hand side depend implicitly on $r_0$.
 All terms except the ``outflux of families'' are calculations with spatial branching processes;
-and we can solve for this outflux using the established form for $q(x)$.
+and we can solve for this outflux using the established form for $q(x)$ in \eqref{eqn:gestalt_q}.
 % Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
@@
 \subsubsection{The genealogy of migrant families}
-\label{ss:migrant_math}
 
@@
 and finally pruning any branches that happen to wander into a patch.
-By our definition of fitness, 
-in this tree the mean number of offspring per individual is $e^{-s_m} < 1$.
+By our definition of fitness, the mean number of offspring per individual is $e^{-s_m}\approx 1-s_m$,
+and the variance is $\xi^2$.
 This pruning removes both branches leading back into the already established patch (which are irrelevant)
@@
 then by Bayes' rule, $e^{-s_m t} = (1-k_e(t))\E[K_t]$.
-It turns out that if the distribution of offspring numbers is not too heavy-tailed 
-(see \citet{jagers1975branching} for details),
+It turns out that if the distribution of offspring numbers is not too heavy-tailed (see \citet{jagers1975branching} for details),
 then $K_t$ has a limiting distribution: $K_t \dconv K$,
-and 
-\begin{align} \label{eqn:EK_limit}
-    e^{-s_m t}/(1-k_e(t)) \to \E[K] \quad \text{as} \quad t \to \infty .
-\end{align}
+and $e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
 In other words, the chance of survival to time $t$ is a constant multiplied by $e^{-s_m t}$,
@@
 then this gives the following decomposition:
-\begin{align} \label{eqn:Zt_decomposition}
+\begin{align}
   \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
@@
 that arose from the main branch in generation $s$,
-and each $W_s$ is an independent copy of the offspring distribution $W$, conditioned on $W_s \ge 1$.
-Since each subfamily $Z^{(s,k)}$ is a branching process with mean offspring number less than one,
-most of these die out quickly,
-and so $Z_t$ is composed of a cluster of recently split subfamilies.
-If we take the expectation of \eqref{eqn:Zt_decomposition},
-we get the limiting mean family size conditioned on nonextinction:
-\begin{align} \label{eqn:mean_K}
-    \E[K] = 1 + \E[W \given W \ge 1] /(e^{s_m}-1) \approx \E[W \given W \ge 1] / s_m .
-\end{align}
+and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.
 
@@
 
-%%%% %%%%
-\subsubsection{Heuristics}
-\label{ss:heuristics}
-
-The parallel form suggested by \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
-suggests that we can find $\migrate(x)$ by multiplying the mean density $q(x)$
-by the ratio of the second terms in the two equations,
-i.e.
-\begin{align} \label{eqn:ratio_gestalt}
-  \migrate(x) 
-    &= \rho q(x) \times 
-    \frac{
-        \text{ ( probability a family establishes in patch at $x$ ) }
-    }{
-        \text{ ( mean occupation density of a family at $x$ ) }
-    } .
-\end{align}
-To make this precise we need to understand what we mean by ``outflux of families'';
-we do this below in \nameref{ss:hitting_occupation},
-and provide a heuristic argument here.
-
-Both numerator and denominator in \eqref{eqn:ratio_gestalt} 
-count only families that have arrived at the new patch,
-so we can consider a family conditioned on having successfully transited between the patches.
-To obtain simply interpretable expressions,
-we assume that once any member of the family arrives at the patch,
-all other members are close by, and so have roughly equal opportunity
-to establish in the new patch.
-
-Suppose the family arrives with $K$ individuals,
-i.e.\ there are $K$ individuals that manage to establish
-The probability that at least one individual leaves a lineage that establishes in the patch
-is $1-(1-p_e)^K$.
-On the other hand, in the absence of the patch of new habitat,
-each of the $K$ newly arrived individuals would leave behind a lineage with 
-mean total size $\int_0^\infty e^{-t s_m} dt = 1/s_m$,
-roughly half of which would stay in the new patch.
-Therefore, the expected contribution of the family to the occupation density is $K/2s_m$.
-(Readers who feel uncomfortable about the geometry of the patch,
-or the distinction between occupation density and number of individuals
-are encouraged to take a detour through \nameref{ss:hitting_occupation}, below.)
-
-This suggests that 
-\begin{align} \label{eqn:ratio_K}
-  \migrate(x) 
-    &= \rho q(x) \times 
-    \frac{
-        \E[1-(1-p_e)^K]
-    }{
-        \E[K]/2s_m
-    } .
-\end{align}
-There are now two extremes to consider.
-If $\P\{K>1/p_e\}$ is small, then each family has a small chance of establishment,
-and the numerator is approximately $p_e \E[K]$.
-On the other hand, if $p_e$ is not too small,
-then the numerator would be close to 1, as any family that arrives will likely establish.
-Combining these, and absorbing the term $\E[W|W>1]$ from equation \eqref{eqn:mean_K} 
-into the constant $C$,
-\begin{align} \label{eqn:migrate}
-  \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m} 
-%     \begin{cases}
-%         q(x) \times 4 \rho s_m s_p / \xi^2  \qquad & \text{ if } s_p < s_m \\
-%         q(x) \times 2 \rho s_m^2 \qquad & \text{ if } s_p > s_m 
-%     \end{cases} .
-\end{align}
-When we use this below, we set $p_e = 2 s_p / \xi^2$.
-Note that the function increases for small $s_m$, 
-but decreases if $s_m > |y|^2 / 2 \sigma^2$, 
-i.e.\ where the condition holds.
-\linelabel{rr:sm_condition}
-
-%%%%%%%%%
-\subsection{A review, and a test, of the assumptions}
-\label{ss:assumptions}
-
-A quick review of the parameter regime we are studying here:
-First, we assume that 
-\textbf{(a)} each patch is large enough to support a stable population of $B$ alleles
-($A^{1/d} > (\sigma/\sqrt{2s_p}) \tan^{-1}(\sqrt{s_m/s_p})$).
-Next, in using equation \eqref{eqn:eqfreq}, we assume that 
-\textbf{(b)} the inter-patch distance is at least the characteristic length
-($R > \sigma/\sqrt{s_m}$),
-and that
-\textbf{(c)} local drift is not so strong that the equilibirum frequency is at least approximately attained
-(using Wright's effective neighborhood size, $s_m \gg 1/(\rho \sigma^d)$).
-The last requirement is necessary because if the $B$ allele fixes in large neighborhoods where it is deleterious,
-we cannot approximate its dynamics via a branching process.
-We also neglect the time for migration-selection equilibrium to be reached,
-which has the effect of over-estimating $\migrate$,
-but don't know a good way to estimate this contribution.
-As discussed above, we also assume that migration to a new patch takes place over a number of generations;
-if there are sufficient rare, long-distance migration events
-that would move between patches in a single hop,
-this would require a different analysis.
-
-To test the robustness of our results to the various approximations we used,
-we used individual-based simulations on a one-dimensional lattice of 501 demes,
-run for 25,000 generations.
-More details are given below in \nameref{ss:simulations},
-and the number of simulations used and parameter values are given in supplementary tables \ref{stab:mutation_params} and \ref{stab:migration_params}.
-To estimate the mean time until adaptation by mutation,
-we used one centrally located patch of 99 demes
-and a mutation rate of $\mu=10^{-5}$,
-and to estimate the mean time until adaptation by migration,
-we used two centrally located patches of 99 demes
-separated by a varying number of demes,
-with one patch initialized to have the $B$ allele at frequency 0.8.
-In each case, we measured the ``time to adaptation'' as the amount of time until at least 100 $B$ alleles
-were present in the previously unadapted patch.
-Figure~\ref{fig:sim_times}
-summarizes how the results compare to theory,
-excluding parameter combinations where a majority of simulations did not adapt by 25,000 generations
-or the patch was not large enough to support a stable population of $B$ alleles.
-Supplementary figures \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times} 
-show all times, 
-and depictions of typical simulation runs are shown in supplementary figures
-\ref{sfig:sims_1}, \ref{sfig:sims_2}, \ref{sfig:sims_3}, \ref{sfig:sims_4}, \ref{sfig:sims_5}, \ref{sfig:sims_6}, \ref{sfig:sims_7}, and \ref{sfig:sims_8}.
-
-The agreement is reasonable for both.
-The theoretical value $1/\mutrate$ underestimates the mean time to mutation
-by a factor of around 2 that depends increases with $s_m$.
-This is to be expected for two reasons:
-First, we compute the time to reach 100 $B$ alleles, while theory predicts
-the time until the the progenitor of those 100 $B$ alleles arose.
-Second, the expression for $\mutrate$ neglects effects near the boundary of the patch,
-and the larger $s_m$ is, 
-the harder it is for mutations that arise near the edge of the patch to establish.
-The theoretical values $1/\migrate$ again has a margin of error of a factor of about 2,
-with the exception of the weakest disadvantage off the patch, $s_m=10^{-4}$,
-which adapt much more quickly than predicted by theory.
-However, these simulations fall outside the region where we expect the model to do well,
-since $R < \sigma/\sqrt{s_m} \approx 100$.
-A glance at supplementary figure \ref{sfig:sims_8} shows more:
-since $s_m$ is so small, the second patch is reached by migrants before the migration-selection eqilibrium is reached,
-and at migration-selection equilibrium, the two patches act as a single meta-patch.
-
-\begin{figure}[ht!]
-  \begin{center}
-      % made in analyze-sims.R
-      \includegraphics{times-predicted-observed}
-  \end{center}
-  \caption{
-      \textbf{(left)} 
-      Time until adaptation by mutation in an isolated patch for different parameter combinations,
-      as observed in simulations (vertical axis),
-      and as predicted from equation~\eqref{eqn:mutrate} (horizontal axis).
-      \textbf{(right)} 
-      Time until adaptation by migration from an already occupied patch to a new patch,
-      as observed in simulations (vertical axis),
-      and as predicted from equation~\eqref{eqn:migrate}, with $C=5$ (horizontal axis).
-      For each,
-      parameter combinations and numbers of simulations are in supplemental tables \ref{stab:migration_params} and \ref{stab:mutation_params};
-      shown are the median (points) and interquartile range (lines)
-      of the time until the patch had at least 100 adapted individuals,
-      for each unique parameter combination.
-  }   \label{fig:sim_times}
-\end{figure}
+This suggests that we add in the spatial motion by following only the motion of the ``trunk''
+(i.e.\ the red line in figure \ref{fig:branching_decomp}).
+Its motion is approximately Brownian, with variance $\sigma^2$ per generation.
+(Recall $\sigma$ is the dispersal distance.)
+We can use these two ingredients, Brownian motion and a branching process, 
+to compute the terms of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
+
+%%%%%%% %%%%%%%%%%%
+\subsubsection{The outflux of migrant alleles}
+
+We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
+as the sum of independent migrant families, each of whose dynamics are
+given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
+Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
+Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
+such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ that initiate new migrant families
+is $\gamma(x)$.
+To make expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} more concrete,
+we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
+of the outflux $\gamma$ at that point.
+By examining the constituent terms,
+it will turn out that equation \eqref{eqn:gestalt_migrate} 
+is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.
+
+First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
+Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
+and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
+The \emph{mean occupation density} of $Z$ at location $y$
+can be thought of 
+as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
+It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
+i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
+gives the expected total number of individuals ever living in $S$:
+\begin{align} \label{eqn:occupation_density_defn}
+    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
+\end{align}
+% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
+Since reproduction and spatial motion are independent,
+$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
+multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
+i.e.\
+\begin{align}
+    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
+\end{align}
+The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
+is the sum of mean occupation densities of a constant outflux of branching processes
+from $\partial B(r_0)$:
+\begin{align} \label{eqn:occupation_integral}
+    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
+         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
+\end{align}
+
+This form we can now compare to 
+the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
+Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
+We assume this new patch is circular, 
+has relatively small area $A$ and is centered at position $y$.
+% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
+% multiplied by the probability that the family reaches the patch, given survival.
+% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
+It would be possible to analyze this probability directly,
+as in \citet{barton1987establishment};
+but here we take a simpler route, 
+approximating this by the chance that the trunk hits the new patch,
+multiplied by the chance that at least one member of the family escapes demographic stochasticity
+and successfully establishes in the new patch.
+Write $h_A(x,y)$ for the probability that the Brownian trunk
+hits the new patch centered at $y$ before the family dies out,
+and $f(A)$ for the chance that the family manages to establish in the new patch,
+given that it successfully arrives.
+As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
+\begin{align}\label{eqn:migrate_integral}
+  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
+\end{align}
+
+Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
+while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
+Since the expected total time spent in the patch does not depend on the distance at which the family started
+if we know that it survived to hit the patch,
+it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
+Following this intuition,
+and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
+it turns out (see section~\nameref{apx:hitting_occupation}) that
+\begin{align} 
+    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
+\end{align}
+where $g(A)$ is the total expected occupation time of the patch given the family arrives.
+Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
+Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
+that (at least asymptotically) does not depend on the distance between $x$ and $y$,
+we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
+The probability $f(A)$ that the family does not die out once it arrives at the new patch is
+approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
+if each member arriving has chance $p_e$ of establishing, independently.
+% If we let $w$ be the width of the patch, then the remaining term is
+% \begin{align}
+%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
+% \end{align}
+Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{apx:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
+we obtain that
+\begin{align}
+  \label{eqn:migrate} 
+  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
+  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
+\end{align}
+This does not depend on $A$, because although
+$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
+the dependence is very weak.
+For instance, the width of a circular patch only very weakly affects the chance 
+of establishment of a new migrant that appears on its edge,
+as long as the patch is large enough.
+For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{apx:hitting_occupation}.
 
@@
 for the effective rates of adaptation by new mutation and by migration.
-We work in one dimension, as the square root term appearing for two dimensions is relatively unimportant 
-(as seen in figure~\ref{fig:sim_occupation_freqs} by the fact that the $d=2$ curve is nearly straight).
-\linelabel{rr:sqrt_term}
-
-Let $\gamma = \min(1,s_m/p_e)$.
+We work in one dimension, as the square root term appearing for two dimensions is relatively unimportant (see figure~\ref{fig:sim_occupation_freqs}).
+
 An important point to notice is that effective migration and mutation rates will be on the same order if 
-$A \mu \approx 2 \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
+$A \mu \approx 2 s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
 where $R$ is the distance between the patches,
@@
 In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
-is $2 \rho \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
+is $2 \rho s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
 which depends fairly strongly on the selective disadvantage $s_m$ between patches.
@@
 Equivalently, the rates are roughly equal if
-$R/\sigma = \log(2 \gamma s_m/(A \mu))/\sqrt{2s_m}$,
+$R/\sigma = \log(2 s_m/(A \mu))/\sqrt{2s_m}$,
 which gives the critical gap size past which adaptation will be mostly parallel
@@
 
-If we take the mutational influx in a patch to be one per 10,000 generations ($\rho A \mu = .0001$),
-$s_m>p_e$ so that $\gamma=1$,
+If we take the mutational influx in a patch to be one per 10,000 generations ($\rho A \mu = .0001$)
 and the population density to be 1 ($\rho =1$, so $-\log(A\mu)\approx 9.2$)
@@
 If $s_m=.001$, then with this higher mutational influx this calculation predicts that mutation will \emph{always} be faster than migration
--- however, this should be taken very cautiously,
-since as discussed above, this model does not hold if $R$ is of the same order as $\sigma$
-or if $s_m$ is small enough that local drift is more important.
-\linelabel{rr:separated_patches}
-
-% One way to visualize this is by a phase diagram.
-% In figure \ref{fig:phase_diagram},
-% we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
-% as a function of various parameters. 
-% This gives us an idea of not only the parameter space where parallel adaptation is likely ($\Tmig \ll \Tmut$),
-% but also where adaptation by any means takes a very long time ($\Tmig$ and $\Tmut$ both very large),
-% an important check in applications of these results.
-% 
-% \begin{figure}[ht]
-%   \begin{center}
-%     \includegraphics{phase-diagram-log}
-%   \end{center}
-%   \caption{
-%   Shaded contour plots of
-%   mean time until adaptation in a new patch occurs by migration ($\Tmig$, blue) or by mutation ($\Tmut$, red),
-%   as a function of population density ($\rho$), 
-%   distance to an already adapted patch in units of dispersal distance ($R/\sigma$),
-%   strength of deleterious selection ($s_m$),
-%   and mutation rate scaled by patch area ($\mu A$).
-%   The right-hand plot shows that $\migrate$ is relatively insensitive to $s_m$ (thanks to the square root).
-%   The black, dotted line shows values where $\Tmig=\Tmut$.
-%   Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=20$, and $\xi^2=1$.
-%   Note that population density $\rho$ and mutation influx $A\mu$ are shown on a log scale.
-%   \label{fig:phase_diagram}
-%   }
-% \end{figure}
+-- however, this is misleading, 
+since with $s_m=.001$, for the assumption of well-separated patches to hold we would need $R/\sigma > 1/\sqrt{2s_m} = 22.4$.
+The correct conclusion is that convergent adaptation is much more likely than shared adaptation
+between patches separated by at least 22.4$\sigma$;
+for shorter separations the model breaks down.
+
+One way to visualize this is by a phase diagram.
+In figure \ref{fig:phase_diagram},
+we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
+as a function of various parameters. 
+This gives us an idea of not only the parameter space where parallel adaptation is likely ($\Tmig \ll \Tmut$),
+but also where adaptation by any means takes a very long time ($\Tmig$ and $\Tmut$ both very large),
+an important check in applications of these results.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics{phase-diagram-log}
+  \end{center}
+  \caption{
+  Shaded contour plots of
+  mean time until adaptation in a new patch occurs by migration ($\Tmig$, blue) or by mutation ($\Tmut$, red),
+  as a function of population density ($\rho$), 
+  distance to an already adapted patch in units of dispersal distance ($R/\sigma$),
+  strength of deleterious selection ($s_m$),
+  and mutation rate scaled by patch area ($\mu A$).
+  The right-hand plot shows that $\migrate$ is relatively insensitive to $s_m$ (thanks to the square root).
+  The black, dotted line shows values where $\Tmig=\Tmut$.
+  Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=20$, and $\xi^2=1$.
+  Note that population density $\rho$ and mutation influx $A\mu$ are shown on a log scale.
+  \label{fig:phase_diagram}
+  }
+\end{figure}
 
@@
 Both these rates scale linearly with population density ($\rho$) 
-and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$, for $p_e>s_m$),
+and the probability of establishment of a rare allele ($p_e\approx 2 s_b/\xi^2$),
 so that the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
@@
 Furthermore, the probability that the second adaptation is a new mutation,
-i.e.\ the probability of parallel adaptation, 
-with $\gamma = \min(1,s_m/p_e)$,
-is 
+i.e.\ the probability of parallel adaptation, is 
 \begin{equation} \label{eqn:parallel_prob}
-    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
+    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
 \end{equation}
 so that the probability of parallel mutation should increase
-approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{s_m}$. \revpoint{2}{13}
-
-\begin{figure}[ht!]
-  \begin{center}
-      % made in analyze-sims.R
-      \includegraphics{prob-mutation-compared}
-  \end{center}
-  \caption{
-      Probability that two patches separated by distance $R$
-      evolve by independent mutations (i.e.\ in parallel),
-      as predicted by theory (colors, contours)
-      and in simulations (points).
-      \textbf{With $C=5$.}
-      XXX: more info
-  }   \label{fig:sim_probs}
-\end{figure}
+approximately logistically with the distance between the patches, at rate $\sqrt{s_m} /\sigma$.
 
@@
 and that these patches are at distances $R_1, \ldots, R_k$ away from this unadapted patch.
-The total rate of adaptation through migration, from equation \eqref{eqn:migrate}
-if patches do not interfere with each other, \linelabel{rr:interference}
-is
+The total rate of adaptation through migration, from equation \eqref{eqn:migrate}, is
 \begin{equation} \label{eqn:total_migrate}
@@
 Equations \eqref{eqn:total_migrate} and \eqref{eqn:migprob} specify the transition rates of the process.
-
 Since the compound parameter $s_p \rho / \xi^2$ is common to both migration and mutation rates,
@@
 the sets of patches sharing the genetic basis of their adaptive response.
+
 Also note that we can rescale time by the typical patch size, and introduce a parameter, say $a_k = A_k/\bar A$,
@@
 i.e.\ that it has exponential tails.
-This suggests one further approximation, 
-that the random lifetime of a family is exponentially distributed with mean $1/s_m$;
-equation \eqref{eqn:mean_K} shows this introduces a constant into the equations below. \linelabel{rr:EK}
+Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
+supposing that the random lifetime of a family is exponentially distributed with mean $1/s_m$.
 Once we do this, 
@@
 selective benefit within the patches or (perhaps surprisingly) the population density, 
-although these two factors will determine the time-scale over which adaptation will occur
-(and note that population density could affect the selection coefficients). \linelabel{rr:popdensity}
+although these two factors will determine the time-scale over which adaptation will occur. 
 This is in contrast with models of panmictic populations
@@
 This is counteracted, however, by the increased difficulty with which the mutation would establish
-once arriving in a new patch,
-if the beneficial effect is also recessive. \linelabel{rr:dominance}
+once arriving in a new patch. 
 As such it is not clear what our intuition should be 
@@
 We have also ignored the possibility of very long distance migration,
-instead focusing on local dispersal (and using Gaussians via the central limit theorem).
+instead focusing on local dispersal (and using a Gaussian model).
 However, dispersal distributions can be very heavy tailed, 
 with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
-In addition, over long time-scales, very rare chance events (mice carried off by hurricanes and the like; \citet{censky1998overwater,nathan2008mechanisms})
+In addition, over long time-scales, very rare chance events (mice carried off by hurricanes and the like \citet{censky1998overwater,nathan2008mechanisms})
 could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
@@
 \subsection*{Acknowledgements}
-We would like to heartily thank Joachim Hermisson for a number of insightful suggestions,
+We would like to heartily thank Joachim Hermisson for a careful reading of the paper,
 and catching the algebra error present in the first version of this paper on the bioRxiv,
-as well as the two anonymous reviewers.
-We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
+and also to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
 as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
@@
 \section*{Materials \& Methods}
-
-%%%%%%%
-\subsection{Simulation methods}
-\label{ss:simulations}
-
-First we briefly describe the simulations we used for illustration and validation
-(the \texttt{R} code used is provided as a supplement).
-We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
-Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
-locally beneficial alleles were more likely to reproduce.
-Each then produced a random, Poisson distributed number of offspring that each, independently, 
-either remained in the same location with probabilty $0.8$ or
-migrated a random number of steps in a uniformly chosen cardinal direction;
-the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
-\linelabel{rr:sim_details}
-Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
-(Although we described the simulation in terms of individuals,
-we kept track only of total numbers in an equivalent way.)
-
-The reproduction rate in simulations for type $b$ alleles was 30\%; 
-this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
-where the value of $s$ is either $s_m$ or $s_p$ depending on the individual's location.
-This determines the values of $s_m$ and $s_p$ reported in the figures, 
-and do not depend on the basic rate of reproduction.
-However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
-we computed the rate of intrinsic growth, 
-i.e.\ the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
-in the absence of migration.
-(The resulting values are close to the first notion of $s$,
-but give better agreement with theory,
-which uses the second definition.)
-
-To sample lineages, we first simulated forwards in time the population dynamics,
-then sampled lineages moving back through time
-by, in each generation,
-moving each lineage to a new deme with probability proportional to the reverse migration probability
-weighted by the number of $B$ alleles in that deme in the previous time step.
-If more than one lineage was found in a deme with $n$ $B$ alleles,
-then each lineage picked a label uniformly from $1 \ldots n$,
-and those picking the same label coalesced.
-Since reproduction is Poisson, this correctly samples from the distribution of lineages given the population dynamics.
 
@@
 \subsection[Hitting and Occupation]{Hitting and occupation}
-\label{ss:hitting_occupation}
-
-Here we make a more precise argument to back up expression \eqref{eqn:twocase_migrate} for the migration rate.
-The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
-since it relies only on a decomposition of the migrant families upon hitting the new patch;
-but it is also imprecise in a number of subtle ways
-that are difficult to formalize.
-The argument we make here therefore takes a somewhat different tack,
-supposing that it suffices to model the spatial movement of a migrant family
-by following only the motion of the ``trunk''
-(i.e.\ the red line in figure \ref{fig:branching_decomp}),
-and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
-(Recall $\sigma$ is the dispersal distance.)
-We then use facts about Brownian motion and branching processes, 
-to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
-\linelabel{rr:moved_argument}
-
-We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
-as the sum of independent migrant families, each of whose dynamics are
-given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
-Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
-Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
-such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ 
-that initiate new migrant families is $\gamma(x)$.
-Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``ouflux of families'',
-when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
-By examining the constituent terms,
-it will turn out that equation \eqref{eqn:gestalt_migrate} 
-is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.
-
-First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
-Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
-and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
-The \emph{mean occupation density} of $Z$ at location $y$
-can be thought of  informally \linelabel{rr:density}
-as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
-It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
-i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
-gives the expected total number of individuals ever living in $S$:
-\begin{align} \label{eqn:occupation_density_defn}
-    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
-\end{align}
-% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
-Since reproduction and spatial motion are independent,
-$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
-multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
-i.e.\
-\begin{align}
-    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
-\end{align}
-The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
-is the sum of mean occupation densities of a constant outflux of branching processes
-from $\partial B(r_0)$:
-\begin{align} \label{eqn:occupation_integral}
-    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
-         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
-\end{align}
-
-This form we can now compare to 
-the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
-Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
-We assume this new patch is circular, 
-has relatively small area $A$ and is centered at position $y$.
-% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
-% multiplied by the probability that the family reaches the patch, given survival.
-% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
-It would be possible to analyze this probability directly,
-as in \citet{barton1987establishment};
-but here we take a simpler route, 
-approximating this by the chance that the trunk hits the new patch,
-multiplied by the chance that at least one member of the family escapes demographic stochasticity
-and successfully establishes in the new patch.
-Write $h_A(x,y)$ for the probability that the Brownian trunk
-hits the new patch centered at $y$ before the family dies out,
-and $f(A)$ for the chance that the family manages to establish in the new patch,
-given that it successfully arrives.
-As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
-\begin{align}\label{eqn:migrate_integral}
-  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
-\end{align}
-Recall that $\rho$ appears here but not in \eqref{eqn:occupation_integral}
-because $q(x)$ is the local \emph{proportion} of $B$ alleles.
-
-Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
-while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
-Since the expected total time spent in the patch does not depend on the distance at which the family started
-if we know that it survived to hit the patch,
-it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
-Following this intuition,
-and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-it turns out (see section~\nameref{ss:hitting_occupation}) that
-\begin{align} 
-    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
-\end{align}
-where $g(A)$ is the total expected occupation time of the patch given the family arrives.
-Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
-Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
-that (at least asymptotically) does not depend on the distance between $x$ and $y$,
-we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
-The probability $f(A)$ that the family does not die out once it arrives at the new patch is
-approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
-if each member arriving has chance $p_e$ of establishing, independently.
-\plr{And if $p_e$ is small?}
-% If we let $w$ be the width of the patch, then the remaining term is
-% \begin{align}
-%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
-% \end{align}
-Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{ss:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
-we obtain that XXX FIX
-\begin{align}
-  \label{eqn:orig_migrate} 
-  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
-  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
-\end{align}
-This does not depend on $A$, because although
-$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
-the dependence is very weak.
-For instance, the width of a circular patch only very weakly affects the chance 
-of establishment of a new migrant that appears on its edge,
-as long as the patch is large enough.
-For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{ss:hitting_occupation}.
+\label{apx:hitting_occupation}
 
@@
 \begin{align}
-    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
-           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
-           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
-           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P\{ \tau_S \in dt \} \\
+    &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P\{ \tau_S \in dt \} \\
+    &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+    &\approx \frac{1}{\E[K]} \int_0^\infty \P\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
     &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
@@
 
-%%%%%%%
-\section{More information about the simulations}
-\label{apx:parameter_tables}
-
-Here we provide additional details regarding the simulations we used
-to validate the theory in figures \ref{fig:sim_times} and \ref{fig:sim_probs}.
-
-% table of parameters, produced by analyze-sims.R
-\input{mutation-parameters-table.tex}
-
-% table of parameters, produced by analyze-sims.R
-\input{migration-parameters-table.tex}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics{mutation-times-predicted}
-  \end{center}
-  \caption{
-    The same data shown in the left panel of figure~\ref{fig:sim_times},
-    but all times shown (not just the interquartile ranges),
-    and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
-  } \label{sfig:sim_migration_times}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics{migration-time-predicted}
-  \end{center}
-  \caption{
-    The same data shown in the right panel of figure~\ref{fig:sim_times},
-    but all times shown (not just the interquartile ranges),
-    and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
-    The left panel has the predicted time to adaptation on the horizontal axis as in figure~\ref{fig:sim_times},
-    and the right panel has, for comparison, the raw distance between patches 
-    (which predicts time to adaptation, but not as well).
-  } \label{sfig:sim_mutation_times}
-\end{figure}
-
-% example trace plots
-
-% sm=0.1
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/18885-r1-501-sb0_01-sm-0_1-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/56325-r1-501-sb0_01-sm-0_1-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/28432-r1-501-sb0_01-sm-0_1-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.1$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_1}
-\end{figure}
-
-% sm=0.01
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/43099-r1-501-sb0_01-sm-0_01-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/69787-r1-501-sb0_01-sm-0_01-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/11821-r1-501-sb0_01-sm-0_01-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.01$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_2}
-\end{figure}
-
-% sm=0.001
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/59611-r1-501-sb0_01-sm-0_001-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/93713-r1-501-sb0_01-sm-0_001-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/29850-r1-501-sb0_01-sm-0_001-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_3}
-\end{figure}
-
-% sm=0.0001
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/15449-r1-501-sb0_01-sm-1e-04-N1600-pophistory-run}
-    \includegraphics{example-mutation-sims/5582-r1-501-sb0_01-sm-1e-04-N400-pophistory-run}
-    \includegraphics{example-mutation-sims/24639-r1-501-sb0_01-sm-1e-04-N25-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_4}
-\end{figure}
-
-% sm=0.1, R=80
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-migration-sims/89826-r1-501-sb0_01-sm-0_1-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/76178-r1-501-sb0_01-sm-0_1-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.1$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_5}
-\end{figure}
-
-% sm=0.01, R=80
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-migration-sims/42080-r1-501-sb0_01-sm-0_01-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/82698-r1-501-sb0_01-sm-0_01-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.01$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_6}
-\end{figure}
-
-% sm=0.001, R=80
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-migration-sims/88079-r1-501-sb0_01-sm-0_001-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/3464-r1-501-sb0_01-sm-0_001-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/37774-r1-501-sb0_01-sm-0_001-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_7}
-\end{figure}
-
-% sm=0.0001, R=80
-\begin{figure}
-  \begin{center}
-    \includegraphics{example-migration-sims/26044-r1-501-sb0_01-sm-1e-04-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/85750-r1-501-sb0_01-sm-1e-04-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/4111-r1-501-sb0_01-sm-1e-04-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations with $s_m=0.0001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
-    On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
-    and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
-    dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_8}
-\end{figure}
-
 \ifsubmissionversion
@@
 \fi
-
-%%%% Reviews
-\includereviews
 

### 2014-11-29 Spatial_adaptation/standing-variation-fns.R

@@ -22,25 +22,30 @@ everything <- function (mu, rho, sb, sd, sigma) {
         ) )
 }
 
-meanTime <- function (mu, rho, sb, sd, sigma) {
+meanTime <- function (mu, rho, sb, sd, sigma,include.new=TRUE) {
     # mean time til adaptation of a point;
     # E[\tau] is int_0^\infty exp(-lambda0 pi v^2 t^2 - lambda pi v^2 t^3) dt
     if (sd==0) {
         return( list(value=0) )
-    } else if (sd==1) {
+    } else if (sd==1) {  ## no standing var.
         f <- function (t) {
+        	stopifnot(include.new)
             exp( - 4 * mu * rho * sb^2 * sigma^2 * t^3 / 3) 
         }
     } else { 
         f <- function (t) {
-            exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/log(1/(1-sd)) + t/3) )
-        }
+            exp( - 4 * mu * rho * sb^2 * sigma^2 * t^2 * (1/log(1/(1-sd)) + include.new*t/3) )
+        }   
     }
+
     xx <- 2^(-30:30)
     yy <- sapply(xx, function(x) { f(x) } )
     scale <- xx[ which.min( abs(10-yy*xx) ) ]
     return( integrate(function(x) { f(x*scale)*scale }, 0, Inf) )
 }
+
+
+
 
 meanNewDensity <- function (mu, rho, sb, sd, sigma) {
     # mean density of patches arising from new mutations

### 2014-11-30 patchy-review-responses.tex
most of the comments

@@
 unfortunately.
-However, we have pointed out in \nameref{ss:assumptions} that we're missing this point,
+However, we have pointed out in the section ``\nameref{ss:assumptions}'' that we're missing this point,
 and discuss there how the time lag to reach this equilbrium contributes to discrepancies between theory and simulations.
@@
 \reply
-Good point, but $A$ is what shows up naturally in $\mutrate$, so we've left it as-is.
+Good point; we already have $w$ defined for this purpose;
+and $A$ appears naturally in $\mutrate$,
+but we've tried to make things more intuitive.
 
@@
 \reply
-See text added in \nameref{ss:assumptions}.
+See text added in ``\nameref{ss:assumptions}''.
 
@@
  \reply
+As discussed above, we've removed a big chunk of the section on haplotypes,
+which should help make this section feel less heavy.
+However, we don't quite see the reviewer's assertion that $R \gg w$ is important here,
+and we don't see how to make the exposition much simpler with these assumptions.
 
@@
 The difference may be that Barton considers LD and we consider a haplotype unbroken by recombination.
-On the balance, and considering the other reviewer's wise comments,
-we've cut most of this from \nameref{ss:patch_haplotype}.
+There are also other assumptions in Barton's paper that it's applicability unclear.
+On the balance, and considering the other reviewers' comments,
+we've cut this section and discussed it a bit in the discussion. 
+\llname{rr:haplotype_discussion}
 
@@
 \reply
-Good points; we meant that it was less sensitive to $s_m$ than we'd otherwise think;
+Good points; we meant that it was less sensitive to $s_m$ than if it didn't have the square root;
 we've replaced this figure with a different one anyhow.
@@
 \reply
+Good point; fixed.
 
@@
 \reply
+We've put the scaling directly back into the equations here, so this isn't an issue.
 
@@
 
+\reply
+Yep, sorry about that
+We've fixed this (and changed the figure).
+
 \point{1c.}{
@@
 
+\reply
+We've done this; see discussion above.
+
 \point{2.}{
@@
 }
+
+\reply
+Thanks very much; see below for responses.
 
@@
 }
+
+\reply
+We have dropped it.
 
@@
     \begin{itemize}
-        \item equation (3) for the expected mutant density is only asymptotic. Therefore, we probably need to require $|y| \gg \sigma / \sqrt(2s_m)$.
+        \item equation (3) for the expected mutant density is only asymptotic. Therefore, we probably need to require $|y| \gg \sigma / \sqrt{2s_m}$.
 
@@
 \reply
+As discssed above, we've added this discussion to ``\nameref{ss:assumptions}''.
+A few more notes:
+It's not clear to us why the second point requires the patch width to be wider than $1/s_m$, shouldn't it be $\sigma/\sqrt{s_m}$
+(which is covered elsewhere)?
+The appearance of $r_0$ is purely to make families well-defined, and don't think it needs to be discussed here.
 
@@
 Good point; we didn't notice that after correcting $\migrate$ the parameter regime on the right didn't make sense.
-We have substituted XXX for these.
+We have substituted figures \ref{fig:sim_times} and \ref{fig:sim_prob} for this one.
 
@@
 \reply
+After scratching our heads and peering at Barton 1979 (as suggested by another reviewer),
+we've seen the wisdom of this
+and cut most of this section.
 

### 2014-11-30 patchy-selection-paper.tex
most of the comments

--- 2014-11-28 patchy-selection-paper.tex
@@
 and by migration from an already colonized patch, \eqref{eqn:migrate},
-it seems helpful to step back and review the assumptions we have made and the parameter regime
-we expect the results to apply to.
-For the mutation rate,
-our results will hold best if $s_m \approx s_p$,
+it seems helpful to step back and review the assumptions 
+underlying the asymptotic results we have used.
+Our results should hold exactly in the limit of 
+large, circular patches that are much further apart than they are wide,
+large population density,
+and small selection coefficients of equal magnitude.
+More concretely, the mutation rate results will hold best if $s_m \approx s_p$,
 if the patch diameter is large relative to $\sigma/\sqrt{\min(s_m,s_p)}$,
-the patch is not too far from circular.
+and the patch is not too far from circular.
 If $s_m \ll s_p$ then \eqref{eqn:mutrate} will underestimate the rate,
-and conversely if $s_m \gg s_p$.
+and will overestimate if $s_m \gg s_p$.
 As for the migration rate,
@@
 ($A^{1/d} > (\sigma/\sqrt{2s_p}) \tan^{-1}(\sqrt{s_m/s_p})$).
+The geometry and size of the patch will also affect the approximation of equation \eqref{eqn:gA_1D}.
 Next, in using equation \eqref{eqn:eqfreq}, we assume that 
-\textbf{(b)} the inter-patch distance is at least the characteristic length
+\textbf{(b)} the inter-patch distance is greater than the characteristic length
 ($R > \sigma/\sqrt{s_m}$),
@@
 
-%%%%%%% %%%%%%%%%%
-\subsection[Haplotypes Shared Within a Patch]{Length of the shared haplotype within a patch}
-\label{ss:patch_haplotype}
-
-The other key issue in thinking about the population genetic signal of adaptation in spatial patches 
-is the process by which haplotypes are whittled down within patches, 
-which plays into a number of aspects of haplotype sharing. 
-The initial haplotype that sweeps within a patch 
-will be dispersed over time by recombination.
-Likewise, the haplotype that is shared between patches coadapted by migration
-(equation~\ref{eqn:haplotype_length}) will also break down. 
-However, a long time after the initial sweep,
-we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
-than with individuals elsewhere,
-since selection against migrants 
-decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}.
-
-Finally we can consider the expected length of haplotype shared within a patch. 
-To do this, we compare the rate of recombination between $B$ and $b$ haplotypes (equation~\ref{eqn:rec_rate_within})
-to the rate of coalescence of $B$ haplotypes within a patch. 
-We will assume that mixing within the patch is fast compared to the rate of drift, 
-so that the rate of coalescence of $B$ alleles within a patch is 
-$\xi^2/(A\rho)$,
-the ``variance effective population size'' of the patch
-(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1975pain,barton2002neutral}).
-The rate of recombination and coalescence will be on the same order if
-$\xi^2 /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$.
-Hence the genetic length of the haplotype shared between pairs of
-individuals within the same patches at equilibrium is on the order of
-\begin{align} \label{eqn:within_haplotype_length} 
-  r = \frac{\sqrt{s} \xi^2}{\sigma \rho A^{(d-1)/d}}  .
-\end{align}
-This does not depend on $A$ in one dimension,
-but decreases with $\sqrt{A}$ in two dimensions.
-In applications, this can be compared to equation~\eqref{eqn:haplotype_length}
-for the length of shared haplotype between patches.
-
 %%%%%%%%%%
@@
 
+\paragraph{Decay of shared haplotypes} \nameref{rr:haplotype_discussion}
+To provide context for the results on shared haplotype length in \nameref{ss:haplotype_length}
+it is important to also understand
+the process by which haplotypes are whittled down within patches.
+The initial haplotype that sweeps within a patch 
+will be dispersed over time by recombination.
+Likewise, the haplotype that is shared between patches coadapted by migration
+(equation~\ref{eqn:haplotype_length}) will also break down. 
+However, a long time after the initial sweep,
+we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
+than with individuals elsewhere,
+since selection against migrants 
+decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
+The literature on clines \citep[e.g.]{barton1979geneflow} has important information,
+but a good deal more work is needed
+to provide robust estimates for these processes.
+
 \paragraph{Standing variation} 
@@
 in $d=1$,
-\begin{align}
+\begin{align} \label{eqn:gA_1D}
     g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
     % &= \int_0^{2w} G_{s_m}(0,y) dy \\
-  &= \int_0^{2w} \frac{ e^{- y \sqrt{2s_m} }}{\sqrt{2s_m}} dy \\
-  &= (1-e^{-2w\sqrt{2s_m}})/(2s_m)
+  &= \int_0^{2w/\sigma} \frac{ e^{- y \sqrt{2s_m}/\sigma }}{\sqrt{2s_m}} dy \\
+  &= (1-e^{-2w\sqrt{2s_m/\sigma}})/(2s_m)
 \end{align}
@@
     % &= \int_0^{2w} G_{s_m}(0,y) dy \\
-  &= \int_0^{2w} 2 y K_0(y \sqrt{2s_m}) dy \\
-  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}} y K_0(y) dy \\
-  &= \frac{1}{2s_m}\left(1- 2w\sqrt{2s_m} K_1(2w\sqrt{2s_m}) \right)  .
+  &= \int_0^{2w/\sigma} 2 y K_0(y \sqrt{2s_m}) dy \\
+  &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}/\sigma} y K_0(y) dy \\
+  &= \frac{1}{2s_m}\left(1- \frac{2w\sqrt{2s_m}}{\sigma} K_1(2w\sqrt{2s_m}/\sigma) \right)  .
 \end{align}
@@
 for any $y$.  
-Take $y=(r,0)$.
-Note that after scaling by $\sqrt{2s_m}/\sigma$,
-\begin{align}
-    \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\|} }{ \|y\|^{-(d-1)/2} e^{-\|y\|} } \\
-        &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r (\|\hat x - x/r\| - 1)} .
+Take $\hat x = (1,0$ and $y=r \hat x$.  Then
+\begin{align}
+  \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\| \sqrt{2s_m}/\sigma } }{ \|y\|^{-(d-1)/2} e^{-\|y\| \sqrt{2s_m}/\sigma } } \\
+  &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r \sqrt{2s_m}/\sigma  (\|\hat x - x/r\| - 1)} .
 \end{align}
@@
 \begin{align}
-    \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \cos \theta} .
+  \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} .
 \end{align}
@@
   \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-            &= C' \int_0^{2 \pi} e^{-2 r_0 \cos \theta} d\theta \\
-                         &= 2 \pi C' I_0(-2r_0),
+  &= C' \int_0^{2 \pi} e^{-2 r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
+  &= 2 \pi C' I_0(-2r_0\sqrt{2 s_m}/\sigma),
 \end{align}
 where $I_0$ is a modified Bessel function of order zero of the first kind.
-The effect of the scaling is to express $r_0$ in natural units, i.e.\ changing the result to $2 \pi C' I_0( - 2 \sqrt{2s_m} r_0 / \sigma ) $.
 

### 2014-11-30 patchy-selection-paper.tex
updated with more sims

@@
 
-Here we make a more precise argument to back up expression \eqref{eqn:twocase_migrate} for the migration rate.
+Here we make a more precise argument to back up expression \eqref{eqn:migrate} for the migration rate.
 The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
@@
 that are difficult to formalize.
-The argument we make here therefore takes a somewhat different tack,
+Here we take a somewhat different tack,
 supposing that it suffices to model the spatial movement of a migrant family
@@
 we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
-The probability $f(A)$ that the family does not die out once it arrives at the new patch is
-approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
-if each member arriving has chance $p_e$ of establishing, independently.
-\plr{And if $p_e$ is small?}
-% If we let $w$ be the width of the patch, then the remaining term is
-% \begin{align}
-%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
-% \end{align}
-Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{ss:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
-we obtain that XXX FIX
-\begin{align}
-  \label{eqn:orig_migrate} 
-  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
-  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
-\end{align}
-This does not depend on $A$, because although
-$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
+
+The result will not depend on $A$, 
+because although $f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
 the dependence is very weak.
@@
 \end{align}
+Once we show that $g(A) \approx 1/(2 s_m)$, 
+we will have arrived at the result, equation \eqref{eqn:ratio_K}.
+
 The function $g(A)$ is the expected amount of time that a Brownian motion begun on the edge of a disk of area $A$

### 2014-11-30 analyze-sims.R

@@ -4,7 +4,7 @@ source("sim-patchy-selection-fns.R")
 
 maxtime <- 25000  # number of gens these were run for
 
-mutsims <- read.table("mutation-sims-info.tsv",sep='\t',header=TRUE)
+mutsims <- read.table("mutation-sims-info.tsv", sep='\t', header=TRUE )
 
 names(mutsims)[names(mutsims)=="N"] <- "N.name"
 mutsims$N.name <- factor( mutsims$N.name, levels=levels(mutsims$N.name)[order(as.numeric(gsub("N-","",levels(mutsims$N.name))))])
@@ -103,6 +103,13 @@ migsims$adapted <- (
     ( migsims$hit100.2 < 24000 ) &
     ( migsims$time2 >= migsims$hit100.2 ) )
 
+migsims$valid <- with( migsims, 
+        ( R*sqrt(abs(sm))/sigma > 1 ) & 
+        ( size2 > (sigma/sqrt(abs(sm)))*atan(sqrt(abs(sm/gb))) ) &
+        ( abs(sm) * 2 * sigma * N > 1 )
+    )
+
+
 pdf(file="migration-time-predicted.pdf",width=9, height=5, pointsize=10)
 layout(t(1:2))
 with( subset(migsims,adapted), {
@@ -183,12 +190,12 @@ par(mar=c(4,4,1,1)+.1)
                     )
             points( xx,
                     tapply(adapttime,paramstring,median,na.rm=TRUE), 
-                    pch=20+sm.pch, bg=mu.cols, lwd=2, col=N.cols
+                    pch=21+(sm.pch%%5), bg=mu.cols, lwd=2, col=N.cols
             ) 
             abline(0,1,untf=TRUE)
             legend("bottomright", cex=0.5, pt.cex=1,
                 legend=c(sm.labs,N.labs,mu.labs),
-                pch=c( 20+(1:nlevels(sm.name)), rep(1,nlevels(N.name)), rep(21,nlevels(mu.name)) ), 
+                pch=c( 21+((1:nlevels(sm.name))%%5), rep(1,nlevels(N.name)), rep(21,nlevels(mu.name)) ), 
                 col=c("black",N.col.pal)[c(rep(1,nlevels(sm.name)),1+1:nlevels(N.name),rep(1,nlevels(mu.name)))],
[…]

### 2014-12-06 patchy-selection-paper.tex
small changes throughout text

@@
 
+   \newif\iflatexml\latexmlfalse %%%graham added
+
+%We should cite Papers by Slatkin and Wiehe and this, on spatial hitchhiking http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3176130/
 %%%%%
@@
 % set this to show line numbers and include responses to reviews or not
-\newif\ifreviewresponses
-\reviewresponsestrue  % include them
+\newif\ifreviewresponses\reviewresponsestrue  % include them
 % \reviewresponsesfalse  % don't include them
@@
 study the fine-scale genealogy of alleles that transit between patches,
-obtaining along the way other useful information about the genealogy of $B$ alleles.
+obtaining along the way other useful information about the genealogy
+of $B$ alleles. \gc{Doing this we arrive at \eqref{eqn:migrate} for
+  the rate at which patches adapt via migration. } %%Thought it good
+                                %%to signpost again where we are
+                                %%going.
 
@@
 From simulations, $C$ seems to be not too far from 1 in practice (see below).
-As J.\ Hermisson pointed out in an earlier draft,
+As J.\ Hermisson pointed out \gc{in comments} on an earlier draft,
 this functional form has a simple intuition:
@@
 $|x|$ must be larger than only 5--10 multiples of $\sigma$.
-\linelabel{rr:q_form}
+\linelabel{rr:q_form}  \gc{OR WE could just personal ref Hermisson here.}
 
@@
 When we use this below, we set $p_e = 2 s_p / \xi^2$.
-Note that the function increases for small $s_m$, 
-but decreases if $s_m > |y|^2 / 2 \sigma^2$, 
-i.e.\ where the condition holds.
+\gc{Note that the approximation has the undesirable property that it is
+non-monotonic in $s_m$, when we expect the rate of adaptation via
+migration to be a decreasing function of $s_m$. However, it only increases for small $s_m$, 
+and is a decreasing function of $s_m$ for $s_m > |y|^2 / 2 \sigma^2$, 
+i.e.\ the parameter region that we are interested in.}
 \linelabel{rr:sm_condition}
+%%GRAHAM felt like we needed to be clearer on this point.
 
@@
 and small selection coefficients of equal magnitude.
-More concretely, the mutation rate results will hold best if $s_m \approx s_p$,
+
+%%Graham brokke this paragraph up.
+\gc{To be more precise,} the mutation rate results will hold best if $s_m \approx s_p$,
 if the patch diameter is large relative to $\sigma/\sqrt{\min(s_m,s_p)}$,
@@
 we used individual-based simulations on a one-dimensional lattice of 501 demes,
-run for 25,000 generations.
+run for 25,000 generations. \gc{We simulated $\rho$ diploid
+  individuals per patch, with nearest neighbor migration and
+  multiplicative selection??!? }
 More details are given below in \nameref{ss:simulations},
-and the number of simulations used and parameter values are given in supplementary tables \ref{stab:mutation_params} and \ref{stab:migration_params}.
+and the number of simulations used and parameter values are given in
+supplementary tables \ref{stab:mutation_params} and
+\ref{stab:migration_params}. 
 To estimate the mean time until adaptation by mutation,
@@
 
-Let $\gamma = \min(1,s_m/p_e)$.
+%%felt like we needed a better intro to this paragraph
+\gc{We first consider the order of magnitude that our parameters need to
+be on in order for adaptation via mutation or migration to dominant.} Let $\gamma = \min(1,s_m/p_e)$.
 An important point to notice is that effective migration and mutation rates will be on the same order if 
@@
 
-Taking these approximations at face value, 
-and assuming both rates are small, 
+\gc{We can go beyond these orders of magnitude calculations to find
+  the probability of parallel adaptation if we are will to take our approximations at face value, 
+and assuming both rates are small. Doing so } 
 the time until the first of the two patches is colonized by the $B$
@@
 share a fairly long stretch of haplotype. 
-Pairs of adapted individuals within a patch will share a haplotype of mean genetic length of
+Pairs of adapted individuals within a patch will \gc{initially} share a haplotype of mean genetic length of
 about $\log(\rho A s_p)/s_p$ around the selected site, 
@@
 This association gets slowly whittled down by recombination during interbreeding at the edge of the patch,
-but there will always be longer LD nearby to the selected site.
+but there will always be longer LD nearby to the selected site
+\gc{CITATION, BARTON?}.
 
@@
 and to convert it to a speed this is divided by $1/2s_m$.
-This intuitive picture can be expanded by a central limit theorem:
-for large $R$, the deviation of $\tau$ about its mean is asymptotically Gaussian,
-and the haplotype halflength $L$ is exponential.
+
+\gc{In conclusion, the length of shared haplotype drops off quickly with the
+distance between the patches. When $s_m$ is small we expect shorter
+haplotypes as the allele is only weakly deleterious and so 
+traverses the distance between patches relatively slowly compared to
+strongly deleterious alleles. Therefore, when adaptation by migration
+is most likely (small $s_m$) 
+we expect shorter
+shared haplotypes.}
+
+%GRAHAM commented out this This intuitive picture can be expanded by a central limit theorem:
+%for large $R$, the deviation of $\tau$ about its mean is asymptotically Gaussian,
+%and the haplotype halflength $L$ is exponential.
 
@@
 
+\gc{DELETE THIS: 
 Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
@@
 then it will be difficult to identify shared haplotypes
-(how recently depends on the background level of LD in the population).
+(how recently depends on the background level of LD in the population).}
 

### 2014-12-09 patchy-review-responses.tex
small changes throughout response to reviews, and application section

@@
 
+We agree with this comment, to test the validity of our model we
+coded an individual level model of migration and selection. 
 We have done a large number of simulations, and compared the results to the theory;
@@
 
-We sympathize entirely: we felt the same way ourselves,
-and went around in similar circles for quite some time
+We have rewritten the method section to simplify the presentation. 
+Before writing the paper, we had initially explored a number of simpler ways of developing the
+argument (along similar the lines to those suggested by the reviewers), 
+and went around in circles for quite some time
 before settling on the argument given in the first draft.
@@
 along the lines of that suggested by the reviewers.
-(We do think the more formal argument is more solid, however,
+We do think the more formal argument is more solid, however,
 and therefore necessary to keep; 
-as discussed below, various seemingly reasonable arguments
-give different answers for subtle reasons.)
+as discussed below in our response to the reviewers, various seemingly reasonable arguments
+give different answers for subtle reasons.
 
@@
     we were neglecting the case that $s_m \ll s_p$.
-    This omission was pure laziness on our part,
+    This omission was sloppy on our part,
     which we have rectified; see the new equation~\eqref{eqn:twocase_migrate}.
@@
     but this seems beside the point.
+\gc{Graham is not sure what you mean by ``we were running tests in regions'' }
 
@@
 \reply
-Good idea, but we've decided to cut this figure.
+Good idea, but we've decided to cut this figure in the resubmission.
 
@@
 We don't know of even a poor approximation to the time to reach migration-selection equilibrium,
-unfortunately.
-However, we have pointed out in the section ``\nameref{ss:assumptions}'' that we're missing this point,
+unfortunately. However, we now pointed out in the section ``\nameref{ss:assumptions}'' that we're missing this point,
 and discuss there how the time lag to reach this equilbrium contributes to discrepancies between theory and simulations.
+
+\gc{seems to graham that we should spread to migration-selection
+  equilibrium as the fisher-kpp. Suggests that patch width$/sqrt{s}
+  \sigma$ wouldn't be a bad guess.}
 
@@
 \reply
-See text added in ``\nameref{ss:assumptions}''.
+See text added in Section ``\nameref{ss:assumptions}''.
 
@@
 \reply
+At the moment it is only known that the same MC1R alleles do not
+contribute to color polymorphism on a number of other patches. We know
+from personal discussion with Michael Nachman that another color
+polymorphism allele, in a different gene, has been mapped on another
+patch. We are currently talking to Nachman about a broader-scale
+effort to identify other causal alleles motivated in part by the work described
+in the paper. However, this work is too preliminary to offer any thing
+certain.
+
+ There is no data about the genetic map in this species, but
+as a rough rule of thumb 1cM is roughly 1Mb in a number of mammalian
+species, we have added this guide to the figure caption \llname{rr:cMMb}
 
@@
 Good points; we meant that it was less sensitive to $s_m$ than if it didn't have the square root;
-we've replaced this figure with a different one anyhow.
+in this draft we've replaced this figure with a different one (so
+disgarded this comment).
 
@@
 \reply
-We've put the scaling directly back into the equations here, so this isn't an issue.
+We've put the scaling directly back into the equations here in this
+new draft, so this isn't an issue.
 
@@
 \reply
-We have dropped it.
+We have dropped this section.
 
@@
 \reply
-We've tried to make it more clear here 
-\llname{rr:longtailed}
-exactly what we sweep under the rug later.
+We've made this more clear here 
+\llname{rr:longtailed}.
 
@@
 \reply
-Fixed? \llname{rr:awkward}
+Fixed \llname{rr:awkward}
 
@@
 \reply
-As discssed above, we've added this discussion to ``\nameref{ss:assumptions}''.
+As discssed above, we've added this discussion to the Section ``\nameref{ss:assumptions}''.
 A few more notes:

### 2014-12-09 patchy-selection-paper.tex
small changes throughout response to reviews, and application section

@@
 shared between patches due to adaptation via migration as a function
-of the distance between environmental patches. 
+of the distance between environmental patches. As a rough rule of
+thumb $1$cM is approximately $1$Mb in a number of mammalian species. \linelabel{rr:cMMb}
   \label{fig:mice_prob_parallel}
@@
 
-\gc{DELETE THIS: 
-Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
-we should expect the haplotype to be rapidly whittled down within
-patches after establishment (equation \ref{eqn:rec_rate_within}). We
-estimate that the effective recombination rate is no less than 50\% of the actual recombination rate.
-Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare 
-(M.~Nachman, personal communication),
-and the variance in offspring number to be $\xi^2 = 3$ (mostly to get round numbers),
-then the shared haplotype length within patches from equation \eqref{eqn:within_haplotype_length}
-is only expected to be on the order of hundreds or thousands of bases
-($10^{-3}$--$10^{-4}$cM).
-This suggests that unless the migration has happened fairly recently
-then it will be difficult to identify shared haplotypes
-(how recently depends on the background level of LD in the population).}
+%\gc{DELETE THIS: 
+% Furthermore, given the relatively small radius of the patches of dark substrate (often $<10$km),
+% we should expect the haplotype to be rapidly whittled down within
+% patches after establishment (equation \ref{eqn:rec_rate_within}). We
+% estimate that the effective recombination rate is no less than 50\% of the actual recombination rate.
+% Indeed, if we take the population density to be roughly $\rho \approx 10$ per hectare 
+% (M.~Nachman, personal communication),
+% and the variance in offspring number to be $\xi^2 = 3$ (mostly to get round numbers),
+% then the shared haplotype length within patches from equation \eqref{eqn:within_haplotype_length}
+% is only expected to be on the order of hundreds or thousands of bases
+% ($10^{-3}$--$10^{-4}$cM).
+% This suggests that unless the migration has happened fairly recently
+% then it will be difficult to identify shared haplotypes
+% (how recently depends on the background level of LD in the population).}
 

### 2014-12-09 patchy-review-responses.tex
small changes throughout response to reviews, and discussion

@@
 \vskip 1em
+
+We thank the editor and the reviewers for their helpful comment. We
+have substantially revised our paper in response ot these comments,
+and think that it is substantially improved as a consequence. Below we
+have a detailed response point-by-point to the reviewers comments. We
+first address the consensus points highlighted by the editor.
 
@@
     The main point of note was that, indeed as Reviewer 3 pointed out,
-    we were running tests in regions where $s_m$ was too small for the theory to apply,
+    we were \gc{originally} running \gc{simulations} in regions where $s_m$ was too small for the theory to apply,
     resulting in $\migrate$ that was not monotonic with $s_m$.
@@
     but this seems beside the point.
-\gc{Graham is not sure what you mean by ``we were running tests in regions'' }
 
@@
 as a rough rule of thumb 1cM is roughly 1Mb in a number of mammalian
-species, we have added this guide to the figure caption \llname{rr:cMMb}
+species, we have added this guide to the physical scale of the shared haplotype in the figure caption.
 

### 2014-12-09 patchy-selection-paper.tex
small changes throughout response to reviews, and discussion

@@
 of the distance between environmental patches. As a rough rule of
-thumb $1$cM is approximately $1$Mb in a number of mammalian species. \linelabel{rr:cMMb}
+thumb $1$cM is approximately $1$Mb in a number of mammalian species. 
   \label{fig:mice_prob_parallel}
@@
 The literature on clines \citep[e.g.]{barton1979geneflow} has important information,
-but a good deal more work is needed
+but more work is needed
 to provide robust estimates for these processes.
-
+\gc{We note that questions about the genomic length-scale of signals of
+sweeps shared by migration have also been addressed in discrete
+population settings \citep{REFSCOMMENTEDBELOW}. This work has shown
+that the length of shared sweept haplotype is often significantly
+shorter than the sweep with each patch, resulting in a pattern of
+shoulders ofelevated $F_{ST}$ between adapted populations some distance away from
+the shared selected allele. It would be of interest to see how similar
+patterns can arise in a continuous population setting, as a way of
+uniting these results. }
+
+%Slatkin and Wiehe and this paper on spatial hitchhiking http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3176130/
 \paragraph{Standing variation} 

### 2014-12-09 patchy-selection-paper.tex
added section infront of each section ref

@@
 
+\gc{GENERAL GRAHAM comment, every place where we give a section name nameref
+I've made sure that we say section nameref as otherwise it looks weird  }
+
 %%%%%%%%%%% %%%%%%%%%%% 
@@
 To be clear about the assumptions (especially regarding scaling) implicit here,
-we provide a derivation below in \nameref{apx:eqfreq}, 
-as well as justification for the asymptotics below in \nameref{apx:asymptotics}.
+we provide a derivation below in section \nameref{apx:eqfreq}, 
+as well as justification for the asymptotics below in section \nameref{apx:asymptotics}.
 In one dimension, the equation can be solved to give $q(x)$ 
@@
 To make this precise we need to understand what we mean by ``outflux of families'';
-we do this below in \nameref{ss:hitting_occupation},
+we do this below in section \nameref{ss:hitting_occupation},
 and provide a heuristic argument here.
@@
 or the distinction between occupation density and number of individuals
-are encouraged to take a detour through \nameref{ss:hitting_occupation}, below.)
+are encouraged to take a detour through section \nameref{ss:hitting_occupation}, below.)
 
@@
   multiplicative selection??!? }
-More details are given below in \nameref{ss:simulations},
+More details are given below in section \nameref{ss:simulations},
 and the number of simulations used and parameter values are given in
@@
 \paragraph{Decay of shared haplotypes} \nameref{rr:haplotype_discussion}
-To provide context for the results on shared haplotype length in \nameref{ss:haplotype_length}
+To provide context for the results on shared haplotype length in
+section \nameref{ss:haplotype_length}
 it is important to also understand
@@
 Here we make a more precise argument to back up expression \eqref{eqn:migrate} for the migration rate.
-The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
+The argument made above in section \nameref{ss:heuristics} applies to general migration mechanisms,
 since it relies only on a decomposition of the migrant families upon hitting the new patch;

### 2014-12-17 patchy-review-responses.tex
peter did final(?) pass through

@@
 We thank the editor and the reviewers for their helpful comment. We
-have substantially revised our paper in response ot these comments,
+have substantially revised our paper in response to these comments,
 and think that it is substantially improved as a consequence. Below we
@@
 We agree with this comment, to test the validity of our model we
-coded an individual level model of migration and selection. 
-We have done a large number of simulations, and compared the results to the theory;
-the main comparison occurs in \ref{fig:sim_times} and figures~\ref{fig:sim_probs};
+ran an additional 21,873 simulations using our individual-based model of migration and selection,
+and compared the results to the theory.
+The main comparison occurs in \ref{fig:sim_times} and figures~\ref{fig:sim_probs};
 we have also added in supplementary section \ref{apx:parameter_tables} 
@@
     This omission was sloppy on our part,
-    which we have rectified; see the new equation~\eqref{eqn:twocase_migrate}.
+    which we have rectified; see the new equation~\eqref{eqn:migrate}.
 
@@
     The main point of note was that, indeed as Reviewer 3 pointed out,
-    we were \gc{originally} running \gc{simulations} in regions where $s_m$ was too small for the theory to apply,
+    we were originally running simulations in regions where $s_m$ was too small for the theory to apply,
     resulting in $\migrate$ that was not monotonic with $s_m$.
@@
 \end{itemize}
+
+We hope you find the revised manuscript as improved as we do.
+
+\noindent \hspace{4em}
+\begin{minipage}{3in}
+\noindent
+{\bf Sincerely,}
+
+\vskip 2em
+
+{\bf Peter Ralph, and\\
+Graham Coop.}\\
+\end{minipage}
+
+\vskip 4em
 
@@
 \reply
-The factor of $\ell$ shouldn't be there:
-suppose that $X$ is a Poisson process of ``centers'' of rate 1, 
+We don't think the factor of $\ell$ should be there:
+for example, suppose that $X$ is a Poisson process of ``centers'' of rate 1, 
 and that around each center of $X$ we distribute $n$ points uniformly within a circle of radius $\ell$;
@@
 \reply
-As discussed above, we have incorporated this good point; see around equation~\eqref{eqn:twocase_migrate}.
+As discussed above, we have incorporated this good point; see around equation~\eqref{eqn:migrate}.
 
@@
 \reply
-We don't know of even a poor approximation to the time to reach migration-selection equilibrium,
-unfortunately. However, we now pointed out in the section ``\nameref{ss:assumptions}'' that we're missing this point,
+We don't know of an approximation to the time to reach migration-selection equilibrium
+that isn't too hand-wavy,
+unfortunately. 
+However, we now point out in the section ``\nameref{ss:assumptions}'' that we're missing this factor,
 and discuss there how the time lag to reach this equilbrium contributes to discrepancies between theory and simulations.
-
-\gc{seems to graham that we should spread to migration-selection
-  equilibrium as the fisher-kpp. Suggests that patch width$/sqrt{s}
-  \sigma$ wouldn't be a bad guess.}
 
@@
     at the patch is successful is approximately $ 2s_p \E [K] /\xi^2$, the authors have
-    implicitly assumed that this quantity is $<\ll 1$. But this implies that the
+    implicitly assumed that this quantity is $\ll 1$. But this implies that the
     fluctuations in the number of B individuals arriving at the patch can be
@@
 \point{}{
-    However, I think we should also care about the opposite limit, $s_m \ll s p$. In
+    However, I think we should also care about the opposite limit, $s_m \ll s_p$. In
     this case, the fluctuations do matter -- arrivals typically occur in clumps
@@
 The difference may be that Barton considers LD and we consider a haplotype unbroken by recombination.
-There are also other assumptions in Barton's paper that it's applicability unclear.
+There are also other assumptions in Barton's paper that makes its applicability unclear.
 On the balance, and considering the other reviewers' comments,
@@
 %%%%%%%%
-\point{In “Applications”,}{the authors spend a lot of time introducing the mice
+\point{In ``Applications'',}{the authors spend a lot of time introducing the mice
     and calculating the values for Fig. 6, but then they don’t show where the
@@
     \begin{itemize}
-        \item[A.]  I think “Simulation methods” should go in the “Materials and meth-
-            ods” section, or at least not right between “Establishment of a locally
-            adaptive allele due to mutational influx” and “Establishment of a lo-
-            cally adaptive allele due to mutational influx”.
+        \item[A.]  I think ``Simulation methods'' should go in the ``Materials and methods'' section, or at least not right between ``Establishment of a locally
+            adaptive allele due to mutational influx'' and ``Establishment of a lo-
+            cally adaptive allele due to mutational influx''.
 
-        \item[B.] The sentence on p7 beginning “By our definition of fitness. . . ”.
+        \item[B.] The sentence on p7 beginning ``By our definition of fitness. . . ''.
 
-        \item[C.]  The paragraph starting “Also note that. . . ” on p11.
+        \item[C.]  The paragraph starting ``Also note that. . . '' on p11.
     \end{itemize}
@@
 Good points; we meant that it was less sensitive to $s_m$ than if it didn't have the square root;
-in this draft we've replaced this figure with a different one (so
-disgarded this comment).
+in this draft we've replaced this figure with a different one anyhow.
 
 %%%%%%%%
-\point{Top of p11 (just above “Multiple patches”):}{should be $\sqrt{2s_m}/\sigma$.
+\point{Top of p11 (just above ``Multiple patches''):}{should be $\sqrt{2s_m}/\sigma$.
 }
@@
 \reply
-We've made this more clear here 
+We've made this more clear
 \llname{rr:longtailed}.
@@
 \reply
-We have added the word ``informally''. \llname{rr:density}
+Rephrased. \llname{rr:density}
 
@@
 In practice it could provide a good reality check.
-\plr{alternatively, compute this for the mice example?}
+% \plr{alternatively, compute this for the mice example?}
 
@@
 Good point; we didn't notice that after correcting $\migrate$ the parameter regime on the right didn't make sense.
-We have substituted figures \ref{fig:sim_times} and \ref{fig:sim_prob} for this one.
+We have substituted figures \ref{fig:sim_times} and \ref{fig:sim_probs} for this one.
 

### 2014-12-17 patchy-selection-paper.tex
peter did final(?) pass through

--- 2014-12-09 patchy-selection-paper.tex
@@
 
-   \newif\iflatexml\latexmlfalse %%%graham added
-
-%We should cite Papers by Slatkin and Wiehe and this, on spatial hitchhiking http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3176130/
+\newif\iflatexml\latexmlfalse %%%graham added; remove with latexml package
+
 %%%%%
@@
 
+% make references to named sections look better
+\newcommand{\secref}[1]{{\emph{\nameref{#1}}}}
+
 % short title:  Convergent Evolution on Patchy Landscapes
@@
 \abstract{
-Species often encounter, and adapt to, many patches of locally similar environmental conditions
-across their range. 
-Such adaptation can occur through convergent evolution as different alleles arise and spread in different patches, 
-or through the spread of alleles by migration
+Species often encounter, and adapt to, many patches of locally similar environmental conditions across their range. 
+Such adaptation can occur through convergent evolution if different alleles arise and spread in different patches, 
+or through the spread of shared alleles by migration
 acting to synchronize adaptation across the species.
-The tension between the two reflects the degree of 
-constraint imposed on evolution by the underlying genetic
-architecture versus how effectively selection acts to inhibit the geographic spread of locally
-adapted alleles.
+The tension between the two reflects 
+the degree of constraint imposed on evolution by the underlying genetic architecture 
+versus how effectively selection and geographic isolation act to inhibit the geographic spread of locally adapted alleles.
 This paper studies a model of the balance between these two routes to adaptation in continuous environments with patchy selection pressures.
@@
 How can we tell which has occurred, i.e.\ what population genetic
-signal is left by the spread of migrant alleles and how long does this
-signal persist for?
-To answer these questions we decompose the migration--selection equilibrium
-surrounding an already adapted patch
-into families of migrant alleles,
+signal is left by the spread of migrant alleles?
+To answer these questions we examine the family structure underlying migration--selection equilibrium
+surrounding an already adapted patch,
 in particular treating those rare families that reach new patches
@@
 For instance, rock pocket mice are often found with a coat color that matches the rocks they live in,
-and the difference in coloration is known to be controlled genetically.
+these color differences are controlled genetically,
+and mice that don't match the local rock color are more likely to be eaten by predators.
 Sometimes, similar genetic changes have occurred independently in different patches,
@@
 However, the genetic basis could also be shared if
-rare migrants may carry the locally beneficial genotypes between nearby patches,
+migrants carry the locally beneficial genotypes between nearby patches,
 despite being at a disadvantage between the patches.
-We use a mathematical model of random migration to determine how quickly adaptation is expected to occur through new mutation or migration from other patches,
+We use a mathematical model of random migration to determine how quickly adaptation is expected to occur 
+through new mutation and through migration from other patches,
 and study in more detail what we would expect successful migrations between patches to look like.
@@
 The convergent evolution of similar phenotypes 
-in response to shared selection pressures is a testament to the power of selection 
-to repeatedly sculpt phenotypic variation. 
+in response to similar selection pressures is a testament to the power that selection 
+has to sculpt phenotypic variation. 
 In some cases this convergence extends to the molecular level, 
@@
 \citep{stern2013genetic,zhen2012parallel,martin2013repeated}. 
-Convergent adaptation can also occur within species, if individuals within a species 
-adapt to the same environment, in parallel, through different genetic changes. 
+Convergent adaptation also occurs within species, if different individuals
+adapt to the same environment through different genetic changes. 
 There are a growing number of examples of this in a range of well studied organisms and phenotypes \citep{Arendt:08}.
@@
 Standing variation for a trait can also increase the probability of convergence, 
-as standing variation for an allele increases the probability that the selective sweep will be \emph{soft}
+as this increases the probability that the selective sweep will be \emph{soft}
 (beginning from a base of multiple copies),
-leading to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}.
+which whether or not the copies derive from the same mutation,
+leads to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}.
 
@@
 that have repeatedly adapted to patches of soil with high concentrations of heavy metals
-(e.g.\ serpentine outcrops and mine tailings) \citep{kruckeberg1951intraspecific,turner2010serpentine,macnair1991evolution,schat1996identical};
+(e.g.\ serpentine outcrops and mine tailings) \citep{kruckeberg1951intraspecific,macnair1991evolution,schat1996identical,turner2010serpentine};
 the alleles conferring heavy metal tolerance are thought to be locally restricted 
@@
 One of the best-known examples is the rock pocket mouse (\textit{Chaetodipus intermedius}),
-which is on a number of black lava flows has much darker pelage than on intervening,
-much lighter soil \citep{DiceBlossom1937}.
+which on a number of black lava flows has much darker pelage than on intervening areas of lighter rock \citep{DiceBlossom1937}.
 Strong predator-mediated selection appears to favour such crypsis \citep{kaufman1974adaptive},
@@
 when a set of alleles provide an adaptive benefit in geographically localized patches, 
-separated by (inhabited) regions where the alleles are deleterious.
+separated by inhabited regions where the alleles are deleterious.
 The main questions are:
-Under what conditions is it likely that each patch adapts in parallel, through new mutation,
+Under what conditions is it likely that each patch adapts in parallel, 
+i.e.\ convergently through new mutation,
 and when is it likely that migration carries these alleles between patches?
@@
 using a variety of known results and new methods.
-In the section \nameref{ss:patchymutation} we develop a simple approximation,
+In the section \secref{ss:patchymutation} we develop a simple approximation,
 equation \eqref{eqn:mutrate}, for the rate at which patches become established by new mutations. 
-The most important conceptual work of the paper occurs in the section \nameref{ss:patchymigration},
+The most important conceptual work of the paper occurs in the section \secref{ss:patchymigration},
 where we develop an understanding of the process by which alleles move from an existing patch 
 to colonize a novel patch, culminating in equation \eqref{eqn:migrate} for the rate at which this happens.
-We combine these two results in the section \nameref{ss:probparallel} 
+We combine these two results in the section \secref{ss:probparallel} 
 to discuss to the probability of parallel adaptation, equation \eqref{eqn:parallel_prob}.
 To understand the genomic signal left by adaptations shared by migration, 
-in the section \nameref{ss:haplotype_length}, 
+in the section \secref{ss:haplotype_length}, 
 we approximate the time it will take for an allele to transit between patches,
@@
 and use this to approximate the length of haplotype that hitchhikes with it (equation \eqref{eqn:haplotype_length}). 
-We then quantify how quickly this haplotype decays within patches (equation \eqref{eqn:rec_rate_within}) 
-and use this to obtain properties about the equilibrium distribution of haplotype lengths 
-(equation \eqref{eqn:within_haplotype_length}), in the section \nameref{ss:patch_haplotype}. 
-Finally, in the section \nameref{ss:applications} we apply this work to understand the geographic scale of
+Finally, in the section \secref{ss:applications} we apply this work to understand the geographic scale of
 convergent evolution in \textit{Chaetodipus intermedius}.
@@
 We assume that this mutational change occurs at a (low) rate of $\mu$ per chromosome per generation,
-and that this initially rare new allele has fitness $1+s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches.
-Outside of these patches we assume that the new allele has a relative fitness
-$1-s_m$  when rare in the intervening areas, with $s_p$ and $s_m$ both positive.
+and that this initially rare new allele has fitness advantage $s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches.
+Outside of these patches we assume that the new allele has a fitness disadvantage of 
+$s_m$  when rare in the intervening areas, with $s_p$ and $s_m$ both positive.
 (Here we define ``fitness'' to be the intrinsic growth rate of the type when rare.)
-To simplify matters, we assume that the disadvantage $s_m$ 
+We assume that the disadvantage $s_m$ 
 is sufficiently large that, on the relevant timescale,
@@
 or if $p(x)$ is close to $p_e$ at some point and is symmetric about the edge of the patch. \linelabel{rr:p_estab}
-We examine this more generally via exact calculation of $p(x)$ in the section \nameref{apx:establishment_sims}.
+We examine this more generally via exact calculation of $p(x)$ in the section \secref{apx:establishment_sims}.
 
@@
 \end{align}
-i.e.\ the product of mutational influx in the patch ($\rho A\mu$) and the probability of establishment ($2s_p/\xi^2$).
+i.e.\ the product of mutational influx in the patch ($\rho A\mu$) and the probability of establishment ($p_e \approx 2s_p/\xi^2$).
 If this rate is low, then the time (in generations) until a mutation arises that
@@
 To determine this, 
-study the fine-scale genealogy of alleles that transit between patches,
-obtaining along the way other useful information about the genealogy
-of $B$ alleles. \gc{Doing this we arrive at \eqref{eqn:migrate} for
-  the rate at which patches adapt via migration. } %%Thought it good
-                                %%to signpost again where we are
-                                %%going.
+we study the fine-scale genealogy of alleles that transit between patches,
+obtaining along the way other useful information about the genealogy of $B$ alleles. 
+Doing this we arrive at \eqref{eqn:migrate} for the rate at which 
+an allele established in one patch spreads to a neighboring one by migration.
 
@@
 but only very few, and rarely, far away from the patch where $B$ is established.
-Denote the expected frequency of allele $B$ at location $x$ relative to the patch by $q(x)$,
+Denote the expected frequency of allele $B$ at displacement $x$ relative to the patch by $q(x)$,
 and assume that the shape of the patch is at least roughly circular.
@@
 From simulations, $C$ seems to be not too far from 1 in practice (see below).
-As J.\ Hermisson pointed out \gc{in comments} on an earlier draft,
+As J.\ Hermisson pointed out in comments on an earlier draft,
 this functional form has a simple intuition:
@@
 The ``renewal'' aspect of this same argument suggests that for equation~\eqref{eqn:eqfreq} to hold,
-$|x|$ must be larger than only 5--10 multiples of $\sigma$.
-\linelabel{rr:q_form}  \gc{OR WE could just personal ref Hermisson here.}
+$|x|$ must be larger than a few multiples of $\sigma$.
+\linelabel{rr:q_form}
 
@@
 To be clear about the assumptions (especially regarding scaling) implicit here,
-we provide a derivation below in \nameref{apx:eqfreq}, 
-as well as justification for the asymptotics below in \nameref{apx:asymptotics}.
+below provide a derivation of \eqref{eqn:eqfreq} in section \secref{apx:eqfreq}, 
+as well as justification for the asymptotics below in section \secref{apx:asymptotics}.
 In one dimension, the equation can be solved to give $q(x)$ 
@@
   Mean occupation frequencies of individual demes (grey dots)
-  and averaged over demes at similar distances (black circles), on
-  a one-dimensional grid of 201 demes, and
+  and averaged over demes at similar distances (black circles) from simulations, 
+  on a one-dimensional grid of 201 demes, and
   a two-dimensional grid of $101\times 101$ demes.
@@
   $\sigma = 0.95$ and 0.74 times the deme spacing, respectively.
+  More details are given below in \secref{ss:simulations}.
    }   \label{fig:sim_occupation_freqs}
@@
 will lead us eventually to the expression \eqref{eqn:migrate} for the rate of effective migration,
-an approximation that is valid if the distance $R$ is large (at least five times $\sigma/\sqrt{2s_m}$, say).
+an approximation that is valid if the distance $R$ is large (at least several times $\sigma/\sqrt{2s_m}$).
 
@@
 and if the individual arrives to a patch where the $B$ allele is favored, 
-we drop the individual from our family.
+we drop the individual from the family.
 Since the $B$ allele is uniformly deleterious elsewhere,
@@
 \end{align}
-In other words, the chance of survival to time $t$ is a constant multiplied by $e^{-s_m t}$,
+In other words, the chance of survival to time $t$ is asymptotically a constant multiplied by $e^{-s_m t}$,
 and conditional on survival, the family size is given by a fixed distribution.
@@
 If we take the expectation of \eqref{eqn:Zt_decomposition},
+since $\E[Z_t] = e^{-s_m t}$,
 we get the limiting mean family size conditioned on nonextinction:
 \begin{align} \label{eqn:mean_K}
-    \E[K] = 1 + \E[W \given W \ge 1] /(e^{s_m}-1) \approx \E[W \given W \ge 1] / s_m .
+    \E[K] 
+    % = 1 + \E[W-1 \given W \ge 1] \lim_{t \to \infty} \sum_{u=0}^t e^{-s_m u} = 1 + \E[W-1 \given W \ge 1] \lim_{t \to \infty} \frac{1-e^{-s_m (t+1)}}{1-e^{-s_m}}
+    = 1 + \E[W-1 \given W \ge 1] /(1-e^{-s_m}) \approx \E[W-1 \given W \ge 1] / s_m .
 \end{align}
@@
 To make this precise we need to understand what we mean by ``outflux of families'';
-we do this below in \nameref{ss:hitting_occupation},
+we do this below in section \secref{ss:hitting_occupation},
 and provide a heuristic argument here.
@@
 
-Suppose the family arrives with $K$ individuals,
-i.e.\ there are $K$ individuals that manage to establish
+Suppose the family arrives with $K$ individuals.
 The probability that at least one individual leaves a lineage that establishes in the patch
@@
 each of the $K$ newly arrived individuals would leave behind a lineage with 
-mean total size $\int_0^\infty e^{-t s_m} dt = 1/s_m$,
+mean total size $\sum_{t=0}^\infty e^{-t s_m} \approx 1/s_m$,
 roughly half of which would stay in the new patch.
@@
 or the distinction between occupation density and number of individuals
-are encouraged to take a detour through \nameref{ss:hitting_occupation}, below.)
+are encouraged to take a detour through section \secref{ss:hitting_occupation}, below.)
 
@@
 When we use this below, we set $p_e = 2 s_p / \xi^2$.
-\gc{Note that the approximation has the undesirable property that it is
+Note that the approximation has the undesirable property that it is
 non-monotonic in $s_m$, when we expect the rate of adaptation via
@@
 and is a decreasing function of $s_m$ for $s_m > |y|^2 / 2 \sigma^2$, 
-i.e.\ the parameter region that we are interested in.}
+which is the parameter region that we are restricted to by other assumptions.
 \linelabel{rr:sm_condition}
-%%GRAHAM felt like we needed to be clearer on this point.
 
@@
 it seems helpful to step back and review the assumptions 
-underlying the asymptotic results we have used.
+underlying the asymptotic results we have used, or will use below.
 Our results should hold exactly in the limit of 
@@
 
-%%Graham brokke this paragraph up.
-\gc{To be more precise,} the mutation rate results will hold best if $s_m \approx s_p$,
+To be more precise, the mutation rate results will hold best if $s_m \approx s_p$,
 if the patch diameter is large relative to $\sigma/\sqrt{\min(s_m,s_p)}$,
@@
 we assume that 
-\textbf{(a)} each patch is large enough to support a stable population of $B$ alleles
+each patch is large enough to support a stable population of $B$ alleles
 ($A^{1/d} > (\sigma/\sqrt{2s_p}) \tan^{-1}(\sqrt{s_m/s_p})$).
@@
 Next, in using equation \eqref{eqn:eqfreq}, we assume that 
-\textbf{(b)} the inter-patch distance is greater than the characteristic length
+the inter-patch distance is large relative to the characteristic length
 ($R > \sigma/\sqrt{s_m}$),
 and that
-\textbf{(c)} local drift is not so strong that the equilibirum frequency is at least approximately attained
+local drift is not so strong that the equilibrium frequency is at least approximately attained
 (using Wright's effective neighborhood size, $s_m \gg 1/(\rho \sigma^d)$).
@@
 we cannot approximate its dynamics via a branching process.
-We also neglect the time for migration-selection equilibrium to be reached,
-which has the effect of over-estimating $\migrate$.
+We also neglect the time for migration-selection equilibrium to be reached.
 As discussed above, we also assume that migration to a new patch takes place over a number of generations;
@@
 we used individual-based simulations on a one-dimensional lattice of 501 demes,
-run for 25,000 generations. \gc{We simulated $\rho$ diploid
-  individuals per patch, with nearest neighbor migration and
-  multiplicative selection??!? }
-More details are given below in \nameref{ss:simulations},
+with $\rho$ haploid individuals per deme,
+dispersal to nearby demes with $\sigma=0.95$,
+and run for 25,000 generations. 
+More details are given below in section \secref{ss:simulations},
 and the number of simulations used and parameter values are given in
@@
 First, we compute the time to reach 100 $B$ alleles, while theory predicts
-the time until the the progenitor of those 100 $B$ alleles arose.
+the time until the progenitor of those 100 $B$ alleles arose.
 Second, the expression for $\mutrate$ neglects effects near the boundary of the patch,
@@
 A glance at supplementary figure \ref{sfig:sims_8} shows more:
-since $s_m$ is so small, the second patch is reached by migrants before the migration-selection eqilibrium is reached,
+since $s_m$ is so small, the second patch is reached by migrants before the migration-selection equilibrium is reached,
 and at migration-selection equilibrium, the two patches act as a single meta-patch.
@@
 %%felt like we needed a better intro to this paragraph
-\gc{We first consider the order of magnitude that our parameters need to
-be on in order for adaptation via mutation or migration to dominant.} Let $\gamma = \min(1,s_m/p_e)$.
-An important point to notice is that effective migration and mutation rates will be on the same order if 
+We first consider the order of magnitude that our parameters need to
+be on in order for adaptation via mutation or migration to dominate.
+Let $\gamma = \min(1,s_m/p_e)$.
+Effective migration and mutation rates will be on the same order if 
 $A \mu \approx 2 \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
@@
 If $s_m=.001$, then with this higher mutational influx this calculation predicts that mutation will \emph{always} be faster than migration
--- however, this should be taken very cautiously,
+-- however, this should be taken with caution,
 since as discussed above, this model does not hold if $R$ is of the same order as $\sigma$
@@
 
-\gc{We can go beyond these orders of magnitude calculations to find
-  the probability of parallel adaptation if we are will to take our approximations at face value, 
-and assuming both rates are small. Doing so } 
-the time until the first of the two patches is colonized by the $B$
+We can go beyond these orders of magnitude calculations to find
+the probability of parallel adaptation if we are willing to take our approximations at face value. 
+Doing so, the time until the first of the two patches is colonized by the $B$
 allele will be approximately exponentially distributed with mean $1/(2
@@
       Probability that two patches 
-      evolve by independent mutations (i.e.\ in parallel),
+      evolve by independent mutations (i.e.\ convergently),
       as a function of their geographic separation ($R$)
@@
 \begin{equation} \label{eqn:total_migrate}
-  \sum_{i=1}^k \migrate(R_i) = C \frac{ 4 \rho s_p s_m }{\xi^2} \sum_{i=1}^{k} \left(R_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m} /\sigma\right).
+    \sum_{i=1}^k \migrate(R_i) = C \rho s_m \min(s_m,p_e) \sum_{i=1}^{k} \left(R_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m} /\sigma\right).
 \end{equation}
@@
 \begin{align}   \label{eqn:migprob}
-  \frac{\left(R_i \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m}
-  /\sigma\right)} {\sum_{j=1}^{k}  \left(R_i \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m}
-    /\sigma\right) } .
+  \frac{
+      \left(R_i \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m} 
+      /\sigma\right)
+  } {
+      \sum_{j=1}^{k}  \left(R_j \right)^{-(d-1)/2} \exp\left(- R_j \sqrt{2 s_m}
+    /\sigma\right) 
+} .
 \end{align}
@@
 
-Since the compound parameter $s_p \rho / \xi^2$ is common to both migration and mutation rates,
+Since the compound parameter $s_p \rho$ is common to both migration and mutation rates,
 it functions only as a time scaling of the process, 
@@
 share a fairly long stretch of haplotype. 
-Pairs of adapted individuals within a patch will \gc{initially} share a haplotype of mean genetic length of
+Pairs of adapted individuals within a patch will initially share a haplotype of mean genetic length of
 about $\log(\rho A s_p)/s_p$ around the selected site, 
@@
 This association gets slowly whittled down by recombination during interbreeding at the edge of the patch,
-but there will always be longer LD nearby to the selected site
-\gc{CITATION, BARTON?}.
+but there will always be longer LD nearby to the selected site \citep{barton1979geneflow}.
 
@@
 
-\gc{In conclusion, the length of shared haplotype drops off quickly with the
+In conclusion, the length of shared haplotype drops off quickly with the
 distance between the patches. When $s_m$ is small we expect shorter
-haplotypes as the allele is only weakly deleterious and so 
-traverses the distance between patches relatively slowly compared to
-strongly deleterious alleles. Therefore, when adaptation by migration
-is most likely (small $s_m$) 
-we expect shorter
-shared haplotypes.}
+haplotypes as the allele is only weakly deleterious and so successful migrants
+traverse the distance between patches relatively slowly 
+compared to successfully migrating strongly deleterious alleles. 
+Therefore, when adaptation by migration is most likely (small $s_m$) 
+we expect shorter shared haplotypes.
 
@@
 In the other direction, the dark MC1R allele was at 34\% at a site with light substrate 10km to the east of the flow (O'Neill).
-These numbers give us a sense of a plausible range of characteristic lengths.
+Note that such apparent asymmetry is expected,
+since as noted above the migration--selection equilibrium can be highly stochastic.
+These numbers give us a sense of a plausible range of the parameters.
 Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
@@
 of the distance between environmental patches. As a rough rule of
-thumb $1$cM is approximately $1$Mb in a number of mammalian species. \linelabel{rr:cMMb}
+thumb $1$cM is approximately $10^6$ bases in a number of mammalian species. 
   \label{fig:mice_prob_parallel}
@@
 This is in contrast with models of panmictic populations
-\citep{softsweepsI, MesserPetrov, Wilson-softsweep} 
+\citep{softsweepsI, MesserPetrov, wilson2014selective} 
 and geographically spread populations adapting to homogeneous selection pressures \citep{ralph2010parallel}, 
@@
 aids establishment of new alleles and the spread of existing
-alleles (but see \citet{Wilson-softsweep} for the complications of varying population sizes).
+alleles (but see \citet{wilson2014selective} for the complications of varying population sizes).
 
@@
 
-\paragraph{Decay of shared haplotypes} \nameref{rr:haplotype_discussion}
-To provide context for the results on shared haplotype length in \nameref{ss:haplotype_length}
-it is important to also understand
-the process by which haplotypes are whittled down within patches.
-The initial haplotype that sweeps within a patch 
-will be dispersed over time by recombination.
-Likewise, the haplotype that is shared between patches coadapted by migration
-(equation~\ref{eqn:haplotype_length}) will also break down. 
-However, a long time after the initial sweep,
-we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
-than with individuals elsewhere,
-since selection against migrants 
-decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
-The literature on clines \citep[e.g.]{barton1979geneflow} has important information,
-but a good deal more work is needed
-to provide robust estimates for these processes.
-
 \paragraph{Standing variation} 
@@
 have appeared independently or not.
-However, especially absolute if rates of adaptation by new mutation are very low,
+However, especially absolute if rates of adaptation by new mutation are low,
 any adaptation that does occur may have to make use of standing variation.
 The case of a panmictic population was studied by \citet{softsweepsI},
-and we study the case of a continuous, spatial population in an upcoming paper.
+and we study the case of a continuous, spatial population in an upcoming paper \citep{ralph2014standing}.
 If parallelism in local adaptation of the sort we study here is due to standing variation
@@
 then the dynamics of adaptation should not depend strongly on migration patterns
-(but note that the initial spatial distribution of standing variation may).
+(but the initial spatial distribution of standing variation may).
 
@@
 in a theoretical context.
-It is simliarly unclear how the model should extend to a polygenic trait.
+It is similarly unclear how the model should extend to a polygenic trait.
+
+\paragraph{Decay of shared haplotypes} \linelabel{rr:haplotype_discussion}
+To provide context for the results on shared haplotype length in
+section \secref{ss:haplotype_length}
+it is important to also understand
+the process by which haplotypes are whittled down within patches.
+The initial haplotype that sweeps within a patch 
+will be dispersed over time by recombination.
+Likewise, the haplotype that is shared between patches coadapted by migration
+(equation~\ref{eqn:haplotype_length}) will also break down. 
+However, a long time after the initial sweep,
+we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
+than with individuals elsewhere,
+since selection against migrants 
+decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
+The literature on clines \citep[e.g.][]{barton1979geneflow} has important information,
+but more work is needed
+to provide robust estimates for these processes.
+Questions about the genomic length-scale of signals of
+sweeps shared by migration have also been addressed in discrete
+population settings \citep{slatkin1998hitchhiking,kim2011hitchhiking}, reviewed in \citet{barton2000genetic}. 
+This work has shown that the length of the shared swept haplotype is often significantly
+shorter than the sweep with each patch, resulting in a pattern of
+shoulders of elevated $F_{ST}$ between adapted populations some distance away from
+the shared selected allele. It would be of interest to see how similar
+patterns can arise in a continuous population setting, as a way of
+uniting these results.
 
@@
 We have also ignored the possibility of very long distance migration,
-instead focusing on local dispersal (and using Gaussians via the central limit theorem).
+instead focusing on local dispersal (and using Gaussian and appealing to the central limit theorem).
 However, dispersal distributions can be very heavy tailed, 
@@
 could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
-Such tail events could greatly increase the probability of parallel adaptation above that predicted by our model. 
+Such tail events could greatly increase the probability of shared adaptation above that predicted by our model. 
 Furthermore, if adaptive alleles do move between distant patches via rare, long distance migration 
@@
 (the \texttt{R} code used is provided as a supplement).
-We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
+We simulated forward-time dynamics of the number of alleles of each type in a rectangular grid (either one- or two-dimensional) of demes with fixed size $N$.
 Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
@@
 Each then produced a random, Poisson distributed number of offspring that each, independently, 
-either remained in the same location with probabilty $1-m$ or
+either remained in the same location with probability $1-m$ or
 migrated a random number of steps with probability $m$
@@
 
-The reproduction rate in simulations for type $b$ alleles was 30\%; 
+The probability of reproduction in each generation in simulations for type $b$ alleles was 30\%; 
 this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
@@
 weighted by the number of $B$ alleles in that deme in the previous time step.
-If more than one lineage was found in a deme with $n$ $B$ alleles,
+If more than one lineage was found in a deme with $n$ alleles of type $B$,
 then each lineage picked a label uniformly from $1 \ldots n$,
@@
 copies of a new mutant allele are approximately independent and experience a uniform selective benefit;
-and can therefore be treated as a branching process (a spatial branching process in this case).
+and can therefore be treated as a branching process.
 Furthermore, whether or not a new, beneficial mutation establishes or is lost to demographic stochasticity
-is determined by this phase where it is rare.
+is determined by this initial phase where it is rare.
 Fortunately, the probability that a branching process does not establish
@@
 One route to the ``equilibrium frequency'' of the allele outside the range where it is advantageous is as follows;
-see \citet{slatkin1973geneflow} or \citet{barton1987establishment,pollak1966survival} (or \citet{KPP1937book} or \citet{fisher1937wave} or \citet{haldane1948theory})
+see \citet{slatkin1973geneflow} or \citet{barton1987establishment}
+(or \citet{KPP1937book} or \citet{fisher1937wave} or \citet{haldane1948theory})
 for other arguments in equivalent models,
-or \citet{etheridge2000introduction} and/or \citet{dawson1993measurevalued} for general theory used below.
-Following is a description of a discrete model (which is nearly the one we simulate from).
+and see \citet{etheridge2000introduction} and/or \citet{dawson1993measurevalued} for a general framework for the stochastic processes below.
+Following is a description of a discrete, haploid model (which is nearly the one we simulate from).
 Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
 and that selection (for or against) the allele is given by the function $s(x)$, with $x$ denoting the spatial location.
-Each individual at location $x$ reproduces after a random lifetime,
-dying, and then leaving behind in the same deme a random number of offspring with distribution given by $X$;
-each offspring then migrates independently to a new location chosen randomly from the distribution given by $x+R$,
-where she replaces a randomly chosen individual.
-If $x+R$ is outside of the range, then she perishes.
+Each individual at location $x$ reproduces at random, exponentially distributed intervals,
+producing a random number of offspring with distribution given by $X$
+who then all migrates to a new location chosen randomly from the distribution given by $x+R$,
+where they replace randomly chosen individuals.
+If $x+R$ is outside of the range, then they perish.
 Each individual's time until reproduction is exponentially distributed:
@@
 For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the random number of mutant alleles added if $X$ new offspring carrying mutant alleles
-replace randomly chosen individuals in a deme where the mutant allele is at frequency $\phi$ (i.e.~hypergeometric with parameters $(X,\phi)$);
-similarly, $\widetilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,1-\phi)$).
+replace randomly chosen individuals in a deme where the mutant allele is at frequency $\phi$ (i.e.~hypergeometric with parameters $(X,1-\phi)$);
+similarly, $\widetilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,\phi)$).
 (We like to think of $\Phi^N_t$ as a measure, but it does not hurt to think of $\Phi^N$ as a vector;
@@
 that $\var[R] = (\sigma \epsilon)^2$, and that the strength of selection is $s(x) \epsilon$,
-and suppose that $\Phi_t(x)$ is deterministic (and sufficiently differentiable) and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
+suppose that $\Phi_t(x)$ is deterministic and twice differentiable,
+and let $\xi(t,x) = \Phi_{t/\epsilon}(x)$;
 then the previous equation with $f(\Phi) = \Phi(x)$ converges to the familiar form:
@@
 
-In one dimension, we are done (see section~\ref{apx:elliptic_integrals});
+In one dimension, we are done (and discuss exact solutions in Appendix~\ref{apx:elliptic_integrals});
 in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
@@
 \begin{align}
-    u(r) = C' (r-r_0)^{(2-d)/2} K_{(2-d)/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
-\end{align}
-where $C'$ and $r_0$ are chosen to match boundary conditions.
+    u(r) = C' (r-r_1)^{(2-d)/2} K_{(2-d)/2} \left( (r-r_1) \sqrt{2s}/\sigma \right) ,
+\end{align}
+where $C'$ and $r_1$ are chosen to match boundary conditions.
 Asymptotics of Bessel functions (\citet{gradshteyn2007table}, 8.451.6) then imply that
@@
 Here we make a more precise argument to back up expression \eqref{eqn:migrate} for the migration rate.
-The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
+The argument made above in section \secref{ss:heuristics} applies to general migration mechanisms,
 since it relies only on a decomposition of the migrant families upon hitting the new patch;
@@
 Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
-Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
-such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ 
-that initiate new migrant families is $\gamma(x)$.
-Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``ouflux of families'',
+Denote by $\gamma(x)$ the mean rate of outflux of migrant families from a point $x \in \partial B(r_0)$,
+i.e.\ the time-averaged proportion of individuals near a point $x$ in $\partial B(r_0)$ 
+that are the founders of new migrant families.
+Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``outflux of families'',
 when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
-By examining the constituent terms,
+However, examining the constituent terms,
 it will turn out that equation \eqref{eqn:gestalt_migrate} 
@@
 First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
-Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
+Suppose that $Z$ is one such spatial branching process as above in section \secref{ss:migrant_math}, 
+started at time 0 with a single individual at $x$,
 and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
@@
 can be thought of  informally \linelabel{rr:density}
-as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
+as the expected total number of offspring of a family beginning at $x$ that ever live near $y$.
 It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
-i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
+i.e.\ the function $u(x,y)$ that, integrated over a set of locations $S$, 
 gives the expected total number of individuals ever living in $S$:
@@
 \begin{align}
-    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
+    u(x,y) = \int_0^\infty e^{-s_m t} p_t(x,y) dt .
 \end{align}
@@
 Recall that $\rho$ appears here but not in \eqref{eqn:occupation_integral}
-because $q(x)$ is the local \emph{proportion} of $B$ alleles.
+because both $q(x)$ and $\gamma(x)$ are local \emph{proportions} of $B$ alleles.
 
@@
 Following this intuition,
-and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-it turns out (see section~\nameref{ss:hitting_occupation}) that
+and using $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
+we show below that
 \begin{align} 
@@
 Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
-Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
+Since the integrands in expressions \eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral} only differ by a factor
 that (at least asymptotically) does not depend on the distance between $x$ and $y$,
@@
 as long as the patch is large enough.
-For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{ss:hitting_occupation}.
 
@@
 since the marginal distribution of a single lineage is Brownian,
+letting $u(x,S) = \int_S u(x,y) dy$,
 \begin{align}
@@
 $\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
-and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\min(t,\tau_\dagger)}$,
+and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\tau_S}$ conditioned on $\tau_S < \tau_\dagger$,
 by the strong Markov property this is equal to
@@
 Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
-(In the main body, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
+(Before, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
 If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
@@
     % &= \int_0^{2w} G_{s_m}(0,y) dy \\
+    \begin{split}
       &= \int_0^{2w/\sigma} \frac{ e^{- y \sqrt{2s_m}/\sigma }}{\sqrt{2s_m}} dy \\
-  &= (1-e^{-2w\sqrt{2s_m/\sigma}})/(2s_m)
+      &= (1-e^{-2w\sqrt{2s_m}/\sigma})/(2s_m)
+    \end{split}
 \end{align}
@@
   &= \frac{1}{s_m}\int_0^{2w\sqrt{2s_m}/\sigma} y K_0(y) dy \\
-  &= \frac{1}{2s_m}\left(1- \frac{2w\sqrt{2s_m}}{\sigma} K_1(2w\sqrt{2s_m}/\sigma) \right)  .
-\end{align}
+  &= \frac{1}{2s_m}\left(1- \frac{2w\sqrt{2s_m}}{\sigma} K_1(2w\sqrt{2s_m}/\sigma) \right)  ,
+\end{align}
+where $K_0$ and $K_1$ are modified Bessel functions of the second kind.
 In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.
@@
 for any $y$.  
-Take $\hat x = (1,0$ and $y=r \hat x$.  Then
+Take $\hat x = (1,0)$ and $y=r \hat x$.  Then
 \begin{align}
@@
 i.e.\ the Fisher-KPP equation with piecewise constant selection.
-Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
+Let $\xi(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
 so that 
@@
 
+\clearpage
+
 \begin{figure}
@@
     and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
-    The left panel has the predicted time to adaptation on the horizontal axis as in figure~\ref{fig:sim_times},
-    and the right panel has, for comparison, the raw distance between patches 
+    The upper panel has the predicted time to adaptation on the horizontal axis as in figure~\ref{fig:sim_times},
+    and the lower panel has, for comparison, the raw distance between patches 
     (which predicts time to adaptation, but not as well).
@@
   \caption{
-    Randomly chosen simulations with $s_m=0.1$, $\sigma\approx 1$, and $\rho$ varying.
+    Randomly chosen simulations of adaptation by new mutation
+    with $s_m=0.1$, $\sigma\approx 1$, and $\rho$ varying.
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
   \caption{
-    Randomly chosen simulations with $s_m=0.01$, $\sigma\approx 1$, and $\rho$ varying.
+    Randomly chosen simulations of adaptation by new mutation
+    with $s_m=0.01$, $\sigma\approx 1$, and $\rho$ varying.
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
   \caption{
-    Randomly chosen simulations with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
+    Randomly chosen simulations of adaptation by new mutation
+    with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
   \caption{
-    Randomly chosen simulations with $s_m=0.1$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    Randomly chosen simulations of adaptation by migration
+    with $s_m=0.1$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
   \caption{
-    Randomly chosen simulations with $s_m=0.01$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    Randomly chosen simulations of adaptation by migration
+    with $s_m=0.01$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
   \caption{
-    Randomly chosen simulations with $s_m=0.001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    Randomly chosen simulations of adaptation by migration
+    with $s_m=0.001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
   \caption{
-    Randomly chosen simulations with $s_m=0.0001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+    Randomly chosen simulations of adaptation by migration
+    with $s_m=0.0001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
 This therefore describes the movement of a lineage backwards in time.
-After the same large-population, continuous-space limit that underlies the rest of this paper (section~\nameref{apx:eqfreq}),
+After the same large-population, continuous-space limit that underlies the rest of this paper (section~\secref{apx:eqfreq}),
 if the assumption that fluctuations in $q(x)$ are small,

### 2014-12-17 analyze-sims.R

@@ -28,20 +28,21 @@ mutsims$adapted <- (
     ( mutsims$time1 >= mutsims$hit100.1 ) )
 
 mutsims$adapttime <- mutsims$hit100.1
-pdf(file="mutation-times-predicted.pdf", width=5, height=5, pointsize=10)
+pdf(file="mutation-times-predicted.pdf", width=4, height=5, pointsize=10)
 with( subset(mutsims,adapted), 
         plot( muttime, adapttime, log='xy', xlab="mean time to adaptation by mutation", ylab="time to hit 100 in patch",
-                col=sm.name, pch=as.numeric(N.name) )
+                xlim=c(120,120000), col=sm.name, pch=as.numeric(N.name) )
         )
 abline(0,1,untf=TRUE)
 # abline(coef(muttime.lm),col='blue',lwd=2,untf=TRUE)
 for ( smval in 1:nlevels(mutsims$sm.name) ) {
-    with( subset(mutsims,adapted & as.numeric(sm.name)==smval), 
+    with( subset(mutsims,adapted & as.numeric(sm.name)==smval),  {
+        muttime.fac <- factor(format(muttime,digits=1))
         points( 
-            sort(unique(muttime)), 
-            tapply(adapttime,factor(muttime,levels=sort(unique(muttime))),mean,na.rm=TRUE), 
+            tapply( muttime, muttime.fac, mean, na.rm=TRUE ),
+            tapply(adapttime, muttime.fac, mean,na.rm=TRUE), 
             cex=3, pch=20 )
-    )
+    } )
 }
 legend("bottomright", legend=c(levels(mutsims$sm.name),levels(mutsims$N.name)), pch=c(rep(1,nlevels(mutsims$sm.name)),1:nlevels(mutsims$N.name)), col=c(1:nlevels(mutsims$sm.name),rep(1,nlevels(mutsims$N.name))) )
 dev.off()
@@ -110,19 +111,21 @@ migsims$valid <- with( migsims,
     )
 
 
-pdf(file="migration-time-predicted.pdf",width=9, height=5, pointsize=10)
-layout(t(1:2))
+pdf(file="migration-time-predicted.pdf",width=4, height=8, pointsize=10)
+layout((1:2))
 with( subset(migsims,adapted), {
-        plot( migtime, hit100.2, col=sm.name, pch=as.numeric(N.name), log='xy', xlim=c(10,1e8), ylab="time to hit 100 in second patch", xlab="mean migration time") 
[…]

### 2014-12-18 patchy-review-responses.tex
one small change to response to reviews
resubmitted

@@
 
-We thank the editor and the reviewers for their helpful comment. We
+We thank the editor and the reviewers for their helpful comments. We
 have substantially revised our paper in response to these comments,
@@
 \end{itemize}
+
+In addition to the format required by PLoS, we have again provided a version of the manuscript with all the figures in the right places.  
+Also note that in both pdfs, references and sections names are hyperlinks to the correct locations in the manuscript.
 

### 2014-12-18 patchy-selection-paper.tex
resubmitted

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{etoolbox}
 
 \setcounter{secnumdepth}{0}
+
+% turn this on to doublespace and put all figures at the very end.
+\newif\ifsubmissionversion
+% \submissionversiontrue
+\submissionversionfalse
 
@@
 % Responses to reviews:
-\input{review-response-commands}
 % set this to show line numbers and include responses to reviews or not
-\newif\ifreviewresponses\reviewresponsestrue  % include them
+\newif\ifreviewresponses
+\reviewresponsestrue  % include them
 % \reviewresponsesfalse  % don't include them
-\ifreviewresponses\else\renewcommand{\linelabel}[1]{}\fi
 \iflatexml\reviewresponsesfalse\fi  % they don't work so well in html
@@
 \newcommand{\responsefile}{patchy-review-responses.tex}  % name of the review reponses file
+
+% override above choices (see makefile)
+\ifdef{\submissionnofigs}{  % for submission, no figures
+  \reviewresponsesfalse
+  \submissionversiontrue
+}{
+}
+\ifdef{\submissionfigs}{  % for reviewers, with figs
+  \reviewresponsestrue
+  \submissionversionfalse
+}{
+}
+\ifdef{\arxiv}{  % for general use
+  \reviewresponsesfalse
+  \submissionversionfalse
+}{
+}
+
+\ifreviewresponses
+  \input{review-response-commands}
+\else
+  \newcommand{\linelabel}[1]{}
+  \newcommand{\revpoint}[2]{}
+  \newcommand{\includereviews}{}
+\fi
 
@@
 
-% turn this on to doublespace and put all figures at the very end.
-\newif\ifsubmissionversion
-% \submissionversiontrue
-\submissionversionfalse
-
 \ifsubmissionversion
-    \usepackage{endfloat}
+    \usepackage[tablesfirst]{endfloat}
     \usepackage{cite}
@@
 
-% \newcommand{\var}{\mathop{\mbox{Var}}}
-% \newcommand{\Exp}{\mathop{\mbox{Exp}}}
 \DeclareMathOperator{\var}{Var}
@@
 and do not discuss the (interesting) cases where rare, long-distance dispersal events
-are more important \citep[see e.g.][for discussion]{levin2003ecology,ralph2010parallel,hallatschek2014acceleration}.  \linelabel{rr:longtailed}
+are more important (see e.g.\ \citet{levin2003ecology,ralph2010parallel,hallatschek2014acceleration} for discussion).  \linelabel{rr:longtailed}
 %We also aim to describe the scenario where it is sensible to think of large, discrete patches;
@@
 decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
-The literature on clines \citep[e.g.][]{barton1979geneflow} has important information,
+The literature on clines (e.g.\ \citet{barton1979geneflow}) has important information,
 but more work is needed
@@
 i.e.\ the Fisher-KPP equation with piecewise constant selection.
-Let $\xi(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
+Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
 so that 
@@
 
+\section*{Supplemental material}
+
+% restart figure and table numbering
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{table}{0}
+
 \begin{figure}
@@
     \includegraphics{example-migration-sims/76178-r1-501-sb0_01-sm-0_1-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
+    \includegraphics{example-migration-sims/95095-r1-501-sb0_01-sm-0_1-N100-pophistory-run}
   \end{center}
@@
         % made in patchy-sim-writeup-plots.R
-        \includegraphics{example-equilibrium}
+        % \includegraphics{example-equilibrium}
     \end{center}

### 2014-12-21 patchy-selection-paper.tex
hours of latex-ing later, resubmitted

@@
 \usepackage{etoolbox}
+
+% make supp figures separate
+\usepackage{caption}
+\DeclareCaptionType{sfigure}[Figure]
 
@@
 \ifsubmissionversion
-    \usepackage[tablesfirst]{endfloat}
+    \usepackage[nolists,nomarkers,nohead]{endfloat}
     \usepackage{cite}
+    \DeclareDelayedFloat{sfigure}[sss]{Supplementary Figure}
     % Use the PLoS provided bibtex style
@@
 
-\ifsubmissionversion
-% \linespread{1.6}
-%\usepackage{setspace} 
-%\doublespacing
-\fi
-
 % Remove brackets from numbering in List of References
@@
 
-\pagestyle{myheadings}
+% \pagestyle{myheadings}
 %% ** EDIT HERE **
@@
 %%%%%%%%%%%
-\abstract{
+\section*{Abstract}
 Species often encounter, and adapt to, many patches of locally similar environmental conditions across their range. 
@@
 
-\textbf{Author Summary:}
+\section*{Author Summary}
 Often, a large species range will include patches where the species differs because it has adapted to locally differing conditions.
@@
 and more generally in understanding how species adapt to patchy, heterogeneous landscapes.
-}
 
@@
 
-\subsection*{Acknowledgements}
-We would like to heartily thank Joachim Hermisson for a number of insightful suggestions,
-and catching the algebra error present in the first version of this paper on the bioRxiv,
-as well as the two anonymous reviewers.
-We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
-as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
-for helpful comments and discussion of the manuscript.
-
 %%%%%
@@
 the transition is fairly symmetrical about the edge of the patch, even if the edge of the patch is not sharp.
-The fit is equally good for other parameter values,
-even if migration can move further than one deme and offspring numbers are not Poisson (results not shown).
+Additional experimentation indicated that the fit remains equally good for other parameter values,
+even if migration can move further than one deme and offspring numbers are not Poisson.
 This lends credence to our approximation that the integral of $p(x)$ over the entire range
@@
 
+\section*{Acknowledgements}
+We would like to heartily thank Joachim Hermisson for a number of insightful suggestions,
+and catching the algebra error present in the first version of this paper on the bioRxiv,
+as well as the two anonymous reviewers.
+We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
+as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
+for helpful comments and discussion of the manuscript.
+
 \bibliography{standing_patches_refs}
 
+\ifsubmissionversion
+\clearpage
+\renewcommand*\listfigurename{Figure Legends}
+\listoffigures
+\fi
+
+\clearpage
+
+% endfloat uses different names
+\ifsubmissionversion
+  \renewcommand{\thesfigure}{S\arabic{sfigure}}
+  \renewcommand{\thepostsfigure}{S\arabic{postsfigure}}
+  % \renewcommand{\thefigure}{S\arabic{figure}}
+  % \renewcommand{\thepostfigure}{S\arabic{postfig}}
+  % \setcounter{postfig}{0}
+  \renewcommand{\thetable}{S\arabic{table}}
+  \renewcommand{\theposttable}{S\arabic{posttable}}
+  \setcounter{posttable}{0}
+\fi
+
+%% list of supplementary figures
+\ifsubmissionversion
+  \renewcommand*\listsfigurename{Supplementary Figure Legends}
+  \listofsfigures
+  \renewcommand*\listtablename{Supplementary Table Legends}
+  \listoftables
+  \pagebreak
+\fi
+
 \appendix
+\section{Supporting Information}
 
@@
 %%%%%%%
-\section{The exact solution to the equilibrium frequency}
+\subsection{The exact solution to the equilibrium frequency}
 \label{apx:elliptic_integrals}
@@
 %%%%%%%
-\section{More information about the simulations}
+\subsection{More information about the simulations}
 \label{apx:parameter_tables}
@@
 
+%%%
+% Tables
+
 % table of parameters, produced by analyze-sims.R
@@
 
-\clearpage
-
-\section*{Supplemental material}
-
-% restart figure and table numbering
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{figure}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{table}{0}
-
-\begin{figure}
+%%%
+% Supplementary Figures
+
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sim_migration_times}
-\end{figure}
-
-\begin{figure}
+\end{sfigure}
+
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sim_mutation_times}
-\end{figure}
+\end{sfigure}
 
@@
 % sm=0.1
-\begin{figure}
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sims_1}
-\end{figure}
+\end{sfigure}
 
 % sm=0.01
-\begin{figure}
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sims_2}
-\end{figure}
+\end{sfigure}
 
 % sm=0.001
-\begin{figure}
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sims_3}
-\end{figure}
+\end{sfigure}
 
 % sm=0.0001
-\begin{figure}
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sims_4}
-\end{figure}
+\end{sfigure}
 
 % sm=0.1, R=80
-\begin{figure}
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sims_5}
-\end{figure}
+\end{sfigure}
 
 % sm=0.01, R=80
-\begin{figure}
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sims_6}
-\end{figure}
+\end{sfigure}
 
 % sm=0.001, R=80
-\begin{figure}
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sims_7}
-\end{figure}
+\end{sfigure}
 
 % sm=0.0001, R=80
-\begin{figure}
+\begin{sfigure}
   \begin{center}
@@
   } \label{sfig:sims_8}
-\end{figure}
-
-\ifsubmissionversion
-\processdelayedfloats
-\fi
+\end{sfigure}
 

### 2014-12-23 patchy-review-responses.tex
accepted resubmission

@@
 
+  \item In response to a request by the PLoS Genetics staff to ``either amend the supplementary material to include the referenced data or remove as many of the references [to ``results not shown''] as possible.''
+    we have changed the sentence 
+    ``The fit is equally good for other parameter values,
+    even if migration can move further than one deme and offspring numbers are not Poisson (results not shown).'' 
+    to 
+    ``
+    Additional experimentation indicated that the fit remains equally good for other parameter values,
+    even if migration can move further than one deme and offspring numbers are not Poisson.
+    ''
+    The additional experimentation is for an upcoming paper, 
+    but one that is in an early enough draft that it does not seem useful to cite it.
+    We are happy to remove the sentence if that would be better.
+
 \end{itemize}
@@
 In addition to the format required by PLoS, we have again provided a version of the manuscript with all the figures in the right places.
+Page and line numbers in the response to reviewers below refer to this document.
 Also note that in both pdfs, references and sections names are hyperlinks to the correct locations in the manuscript.

### 2014-12-23 patchy-selection-paper.tex
accepted resubmission

@@
 In one dimension, the equation can be solved to give $q(x)$ 
-in terms of Jacobi elliptic functions; see appendix \ref{apx:elliptic_integrals}.
+in terms of Jacobi elliptic functions; see the Supporting Information.
 
@@
 so that the probability of parallel mutation should increase
-approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{s_m}$. \revpoint{2}{13}
+approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{2 s_m}$. \revpoint{2}{13}
 
@@
 
-In one dimension, we are done (and discuss exact solutions in Appendix~\ref{apx:elliptic_integrals});
+In one dimension, we are done (and discuss exact solutions in the Supporting Information);
 in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
@@
 \appendix
-\section{Supporting Information}
+\section*{Supporting Information}
 
@@
 %%%%%%%
-\subsection{The exact solution to the equilibrium frequency}
+\subsection*{The exact solution to the equilibrium frequency}
 \label{apx:elliptic_integrals}
@@
 
-\subsection{Doing the integrals}
+\subsection*{Doing the integrals}
 \label{apx:some_integrals}
@@
 %%%%%%%
-\subsection{More information about the simulations}
+\subsection*{More information about the simulations}
 \label{apx:parameter_tables}
@@
 
-\section{Temporary work: hitting probs}
+\section*{Temporary work: hitting probs}
 

### 2015-02-26 patchy-review-responses-second-round.tex
added reviews, second round

% \documentclass{article}
% \usepackage{fullpage}
% \input{review-response-commands}
% 
% \begin{document}

\section*{From the editor}

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
  {\it PLoS Genetics}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    Molecular and Computational Biology\\
    University of Southern California, \\
    \emph{and} Evolution and Ecology\\
    University of California at Davis\\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

\begin{quote}
    Both reviewer 1 and reviewer 2 still question your main result presented in Eq. 1 and I do agree with them that there is definitely something wrong with the dimensions. This was already pointed in the first round of reviews and both reviewers feel that this issue has not been satisfactorily addressed. Both reviewer also argued in the first round that the derivations in the paper are overly complicated and that the problem can in fact be intuitively understood via much simpler arguments (both came up with very similar approaches independently). You stated in your replies that you prefer to stick with your derivation because you consider it more solid. However, I am not even fully convinced that your Eq. 11 is more correct than the reviewers’ intuitive results. Reviewer 1 points out that there are quite substantial deviations in your numerical results (Fig. 4) and reviewer 2 stresses that the 1D simulations won’t even be able to catch the potential problems associated with the boundary of the target area. Clearly, something is not yet fully understood here and I don’t see how the paper can be accepted without having settled these issues.
\end{quote}

\begin{quote}
    It would be easy to simply reject the paper at this point. However, as I expressed earlier, I still think that this is an extremely interesting problem that could be of great relevance for future studies, given that the authors can convincingly resolve the issues raised above. The dimensionality problem clearly needs to be resolved, and it needs to be explained how their result and the intuitive results from the reviewers compare with each other.
\end{quote}

\begin{quote}
    I also think that presenting and discussing the intuitive results really would be important for the paper, as originally suggested by reviewers 1 and 2. There is a tendency in population genetics to derive complicated “exact” results that are nonetheless of only limited usefulness in reality because it’s unclear how realistic the underlying models are in the first place. I won’t hide here that I strongly prefer a heuristic approach that reveals what’s actually going on, over a lengthy technical calculation that may be formally more correct, yet it’s unclear whether this really matters in reality. I feel the majority of the audience of PLoS Genetics is probably with me on this.
\end{quote}

\begin{quote}
    Should you decide to tackle these problems in a major revision, please also address the remaining comments raised by the reviewers. Sorry that I can’t bring you better news at this point but I think you will agree that the reviews are pretty clear in what needs to be done in order to get this paper accepted in PLoS Genetics.
\end{quote}

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf Peter Ralph, and\\
Graham Coop.}\\
\end{minipage}

\vskip 4em

%%%%%%%%%%%%%%
\reviewersection

%%%%%%
\point{}{
    The issue of dimensions and geometry of the target patch is the main difficulty I have with the paper. I just don't see how your equation 11 can be correct. It has dimensions of 1/time/area when it should be 1/time. Somehow the target area needs to play a role. Reviewer 2 has raised similar issues.
}

\reply

%%%%%%
\point{}{
Regarding the simulations presented in Fig 4, I don't think they strongly support the functional form of $\migrate$. Points with very similar observed migration time have predicted migration times that differ by a factor of a 100 (the more extreme are outside the validity of the approximations, but still the I don't think this figure is strong evidence that the math is correct.)
}

\reply

%%%%%%
\point{eq.~S6 is wrong}{
    as easily seen with $\Theta=0$ \ldots
    so eq.~S6 has an extra 2, resulting from a wrong expansion of $\sqrt{1+x}$,
    and so eq.~57--61 are also wrong.
}

\reply

%%%%%%
\point{time in transit}{
    One can obtain these results by a simple saddle point approximation:
    \begin{align*}
        \P\{ \text{getting to $x$ in time $t$} \} &= \P\{\text{diffusion to $x$ in time $t$}\} \P\{\text{surviving}\} \\
                                                  &= \frac{1}{(2\pi\sigma^2)^{d/2}} e^{-\frac{x^2}{2\sigma^2 t}-st}
    \end{align*}
    the exponent is minimal when
    \[
        \frac{x^2}{2\sigma^2 t^2} = s \quad \Rightarrow \quad t^* = \frac{x}{\sigma\sqrt{2s}},
    \]
    this gives your eq.~17.
    The variance is simply the inverse of the second derivative of the exponent
    \[
        \frac{1}{\var[\tau]} = \frac{x^2}{\sigma^2 (t^*)^3} = \frac{\sigma}{x(2s)^{3/2}} .
    \]
}

\reply

%%%%%%
\point{now to the thornier issues of dimension, geometry, and fluxes}{
    \begin{itemize}
        \item [with $\ell = \sigma/\sqrt{2s}$, $r$ the radius of the patch, $R$ the distance between patches, and $\ell \ll r \ll R$]
        \item the average density is $q(x) \sim e^{-x\sqrt{2s}/\sigma} = e^{-x/\ell}$
        \item but fluctuates far from the source as [families of size $K$, with local density $k/\ell^d$]
        \item families of size $1/s$ live for $1/s$ generations and hence contribute $1/s^2$ to the time-averaged density $q(x)$
        \item \[ q(x) = \lambda_\text{arr}(x) / s^2 \ell^d, \]
            where $\lambda_\text{arr}(x)$ is the rate at which families arrive that overlap with $x$
        \item \[ \lambda_\text{arr}(x) = q(x) s^2 \ell^d \]
            and
            \[ \Lambda_\text{arr}(x) = \lambda(x) \ell^d = q(x) s^2 \]
            would be an ``arrival density''
        \item to get at $\migrate$, arrival needs to be integrated over the relevant area. In $d=1$ this is simply a stretch of length $\ell$,
            in $d=2$ it should be $\sqrt{\ell^3 r}$,
            hence one should get
            \[ \migrate \sim \rho q(x) s^2 \ell^d \sqrt{r/\ell} (1-e^-{p_0/s}) \]
        \item result does depend on target geometry
        \item formulation in terms of fluxes might be easier
    \end{itemize}
}

\reply

%%%%%%%%%%%%%%
\reviewersection

\begin{quote}
I think that the manuscript has been improved, but I don't think that all the reviewers' points have been adequately addressed. As I wrote in my previous review, I really like the authors' approach of giving simple, intuitive explanations. But writing this kind of paper actually requires a much better understanding of the model and dynamics than one needs to write a paper with the standard formal derivations, and I'm not sure that the authors are there yet.
\end{quote}

%%%%%%
\point{1.}{
I still think that the main result (11) is dimensionally inconsistent -- it needs to be integrated over the boundary layer of the patch. In their response to my original review, the authors just referred to their response to a similar point by reviewer 1, but I don't think that this addresses my concern -- reviewer 1 was talking about integrating over a different area, the one occupied by a family, which I agree is not correct. (Note that the 1D simulations aren't going to be able to catch this issue, since you can't play with the length of the boundary.)
}

\reply

%%%%%%
\point{2.}{
I do think that reviewer 1 raises a good point though. Looking back through the manuscript, I can't find where the authors show that the family is localized within a lengthscale ~ sigma/\sqrt{s_m} << A^{1/d}. This is important, since it's needed to justify the approximation that the whole family arrives at the patch together. This seems like a basic fact about subcritical branching random walks/Brownian motions, so there's probably a formal proof of it somewhere in the literature, but I think that it's straightforward to give an intuitive explanation.
}

\reply

%%%%%%
\point{3.}{
Speaking of the literature, what's known about hitting probabilities for subcritical branching random walks/Brownian motions?
}

\reply

%%%%%%
\point{4.}{
I still think that the derivation of the main result is too complicated, and could be made much simpler along the lines of the suggestion in my previous review. The extra complications introduced in the current version do not actually make the result any more rigorous -- they just take a less direct route to introducing the same approximations.
}

\reply

%%%%%%
\point{Minor, 1.}{
I don't totally understand the simulation methods. The authors say that individuals first chose whether or not to reproduce with probability depending on their fitness, and then produced a Poisson-distributed number of offspring. So the selection is all in whether or not individuals reproduce, not in the number of offspring conditioned on reproduction? What exactly was the function for the probability of reproduction, and what was the mean of the offspring distribution? I'm just wondering if there's a detail here that would account for the discrepancy between the analysis and the simulations. Also, it would be helpful to say exactly what the value of sigma is in the chosen migration model.
}

\reply

%%%%%%
\point{Minor, 2.}{
I'm not sure that the description of the valid parameter range in "A review, and a test of the assumptions" is entirely correct. For the mutation results, it's fine if $s_m$ and $s_p$ have very different magnitudes, as long as the patches are sufficiently big/migration is sufficiently weak. What precisely is the condition here?
}

\reply

%%%%%%
\point{Minor, 3.}{
In the Applications, I understand that there's still a lot to learn about this system, but the authors need to at least say how far apart the actual lava flow patches are, so we have a sense of what their estimate of "tens to a few hundred kilometers" means for mutation vs migration.
}

\reply

%%%%%%
\point{Minor, 4.}{
In their response to my comments, the authors said that they did not see why the assumption $R \gg w$ was neceessary. I think that it is needed in order to talk sensibly about a single "distance to a patch" $R$: if $w \simgt R$, then generally some parts of the patch will be much closer than others, right?
}

\reply

%%%%%%
\point{Minor, 5.}{
I still think that including any sub-leading terms in the length of the shared haplotype is falsely precise.
}

\reply

%%%%%%%%%%%%%%
\reviewersection

\begin{quote}
I appreciate the careful and thorough revision - and especially the exhaustive simulation work. It is an impressive and truly content-rich manuscript. I have only a few minor points:
\end{quote}

%%%%%
\point{Figure 1:}{
    point panel, $s_m = -0.02$ appears twice and $s_b$ should rather be $s_p$?
}

\reply

%%%%%
\revpoint{Figure 4:}{
    The new figure is a bit hard to read: the grey symbols are hardly visible, square symbols appear with two meanings, the colours for the filling do not match the legend ... Maybe reduce the number of parameter combinations shown (drop some intermediate values) to focus on the "main message" and get a clearer picture? - Details can still be given in the supplement.
}

\reply

%%%%%
\point{Figure 4:}{
    Mention all simulation parameters in the legend (eg: what is $p_d$?)
}

\reply

%%%%%
\point{point 169:}{
}{
    my name in acknowledgements is enough, cut here
}

\point

%%%%%
\revpoint{line 176:}{
    *we* provide (?)
}

\point

%%%%%
\revpoint{line 262:}{
    cut "depends" (?)
}

\reply

%%%%%
\revpoint{line 263:}{
    time to reach 100 B alleles: simulations should show how long this takes $\to$ can this explain anything or is it irrelevant? (Unclear in Fig 4 since value for $s_p$ is not provided).
}

\reply

### 2015-02-28 patchy-selection-paper.tex
fixing geiger ref

@@
 % \newcommand{\responsefile}{pbio-reviews-19sept12-responses.tex}  % name of the review reponses file
-\newcommand{\responsefile}{patchy-review-responses.tex}  % name of the review reponses file
+\newcommand{\responsefile}{patchy-review-responses-second-round.tex}  % name of the review reponses file
 
@@
 We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} (see also \citet{chauvin1991growing}). 
+% (lyons1995conceptual deals with conditioned on eventual nonextinction)
 When we condition on survival until $t$, we condition on existence of
@@
 Given this lineage, nothing distinguishes any of the other reproduction events --
-each individual along the trunk lineage must give birth to at least one offspring,
+each individual along the trunk lineage must give birth to at least one offspring that survives until time $t$,
 and all other reproduction events are unconditioned.
@@
 that arose from the main branch in generation $s$,
-and each $W_s$ is an independent copy of the offspring distribution $W$, conditioned on $W_s \ge 1$.
+and each $W_s$ is independent with distribution $\P\{ W_s = k\} = \P\{ W = k \} (1-k_e(t-s)^k)/(1-k_e(t-s+1))$. \linelabel{ll:branching_process_fix}
 Since each subfamily $Z^{(s,k)}$ is a branching process with mean offspring number less than one,
@@
     \E[K] 
-    % = 1 + \E[W-1 \given W \ge 1] \lim_{t \to \infty} \sum_{u=0}^t e^{-s_m u} = 1 + \E[W-1 \given W \ge 1] \lim_{t \to \infty} \frac{1-e^{-s_m (t+1)}}{1-e^{-s_m}}
     = 1 + \E[W-1 \given W \ge 1] /(1-e^{-s_m}) \approx \E[W-1 \given W \ge 1] / s_m .
@@
 \begin{align} \label{eqn:gA_1D}
+  \begin{split}
     g(A) % &= \E^w[ \int_0^\tau \one_A(B_t) dt ] 
     % &= \int_0^{2w} G_{s_m}(0,y) dy \\
-    \begin{split}
       &= \int_0^{2w/\sigma} \frac{ e^{- y \sqrt{2s_m}/\sigma }}{\sqrt{2s_m}} dy \\
@@
     r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
-            &= - 2 r_0 \cos \theta + O(1/r) .
+            &= - r_0 \cos \theta + O(1/r) .
 \end{align}
@@
 \begin{align}
-  \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-2 r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} .
+  \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} .
 \end{align}
@@
   \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-  &= C' \int_0^{2 \pi} e^{-2 r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
-  &= 2 \pi C' I_0(-2r_0\sqrt{2 s_m}/\sigma),
+  &= C' \int_0^{2 \pi} e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
+  &= 2 \pi C' I_0(-r_0\sqrt{2 s_m}/\sigma),
 \end{align}

### 2015-05-26 patchy-review-responses-second-round.tex

@@
 %%%%%%
-\point{eq.~S6 is wrong}{
+\point{eq.~56 is wrong}{
     as easily seen with $\Theta=0$ \ldots
-    so eq.~S6 has an extra 2, resulting from a wrong expansion of $\sqrt{1+x}$,
+    so eq.~56 has an extra 2, resulting from a wrong expansion of $\sqrt{1+x}$,
     and so eq.~57--61 are also wrong.
@@
 \reply
+Thanks for catching that mistake.  We removed the extra factor of 2 (section~\nameref{apx:outflux}).
 
@@
 \reply
+true, depends if you think ``saddle point approximation'' or ``laplace transform'' is a simpler idea?  this way avoids writing down more integrals; I hate to add more equations...
 
@@
     \begin{itemize}
-        \item [with $\ell = \sigma/\sqrt{2s}$, $r$ the radius of the patch, $R$ the distance between patches, and $\ell \ll r \ll R$]
+        \item {} [with $\ell = \sigma/\sqrt{2s}$, $r$ the radius of the patch, $R$ the distance between patches, and $\ell \ll r \ll R$]
         \item the average density is $q(x) \sim e^{-x\sqrt{2s}/\sigma} = e^{-x/\ell}$
@@
 
+perhaps the reviewer is mistaking a different set of terminology and basic concepts for a lack of understanding?
+in this process we have apparently too quickly judging physics-style arguments to be incorrect, some of the time.
+
 %%%%%%
-\point{1.}{
+\point{}{
 I still think that the main result (11) is dimensionally inconsistent -- it needs to be integrated over the boundary layer of the patch.
@@
 %%%%%%
-\point{2.}{
+\point{}{
 I do think that reviewer 1 raises a good point though.
-Looking back through the manuscript, I can't find where the authors show that the family is localized within a lengthscale ~ sigma/\sqrt{s_m} << A^{1/d}.
+Looking back through the manuscript, I can't find where the authors show that the family is localized within a length scale $\sim \sigma/\sqrt{s_m} \ll A^{1/d}$.
 This is important, since it's needed to justify the approximation that the whole family arrives at the patch together.
@@
 %%%%%%
-\point{3.}{
+\point{}{
 Speaking of the literature, what's known about hitting probabilities for subcritical branching random walks/Brownian motions?
@@
 %%%%%%
-\point{4.}{
+\point{}{
 I still think that the derivation of the main result is too complicated, and could be made much simpler along the lines of the suggestion in my previous review.
@@
 %%%%%%
-\point{Minor, 1.}{
+\point{}{
 I don't totally understand the simulation methods.
@@
 %%%%%%
-\point{Minor, 2.}{
+\point{}{
 I'm not sure that the description of the valid parameter range in "A review, and a test of the assumptions" is entirely correct.
@@
 %%%%%%
-\point{Minor, 3.}{
+\point{}{
 In the Applications, I understand that there's still a lot to learn about this system, but the authors need to at least say how far apart the actual lava flow patches are, so we have a sense of what their estimate of "tens to a few hundred kilometers" means for mutation vs migration.
@@
 %%%%%%
-\point{Minor, 4.}{
+\point{}{
 In their response to my comments, the authors said that they did not see why the assumption $R \gg w$ was neceessary.
-I think that it is needed in order to talk sensibly about a single "distance to a patch" $R$: if $w \simgt R$, then generally some parts of the patch will be much closer than others, right?
+I think that it is needed in order to talk sensibly about a single "distance to a patch" $R$: if $w \gtrsim R$, then generally some parts of the patch will be much closer than others, right?
 }
@@
 %%%%%%
-\point{Minor, 5.}{
+\point{}{
 I still think that including any sub-leading terms in the length of the shared haplotype is falsely precise.
@@
 %%%%%
-\revpoint{Figure 4:}{
+\point{Figure 4:}{
     The new figure is a bit hard to read: the grey symbols are hardly visible, square symbols appear with two meanings, the colours for the filling do not match the legend ...
@@
 
-\point
+\reply
 
 %%%%%
-\revpoint{line 176:}{
+\point{line 176:}{
     *we* provide (?)
@@
 
-\point
+\reply
 
 %%%%%
-\revpoint{line 262:}{
+\point{line 262:}{
     cut "depends" (?)
@@
 %%%%%
-\revpoint{line 263:}{
+\point{line 263:}{
     time to reach 100 B alleles: simulations should show how long this takes $\to$ can this explain anything or is it irrelevant? (Unclear in Fig 4 since value for $s_p$ is not provided).

### 2015-05-26 patchy-review-responses-second-round.tex
.

@@
         \item the average density is $q(x) \sim e^{-x\sqrt{2s}/\sigma} = e^{-x/\ell}$
-        \item but fluctuates far from the source as [families of size $K$, with local density $k/\ell^d$]
-        \item families of size $1/s$ live for $1/s$ generations and hence contribute $1/s^2$ to the time-averaged density $q(x)$
+        \item but fluctuates far from the source as 
+          [families of size $K$, fluctuations of duration $1/s$]
+        \item local density of a family: $K/\ell^d$
+        \item families of size $K$ live for $K$ generations and hence contribute $K^2$ to the time-averaged density $q(x)$, which is hence dominated by the largest families, $K \sim 1/s$, but families reach this size only with probability $s$
         \item \[ q(x) = \lambda_\text{arr}(x) / s^2 \ell^d, \]

### 2015-05-26 pgen-reviews/revresponse.R

@@ -1,0 +1,45 @@
+ninds <- 1e3
+ngens <- 6e2
+
+# reviewer's assertion: families of size K live for K generations and so contribute K^2 to the total
+# my assertion:
+#  E[ sum_{t \ge 0} Z_t | Z_0=K ] = K sum_{t \ge 0} (1-s)^t = K / s 
+# oh: but time TO HIT K starting from 1 is proportional to K, given that you do
+#      and total contribution in this case is indeed K^2
+
+f <- function (ninit,s) {
+    gens <- matrix( 0, nrow=ninds, ncol=ngens )
+    gens[,1] <- ninit
+    for (k in 2:ncol(gens)) {
+        gens[,k] <- rpois( nrow(gens), gens[,k-1]*(1-s) )
+    }
+
+    # this does total time
+    if (FALSE) {
+        t.zero <- 1+rowSums(gens!=0)
+        hist(t.zero,main=paste("Z0=", ninit, "s=",s),xlim=range(t.zero,ninit))
+        abline(v=ninit, col='red',lwd=2)
+        abline(v=mean(t.zero),lwd=2,col='green')
+    }
+
+    # layout(t(1:2))
+    # plot(colMeans(ttt,na.rm=TRUE))
+    # matplot(t(gens[apply(gens,1,max)>10,1:100]),type='l')
+
+    # and this does total contribution
+    total.contrib <- rowSums(gens)
+    hist(total.contrib,main=paste("Z0=", ninit, "s=",s),xlim=range(total.contrib,ninit^2))
+    abline(v=ninit^2, col='red',lwd=2)
+    abline(v=mean(total.contrib),lwd=2,col='green')
+    abline(v=ninit/s,lwd=2,col='purple',lty=2)
+}
+
+
+layout(matrix(1:6,nrow=2))
+
+f(20,.02)
+f(20,.2)
+f(40,.02)
+f(40,.2)
+f(80,.02)
+f(80,.2)

### 2015-05-27 patchy-selection-paper.tex
fixing up integration over S

@@
 This decomposition is illustrated in figure~\ref{fig:branching_decomp}.
-In terms of these families, the equilibrium frequency above is then
+The opportunity for adaptation depends on the number $B$ alleles present at migration-selection balance:
+in terms of these families, the time-averaged equilibrium number of $B$ alleles
+present in the area of the new patch \emph{without} their selective advantage is,
+writing $q(S) = \int_S q(y) dy$,
+XXX PHRASE BETTER XXX
 \begin{align}
@@
     \begin{split}
-        \rho q(x) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $x$ ) } ,
+        \rho q(S) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation in new patch per family ) } ,
 \end{split}
@@
 and the quantity we wish to compute,
-the effective rate at which \emph{families of} migrant $B$ alleles establish in a new patch at displacement $x$ from the original,
+the effective rate at which \emph{families of} migrant $B$ alleles establish in the new patch
 is
@@
     \begin{split}
-        \migrate(x) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in patch at $x$ ) }
+        \migrate(S) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in new patch ) } .
     \end{split}
@@
 The parallel form suggested by \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
-suggests that we can find $\migrate(x)$ by multiplying the mean density $q(x)$
+suggests that we can find $\migrate(S)$ by multiplying the mean density $q(S)$
 by the ratio of the second terms in the two equations,
@@
 \begin{align} \label{eqn:ratio_gestalt}
-  \migrate(x) 
-    &= \rho q(x) \times 
+  \migrate(S) 
+    &= \rho q(S) \times 
     \frac{
-        \text{ ( probability a family establishes in patch at $x$ ) }
+        \text{ ( probability a family establishes in $S$ ) }
     }{
-        \text{ ( mean occupation density of a family at $x$ ) }
-    } .
-\end{align}
+        \text{ ( mean occupation density of a family in $S$ ) }
+    } ,
+\end{align}
+where again the ``occupation density'' is computed \emph{without} the selective advantage.
 To make this precise we need to understand what we mean by ``outflux of families'';
@@
 \begin{align} \label{eqn:ratio_K}
-  \migrate(x) 
-    &= \rho q(x) \times 
+  \migrate(S) 
+    &= \rho q(S) \times 
     \frac{
@@
 then the numerator would be close to 1, as any family that arrives will likely establish.
+
+On the other hand, XXX WHAT IS $q(S)$ XXX.
+
 Combining these, and absorbing the term $\E[W|W>1]$ from equation \eqref{eqn:mean_K} 
@@
 Denote by $\gamma(x)$ the mean rate of outflux of migrant families from a point $x \in \partial B(r_0)$,
-i.e.\ the time-averaged proportion of individuals near a point $x$ in $\partial B(r_0)$ 
+i.e.\ the time-averaged density of individuals near a point $x$ in $\partial B(r_0)$ 
 that are the founders of new migrant families.
@@
 when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
-However, examining the constituent terms,
-it will turn out that equation \eqref{eqn:gestalt_migrate} 
-is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.
+However,
+it will turn out that the integrand of equation \eqref{eqn:gestalt_migrate} 
+is well-approximated by a constant multiple of the integrand of \eqref{eqn:gestalt_q}.
 
@@
 started at time 0 with a single individual at $x$,
-and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
-The \emph{mean occupation density} of $Z$ at location $y$
+and write $S$ for the new patch, which has area $A$ and is centered at $y$.
+The \emph{mean occupation measure} of $Z$ in the region $S$
+which we denote $u(x,S)$,
+% and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
+% The \emph{mean occupation density} of $Z$ at location $y$
 can be thought of informally \linelabel{rr:density}
-as the expected total number of offspring of a family beginning at $x$ that ever live near $y$.
-It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
-i.e.\ the function $u(x,y)$ that, integrated over a set of locations $S$, 
-gives the expected total number of individuals ever living in $S$:
-\begin{align} \label{eqn:occupation_density_defn}
-    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
-\end{align}
-% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
-Since reproduction and spatial motion are independent,
-$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
-multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
-i.e.\
-\begin{align}
-    u(x,y) = \int_0^\infty e^{-s_m t} p_t(x,y) dt .
-\end{align}
-The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
-is the sum of mean occupation densities of a constant outflux of branching processes
+as the expected total number of offspring of a family beginning at $x$ that ever live in $S$.
+% It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
+% i.e.\ the function $u(x,y)$ that, integrated over a set of locations $S$, 
+% gives the expected total number of individuals ever living in $S$:
+% \begin{align} \label{eqn:occupation_density_defn}
+%     \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
+% \end{align}
+% % Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
+% Since reproduction and spatial motion are independent,
+% $u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
+% multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
+% i.e.\
+% \begin{align}
+%     u(x,y) = \int_0^\infty e^{-s_m t} p_t(x,y) dt .
+% \end{align}
+Let $q(S) = \int_S q(y) dy$ denote the total equilibrium frequency in $S$.
+This is decomposed in expression \eqref{eqn:gestalt_q} 
+as the sum of mean occupation densities of a constant outflux of branching processes
 from $\partial B(r_0)$:
 \begin{align} \label{eqn:occupation_integral}
-    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
-         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
-\end{align}
+    \rho q(S) &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  . 
+\end{align}
+
+Now we will decompose $u(x,S)$ under the assumption that 
+the marginal distribution of the spatial motion of a single lineage is Brownian.
+Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
+and $\tau_\dagger$ an independent Exponential($s_m$) time.
+The mean occupation time of $Z$ spent in a region $S$ is,
+\begin{align}
+    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
+           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
+        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
+        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
+        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
+\end{align}
+where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
+and likewise $\E^x$.
+If we define
+$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
+and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\tau_S}$ conditioned on $\tau_S < \tau_\dagger$,
+by the strong Markov property this is equal to
+\begin{align}
+    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
+           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] \\
+           &= \P^x\{ \tau_S < \tau_\dagger \}  g(A)  ,
+\end{align}
+where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
+% provided $S$ is not pathological.
+and $g(A)$ is defined to be the latter expectation,
+which does not depend on $x$ since $S$ is circular with area $A$.
 
@@
 Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
-We assume this new patch is circular, 
-has relatively small area $A$ and is centered at position $y$.
 % The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
@@
 and successfully establishes in the new patch.
-Write $h_A(x,y)$ for the probability that the Brownian trunk
-hits the new patch centered at $y$ before the family dies out,
-and $f(A)$ for the chance that the family manages to establish in the new patch,
+Write $h(x,S)$ for the probability that the Brownian trunk hits the patch $S$
+before the family dies out,
+and $f(S)$ for the chance that the family manages to establish in the new patch,
 given that it successfully arrives.
-As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
+(This is independent of $x$ in the limit as the distance from $x$ to $S$ grows.)
+As for $q(S)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
 \begin{align}\label{eqn:migrate_integral}
-  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
-\end{align}
-Recall that $\rho$ appears here but not in \eqref{eqn:occupation_integral}
-because both $q(x)$ and $\gamma(x)$ are local \emph{proportions} of $B$ alleles.
-
-Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
-while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
-Since the expected total time spent in the patch does not depend on the distance at which the family started
-if we know that it survived to hit the patch,
-it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
-Following this intuition,
-and using $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-we show below that
+  \migrate(S) &= \int_{\partial B(r_0)} \gamma(x) h(x,S) f(S) dx .
+\end{align}
+
+XXX CHECK FACTORS OF $\rho$ BELOW XXX
+
+If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
+and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
+then 
 \begin{align}
-    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
-\end{align}
-where $g(A)$ is the total expected occupation time of the patch given the family arrives.
-Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
+    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
+           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
+           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
+    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
+\end{align}
+Therefore, we have that
+\begin{align} \label{eqn:boils_downto}
+    h(x,S) \E[K] 
+    \approx \P^x[ \tau_S < \tau_\dagger ] 
+    = u(x,S) / g(A) .
+\end{align}
+
+% Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
+% while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
+% Since the expected total time spent in the patch does not depend on the distance at which the family started
+% if we know that it survived to hit the patch,
+% it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
+% Following this intuition,
+% and using $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
+% we show below that
+% \begin{align} 
+%     h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
+% \end{align}
+% where $g(A)$ is the total expected occupation time of the patch given the family arrives.
+% Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
 Since the integrands in expressions \eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral} only differ by a factor
@@
 we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
-
+XXX NOT RIGHT XXX
 The result will not depend on $A$, 
@@
 
-Consider a single migrant family $Z$ beginning at $x$,
-and write $S$ for the new patch, which has area $A$ and is centered at $y$.
-Let $B_t$ be a Brownian motion with variance $\sigma^2$, 
-and $\tau_\dagger$ an independent Exponential($s_m$) time.
-The mean occupation time of $Z$ spent in a region $S$ is,
-since the marginal distribution of a single lineage is Brownian,
-letting $u(x,S) = \int_S u(x,y) dy$,
+By equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
 \begin{align}
-    u(x,S) &= \E\left[\int_0^\infty Z_t(S) dt \right] \\
-           &= \int_0^\infty \E[Z_t] p_t(x,S) dt \\
-        &= \int_0^\infty e^{-s_m t} p_t(x,S) dt \\
-        &= \int_0^\infty \P^x\{ \tau_\dagger > t \; \& \; B_t \in S \} dt \\
-        &= \E^x\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right]  ,
-\end{align}
-where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
-and likewise $\E^x$.
-If we define
-$\tau_S$ to be the hitting time of $S$ by the Brownian motion $B$,
-and $\mu_S(x)$ to be the hitting distribution of $\partial S$ by $B_{\tau_S}$ conditioned on $\tau_S < \tau_\dagger$,
-by the strong Markov property this is equal to
-\begin{align}
-    u(x,S) % &= \E^x[ \int_0^{\tau_\dagger} \one_S(B_t) dt ] \\
-           &= \P^x\{ \tau_S < \tau_\dagger \}  \E^{\mu_S(x)}\left[ \int_0^{\tau_\dagger} \one_S(B_t) dt \right] ,
-\end{align}
-where now $\E^\mu$ denotes expectations for Brownian motion for which the distribution of $B_0$ is $\mu$, 
-provided $S$ is not pathological.
-If $S$ is circular with area $A$, the latter expectation will not depend on $x$;
-call this $g(A)$.
-
-Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
-(Before, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
-If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
-and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
-then this is equal to
-\begin{align}
-    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
-           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
-           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
-           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
-    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
-\end{align}
-Therefore, we have that
-\begin{align} \label{eqn:boils_downto}
-    h(x,S) \E[K] 
-    \approx \P^x[ \tau_S < \tau_\dagger ] 
-    = u(x,S) / g(A) .
-\end{align}
-
-If we let $q(S) = \int_S q(y) dy$ be the total equilibrium occupation of $S$,
-then by equations~\eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral},
-\begin{align}
-    q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
+    \rho q(S)  &= \int_{\partial B(r_0)} \gamma(x) u(x,S) dx  \\
           &\approx g(A) \E[K] \int_{\partial B(r_0)} \gamma(x) h(x,S) dx \\
-          &= \frac{ g(A) \E[K] }{ \rho f(A) } \migrate(y) . \label{eqn:q_migrate_relation}
+          &= \frac{ g(A) \E[K] }{ f(A) } \migrate(S) . \label{eqn:q_migrate_relation}
 \end{align}
@@
 for helpful comments and discussion of the manuscript.
+This work was supported by grants from the
+National Science Foundation under Grant No. 1262645 to PLR and GC and by the
+National Institute of General Medical Sciences of the National
+Institutes of Health under award numbers NIH RO1GM83098 and
+RO1GM107374 to GC.
 

### 2015-08-24 patchy-selection-paper.tex
fixed formula in analyze-sims

@@
       \textbf{(left)} 
+      % XXX HERE
       Time until adaptation by mutation in an isolated patch for different parameter combinations,

### 2015-08-24 analyze-sims.R

@@ -4,6 +4,16 @@ source("sim-patchy-selection-fns.R")
 
 maxtime <- 25000  # number of gens these were run for
 
+###
+# mutation:
+#
+# All simulations also used a linear grid of 501 demes with a patch of 99 demes in the center, the migration
+# model described in Simulation methods, μ = 10 −5 , and s p = .0023 (calculated as the growth rate as
+# described in the text). T adapt is the mean time until 100 B alleles were present in the patch, and p adapted is
+# the proportion of the simulations that adapted by the 25,000 generations. The simulation began with no B
+# alleles.
+
+## made with collate-sims.R
 mutsims <- read.table("mutation-sims-info.tsv", sep='\t', header=TRUE )
 
 names(mutsims)[names(mutsims)=="N"] <- "N.name"
@@ -79,7 +89,14 @@ if (interactive()) {
 
 ################
 # migration
+#
+# "All simulations also used a linear grid of 501 demes with two patches of 99 demes each, separated by R demes
+# in the center; the migration model described in Simulation methods, μ = 10 −5 , and s p = .0023 (calculated
+# as the growth rate as described in the text). T adapt is the mean time until 100 B alleles were present in the
+# patch, and p adapted is the proportion of the simulations that adapted by the 25,000 generations. At the start
+# of the simulation, one patch was initialized with a frequency of 0.8 B alleles, which were absent elsewhere.
 
+## made with collate-sims.R
 migsims <- read.table("migration-sims-info.tsv",sep='\t',header=TRUE)
 
 names(migsims)[names(migsims)=="N"] <- "N.name"
@@ -98,7 +115,8 @@ migsims$paramstring <- with(migsims, factor( paste(sm.name,R.name,N.name,dims,se
 
 migsims$adjust <- 0  # with( migsims, log(N)/gb + size1/2 / ( sigma * sqrt(gb) ) )
 # migsims$migtime <- with( migsims, 0.5/( 4 * N * gb * abs(sm) * exp(- R * sqrt(2*abs(sm))/sigma ) ) + adjust )
[…]

### 2015-08-24 patchy-selection-paper.tex
.

@@
 This decomposition is illustrated in figure~\ref{fig:branching_decomp}.
-The opportunity for adaptation depends on the number $B$ alleles present at migration-selection balance:
-in terms of these families, the time-averaged equilibrium number of $B$ alleles
-present in the area of the new patch \emph{without} their selective advantage is,
-writing $q(S) = \int_S q(y) dy$,
-XXX PHRASE BETTER XXX
+The opportunity for adaptation depends on how often these families of $B$ alleles encounter the new patch under
+migration-selection equilibrium.
+Suppose that the area occupied by the new patch is $S$.
+We can learn about the rate of arrival of families at $S$ from
+the time-averaged number of $B$ alleles expected in $S$ were it not a patch
+(i.e., if $B$ was still maladaptive in $S$),
+which from equation \eqref{eqn:eqfreq}
+is 
 \begin{align}
@@
 \end{align}
-and the quantity we wish to compute,
-the effective rate at which \emph{families of} migrant $B$ alleles establish in the new patch
+where $q(S) = \int_S q(y) dy$.
+The quantity we wish to compute,
+the effective rate at which \emph{families of} migrant $B$ alleles establish in the new patch,
 is

### 2015-08-27 patchy-selection-paper.tex
.

@@
 \DeclareMathOperator{\sn}{sn}
+\DeclareMathOperator{\erf}{erf}
 \renewcommand{\P}{\mathbb{P}}
@@
 then the numerator would be close to 1, as any family that arrives will likely establish.
+
+The remaining term is $q(S)$, the integral of equation \eqref{eqn:eqfreq} over the patch.
+Suppose the new patch has width $w$, and is at distance $x$ from the original patch.
+If the new patch is small ($w \ll \sigma/\sqrt{2s_m}$), 
+then $q(S)$ is approximately $q(x)$ multiplied by the area of the new patch.
+If the new patch is a long rectangle with width $u$ with long axis pointed towards the old patch,
+$q(S)$ is approximately $u C \sqrt{\sigma/\sqrt{2s_m}} \erf(\sqrt{x\sqrt{2s_m}/\sigma})$,
+where $\erf()$ is the error function.
 

### 2015-08-29 patchy-review-responses-second-round.tex
fixing up integration over S
edits

--- 2015-05-26 patchy-review-responses-second-round.tex
@@
 
+\reply
+
+The main change,
+which addresses most of the concerns,
+is that we uncovered an error in the derivation of equation \eqref{eqn:migrate},
+that supplied the missing factor of area noted by the reviewers.
+The error occurred in the transition between section \nameref{ss:heuristics} 
+and section \nameref{ss:hitting_occupation}
+because in the former we had tried to keep things simple
+by not including shape of the patch in
+gestalt equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate},
+but then (more correctly) treated the shape in the latter;
+effectively we were treating the patch as very small in one place
+and large in another.
+We've rewritten the argument in \nameref{ss:heuristics} 
+to deal explicitly with the integral over the whole patch,
+which also necessitated adding a new (short) appendix \nameref{apx:qS}.
+Section \nameref{ss:hitting_occupation} required very little change.
+
 %%%%%%%%%%%%%%
@@
     The issue of dimensions and geometry of the target patch is the main difficulty I have with the paper.
-I just don't see how your equation 11 can be correct.
+I just don't see how your equation \eqref{eqn:migrate} can be correct.
 It has dimensions of 1/time/area when it should be 1/time.
@@
 \reply
+Now that we've found the problem in the derivation,
+we heartily agree;
+the missing factor was, indeed, a multiplicative factor of area.
+(See above for more on this.)
 
@@
 Regarding the simulations presented in Fig 4, I don't think they strongly support the functional form of $\migrate$.
-Points with very similar observed migration time have predicted migration times that differ by a factor of a 100 (the more extreme are outside the validity of the approximations, but still the I don't think this figure is strong evidence that the math is correct.)
+Points with very similar observed migration time have predicted migration times that differ by a factor of a 100 (the more extreme are outside the validity of the approximations, but still I don't think this figure is strong evidence that the math is correct.)
 }
@@
 \reply
+We think the main issue here was inclusion of points outside the validity of the approximations,
+since of the parameters falling in the range of validity discussed in \nameref{ss:assumptions},
+the largest discrepancy was a factor of 7, not 100.
+We've re-made this figure only using parameter combinations that fall within the range of validity
+($R > \sigma/\sqrt{s_m}$, patch width larger than $\sigma/\sqrt{s_m} \atan(\sqrt{s_m/s_b})$, and $2 s_m \sigma N > 1$);
+%         ( R*sqrt(abs(sm))/sigma > 1 ) & 
+%         ( size2 > (sigma/sqrt(abs(sm)))*atan(sqrt(abs(sm/gb))) ) &
+%         ( abs(sm) * 2 * sigma * N > 1 )
+the figures that include all simulation results are still included in the supplement
+(and hopefully will inspire theory that is applicable in the other regions).
 
@@
 \reply
-true, depends if you think ``saddle point approximation'' or ``laplace transform'' is a simpler idea?  this way avoids writing down more integrals; I hate to add more equations...
+This is a nice point;
+perhaps which approach should be included
+comes down to personal taste
+as to whether ``saddle point approximation'' or ``Laplace transform'' is a simpler idea?  
+We've left this as-is.
 

### 2015-08-29 patchy-selection-paper.tex
edits

@@
 \DeclareMathOperator{\erf}{erf}
+\DeclareMathOperator{\atan}{atan}
 \renewcommand{\P}{\mathbb{P}}
@@
 Therefore, the expected contribution of the family to the occupation density is $K/2s_m$.
-(Readers who feel uncomfortable about the geometry of the patch,
-or the distinction between occupation density and number of individuals
-are encouraged to take a detour through section \secref{ss:hitting_occupation}, below.)
+(For more careful consideration of the geometry of the patch
+and the distinction between occupation density and number of individuals,
+see section \secref{ss:hitting_occupation}, below.)
 
@@
       for each unique parameter combination.
+      Only parameter combinations to which the theory here is applicable were used
+      (precisely, if
+      $R > \sigma/\sqrt{s_m}$, 
+      patch width larger than $\sigma/\sqrt{s_m} \atan(\sqrt{s_m/s_p})$, 
+      and $s_m > 2 \sigma \rho$);
+      the remaining combinations are shown in figures
+      \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times}.
   }   \label{fig:sim_times}
@@
 and write $S$ for the new patch, which has area $A$ and is centered at $y$.
-The \emph{mean occupation measure} of $Z$ in the region $S$
+The \emph{mean occupation measure} of $Z$ in the region $S$,
 which we denote $u(x,S)$,
@@
 and $g(A)$ is defined to be the latter expectation,
-which does not depend on $x$ since $S$ is circular with area $A$.
+which does not depend on $x$ if $S$ is circular (with area $A$).
 
@@
 
-XXX CHECK FACTORS OF $\rho$ BELOW XXX
-
 If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
@@
 we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
-XXX NOT RIGHT XXX
-The result will not depend on $A$, 
+This factor will not depend on $A$, 
 because although $f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
@@
 
-%%%%%%% %%%%%%%%%%%%
-\subsection[Outflux of Families]{The outflux of families}
-\label{apx:outflux}
-
-It is of interest to have an expression for the ``outflux of families'',
-$\gamma(r_0)$, 
-defined to be the mean proportion of $B$ alleles at distance $r_0$
-whose lineage traces back to the original patch without venturing further away than $r_0$.
-By expression \eqref{eqn:occupation_integral},
-\[
-    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
-\]
-for any $y$.  
-Take $\hat x = (1,0)$ and $y=r \hat x$.  Then
-\begin{align}
-  \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\| \sqrt{2s_m}/\sigma } }{ \|y\|^{-(d-1)/2} e^{-\|y\| \sqrt{2s_m}/\sigma } } \\
-  &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r \sqrt{2s_m}/\sigma  (\|\hat x - x/r\| - 1)} .
-\end{align}
-Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
-so that the expression in the exponent is to, first order in $1/r$,
-\begin{align}
-    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
-            &= - r_0 \cos \theta + O(1/r) .
-\end{align}
-This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
-\begin{align}
-  \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} .
-\end{align}
-In one dimension we are done.
-In two dimensions,
-by bounded convergence and \citet{gradshteyn2007table} (3.339),
-\begin{align}
-  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-  &= C' \int_0^{2 \pi} e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
-  &= 2 \pi C' I_0(-r_0\sqrt{2 s_m}/\sigma),
-\end{align}
-where $I_0$ is a modified Bessel function of order zero of the first kind.
-
 %%%%%%%%%%%
@@
 This work was supported by grants from the
-National Science Foundation under Grant No. 1262645 to PLR and GC and by the
+National Science Foundation under Grant No.\ 1262645 to PLR and GC and by the
 National Institute of General Medical Sciences of the National
@@
 for the length of shared haplotype between patches.
+
+%%%%%%% %%%%%%%%%%%%
+\subsection[Outflux of Families]{The outflux of families}
+\label{apx:outflux}
+
+It is of interest to have an expression for the ``outflux of families'',
+$\gamma(r_0)$, 
+defined to be the mean proportion of $B$ alleles at distance $r_0$
+whose lineage traces back to the original patch without venturing further away than $r_0$.
+By expression \eqref{eqn:occupation_integral},
+\[
+    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
+\]
+for any $y$.  
+Take $\hat x = (1,0)$ and $y=r \hat x$.  Then
+\begin{align}
+  \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\| \sqrt{2s_m}/\sigma } }{ \|y\|^{-(d-1)/2} e^{-\|y\| \sqrt{2s_m}/\sigma } } \\
+  &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r \sqrt{2s_m}/\sigma  (\|\hat x - x/r\| - 1)} .
+\end{align}
+Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
+so that the expression in the exponent is to, first order in $1/r$,
+\begin{align}
+    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
+            &= - r_0 \cos \theta + O(1/r) .
+\end{align}
+This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
+\begin{align}
+  \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} .
+\end{align}
+In one dimension we are done.
+In two dimensions,
+by bounded convergence and \citet{gradshteyn2007table} (3.339),
+\begin{align}
+  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
+  &= C' \int_0^{2 \pi} e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
+  &= 2 \pi C' I_0(-r_0\sqrt{2 s_m}/\sigma),
+\end{align}
+where $I_0$ is a modified Bessel function of order zero of the first kind.

### 2015-08-30 patchy-selection-paper.tex
.

--- 2015-08-24 patchy-selection-paper.tex
@@
 \newcommand{\given}{\,\vert\,}
+\newcommand{\st}{\,\colon\,}
 
@@
 
-On the other hand, XXX WHAT IS $q(S)$ XXX.
+The remaining term is $q(S)$, the integral of equation \eqref{eqn:eqfreq} over the patch.
+Suppose the new patch is at distance $x$ from the original patch.
+If the new patch has width smaller than $\sigma/\sqrt{2s_m}$, 
+then $q(S)$ is approximately $q(x)$ multiplied by the area of the new patch.
+If $S$ is larger, then we should use the area of the closest strip
+of width $\sigma/\sqrt{2s_m}$ to the original patch;
+see section \ref{apx:qS} for more details.
+In either case, denote this area $A$.
 
@@
 \begin{align} \label{eqn:migrate}
-  \migrate(x) &\approx C \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m} 
+  \migrate(x) &\approx C A \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m}  ,
 %     \begin{cases}
@@
 \end{align}
+where $A$ is the area of the new patch if that is no larger than $\sigma/\sqrt{2s_m}$.
 When we use this below, we set $p_e = 2 s_p / \xi^2$.
@@
 In either case, $g(A) \approx 1/(2s_m)$, which is the approximation we use in the main text.
+
+%%%%%%% %%%%%%%%%%%%
+\subsection[The Integral $q(S)$]{The integral $q(S)$}
+\label{apx:qS}
+
+In the developement above
+we need to approximate the integral of equation \eqref{eqn:eqfreq}
+across the area occupied by the new patch,
+$q(S) = \int_S q(z) dz$.
+The precise answer depends on the shape and orientation of the patches;
+but we aim for a usable approximation,
+primarily in terms of the shortest distance from the old patch to the new patch, denoted $x$,
+and assuming $x$ is large enough we can take equation \eqref{eqn:eqfreq} as an equality.
+
+Since $q(x)$ decreases as $|x|$ increases,
+\begin{align}
+    q(S) \le |S| q(x) % C \left( \frac{x\sqrt{2s_m}}{\sigma} \right)^{(d-1)/2} e^{-x\sqrt{2s_m}/\sigma} ,
+\end{align}
+where $|S|$ is the area of $S$.
+This will be a good approximation if $S$ is small.
+
+In one dimension, if $S$ has length $\ell$,
+\begin{align}
+    q(S) &= \int_x^{x+\ell} C e^{-u\sqrt{2s_m}/\sigma} du \\
+         &= C e^{-x\sqrt{2s_m}/\sigma} \frac{\sigma}{\sqrt{2s_m}} \left( 1 - e^{-\ell\sqrt{2s_m}/\sigma} \right) \\
+         &\le C e^{-x\sqrt{2s_m}/\sigma} \frac{\sigma}{\sqrt{2s_m}} .
+\end{align}
+
+In two dimensions, 
+suppose that $T$ is a rectangle enclosing $S$ with one axis aligned towards the original patch and transverse width $w$.
+Then
+\begin{align}
+    q(S) &\le \int_T q(x) dx \\
+        &\le w \int_x^\infty q(r) r dr \\
+        &= w C \int_x^\infty \left( \frac{r\sqrt{2s_m}}{\sigma} \right)^{-1/2} e^{-r\sqrt{2s_m}/\sigma} dr \\
+        &\le w \frac{\sigma}{\sqrt{2s_m}} C \sqrt{\pi} \left( \frac{x\sqrt{2s_m}}{\sigma} \right)^{-1/2} e^{-x\sqrt{2s_m}/\sigma} dr .
+\end{align}
+If we absorb $\sqrt{\pi}$ into the constant $C$,
+this is of the form $(\text{width}) \times (\sigma/\sqrt{2s_m}) \times q(x)$,
+i.e., roughly the area, after replacing the length by $\sigma/\sqrt{2s_m}$.
+
+In both cases,
+the upper bound is $q(S) \le A q(x)$,
+where $A$ is the area of the parts of $S$ that are no more than $x+\sigma/\sqrt{2s_m}$ away from the original patch.
+Lower bounds could be obtained,
+but we stick with this rough approximation.
 

### 2015-08-30 patchy-review-responses-second-round.tex
.

@@
 
+We also turned up a minor bug in the discussion of conditioned branching processes:
+the exact picture is not quite so simple,
+but the approximate picture is still the same --
+see around \llname{ll:branching_process_fix}.
+
 %%%%%%%%%%%%%%
@@
 \reply
-Thanks for catching that mistake.  We removed the extra factor of 2 (section~\nameref{apx:outflux}).
+Thanks for catching that mistake.  
+We removed the extra factor of 2 (section~\secref{apx:outflux}).
 
@@
 as to whether ``saddle point approximation'' or ``Laplace transform'' is a simpler idea?  
-We've left this as-is.
+We decided to stick with the Laplace transform.
 
@@
             hence one should get
-            \[ \migrate \sim \rho q(x) s^2 \ell^d \sqrt{r/\ell} (1-e^-{p_0/s}) \]
+            \[ \migrate \sim \rho q(x) s^2 \ell^d \sqrt{r/\ell} (1-e^{-p_e/s}) \]
         \item result does depend on target geometry
-        \item formulation in terms of fluxes might be easier
     \end{itemize}
+}
+
+\reply
+As discussed above,
+we've debugged the issue about dimensions,
+so that the argument now correctly does integrate over the new patch.
+The result does depend on target geometry, just as the reviewer points out.
+The other part of this sketch
+makes further simplifications to \eqref{eqn:migrate}
+based on
+
+\point{fluxes}{
 }

### 2015-08-30 patchy-selection-paper.tex
.

@@
 that arose from the main branch in generation $s$,
-and each $W_s$ is independent with distribution $\P\{ W_s = k\} = \P\{ W = k \} (1-k_e(t-s)^k)/(1-k_e(t-s+1))$. \linelabel{ll:branching_process_fix}
+and each $W_s$ is independent with distribution $\P\{ W_s = k\} = \P\{ W = k \} (1-k_e(t-s)^k)/(1-k_e(t-s+1))$, 
+where $W$ is the unconditioned offspring number distribution.  \linelabel{ll:branching_process_fix}
 Since each subfamily $Z^{(s,k)}$ is a branching process with mean offspring number less than one,
@@
     \E[K] 
-    = 1 + \E[W-1 \given W \ge 1] /(1-e^{-s_m}) \approx \E[W-1 \given W \ge 1] / s_m .
+    % = 1 + \E[W-1 \given W \ge 1] /(1-e^{-s_m}) 
+    = 1 + \sum_{s=0}^\infty \E[W_s-1] e^{-s_m t}
+    \approx \E[W-1 \given W \ge 1] / s_m .
 \end{align}
@@
 
-Combining these, and absorbing the term $\E[W|W>1]$ from equation \eqref{eqn:mean_K} 
+Combining these, and absorbing the term $\E[W-1|W>1]$ from equation \eqref{eqn:mean_K} 
 into the constant $C$,
@@
 but we stick with this rough approximation.
+
+%%%%%%% %%%%%%%%%%%%
+\subsection[Outflux of Families]{The outflux of families}
+\label{apx:outflux}
+
+It is of interest to have an expression for the ``outflux of families'',
+$\gamma(r_0)$, 
+defined to be the mean density of $B$ alleles at distance $r_0$
+whose lineage traces back to the original patch without venturing further away than $r_0$.
+By expression \eqref{eqn:occupation_integral},
+taking $S$ to be a small area located at $y$,
+\[
+    \gamma(r_0) = \frac{\rho q(y)}{\int_{B(r_0)} u(x,y) dx}
+\]
+for any $y$.  
+Take $\hat x = (1,0)$ and $y=r \hat x$.  Then
+\begin{align}
+  \frac{ q(y) }{ u(x,y) } &\simeq C' \frac{ \|y\|^{-(d-1)/2} e^{-\|y\| \sqrt{2s_m}/\sigma } }{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\| \sqrt{2s_m}/\sigma } } \\
+  &= C' \| \hat x - x/r \|^{(d-1)/2} e^{ r \sqrt{2s_m}/\sigma  (\|\hat x - x/r\| - 1)} .
+\end{align}
+Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
+so that the expression in the exponent is to, first order in $1/r$,
+\begin{align}
+    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
+            &= - r_0 \cos \theta + O(1/r) .
+\end{align}
+This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
+\begin{align}
+\frac{q(y)}{u(x,y)} \longrightarrow C' e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} \qquad \text{ as } r \to \infty .
+\end{align}
+In one dimension we are done (since $\cos \theta=1$).
+In two dimensions,
+by bounded convergence and \citet{gradshteyn2007table} (3.339),
+\begin{align}
+  \int_{B(r_0)} \frac{q(y)}{u(x,y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
+  &= C' \int_0^{2 \pi} e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
+  &= 2 \pi C' I_0(-r_0\sqrt{2 s_m}/\sigma),
+\end{align}
+where $I_0$ is a modified Bessel function of order zero of the first kind.
 
@@
 for the length of shared haplotype between patches.
-
-%%%%%%% %%%%%%%%%%%%
-\subsection[Outflux of Families]{The outflux of families}
-\label{apx:outflux}
-
-It is of interest to have an expression for the ``outflux of families'',
-$\gamma(r_0)$, 
-defined to be the mean proportion of $B$ alleles at distance $r_0$
-whose lineage traces back to the original patch without venturing further away than $r_0$.
-By expression \eqref{eqn:occupation_integral},
-\[
-    \gamma(r_0) = \frac{q(y)}{\int_{B(r_0)} u(x,y) dx}
-\]
-for any $y$.  
-Take $\hat x = (1,0)$ and $y=r \hat x$.  Then
-\begin{align}
-  \frac{ u(x,y) }{ q(y) } &\simeq C' \frac{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\| \sqrt{2s_m}/\sigma } }{ \|y\|^{-(d-1)/2} e^{-\|y\| \sqrt{2s_m}/\sigma } } \\
-  &= C' \| \hat x - x/r \|^{-(d-1)/2} e^{-r \sqrt{2s_m}/\sigma  (\|\hat x - x/r\| - 1)} .
-\end{align}
-Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
-so that the expression in the exponent is to, first order in $1/r$,
-\begin{align}
-    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
-            &= - r_0 \cos \theta + O(1/r) .
-\end{align}
-This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
-\begin{align}
-  \frac{u(x,y)}{q(y)} \longrightarrow_{r \to \infty} C' e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} .
-\end{align}
-In one dimension we are done.
-In two dimensions,
-by bounded convergence and \citet{gradshteyn2007table} (3.339),
-\begin{align}
-  \int_{B(r_0)} \frac{u(x,y)}{q(y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-  &= C' \int_0^{2 \pi} e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
-  &= 2 \pi C' I_0(-r_0\sqrt{2 s_m}/\sigma),
-\end{align}
-where $I_0$ is a modified Bessel function of order zero of the first kind.

### 2015-08-30 patchy-review-responses-second-round.tex
almost through reviews

@@
         \item {} [with $\ell = \sigma/\sqrt{2s}$, $r$ the radius of the patch, $R$ the distance between patches, and $\ell \ll r \ll R$]
-        \item the average density is $q(x) \sim e^{-x\sqrt{2s}/\sigma} = e^{-x/\ell}$
-        \item but fluctuates far from the source as [families of size $K$, with local density $k/\ell^d$]
+        \item the average density is $q(x) \sim e^{-x/\ell}$
+        \item but fluctuates far from the source as [families of size $K$, with local density $K/\ell^d$]
         \item families of size $1/s$ live for $1/s$ generations and hence contribute $1/s^2$ to the time-averaged density $q(x)$
@@
 so that the argument now correctly does integrate over the new patch.
-The result does depend on target geometry, just as the reviewer points out.
-The other part of this sketch
-makes further simplifications to \eqref{eqn:migrate}
-based on
+The result does depend on target geometry, just as the reviewer points out (thanks!);
+this is now discussed in section \secref{apx:qS}.
+The other part of this argument, as we understand it,
+modifies terms in \eqref{eqn:migrate}
+by specifying a form for $A$, the relevant area of the new patch,
+and postulating a specific form for $K$.
+We've opted not to make these changes,
+since these require additional assumptions,
+and it feels easier on the reader to not require them to parse such assumptions.
 
-\point{fluxes}{
-}
+\point{Formulation in terms of fluxes}{(suggested by reviewer 2) might be easier.}
 
 \reply
+Perhaps; we've gone the even easier route of working with $q(S)$ and $\migrate(S)$ rather than densities.
 
-%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%  Reviewer 2
 \reviewersection
-
-\begin{quote}
-I think that the manuscript has been improved, but I don't think that all the reviewers' points have been adequately addressed.
-As I wrote in my previous review, I really like the authors' approach of giving simple, intuitive explanations.
-But writing this kind of paper actually requires a much better understanding of the model and dynamics than one needs to write a paper with the standard formal derivations, and I'm not sure that the authors are there yet.
-\end{quote}
-
-perhaps the reviewer is mistaking a different set of terminology and basic concepts for a lack of understanding?
-in this process we have apparently too quickly judging physics-style arguments to be incorrect, some of the time.
 
@@
 \reply
+As discussed above, we fixed this.
+And, apologies for not catching on in the first round of reviews.
 
@@
 \reply
+Good point; this works out nicely: \revref.
 
@@
 \reply
+In general, we could write down differential equations
+that tell us what we want to know.
+There are at least two approaches: 
+ignore the old patch, and simply ask for the probability of establishment;
+or include the old patch, and ask about the transform of the occupation time in the new patch,
+$\E[\exp(-\int_0^t \langle Z_s, p_e \rangle ds)]$.
+For instance,
+if $Z$ is a continuous-time Markovian branching process
+whose spatial motion has generator $A$,
+that when at location $x$, branches at rate $r(x)$,
+producing a number of 
+offspring whose distribution has generating function $\Psi_x(u)=\E^x[u^W]$.
+Then if $f(x)$ denotes the probability of nonextinction
+of a process beginning with a single individual at $x$,
+\begin{align*}
+    Af(x) + r(x) (\Psi_x(f(x))-f(x)) = 0 .
+\end{align*}
+This is the approach taken by \citep{barton1987establishment};
+we chose not to go this route because thinking explicitly about the decomposition into families
+seems much more interesting and useful.
+There is a large and diverse literature on branching Brownian motion,
+but the majority of it is devoted to critical or supercritical Brownian motion.
+We haven't been able to dig up anything that gives the sort of asymptotics that we need;
+but it may well be out there; pointers are welcome.
 
@@
 \reply
+It's true, that in the case that $p_e < s_m$ we get the same result
+as what is obtained ignoring correlations in arrivals
+(e.g., by the Fick's law argument of the previous review).
+To our taste, the "complications" make the paper more digestible,
+since it is more clear what's going on,
+especially to those who are less familiar with branching processes.
+
+That said, we do go beyond these approximations, since we include the case $p_e > s_m$,
+where family structure matters --
+so, we're not clear how to address this.
 
@@
 \reply
+Whoops -- we had erroneously left in that statement that the number of offspring each step was Poisson;
+in fact, the number of offspring produced in a generation is either 0 or 1.
+The exact function is given in the next paragraph \revref,
+and as detailed there, we don't naively assume
+that whatever we labeled \texttt{sm} in the code
+is actually what we mean by $s_m$ in the theory.
+We've also added a note about what $\sigma$ is \llname{rr:sim_details}.
 
@@
 \reply
+Good point; we've been more precise about these conditions. \revref
 
@@
 \reply
+That's a trickier question than it at first seems,
+but we've added a rough idea \revref.
 
@@
 \point{}{
-In their response to my comments, the authors said that they did not see why the assumption $R \gg w$ was neceessary.
+In their response to my comments, the authors said that they did not see why the assumption $R \gg w$ was necessary.
 I think that it is needed in order to talk sensibly about a single "distance to a patch" $R$: if $w \gtrsim R$, then generally some parts of the patch will be much closer than others, right?
@@
 \reply
+Ah, right.
+By "distance to a patch" we mean the length of the shortest path beginning in one and ending in the other.
+We've clarified this \revref.
+Given this, and the caveats around how to calculate $A$ in \eqref{eqn:migrate},
+we still don't think the assumption is necessary?
 
@@
 \reply
+We would agree, but it's not certain which of the various terms 
+in \eqref{eqn:haplotype_length} and the subsequent equation for $\var[L]$
+are the leading ones,
+depending on how $s_m$ compares to $R/\sigma$.
 
@@
 \reply
+We intended for $s_m$ to appear twice, since the selection coefficient is -0.02 on both sides.  Good point about $s_b$; fixed.
 
@@
 \reply
+Good catch about the colors and the squares; we've fixed that.
+Also, as noted above, we've dropped the simulations that don't fall within the range of validity;
+hopefully this unclutters things sufficiently?
 
@@
 \point{Figure 4:}{
-    Mention all simulation parameters in the legend (eg: what is $p_d$?)
+    Mention all simulation parameters in the legend.
 }
@@
 \reply
+This is done. \revref
 
@@
 \reply
+OK as revised? \revref
 
@@
 \reply
+Yes!  Thanks. \revref
 
@@
 \reply
+Also done.
 

### 2015-08-30 patchy-selection-paper.tex
almost through reviews

@@
 
-Now suppose that there are two patches, each of area $A$ and separated by distance $R$. 
+Now suppose that there are two patches, each of area $A$ and separated by distance $R$
+(i.e., the shortest distance between the two is $R$). \revpoint{2}{9}
 If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
@@
 To be clear about the assumptions (especially regarding scaling) implicit here,
-below provide a derivation of \eqref{eqn:eqfreq} in section \secref{apx:eqfreq}, 
+below we provide a derivation of \eqref{eqn:eqfreq} in section \secref{apx:eqfreq},  \revpoint{3}{5}
 as well as justification for the asymptotics below in section \secref{apx:asymptotics}.
@@
 that arose from the main branch in generation $s$,
-and each $W_s$ is independent with distribution $\P\{ W_s = k\} = \P\{ W = k \} (1-k_e(t-s)^k)/(1-k_e(t-s+1))$, 
+and each $W_s$ is independent with distribution 
+%% equation (2.3) in Geiger:
+%% $\P\{ W_s = k\} = \sum_{j=k}^\infty \P\{ W = j \} k_e(t-s-1)^{j-k} (1-k_e(t-s-1))/(1-k_e(t-s))$, 
+%% and equation (2.5) in Geiger:
+that tends to $\P\{W_s=k\} \simeq \P\{W\ge k\}/\E[W]$ as $s \to \infty$,
 where $W$ is the unconditioned offspring number distribution.  \linelabel{ll:branching_process_fix}
+In particular, $\E[W_s-1] \sim \E[W(W-1)]/\E[W]$;
+if $W$ is Poisson then this is equal to $e^{-s_m}$.
 Since each subfamily $Z^{(s,k)}$ is a branching process with mean offspring number less than one,
@@
     \E[K] 
-    % = 1 + \E[W-1 \given W \ge 1] /(1-e^{-s_m}) 
     = 1 + \sum_{s=0}^\infty \E[W_s-1] e^{-s_m t}
-    \approx \E[W-1 \given W \ge 1] / s_m .
-\end{align}
+    \approx \E[W(W-1)] / ( \E[W] s_m  )
+\end{align}
+
+This decomposition also tells us
+that the $K$ individuals in a family
+have most recent common ancestor on order of $1/s_m$ generations ago,
+and so are spread out over a region of order $\sigma/\sqrt{s_m}$.  \revpoint{2}{2}
 
@@
 
-Combining these, and absorbing the term $\E[W-1|W>1]$ from equation \eqref{eqn:mean_K} 
+Combining these, and absorbing the term $\E[W(W-1)]/\E[W]$ from equation \eqref{eqn:mean_K} 
 into the constant $C$,
@@
 
-To be more precise, the mutation rate results will hold best if $s_m \approx s_p$,
+To be more precise, the mutation rate results most obviously apply if $s_m \approx s_p$,
 if the patch diameter is large relative to $\sigma/\sqrt{\min(s_m,s_p)}$,
 and the patch is not too far from circular.
-If $s_m \ll s_p$ then \eqref{eqn:mutrate} will underestimate the rate,
-and will overestimate if $s_m \gg s_p$.
+If $s_m \ll s_p$ then a strip of width $\sigma/\sqrt{s_m}$
+should be added to the outside of the patch in computing $A$ for equation \eqref{eqn:mutrate},
+and if $s_m \gg s_p$, a strip of width $\sigma/s_p$ should be subtracted from the patch. \revpoint{2}{6}
 As for the migration rate,
@@
 The theoretical value $1/\mutrate$ underestimates the mean time to mutation
-by a factor of around 2 that depends increases with $s_m$.
+by a factor of around 2 that increases slightly with $s_m$.
 This is to be expected for two reasons:
@@
       \textbf{(left)} 
-      % XXX HERE
-      Time until adaptation by mutation in an isolated patch for different parameter combinations,
+      Time until adaptation by mutation in an isolated patch for different combinations
+      of selection coefficient on patches ($s_m$), 
+      population density ($\rho$),
+      mutation rate ($\mu$),
+      and distance between patches ($R$), \revpoint{3}{3}
       as observed in simulations (vertical axis),
       and as predicted from equation~\eqref{eqn:mutrate} (horizontal axis).
+      In all simulations, $s_p=0.01$.
       \textbf{(right)} 
@@
       the remaining combinations are shown in figures
-      \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times}.
+      S\ref{sfig:sim_migration_times} and S\ref{sfig:sim_mutation_times}.
   }   \label{fig:sim_times}
@@
 presumably to avoid visual predators.
+These regions of darker substrate are typically tens of kilometers wide and hundreds of kilometers apart. \revpoint{5}{7}
 \citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
@@
 We simulated forward-time dynamics of the number of alleles of each type in a rectangular grid (either one- or two-dimensional) of demes with fixed size $N$.
-Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
+Each generation, each individual independently chose to reproduce or not with a probability $r$ depending on her type and location in the grid;
 locally beneficial alleles were more likely to reproduce.
-Each then produced a random, Poisson distributed number of offspring that each, independently, 
-either remained in the same location with probability $1-m$ or
+Each extant individual then either remained in the same location with probability $1-m$ or
 migrated a random number of steps with probability $m$
@@
 the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
+This gave us values of $\sigma=0.95$ deme spacings in 1D
+and $\sigma=0.74$ in 2D.
 \linelabel{rr:sim_details}
@@
 
-The probability of reproduction in each generation in simulations for type $b$ alleles was 30\%; 
+The base probability of reproduction in each generation in simulations for type $b$ alleles was $r=0.3$;
 this was then multiplied by $1+s$ to get the probability of reproduction for type $B$,
@@
 This determines the values of $s_m$ and $s_p$ reported in the figures, 
-and do not depend on the basic rate of reproduction.
+and do not depend on the basic rate of reproduction. \revpoint{2}{5}
 However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
@@
 \section*{Acknowledgements}
-We would like to heartily thank Joachim Hermisson for a number of insightful suggestions,
-and catching the algebra error present in the first version of this paper on the bioRxiv,
-as well as the two anonymous reviewers.
+We would like to heartily thank Joachim Hermisson, Richard Neher, and an anonymous reviewer
+for a number of insightful suggestions,
+and catching several errors present in earlier drafts. \revpoint{3}{4}
 We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 

### 2015-08-30 analyze-sims.R

@@ -190,16 +190,16 @@ layout(t(1:2))
 par(mar=c(4,4,1,1)+.1)
 # MUTATION
     # remove simulation configurations where a substantial portion haven't yet adapted
-    usethese <- with( mutsims, adapted & ( muttime < maxtime/10 ) ) 
+    usethese <- with( mutsims, adapted & ( muttime < maxtime/10 ) & ( abs(sm) > 1e-4 ) )   # remove sm=1e-4 to reduce number of plotting types
     with( droplevels(subset(mutsims,usethese)), {
             sm.vals <- (levels(droplevels(sm.name))); sm.col.pal <- sequential_hcl(length(sm.vals))
             sm.pch <- match( levels(sm.name)[tapply(sm.name,paramstring,"[",1)], sm.vals )
             sm.cols <- sm.col.pal[ sm.pch ]
             sm.labs <- parse( text=gsub("sm-","s[m]==",levels(sm.name)) )
-            N.vals <- (levels(droplevels(N.name))); N.col.pal <- terrain_hcl(length(N.vals))
+            N.vals <- (levels(droplevels(N.name))); N.col.pal <- terrain_hcl(length(N.vals),l=c(40,60))
             N.cols <- N.col.pal[ match( levels(N.name)[tapply(N.name,paramstring,"[",1)], N.vals ) ]
             N.labs <- parse( text=gsub("N-","rho ==",levels(N.name)) )
-            mu.vals <- (levels(droplevels(mu.name))); mu.col.pal <- sequential_hcl(length(mu.vals))
+            mu.vals <- (levels(droplevels(mu.name))); mu.col.pal <- sequential_hcl(length(mu.vals),l=c(20,60))
             mu.pch <- match( levels(mu.name)[tapply(mu.name,paramstring,"[",1)], mu.vals )
             mu.cols <- mu.col.pal[ mu.pch ]
             mu.labs <- parse( text=gsub("mu","mu == ",levels(mu.name)) )
@@ -223,7 +223,7 @@ par(mar=c(4,4,1,1)+.1)
                 legend=c(sm.labs,N.labs,mu.labs),
                 pch=c( 21+((1:nlevels(sm.name))%%5), rep(1,nlevels(N.name)), rep(21,nlevels(mu.name)) ), 
[…]

### 2015-08-30 patchy-sim-writeup-plots.R

@@ -70,7 +70,7 @@ sliced <- pophist$pophist[,,2,timeslice]/pophist$pop$params$N
 # plot
 matplot( sliced, type='l', xlab='deme number (space)', ylab='allele frequency', col=adjustcolor(2:10,.5) )
 abline(v=0.5 + which(diff(as.vector(pophist$pop$params$s))!=0), lty=2 )
-text( mean(which(as.vector(pophist$pop$params$s)>0)), .05, labels=as.expression(substitute(s[b]==sb,list(sb=max(pophist$pop$params$s)))) )
+text( mean(which(as.vector(pophist$pop$params$s)>0)), .05, labels=as.expression(substitute(s[p]==sb,list(sb=max(pophist$pop$params$s)))) )
 text( c(.15,.85)*dim(pophist$pophist)[2], .900, labels=as.expression(substitute(s[m]==sm,list(sm=min(pophist$pop$params$s)))) )
 lines( rowMeans(pophist$pophist[,,2,min(timeslice):max(timeslice)])/pophist$pop$params$N, lwd=2 )
 # done

### 2015-08-30 patchy-review-responses-second-round.tex
.

@@
 \begin{quote}
-    Both reviewer 1 and reviewer 2 still question your main result presented in Eq.
-1 and I do agree with them that there is definitely something wrong with the dimensions.
-This was already pointed in the first round of reviews and both reviewers feel that this issue has not been satisfactorily addressed.
-Both reviewer also argued in the first round that the derivations in the paper are overly complicated and that the problem can in fact be intuitively understood via much simpler arguments (both came up with very similar approaches independently).
-You stated in your replies that you prefer to stick with your derivation because you consider it more solid.
-However, I am not even fully convinced that your Eq.
-11 is more correct than the reviewers’ intuitive results.
-Reviewer 1 points out that there are quite substantial deviations in your numerical results (Fig.
-4) and reviewer 2 stresses that the 1D simulations won’t even be able to catch the potential problems associated with the boundary of the target area.
-Clearly, something is not yet fully understood here and I don’t see how the paper can be accepted without having settled these issues.
+    Both reviewer 1 and reviewer 2 still question your main result presented in Eq. 11 and I do agree with them that there is definitely something wrong with the dimensions. 
+    % This was already pointed in the first round of reviews and both reviewers feel that this issue has not been satisfactorily addressed. 
 \end{quote}
@@
 \begin{quote}
-    It would be easy to simply reject the paper at this point.
-However, as I expressed earlier, I still think that this is an extremely interesting problem that could be of great relevance for future studies, given that the authors can convincingly resolve the issues raised above.
-The dimensionality problem clearly needs to be resolved, and it needs to be explained how their result and the intuitive results from the reviewers compare with each other.
+    Both reviewer also argued in the first round that the derivations in the paper are overly complicated and that the problem can in fact be intuitively understood via much simpler arguments (both came up with very similar approaches independently).
+    You stated in your replies that you prefer to stick with your derivation because you consider it more solid. 
+    However, I am not even fully convinced that your Eq. 11 is more correct than the reviewers' intuitive results. 
 \end{quote}
@@
 \begin{quote}
+    Reviewer 1 points out that there are quite substantial deviations in your numerical results (Fig.
+4) and reviewer 2 stresses that the 1D simulations won't even be able to catch the potential problems associated with the boundary of the target area.
+Clearly, something is not yet fully understood here and I don't see how the paper can be accepted without having settled these issues.
+\end{quote}
+
+% \begin{quote}
+    % It would be easy to simply reject the paper at this point.
+However, as I expressed earlier, I still think that this is an extremely interesting problem that could be of great relevance for future studies, given that the authors can convincingly resolve the issues raised above.
+The dimensionality problem clearly needs to be resolved, and it needs to be explained how their result and the intuitive results from the reviewers compare with each other.
+% \end{quote}
+
+\begin{quote}
     I also think that presenting and discussing the intuitive results really would be important for the paper, as originally suggested by reviewers 1 and 2.
-There is a tendency in population genetics to derive complicated “exact” results that are nonetheless of only limited usefulness in reality because it’s unclear how realistic the underlying models are in the first place.
-I won’t hide here that I strongly prefer a heuristic approach that reveals what’s actually going on, over a lengthy technical calculation that may be formally more correct, yet it’s unclear whether this really matters in reality.
+There is a tendency in population genetics to derive complicated ``exact'' results that are nonetheless of only limited usefulness in reality because it's unclear how realistic the underlying models are in the first place.
+I won't hide here that I strongly prefer a heuristic approach that reveals what's actually going on, over a lengthy technical calculation that may be formally more correct, yet it's unclear whether this really matters in reality.
 I feel the majority of the audience of PLoS Genetics is probably with me on this.
@@
 
-\begin{quote}
-    Should you decide to tackle these problems in a major revision, please also address the remaining comments raised by the reviewers.
-Sorry that I can’t bring you better news at this point but I think you will agree that the reviews are pretty clear in what needs to be done in order to get this paper accepted in PLoS Genetics.
-\end{quote}
+% \begin{quote}
+%     Should you decide to tackle these problems in a major revision, please also address the remaining comments raised by the reviewers.
+Sorry that I can't bring you better news at this point but I think you will agree that the reviews are pretty clear in what needs to be done in order to get this paper accepted in PLoS Genetics.
+% \end{quote}
 
@@
 to deal explicitly with the integral over the whole patch,
-which also necessitated adding a new (short) appendix \nameref{apx:qS}.
+which also necessitated adding a new (short) section \nameref{apx:qS}.
 Section \nameref{ss:hitting_occupation} required very little change.
@@
 \reply
-This is done. \revref
+Done.
 
@@
 \reply
-Also done.
+Also done, thanks.
 
@@
 \reply
+In all the simulations, we fixed $s_p=0.01$ (now stated),
+which measured in terms of instantaneous growth is a selection coefficient of about 0.002.
+It's not at all straightforward to determine when the originating mutation arose,
+since this would require tracing lineages back -- doable, but a lot more work.
+Also, it seems that the second issue -- not really treating the boundary correctly
+-- is more significant than the time the numbers take to go from 1 to 100.
+Taking these two together,
+and not wanting to further complicate the paper,
+we decided to not try to pursue this farther.

### 2015-08-30 patchy-selection-paper.tex
.

@@
       mutation rate ($\mu$),
-      and distance between patches ($R$), \revpoint{3}{3}
+      and distance between patches ($R$), 
       as observed in simulations (vertical axis),

### 2015-09-04 patchy-selection-paper.tex
added patchy info from michael
edited through

@@
 \newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
-\newcommand{\given}{\,\vert\,}
+\newcommand{\given}{\,\mid\,}
 \newcommand{\st}{\,\colon\,}
@@
 Which is more likely to occur, as a function of distance between the patches?
-How can we tell which has occurred, i.e.\ what population genetic
+How can we tell which has occurred, i.e., what population genetic
 signal is left by the spread of migrant alleles?
@@
 Such evolution of convergent phenotypes is favored by higher mutational input, 
-i.e.\ higher total mutational rate and/or population size \citep{softsweepsII}.
+i.e., higher total mutational rate and/or population size \citep{softsweepsII}.
 The geographic distribution of populations can also affect the probability of parallel mutation within a species:
@@
 (beginning from a base of multiple copies),
-which whether or not the copies derive from the same mutation,
-leads to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}.
+which 
+leads to genetic patterns similar to convergent adaptation \citep{orr2001sieve,softsweepsI}
+whether or not the copies derive from the same mutation,
 
@@
 that have repeatedly adapted to patches of soil with high concentrations of heavy metals
-(e.g.\ serpentine outcrops and mine tailings) \citep{kruckeberg1951intraspecific,macnair1991evolution,schat1996identical,turner2010serpentine};
+(e.g., serpentine outcrops and mine tailings) \citep{kruckeberg1951intraspecific,macnair1991evolution,schat1996identical,turner2010serpentine};
 the alleles conferring heavy metal tolerance are thought to be locally restricted 
@@
 Similarly, across the American Southwest, a variety of species of animals have developed locally adaptive cryptic coloration
-to particular substrates, e.g.\ dark rock outcrops or white sand dunes
+to particular substrates, e.g., dark rock outcrops or white sand dunes
 \citep{benson1933concealing}.
@@
 at least two distinct genetic changes are responsible for the dark pigmentation found on different outcrops \citep{hoekstra2003different}. 
-Similar situations have been demonstrated in other small rodent systems \citep{steiner2009genetic,kingsley2009melanism,dice1940ecologic}
+Similar situations have been demonstrated in other small rodent systems \citep{dice1940ecologic,steiner2009genetic,kingsley2009melanism}
 and in lizards \citep{rosenblum2010molecular}.
@@
 Under what conditions is it likely that each patch adapts in parallel, 
-i.e.\ convergently through new mutation,
+i.e., convergently through new mutation,
 and when is it likely that migration carries these alleles between patches?
 How easy will it be to detect adaptive alleles that are shared by
-migration, i.e.\ over what genomic scale will a population genetic
+migration, i.e., over what genomic scale will a population genetic
 signal be visible?
@@
 We combine these two results in the section \secref{ss:probparallel} 
-to discuss to the probability of parallel adaptation, equation \eqref{eqn:parallel_prob}.
+to discuss the probability of parallel adaptation, equation \eqref{eqn:parallel_prob}.
 To understand the genomic signal left by adaptations shared by migration, 
 in the section \secref{ss:haplotype_length}, 
-we approximate the time it will take for an allele to transit between patches,
+we study the time it will take for an allele to transit between patches,
 (equation \eqref{eqn:mean_tau}),
-and use this to approximate the length of haplotype that hitchhikes with it (equation \eqref{eqn:haplotype_length}). 
+and thus the length of haplotype that hitchhikes with it (equation \eqref{eqn:haplotype_length}). 
 Finally, in the section \secref{ss:applications} we apply this work to understand the geographic scale of
@@
 or as general as a knockout of a gene or one of a set of genes.
-We assume that this mutational change occurs at a (low) rate of $\mu$ per chromosome per generation,
-and that this initially rare new allele has fitness advantage $s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches.
-Outside of these patches we assume that the new allele has a fitness disadvantage of 
-$s_m$  when rare in the intervening areas, with $s_p$ and $s_m$ both positive.
+This change occurs at a (low) rate of $\mu$ per chromosome per generation,
+and has fitness advantage $s_p$ relative to the unmutated type ($b$) in these ``poor'' habitat patches.
+Outside of these patches the new allele has a fitness disadvantage of 
+$s_m$ in the intervening areas, with $s_p$ and $s_m$ both positive.
 (Here we define ``fitness'' to be the intrinsic growth rate of the type when rare.)
@@
 We are interested in the establishment of mutations in the ``poor'' patches by either
-migration or mutation, and so will focus on whether the allele
+migration or mutation, which depends on whether the allele
 can escape initial loss by drift when rare. 
@@
 
-Also, to avoid introducing more parameters,
-we assume population density $\rho$ is equal across the range (even in the ``poor'' patches).
-If this is not the case, in most equations below, the population density that should be used is the density
-on the poor patches; see \citet{lenormand2002limits} for a more thorough approach.
-We further assume that the variance in offspring number is $\xi^2$, 
+We also assume population density $\rho$ is constant across the range (even in the ``poor'' patches).
+% If this is not the case, in most equations below, the population density that should be used is the density
+% on the poor patches; see \citet{lenormand2002limits} for a more thorough approach.
+The variance in offspring number is $\xi^2$, 
 and that the mean squared distance between parent and child is $\sigma^2$
-(i.e.\ $\sigma$ is the dispersal distance). 
+(i.e., $\sigma$ is the dispersal distance). 
 We will deal with migration by immediately appealing to the central limit theorem,
@@
 and do not discuss the (interesting) cases where rare, long-distance dispersal events
-are more important (see e.g.\ \citet{levin2003ecology,ralph2010parallel,hallatschek2014acceleration} for discussion).  \linelabel{rr:longtailed}
+are more important (see e.g., \citet{levin2003ecology,ralph2010parallel,hallatschek2014acceleration} for discussion).  \linelabel{rr:longtailed}
 %We also aim to describe the scenario where it is sensible to think of large, discrete patches;
@@
 migration in models of spatially varying selection.
-\citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
+\citet{haldane1948theory}, \citet{fisher1950frequencies}, and \citet{slatkin1973geneflow} first analyzed
 deterministic models of of this sort.
@@
 
-Consider first a single, isolated poor habitat patch in which no $B$ allele has yet become established. 
-We first compute the time scale on which a new $B$ mutations will appear and fix in such a patch,
-if the rate of mutation from $b$ to $B$ is $\mu$ per individual per generation.
+We first compute the time scale on which a new $B$ mutations will appear and fix in 
+a single, isolated poor habitat patch in which no $B$ allele has yet become established. 
 As we are interested in patches where local adaptation can occur,
-we will assume that our patch is larger than the cutoff for local establishment 
+we will assume that the patch is larger than the cutoff for local establishment 
 mentioned above.
@@
 but results will be more interpretable if we proceed with a simple approximation.
-The total influx per generation of mutations that successfully establish is given by the
+The total influx per generation of mutations that successfully establish is 
+the product of population density $\rho$, mutation rate $\mu$, and the
 integral of $p(x)$ over the entire species range:
 \begin{equation}
-  \mutrate = \rho \mu \int p(x) dx \label{eqn:exactmutrate}
+  \mutrate = \rho \mu \int p(x) dx \label{eqn:exactmutrate} .
 \end{equation}
 % $\int \rho \mu p(x) dx$,
-and hence depends in a complicated way on the patch geometry and selection coefficients,
+This depends in a complicated way on the patch geometry and selection coefficients,
 but still scales linearly with the mutational influx density $\rho \mu$.
@@
 \end{align}
-i.e.\ the product of mutational influx in the patch ($\rho A\mu$) and the probability of establishment ($p_e \approx 2s_p/\xi^2$).
+i.e., the product of mutational influx in the patch ($\rho A\mu$) and the probability of establishment ($p_e \approx 2s_p/\xi^2$).
 If this rate is low, then the time (in generations) until a mutation arises that
@@
 
-Now suppose that there are two patches, each of area $A$ and separated by distance $R$
+Now suppose that there are two patches separated by distance $R$
 (i.e., the shortest distance between the two is $R$). \revpoint{2}{9}
@@
 obtaining along the way other useful information about the genealogy of $B$ alleles. 
-Doing this we arrive at \eqref{eqn:migrate} for the rate at which 
+Doing this we arrive at equation \eqref{eqn:migrate} for the rate at which 
 an allele established in one patch spreads to a neighboring one by migration.
@@
 Migration--selection balance ensures that 
-there will be some number of $B$ alleles present in the regions where they are disadvantageous,
-but only very few, and rarely, far away from the patch where $B$ is established.
+there will be some $B$ alleles present in the regions where they are disadvantageous,
+but only rarely, far away from the patch where $B$ is established.
 Denote the expected frequency of allele $B$ at displacement $x$ relative to the patch by $q(x)$,
@@
 
-From simulations, $C$ seems to be not too far from 1 in practice (see below).
 As J.\ Hermisson pointed out in comments on an earlier draft,
@@
 These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
-To be clear about the assumptions (especially regarding scaling) implicit here,
+To be clear about the assumptions implicit here,
 below we provide a derivation of \eqref{eqn:eqfreq} in section \secref{apx:eqfreq},  \revpoint{3}{5}
@@
 This expected frequency $q(x)$ is the \emph{time-averaged} occupation frequency,
-i.e.\ the total number of $B$ alleles found across $T$ generations near location $x$, per unit area, divided by $T$.
+i.e., the total number of $B$ alleles found across $T$ generations near location $x$, per unit area, divided by $T$.
 If, for instance, $q(x)=.01$ and the population density is $\rho=10$ individuals per unit area, 
@@
 The relevant time scale on which $B$ alleles migrate between patches is given by the rate of appearance of such families.
-To compute the rate of establishment by such migration,
-we argue that these families (suitably defined) can be approximated by
-independent branching processes whose spatial motion is described by Brownian motion.
-Properties of these, combined with equation \eqref{eqn:eqfreq},
-will lead us eventually to the expression \eqref{eqn:migrate} for the rate of effective migration,
-an approximation that is valid if the distance $R$ is large (at least several times $\sigma/\sqrt{2s_m}$).
+% To compute the rate of establishment by such migration,
+% we argue that these families (suitably defined) can be approximated by
+% independent branching processes whose spatial motion is described by Brownian motion.
+% Properties of these, combined with equation \eqref{eqn:eqfreq},
+% will lead us eventually to the expression \eqref{eqn:migrate} for the rate of effective migration,
+% an approximation that is valid if the distance $R$ is large (at least several times $\sigma/\sqrt{2s_m}$).
 
@@
 
-To make the argument, we decompose the genealogy of $B$ alleles into families in the following way.
+This suggests decomposing the genealogy of $B$ alleles into families in the following way.
 First, consider the genealogy of all $B$ alleles that were alive at any time outside of the original patch.  
@@
 by a collection of independent spatial branching processes
-whose members are ignored if they enter the original patch.
+whose members are ignored if they enter the original patch,
+illustrated in figure~\ref{fig:branching_decomp}.
 (This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
-This decomposition is illustrated in figure~\ref{fig:branching_decomp}.
-The opportunity for adaptation depends on how often these families of $B$ alleles encounter the new patch under
-migration-selection equilibrium.
+The opportunity for adaptation depends on how often these families of $B$ alleles encounter the new patch.
 Suppose that the area occupied by the new patch is $S$.
@@
 but their product does not.
-All terms except the ``outflux of families'' are calculations with spatial branching processes;
-and we can solve for this outflux using the established form for $q(x)$.
-% Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
-% -- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.
 
@@
 we therefore need to look more closely at the rare families of $B$ alleles 
-far from the original patch where it is already established.
-Consider one such family,
-that lives far from the original patch.
-As noted above, we will model this as a branching process,
-i.e.\ by assuming that each individual
+far from the original patch.
+It is reasonable to model each such family as a branching process,
+i.e., assuming that each individual
 migrates, then reproduces, independently of all other family members.
@@
 as depicted in Figure~\ref{fig:branching_decomp}.
-We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} (see also \citet{chauvin1991growing}). 
+We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary}.
+% (see also \citet{chauvin1991growing}). 
 % (lyons1995conceptual deals with conditioned on eventual nonextinction)
@@
 where $W$ is the unconditioned offspring number distribution.  \linelabel{ll:branching_process_fix}
-In particular, $\E[W_s-1] \sim \E[W(W-1)]/\E[W]$;
+In particular, $\E[W_s-1] \simeq \E[W(W-1)]/\E[W]$;
 if $W$ is Poisson then this is equal to $e^{-s_m}$.
@@
 
-This decomposition also tells us
-that the $K$ individuals in a family
-have most recent common ancestor on order of $1/s_m$ generations ago,
-and so are spread out over a region of order $\sigma/\sqrt{s_m}$.  \revpoint{2}{2}
+Subfamilies typically live for $1/s_m$ generations,
+thus spreading across an area of width $\sigma/\sqrt{s_m}$.
+The oldest subfamily in a surviving migrant family 
+is of order $\log(1/s_m)/s_m$ generations,
+so the entire family is spread slightly wider.
+\revpoint{2}{2}
+% If $\P\{ T_k > k \} = e^{-sk}$
+% then 
+% \begin{align}
+%   \P\{ max_{k \ge n} (T_k - k) \le 0 \} &\simeq \exp(-e^{-sn}/s) \\
+%       &= \exp( - \exp( -s(n+\log(s)/s) ) ) \\
+% \end{align}
+% which is Gumbel with mean $(\log(s)+\gamma)/s$ and variance $\pi^2/(6s^2)$.
 
@@
 The remaining term is $q(S)$, the integral of equation \eqref{eqn:eqfreq} over the patch.
-Suppose the new patch is at distance $x$ from the original patch.
-If the new patch has width smaller than $\sigma/\sqrt{2s_m}$, 
-then $q(S)$ is approximately $q(x)$ multiplied by the area of the new patch.
+Since the patches are at distance $R$,
+if the new patch has width smaller than $\sigma/\sqrt{2s_m}$, 
+then $q(S)$ is approximately $q(R)$ multiplied by the area of the new patch.
 If $S$ is larger, then we should use the area of the closest strip
@@
 see section \ref{apx:qS} for more details.
-In either case, denote this area $A$.
+In either case, denote this area $A'$;
+for numerical examples we compute $q(S)$ directly.
 
@@
 \begin{align} \label{eqn:migrate}
-  \migrate(x) &\approx C A \rho s_m \min(s_m,p_e) \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |y| > \sigma/\sqrt{2 s_m}  ,
-%     \begin{cases}
-%         q(x) \times 4 \rho s_m s_p / \xi^2  \qquad & \text{ if } s_p < s_m \\
-%         q(x) \times 2 \rho s_m^2 \qquad & \text{ if } s_p > s_m 
-%     \end{cases} .
-\end{align}
-where $A$ is the area of the new patch if that is no larger than $\sigma/\sqrt{2s_m}$.
+  \migrate(R) &\approx C A' \rho s_m \min(s_m,p_e) \left( |R| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |R| \sqrt{2 s_m} / \sigma)  \qquad \text{for } |R| > \sigma/\sqrt{2 s_m}  ,
+\end{align}
+where $A'$ is the area of the new patch no further than $R+\sigma/\sqrt{2s_m}$
+from the old patch.
 When we use this below, we set $p_e = 2 s_p / \xi^2$.
@@
 migration to be a decreasing function of $s_m$. However, it only increases for small $s_m$, 
-and is a decreasing function of $s_m$ for $s_m > |y|^2 / 2 \sigma^2$, 
+and is a decreasing function of $s_m$ for $s_m > |R|^2 / 2 \sigma^2$, 
 which is the parameter region that we are restricted to by other assumptions.
@@
 Our results should hold exactly in the limit of 
-large, circular patches that are much further apart than they are wide,
+large, circular patches sufficiently far apart,
 large population density,
@@
 should be added to the outside of the patch in computing $A$ for equation \eqref{eqn:mutrate},
-and if $s_m \gg s_p$, a strip of width $\sigma/s_p$ should be subtracted from the patch. \revpoint{2}{6}
+and if $s_m \gg s_p$, a strip of width $\sigma/\sqrt{s_p}$ should be subtracted from the patch. \revpoint{2}{6}
 As for the migration rate,
@@
 summarizes how the results compare to theory,
-excluding parameter combinations where a majority of simulations did not adapt by 25,000 generations
-or the patch was not large enough to support a stable population of $B$ alleles.
+excluding parameter combinations 
+that violate the assumptions discussed above, 
+or where a majority of simulations did not adapt by 25,000 generations.
 Supplementary figures \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times} 
@@
 the harder it is for mutations that arise near the edge of the patch to establish.
-The theoretical values $1/\migrate$ again has a margin of error of a factor of about 2,
-with the exception of the weakest disadvantage off the patch, $s_m=10^{-4}$,
-which adapt much more quickly than predicted by theory.
-However, these simulations fall outside the region where we expect the model to do well,
-since $R < \sigma/\sqrt{s_m} \approx 100$.
-A glance at supplementary figure \ref{sfig:sims_8} shows more:
-since $s_m$ is so small, the second patch is reached by migrants before the migration-selection equilibrium is reached,
-and at migration-selection equilibrium, the two patches act as a single meta-patch.
+The theoretical values $1/\migrate$ again has a margin of error of a factor of about 2.
+% with the exception of the weakest disadvantage off the patch, $s_m=10^{-4}$,
+% which adapt much more quickly than predicted by theory.
+% However, these simulations fall outside the region where we expect the model to do well,
+% since $R < \sigma/\sqrt{s_m} \approx 100$.
+Supplementary figure \ref{sfig:sims_8} shows simulations at more parameter values.
+% since $s_m$ is so small, the second patch is reached by migrants before the migration-selection equilibrium is reached,
+% and at migration-selection equilibrium, the two patches act as a single meta-patch.
 
@@
 
-We will assume that once a $B$ allele is introduced (by migration or mutation) 
+We assume that once a $B$ allele is introduced (by migration or mutation) 
 it becomes established in the poor habitat patch rapidly
@@
 for the effective rates of adaptation by new mutation and by migration.
-We work in one dimension, as the square root term appearing for two dimensions is relatively unimportant 
-(as seen in figure~\ref{fig:sim_occupation_freqs} by the fact that the $d=2$ curve is nearly straight).
+We work in one dimension, as the square root term appearing for two dimensions is relatively unimportant.
+% (as seen in figure~\ref{fig:sim_occupation_freqs} by the fact that the $d=2$ curve is nearly straight).
 \linelabel{rr:sqrt_term}
@@
 be on in order for adaptation via mutation or migration to dominate.
-Let $\gamma = \min(1,s_m/p_e)$.
+Let $\gamma = \min(1,s_m/p_e)$
+and $w = A/A'$ 
+-- since $A$ is the area of the not-yet-adapted patch and $A'$ is the area of its closest $\sigma/\sqrt{2s_m}$ strip,
+$w$ is approximately the width of the patch in units of $\sigma/\sqrt{2s_m}$.
 Effective migration and mutation rates will be on the same order if 
-$A \mu \approx 2 \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
-where $R$ is the distance between the patches,
-and $A$ is the area of the not-yet-adapted patch.
+$A \mu \approx 2 A' \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
+where $R$ is the distance between the patches.
 In other words, the migrational analogue of ``mutational influx'' ($\mu \rho A$) 
-is $2 \rho \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
+is $2 \rho A' \gamma s_m \exp(- R \sqrt{2 s_m} / \sigma)$,
 which depends fairly strongly on the selective disadvantage $s_m$ between patches.
-% Note in particular that if $2 s_m < \mu A$ then migration is always slower than new mutation.
 Equivalently, the rates are roughly equal if
-$R/\sigma = \log(2 \gamma s_m/(A \mu))/\sqrt{2s_m}$,
+$R/\sigma = \log(2 \gamma s_m/(w \mu))/\sqrt{2s_m}$,
 which gives the critical gap size past which adaptation will be mostly parallel
-in terms of selection coefficients, patch size, and mutation rate.
+in terms of selection coefficients, patch width, and mutation rate.
 
 % cdist <- function (sm,z) { (-log(z)+log(2*sm))/sqrt(2*sm) }
+% cdist(c(.05,.001),z=1e-5)
 % cdist(c(.05,.001),z=1e-4)
+% cdist(c(.05,.001),z=1e-3)
 % cdist(c(.05,.001),z=1e-2)
@@
 
-If we take the mutational influx in a patch to be one per 10,000 generations ($\rho A \mu = .0001$),
-$s_m>p_e$ so that $\gamma=1$,
-and the population density to be 1 ($\rho =1$, so $-\log(A\mu)\approx 9.2$)
+If we take 
+$\mu=10^{-5}$,
+the patch width to be ten times the length scale $\sigma/\sqrt{2s_m}$
+so $w=10$ (and $-\log(w\mu)\approx 9.2$),
+and $s_m>p_e$ so that $\gamma=1$,
 then adaptation is mostly parallel for patches separated by gaps larger than $R/\sigma > (9.2+\log(2s_m))/\sqrt{2s_m}$.
@@
 then for convergence to be likely,
-the distance between patches must be at least 21.8 times the dispersal distance (i.e.\ $R/\sigma > 21.8$).
+the distance between patches must be at least 21.8 times the dispersal distance (i.e., $R/\sigma > 21.8$).
 If $s_m$ is much smaller, say $s_m = .001$, 
 the required distance increases to $R/\sigma > 67$.
-On the other hand, if the separation is much smaller than this transitional value, 
-then adaptations are likely to be shared between patches.
-
-If the mutational influx is higher -- say, $\rho A \mu = .01$ --
+
+If the mutation rate is higher -- say, $\mu=10^{-3}$ --
 the required separation between patches is only reduced to $R/\sigma > 7.3$ with $s_m=.05$.
@@
 
-We can go beyond these orders of magnitude calculations to find
+We can go beyond these rough calculations to find
 the probability of parallel adaptation if we are willing to take our approximations at face value. 
@@
 Furthermore, the probability that the second adaptation is a new mutation,
-i.e.\ the probability of parallel adaptation, 
+i.e., the probability of parallel adaptation, 
 with $\gamma = \min(1,s_m/p_e)$,
@@
 \begin{equation} \label{eqn:parallel_prob}
-    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ A \mu / (2s_m) }{ A \mu / (2s_m) + C \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
+    \frac{\mutrate}{\mutrate+\migrate(R)} = \frac{ w \mu / (2s_m) }{ w \mu / (2s_m) + C \gamma \left(R \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2}\exp\left(- \sqrt{2 s_m} R / \sigma \right) },  
 \end{equation}
@@
       Probability that two patches 
-      evolve by independent mutations (i.e.\ convergently),
+      evolve by independent mutations (i.e., convergently),
       as a function of their geographic separation ($R$)
@@
 \subsubsection{Multiple patches}
-Above we worked with two patches separated by distance $R$;
-but these calculations apply as well to multiple patches
-in which the $B$ allele can facilitate local adaptation. 
-Under our assumptions of selective exclusion and rapid establishment, 
-each patch will adapt through establishment of a single copy of the $B$ allele, 
-either by migration or mutation.
-If we imagine that the times between successive establishments are the result of many nearly-independent attempts
+
+Suppose that instead of patches there are many,
+each patch adapting through either migration or mutation.
+If the times between successive establishments are the result of many nearly-independent attempts
 with small probability of success,
@@
 \begin{equation} \label{eqn:total_migrate}
-    \sum_{i=1}^k \migrate(R_i) = C \rho s_m \min(s_m,p_e) \sum_{i=1}^{k} \left(R_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m} /\sigma\right).
+    \sum_{i=1}^k \migrate(R_i) = C \rho A' s_m \min(s_m,p_e) \sum_{i=1}^{k} \left(R_i \sqrt{2 s_m} /\sigma \right)^{-(d-1)/2} \exp\left(- R_i \sqrt{2 s_m} /\sigma\right).
 \end{equation}
@@
 \end{align}
-Equations \eqref{eqn:total_migrate} and \eqref{eqn:migprob} specify the transition rates of the process.
 
@@
 the sets of patches sharing the genetic basis of their adaptive response.
-Also note that we can rescale time by the typical patch size, and introduce a parameter, say $a_k = A_k/\bar A$,
+We can also rescale time by the typical patch size, and introduce a parameter, say $a_k = A_k/\bar A$,
 making the properties (other than the time scaling) of the discrete model independent of the \emph{numerical sizes} of the demes themselves.
@@
 
-The resulting spatially discrete process is quite general,
-but if we oversimplify back to the island model, it is tractable:
-suppose the probability of adaptation by mutation is the same for each patch
-and the probability of migration is the same between each pair of patches.
+We can say more if we suppose that all patches have the same area and are the same distance from each other
+(i.e., the ``island model'').
 The resulting process is a continuous-time version of the ``Chinese restaurant process''
-described in \citet{aldous1985exchangeability} and \citet{pitman1995partitions},
+described in \citet{aldous1985exchangeability}, % and \citet{pitman1995partitions},
 and so the final partition of types across demes has the Ewens
-sampling distribution with parameter $A \mu / (4 s_m \exp (-\sqrt{2 s_m}R/\sigma))$.
+sampling distribution with parameter $w \mu / (4 s_m \exp (-\sqrt{2 s_m}R/\sigma))$.
 Now we can compute most properties we might want about the process.
@@
 and the length of the haplotype shared between this lineage and one in the original patch.
-The entire time that the lineage is outside the region where the $B$ allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
+The time that the lineage is outside the region where the $B$ allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
 the haplotype that accompanies it is broken down rapidly. 
@@
 since most others with which it recombines have similar haplotypic backgrounds. 
-We first discuss the transit of the lineage, and its accompanying haplotype, and return
-to how haplotypes are whittled down within established patches in the next section.
+% We first discuss the transit of the lineage, and its accompanying haplotype, and return
+% to how haplotypes are whittled down within established patches in the next section.
 
@@
 the time spent by the trunk lineage of the first successful migrant ``in transit'' between the patches,
-i.e.\ $\tau$ conditioned on $\tau < \infty$.
+i.e., $\tau$ conditioned on $\tau < \infty$.
 In so doing, we neglect 
@@
 (so that faster migrants would be more likely to have arrived first).
-These factors should have a minor impact if the rest of our framework fits.
-
-Our first obstacle is that we do not know $1-k_e(t)$, 
-the probability a family survives until $t$, exactly.
-However, we do know that $1-k_e(t) \simeq e^{-s_m t} / \E[K]$,
-i.e.\ that it has exponential tails.
-This suggests one further approximation, 
-that the random lifetime of a family is exponentially distributed with mean $1/s_m$;
-equation \eqref{eqn:mean_K} shows this introduces a constant into the equations below. \linelabel{rr:EK}
-Once we do this, 
-we can use standard results on hitting times of $d$-dimensional Brownian motion
+% These factors should have a minor impact if the rest of our framework fits.
+
+Since each generation spent in transit provides an opportunity for recombination,
+if recombination is Poisson, the length of the haplotype (in Morgans)
+initially shared between populations on each side of the selected locus is exponentially distributed
+with mean $\tau$. 
+Therefore, if $L$ is the length of hitchhiking segment on, say, the right of
+the selected locus, then
+\begin{equation} \label{eqn:haplen_cdf2}
+\P\{L>\ell\} = \E[e^{-\ell \tau}|\tau<\infty] .
+%  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
+\end{equation}
+
+Computing this depends on $1-k_e(t)$, 
+the probability a family survives until $t$.
+Since $1-k_e(t) \simeq e^{-s_m t} / \E[K]$,
+we approximate the lifetime distribution of a family by an exponential with mean $1/s_m$.
+% i.e., that it has exponential tails.
+% This suggests one further approximation, 
+% that the random lifetime of a family is exponentially distributed with mean $1/s_m$;
+% equation \eqref{eqn:mean_K} shows this introduces a constant into the equations below. 
+\linelabel{rr:EK}
+We can then use standard results on hitting times of $d$-dimensional Brownian motion
 that is killed at rate $s_m$ (see \citet{borodin2002handbook} 2.2.0.1 and 4.2.0.1).
@@
 $R$ from the already adapted patch, then 
-we know that the Laplace transform of the transit time $\tau$ is
 \begin{align}
@@
 We are interested in lineages that manage to reach the patch before being killed,
-i.e.\ having $\tau < \infty$,
+i.e., having $\tau < \infty$,
 which occurs with probability
@@
 To keep the expressions simple, in the remainder of this section we only compute quantities for one dimension.
-By Bayes' rule, the Laplace transform of $\tau$ for a successful lineage is
+By Bayes' rule, 
 \begin{equation} \label{eqn:haplen_cdf}
@@
 \begin{equation} 
-  \E[\tau] = (R\sqrt{2s_m}/\sigma)\times 1/(2s_m) \label{eqn:mean_tau}
+  \E[\tau \given \tau<\infty ] = (R\sqrt{2s_m}/\sigma)\times 1/(2s_m) \label{eqn:mean_tau}
 \end{equation}
-and that $\var[\tau] = (R\sqrt{2s_m}/\sigma) \times 1/(2s_m)^2$.
+and that $\var[\tau \given \tau<\infty ] = (R\sqrt{2s_m}/\sigma) \times 1/(2s_m)^2$.
 
@@
 % Above, we approximated this by the time $\tau$ it takes the trunk lineage to hit the new patch;
-% now that we are studying a successful migrant family means we need to condition on the establishment occurring, i.e.\ on $\tau < \infty$.
+% now that we are studying a successful migrant family means we need to condition on the establishment occurring, i.e., on $\tau < \infty$.
 % By the above arguments, and equation~\eqref{eqn:estab_prob}, we have that the conditional probability density is
@@
 
-Since each generation spent in transit provides an opportunity for recombination,
-if recombination is Poisson, the length of the haplotype (in Morgans)
-initially shared between populations on each side of the selected locus is exponentially distributed
-with mean $\tau$, so that if $L$ is the length of hitchhiking segment on, say, the right of
-the selected locus, then
-\begin{equation} \label{eqn:haplen_cdf2}
-\P\{L>\ell\} = \E[e^{-\ell \tau}|\tau<\infty] ,
-%  &= \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
-\end{equation}
-and so (one minus) equation~\eqref{eqn:haplen_cdf} in fact gives the cumulative distribution function of $L$.
-The form of this distribution function implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
+The form of equation \eqref{eqn:haplen_cdf} 
+implies that if $Y$ is an exponential random variable with rate $R\sqrt{2}/\sigma$,
 then $L$ has the same distribution as $(Y + \sqrt{s_m})^2 - s_m$.
@@
 
-As a general rule of thumb, equation~\eqref{eqn:mean_tau} means that families who successfully establish
-move at speed $\sqrt{s_m} \sigma$ towards the new patch 
+As a rule of thumb, equation~\eqref{eqn:mean_tau} means that families who successfully establish
+move at speed $\sqrt{2s_m} \sigma$ towards the new patch 
 -- if $s_m$, the strength of the selection against them, is smaller, the need to move quickly is less imperative.
 Then, equation~\eqref{eqn:haplotype_length} means that the haplotype length
-is roughly the length one would expect given the mean transit time.
+is roughly the length one would expect given the mean transit time,
+so more weakly deleterious transiting alleles arrive with them shorter haplotypes.
 Note that we have become used to seeing $\sigma$ divided by $\sqrt{2s_m}$, rather than multiplied;
@@
 and to convert it to a speed this is divided by $1/2s_m$.
-
-In conclusion, the length of shared haplotype drops off quickly with the
-distance between the patches. When $s_m$ is small we expect shorter
-haplotypes as the allele is only weakly deleterious and so successful migrants
-traverse the distance between patches relatively slowly 
-compared to successfully migrating strongly deleterious alleles. 
-Therefore, when adaptation by migration is most likely (small $s_m$) 
-we expect shorter shared haplotypes.
-
-%GRAHAM commented out this This intuitive picture can be expanded by a central limit theorem:
-%for large $R$, the deviation of $\tau$ about its mean is asymptotically Gaussian,
-%and the haplotype halflength $L$ is exponential.
 
@@
 % > about 10 animals with food provided ad libitum, so I think this is an upper
-% > density (i.e.\ 40/hectare).  As for dispersal distance, I'm less sure.
+% > density (i.e., 40/hectare).  As for dispersal distance, I'm less sure.
 % > Studies on closely related species (genus Perognathus) in similar habitats
@@
 % > Michael
+%
+% 
+% On Mon, Aug 31, 2015 at 1:42 PM, Michael Nachman <mnachman@berkeley.edu> wrote:
+% > Peter:
+% >
+% > I think this is better, but I would even be a little more cautious, e.g.:
+% > "SOME of the largest patches of
+% > dark rock range from 10km to 100km wide and 50-400km from each other, and
+% > melanic populations of C. intermedius have been found on many of these
+% > (but not all) [Benson, Dice & Blossom]. However, patches of all sizes
+% > occur across all scales in a heterogeneous manner, so it is not clear
+% > to what degree these should be treated as isolated from each other."
+% >
+% >
+% > Note that I changed the numbers a little.  For example, Armendaris and
+% > Carrizozo are less than 100 km from each other, and both are big.
 
@@
 and generally have a light pelage similar to the predominant rock color.
-However, in some regions these mice live on darker substrates (e.g.\ old lava flows),
+However, in some regions these mice live on darker substrates (e.g., old lava flows),
 and in these patches have much darker pigmentation, 
 presumably to avoid visual predators.
-These regions of darker substrate are typically tens of kilometers wide and hundreds of kilometers apart. \revpoint{5}{7}
+Some of the largest patches of dark rock range from 10km to 100km wide
+and lie 50--400km from each other,
+and dark-colored populations of \emph{C.\ intermedius} have been found on many of these.
+(However, patches of all sizes occur across all scales in a heterogeneous manner.)
+\revpoint{5}{7}
 \citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
@@
 and has a dominant or partially dominant effect depending on the measure of coat color.
-These areas of dark rock are separated from each other by a range of distances, 
-with some being separated from any other by hundreds of kilometers of light colored substrate. 
-The Pinacate allele is not present in a number of other populations of dark colored rock pocket mouse, 
+The Pinacate allele is not present in a number of other populations with dark pelage,
 suggesting these populations have adapted in parallel \citep{nachman2003genetic,hoekstra2003different}.
@@
 despite intervening light habitat.
-We apply our theory to this example,
-in particular investigating how the probability of parallel mutation depends on 
-the distance between dark outcrops.
 
@@
 it is likely that not all of these changes are needed for a population to begin to adapt. 
-Also there a number of genes besides MC1R at which adaptive changes affecting pigmentation 
+Also, there a number of genes besides MC1R at which adaptive changes affecting pigmentation 
 have been identified in closely related species and more broadly across vertebrates \citep{Hoekstra:06}.
-To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$, 
-where the adaptive mutation has to occur at a single base pair; 
-and a high mutation rate $\mu= 10^{-5}$, equivalent to the mutation
-being able to arise at any of a thousand base pairs.  
+To span a range of plausible values, we use a low mutation rate of $\mu= 10^{-8}$
+(a single base pair),
+and a high mutation rate $\mu= 10^{-5}$
+(a kilobase).
 Finally, we set $A=100\text{km}^2$ (roughly the size of the Pinacate
@@
 either a new mutation or migrant allele to be large. 
-We can use equation \eqref{eqn:haplotype_length} to calculate how long the founding
-haplotype shared between populations is expected to be,
-as shown in Figure~\ref{fig:mice_prob_parallel}. 
+Figure~\ref{fig:mice_prob_parallel} also shows how long the founding haplotype 
+shared between populations is expected to be,
+from equation \eqref{eqn:haplotype_length}.
 The initial length can be quite long between geographically close patches (tens of kilometers). 
@@
 and geographically spread populations adapting to homogeneous selection pressures \citep{ralph2010parallel}, 
-where the probability of multiple, independently arising adaptive alleles increases with the population size/density. 
+where the probability of multiple, independently arising adaptive alleles increases with the population size.
 However, in all of these models the dependence on the beneficial selection
@@
 of adaptations between patches easier, 
-it will also makes it harder to spot the sharing,
+it also makes it harder to spot such sharing,
 since the lucky alleles that manage to colonize new patches move slower,
@@
 -- the analogue of the mutational influx $\mu \rho A$ --
-is $2 \rho s_m \exp(- R \sqrt{2 s_m} / \sigma)$.
+is $2 \rho A' s_m \exp(- R \sqrt{2 s_m} / \sigma)$.
 Equivalently, the critical gap size between patches past which adaptation is likely independent
-is $R = (\sigma/\sqrt{2s_m}) \log(2s_m/(A \mu))$.
+is $R = (\sigma/\sqrt{2s_m}) \log(2s_m/(w \mu))$.
 Finally, successfully transiting migrant lineages move at rate $\sigma \sqrt{2s_m}$,
@@
 We have focused on relative rates of adaptation,
-since we are thinking of applications in which we know that adaptation has occurred,
-and the question is whether or not adaptations in distinct patches
+since in applications where adaptation has occurred,
+the question is whether adaptations in distinct patches
 have appeared independently or not.
-However, especially absolute if rates of adaptation by new mutation are low,
-any adaptation that does occur may have to make use of standing variation.
+However, 
+any adaptation that does occur may have to make use of standing variation,
+if mutation rates are low.
 The case of a panmictic population was studied by \citet{softsweepsI},
-and we study the case of a continuous, spatial population in an upcoming paper \citep{ralph2014standing}.
+and we study the case of a continuous, spatial population in \citet{ralph2014standing}.
 If parallelism in local adaptation of the sort we study here is due to standing variation
@@
 Likewise, the haplotype that is shared between patches coadapted by migration
-(equation~\ref{eqn:haplotype_length}) will also break down. 
+will also break down
+(equation~\ref{eqn:haplotype_length}).
 However, a long time after the initial sweep,
@@
 decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
-The literature on clines (e.g.\ \citet{barton1979geneflow}) has important information,
+The literature on clines (e.g., \citet{barton1979geneflow}) has important information,
 but more work is needed
@@
 This work has shown that the length of the shared swept haplotype is often significantly
-shorter than the sweep with each patch, resulting in a pattern of
+shorter than the sweep within each patch, resulting in a pattern of
 shoulders of elevated $F_{ST}$ between adapted populations some distance away from
@@
 We have also ignored the possibility of very long distance migration,
-instead focusing on local dispersal (and using Gaussian and appealing to the central limit theorem).
+instead focusing on local dispersal (hence Gaussian by the central limit theorem).
 However, dispersal distributions can be very heavy tailed, 
@@
 Real habitat geometry can be much more complex,
-e.g.\ with archipelagos of patches of varying sizes, 
+e.g., with archipelagos of patches of varying sizes, 
 or patches connected by long, skinny corridors, for instance.
@@
 locally beneficial alleles were more likely to reproduce.
-Each extant individual then either remained in the same location with probability $1-m$ or
-migrated a random number of steps with probability $m$
+Each extant individual then either remained in the same location with probability $1-m$
+or else migrated a random number of steps 
 in a uniformly chosen cardinal direction; 
@@
 the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
+In 2D, diagonal steps were also used.
 This gave us values of $\sigma=0.95$ deme spacings in 1D
@@
 we computed the rate of intrinsic growth, 
-i.e.\ the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
+i.e., the $s$ so that the numbers of $B$ alleles when rare would change by $e^{st}$ after $t$ generations
 in the absence of migration.
@@
 
-To sample lineages, we first simulated forwards in time the population dynamics,
-then sampled lineages moving back through time
+To sample lineages, we first simulated the population dynamics forwards in time,
+then sampled lineages back through time
 by, in each generation,
@@
 
-When rare, under at least some demographic models,
+When rare, 
 copies of a new mutant allele are approximately independent and experience a uniform selective benefit;
@@
 is determined by this initial phase where it is rare.
-Fortunately, the probability that a branching process does not establish
-(i.e.\ dies out)
+Fortunately, the probability that a branching process dies out
 can be found as a fixed point of the generating function of the process \citep{jagers1975branching}.
@@
     Note that the approximation used in the text is here $p_e \approx 2 s_p / \xi^2 = 0.1$,
-    agreeing with the asymptotic value on the right.
+    agreeing with the value in the center of the patch.
     \label{fig:prob_estab_calcs}
@@
 and see \citet{etheridge2000introduction} and/or \citet{dawson1993measurevalued} for a general framework for the stochastic processes below.
-Following is a description of a discrete, haploid model (which is nearly the one we simulate from).
+
 Suppose that the population is composed of a finite number of small demes of equal size $N$ arranged in a regular grid,
@@
 producing a random number of offspring with distribution given by $X$
-who then all migrates to a new location chosen randomly from the distribution given by $x+R$,
+who then all migrate to a new location chosen randomly from the distribution given by $x+R$,
 where they replace randomly chosen individuals.
@@
 Each individual's time until reproduction is exponentially distributed:
-the reproduction rate is 1 if it carries the original allele, or with rate $1+s(x)$ if it carries the mutant allele.
+the reproduction rate is 1 if it carries the original allele, or is $1+s(x)$ if it carries the mutant allele.
 Suppose that the number of offspring $X$ has mean $\mu$; the variance of $X$ will not enter into the formula
@@
 Also suppose that the migration displacement $R$ has mean zero and variance $\sigma^2$;
-if we are in more than one dimension, we mean that the components of the dispersal distance are uncorrelated
+in more than one dimension, we mean that the components of the dispersal distance are uncorrelated
 and each have variance $\sigma^2$.
-We also assume that the distribution of $R$ makes the associated random walk irreducible and aperiodic.
 
@@
 For $0\le \phi \le 1$, we also denote by $\bar X_\phi$ the random number of mutant alleles added if $X$ new offspring carrying mutant alleles
-replace randomly chosen individuals in a deme where the mutant allele is at frequency $\phi$ (i.e.~hypergeometric with parameters $(X,1-\phi)$);
-similarly, $\widetilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,\phi)$).
+replace randomly chosen individuals in a deme where the mutant allele is at frequency $\phi$ (i.e.~hypergeometric with parameters $(X,N\phi,N(1-\phi))$);
+similarly, $\widetilde X_\phi$ is the number lost if the new offspring do not carry the allele (i.e.~hypergeometric with parameters $(X,N(1-\phi),N\phi)$).
 (We like to think of $\Phi^N_t$ as a measure, but it does not hurt to think of $\Phi^N$ as a vector;
@@
 One reason for this is that at finite $N$,
-the process $\Phi_t$ is an irreducible, aperiodic finite-state Markov chain with absorbing states at 0 and 1;
+the process $\Phi_t$ is an irreducible finite-state Markov chain with absorbing states at 0 and 1;
 therefore, the inevitable outcome is extinction of one type or another,
@@
 
-If $q(r)$ is a radially symmetric equilibrium frequency and $\xi(t,x)=q(|x|)$,
-then $\partial_t \xi(t,x) = 0$.
-So, using \eqref{eqn:radial_fkpp}, a radially symmetric solution 
-with $s(r) = -s < 0$ for all $r>r_0$ 
-solves, for $r_0 < r < \infty$,
+A radially symmetric equilibrium frequency $\xi(t,x)=q(|x|)$,
+with $s(r) = -s < 0$ for all $r>r_0$,
+solves for $r_0 < r < \infty$,
 \begin{align} \label{eqn:radial_eqfreq_diffeq}
@@
 since it relies only on a decomposition of the migrant families upon hitting the new patch;
-but it is also imprecise in a number of subtle ways
+but it is also imprecise in subtle ways
 that are difficult to formalize.
@@
 by following only the motion of the ``trunk''
-(i.e.\ the red line in figure \ref{fig:branching_decomp}),
+(i.e., the red line in figure \ref{fig:branching_decomp}),
 and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
@@
 Denote by $\gamma(x)$ the mean rate of outflux of migrant families from a point $x \in \partial B(r_0)$,
-i.e.\ the time-averaged density of individuals near a point $x$ in $\partial B(r_0)$ 
+i.e., the time-averaged density of individuals near a point $x$ in $\partial B(r_0)$ 
 that are the founders of new migrant families.
-Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``outflux of families'',
+Expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``outflux of families'',
 when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
@@
 started at time 0 with a single individual at $x$,
-and write $S$ for the new patch, which has area $A$ and is centered at $y$.
+and write $S$ for the new patch.
 The \emph{mean occupation measure} of $Z$ in the region $S$,
@@
 % It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
-% i.e.\ the function $u(x,y)$ that, integrated over a set of locations $S$, 
+% i.e., the function $u(x,y)$ that, integrated over a set of locations $S$, 
 % gives the expected total number of individuals ever living in $S$:
@@
 % multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
-% i.e.\
+% i.e.,
 % \begin{align}
@@
 \end{align}
-where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e.\ $B_0=x$),
+where $\P^x$ gives probabilities for Brownian motion begun at $x$ (i.e., $B_0=x$),
 and likewise $\E^x$.
@@
 given that it successfully arrives.
-(This is independent of $x$ in the limit as the distance from $x$ to $S$ grows.)
+(This is approximately independent of $x$.)
 As for $q(S)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
@@
 
-% Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
-% while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
-% Since the expected total time spent in the patch does not depend on the distance at which the family started
-% if we know that it survived to hit the patch,
-% it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
-% Following this intuition,
-% and using $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-% we show below that
-% \begin{align} 
-%     h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
-% \end{align}
-% where $g(A)$ is the total expected occupation time of the patch given the family arrives.
-% Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
 Since the integrands in expressions \eqref{eqn:occupation_integral} and \eqref{eqn:migrate_integral} only differ by a factor
-that (at least asymptotically) does not depend on the distance between $x$ and $y$,
+that (at least asymptotically) does not depend on the distance between $x$ and $S$,
 we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
@@
 across the area occupied by the new patch,
-$q(S) = \int_S q(z) dz$.
+$q(S) = \int_S q(x) dx$.
 The precise answer depends on the shape and orientation of the patches;
 but we aim for a usable approximation,
-primarily in terms of the shortest distance from the old patch to the new patch, denoted $x$,
-and assuming $x$ is large enough we can take equation \eqref{eqn:eqfreq} as an equality.
-
-Since $q(x)$ decreases as $|x|$ increases,
+primarily in terms of the shortest distance from the old patch to the new patch, denoted $R$,
+and assuming $R$ is large enough we can take equation \eqref{eqn:eqfreq} as an equality.
+
+Since $q(R)$ decreases as $|x|$ increases,
 \begin{align}
-    q(S) \le |S| q(x) % C \left( \frac{x\sqrt{2s_m}}{\sigma} \right)^{(d-1)/2} e^{-x\sqrt{2s_m}/\sigma} ,
-\end{align}
-where $|S|$ is the area of $S$.
+    q(S) \le A q(R) % C \left( \frac{R\sqrt{2s_m}}{\sigma} \right)^{(d-1)/2} e^{-R\sqrt{2s_m}/\sigma} ,
+\end{align}
+where $A$ is the area of $S$.
 This will be a good approximation if $S$ is small.
@@
 \begin{align}
-    q(S) &= \int_x^{x+\ell} C e^{-u\sqrt{2s_m}/\sigma} du \\
-         &= C e^{-x\sqrt{2s_m}/\sigma} \frac{\sigma}{\sqrt{2s_m}} \left( 1 - e^{-\ell\sqrt{2s_m}/\sigma} \right) \\
-         &\le C e^{-x\sqrt{2s_m}/\sigma} \frac{\sigma}{\sqrt{2s_m}} .
+    q(S) &= \int_R^{R+\ell} C e^{-u\sqrt{2s_m}/\sigma} du \\
+         &= C e^{-R\sqrt{2s_m}/\sigma} \frac{\sigma}{\sqrt{2s_m}} \left( 1 - e^{-\ell\sqrt{2s_m}/\sigma} \right) \\
+         &\le C e^{-R\sqrt{2s_m}/\sigma} \frac{\sigma}{\sqrt{2s_m}} .
 \end{align}
@@
     q(S) &\le \int_T q(x) dx \\
-        &\le w \int_x^\infty q(r) r dr \\
-        &= w C \int_x^\infty \left( \frac{r\sqrt{2s_m}}{\sigma} \right)^{-1/2} e^{-r\sqrt{2s_m}/\sigma} dr \\
-        &\le w \frac{\sigma}{\sqrt{2s_m}} C \sqrt{\pi} \left( \frac{x\sqrt{2s_m}}{\sigma} \right)^{-1/2} e^{-x\sqrt{2s_m}/\sigma} dr .
+        &\le w \int_R^\infty q(r) r dr \\
+        &= w C \int_R^\infty \left( \frac{r\sqrt{2s_m}}{\sigma} \right)^{-1/2} e^{-r\sqrt{2s_m}/\sigma} dr \\
+        &\le w \frac{\sigma}{\sqrt{2s_m}} C \sqrt{\pi} \left( \frac{R\sqrt{2s_m}}{\sigma} \right)^{-1/2} e^{-R\sqrt{2s_m}/\sigma} dr .
 \end{align}
 If we absorb $\sqrt{\pi}$ into the constant $C$,
-this is of the form $(\text{width}) \times (\sigma/\sqrt{2s_m}) \times q(x)$,
+this is of the form $(\text{width}) \times (\sigma/\sqrt{2s_m}) \times q(R)$,
 i.e., roughly the area, after replacing the length by $\sigma/\sqrt{2s_m}$.
@@
 In both cases,
-the upper bound is $q(S) \le A q(x)$,
-where $A$ is the area of the parts of $S$ that are no more than $x+\sigma/\sqrt{2s_m}$ away from the original patch.
-Lower bounds could be obtained,
-but we stick with this rough approximation.
-
-%%%%%%% %%%%%%%%%%%%
-\subsection[Outflux of Families]{The outflux of families}
-\label{apx:outflux}
-
-It is of interest to have an expression for the ``outflux of families'',
-$\gamma(r_0)$, 
-defined to be the mean density of $B$ alleles at distance $r_0$
-whose lineage traces back to the original patch without venturing further away than $r_0$.
-By expression \eqref{eqn:occupation_integral},
-taking $S$ to be a small area located at $y$,
-\[
-    \gamma(r_0) = \frac{\rho q(y)}{\int_{B(r_0)} u(x,y) dx}
-\]
-for any $y$.  
-Take $\hat x = (1,0)$ and $y=r \hat x$.  Then
-\begin{align}
-  \frac{ q(y) }{ u(x,y) } &\simeq C' \frac{ \|y\|^{-(d-1)/2} e^{-\|y\| \sqrt{2s_m}/\sigma } }{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\| \sqrt{2s_m}/\sigma } } \\
-  &= C' \| \hat x - x/r \|^{(d-1)/2} e^{ r \sqrt{2s_m}/\sigma  (\|\hat x - x/r\| - 1)} .
-\end{align}
-Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
-so that the expression in the exponent is to, first order in $1/r$,
-\begin{align}
-    r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
-            &= - r_0 \cos \theta + O(1/r) .
-\end{align}
-This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
-\begin{align}
-\frac{q(y)}{u(x,y)} \longrightarrow C' e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} \qquad \text{ as } r \to \infty .
-\end{align}
-In one dimension we are done (since $\cos \theta=1$).
-In two dimensions,
-by bounded convergence and \citet{gradshteyn2007table} (3.339),
-\begin{align}
-  \int_{B(r_0)} \frac{q(y)}{u(x,y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-  &= C' \int_0^{2 \pi} e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
-  &= 2 \pi C' I_0(-r_0\sqrt{2 s_m}/\sigma),
-\end{align}
-where $I_0$ is a modified Bessel function of order zero of the first kind.
+the upper bound is $q(S) \le A' q(x)$,
+where $A'$ is the area of the parts of $S$ that are no more than $R+\sigma/\sqrt{2s_m}$ away from the original patch.
+Lower bounds could be obtained along similar lines.
+
+% %%%%%%% %%%%%%%%%%%%
+% \subsection[Outflux of Families]{The outflux of families}
+% \label{apx:outflux}
+% 
+% It is of interest to have an expression for the ``outflux of families'',
+% $\gamma(r_0)$, 
+% defined to be the mean density of $B$ alleles at distance $r_0$
+% whose lineage traces back to the original patch without venturing further away than $r_0$.
+% By expression \eqref{eqn:occupation_integral},
+% taking $S$ to be a small area located at $y$,
+% \[
+%     \gamma(r_0) = \frac{\rho q(y)}{\int_{B(r_0)} u(x,y) dx}
+% \]
+% for any $y$.  
+% Take $\hat x = (1,0)$ and $y=r \hat x$.  Then
+% \begin{align}
+%   \frac{ q(y) }{ u(x,y) } &\simeq C' \frac{ \|y\|^{-(d-1)/2} e^{-\|y\| \sqrt{2s_m}/\sigma } }{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\| \sqrt{2s_m}/\sigma } } \\
+%   &= C' \| \hat x - x/r \|^{(d-1)/2} e^{ r \sqrt{2s_m}/\sigma  (\|\hat x - x/r\| - 1)} .
+% \end{align}
+% Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
+% so that the expression in the exponent is to, first order in $1/r$,
+% \begin{align}
+%     r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
+%             &= - r_0 \cos \theta + O(1/r) .
+% \end{align}
+% This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
+% \begin{align}
+% \frac{q(y)}{u(x,y)} \longrightarrow C' e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} \qquad \text{ as } r \to \infty .
+% \end{align}
+% In one dimension we are done (since $\cos \theta=1$).
+% In two dimensions,
+% by bounded convergence and \citet{gradshteyn2007table} (3.339),
+% \begin{align}
+%   \int_{B(r_0)} \frac{q(y)}{u(x,y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
+%   &= C' \int_0^{2 \pi} e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
+%   &= 2 \pi C' I_0(-r_0\sqrt{2 s_m}/\sigma),
+% \end{align}
+% where $I_0$ is a modified Bessel function of order zero of the first kind.
 
@@
 Here we describe how to solve \eqref{eqn:radial_eqfreq_diffeq} exactly in one dimension,
-i.e.\ the Fisher-KPP equation with piecewise constant selection.
+i.e., the Fisher-KPP equation with piecewise constant selection.
 Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
@@
 (We have then that $\partial_x^2 \phi = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
-Also note that this implies ``conservation of energy'', i.e.\ $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
+Also note that this implies ``conservation of energy'', i.e., $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
 Rearranging, we get that $dx = d\phi / \sqrt{-V(\phi)}$, so where $\phi(x)$ is monotone, the inverse is
@@
     and on the right are twenty-five curves showing the frequencies of $B$ at evenly spaced time points
-    (i.e.\ each line represents a vertical slice through the plot on the left);
+    (i.e., each line represents a vertical slice through the plot on the left);
     dotted black lines indicate the patches where $B$ is advantageous.
@@
 Explicit solutions are not available in the general case,
-but if we specialize to a completely symmetric model (i.e.\ the ``island'' model),
+but if we specialize to a completely symmetric model (i.e., the ``island'' model),
 then the mathematical picture is a pretty one.
@@
 in a patch at migration-selection balance, 
-i.e.\ with the local proportion of $B$ alleles near $x$ equal to $q(x)$,
+i.e., with the local proportion of $B$ alleles near $x$ equal to $q(x)$,
 and suppose that the lineage stays close enough to the patch 
@@
 is equal to this influx divided by the local abundance of $B$ alleles,
-i.e.\ $dt \times q(y) (1+s(y)) m(y,x) / q(x)$.
+i.e., $dt \times q(y) (1+s(y)) m(y,x) / q(x)$.
 This therefore describes the movement of a lineage backwards in time.
@@
 if the assumption that fluctuations in $q(x)$ are small,
-this motion should converge to a diffusion process (as assumed by e.g.\ \citet{hallatschek2008surfing}),
-notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
+this motion should converge to a diffusion process (as assumed by e.g., \citet{hallatschek2008surfing}),
+notwithstanding some technical difficulties (e.g., discontinuity of $s(x)$).
 This diffusion, when at $x$, 
@@
 on the proportion of the patch in which $q(x)$ is intermediate,
-i.e.\ where both $B$ and $b$ are present.
+i.e., where both $B$ and $b$ are present.
 This region where $q(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.

### 2015-09-04 Lava_flow_mice_app.R

@@ -5,35 +5,21 @@ A=100
 
 pdf(file="Lava_flow_mice_prob_parallel.pdf",pointsize=10,width=5,height=4)
 layout((1:2))
-par(mar=c(4,3.5,1,1)+.1)
+par(mar=c(4,3.5,1,1)+.1, mgp=c(2.5,1,0))
 R = seq(5,600,length=10000) ##distance in Km
 
-#layout(t(1:2))
-mu = 1e-8
-prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
-plot(R, prob.parallel,type="l",lwd=3, xlab="", ylab="",log="x") 
-mtext(side=2,line=2.5,"prob parallel mutation")
-mu=1e-5
-prob.parallel <- mu*A/(mu*A + (1/char.length[1])* exp(-R/char.length[1]))  ##assuming sigma=1
+plot(0,type='n', log='x', xlim=range(R), ylim=c(0,1), xlab="Distance between patches (km)", ylab="prob parallel mutations")
+for (cl in char.length) {
+    mu = 1e-8
+    prob.parallel <- mu*A/(mu*A + (sqrt(A)*min(sqrt(A),cl))*(1/cl)* exp(-R/cl))  ##assuming sigma=1
+    lines(R, prob.parallel, lwd=3, lty=match(cl,char.length))
 
-lines(R, prob.parallel,lwd=3,col="red")
+    mu=1e-5
+    prob.parallel <- mu*A/(mu*A + (1/cl)* exp(-R/cl))  ##assuming sigma=1
+    lines(R, prob.parallel, lwd=3, col="red", lty=match(cl,char.length))
+}
 
-
-R = seq(5,600,length=1000) ##distance in Km
-mu = 1e-8
-prob.parallel <- mu*A/(mu*A + (1/char.length[2])* exp(-R/char.length[2]))  ##assuming sigma=1
-lines(R, prob.parallel,lty=2,lwd=3)
-
-mu=1e-5
-prob.parallel <- mu*A/(mu*A + (1/char.length[2])* exp(-R/char.length[2]))  ##assuming sigma=1
-
-lines(R, prob.parallel,lwd=3,col="red",lty=2)
-
[…]

### 2015-09-04 generate-patchy-run.R

@@ -83,6 +83,7 @@ source("sim-patchy-selection-fns.R")
 source("lineages.R")
 
 # postcompute
+#  didn't mean to have sqrt(n) below, meant n/sqrt(2); but that's what we simulated with...
 if (is.null(params$migrsteps)) {
     if (min(params$range)>1) { # 2D
         params$migrsteps <- c( lapply( 1:5, function (n) { c( 2^(-(n+2)), n, 0 ) } ),  ## 2D

### 2015-09-05 patchy-review-responses-second-round.tex
edited through
finished responses

@@
 
-\begin{quote}
-    Both reviewer 1 and reviewer 2 still question your main result presented in Eq. 11 and I do agree with them that there is definitely something wrong with the dimensions. 
-    % This was already pointed in the first round of reviews and both reviewers feel that this issue has not been satisfactorily addressed. 
-\end{quote}
+We would like to submit
+a revised version of our manuscript,
+``Convergent Evolution During Local Adaptation to Patchy Landscapes'',
+for consideration for publication in PLoS Genetics.
+We have dealt with all the reviewer's comments
+(which were useful),
+as detailed below,
+and have further edited the text for readability.
 
-\begin{quote}
-    Both reviewer also argued in the first round that the derivations in the paper are overly complicated and that the problem can in fact be intuitively understood via much simpler arguments (both came up with very similar approaches independently).
-    You stated in your replies that you prefer to stick with your derivation because you consider it more solid. 
-    However, I am not even fully convinced that your Eq. 11 is more correct than the reviewers' intuitive results. 
-\end{quote}
+The main change was that we found where
+the missing factor of area pointed out by the reviewers
+had slipped through the cracks in our argument;
+this required fairly minor re-writing,
+and improved somewhat the fit of the theory to simulations.
 
-\begin{quote}
-    Reviewer 1 points out that there are quite substantial deviations in your numerical results (Fig.
-4) and reviewer 2 stresses that the 1D simulations won't even be able to catch the potential problems associated with the boundary of the target area.
-Clearly, something is not yet fully understood here and I don't see how the paper can be accepted without having settled these issues.
-\end{quote}
+We also turned up a minor bug in the discussion of conditioned branching processes:
+the exact picture is not quite so simple as was previously implied,
+but the approximate picture is still the same --
+see around \llname{ll:branching_process_fix}.
 
-% \begin{quote}
-    % It would be easy to simply reject the paper at this point.
-However, as I expressed earlier, I still think that this is an extremely interesting problem that could be of great relevance for future studies, given that the authors can convincingly resolve the issues raised above.
-The dimensionality problem clearly needs to be resolved, and it needs to be explained how their result and the intuitive results from the reviewers compare with each other.
-% \end{quote}
+Finally, we have further endeavored to bring the intuitive picture to the fore
+and otherwise cut down on needless words.
+In particular, we have 
+re-written parts of the section \secref{ss:heuristics},
+substantially shortened the discussion of branching processes
+in the section \secref{ss:migrant_math},
+and cut entirely the trailing section on ``outflux of migrant families''.
 
-\begin{quote}
-    I also think that presenting and discussing the intuitive results really would be important for the paper, as originally suggested by reviewers 1 and 2.
-There is a tendency in population genetics to derive complicated ``exact'' results that are nonetheless of only limited usefulness in reality because it's unclear how realistic the underlying models are in the first place.
-I won't hide here that I strongly prefer a heuristic approach that reveals what's actually going on, over a lengthy technical calculation that may be formally more correct, yet it's unclear whether this really matters in reality.
-I feel the majority of the audience of PLoS Genetics is probably with me on this.
-\end{quote}
+We hope you find the paper as improved as we do.
 
-% \begin{quote}
-%     Should you decide to tackle these problems in a major revision, please also address the remaining comments raised by the reviewers.
-Sorry that I can't bring you better news at this point but I think you will agree that the reviews are pretty clear in what needs to be done in order to get this paper accepted in PLoS Genetics.
-% \end{quote}
+\vspace{1in}
 
@@
 
-\vskip 4em
+\pagebreak
+
+\textbf{Response to Letter from the Guest Editor:}
+
+\begin{quote}
+    \it
+    Both reviewer 1 and reviewer 2 still question your main result presented in Eq. 11 and I do agree with them that there is definitely something wrong with the dimensions. 
+    % This was already pointed in the first round of reviews and both reviewers feel that this issue has not been satisfactorily addressed. 
+\end{quote}
 
 \reply
-
-The main change,
-which addresses most of the concerns,
-is that we uncovered an error in the derivation of equation \eqref{eqn:migrate},
-that supplied the missing factor of area noted by the reviewers.
-The error occurred in the transition between section \nameref{ss:heuristics} 
-and section \nameref{ss:hitting_occupation}
-because in the former we had tried to keep things simple
-by not including shape of the patch in
-gestalt equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate},
-but then (more correctly) treated the shape in the latter;
+Thanks to the reviewers for sticking to this;
+on revision we uncovered the error in the derivation of equation \eqref{eqn:migrate},
+that supplied the missing factor of area.
+The error occurred in the transition between section \secref{ss:heuristics} 
+and section \secref{ss:hitting_occupation}
+because in the former section we had tried to keep things simple
+by working with densities in gestalt equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate},
+but then (more correctly) treated the shape in the latter section;
 effectively we were treating the patch as very small in one place
 and large in another.
-We've rewritten the argument in \nameref{ss:heuristics} 
+We've rewritten the argument (section \secref{ss:heuristics})
 to deal explicitly with the integral over the whole patch,
-which also necessitated adding a new (short) section \nameref{apx:qS}.
-Section \nameref{ss:hitting_occupation} required very little change.
+which also necessitated adding a new (short) section (\secref{apx:qS})
+that deals with the geometry of the patch.
+Section \secref{ss:hitting_occupation} required very little change.
 
-We also turned up a minor bug in the discussion of conditioned branching processes:
-the exact picture is not quite so simple,
-but the approximate picture is still the same --
-see around \llname{ll:branching_process_fix}.
+\begin{quote}
+    \it
+    Both reviewers also argued in the first round that the derivations in the paper are overly complicated and that the problem can in fact be intuitively understood via much simpler arguments (both came up with very similar approaches independently).
+    % You stated in your replies that you prefer to stick with your derivation because you consider it more solid. 
+    % However, I am not even fully convinced that your Eq. 11 is more correct than the reviewers' intuitive results. 
+    [\ldots] I also think that presenting and discussing the intuitive results really would be important for the paper, as originally suggested by reviewers 1 and 2. 
+\end{quote}
+
+\reply
+We agree; the main argument around these results (in section \secref{ss:heuristics})
+we think follows the arguments of the reviewers closely
+(and arrives at the same answer; see below).
+We also still provide a more detailed argument towards the end of the paper;
+in our view, this gives the best of both worlds.
+
+% \begin{quote}
+%     Reviewer 1 points out that there are quite substantial deviations in your numerical results (Fig.
+4) and reviewer 2 stresses that the 1D simulations won't even be able to catch the potential problems associated with the boundary of the target area.
+Clearly, something is not yet fully understood here and I don't see how the paper can be accepted without having settled these issues.
+% \end{quote}
+
+% \begin{quote}
+    % It would be easy to simply reject the paper at this point.
+However, as I expressed earlier, I still think that this is an extremely interesting problem that could be of great relevance for future studies, given that the authors can convincingly resolve the issues raised above.
+The dimensionality problem clearly needs to be resolved, and it needs to be explained how their result and the intuitive results from the reviewers compare with each other.
+% \end{quote}
+
+% \begin{quote}
+%     I also think that presenting and discussing the intuitive results really would be important for the paper, as originally suggested by reviewers 1 and 2.
+There is a tendency in population genetics to derive complicated ``exact'' results that are nonetheless of only limited usefulness in reality because it's unclear how realistic the underlying models are in the first place.
+I won't hide here that I strongly prefer a heuristic approach that reveals what's actually going on, over a lengthy technical calculation that may be formally more correct, yet it's unclear whether this really matters in reality.
+I feel the majority of the audience of PLoS Genetics is probably with me on this.
+% \end{quote}
+
+% \begin{quote}
+%     Should you decide to tackle these problems in a major revision, please also address the remaining comments raised by the reviewers.
+Sorry that I can't bring you better news at this point but I think you will agree that the reviews are pretty clear in what needs to be done in order to get this paper accepted in PLoS Genetics.
+% \end{quote}
 
@@
 We think the main issue here was inclusion of points outside the validity of the approximations,
-since of the parameters falling in the range of validity discussed in \nameref{ss:assumptions},
+since of the parameters falling in the range of validity discussed in \secref{ss:assumptions},
 the largest discrepancy was a factor of 7, not 100.
@@
 Thanks for catching that mistake.  
-We removed the extra factor of 2 (section~\secref{apx:outflux}).
+% We removed the extra factor of 2 (section~\secref{apx:outflux}).
+However, in the interest of making the paper less complex, we decided to remove that section.
 
@@
 \reply
-This is a nice point;
-perhaps which approach should be included
-comes down to personal taste
-as to whether ``saddle point approximation'' or ``Laplace transform'' is a simpler idea?  
-We decided to stick with the Laplace transform.
+This is a nice point,
+and we've added a note to this effect \revref.
+We also rewrote this section somewhat to flow more naturally:
+since $\P\{L>\ell\}=\E[e^{-\ell\tau}]$ is more properly the quantity we're interested in,
+we start with it.
 
 %%%%%%
-\point{now to the thornier issues of dimension, geometry, and fluxes}{
+\point{now to the thornier issues}{
     \begin{itemize}
@@
             hence one should get
-            \[ \migrate \sim \rho q(x) s^2 \ell^d \sqrt{r/\ell} (1-e^{-p_e/s}) \]
+            \[ \migrate \sim \rho q(R) s^2 \ell^d \sqrt{r/\ell} (1-e^{-p_e/s}) \]
         \item result does depend on target geometry
@@
 \reply
-As discussed above,
-we've debugged the issue about dimensions,
-so that the argument now correctly does integrate over the new patch.
-The result does depend on target geometry, just as the reviewer points out (thanks!);
-this is now discussed in section \secref{apx:qS}.
-The other part of this argument, as we understand it,
-modifies terms in \eqref{eqn:migrate}
-by specifying a form for $A$, the relevant area of the new patch,
-and postulating a specific form for $K$.
-We've opted not to make these changes,
-since these require additional assumptions,
-and it feels easier on the reader to not require them to parse such assumptions.
+As discussed above, we fixed the bug in our argument here.
+Now, \eqref{eqn:migrate} agrees with the reviewer's:
+Both have $\rho q(R)$.
+We have $A'$ for the area of the nearest $\ell=\sigma/\sqrt{2s_m}$ portion of the patch;
+this corresponds above to $\ell^{d} \sqrt{r/\ell}$
+since (see figure in reviewer report) assuming a circular patch, the transverse dimension of $A'$ is $\sqrt{r\ell}$.
+The remaining term is $\E[1-(1-p_e)^K]/(\E[K]/2s_m)$;
+we approximate this by $s_m \min(s_m,p_e)$; 
+above it is $s^2 (1-e^{-p_e/s})$;
+the two agree to first order (and they are both first order approximations).
+The result does, indeed, depend on target geometry;
+we've added a subfigure to figure \ref{fig:branching_decomp} to illustrate this point.
 
@@
 Good point; this works out nicely: \revref.
+We have looked a fair bit for a formal proof in the literature and not turned it up.
 
@@
 For instance,
-if $Z$ is a continuous-time Markovian branching process
+suppose $Z$ is a continuous-time Markovian branching process
 whose spatial motion has generator $A$,
@@
 we chose not to go this route because thinking explicitly about the decomposition into families
-seems much more interesting and useful.
+seems more interesting and useful.
 There is a large and diverse literature on branching Brownian motion,
@@
 \reply
+We have gone back over the section surrounding \eqref{eqn:migrate},
+aiming for maximal simplicity along these lines.
 It's true, that in the case that $p_e < s_m$ we get the same result
@@
 (e.g., by the Fick's law argument of the previous review).
-To our taste, the "complications" make the paper more digestible,
-since it is more clear what's going on,
-especially to those who are less familiar with branching processes.
-
-That said, we do go beyond these approximations, since we include the case $p_e > s_m$,
+However, we do go beyond these approximations, since we include the case $p_e > s_m$,
 where family structure matters --
-so, we're not clear how to address this.
+so, we think we've addressed this,
+and hope we're not missing further simplifications.
 
@@
 Whoops -- we had erroneously left in that statement that the number of offspring each step was Poisson;
-in fact, the number of offspring produced in a generation is either 0 or 1.
+in fact, the number of offspring produced in a generation is either 0 or 1
+(but, individuals may live for many generations).
 The exact function is given in the next paragraph \revref,
@@
 \reply
-That's a trickier question than it at first seems,
-but we've added a rough idea \revref.
+That's a trickier question than it seems --
+in particular, there doesn't exist a map of rock color
+(just rock type, and the map to color is not one-to-one).
+This is further complicated by the fact the mice live on rocky outcrops (only),
+not on sandy or otherwise flat areas.
+After consulting with Michael Nachman,
+we've added a rough idea the relevant scales \revref.
 
@@
 \reply
-Ah, right.
+Ah: now we see what the misunderstanding was.
 By "distance to a patch" we mean the length of the shortest path beginning in one and ending in the other.
 We've clarified this \revref.
-Given this, and the caveats around how to calculate $A$ in \eqref{eqn:migrate},
-we still don't think the assumption is necessary?
+Given this, and the discussion of how to calculate $A'$,
+we still don't think the assumption is necessary.
 
@@
 are the leading ones,
-depending on how $s_m$ compares to $R/\sigma$.
+depending on how $s_m$ compares to $\sigma/R$.
 

### 2015-09-05 patchy-selection-paper.tex
finished responses

@@
 \iflatexml\reviewresponsesfalse\fi  % they don't work so well in html
-% \newcommand{\responsefile}{pbio-reviews-19sept12-responses.tex}  % name of the review reponses file
-\newcommand{\responsefile}{patchy-review-responses-second-round.tex}  % name of the review reponses file
+% \newcommand{\responsefile}{pbio-reviews-19sept12-responses.tex}  % name of the review responses file
+\newcommand{\responsefile}{patchy-review-responses-second-round.tex}  % name of the review responses file
 
@@
 \newcommand{\deq}{\stackrel{\scriptscriptstyle{d}}{=}}
-\newcommand{\given}{\,\mid\,}
+\newcommand{\given}{\mid}
 \newcommand{\st}{\,\colon\,}
@@
 Now suppose that there are two patches separated by distance $R$
-(i.e., the shortest distance between the two is $R$). \revpoint{2}{9}
+(i.e., the shortest distance between the two is $R$). \revpoint{2}{8}
 If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
@@
 by first forming a tree in which each individual has a random number of offspring,
-then recursively assigning spatial locations to each individual in the tree,
-and finally pruning any branches that happen to wander into a patch.
-By our definition of fitness, 
-in this tree the mean number of offspring per individual is $e^{-s_m} < 1$.
+then assigning spatial locations to each individual in the tree,
+and finally pruning any branches that wander into a patch.
 This pruning removes both branches leading back into the already established patch (which are irrelevant)
@@
 Since we are concerned with the rare migrant families that transit quickly between patches, 
-back migrants to the already established patch will contribute little, since they are going the wrong direction;
-so to simplify the analysis we ignore this detail.  % but could add this detail in an appendix
+we ignore back migrants to the already established patch. 
+% will contribute little, since they are going the wrong direction;
+% so to simplify the analysis we ignore this detail.  % but could add this detail in an appendix
 
@@
 which dies at a random time whose distribution is given by $k_e(t)$,
-and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
+surrounded by ``branches'',
+% and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
 as depicted in Figure~\ref{fig:branching_decomp}.
-We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary}.
+We can understand this with a decomposition described by \citet{geiger1999elementary}.
 % (see also \citet{chauvin1991growing}). 
 % (lyons1995conceptual deals with conditioned on eventual nonextinction)
-When we condition on survival until $t$, we condition on existence of
-at least one trunk lineage surviving from time $0$ to time $t$ (the long red
-lineage in Figure \ref{fig:branching_decomp}).
-Given this lineage, nothing distinguishes any of the other reproduction events --
-each individual along the trunk lineage must give birth to at least one offspring that survives until time $t$,
-and all other reproduction events are unconditioned.
-In other words, the genealogy of a family that survives until $t$
+When we condition on survival until $t$, we condition on survival of at least one trunk lineage
+(the long red lineage in Figure \ref{fig:branching_decomp}).
+Given this lineage, the other reproduction events are undistinguished --
+% each individual along the trunk lineage must give birth to at least one offspring that survives until time $t$,
+% and all other reproduction events are unconditioned.
+% In other words, 
+the genealogy of a family that survives until $t$
 can be decomposed into a trunk that lasts until $t$
-and that sprouts independent branches that may or may not survive until $t$.
-Specifically, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
+and that sprouts independent, unconditioned branches that may or may not survive until $t$.
+Concretely, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
 % (so $\P\{Z_t = 0\} = k_e(t)$),
-then this gives the following decomposition:
+then
 \begin{align} \label{eqn:Zt_decomposition}
@@
   \end{center}
-\caption{Cartoon of the decomposition described in the text. 
+  \caption{
+      \textbf{(a)} 
+      Cartoon of the decomposition described in the text. 
     Migrants families that get further than $r_0$ from the already-established patch (grey circle)
@@
     (and hence should be much wigglier than in this cartoon, which confounds space and time).
+      \textbf{(b)} 
+      Diagram of the area $A'$ used in equation \eqref{eqn:migrate}:
+      $R$ is the shortest distance from the old patch to the new,
+      and $A'$ is the area of the portion of the new patch that is no more than distance $R+\sigma/\sqrt{2s_m}$
+      from the old patch.
 } \label{fig:branching_decomp}
@@
 of width $\sigma/\sqrt{2s_m}$ to the original patch;
-see section \ref{apx:qS} for more details.
+see figure \ref{fig:branching_decomp}, and section \secref{apx:qS} for more details.
 In either case, denote this area $A'$;
@@
 \begin{equation} \label{eqn:haplen_cdf}
-\E[e^{-\ell \tau}|\tau<\infty]  = \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
+    % \E[e^{-\ell \tau}|\tau<\infty]  =
+    \P\{ L > \ell \}
+    = 
+    \exp\left\{{-\frac{R}{\sigma}\left(\sqrt{2(\ell+s_m)} - \sqrt{2s_m}\right)}\right\} .
 \end{equation}
@@
 and that $\var[\tau \given \tau<\infty ] = (R\sqrt{2s_m}/\sigma) \times 1/(2s_m)^2$.
+(A saddle point approximation provides an alternative route to these expressions.) \revpoint{1}{4}
 
@@
 (However, patches of all sizes occur across all scales in a heterogeneous manner.)
-\revpoint{5}{7}
+\revpoint{2}{7}
 \citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
@@
 
-In the developement above
+In the development above
 we need to approximate the integral of equation \eqref{eqn:eqfreq}

### 2015-09-05 analyze-sims.R

@@ -38,7 +38,7 @@ mutsims$adapted <- (
     ( mutsims$time1 >= mutsims$hit100.1 ) )
 mutsims$adapttime <- mutsims$hit100.1
 
-pdf(file="mutation-times-predicted.pdf", width=4, height=5, pointsize=10)
+pdf(file="mutation-times-predicted.pdf", width=6, height=5, pointsize=10)
 with( subset(mutsims,adapted), 
         plot( muttime, adapttime, log='xy', xlab="mean time to adaptation by mutation", ylab="time to hit 100 in patch",
                 xlim=c(120,120000), col=sm.name, pch=as.numeric(N.name) )
@@ -130,7 +130,7 @@ migsims$valid <- with( migsims,
     )
 
 
-pdf(file="migration-time-predicted.pdf",width=4, height=8, pointsize=10)
+pdf(file="migration-time-predicted.pdf",width=6, height=8, pointsize=10)
 layout((1:2))
 with( subset(migsims,adapted), {
         plot( migtime, hit100.2, col=sm.name, pch=as.numeric(N.name), log='xy', 

### 2015-09-08 patchy-review-responses-second-round.tex
done with revisions and replies

@@
 \reply
-Good point; this works out nicely: \revref.
+Good point; we've added this, and it works out nicely: \revref.
 We have looked a fair bit for a formal proof in the literature and not turned it up.
+There are general results that say that for small $s$, 
+a subcritical branching process looks nearly like the critical case,
+where for instance $\P\{Z_n s/(1-(1-s)^{-n}) \le u \given Z_n>0\} \simeq 1-e^{-2u/\var[Z_1]}$
+\citep{jagers1975branching},
+but it doesn't seem like invoking these results would simplify the exposition.
 

### 2015-09-11 patchy-selection-paper.tex
moved notes on lineage movement to clinal-lineages

@@
   \begin{center}
+    % made in patchy-sim-writeup-plots.R
     \includegraphics{sim-occupation-freqs}
@@
 for helpful comments and discussion of the manuscript.
+\ifsubmissionversion
+% PLoS puts this somewhere else
+\else
 This work was supported by grants from the
@@
 RO1GM107374 to GC.
+\fi
 
@@
 % endfloat uses different names
+\renewcommand{\thesfigure}{S\arabic{sfigure}}
+\renewcommand{\thetable}{S\arabic{table}}
 \ifsubmissionversion
-  \renewcommand{\thesfigure}{S\arabic{sfigure}}
   \renewcommand{\thepostsfigure}{S\arabic{postsfigure}}
@@
   % \setcounter{postfig}{0}
-  \renewcommand{\thetable}{S\arabic{table}}
   \renewcommand{\theposttable}{S\arabic{posttable}}

### 2015-09-11 patchy-sim-writeup-plots.R

@@ -52,6 +52,7 @@ par(mar=c(5,4,1,1)+.1)
 # averaged
 plot( 1, 1, xlim=range(unlist(lapply(plotinfo,"[[","patchloc"))), ylim=range(unlist(lapply(plotinfo,"[[","obs.freqs"))), 
         log='y', xlab='distance from patch', ylab='allele frequency', type='n' )
+mtext("A",side=3,line=0.1,cex=1.5,adj=-0.3,padj=1)
 abline(v=0,lty=2)
 lapply( plotinfo, function (x) { with(x, {
         lines( patchloc, obs.freqs, pch=20, cex=.5, col=grey(.7) )
@@ -69,6 +70,7 @@ timeslice <- floor( seq( .05*dim(pophist$pophist)[4], dim(pophist$pophist)[4], l
 sliced <- pophist$pophist[,,2,timeslice]/pophist$pop$params$N
 # plot
 matplot( sliced, type='l', xlab='deme number (space)', ylab='allele frequency', col=adjustcolor(2:10,.5) )
+mtext("B",side=3,line=0.1,cex=1.5,adj=-0.3,padj=1)
 abline(v=0.5 + which(diff(as.vector(pophist$pop$params$s))!=0), lty=2 )
 text( mean(which(as.vector(pophist$pop$params$s)>0)), .05, labels=as.expression(substitute(s[p]==sb,list(sb=max(pophist$pop$params$s)))) )
 text( c(.15,.85)*dim(pophist$pophist)[2], .900, labels=as.expression(substitute(s[m]==sm,list(sm=min(pophist$pop$params$s)))) )

### 2015-10-15 patchy-selection-paper.tex
utf8
new sentence in abstract

@@
 \usepackage{etoolbox}
+\usepackage[utf8]{inputenc}
 
@@
 as spatial branching processes.
+A main result is that locally adapted alleles in different patches
+are likely to be of independent origin
+if the distance between the patches is larger than
+$(\sigma/s_m)\log(Cs_m/\mu)$, 
+where $\sigma$ is the dispersal distance, 
+$s_m$ is the selective disadvantage of these alleles between patches,
+$\mu$ is the mutation rate, and $C$ is a constant depending on the probability of establishment and patch geometry.
 This provides a way to understand the role of geographic separation between patches 

### 2015-10-17 patchy-selection-paper-for-PLoS.tex
moving over to plos

--- 2015-09-05 patchy-selection-paper.tex
@@
-\documentclass{article}
-\usepackage{fullpage}
+% Template for PLoS
+% Version 3.1 February 2015
+%
+% To compile to pdf, run:
+% latex plos.template
+% bibtex plos.template
+% latex plos.template
+% latex plos.template
+% dvipdf plos.template
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
+% the final text of your manuscript.
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig." instead of "Figure".
+% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - tabs/spacing/line breaks within cells to alter layout or alignment
+% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
+% - colors, shading, or graphic objects
+% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
+%
+% Please be sure to include all portions of an equation in the math environment.
+%
+% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% fixltx2e package for \textsubscript
+\usepackage{fixltx2e}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
 \usepackage{amsmath,amssymb}
-\usepackage{marginnote}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage[hidelinks]{hyperref}
-\usepackage{etoolbox}
-
-% make supp figures separate
-\usepackage{caption}
-\DeclareCaptionType{sfigure}[Figure]
-
-\setcounter{secnumdepth}{0}
-
-% turn this on to doublespace and put all figures at the very end.
-\newif\ifsubmissionversion
-% \submissionversiontrue
-\submissionversionfalse
-
-\newif\iflatexml\latexmlfalse %%%graham added; remove with latexml package
-
-%%%%%
-% Responses to reviews:
-% set this to show line numbers and include responses to reviews or not
-\newif\ifreviewresponses
-\reviewresponsestrue  % include them
-% \reviewresponsesfalse  % don't include them
-\iflatexml\reviewresponsesfalse\fi  % they don't work so well in html
-% \newcommand{\responsefile}{pbio-reviews-19sept12-responses.tex}  % name of the review responses file
-\newcommand{\responsefile}{patchy-review-responses-second-round.tex}  % name of the review responses file
-
-% override above choices (see makefile)
-\ifdef{\submissionnofigs}{  % for submission, no figures
-  \reviewresponsesfalse
-  \submissionversiontrue
-}{
-}
-\ifdef{\submissionfigs}{  % for reviewers, with figs
-  \reviewresponsestrue
-  \submissionversionfalse
-}{
-}
-\ifdef{\arxiv}{  % for general use
-  \reviewresponsesfalse
-  \submissionversionfalse
-}{
-}
-
-\ifreviewresponses
-  \input{review-response-commands}
-\else
-  \newcommand{\linelabel}[1]{}
-  \newcommand{\revpoint}[2]{}
-  \newcommand{\includereviews}{}
-\fi
-
-% % Text layout
-% \topmargin 0.0cm
-% \oddsidemargin 0.5cm
-% \evensidemargin 0.5cm
-% \textwidth 16cm 
-% \textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% rotating package for sideways tables
+\usepackage{rotating}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
 % Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-\ifsubmissionversion
-    \usepackage[nolists,nomarkers,nohead]{endfloat}
-    \usepackage{cite}
-    \DeclareDelayedFloat{sfigure}[sss]{Supplementary Figure}
-    % Use the PLoS provided bibtex style
-    \bibliographystyle{plos2009}
-    \newcommand{\citep}[1]{\cite{#1}}
-    \newcommand{\citet}[1]{\cite{#1}}
-\else
-    \usepackage{natbib}
-    \bibliographystyle{abbrvnat}
-\fi
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
 
@@
 
-% \pagestyle{myheadings}
-%% ** EDIT HERE **
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
+
+%% Include all macros below
+
+\newcommand{\citep}[1]{\cite{#1}}
+\newcommand{\citet}[1]{\cite{#1}}
 
@@
 
-\newcommand{\gc}[1]{{\it\color{green}(#1)} }
-\newcommand{\plr}[1]{{\it\color{blue}(#1)}}
-
 % make references to named sections look better
@@
 
-% short title:  Convergent Evolution on Patchy Landscapes
-\title 
-{
+%% END MACROS SECTION
+
+\begin{document}
+\vspace*{0.35in}
+
+% Title must be 250 characters or less.
+% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
+\begin{flushleft}
+{\Large
+\textbf\newline{
 Convergent Evolution During Local Adaptation to Patchy Landscapes
 }
-
-\author{Peter L. Ralph$^1$ \\ email: pralph@usc.edu  \and Graham Coop$^2$ \\ email: gmcoop@ucdavis.edu }
-
-\begin{document}
-
-\maketitle
-\begin{center}
-
-$^1$ Computational Biology and Bioinformatics \\ 
-University of Southern California, Los Angeles, CA \\
-$^2$Center for Population Biology \& Department of Evolution and Ecology \\ 
-University of California -- Davis, Davis, CA, 95616
-\end{center}
-
-% Keywords
-% convergent evolution; parallel evolution; MC1r; local adaptation; landscape genetics; branching processes; reaction-diffusion equations
-
-%%%%%%%%%%%
+}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Peter L.\ Ralph\textsuperscript{1,\Yinyang,*},
+Graham Coop\textsuperscript{2,\Yinyang}
+\\
+\bigskip
+\bf{1} Computational Biology \& Bioinformatics, University of Southern California, Los Angeles, California, USA
+\\
+\bf{2} Evolution \& Ecology, University of California, Davis, California, USA
+\\
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+\Yinyang These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+% \ddag These authors also contributed equally to this work.
+
+% Current address notes
+% \textcurrency a Insert current address of first author with an address update
+% \textcurrency b Insert current address of second author with an address update
+% \textcurrency c Insert current address of third author with an address update
+
+% Deceased author note
+% \dag Deceased
+
+% Group/Consortium Author Note
+% \textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* pralph@usc.edu
+
+\end{flushleft}
+% Please keep the abstract below 300 words
 \section*{Abstract}
-Species often encounter, and adapt to, many patches of locally similar environmental conditions across their range. 
-Such adaptation can occur through convergent evolution if different alleles arise and spread in different patches, 
+Species often encounter, and adapt to, 
+many patches of similar environmental conditions across their range. 
+Such adaptation can occur through convergent evolution 
+if different alleles arise in different patches, 
 or through the spread of shared alleles by migration
@@
 The tension between the two reflects 
-the degree of constraint imposed on evolution by the underlying genetic architecture 
-versus how effectively selection and geographic isolation act to inhibit the geographic spread of locally adapted alleles.
-This paper studies a model of the balance between these two routes to adaptation in continuous environments with patchy selection pressures.
+the constraint imposed on evolution by the underlying genetic architecture 
+versus how effectively selection and geographic isolation 
+act to inhibit the geographic spread of locally adapted alleles.
+This paper studies the balance between these two routes to adaptation 
+in a model of continuous environments with patchy selection pressures.
 We address the following questions: 
-How long does it take for a novel, locally adapted allele to appear in a
-patch of habitat where it is favored through mutation?
+How long does it take for a novel allele to appear in a
+patch where it is locally adapted through mutation?
 Or, through migration from another, already adapted patch?
 Which is more likely to occur, as a function of distance between the patches?
-How can we tell which has occurred, i.e., what population genetic
-signal is left by the spread of migrant alleles?
-To answer these questions we examine the family structure underlying migration--selection equilibrium
+What population genetic signal is left by the spread of migrant alleles?
+To answer these questions we examine the family structure 
+underlying migration--selection equilibrium
 surrounding an already adapted patch,
-in particular treating those rare families that reach new patches
+treating those rare families that reach new patches
 as spatial branching processes.
+A main result is that locally adapted alleles in different patches
+are likely to be of independent origin
+if the distance between the patches is larger than
+$(\sigma/s_m)\log(Cs_m/\mu)$, 
+where $\sigma$ is the dispersal distance, 
+$s_m$ is the selective disadvantage of these alleles between patches,
+$\mu$ is the mutation rate, 
+and $C$ depends on the probability of establishment and patch geometry.
 This provides a way to understand the role of geographic separation between patches 
@@
 
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
 \section*{Author Summary}
@@
 and more generally in understanding how species adapt to patchy, heterogeneous landscapes.
+
+\linenumbers
 
@@
 
-These asymptotics fit simulations quite well, as shown in Figure~\ref{fig:sim_occupation_freqs}.
+These asymptotics fit simulations quite well, as shown in Fig~\ref{fig:sim_occupation_freqs}.
 To be clear about the assumptions implicit here,
@@
 In one dimension, the equation can be solved to give $q(x)$ 
-in terms of Jacobi elliptic functions; see the Supporting Information.
+in terms of Jacobi elliptic functions \citep{NEQwiki}; 
+see the Supporting Information.
 
@@
   \begin{center}
-    \includegraphics{sim-occupation-freqs}
+    % made in patchy-sim-writeup-plots.R
+    % \includegraphics{sim-occupation-freqs}
   \end{center}
   \caption{
+      \textbf{Simulated migration-selection balance.}
   \textbf{(A)} 
@@
 alleles, interspersed by long periods of absence. 
-An example of one of these rare excursions is shown in Figure~\ref{fig:sim_snapshots}.
+An example of one of these rare excursions is shown in Fig~\ref{fig:sim_snapshots}.
 The relevant time scale on which $B$ alleles migrate between patches is given by the rate of appearance of such families.
@@
   \caption{
-  A segment of the temporal dynamics of the same simulation as Figure~\ref{fig:sim_occupation_freqs}B.
+      \textbf{A segment of the temporal dynamics of the same simulation as Fig~\ref{fig:sim_occupation_freqs}B.}
   Space is again shown across the horizontal axis, with time on the vertical axis; 
@@
 whose members are ignored if they enter the original patch,
-illustrated in figure~\ref{fig:branching_decomp}.
+illustrated in Fig~\ref{fig:branching_decomp}.
 (This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
@@
 %%%% %%%%
-\subsubsection{The genealogy of migrant families}
+\paragraph{The genealogy of migrant families}
 \label{ss:migrant_math}
@@
 % and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
-as depicted in Figure~\ref{fig:branching_decomp}.
+as depicted in Fig~\ref{fig:branching_decomp}.
 We can understand this with a decomposition described by \citet{geiger1999elementary}.
@@
 When we condition on survival until $t$, we condition on survival of at least one trunk lineage
-(the long red lineage in Figure \ref{fig:branching_decomp}).
+(the long red lineage in Fig \ref{fig:branching_decomp}).
 Given this lineage, the other reproduction events are undistinguished --
@@
   \caption{
+      \textbf{Cartoon of the decomposition described in the text.}
       \textbf{(a)} 
-      Cartoon of the decomposition described in the text. 
     Migrants families that get further than $r_0$ from the already-established patch (grey circle)
@@
     and those that did not survive until $t$ (in grey).
-    Figure~\ref{fig:sim_snapshots} has a depiction of such a family, from a simulation. 
+    Fig~\ref{fig:sim_snapshots} has a depiction of such a family, from a simulation. 
     Note that the spatial motions of lineages are independent random walks
@@
 %%%% %%%%
-\subsubsection{Heuristics}
+\paragraph{Heuristics}
 \label{ss:heuristics}
@@
 of width $\sigma/\sqrt{2s_m}$ to the original patch;
-see figure \ref{fig:branching_decomp}, and section \secref{apx:qS} for more details.
+see Fig \ref{fig:branching_decomp}, and section \secref{apx:qS} for more details.
 In either case, denote this area $A'$;
@@
 were present in the previously unadapted patch.
-Figure~\ref{fig:sim_times}
+Fig~\ref{fig:sim_times}
 summarizes how the results compare to theory,
@@
 or where a majority of simulations did not adapt by 25,000 generations.
-Supplementary figures \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times} 
+\ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times} Fig
 show all times, 
-and depictions of typical simulation runs are shown in supplementary figures
-\ref{sfig:sims_1}, \ref{sfig:sims_2}, \ref{sfig:sims_3}, \ref{sfig:sims_4}, \ref{sfig:sims_5}, \ref{sfig:sims_6}, \ref{sfig:sims_7}, and \ref{sfig:sims_8}.
+and depictions of typical simulation runs are shown in 
+\ref{sfig:sims_1}, \ref{sfig:sims_2}, \ref{sfig:sims_3}, \ref{sfig:sims_4}, \ref{sfig:sims_5}, \ref{sfig:sims_6}, \ref{sfig:sims_7}, and \ref{sfig:sims_8} Fig.
 
@@
 % since $R < \sigma/\sqrt{s_m} \approx 100$.
-Supplementary figure \ref{sfig:sims_8} shows simulations at more parameter values.
+\ref{sfig:sims_8} Fig shows simulations at more parameter values.
 % since $s_m$ is so small, the second patch is reached by migrants before the migration-selection equilibrium is reached,
@@
   \caption{
+      \textbf{Observed versus predicted time until adaptation.}
       \textbf{(left)} 
@@
       and $s_m > 2 \sigma \rho$);
-      the remaining combinations are shown in figures
-      S\ref{sfig:sim_migration_times} and S\ref{sfig:sim_mutation_times}.
+      the remaining combinations are shown in 
+      S\ref{sfig:sim_migration_times} and S\ref{sfig:sim_mutation_times} Fig.
   }   \label{fig:sim_times}
@@
 We work in one dimension, as the square root term appearing for two dimensions is relatively unimportant.
-% (as seen in figure~\ref{fig:sim_occupation_freqs} by the fact that the $d=2$ curve is nearly straight).
+% (as seen in Fig~\ref{fig:sim_occupation_freqs} by the fact that the $d=2$ curve is nearly straight).
 \linelabel{rr:sqrt_term}
@@
 \linelabel{rr:separated_patches}
-
-% One way to visualize this is by a phase diagram.
-% In figure \ref{fig:phase_diagram},
-% we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
-% as a function of various parameters. 
-% This gives us an idea of not only the parameter space where parallel adaptation is likely ($\Tmig \ll \Tmut$),
-% but also where adaptation by any means takes a very long time ($\Tmig$ and $\Tmut$ both very large),
-% an important check in applications of these results.
-% 
-% \begin{figure}[ht]
-%   \begin{center}
-%     \includegraphics{phase-diagram-log}
-%   \end{center}
-%   \caption{
-%   Shaded contour plots of
-%   mean time until adaptation in a new patch occurs by migration ($\Tmig$, blue) or by mutation ($\Tmut$, red),
-%   as a function of population density ($\rho$), 
-%   distance to an already adapted patch in units of dispersal distance ($R/\sigma$),
-%   strength of deleterious selection ($s_m$),
-%   and mutation rate scaled by patch area ($\mu A$).
-%   The right-hand plot shows that $\migrate$ is relatively insensitive to $s_m$ (thanks to the square root).
-%   The black, dotted line shows values where $\Tmig=\Tmut$.
-%   Default parameter values were $A=100$, $\rho=10$, $\mu=10^{-8}$, $s_p=0.1$, $s_m=0.01$, $R/\sigma=20$, and $\xi^2=1$.
-%   Note that population density $\rho$ and mutation influx $A\mu$ are shown on a log scale.
-%   \label{fig:phase_diagram}
-%   }
-% \end{figure}
 
@@
 
-We tested this using the same simulations as figure \ref{fig:sim_times},
+We tested this using the same simulations as Fig \ref{fig:sim_times},
 by using the empirical distributions of the respective times to adaptation
@@
 The results are compared to \eqref{eqn:parallel_prob}
-in figure~\ref{fig:sim_probs}.
+in Fig~\ref{fig:sim_probs}.
 
@@
       % made in analyze-sims.R
-      \includegraphics{prob-mutation-compared}
+      % \includegraphics{prob-mutation-compared}
   \end{center}
   \caption{
-      Probability that two patches 
+      \textbf{Probability that two patches 
       evolve by independent mutations (i.e., convergently),
       as a function of their geographic separation ($R$)
-      and the strength of selection between patches ($s_m$).
+      and the strength of selection between patches ($s_m$).}
       The background shades of grey and contours
       give the probability predicted by theory (equation~\eqref{eqn:parallel_prob}, with $C=5$).
-      Points correspond to parameter combinations of the simulations in figure \ref{fig:sim_times},
+      Points correspond to parameter combinations of the simulations in Fig \ref{fig:sim_times},
       with the shade of grey and the red numeric label indicating the same probability
@@
 %%%%%%%%%%%%
-\subsubsection{Multiple patches}
+\paragraph{Multiple patches}
 
@@
   \begin{center}
-      \includegraphics{sim-transit}
+      % \includegraphics{sim-transit}
   \end{center}
   \caption{
-  \textbf{Adaptation by migration:}
+  \textbf{Simulation of adaptation by migration.}
   An initially unadapted patch (10 demes at the left end of the range) is colonized around generation 5000;
@@
   between the patches and coalesces with a lineage in the unadapted patch.
-  In the left panel, as in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
+  In the left panel, as in Fig~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
   other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
@@
 in which case the haplotype has literally hitchhiked across geography!
-Figure~\ref{fig:lineagesmotion} shows a simulation of such an event,
+Fig~\ref{fig:lineagesmotion} shows a simulation of such an event,
 including the lineage that founds an adaptive population on a second patch,
 and the length of the haplotype shared between this lineage and one in the original patch.
-The time that the lineage is outside the region where the $B$ allele is common (dark red in Figure~\ref{fig:lineagesmotion}), 
+The time that the lineage is outside the region where the $B$ allele is common (dark red in Fig~\ref{fig:lineagesmotion}), 
 the haplotype that accompanies it is broken down rapidly. 
@@
   \begin{center}
-    \includegraphics{Lava_flow_mice_prob_parallel}
+    % \includegraphics{Lava_flow_mice_prob_parallel}
   \end{center}
   \caption{
+      \textbf{Results for the rock pocket mouse.}
 Top panel: The probability of parallel mutation as a function of the distance
@@
 patch). In
-Figure~\ref{fig:mice_prob_parallel} we show the dependence of the probability
+Fig~\ref{fig:mice_prob_parallel} we show the dependence of the probability
 of parallel mutation on the distance between lava flow patches using these parameters, 
@@
 either a new mutation or migrant allele to be large. 
-Figure~\ref{fig:mice_prob_parallel} also shows how long the founding haplotype 
+Fig~\ref{fig:mice_prob_parallel} also shows how long the founding haplotype 
 shared between populations is expected to be,
@@
 and where it has a linear transition.
-These solutions are shown, and parameters described, in figure~\ref{fig:prob_estab_calcs}.
+These solutions are shown, and parameters described, in Fig~\ref{fig:prob_estab_calcs}.
 
@@
     \end{center}
-    \caption{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
-    in an abrupt transition (left) and a gradual transition (right).
+    \caption{
+        \textbf{The probability of establishment of a new mutant allele as a function of distance from the edge of the region where it is beneficial,
+        in an abrupt transition (left) and a gradual transition (right).}
     The allele is deleterious to the left and beneficial to the right (with selection coefficients $\mp 0.05$ respectively);
@@
 by following only the motion of the ``trunk''
-(i.e., the red line in figure \ref{fig:branching_decomp}),
+(i.e., the red line in Fig \ref{fig:branching_decomp}),
 and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
@@
 as the sum of independent migrant families, each of whose dynamics are
-given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
+given by a spatial branching process (as depicted in Fig \ref{fig:branching_decomp}).
 Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
@@
 
-\section*{Acknowledgements}
-We would like to heartily thank Joachim Hermisson, Richard Neher, and an anonymous reviewer
-for a number of insightful suggestions,
-and catching several errors present in earlier drafts. \revpoint{3}{4}
-We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
-as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
-for helpful comments and discussion of the manuscript.
-This work was supported by grants from the
-National Science Foundation under Grant No.\ 1262645 to PLR and GC and by the
-National Institute of General Medical Sciences of the National
-Institutes of Health under award numbers NIH RO1GM83098 and
-RO1GM107374 to GC.
-
-\bibliography{standing_patches_refs}
-
-\ifsubmissionversion
-\clearpage
-\renewcommand*\listfigurename{Figure Legends}
-\listoffigures
-\fi
-
-\clearpage
-
-% endfloat uses different names
-\ifsubmissionversion
-  \renewcommand{\thesfigure}{S\arabic{sfigure}}
-  \renewcommand{\thepostsfigure}{S\arabic{postsfigure}}
-  % \renewcommand{\thefigure}{S\arabic{figure}}
-  % \renewcommand{\thepostfigure}{S\arabic{postfig}}
-  % \setcounter{postfig}{0}
-  \renewcommand{\thetable}{S\arabic{table}}
-  \renewcommand{\theposttable}{S\arabic{posttable}}
-  \setcounter{posttable}{0}
-\fi
-
-%% list of supplementary figures
-\ifsubmissionversion
-  \renewcommand*\listsfigurename{Supplementary Figure Legends}
-  \listofsfigures
-  \renewcommand*\listtablename{Supplementary Table Legends}
-  \listoftables
-  \pagebreak
-\fi
-
-\appendix
+% \ifsubmissionversion
+% \clearpage
+% \renewcommand*\listfigurename{Figure Legends}
+% \listoffigures
+% \fi
+% 
+% \clearpage
+% 
+% % endfloat uses different names
+% \renewcommand{\thesfigure}{S\arabic{sfigure}}
+% \renewcommand{\thetable}{S\arabic{table}}
+% \ifsubmissionversion
+%   \renewcommand{\thepostsfigure}{S\arabic{postsfigure}}
+%   % \renewcommand{\thefigure}{S\arabic{figure}}
+%   % \renewcommand{\thepostfigure}{S\arabic{postfig}}
+%   % \setcounter{postfig}{0}
+%   \renewcommand{\theposttable}{S\arabic{posttable}}
+%   \setcounter{posttable}{0}
+% \fi
+% 
+% %% list of supplementary figures
+% \ifsubmissionversion
+%   \renewcommand*\listsfigurename{Supplementary Figure Legends}
+%   \listofsfigures
+%   \renewcommand*\listtablename{Supplementary Table Legends}
+%   \listoftables
+%   \pagebreak
+% \fi
+% 
+% \appendix
+% \section*{Supporting Information}
+% 
+% \setcounter{secnumdepth}{2}
+
 \section*{Supporting Information}
 
-\setcounter{secnumdepth}{2}
-
 %%%%%%%
-\subsection*{The exact solution to the equilibrium frequency}
+\subsection*{S1 Text}
+% \subsection*{The exact solution to the equilibrium frequency}
 \label{apx:elliptic_integrals}
 
+\textbf{
 Here we describe how to solve \eqref{eqn:radial_eqfreq_diffeq} exactly in one dimension,
 i.e., the Fisher-KPP equation with piecewise constant selection.
+}
 Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
@@
 
-\subsection*{Doing the integrals}
+\paragraph{Doing the integrals}
 \label{apx:some_integrals}
@@
 
-% Presumably, this gets the same answer as \citep{slatkin1973geneflow}, who finds that in the solvable cases,
-% for $x$ in the last interval (which ends in $\infty$), that
-% \begin{align*}
-%     \phi(x) &= \frac{3}{2}\left( 1 - \tanh^2 ( x \sqrt{2s} / 2 \sigma + C ) \right) \\
-%         &\simeq C' \exp\left( - \frac{ x \sqrt{2s} }{ \sigma } \right) .
-% \end{align*}
-
-%%%%%%%
-\subsection*{More information about the simulations}
-\label{apx:parameter_tables}
-
-Here we provide additional details regarding the simulations we used
-to validate the theory in figures \ref{fig:sim_times} and \ref{fig:sim_probs}.
+% %%%%%%%
+% \subsection*{More information about the simulations}
+% \label{apx:parameter_tables}
+% 
+% In tables \ref{sfig:sim_migration_times}
+% and \ref{sfig:sim_mutation_times}
+% we provide additional details regarding the simulations we used
+% to validate the theory in \ref{fig:sim_times} and \ref{fig:sim_probs} Fig.
 
@@
 
+\subsection*{S1 Table}
+\label{stab:mutation_params}
+
 % table of parameters, produced by analyze-sims.R
-\input{mutation-parameters-table.tex}
+%\input{mutation-parameters-table.tex}
+
+\textbf{Parameter values used in estimates of mean time to adaptation by mutation of figure~\ref{fig:sim_times}.}
+All simulations also used a linear grid of 501 demes with a patch of 99 demes in the center, 
+the migration model described in \nameref{ss:simulations},
+$\mu=10^{-5}$, and $s_p=.0023$ (calculated as the growth rate as described in the text).
+$T_\text{adapt}$ is the mean time until 100 $B$ alleles were present in the patch,
+and $p_\text{adapted}$ is the proportion of the simulations that adapted by the 25,000 generations.
+The simulation began with no $B$ alleles.
+
+\subsection*{S1 Table}
+\label{stab:migration_params}
 
 % table of parameters, produced by analyze-sims.R
-\input{migration-parameters-table.tex}
+%\input{migration-parameters-table.tex}
+
+\textbf{Parameter values used in estimates of mean time to adaptation by migration of figure~\ref{fig:sim_times}.}
+All simulations also used a linear grid of 501 demes with two patches of 99 demes each,
+separated by $R$ demes in the center;
+the migration model described in \nameref{ss:simulations},
+$\mu=10^{-5}$, and $s_p=.0023$ (calculated as the growth rate as described in the text).
+$T_\text{adapt}$ is the mean time until 100 $B$ alleles were present in the patch,
+and $p_\text{adapted}$ is the proportion of the simulations that adapted by the 25,000 generations.
+At the start of the simulation, one patch was initialized with a frequency of 0.8 $B$ alleles, which were absent elsewhere.
 
@@
 
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{mutation-times-predicted}
-  \end{center}
-  \caption{
-    The same data shown in the left panel of figure~\ref{fig:sim_times},
+\subsection*{S1 Fig}
+\label{sfig:sim_migration_times}
+
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{mutation-times-predicted}
+%   \end{center}
+%   \caption{
+      \textbf{Observed versus predicted times to adaptation by mutation, 
+      including parameters where the model is not predicted to hold.}
+    The same data shown in the left panel of Fig~\ref{fig:sim_times},
     but all times shown (not just the interquartile ranges),
     and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
-  } \label{sfig:sim_migration_times}
-\end{sfigure}
-
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{migration-time-predicted}
-  \end{center}
-  \caption{
-    The same data shown in the right panel of figure~\ref{fig:sim_times},
+%   } \label{sfig:sim_migration_times}
+% \end{sfigure}
+
+\subsection*{S2 Fig}
+\label{sfig:sim_mutation_times}
+
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{migration-time-predicted}
+%   \end{center}
+%   \caption{
+      \textbf{Observed versus predicted times to adaptation by migration, 
+      including parameters where the model is not predicted to hold.}
+    The same data shown in the right panel of Fig~\ref{fig:sim_times},
     but all times shown (not just the interquartile ranges),
     and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
-    The upper panel has the predicted time to adaptation on the horizontal axis as in figure~\ref{fig:sim_times},
+    The upper panel has the predicted time to adaptation on the horizontal axis as in Fig~\ref{fig:sim_times},
     and the lower panel has, for comparison, the raw distance between patches 
     (which predicts time to adaptation, but not as well).
-  } \label{sfig:sim_mutation_times}
-\end{sfigure}
+%   } \label{sfig:sim_mutation_times}
+% \end{sfigure}
 
@@
 
-% sm=0.1
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/18885-r1-501-sb0_01-sm-0_1-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/56325-r1-501-sb0_01-sm-0_1-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/28432-r1-501-sb0_01-sm-0_1-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations of adaptation by new mutation
-    with $s_m=0.1$, $\sigma\approx 1$, and $\rho$ varying.
+\subsection*{S3 Fig}
+\label{sfig:sims_1}
+
+% % sm=0.1
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{example-mutation-sims/18885-r1-501-sb0_01-sm-0_1-N1200-pophistory-run}
+%     % \includegraphics{example-mutation-sims/56325-r1-501-sb0_01-sm-0_1-N600-pophistory-run}
+%     % \includegraphics{example-mutation-sims/28432-r1-501-sb0_01-sm-0_1-N50-pophistory-run}
+%   \end{center}
+%   \caption{
+\textbf{Randomly chosen simulations of adaptation by new mutation
+with $s_m=0.1$, $\sigma\approx 1$, and $\rho$ varying.}
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
     dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_1}
-\end{sfigure}
-
-% sm=0.01
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/43099-r1-501-sb0_01-sm-0_01-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/69787-r1-501-sb0_01-sm-0_01-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/11821-r1-501-sb0_01-sm-0_01-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations of adaptation by new mutation
-    with $s_m=0.01$, $\sigma\approx 1$, and $\rho$ varying.
+%   } \label{sfig:sims_1}
+% \end{sfigure}
+
+\subsection*{S4 Fig}
+\label{sfig:sims_2}
+
+% % sm=0.01
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{example-mutation-sims/43099-r1-501-sb0_01-sm-0_01-N1200-pophistory-run}
+%     % \includegraphics{example-mutation-sims/69787-r1-501-sb0_01-sm-0_01-N600-pophistory-run}
+%     % \includegraphics{example-mutation-sims/11821-r1-501-sb0_01-sm-0_01-N50-pophistory-run}
+%   \end{center}
+%   \caption{
+\textbf{Randomly chosen simulations of adaptation by new mutation
+with $s_m=0.01$, $\sigma\approx 1$, and $\rho$ varying.}
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
     dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_2}
-\end{sfigure}
-
-% sm=0.001
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/59611-r1-501-sb0_01-sm-0_001-N1200-pophistory-run}
-    \includegraphics{example-mutation-sims/93713-r1-501-sb0_01-sm-0_001-N600-pophistory-run}
-    \includegraphics{example-mutation-sims/29850-r1-501-sb0_01-sm-0_001-N50-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations of adaptation by new mutation
-    with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
+%   } \label{sfig:sims_2}
+% \end{sfigure}
+
+\subsection*{S5 Fig}
+\label{sfig:sims_3}
+
+% % sm=0.001
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{example-mutation-sims/59611-r1-501-sb0_01-sm-0_001-N1200-pophistory-run}
+%     % \includegraphics{example-mutation-sims/93713-r1-501-sb0_01-sm-0_001-N600-pophistory-run}
+%     % \includegraphics{example-mutation-sims/29850-r1-501-sb0_01-sm-0_001-N50-pophistory-run}
+%   \end{center}
+%   \caption{
+\textbf{Randomly chosen simulations of adaptation by new mutation
+with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.}
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
     dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_3}
-\end{sfigure}
-
-% sm=0.0001
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{example-mutation-sims/15449-r1-501-sb0_01-sm-1e-04-N1600-pophistory-run}
-    \includegraphics{example-mutation-sims/5582-r1-501-sb0_01-sm-1e-04-N400-pophistory-run}
-    \includegraphics{example-mutation-sims/24639-r1-501-sb0_01-sm-1e-04-N25-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations of adaptation by new mutation
-    with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.
+%   } \label{sfig:sims_3}
+% \end{sfigure}
+
+\subsection*{S6 Fig}
+\label{sfig:sims_4}
+
+% % sm=0.0001
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{example-mutation-sims/15449-r1-501-sb0_01-sm-1e-04-N1600-pophistory-run}
+%     % \includegraphics{example-mutation-sims/5582-r1-501-sb0_01-sm-1e-04-N400-pophistory-run}
+%     % \includegraphics{example-mutation-sims/24639-r1-501-sb0_01-sm-1e-04-N25-pophistory-run}
+%   \end{center}
+%   \caption{
+\textbf{Randomly chosen simulations of adaptation by new mutation
+with $s_m=0.001$, $\sigma\approx 1$, and $\rho$ varying.}
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
     dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_4}
-\end{sfigure}
-
-% sm=0.1, R=80
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{example-migration-sims/89826-r1-501-sb0_01-sm-0_1-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/76178-r1-501-sb0_01-sm-0_1-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/95095-r1-501-sb0_01-sm-0_1-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations of adaptation by migration
-    with $s_m=0.1$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+%   } \label{sfig:sims_4}
+% \end{sfigure}
+
+\subsection*{S7 Fig}
+\label{sfig:sims_5}
+
+% % sm=0.1, R=80
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{example-migration-sims/89826-r1-501-sb0_01-sm-0_1-N4000-pophistory-run}
+%     % \includegraphics{example-migration-sims/76178-r1-501-sb0_01-sm-0_1-N1000-pophistory-run}
+%     % \includegraphics{example-migration-sims/95095-r1-501-sb0_01-sm-0_1-N100-pophistory-run}
+%   \end{center}
+%   \caption{
+\textbf{Randomly chosen simulations of adaptation by migration
+with $s_m=0.1$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.}
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
     dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_5}
-\end{sfigure}
-
-% sm=0.01, R=80
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{example-migration-sims/42080-r1-501-sb0_01-sm-0_01-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/82698-r1-501-sb0_01-sm-0_01-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations of adaptation by migration
-    with $s_m=0.01$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+%   } \label{sfig:sims_5}
+% \end{sfigure}
+
+\subsection*{S8 Fig}
+\label{sfig:sims_6}
+
+% % sm=0.01, R=80
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{example-migration-sims/42080-r1-501-sb0_01-sm-0_01-N4000-pophistory-run}
+%     % \includegraphics{example-migration-sims/82698-r1-501-sb0_01-sm-0_01-N1000-pophistory-run}
+%     % \includegraphics{example-migration-sims/97545-r1-501-sb0_01-sm-0_01-N100-pophistory-run}
+%   \end{center}
+%   \caption{
+\textbf{Randomly chosen simulations of adaptation by migration
+with $s_m=0.01$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.}
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
     dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_6}
-\end{sfigure}
-
-% sm=0.001, R=80
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{example-migration-sims/88079-r1-501-sb0_01-sm-0_001-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/3464-r1-501-sb0_01-sm-0_001-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/37774-r1-501-sb0_01-sm-0_001-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations of adaptation by migration
-    with $s_m=0.001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+%   } \label{sfig:sims_6}
+% \end{sfigure}
+
+\subsection*{S9 Fig}
+\label{sfig:sims_7}
+
+% % sm=0.001, R=80
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{example-migration-sims/88079-r1-501-sb0_01-sm-0_001-N4000-pophistory-run}
+%     % \includegraphics{example-migration-sims/3464-r1-501-sb0_01-sm-0_001-N1000-pophistory-run}
+%     % \includegraphics{example-migration-sims/37774-r1-501-sb0_01-sm-0_001-N100-pophistory-run}
+%   \end{center}
+%   \caption{
+\textbf{Randomly chosen simulations of adaptation by migration
+with $s_m=0.001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.}
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
     dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_7}
-\end{sfigure}
-
-% sm=0.0001, R=80
-\begin{sfigure}
-  \begin{center}
-    \includegraphics{example-migration-sims/26044-r1-501-sb0_01-sm-1e-04-N4000-pophistory-run}
-    \includegraphics{example-migration-sims/85750-r1-501-sb0_01-sm-1e-04-N1000-pophistory-run}
-    \includegraphics{example-migration-sims/4111-r1-501-sb0_01-sm-1e-04-N100-pophistory-run}
-  \end{center}
-  \caption{
-    Randomly chosen simulations of adaptation by migration
-    with $s_m=0.0001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.
+%   } \label{sfig:sims_7}
+% \end{sfigure}
+
+\subsection*{S10 Fig}
+\label{sfig:sims_8}
+
+% % sm=0.0001, R=80
+% \begin{sfigure}
+%   \begin{center}
+%     % \includegraphics{example-migration-sims/26044-r1-501-sb0_01-sm-1e-04-N4000-pophistory-run}
+%     % \includegraphics{example-migration-sims/85750-r1-501-sb0_01-sm-1e-04-N1000-pophistory-run}
+%     % \includegraphics{example-migration-sims/4111-r1-501-sb0_01-sm-1e-04-N100-pophistory-run}
+%   \end{center}
+%   \caption{
+\textbf{Randomly chosen simulations of adaptation by migration
+with $s_m=0.0001$, $R=80$, $\sigma\approx 1$, and $\rho$ varying.}
     On the left of each is a space-time heatmap of the local frequency of $B$ alleles;
@@
     dotted black lines indicate the patches where $B$ is advantageous.
-  } \label{sfig:sims_8}
-\end{sfigure}
-
-%%%% Reviews
-\includereviews
+%   } \label{sfig:sims_8}
+% \end{sfigure}
+
+\section*{Acknowledgements}
+We would like to heartily thank Joachim Hermisson, Richard Neher, and an anonymous reviewer
+for a number of insightful suggestions,
+and catching several errors present in earlier drafts. \revpoint{3}{4}
+We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 
+as well as Gideon Bradburd, and Simon Aeschbacher and the rest of the Coop lab
+for helpful comments and discussion of the manuscript.
+
+\nolinenumbers
+
+% \bibliography{standing_patches_refs}
+\begin{thebibliography}{68}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi: #1}\else
+  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+
+\bibitem[Aldous(1985)]{aldous1985exchangeability}
+D.~J. Aldous.
+\newblock Exchangeability and related topics.
+\newblock In \emph{École d'été de probabilités de {S}aint-{F}lour,
+  {XIII}---1983}, volume 1117 of \emph{Lecture Notes in Math.}, pages 1--198.
+  Springer, Berlin, 1985.
+\newblock URL
+  \url{http://www.springerlink.com/content/c31v17440871210x/fulltext.pdf}.
+
+\bibitem[Allred and Beck(1963)]{allred1963range}
+D.~M. Allred and D.~E. Beck.
+\newblock Range of movement and dispersal of some rodents at the {Nevada}
+  atomic test site.
+\newblock \emph{Journal of Mammalogy}, 44\penalty0 (2):\penalty0 pp. 190--200,
+  1963.
+\newblock ISSN 00222372.
+\newblock URL \url{http://www.jstor.org/stable/1377452}.
+
+\bibitem[Arendt and Reznick(2008)]{Arendt:08}
+J.~Arendt and D.~Reznick.
+\newblock Convergence and parallelism reconsidered: what have we learned about
+  the genetics of adaptation?
+\newblock \emph{Trends Ecol Evol}, 23\penalty0 (1):\penalty0 26--32, Jan. 2008.
+\newblock \doi{10.1016/j.tree.2007.09.011}.
+
+\bibitem[Barton(1979)]{barton1979geneflow}
+N.~Barton.
+\newblock Gene flow past a cline.
+\newblock \emph{Heredity}, 43\penalty0 (3):\penalty0 333--339, Dec. 1979.
+\newblock ISSN 0018-067X.
+\newblock URL \url{http://dx.doi.org/10.1038/hdy.1979.86}.
+
+\bibitem[Barton(1987)]{barton1987establishment}
+N.~H. Barton.
+\newblock The probability of establishment of an advantageous mutant in a
+  subdivided population.
+\newblock \emph{Genetics Research}, 50\penalty0 (1):\penalty0 35--40, 1987.
+\newblock \doi{10.1017/S0016672300023314}.
+\newblock URL \url{http://dx.doi.org/10.1017/S0016672300023314}.
+
+\bibitem[Barton(2000)]{barton2000genetic}
+N.~H. Barton.
+\newblock Genetic hitchhiking.
+\newblock \emph{Philos Trans R Soc Lond B Biol Sci}, 355\penalty0
+  (1403):\penalty0 1553--1562, Nov. 2000.
+\newblock \doi{10.1098/rstb.2000.0716}.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC1692896/}.
+
+\bibitem[Barton et~al.(2013)Barton, Etheridge, Kelleher, and
+  Véber]{barton2013genetic}
+N.~H. Barton, A.~M. Etheridge, J.~Kelleher, and A.~Véber.
+\newblock Genetic hitchhiking in spatially extended populations.
+\newblock \emph{Theoretical Population Biology}, \penalty0 (0):\penalty0 --,
+  2013.
+\newblock ISSN 0040-5809.
+\newblock \doi{10.1016/j.tpb.2012.12.001}.
+\newblock URL
+  \url{http://www.sciencedirect.com/science/article/pii/S0040580912001359}.
+
+\bibitem[Benson(1933)]{benson1933concealing}
+S.~B. Benson.
+\newblock \emph{Concealing coloration among some desert rodents of the
+  southwestern United States}.
+\newblock Number v. 40 in University of California publications in zoology.
+  University of California Press, 1933.
+\newblock URL \url{http://books.google.com/books?id=Lis-AQAAIAAJ}.
+
+\bibitem[Borodin and Salminen(2002)]{borodin2002handbook}
+A.~N. Borodin and P.~Salminen.
+\newblock \emph{Handbook of Brownian motion: facts and formulae}.
+\newblock Springer, 2002.
+
+\bibitem[Cantrell and Cosner(1989)]{cantrell1989diffusive}
+R.~S. Cantrell and C.~Cosner.
+\newblock Diffusive logistic equations with indefinite weights: population
+  models in disrupted environments.
+\newblock \emph{Proceedings of the Royal Society of Edinburgh, Section: A
+  Mathematics}, 112\penalty0 (3-4):\penalty0 293--318, 0 1989.
+\newblock ISSN 1473-7124.
+\newblock \doi{10.1017/S030821050001876X}.
+\newblock URL \url{http://journals.cambridge.org/article_S030821050001876X}.
+
+\bibitem[Cantrell and Cosner(1991)]{cantrell1991diffusive}
+R.~S. Cantrell and C.~Cosner.
+\newblock Diffusive logistic equations with indefinite weights: population
+  models in disrupted environments. {II}.
+\newblock \emph{SIAM J. Math. Anal.}, 22\penalty0 (4):\penalty0 1043--1064,
+  1991.
+\newblock ISSN 0036-1410.
+\newblock \doi{10.1137/0522068}.
+\newblock URL \url{http://dx.doi.org/10.1137/0522068}.
+
+\bibitem[Censky et~al.(1998)Censky, Hodge, and Dudley]{censky1998overwater}
+E.~J. Censky, K.~Hodge, and J.~Dudley.
+\newblock Over-water dispersal of lizards due to hurricanes.
+\newblock \emph{Nature}, 395\penalty0 (6702):\penalty0 556--556, Oct. 1998.
+\newblock ISSN 00280836.
+\newblock URL \url{http://dx.doi.org/10.1038/26886}.
+
+\bibitem[Cherry and Wakeley(2003)]{cherry2003diffusion}
+J.~L. Cherry and J.~Wakeley.
+\newblock A diffusion approximation for selection and drift in a subdivided
+  population.
+\newblock \emph{Genetics}, 163\penalty0 (1):\penalty0 421--428, Jan. 2003.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/12586727}.
+
+\bibitem[Conley(1975)]{conley1975application}
+C.~Conley.
+\newblock An application of {Wazewski}'s method to a non-linear boundary value
+  problem which arises in population genetics.
+\newblock \emph{Journal of Mathematical Biology}, 2\penalty0 (3):\penalty0
+  241--249, 1975.
+\newblock ISSN 0303-6812.
+\newblock \doi{10.1007/BF00277153}.
+\newblock URL \url{http://dx.doi.org/10.1007/BF00277153}.
+
+\bibitem[Dawson(1993)]{dawson1993measurevalued}
+D.~A. Dawson.
+\newblock Measure-valued {Markov} processes.
+\newblock In \emph{École d'Été de Probabilités de Saint-Flour XXI---1991},
+  volume 1541 of \emph{Lecture Notes in Math.}, pages 1--260. Springer, Berlin,
+  1993.
+
+\bibitem[Dice(1940)]{dice1940ecologic}
+L.~R. Dice.
+\newblock Ecologic and genetic variability within species of
+  \textit{{Peromyscus}}.
+\newblock \emph{The American Naturalist}, 74\penalty0 (752):\penalty0 pp.
+  212--221, 1940.
+\newblock ISSN 00030147.
+\newblock URL \url{http://www.jstor.org/stable/2457573}.
+
+\bibitem[Dice and Blossom(1937)]{DiceBlossom1937}
+L.~R. Dice and P.~M. Blossom.
+\newblock \emph{Studies of Mammalian Ecology in Southwestern North America With
+  Special Attention to the Colors of Desert Mammals}.
+\newblock Carnegie Institution, 1937.
+
+\bibitem[Donnelly and Joyce(1989)]{donnelly1989continuity}
+P.~Donnelly and P.~Joyce.
+\newblock Continuity and weak convergence of ranked and size-biased
+  permutations on the infinite simplex.
+\newblock \emph{Stochastic Process. Appl.}, 31\penalty0 (1):\penalty0 89--103,
+  1989.
+\newblock ISSN 0304-4149.
+\newblock URL
+  \url{http://www.sciencedirect.com/science/article/pii/030441498990104X}.
+
+\bibitem[Etheridge(2000)]{etheridge2000introduction}
+A.~M. Etheridge.
+\newblock \emph{An introduction to superprocesses}, volume~20 of
+  \emph{University Lecture Series}.
+\newblock American Mathematical Society, Providence, RI, 2000.
+\newblock ISBN 0-8218-2706-5.
+
+\bibitem[Fisher(1937)]{fisher1937wave}
+R.~Fisher.
+\newblock {The wave of advance of advantageous genes}.
+\newblock \emph{Ann. Eugenics}, 7:\penalty0 353--369, 1937.
+\newblock URL
+  \url{http://digital.library.adelaide.edu.au/coll/special/fisher/152.pdf}.
+
+\bibitem[Fisher(1930)]{fisher1930genetical}
+R.~A. Fisher.
+\newblock \emph{The genetical theory of natural selection}.
+\newblock Oxford University Press, Oxford, 1930.
+\newblock ISBN 0-19-850440-3.
+\newblock URL \url{http://www.archive.org/details/geneticaltheoryo031631mbp}.
+
+\bibitem[Fisher(1950)]{fisher1950frequencies}
+R.~A. Fisher.
+\newblock Gene frequencies in a cline determined by selection and diffusion.
+\newblock \emph{Biometrics}, 6\penalty0 (4):\penalty0 pp. 353--361, 1950.
+\newblock ISSN 0006341X.
+\newblock URL \url{http://www.jstor.org/stable/3001780}.
+
+\bibitem[French et~al.(1968)French, Tagami, and Hayden]{french1968dispersal}
+N.~R. French, T.~Y. Tagami, and P.~Hayden.
+\newblock Dispersal in a population of desert rodents.
+\newblock \emph{Journal of Mammalogy}, 49\penalty0 (2):\penalty0 pp. 272--280,
+  1968.
+\newblock ISSN 00222372.
+\newblock URL \url{http://www.jstor.org/stable/1377984}.
+
+\bibitem[Geiger(1999)]{geiger1999elementary}
+J.~Geiger.
+\newblock Elementary new proofs of classical limit theorems for
+  {Galton}-{Watson} processes.
+\newblock \emph{Journal of Applied Probability}, 36\penalty0 (2):\penalty0 pp.
+  301--309, 1999.
+\newblock URL \url{http://www.jstor.org/stable/3215457}.
+
+\bibitem[Gradshteyn and Ryzhik(2007)]{gradshteyn2007table}
+I.~S. Gradshteyn and I.~M. Ryzhik.
+\newblock \emph{Table of integrals, series, and products}.
+\newblock Elsevier/Academic Press, Amsterdam, seventh edition, 2007.
+\newblock ISBN 978-0-12-373637-6; 0-12-373637-4.
+\newblock Translated from the Russian, Translation edited and with a preface by
+  Alan Jeffrey and Daniel Zwillinger.
+
+\bibitem[Haldane(1927)]{haldane1927mathematical}
+J.~B.~S. Haldane.
+\newblock A mathematical theory of natural and artificial selection, part {V}:
+  Selection and mutation.
+\newblock \emph{Mathematical Proceedings of the Cambridge Philosophical
+  Society}, 23\penalty0 (07):\penalty0 838--844, 7 1927.
+\newblock ISSN 1469-8064.
+\newblock \doi{10.1017/S0305004100015644}.
+\newblock URL \url{http://journals.cambridge.org/article_S0305004100015644}.
+
+\bibitem[Haldane(1948)]{haldane1948theory}
+J.~B.~S. Haldane.
+\newblock The theory of a cline.
+\newblock \emph{J Genet}, 48\penalty0 (3):\penalty0 277--284, Jan. 1948.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/18905075}.
+
+\bibitem[Hallatschek and Fisher(2014)]{hallatschek2014acceleration}
+O.~Hallatschek and D.~S. Fisher.
+\newblock Acceleration of evolutionary spread by long-range dispersal.
+\newblock \emph{Proc Natl Acad Sci U S A}, 111\penalty0 (46):\penalty0
+  4911--4919, Nov. 2014.
+\newblock \doi{10.1073/pnas.1404663111}.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/25368183}.
+
+\bibitem[Hermisson and Pennings(2005)]{softsweepsI}
+J.~Hermisson and P.~S. Pennings.
+\newblock Soft sweeps: molecular population genetics of adaptation from
+  standing genetic variation.
+\newblock \emph{Genetics}, 169\penalty0 (4):\penalty0 2335--2352, Apr. 2005.
+\newblock \doi{10.1534/genetics.104.036947}.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/15716498}.
+
+\bibitem[Hoekstra(2006)]{Hoekstra:06}
+H.~E. Hoekstra.
+\newblock {Genetics, development and evolution of adaptive pigmentation in
+  vertebrates}.
+\newblock \emph{Heredity (Edinb)}, 97\penalty0 (3):\penalty0 222--234, Sep
+  2006.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/16823403}.
+
+\bibitem[Hoekstra and Nachman(2003)]{hoekstra2003different}
+H.~E. Hoekstra and M.~W. Nachman.
+\newblock Different genes underlie adaptive melanism in different populations
+  of rock pocket mice.
+\newblock \emph{Mol. Ecol.}, 12:\penalty0 1185--1194, May 2003.
+\newblock URL
+  \url{http://www3.interscience.wiley.com/journal/118890526/abstract}.
+
+\bibitem[Hoekstra et~al.(2004)Hoekstra, Drumm, and
+  Nachman]{hoekstra2004ecological}
+H.~E. Hoekstra, K.~E. Drumm, and M.~W. Nachman.
+\newblock Ecological genetics of adaptive color polymorphism in pocket mice:
+  geographic variation in selected and neutral genes.
+\newblock \emph{Evolution}, 58\penalty0 (6):\penalty0 1329--1341, June 2004.
+\newblock URL
+  \url{http://onlinelibrary.wiley.com/doi/10.1111/j.0014-3820.2004.tb01711.x/abstract}.
+
+\bibitem[Hoekstra et~al.(2005)Hoekstra, Krenz, and Nachman]{hoekstra2004local}
+H.~E. Hoekstra, J.~G. Krenz, and M.~W. Nachman.
+\newblock Local adaptation in the rock pocket mouse (\textit{{Chaetodipus}
+  intermedius}): natural selection and phylogenetic history of populations.
+\newblock \emph{Heredity}, 94\penalty0 (2):\penalty0 217--228, Nov. 2005.
+\newblock ISSN 0018067X.
+\newblock URL \url{http://dx.doi.org/10.1038/sj.hdy.6800600}.
+
+\bibitem[Jagers(1975)]{jagers1975branching}
+P.~Jagers.
+\newblock \emph{Branching processes with biological applications}.
+\newblock Wiley Series in Probability and Statistics: Applied Probability and
+  Statistics Section Series. Wiley, 1975.
+\newblock ISBN 9780471436522.
+
+\bibitem[Kaufman(1974)]{kaufman1974adaptive}
+D.~W. Kaufman.
+\newblock Adaptive coloration in \textit{{Peromyscus} polionotus}: Experimental
+  selection by owls.
+\newblock \emph{Journal of Mammalogy}, 55\penalty0 (2):\penalty0 pp. 271--283,
+  1974.
+\newblock ISSN 00222372.
+\newblock URL \url{http://www.jstor.org/stable/1378997}.
+
+\bibitem[Kim and Maruki(2011)]{kim2011hitchhiking}
+Y.~Kim and T.~Maruki.
+\newblock Hitchhiking effect of a beneficial mutation spreading in a subdivided
+  population.
+\newblock \emph{Genetics}, 189\penalty0 (1):\penalty0 213--226, Sept. 2011.
+\newblock \doi{10.1534/genetics.111.130203}.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3176130/}.
+
+\bibitem[Kingsley et~al.(2009)Kingsley, Manceau, Wiley, and
+  Hoekstra]{kingsley2009melanism}
+E.~P. Kingsley, M.~Manceau, C.~D. Wiley, and H.~E. Hoekstra.
+\newblock Melanism in \textit{{Peromyscus}} is caused by independent mutations
+  in agouti.
+\newblock \emph{PLoS ONE}, 4:\penalty0 e6435, 2009.
+\newblock URL \url{http://dx.doi.org/10.1371%2Fjournal.pone.0006435}.
+
+\bibitem[Kolmogorov et~al.(1991)Kolmogorov, Petrovskii, and
+  Piscunov]{KPP1937book}
+A.~Kolmogorov, I.~Petrovskii, and N.~Piscunov.
+\newblock A study of the equation of diffusion with increase in the quantity of
+  matter, and its application to a biological problem.
+\newblock In \emph{Selected Works of A.N. Kolmogorov: Mathematics and
+  mechanics}, volume~25 of \emph{Mathematics and its Applications (Soviet
+  Series)}, pages 1--25. Kluwer Academic Publishers Group, Dordrecht, 1991.
+
+\bibitem[Kondrashov(2003)]{Kondrashov:03}
+A.~S. Kondrashov.
+\newblock {Accumulation of {Dobzhansky}--{Muller} incompatibilities within a
+  spatially structured population}.
+\newblock \emph{Evolution}, 57\penalty0 (1):\penalty0 151--153, Jan 2003.
+
+\bibitem[Kruckeberg(1951)]{kruckeberg1951intraspecific}
+A.~R. Kruckeberg.
+\newblock Intraspecific variability in the response of certain native plant
+  species to serpentine soil.
+\newblock \emph{American Journal of Botany}, 38\penalty0 (6):\penalty0 pp.
+  408--419, 1951.
+\newblock URL \url{http://www.jstor.org/stable/2438248}.
+
+\bibitem[Lenormand(2002)]{lenormand2002limits}
+T.~Lenormand.
+\newblock Gene flow and the limits to natural selection.
+\newblock \emph{Trends in Ecology \& Evolution}, 17\penalty0 (4):\penalty0 183
+  -- 189, 2002.
+\newblock ISSN 0169-5347.
+\newblock \doi{DOI: 10.1016/S0169-5347(02)02497-7}.
+\newblock URL
+  \url{http://www.sciencedirect.com/science/article/pii/S0169534702024977}.
+
+\bibitem[Levin et~al.(2003)Levin, Muller-Landau, Nathan, and
+  Chave]{levin2003ecology}
+S.~A. Levin, H.~C. Muller-Landau, R.~Nathan, and J.~Chave.
+\newblock The ecology and evolution of seed dispersal: a theoretical
+  perspective.
+\newblock \emph{Annu. Rev. Ecol. Evol. Syst.}, 34:\penalty0 575--604, 2003.
+\newblock URL
+  \url{http://arjournals.annualreviews.org/doi/pdf/10.1146/annurev.ecolsys.34.011802.132428}.
+
+\bibitem[Lou and Yanagida(2006)]{lou2006minimization}
+Y.~Lou and E.~Yanagida.
+\newblock Minimization of the principal eigenvalue for an elliptic boundary
+  value problem with indefinite weight, and applications to population
+  dynamics.
+\newblock \emph{Japan J. Indust. Appl. Math.}, 23\penalty0 (3):\penalty0
+  275--292, 2006.
+\newblock ISSN 0916-7005.
+\newblock URL \url{http://projecteuclid.org/euclid.jjiam/1197390801}.
+
+\bibitem[Macnair(1991)]{macnair1991evolution}
+M.~R. Macnair.
+\newblock Why the evolution of resistance to anthropogenic toxins normally
+  involves major gene changes: the limits to natural selection.
+\newblock \emph{Genetica}, 84\penalty0 (3):\penalty0 213--219, 1991.
+\newblock ISSN 0016-6707.
+\newblock \doi{10.1007/BF00127250}.
+\newblock URL \url{http://dx.doi.org/10.1007/BF00127250}.
+
+\bibitem[Martin and Orgogozo(2013)]{martin2013repeated}
+A.~Martin and V.~Orgogozo.
+\newblock The loci of repeated evolution: a catalog of genetic hotspots of
+  phenotypic variation.
+\newblock \emph{Evolution}, 67\penalty0 (5):\penalty0 1235--1250, 2013.
+\newblock ISSN 1558-5646.
+\newblock \doi{10.1111/evo.12081}.
+\newblock URL \url{http://dx.doi.org/10.1111/evo.12081}.
+
+\bibitem[Maruyama(1970)]{maruyama1970fixation}
+T.~Maruyama.
+\newblock On the fixation probability of mutant genes in a subdivided
+  population.
+\newblock \emph{Genetics Research}, 15\penalty0 (02):\penalty0 221--225, 1970.
+\newblock \doi{10.1017/S0016672300001543}.
+\newblock URL \url{http://dx.doi.org/10.1017/S0016672300001543}.
+
+\bibitem[{Maynard Smith} and Haigh(1974)]{maynardsmith1974hitchhiking}
+J.~{Maynard Smith} and J.~Haigh.
+\newblock The hitch-hiking effect of a favourable gene.
+\newblock \emph{Genet Res}, 23\penalty0 (1):\penalty0 23--35, Feb. 1974.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/4407212}.
+
+\bibitem[Messer and Petrov(2013)]{MesserPetrov}
+P.~W. Messer and D.~A. Petrov.
+\newblock {{P}opulation genomics of rapid adaptation by soft selective sweeps}.
+\newblock \emph{Trends Ecol. Evol. (Amst.)}, 28\penalty0 (11):\penalty0
+  659--669, Nov 2013.
+
+\bibitem[Nachman et~al.(2003)Nachman, Hoekstra, and
+  D'Agostino]{nachman2003genetic}
+M.~W. Nachman, H.~E. Hoekstra, and S.~L. D'Agostino.
+\newblock The genetic basis of adaptive melanism in pocket mice.
+\newblock \emph{Proc Natl Acad Sci U S A}, 100\penalty0 (9):\penalty0
+  5268--5273, Apr. 2003.
+\newblock \doi{10.1073/pnas.0431157100}.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/12704245}.
+
+\bibitem[Nagylaki(1975)]{nagylaki1975conditions}
+T.~Nagylaki.
+\newblock Conditions for the existence of clines.
+\newblock \emph{Genetics}, 80\penalty0 (3):\penalty0 595--615, 1975.
+\newblock URL \url{http://www.genetics.org/content/80/3/595.abstract}.
+
+\bibitem[Nathan et~al.(2008)Nathan, Schurr, Spiegel, Steinitz, Trakhtenbrot,
+  and Tsoar]{nathan2008mechanisms}
+R.~Nathan, F.~M. Schurr, O.~Spiegel, O.~Steinitz, A.~Trakhtenbrot, and
+  A.~Tsoar.
+\newblock Mechanisms of long-distance seed dispersal.
+\newblock \emph{Trends in Ecology \& Evolution}, 23\penalty0 (11):\penalty0 638
+  -- 647, 2008.
+\newblock ISSN 0169-5347.
+\newblock \doi{http://dx.doi.org/10.1016/j.tree.2008.08.003}.
+\newblock URL
+  \url{http://www.sciencedirect.com/science/article/pii/S0169534708002723}.
+
+\bibitem[NEQwiki(2013)]{NEQwiki}
+NEQwiki.
+\newblock {NEQwiki}, the nonlinear equations encyclopedia, 2013.
+\newblock URL
+  \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
+\newblock Accessed December 17, 2014.
+
+\bibitem[Orr and Betancourt(2001)]{orr2001sieve}
+H.~A. Orr and A.~J. Betancourt.
+\newblock Haldane's sieve and adaptation from the standing genetic variation.
+\newblock \emph{Genetics}, 157\penalty0 (2):\penalty0 875--884, Feb. 2001.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/11157004}.
+
+\bibitem[Pennings and Hermisson(2006)]{softsweepsII}
+P.~S. Pennings and J.~Hermisson.
+\newblock Soft sweeps {II} - molecular population genetics of adaptation from
+  recurrent mutation or migration.
+\newblock \emph{Mol Biol Evol}, page msj117, 2006.
+\newblock \doi{10.1093/molbev/msj117}.
+\newblock URL
+  \url{http://mbe.oxfordjournals.org/cgi/content/abstract/msj117v1}.
+
+\bibitem[Pollak(1966)]{pollak1966survival}
+E.~Pollak.
+\newblock On the survival of a gene in a subdivided population.
+\newblock \emph{Journal of Applied Probability}, 3\penalty0 (1):\penalty0
+  142--155, 1966.
+\newblock ISSN 00219002.
+\newblock URL \url{http://www.jstor.org/stable/3212043}.
+
+\bibitem[Ralph and Coop(2010)]{ralph2010parallel}
+P.~Ralph and G.~Coop.
+\newblock Parallel adaptation: One or many waves of advance of an advantageous
+  allele?
+\newblock \emph{Genetics}, 186\penalty0 (2):\penalty0 647--668, 2010.
+\newblock \doi{10.1534/genetics.110.119594}.
+\newblock URL \url{http://www.genetics.org/cgi/content/abstract/186/2/647}.
+
+\bibitem[Ralph and Coop(2015)]{ralph2014standing}
+P.~L. Ralph and G.~Coop.
+\newblock The role of standing variation in geographic convergent adaptation.
+\newblock \emph{The American Naturalist}, 2015.
+\newblock \doi{10.1101/009803}.
+
+\bibitem[Reynolds and Rhodes(2009)]{reynolds2009levy}
+A.~M. Reynolds and C.~J. Rhodes.
+\newblock The {L{é}vy} flight paradigm: random search patterns and mechanisms.
+\newblock \emph{Ecology}, 90\penalty0 (4):\penalty0 877--887, Apr. 2009.
+\newblock URL \url{http://www.esajournals.org/doi/abs/10.1890/08-0153.1}.
+
+\bibitem[Rosenblum et~al.(2010)Rosenblum, Römpler, Schöneberg, and
+  Hoekstra]{rosenblum2010molecular}
+E.~B. Rosenblum, H.~Römpler, T.~Schöneberg, and H.~E. Hoekstra.
+\newblock Molecular and functional basis of phenotypic convergence in white
+  lizards at {W}hite {S}ands.
+\newblock \emph{Proceedings of the National Academy of Sciences}, 107\penalty0
+  (5):\penalty0 2113--2117, 2010.
+\newblock \doi{10.1073/pnas.0911042107}.
+\newblock URL \url{http://www.pnas.org/content/107/5/2113.abstract}.
+
+\bibitem[Schat et~al.(1996)Schat, Vooijs, and Kuiper]{schat1996identical}
+H.~Schat, R.~Vooijs, and E.~Kuiper.
+\newblock Identical major gene loci for heavy metal tolerances that have
+  independently evolved in different local populations and subspecies of
+  \textit{{Silene} vulgaris}.
+\newblock \emph{Evolution}, 50\penalty0 (5):\penalty0 pp. 1888--1895, 1996.
+\newblock URL \url{http://www.jstor.org/stable/2410747}.
+
+\bibitem[Sexton et~al.(2013)Sexton, Hangartner, and
+  Hoffmann]{sexton2013genetic}
+J.~P. Sexton, S.~B. Hangartner, and A.~A. Hoffmann.
+\newblock Genetic isolation by environment or distance: Which pattern of gene
+  flow is most common?
+\newblock \emph{Evolution}, 2013.
+\newblock ISSN 1558-5646.
+\newblock \doi{10.1111/evo.12258}.
+\newblock URL \url{http://dx.doi.org/10.1111/evo.12258}.
+
+\bibitem[Slatkin(1973)]{slatkin1973geneflow}
+M.~Slatkin.
+\newblock Gene flow and selection in a cline.
+\newblock \emph{Genetics}, 75\penalty0 (4):\penalty0 733--756, 1973.
+\newblock URL \url{http://www.genetics.org/cgi/content/abstract/75/4/733}.
+
+\bibitem[Slatkin and Wiehe(1998)]{slatkin1998hitchhiking}
+M.~Slatkin and T.~Wiehe.
+\newblock Genetic hitch-hiking in a subdivided population.
+\newblock \emph{Genet Res}, 71\penalty0 (2):\penalty0 155--160, Apr. 1998.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/9717437}.
+
+\bibitem[Steiner et~al.(2009)Steiner, Rompler, Boettger, Schoneberg, and
+  Hoekstra]{steiner2009genetic}
+C.~C. Steiner, H.~Rompler, L.~M. Boettger, T.~Schoneberg, and H.~E. Hoekstra.
+\newblock The genetic basis of phenotypic convergence in beach mice: Similar
+  pigment patterns but different genes.
+\newblock \emph{Mol Biol Evol}, 26\penalty0 (1):\penalty0 35--45, 2009.
+\newblock \doi{10.1093/molbev/msn218}.
+\newblock URL \url{http://mbe.oxfordjournals.org/cgi/content/abstract/26/1/35}.
+
+\bibitem[Stern(2013)]{stern2013genetic}
+D.~L. Stern.
+\newblock The genetic causes of convergent evolution.
+\newblock \emph{Nat Rev Genet}, 14\penalty0 (11):\penalty0 751--764, Nov. 2013.
+\newblock \doi{10.1038/nrg3483}.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/24105273}.
+
+\bibitem[Turner et~al.(2010)Turner, Bourne, Von~Wettberg, Hu, and
+  Nuzhdin]{turner2010serpentine}
+T.~L. Turner, E.~C. Bourne, E.~J. Von~Wettberg, T.~T. Hu, and S.~V. Nuzhdin.
+\newblock Population resequencing reveals local adaptation of
+  \textit{{Arabidopsis} lyrata} to serpentine soils.
+\newblock \emph{Nat. Genet.}, 42\penalty0 (3):\penalty0 260--3, Jan. 2010.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/20101244}.
+
+\bibitem[Wilson et~al.(2014)Wilson, Petrov, and Messer]{wilson2014selective}
+B.~A. Wilson, D.~A. Petrov, and P.~W. Messer.
+\newblock Soft selective sweeps in complex demographic scenarios.
+\newblock \emph{Genetics}, 198\penalty0 (2):\penalty0 669--684, Oct. 2014.
+\newblock \doi{10.1534/genetics.114.165571}.
+\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/25060100}.
+
+\bibitem[Zhen et~al.(2012)Zhen, Aardema, Medina, Schumer, and
+  Andolfatto]{zhen2012parallel}
+Y.~Zhen, M.~L. Aardema, E.~M. Medina, M.~Schumer, and P.~Andolfatto.
+\newblock Parallel molecular evolution in an herbivore community.
+\newblock \emph{Science}, 337\penalty0 (6102):\penalty0 1634--1637, 2012.
+\newblock \doi{10.1126/science.1226630}.
+\newblock URL \url{http://www.sciencemag.org/content/337/6102/1634.abstract}.
+
+\end{thebibliography}
 
 \end{document}
-
-%%%%%%%% END %%%%%%%%%%%%%%%
-%%%%%%%% %%% %%%%%%%%%%%%%%%
-
-\begin{figure}[ht!]
-    \begin{center}
-        % made in patchy-sim-writeup-plots.R
-        % \includegraphics{example-equilibrium}
-    \end{center}
-    \label{fig:example_equil}
-    \caption{
-    Snapshots of a simulation of allelic frequency dynamics around a patch where the allele is beneficial:
-    the focal allele is benefical in the middle region and deleterious outside.
-    The dynamics were simulated for 10000 generations across demes with 1000 individuals in each
-    and random dispersal to nearby demes with a mean dispersal distance of $\sigma \approx 2$ demes.
-    Colored lines show allele frequencies across space in 50 uniformly spaced generations;
-    the solid line shows the mean across all generations.
-    Below is the same plot, except on a log scale.
-    }
-\end{figure}
-
-\section*{Temporary work: hitting probs}
-
-Now consider the probability that a migrant family from $x$ will ever establish in the new patch,
-which we assume has relatively small area $A$ and is centered at position $y$.
-The probability that this family reaches $A$ by time $t$ is $(1-k_e(t))$
-multiplied by the probability that the family reaches the patch, given survival.
-If $t$ is large then the previous section argues that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
-and the probability of reaching the patch by $t$ approximately the chance that a Brownian path
-will have hit the patch by then.
-If $w = A^{1/d}$ is the width of the patch, then 
-this is the chance that a $d$-dimensional Bessel process begun at $\|y-x\|$ hits $w$ by time $t$.
-Let $\tau_w$ be the hitting time of $w$,
-so, integrating by parts, and using Borodin \& Salminen (2.2.0.1 and 4.2.0.1)
-\begin{align}
-    \int_0^\infty (1-k_e(t)) \P^x\{ \tau_w \le t \} dt  
-    &\approx \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \le t \} dt  / \E[K] \\
-    &= \frac{1}{s_m \E[K]} \int_0^\infty e^{-s_m t} \P^x\{ \tau_w \in dt \} \\
-    &= \frac{1}{s_m \E[K]} \E^x[e^{-s_m \tau_w}] \\
-    &= 
-    \frac{1}{s_m \E[K]} \times
-    \begin{cases}
-        e^{-(x-w) \sqrt{2s_m}} \qquad & d=1 \\
-        \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } \qquad & d=2 
-    \end{cases}
-\end{align}
-Note that the $d=2$ case
-\begin{align}
-    \frac{ K_0(x\sqrt{2s_m}) }{ K_0(w\sqrt{2s_m}) } = \sqrt{\frac{w}{x}} e^{-(x-w)\sqrt{2s_m}} ( 1 + O(1/w) ).
-\end{align}
-
-\section{Patchy selection} 
-\label{ss:discretedemes}
-
-We have been working in a continuous model of geographic space; 
-in this section we move to a discrete model of isolated populations exchanging rare migrants,
-but in section \ref{ss:patchyspace} discuss how a model with {\em patchy} selection can be approximated by such a discrete model.
-
-Suppose now that we have a set of $L$ discrete populations of individuals that may reproduce or migrate,
-and that migration and mutation are rare, so that the waiting times until either event is approximately exponential.
-Suppose that for each pair of populations, labeled $i$ and $j$, the mean number of migrants that travel from $i$ to $j$ per generation
-is $M_{ij}$ and that the mean number of new mutations appearing in population $i$ per generation is $\mu_i$.
-
-{\tt Include cartoon here.}
-
-If we suppose that fixation occurs on a faster time scale than mutation or migration,
-so that it is very unlikely that two different migrants or mutants begin to fix locally in the same population,
-then we can treat the process of extinction or local fixation as instantaneous 
-(a limit previous studied by \cite{Slatkin:81}).
-In this limit we may also ignore swamping effects of inmigration of nonmutant alleles.
-
-In this case, we have the following picture.
-The population at first starts out with no selected alleles. 
-At some later time point suppose that there are $T$ different alleles present in the population, 
-that the $k^\mathrm{th}$ allele is occupying demes $\{i^k_1, \ldots, i^k_{n_k}\}$, for $1\le k \le T$.
-Let $I$ denote the set of all adapted demes, 
-and let $J = \{j_1, \ldots, j_\ell\}$ denote the demes that have not yet adapted, with $\ell = L - \sum_k n_k$.
-Then we are assuming that two types of event can happen which could change the state:
-either at rate $\mu_j p_f$, an unadapted deme $j$ produces a new mutant, which fixes in this deme;
-or at rate $M_{ij} p_f$, an adapted deme $i$ sends a migrant to unadapted deme $j$, which fixes in that deme.
-Therefore, under these assumptions, the probability of fixation $p_f$ (and hence the selection coefficient)
-only enter as a time scaling: for instance,
-if we define $R = \sum_{j \in J} \left( \mu_j + \sum_{i \in I} M_{ij} \right)$,
-then the total rate of such events is $p_f R$.
-The probability that the next event results in, say, adaptation of deme $j$
-is 
-\[
- \left( \mu_j + \sum_{i \in I} M_{ij} \right) / R,
-\]
-and the probability that deme $j$ adapts through migration from deme $i$, given that deme $j$ is the next to adapt, is
-$M_ij / R$,
-while the probability that it adapts through a new mutation is $\mu_j / R$.
-
-The main point here is that the final pattern,
-and hence the likely extent of parallel adaptation,
-is independent of the strength of selection.
-This is in contrast to the continuous case, however, this observation is dependent on the assumption of weak migration
-and fails to hold if different migrant lineages interact.
-This can be viewed as a continuous-time Markov process XXX.
-Explicit solutions are not available in the general case,
-but if we specialize to a completely symmetric model (i.e., the ``island'' model),
-then the mathematical picture is a pretty one.
-
-\subsection{Symmetric patches}
-
-Indeed, suppose that migration and mutation are symmetric: $M_{ij}=m$ and $\mu_j = \mu$ for all $j$ and $i\neq j$.
-Suppose at some time the $k^\mathrm{th}$ allele is occupying $n_k$ demes, 
-with $n = \sum_k n_k < L$.
-Then at rate $(L-n) N \mu p_f + n (L-n) N m p_f$,
-either a migration or mutation event happens.
-With probability $\mu/(\mu + n m)$, it is a mutation,
-and a new deme is randomly chosen and assigned a new allele.
-With probability $n_k m / (\mu + n m)$,
-the $k^\mathrm{th}$ type sends a successful migrant to a randomly chosen new island,
-which is assigned the $k^\mathrm{th}$ allele.
-
-This process is a continuous-time version of the ``Chinese restaurant process''
-described in \citet{aldous1985exchangeability} and \citet{pitman1995partitions},
-and so the final partition of types across demes has the Ewens sampling distribution with parameter $\mu/m$.
-Again note that the selection coefficient, which is implicit in the probability $p_f$,
-does not enter into the final distribution.
-
-Now we can compute most properties we might want about the process.
-For instance, the expected number of distinct mutational origins is
-\begin{equation}
-    \E\left[ \mbox{ \# of independent mutations } \right] = 
-            1 + \frac{\mu}{m} \sum_{k=2}^L \frac{1}{k+(\mu/m)}. \label{discrete_expected}
-\end{equation}
-The probability that there is only a single successful mutation is
-\begin{equation}
-\P \left\{ \mbox{only one mutational origin} \right\} = 
-            \prod_{k=1}^{L-1} \frac{ k }{ k+(\mu/m)} .
-\end{equation}
-More generally, suppose we sample a single deme, and are interested in the total number of demes $S$
-(including the sampled one) that share the same fixed mutation as our sampled deme.
-Then $S$ has distribution
-\begin{equation} \label{eqn:Sdistrn}
-\P\{ S=s \} = \frac{ (s-1)! (\mu/m)^{L-s} }{ \prod_{k=1}^{L-1} (k+(\mu/m)) } .
-\end{equation}
-Furthermore, if $L$ is large, then $S/L$ has probability density approximately
-\begin{equation} \label{eqn:betadistrn}
-   (\mu/m) (1-x)^{(\mu/m)-1}
-\end{equation}
-namely, is a $\mathrm{Beta}((\mu/m), 1)$ distribution--- see \cite{donnelly1989continuity} and \cite{perman1992sizebiased}.
-
-% In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
-% for a model with 20 interconnected demes, across different parameter values.  
-% This indicates that we need the population-scaled mutation and migration rates to be within a factor of about 100 of each other for parallel adaptation to leave an interesting pattern.
-If migration is faster, then a single type is likely to take over, while if migration is weaker, each deme is likely to come up with its own type.
-
-The high connectedness of the discrete deme island model means that the expected number of distinct alleles, (Equation \eqref{discrete_expected}), 
-grows with the $\log$ of the number of demes.
-This strongly contrasts with the continuous spatial model where the local nature of dispersal means that doubling the species range will double the number of mutations expected.
-Furthermore, while in the continuous model the areas occupied by distinct mutations are similar in area, in this strong-selection discrete island model, 
-only a few types tend to dominate even as the number of demes (and thus, distinct types) grows large (see Equations \eqref{eqn:Sdistrn} and \eqref{eqn:betadistrn}).
-
-A more general model would allow migration only along edges of some graph connecting the demes.
-In the low migration limit such a model still produces a partition distribution independent of the selection coefficient,
-but it does not in general have the Ewens distribution.
-Another extension would be to include the time alleles need to achieve an intermediate frequency,
-along the lines of \citet{Navarro:03}, which would reintroduce dependence on the selection coefficient.
-
-\texttt{XXX omit this? XXX}
-We have treated the time during which the selected allele is at intermediate frequency in a deme (about $\log(\sigma^2 \rho)/s_p$ generations) as negligible, 
-which ignores migration events that occur while the mutation is spreading through the population. 
-we now sketch of a more realistic approximation.
-For each occupied--unoccupied pair of demes, if the number of mutant alleles in the occupied deme is $n(t)>0$,
-then at time $t$, migrants from the occupied deme pass the selected allele to the unoccupied deme at rate $2 n(t) m p_f(s_p)$,
-while newly arisen mutants arise and fix in the unoccupied deme at constant rate $2 N \mu p_f(s_p)$.
-The selected type should grow approximately as $n(t)=\exp(s(t-t_0))$.
-Of course, to follow this model through, we will have to account for multiple types within a single deme and other complications,
-but we can at least make a few observations.
-One is that the final distribution of types is no longer independent of $s_p$, which enters through the dynamics $n(t)$.
-However, it is clear that increasing the selection coefficient $s$ will decrease the number of types (independent mutational origins),
-while on the other hand, increasing the deme size $N$ will increase the number of types,
-and that the case presented above is at the extreme (see also \cite{Navarro:03} for deterministic approximations).
-
-%%%%%%% %%%%%%%%%%
-\subsection[Haplotypes Shared Within a Patch]{Length of the shared haplotype within a patch}
-\label{ss:patch_haplotype}
-
-The other key issue in thinking about the population genetic signal of adaptation in spatial patches 
-is the process by which haplotypes are whittled down within patches, 
-which plays into a number of aspects of haplotype sharing. 
-The initial haplotype that sweeps within a patch 
-will be dispersed over time by recombination.
-Likewise, the haplotype that is shared between patches coadapted by migration
-(equation~\ref{eqn:haplotype_length}) will also break down. 
-However, a long time after the initial sweep,
-we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
-than with individuals elsewhere,
-since selection against migrants 
-decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}. 
-We first develop a simple approximation to the rate 
-at which recombination breaks down haplotypes within patches.
-
-Refer again to Figure~\ref{fig:lineagesmotion}. 
-When a lineage is in a region with high frequency of $B$ alleles (dark red), 
-it is very unlikely to recombine off the $B$ background. 
-However, when it wanders out of the patch,
-into regions with lower frequency of $B$ alleles,
-when it recombines it will
-usually be on to the background of the $b$ allele.
-
-A detailed treatment of this is well beyond the scope of this paper,
-but we can get a good idea of the genomic scale of these effects through heuristic arguments.
-To understand the probability of a recombination moving our lineage
-off of the B background we need to understand what fraction of the
-time our lineage spends in geographic regions where $b$ is common. 
-We will follow the lineage of a single $B$ allele
-in a patch at migration-selection balance, 
-i.e., with the local proportion of $B$ alleles near $x$ equal to $q(x)$,
-and suppose that the lineage stays close enough to the patch 
-that fluctuations about $q(x)$ are small. 
-(Note that this assumption differs from
-previous sections, but here our focus is on a typical lineage 
-in the patch rather than the rare lineages that make it far from the patch).
-If we suppose that type $B$ individuals at location $y$ reproduce at rate $1+s(y)$,
-and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
-then the total rate of influx of $B$ alleles from $y$ into $x$ is $\rho q(y) (1+s(y)) m(y,x)$.
-As argued by \citet{hudson1988coalescent}, 
-this implies that the chance that a $B$ allele at $x$
-has been inherited from a parent at $y$ in a time interval $dt$
-is equal to this influx divided by the local abundance of $B$ alleles,
-i.e., $dt \times q(y) (1+s(y)) m(y,x) / q(x)$.
-This therefore describes the movement of a lineage backwards in time.
-After the same large-population, continuous-space limit that underlies the rest of this paper (section~\secref{apx:eqfreq}),
-if the assumption that fluctuations in $q(x)$ are small,
-this motion should converge to a diffusion process (as assumed by e.g., \citet{hallatschek2008surfing}),
-notwithstanding some technical difficulties (e.g., discontinuity of $s(x)$).
-This diffusion, when at $x$, 
-is pulled toward regions with more $B$ alleles
-by a mean displacement term $(1+s(x))\frac{d}{d\,x} \log( q(x) (1+s(x)) )$,
-where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
-and moves at speed $\sqrt{1+s(x)}$ (which is nearly constant, since $s(x)$ is small).
-In It\=o differential notation, 
-the location $X_t$ of a $B$ lineage $t$ generations ago satisfies
-\begin{align}
-    d X_t &= (1+s(X_t)) \frac{d}{dx} \log( q(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,
-\end{align}
-where $B_t$ is a standard Brownian motion.
-Furthermore, it is easy to verify that the stationary distribution of this process has density
-\begin{align}
-  \pi(x) = \frac{ q(x)^2 (1+s(x)) }{ \int_y q(y)^2 (1+s(y)) dy } . \label{eqn:lineagestatdist}
-\end{align}
-A lineage at a locus at genetic distance $R$ from a $B$ allele
-moves as a $B$ lineage until it recombines onto a $b$ background,
-which occurs when at $x$ with rate $r (1-q(x)) (1+s(x)/2)$.
-The Feynman-Kac formula could then be used to write down differential equations
-solved by various moments of the haplotype length as a function of the sampling location of the lineage;
-instead, we only use the above as a guide for rough approximation.
-
-Based on this picture of
-the lineage bouncing around in the patch,
-we propose a simple heuristic to describe the probability of recombination. 
-If recombination is on a slower time scale than mixing within the patch,
-then the time until the last recombination is the integral of $r (1-q(x)) (1+s(x)/2)$ 
-against the stationary distribution (equation~\ref{eqn:lineagestatdist}).
-Within the patch, $q(x) \approx 1$,
-so recombination off the $B$ background is unlikely to occur,
-and so the rate of recombination depends to a first approximation 
-on the proportion of the patch in which $q(x)$ is intermediate,
-i.e., where both $B$ and $b$ are present.
-This region where $q(x)$ is intermediate has width of order $\sigma/\sqrt{s}$, where $s = \min(s_p,s_m)$.
-As we increase the size of a patch, the area on which $q(x)$ is intermediate scales with the perimeter of the patch.
-% Therefore, the proportion of the time a lineage at stationarity is found in the intermediate regions
-% multiplied by $q(x)$ in these intermediate regions, where we assume $q(x) \approx 1/2$. 
-The per generation rate of recombination of a locus linked to the $b$ background is thus
-reduced by a factor roughly equal to the proportion of the patch lying within distance $\sigma/\sqrt{s}$ of the boundary,
-so the effective recombination rate is
-% \plr{Note to ourselves: If a $d$-dimensional object with area A has radius $r$ (in any sense)
-% and you shrink it to have radius $r_0$,
-% then the resulting object has area $A \times (r_0/r)^d$;
-% here we have $r_0 = r - \sigma/\sqrt{s}$,
-% in $d=1$ the radius is $A/2$ and in $d=2$ it is $\sqrt{A/\pi}$.}
-\begin{align}
-r_\text{eff} &=
-    \begin{cases}
-    r \frac{ \sigma /\sqrt{s} }{ \sigma/\sqrt{s} + A }    \qquad & d=1 \\
-   r \frac{\pi  \sigma /\sqrt{s}  \left( \sigma /\sqrt{s}+
-  \sqrt{A/\pi} \right)}{A+\pi \sigma/\sqrt{s} \left( \sigma/\sqrt{s} +
-  \sqrt{A/\pi} \right) }    \qquad & d=2 .
-\end{cases}
-\label{eqn:rec_rate_within}   %(\sigma/\sqrt{s})A^{-1/d}
-\end{align}
-It follows that $T$ generations after the founding of a patch we expect a
-haplotype of genetic length $1/(r_\text{eff}T)$ Morgans to remain linked to
-the $B$ allele in a typical lineage. 
-Note that the length of haplotype
-decays at a rate roughly proportional to $\sqrt{s}/ \sigma$, as the narrower the cline 
-the less opportunity for recombination to move our lineage off the $B$ background.
-
-Finally we can consider the expected length of haplotype shared within a patch. 
-To do this, we compare the rate of recombination between $B$ and $b$ haplotypes (equation~\ref{eqn:rec_rate_within})
-to the rate of coalescence of $B$ haplotypes within a patch. 
-We will assume that mixing within the patch is fast compared to the rate of drift, 
-so that the rate of coalescence of $B$ alleles within a patch is 
-$\xi^2/(A\rho)$,
-the ``variance effective population size'' of the patch
-(this bypasses the well-known difficulties to spatial coalescence \citep{felsenstein1975pain,barton2002neutral}).
-The rate of recombination and coalescence will be on the same order if
-$\xi^2 /(A\rho) \approx r (\sigma/\sqrt{s}) A^{-1/d}$.
-Hence the genetic length of the haplotype shared between pairs of
-individuals within the same patches at equilibrium is on the order of
-\begin{align} \label{eqn:within_haplotype_length} 
-  r = \frac{\sqrt{s} \xi^2}{\sigma \rho A^{(d-1)/d}}  .
-\end{align}
-This does not depend on $A$ in one dimension,
-but decreases with $\sqrt{A}$ in two dimensions.
-In applications, this can be compared to equation~\eqref{eqn:haplotype_length}
-for the length of shared haplotype between patches.

### 2015-10-17 patchy-selection-paper.tex
moving over to plos

@@
 \usepackage{caption}
-\DeclareCaptionType{sfigure}[Figure]
+\DeclareCaptionType{sfigure}[Fig]
 
@@
 
-%%%%%
-% Responses to reviews:
-% set this to show line numbers and include responses to reviews or not
-\newif\ifreviewresponses
-\reviewresponsestrue  % include them
-% \reviewresponsesfalse  % don't include them
-\iflatexml\reviewresponsesfalse\fi  % they don't work so well in html
-% \newcommand{\responsefile}{pbio-reviews-19sept12-responses.tex}  % name of the review responses file
-\newcommand{\responsefile}{patchy-review-responses-second-round.tex}  % name of the review responses file
-
-% override above choices (see makefile)
-\ifdef{\submissionnofigs}{  % for submission, no figures
-  \reviewresponsesfalse
-  \submissionversiontrue
-}{
-}
-\ifdef{\submissionfigs}{  % for reviewers, with figs
-  \reviewresponsestrue
-  \submissionversionfalse
-}{
-}
-\ifdef{\arxiv}{  % for general use
-  \reviewresponsesfalse
-  \submissionversionfalse
-}{
-}
-
-\ifreviewresponses
-  \input{review-response-commands}
-\else
+%  %%%%%
+%  % Responses to reviews:
+%  % set this to show line numbers and include responses to reviews or not
+%  \newif\ifreviewresponses
+%  \reviewresponsestrue  % include them
+%  % \reviewresponsesfalse  % don't include them
+%  \iflatexml\reviewresponsesfalse\fi  % they don't work so well in html
+%  % \newcommand{\responsefile}{pbio-reviews-19sept12-responses.tex}  % name of the review responses file
+%  % \newcommand{\responsefile}{patchy-review-responses-second-round.tex}  % name of the review responses file
+%  
+%  % override above choices (see makefile)
+%  \ifdef{\submissionnofigs}{  % for submission, no figures
+%    \reviewresponsesfalse
+%    \submissionversiontrue
+%  }{
+%  }
+%  \ifdef{\submissionfigs}{  % for reviewers, with figs
+%    \reviewresponsestrue
+%    \submissionversionfalse
+%  }{
+%  }
+%  \ifdef{\arxiv}{  % for general use
+%    \reviewresponsesfalse
+%    \submissionversionfalse
+%  }{
+%  }
+%  
+%  \ifreviewresponses
+%    \input{review-response-commands}
+%  \else
     \newcommand{\linelabel}[1]{}
@@
     \newcommand{\includereviews}{}
-\fi
+%  \fi
 
@@
   \caption{
-  A segment of the temporal dynamics of the same simulation as Figure~\ref{fig:sim_occupation_freqs}B.
+  A segment of the temporal dynamics of the same simulation as Fig~\ref{fig:sim_occupation_freqs}B.
   Space is again shown across the horizontal axis, with time on the vertical axis; 
@@
 % One way to visualize this is by a phase diagram.
-% In figure \ref{fig:phase_diagram},
+% In Fig \ref{fig:phase_diagram},
 % we give a contour plot for the time until adaptation by migration ($\Tmig = 1/\migrate$) and by new mutation ($\Tmut=1/\mutrate$)
@@
       give the probability predicted by theory (equation~\eqref{eqn:parallel_prob}, with $C=5$).
-      Points correspond to parameter combinations of the simulations in figure \ref{fig:sim_times},
+      Points correspond to parameter combinations of the simulations in Fig \ref{fig:sim_times},
       with the shade of grey and the red numeric label indicating the same probability
@@
 
-Here we provide additional details regarding the simulations we used
-to validate the theory in figures \ref{fig:sim_times} and \ref{fig:sim_probs}.
+In tables \ref{sfig:sim_migration_times}
+and \ref{sfig:sim_mutation_times}
+we provide additional details regarding the simulations we used
+to validate the theory in \ref{fig:sim_times} and \ref{fig:sim_probs} Fig.
 
@@
   \caption{
-    The same data shown in the left panel of figure~\ref{fig:sim_times},
+    The same data shown in the left panel of Fig~\ref{fig:sim_times},
     but all times shown (not just the interquartile ranges),
@@
   \caption{
-    The same data shown in the right panel of figure~\ref{fig:sim_times},
+    The same data shown in the right panel of Fig~\ref{fig:sim_times},
     but all times shown (not just the interquartile ranges),
     and including those parameter values at which most of the simulations did not adapt by 25,000 generations.
-    The upper panel has the predicted time to adaptation on the horizontal axis as in figure~\ref{fig:sim_times},
+    The upper panel has the predicted time to adaptation on the horizontal axis as in Fig~\ref{fig:sim_times},
     and the lower panel has, for comparison, the raw distance between patches 
@@
 
-% In Figure \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
+% In Fig \ref{fig:discbyratio} we plot mean number of types and size (number of demes occupied) of a sampled type,
 % for a model with 20 interconnected demes, across different parameter values.  
@@
 
-Refer again to Figure~\ref{fig:lineagesmotion}. 
+Refer again to Fig~\ref{fig:lineagesmotion}. 
 When a lineage is in a region with high frequency of $B$ alleles (dark red), 

### 2015-10-19 patchy-selection-paper-for-PLoS.tex
.

@@
 %%%
-\subsection{Model of a patchy landscape}
+\subsection*{Model of a patchy landscape}
 \label{ss:patchyspace}
@@
 %%%%%
-\subsection[Establishment by Mutation]{Establishment of a locally adaptive allele due to mutational influx}
+% \subsection[Establishment by Mutation]{Establishment of a locally adaptive allele due to mutational influx}
+\subsection*{Establishment of a locally adaptive allele due to mutational influx}
 \label{ss:patchymutation}
@@
 %%%%%%%
-\subsection[Establishment by Migration]{Establishment of a locally adaptive allele due to migrational influx}
+% \subsection[Establishment by Migration]{Establishment of a locally adaptive allele due to migrational influx}
+\subsection*{Establishment of a locally adaptive allele due to migrational influx}
 \label{ss:patchymigration}
@@
 Now suppose that there are two patches separated by distance $R$
-(i.e., the shortest distance between the two is $R$). \revpoint{2}{8}
+(i.e., the shortest distance between the two is $R$).
 If the $B$ allele has arisen and become established in the first patch, but has not yet appeared in the second,
@@
 To be clear about the assumptions implicit here,
-below we provide a derivation of \eqref{eqn:eqfreq} in section \secref{apx:eqfreq},  \revpoint{3}{5}
+below we provide a derivation of \eqref{eqn:eqfreq} in section \secref{apx:eqfreq},
 as well as justification for the asymptotics below in section \secref{apx:asymptotics}.
@@
   \begin{center}
-    \includegraphics{sim-snapshots}
+    % \includegraphics{sim-snapshots}
   \end{center}
@@
 so the entire family is spread slightly wider.
-\revpoint{2}{2}
 % If $\P\{ T_k > k \} = e^{-sk}$
@@
   \begin{center}
-    \includegraphics{branching-concept}
+    % \includegraphics{branching-concept}
   \end{center}
@@
 %%%%%%%%%
-\subsection{A review, and a test, of the assumptions}
+\subsection*{A review, and a test, of the assumptions}
 \label{ss:assumptions}
@@
 should be added to the outside of the patch in computing $A$ for equation \eqref{eqn:mutrate},
-and if $s_m \gg s_p$, a strip of width $\sigma/\sqrt{s_p}$ should be subtracted from the patch. \revpoint{2}{6}
+and if $s_m \gg s_p$, a strip of width $\sigma/\sqrt{s_p}$ should be subtracted from the patch.
 As for the migration rate,
@@
       % made in analyze-sims.R
-      \includegraphics{times-predicted-observed}
+      % \includegraphics{times-predicted-observed}
   \end{center}
@@
 %%%%%%%%%
-\subsection[Probability of Parallel Adaptation]{The probability of parallel adaptation between patches.} 
+% \subsection[Probability of Parallel Adaptation]{The probability of parallel adaptation between patches.} 
+\subsection*{The probability of parallel adaptation between patches.} 
 \label{ss:probparallel}
@@
 so that the probability of parallel mutation should increase
-approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{2 s_m}$. \revpoint{2}{13}
+approximately logistically with the distance between the patches, on the spatial scale $\sigma/\sqrt{2 s_m}$.
 
@@
 %%%%%%%%%
-\subsection[Haplotypes Shared Between Patches]{Length of the hitchhiking haplotype}
+% \subsection[Haplotypes Shared Between Patches]{Length of the hitchhiking haplotype}
+\subsection*{Length of the hitchhiking haplotype}
 \label{ss:haplotype_length}
@@
 and that $\var[\tau \given \tau<\infty ] = (R\sqrt{2s_m}/\sigma) \times 1/(2s_m)^2$.
-(A saddle point approximation provides an alternative route to these expressions.) \revpoint{1}{4}
+(A saddle point approximation provides an alternative route to these expressions.)
 
@@
 
-% In Appendix~\ref{apx:limiting_dist_of_HH_haplotype}, 
-% we also give some expressions for the limiting distribution of the length of shared haplotype.
-
-% 
-% \section{Limiting distribution of the transit time and hitchhiking haplotype}
-% \label{apx:limiting_dist_of_HH_haplotype}
-% 
-% In the above we already assumed that the inter-patch distance $R$ is large relative to the migration-selection cline width $\sigma/\sqrt{s_m}$;
-% also taking the large-$R$ limit, obtains a more intuitive description of the process
-% via a central limit theorem for $\tau$.
-% Define $\eta_\tau$ to be the deviation of $\tau$ from its mean, normalized by its standard deviation,
-% so that $\tau = R/\sigma(2s_m)^{-1/2} + \eta_\tau \sqrt{R/\sigma} (2s_m)^{-3/4}$.
-% Plugging this into equation~\eqref{eqn:haplen_cdf},
-% after some algebra (and using that $\sqrt{R+\epsilon} = \sqrt{R} + \frac{\epsilon}{2\sqrt{R}} - \frac{\epsilon^2}{2\sqrt{R}^3} + O(\epsilon^3)$) we obtain that
-% \begin{align}
-%   \E[e^{-\ell \eta}] &= \exp\left\{ \frac{\ell^2}{2} + O\left(\frac{1}{R}\right) \right\}.
-% \end{align}
-% Since the Laplace transform uniquely determines a distribution \citep{durrett1996probability},
-% this implies that $\eta_\tau$ converges in distribution to a standard Gaussian random variable as $R \to \infty$.
-% In other words, we now know that the distribution of $\tau$ about its mean is approximately Gaussian.
-% Since the standard deviation of $\tau$ is of lower order than its mean,
-% a similar calculation shows that for large $R$, the haplotype halflength $L$ is approximately exponentially distributed, 
-% with mean $\sqrt{2s_m} \sigma/R$.
-
-% \begin{align}
-%   \E[e^{-\ell \eta}] &= \E\left[ \exp\left\{-\ell\left(\tau - \frac{R}{\sigma\sqrt{2s_m}}\right)\frac{(2s_m)^{3/4}\sqrt{\sigma}}{\sqrt{R}} \right\} \right] \\
-%   &= \E\left[ \exp\left\{ - \left(\frac{\ell (2s_m)^{3/4} \sqrt{\sigma}}{\sqrt{R}}\right) \tau \right\} \right] \exp\left( (2s_m)^{1/4} \sqrt{\frac{R}{\sigma}} \right) \\
-% \end{align}
-
 %%%%%%%%%%
-\subsection{Applications} 
+\subsection*{Applications} 
 \label{ss:applications}
@@
 (However, patches of all sizes occur across all scales in a heterogeneous manner.)
-\revpoint{2}{7}
 \citet{nachman2003genetic} demonstrated that on one of these flows (Pinacate), 
@@
 This determines the values of $s_m$ and $s_p$ reported in the figures, 
-and do not depend on the basic rate of reproduction. \revpoint{2}{5}
+and do not depend on the basic rate of reproduction.
 However, to obtain values for $s_p$ and $s_m$ when comparing theory to simulation,
@@
 
-\subsection[Probability of Establishment]{Numerical calculation of the probability of establishment}
+% \subsection[Probability of Establishment]{Numerical calculation of the probability of establishment}
+\subsection*{Numerical calculation of the probability of establishment}
 \label{apx:establishment_sims}
@@
         % made in patchy-selection-plots.R
-        \includegraphics{prob-establishment}
+        % \includegraphics{prob-establishment}
     \end{center}
@@
 %%%%%%%%%%%%
-\subsection[Equilibrium Frequency]{The equilibrium frequency}
+% \subsection[Equilibrium Frequency]{The equilibrium frequency}
+\subsection*{The equilibrium frequency}
 \label{apx:eqfreq}
@@
 
-In one dimension, we are done (and discuss exact solutions in the Supporting Information);
+In one dimension, we are done (and discuss exact solutions in S1 Text);
 in higher dimensions, we are more interested in the mean frequency at a given distance $r$ from a patch.
@@
 %%%%%%%%%%%
-\subsection[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
+% \subsection[Asymptotics]{Asymptotic solution for the equilibrium frequency} 
+\subsection*{Asymptotic solution for the equilibrium frequency} 
 \label{apx:asymptotics}
@@
 %%%%%%%%% %%%%%%%%%%%%%%
-\subsection[Hitting and Occupation]{Hitting and occupation}
+% \subsection[Hitting and Occupation]{Hitting and occupation}
+\subsection*{Hitting and occupation}
 \label{ss:hitting_occupation}
@@
 %%%%%%% %%%%%%%%%%%%
-\subsection[The Integral $q(S)$]{The integral $q(S)$}
+% \subsection[The Integral $q(S)$]{The integral $q(S)$}
+\subsection*{The integral $q(S)$}
 \label{apx:qS}
@@
 Lower bounds could be obtained along similar lines.
-
-% %%%%%%% %%%%%%%%%%%%
-% \subsection[Outflux of Families]{The outflux of families}
-% \label{apx:outflux}
-% 
-% It is of interest to have an expression for the ``outflux of families'',
-% $\gamma(r_0)$, 
-% defined to be the mean density of $B$ alleles at distance $r_0$
-% whose lineage traces back to the original patch without venturing further away than $r_0$.
-% By expression \eqref{eqn:occupation_integral},
-% taking $S$ to be a small area located at $y$,
-% \[
-%     \gamma(r_0) = \frac{\rho q(y)}{\int_{B(r_0)} u(x,y) dx}
-% \]
-% for any $y$.  
-% Take $\hat x = (1,0)$ and $y=r \hat x$.  Then
-% \begin{align}
-%   \frac{ q(y) }{ u(x,y) } &\simeq C' \frac{ \|y\|^{-(d-1)/2} e^{-\|y\| \sqrt{2s_m}/\sigma } }{ \|y-x\|^{-(d-1)/2} e^{-\|y-x\| \sqrt{2s_m}/\sigma } } \\
-%   &= C' \| \hat x - x/r \|^{(d-1)/2} e^{ r \sqrt{2s_m}/\sigma  (\|\hat x - x/r\| - 1)} .
-% \end{align}
-% Now let $x = (r_0 \cos \theta, r_0 \sin \theta)$,
-% so that the expression in the exponent is to, first order in $1/r$,
-% \begin{align}
-%     r (\|\hat x - x/r\| - 1) &= r \left( \sqrt{ \left( 1 - \frac{r_0}{r} \cos \theta \right)^2 + \frac{r_0^2}{r^2} \sin^2 \theta } - 1 \right ) \\
-%             &= - r_0 \cos \theta + O(1/r) .
-% \end{align}
-% This implies that, still with $r$, $r_0$, and $\theta$ defined as above,
-% \begin{align}
-% \frac{q(y)}{u(x,y)} \longrightarrow C' e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} \qquad \text{ as } r \to \infty .
-% \end{align}
-% In one dimension we are done (since $\cos \theta=1$).
-% In two dimensions,
-% by bounded convergence and \citet{gradshteyn2007table} (3.339),
-% \begin{align}
-%   \int_{B(r_0)} \frac{q(y)}{u(x,y)} dx &\to \gamma(r_0) \quad \text{as } y \to \infty\\
-%   &= C' \int_0^{2 \pi} e^{-r_0 \frac{\sqrt{2s_m}}{\sigma} \cos \theta} d\theta \\
-%   &= 2 \pi C' I_0(-r_0\sqrt{2 s_m}/\sigma),
-% \end{align}
-% where $I_0$ is a modified Bessel function of order zero of the first kind.
 
@@
 }
-Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
-so that 
-\[
-\partial_t p(t,x) = \frac{\mu}{2} \sigma^2 \partial_x^2 p(t,x) + \mu s(x) p(t,x) (1-p(t,x)) .
-\]
-The stable distribution $\phi(x) = \lim_{t\to\infty} p(t,x)$ then solves
-\begin{align} \label{eqn:definingphi}
-    \partial_x^2 \phi(x) = - 2 s(x) \phi(x) (1-\phi(x)) /\sigma^2,
-\end{align}
-with appropriate boundary conditions.
-First rescale space by $\sigma/\sqrt{2}$ so the $\sigma^2/2$ term disappears.
-If we now assume that $s(x)$ is piecewise constant,
-$s(x) = s_i$ for $x \in [x_i,x_{i+1})$, with $x_0=-\infty$ and $x_{n+2}=\infty$,
-then the equation is integrable: if we multiply through by $2\partial_x \phi(x)$ and integrate, then we get that
-\begin{align} \label{eqn:conservation}
-    ( \partial_x \phi(x) )^2  &= - \int^{x} 2 s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
-        &= - s_i \phi(x)^2 \left( 1 - \frac{2}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) ,
-\end{align}
-if we define
-\begin{align} 
-  K_i = - s_i \phi(x_i)^2 \left( 1 - \frac{2}{3} \phi(x_i) \right) - ( \partial_x \phi(x_i) )^2  .
-\end{align}
-For ease of reference, we define
-\[
-        V_i(\phi) = s_i \phi^2 \left( 1 - \frac{2}{3} \phi \right) + K_i .
-\]
-Note that $V_i'(0)=V_i'(1)=0$, that $V_i(0)=K_i$ and $V_i(1) = K_i+s_i/3$.  
-We will always have that $V(\phi) \le 0$.
-(We have then that $\partial_x^2 \phi = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
-Also note that this implies ``conservation of energy'', i.e., $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
-Rearranging, we get that $dx = d\phi / \sqrt{-V(\phi)}$, so where $\phi(x)$ is monotone, the inverse is
-\[
-    x(\phi) = x(\phi^*) \pm \int_{\phi^*}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } .
-\]
-For each $i$ then define the elliptic function
-\begin{align}  \label{eqn:elliptic_function}
-    F_i(\phi) = \int_{\phi_i^*}^\phi \frac{ d\psi }{ \sqrt{ -V_i(\psi) } } ,
-\end{align}
-where take the positive branch of the square root, and $\phi_i^*$ will be chosen later.
-Then we have that
-$x(\phi) - x(\phi_0) = \pm( F(\phi) - F(\phi_0))$,
-or for an appropriate $x_0$,
-\[
-    \phi(x) = F^{-1}\left( F(\phi(x_0)) \pm (x - x_0) \right).
-\]
-
-We clearly want $\lim_{x \to \infty} \partial_x \phi(\pm x) = 0$, 
-and $\lim_{x \to \infty} \phi(\pm x)$ to be zero or one depending on the sign of $s_0$ and $s_{n+1}$.
-Since $(\partial_x \phi(x))^2 = -V(x)$, this implies that if $s_0<0$, then $K_0 = 0$,
-while if $s_0>0$ then $K_0 = s_0/6$; and likewise for $K_{n+1}$.
-
-We also require that $\phi(x)$ and $\phi'(x)$ are continuous.
-Continuity of $\phi'(x)$ is equivalent to $V_i(x_{i+1}) = V_{i+1}(x_{i+1})$,
-which we can rearrange to find an equation for $K_{i+1}$ in terms of $K_i$ and $\phi(x_{i+1})$:
-\[
-    K_{i+1} - K_i = (s_i - s_{i+1}) \phi(x_{i+1})^2 (1-2\phi(x_{i+1})/3) .
-\]
-What about the frequency at the points the selection changes, $\phi(x_i)$?  
-Well, if $\phi(x)$ is monotone on $[x_i,x_{i+1})$ then we can without loss of generality take $\phi_i^*=0$ or $1$ depending on the sign of $s_i$.
-Otherwise, let $\phi_i^*$ be the (unique) root of $V_i$ in $[x_i,x_{i+1})$, so $V_i(\phi_i^*)=0$.
-In this case, $\phi_i^*$ is the maximum or minimum of $\phi$ in the interval:
-if $s>0$, then $\phi_i = \max\{ \phi(x) : x \in [x_i,x_{i+1})\}$. 
-Recall we defined $F_i$ using $\phi_i^*$; now using the fact that $\phi$ is monotone with the opposite sign on either side of $\phi_i^*$,
-$x_{i+1} - x_i = F_i(\phi(x_{i+1})) - F_i(\phi_i^*) + F_i(\phi(x_{i})) - F_i(\phi_i^*)$,
-and that $F(\phi_i^*) = 0$,
-we know that the length of the $i$th stretch is
-\[
-    x_{i+1} - x_i = \pm \left( F_i(\phi(x_i)) + F_i(\phi(x_{i+1})) \right).
-\]
-Note that all the $\pm{}$'s are easily relatable to the signs of $s_i$.
-If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be able to solve the equations for $\phi(x_i)$ and $K_i$ recursively upwards.
-In some cases, such as \citet{slatkin1973geneflow}, we can infer $\phi(0)$ and $\partial_x \phi(0)$ by spatial symmetry.
-In other cases, we are only given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
-
-\paragraph{Doing the integrals}
-\label{apx:some_integrals}
-
-An important ingredient in the above method is the integral \eqref{eqn:elliptic_function},
-to which a method for solving the Korteweg-de Vries equation can be applied \citep{NEQwiki}.
-Recall that $V_i(\phi) =  s_i \phi^2(1-2\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries;
-since we're just working within an interval with $s$ constant, we can rescale space by $\sqrt{2 s_i}/\sigma$,
-so that $s$ and $\sigma$ drop from the equation.
-We then want to integrate
-\[
-    \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
-         \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } ,
-\]
-over a domain where $V$ is always negative.
-Let $\phi^2(1-2\phi/3)+K = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
-and change variables first to $y^2=(\alpha-\phi)$, 
-and then to $x = y/\sqrt{\alpha-\beta}$, so that
-\begin{align*}
-    \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } 
-        = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-y^2) } } \\
-        = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
-\end{align*}
-with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
-Now Jacobi's incomplete elliptic integral of the first kind is defined by
-\[
-    F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) }} ,
-\]
-and the Jacobian elliptic function $\sn(x;k)$ is the inverse: $F(\sn(x;k);k) = x$.
-As $k \to 0$, $\sn(x;k) \to \sin(x)$, while as $k \to 1$, $\sn(x;k) \to \sinh(x)$.
+% Let $p(t,x)$ be the proportion of the allele under spatially varying selection $s(x)$ in one dimension,
+% so that 
+% \[
+% \partial_t p(t,x) = \frac{\mu}{2} \sigma^2 \partial_x^2 p(t,x) + \mu s(x) p(t,x) (1-p(t,x)) .
+% \]
+% The stable distribution $\phi(x) = \lim_{t\to\infty} p(t,x)$ then solves
+% \begin{align} \label{eqn:definingphi}
+%     \partial_x^2 \phi(x) = - 2 s(x) \phi(x) (1-\phi(x)) /\sigma^2,
+% \end{align}
+% with appropriate boundary conditions.
+% First rescale space by $\sigma/\sqrt{2}$ so the $\sigma^2/2$ term disappears.
+% If we now assume that $s(x)$ is piecewise constant,
+% $s(x) = s_i$ for $x \in [x_i,x_{i+1})$, with $x_0=-\infty$ and $x_{n+2}=\infty$,
+% then the equation is integrable: if we multiply through by $2\partial_x \phi(x)$ and integrate, then we get that
+% \begin{align} \label{eqn:conservation}
+%     ( \partial_x \phi(x) )^2  &= - \int^{x} 2 s(x) \phi(x) (1-\phi(x)) \partial_x \phi(x) dx \\
+%         &= - s_i \phi(x)^2 \left( 1 - \frac{2}{3} \phi(x) \right) - K_i \quad \mbox{for } x \in [x_i,x_{i+1}) ,
+% \end{align}
+% if we define
+% \begin{align} 
+%   K_i = - s_i \phi(x_i)^2 \left( 1 - \frac{2}{3} \phi(x_i) \right) - ( \partial_x \phi(x_i) )^2  .
+% \end{align}
+% For ease of reference, we define
+% \[
+%         V_i(\phi) = s_i \phi^2 \left( 1 - \frac{2}{3} \phi \right) + K_i .
+% \]
+% Note that $V_i'(0)=V_i'(1)=0$, that $V_i(0)=K_i$ and $V_i(1) = K_i+s_i/3$.  
+% We will always have that $V(\phi) \le 0$.
+% (We have then that $\partial_x^2 \phi = - \partial_\phi V(\phi)$, the equation of motion of a particle in potential $V$.
+% Also note that this implies ``conservation of energy'', i.e., $( \partial_x \phi(x) )^2 + V(\phi(x))$ is constant.)
+% Rearranging, we get that $dx = d\phi / \sqrt{-V(\phi)}$, so where $\phi(x)$ is monotone, the inverse is
+% \[
+%     x(\phi) = x(\phi^*) \pm \int_{\phi^*}^\phi \frac{ d\psi }{ \sqrt{ -V(\psi) } } .
+% \]
+% For each $i$ then define the elliptic function
+% \begin{align}  \label{eqn:elliptic_function}
+%     F_i(\phi) = \int_{\phi_i^*}^\phi \frac{ d\psi }{ \sqrt{ -V_i(\psi) } } ,
+% \end{align}
+% where take the positive branch of the square root, and $\phi_i^*$ will be chosen later.
+% Then we have that
+% $x(\phi) - x(\phi_0) = \pm( F(\phi) - F(\phi_0))$,
+% or for an appropriate $x_0$,
+% \[
+%     \phi(x) = F^{-1}\left( F(\phi(x_0)) \pm (x - x_0) \right).
+% \]
+% 
+% We clearly want $\lim_{x \to \infty} \partial_x \phi(\pm x) = 0$, 
+% and $\lim_{x \to \infty} \phi(\pm x)$ to be zero or one depending on the sign of $s_0$ and $s_{n+1}$.
+% Since $(\partial_x \phi(x))^2 = -V(x)$, this implies that if $s_0<0$, then $K_0 = 0$,
+% while if $s_0>0$ then $K_0 = s_0/6$; and likewise for $K_{n+1}$.
+% 
+% We also require that $\phi(x)$ and $\phi'(x)$ are continuous.
+% Continuity of $\phi'(x)$ is equivalent to $V_i(x_{i+1}) = V_{i+1}(x_{i+1})$,
+% which we can rearrange to find an equation for $K_{i+1}$ in terms of $K_i$ and $\phi(x_{i+1})$:
+% \[
+%     K_{i+1} - K_i = (s_i - s_{i+1}) \phi(x_{i+1})^2 (1-2\phi(x_{i+1})/3) .
+% \]
+% What about the frequency at the points the selection changes, $\phi(x_i)$?  
+% Well, if $\phi(x)$ is monotone on $[x_i,x_{i+1})$ then we can without loss of generality take $\phi_i^*=0$ or $1$ depending on the sign of $s_i$.
+% Otherwise, let $\phi_i^*$ be the (unique) root of $V_i$ in $[x_i,x_{i+1})$, so $V_i(\phi_i^*)=0$.
+% In this case, $\phi_i^*$ is the maximum or minimum of $\phi$ in the interval:
+% if $s>0$, then $\phi_i = \max\{ \phi(x) : x \in [x_i,x_{i+1})\}$. 
+% Recall we defined $F_i$ using $\phi_i^*$; now using the fact that $\phi$ is monotone with the opposite sign on either side of $\phi_i^*$,
+% $x_{i+1} - x_i = F_i(\phi(x_{i+1})) - F_i(\phi_i^*) + F_i(\phi(x_{i})) - F_i(\phi_i^*)$,
+% and that $F(\phi_i^*) = 0$,
+% we know that the length of the $i$th stretch is
+% \[
+%     x_{i+1} - x_i = \pm \left( F_i(\phi(x_i)) + F_i(\phi(x_{i+1})) \right).
+% \]
+% Note that all the $\pm{}$'s are easily relatable to the signs of $s_i$.
+% If we knew $\phi(x_1)$ and $\phi'(x_1)$, then we'd be able to solve the equations for $\phi(x_i)$ and $K_i$ recursively upwards.
+% In some cases, such as \citet{slatkin1973geneflow}, we can infer $\phi(0)$ and $\partial_x \phi(0)$ by spatial symmetry.
+% In other cases, we are only given $\phi(-\infty)$ and $\phi(\infty)$, and have to work inwards from the ends.
+% 
+% 
+% \paragraph{Doing the integrals}
+% \label{apx:some_integrals}
+% 
+% An important ingredient in the above method is the integral \eqref{eqn:elliptic_function},
+% to which a method for solving the Korteweg-de Vries equation can be applied \citep{NEQwiki}.
+% Recall that $V_i(\phi) =  s_i \phi^2(1-2\phi/3) + K_i$, with $K_i = V_i(0)$ chosen to match $V$ at the boundaries;
+% since we're just working within an interval with $s$ constant, we can rescale space by $\sqrt{2 s_i}/\sigma$,
+% so that $s$ and $\sigma$ drop from the equation.
+% We then want to integrate
+% \[
+%     \int \frac{ d\phi }{ \sqrt{-V(\phi)} } = 
+%          \int \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } ,
+% \]
+% over a domain where $V$ is always negative.
+% Let $\phi^2(1-2\phi/3)+K = (\alpha-\phi)(\phi-\beta)(\phi-\gamma)$,
+% and change variables first to $y^2=(\alpha-\phi)$, 
+% and then to $x = y/\sqrt{\alpha-\beta}$, so that
+% \begin{align*}
+%     \frac{ d\phi }{ \sqrt{ \phi^2 (1-2\phi/3) + K } } 
+%         = \frac{ - 2 dy }{ \sqrt{ (\alpha-\beta-y^2) (\alpha-\gamma-y^2) } } \\
+%         = \frac{ - 2 dx }{ \sqrt{\alpha-\gamma} \sqrt{ (1-x^2) (1-S^2 x^2) } } ,
+% \end{align*}
+% with $S^2 = (\alpha-\beta)/(\alpha-\gamma)$.
+% Now Jacobi's incomplete elliptic integral of the first kind is defined by
+% \[
+%     F(x;k) = \int_0^x \frac{dt}{\sqrt{ (1-t^2)(1-k^2t^2) }} ,
+% \]
+% and the Jacobian elliptic function $\sn(x;k)$ is the inverse: $F(\sn(x;k);k) = x$.
+% As $k \to 0$, $\sn(x;k) \to \sin(x)$, while as $k \to 1$, $\sn(x;k) \to \sinh(x)$.
+% 
+%
 
@@
 
+\subsection*{Text S1}
+
 \section*{Acknowledgements}
@@
 for a number of insightful suggestions,
-and catching several errors present in earlier drafts. \revpoint{3}{4}
+and catching several errors present in earlier drafts. 
 We would also like to acknowledge Michael Nachman, Steve Evans, and Yaniv Brandvain for useful discussions, 

### 2015-10-19 patchy-selection-paper-for-PLoS.tex
final whatnot for submission

@@
 as spatial branching processes.
-A main result is that locally adapted alleles in different patches
-are likely to be of independent origin
-if the distance between the patches is larger than
-$(\sigma/s_m)\log(Cs_m/\mu)$, 
-where $\sigma$ is the dispersal distance, 
-$s_m$ is the selective disadvantage of these alleles between patches,
-$\mu$ is the mutation rate, 
-and $C$ depends on the probability of establishment and patch geometry.
+A main result is that patches further apart than a critical distance will
+likely evolve independent locally adapted alleles;
+this distance is proportional to the spatial scale of selection
+($\sigma/\sqrt{s_m}$, where $\sigma$ is the dispersal distance and
+$s_m$ is the selective disadvantage of these alleles between patches),
+and depends linearly on $\log(s_m/\mu)$, where $\mu$ is the mutation rate.
 This provides a way to understand the role of geographic separation between patches 
@@
 First we briefly describe the simulations we used for illustration and validation
-(the \texttt{R} code used is provided as a supplement).
+(the \texttt{R} code used is provided in S1 Scripts and at \url{http://github.com/petrelharp/spatial_selection}).
 We simulated forward-time dynamics of the number of alleles of each type in a rectangular grid (either one- or two-dimensional) of demes with fixed size $N$.
@@
 \textbf{
-Here we describe how to solve \eqref{eqn:radial_eqfreq_diffeq} exactly in one dimension,
+Description of how to solve \eqref{eqn:radial_eqfreq_diffeq} exactly in one dimension,
 i.e., the Fisher-KPP equation with piecewise constant selection.
@@
 
+\subsection*{S1 Scripts}
+
+\textbf{Scripts used in producing this manuscript.}
+These can also be found at \url{http://github.com/petrelharp/spatial_selection}.
+
 \subsection*{S1 Table}
@@
 
+% % latex table generated in R 3.1.2 by xtable 1.7-4 package
+% % Wed Dec 17 14:13:09 2014
+% % \begin{table}[ht]
+% \begin{center}
+% {\tiny
+% \begin{tabular}{|rrrrrrr||rrrrrrr|}
+%   \hline
+% $\mu$ & $\rho$ & $s_p$ & $s_m$ & $T_\text{adapt}$ & $p_\text{adapted}$ & n & $\mu$ & $\rho$ & $s_p$ & $s_m$ & $T_\text{adapt}$ & $p_\text{adapted}$ & n \\ 
+%   \hline
+% 1e-05 & 25 & 0.01 & -0.10 & 19630 & 0.06 & 100 & 1e-05 & 600 & 0.01 & -0.10 & 1082 & 1.00 & 40 \\ 
+%   1e-05 & 25 & 0.01 & -0.01 & 15533 & 0.55 & 100 & 1e-05 & 600 & 0.01 & -0.01 & 924 & 1.00 & 40 \\ 
+%   1e-05 & 25 & 0.01 & -0.00 & 14353 & 0.72 & 100 & 1e-05 & 600 & 0.01 & -0.01 & 828 & 1.00 & 40 \\ 
+%   1e-05 & 25 & 0.01 & -0.00 & 13282 & 0.74 & 100 & 1e-05 & 600 & 0.01 & -0.00 & 846 & 1.00 & 40 \\ 
+%   1e-05 & 50 & 0.01 & -0.10 & 11587 & 0.33 & 40 & 1e-05 & 600 & 0.01 & -0.00 & 966 & 1.00 & 40 \\ 
+%   1e-05 & 50 & 0.01 & -0.10 & 11586 & 0.38 & 40 & 1e-05 & 800 & 0.01 & -0.10 & 802 & 1.00 & 40 \\ 
+%   1e-05 & 50 & 0.01 & -0.01 & 9419 & 0.88 & 40 & 1e-05 & 800 & 0.01 & -0.10 & 1182 & 1.00 & 40 \\ 
+%   1e-05 & 50 & 0.01 & -0.01 & 8848 & 0.82 & 40 & 1e-05 & 800 & 0.01 & -0.01 & 846 & 1.00 & 40 \\ 
+%   1e-05 & 50 & 0.01 & -0.00 & 6529 & 0.95 & 40 & 1e-05 & 800 & 0.01 & -0.01 & 607 & 1.00 & 40 \\ 
+%   1e-05 & 50 & 0.01 & -0.00 & 9986 & 0.90 & 40 & 1e-05 & 800 & 0.01 & -0.00 & 646 & 1.00 & 40 \\ 
+%   1e-05 & 100 & 0.01 & -0.10 & 7425 & 0.91 & 316 & 1e-05 & 800 & 0.01 & -0.00 & 507 & 1.00 & 40 \\ 
+%   1e-06 & 100 & 0.01 & -0.10 & 21010 & 0.23 & 217 & 1e-05 & 1000 & 0.01 & -0.10 & 615 & 1.00 & 256 \\ 
+%   1e-05 & 100 & 0.01 & -0.03 & 6036 & 0.98 & 195 & 2.4e-06 & 1000 & 0.01 & -0.10 & 5309 & 0.96 & 257 \\ 
+%   1e-06 & 100 & 0.01 & -0.03 & 19859 & 0.31 & 194 & 1e-05 & 1000 & 0.01 & -0.03 & 523 & 1.00 & 194 \\ 
+%   1e-05 & 100 & 0.01 & -0.01 & 5408 & 0.98 & 315 & 1e-06 & 1000 & 0.01 & -0.03 & 6425 & 0.95 & 195 \\ 
+%   1e-06 & 100 & 0.01 & -0.01 & 20606 & 0.28 & 216 & 1e-05 & 1000 & 0.01 & -0.01 & 449 & 1.00 & 254 \\ 
+%   1e-05 & 100 & 0.01 & -0.00 & 4745 & 0.99 & 195 & 2.4e-06 & 1000 & 0.01 & -0.01 & 4806 & 0.96 & 256 \\ 
+%   1e-06 & 100 & 0.01 & -0.00 & 19564 & 0.33 & 195 & 1e-05 & 1000 & 0.01 & -0.00 & 460 & 1.00 & 193 \\ 
+%   1e-05 & 100 & 0.01 & -0.00 & 4894 & 0.99 & 315 & 1e-06 & 1000 & 0.01 & -0.00 & 5203 & 0.98 & 194 \\ 
+%   1e-06 & 100 & 0.01 & -0.00 & 19417 & 0.33 & 216 & 1e-05 & 1000 & 0.01 & -0.00 & 444 & 1.00 & 255 \\ 
+%   1e-05 & 100 & 0.01 & -0.00 & 4940 & 0.99 & 316 & 2.4e-06 & 1000 & 0.01 & -0.00 & 4037 & 1.00 & 257 \\ 
+%   1e-06 & 100 & 0.01 & -0.00 & 18025 & 0.40 & 217 & 1e-05 & 1000 & 0.01 & -0.00 & 404 & 1.00 & 216 \\ 
+%   1e-05 & 200 & 0.01 & -0.10 & 4296 & 0.97 & 40 & 1e-06 & 1000 & 0.01 & -0.00 & 4567 & 1.00 & 216 \\ 
+%   1e-05 & 200 & 0.01 & -0.10 & 3736 & 1.00 & 40 & 1e-05 & 1200 & 0.01 & -0.10 & 501 & 1.00 & 40 \\ 
+%   1e-05 & 200 & 0.01 & -0.01 & 2943 & 1.00 & 40 & 1e-05 & 1200 & 0.01 & -0.10 & 460 & 1.00 & 40 \\ 
+%   1e-05 & 200 & 0.01 & -0.01 & 2402 & 1.00 & 40 & 1e-05 & 1200 & 0.01 & -0.01 & 336 & 1.00 & 40 \\ 
+%   1e-05 & 200 & 0.01 & -0.00 & 1899 & 1.00 & 40 & 1e-05 & 1200 & 0.01 & -0.01 & 431 & 1.00 & 40 \\ 
+%   1e-05 & 200 & 0.01 & -0.00 & 1992 & 1.00 & 40 & 1e-05 & 1200 & 0.01 & -0.00 & 390 & 1.00 & 40 \\ 
+%   1e-05 & 400 & 0.01 & -0.10 & 1740 & 1.00 & 100 & 1e-05 & 1200 & 0.01 & -0.00 & 372 & 1.00 & 40 \\ 
+%   1e-05 & 400 & 0.01 & -0.01 & 1471 & 1.00 & 100 & 1e-05 & 1600 & 0.01 & -0.10 & 299 & 1.00 & 100 \\ 
+%   1e-05 & 400 & 0.01 & -0.00 & 1446 & 1.00 & 100 & 1e-05 & 1600 & 0.01 & -0.01 & 462 & 0.99 & 100 \\ 
+%   1e-05 & 400 & 0.01 & -0.00 & 1224 & 1.00 & 100 & 1e-05 & 1600 & 0.01 & -0.00 & 241 & 1.00 & 100 \\ 
+%   1e-05 & 600 & 0.01 & -0.10 & 1276 & 1.00 & 40 & 1e-05 & 1600 & 0.01 & -0.00 & 274 & 1.00 & 100 \\ 
+%    \hline
+% \end{tabular}
+% }
+%\caption
+{
     \textbf{Parameter values used in estimates of mean time to adaptation by mutation of figure~\ref{fig:sim_times}.}
@@
             The simulation began with no $B$ alleles.
-
-\subsection*{S1 Table}
+            } 
+% \label{stab:mutation_params}
+% \end{center}
+%\end{table}
+
+\subsection*{S2 Table}
  \label{stab:migration_params}
@@
 
+% latex table generated in R 3.1.2 by xtable 1.7-4 package
+% Wed Dec 17 14:14:44 2014
+%\begin{table}[ht]
+%\begin{center}
+%{\tiny
+%\begin{tabular}{|rrrrrrr||rrrrrrr|}
+%  \hline
+%R & $\rho$ & $s_p$ & $s_m$ & $T_\text{adapt}$ & $p_\text{adapted}$ & n & R & $\rho$ & $s_p$ & $s_m$ & $T_\text{adapt}$ & $p_\text{adapted}$ & n \\ 
+%  \hline
+%100 & 100 & 0.01 & -1e-01 & 24975 & 0.00 & 300 & 60 & 1000 & 0.01 & -3e-02 & 22224 & 0.21 & 200 \\ 
+%  10 & 100 & 0.01 & -1e-01 & 24975 & 0.00 & 40 & 100 & 1000 & 0.01 & -1e-02 & 18098 & 0.55 & 299 \\ 
+%  120 & 100 & 0.01 & -1e-01 & 24975 & 0.00 & 295 & 10 & 1000 & 0.01 & -1e-02 & 14 & 1.00 & 100 \\ 
+%  160 & 100 & 0.01 & -1e-01 & 24975 & 0.00 & 40 & 120 & 1000 & 0.01 & -1e-02 & 23242 & 0.14 & 297 \\ 
+%  20 & 100 & 0.01 & -1e-01 & 24975 & 0.00 & 40 & 160 & 1000 & 0.01 & -1e-02 & 24762 & 0.01 & 100 \\ 
+%  40 & 100 & 0.01 & -1e-01 & 24975 & 0.00 & 40 & 20 & 1000 & 0.01 & -1e-02 & 136 & 1.00 & 100 \\ 
+%  60 & 100 & 0.01 & -1e-01 & 24975 & 0.00 & 300 & 40 & 1000 & 0.01 & -1e-02 & 874 & 1.00 & 99 \\ 
+%  80 & 100 & 0.01 & -1e-01 & 24975 & 0.00 & 40 & 60 & 1000 & 0.01 & -1e-02 & 3485 & 1.00 & 300 \\ 
+%  100 & 100 & 0.01 & -3e-02 & 24975 & 0.00 & 200 & 80 & 1000 & 0.01 & -1e-02 & 9204 & 0.96 & 100 \\ 
+%  120 & 100 & 0.01 & -3e-02 & 24975 & 0.00 & 199 & 100 & 1000 & 0.01 & -3e-03 & 4168 & 1.00 & 200 \\ 
+%  60 & 100 & 0.01 & -3e-02 & 24903 & 0.01 & 200 & 120 & 1000 & 0.01 & -3e-03 & 6304 & 1.00 & 200 \\ 
+%  100 & 100 & 0.01 & -1e-02 & 24354 & 0.04 & 300 & 60 & 1000 & 0.01 & -3e-03 & 1273 & 1.00 & 199 \\ 
+%  10 & 100 & 0.01 & -1e-02 & 1705 & 0.95 & 40 & 100 & 1000 & 0.01 & -1e-03 & 2813 & 1.00 & 299 \\ 
+%  120 & 100 & 0.01 & -1e-02 & 24802 & 0.01 & 299 & 10 & 1000 & 0.01 & -1e-03 & 12 & 1.00 & 100 \\ 
+%  160 & 100 & 0.01 & -1e-02 & 24975 & 0.00 & 40 & 120 & 1000 & 0.01 & -1e-03 & 4002 & 1.00 & 300 \\ 
+%  20 & 100 & 0.01 & -1e-02 & 3601 & 0.95 & 40 & 160 & 1000 & 0.01 & -1e-03 & 7386 & 1.00 & 100 \\ 
+%  40 & 100 & 0.01 & -1e-02 & 9463 & 0.88 & 40 & 20 & 1000 & 0.01 & -1e-03 & 121 & 1.00 & 100 \\ 
+%  60 & 100 & 0.01 & -1e-02 & 18674 & 0.45 & 299 & 40 & 1000 & 0.01 & -1e-03 & 507 & 1.00 & 100 \\ 
+%  80 & 100 & 0.01 & -1e-02 & 23899 & 0.12 & 40 & 60 & 1000 & 0.01 & -1e-03 & 1115 & 1.00 & 300 \\ 
+%  100 & 100 & 0.01 & -3e-03 & 14601 & 0.80 & 200 & 80 & 1000 & 0.01 & -1e-03 & 1941 & 1.00 & 100 \\ 
+%  120 & 100 & 0.01 & -3e-03 & 18645 & 0.53 & 199 & 100 & 1000 & 0.01 & -1e-04 & 2528 & 1.00 & 300 \\ 
+%  60 & 100 & 0.01 & -3e-03 & 5424 & 0.98 & 200 & 10 & 1000 & 0.01 & -1e-04 & 11 & 1.00 & 99 \\ 
+%  100 & 100 & 0.01 & -1e-03 & 8992 & 0.96 & 299 & 120 & 1000 & 0.01 & -1e-04 & 3590 & 1.00 & 298 \\ 
+%  10 & 100 & 0.01 & -1e-03 & 522 & 1.00 & 40 & 160 & 1000 & 0.01 & -1e-04 & 5648 & 1.00 & 100 \\ 
+%  120 & 100 & 0.01 & -1e-03 & 11375 & 0.93 & 300 & 20 & 1000 & 0.01 & -1e-04 & 122 & 1.00 & 100 \\ 
+%  160 & 100 & 0.01 & -1e-03 & 16930 & 0.68 & 40 & 40 & 1000 & 0.01 & -1e-04 & 498 & 1.00 & 100 \\ 
+%  20 & 100 & 0.01 & -1e-03 & 1518 & 0.97 & 40 & 60 & 1000 & 0.01 & -1e-04 & 1051 & 1.00 & 298 \\ 
+%  40 & 100 & 0.01 & -1e-03 & 2234 & 1.00 & 40 & 80 & 1000 & 0.01 & -1e-04 & 1796 & 1.00 & 100 \\ 
+%  60 & 100 & 0.01 & -1e-03 & 4277 & 0.98 & 300 & 10 & 4000 & 0.01 & -1e-01 & 0 & 1.00 & 20 \\ 
+%  80 & 100 & 0.01 & -1e-03 & 5328 & 1.00 & 40 & 160 & 4000 & 0.01 & -1e-01 & 24975 & 0.00 & 20 \\ 
+%  100 & 100 & 0.01 & -1e-04 & 6867 & 0.99 & 299 & 20 & 4000 & 0.01 & -1e-01 & 4406 & 1.00 & 20 \\ 
+%  10 & 100 & 0.01 & -1e-04 & 461 & 1.00 & 40 & 40 & 4000 & 0.01 & -1e-01 & 24975 & 0.00 & 20 \\ 
+%  120 & 100 & 0.01 & -1e-04 & 9347 & 0.99 & 300 & 80 & 4000 & 0.01 & -1e-01 & 24975 & 0.00 & 20 \\ 
+%  160 & 100 & 0.01 & -1e-04 & 12731 & 0.90 & 40 & 10 & 4000 & 0.01 & -1e-02 & 3746 & 0.85 & 20 \\ 
+%  20 & 100 & 0.01 & -1e-04 & 860 & 1.00 & 40 & 160 & 4000 & 0.01 & -1e-02 & 24975 & 0.00 & 20 \\ 
+%  40 & 100 & 0.01 & -1e-04 & 2031 & 1.00 & 40 & 20 & 4000 & 0.01 & -1e-02 & 58 & 1.00 & 20 \\ 
+%  60 & 100 & 0.01 & -1e-04 & 3529 & 0.99 & 300 & 40 & 4000 & 0.01 & -1e-02 & 362 & 1.00 & 20 \\ 
+%  80 & 100 & 0.01 & -1e-04 & 4839 & 1.00 & 40 & 80 & 4000 & 0.01 & -1e-02 & 4025 & 1.00 & 20 \\ 
+%  100 & 1000 & 0.01 & -1e-01 & 24975 & 0.00 & 300 & 10 & 4000 & 0.01 & -1e-03 & 6244 & 0.75 & 20 \\ 
+%  10 & 1000 & 0.01 & -1e-01 & 1004 & 0.99 & 100 & 160 & 4000 & 0.01 & -1e-03 & 4399 & 1.00 & 20 \\ 
+%  120 & 1000 & 0.01 & -1e-01 & 24975 & 0.00 & 298 & 20 & 4000 & 0.01 & -1e-03 & 52 & 1.00 & 20 \\ 
+%  160 & 1000 & 0.01 & -1e-01 & 24975 & 0.00 & 100 & 40 & 4000 & 0.01 & -1e-03 & 281 & 1.00 & 20 \\ 
+%  20 & 1000 & 0.01 & -1e-01 & 18403 & 0.42 & 100 & 80 & 4000 & 0.01 & -1e-03 & 1209 & 1.00 & 20 \\ 
+%  40 & 1000 & 0.01 & -1e-01 & 24965 & 0.00 & 100 & 10 & 4000 & 0.01 & -1e-04 & 3746 & 0.85 & 20 \\ 
+%  60 & 1000 & 0.01 & -1e-01 & 24975 & 0.00 & 300 & 160 & 4000 & 0.01 & -1e-04 & 3798 & 1.00 & 20 \\ 
+%  80 & 1000 & 0.01 & -1e-01 & 24975 & 0.00 & 100 & 20 & 4000 & 0.01 & -1e-04 & 52 & 1.00 & 20 \\ 
+%  100 & 1000 & 0.01 & -3e-02 & 24975 & 0.00 & 200 & 40 & 4000 & 0.01 & -1e-04 & 279 & 1.00 & 20 \\ 
+%  120 & 1000 & 0.01 & -3e-02 & 24975 & 0.00 & 200 & 80 & 4000 & 0.01 & -1e-04 & 1039 & 1.00 & 20 \\ 
+%   \hline
+%\end{tabular}
+%}
+%%\caption
+{
     \textbf{Parameter values used in estimates of mean time to adaptation by migration of figure~\ref{fig:sim_times}.}
@@
             At the start of the simulation, one patch was initialized with a frequency of 0.8 $B$ alleles, which were absent elsewhere.
+            } 
+% \label{stab:migration_params}
+% \end{center}
+%\end{table}
 
@@
 
-\subsection*{Text S1}
-
 \section*{Acknowledgements}
@@
 % \bibliography{standing_patches_refs}
-\begin{thebibliography}{68}
-\providecommand{\natexlab}[1]{#1}
-\providecommand{\url}[1]{\texttt{#1}}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi: #1}\else
-  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
-
-\bibitem[Aldous(1985)]{aldous1985exchangeability}
-D.~J. Aldous.
-\newblock Exchangeability and related topics.
-\newblock In \emph{École d'été de probabilités de {S}aint-{F}lour,
-  {XIII}---1983}, volume 1117 of \emph{Lecture Notes in Math.}, pages 1--198.
-  Springer, Berlin, 1985.
-\newblock URL
-  \url{http://www.springerlink.com/content/c31v17440871210x/fulltext.pdf}.
-
-\bibitem[Allred and Beck(1963)]{allred1963range}
-D.~M. Allred and D.~E. Beck.
-\newblock Range of movement and dispersal of some rodents at the {Nevada}
-  atomic test site.
-\newblock \emph{Journal of Mammalogy}, 44\penalty0 (2):\penalty0 pp. 190--200,
-  1963.
-\newblock ISSN 00222372.
-\newblock URL \url{http://www.jstor.org/stable/1377452}.
-
-\bibitem[Arendt and Reznick(2008)]{Arendt:08}
-J.~Arendt and D.~Reznick.
+\begin{thebibliography}{10}
+
+\bibitem{stern2013genetic}
+Stern DL.
+\newblock The genetic causes of convergent evolution.
+\newblock Nat Rev Genet. 2013 Nov;14(11):751--764.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/24105273}.
+
+\bibitem{zhen2012parallel}
+Zhen Y, Aardema ML, Medina EM, Schumer M, Andolfatto P.
+\newblock Parallel Molecular Evolution in an Herbivore Community.
+\newblock Science. 2012;337(6102):1634--1637.
+\newblock Available from:
+  \url{http://www.sciencemag.org/content/337/6102/1634.abstract}.
+
+\bibitem{martin2013repeated}
+Martin A, Orgogozo V.
+\newblock The loci of repeated evolution: a catalog of genetic hotspots of
+  phenotypic variation.
+\newblock Evolution. 2013;67(5):1235--1250.
+\newblock Available from: \url{http://dx.doi.org/10.1111/evo.12081}.
+
+\bibitem{Arendt:08}
+Arendt J, Reznick D.
 \newblock Convergence and parallelism reconsidered: what have we learned about
   the genetics of adaptation?
-\newblock \emph{Trends Ecol Evol}, 23\penalty0 (1):\penalty0 26--32, Jan. 2008.
-\newblock \doi{10.1016/j.tree.2007.09.011}.
-
-\bibitem[Barton(1979)]{barton1979geneflow}
-N.~Barton.
-\newblock Gene flow past a cline.
-\newblock \emph{Heredity}, 43\penalty0 (3):\penalty0 333--339, Dec. 1979.
-\newblock ISSN 0018-067X.
-\newblock URL \url{http://dx.doi.org/10.1038/hdy.1979.86}.
-
-\bibitem[Barton(1987)]{barton1987establishment}
-N.~H. Barton.
+\newblock Trends Ecol Evol. 2008 Jan;23(1):26--32.
+
+\bibitem{softsweepsII}
+Pennings PS, Hermisson J.
+\newblock Soft Sweeps {II} - Molecular Population Genetics of Adaptation from
+  Recurrent Mutation or Migration.
+\newblock Mol Biol Evol. 2006;p. msj117.
+\newblock Available from:
+  \url{http://mbe.oxfordjournals.org/cgi/content/abstract/msj117v1}.
+
+\bibitem{ralph2010parallel}
+Ralph P, Coop G.
+\newblock Parallel Adaptation: One or Many Waves of Advance of an Advantageous
+  Allele?
+\newblock Genetics. 2010;186(2):647--668.
+\newblock Available from:
+  \url{http://www.genetics.org/cgi/content/abstract/186/2/647}.
+
+\bibitem{orr2001sieve}
+Orr HA, Betancourt AJ.
+\newblock Haldane's sieve and adaptation from the standing genetic variation.
+\newblock Genetics. 2001 Feb;157(2):875--884.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/11157004}.
+
+\bibitem{softsweepsI}
+Hermisson J, Pennings PS.
+\newblock Soft sweeps: molecular population genetics of adaptation from
+  standing genetic variation.
+\newblock Genetics. 2005 Apr;169(4):2335--2352.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/15716498}.
+
+\bibitem{slatkin1973geneflow}
+Slatkin M.
+\newblock Gene Flow and Selection in a Cline.
+\newblock Genetics. 1973;75(4):733--756.
+\newblock Available from:
+  \url{http://www.genetics.org/cgi/content/abstract/75/4/733}.
+
+\bibitem{kruckeberg1951intraspecific}
+Kruckeberg AR.
+\newblock Intraspecific Variability in the Response of Certain Native Plant
+  Species to Serpentine Soil.
+\newblock American Journal of Botany. 1951;38(6):pp. 408--419.
+\newblock Available from: \url{http://www.jstor.org/stable/2438248}.
+
+\bibitem{macnair1991evolution}
+Macnair MR.
+\newblock Why the evolution of resistance to anthropogenic toxins normally
+  involves major gene changes: the limits to natural selection.
+\newblock Genetica. 1991;84(3):213--219.
+\newblock Available from: \url{http://dx.doi.org/10.1007/BF00127250}.
+
+\bibitem{schat1996identical}
+Schat H, Vooijs R, Kuiper E.
+\newblock Identical Major Gene Loci for Heavy Metal Tolerances that Have
+  Independently Evolved in Different Local Populations and Subspecies of
+  \textit{{Silene} vulgaris}.
+\newblock Evolution. 1996;50(5):pp. 1888--1895.
+\newblock Available from: \url{http://www.jstor.org/stable/2410747}.
+
+\bibitem{turner2010serpentine}
+Turner TL, Bourne EC, Von~Wettberg EJ, Hu TT, Nuzhdin SV.
+\newblock Population resequencing reveals local adaptation of
+  \textit{{Arabidopsis} lyrata} to serpentine soils.
+\newblock Nat Genet. 2010 Jan;42(3):260--3.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/20101244}.
+
+\bibitem{benson1933concealing}
+Benson SB.
+\newblock Concealing coloration among some desert rodents of the southwestern
+  United States.
+\newblock No. v. 40 in University of California publications in zoology.
+  University of California Press; 1933.
+\newblock Available from: \url{http://books.google.com/books?id=Lis-AQAAIAAJ}.
+
+\bibitem{DiceBlossom1937}
+Dice LR, Blossom PM.
+\newblock Studies of Mammalian Ecology in Southwestern North America With
+  Special Attention to the Colors of Desert Mammals.
+\newblock Carnegie Institution; 1937.
+
+\bibitem{kaufman1974adaptive}
+Kaufman DW.
+\newblock Adaptive Coloration in \textit{{Peromyscus} polionotus}: Experimental
+  Selection by Owls.
+\newblock Journal of Mammalogy. 1974;55(2):pp. 271--283.
+\newblock Available from: \url{http://www.jstor.org/stable/1378997}.
+
+\bibitem{hoekstra2003different}
+Hoekstra HE, Nachman MW.
+\newblock Different genes underlie adaptive melanism in different populations
+  of rock pocket mice.
+\newblock Mol Ecol. 2003 May;12:1185--1194.
+\newblock Available from:
+  \url{http://www3.interscience.wiley.com/journal/118890526/abstract}.
+
+\bibitem{dice1940ecologic}
+Dice LR.
+\newblock Ecologic and Genetic Variability within Species of
+  \textit{{Peromyscus}}.
+\newblock The American Naturalist. 1940;74(752):pp. 212--221.
+\newblock Available from: \url{http://www.jstor.org/stable/2457573}.
+
+\bibitem{steiner2009genetic}
+Steiner CC, Rompler H, Boettger LM, Schoneberg T, Hoekstra HE.
+\newblock The Genetic Basis of Phenotypic Convergence in Beach Mice: Similar
+  Pigment Patterns but Different Genes.
+\newblock Mol Biol Evol. 2009;26(1):35--45.
+\newblock Available from:
+  \url{http://mbe.oxfordjournals.org/cgi/content/abstract/26/1/35}.
+
+\bibitem{kingsley2009melanism}
+Kingsley EP, Manceau M, Wiley CD, Hoekstra HE.
+\newblock Melanism in \textit{{Peromyscus}} is caused by independent mutations
+  in agouti.
+\newblock PLoS ONE. 2009;4:e6435.
+\newblock Available from:
+  \url{http://dx.doi.org/10.1371%2Fjournal.pone.0006435}.
+
+\bibitem{rosenblum2010molecular}
+Rosenblum EB, Römpler H, Schöneberg T, Hoekstra HE.
+\newblock Molecular and functional basis of phenotypic convergence in white
+  lizards at {W}hite {S}ands.
+\newblock Proceedings of the National Academy of Sciences.
+  2010;107(5):2113--2117.
+\newblock Available from:
+  \url{http://www.pnas.org/content/107/5/2113.abstract}.
+
+\bibitem{levin2003ecology}
+Levin SA, Muller-Landau HC, Nathan R, Chave J.
+\newblock The Ecology and Evolution of Seed Dispersal: a theoretical
+  perspective.
+\newblock Annu Rev Ecol Evol Syst. 2003;34:575--604.
+\newblock Available from:
+  \url{http://arjournals.annualreviews.org/doi/pdf/10.1146/annurev.ecolsys.34.011802.132428}.
+
+\bibitem{hallatschek2014acceleration}
+Hallatschek O, Fisher DS.
+\newblock Acceleration of evolutionary spread by long-range dispersal.
+\newblock Proc Natl Acad Sci U S A. 2014 Nov;111(46):4911--4919.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/25368183}.
+
+\bibitem{haldane1948theory}
+Haldane JBS.
+\newblock The theory of a cline.
+\newblock J Genet. 1948 Jan;48(3):277--284.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/18905075}.
+
+\bibitem{fisher1950frequencies}
+Fisher RA.
+\newblock Gene Frequencies in a Cline Determined by Selection and Diffusion.
+\newblock Biometrics. 1950;6(4):pp. 353--361.
+\newblock Available from: \url{http://www.jstor.org/stable/3001780}.
+
+\bibitem{nagylaki1975conditions}
+Nagylaki T.
+\newblock Conditions for the existence of clines.
+\newblock Genetics. 1975;80(3):595--615.
+\newblock Available from:
+  \url{http://www.genetics.org/content/80/3/595.abstract}.
+
+\bibitem{conley1975application}
+Conley C.
+\newblock An application of {Wazewski}'s method to a non-linear boundary value
+  problem which arises in population genetics.
+\newblock Journal of Mathematical Biology. 1975;2(3):241--249.
+\newblock Available from: \url{http://dx.doi.org/10.1007/BF00277153}.
+
+\bibitem{lenormand2002limits}
+Lenormand T.
+\newblock Gene flow and the limits to natural selection.
+\newblock Trends in Ecology \& Evolution. 2002;17(4):183 -- 189.
+\newblock Available from:
+  \url{http://www.sciencedirect.com/science/article/pii/S0169534702024977}.
+
+\bibitem{barton1987establishment}
+Barton NH.
 \newblock The probability of establishment of an advantageous mutant in a
   subdivided population.
-\newblock \emph{Genetics Research}, 50\penalty0 (1):\penalty0 35--40, 1987.
-\newblock \doi{10.1017/S0016672300023314}.
-\newblock URL \url{http://dx.doi.org/10.1017/S0016672300023314}.
-
-\bibitem[Barton(2000)]{barton2000genetic}
-N.~H. Barton.
-\newblock Genetic hitchhiking.
-\newblock \emph{Philos Trans R Soc Lond B Biol Sci}, 355\penalty0
-  (1403):\penalty0 1553--1562, Nov. 2000.
-\newblock \doi{10.1098/rstb.2000.0716}.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC1692896/}.
-
-\bibitem[Barton et~al.(2013)Barton, Etheridge, Kelleher, and
-  Véber]{barton2013genetic}
-N.~H. Barton, A.~M. Etheridge, J.~Kelleher, and A.~Véber.
-\newblock Genetic hitchhiking in spatially extended populations.
-\newblock \emph{Theoretical Population Biology}, \penalty0 (0):\penalty0 --,
-  2013.
-\newblock ISSN 0040-5809.
-\newblock \doi{10.1016/j.tpb.2012.12.001}.
-\newblock URL
-  \url{http://www.sciencedirect.com/science/article/pii/S0040580912001359}.
-
-\bibitem[Benson(1933)]{benson1933concealing}
-S.~B. Benson.
-\newblock \emph{Concealing coloration among some desert rodents of the
-  southwestern United States}.
-\newblock Number v. 40 in University of California publications in zoology.
-  University of California Press, 1933.
-\newblock URL \url{http://books.google.com/books?id=Lis-AQAAIAAJ}.
-
-\bibitem[Borodin and Salminen(2002)]{borodin2002handbook}
-A.~N. Borodin and P.~Salminen.
-\newblock \emph{Handbook of Brownian motion: facts and formulae}.
-\newblock Springer, 2002.
-
-\bibitem[Cantrell and Cosner(1989)]{cantrell1989diffusive}
-R.~S. Cantrell and C.~Cosner.
+\newblock Genetics Research. 1987;50(1):35--40.
+\newblock Available from: \url{http://dx.doi.org/10.1017/S0016672300023314}.
+
+\bibitem{pollak1966survival}
+Pollak E.
+\newblock On the Survival of a Gene in a Subdivided Population.
+\newblock Journal of Applied Probability. 1966;3(1):142--155.
+\newblock Available from: \url{http://www.jstor.org/stable/3212043}.
+
+\bibitem{haldane1927mathematical}
+Haldane JBS.
+\newblock A Mathematical Theory of Natural and Artificial Selection, Part {V}:
+  Selection and Mutation.
+\newblock Mathematical Proceedings of the Cambridge Philosophical Society. 1927
+  7;23(07):838--844.
+\newblock Available from:
+  \url{http://journals.cambridge.org/article_S0305004100015644}.
+
+\bibitem{fisher1930genetical}
+Fisher RA.
+\newblock The genetical theory of natural selection.
+\newblock Oxford: Oxford University Press; 1930.
+\newblock Available from:
+  \url{http://www.archive.org/details/geneticaltheoryo031631mbp}.
+
+\bibitem{maruyama1970fixation}
+Maruyama T.
+\newblock On the fixation probability of mutant genes in a subdivided
+  population.
+\newblock Genetics Research. 1970;15(02):221--225.
+\newblock Available from: \url{http://dx.doi.org/10.1017/S0016672300001543}.
+
+\bibitem{cherry2003diffusion}
+Cherry JL, Wakeley J.
+\newblock A diffusion approximation for selection and drift in a subdivided
+  population.
+\newblock Genetics. 2003 Jan;163(1):421--428.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/12586727}.
+
+\bibitem{cantrell1989diffusive}
+Cantrell RS, Cosner C.
 \newblock Diffusive logistic equations with indefinite weights: population
   models in disrupted environments.
-\newblock \emph{Proceedings of the Royal Society of Edinburgh, Section: A
-  Mathematics}, 112\penalty0 (3-4):\penalty0 293--318, 0 1989.
-\newblock ISSN 1473-7124.
-\newblock \doi{10.1017/S030821050001876X}.
-\newblock URL \url{http://journals.cambridge.org/article_S030821050001876X}.
-
-\bibitem[Cantrell and Cosner(1991)]{cantrell1991diffusive}
-R.~S. Cantrell and C.~Cosner.
-\newblock Diffusive logistic equations with indefinite weights: population
-  models in disrupted environments. {II}.
-\newblock \emph{SIAM J. Math. Anal.}, 22\penalty0 (4):\penalty0 1043--1064,
-  1991.
-\newblock ISSN 0036-1410.
-\newblock \doi{10.1137/0522068}.
-\newblock URL \url{http://dx.doi.org/10.1137/0522068}.
-
-\bibitem[Censky et~al.(1998)Censky, Hodge, and Dudley]{censky1998overwater}
-E.~J. Censky, K.~Hodge, and J.~Dudley.
-\newblock Over-water dispersal of lizards due to hurricanes.
-\newblock \emph{Nature}, 395\penalty0 (6702):\penalty0 556--556, Oct. 1998.
-\newblock ISSN 00280836.
-\newblock URL \url{http://dx.doi.org/10.1038/26886}.
-
-\bibitem[Cherry and Wakeley(2003)]{cherry2003diffusion}
-J.~L. Cherry and J.~Wakeley.
-\newblock A diffusion approximation for selection and drift in a subdivided
-  population.
-\newblock \emph{Genetics}, 163\penalty0 (1):\penalty0 421--428, Jan. 2003.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/12586727}.
-
-\bibitem[Conley(1975)]{conley1975application}
-C.~Conley.
-\newblock An application of {Wazewski}'s method to a non-linear boundary value
-  problem which arises in population genetics.
-\newblock \emph{Journal of Mathematical Biology}, 2\penalty0 (3):\penalty0
-  241--249, 1975.
-\newblock ISSN 0303-6812.
-\newblock \doi{10.1007/BF00277153}.
-\newblock URL \url{http://dx.doi.org/10.1007/BF00277153}.
-
-\bibitem[Dawson(1993)]{dawson1993measurevalued}
-D.~A. Dawson.
-\newblock Measure-valued {Markov} processes.
-\newblock In \emph{École d'Été de Probabilités de Saint-Flour XXI---1991},
-  volume 1541 of \emph{Lecture Notes in Math.}, pages 1--260. Springer, Berlin,
-  1993.
-
-\bibitem[Dice(1940)]{dice1940ecologic}
-L.~R. Dice.
-\newblock Ecologic and genetic variability within species of
-  \textit{{Peromyscus}}.
-\newblock \emph{The American Naturalist}, 74\penalty0 (752):\penalty0 pp.
-  212--221, 1940.
-\newblock ISSN 00030147.
-\newblock URL \url{http://www.jstor.org/stable/2457573}.
-
-\bibitem[Dice and Blossom(1937)]{DiceBlossom1937}
-L.~R. Dice and P.~M. Blossom.
-\newblock \emph{Studies of Mammalian Ecology in Southwestern North America With
-  Special Attention to the Colors of Desert Mammals}.
-\newblock Carnegie Institution, 1937.
-
-\bibitem[Donnelly and Joyce(1989)]{donnelly1989continuity}
-P.~Donnelly and P.~Joyce.
-\newblock Continuity and weak convergence of ranked and size-biased
-  permutations on the infinite simplex.
-\newblock \emph{Stochastic Process. Appl.}, 31\penalty0 (1):\penalty0 89--103,
-  1989.
-\newblock ISSN 0304-4149.
-\newblock URL
-  \url{http://www.sciencedirect.com/science/article/pii/030441498990104X}.
-
-\bibitem[Etheridge(2000)]{etheridge2000introduction}
-A.~M. Etheridge.
-\newblock \emph{An introduction to superprocesses}, volume~20 of
-  \emph{University Lecture Series}.
-\newblock American Mathematical Society, Providence, RI, 2000.
-\newblock ISBN 0-8218-2706-5.
-
-\bibitem[Fisher(1937)]{fisher1937wave}
-R.~Fisher.
-\newblock {The wave of advance of advantageous genes}.
-\newblock \emph{Ann. Eugenics}, 7:\penalty0 353--369, 1937.
-\newblock URL
-  \url{http://digital.library.adelaide.edu.au/coll/special/fisher/152.pdf}.
-
-\bibitem[Fisher(1930)]{fisher1930genetical}
-R.~A. Fisher.
-\newblock \emph{The genetical theory of natural selection}.
-\newblock Oxford University Press, Oxford, 1930.
-\newblock ISBN 0-19-850440-3.
-\newblock URL \url{http://www.archive.org/details/geneticaltheoryo031631mbp}.
-
-\bibitem[Fisher(1950)]{fisher1950frequencies}
-R.~A. Fisher.
-\newblock Gene frequencies in a cline determined by selection and diffusion.
-\newblock \emph{Biometrics}, 6\penalty0 (4):\penalty0 pp. 353--361, 1950.
-\newblock ISSN 0006341X.
-\newblock URL \url{http://www.jstor.org/stable/3001780}.
-
-\bibitem[French et~al.(1968)French, Tagami, and Hayden]{french1968dispersal}
-N.~R. French, T.~Y. Tagami, and P.~Hayden.
-\newblock Dispersal in a population of desert rodents.
-\newblock \emph{Journal of Mammalogy}, 49\penalty0 (2):\penalty0 pp. 272--280,
-  1968.
-\newblock ISSN 00222372.
-\newblock URL \url{http://www.jstor.org/stable/1377984}.
-
-\bibitem[Geiger(1999)]{geiger1999elementary}
-J.~Geiger.
-\newblock Elementary new proofs of classical limit theorems for
-  {Galton}-{Watson} processes.
-\newblock \emph{Journal of Applied Probability}, 36\penalty0 (2):\penalty0 pp.
-  301--309, 1999.
-\newblock URL \url{http://www.jstor.org/stable/3215457}.
-
-\bibitem[Gradshteyn and Ryzhik(2007)]{gradshteyn2007table}
-I.~S. Gradshteyn and I.~M. Ryzhik.
-\newblock \emph{Table of integrals, series, and products}.
-\newblock Elsevier/Academic Press, Amsterdam, seventh edition, 2007.
-\newblock ISBN 978-0-12-373637-6; 0-12-373637-4.
-\newblock Translated from the Russian, Translation edited and with a preface by
-  Alan Jeffrey and Daniel Zwillinger.
-
-\bibitem[Haldane(1927)]{haldane1927mathematical}
-J.~B.~S. Haldane.
-\newblock A mathematical theory of natural and artificial selection, part {V}:
-  Selection and mutation.
-\newblock \emph{Mathematical Proceedings of the Cambridge Philosophical
-  Society}, 23\penalty0 (07):\penalty0 838--844, 7 1927.
-\newblock ISSN 1469-8064.
-\newblock \doi{10.1017/S0305004100015644}.
-\newblock URL \url{http://journals.cambridge.org/article_S0305004100015644}.
-
-\bibitem[Haldane(1948)]{haldane1948theory}
-J.~B.~S. Haldane.
-\newblock The theory of a cline.
-\newblock \emph{J Genet}, 48\penalty0 (3):\penalty0 277--284, Jan. 1948.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/18905075}.
-
-\bibitem[Hallatschek and Fisher(2014)]{hallatschek2014acceleration}
-O.~Hallatschek and D.~S. Fisher.
-\newblock Acceleration of evolutionary spread by long-range dispersal.
-\newblock \emph{Proc Natl Acad Sci U S A}, 111\penalty0 (46):\penalty0
-  4911--4919, Nov. 2014.
-\newblock \doi{10.1073/pnas.1404663111}.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/25368183}.
-
-\bibitem[Hermisson and Pennings(2005)]{softsweepsI}
-J.~Hermisson and P.~S. Pennings.
-\newblock Soft sweeps: molecular population genetics of adaptation from
-  standing genetic variation.
-\newblock \emph{Genetics}, 169\penalty0 (4):\penalty0 2335--2352, Apr. 2005.
-\newblock \doi{10.1534/genetics.104.036947}.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/15716498}.
-
-\bibitem[Hoekstra(2006)]{Hoekstra:06}
-H.~E. Hoekstra.
-\newblock {Genetics, development and evolution of adaptive pigmentation in
-  vertebrates}.
-\newblock \emph{Heredity (Edinb)}, 97\penalty0 (3):\penalty0 222--234, Sep
-  2006.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/16823403}.
-
-\bibitem[Hoekstra and Nachman(2003)]{hoekstra2003different}
-H.~E. Hoekstra and M.~W. Nachman.
-\newblock Different genes underlie adaptive melanism in different populations
-  of rock pocket mice.
-\newblock \emph{Mol. Ecol.}, 12:\penalty0 1185--1194, May 2003.
-\newblock URL
-  \url{http://www3.interscience.wiley.com/journal/118890526/abstract}.
-
-\bibitem[Hoekstra et~al.(2004)Hoekstra, Drumm, and
-  Nachman]{hoekstra2004ecological}
-H.~E. Hoekstra, K.~E. Drumm, and M.~W. Nachman.
-\newblock Ecological genetics of adaptive color polymorphism in pocket mice:
-  geographic variation in selected and neutral genes.
-\newblock \emph{Evolution}, 58\penalty0 (6):\penalty0 1329--1341, June 2004.
-\newblock URL
-  \url{http://onlinelibrary.wiley.com/doi/10.1111/j.0014-3820.2004.tb01711.x/abstract}.
-
-\bibitem[Hoekstra et~al.(2005)Hoekstra, Krenz, and Nachman]{hoekstra2004local}
-H.~E. Hoekstra, J.~G. Krenz, and M.~W. Nachman.
-\newblock Local adaptation in the rock pocket mouse (\textit{{Chaetodipus}
-  intermedius}): natural selection and phylogenetic history of populations.
-\newblock \emph{Heredity}, 94\penalty0 (2):\penalty0 217--228, Nov. 2005.
-\newblock ISSN 0018067X.
-\newblock URL \url{http://dx.doi.org/10.1038/sj.hdy.6800600}.
-
-\bibitem[Jagers(1975)]{jagers1975branching}
-P.~Jagers.
-\newblock \emph{Branching processes with biological applications}.
-\newblock Wiley Series in Probability and Statistics: Applied Probability and
-  Statistics Section Series. Wiley, 1975.
-\newblock ISBN 9780471436522.
-
-\bibitem[Kaufman(1974)]{kaufman1974adaptive}
-D.~W. Kaufman.
-\newblock Adaptive coloration in \textit{{Peromyscus} polionotus}: Experimental
-  selection by owls.
-\newblock \emph{Journal of Mammalogy}, 55\penalty0 (2):\penalty0 pp. 271--283,
-  1974.
-\newblock ISSN 00222372.
-\newblock URL \url{http://www.jstor.org/stable/1378997}.
-
-\bibitem[Kim and Maruki(2011)]{kim2011hitchhiking}
-Y.~Kim and T.~Maruki.
-\newblock Hitchhiking effect of a beneficial mutation spreading in a subdivided
-  population.
-\newblock \emph{Genetics}, 189\penalty0 (1):\penalty0 213--226, Sept. 2011.
-\newblock \doi{10.1534/genetics.111.130203}.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3176130/}.
-
-\bibitem[Kingsley et~al.(2009)Kingsley, Manceau, Wiley, and
-  Hoekstra]{kingsley2009melanism}
-E.~P. Kingsley, M.~Manceau, C.~D. Wiley, and H.~E. Hoekstra.
-\newblock Melanism in \textit{{Peromyscus}} is caused by independent mutations
-  in agouti.
-\newblock \emph{PLoS ONE}, 4:\penalty0 e6435, 2009.
-\newblock URL \url{http://dx.doi.org/10.1371%2Fjournal.pone.0006435}.
-
-\bibitem[Kolmogorov et~al.(1991)Kolmogorov, Petrovskii, and
-  Piscunov]{KPP1937book}
-A.~Kolmogorov, I.~Petrovskii, and N.~Piscunov.
-\newblock A study of the equation of diffusion with increase in the quantity of
-  matter, and its application to a biological problem.
-\newblock In \emph{Selected Works of A.N. Kolmogorov: Mathematics and
-  mechanics}, volume~25 of \emph{Mathematics and its Applications (Soviet
-  Series)}, pages 1--25. Kluwer Academic Publishers Group, Dordrecht, 1991.
-
-\bibitem[Kondrashov(2003)]{Kondrashov:03}
-A.~S. Kondrashov.
-\newblock {Accumulation of {Dobzhansky}--{Muller} incompatibilities within a
-  spatially structured population}.
-\newblock \emph{Evolution}, 57\penalty0 (1):\penalty0 151--153, Jan 2003.
-
-\bibitem[Kruckeberg(1951)]{kruckeberg1951intraspecific}
-A.~R. Kruckeberg.
-\newblock Intraspecific variability in the response of certain native plant
-  species to serpentine soil.
-\newblock \emph{American Journal of Botany}, 38\penalty0 (6):\penalty0 pp.
-  408--419, 1951.
-\newblock URL \url{http://www.jstor.org/stable/2438248}.
-
-\bibitem[Lenormand(2002)]{lenormand2002limits}
-T.~Lenormand.
-\newblock Gene flow and the limits to natural selection.
-\newblock \emph{Trends in Ecology \& Evolution}, 17\penalty0 (4):\penalty0 183
-  -- 189, 2002.
-\newblock ISSN 0169-5347.
-\newblock \doi{DOI: 10.1016/S0169-5347(02)02497-7}.
-\newblock URL
-  \url{http://www.sciencedirect.com/science/article/pii/S0169534702024977}.
-
-\bibitem[Levin et~al.(2003)Levin, Muller-Landau, Nathan, and
-  Chave]{levin2003ecology}
-S.~A. Levin, H.~C. Muller-Landau, R.~Nathan, and J.~Chave.
-\newblock The ecology and evolution of seed dispersal: a theoretical
-  perspective.
-\newblock \emph{Annu. Rev. Ecol. Evol. Syst.}, 34:\penalty0 575--604, 2003.
-\newblock URL
-  \url{http://arjournals.annualreviews.org/doi/pdf/10.1146/annurev.ecolsys.34.011802.132428}.
-
-\bibitem[Lou and Yanagida(2006)]{lou2006minimization}
-Y.~Lou and E.~Yanagida.
+\newblock Proceedings of the Royal Society of Edinburgh, Section: A
+  Mathematics. 1989 0;112(3-4):293--318.
+\newblock Available from:
+  \url{http://journals.cambridge.org/article_S030821050001876X}.
+
+\bibitem{lou2006minimization}
+Lou Y, Yanagida E.
 \newblock Minimization of the principal eigenvalue for an elliptic boundary
@@
   dynamics.
-\newblock \emph{Japan J. Indust. Appl. Math.}, 23\penalty0 (3):\penalty0
-  275--292, 2006.
-\newblock ISSN 0916-7005.
-\newblock URL \url{http://projecteuclid.org/euclid.jjiam/1197390801}.
-
-\bibitem[Macnair(1991)]{macnair1991evolution}
-M.~R. Macnair.
-\newblock Why the evolution of resistance to anthropogenic toxins normally
-  involves major gene changes: the limits to natural selection.
-\newblock \emph{Genetica}, 84\penalty0 (3):\penalty0 213--219, 1991.
-\newblock ISSN 0016-6707.
-\newblock \doi{10.1007/BF00127250}.
-\newblock URL \url{http://dx.doi.org/10.1007/BF00127250}.
-
-\bibitem[Martin and Orgogozo(2013)]{martin2013repeated}
-A.~Martin and V.~Orgogozo.
-\newblock The loci of repeated evolution: a catalog of genetic hotspots of
-  phenotypic variation.
-\newblock \emph{Evolution}, 67\penalty0 (5):\penalty0 1235--1250, 2013.
-\newblock ISSN 1558-5646.
-\newblock \doi{10.1111/evo.12081}.
-\newblock URL \url{http://dx.doi.org/10.1111/evo.12081}.
-
-\bibitem[Maruyama(1970)]{maruyama1970fixation}
-T.~Maruyama.
-\newblock On the fixation probability of mutant genes in a subdivided
+\newblock Japan J Indust Appl Math. 2006;23(3):275--292.
+\newblock Available from:
+  \url{http://projecteuclid.org/euclid.jjiam/1197390801}.
+
+\bibitem{NEQwiki}
+NEQwiki. {NEQwiki}, the nonlinear equations encyclopedia; 2013.
+\newblock Accessed December 17, 2014.
+\newblock Available from:
+  \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
+
+\bibitem{jagers1975branching}
+Jagers P.
+\newblock Branching processes with biological applications.
+\newblock Wiley Series in Probability and Statistics: Applied Probability and
+  Statistics Section Series. Wiley; 1975.
+
+\bibitem{geiger1999elementary}
+Geiger J.
+\newblock Elementary New Proofs of Classical Limit Theorems for
+  {Galton}-{Watson} Processes.
+\newblock Journal of Applied Probability. 1999;36(2):pp. 301--309.
+\newblock Available from: \url{http://www.jstor.org/stable/3215457}.
+
+\bibitem{aldous1985exchangeability}
+Aldous DJ.
+\newblock Exchangeability and related topics.
+\newblock In: École d'été de probabilités de {S}aint-{F}lour,
+  {XIII}---1983. vol. 1117 of Lecture Notes in Math. Berlin: Springer; 1985. p.
+  1--198.
+\newblock Available from:
+  \url{http://www.springerlink.com/content/c31v17440871210x/fulltext.pdf}.
+
+\bibitem{donnelly1989continuity}
+Donnelly P, Joyce P.
+\newblock Continuity and weak convergence of ranked and size-biased
+  permutations on the infinite simplex.
+\newblock Stochastic Process Appl. 1989;31(1):89--103.
+\newblock Available from:
+  \url{http://www.sciencedirect.com/science/article/pii/030441498990104X}.
+
+\bibitem{maynardsmith1974hitchhiking}
+{Maynard Smith} J, Haigh J.
+\newblock The hitch-hiking effect of a favourable gene.
+\newblock Genet Res. 1974 Feb;23(1):23--35.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/4407212}.
+
+\bibitem{barton2013genetic}
+Barton NH, Etheridge AM, Kelleher J, Véber A.
+\newblock Genetic hitchhiking in spatially extended populations.
+\newblock Theoretical Population Biology. 2013;(0):--.
+\newblock Available from:
+  \url{http://www.sciencedirect.com/science/article/pii/S0040580912001359}.
+
+\bibitem{barton1979geneflow}
+Barton N.
+\newblock Gene flow past a cline.
+\newblock Heredity. 1979 Dec;43(3):333--339.
+\newblock Available from: \url{http://dx.doi.org/10.1038/hdy.1979.86}.
+
+\bibitem{borodin2002handbook}
+Borodin AN, Salminen P.
+\newblock Handbook of Brownian motion: facts and formulae.
+\newblock Springer; 2002.
+
+\bibitem{nachman2003genetic}
+Nachman MW, Hoekstra HE, D'Agostino SL.
+\newblock The genetic basis of adaptive melanism in pocket mice.
+\newblock Proc Natl Acad Sci U S A. 2003 Apr;100(9):5268--5273.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/12704245}.
+
+\bibitem{hoekstra2004local}
+Hoekstra HE, Krenz JG, Nachman MW.
+\newblock Local adaptation in the rock pocket mouse (\textit{{Chaetodipus}
+  intermedius}): natural selection and phylogenetic history of populations.
+\newblock Heredity. 2005 Nov;94(2):217--228.
+\newblock Available from: \url{http://dx.doi.org/10.1038/sj.hdy.6800600}.
+
+\bibitem{hoekstra2004ecological}
+Hoekstra HE, Drumm KE, Nachman MW.
+\newblock Ecological genetics of adaptive color polymorphism in pocket mice:
+  geographic variation in selected and neutral genes.
+\newblock Evolution. 2004 Jun;58(6):1329--1341.
+\newblock Available from:
+  \url{http://onlinelibrary.wiley.com/doi/10.1111/j.0014-3820.2004.tb01711.x/abstract}.
+
+\bibitem{french1968dispersal}
+French NR, Tagami TY, Hayden P.
+\newblock Dispersal in a Population of Desert Rodents.
+\newblock Journal of Mammalogy. 1968;49(2):pp. 272--280.
+\newblock Available from: \url{http://www.jstor.org/stable/1377984}.
+
+\bibitem{allred1963range}
+Allred DM, Beck DE.
+\newblock Range of Movement and Dispersal of Some Rodents at the {Nevada}
+  Atomic Test Site.
+\newblock Journal of Mammalogy. 1963;44(2):pp. 190--200.
+\newblock Available from: \url{http://www.jstor.org/stable/1377452}.
+
+\bibitem{Hoekstra:06}
+Hoekstra HE.
+\newblock {Genetics, development and evolution of adaptive pigmentation in
+  vertebrates}.
+\newblock Heredity (Edinb). 2006 Sep;97(3):222--234.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/16823403}.
+
+\bibitem{MesserPetrov}
+Messer PW, Petrov DA.
+\newblock {{P}opulation genomics of rapid adaptation by soft selective sweeps}.
+\newblock Trends Ecol Evol (Amst). 2013 Nov;28(11):659--669.
+
+\bibitem{wilson2014selective}
+Wilson BA, Petrov DA, Messer PW.
+\newblock Soft selective sweeps in complex demographic scenarios.
+\newblock Genetics. 2014 Oct;198(2):669--684.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/25060100}.
+
+\bibitem{ralph2014standing}
+Ralph PL, Coop G.
+\newblock The role of standing variation in geographic convergent adaptation.
+\newblock The American Naturalist. 2015;.
+
+\bibitem{slatkin1998hitchhiking}
+Slatkin M, Wiehe T.
+\newblock Genetic hitch-hiking in a subdivided population.
+\newblock Genet Res. 1998 Apr;71(2):155--160.
+\newblock Available from: \url{http://www.ncbi.nlm.nih.gov/pubmed/9717437}.
+
+\bibitem{kim2011hitchhiking}
+Kim Y, Maruki T.
+\newblock Hitchhiking effect of a beneficial mutation spreading in a subdivided
   population.
-\newblock \emph{Genetics Research}, 15\penalty0 (02):\penalty0 221--225, 1970.
-\newblock \doi{10.1017/S0016672300001543}.
-\newblock URL \url{http://dx.doi.org/10.1017/S0016672300001543}.
-
-\bibitem[{Maynard Smith} and Haigh(1974)]{maynardsmith1974hitchhiking}
-J.~{Maynard Smith} and J.~Haigh.
-\newblock The hitch-hiking effect of a favourable gene.
-\newblock \emph{Genet Res}, 23\penalty0 (1):\penalty0 23--35, Feb. 1974.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/4407212}.
-
-\bibitem[Messer and Petrov(2013)]{MesserPetrov}
-P.~W. Messer and D.~A. Petrov.
-\newblock {{P}opulation genomics of rapid adaptation by soft selective sweeps}.
-\newblock \emph{Trends Ecol. Evol. (Amst.)}, 28\penalty0 (11):\penalty0
-  659--669, Nov 2013.
-
-\bibitem[Nachman et~al.(2003)Nachman, Hoekstra, and
-  D'Agostino]{nachman2003genetic}
-M.~W. Nachman, H.~E. Hoekstra, and S.~L. D'Agostino.
-\newblock The genetic basis of adaptive melanism in pocket mice.
-\newblock \emph{Proc Natl Acad Sci U S A}, 100\penalty0 (9):\penalty0
-  5268--5273, Apr. 2003.
-\newblock \doi{10.1073/pnas.0431157100}.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/12704245}.
-
-\bibitem[Nagylaki(1975)]{nagylaki1975conditions}
-T.~Nagylaki.
-\newblock Conditions for the existence of clines.
-\newblock \emph{Genetics}, 80\penalty0 (3):\penalty0 595--615, 1975.
-\newblock URL \url{http://www.genetics.org/content/80/3/595.abstract}.
-
-\bibitem[Nathan et~al.(2008)Nathan, Schurr, Spiegel, Steinitz, Trakhtenbrot,
-  and Tsoar]{nathan2008mechanisms}
-R.~Nathan, F.~M. Schurr, O.~Spiegel, O.~Steinitz, A.~Trakhtenbrot, and
-  A.~Tsoar.
+\newblock Genetics. 2011 Sep;189(1):213--226.
+\newblock Available from:
+  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3176130/}.
+
+\bibitem{barton2000genetic}
+Barton NH.
+\newblock Genetic hitchhiking.
+\newblock Philos Trans R Soc Lond B Biol Sci. 2000 Nov;355(1403):1553--1562.
+\newblock Available from:
+  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC1692896/}.
+
+\bibitem{reynolds2009levy}
+Reynolds AM, Rhodes CJ.
+\newblock The {L{é}vy} flight paradigm: random search patterns and mechanisms.
+\newblock Ecology. 2009 Apr;90(4):877--887.
+\newblock Available from:
+  \url{http://www.esajournals.org/doi/abs/10.1890/08-0153.1}.
+
+\bibitem{censky1998overwater}
+Censky EJ, Hodge K, Dudley J.
+\newblock Over-water dispersal of lizards due to hurricanes.
+\newblock Nature. 1998 Oct;395(6702):556--556.
+\newblock Available from: \url{http://dx.doi.org/10.1038/26886}.
+
+\bibitem{nathan2008mechanisms}
+Nathan R, Schurr FM, Spiegel O, Steinitz O, Trakhtenbrot A, Tsoar A.
 \newblock Mechanisms of long-distance seed dispersal.
-\newblock \emph{Trends in Ecology \& Evolution}, 23\penalty0 (11):\penalty0 638
-  -- 647, 2008.
-\newblock ISSN 0169-5347.
-\newblock \doi{http://dx.doi.org/10.1016/j.tree.2008.08.003}.
-\newblock URL
+\newblock Trends in Ecology \& Evolution. 2008;23(11):638 -- 647.
+\newblock Available from:
   \url{http://www.sciencedirect.com/science/article/pii/S0169534708002723}.
 
-\bibitem[NEQwiki(2013)]{NEQwiki}
-NEQwiki.
-\newblock {NEQwiki}, the nonlinear equations encyclopedia, 2013.
-\newblock URL
-  \url{http://www.primat.mephi.ru/wiki/ow.asp?Korteweg-de_Vries_equation}.
-\newblock Accessed December 17, 2014.
-
-\bibitem[Orr and Betancourt(2001)]{orr2001sieve}
-H.~A. Orr and A.~J. Betancourt.
-\newblock Haldane's sieve and adaptation from the standing genetic variation.
-\newblock \emph{Genetics}, 157\penalty0 (2):\penalty0 875--884, Feb. 2001.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/11157004}.
-
-\bibitem[Pennings and Hermisson(2006)]{softsweepsII}
-P.~S. Pennings and J.~Hermisson.
-\newblock Soft sweeps {II} - molecular population genetics of adaptation from
-  recurrent mutation or migration.
-\newblock \emph{Mol Biol Evol}, page msj117, 2006.
-\newblock \doi{10.1093/molbev/msj117}.
-\newblock URL
-  \url{http://mbe.oxfordjournals.org/cgi/content/abstract/msj117v1}.
-
-\bibitem[Pollak(1966)]{pollak1966survival}
-E.~Pollak.
-\newblock On the survival of a gene in a subdivided population.
-\newblock \emph{Journal of Applied Probability}, 3\penalty0 (1):\penalty0
-  142--155, 1966.
-\newblock ISSN 00219002.
-\newblock URL \url{http://www.jstor.org/stable/3212043}.
-
-\bibitem[Ralph and Coop(2010)]{ralph2010parallel}
-P.~Ralph and G.~Coop.
-\newblock Parallel adaptation: One or many waves of advance of an advantageous
-  allele?
-\newblock \emph{Genetics}, 186\penalty0 (2):\penalty0 647--668, 2010.
-\newblock \doi{10.1534/genetics.110.119594}.
-\newblock URL \url{http://www.genetics.org/cgi/content/abstract/186/2/647}.
-
-\bibitem[Ralph and Coop(2015)]{ralph2014standing}
-P.~L. Ralph and G.~Coop.
-\newblock The role of standing variation in geographic convergent adaptation.
-\newblock \emph{The American Naturalist}, 2015.
-\newblock \doi{10.1101/009803}.
-
-\bibitem[Reynolds and Rhodes(2009)]{reynolds2009levy}
-A.~M. Reynolds and C.~J. Rhodes.
-\newblock The {L{é}vy} flight paradigm: random search patterns and mechanisms.
-\newblock \emph{Ecology}, 90\penalty0 (4):\penalty0 877--887, Apr. 2009.
-\newblock URL \url{http://www.esajournals.org/doi/abs/10.1890/08-0153.1}.
-
-\bibitem[Rosenblum et~al.(2010)Rosenblum, Römpler, Schöneberg, and
-  Hoekstra]{rosenblum2010molecular}
-E.~B. Rosenblum, H.~Römpler, T.~Schöneberg, and H.~E. Hoekstra.
-\newblock Molecular and functional basis of phenotypic convergence in white
-  lizards at {W}hite {S}ands.
-\newblock \emph{Proceedings of the National Academy of Sciences}, 107\penalty0
-  (5):\penalty0 2113--2117, 2010.
-\newblock \doi{10.1073/pnas.0911042107}.
-\newblock URL \url{http://www.pnas.org/content/107/5/2113.abstract}.
-
-\bibitem[Schat et~al.(1996)Schat, Vooijs, and Kuiper]{schat1996identical}
-H.~Schat, R.~Vooijs, and E.~Kuiper.
-\newblock Identical major gene loci for heavy metal tolerances that have
-  independently evolved in different local populations and subspecies of
-  \textit{{Silene} vulgaris}.
-\newblock \emph{Evolution}, 50\penalty0 (5):\penalty0 pp. 1888--1895, 1996.
-\newblock URL \url{http://www.jstor.org/stable/2410747}.
-
-\bibitem[Sexton et~al.(2013)Sexton, Hangartner, and
-  Hoffmann]{sexton2013genetic}
-J.~P. Sexton, S.~B. Hangartner, and A.~A. Hoffmann.
+\bibitem{cantrell1991diffusive}
+Cantrell RS, Cosner C.
+\newblock Diffusive logistic equations with indefinite weights: population
+  models in disrupted environments. {II}.
+\newblock SIAM J Math Anal. 1991;22(4):1043--1064.
+\newblock Available from: \url{http://dx.doi.org/10.1137/0522068}.
+
+\bibitem{sexton2013genetic}
+Sexton JP, Hangartner SB, Hoffmann AA.
 \newblock Genetic isolation by environment or distance: Which pattern of gene
   flow is most common?
-\newblock \emph{Evolution}, 2013.
-\newblock ISSN 1558-5646.
-\newblock \doi{10.1111/evo.12258}.
-\newblock URL \url{http://dx.doi.org/10.1111/evo.12258}.
-
-\bibitem[Slatkin(1973)]{slatkin1973geneflow}
-M.~Slatkin.
-\newblock Gene flow and selection in a cline.
-\newblock \emph{Genetics}, 75\penalty0 (4):\penalty0 733--756, 1973.
-\newblock URL \url{http://www.genetics.org/cgi/content/abstract/75/4/733}.
-
-\bibitem[Slatkin and Wiehe(1998)]{slatkin1998hitchhiking}
-M.~Slatkin and T.~Wiehe.
-\newblock Genetic hitch-hiking in a subdivided population.
-\newblock \emph{Genet Res}, 71\penalty0 (2):\penalty0 155--160, Apr. 1998.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/9717437}.
-
-\bibitem[Steiner et~al.(2009)Steiner, Rompler, Boettger, Schoneberg, and
-  Hoekstra]{steiner2009genetic}
-C.~C. Steiner, H.~Rompler, L.~M. Boettger, T.~Schoneberg, and H.~E. Hoekstra.
-\newblock The genetic basis of phenotypic convergence in beach mice: Similar
-  pigment patterns but different genes.
-\newblock \emph{Mol Biol Evol}, 26\penalty0 (1):\penalty0 35--45, 2009.
-\newblock \doi{10.1093/molbev/msn218}.
-\newblock URL \url{http://mbe.oxfordjournals.org/cgi/content/abstract/26/1/35}.
-
-\bibitem[Stern(2013)]{stern2013genetic}
-D.~L. Stern.
-\newblock The genetic causes of convergent evolution.
-\newblock \emph{Nat Rev Genet}, 14\penalty0 (11):\penalty0 751--764, Nov. 2013.
-\newblock \doi{10.1038/nrg3483}.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/24105273}.
-
-\bibitem[Turner et~al.(2010)Turner, Bourne, Von~Wettberg, Hu, and
-  Nuzhdin]{turner2010serpentine}
-T.~L. Turner, E.~C. Bourne, E.~J. Von~Wettberg, T.~T. Hu, and S.~V. Nuzhdin.
-\newblock Population resequencing reveals local adaptation of
-  \textit{{Arabidopsis} lyrata} to serpentine soils.
-\newblock \emph{Nat. Genet.}, 42\penalty0 (3):\penalty0 260--3, Jan. 2010.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/20101244}.
-
-\bibitem[Wilson et~al.(2014)Wilson, Petrov, and Messer]{wilson2014selective}
-B.~A. Wilson, D.~A. Petrov, and P.~W. Messer.
-\newblock Soft selective sweeps in complex demographic scenarios.
-\newblock \emph{Genetics}, 198\penalty0 (2):\penalty0 669--684, Oct. 2014.
-\newblock \doi{10.1534/genetics.114.165571}.
-\newblock URL \url{http://www.ncbi.nlm.nih.gov/pubmed/25060100}.
-
-\bibitem[Zhen et~al.(2012)Zhen, Aardema, Medina, Schumer, and
-  Andolfatto]{zhen2012parallel}
-Y.~Zhen, M.~L. Aardema, E.~M. Medina, M.~Schumer, and P.~Andolfatto.
-\newblock Parallel molecular evolution in an herbivore community.
-\newblock \emph{Science}, 337\penalty0 (6102):\penalty0 1634--1637, 2012.
-\newblock \doi{10.1126/science.1226630}.
-\newblock URL \url{http://www.sciencemag.org/content/337/6102/1634.abstract}.
+\newblock Evolution. 2013;Available from:
+  \url{http://dx.doi.org/10.1111/evo.12258}.
+
+\bibitem{Kondrashov:03}
+Kondrashov AS.
+\newblock {Accumulation of {Dobzhansky}--{Muller} incompatibilities within a
+  spatially structured population}.
+\newblock Evolution. 2003 Jan;57(1):151--153.
+
+\bibitem{KPP1937book}
+Kolmogorov A, Petrovskii I, Piscunov N.
+\newblock A study of the equation of diffusion with increase in the quantity of
+  matter, and its application to a biological problem.
+\newblock In: Selected Works of A.N. Kolmogorov: Mathematics and mechanics.
+  vol.~25 of Mathematics and its Applications (Soviet Series). Dordrecht:
+  Kluwer Academic Publishers Group; 1991. p. 1--25.
+
+\bibitem{fisher1937wave}
+Fisher R.
+\newblock {The wave of advance of advantageous genes}.
+\newblock Ann Eugenics. 1937;7:353--369.
+\newblock Available from:
+  \url{http://digital.library.adelaide.edu.au/coll/special/fisher/152.pdf}.
+
+\bibitem{etheridge2000introduction}
+Etheridge AM.
+\newblock An introduction to superprocesses. vol.~20 of University Lecture
+  Series.
+\newblock Providence, RI: American Mathematical Society; 2000.
+
+\bibitem{dawson1993measurevalued}
+Dawson DA.
+\newblock Measure-valued {Markov} processes.
+\newblock In: École d'Été de Probabilités de Saint-Flour XXI---1991. vol.
+  1541 of Lecture Notes in Math. Berlin: Springer; 1993. p. 1--260.
+
+\bibitem{gradshteyn2007table}
+Gradshteyn IS, Ryzhik IM.
+\newblock Table of integrals, series, and products.
+\newblock Seventh ed. Elsevier/Academic Press, Amsterdam; 2007.
+\newblock Translated from the Russian, Translation edited and with a preface by
+  Alan Jeffrey and Daniel Zwillinger.
 

### 2015-10-19 patchy-sim-writeup-plots.R

@@ -3,6 +3,11 @@ source("sim-patchy-selection-fns.R")
 source("lineages.R")
 require(gsl)
 
+plotfns <- c(
+             pdf=function (...) { pdf(...,pointsize=10) },
+             tiff=function (...) { tiff(...,pointsize=10,res=288,units="in") }
+         )
+
 # yes, the parentheses differ in the exponents for x:
 expl.approx <- function (x,d) { sqrt(pi/2) * x^((1-d)/2) * exp(-x) }
 bessel.approx <- function (x,d) { x^(1-(d/2)) * bessel_Knu( nu=1-d/2, x=x ) } 
@@ -46,37 +51,42 @@ plotdecay <- function (pophist, shift=FALSE) {
 plotinfo <- lapply( c("3994-1000-1000-pophistory-run.Rdata","7705124-r101-101-sb0.05-sm-0.005-pophistory-run.Rdata"), function (run.name) {
     load(run.name); plotdecay(pophist) } )
 
-pdf(file="sim-occupation-freqs.pdf",width=6,height=3,pointsize=10)
-layout(t(1:2))
-par(mar=c(5,4,1,1)+.1)
-# averaged
-plot( 1, 1, xlim=range(unlist(lapply(plotinfo,"[[","patchloc"))), ylim=range(unlist(lapply(plotinfo,"[[","obs.freqs"))), 
-        log='y', xlab='distance from patch', ylab='allele frequency', type='n' )
-mtext("A",side=3,line=0.1,cex=1.5,adj=-0.3,padj=1)
-abline(v=0,lty=2)
-lapply( plotinfo, function (x) { with(x, {
-        lines( patchloc, obs.freqs, pch=20, cex=.5, col=grey(.7) )
-        lines( plotlocs, bessel.const * bessel.approx( plotlocs*theory.decay, d=dimension ), col=c('red','green')[dimension] )
-        points( plotlocs, tapply(obs.freqs,patchdist,mean), pch=20, cex=.5 )
-        if (FALSE) {  lines( plotlocs, bessel.shift.const * bessel.approx( (plotlocs+bessel.shift$minimum)*theory.decay, d=dimension ), lwd=2, lty=2 ) }
-} ) } )
-legend('topright',legend=paste("d=",c(1,2)),lty=1,col=c('red','green'))
-# and, snapshots
-run.name <- "3975-1-10000-pophistory-run.Rdata"
[…]

### 2015-10-19 patchy-selection-paper.tex

@@
 as spatial branching processes.
-A main result is that locally adapted alleles in different patches
-are likely to be of independent origin
-if the distance between the patches is larger than
-$(\sigma/s_m)\log(Cs_m/\mu)$, 
-where $\sigma$ is the dispersal distance, 
-$s_m$ is the selective disadvantage of these alleles between patches,
-$\mu$ is the mutation rate, and $C$ is a constant depending on the probability of establishment and patch geometry.
+A main result is that patches further apart than a critical distance will
+likely evolve independent locally adapted alleles;
+this distance is proportional to the spatial scale of selection
+($\sigma/\sqrt{s_m}$, where $\sigma$ is the dispersal distance and
+$s_m$ is the selective disadvantage of these alleles between patches),
+and depends linearly on $\log(s_m/\mu)$, where $\mu$ is the mutation rate.
 This provides a way to understand the role of geographic separation between patches 

### 2015-10-19 patchy-selection-paper-for-PLoS.tex
final?

@@
 When we condition on survival until $t$, we condition on survival of at least one trunk lineage
-(the long red lineage in Fig \ref{fig:branching_decomp}).
+(the long red lineage in Fig~\ref{fig:branching_decomp}).
 Given this lineage, the other reproduction events are undistinguished --
@@
 of width $\sigma/\sqrt{2s_m}$ to the original patch;
-see Fig \ref{fig:branching_decomp}, and section \secref{apx:qS} for more details.
+see Fig~\ref{fig:branching_decomp}, and section \secref{apx:qS} for more details.
 In either case, denote this area $A'$;
@@
       the remaining combinations are shown in 
-      S\ref{sfig:sim_migration_times} and S\ref{sfig:sim_mutation_times} Fig.
+      \ref{sfig:sim_migration_times} and \ref{sfig:sim_mutation_times} Fig.
   }   \label{fig:sim_times}
@@
 
-We tested this using the same simulations as Fig \ref{fig:sim_times},
+We tested this using the same simulations as Fig~\ref{fig:sim_times},
 by using the empirical distributions of the respective times to adaptation
@@
       give the probability predicted by theory (equation~\eqref{eqn:parallel_prob}, with $C=5$).
-      Points correspond to parameter combinations of the simulations in Fig \ref{fig:sim_times},
+      Points correspond to parameter combinations of the simulations in Fig~\ref{fig:sim_times},
       with the shade of grey and the red numeric label indicating the same probability
@@
 by following only the motion of the ``trunk''
-(i.e., the red line in Fig \ref{fig:branching_decomp}),
+(i.e., the red line in Fig~\ref{fig:branching_decomp}),
 and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
@@
 as the sum of independent migrant families, each of whose dynamics are
-given by a spatial branching process (as depicted in Fig \ref{fig:branching_decomp}).
+given by a spatial branching process (as depicted in Fig~\ref{fig:branching_decomp}).
 Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
@@
 %
-% \appendix
-% \section*{Supporting Information}
-% 
-% \setcounter{secnumdepth}{2}
+
+\appendix
+\renewcommand{\thesubsection}{S\arabic{subsection}}
+\setcounter{secnumdepth}{2}
 
@@
 
-\subsection*{S1 Fig}
+% S1
+\subsection{Fig}
 \label{sfig:sim_migration_times}
@@
 
-\subsection*{S2 Fig}
+% S2 
+\subsection{Fig}
 \label{sfig:sim_mutation_times}
@@
 % example trace plots
-
-\subsection*{S3 Fig}
+% S3 
+\subsection{Fig}
 \label{sfig:sims_1}
@@
 
-\subsection*{S4 Fig}
+% S4 
+\subsection{Fig}
 \label{sfig:sims_2}
@@
 
-\subsection*{S5 Fig}
+% S5 
+\subsection{Fig}
 \label{sfig:sims_3}
@@
 
-\subsection*{S6 Fig}
+% S6 
+\subsection{Fig}
 \label{sfig:sims_4}
@@
 
-\subsection*{S7 Fig}
+% S7 
+\subsection{Fig}
 \label{sfig:sims_5}
@@
 
-\subsection*{S8 Fig}
+% S8 
+\subsection{Fig}
 \label{sfig:sims_6}
@@
 % \end{sfigure}
-
-\subsection*{S9 Fig}
+% S9 
+\subsection{Fig}
 \label{sfig:sims_7}
@@
 % \end{sfigure}
-
-\subsection*{S10 Fig}
+% S10 
+\subsection{Fig}
 \label{sfig:sims_8}
